### 2010-02-17 doc/thesis/personal.tex
Merge ssh://oldcub.ath.cx/home/micahc/brainid

\documentclass[12pt]{article}
\usepackage[left=1in,top=1in,right=1in,nohead]{geometry}

\usepackage[]{times}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
%\usepackage{setspace}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
:Title: Personal Statement

\end{comment}

\author{Micah Chambers}
\title{Personal Statement}

\begin{document}

%Why are you fascinated by your research area?

%What examples of leadership skills and unique
% characteristics do you bring to your chosen field?
%What personal and individual strengths do you have
% to make you a qualified applicant?

%How will receiving the fellowship contribute to your
% career goals.
\Large
\begin{multicols}{2}
\begin{flushleft}
Micah Chambers
\end{flushleft}

\begin{flushright}
Personal Statement
\end{flushright}
\end{multicols}
\normalsize

Among biomedical research areas, Neurology is probably
audacious, and in my opinion the most interesting. While many
biological systems have a thousand or more variables, the brain
has more than 100 billion. In fact, our brains are so complex, super
computers can only simulate
1 ten-millionth of the neurons. It would take another
30 years of Moore's law to scale raw computing
power to the level of a single human brain.
So the question for Neuroscientists is this: given that we cannot
simulate an entire brain, what can we do? Surprisingly, the answer is plenty.
While neuron-level simulation is a long way off,
its possible that such an analysis may not be necessary, or even
productive. By studying small \emph{regions} of the brain, 
we can still learn a great deal about how the brain
works, and how at times it doesn't work.

One reason that neuroscience is such an
exciting field is that it is still wide open.
For instance, we still have no way to accurately
detect small activations, no way to generically
decode activation patterns, and no standard
way to determine causality. As an engineer, problem solving
is a passion; and neuroscience certainly provides plenty of problems.
Overcoming these difficulties requires both new techniques
and new ways of applying old ones.  Designing novel methods to present
data is also crucial; the traditional "activation map" is intuitive, 
but how does one show data like causality
- or switching - in an intuitive manner that doesn't 
oversimplify? All these
problems have solutions, and it is a fascinating
endeavor to search for them.

Neuroscience is more than interesting though, it
can actually make a difference in people's lives. 
MRI, MEG and EEG can all lead to a better 
understanding of why people suffer from various conditions.
Though the benefits aren't immediate,
the ability to diagnose and understand neurological conditions
could eventually lead to methods of prevention or treatment.
Further, many conditions long thought
to be localized may in fact stem from neurological conditions;
after all, the nervous system affects every part of the body.  
Of course neuroimaging doesn't just benefit the ill, it
benefits all of humanity by leading us to a 
greater understanding of
ourselves. Why do some people have a propensity for
alcoholism? What conditions would give one person 
photographic memory and another virtually no memory? 
The answer to these questions all reside somewhere within 
the brain, waiting for use to find them. The Chinese General,
Sun Tzu, said that if you know your enemies and you
know yourself, you will not be imperiled by a hundred battles;
this is the ultimate goal of neuroscience: to know the things
that harm us, and to know ourselves.

%\section*{Why not}
\bigskip
The fact that engineering is not an end unto itself
is actually one of the reasons I find it so fascinating. 
Engineering is about designing a solution to a problem, but this
often requires learning a new field well enough
to propose solutions for it. Throughout my career, I have
attempted to broaden my horizons: as an undergraduate
I minored in
math, economics and computer science because I enjoyed considering
different approaches to problems. When I was working 
as a computer engineer in Maryland, I took a 6 hour course
in Japanese at George Mason University, something 
I hope to take up again when my schedule permits it.
In Egypt, I took an Egyptian studies course;
learning about the countries' long history, its people
and even it's politics. When working in a field as diverse as Biomedical
Engineering, it is important to be have broad experiences
because at any point these experiences may become relevant.
For instance, many of the analysis tools used for neural
connectivity and
activation originated in economics, and neural networks were
first trained not by software engineers but by sociologists.

%\section*{teamwork}
\bigskip
Last summer, I also took a robotics course in Egypt (through
Virginia Tech), and
halfway through the semester our team had still 
not been given the instructions for our semester project.
Finally, as a team, we decided on our own final project.
As the most experienced electronics engineer on the team,
I took the lead of the electronics group. I mention this
because my experiences with that team exemplify two
important leadership qualities: cooperation and management.

There is a definite leadership quality to cooperation; 
most people spend their entire lives as 
subordinates to \emph{someone}, but this shouldn't preclude leadership.
Willingness to speak up in a group setting, and at times partake
in vigorous discussion are important parts of teamwork. Of course
willingness to make concessions and admit fault are imperative as well.
The robotics team discussed what
type of project we wanted to complete: we debated the merits of
easier goals versus ambitious ones and we did so without a specific
leader. Being able to take part in such a team is, in of itself,
a leadership skill. I had a similar experience in my Operating
Systems class: because there were only three of us on the team,
we were able to divide responsibilities without specifying a leader.
Of course, this presupposes everyone is willing to discuss
disputes so they may be resolved properly. Success of 
such groups often depends on group dynamics, so this type of
group isn't always possible.  Thus there are times when one person
ought to be in charge. When I took the lead of the electronics
team in Egypt, I did so because everyone agreed I was the best 
equipped for the task. I generally knew how the parts went together,
and I was able to delegate tasks because of that. During my
final semester as an undergraduate, I took a Bioinformatics course,
in which we had a semester long project.
I started working on the project early 
and although I wasn't the assigned leader, I 
eventually began delegating tasks. As the only team member
with a working knowledge of the project, it was my duty to
step up.  As a TA for embedded systems last spring, I acted
as the project manager for five teams of four. In that case, I
delegated very little, but I provided the team with direction
and incentives to keep on top of their work.
In the end, the type of leadership that is necessary depends on the 
circumstances, and my diverse experience has definitely
helped me understand leadership better.

Concluding, I believe that I am uniquely qualified to 
research Biomedical Imaging. Besides taking courses in
physiology and medical imaging, I have leadership
experience which will help me work in a team,
and diverse interests which which help me learn new fields.
%There are two primary ways that this fellowship will aid me.
%First it will be 
%helpful to have access to the computing resources. Every
%type of neural analysis requires computers.
%Testing new algorithms is imperative, but it is time consuming
%even on fast machines; thus is would be useful to have
%access to the TeraGrid. 
%Secondly this fellowship would allow me to spend more
%time developing a broad knowledge base, rather than having to work to pay
%for school. As I mentioned, advances in neuroscience often come from
%diverse fields, and thus requires a diverse knowledge set.
%Studying how other fields solve similar problems is crucial
%in Bioimaging, so I prefer to read 
%not just Engineering or Neuroscience journals, but 
%economics and biology journals as well. Having this fellowship
%would certainly give me more time to do that.

\end{document}

### 2010-02-17 doc/thesis/proposal.tex
Merge ssh://oldcub.ath.cx/home/micahc/brainid

\documentclass[12pt]{article}
\usepackage[left=1in,top=1in,right=1in,nohead]{geometry}

\usepackage[]{times}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{mdwlist}
%\usepackage{setspace}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
:Title: Personal Statement

\end{comment}

\author{Micah Chambers}
\title{Research Experience}

%What are all of your applicable experiences?
%For each experience, what were the key questions, methodology, findings, and conclusions?
%Did you work in a team and/or independently?
%How did you assist in the analysis of results?

\begin{document}
\Large
\begin{center}
Automatic Neural Causality Analysis Using Granger Causality Mapping and Mutual Information

\large
By: Micah Chambers
\end{center}
\normalsize

Recently there has been great interest in the area of causality
mapping of Neural Networks. While this development
is crucial to understanding how the brain works,
there are still many problems with the existing techniques.
Here I propose a new processing pipeline and a new method of performing
causality mapping that makes fewer assumptions 
and is still computationally tractable. The goal then is to
develop a set of tools that will:

\begin{enumerate*}
\item Find independent regions to input into the Granger Causality Mapping (GCM) algorithm.
\item Calculate the causality within the independent regions.
\item Create a dependency tree for intuitive display of causality.
\end{enumerate*}

%\subsubsection*{Background}
For the past ten years the primary method of analysis of Functional
MRI has been Statistical Parametric Mapping (SPM) using 
the General Linear Model (GLM). While this method
has been effective, in the interest of computational tractability
it makes certain assumptions that do not hold. For this reason, such
methods must be extremely conservative to avoid false positives, which means
that SPM is incapable of detecting small activation regions or regions of low 
activation. Additionally, SPM
only measures how well the data fits the GLM and not the actual activation
level. Finally, SPM is a univariate model, meaning it cannot determine
inter-voxel causality.  Because of these limitations, there is currently 
a lot of work being done to find the successor to SPM.

Recently, "Dynamic Causal Modeling" (DCM) has been introduced, which attempts 
model activation of neural pathways.
Unfortunately DCM has not become anywhere near as ubiquitous a SPM, in no
small part because it is significantly more difficult to use. Because DCM works by 
testing FMRI data against a proposed network, the researcher must propose
a functional layout to perform DCM, which is often difficult.

Another recent development, Vector Autoregression has many desirable
features; it is able to infer dependencies, it requires no a priori knowledge of the system,
and versions exist for nonlinear systems \cite{NVAR}. 
Vector Autoregression is actually a generalized version of Granger Causality Mapping (GCM),
which Zhou et. al. \cite{granger} recently showed to be effective for
finding causal neural networks. The downside is computation time, which can
be extreme when performed on every pair of voxels. Previous studies have focused on small
regions, region-wise causality, or Principal Component Analysis (PCA)
to circumvent the computation issues. On the other hand, I propose
performing full, voxel-level GCM by minimizing the number of unrelated voxels
that are processed, without resorting to linear or Gaussian assumptions.

\bigskip
%\subsubsection*{Proposed Solution}
To minimize the number of non-related voxels that are processed by GCM, I propose
a preprocessing stage that breaks voxels up into independent blocks. Ordinarily
GCM is $O(N^2TD)$ where $N$ is the number of voxels, $T$ is the
number of timepoints, and $D$ is the number of time delays allowed. By breaking
the brain into independent blocks, computation only has to be performed within each
independent block, rather than between every voxel. 
Thus instead of the algorithm being $O(N^2TD)$ it will be $O(P^2TD)$ where $P \ll N$. 
To determine which regions are independent, I propose
the use of mutual information. Unlike PCA or clustering algorithms, 
Mutual Information is based
on Bayesian statistics and makes no assumptions about the underlying distributions.
Moreover, it has recently been demonstrated that mutual information works for larger 
regions of the brain \cite{regionalMI}. Therefore, it is possible to break the
region of interest up into blocks and perform mutual information between the blocks,
rather than individual voxels. Of course there is a trade off between larger blocks
and smaller blocks: larger blocks may miss 1 or 2 voxels that are interdependent with
outside voxels, whereas smaller blocks may be intractable. With the mutual
information between blocks known, a threshold can then be applied, and blocks with 
mutual information combined into a single independent block.
Once a set of independent blocks has been found, GCM may be applied to the internals
of each block without fear of missing voxel-wise causality.

The final step of the pipeline is to build a causation map, which can be displayed
to the user. One inherent problem with any technique measuring causation is 
intermediate terms. For instance, if $A = f(B)$, $B = g(C, D, E, F)$, then
$A$ will appear to be a function of $C$ even if it only a function of $B$.
To combat this problem, and
provide the simplest, most likely network, the algorithm will attempt to account
for intermediate causality. Thus, the algorithm will sever weak connections
when a stronger (albeit longer) path exists. The ultimate output from the algorithm
will be a 
causality graph that can then be simplified for statistical or graphical purposes.
Two potential confounds are regional changes in blood flow and slice timing.
Slice timing will result in causation during resting state, and can be fixed by
accounting for such delays in the model. 
Regional blood flow is more difficult; however, it may be possible to account for
by introducing bias in regions near larger blood vessels. 
Depending on the size of these effect, using time-series subtraction 
may be helpful.

\bigskip
%\subsubsection*{Conclusion}
Concluding, the proposed method will be a comprehensive set of tools to develop
voxel-wise causality maps from FMRI images. The end goal is
a better understanding of how regions of the brain communicate. 
The scientific benefits of understanding neural connections is manifold, both for
clinical and research purposes. In particular, those with connection disorders 
would benefit greatly from this research.
Of course, understanding why one part of the brain triggers another is 
an important step in understanding sentience.
Many of the techniques I have included here have a history of success in the
investigation of neural networks; thus I believe these goals are feasible. 
By building on the knowledge gained from previous studies it
should be possible to avoid pitfalls, and lay the groundwork for future research.

\small
\bibliographystyle{ieeetr}
\bibliography{savedrecs}
\end{document}

### 2010-02-17 doc/thesis/research.tex
Merge ssh://oldcub.ath.cx/home/micahc/brainid

\documentclass[12pt]{article}
\usepackage[left=1in,top=1in,right=1in,nohead]{geometry}

\usepackage[]{times}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
%\usepackage{setspace}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
:Title: Personal Statement

\end{comment}

\author{Micah Chambers}
\title{Research Experience}

%What are all of your applicable experiences?
%For each experience, what were the key questions, methodology, findings, and conclusions?
%Did you work in a team and/or independently?
%How did you assist in the analysis of results?

\begin{document}
\Large
\begin{multicols}{2}
\begin{flushleft}
Micah Chambers
\end{flushleft}

\begin{flushright}
Research Experience
\end{flushright}
\end{multicols}
\normalsize

\subsubsection*{Masters Thesis}
My primary experience in research comes from my masters program at Virginia 
Tech. As a graduate student, beginning in fall 2008, I worked
with Dr. Chris Wyatt on the analysis of FMRI timeseries. The final goal
of the research, which is not yet complete, is to determine the hidden
flow-inducing signal to brain regions, and use
that as input to an Artificial Neural Network (ANN). The reason we are 
performing this step rather than directly using an ANN
is that ANNs of the BOLD signal would
necessarily be more complicated and thus less comparable between
individuals. Because every patient would have a completely
different ANN, it would be difficult to distinguish
pathologies based on network weights.
Additionally, it is important to minimize the number of
parameters when performing
such pattern recognition techniques, to avoid the bias-variance
dilemma. Preprocessing the data to flow-inducing signal thus
removes degrees  of freedom (such as cerbral blood flow, differences in 
blood vessel compliance, etc.) from the neural network.

To determine the underlying
state of the system we have been using particle filters.
Particle filters make no Gaussian assumptions and actually
are able to estimate the noise distribution in parameters. 
The result of a particle filter is joint 
probability distribution function (PDF) of the parameters
defined by a mixture of impulses. 
Thus, if there are several highly likely sets of parameters that could
give equally valid results, all of them are part of the final
distribution.  This is helpful because it gives a far more
complete picture of the system, and could lead to insight
into the source of noise in the system and how parameters interact. 
Additionally having a full probability
distribution for parameters gives a much better impression of
model confidence.

Though the research is not yet complete, the particle filter
has been robust to noise up to an SNR of 2. In fact, we have
been able to correctly estimate 25 model parameters in the face
of such noise. Currently we are working to
validate the particle filter on real data; something 
previous research failed to do. 
We used a basic particle filter library, that I then modified
for our purposes. Most of this project I have done myself,
with guidance from Dr. Wyatt for direction and concepts.

\subsubsection*{Undergraduate Research}
As an undergraduate student at Virginia Tech, I took part in two
undergraduate research opportunities. 
The first was during the summer after my sophomore year; 
I worked on something called "wireless sensor networks". 
Two of us in the lab worked together for the summer, which 
ended up consisting of a large amounts of troubleshooting
software and not very much "research". I did, however
get exposed to reading papers, and I learned the basics 
of NesC, a C like language used for programming MOTES (the
nodes in wireless sensor networks).

My second chance at undergraduate research was significantly
better in that I learned a great deal more.
For the first half of the summer, 2007, I worked on the
sonar system for Virginia Tech's Autonomous Underwater vehicle.
Every year in San Diego the  
Association for Unmanned Vehicle Systems International (AUVSI)
has a competition for Autonomous Underwater Vehicles. For
this project I worked alone, and developed dsPIC code to detect
a particular pinger among several based on ping period.
Ultimately the board was able to detect underwater pings of 
particular frequencies, however due to on-board noise, the
hydrophones all detected incoming sound at the same time.
Because of this, our board could detect the pinger, 
but it couldn't properly detect its location. 
I wrote the code to perform beam-forming; however, because 
of the interference beam-forming was impossible.

Though ultimately the project didn't work, I learned a good deal
about beam-forming, underwater sound, and, naturally, the importance
of testing individual components.

\end{document}

### 2010-02-17 doc/thesis/plot.py

@@ -1,0 +1,53 @@
+#!/usr/bin/python
+
+import pylab as P
+import sys
+from nifti import *
+
+#t = arange(0.0, 2.0, 0.01)
+#s = sin(2*pi*t)
+#plot(t, s, linewidth=1.0)
+#
+#xlabel('time (s)')
+#ylabel('voltage (mV)')
+#title('About as simple as it gets, folks')
+#grid(True)
+#show()
+
+images = list();
+for arg in sys.argv[1:]:
+    images.append(image.NiftiImage(arg))
+
+leg = []
+
+data = sys.stdin.readlines()
+time = []
+for lineno in range(0,len(data)):
+    data[lineno] = data[lineno].split()
+    time.append(float(data[lineno][0])/2)
+    data[lineno] = float(data[lineno][1])*.0002-.0004
+
+for line in data:
+    print line
+
+for nifti in images:
+    print nifti
+    for iter in range(0,nifti.extent[0]):
+        P.plot(nifti.data[:, 0,0,iter])
+#        leg = leg + [nifti.filename+":"+str(iter)]
+#P.legend(sys.argv[1:])
+P.plot(time, data)
+#P.legend(["BOLD", "With Noise SNR = 2", "Input"])
+
+P.xlabel("Time, Seconds")
+P.ylabel("BOLD Signal %")
+P.show()
+#
+#names = ['Ts', 'Tf', 'epsilon', 'T0', 'alpha', 'E_0', 'V0', 'Vt', 'Qt', 'St', 'Ft']
+#for i in range(0,11):
+#    P.subplot(6, 2, i+1)
+#    line1 = P.plot(truestat.data[:, 0, i, 0])
+#    line2 = P.plot(eststat.data[:, 0, i, 0])
+#    P.ylabel(names[i])
+#
+#P.show()

### 2010-02-17 doc/thesis/plotout.py

@@ -1,0 +1,116 @@
+#!/usr/bin/python
+from math import sqrt
+import sys
+import getopt
+import pylab as P
+from nifti import *
+
+def main(argv):                         
+    truebold = "bold"
+    truestate = "state"
+    estbold = "particle-bold"
+    eststate = "particle-state"
+    saveloc = ""
+    try:                                
+        opts, args = getopt.getopt(argv, "hb:s:y:x:p:", ["help", "bold=", 
+                    "state=", "estbold=", "eststate=", "prefix="])
+    except getopt.GetoptError:
+        usage()
+        sys.exit(2)
+
+    for opt, arg in opts:
+        if opt in ("-h", "--help"):
+            usage();
+            sys.exit();
+        elif opt in ("-b", "--bold"):
+            truebold = arg;
+        elif opt in ("-s", "--state"):
+            truestate = arg;
+        elif opt in ("-x", "--eststate"):
+            eststate = arg;
+        elif opt in ("-y", "--estbold"):
+            estbold = arg;
+        elif opt in ("-p", "--prefix"):
+            saveloc = arg
+
+    truebold = image.NiftiImage(truebold)
+    try: 
+        truestate = image.NiftiImage(truestate)
+    except  getopt.GetoptError:
+        print "Hoops"
+    
+    estbold = image.NiftiImage(estbold)
+    eststate = image.NiftiImage(eststate)
+    print eststate
+    print estbold
+    
+    leg = []
+    time = range(0, 1199, 2)
+    
+    for iter in range(0,truebold.extent[0]):
+        P.plot(time, truebold.data[:, 0,0,iter])
+        leg = leg + ["Simulated BOLD"]
+    
+    for iter in range(0,estbold.extent[0]):
+        bold = estbold.data[:, 0, 0, iter]
+        error = estbold.data[:, 1, 0, iter]
+        for errno in range(0, len(error)):
+            error[errno] = 2*sqrt(error[errno])
+        if(estbold.extent[2] == 2):
+            P.errorbar(time, bold, yerr=error, fmt='r', elinewidth=.5, ecolor='g')
+        else:
+            P.plot(time, estbold.data[:, 0,0,iter], 'b')
+        leg = leg + ["Estimated BOLD"]
+        
+    P.xlabel("Time, Seconds")
+    P.ylabel("BOLD Signal %")
[…]

### 2010-02-17 code/calc_parammap.cpp
changes to how spline is calculated in preprocessing stages.

@@ -261,7 +261,7 @@ int main(int argc, char* argv[])
     } else {
         *output << "De-trending, then dividing by mean" << endl;
         method = BoldPF::DIRECT;
-        inImage = deSpline(inImage, 10.0, input, a_timestep());
+        inImage = deSpline(inImage, 6, input, a_timestep());
     }
 
     /* Save detrended image */

### 2010-02-17 code/include/segment.h

@@ -21,7 +21,7 @@ Image4DType::Pointer pruneFMRI(const Image4DType::Pointer fmri_img,
             unsigned int remove);
 
 Image4DType::Pointer deSpline(const Image4DType::Pointer fmri_img,
-            double min_delay, std::vector<Activation>& stim, double dt);
+            unsigned int numknots, std::vector<Activation>& stim, double dt);
 
 Image4DType::Pointer summ(const Image4DType::Pointer fmri_img, 
             const Label3DType::Pointer labelmap, std::list<LabelType>& sections);
@@ -54,7 +54,7 @@ double get_average(const Image4DType::Pointer fmri_img,
         const Label3DType::Pointer labelmap);
 
 Image4DType::Pointer getspline(const Image4DType::Pointer fmri_img,
-            const std::vector< std::vector<unsigned int> >& knots);
+            const std::vector<unsigned int>& knots);
 
 Image3DType::Pointer get_average(const Image4DType::Pointer fmri_img);
 Image3DType::Pointer get_variance(const Image4DType::Pointer fmri_img);

### 2010-02-17 code/segment.cpp

@@ -6,6 +6,7 @@
 #include <cstdio>
 #include <iomanip>
 #include <vector>
+#include <list>
 
 #include <itkImageFileWriter.h>
 #include <itkGDCMImageIO.h>
@@ -449,7 +450,7 @@ int detrend_avg(const Image4DType::Pointer fmri_img, Image4DType::IndexType inde
 };
 
 int detrend_avg(const Image4DType::Pointer fmri_img, Image4DType::IndexType index, 
-            const std::vector< std::vector<unsigned int> >& knots, 
+            const std::vector< unsigned int >& knots, 
             Image4DType::Pointer output)
 {
     gsl_interp_accel *acc = gsl_interp_accel_alloc();
@@ -459,15 +460,13 @@ int detrend_avg(const Image4DType::Pointer fmri_img, Image4DType::IndexType inde
     double level[knots.size()];
 
     for(unsigned int ii = 0 ; ii < knots.size() ; ii++) {
-        xpos[ii] = 0;
+        xpos[ii] = knots[ii];
         level[ii] = 0;
-        for(unsigned int jj = 0 ; jj < knots[ii].size() ; jj++) {
-            xpos[ii] += knots[ii][jj];
-            index[3] = knots[ii][jj];
+        for(unsigned int jj = -1 ; jj <= 1 ; jj++) {
+            index[3] = knots[ii];
             level[ii] += fmri_img->GetPixel(index);
         }
-        xpos[ii] /= knots[ii].size();
-        level[ii] /= knots[ii].size();
+        level[ii] /= 3.;
     }
 
 //    for(unsigned int ii = 0 ; ii < knots.size() ; ii++){
@@ -486,61 +485,100 @@ int detrend_avg(const Image4DType::Pointer fmri_img, Image4DType::IndexType inde
     return 0;
 }
 
-void getknots(std::vector< std::vector<unsigned int> >& knots, double min_delay, 
-            std::vector<Activation>& stim, double dt, unsigned int length)
+bool compare_l(Activation A, Activation B)
 {
-    //simplify activation vector to only level changes, and filter our
-    //times where level changes rapidly
-    double prev_l = stim[0].level;
-    double prev_t = -min_delay;
-    //generates list of [0, a), [b, c), [d, e) periods
-    for(unsigned int i = 0; i < stim.size() ; i++) { 
-        if(stim[i].level != prev_l) { //trigger
-            if(stim[i].time - prev_t > min_delay) {
-                //translate times to elements
[…]

### 2010-02-17 code/segment.cpp
change in preprocessing.

@@ -499,6 +499,7 @@ bool compare_t(Activation A, Activation B)
 void getknots(std::vector<double>& knots, unsigned int num, 
             std::vector<Activation>& stim)
 {
+    const double RANGE = 20;
     std::list<Activation> fifo;
     std::list<Activation> mins;
     std::vector<Activation>::iterator it = stim.begin();
@@ -506,24 +507,19 @@ void getknots(std::vector<double>& knots, unsigned int num,
     double sum_prev_prev = 1;
     double sum_prev = 0;
     double sum = 0;
-    double time_prev = it->time;
-    while(it->time - frame_begin < 10) {
+    double time_prev = RANGE;
+    while(it->time - frame_begin < RANGE) {
         sum += ((it+1)->time-it->time)*(it->level);
         fifo.push_front(*it);
         it++;
     }
-    sum -= (it-1)->level*(it->time-10);
+    sum -= (it-1)->level*(it->time-RANGE);
 
     Activation tmp;
 
     while(fifo.size() != 1 || it != stim.end()) {
-        printf("Sums: %f %f %f\n", sum_prev_prev, sum_prev, sum);
-        for(std::list<Activation>::iterator tmpit = fifo.begin(); 
-                    tmpit != fifo.end(); tmpit++) {
-            printf("\t%f, %f\n", tmpit->time, tmpit->level);
-        }
-
-        if(sum > sum_prev && sum_prev < sum_prev_prev) {
+        if(sum > sum_prev && sum_prev <= sum_prev_prev) {
+            printf("Adding: %f %f\n", time_prev, sum_prev);
             tmp.time = time_prev;
             tmp.level = sum_prev;
             mins.push_back(tmp);
@@ -531,12 +527,13 @@ void getknots(std::vector<double>& knots, unsigned int num,
 
         sum_prev_prev = sum_prev;
         sum_prev = sum;
-        time_prev = fifo.back().time+10;
+        time_prev = fifo.back().time+RANGE;
 
         std::list<Activation>::reverse_iterator sec = ++fifo.rbegin();
-        if(it != stim.end() && it->time - fifo.back().time - 10 < 
-                    sec->time - fifo.back().time) {
-            double delta = it->time - fifo.back().time - 10;
+        if(fifo.size() <= 1 || ( it != stim.end() && 
+                    it->time - fifo.back().time - RANGE < 
+                    sec->time - fifo.back().time ) ) {
[…]

### 2010-02-17 code/calc_parammap.cpp
cleanup to knot code. Pruning thesis.

@@ -261,7 +261,7 @@ int main(int argc, char* argv[])
     } else {
         *output << "De-trending, then dividing by mean" << endl;
         method = BoldPF::DIRECT;
-        inImage = deSpline(inImage, 6, input, a_timestep());
+        inImage = deSpline(inImage, 15, input, a_timestep());
     }
 
     /* Save detrended image */

### 2010-02-17 code/segment.cpp

@@ -453,24 +453,27 @@ int detrend_avg(const Image4DType::Pointer fmri_img, Image4DType::IndexType inde
             const std::vector< unsigned int >& knots, 
             Image4DType::Pointer output)
 {
-    gsl_interp_accel *acc = gsl_interp_accel_alloc();
-    gsl_spline *spline = gsl_spline_alloc(gsl_interp_cspline, knots.size());
+    static gsl_interp_accel *acc = gsl_interp_accel_alloc();
+    static gsl_spline *spline = gsl_spline_alloc(gsl_interp_cspline, knots.size());
 
     double xpos[knots.size()];
     double level[knots.size()];
-
     for(unsigned int ii = 0 ; ii < knots.size() ; ii++) {
         xpos[ii] = knots[ii];
         level[ii] = 0;
-        for(unsigned int jj = -1 ; jj <= 1 ; jj++) {
-            index[3] = knots[ii];
+        for(int jj = -1 ; jj <= 1; jj++) {
+            index[3] = knots[ii]+jj;
             level[ii] += fmri_img->GetPixel(index);
+//            printf("|%f|", level[ii]);
         }
         level[ii] /= 3.;
+        
+//        printf("Index: %lu %lu %lu %lu %f\n", index[0],index[1],index[2],index[3],
+//                    fmri_img->GetPixel(index));
     }
 
 //    for(unsigned int ii = 0 ; ii < knots.size() ; ii++){
-//        fprintf(stderr, "%f => %f\n", xpos[ii], level[ii]);
+//        printf("%f => %f\n", xpos[ii], level[ii]);
 //    }
 
     itk::ImageLinearIteratorWithIndex< Image4DType > 
@@ -496,7 +499,7 @@ bool compare_t(Activation A, Activation B)
 }
 
 
-void getknots(std::vector<double>& knots, unsigned int num, 
+void getknots(std::list<double>& knots, unsigned int num, 
             std::vector<Activation>& stim)
 {
     const double RANGE = 20;
@@ -519,7 +522,6 @@ void getknots(std::vector<double>& knots, unsigned int num,
 
     while(fifo.size() != 1 || it != stim.end()) {
         if(sum > sum_prev && sum_prev <= sum_prev_prev) {
-            printf("Adding: %f %f\n", time_prev, sum_prev);
             tmp.time = time_prev;
             tmp.level = sum_prev;
             mins.push_back(tmp);
[…]

### 2010-02-17 doc/thesis/thesis.tex
added thesis.tex

\documentclass{article}
\usepackage[left=1cm,top=2cm,right=1cm,nohead,nofoot]{geometry}

%\usepackage{times}
%\usepackage{tikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
\end{comment}

\begin{document}

\section*{Introduction to Modeling}
The short term goal of current neurological models is to
learn something about the system being modeled.
While control (for instance in prosthetics), or predication
(such as BlueBrain) may some day be feasible, in the
short run we still have a great deal to learn about how our
brain works. By learning about specific systems (the dynamics
of a single person's brain), we may be able to diagnose diseases,
find their causes and even someday rectify the cause. Of course
by learning about various specific systems we can also learn 
about the general way our brains work. The reason then, for 
studying the BOLD model then is
to be able to derive the various system parameters and understand
why one brain region (say a voxel) reacts in one way, when another
acts completely differently.

While the General Linear Model has been extremely effective at
gleaning information about which voxels are "active" and which
are not, its purpose is only to say which regions are "statistically
significant". However, considering that voxels consist of millions
of neurons, the reality is that a
region could be active, but at some other level or time constant
than expected; and be completely missed by the linear model. 
Studies showing activation
in Cadavers of Phantoms highlight the inherant problem with the 
General Linear Model: a problem that mandates multiple
comparison testing with very high thresholds to ensure any 
sort of face validity \citep{drift}.
By moving to models that actually "learn" the 
parameters of the model, model error will be drastically
reduced allowing for much more reasonable statistical thresholds.

Of course, it is well known that many of the assumptions 
that the GLM makes about BOLD activation do not hold. 
The General Linear Model ignores
the fact that the BOLD response is nonlinear, contains non-gaussian
noise and most importantly that brain networks are 
so called "Small World Networks" \citep{smallworld}, \citep{noise}. 
Merely accounting for
nonlinearities in the time-series of the BOLD response can
result in significantly better estimation as shown in \citep{nonlinearmodels}.
But even more crucial is the recent movements away from mass-
univariate models. Current mass-univariate models 
assume uniform connectivity from input to every voxel; yet
this is obviously not the case. While
violating linearity assumptions or Gaussian assumptions 
requires us to be more conservative, the lack of any method
to switch brain circuits on and off means that any model
we make for brain regions would not be Turing complete;
in effect implying that humans are incapable of 
thought \citep{turing}.

So called Dynamic Causal Modeling is the beginning of 
the next phase of neurology. DCM is the first 
brain study to show any significant connection between 
diffusion tensor imaging and actual function layout in
the human brain \citep{dti_dcm}. By incorporating connections
between regions and a realistic activation model, DCM
corrects two of the largest problems with the GLM. While
there are other techniques that may be useful in the future,
they either lack physiological analogs (Artificial Neural
Networks) or are extremely computationally expensive 
(mutual information).  Given that much of the 
potential of the GLM has been exhausted, and that DCM 
is one of the first well defined methods capable of learning 
complicated brain circuits DCM is crucial to the future
of FMRI and our understanding of the human brain.

\section*{Results}
The particle filter shows great promise in being able 
to learn a variety of different regional activation
parameters. We have performed tests with a random binary
pulse train as input to a simulated region and then tested
the particle filter's ability to estimate the signal
parameters. Additionally random gaussian noise with an
Signal-To-Noise ratio of 2.0 was added. Figure 1 shows
the pulse train, the clean signal and the noisy signal, which
was the input to the particle filter. A random set of 
system parameters was chosen to simulate the region,
which the particle filter had no knowledge of other than
a mean and rough variance.
The tests were performed using 10,000 particles and took
approximately 4 minutes to compute.

%figure 1
\begin{figure}
\includegraphics[width=8in,height=4in]{noise.png}
\caption{Simulated time-series for a single region. The red line
shows the input stimulus, the blue line shows the base BOLD 
response, and the green line is the BOLD response with Gaussian
white noise added at SNR of 2}
\label{fig:noise}
\end{figure}

%figure 2
\begin{figure}
\includegraphics[width=8in,height=4in]{bold.png}
\caption{Particle Filter Results. The blue line is the "true"
(simulated) BOLD response, the red line
is the output of the particle filter with 2 standard deviations
in green. Note that the error bar for time 0 is outside the scope
of the image and is approximately $\pm .003$.}
\label{fig:bold}
\end{figure}

%figure 3
\begin{figure}
\includegraphics[width=8in,height=4in]{state.png}
\caption{Particle Filter Results. The blue line is the "true"
parameter and the green line is the estimated value for the parameter.
Note that the timescale for $S_t$ is different to highlight an typical
stimulus response. Again 2 standard deviations are shown.}
\label{fig:state}
\end{figure}

Figure 2 shows the estimated BOLD response versus the simulated
data (without noise). It is clear that as the estimated BOLD
response converges to the true BOLD as time progresses. This
is because more and more potential states are eliminated by system's
response to new input sequences. Convergence depends highly on 
the input signal and the weighting function thus choice of both
are extremely important. We have found that an impulse train
(emulated with .1 second width square waves) gives a very good
response. It is also important to choose an input with
some longer hold times, at least as long as than the 
expected time constants, which in this case was around 
8 seconds. The weighting function we have chosen is based
on the exponential distribution, with a variance equal to
the RMS of the signal. While it is important to converge
by the end of the time-series, it is often the case that 
converging too fast will harm the state estimation by over-
emphasizing the measurements of single time points. The rate
of convergence is well within control of the user of the 
particle filter based on the weighting function and the frequency
of re-sampling. Note that resampling can cause quantization
errors due to the nature of estimating a long tailed 
distribution with a finite number of samples.

Figure 3 shows the estimated versus simulated values for
several key parameters. Notice how the variance drops as
the particle filter continues. The parameters shown are
$V_0$ which is a scaling parameter, $\tau_s$, and $\tau_f$
which are both time constants and $s_t$ which is a hidden
state variable, estimating the flow inducing signal. Note that
even though some of the parameters don't converge to the exact
value, that the estimated $s_t$ still matches the true
$s_t$ relatively well. It is often the cast that although a
few parameters don't converge to their true value, one 
parameter may pick up the slack for another. It is of
note that $\tau_s$ and $\tau_f$ do converge at least close
to their true values, because they cannot be determined
from steady state.

\bibliographystyle{abbrvnat}
\bibliography{proposal_micah}

\end{document}

### 2010-02-17 code/BoldModel.cpp
added mask to calc_parammap. Check for errors during resampling. Increased variance in drift param.

@@ -42,7 +42,7 @@ BoldModel::BoldModel(aux::vector stddev, bool expweight,
     }
     
     if(drift.size() != MEAS_SIZE) 
-        drift = aux::vector(MEAS_SIZE, 1e-100);
+        drift = aux::vector(MEAS_SIZE, 1e-10);
     for(unsigned int i = 0; i < MEAS_SIZE ; i++)
         defaultstate[STATE_SIZE-MEAS_SIZE+i] = drift[i];
 

### 2010-02-17 code/calc_parammap.cpp

@@ -49,6 +49,18 @@ struct callback_data
     Image4DType::IndexType pos;
 };
 
+bool checkmask(Label4DType::Pointer maskimg, Image4DType::PointType point)
+{
+    Image4DType::IndexType index;
+    maskimg->TransformPhysicalPointToIndex(point, index);
+    if(maskimg->GetRequestedRegion().IsInside(index) && 
+                    maskimg->GetPixel(index) > 0) {
+        return true;
+    }
+
+    return false;
+}
+
 int callback(BoldPF* bold, void* data)
 {
     boost::mpi::communicator world;
@@ -93,7 +105,51 @@ void fillvector(std::vector< aux::vector >& output, Image4DType* input,
         }
     }
 }
-            
+
+Image4DType::Pointer preprocess_help(Image4DType::Pointer input, 
+            std::vector<Activation>& stim, double sampletime, unsigned int erase,
+            bool nospline)
+{
+    boost::mpi::communicator world;
+    const unsigned int rank = world.rank();
+    /* Set up measurements image */
+    //*output << "Conditioning FMRI Image" << endl;
+    //remove first 2 time step, since they are typically polluted
+    input = pruneFMRI(input, stim, sampletime, erase);
+
+    //calculate %difference, which is used normally for the bold signal
+    // or the modified % difference (with spline rather than mean)
+    if(nospline) {
+//        *output << "Changing Image to %difference" << endl;
+        SubF4::Pointer sub = SubF4::New();   
+        DivF4::Pointer div = DivF4::New();
+        Image4DType::Pointer mean = extrude(Tmean(input),
+                    input->GetRequestedRegion().GetSize()[3]);
+        sub->SetInput1(input);
+        sub->SetInput2(mean);
+        div->SetInput1(sub->GetOutput());
+        div->SetInput2(mean);
+        div->Update();
+        input = div->GetOutput();
+    } else {
+//        *output << "De-trending, then dividing by mean" << endl;
+        input = deSpline(input, 15, stim, sampletime);
+    }
+
+    /* Save detrended image */
+    if(rank == 0) try {
+        itk::ImageFileWriter<Image4DType>::Pointer out = 
+                    itk::ImageFileWriter<Image4DType>::New();
+        out->SetInput(input);
+        out->SetFileName("pfilter_input.nii.gz");
+//        cout << "Writing: " << tmp << endl;
+        out->Update();
+    } catch(itk::ExceptionObject) {
[…]

### 2010-02-17 code/include/BoldModel.hpp

@@ -58,8 +58,8 @@ public:
         return (name < (int)GVAR_SIZE) ? name : index*LVAR_SIZE + name;
     };
 
-    double getA1() {return A1;};
-    double getA2() {return A2;};
+    static double getA1() {return A1;};
+    static double getA2() {return A2;};
 
 private:
     //the standard deviations for the parameters theta, which are

### 2010-02-17 code/include/RegularizedParticleResamplerMod.hpp

@@ -174,14 +174,30 @@ RegularizedParticleResamplerMod<NT, KT>::resample_help(
     double weight = 0;
 
     /* standardise particles */
+    for(unsigned int i = 0 ; i < model->getMeasurementSize() ; i++) {
+        unsigned int tmp = model->getStateSize()-i-1;
+        if(covariance(tmp, tmp) < 1e-100) {
+            std::cerr << "Warning covariance of drift (" << tmp 
+                        << ") dropped to 0, resetting" << std::endl;
+            covariance(tmp, tmp) = 1e-10;
+        }
+    }
     aux::matrix sd(covariance);
     aux::vector diag_v(sd.size1());
-    #ifndef NDEBUG
     int err = lapack::syev('V', 'U', sd, diag_v);
-    #endif
-    assert(err == 0);
-    bool verbose = false;
+    if(err != 0 ) {
+        std::cerr << "Hmm the lapack::syev returned a non-zero error" << std::endl;
+        std::cerr << "Covariance matrix: " << std::endl;
+        outputMatrix(std::cerr, covariance);
+        std::cerr << "Standard deviation: " << std::endl;
+        outputMatrix(std::cerr, sd);
+        std::cerr << "Diagonalized result: " << std::endl;
+        outputVector(std::cerr, diag_v);
+        std::cerr << std::endl;
+        exit(-1);
+    }
 
+    bool verbose = false;
     for(unsigned int i = 0 ; i<diag_v.size() ; i++) {
         if(diag_v(i) < 0) {
             if(rank == 0) {
@@ -224,7 +240,7 @@ RegularizedParticleResamplerMod<NT, KT>::resample_help(
         out.add(x, weight);
     }
     if(badcount != 0) {
-        std::cerr << "Reweighted " << badcount << "/" << p.getSize() << "particles"
+        std::cerr << "Reset " << badcount << "/" << p.getSize() << "particles"
                     << std::endl;
     }
 };

### 2010-02-17 code/segment.cpp

@@ -1304,6 +1304,7 @@ Image4DType::Pointer deSpline(const Image4DType::Pointer fmri_img,
     Image4DType::Pointer spline = getspline(fmri_img, knots_final);
     
     {
+    printf("Writing spline");
     itk::ImageFileWriter< Image4DType >::Pointer writer = 
                 itk::ImageFileWriter< Image4DType >::New();
     writer->SetInput(spline);
@@ -1317,6 +1318,7 @@ Image4DType::Pointer deSpline(const Image4DType::Pointer fmri_img,
     Image4DType::Pointer avg = extrude(Tmean(fmri_img), 
                 fmri_img->GetRequestedRegion().GetSize()[3]);
     {
+    printf("Writing fmri_img");
     itk::ImageFileWriter< Image4DType >::Pointer writer = 
                 itk::ImageFileWriter< Image4DType >::New();
     writer->SetInput(fmri_img);
@@ -1324,6 +1326,7 @@ Image4DType::Pointer deSpline(const Image4DType::Pointer fmri_img,
     writer->Update();
     }
     {
+    printf("Writing avg");
     itk::ImageFileWriter< Image4DType >::Pointer writer = 
                 itk::ImageFileWriter< Image4DType >::New();
     writer->SetInput(avg);

### 2010-02-17 doc/thesis/thesis.tex
changed default output for calc_parammap. Eval now uses spacing/orientation.

@@
 \begin{document}
+\section*{Introduction}
+Nuclear Magnetic Resonance Imaging is a powerful tool in the analysis
+of both brain struction as well as activity. In particular Functional
+MRI gives very high spatial resolution at the cost of rather limited
+temporal resolution. Whereas other methods
+of analyzing neural signals can be invasive or ill-posed, 
+FMRI is relatively easy, and its analysis straight forward. Because of
+these benefits, FMRI has great potential for studying human cognition.
+Unfortunately, in order to achieve the speeds necessary to catch the
+Blood Oxygen Level Dependent (BOLD) signal, very little time can
+be spent allowing NMR spin to settle out, nor can much time be spent
+gathering a large region in K-Space. Because of these restrictions,
+FMRI is plagued with unidentified noise (cite) and offers a temporal
+resolution on the order of seconds rather than milliseconds.
+
+Traditional methods of analyzing FMRI images are ad-hoc and in many
+ways arbitrary. Data is often thrown out due to a lack of activation (cite),
+and there is no standard method of removing the low frequency drift
+present in all such images. Often times "activations" are looked at
+individually by taking the percent difference between the "activated"
+signal and the signal just before a stimulus was applied. Another popular
+method is a direct highpass filter, but this can attenuate actual
+signal, and introduce high frequency artifacts (cite).
+
+The traditional method of analyzing FMRI data is to use the average
+over multiple runs,
 

### 2010-02-17 code/BoldModel.cpp

@@ -12,8 +12,8 @@
 
 BoldModel::BoldModel(aux::vector stddev, bool expweight, 
             size_t sections, aux::vector drift) :
-            sigma(stddev), GVAR_SIZE(4), LVAR_SIZE(7), SIMUL_STATES(sections),
-            STATE_SIZE(4+7*sections + sections),
+            sigma(stddev), SIMUL_STATES(sections),
+            STATE_SIZE(GVAR_SIZE+LVAR_SIZE*sections + sections),
             MEAS_SIZE(sections),
             INPUT_SIZE(1)//, segments(sections)
 {
@@ -349,3 +349,79 @@ bool BoldModel::reweight(aux::vector& checkme, double& weightout) const
     return false;
 }
 
+aux::vector BoldModel::defmu(unsigned int simul)
+{
+    aux::vector ret(simul+simul*LVAR_SIZE+GVAR_SIZE);
+    for(unsigned int ii = 0 ; ii < simul; ii++) {
+        ret[indexof(TAU_S, ii)] = 4.98;
+        ret[indexof(TAU_F, ii)] = 8.31;
+        ret[indexof(EPSILON, ii)] = 0.069;
+        ret[indexof(TAU_0, ii)] = 8.38;
+        ret[indexof(ALPHA, ii)] = .189;
+        ret[indexof(E_0, ii)] = .635;
+        ret[indexof(V_0, ii)] = 1.49e-2;
+
+        ret[indexof(V_T,ii)] = 1;
+        ret[indexof(Q_T,ii)] = 1;
+        ret[indexof(S_T,ii)] = 0;
+        ret[indexof(F_T,ii)]= 1;
+    }
+    
+    for(unsigned int i = 0; i < simul; i++)
+        ret[GVAR_SIZE+simul*LVAR_SIZE+i] = 0;
+    return ret;
+}
+
+aux::vector BoldModel::defvar(unsigned int simul)
+{
+    aux::vector ret(simul+simul*LVAR_SIZE+GVAR_SIZE, 0);
+    
+    for(unsigned int ii = 0 ; ii < simul ; ii++) {
+        //set the variances for all the variables to 3*sigma
+        ret(indexof(TAU_S  ,ii)) = 1.07*1.07;
+        ret(indexof(TAU_F  ,ii)) = 1.51*1.51;
[…]

### 2010-02-17 code/calc_parammap.cpp

@@ -310,7 +310,7 @@ int main(int argc, char* argv[])
     aux::vector a_values(2);
     a_values[0] = BoldModel::getA1();
     a_values[1] = BoldModel::getA2();
-
+       
     /* Calculate parameters for every voxel */
     for(index3[0] = 0 ; index3[0] < xlen ; index3[0]++) {
         for(index3[1] = 0 ; index3[1] < ylen ; index3[1]++) {
@@ -349,10 +349,11 @@ int main(int argc, char* argv[])
                 
                 }
 
+                const int SIMUL = 1;
                 //set the output to a standard -1 if BoldPF failed
                 if(result != BoldPF::DONE) {
-                    mu = aux::vector(BASICPARAMS, -1);
-                    var = aux::vector(BASICPARAMS, -1);
+                    mu = BoldModel::defmu(SIMUL);
+                    var = 4*BoldModel::defvar(SIMUL);
                 }
                 //write the calculated expected value/variance of parameters
                 writeVector<double, aux::vector>(paramMuImg, 3, mu, index4);
@@ -379,6 +380,7 @@ int main(int argc, char* argv[])
     
     //write final output
     if(rank == 0) {
+        paramMuImg->CopyInformation(inImage);
         itk::ImageFileWriter<Image4DType>::Pointer out1 = 
                     itk::ImageFileWriter<Image4DType>::New();
         out1->SetInput(paramMuImg);
@@ -386,6 +388,7 @@ int main(int argc, char* argv[])
         cout << "Writing: param_exp.nii.gz" << endl;
         out1->Update();
 
+        paramVarImg->CopyInformation(inImage);
         itk::ImageFileWriter<Image4DType>::Pointer out2 = 
                     itk::ImageFileWriter<Image4DType>::New();
         out2->SetInput(paramVarImg);
@@ -393,6 +396,7 @@ int main(int argc, char* argv[])
         cout << "Writing: param_var.nii.gz" << endl;
         out2->Update();
         
+        measMuImg->CopyInformation(inImage);
         itk::ImageFileWriter<Image4DType>::Pointer out3 = 
                     itk::ImageFileWriter<Image4DType>::New();
         out3->SetInput(measMuImg);

### 2010-02-17 code/include/BoldModel.hpp

@@ -54,12 +54,16 @@ public:
     aux::vector estMeasMean(aux::DiracMixturePdf& in) const;
 
     //goes to the index of the given state
-    inline size_t indexof(int name, int index) const {
+    static inline size_t indexof(int name, int index) {
         return (name < (int)GVAR_SIZE) ? name : index*LVAR_SIZE + name;
     };
 
     static double getA1() {return A1;};
     static double getA2() {return A2;};
+
+    static aux::vector defmu(unsigned int);
+    static aux::vector defvar(unsigned int);
+    static aux::symmetric_matrix defcov(unsigned int);
 
 private:
     //the standard deviations for the parameters theta, which are
@@ -83,8 +87,8 @@ private:
     aux::vector sigma;
     
     //Constants
-    const unsigned int GVAR_SIZE;
-    const unsigned int LVAR_SIZE;
+    static const unsigned int GVAR_SIZE = 4;
+    static const unsigned int LVAR_SIZE = 7;
     const unsigned int SIMUL_STATES;
     const unsigned int STATE_SIZE;
 

### 2010-02-17 code/include/tools.h

@@ -91,7 +91,7 @@ Image3DType::Pointer get_rms(Image4DType::Pointer in);
 
 //write a vector to a dimension of an image
 template <class T, class Vector>
-void writeVector(typename itk::OrientedImage< T, 4 >::Pointer out, int dir, 
+int writeVector(typename itk::OrientedImage< T, 4 >::Pointer out, int dir, 
             const Vector& input, 
             typename itk::OrientedImage< T, 4 >::IndexType start)
 {
@@ -105,6 +105,7 @@ void writeVector(typename itk::OrientedImage< T, 4 >::Pointer out, int dir,
         it.Set(input[i]);
         ++it;
     }
+    return i;
 };
 
 //dir1 should be the first matrix dimension, dir2 the second

### 2010-02-17 code/parammap_eval.cpp

@@ -175,10 +175,11 @@ void simulate(std::list<double>& sim, const std::vector<double>& params,
 /* Main Function */
 int main(int argc, char* argv[])
 {
-    vul_arg<string> a_params1(0, "4D param file, in the order: TAU_0, ALPHA,"
-                "E_0, V_0, TAU_S, TAU_F, EPSILON");
+    vul_arg<string> a_params1(0, "4D param file, should be smaller image, "
+                "in the order: TAU_0, ALPHA, E_0, V_0, TAU_S, TAU_F, EPSILON");
     vul_arg<string> a_params2(0, "4D param file, in the order: TAU_0, ALPHA,"
                 "E_0, V_0, TAU_S, TAU_F, EPSILON");
+    vul_arg<string> a_mask(0, "3 or 4D mask file (vol 0 will be used)");
     vul_arg<string> a_output(0, "3D Output Image with MSE");
     
     vul_arg<double> a_shortstep("-u", "micro-Step size", 1/128.);
@@ -194,6 +195,7 @@ int main(int argc, char* argv[])
     
     vul_arg_display_usage("No Warning, just echoing");
 
+    Label4DType::Pointer maskImage;
     Image4DType::Pointer paramImage1;
     Image4DType::Pointer paramImage2;
     Image3DType::Pointer output;
@@ -223,21 +225,22 @@ int main(int argc, char* argv[])
         fprintf(stderr, "Error opening %s\n", a_params2().c_str());
         exit(-1);
     }
+    
+    try {
+    itk::ImageFileReader<Label4DType>::Pointer reader;
+    reader = itk::ImageFileReader<Label4DType>::New();
+    reader->SetImageIO(itk::modNiftiImageIO::New());
+    reader->SetFileName( a_mask() );
+    reader->Update();
+    maskImage = reader->GetOutput();
+    } catch(itk::ExceptionObject) {
+        fprintf(stderr, "Error opening %s\n", a_mask().c_str());
+        exit(-1);
+    }
 
     unsigned int xlen = paramImage1->GetRequestedRegion().GetSize()[0];
     unsigned int ylen = paramImage1->GetRequestedRegion().GetSize()[1];
     unsigned int zlen = paramImage1->GetRequestedRegion().GetSize()[2];
-
-    if(paramImage1->GetRequestedRegion().GetSize()[0] != 
-                paramImage2->GetRequestedRegion().GetSize()[0] ||
[…]

### 2010-03-01 code/BoldModel.cpp
bug fixes.

@@ -195,8 +195,7 @@ void BoldModel::generateComponent(gsl_rng* rng, aux::vector& fillme,
     }
     
     for(size_t i = STATE_SIZE - MEAS_SIZE; i < STATE_SIZE; i++) {
-        if(theta_mu[i] != 0) 
-            fillme[i] = gsl_ran_gaussian(rng, k_sigma[i]) + theta_mu[i];
+        fillme[i] = gsl_ran_gaussian(rng, k_sigma[i]) + theta_mu[i];
     }
 }
     
@@ -298,6 +297,9 @@ void BoldModel::generatePrior(aux::DiracMixturePdf& x0, int samples,
     unsigned int count = 0;
     double k_sigma[STATE_SIZE]; 
     double theta_mu[STATE_SIZE];
+    std::cout << "k_sigma: " << k_sigma << "\n";
+    std::cout << "theta_mu: " << theta_mu << "\n";
+
     for(unsigned int i = 0 ; i < STATE_SIZE-MEAS_SIZE; i++) {
         if(indexof(S_T, count) == i) {
             count++;

### 2010-03-01 code/include/tools.h

@@ -33,6 +33,8 @@ Image3DType::Pointer Tmean(const Image4DType::Pointer fmri_img);
 Image4DType::Pointer extrude(const Image3DType::Pointer input, unsigned int len);
 Image3DType::Pointer mse(const Image4DType::Pointer input1,
             const Image4DType::Pointer input2);
+Image4DType::Pointer pctDiff(const Image4DType::Pointer input1,
+            const Image4DType::Pointer input2);
 
 //dir1 should be the direction of several separate series
 //dir2 should be the direction that you want to get rms of

### 2010-03-01 code/setorientation.cpp

@@ -62,8 +62,8 @@ int main (int argc, char** argv)
     cin.clear();
     
     printf("New Direction? (16 numbers) ctr-d to skip ");
-    for(int i = 0 ; i < 3 ; i++) {
-        for(int j = 0 ; j < 3 ; j++) {
+    for(int i = 0 ; i < 4 ; i++) {
+        for(int j = 0 ; j < 4 ; j++) {
             cin >> direction(i,j);
         }
     }
@@ -71,6 +71,7 @@ int main (int argc, char** argv)
         img->SetDirection(direction);
     cin.clear();
     
+    printf("Writing\n");
     ImageWriterType::Pointer writer = ImageWriterType::New();
     writer->SetImageIO(itk::modNiftiImageIO::New());
     writer->SetFileName( argv[2] );

### 2010-03-01 code/mse.cpp
forgot to add some stuff

@@ -1,0 +1,86 @@
+#include <itkOrientedImage.h>
+#include <itkImageFileWriter.h>
+#include <itkImageFileReader.h>
+#include <itkImageLinearIteratorWithIndex.h>
+#include <itkMetaDataObject.h>
+
+#include <indii/ml/aux/vector.hpp>
+#include <indii/ml/aux/matrix.hpp>
+
+#include "tools.h"
+#include "modNiftiImageIO.h"
+
+#include <cmath>
+#include <iostream>
+#include <string>
+
+#include <vul/vul_arg.h>
+
+using namespace std;
+
+typedef itk::ImageFileReader< Image4DType >  ImageReaderType;
+typedef itk::ImageFileWriter< Image4DType >  WriterType;
+
+typedef itk::ImageLinearIteratorWithIndex<Image4DType> ImgIter;
+
+namespace aux = indii::ml::aux;
+
+/* Main Function */
+int main(int argc, char* argv[])
+{
+    vul_arg<string> a_input1(0, "4D param file timeseries file");
+    vul_arg<string> a_input2(0, "4D param file timeseries file");
+    vul_arg<string> a_output(0, "3D Output Image with MSE");
+    
+    vul_arg_parse(argc, argv);
+    
+    vul_arg_display_usage("No Warning, just echoing");
+
+    Image4DType::Pointer input1;
+    Image4DType::Pointer input2;
+    Image3DType::Pointer output;
+
+    std::vector<Activation> input;
+
+    /* Open up the input */
+    try {
+    ImageReaderType::Pointer reader;
+    reader = ImageReaderType::New();
+    reader->SetImageIO(itk::modNiftiImageIO::New());
+    reader->SetFileName( a_input1() );
+    reader->Update();
+    input1 = reader->GetOutput();
+    } catch(itk::ExceptionObject) {
+        fprintf(stderr, "Error opening %s\n", a_input1().c_str());
+        exit(-1);
+    }
+    try {
+    ImageReaderType::Pointer reader;
+    reader = ImageReaderType::New();
+    reader->SetImageIO(itk::modNiftiImageIO::New());
+    reader->SetFileName( a_input2() );
+    reader->Update();
+    input1 = reader->GetOutput();
+    } catch(itk::ExceptionObject) {
+        fprintf(stderr, "Error opening %s\n", a_input2().c_str());
+        exit(-1);
+    }
+    
+    output = mse(input1, input2);
+    
+    //write final output
[…]

### 2010-03-01 code/parammap_diff.cpp

@@ -1,0 +1,89 @@
+#include <itkOrientedImage.h>
+#include <itkImageFileWriter.h>
+#include <itkImageFileReader.h>
+#include <itkImageLinearIteratorWithIndex.h>
+#include <itkMetaDataObject.h>
+
+#include <indii/ml/aux/vector.hpp>
+#include <indii/ml/aux/matrix.hpp>
+
+#include "tools.h"
+#include "modNiftiImageIO.h"
+
+#include <cmath>
+#include <iostream>
+#include <string>
+
+#include <vul/vul_arg.h>
+
+using namespace std;
+
+typedef itk::ImageFileReader< Image4DType >  ImageReaderType;
+typedef itk::ImageFileWriter< Image4DType >  WriterType;
+
+typedef itk::ImageLinearIteratorWithIndex<Image4DType> ImgIter;
+            
+/* Main Function */
+int main(int argc, char* argv[])
+{
+    vul_arg<string> a_params1(0, "4D param file, should be smaller image, "
+                "in the order: TAU_0, ALPHA, E_0, V_0, TAU_S, TAU_F, EPSILON");
+    vul_arg<string> a_params2(0, "4D param file, in the order: TAU_0, ALPHA,"
+                "E_0, V_0, TAU_S, TAU_F, EPSILON");
+    vul_arg<string> a_output(0, "4D Output Image with percent diff");
+    
+    vul_arg<bool> a_invert("-i", "Show 1/(percent difference), thus bright "
+                "spots will be good, and dark spots are bad", false);
+    
+    vul_arg_parse(argc, argv);
+    
+    vul_arg_display_usage("No Warning, just echoing");
+
+    Image4DType::Pointer paramImage1;
+    Image4DType::Pointer paramImage2;
+    Image4DType::Pointer output;
+
+    /* Open up the input */
+    try {
+    ImageReaderType::Pointer reader;
+    reader = ImageReaderType::New();
+    reader->SetImageIO(itk::modNiftiImageIO::New());
+    reader->SetFileName( a_params1() );
+    reader->Update();
+    paramImage1 = reader->GetOutput();
+    } catch(itk::ExceptionObject) {
+        fprintf(stderr, "Error opening %s\n", a_params1().c_str());
+        exit(-1);
+    }
+    try {
+    ImageReaderType::Pointer reader;
+    reader = ImageReaderType::New();
+    reader->SetImageIO(itk::modNiftiImageIO::New());
[…]

### 2010-03-01 code/test.cc

@@ -1,0 +1,66 @@
+#include "itkOrientedImage.h"
+#include "itkImageFileWriter.h"
+#include "itkMetaDataObject.h"
+#include "modNiftiImageIO.h"
+
+#include <cstdio>
+
+using namespace std;
+
+typedef itk::OrientedImage<unsigned short, 4> Image4DType;
+
+Image4DType::Pointer init4DImage(size_t xlen, size_t ylen, 
+            size_t zlen, size_t tlen)
+{
+    Image4DType::Pointer out = Image4DType::New();
+    Image4DType::RegionType out_region;
+    Image4DType::IndexType out_index;
+    Image4DType::SizeType out_size;
+
+    out_size[0] = xlen;
+    out_size[1] = ylen;
+    out_size[2] = zlen;
+    out_size[3] = tlen; 
+    
+    out_index[0] = 0;
+    out_index[1] = 0;
+    out_index[2] = 0;
+    out_index[3] = 0;
+    
+    out_region.SetSize(out_size);
+    out_region.SetIndex(out_index);
+
+    itk::EncapsulateMetaData<string>(out->GetMetaDataDictionary(), "VERSION", 
+                BRAINID_VERSION);
+    
+    out->SetRegions( out_region );
+    out->Allocate();
+    return out;
+}
+
+
+int main (int argc, char** argv)
+{
+    Image4DType::Pointer newimage = init4DImage(9,5,7,3);
+
+    for(int xx = 0 ; xx < 9; xx++) {
+        for(int yy = 0 ; yy < 5; yy++) {
+            for(int zz = 0 ; zz < 7; zz++) {
+                for(int tt = 0 ; tt < 3; tt++) {
+                    Image4DType::IndexType index = {{xx,yy,zz,tt}};
+                    unsigned short num = xx+10*yy + zz*100+tt*1000;
+                    newimage->SetPixel(index, num);
+                }
+            }
+        }
+    }
+
+    itk::ImageFileWriter<Image4DType>::Pointer write = itk::ImageFileWriter<Image4DType>::New();
+    write->SetInput(newimage);
+    write->SetFileName(argv[1]);
+    write->Update();
+    
+    return 0;
+}
+
+

### 2010-03-01 code/test.cc
oops. remove a file that shouldn't have been there

@@ -1,66 +1,0 @@
-#include "itkOrientedImage.h"
-#include "itkImageFileWriter.h"
-#include "itkMetaDataObject.h"
-#include "modNiftiImageIO.h"
-
-#include <cstdio>
-
-using namespace std;
-
-typedef itk::OrientedImage<unsigned short, 4> Image4DType;
-
-Image4DType::Pointer init4DImage(size_t xlen, size_t ylen, 
-            size_t zlen, size_t tlen)
-{
-    Image4DType::Pointer out = Image4DType::New();
-    Image4DType::RegionType out_region;
-    Image4DType::IndexType out_index;
-    Image4DType::SizeType out_size;
-
-    out_size[0] = xlen;
-    out_size[1] = ylen;
-    out_size[2] = zlen;
-    out_size[3] = tlen; 
-    
-    out_index[0] = 0;
-    out_index[1] = 0;
-    out_index[2] = 0;
-    out_index[3] = 0;
-    
-    out_region.SetSize(out_size);
-    out_region.SetIndex(out_index);
-
-    itk::EncapsulateMetaData<string>(out->GetMetaDataDictionary(), "VERSION", 
-                BRAINID_VERSION);
-    
-    out->SetRegions( out_region );
-    out->Allocate();
-    return out;
-}
-
-
-int main (int argc, char** argv)
-{
-    Image4DType::Pointer newimage = init4DImage(9,5,7,3);
-
-    for(int xx = 0 ; xx < 9; xx++) {
-        for(int yy = 0 ; yy < 5; yy++) {
-            for(int zz = 0 ; zz < 7; zz++) {
-                for(int tt = 0 ; tt < 3; tt++) {
-                    Image4DType::IndexType index = {{xx,yy,zz,tt}};
-                    unsigned short num = xx+10*yy + zz*100+tt*1000;
-                    newimage->SetPixel(index, num);
-                }
-            }
-        }
-    }
-
-    itk::ImageFileWriter<Image4DType>::Pointer write = itk::ImageFileWriter<Image4DType>::New();
-    write->SetInput(newimage);
-    write->SetFileName(argv[1]);
-    write->Update();
-    
-    return 0;
-}
-
-

### 2010-03-01 code/tools.cc
woot fixed a dumb bug in percent difference.

@@ -37,7 +37,7 @@ Image4DType::Pointer pctDiff(const Image4DType::Pointer input1,
 
     Image4DType::SpacingType space = input2->GetSpacing();
     space[3] = input1->GetSpacing()[3];
-    input2->SetSpacing(space[3]);
+    input2->SetSpacing(space);
 
     while(!iter1.IsAtEnd()) {
         while(!iter1.IsAtEndOfLine()) {
@@ -48,9 +48,7 @@ Image4DType::Pointer pctDiff(const Image4DType::Pointer input1,
             if(input2->GetRequestedRegion().IsInside(index)) {
                 printf("%li %li %li %li\n", iter1.GetIndex()[0], iter1.GetIndex()[1],
                             iter1.GetIndex()[2], iter1.GetIndex()[3]);
-                itero.Set(fabs((iter1.Get() - input2->GetPixel(index))/iter1.Get()));
-                if(itero.Get() == 1) 
-                    printf("WTF: %f %f -> 0?\n", iter1.Get(), input2->GetPixel(index));
+                itero.Set(fabs((input2->GetPixel(index)-iter1.Get())/input2->GetPixel(index)));
             }
             ++iter1; ++itero;
         }

### 2010-03-01 code/BoldModel.cpp
various small debugging changes to find a problem with q_t's covariance hitting NAN

@@ -118,7 +118,7 @@ int BoldModel::transition(aux::vector& dustin, const double time,
      * inf - inf = nan, so nan or inf in any member 
      * will cause this to fail
     */
-    if(isnan(tmp-tmp)) {
+    if(isnan(tmp) || isinf(tmp)) {
         dustin = defaultvector;
         return -1;
     }

### 2010-03-01 code/calc_parammap.cpp

@@ -74,17 +74,20 @@ int callback(BoldPF* bold, void* data)
     aux::vector mu = bold->getDistribution().getDistributedExpectation();
     aux::vector meas =  bold->getModel().measure(mu);
 
+    outputVector(std::cout, mu);
+    std::cout << "\n";
+    outputVector(std::cout, meas);
+    std::cout << "\n";
+
     //add DC term
     if(cdata->method == BoldPF::DIRECT) {
         for(unsigned int i = 0 ; i < meas.size(); i++) {
-            std::cout << ":" << mu[mu.size()-meas.size()+i] << ": ";
             meas[i] -= mu[mu.size()-meas.size()+i];
         }
     }
 
     if(rank == 0) 
          cdata->image->SetPixel(cdata->pos, meas[0]);
-    std::cout << "time: " << bold->getContTime() << "\n";
     return 0;
 }
 
@@ -351,8 +354,8 @@ int main(int argc, char* argv[])
                     fillvector(meas, inImage, index4, a_delta());
 
                     //create the bold particle filter
-                    BoldPF boldpf(meas, input, rms->GetPixel(index3), a_timestep(),
-                            &ofile, a_num_particles()*(1<<i), 1./a_divider(), method);
+                    BoldPF boldpf(meas, input, rms->GetPixel(index3)/5., a_timestep(),
+                            output, a_num_particles()*(1<<i), 1./a_divider(), method);
                     
                     //create the callback function
                     cbd.pos = index4;

### 2010-03-01 code/include/BoldPF.h

@@ -189,7 +189,7 @@ int BoldPF::run(void* pass = NULL)
                 << " disctime_l: " << disctime_l 
                 << " dt_l: " << dt_l << " dt_s: " << dt_s 
                 << " measure size: " << measure.size()
-                << " stim size: " << stim.size() << endl;
+                << " stim size: " << stim.size() << "\n";
     
     if(call_points.start) 
         callback(this, pass);
@@ -202,23 +202,23 @@ int BoldPF::run(void* pass = NULL)
      while(status == RUNNING && disctime_s*dt_s < dt_l*measure.size()) {
         /* time */
         conttime = disctime_s*dt_s;
-        *debug << "."; 
+//        *debug << "."; 
         
         /* Update Input if there is any*/
         while(stim_index < stim.size() && stim[stim_index].time <= conttime) {
             model->setinput(aux::vector(1, stim[stim_index].level));
             stim_index++;
-            *debug << conttime;
+ //           *debug << conttime;
         }
         
 
         /* Check to see if it is time to update */
         if(conttime >= disctime_l*dt_l) { 
             //acquire the latest measurement
-            *debug << "Measuring at " <<  conttime << endl;
+            *debug << "Measuring at " <<  conttime << "\n";
             aux::vector meas(measure[disctime_l]);
             outputVector(*debug, meas);
-            *debug << endl;
+            *debug << "\n";
 
             //step forward in time, with measurement
             filter->filter(conttime, meas);
@@ -232,36 +232,44 @@ int BoldPF::run(void* pass = NULL)
             //for instance if all the particles go to an unreasonable value like 
             //inf/nan/neg
             if(isnan(ess) || isinf(ess)) {
-                *debug << endl << "Error! ESS was " << ess << endl;
+                *debug << "\n" << "Error! ESS was " << ess << "\n";
                 status = ERROR;
                 break;
             } 
             
             //time to resample
             if(ess < ESS_THRESH) {
-                *debug << " ESS: " << ess << ", Stratified Resampling" << endl;
+                *debug << " ESS: " << ess << ", Stratified Resampling\n";
                 aux::symmetric_matrix statecov
                             = filter->getFilteredState().getDistributedCovariance();
+                aux::vector tmpmu 
+                            = filter->getFilteredState().getDistributedExpectation();
                 filter->resample(&resampler);
                 
-                *debug << " ESS: " << ess << ", Regularized Resampling" << endl << endl;
[…]

### 2010-03-01 doc/presentation/fmri_overview.tex
Merge branch 'master' of ssh://oldcub.ath.cx/home/micahc/brainid

%Micah Chambers

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Darmstadt}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}
\usefonttheme[onlylarge]{structurebold}
\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
\setbeamertemplate{navigation symbols}{}
\setbeamercovered{transparent}

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{graphics}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage[small]{caption}

\title{A Crash Course in FMRI Imaging}

\subtitle{Or More than you ever wanted to know about FMRI}

\author{Micah Chambers}

\institute {
  Bradley Department of Electrical and Computer Engineering\\
  Virginia Tech University}

\subject{Medical Imaging}

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

\pgfdeclareimage[height=0.5cm]{university-logo}{VT_marn_she_invent_cmyk}
\logo{\pgfuseimage{university-logo}}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}
\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section{About FMRI}
\subsection{FMRI Basics}
\begin{frame}{Pulse Sequence}
  % - A title should summarize the slide in an understandable fashion
  %   for anyone how does not follow everything on the slide itself.

  \begin{itemize}
    \item $T2*$ Weighted Image, acquired either line by line or 
         with spiral sequence.
    \item One or more more slices are captured in a single reptition time (TR).
    \item Echo Planar Pulse Sequence, slices acquired in less than a
        tenth of a second, depending on resolution, etc.
    \item A full volume can usually be acquired in 2 to 4 seconds.
  \end{itemize}
  \begin{figure}
  \includegraphics[scale=.23]{epi}
  \caption{
    \tiny
    \href{http://airto.hosted.ats.ucla.edu/BMCweb/BMC_BIOS/MarkCohen/Papers/EPI-fMRI.html}
    {http://airto.hosted.ats.ucla.edu/} }
  \end{figure}
\end{frame}

\subsection{The Bold Response}
\begin{frame}{Visualization of the BOLD Response}
\begin{figure}
\includegraphics[scale=.23]{model}
\caption{
    \tiny
    \cite{ISI:000189252300007}
}
\end{figure}
\end{frame}

\begin{frame}{Equations}
  \begin{itemize}
    \item Normalized Cerebral Blood Flow:
    $$\ddot{f}(t) = \epsilon u(t) - \dot{f}(t)/\tau_s - (f(t)/\tau_f - 1)$$
    \item Normalized Cerebral Blood Volume:
    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
    \item Normalized Deoxyhaemoglobin Content:
    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
    \item Hemodynamic Response - BOLD Signal
    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
  \end{itemize}
\end{frame}

\subsection{Practical Concerns}
\begin{frame}{Pros and Cons}
  \begin{itemize}
    \item Benefits
    \begin{itemize}
        \item In Comparison to EEG/MEG very good resolution.
        \item Unlike MEG or EEG Image reconstruction is a known/solved problem.
    \end{itemize}
    \item Drawbacks
    \begin{itemize}
      \item Measures the result of a long chain of activations
      \begin{itemize}
        \item Exact variables and parameters are unknown and are
            difficult to calculate.
        \item Significant Amount of Lag between activation
            and a measurable output - can be as much as 8 seconds.
      \end{itemize}
      \item Slow Temporal Resolution
      \item Extremely Low Signal-to-Noise Ratio 
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Drawbacks Continued}
   \begin{itemize}
      \item Non-Gaussian, non-stationary Noise, Caused by:
      \begin{itemize}
        \item Physiological Interference
        \item Interference from the scanner
        \item Other, as yet unknown, factors
        \item See \cite{ISI:000080539300008} and \cite{ISI:000246766700003}
      \end{itemize}
  \end{itemize}
\begin{figure}
    \includegraphics[scale=.15]{timeseries.png}
\end{figure}
\end{frame}

\section{Processing}
\subsection{Preprocessing}
\begin{frame}{Preprocessing}
\begin{itemize}
    \item Gaussian Filters especially when using the General Linear Model
    \item Removing Drift
    \begin{itemize}
        \item Extremely necessary, otherwise p value must be set prohibitively low. \cite{ISI:000246766700003}
        \item High Pass filters
        \item Various De-Trending methods 
        \begin{itemize}
            \item Polynomial fit
            \item Spline Fit
            \item Subtracting the Mean from the Timeseries
            \item Wavelets
        \end{itemize}
    \end{itemize}
     \item Detrended signals are often normalized using \%Difference.
    \item per-activation normalization
    \begin{itemize}
        \item Average signal for a few samples before activation \emph{should} begin.
    \end{itemize}
\end{itemize}
\end{frame}

\subsection{Linear Processing}
\begin{frame}{The General Linear Model}
\begin{itemize}
    \item Uses linear methods to extract useful information
    \item Can be effective for certain types of signal
    \item Used in many of the classical studies to determine regions
        associated with various tasks.
    \item Limited in dealing with complicated tasks.
    \item Utilizes Statistical Parameter Maps used to create clusters, then find 
        clusters not likely to have arisen randomly. (Random Field Theory)
    \item SPM is the standard tool for performing linear analysis.
\end{itemize}
\end{frame}

\subsection{Nonlinear Processing}
\begin{frame}{Nonlinear Analysis}
\begin{itemize}
    \item Use nonlinear state evolution models to determine activation.
    \item Removal of drift still necessary.
    \item Typically Rely on some sort of gradient decent technique to determine
        model parameters.
    \item Neural Networks
    \begin{itemize}
        \item Good at gathering disparate low SNR signals into meaningful information. 
        \item The method typically used by people studying visual cortex to "read" images from 
            a persons brain.
        \item Needs some method of removing irrelevant voxel data, to prevent over-fitting.
    \end{itemize}
    \item Particle Filters
    \begin{itemize}
        \item Based on real biological systems - recall the BOLD state equations
        \item Able to determine underlying state variables
        \item Tolerant of Noisy readings
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{BOLD Model}
\begin{itemize}
    \item The Balloon Model proposed by \cite{ISI:000073759600002} is the basic model.
    \item There are more complicated versions of the BOLD model:
    \begin{itemize}
        \item \cite{ISI:000240969200015} Reviews several existing models
        \item \cite{Buxton2004S220} Pioneered the Balloon model which was shown in the beginning.
        \item \cite{ISI:000234015300018} Adds interesting neural activation and a habituation model
        \item Some models loosen the link between CMRO2 (oxygen metabolism)
            and Cerebral Blood Flow - likely due to several papers that report
            such a decoupling.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Post Stimulus Undershoot Controversy}
\begin{itemize}
    \item The delinking of CMRO2 from CBF is postulated based on the "Post Stimulus Undershoot"
    \begin{itemize}
        \item In 2008 several papers claimed prove CBV and CBF both returned
            to normal before the BOLD signal, implying a decoupling of CBV from the BOLD signal.
        \item In November of 2008, \cite{ISI:000265938700001} went out of the way to refute such a de-coupling,
            claiming that other experiments were catching Arterial CBV rather than Venous CBV.
        \item In \cite{ISI:000240969200015} (2004), it was shown that merely adding volume viscosity to
            the BOLD equations was sufficient to account for much of the undershoot.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Model Comparisons}
\begin{figure}
\includegraphics[scale=.165]{demeux}
\end{figure}
\end{frame}

\section{Conclusion}
\begin{frame}{The Future of FMRI}
\begin{itemize}
    \item Extremely Powerful tool.
    \item Better Models are increasing FMRI's flexibility.
    \item Combining FMRI with other methods of imaging will add 
        extensive knowledge to the understanding of how the human brain
        works.
    \item Much of the newer research is based around non-linear non-Gaussian
        models.
\end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section{References}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
  \bibliographystyle{abbrvnat}
  \bibliography{references}
\end{frame}

\end{document}

### 2010-03-01 doc/presentation/thesis_pres.tex
Merge branch 'master' of ssh://oldcub.ath.cx/home/micahc/brainid

--- 2010-03-01 doc/presentation/fmri_overview.tex
@@
 \documentclass{beamer}
+\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5,debug]{beamerposter}
 
@@
 {
-  \usetheme{Darmstadt}
+  \usetheme{VT}
   % or ...
@@
 }
+
 \usefonttheme[onlylarge]{structurebold}
+\usecolortheme{crane}
 \setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
@@
 
-\title{A Crash Course in FMRI Imaging}
+\title{BOLD Parameter Estimation using Sequential Monte Carlo Methods}
 
-\subtitle{Or More than you ever wanted to know about FMRI}
+%\subtitle{}
 
@@
 
-\pgfdeclareimage[height=0.5cm]{university-logo}{VT_marn_she_invent_cmyk}
+\pgfdeclareimage[width=1.5cm]{university-logo}{logo}
 \logo{\pgfuseimage{university-logo}}
@@
 
-\section{About FMRI}
-\subsection{FMRI Basics}
-\begin{frame}{Pulse Sequence}
-  % - A title should summarize the slide in an understandable fashion
-  %   for anyone how does not follow everything on the slide itself.
-
-  \begin{itemize}
-    \item $T2*$ Weighted Image, acquired either line by line or 
-         with spiral sequence.
-    \item One or more more slices are captured in a single reptition time (TR).
-    \item Echo Planar Pulse Sequence, slices acquired in less than a
-        tenth of a second, depending on resolution, etc.
-    \item A full volume can usually be acquired in 2 to 4 seconds.
-  \end{itemize}
-  \begin{figure}
-  \includegraphics[scale=.23]{epi}
-  \caption{
-    \tiny
-    \href{http://airto.hosted.ats.ucla.edu/BMCweb/BMC_BIOS/MarkCohen/Papers/EPI-fMRI.html}
-    {http://airto.hosted.ats.ucla.edu/} }
-  \end{figure}
-\end{frame}
-
+\section{FMRI Review}
 \subsection{The Bold Response}
-\begin{frame}{Visualization of the BOLD Response}
+\begin{frame}{The BOLD Response}
 \begin{figure}
@@
 
+\begin{frame}{Purpose/Usefullness}
+  \begin{itemize}
+    \item blah
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{BOLD Signal Properties}
+  \begin{itemize}
+    \item ...
+    \begin{itemize}
+      \item Exact variables and parameters are unknown and are
+          difficult to calculate.
+      \item Significant Amount of Lag between activation
+          and a measurable output - can be as much as 8 seconds.
+    \end{itemize}
+    \item Slow Temporal Resolution
+    \item Noise characterized by brownian motion
+    \item fft of mri signal, with and without DC
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Preprocessing}
+  \begin{itemize}
+    \item High Pass Filter and Low Pass Filter
+    \item Wavelet Detrending
+    \item ...
+    \item Spline
+  \end{itemize}
+\end{frame}
+
+\section{Statistical Parametric Mapping}
+\begin{frame}{Method}
+  \begin{itemize}
+    \item
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Results}
+  \begin{itemize}
+    \item ...
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Limitations}
+  \begin{itemize}
+    \item ...
+  \end{itemize}
+\end{frame}
+
+\section{Nonlinear Regression}
 \begin{frame}{Equations}
@@
 
-\subsection{Practical Concerns}
-\begin{frame}{Pros and Cons}
-  \begin{itemize}
-    \item Benefits
-    \begin{itemize}
-        \item In Comparison to EEG/MEG very good resolution.
-        \item Unlike MEG or EEG Image reconstruction is a known/solved problem.
-    \end{itemize}
-    \item Drawbacks
-    \begin{itemize}
-      \item Measures the result of a long chain of activations
-      \begin{itemize}
-        \item Exact variables and parameters are unknown and are
-            difficult to calculate.
-        \item Significant Amount of Lag between activation
-            and a measurable output - can be as much as 8 seconds.
-      \end{itemize}
-      \item Slow Temporal Resolution
-      \item Extremely Low Signal-to-Noise Ratio 
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Drawbacks Continued}
-   \begin{itemize}
-      \item Non-Gaussian, non-stationary Noise, Caused by:
-      \begin{itemize}
-        \item Physiological Interference
-        \item Interference from the scanner
-        \item Other, as yet unknown, factors
-        \item See \cite{ISI:000080539300008} and \cite{ISI:000246766700003}
-      \end{itemize}
-  \end{itemize}
-\begin{figure}
-    \includegraphics[scale=.15]{timeseries.png}
-\end{figure}
-\end{frame}
-
-\section{Processing}
-\subsection{Preprocessing}
-\begin{frame}{Preprocessing}
-\begin{itemize}
-    \item Gaussian Filters especially when using the General Linear Model
-    \item Removing Drift
-    \begin{itemize}
-        \item Extremely necessary, otherwise p value must be set prohibitively low. \cite{ISI:000246766700003}
-        \item High Pass filters
-        \item Various De-Trending methods 
-        \begin{itemize}
-            \item Polynomial fit
-            \item Spline Fit
-            \item Subtracting the Mean from the Timeseries
-            \item Wavelets
-        \end{itemize}
-    \end{itemize}
-     \item Detrended signals are often normalized using \%Difference.
-    \item per-activation normalization
-    \begin{itemize}
-        \item Average signal for a few samples before activation \emph{should} begin.
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\subsection{Linear Processing}
-\begin{frame}{The General Linear Model}
-\begin{itemize}
-    \item Uses linear methods to extract useful information
-    \item Can be effective for certain types of signal
-    \item Used in many of the classical studies to determine regions
-        associated with various tasks.
-    \item Limited in dealing with complicated tasks.
-    \item Utilizes Statistical Parameter Maps used to create clusters, then find 
-        clusters not likely to have arisen randomly. (Random Field Theory)
-    \item SPM is the standard tool for performing linear analysis.
-\end{itemize}
-\end{frame}
-
-\subsection{Nonlinear Processing}
-\begin{frame}{Nonlinear Analysis}
-\begin{itemize}
-    \item Use nonlinear state evolution models to determine activation.
-    \item Removal of drift still necessary.
-    \item Typically Rely on some sort of gradient decent technique to determine
-        model parameters.
-    \item Neural Networks
-    \begin{itemize}
-        \item Good at gathering disparate low SNR signals into meaningful information. 
-        \item The method typically used by people studying visual cortex to "read" images from 
-            a persons brain.
-        \item Needs some method of removing irrelevant voxel data, to prevent over-fitting.
-    \end{itemize}
-    \item Particle Filters
-    \begin{itemize}
-        \item Based on real biological systems - recall the BOLD state equations
-        \item Able to determine underlying state variables
-        \item Tolerant of Noisy readings
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{BOLD Model}
+\begin{frame}{Alternative Equations}
 \begin{itemize}
@@
 
-\begin{frame}{Post Stimulus Undershoot Controversy}
-\begin{itemize}
-    \item The delinking of CMRO2 from CBF is postulated based on the "Post Stimulus Undershoot"
-    \begin{itemize}
-        \item In 2008 several papers claimed prove CBV and CBF both returned
-            to normal before the BOLD signal, implying a decoupling of CBV from the BOLD signal.
-        \item In November of 2008, \cite{ISI:000265938700001} went out of the way to refute such a de-coupling,
-            claiming that other experiments were catching Arterial CBV rather than Venous CBV.
-        \item In \cite{ISI:000240969200015} (2004), it was shown that merely adding volume viscosity to
-            the BOLD equations was sufficient to account for much of the undershoot.
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
 \begin{frame}{Model Comparisons}
@@
 
-\section{Conclusion}
-\begin{frame}{The Future of FMRI}
+\section{Parameter Identification}
+\begin{frame}{Particle Filters}
 \begin{itemize}
-    \item Extremely Powerful tool.
-    \item Better Models are increasing FMRI's flexibility.
-    \item Combining FMRI with other methods of imaging will add 
-        extensive knowledge to the understanding of how the human brain
-        works.
-    \item Much of the newer research is based around non-linear non-Gaussian
-        models.
+    \item Why Particle Filters
+    \item Goal: $\epsilon$
+    \item Entire brain
 \end{itemize}
@@
 
-% All of the following is optional and typically not needed. 
-\appendix
-\section{References}
+\begin{frame}{Single Timeseries Results}
+\end{frame}
+
+\begin{frame}{Simulation}
+\end{frame}
+
+\begin{frame}{Simulation Results}
+\end{frame}
+
+\section{Conclusion}
 

### 2010-03-01 doc/thesis/thesis.tex
Merge branch 'master' of ssh://oldcub.ath.cx/home/micahc/brainid

--- 2010-02-17 doc/thesis/thesis.tex
@@
 \begin{document}
+\section*{Introduction}
+Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
+neural activity. At the cost of rather limited temporal resolution, FMRI gives
+high spatial for relatively cheap. Whereas other methods
+of analyzing neural signals can be invasive or ill-posed, 
+FMRI is relatively quick and easy, and its analysis straight forward. Because of
+these benefits, FMRI has great potential for studying human cognition.
+Unfortunately, FMRI does suffer from severe noise and, as previously
+mentioned, low temporal resolution. Because of the extensive noise, 
+and relative unpredictability of the absolute signal level, a large amount 
+of conditioning is necessary to obtain "valid" activation. In
+fact, typical activation detection does little to account for variations
+in the shape of the signal, but instead is a direct T-test between the measured signal
+and the expected shape of the signal. Because of these limitations,
+multiple comparison tests are imperative (ref fish), and active regions
+must be very large to be detected with any confidence. This paper
+will introduce a more robust method of analyzing FMRI signals for 
+activation, by means of estimating the BOLD parameters.
+
+\section*{General Linear Model}
+
+\section*{BOLD Model}
+
+\subsection*{Preprocessing}
+After FMRI data has been acquired it is always necessary to modify the
+data in some way to make different runs comparable. Because FMRI signal
+levels are essentially unit-less, at the very least it is necessary to convert
+the data into \% difference from the baseline. Finding the baseline is the
+first hurdle in dealing with FMRI. Various methods exist for doing this,
+but they are all ad-hoc. It is common to treat each stimulus pulse independently,
+ignoring the long decay of the BOLD signal. By treating each input separately,
+the "base" for calculating the percent difference is merely the signal level 
+before the stimulus is applied. The problem with this is that the BOLD signal
+can be heavily delayed, and fall times can be as long as 10 or 15 seconds,
+meaning that the assumption is clearly a false one. Another popular method 
+of dealing with this problem, is to put all the time series through a high
+pass filter. This will of course remove the DC component of the signal, and 
+some amount of so called "drift". The problem with this method is that it is
+not adaptive to the input. Huge variations in drift frequencies can exist 
+in a single time-series. Thus, a single cutoff frequency could miss a significant
+drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
+set too high.
+
+\subsection*{
 

### 2010-03-01 code/include/segment.h

@@ -21,19 +21,12 @@ Image4DType::Pointer pruneFMRI(const Image4DType::Pointer fmri_img,
             unsigned int remove);
 
 Image4DType::Pointer deSplineByStim(const Image4DType::Pointer fmri_img,
-            unsigned int numknots, std::vector<Activation>& stim, double dt);
+Image4DType::Pointer getspline(const Image4DType::Pointer fmri_img,
+            const std::vector<unsigned int>& knots);
 
-Image4DType::Pointer deSplineBlind(const Image4DType::Pointer fmri_img,
-            unsigned int numknots);
-
-Image4DType::Pointer splitByRegion(const Image4DType::Pointer fmri_img,
-            const Label3DType::Pointer labelmap, int label);
-
-Image4DType::Pointer read_dicom(std::string directory, double skip = 0);
-
-double get_average(const Image4DType::Pointer fmri_img, 
-        const Label3DType::Pointer labelmap);
-
+Image3DType::Pointer get_average(const Image4DType::Pointer fmri_img);
+Image3DType::Pointer get_variance(const Image4DType::Pointer fmri_img);
+>>>>>>> 23c4099152aa6e288a616de086715af8ca74a830
 Image3DType::Pointer extract(Image4DType::Pointer input, size_t index);
 
 std::list<LabelType> getlabels(Label3DType::Pointer labelmap);

### 2010-03-01 doc/thesis/plot.py

@@ -1,53 +1,0 @@
-#!/usr/bin/python
-
-import pylab as P
-import sys
-from nifti import *
-
-#t = arange(0.0, 2.0, 0.01)
-#s = sin(2*pi*t)
-#plot(t, s, linewidth=1.0)
-#
-#xlabel('time (s)')
-#ylabel('voltage (mV)')
-#title('About as simple as it gets, folks')
-#grid(True)
-#show()
-
-images = list();
-for arg in sys.argv[1:]:
-    images.append(image.NiftiImage(arg))
-
-leg = []
-
-data = sys.stdin.readlines()
-time = []
-for lineno in range(0,len(data)):
-    data[lineno] = data[lineno].split()
-    time.append(float(data[lineno][0])/2)
-    data[lineno] = float(data[lineno][1])*.0002-.0004
-
-for line in data:
-    print line
-
-for nifti in images:
-    print nifti
-    for iter in range(0,nifti.extent[0]):
-        P.plot(nifti.data[:, 0,0,iter])
-#        leg = leg + [nifti.filename+":"+str(iter)]
-#P.legend(sys.argv[1:])
-P.plot(time, data)
-#P.legend(["BOLD", "With Noise SNR = 2", "Input"])
-
-P.xlabel("Time, Seconds")
-P.ylabel("BOLD Signal %")
-P.show()
-#
-#names = ['Ts', 'Tf', 'epsilon', 'T0', 'alpha', 'E_0', 'V0', 'Vt', 'Qt', 'St', 'Ft']
-#for i in range(0,11):
-#    P.subplot(6, 2, i+1)
-#    line1 = P.plot(truestat.data[:, 0, i, 0])
-#    line2 = P.plot(eststat.data[:, 0, i, 0])
-#    P.ylabel(names[i])
-#
-#P.show()

### 2010-03-01 doc/thesis/plotout.py

@@ -1,116 +1,0 @@
-#!/usr/bin/python
-from math import sqrt
-import sys
-import getopt
-import pylab as P
-from nifti import *
-
-def main(argv):                         
-    truebold = "bold"
-    truestate = "state"
-    estbold = "particle-bold"
-    eststate = "particle-state"
-    saveloc = ""
-    try:                                
-        opts, args = getopt.getopt(argv, "hb:s:y:x:p:", ["help", "bold=", 
-                    "state=", "estbold=", "eststate=", "prefix="])
-    except getopt.GetoptError:
-        usage()
-        sys.exit(2)
-
-    for opt, arg in opts:
-        if opt in ("-h", "--help"):
-            usage();
-            sys.exit();
-        elif opt in ("-b", "--bold"):
-            truebold = arg;
-        elif opt in ("-s", "--state"):
-            truestate = arg;
-        elif opt in ("-x", "--eststate"):
-            eststate = arg;
-        elif opt in ("-y", "--estbold"):
-            estbold = arg;
-        elif opt in ("-p", "--prefix"):
-            saveloc = arg
-
-    truebold = image.NiftiImage(truebold)
-    try: 
-        truestate = image.NiftiImage(truestate)
-    except  getopt.GetoptError:
-        print "Hoops"
-    
-    estbold = image.NiftiImage(estbold)
-    eststate = image.NiftiImage(eststate)
-    print eststate
-    print estbold
-    
-    leg = []
-    time = range(0, 1199, 2)
-    
-    for iter in range(0,truebold.extent[0]):
-        P.plot(time, truebold.data[:, 0,0,iter])
-        leg = leg + ["Simulated BOLD"]
-    
-    for iter in range(0,estbold.extent[0]):
-        bold = estbold.data[:, 0, 0, iter]
-        error = estbold.data[:, 1, 0, iter]
-        for errno in range(0, len(error)):
-            error[errno] = 2*sqrt(error[errno])
-        if(estbold.extent[2] == 2):
-            P.errorbar(time, bold, yerr=error, fmt='r', elinewidth=.5, ecolor='g')
-        else:
-            P.plot(time, estbold.data[:, 0,0,iter], 'b')
-        leg = leg + ["Estimated BOLD"]
-        
-    P.xlabel("Time, Seconds")
-    P.ylabel("BOLD Signal %")
[…]

### 2010-03-01 scripts/getbuildtest.py

@@ -17,7 +17,7 @@ GSL_URL="ftp://ftp.gnu.org/gnu/gsl/gsl-1.9.tar.gz"
 BOOST_URL="http://voxel.dl.sourceforge.net/sourceforge/boost/boost_1_38_0.tar.gz"
 BINDINGS_URL="http://mathema.tician.de/news.tiker.net/download/software/boost-numeric-bindings/boost-numeric-bindings-20081116.tar.gz"
 #ITK_URL="http://voxel.dl.sourceforge.net/sourceforge/itk/InsightToolkit-3.12.0.tar.gz"
-ITK_URL="http://voxel.dl.sourceforge.net/sourceforge/itk/InsightToolkit-3.14.0.tar.gz"
+ITK_URL="http://voxel.dl.sourceforge.net/sourceforge/itk/InsightToolkit-3.16.0.tar.gz"
 #OPENMPI_URL="http://www.open-mpi.org/software/ompi/v1.3/downloads/openmpi-1.3.1.tar.gz"
 OPENMPI_URL="http://www.open-mpi.org/software/ompi/v1.2/downloads/openmpi-1.2.9.tar.gz"
 LAPACK_URL="http://www.netlib.org/lapack/lapack-3.2.1.tgz"

### 2010-03-01 scripts/plot.py

@@ -2,7 +2,7 @@
 
 import pylab as P
 import sys
-from nifti import *
+import nibabel 
 
 #t = arange(0.0, 2.0, 0.01)
 #s = sin(2*pi*t)
@@ -16,15 +16,16 @@ from nifti import *
 
 images = list();
 for arg in sys.argv[1:]:
-    images.append(image.NiftiImage(arg))
+    img = nibabel.load(arg)
+    images.append(img)
 
 leg = []
 
-for nifti in images:
-    print nifti
-    for iter in range(0,nifti.extent[0]):
-        P.plot(nifti.data[:, 0,0,iter])
-        leg = leg + [nifti.filename+":"+str(iter)]
+for i in range(0, len(images)):
+    print images[i]
+    for iter in range(0,images[i].get_shape()[3]):
+        P.plot(images[i].get_data()[:, 0,0,iter])
+        leg = leg + [sys.argv[i+1]+":"+str(iter)]
 #P.legend(sys.argv[1:])
 P.legend(leg)
 

### 2010-03-01 code/include/segment.h
fixed some sort of stupid problem with git, repo and a little work on presentation.

@@ -21,12 +21,19 @@ Image4DType::Pointer pruneFMRI(const Image4DType::Pointer fmri_img,
             unsigned int remove);
 
 Image4DType::Pointer deSplineByStim(const Image4DType::Pointer fmri_img,
-Image4DType::Pointer getspline(const Image4DType::Pointer fmri_img,
-            const std::vector<unsigned int>& knots);
+            unsigned int numknots, std::vector<Activation>& stim, double dt);
 
-Image3DType::Pointer get_average(const Image4DType::Pointer fmri_img);
-Image3DType::Pointer get_variance(const Image4DType::Pointer fmri_img);
->>>>>>> 23c4099152aa6e288a616de086715af8ca74a830
+Image4DType::Pointer deSplineBlind(const Image4DType::Pointer fmri_img,
+            unsigned int numknots);
+
+Image4DType::Pointer splitByRegion(const Image4DType::Pointer fmri_img,
+            const Label3DType::Pointer labelmap, int label);
+
+Image4DType::Pointer read_dicom(std::string directory, double skip = 0);
+
+double get_average(const Image4DType::Pointer fmri_img, 
+        const Label3DType::Pointer labelmap);
+
 Image3DType::Pointer extract(Image4DType::Pointer input, size_t index);
 
 std::list<LabelType> getlabels(Label3DType::Pointer labelmap);

### 2010-03-01 doc/presentation/thesis_pres.tex
fixed some sort of stupid problem with git, repo and a little work on presentation.
better %diff calculation. BoldPF now uses its own covariance prior. Expweighting added as default. Bold model is a little more verbose.

@@
 \usepackage{graphics}
-\usepackage{times}
+%\usepackage{times}
 \usepackage[T1]{fontenc}
 \usepackage[small]{caption}
+\usepackage{algorithmic}
 
@@
 
-\institute {
-  Bradley Department of Electrical and Computer Engineering\\
-  Virginia Tech University}
+\institute {Virginia Tech Bioimaging Systems Lab}
 
@@
 \section{FMRI Review}
-\subsection{The Bold Response}
 \begin{frame}{The BOLD Response}
@@
     \tiny
-    \cite{ISI:000189252300007}
+    \cite{nonlinearfiltering}
 }
@@
 
-\begin{frame}{Purpose/Usefullness}
-  \begin{itemize}
-    \item blah
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{BOLD Signal Properties}
-  \begin{itemize}
-    \item ...
   \begin{itemize}
@@
         and a measurable output - can be as much as 8 seconds.
-    \end{itemize}
     \item Slow Temporal Resolution
-    \item Noise characterized by brownian motion
-    \item fft of mri signal, with and without DC
+    \item Noise characterized by brownian motion, which clashes with low 
+        frequency elements.
+    \begin{center}
+    \includegraphics[width=10cm,height=4cm]{signal_fft.pdf}
+    \end{center}
   \end{itemize}
@@
   \begin{itemize}
-    \item High Pass Filter and Low Pass Filter
-    \item Wavelet Detrending
-    \item ...
-    \item Spline
+    \item Low Pass Filter (Gaussian Filter, not recommended)
+    \item Drift Removal (not always performed)
+    \begin{itemize}
+        \item High Pass Filter
+        \item Linear 
+        \item Quadratic
+        \item Wavelet 
+        \item Spline (Which I am using)
     \end{itemize}
+  \end{itemize}
+  \begin{figure}
+    \includegraphics[width=10cm]{detrending_effect.png}
+    \caption{
+        \tiny
+        \cite{detrending}
+    }
+  \end{figure}
 \end{frame}
@@
 \begin{frame}{Method}
-  \begin{itemize}
-    \item
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Results}
-  \begin{itemize}
-    \item ...
-  \end{itemize}
+  \begin{figure}
+    \includegraphics[width=11cm]{glm_pipeline.png}
+    \caption{
+        \tiny
+        \cite{spm_pipeline}
+    }
+  \end{figure}
 \end{frame}
@@
 \begin{frame}{Limitations}
+\begin{columns}
+  \column{2.5in}
   \begin{itemize}
-    \item ...
+    \item Linear, for a signal which is known to be nonlinear
+    \item Essentially the weighted sum of a set of "expected" responses.
+    \item Parametric
+    \begin{itemize}
+        \item Forced to make assumptions about underlying distributions
+        \item No time-scaling.
     \end{itemize}
+  \end{itemize}
+  \column{3in}
+  \begin{figure}
+    \includegraphics[width=2.5in]{glm.png}
+    \caption{
+        \tiny
+        \cite{spm_pipeline}
+    }
+  \end{figure}
+\end{columns}
 \end{frame}
@@
 
-\begin{frame}{Alternative Equations}
-\begin{itemize}
-    \item The Balloon Model proposed by \cite{ISI:000073759600002} is the basic model.
-    \item There are more complicated versions of the BOLD model:
-    \begin{itemize}
-        \item \cite{ISI:000240969200015} Reviews several existing models
-        \item \cite{Buxton2004S220} Pioneered the Balloon model which was shown in the beginning.
-        \item \cite{ISI:000234015300018} Adds interesting neural activation and a habituation model
-        \item Some models loosen the link between CMRO2 (oxygen metabolism)
-            and Cerebral Blood Flow - likely due to several papers that report
-            such a decoupling.
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Model Comparisons}
+\begin{frame}{Model Comparison}
   \begin{figure}
-\includegraphics[scale=.165]{demeux}
+    \includegraphics[width=5in,height=2.4in]{comparison.png}
+    \caption{
+        \tiny
+        \cite{comparison}
+    }
   \end{figure}
@@
 \begin{itemize}
-    \item Why Particle Filters
-    \item Goal: $\epsilon$
-    \item Entire brain
+    \item Non-parametric, no assumptions are violated
+    \item Model based, fit parameters to input, constrained by physical variables
+    \item Fits a mixture PDF to the posterior of all parameters
+    \item Non-trivial computation cost
+    \item I use a Regularized Particle Filter
+    \begin{enumerate}
+        \item Regularized Re-sampling prevents particles from de-generating into a 
+            small number of unique particles
+        \item Allows distributions to move more freely
+    \end{enumerate}
 \end{itemize}
@@
 
-\begin{frame}{Single Timeseries Results}
+\begin{frame}{Particle Filter}
+    \begin{columns}
+    \column{1.8in}
+    \begin{itemize}
+        \item $S_t = \{p_{0,t}, ... , p_{N,t}\}$, the set of particles
+        \item $w_{i,t}$, weight of particle $p_{i,t}$
+        \item $y_t$, measurement at time $t$, there is not a $y_t$ for every $t$.
+        \item $f(p_{i,t},y_t)$, weighting function 
+        \item $s(p_{i,t})$, step function 
+    \end{itemize}
+    \column{3.2in}
+    \begin{algorithmic}
+     \STATE Draw $S_0$ from prior distribution 
+     \FOR{$t = 0:t_{step}:t_{end}$}
+        \FOR{each $p_{i,t-1} \in S_{t-1}$}
+           \STATE $p_{i,t} = s(p_{i,t-1})$ 
+           \IF{There is a measurement at time $t$}
+               \FOR{every $p_{i,t}$}
+                    \STATE $w_{i,t} = w_{i,t-1}f(p_{i,t},y_t)$
+               \ENDFOR 
+               \STATE Resample if weights are unevenly distributed\\
+           \ENDIF
+        \ENDFOR
+     \ENDFOR
+%                    \item If a weights are not evenly distributed enough, then resample
+     \end{algorithmic}
+     \end{columns}
 \end{frame}
 
-\begin{frame}{Simulation}
+\subsection{Single-Region}
+\begin{frame}{Single Timeseries Results}
+  \begin{figure}
+    \includegraphics[height=3in]{noise.png}
+  \end{figure}
 \end{frame}
 
-\begin{frame}{Simulation Results}
+\begin{frame}{Single Timeseries Results, Measurement Convergence}
+  \begin{figure}
+    \includegraphics[height=3in]{bold.png}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Single Timeseries Results, State Convergence}
+  \begin{figure}
+    \includegraphics[height=3in]{state.png}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Factors Affecting Convergence}
+    \begin{enumerate}
+        \item Weighting function
+            \begin{itemize}
+                \item Needs to be continuous and defined for any input, should go to 0
+                \item Too wide a weighting function results in under-sensitivity, 
+                            slow or no convergence
+                \item Too thin a weighting function reduces robustness to noise
+            \end{itemize}
+        \item How often re-sampling is done, re-sampling should be minimized
+            \begin{itemize}
+                \item Stratified Resampling can result in truncated tails on posterior
+                \item Regularized Resampling can result in reduced robustness to noise 
+            \end{itemize}
+        \item Number of particles
+            \begin{itemize}
+                \item More particles give higher fidelity of posterior
+            \end{itemize}
+    \end{enumerate}
+\end{frame}
+
+\subsection{Multi-Region}
+\begin{frame}{Parameter Map Generation/Simulation}
+\begin{columns}
+  \column{2.5in}
+  \begin{enumerate}
+    \item Generate a parameter map, with a set of parameters for each voxel
+    \item Simulate every set of parameters, and use as input to possum
+    \item Perform preprocessing (de-trend and normalize)
+    \item Run particle filter on every grey matter voxel in image, generating a new
+            parameter map
+  \end{enumerate}
+  \column{3in}
+  \begin{figure}
+    \includegraphics[width=2.8in]{equation.png}
+    \caption{
+    }
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Simulation Results, $\tau_0$}
+  \begin{figure}
+    \includegraphics[height=3in]{results.png}
+    \caption{Percent Difference from "Actual"}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Simulation Results, multiple}
+  \begin{figure}
+    \includegraphics[height=3in]{results2.png}
+    \caption{Percent Difference from "Actual"}
+  \end{figure}
 \end{frame}

### 2010-03-01 code/BoldModel.cpp

@@ -47,12 +47,29 @@ BoldModel::BoldModel(aux::vector stddev, bool expweight,
         defaultstate[STATE_SIZE-MEAS_SIZE+i] = drift[i];
 
     std::cerr << "Sizes:" << std::endl;
-    std::cerr << this->GVAR_SIZE        << std::endl;
-    std::cerr << this->LVAR_SIZE        << std::endl;
-    std::cerr << this->SIMUL_STATES     << std::endl;
-    std::cerr << this->STATE_SIZE       << std::endl;
-    std::cerr << this->MEAS_SIZE        << std::endl;
-    std::cerr << this->INPUT_SIZE       << std::endl;
+    std::cerr << "GVAR_SIZE    " << this->GVAR_SIZE    << std::endl;
+    std::cerr << "LVAR_SIZE    " << this->LVAR_SIZE    << std::endl;
+    std::cerr << "SIMUL_STATES " << this->SIMUL_STATES << std::endl;
+    std::cerr << "STATE_SIZE   " << this->STATE_SIZE   << std::endl;
+    std::cerr << "MEAS_SIZE    " << this->MEAS_SIZE    << std::endl;
+    std::cerr << "INPUT_SIZE   " << this->INPUT_SIZE   << std::endl;
+
+    for(unsigned int i = 0 ; i < 3 ; i++) {
+        std::cerr << "indexof(TAU_0, " << i << ") " << indexof(TAU_0, i) << std::endl;
+        std::cerr << "indexof(ALPHA, " << i << ") " << indexof(ALPHA, i) << std::endl;
+        std::cerr << "indexof(E_0  , " << i << ") " << indexof(E_0  , i) << std::endl;
+        std::cerr << "indexof(V_0  , " << i << ") " << indexof(V_0  , i) << std::endl;
+        std::cerr << "indexof(TAU_S, " << i << ") " << indexof(TAU_S, i) << std::endl;
+        std::cerr << "indexof(TAU_F, " << i << ") " << indexof(TAU_F, i) << std::endl;
+        std::cerr << "indexof(EPSILON, " << i << ") " << indexof(EPSILON, i) << std::endl;
+        std::cerr << "indexof(V_T  , " << i << ") " << indexof(V_T  , i) << std::endl;
+        std::cerr << "indexof(Q_T  , " << i << ") " << indexof(Q_T  , i) << std::endl;
[…]

### 2010-03-01 code/calc_parammap.cpp

@@ -66,18 +66,12 @@ int callback(BoldPF* bold, void* data)
 {
     boost::mpi::communicator world;
     const unsigned int rank = world.rank();
-//    const unsigned int size = world.size();
     
     callback_data* cdata = (struct callback_data*)data;
     
     cdata->pos[3] = bold->getDiscTimeL();
     aux::vector mu = bold->getDistribution().getDistributedExpectation();
     aux::vector meas =  bold->getModel().measure(mu);
-
-    outputVector(std::cout, mu);
-    std::cout << "\n";
-    outputVector(std::cout, meas);
-    std::cout << "\n";
 
     //add DC term
     if(cdata->method == BoldPF::DIRECT) {
@@ -181,6 +175,7 @@ int main(int argc, char* argv[])
     vul_arg<string> a_input(0, "4D timeseries file");
     vul_arg<string> a_mask(0, "3D mask file");
     
+    vul_arg<bool> a_dc("-c", "Calculate DC gain as a state variable", false);
     vul_arg<bool> a_delta("-l", "Use deltas between measurements, this precludes"
                 "the drift option", false);
     vul_arg<bool> a_smart("-S", "Use \"smart\" knots based on less active regions"
@@ -190,7 +185,7 @@ int main(int argc, char* argv[])
     vul_arg<unsigned> a_divider("-d", "Intermediate Steps between samples.", 128);
     vul_arg<string> a_stimfile("-s", "file containing \"<time> <value>\""
                 "pairs which give the time at which input changed", "");
-    vul_arg<bool> a_expweight("-e", "Use exponential weighting function", false);
+    vul_arg<bool> a_expweight("-e", "Use exponential weighting function", true);
     vul_arg<double> a_timestep("-t", "TR (timesteps in 4th dimension)", 2);
     
     vul_arg_parse(argc, argv);
@@ -303,6 +298,7 @@ int main(int argc, char* argv[])
     
     unsigned int method = BoldPF::DIRECT;
     if(a_delta()) method = BoldPF::DELTA;
+    else if(a_dc()) method = BoldPF::DC;
 
     //callback variables, to fill in 
     BoldPF::CallPoints callpoints;
@@ -355,7 +351,8 @@ int main(int argc, char* argv[])
 
                     //create the bold particle filter
                     BoldPF boldpf(meas, input, rms->GetPixel(index3)/5., a_timestep(),
-                            output, a_num_particles()*(1<<i), 1./a_divider(), method);
[…]

### 2010-03-01 code/include/BoldPF.h

@@ -38,7 +38,7 @@ public:
     /* Types */
     typedef indii::ml::filter::ParticleFilter<double> Filter;
     enum Status {ERROR=-1, UNSTARTED=0, RUNNING=1, PAUSED=2, DONE=3};
-    enum Method {DIRECT, DELTA, PROCESS};
+    enum Method {DIRECT, DELTA, DC};
     
     /* Callback types */
     struct CallPoints{
@@ -61,7 +61,7 @@ public:
                 const std::vector<Activation>& activations,  double weightvar,
                 double longstep, std::ostream* output, 
                 unsigned int numparticles = 1000, double shortstep = 1./64,
-                unsigned int method = DIRECT);
+                unsigned int method = DIRECT, bool exp = false);
 
     /* Destructor */
     ~BoldPF();
@@ -118,6 +118,8 @@ private:
     RegularizedParticleResamplerMod<aux::Almost2Norm, aux::AlmostGaussianKernel>*
                 resampler_reg;
 
+    void generatePrior(aux::DiracMixturePdf& out, double scale, int count);
+
     //input and measurement vectors
     std::vector<aux::vector> measure;
     std::vector<Activation> stim;
@@ -162,6 +164,7 @@ int BoldPF::getStatus()
 {
     return status;
 };
+    
 
 /* Run - runs the particle filter
  * pass - variable to pass to callback function
@@ -250,13 +253,8 @@ int BoldPF::run(void* pass = NULL)
                 try {
                     filter->setFilteredState( resampler_reg->resample(
                                 filter->getFilteredState(), statecov) );
-                } catch(int err) {
-                    aux::vector mu2 = filter->getFilteredState().
-                                getDistributedExpectation();
-                    aux::vector cov2 = filter->getFilteredState().
-                                getDistributedExpectation();
-                    if(rank == 0) {
-                        *debug << "Error: " << err << endl;
+                } catch(...) {
+                    if(rank == 0){
                         *debug << "Ess: " << ess << endl;
                         *debug << "Mu: ";
                         outputVector(*debug, tmpmu);
@@ -264,9 +262,8 @@ int BoldPF::run(void* pass = NULL)
                         outputMatrix(*debug, statecov);
                         *debug << endl;
                     }
[…]

### 2010-03-01 code/include/RegularizedParticleResamplerMod.hpp

@@ -174,14 +174,14 @@ RegularizedParticleResamplerMod<NT, KT>::resample_help(
     double weight = 0;
 
     /* standardise particles */
-    for(unsigned int i = 0 ; i < model->getMeasurementSize() ; i++) {
-        unsigned int tmp = model->getStateSize()-i-1;
-        if(covariance(tmp, tmp) < 1e-200) {
-            std::cerr << "Warning covariance of drift (" << tmp 
-                        << ") dropped to 0, resetting" << std::endl;
-            covariance(tmp, tmp) = 1e-100;
-        }
-    }
+//    for(unsigned int i = 0 ; i < model->getMeasurementSize() ; i++) {
+//        unsigned int tmp = model->getStateSize()-i-1;
+//        if(covariance(tmp, tmp) < 1e-200) {
+//            std::cerr << "Warning covariance of drift (" << tmp 
+//                        << ") dropped to 0, resetting" << std::endl;
+//            covariance(tmp, tmp) = 1e-100;
+//        }
+//    }
     aux::matrix sd(covariance);
     aux::vector diag_v(sd.size1());
     int err = lapack::syev('V', 'U', sd, diag_v);

### 2010-03-01 code/tools.cc

@@ -48,7 +48,7 @@ Image4DType::Pointer pctDiff(const Image4DType::Pointer input1,
             if(input2->GetRequestedRegion().IsInside(index)) {
                 printf("%li %li %li %li\n", iter1.GetIndex()[0], iter1.GetIndex()[1],
                             iter1.GetIndex()[2], iter1.GetIndex()[3]);
-                itero.Set(fabs((input2->GetPixel(index)-iter1.Get())/input2->GetPixel(index)));
+                itero.Set((input2->GetPixel(index)-iter1.Get())/input2->GetPixel(index));
             }
             ++iter1; ++itero;
         }
@@ -151,11 +151,6 @@ void outputMatrix(std::ostream& out, indii::ml::aux::matrix mat)
   }
 }
 
-void fftline()
-{
-
-}
-
 unsigned int round_power_2(unsigned int in) 
 {
     unsigned int count = 0;
@@ -227,7 +222,7 @@ itk::OrientedImage<double, 4>::Pointer fft_image(
 
     //Round up the nearest power of 2 for image length
     //and create a temporary image for FFT's
-    lineSize[0] = round_power_2(inSize[3]);
+    lineSize[0] = round_power_2(inSize[3])*4;
     Real4DT::SizeType timeSizeOut = {{1, 1, 1, lineSize[0]}};
     Real1DT::Pointer working = Real1DT::New();
     working->SetRegions(lineSize);

### 2010-03-01 doc/thesis/thesis.tex
Merge ssh://oldcub.ath.cx/home/micahc/brainid

@@
 Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
-neural activity. At the cost of rather limited temporal resolution, FMRI gives
-high spatial for relatively cheap. Whereas other methods
-of analyzing neural signals can be invasive or ill-posed, 
-FMRI is relatively quick and easy, and its analysis straight forward. Because of
-these benefits, FMRI has great potential for studying human cognition.
-Unfortunately, FMRI does suffer from severe noise and, as previously
-mentioned, low temporal resolution. Because of the extensive noise, 
-and relative unpredictability of the absolute signal level, a large amount 
-of conditioning is necessary to obtain "valid" activation. In
+neural activity. Despite its rather limited temporal resolution, FMRI still
+is the best way of measuring neural activity deep in the brain. 
+Whereas other methods
+of analyzing neural signals can be invasive or difficult to acquire, 
+FMRI is relatively quick and cheap, and its analysis is straight forward.
+Because of these benefits, FMRI continues to have great potential for
+studying human cognition. Despite the fact that FMRI is so widely used,
+the choice of analysis tools is still relatively limited. Every widely
+available analysis tool is based on parametric modeling, which \emph{requires}
+prior knowledge of the noise, and depends on that knowledge being accurate.
+This limitation casts a long shadow over any experiment that claims to be
+able to reject the null hypothesis, no matter the P-value. While obviously
+there is a correlation between the BOLD signal and actual activity, its
+possible the current parametric tests are over-estimating the actual amount
+of activity. Its worht noting that the opposite could also be true, parametric
+based t-tests may in fact underestimate activity.
+
 fact, typical activation detection does little to account for variations
@@
 
-\section*{General Linear Model}
+\subsection*{FMRI?}
 
-\section*{BOLD Model}
+\subsection*{BOLD Physiology}
+The physiology that leads to the BOLD model has been well studied over the 
+past decade and a half. While some uncertainty remains, for instance in
+relation to the BOLD post-stimulus undershoot <papers relating to back and
+forth on post-stimulus undershoot>, generally the model fits reality very 
+well. As with every piece of tissue in the body, the brain requires oxygen
+to extract energy from glucose in the blood. This process of removing oxygen
+from red blood cells sets into motion a chain of events
 
-\subsection*{Preprocessing}
-After FMRI data has been acquired it is always necessary to modify the
-data in some way to make different runs comparable. Because FMRI signal
-levels are essentially unit-less, at the very least it is necessary to convert
-the data into \% difference from the baseline. Finding the baseline is the
-first hurdle in dealing with FMRI. Various methods exist for doing this,
-but they are all ad-hoc. It is common to treat each stimulus pulse independently,
-ignoring the long decay of the BOLD signal. By treating each input separately,
-the "base" for calculating the percent difference is merely the signal level 
-before the stimulus is applied. The problem with this is that the BOLD signal
-can be heavily delayed, and fall times can be as long as 10 or 15 seconds,
-meaning that the assumption is clearly a false one. Another popular method 
-of dealing with this problem, is to put all the time series through a high
-pass filter. This will of course remove the DC component of the signal, and 
-some amount of so called "drift". The problem with this method is that it is
-not adaptive to the input. Huge variations in drift frequencies can exist 
-in a single time-series. Thus, a single cutoff frequency could miss a significant
-drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
-set too high.
+inherently alters the local ratio of Oxygenated Hemoglobin
+to De-oxygenated hemoglobin, which of course changes the atomic makeup of the
+region. Of course MRI's contrast is based on the nuclear spin excitability of
+atoms, and so change in the amount of oxygen changes the local oxygen concentration of 
+The changes in the FMRI signal result from the changes in the ratio
+of Oxygenated Hemaglobin to Deoxygenated Hemaglobin.
 
-\subsection*{
+\section*{Current Techniques}
+\subsection*{Basic T-Test}
+\subsection*{General Linear Model}
 
-\section*{Introduction to Modeling}
+\section*{Proposed Approach}
+\subsection{Modeling}
 The short term goal of current neurological models is to
@@
 reduced allowing for much more reasonable statistical thresholds.
+\subsection{Particle Filters}
+
+\section*{Methods}
+\subsection*{Preprocessing}
+After FMRI data has been acquired it is always necessary to modify the
+data in some way to make different runs comparable. Because FMRI signal
+levels are essentially unit-less, at the very least it is necessary to convert
+the data into \% difference from the baseline. Finding the baseline is the
+first hurdle in dealing with FMRI. Various methods exist for doing this,
+but they are all ad-hoc. It is common to treat each stimulus pulse independently,
+ignoring the long decay of the BOLD signal. By treating each input separately,
+the "base" for calculating the percent difference is merely the signal level 
+before the stimulus is applied. The problem with this is that the BOLD signal
+can be heavily delayed, and fall times can be as long as 10 or 15 seconds,
+meaning that the assumption is clearly a false one. Another popular method 
+of dealing with this problem, is to put all the time series through a high
+pass filter. This will of course remove the DC component of the signal, and 
+some amount of so called "drift". The problem with this method is that it is
+not adaptive to the input. Huge variations in drift frequencies can exist 
+in a single time-series. Thus, a single cutoff frequency could miss a significant
+drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
+set too high.
+
+\section*{Introduction to Modeling}
 

### 2010-03-01 scripts/bview.py

@@ -1,0 +1,183 @@
+#!/usr/bin/python
+
+from matplotlib.widgets import Slider, Button, RadioButtons
+import pylab
+import sys
+import nibabel
+import time
+import numpy
+
+global g_axslice
+global g_sliceSlide
+global g_timeSlide
+global g_dir
+
+def getequivindex(point, targetimg, img2):
+#    trans = [targetimg.get_header()['srow_x'], targetimg.get_header()['srow_x'], \
+#                targetimg.get_header()['srow_x']]
+#    nibabel.orientations.apply_orientation(point, trans)
+#    print targetimg, img2
+    head1 = targetimg.get_header()
+    forward = numpy.matrix([head1['srow_x'][0:3], head1['srow_y'][0:3], head1['srow_z'][0:3]])
+    point = forward*(numpy.matrix(point).T)
+    point = point + numpy.matrix([head1['qoffset_x'], head1['qoffset_y'], head1['qoffset_z']]).T
+#    print point
+    #going to other ...
+    head2 = img2.get_header()
+    reverse = numpy.matrix([head2['srow_x'][0:3], head2['srow_y'][0:3], head2['srow_z'][0:3]]).I
+#    print reverse
+    point = point - numpy.matrix([head2['qoffset_x'], head2['qoffset_y'], head2['qoffset_z']]).T
+#    print point
+    point = reverse*point
+#    print point
+    return [round(num) for num in point[:,0]]
+
+
+def imshowslice(dir, slnum, time):
+    if dir == 'x-y':
+        pylab.subplot(111)
+        pylab.imshow(images[0].get_data()[:,:,slnum,time])
+    elif dir == 'y-z':
+        pylab.subplot(111)
+        pylab.imshow(images[0].get_data()[slnum,:,:,time])
+    elif dir == 'x-z':
+        pylab.subplot(111)
+        pylab.imshow(images[0].get_data()[:,slnum,:,time])
+
+
+#Input
+images = list()
+for arg in sys.argv[1:]:
+    print arg
+    images.append(nibabel.load(arg))
+
+#Initialize
+g_dir = 'x-y'
+fig3d = pylab.figure()
+ax = fig3d.add_subplot(111)
+pylab.subplots_adjust(left=0.25, bottom=0.25)
+imshowslice(g_dir, 0, 0)
+
+# - - - - - - - - - -
+# Initialize sliders
+# - - - - - - - - - - 
+axcolor = 'lightgoldenrodyellow'
[…]

### 2010-03-01 scripts/bview.py
"finished" with bview.py

@@ -35,14 +35,13 @@ def getequivindex(point, targetimg, img2):
 
 def imshowslice(dir, slnum, time):
     if dir == 'x-y':
-        pylab.subplot(111)
-        pylab.imshow(images[0].get_data()[:,:,slnum,time])
+        data = [images[0].get_data()[:,i,slnum,time] for i in range(images[0].get_shape()[1])]
     elif dir == 'y-z':
-        pylab.subplot(111)
-        pylab.imshow(images[0].get_data()[slnum,:,:,time])
+        data = [images[0].get_data()[slnum,:,i,time] for i in range(images[0].get_shape()[2])]
     elif dir == 'x-z':
-        pylab.subplot(111)
-        pylab.imshow(images[0].get_data()[:,slnum,:,time])
+        data = [images[0].get_data()[:,slnum,i,time] for i in range(images[0].get_shape()[2])]
+    pylab.subplot(111)
+    pylab.imshow(data,cmap='gray', origin='lower')
 
 
 #Input
@@ -116,15 +115,15 @@ def mousecall(event):
             #get position
             pos = getequivindex([x,y,z], images[0], images[i])
             print pos
-            if pos[0] < 0 or pos[0] > images[i].get_header()['dim'][1]:
+            if pos[0] < 0 or pos[0] >= images[i].get_header()['dim'][1]:
                 print "Skipping, point not in image: " + sys.argv[i+1]
                 continue
             
-            if pos[1] < 0 or pos[1] > images[i].get_header()['dim'][2]:
+            if pos[1] < 0 or pos[1] >= images[i].get_header()['dim'][2]:
                 print "Skipping, point not in image: " + sys.argv[i+1]
                 continue
             
-            if pos[2] < 0 or pos[2] > images[i].get_header()['dim'][3]:
+            if pos[2] < 0 or pos[2] >= images[i].get_header()['dim'][3]:
                 print "Skipping, point not in image: " + sys.argv[i+1]
                 continue
 

### 2010-03-02 doc/thesis/thesis.tex
little cleanup in bview. more thesis work.

@@
 Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
-neural activity. Despite its rather limited temporal resolution, FMRI still
-is the best way of measuring neural activity deep in the brain. 
+of neural activity. Despite its rather limited temporal resolution, FMRI is still
+the best way of measuring neural activity deep in the brain. 
 Whereas other methods
@@
 FMRI is relatively quick and cheap, and its analysis is straight forward.
-Because of these benefits, FMRI continues to have great potential for
-studying human cognition. Despite the fact that FMRI is so widely used,
+Because of these benefits, FMRI continues to be crucial to the study of human 
+cognition. Despite the fact that FMRI is so widely used,
 the choice of analysis tools is still relatively limited. Every widely
-available analysis tool is based on parametric modeling, which \emph{requires}
-prior knowledge of the noise, and depends on that knowledge being accurate.
-This limitation casts a long shadow over any experiment that claims to be
+available analysis tool is based on parametric modeling, which requires
+prior knowledge of the noise distribution. Indeed, the linear methods
+usually applied are known not to be robust to non-gaussian, non-white
+noise, which is why so many pre-processing steps are necessary.
+These limitations cast a long shadow over any experiment that claims to be
 able to reject the null hypothesis, no matter the P-value. While obviously
@@
 possible the current parametric tests are over-estimating the actual amount
-of activity. Its worht noting that the opposite could also be true, parametric
-based t-tests may in fact underestimate activity.
+of activity. Its also possible that the current methods underestimate
+actual activity. In this paper we propose a robust model-based approach to 
+activation detection that makes no assumptions about the underlying noise
+model.
 
@@
 
-\subsection*{FMRI?}
+\subsection*{FMRI}
+Before going further, it is necessary to discuss Functional MRI,
+and what it measures. All MRI works by exciting nuclear spin away
+from a steady state controlled by a large electromagnet. Different 
+nuclei respond to RF impulses based on their natural frequency
+of oscillation. Therefore contrast in MRI is based on the aggregate
+nuclear composition of a given region. Thus, regions with more oxygen
+will respond differently from regions with proportionately less 
+oxygen. This exact difference is what causes the changes over time
+in the MR signal in functional MRI. Unfortunately, it takes a significant
+amount of time for nuclear spin to return to equilibrium. Thus instead
+of waiting, a type of imaging called Echo Planar Imaging (EPI) is used,
+wherein a single excitation is used, and then as quickly as possible
+every voxel (3 Dimensional Pixel) in a particular region is read. 
+Because "reading" a particular region requires the switching on of a
+gradient magnetic field, it takes a decent amount of time to perform
+measurements. Problematically, the signal from the original excitation
+drops exponentially, thus the signal to noise ratio can be very low
+even at modest resolutions. Eventually the signal becomes unreadable,
+and a new excitation must be applied. As more measurements are made, and thus
+as the spatial resolution increases, the rate at which entire volumes
+can be acquired drops. Typically volumes can be acquired every two to
+three seconds, at resolutions of about 64 by 64 by 32, but their signal
+to noise ratio is often poor. Other imaging modalities such as EEG and
+MEG are capable of much faster measurements, but are unable to isolate
+measurement locations so precisely as FMRI. As with all biological systems,
+noise is much higher than in human designed systems, and yet these 
+biological systems are capable of performing amazing feats.
 
@@
 The physiology that leads to the BOLD model has been well studied over the 
-past decade and a half. While some uncertainty remains, for instance in
+past decade and a half. While some uncertainty remains (for instance in
 relation to the BOLD post-stimulus undershoot <papers relating to back and
-forth on post-stimulus undershoot>, generally the model fits reality very 
+forth on post-stimulus undershoot>) generally the model fits reality very 
 well. As with every piece of tissue in the body, the brain requires oxygen
 to extract energy from glucose in the blood. This process of removing oxygen
-from red blood cells sets into motion a chain of events
+from red blood cells sets into motion a chain of events that locally 
+alters the composition of the blood. Inactive neurons can be thought of
+as a slingshot cocked and ready to fire. As soon as a signal moves up
+the axon and causes the neuron to fire (and thus changes the state
+of a membrane at that location), ions quickly move across the altered
+membrane to compensate for a high charge imbalance between the interior
+and exterior of the cell. This process, similar to allowing a stretched
+rubber band to contract leaves the system at a lower energy state. This cannot
+last though, because the neuron needs to be ready to fire again rather quickly.
+To return the cell to a ready position the membrane becomes impermeable
+to the ions again, and then begins pumping ions back into the cell. This
+process takes a large amount of energy, whereas the actual firing takes
+very little energy. Thus, after firing, glucose is burned, removing oxygen
+from the blood and thus causing a dip in the ratio of oxygenated hemoglobin to 
+de-oxygenated hemoglobin. This is the first potentially measurable effect
+that MRI can see, although it usually lasts fewer than 2 seconds making it
+difficult for FMRI to catch. As a result of the decreased amount of local oxygen,
+the capillaries compensate by increasing blood flow to that region. Because,
+of the quickly increased blood flow into local capillaries, the blood
+volume increases in addition to the local oxygen content. This leads to a Windkessel
+effect which further lags the normalization of oxygen content. The effect of
+all this is an overshoot in the oxygen content above the initial level. After
+the work of recharging is done, there may be a prolonged undershoot lasting
+as much as 90 seconds \cite{spatial_poststim}, though the reason for this
+is debated \cite{origin_poststim}.
 
-inherently alters the local ratio of Oxygenated Hemoglobin
-to De-oxygenated hemoglobin, which of course changes the atomic makeup of the
-region. Of course MRI's contrast is based on the nuclear spin excitability of
-atoms, and so change in the amount of oxygen changes the local oxygen concentration of 
-The changes in the FMRI signal result from the changes in the ratio
-of Oxygenated Hemaglobin to Deoxygenated Hemaglobin.
+All these effects are generally accepted as the cause of the BOLD signal, 
+but FMRI doesn't detect this happening in one neuron, but rather as the 
+aggregate over hundreds or millions of cells. Though local neurons act
+"together" (i.e. around the same time), the density of neurons, the
+density of capillaries, and slight differences in activation across 
+a particular voxel can all lead to signal attenuation or noise. 
+A particularly insidious type of noise present in FMRI is a low frequncey
+drift, characterized by a Weiner process. Though not present in all
+regions, it is prevalent enough to cause problems \cite{detrend}. It is still not
+clear where exactly this noise comes although it is possible it is 
+the result of magnets heating up, or some distortion in magnetic
+fields. It is clear that this drift signal is not due to a true physiological
+effects however, given its presence in cadavers and phantoms\cite{drift}.
 
@@
 \subsection*{Basic T-Test}
+The most basic method of analyzing FMRI data is through a basic T-test
+between "resting state" and "active state" samples. This is done by 
+taking the average and variance of the inactive period, and the 
+active period separately then treating them both as gaussian distributions.
+If they are in fact Gaussian distributions, then a basic t-test will
+give the likelihood that the samples came from the same distribution
+(the p-value). Of course, this test is fraught with problems; even if
+the drift mentioned earlier has been removed, there is little reason
+to believe that the noise is Gaussian, or even stable. Additionally, 
+even if the noise were Gaussian, a t-test with a p-value of .05 over
+5000 or more samples is on average going to generate $.05*5000$ false
+positives. To compensate for this, bonferoni correction, also known as
+multiple comparison tests are performed; essentially p-values are 
+divided by the number of independent
+tests being run. This, however, leads to extremely low p-values, so
+low that it would be impossible for any biological system to satisfy. To
+compensate, a Gaussian kernel is applied to the image, thus reducing
+variance (and thus separating the active and inactive distributions)
+as well as decreasing the effective number of voxels. Since t-tests are
+now no longer being applied to n <I need to define n> independent voxels,
+the factor by which the p-value must be divided by can be decreased.
+<Do I need to mathematically define all this?> The derivation and application
+of random field theory, and its use can be found in various papers \cite{univ_mult_rft}.
+
 \subsection*{General Linear Model}
+
+\subsection*{Whats wrong with these techniques}
+There are a few problems with the techniques mentioned in the 
+previous sections. First, they essentially ignore prior knowledge about
+the system. The physiological models for the BOLD signal are quite good
+and based in physics. Thus the parameters may be constrained by reality,
+which crucially reduces model variance. Additionally, lacking real world
+parameters means that the parameters have no clear physiological meaning,
+so they are less useful in identifying pathologies.
+
+Secondly, these methods are still merely t-tests, which notoriously
+lack robustness to non-Gaussian noise. While different techniques 
+exist for imposing Gaussianity, those techniques are incapable of
+discriminating noise from signal, otherwise analyzing FMRI would be
+relatively simple. There is no way to know how much signal is removed
+by various smoothing techniques, or even if entire regions have been
+smoothed into oblivion. Instead of extensively filtering data to 
+remove noise, the analysis method itself must be robust to noise,
+which is why we propose here the use of particle filters.
 

### 2010-03-02 scripts/bview.py

@@ -13,23 +13,15 @@ global g_timeSlide
 global g_dir
 
 def getequivindex(point, targetimg, img2):
-#    trans = [targetimg.get_header()['srow_x'], targetimg.get_header()['srow_x'], \
-#                targetimg.get_header()['srow_x']]
-#    nibabel.orientations.apply_orientation(point, trans)
-#    print targetimg, img2
     head1 = targetimg.get_header()
     forward = numpy.matrix([head1['srow_x'][0:3], head1['srow_y'][0:3], head1['srow_z'][0:3]])
     point = forward*(numpy.matrix(point).T)
     point = point + numpy.matrix([head1['qoffset_x'], head1['qoffset_y'], head1['qoffset_z']]).T
-#    print point
     #going to other ...
     head2 = img2.get_header()
     reverse = numpy.matrix([head2['srow_x'][0:3], head2['srow_y'][0:3], head2['srow_z'][0:3]]).I
-#    print reverse
     point = point - numpy.matrix([head2['qoffset_x'], head2['qoffset_y'], head2['qoffset_z']]).T
-#    print point
     point = reverse*point
-#    print point
     return [round(num) for num in point[:,0]]
 
 
@@ -143,13 +135,6 @@ def mousecall(event):
         pylab.show()
 
 cid = fig3d.canvas.mpl_connect('button_press_event', mousecall)
-
-#resetax = pylab.axes([0.8, 0.025, 0.1, 0.04])
-#button = Button(resetax, 'Reset', color=axcolor, hovercolor='0.975')
-#def reset(event):
-#    sfreq.reset()
-#    samp.reset()
-#button.on_clicked(reset)
 
 #!======================
 #! Interactive - Direction

### 2010-03-03 doc/thesis/thesis.tex
some more work on the thesis

@@
 \section*{Current Techniques}
-\subsection*{Basic T-Test}
+\subsection*{Basic Statistical Parametric Mapping}
 The most basic method of analyzing FMRI data is through a basic T-test
@@
 \subsection*{General Linear Model}
+The most used form of FMRI analysis is still based on Statistical Parametric
+Mapping, but is able to account for several different levels or types
+of activation. By adding a General Linear Model to the analysis, the
+output signal timeseries (what FMRI detects) is regressed over the weighted
+sum of various confound timeseries. The equation for a general linear
+model is then
+\begin{equation}
+Y(t) = X(t)\beta + \epsilon(t)
+\end{equation}
+where $Y(t)$ is the smoothed or detrended timeseries of measurements,
+$X(t)$ is a row vector of stimuli, $\beta$ is a column vector of weights,
+and $\epsilon$ is the error. Thus for every time, the measurement is
+assumed to be a weighted sum of the inputs plus some error. The calculation
+of $\beta$ then is merely a gradient descent search to minimize the
+mean squared error.
+
+<Image of GLM>
+
+Of course, a square wave input is not going to result in a square wave
+in the activation of brain regions. Thus, various methods are used to 
+smooth $X(t)$ through time, and bandlimit the input. The best technique
+is convolving the stimulus input with a hemodynamic response function,
+which mimicks the basic shape of BOLD activation, including a delay
+due to rise time and fall time. This hemodynamic signal is static however,
+so every region of the brain still gets matched to the same signal,
+although the weights of various stimulus or confounds (such as attention)
+are allowed to vary.
+
+<Image of Hemodynamic Response Function>
+
+Ultimately, activation due to a particular stimuli is decided by the 
+$\beta$ value corresponding to that stimuli's column of $X(t)$.
+<Need to check this>
+The null hypothesis as to whether the outcome was random is then
+based on a t-test of the $\epsilon(t)$ timeseries.
 
@@
 exist for imposing Gaussianity, those techniques are incapable of
-discriminating noise from signal, otherwise analyzing FMRI would be
-relatively simple. There is no way to know how much signal is removed
+discriminating noise from signal. There is no way to know how much 
+signal is removed
 by various smoothing techniques, or even if entire regions have been
 smoothed into oblivion. Instead of extensively filtering data to 
-remove noise, the analysis method itself must be robust to noise,
-which is why we propose here the use of particle filters.
+remove noise, the analysis method itself must be robust arbitrary
+types of noise, which is why we propose here the use of particle filters.
 
 \section*{Proposed Approach}
-\subsection{Modeling}
-The short term goal of current neurological models is to
-learn something about the system being modeled.
-While control (for instance in prosthetics), or predication
-(such as BlueBrain) may some day be feasible, in the
-short run we still have a great deal to learn about how our
-brain works. By learning about specific systems (the dynamics
-of a single person's brain), we may be able to diagnose diseases,
-find their causes and even someday rectify the cause. Of course
-by learning about various specific systems we can also learn 
-about the general way our brains work. The reason then, for 
-studying the BOLD model then is
-to be able to derive the various system parameters and understand
-why one brain region (say a voxel) reacts in one way, when another
-acts completely differently.
+\subsection{Goal}
+The ultimate goal of modeling neural activation is to learn something
+new about how the brain works. The General Linear Model, and 
+SPM's goal is to find a correlation between input, or what a 
+person is thinking, and which regions are active. This is an 
+important task, but it misses the potential of FMRI. FMRI's 
+greatest weakeness as a tool for studying neural activity is that
+it is connected to neurons through a long chain of events. But
+what if, in addition to seeing which brain regions are active
+for which stimuli, we could also study pathologies such as poor
+blood flow, or derive physical parameters such as blood vessel
+compliance? By using a realistic physical model as the basis for
+discovering the relation between the outside world and FMRI 
+signal, the path to detecting activation gives all those other
+things for free, precisely \emph{because} FMRI is the result of such
+a long chain of events. It is actually quite common to use models to derive
+the values of physiological parameters. For instance, blood volume
+is often measured by injecting a dye into the blood stream and then
+assuming that the dye is removed at a rate proportional to the
+concentration in the blood. By performing a simple regression it is
+possible to estimate blood volume quite accurately. While for
+this particular project I don't claim that the results will give 
+physiologically plausible results, with some tuning of the algorithm
+it would very possible.
+
+While the General Linear Model has historically done well with
+detecting activation, it is limited to scaling the entire input
+signal up or down. This leaves out even basic effects such as 
+longer or shorter time constants, as well as nonlinear effects which
+are known to be present in the BOLD system. 
+**
 

### 2010-03-15 doc/thesis/thesis.tex
work on particle filters+

@@
 
+A large number of models have been proposed for the BOLD signal with varying
+amounts of complexity. The simplest model is the so called "Balloon Model"
+which first proposed the windkessel effect as a factor the BOLD response.
+The model we will use is the model proposed by Buxton Et. Al. and later used
+in Riera et. al. \cite{Riera}. The model has four state variables, $s$, $f$,
+$v$, $q$, representing flow inducing signal, cerebral blood flow, cerebral
+blood volume and deoxyhemaglobin to hemaglobin ratio, respectively. The 
+state variables change over time, given by the state evolution equations:
+\begin{equation}
+\dot{s} = \epsilon u -  
+\end{equation}
+
+\begin{equation}
+\dot{f} = s
+\end{equation}
+
+\begin{equation}
+\dot{v} = 
+\end{equation}
+
+\begin{equation}
+\dot{q} = 
+\end{equation}
+
+Additional papers have added such things as
+....
+
 All these effects are generally accepted as the cause of the BOLD signal, 
@@
 Mapping, but is able to account for several different levels or types
-of activation. By adding a General Linear Model to the analysis, the
+of stimulus. By adding a General Linear Model to the analysis, the
 output signal timeseries (what FMRI detects) is regressed over the weighted
-sum of various confound timeseries. The equation for a general linear
+sum of the various confound's timeseries. The equation for a general linear
 model is then
@@
 due to rise time and fall time. This hemodynamic signal is static however,
-so every region of the brain still gets matched to the same signal,
-although the weights of various stimulus or confounds (such as attention)
-are allowed to vary.
+so every region of the brain gets the same design matrices (X(t)), 
+although the weights of various stimulus or confounds are allowed to vary.
 
@@
 previous sections. First, they essentially ignore prior knowledge about
-the system. The physiological models for the BOLD signal are quite good
-and based in physics. Thus the parameters may be constrained by reality,
-which crucially reduces model variance. Additionally, lacking real world
-parameters means that the parameters have no clear physiological meaning,
-so they are less useful in identifying pathologies.
+the system. Although the most advanced form of the general linear model includes
+a "Hemodynamic Response Function," that hemodynamic response function is
+static across every region of the brain. It is well known that capillary beds
+are not uniform and so blood perfusion cannot possible be static across the
+brain. Thus, if extra information were available a-priori, that information
+could not be incorporated without modifications to the General Linear Model.
+Similarly heart rate could not be added either. It would obviously be advantageous
+to have true physiological parameters as entry points for these various other
+model parameters. The physiological models for the BOLD signal are quite good
+and based on realistic physics. While the exact connection between a stimulus
+and the flow inducing signal is not precisely known, model fits are actually
+quite good \cite{ModelCompareison}. Regardless, being based on some real 
+physiological 
+parameter would allow for the establishment of reasonable priors and decrease
+model variance without breaking a sweat. Of course, using real parameters 
+has the additional
+bonus of potentially providing information about physical pathologies. It
+is quite possible that physical properties such as decreased compliance of
+blood vessels could indicate a neurological condition that is not easily
+seen in a T1 or T2 map. In essence, this could make FMRI a much more 
+useful clinical tool than it is now. The other problem with linear models
+is that they are a linear fit to a nonlinear signal. It is not uncommon
+for data to be thrown out in FMRI studies because no significant activation
+has been seen. However, if, for whatever reason, the BOLD response
+was acting more nonlinear than in other patients it would be completely
+possible for SPM to miss that activation.
 
-Secondly, these methods are still merely t-tests, which notoriously
+<Image with two different $\alpha$s>
+
+<image comparing the results of 10\% changes in various signals>
+
+Secondly, these methods are still based on t-tests, which notoriously
 lack robustness to non-Gaussian noise. While different techniques 
@@
 discriminating noise from signal. There is no way to know how much 
-signal is removed
-by various smoothing techniques, or even if entire regions have been
-smoothed into oblivion. Instead of extensively filtering data to 
-remove noise, the analysis method itself must be robust arbitrary
-types of noise, which is why we propose here the use of particle filters.
+signal is removed by various smoothing techniques, or even if entire 
+regions have been smoothed into oblivion. Instead of extensively filtering data to 
+remove noise, the analysis method itself must be robust a wide range of noise,
+which is why we propose here the use of particle filters.
 
 \section*{Proposed Approach}
-\subsection{Goal}
-The ultimate goal of modeling neural activation is to learn something
-new about how the brain works. The General Linear Model, and 
+\subsection*{Goal}
+The ultimate goal of this project is to provide a new set of tools
+for analyzing FMRI data. Whereas SPM techniques have been highly 
+successful at finding macroscopic regions of activation, linear 
+modeling can carry significant bias error due to lack of model
+flexibility. While adding parameters can significantly increase
+error due to model variance, this effect is mitigated by the fact
+that we plan to use a model that is based on first principals. The
+purpose of this paper is thus to evaluate the potential of using
+a particle filter along with the BOLD model to derive physical 
+parameters. In so doing, we hope to be able to show that neuronal
+efficacy, $\epsilon$ is a sutiable variable for estimating voxel 
+activation from a standard FMRI image. We also hope to show that 
+estimated posterior distribution of the parameters, derived from
+the particle filter, is able to provide an accurate measure of the
+confidence interval.
+
+\subsection*{Introduction to Particle Filters}
+Particle filters, also known as Sequential Monte-Carlo methods (SMC),
+are essentially an on-line version of the Markov-Chain Monte Carlo (MCMC)
+algorithm. Similar to MCMC, Particle Filters assume the Markov Assumption
+holds, in essence that 
+\begin{equation}
+P(X_t | X_{t-1}, X_{t-2}, ... X_{0}) = P(X_t | X_{t-1})$
+\end{equation}
+for a time evolving parameter, $X_t$. This is actually a very good
+assumption in most biological systems, since chemicals driving physiological
+parameters rarely have memory. Unlike MCMC algorithms, particle filters
+only calculate the posterior distribution:
+\begin{equation}
+P(X_t | y_t, y_{t-1}, ... y_0)
+\end{equation}
+for $t$ measurements, $y_t, ... y_0$. This frees the algorith from having
+to calculate, or even keep track of the values of $X$ at times other than
+$t$, and significantly reducing computation time. This also easily lends
+the particle filter to solving stochastic differential equations (SDE's).
+
+...
+
+\subsection*{My Specific Particle Filte}
+Although particle filters are well defined, significant design decisions
+are application dependent. The most important choice is the weighting function,
+$w(y_t - y(\theta_t))$. Obviously the function should weight $0$ maximally,
+the drop off rate is extremely important. A very thin gaussian probability 
+distribution function has nice properties, but thin tails. As a result, large
+outliers in the measurement vector could easily force all the particles to
+have near 0 weights, thus forcing the particle filter to converge improperly. 
+On the other end of the spectrum, a cauchy pdf may not
+weight particles in the middle enough, preventing the particles to never 
+converge. Of course, the scale factor, or variance, also plays a signifcant
+role in the convergeance rate. The importance of the weighting function
+cannot be overstated, as this is the primary factor in deciding the rate at
+which the particle filter converges. As part of the simulations, I compare
+a gaussian based weighting function with an exponential weightin function.
+In both cases, I scaled the width of the weighting functions automatically based
+on the signal variance.
+
+Dealing with low frequency noise is also a crucial part of the tests. 
+I tested two different methods of dealing with this noise. Various methods
+are used to deal with this problem in other toolsets. The most common method
+is to apply a high-pass filter to the measurement vector. Although this makes
+sense in many applications, in this case it may not be adaptive enough for
+a random walk type noise. On the other hand, it has been shown that a spline
+de-trending method is able to better deal with the low frequency noise
+typically present in FMRI. Thus, my first method
+was to use a spline fit with knots every 80 or 90 measurements to trance
+the basic shape of the drift. I then added an extra DC gain parameter to
+the model to account for the fact that the de-trended signal tended to have 
+a mean of 0, rather than some positive value.
+
+My second method of dealing with low frequency drift was to counteract
+it by weighting particles based on $(y_t-y_{t-1}) - (y(\theta_t)-y(\theta_{t_1}))$ 
+rather than directly on $y_t - y(\theta_t)$. It is generally believed that the noise present in 
+FMRI has approximately gaussian steps, which means that this method only
+has to deal with gaussian white noise. Such a technique is usually not considered a
+good idea because high frequency noise is almost exclusively assumed in
+systems theory. In this case however; the noise is all in the lower frequency
+range, so performing a differentiation like operation, which magnifies high
+frequency signals, is exactly what we want.
+
+... more specifics...
+
+The General Linear Model, and 
 SPM's goal is to find a correlation between input, or what a 

### 2010-03-19 doc/thesis/thesis.tex
worked on particle filter stuff a bit

@@
 parameters. In so doing, we hope to be able to show that neuronal
-efficacy, $\epsilon$ is a sutiable variable for estimating voxel 
+efficacy, $\epsilon$ is a suitable variable for estimating voxel 
 activation from a standard FMRI image. We also hope to show that 
@@
 \subsection*{Introduction to Particle Filters}
-Particle filters, also known as Sequential Monte-Carlo methods (SMC),
-are essentially an on-line version of the Markov-Chain Monte Carlo (MCMC)
-algorithm. Similar to MCMC, Particle Filters assume the Markov Assumption
-holds, in essence that 
+Particle filters, also known as Sequential Monte Carlo (SMC) methods
+are a powerful way of estimating the posterior probability distribution
+of a set of parameters. Unlike Markov Chain Monte Carlo estimation, particle
+filters extend relatively easily to any finite impulse response (FIR) system.
+And unlike Variational Bayes or other gradient descent algorithms, particle
+filters can be run online, with relatively lower cost. The idea of the particle
+filter is to start with a wide mixture PDF of possible parameter sets, 
+and then, as measurements come in, to weight more heavily parameter sets 
+that tend to give good estimations of the measurements. The reliance on
+an initial mixture PDF obviously introduces bias; however, it is often quite
+easy to establish a reasonable range of parameters, especially when the
+model being used has a physical meaning. Suppose either a set or stream
+of measurements $y_t$ are made at discrete times $t = 0, 1, ... $, where
+$y_t$ could be a vector of measurements. Then the goal is to find the set
+of parameters $\hat{\theta}$ that give the best estimate, $\hat{y}_t$, of $y_t$.
+In a biological case, with a model based on first principals, there is 
+presumably some true set of parameters, $\theta$, a true model $f$
+and a true observation equation, $g$ that dictate how the stimulus 
+produce noisy measurements,
 \begin{equation}
-P(X_t | X_{t-1}, X_{t-2}, ... X_{0}) = P(X_t | X_{t-1})$
+\dot{x}(t) = f(t, x(t), u(t), \theta) + \nu_d
+
+y(t) = g(t, x(t), u(t), \theta) + \nu_s
 \end{equation}
-for a time evolving parameter, $X_t$. This is actually a very good
-assumption in most biological systems, since chemicals driving physiological
-parameters rarely have memory. Unlike MCMC algorithms, particle filters
-only calculate the posterior distribution:
+
+Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
+constants, $\nu_d$ is drift noise (such as a Weiner process), $\nu_s$ is additive noise, $y(t)$ is
+the vector of observations. $\nu_d$ may be driven by unknown changes in 
+cerebral blood flow, or blood oxygen content, and is assumed to be zero mean.
+Similarly $\nu_s$, stationary
+noise could be driven by unknown changes due to the heart beat and is also assumed
+to be zero mean. There are also non-biological effects that could drive these
+noise terms such as inherent MR noise.
+
+Although not generally necessary for particle filters, we will make a few
+assumptions based on the particular type of systems faced in biological 
+processes. First, that $\theta$ is time invariant. This assumption is 
+based on the idea that most physiological parameters don't change in the 
+short run unless they are driven by some other stimulus. The obvious exception
+might be something like the heart; however, period between heartbeats vary
+quickly enough that accurately accounting for heart beats necessitates a 
+new set of model parameters, or direct outside information. In short we
+assume no parameters are time varying, because not information exists to
+describe any of theme in that way. Luckily particle filters are capable 
+of dealing with non-white, non-Gaussian noise, suck unanticipated influence
+may be re-factored as noise. Secondly we assume that input cannot directly
+influence the output, which in the case of the BOLD signal is a good assumption.
+%Third, through some sort of preprocessing, we will assume that $\nu_d$ can be
+%decreased or completely removed, since it 
+Finally, $x(t)$ will include $\theta$, the unknown model constants, although
+obviously that will mean that the vector $\dot{x}$ will always have some members
+that are 0. The state space equations are then simplified to:
 \begin{equation}
-P(X_t | y_t, y_{t-1}, ... y_0)
+\dot{x}(t) = f(x(t), u(t)) + \nu_d
+
+y(t) = g(x(t)) + \nu_s
 \end{equation}
-for $t$ measurements, $y_t, ... y_0$. This frees the algorith from having
-to calculate, or even keep track of the values of $X$ at times other than
-$t$, and significantly reducing computation time. This also easily lends
-the particle filter to solving stochastic differential equations (SDE's).
 
-...
+The goal of the particle filter, is then to derive the true value of $x(t)$
+for each time, given known inputs $u(t), u(t-1), ... u(0)$  and noise 
+measurements produced by the system, $y(t), ... , y(0)$ (every time up to
+$t$). To begin with then, the particle filter starts with a prior distribution,
+which is used to build a Dirac Mixture Probability Distribution function:
+
+\begin{equation}
+P(\hat{x}(0)) = \frac{1}{N}\Sigma_0^{N-1} \delta(x_(0) - \hat{x}_i(0))
+\end{equation}
+
+where $\hat{x}(0)$ is an empirical representation of the prior, and $\hat{x}_i(0)$
+is the $i^{th}$ particle, a single possible representation of the system at time 0.
+Initially, since each $\hat{x}_i$ has been drawn form the prior, and no measurements
+have been made, the particles are all equally weighted. When a measurement becomes
+available, that information is then incorporated into the posterior by re-weighting
+the particles .
+
+\begin{equation}
+\dot{y} = -ky
+\end{equation}
+$\theta_i$ would have an estimate for $k$ \emph{and} $y$. Of course in many cases
+the model is a potentially non-linear differential equation with no closed
+form solution. Initially every $\theta_i$ will have the same weight, because
+there isn't yet anything to base a weighting on. Note that this doesn't mean
+that the particles have been evenly distributed, so the distribution is not 
+necessarily "flat" just because all the particles are equaly weighted. When a
+new measurement is available, the particles are then reweighted based on how
+well they predicted the new measurement.
 
@@
 comparison testing with very high thresholds to ensure any 
-sort of face validity \citep{drift}.
+sort of face validity \cite{drift}.
 By moving to models that actually "learn" the 
@@
 noise and most importantly that brain networks are 
-so called "Small World Networks" \citep{smallworld}, \citep{noise}. 
+so called "Small World Networks" \cite{smallworld}, \cite{noise}. 
 Merely accounting for
 nonlinearities in the time-series of the BOLD response can
-result in significantly better estimation as shown in \citep{nonlinearmodels}.
+result in significantly better estimation as shown in \cite{nonlinearmodels}.
 But even more crucial is the recent movements away from mass-
@@
 in effect implying that humans are incapable of 
-thought \citep{turing}.
+thought \cite{turing}.
 
@@
 diffusion tensor imaging and actual function layout in
-the human brain \citep{dti_dcm}. By incorporating connections
+the human brain \cite{dti_dcm}. By incorporating connections
 between regions and a realistic activation model, DCM

### 2010-03-22 doc/thesis/thesis.tex
more work with particle filters.

@@
 are a powerful way of estimating the posterior probability distribution
-of a set of parameters. Unlike Markov Chain Monte Carlo estimation, particle
+of a set of parameters give a timeseries of measurements. Unlike Markov 
+Chain Monte Carlo estimation, particle
 filters extend relatively easily to any finite impulse response (FIR) system.
-And unlike Variational Bayes or other gradient descent algorithms, particle
+Unlike Variational Bayes or simple gradient descent algorithms, particle
 filters can be run online, with relatively lower cost. The idea of the particle
@@
 \dot{x}(t) = f(t, x(t), u(t), \theta) + \nu_d
+\end{equation}
 
+\begin{equation}
 y(t) = g(t, x(t), u(t), \theta) + \nu_s
@@
 Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
-constants, $\nu_d$ is drift noise (such as a Weiner process), $\nu_s$ is additive noise, $y(t)$ is
-the vector of observations. $\nu_d$ may be driven by unknown changes in 
-cerebral blood flow, or blood oxygen content, and is assumed to be zero mean.
-Similarly $\nu_s$, stationary
-noise could be driven by unknown changes due to the heart beat and is also assumed
-to be zero mean. There are also non-biological effects that could drive these
-noise terms such as inherent MR noise.
+constants, $\nu_d$ and $\nu_s$ are bother stationary zero-mean distributions, $u(t)$
+is timeseries of stimuli and
+$y(t)$ is the vector of observations. $\nu_d$, which produces a drift term in $y(t)$
+may be driven by unknown changes in 
+cerebral blood flow, or blood oxygen content. $\nu_s$ will produce standard
+stationary white noise in $y(t)$ and may encapsulate noise by unknown changes due 
+to the heart beat. There are also significant non-biological effects that 
+could drive these noise terms such as inherent MR noise.
 
@@
 short run unless they are driven by some other stimulus. The obvious exception
-might be something like the heart; however, period between heartbeats vary
-quickly enough that accurately accounting for heart beats necessitates a 
+might be something like the heart; however, period between heartbeats vary quickly enough that accurately accounting for them would necessitate an entirely
 new set of model parameters, or direct outside information. In short we
-assume no parameters are time varying, because not information exists to
+assume no parameters are time varying, because not enough information exists to
 describe any of theme in that way. Luckily particle filters are capable 
-of dealing with non-white, non-Gaussian noise, suck unanticipated influence
+of dealing with non-white, non-Gaussian noise, so ananticipated influence
 may be re-factored as noise. Secondly we assume that input cannot directly
@@
 obviously that will mean that the vector $\dot{x}$ will always have some members
-that are 0. The state space equations are then simplified to:
+that are constant and 0. The state space equations are then simplified to:
 \begin{equation}
 \dot{x}(t) = f(x(t), u(t)) + \nu_d
+\end{equation}
 
+\begin{equation}
 y(t) = g(x(t)) + \nu_s
@@
 
-The goal of the particle filter, is then to derive the true value of $x(t)$
+The goal of the particle filter, is then to derive the probably distribution
+of the value of $x(t)$
 for each time, given known inputs $u(t), u(t-1), ... u(0)$  and noise 
@@
 $t$). To begin with then, the particle filter starts with a prior distribution,
-which is used to build a Dirac Mixture Probability Distribution function:
+which is used to build a Dirac Mixture Probability Distribution Function (PDF):
 
 \begin{equation}
-P(\hat{x}(0)) = \frac{1}{N}\Sigma_0^{N-1} \delta(x_(0) - \hat{x}_i(0))
+\hat{x}(0) \sim p(X = x) = \frac{1}{n}\sum_{i = 0}^{n-1}p_i(x)
+\end{equation}
+where $p_i(x)$ is some probability density function centered at $\hat{x}_i(0)$, particle
+$i$ at time $0$. Thus the sum of the density at $x$ from each particles' PDF is the density
+at $x$. If
+the distributions are infinitely thin, then the distribution will be discrete, otherwise
+it will be continuous.
+
+Thus $\hat{x}(0)$ is an empirical representation of the prior, and $\hat{x}_i(0)$
+is the $i^{th}$ particle, a single possible representation of the system at time 0.
+Initially, since each $\hat{x}_i$ has been drawn from the prior, and no measurements
+have been made, the particles are all equally weighted. When a measurement becomes
+available, that information is then incorporated into the posterior by re-weighting
+the particles according to how well they estimated this new measurement. 
+\begin{equation}
+Pr[\hat{x}_i(t) | y(0:t-1)] = \int Pr[\hat{x}_i(t) | \hat{x}_i(t-1), y(0:t-1)]Pr[\hat{x}_i(t-1)
+\end{equation}
+Since $\hat{x}(t) = \hat{x}(t-1) + \int_{t-1}^t f(x(t), u(t))dt + \int_{t-1}^t\nu_d dt$, 
+\begin{equation}
+Pr[\hat{x}_i(t) | \hat{x}_i(t-1)] = Pr[\int{t-1}^t \nu_d]
 \end{equation}
 
-where $\hat{x}(0)$ is an empirical representation of the prior, and $\hat{x}_i(0)$
-is the $i^{th}$ particle, a single possible representation of the system at time 0.
-Initially, since each $\hat{x}_i$ has been drawn form the prior, and no measurements
-have been made, the particles are all equally weighted. When a measurement becomes
-available, that information is then incorporated into the posterior by re-weighting
-the particles .
+of course $\int_{t-1}^t f(x(t), u(t))dt$ may need to be evaluated numerically. Therefore,
+to estimate the posterior probability at $t$, it is just necessary to propagate
+each particle forward in time according to a state space function. To form 
+the posterior probability around the true $x(t)$, each particle is weighted
+based on
+
+This
+then results in the expected value of $\hat{y}(t)$:
+\begin{equation}
+E[\hat{y}_i(t) | \hat{x}_i(t-1)] = E[\hat{x}_i(t)]
+\end{equation}
 
@@
 have near 0 weights, thus forcing the particle filter to converge improperly. 
-On the other end of the spectrum, a cauchy pdf may not
+On the other end of the spectrum, a cauchy PDF may not
 weight particles in the middle enough, preventing the particles to never 

### 2010-03-23 doc/thesis/thesis.tex
guh more on the particle filters, although I'm staring to \emph{really} understand

@@
 \subsection*{Introduction to Particle Filters}
-Particle filters, also known as Sequential Monte Carlo (SMC) methods
+Particle filters, a type of Sequential Monte Carlo (SMC) methods
 are a powerful way of estimating the posterior probability distribution
 of a set of parameters give a timeseries of measurements. Unlike Markov 
-Chain Monte Carlo estimation, particle
-filters extend relatively easily to any finite impulse response (FIR) system.
-Unlike Variational Bayes or simple gradient descent algorithms, particle
-filters can be run online, with relatively lower cost. The idea of the particle
+Chain Monte Carlo estimation, Sequential Monte-Carlo methods are designed
+to be used with parameters that vary with time. Unlike variations of the
+Kalman filter, particle filters do not make the assumption that noise
+is Gaussian. Thus particle filters are often the best solution to bayesian 
+tracking for non-linear, non-gaussian systems. 
+%todo: might need to reference attempts to model a system of our type with
+%unscented kalman filter
+The idea of the particle
 filter is to start with a wide mixture PDF of possible parameter sets, 
@@
 that tend to give good estimations of the measurements. The reliance on
-an initial mixture PDF obviously introduces bias; however, it is often quite
+an initial mixture PDF obviously introduces bias; however, every gradient
+descent must choose starting points and it is often quite
 easy to establish a reasonable range of parameters, especially when the
-model being used has a physical meaning. Suppose either a set or stream
-of measurements $y_t$ are made at discrete times $t = 0, 1, ... $, where
-$y_t$ could be a vector of measurements. Then the goal is to find the set
-of parameters $\hat{\theta}$ that give the best estimate, $\hat{y}_t$, of $y_t$.
-In a biological case, with a model based on first principals, there is 
-presumably some true set of parameters, $\theta$, a true model $f$
-and a true observation equation, $g$ that dictate how the stimulus 
-produce noisy measurements,
+model being used has a physical meaning. Suppose a set or stream of measurements
+are given, $\{y(t), t = 1, 2, 3, ... T\}$, where $T$ is permitted to go
+ to infinity. Then the goal is to find the 
+parameters, $\hat{\theta}$, and underlying state time series, $\hat{x}[0:T]$
+that minimize the difference between $\hat{y}[0:T]$
+and $y[0:T]$. In our case, we will assume that we know the form
+of the model, which is based on first principals, and that
+there is some true $\theta$ and a true time-series of underlying
+state variable, $x[0:T]$ that drives $y[0:T]$. Assuming a model form 
+such as we do here reduces model variance, potentially at the cost of increased
+bias (or systematic) error. We will assume a basic state space model:
+
 \begin{equation}
-\dot{x}(t) = f(t, x(t), u(t), \theta) + \nu_d
+\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_d)
 \end{equation}
@@
 \begin{equation}
-y(t) = g(t, x(t), u(t), \theta) + \nu_s
+y(t) = g(t, x(t), u(t), \theta, \nu_s)
 \end{equation}
@@
 Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
-constants, $\nu_d$ and $\nu_s$ are bother stationary zero-mean distributions, $u(t)$
-is timeseries of stimuli and
-$y(t)$ is the vector of observations. $\nu_d$, which produces a drift term in $y(t)$
-may be driven by unknown changes in 
-cerebral blood flow, or blood oxygen content. $\nu_s$ will produce standard
-stationary white noise in $y(t)$ and may encapsulate noise by unknown changes due 
-to the heart beat. There are also significant non-biological effects that 
-could drive these noise terms such as inherent MR noise.
+constants, $u(t)$ is a stimulus, $y(t)$ an observation, and
+$\nu_x$ and $\nu_y$ are random variates. Obviously any one of these could
+be a vector, so for instance $u(t)$ could encode multiple types of stimuli.
 
@@
 assumptions based on the particular type of systems faced in biological 
-processes. First, that $\theta$ is time invariant. This assumption is 
-based on the idea that most physiological parameters don't change in the 
-short run unless they are driven by some other stimulus. The obvious exception
-might be something like the heart; however, period between heartbeats vary quickly enough that accurately accounting for them would necessitate an entirely
-new set of model parameters, or direct outside information. In short we
+processes. First, the systems are assumed to be time invariant. This 
+assumption is based on the idea that if you froze the system for $\delta t$
+seconds, when unfrozen the system would continue as if nothing happend. 
+Few biological systems are predictible enough for them to be summarized
+by a time varying function. Although the heart may seem like an obvious
+exception, period between heartbeats vary often enough that prediction
+would necessate another state-space model. In short, we
 assume no parameters are time varying, because not enough information exists to
 describe any of theme in that way. Luckily particle filters are capable 
-of dealing with non-white, non-Gaussian noise, so ananticipated influence
+of dealing with non-white, non-Gaussian noise, so unanticipated influence
 may be re-factored as noise. Secondly we assume that input cannot directly
@@
 %decreased or completely removed, since it 
-Finally, $x(t)$ will include $\theta$, the unknown model constants, although
-obviously that will mean that the vector $\dot{x}$ will always have some members
-that are constant and 0. The state space equations are then simplified to:
+Third, we will assume noise is additive.
+Finally, $x(t)$ will encapsulate $\theta$, the unknown model constants, which
+means that the vector $\dot{x}$ will always have members
+that are 0. The results of these assumptions are a simplified version of the
+state space equations:
+
 \begin{equation}
-\dot{x}(t) = f(x(t), u(t)) + \nu_d
+\dot{x}(t) = f(x(t), u(t)) + \nu_x
 \end{equation}
@@
 \begin{equation}
-y(t) = g(x(t)) + \nu_s
+y(t) = g(x(t)) + \nu_y
 \end{equation}
 
-The goal of the particle filter, is then to derive the probably distribution
-of the value of $x(t)$
-for each time, given known inputs $u(t), u(t-1), ... u(0)$  and noise 
-measurements produced by the system, $y(t), ... , y(0)$ (every time up to
-$t$). To begin with then, the particle filter starts with a prior distribution,
+$\nu_x$ will result in something akin to adding a weiner process to $y[0:T]$, and thus
+will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
+in $y[0:T]$. For some tests I will use de-trending methods to reduce the effects of 
+$\nu_x$, the remainder of which gets re-factored into $\nu_y$. Both $\nu_x$ and $\nu_y$
+have biological and non-biological sources. MR can lead to both types of noise, 
+as demonstrated in \cite{drift}. Meanwhile changes in metabolism, heart rate, or
+other biochemical intervention could all lead to either $\nu_x$ or $\nu_y$.
+
+The goal of the particle filter to evolve a probably distribution $\hat{x}(T)$,
+that asymptotically approaches the probability distribution of $x(t)$,
+given known inputs, $u[0:T]$ and noisy readings $y[0:T]$.
+To begin with, the particle filter starts with a prior distribution,
 which is used to build a Dirac Mixture Probability Distribution Function (PDF):
@@
 the particles according to how well they estimated this new measurement.
+
+The estimated posterior of $\hat{x}(t)$, for discrete time, $t$ (before incorporating
+any measurement that might be available at $t$) is then:
 \begin{equation}
-Pr[\hat{x}_i(t) | y(0:t-1)] = \int Pr[\hat{x}_i(t) | \hat{x}_i(t-1), y(0:t-1)]Pr[\hat{x}_i(t-1)
+\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t), \hat{x}(t-1) | y[0:t-1]\} d\hat{x}(t-1)
+\end{equation}
+
+and by bayes theorem $\Pr\{\hat{x}(t), \hat{x}(t-1) | y[0:t-1]\} = 
+\Pr\{\hat{x}(t) | \hat{x}(t-1), y[0:t-1]\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}$,
+therefore 
+\begin{equation}
+\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t) | \hat{x}(t-1) ,y[0:t-1]\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}  d\hat{x}(t-1)
+\end{equation}
+
+Since in this work we will only deal with markov processes, we assume that 
+$\Pr\{\hat{x}(t) | \hat{x}(t-1), y[0:t-1]\} = \Pr\{\hat{x}(t) | \hat{x}(t-1)\}$
+
+thus the posterior becomes:
+\begin{equation}
+\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t) | \hat{x}(t-1)\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}  d\hat{x}(t-1)
+\end{equation}
+
+Therefore in between measurements, there is no need to know previous measurements in
+order to make predictions about the internal state. This is important because it means makes
+particle filters significantly more memory efficient, and although it isn't useful in this
+case, it means a particle filter could theoretically be run indefinitely without the
+need to store old data.
+
+As stated previously, new measurements may be incorporated by reweighting particles.
+To find the updated distribution, $\Pr\{\hat{x}(t) | y[0:t]\}$, bayes theorem is
+again applied:
+\begin{equation}
+\Pr\{\hat{x}(t) | y[0:t]\} = \Pr\{\hat{x}(t) | y[0:t-1], y(t)\} = \frac{\Pr\{y(t) | \hat{x}(t), y[0:t-1]\}\Pr\{\hat{x}(t) | y[0:t-1]\}}
+{\Pr\{y(t) | y[0:t-1]\}}
+\end{equation}
+
+\begin{equation}
+\Pr\{\hat{x}_i(t) | y[0:t-1]\} = \int \Pr\{\hat{x}_i(t) | \hat{x}_i(t-1), y[0:t-1]\}\Pr[\hat{x}_i(t-1)
 \end{equation}
@@
 \begin{equation}
-Pr[\hat{x}_i(t) | \hat{x}_i(t-1)] = Pr[\int{t-1}^t \nu_d]
+\Pr\{\hat{x}_i(t) | \hat{x}_i(t-1)\} = \Pr\{\int{t-1}^t \nu_d\}
 \end{equation}

### 2010-03-23 doc/thesis/thesis.tex
more particle work.

@@
 that tend to give good estimations of the measurements. The reliance on
-an initial mixture PDF obviously introduces bias; however, every gradient
-descent must choose starting points and it is often quite
+an initial mixture PDF can introduce bias; however, this effect can be
+minimized by alterring the initial weights in the mixture pdf. Of course
+every gradient descent must choose starting points and it is often quite
 easy to establish a reasonable range of parameters, especially when the
@@
 processes. First, the systems are assumed to be time invariant. This 
-assumption is based on the idea that if you froze the system for $\delta t$
+assumption is based on the idea that if you froze the system for $\Delta t$
 seconds, when unfrozen the system would continue as if nothing happend. 
@@
 
-$\nu_x$ will result in something akin to adding a weiner process to $y[0:T]$, and thus
+$\nu_x$ will result in something akin to adding a Weiner process to $y[0:T]$, and thus
 will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
-in $y[0:T]$. For some tests I will use de-trending methods to reduce the effects of 
-$\nu_x$, the remainder of which gets re-factored into $\nu_y$. Both $\nu_x$ and $\nu_y$
+in $y[0:T]$. For some of the tests, I will use de-trending methods to reduce the effects of 
+$\nu_x$, the remainder of which should be re-factored into $\nu_y$. Both $\nu_x$ and $\nu_y$
 have biological and non-biological sources. MR can lead to both types of noise, 
@@
 
-The goal of the particle filter to evolve a probably distribution $\hat{x}(T)$,
-that asymptotically approaches the probability distribution of $x(t)$,
-given known inputs, $u[0:T]$ and noisy readings $y[0:T]$.
-To begin with, the particle filter starts with a prior distribution,
-which is used to build a Dirac Mixture Probability Distribution Function (PDF):
+The goal of the particle filter is to evolve a probably distribution 
+$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
+that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
+Considering that $y$ contains measurement noise and noise in $x$ can drive
+changes in $y$, it is clear that $Pr(x(t) | u[0:T]$ is not a single true value
+but a true posterior. 
+To begin with, the particle filter starts with a proposal distribution, and $N_p$
+particles need to be drawn from that distribution, $q(x)$:
+\begin{equation}
+\{[\hat{x}_i(0),w_i] : \hat{x}_i(0) \sim q(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
+\end{equation}
+
+Where $N_p$ is the number of particles or points used to describe the prior 
+using a Mixture PDF.
 
 \begin{equation}
-\hat{x}(0) \sim p(X = x) = \frac{1}{n}\sum_{i = 0}^{n-1}p_i(x)
+Pr(\hat{x}(0) = x(0)) = \sum_{i=1}^{N_p} w_i\delta(x(0) - \hat{x}_i(0) ) dx
 \end{equation}
-where $p_i(x)$ is some probability density function centered at $\hat{x}_i(0)$, particle
-$i$ at time $0$. Thus the sum of the density at $x$ from each particles' PDF is the density
-at $x$. If
-the distributions are infinitely thin, then the distribution will be discrete, otherwise
-it will be continuous.
+
+where
+
+\begin{equation}
+w_i(0) \propto \frac{p(x_i(0))}{q(x_i(0))}
+\end{equation}
+
+is the \emph{normalized} initial weights. Note that initially $p(x_i(0)$ is
+an approximation of the proposal distribution, $q(x)$, thus by dividing
+by $q(x_i(0))$, the bias effect of the proposal distribution is negated.
+For the BOLD signal all the parameters have been studied and have relatively well
+known mean and variance, so a prior bias is actually preferred, so in our
+case we did not initially normalize the particle weights. Note however, that
+when updating weights it is necessary to use this equation, but in that case
+$q$ will not be the prior distribution but $p(\hat{x}(T) | \hat{x}(T-1)$
+
+this the definition of the weight, so when weights are updated it is necessary
+to divide by the proposal distributionto converge
+to the true distribution it is necessary to divide by the proposal distribution.
+
+Where $\delta(x-x_0)$ is 1 when $x = x_0$ (the Kronecker delta function).
+For strictly positive parameters we used a gamma distribution,
+whereas for parameters that could be negative, we used a Gaussian distribution. Both
+distributions utilized standard deviations twice that found in previous studies.
+
+Once the probability, $Pr(\hat{x}(T) | \hat{x}[0:T-1], y[0:T-1]$ has been found
+(initially this is just Mixture approximating the prior since no measurements are 
+available and no previous probabilities are available), its possible to approximate
+the probability for short times between before a measurement is available, by shifting
+the probability according the progression of the state equations. This is only 
+an approximate, since integrating $\nu_x$ should increase uncertainty as more 
+time without a measurement passes. 
+\begin{equation}
+Pr(\hat{x}(T+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i\delta\left(x - (\hat{x}_i(T) + \int_T^{T+\Delta} \dot{\hat{x}}_i(t) dt) \right) dx
+\end{equation}
+
+When a measurement becomes available it is incorporated into the probability,
+which should eventually cause the probability to converge. 
+Since the weights are the probability of of $x(t)$ being exactly $x_i(t)$, 
+the weights are defined as
+\begin{equation}
+  d
+\end{equation}
 

### 2010-03-24 doc/thesis/thesis.tex
nearly done with particle filter part.

@@
 To begin with, the particle filter starts with a proposal distribution, and $N_p$
-particles need to be drawn from that distribution, $q(x)$:
+particles need to be drawn from that distribution, $\alpha(x)$:
 \begin{equation}
-\{[\hat{x}_i(0),w_i] : \hat{x}_i(0) \sim q(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
+\{[\hat{x}_i(0),w_i] : \hat{x}_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
 \end{equation}
@@
 
-where
+Where $\delta(x-x_0)$ is 1 when $x = x_0$ (the Kronecker delta function).
 
-\begin{equation}
-w_i(0) \propto \frac{p(x_i(0))}{q(x_i(0))}
-\end{equation}
-
-is the \emph{normalized} initial weights. Note that initially $p(x_i(0)$ is
-an approximation of the proposal distribution, $q(x)$, thus by dividing
-by $q(x_i(0))$, the bias effect of the proposal distribution is negated.
+If a true prior is preferred, then the weights
+should all be $1/N_p$, and since $\hat{x}_i$ was drawn from the prior, this will
+be an approximation of the prior distribution. If a relatively flat prior is 
+preferred, then each particle's weight could be divided by the density, $\alpha(\hat{x}_i)$,
+which creates a flat prior with support points in the region of $\alpha(x)$. Either
+way, $\alpha(x)$ should be much broader than the true posterior, $Pr(x(0))$, since the
+choice of support points is crucial to the convergence of any sampling importance
+algorithm. 
 For the BOLD signal all the parameters have been studied and have relatively well
-known mean and variance, so a prior bias is actually preferred, so in our
-case we did not initially normalize the particle weights. Note however, that
-when updating weights it is necessary to use this equation, but in that case
-$q$ will not be the prior distribution but $p(\hat{x}(T) | \hat{x}(T-1)$
-
-this the definition of the weight, so when weights are updated it is necessary
-to divide by the proposal distributionto converge
-to the true distribution it is necessary to divide by the proposal distribution.
-
-Where $\delta(x-x_0)$ is 1 when $x = x_0$ (the Kronecker delta function).
-For strictly positive parameters we used a gamma distribution,
+known mean and variance, so a prior bias is actually preferred. In our
+case we did not initially flatten the distributions, but instead used 
+distributions that fit what was known about the parameters.  
+For strictly positive parameters (members of $\hat{x}$) we used a gamma distribution,
 whereas for parameters that could be negative, we used a Gaussian distribution. Both
@@
 
+Note that all the probability's implicitly depend on $u[0:T]$, so those terms 
+will be left off for simplicity.
 Once the probability, $Pr(\hat{x}(T) | \hat{x}[0:T-1], y[0:T-1]$ has been found
@@
 the probability according the progression of the state equations. This is only 
-an approximate, since integrating $\nu_x$ should increase uncertainty as more 
+an approximate, since integrating $\nu_x$ should increase uncertainty as
 time without a measurement passes.
+
 \begin{equation}
@@
 
-When a measurement becomes available it is incorporated into the probability,
-which should eventually cause the probability to converge. 
-Since the weights are the probability of of $x(t)$ being exactly $x_i(t)$, 
-the weights are defined as
+When a measurement becomes available it is incorporated into the probability.
+This process of incorporating new data is called sequential importance sampling,
+and eventually causes the probability to converge. The weight is defined
+as
+
+\begin{equation}
+w_i(T) \propto \frac{Pr(x_i[0:T] | y[0:T])}{q(x_i)}
+\end{equation}
+
+where $q$ is called an \emph{importance density}, which means it decides where
+the support points for $\hat{x}(T)$ are located. To remove the bias due to the
+location of the support points, we divide by $q(\hat{x}_i)$. By dividing by 
+the density of the support points (particles), the location of the points are removed
+from the probability density. As a result the weight is dependent solely based
+on $Pr(x_i[0:T] | y[0:T]$, the probability of the $i^{th}$ particle's measurements
+being different from $y[0:T]$ due to noise alone.
+An example of an importance density would be drawing a large
+number of points from the standard normal, $N(0,1)$ and then weighting each point, $l$ by 
+$1/\beta(l), \beta \sim N(0,1)$. Of course if there is a far off peak in
+the posterior that $q$ does not allocate support points in, there will essentially
+be a quantization error, and that part of the density can't be modeled.
+
+For this work we used $q(x_i) = Pr(x_i(T) | x_i[0:T-1]) = Pr(x_i(T) | x_i(T-1))$,
+based on the Markov assumption, and the belief that the state space model is 
+at least close to correct.  
+%<stopped here>
+This has the added benefit that it is not necessary to
+re-draw points from a $q$ function at every time step, which is faster. 
+this the definition of the weight, so when weights are updated it is necessary
+to divide by the proposal distributionto converge
+to the true distribution it is necessary to divide by the proposal distribution.
+
 \begin{equation}

### 2010-03-24 doc/thesis/weightedhist.py

@@ -1,0 +1,39 @@
+import pylab
+import numpy
+from scipy.stats import norm
+import random
+
+def rolldice(num):
+    sum = 0
+    for i in range(0, num):
+        sum = sum + random.randrange(1, 7)
+    return sum
+
+def two_die(total):
+    if total < 2 or total > 12:
+        return 0
+
+    density = range(1, 7)
+    density.extend(range(5,0,-1))
+    return density[total-2]
+
+def test():
+    bin = [x + 1.5 for x in range(0, 12)]
+    pointsv = [rolldice(2) for i in range(0, 50000)]
+    pylab.hist(pointsv, bin);
+    pylab.show()
+
+    weightv = [1./two_die(point) for point in pointsv]
+    pylab.hist(pointsv, bin, weights=weightv)
+    pylab.show()
+
+def test2():
+    pointsv = pylab.randn(50000);
+    pylab.hist(pointsv);
+    pylab.show()
+
+    weightv = [1./norm.pdf(point) for point in pointsv]
+    pylab.hist(pointsv, 30, weights=weightv)
+    pylab.show()
+
+test2()

### 2010-03-24 doc/thesis/thesis.tex
need to justify factoring importance density

@@
 \usepackage{graphicx}
+\usepackage{cancel}
 
@@
 Considering that $y$ contains measurement noise and noise in $x$ can drive
-changes in $y$, it is clear that $Pr(x(t) | u[0:T]$ is not a single true value
+changes in $y$, it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
 but a true posterior. 
@@
 will be left off for simplicity.
-Once the probability, $Pr(\hat{x}(T) | \hat{x}[0:T-1], y[0:T-1]$ has been found
+Once the probability, $Pr(\hat{x}(T) | \hat{x}[0:T-1], y[0:T-1])$ has been found
 (initially this is just Mixture approximating the prior since no measurements are 
@@
 \begin{equation}
+\label{weightfunc}
 w_i(T) \propto \frac{Pr(x_i[0:T] | y[0:T])}{q(x_i)}
@@
 
-where $q$ is called an \emph{importance density}, which means it decides where
+where $q$ is called an \emph{importance density}, meaning it decides where
 the support points for $\hat{x}(T)$ are located. To remove the bias due to the
 location of the support points, we divide by $q(\hat{x}_i)$. By dividing by 
-the density of the support points (particles), the location of the points are removed
-from the probability density. As a result the weight is dependent solely based
-on $Pr(x_i[0:T] | y[0:T]$, the probability of the $i^{th}$ particle's measurements
+the density of the support points (particles), the effect of the particle 
+distribution may be removed from the posterior density. As a result the weight
+is dependent solely based
+on $Pr(x_i[0:T] | y[0:T])$, the probability of the $i^{th}$ particle's measurements
 being different from $y[0:T]$ due to noise alone.
@@
 based on the Markov assumption, and the belief that the state space model is 
-at least close to correct.  
+able to approximate the true state. The weight is also based on $Pr(x_i[0:K] | y[0:K])$,
+which may be broken up as follows.
+\begin{eqnarray}
+Pr(x[0:T] | y[0:T]) & = & \frac{Pr(y[0:T], x[0:T])}{Pr(y[0:T])} \nonumber \\
+ & = & \frac{Pr(y(T), x[0:T] | y[0:T-1]) \cancel{Pr(y[0:T-1])}}{Pr(y(T) | y[0:T-1]) \cancel{Pr(y[0:T-1])}} \nonumber \\
+ & = & \frac{Pr(y(T)| x[0:T], y[0:T-1]) Pr(x[0:T] | y[0:T-1])}{Pr(y(T) | y[0:T-1]) } \nonumber \\
+ & = & \frac{Pr(y(T)| x[0:T], y[0:T-1]) Pr(x(T) | x[0:T-1], y[0:T-1]) Pr(x[0:T-1] | y[0:T-1])}{Pr(y(T) | y[0:T-1])} \nonumber \\
+\end{eqnarray}
+
+Using the assumption that $y(t)$ is fully constrained by $x(t)$,
+and that $x(t)$ is fully constrained by $x(t-1)$, we are able to
+make the reasonably good assumptions that:
+\begin{equation}
+Pr(y(T) | x[0:T], y[0:T-1]) = Pr(y(T) | x(T))
+\end{equation}
+
+\begin{equation}
+Pr(x(T) | x[0:T], y[0:T-1]) = Pr(x(T) | x(T-1))
+\end{equation}
+
+Additionally, for the particle filter $y(T)$ and $y[0:T-1]$ are 
+given, and therefore constant across all particles. Thus $Pr(x[0:T] | y[0:T]$
+may be simplified to:
+\begin{eqnarray}
+Pr(x[0:T] | y[0:T]) & = & \frac{Pr(y(T)| x[0:T], y[0:T-1]) Pr(x(T) | x[0:T-1], y[0:T-1]) 
+            Pr(x[0:T-1] | y[0:T-1])}{Pr(y(T) | y[0:T-1])} \nonumber \\
+& = & \frac{Pr(y(T)| x(T)) Pr(x(T) | x(T-1)) Pr(x[0:T-1] | y[0:T-1])}{Pr(y(T) | y[0:T-1])} \nonumber \\
+& \propto & Pr(y(T)| x(T)) Pr(x(T) | x(T-1)) Pr(x[0:T-1] | y[0:T-1])
+\end{eqnarray}
+
+Plugging this into \ref{weightfunc} leads to:
+\begin{eqnarray}
+w_i(T) & \propto & \frac{Pr(y(T)| x(T)) Pr(x(T) | x(T-1)) Pr(x[0:T-1] | y[0:T-1])}{q(x(T) | y[0:T])q(x[0:T-1] | y[0:Tjk:wq
+} \nonumber \\
+& \propto & Pr(y(T)| x(T)) Pr(x[0:T-1] | y[0:T-1]) \\
+& \propto & \frac{Pr(y(T)| x(T)) \cancel{Pr(x(T) | x(T-1))} Pr(x[0:T-1] | y[0:T-1])}{\cancel{Pr(x(T) | x(T-1))}} \nonumber \\
+& \propto & Pr(y(T)| x(T)) Pr(x[0:T-1] | y[0:T-1]) \\
+\end{eqnarray}
+
+and of course, since ....wait?
 %<stopped here>

### 2010-03-29 doc/thesis/thesis.tex
need to finish the resampling part of particle filters....

--- 2010-03-24 doc/thesis/thesis.tex
@@
 \usepackage{amssymb}
+\usepackage{algorithmic}
 \usepackage{setspace}
 \usepackage{graphicx}
+\usepackage{cancel}
 
@@
 \begin{document}
-\section*{Introduction}
+\section{Introduction}
 Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
@@
 
-\subsection*{FMRI}
+\subsection{FMRI}
 Before going further, it is necessary to discuss Functional MRI,
@@
 
-\subsection*{BOLD Physiology}
+\subsection{BOLD Physiology}
 The physiology that leads to the BOLD model has been well studied over the 
@@
 
-\section*{Current Techniques}
-\subsection*{Basic Statistical Parametric Mapping}
+\section{Current Techniques}
+\subsection{Basic Statistical Parametric Mapping}
 The most basic method of analyzing FMRI data is through a basic T-test
@@
 
-\subsection*{General Linear Model}
+\subsection{General Linear Model}
 The most used form of FMRI analysis is still based on Statistical Parametric
@@
 
-\subsection*{Whats wrong with these techniques}
+\subsection{Whats wrong with these techniques}
 There are a few problems with the techniques mentioned in the 
@@
 
-\section*{Proposed Approach}
-\subsection*{Goal}
+\section{Proposed Approach}
+\subsection{Goal}
 The ultimate goal of this project is to provide a new set of tools
@@
 
-\subsection*{Introduction to Particle Filters}
+\subsection{Introduction to Particle Filters}
 Particle filters, a type of Sequential Monte Carlo (SMC) methods
@@
 %unscented kalman filter
+\subsubsection{Model}
 The idea of the particle
@@
 \begin{equation}
+\label{stateass}
 \dot{x}(t) = f(x(t), u(t)) + \nu_x
@@
 \begin{equation}
+\label{measass}
 y(t) = g(x(t)) + \nu_y
@@
 
+\subsubsection{Prior}
+
 The goal of the particle filter is to evolve a probably distribution 
@@
 \begin{equation}
-\{[\hat{x}_i(0),w_i] : \hat{x}_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
+\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
 \end{equation}
@@
 using a Mixture PDF. 
-
 \begin{equation}
-Pr(\hat{x}(0) = x(0)) = \sum_{i=1}^{N_p} w_i\delta(x(0) - \hat{x}_i(0) ) dx
+\hat{Pr}(x(0) = \hat{x}) = \sum_{i=1}^{N_p} w_i\delta(\hat{x} - x_i(0) ) dx
 \end{equation}
-
-Where $\delta(x-x_0)$ is 1 when $x = x_0$ (the Kronecker delta function).
+Where $\delta(x-x_0)$ is 1 if and only if $x = x_0$ (the Kronecker delta function).
 
 If a true prior is preferred, then the weights
-should all be $1/N_p$, and since $\hat{x}_i$ was drawn from the prior, this will
+should all be $1/N_p$, and since $x_i$ was drawn from the prior, this will
 be an approximation of the prior distribution. If a relatively flat prior is 
-preferred, then each particle's weight could be divided by the density, $\alpha(\hat{x}_i)$,
+preferred, then each particle's weight could be divided by the density, $\alpha(x_i)$,
 which creates a flat prior with support points in the region of $\alpha(x)$. Either
@@
 For the BOLD signal all the parameters have been studied and have relatively well
-known mean and variance, so a prior bias is actually preferred. In our
-case we did not initially flatten the distributions, but instead used 
-distributions that fit what was known about the parameters.  
-For strictly positive parameters (members of $\hat{x}$) we used a gamma distribution,
-whereas for parameters that could be negative, we used a Gaussian distribution. Both
-distributions utilized standard deviations twice that found in previous studies.
+known mean and variance, so a prior could be very helpful. We ran simulations for
+both normalized and un-normalized priors, although we believe in cases such as this,
+where a good prior exists, it should be used. 
+For strictly positive parameters (members of $x$) we used a gamma distribution,
+whereas for parameters that could be negative, we used a Gaussian distribution. In
+both cases standard deviations twice that found in previous studies were used.
 
-Note that all the probability's implicitly depend on $u[0:T]$, so those terms 
+Note that all the probabilities implicitly depend on $u[0:T]$, so those terms 
 will be left off for simplicity.
-Once the probability, $Pr(\hat{x}(T) | \hat{x}[0:T-1], y[0:T-1]$ has been found
+Once the probability, $\hat{Pr}(x(T) | x[0:T-1], y[0:T-1])$ has been found
 (initially this is just Mixture approximating the prior since no measurements are 
 available and no previous probabilities are available), its possible to approximate
-the probability for short times between before a measurement is available, by shifting
+the probability for short times between times when measurement is available, by shifting
 the probability according the progression of the state equations. This is only 
@@
 \begin{equation}
-Pr(\hat{x}(T+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i\delta\left(x - (\hat{x}_i(T) + \int_T^{T+\Delta} \dot{\hat{x}}_i(t) dt) \right) dx
+\hat{Pr}(x(T+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
 \end{equation}
 
+\subsubsection{Weighting}
 When a measurement becomes available it is incorporated into the probability.
@@
 \begin{equation}
-w_i(T) \propto \frac{Pr(x_i[0:T] | y[0:T])}{q(x_i)}
+\label{weightfunc}
+w_i(T) \propto \frac{\hat{Pr}(x_i[0:T] | y[0:T])}{q(x_i[0:T] | y[0:T])}
 \end{equation}
 
-where $q$ is called an \emph{importance density}, which means it decides where
-the support points for $\hat{x}(T)$ are located. To remove the bias due to the
-location of the support points, we divide by $q(\hat{x}_i)$. By dividing by 
-the density of the support points (particles), the location of the points are removed
-from the probability density. As a result the weight is dependent solely based
-on $Pr(x_i[0:T] | y[0:T]$, the probability of the $i^{th}$ particle's measurements
+where $q$ is called an \emph{importance density}, meaning it decides where
+the support points for $x(T)$ are located. To remove the bias due to the
+location of the support points, we divide by $q(x_i[0:T] | y[0:T])$. By dividing by 
+the posterior density of the support points (particles), the effect of the particle 
+distribution may be removed from the posterior density. As a result the weight
+is dependent solely based
+on $\hat{Pr}(x_i[0:T] | y[0:T])$, the probability of the $i^{th}$ particle's measurements
 being different from $y[0:T]$ due to noise alone.
@@
 $1/\beta(l), \beta \sim N(0,1)$. Of course if there is a far off peak in
-the posterior that $q$ does not allocate support points in, there will essentially
-be a quantization error, and that part of the density can't be modeled.
+the posterior that $q$ does not allocate support points in, there will 
+be a quantization error, and that part of the density can't be modeled. This is why
+it is absolutely necessary that $q$ covers $\hat{Pr}(x_i[0:T] | y[0:T])$.
 
-For this work we used $q(x_i) = Pr(x_i(T) | x_i[0:T-1]) = Pr(x_i(T) | x_i(T-1))$,
+$q(x_i[0:T] | y[0:T])$ may be simplified by assuming that $y(T)$ doesn't contain 
+any information about $x(T-1)$, which is more practical since knowledge of future
+measurements is impractical.
+
+\begin{eqnarray}
+q(x[0:T] | y[0:T]) & = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T]) \nonumber \\
+& = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T-1]) \nonumber \\
+& = & q(x(T) | x(T-1), y[0:T])q(x[0:T-1] | y[0:T-1])
+\end{eqnarray}
+
+In this paper we will use 
+$q(x_i(T) | x_i(T-1), y[0:T]) =  \hat{Pr}(x_i(T) | x_i(T-1))$,
 based on the Markov assumption, and the belief that the state space model is 
-at least close to correct.  
+able to approximate the true state. This means that prior to re-weighting 
+particles, the particles will be distributed the same as the previous time but
+moved forward according to the integration of $f(x(t), u(t))$.
+
+In addition to $q(x_i(T) | x_i[0:T-1], y[0:T])$, the weight is also based on $Pr(x_i[0:K] | y[0:K])$,
+which may be broken up as follows.
+\begin{eqnarray}
+\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y[0:T], x[0:T])}{\hat{Pr}(y[0:T])} \nonumber \\
+ & = & \frac{\hat{Pr}(y(T), x[0:T] | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}}{\hat{Pr}(y(T) | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}} \nonumber \\
+ & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x[0:T] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1]) } \nonumber \\
+ & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
+\end{eqnarray}
+
+Using the assumption that $y(t)$ is fully constrained by $x(t)$ \eqref{measass},
+and that $x(t)$ is fully constrained by $x(t-1)$ \eqref{stateass}, we are able to
+make the reasonably good assumptions that:
+\begin{equation}
+\hat{Pr}(y(T) | x[0:T], y[0:T-1]) = \hat{Pr}(y(T) | x(T))
+\end{equation}
+
+\begin{equation}
+\hat{Pr}(x(T) | x[0:T], y[0:T-1]) = \hat{Pr}(x(T) | x(T-1))
+\end{equation}
+
+Additionally, for the particle filter $y(T)$ and $y[0:T-1]$ are 
+given, and therefore constant across all particles. Thus $\hat{Pr}(x[0:T] | y[0:T])$
+may be simplified to:
+\begin{eqnarray}
+\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) 
+            \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
+& = & \frac{\hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
+& \propto & \hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])
+\end{eqnarray}
+
+Plugging these simplifications into \eqref{weightfunc} leads to:
+\begin{eqnarray}
+\label{weightevolve}
+w_i(T) & \propto & \frac{\hat{Pr}(y(T)| x(T)) \cancel{\hat{Pr}(x(T) | x(T-1))} \hat{Pr}(x[0:T-1] | y[0:T-1])}
+                         {\cancel{\hat{Pr}(x_i(T) | x_i(T-1))}q(x[0:T-1] | y[0:T-1])} \nonumber \\
+& \propto & w_i(T-1)\hat{Pr}(y(T)| x(T)) 
+\end{eqnarray}
+
+Thus, by making the following relatively weak assumptions, evolving a posterior
+density  is easy and requires almost no knowledge of noise distribution.
+\begin{enumerate}
+\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ provide 
+a sufficiently flexible model to encapsulate the true time series.
+\item $E[\nu_x] = 0$ and $E[\nu_y] = 0$, and $\nu_x$ and $\nu_y$ are stationary
+\item The PDF $q(x_i(0))$ (the prior) fully covers $Pr(x_i(0))$
+\item Markov Assumption: $Pr(x(T) | x[0:T]) = Pr(x(T) | x(T-1))$
+\item $q(x[0:T-1] | y[0:T]) = q(x[0:T-1] | y[0:T-1])$
+\end{enumerate}
+
+\subsubsection{Basic Particle Filter Algorithm}
+From the definition of $w_i$, the algorithm sequential importance sampling (SIS) 
+is relatively simple.
+
+\begin{algorithmic}
+\STATE Initialize $N_p$ Particles: 
+        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
+\STATE $T$ = \{Set of Measurement Times\}
+\FOR{$t$ in $T$}
+    \FOR{$i$ in $N_p$}
+        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
+    \ENDFOR
+\ENDFOR
+
+\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+\end{algorithmic}
+The result is then a discrete approximation of the posterior distribution.
+
+\subsubsection{Resampling}
+As a consequence 
+of the wide prior distribution (required for a proper discretization of a continuous
+distribution), there will be many particles with insignificant weights. While this does help
+describe the tails of the distribution very well, it means that only a small portion of the
+computation will be spent describing the most probable region. Ideally every particle would 
+equally decrease the entropy of the distribution, thus the lower the variance of the weights,
+the more efficiently the discrete distribution is in describing the continuous distribution. 
+A common measure of "Particle Degeneracy" is the effective number of particles, described
+in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
+Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
+of each particle. Of course the true weight is unknown, so a heuristic approximating 
+$N_{eff}$ is used:
+\begin{equation}
+N_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
+\end{equation}
+Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
+$N_{eff}$ drops precipitously.  To alleviate this problem
+a common technique known as resampling must be applied. The idea of re-sampling is to 
+draw from the approximate posterior, thus generating a replica of the posterior with 
+a support more suited to the posterior. Thus, if weights are all set to $1/N_p$, and 
+$N_p$ points are drawn from the posterior,
+\begin{equation}
+\hat{x}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
+\end{equation}
+then $\hat{x} \sim x$. Unfortunately this isn't necessarily the truth: since the support is
+still limited to the original particles, the number of unique particles may only go down.
+This effect, often dubbed "particle impoverishment" can result in excessive quantization
+errors in the final distribution.
+
+%<more here>
+
+The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
+or a genetic algorithm. "Fit" versions of $x$ spawn more children nearby which allow
+for more accurate estimation near points of high likelihood. As the variance of the estimated
+$x$'s decrease the radius in which children are spawned also decreases. Eventually the radius
+will approach the width of the $\nu_x$ and $\nu_y$.
+
 %<stopped here>
@@
 to the true distribution it is necessary to divide by the proposal distribution.
-
-\begin{equation}
-  d
-\end{equation}
 
@@
 
-\subsection*{My Specific Particle Filte}
+\subsection{My Specific Particle Filte}
 Although particle filters are well defined, significant design decisions
@@
 
-\section*{Methods}
-\subsection*{Preprocessing}
+\section{Methods}
+\subsection{Preprocessing}
 After FMRI data has been acquired it is always necessary to modify the
@@
 
-\section*{Introduction to Modeling}
+\section{Introduction to Modeling}
 
@@
 
-\section*{Results}
+\section{Results}
 The particle filter shows great promise in being able 

### 2010-03-30 doc/thesis/thesis.tex
need to explain the weighting function.

@@
 \begin{equation}
+\label{neff}
 N_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
@@
 This effect, often dubbed "particle impoverishment" can result in excessive quantization
-errors in the final distribution.
+errors in the final distribution. However, by instead sampling from a smoothed version of
+the discrete distribution, sample impoverishment is prevented. The question then is how to 
+decide on the smoothing kernel. 
+Often times the easiest way to sample from the continous distribution is to break the 
+re-sampling down into two steps. First a member of the discrete distribution is randomly
+selected based on the weights, and then based on the smoothing a nearby state variable 
+is selected. The process of the selection will be defined as:
+\begin{equation}
+x'_i = x_i + h\sigma \epsilon
+\end{equation}
+Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
+and $\epsilon$ is drawn from the chosen kernel.
+It has been proven that when all the elements of the mixture
+have the same weight, as is the case after basic re-sampling, the kernel that minimizes the 
+MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
+Particle Filters, C Musso, N Oudjane and F LeGrand). 
+\begin{equation}
+K = \left\{
+\begin{array}{lr}
+\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
+0 & otherwise
+\end{array}\right.
+\end{equation}
+%<more here>
 
-%<more here>
+If the noise is assumed to be Gaussian then it is possible to further optimize. Obviously,
+its very possible that distribution will deviate from that assumption, however in this
+case assuming a Gaussian still allows a great deal of flexibility, since the distribution
+is a mixture. Thus we let $h$ be defines as:
+\begin{eqnarray}
+h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}
+\end{eqnarray}
+and although it is very possible the underlying noise is non-gaussian, the Gaussian
+may still be a good choice. Its worth noting that as $n_x$, the number of dimensions in
+$x$ go to infinity, the standard deviation based approximation becomes inefective according
+ to (cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
+ tracking, sanjeev arulampalam, simon maskell, neil gordon...). Because our regularized 
+ resampling method is based on particle variance, resampling very often will slow down 
+ convergeance, because the resampling itself can lead to increased variance. Thus, for
+ the most part the resulting algorith is the basic particle filter, except for when the
+ $N_{eff}$ drops very low (say less than 50), at which point regularized resampling is 
+ performed. The resulting algorthm is as follows:
+
+ \begin{algorithmic}
+\STATE Initialize $N_p$ Particles: 
+        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
+\STATE $T$ = \{Set of Measurement Times\}
+\FOR{$t$ in $T$}
+    \FOR{$i$ in $N_p$}
+        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
+    \ENDFOR
+
+    \STATE Calculate $N_{eff}$ with \eqref{neff}
+    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
+        \STATE Calculate empirical $\sigma$ 
+        \STATE Redraw particles using basic resampling
+        \FOR{$i$ in $N_p$}
+            \STATE Draw $\epsilon \sim K$
+            \STATE $x_i = x_i + h \sigma \epsilon$
+        \ENDFOR
+    \ENDIF
+\ENDFOR
+
+\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+
+ \end{algorithmic}
 
@@
 
+\subsubsection{Choosing $\hat{Pr}(y(T) | x(T))$}
+Analysis of Gaussian versus exponential
+
+Discussion of optimality (or lack thereof)
+
+Discussion of using a gaussian mixture pdf to describe a non-gaussian distribution
+
 %<stopped here>
-This has the added benefit that it is not necessary to
-re-draw points from a $q$ function at every time step, which is faster. 
-this the definition of the weight, so when weights are updated it is necessary
-to divide by the proposal distributionto converge
-to the true distribution it is necessary to divide by the proposal distribution.
-
-Thus $\hat{x}(0)$ is an empirical representation of the prior, and $\hat{x}_i(0)$
-is the $i^{th}$ particle, a single possible representation of the system at time 0.
-Initially, since each $\hat{x}_i$ has been drawn from the prior, and no measurements
-have been made, the particles are all equally weighted. When a measurement becomes
-available, that information is then incorporated into the posterior by re-weighting
-the particles according to how well they estimated this new measurement.
-
-The estimated posterior of $\hat{x}(t)$, for discrete time, $t$ (before incorporating
-any measurement that might be available at $t$) is then:
-\begin{equation}
-\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t), \hat{x}(t-1) | y[0:t-1]\} d\hat{x}(t-1)
-\end{equation}
-
-and by bayes theorem $\Pr\{\hat{x}(t), \hat{x}(t-1) | y[0:t-1]\} = 
-\Pr\{\hat{x}(t) | \hat{x}(t-1), y[0:t-1]\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}$,
-therefore 
-\begin{equation}
-\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t) | \hat{x}(t-1) ,y[0:t-1]\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}  d\hat{x}(t-1)
-\end{equation}
-
-Since in this work we will only deal with markov processes, we assume that 
-$\Pr\{\hat{x}(t) | \hat{x}(t-1), y[0:t-1]\} = \Pr\{\hat{x}(t) | \hat{x}(t-1)\}$
-
-thus the posterior becomes:
-\begin{equation}
-\Pr\{\hat{x}(t) | y[0:t-1]\} =  \int \Pr\{\hat{x}(t) | \hat{x}(t-1)\}\Pr\{\hat{x}(t-1) | y[0:t-1]\}  d\hat{x}(t-1)
-\end{equation}
-
-Therefore in between measurements, there is no need to know previous measurements in
-order to make predictions about the internal state. This is important because it means makes
-particle filters significantly more memory efficient, and although it isn't useful in this
-case, it means a particle filter could theoretically be run indefinitely without the
-need to store old data.
-
-As stated previously, new measurements may be incorporated by reweighting particles.
-To find the updated distribution, $\Pr\{\hat{x}(t) | y[0:t]\}$, bayes theorem is
-again applied:
-\begin{equation}
-\Pr\{\hat{x}(t) | y[0:t]\} = \Pr\{\hat{x}(t) | y[0:t-1], y(t)\} = \frac{\Pr\{y(t) | \hat{x}(t), y[0:t-1]\}\Pr\{\hat{x}(t) | y[0:t-1]\}}
-{\Pr\{y(t) | y[0:t-1]\}}
-\end{equation}
-
-\begin{equation}
-\Pr\{\hat{x}_i(t) | y[0:t-1]\} = \int \Pr\{\hat{x}_i(t) | \hat{x}_i(t-1), y[0:t-1]\}\Pr[\hat{x}_i(t-1)
-\end{equation}
-Since $\hat{x}(t) = \hat{x}(t-1) + \int_{t-1}^t f(x(t), u(t))dt + \int_{t-1}^t\nu_d dt$, 
-\begin{equation}
-\Pr\{\hat{x}_i(t) | \hat{x}_i(t-1)\} = \Pr\{\int{t-1}^t \nu_d\}
-\end{equation}
-
-of course $\int_{t-1}^t f(x(t), u(t))dt$ may need to be evaluated numerically. Therefore,
-to estimate the posterior probability at $t$, it is just necessary to propagate
-each particle forward in time according to a state space function. To form 
-the posterior probability around the true $x(t)$, each particle is weighted
-based on
-
-This
-then results in the expected value of $\hat{y}(t)$:
-\begin{equation}
-E[\hat{y}_i(t) | \hat{x}_i(t-1)] = E[\hat{x}_i(t)]
-\end{equation}
-
-\begin{equation}
-\dot{y} = -ky
-\end{equation}
-$\theta_i$ would have an estimate for $k$ \emph{and} $y$. Of course in many cases
-the model is a potentially non-linear differential equation with no closed
-form solution. Initially every $\theta_i$ will have the same weight, because
-there isn't yet anything to base a weighting on. Note that this doesn't mean
-that the particles have been evenly distributed, so the distribution is not 
-necessarily "flat" just because all the particles are equaly weighted. When a
-new measurement is available, the particles are then reweighted based on how
-well they predicted the new measurement.
-
 \subsection{My Specific Particle Filte}

### 2010-03-31 doc/thesis/thesis.tex
finished with particle filter part

@@
 \begin{equation}
-\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_d)
+\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x)
 \end{equation}
@@
 \begin{equation}
-y(t) = g(t, x(t), u(t), \theta, \nu_s)
+y(t) = g(t, x(t), u(t), \theta, \nu_y)
 \end{equation}
@@
 %decreased or completely removed, since it 
-Third, we will assume noise is additive.
+Third, we will assume noise is additive, and that $\nu_x$ may be projected into
+a weiner, or other summing process that is additive with $g$ and $\nu_y$, which
+will be named $\nu_d$.
 Finally, $x(t)$ will encapsulate $\theta$, the unknown model constants, which
@@
 \label{stateass}
-\dot{x}(t) = f(x(t), u(t)) + \nu_x
+\dot{x}(t) = f(x(t), u(t))
 \end{equation}
@@
 \label{measass}
-y(t) = g(x(t)) + \nu_y
+y(t) = g(x(t)) + \nu_y + \nu_d
 \end{equation}
 
-$\nu_x$ will result in something akin to adding a Weiner process to $y[0:T]$, and thus
+Because $\nu_d$ is something akin to and additive Weiner process $y[0:T]$, it 
 will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
 in $y[0:T]$. For some of the tests, I will use de-trending methods to reduce the effects of 
-$\nu_x$, the remainder of which should be re-factored into $\nu_y$. Both $\nu_x$ and $\nu_y$
+$\nu_d$, the remainder of which will be re-factored into $\nu_y$. Both $\nu_d$ and $\nu_y$
 have biological and non-biological sources. MR can lead to both types of noise, 
 as demonstrated in \cite{drift}. Meanwhile changes in metabolism, heart rate, or
-other biochemical intervention could all lead to either $\nu_x$ or $\nu_y$.
+other biochemical intervention could all lead to either $\nu_d$ or $\nu_y$.
 
@@
 the probability according the progression of the state equations. This is only 
-an approximate, since integrating $\nu_x$ should increase uncertainty as
+an approximate, since integrating $\nu_d$ should increase uncertainty as
 time without a measurement passes.
@@
 a sufficiently flexible model to encapsulate the true time series.
-\item $E[\nu_x] = 0$ and $E[\nu_y] = 0$, and $\nu_x$ and $\nu_y$ are stationary
+\item $E[\nu_d] = 0$ and $E[\nu_y] = 0$, and $\nu_x = d\nu_d$, $\nu_y$ are stationary
 \item The PDF $q(x_i(0))$ (the prior) fully covers $Pr(x_i(0))$
@@
 \label{neff}
-N_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
+\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
 \end{equation}
@@
 draw from the approximate posterior, thus generating a replica of the posterior with 
-a support more suited to the posterior. Thus, if weights are all set to $1/N_p$, and 
+a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
 $N_p$ points are drawn from the posterior,
 \begin{equation}
-\hat{x}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
+\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
 \end{equation}
-then $\hat{x} \sim x$. Unfortunately this isn't necessarily the truth: since the support is
-still limited to the original particles, the number of unique particles may only go down.
+then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
+still limited to the original particles, the number of unique particles can only go down.
 This effect, often dubbed "particle impoverishment" can result in excessive quantization
-errors in the final distribution. However, by instead sampling from a smoothed version of
-the discrete distribution, sample impoverishment is prevented. The question then is how to 
-decide on the smoothing kernel. 
+errors in the final distribution. However, there is a solution. Instead of sampling from the
+discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
+that distribution. Because the distribution is continuous, there is no way for a collapse
+of the particles to occur. The question then, is how to decide on the smoothing kernel. 
 Often times the easiest way to sample from the continous distribution is to break the 
@@
 \begin{equation}
-x'_i = x_i + h\sigma \epsilon
+\chi_i = x_i + h\sigma \epsilon
 \end{equation}
@@
 It has been proven that when all the elements of the mixture
-have the same weight, as is the case after basic re-sampling, the kernel that minimizes the 
+have the same weight, as is the case after basic resampling, the kernel that minimizes the 
 MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
@@
 and although it is very possible the underlying noise is non-gaussian, the Gaussian
-may still be a good choice. Its worth noting that as $n_x$, the number of dimensions in
-$x$ go to infinity, the standard deviation based approximation becomes inefective according
- to (cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
- tracking, sanjeev arulampalam, simon maskell, neil gordon...). Because our regularized 
- resampling method is based on particle variance, resampling very often will slow down 
- convergeance, because the resampling itself can lead to increased variance. Thus, for
- the most part the resulting algorith is the basic particle filter, except for when the
- $N_{eff}$ drops very low (say less than 50), at which point regularized resampling is 
- performed. The resulting algorthm is as follows:
+may work, but sub-optimally. It has been proposed that (Monte Carlo Approximations for
+General State-Space Models, markus Hurzeler and Hans R. Kunsch) if the underlying 
+distribution is non-Gaussian, then using this bandwidth will oversmooth. 
+At the same time, as $n_x$, the number of dimensions in
+$x$, goes to infinity, the standard deviation based approximation becomes less effective
+(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
+tracking, sanjeev arulampalam, simon maskell, neil gordon...). In reality over smoothing
+should not be too great an issue because the smoothing is only being applied to find new
+particles. If the distribution is over smoothed then the algorithm may not converge as rapidly;
+however, because the bandwidth is still based on particle variance, which will decay as 
+particles are ruled out, it is still able to converge. In fact over smoothing is preferrable
+to under smoothing, since the latter would result in false negatives, but the previous does
+not result in false anything. Additionally, because of the high dimensionality of our system,
+and limited measurements, it is helpful to have a broader bandwidth to explore with. Because 
+of the potentially wide smoothing factor applied by regularized resampling, performing this
+step at every measurement would allow particles a great deal of mobility. This mobility is
+the opposite of convergence, which is why regularized resampling should only be done when
+$\hat{N}_{eff}$ drops very low (say less than 50). Other than the periodic regularized
+resampling then, the regularized particle filter is nearly identical to the basic sampling
+importance sampling filter (SIS).
 
@@
         \STATE Calculate empirical $\sigma$ 
-        \STATE Redraw particles using basic resampling
+        \STATE $h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}$
+        \STATE Redraw particles using (stratified) basic resampling
         \FOR{$i$ in $N_p$}
@@
 The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. "Fit" versions of $x$ spawn more children nearby which allow
+or a genetic algorithm. Versions of $x$ that are "fit", or give good measurements, spawn more children 
+nearby which allow
 for more accurate estimation near points of high likelihood. As the variance of the estimated
@@
 
-\subsubsection{Choosing $\hat{Pr}(y(T) | x(T))$}
-Analysis of Gaussian versus exponential
+\subsection{Choosing $\hat{Pr}(y(T) | x(T))$}
+Choosing a representation of an unknown distribution is certainly tricky,
+and so the fact that $\hat{Pr}(y(T) | x(T)) = \nu_d + \nu_y$ means that
+there is a significant piece of the algorithm that is based on conjecture.
+Studies of the noise in FMRI typically attribute noise to a Gaussian random
+variable or an additive noise process with Gaussian steps.
 
-Discussion of optimality (or lack thereof)
+\subsubsection{Classical De-trending}
+The non-stationary
+aspect of a Weiner process is difficult to compensate for, and so various methods
+have been developed to compensate for it. \cite{detrend} and \cite{drift} have
+demonstrated that this component is prevalent, and may in fact just be a characteristic
+of FMRI. In some studies, as many as half the voxels benefit from detrending, meaning
+that this is certainly a serious barrier to inference. All the existing methods are performed
+during the preprocessing stage, rather than as an integral part of analyzing the BOLD
+signal. There is no shortage of theories on the "best" method of detrending, however
+a head to head comparison, \cite{detrend}, showed that in most cases subtracting off
+a spline works the best. The benefit of the spline versus wavelets, high pass 
+filtering or other DC removal techniques is that the frequency response is not set.
+A spline is able to move quickly when the signal is moving quickly, and move more
+slowly when the signal moves more slowly. That said, the spline will still remove some
+amount of signal, just like all of these methods.
 
-Discussion of using a gaussian mixture pdf to describe a non-gaussian distribution
+<image of de-spline'd lines with "true" lines>
 
-%<stopped here>
-\subsection{My Specific Particle Filte}
-Although particle filters are well defined, significant design decisions
-are application dependent. The most important choice is the weighting function,
-$w(y_t - y(\theta_t))$. Obviously the function should weight $0$ maximally,
-the drop off rate is extremely important. A very thin gaussian probability 
+\subsubsection{Delta Based Inference}
+I also propose and test a different method of dealing with the so called "drift". 
+Instead of comparing the direct output of the particle filter with the direct
+measurement, the algorith compares the change in signal over that period of time.
+In most signal processing cases this would foolish, but that is because the 
+general assumption that all noise is high frequency noise does not hold up here.
+In fact, every pipelines for the analysis of BOLD signal uses a high pass filter,
+but low poss filters are rarely applied, because it is a well known fact that 
+most of the signal is in the high frequency range. The particle filter is an 
+extremely robust method of inference, and so I would assert that the particle
+filter ought to be given as raw data as possible. While taking direct measurements
+without de-trending would give awful results, using the difference removes the 
+DC component and turns a Weiner process into a Gaussian random variable.
+
+\begin{equation}
+\label{measass}
+\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
+\end{equation}
+
+Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply Gaussian. If
+$\nu_d$ is some other additive process, the difference will still be one of a few stable
+distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
+with a maximum variance of twice the original variance. All the assumptions made originally
+for the particle filter hold, and all of the parameters may be distringuished based on
+the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
+rather than direct signal level.
+
+\subsubsection{Weighting Function}
+Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting funciton,
+is based on an unknown distribution, it is necessary to decide on a function
+that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
+needs to centered at zero and have a scale comparable to the signal levels.
+While a Gaussian function is the natural choice, we also wanted to try a distribution
+with wider tails, so that outliers don't completely destroy particle's weights.
+Therefore, we tried three weighting functions; Gaussian, exponential and 
+the Cauchy distribution. The standard deviations (or scale) was set to 
+$\sigma_y/5$, where $\sigma_y$ is the variance of all $y[0:\infty]$. Of course 
+since the particle filter requires a weighting function to run, this means that
+before the particle filter starts all the measurements have to be in. In cases where
+this is impossible, a heuristic based on a small sample may work just as well.
+
+The shape of the weighting function is extremely important, because it essentially decides
+the rejection rate of particles. A very thin gaussian probability 
 distribution function has nice properties, but thin tails. As a result, large
@@
 have near 0 weights, thus forcing the particle filter to converge improperly. 
-On the other end of the spectrum, a cauchy PDF may not
-weight particles in the middle enough, preventing the particles to never 
-converge. Of course, the scale factor, or variance, also plays a signifcant
-role in the convergeance rate. The importance of the weighting function
+On the other end of the spectrum, a cauchy PDF, has relatively fat tails, and
+may not weight central particles high enough, preventing the particles from 
+converging at a reasonable rate. Exponetials have the benefit of having an 
+extremely smooth drop to zero, and a slope of 1 at the origin. Having a non-zero
+slope at the origin is beneficial because it discriminates all the way up until
+the measured and predicted $y$ are the same. The importance of the weighting function
 cannot be overstated, as this is the primary factor in deciding the rate at
-which the particle filter converges. As part of the simulations, I compare
-a gaussian based weighting function with an exponential weightin function.
-In both cases, I scaled the width of the weighting functions automatically based
-on the signal variance.
+which the particle filter converges.
 
-Dealing with low frequency noise is also a crucial part of the tests. 
-I tested two different methods of dealing with this noise. Various methods
-are used to deal with this problem in other toolsets. The most common method
-is to apply a high-pass filter to the measurement vector. Although this makes
-sense in many applications, in this case it may not be adaptive enough for
-a random walk type noise. On the other hand, it has been shown that a spline
-de-trending method is able to better deal with the low frequency noise
-typically present in FMRI. Thus, my first method
-was to use a spline fit with knots every 80 or 90 measurements to trance
-the basic shape of the drift. I then added an extra DC gain parameter to
-the model to account for the fact that the de-trended signal tended to have 
-a mean of 0, rather than some positive value.
-
-My second method of dealing with low frequency drift was to counteract
-it by weighting particles based on $(y_t-y_{t-1}) - (y(\theta_t)-y(\theta_{t_1}))$ 
-rather than directly on $y_t - y(\theta_t)$. It is generally believed that the noise present in 
-FMRI has approximately gaussian steps, which means that this method only
-has to deal with gaussian white noise. Such a technique is usually not considered a
-good idea because high frequency noise is almost exclusively assumed in
-systems theory. In this case however; the noise is all in the lower frequency
-range, so performing a differentiation like operation, which magnifies high
-frequency signals, is exactly what we want.
-
-... more specifics...
-
-The General Linear Model, and 
-SPM's goal is to find a correlation between input, or what a 
-person is thinking, and which regions are active. This is an 
-important task, but it misses the potential of FMRI. FMRI's 
-greatest weakeness as a tool for studying neural activity is that
-it is connected to neurons through a long chain of events. But
-what if, in addition to seeing which brain regions are active
-for which stimuli, we could also study pathologies such as poor
-blood flow, or derive physical parameters such as blood vessel
-compliance? By using a realistic physical model as the basis for
-discovering the relation between the outside world and FMRI 
-signal, the path to detecting activation gives all those other
-things for free, precisely \emph{because} FMRI is the result of such
-a long chain of events. It is actually quite common to use models to derive
-the values of physiological parameters. For instance, blood volume
-is often measured by injecting a dye into the blood stream and then
-assuming that the dye is removed at a rate proportional to the
-concentration in the blood. By performing a simple regression it is
-possible to estimate blood volume quite accurately. While for
-this particular project I don't claim that the results will give 
-physiologically plausible results, with some tuning of the algorithm
-it would very possible.
-
-While the General Linear Model has historically done well with
-detecting activation, it is limited to scaling the entire input
-signal up or down. This leaves out even basic effects such as 
-longer or shorter time constants, as well as nonlinear effects which
-are known to be present in the BOLD system. 
-**
-
-While the General Linear Model has been extremely effective at
-gleaning information about which voxels are "active" and which
-are not, its purpose is only to say which regions are "statistically
-significant". However, considering that voxels consist of millions
-of neurons, the reality is that a
-region could be active, but at some other level or time constant
-than expected; and be completely missed by the linear model. 
-Studies showing activation
-in Cadavers of Phantoms highlight the inherant problem with the 
-General Linear Model: a problem that mandates multiple
-comparison testing with very high thresholds to ensure any 
-sort of face validity \cite{drift}.
-By moving to models that actually "learn" the 
-parameters of the model, model error will be drastically
-reduced allowing for much more reasonable statistical thresholds.
-\subsection{Particle Filters}
+%... more specifics...
+%
+%The General Linear Model, and 
+%SPM's goal is to find a correlation between input, or what a 
+%person is thinking, and which regions are active. This is an 
+%important task, but it misses the potential of FMRI. FMRI's 
+%greatest weakeness as a tool for studying neural activity is that
+%it is connected to neurons through a long chain of events. But
+%what if, in addition to seeing which brain regions are active
+%for which stimuli, we could also study pathologies such as poor
+%blood flow, or derive physical parameters such as blood vessel
+%compliance? By using a realistic physical model as the basis for
+%discovering the relation between the outside world and FMRI 
+%signal, the path to detecting activation gives all those other
+%things for free, precisely \emph{because} FMRI is the result of such
+%a long chain of events. It is actually quite common to use models to derive
+%the values of physiological parameters. For instance, blood volume
+%is often measured by injecting a dye into the blood stream and then
+%assuming that the dye is removed at a rate proportional to the
+%concentration in the blood. By performing a simple regression it is
+%possible to estimate blood volume quite accurately. While for
+%this particular project I don't claim that the results will give 
+%physiologically plausible results, with some tuning of the algorithm
+%it would very possible. 
+%
+%While the General Linear Model has historically done well with
+%detecting activation, it is limited to scaling the entire input
+%signal up or down. This leaves out even basic effects such as 
+%longer or shorter time constants, as well as nonlinear effects which
+%are known to be present in the BOLD system. 
+%**
+%
+%
+%While the General Linear Model has been extremely effective at
+%gleaning information about which voxels are "active" and which
+%are not, its purpose is only to say which regions are "statistically
+%significant". However, considering that voxels consist of millions
+%of neurons, the reality is that a
+%region could be active, but at some other level or time constant
+%than expected; and be completely missed by the linear model. 
+%Studies showing activation
+%in Cadavers of Phantoms highlight the inherant problem with the 
+%General Linear Model: a problem that mandates multiple
+%comparison testing with very high thresholds to ensure any 
+%sort of face validity \cite{drift}.
+%By moving to models that actually "learn" the 
+%parameters of the model, model error will be drastically
+%reduced allowing for much more reasonable statistical thresholds.
+%\subsection{Particle Filters}
 

### 2010-04-05 doc/thesis/thesis.tex
more work

@@
 but FMRI doesn't detect this happening in one neuron, but rather as the 
-aggregate over hundreds or millions of cells. Though local neurons act
+aggregate over millions of cells. Though local neurons act
 "together" (i.e. around the same time), the density of neurons, the
@@
 
-If the noise is assumed to be Gaussian then it is possible to further optimize. Obviously,
-its very possible that distribution will deviate from that assumption, however in this
-case assuming a Gaussian still allows a great deal of flexibility, since the distribution
-is a mixture. Thus we let $h$ be defines as:
+If the noise is assumed to be Gaussian then it is possible to further optimize. 
+Thus we let $h$ be defines as:
 \begin{eqnarray}
@@
 distribution is non-Gaussian, then using this bandwidth will oversmooth. 
-At the same time, as $n_x$, the number of dimensions in
-$x$, goes to infinity, the standard deviation based approximation becomes less effective
-(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
-tracking, sanjeev arulampalam, simon maskell, neil gordon...). In reality over smoothing
+In reality over smoothing
 should not be too great an issue because the smoothing is only being applied to find new
@@
 particles are ruled out, it is still able to converge. In fact over smoothing is preferrable
-to under smoothing, since the latter would result in false negatives, but the previous does
-not result in false anything. Additionally, because of the high dimensionality of our system,
-and limited measurements, it is helpful to have a broader bandwidth to explore with. Because 
+to under smoothing, since the latter would result in false negatives, but the previous only
+results in a slower decay of the variance. 
+At the same time, as $n_x$, the number of dimensions in
+$x$, goes to infinity, the standard deviation based approximation becomes less effective
+(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
+tracking, sanjeev arulampalam, simon maskell, neil gordon...).  Because of the high dimensionality of our system,
+and limited measurements, it is helpful to have a broader bandwidth to explore the distribution. 
+Nevertheless, because 
 of the potentially wide smoothing factor applied by regularized resampling, performing this
 step at every measurement would allow particles a great deal of mobility. This mobility is
-the opposite of convergence, which is why regularized resampling should only be done when
+the enemy of convergence, which is why regularized resampling should only be done when
 $\hat{N}_{eff}$ drops very low (say less than 50). Other than the periodic regularized
@@
 The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. Versions of $x$ that are "fit", or give good measurements, spawn more children 
-nearby which allow
-for more accurate estimation near points of high likelihood. As the variance of the estimated
-$x$'s decrease the radius in which children are spawned also decreases. Eventually the radius
-will approach the width of the $\nu_x$ and $\nu_y$.
+or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
+nearby which allow for more accurate estimation near points of high likelihood. 
+As the variance of the estimated
+$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
+will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.
 
@@
 and so the fact that $\hat{Pr}(y(T) | x(T)) = \nu_d + \nu_y$ means that
-there is a significant piece of the algorithm that is based on conjecture.
+there is a significant piece of the algorithm that is based primarily conjecture.
 Studies of the noise in FMRI typically attribute noise to a Gaussian random
@@
 The non-stationary
-aspect of a Weiner process is difficult to compensate for, and so various methods
+aspect of a Weiner process as with $\nu_d$ is difficult to compensate for, and so various methods
 have been developed to compensate for it. \cite{detrend} and \cite{drift} have
-demonstrated that this component is prevalent, and may in fact just be a characteristic
+demonstrated that this component is prevalent, and may in fact be a characteristic
 of FMRI. In some studies, as many as half the voxels benefit from detrending, meaning
@@
 Instead of comparing the direct output of the particle filter with the direct
-measurement, the algorith compares the change in signal over that period of time.
+measurement, the algorith compares the change in signal over a single TR. 
 In most signal processing cases this would foolish, but that is because the 
-general assumption that all noise is high frequency noise does not hold up here.
-In fact, every pipelines for the analysis of BOLD signal uses a high pass filter,
+general assumption that all noise is high frequency is not the case here.
+In fact, every pipeline for the analysis of BOLD signal uses a high pass filter,
 but low poss filters are rarely applied, because it is a well known fact that 
-most of the signal is in the high frequency range. The particle filter is an 
+most of the signal is in the high frequency range and most of the noise is actually 
+in the low frequency range. The particle filter is an 
 extremely robust method of inference, and so I would assert that the particle
-filter ought to be given as raw data as possible. While taking direct measurements
+filter ought to be given as \emph{raw} data as possible. While taking direct measurements
 without de-trending would give awful results, using the difference removes the 
@@
 
-Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply Gaussian. If
+Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
 $\nu_d$ is some other additive process, the difference will still be one of a few stable
@@
 the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
-rather than direct signal level.
+rather than string of direct readings.
+
+<frequncy response graphs, highlighting noise frequency range and signal frequency range>
 
@@
 
-%... more specifics...
-%
-%The General Linear Model, and 
-%SPM's goal is to find a correlation between input, or what a 
-%person is thinking, and which regions are active. This is an 
-%important task, but it misses the potential of FMRI. FMRI's 
-%greatest weakeness as a tool for studying neural activity is that
-%it is connected to neurons through a long chain of events. But
-%what if, in addition to seeing which brain regions are active
-%for which stimuli, we could also study pathologies such as poor
-%blood flow, or derive physical parameters such as blood vessel
-%compliance? By using a realistic physical model as the basis for
-%discovering the relation between the outside world and FMRI 
-%signal, the path to detecting activation gives all those other
-%things for free, precisely \emph{because} FMRI is the result of such
-%a long chain of events. It is actually quite common to use models to derive
-%the values of physiological parameters. For instance, blood volume
-%is often measured by injecting a dye into the blood stream and then
-%assuming that the dye is removed at a rate proportional to the
-%concentration in the blood. By performing a simple regression it is
-%possible to estimate blood volume quite accurately. While for
-%this particular project I don't claim that the results will give 
-%physiologically plausible results, with some tuning of the algorithm
-%it would very possible. 
-%
-%While the General Linear Model has historically done well with
-%detecting activation, it is limited to scaling the entire input
-%signal up or down. This leaves out even basic effects such as 
-%longer or shorter time constants, as well as nonlinear effects which
-%are known to be present in the BOLD system. 
-%**
-%
-%
-%While the General Linear Model has been extremely effective at
-%gleaning information about which voxels are "active" and which
-%are not, its purpose is only to say which regions are "statistically
-%significant". However, considering that voxels consist of millions
-%of neurons, the reality is that a
-%region could be active, but at some other level or time constant
-%than expected; and be completely missed by the linear model. 
-%Studies showing activation
-%in Cadavers of Phantoms highlight the inherant problem with the 
-%General Linear Model: a problem that mandates multiple
-%comparison testing with very high thresholds to ensure any 
-%sort of face validity \cite{drift}.
-%By moving to models that actually "learn" the 
-%parameters of the model, model error will be drastically
-%reduced allowing for much more reasonable statistical thresholds.
-%\subsection{Particle Filters}
+Obviously the optimal $\hat{Pr}(y(T) | x(T))$ is the true $Pr(y(T) | x(T))$; however
+since that is unknown, I tested multiple different distributions. As stated previous,
+because the exact scale of $Pr(y(T) | x(T))$ it not even known, the exponential distribution
+has the added benefit of having a negative slope all the way form zeo to infinity. Thus
+even if the scale is completely wrong, particles will still be well differentiated.
+
+<Q-Q plot of real fmri data with Gaussian, DC>
+<Q-Q plot of real fmri data with Gaussian, deltas>
 
 \section{Methods}
+This paper describes two types of experiments; first we will cover
+simulations which have the benefit of a ground truth, then we will
+cover the methods used in the use of the particle filter on real data.
+
 \subsection{Preprocessing}
+As discussed in the section on de-trending, the normal pipeline for analyzing
+FMRI involves a great deal of preprocessing. In this paper we make an effort to
+minimize any type of preprocessing that will degrade the signal. 
 After FMRI data has been acquired it is always necessary to modify the
 data in some way to make different runs comparable. Because FMRI signal
-levels are essentially unit-less, at the very least it is necessary to convert
-the data into \% difference from the baseline. Finding the baseline is the
-first hurdle in dealing with FMRI. Various methods exist for doing this,
-but they are all ad-hoc. It is common to treat each stimulus pulse independently,
-ignoring the long decay of the BOLD signal. By treating each input separately,
-the "base" for calculating the percent difference is merely the signal level 
-before the stimulus is applied. The problem with this is that the BOLD signal
-can be heavily delayed, and fall times can be as long as 10 or 15 seconds,
-meaning that the assumption is clearly a false one. Another popular method 
-of dealing with this problem, is to put all the time series through a high
-pass filter. This will of course remove the DC component of the signal, and 
-some amount of so called "drift". The problem with this method is that it is
+levels are unit-less, at the very least it is necessary to convert
+the data into \% difference from the baseline. This process removes no data
+from signal since it merely subtracting then dividing by a constant. This
+is the signal that was input into the delta based particle filter.
+Of course there are much more advanced ways of performing this task.
+The generally accepted standard is actually to use a high pass filter, although the
+cutoff frequency is application dependent and often applied haphazardly.
+The high pass filter thus removes the DC component of the signal, and 
+some amount of the so called "drift". The problem with this method is that it is
 not adaptive to the input. Huge variations in drift frequencies can exist 
@@
 drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
-set too high.
+set too high. This is why, as I mentioned in the De-trending section, a spline
+based detrending method will generally give better results.
 
-\section{Introduction to Modeling}
+For simulated and real images (tests with multiple time-series), tests were 
+also run with and without Gaussian filtering with sigma of %not sure todo
+were run, since it is standard
+practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
+a spatial filter such as Gaussian filtering increased SNR but can also lead to less
+precision in the output maps.
 
-Of course, it is well known that many of the assumptions 
-that the GLM makes about BOLD activation do not hold. 
-The General Linear Model ignores
-the fact that the BOLD response is nonlinear, contains non-gaussian
-noise and most importantly that brain networks are 
-so called "Small World Networks" \cite{smallworld}, \cite{noise}. 
-Merely accounting for
-nonlinearities in the time-series of the BOLD response can
-result in significantly better estimation as shown in \cite{nonlinearmodels}.
-But even more crucial is the recent movements away from mass-
-univariate models. Current mass-univariate models 
-assume uniform connectivity from input to every voxel; yet
-this is obviously not the case. While
-violating linearity assumptions or Gaussian assumptions 
-requires us to be more conservative, the lack of any method
-to switch brain circuits on and off means that any model
-we make for brain regions would not be Turing complete;
-in effect implying that humans are incapable of 
-thought \cite{turing}.
+\subsection{Simulation}
+We performed two types of simulations. First, we simulated a single BOLD time-series based
+on a random chosen set of model parameters. This process was relatively straight forward
+given the state-space equations for the BOLD signal. After a "true" signal was generated,
+we then added a carrier level, since BOLD is typically measured as a \% difference from the
+base level. Finally, we added Gaussian noise, and a Weiner Process to the clean signal. The
+variance of the Gaussian noise may be expressed in terms of the desired noise SNR, $R$ as:
+\begin{equation}
+var(y_{noisy}) = var(y) / R
+\end{equation}
+Since SNR doesn't have quite the same meaning for a Wiener process based noise, the variance 
+of the Gaussian steps was set to be:
+\begin{equation}
+var(y_{noisy}) = var(y) / (4R)
+\end{equation}
+Once this noisy simulated time series was generated, the exact same particle filter algorithm
+that would later be run on full sized images, was run on this single voxel image. We ran
+a series of tests to determine the convergence rate of the particle filter, the number
+of particles that were required, how weighting functions compared, how different de-trending
+methods compared with each other and, finally the variance of the result. By running the exact
+time-series with different noise realizations, it was possible to determine the model variance.
+As the reader may know, the error of an estimator may be calculated as:
+\begin{equation}
+MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
+\end{equation}
+The variance is an expression of how much the result would change for different noise realizations,
+whereas the bias is an expression of how well the model matches the true underlying model. In
+this case, because the same model is being used in the particle filter and underlying simulation,
+the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
+underlying state space equation, there will be some amount of bias error, but assuming that the
+noise is similar to the noise used in these tests, the model variance will actually be about the
+same. Thus calculating the model variance is extremely helpful in calculating how well determined
+our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
+thousands present in a real image, makes it easier to 
+compare the output with the ground truth, with various parameters.
 
-So called Dynamic Causal Modeling is the beginning of 
-the next phase of neurology. DCM is the first 
-brain study to show any significant connection between 
-diffusion tensor imaging and actual function layout in
-the human brain \cite{dti_dcm}. By incorporating connections
-between regions and a realistic activation model, DCM
-corrects two of the largest problems with the GLM. While
-there are other techniques that may be useful in the future,
-they either lack physiological analogs (Artificial Neural
-Networks) or are extremely computationally expensive 
-(mutual information).  Given that much of the 
-potential of the GLM has been exhausted, and that DCM 
-is one of the first well defined methods capable of learning 
-complicated brain circuits DCM is crucial to the future
-of FMRI and our understanding of the human brain.
+Second we used a modified version of the FSL tool 
+POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
+by creating a random image, smoothing it with a large Gaussian kernel, then thresholding
+the results. Finally connected regions were each given a set of parameters from a finite
+list of randomly chosen parameter sets. The result was a four dimensional (length x width
+x height x parameter) image with spatially varying parameters. Time-series of activation level
+was generated for each set of parameters, then activation levels were fed into POSSUM's 
+function for generating frequency domain data. The patche for POSSUM will be made available.
+For each time-series in the simulate FMRI image, the final \emph{static} parameters are saved
+into a parameter map. This parameter map may then be compared to the map used to generate the 
+simulated data; additionally a new simulation using the calculated parameters may also be 
+generated to test the functional difference between the two maps. This would give an absolute 
+quantitative difference between the two parameter sets irrespective to parameter slopiness.
+So for instance, if $V_0$ is halved, $\epsilon$ doubling may very well give a similar result.
+In this case the \% difference between the parameters will be large in each case, but the functional
+difference between the parameters will not be great. This is obviously a bad situation, which
+is why we wanted to test for it.
+
+\subsection{Real Data}
+Finally, we also performed inference based on real FMRI data. The scanner we used...
+%... more specifics...
+
+The final result from calculating parameters with the real data was similar to that
+from the results from the POSSUM simulated data. The difference being that there was
+no ground truth the check it with.
 
 \section{Results}
-The particle filter shows great promise in being able 
-to learn a variety of different regional activation
-parameters. We have performed tests with a random binary
-pulse train as input to a simulated region and then tested
-the particle filter's ability to estimate the signal
-parameters. Additionally random gaussian noise with an
-Signal-To-Noise ratio of 2.0 was added. Figure 1 shows
-the pulse train, the clean signal and the noisy signal, which
-was the input to the particle filter. A random set of 
-system parameters was chosen to simulate the region,
-which the particle filter had no knowledge of other than
-a mean and rough variance.
-The tests were performed using 10,000 particles and took
-approximately 4 minutes to compute.
+\subsection{Single Time-Series Simulation}
 
-%figure 1
-\begin{figure}
-\includegraphics[width=8in,height=4in]{noise.png}
-\caption{Simulated time-series for a single region. The red line
-shows the input stimulus, the blue line shows the base BOLD 
-response, and the green line is the BOLD response with Gaussian
-white noise added at SNR of 2}
-\label{fig:noise}
-\end{figure}
+Graphs:
 
-%figure 2
-\begin{figure}
-\includegraphics[width=8in,height=4in]{bold.png}
-\caption{Particle Filter Results. The blue line is the "true"
-(simulated) BOLD response, the red line
-is the output of the particle filter with 2 standard deviations
-in green. Note that the error bar for time 0 is outside the scope
-of the image and is approximately $\pm .003$.}
-\label{fig:bold}
-\end{figure}
+For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
+\{100, 500, 1000\} particles
+\begin{enumerate}
+\item Comparison with a linear system with similar number of degrees of freedom
+\item Ground truth vs. Estimated signal during particle filter run
+\item Ground truth vs. Estimated signal with final parameter set
+\item True Parameters vs. Final Parameter Sets
+\item Variance of final parameters when faced with same ground truth, different noise
+\item Variance of final parameters when faced with same ground truth, different noise
+\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
+\item Estimator Variance based on different noise runs
+\end{enumerate}
 
-%figure 3
-\begin{figure}
-\includegraphics[width=8in,height=4in]{state.png}
-\caption{Particle Filter Results. The blue line is the "true"
-parameter and the green line is the estimated value for the parameter.
-Note that the timescale for $S_t$ is different to highlight an typical
-stimulus response. Again 2 standard deviations are shown.}
-\label{fig:state}
-\end{figure}
+%note to self, epsilon should probably be uniform between 0 and something
+\subsection{Simulated Volume}
+\begin{enumerate}
+\item Parameter Map 
+\item Error map of parameters
+\item Histogram of \%errors between parameters
+\item Activation Map based on a single region with high $\epsilon$, compared with linear
+\end{enumerate}
 
-Figure 2 shows the estimated BOLD response versus the simulated
-data (without noise). It is clear that as the estimated BOLD
-response converges to the true BOLD as time progresses. This
-is because more and more potential states are eliminated by system's
-response to new input sequences. Convergence depends highly on 
-the input signal and the weighting function thus choice of both
-are extremely important. We have found that an impulse train
-(emulated with .1 second width square waves) gives a very good
-response. It is also important to choose an input with
-some longer hold times, at least as long as than the 
-expected time constants, which in this case was around 
-8 seconds. The weighting function we have chosen is based
-on the exponential distribution, with a variance equal to
-the RMS of the signal. While it is important to converge
-by the end of the time-series, it is often the case that 
-converging too fast will harm the state estimation by over-
-emphasizing the measurements of single time points. The rate
-of convergence is well within control of the user of the 
-particle filter based on the weighting function and the frequency
-of re-sampling. Note that resampling can cause quantization
-errors due to the nature of estimating a long tailed 
-distribution with a finite number of samples.
+\subsection{FMRI Data}
+....
 
-Figure 3 shows the estimated versus simulated values for
-several key parameters. Notice how the variance drops as
-the particle filter continues. The parameters shown are
-$V_0$ which is a scaling parameter, $\tau_s$, and $\tau_f$
-which are both time constants and $s_t$ which is a hidden
-state variable, estimating the flow inducing signal. Note that
-even though some of the parameters don't converge to the exact
-value, that the estimated $s_t$ still matches the true
-$s_t$ relatively well. It is often the cast that although a
-few parameters don't converge to their true value, one 
-parameter may pick up the slack for another. It is of
-note that $\tau_s$ and $\tau_f$ do converge at least close
-to their true values, because they cannot be determined
-from steady state.
+image comparing epsilon-map with GLM activation map
+
+\section{Conclusion}
 
 \bibliographystyle{abbrvnat}
-\bibliography{proposal_micah}
+\bibliography{references}
 

### 2010-04-06 doc/thesis/thesis.tex
quick addition to thesis. fix to svn version tagging.
latex runs correctly now.

@@
 the work of recharging is done, there may be a prolonged undershoot lasting
-as much as 90 seconds \cite{spatial_poststim}, though the reason for this
+as much as 90 seconds \cite{spatial_undershoot}, though the reason for this
 is debated \cite{origin_poststim}.
@@
 and the flow inducing signal is not precisely known, model fits are actually
-quite good \cite{ModelCompareison}. Regardless, being based on some real 
+quite good \cite{nonlinearmodels}. Regardless, being based on some real 
 physiological 
@@
 \item Estimator Variance based on different noise runs
+\item Final Particle Distribution
 \end{enumerate}

### 2010-04-20 doc/thesis/thesis.tex
some work on the lit review portion

@@
 of neural activity. Despite its rather limited temporal resolution, FMRI is still
-the best way of measuring neural activity deep in the brain. 
+the best way of measuring neural activity for the majority of the brain.
 Whereas other methods
 of analyzing neural signals can be invasive or difficult to acquire, 
-FMRI is relatively quick and cheap, and its analysis is straight forward.
+FMRI is relatively quick and cheap, and its analysis straight forward.
 Because of these benefits, FMRI continues to be crucial to the study of human 
-cognition. Despite the fact that FMRI is so widely used,
-the choice of analysis tools is still relatively limited. Every widely
-available analysis tool is based on parametric modeling, which requires
-prior knowledge of the noise distribution. Indeed, the linear methods
-usually applied are known not to be robust to non-gaussian, non-white
-noise, which is why so many pre-processing steps are necessary.
-These limitations cast a long shadow over any experiment that claims to be
-able to reject the null hypothesis, no matter the P-value. While obviously
-there is a correlation between the BOLD signal and actual activity, its
-possible the current parametric tests are over-estimating the actual amount
-of activity. Its also possible that the current methods underestimate
-actual activity. In this paper we propose a robust model-based approach to 
-activation detection that makes no assumptions about the underlying noise
-model.
-
-fact, typical activation detection does little to account for variations
-in the shape of the signal, but instead is a direct T-test between the measured signal
-and the expected shape of the signal. Because of these limitations,
-multiple comparison tests are imperative (ref fish), and active regions
-must be very large to be detected with any confidence. This paper
-will introduce a more robust method of analyzing FMRI signals for 
-activation, by means of estimating the BOLD parameters.
+cognition. Despite its prevalence, there have been relatively few developments
+in the actual analysis of FMRI images. A steady stream of studies have adapted
+the BOLD model from the initial Baloon model first proposed by \cite{Buxton1998}
+all the way to full fully autonomous system of equations \cite{Riera2004}. And while
+there have been numerous forks in the model, enough in fact to make an entire paper
+studying the differences,\cite{Deneux2006}, it is widely known that all these
+models have quantitatively less bias error than General Linear Model which is
+typically employed today. Then again, depending on the model there may between
+seven \cite{Riera2004} and fifteen \cite{Zheng2002} parameters per voxel to
+be optimized. Clearly there is a significant risk of error due to variance
+with so many degrees of freedom, not to mention a significantly increased
+computation cost. In this thesis I demonstrate the use of a 
+particle filter as a means of addressing these problems.
 
 \subsection{FMRI}
-Before going further, it is necessary to discuss Functional MRI,
-and what it measures. All MRI works by exciting nuclear spin away
-from a steady state controlled by a large electromagnet. Different 
-nuclei respond to RF impulses based on their natural frequency
-of oscillation. Therefore contrast in MRI is based on the aggregate
-nuclear composition of a given region. Thus, regions with more oxygen
-will respond differently from regions with proportionately less 
-oxygen. This exact difference is what causes the changes over time
-in the MR signal in functional MRI. Unfortunately, it takes a significant
-amount of time for nuclear spin to return to equilibrium. Thus instead
-of waiting, a type of imaging called Echo Planar Imaging (EPI) is used,
-wherein a single excitation is used, and then as quickly as possible
-every voxel (3 Dimensional Pixel) in a particular region is read. 
-Because "reading" a particular region requires the switching on of a
-gradient magnetic field, it takes a decent amount of time to perform
-measurements. Problematically, the signal from the original excitation
-drops exponentially, thus the signal to noise ratio can be very low
-even at modest resolutions. Eventually the signal becomes unreadable,
-and a new excitation must be applied. As more measurements are made, and thus
-as the spatial resolution increases, the rate at which entire volumes
-can be acquired drops. Typically volumes can be acquired every two to
-three seconds, at resolutions of about 64 by 64 by 32, but their signal
-to noise ratio is often poor. Other imaging modalities such as EEG and
-MEG are capable of much faster measurements, but are unable to isolate
-measurement locations so precisely as FMRI. As with all biological systems,
-noise is much higher than in human designed systems, and yet these 
-biological systems are capable of performing amazing feats.
+Magnetic Resonance Imaging, MRI is a method of building 3D images
+non-invasively, based on the difference between nuclear spin
+relaxation times in various molecules. Initially the entity being
+imaged is brought into a large magnetic field which aligns the spins
+of molecules in the same direction; radio frequency (RF) signals may
+then be used to excite nuclear spin away from the steady alignment. 
+As the nuclei precess back to their original orientation, they resonate
+at the same RF frequency of their original excitation. Conveniently, the
+excitation of nuclear spins return their original state at different
+rates, called the T1 relaxation time, depending on the properties of 
+the material excited. Additionally, the
+coherence of the spins also decay differently (and quite a bit faster
+than T1) based on the properties of the region that has been excited.
+This gives two primary methods of contrasting substances,
+which is the basis of T1 and T2 weighted images. Additionally, there
+dephasing occurs at two different rates, the T2 relaxation time,
+which is impossible to recover from, and T2$^*$ relaxation, which is
+much faster, but possible to recover from with an inversion pulse.
+Oftentimes T1 relaxation times can be on the order of seconds if 
+a significant excitation pulse is applied. 
+In order to rapidly acquire entire brain images, as is done in Functional 
+MRI, a single large excitation pulse is applied to the entire brain,
+and the entire volume is acquired in a single T1 relaxation period. 
+Because the entire k-space (spatial-frequency) volume is acquired 
+from a single excitation, the signal to noise ration is very low
+in this type  of imaging (Echo Planar Imaging).
+
+Increasing the spatial resolution of EPI imaging necessarily 
+requires more time or faster magnetic field switching. Increasing
+magnet switching rates though is difficult, because it can result in
+more artifacts, or even lower signal to noise ratios. The result is
+that at best FMRI is capable of 1.2 second temporal resolution. 
+Additionally, the means that each voxel of the image will contain 
+the sum of a large amount neurons, capillaries and veins. Thus, the
+FMRI signal, which is sensitive to the chemical composition of 
+materials, is summing up the composition of various types of tissue
+in addition to the blood, whose composition is what we actually care about.
+In particular, the presence of Deoxyhemoglobin, Hemoglobin whose
+oxygen has been used by a metabolic process, has a larger magnetic
+response compared to Oxygenated Hemoglobin. Thus, capillaries near
+very active cells will typically have a higher Deoxyhemoglobin content,
+and regions with lower activity will have a lower Deoxyhamoglobin content.
+Unfortunately, as mentioned previously, blood is only a small part of
+each voxel, which means that a single EPI image doesn't tell much about
+Deoxyhemoglobin content. However assuming blood is the only thing changing
+in the short term, percent difference from a baseline signal \emph{will}
+tell us something. FMRI analysis is thus necessarily performed on the percent change
+from the baseline. Luckily the assumption that tissue content does not change in
+the short run is actually pretty good, although other facts can pollute
+the baseline signal, as we will discuss later.
 
@@
 \end{document}
+
+%trash
+%stream of new developments in modeling the BOLD signal, nothing has have been
+%a plethora of new studies pushing advancements 
+%the choice of analysis tools is still relatively limited. Every widely
+%available analysis tool is based on parametric modeling, which requires
+%prior knowledge of the noise distribution. Indeed, the linear methods
+%usually applied are known not to be robust to non-gaussian, non-white
+%noise, which is why so many pre-processing steps are necessary.
+%These limitations cast a long shadow over any experiment that claims to be
+%able to reject the null hypothesis, no matter the P-value. While obviously
+%there is a correlation between the BOLD signal and actual activity, its
+%possible the current parametric tests are over-estimating the actual amount
+%of activity. Its also possible that the current methods underestimate
+%actual activity. In this paper we propose a robust model-based approach to 
+%activation detection that makes no assumptions about the underlying noise
+%model.

### 2010-04-21 doc/thesis/thesis.tex
more work on the review

@@
 more artifacts, or even lower signal to noise ratios. The result is
-that at best FMRI is capable of 1.2 second temporal resolution. 
+that at \emph{best} FMRI is capable of 1 second temporal resolution. 
 Additionally, the means that each voxel of the image will contain 
@@
 In particular, the presence of Deoxyhemoglobin, Hemoglobin whose
-oxygen has been used by a metabolic process, has a larger magnetic
+oxygen has been used by a metabolic process, has a decreased magnetic
 response compared to Oxygenated Hemoglobin. Thus, capillaries near
-very active cells will typically have a higher Deoxyhemoglobin content,
-and regions with lower activity will have a lower Deoxyhamoglobin content.
+very active cells will typically have a higher Deoxyhemoglobin content and
+lower signal, and regions with lower activity will have a lower 
+Deoxyhamoglobin content and thus higher signal.
 Unfortunately, as mentioned previously, blood is only a small part of
@@
 from the baseline. Luckily the assumption that tissue content does not change in
-the short run is actually pretty good, although other facts can pollute
+the short run is actually pretty good, although other factors can pollute
 the baseline signal, as we will discuss later.
@@
 \subsection{BOLD Physiology}
-The physiology that leads to the BOLD model has been well studied over the 
-past decade and a half. While some uncertainty remains (for instance in
+It is will known that Deoxyhemoglobin acts as a contrast agent in EPI imaging
+\cite{Buxton1999}, \cite{WEISSKOFF1994}, \cite{Ogawa}, so then what causes
+changes Deoxyhemoglobin in a voxel? Obviously increased metabolism will increase
+Deoxyhemoglobin, but changes in blood volume in a voxel will also alter 
+the MR signal depending on the ratio of Deoxyhemoglobin to hemoglobin. 
+This was the impetus
+for the ground breaking research \cite{Buxton1999} and \cite{Mandeville1999}.
+These were the first two attempts to quantitatively account for the shape of the 
+BOLD signal as a consequence of the lag between the cerebral blood flow (CBF) 
+and the cerebral blood volume (CBV). In fact \cite{Buxton1999} went to far as
+to show that a simple, well chosen blood flow waveform, along with a balloon 
+model, could fully account for the BOLD signal. Although complex
+changes in metabolism exist (\cite{Zheng2002}), most recent studies have
+focused on their ability to explain the prolonged BOLD post stimulus 
+undershoot \cite{Donahue2009}, \cite{Yacoub2006}. This is because \cite{Buxton2004}
+and later \cite{Riera2004} showed that the primary signal may be accurately
+be estimated by a simple flow-locked expression of the CMRO2. Comparison
+between the more complicated models including a separate expression for CMRO2
+and the simplified Buxton expression show little actual difference \cite{Deneux2006}.
+In \cite{Buxton1999}, it was shown that small changes in the flow expression could
+result in vastly different recovery times of the baseline BOLD signal, and
+in \cite{Mandeville1999} a cause for delayed CBV recovery was proposed, based
+on differences in compliance of capillaries. \cite{Behzadi2005} provided a simple,
+complete model and explanation for delayed compliance of neural capillaries,
+and those simple changes can result in a significant improvement in the bias error,
+as shown in \cite{Deneux2006}. In the end, it may be impossible to refute the
+possibility of a prolonged period of increased metabolic activity after activation,
+however, much of the BOLD post stimulus undershoot may be explained by a 
+prolonged CBV undershoot, as shown in \cite{Chan2009}. Because of the greatly
+increased complexity of CMRO2 modeling, compared with the relative simplicity
+of the Behzadi model, in most cases it is not worth using the more advanced CMRO2
+model unless the metabolic rate is an end goal.
+
+The physiology that leads to the BOLD signal has been well studied over the 
+past decade and a half, but there is still some question about the origins
+of certain signal features. In the seminal paper \cite{Buxton1999}, the original
+Balloon model was proposed. Buxton et. al showed that all of the features
+of the BOLD signal may be explained by lag between ... . While some uncertainty remains (for instance in
 relation to the BOLD post-stimulus undershoot <papers relating to back and

### 2010-04-23 doc/thesis/thesis.tex
more on the lit review

@@
 cognition. Despite its prevalence, there have been relatively few developments
-in the actual analysis of FMRI images. A steady stream of studies have adapted
-the BOLD model from the initial Baloon model first proposed by \cite{Buxton1998}
+in the actual analysis of FMRI images. A steady stream of studies have built
+on the original BOLD signal derivation first described in \cite{Ogawa}, 
+from the Baloon model first proposed by \cite{Buxton1998}
 all the way to full fully autonomous system of equations \cite{Riera2004}. And while
@@
 models have quantitatively less bias error than General Linear Model which is
-typically employed today. Then again, depending on the model there may between
+typically employed today. Then again, depending on the model there may be between
 seven \cite{Riera2004} and fifteen \cite{Zheng2002} parameters per voxel to
@@
 particle filter as a means of addressing these problems.
+
+FMRI images as a method of detecting neural activation is based on 
+temporal changes of the Blood Oxygen Level Dependent (BOLD) signal.
+The BOLD signal is caused by minute changes in the ratio of Deoxygenated
+Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain.
+Because Deoxygenated hemoglobin is paramagnetic, higher concentrations
+attenuate the signal when using T2 weighted imaging techniques, such 
+as Echo Planar Imaging (EPI) which is used in FMRI. When axons becomes active,
+a large amount of ions quickly flow out of the cell. In order for the action
+potential to be used again, an active pumping process moves ions back into the
+axon. This process of recharging the axon takes a large amout of enengy, which 
+naturally uses oxygen. On a massive scale (cubic millimiter) this activation/recharge
+process is happening all
+the time; however, it happens at a much higher rate when a portion of
+the brain is very active. Thus, blood vessels in a very active area will 
+tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin,
+resulting in lower FMRI signal. However, to compensate for activation, muscles that
+control blood vessels relax in that region to allow more blow flow, which 
+in fact results in a higher concentration of oxygenated hemoglobin. Thus,
+increased activation actually tends to \emph{increase} the MR signal in
+comparison with the base level. It is this overcompensation that is the 
+primary signal detected with FMRI imaging. This cascade of events
+can, as a consequence of increased activity, increase the local metabolism, 
+blood flow, blood volume, and oxygenated hemoglobin; though not necessarily
+in sync. The lag between these
+various factors is what causes many of the complexities of the BOLD signal.
 
@@
 \subsection{BOLD Physiology}
-It is will known that Deoxyhemoglobin acts as a contrast agent in EPI imaging
-\cite{Buxton1999}, \cite{WEISSKOFF1994}, \cite{Ogawa}, so then what causes
-changes Deoxyhemoglobin in a voxel? Obviously increased metabolism will increase
-Deoxyhemoglobin, but changes in blood volume in a voxel will also alter 
-the MR signal depending on the ratio of Deoxyhemoglobin to hemoglobin. 
+It is well known that the two types of hemoglobin act as a contrast agents in 
+EPI imaging
+\cite{Buxton1999}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
+between DeHb/O2Hb and neural activity is non-trivial. Intuitively, increased 
+metabolism will increase Deoxyhemoglobin, however blood vessels are quick
+to compensate by increasing local blood flow. Increased inflow will of course
+preceed increased outflow, and increased inflow is encouraged by loosening 
+capilary beds. Both of these factors drive increased storage capacity.
+Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
+Hemoglobin, increased volume of blood can certainly effect this ratio if 
+metabolism doesn't exactly match the increased inflow of oxygenated blood.
 This was the impetus
-for the ground breaking research \cite{Buxton1999} and \cite{Mandeville1999}.
+for the ground breaking balloon model (\cite{Buxton1999}) and windkessel
+model (\cite{Mandeville1999}). These models derive from first principals
+the increased deoxyhemoglobin ratio and volume of capillaries based on flow.
 These were the first two attempts to quantitatively account for the shape of the 
@@
 and the cerebral blood volume (CBV). In fact \cite{Buxton1999} went to far as
-to show that a simple, well chosen blood flow waveform, along with a balloon 
-model, could fully account for the BOLD signal. Although complex
-changes in metabolism exist (\cite{Zheng2002}), most recent studies have
+to show that a simple, well chosen blood flow waveform coupled with a simple
+cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
+model, could fully account for the BOLD signal.
+
+Although complex
+models for metabolism exist (\cite{Zheng2002}), most recent studies have
 focused on their ability to explain the prolonged BOLD post stimulus 
 undershoot \cite{Donahue2009}, \cite{Yacoub2006}. This is because \cite{Buxton2004}
-and later \cite{Riera2004} showed that the primary signal may be accurately
-be estimated by a simple flow-locked expression of the CMRO2. Comparison
-between the more complicated models including a separate expression for CMRO2
-and the simplified Buxton expression show little actual difference \cite{Deneux2006}.
-In \cite{Buxton1999}, it was shown that small changes in the flow expression could
+and later \cite{Riera2004} showed that the main portion of the signal may be 
+accurately estimated by a simple blood flow locked expression of the CMRO2. 
+Although \cite{Deneux2006} did not deal extensively with prolonged post
+stimulus undershoot, the comparisons made in that publication showed minimal
+improvement from separate expressions of CMRO2. For some time there was 
+debate regarding the cause for the post-stimulus undershoot often seen in 
+FMRI studies. The problem stemmed from the fact that there were many studies
+showing CBV and CBF returning to baseline well before the BOLD signal. In fact
+this result seemed to be confirmed by several studes: \cite{Frahm2008}, \cite{Donahue2008}, 
+\cite{Lu2004}.
+
+%here
+
+In fact, the original Buxton
+paper demonstrated how minimal changes in the flow expression could
 result in vastly different recovery times of the baseline BOLD signal, and
-in \cite{Mandeville1999} a cause for delayed CBV recovery was proposed, based
-on differences in compliance of capillaries. \cite{Behzadi2005} provided a simple,
-complete model and explanation for delayed compliance of neural capillaries,
-and those simple changes can result in a significant improvement in the bias error,
-as shown in \cite{Deneux2006}. In the end, it may be impossible to refute the
+in \cite{Mandeville1999} a cause for delayed CBF recovery was proposed, based
+on hysterous in the compliance of capillaries. \cite{Behzadi2005} provided
+additional incite into this possibility, going so far as proposing a simple
+yet complete model and explanation for delayed compliance of neural capillaries.
+Those simple changes can result in error levels on par with the more 
+complex CMRO2 inclusive models, which was one of the major findings of \cite{Deneux2006}. 
+This was reiterated in \cite{Chan2009}, which concluded that many of
+the studies claiming a CBV recovery well before the BOLD signal 
+(\cite{Frahm1996}, \cite{Kruger1996}, \cite{Lu2004}) 
+could actually have been detecting arterial blood flow, which does not significantly effect the
+BOLD signal. 
+In the end, it may be impossible to refute the
 possibility of a prolonged period of increased metabolic activity after activation,

### 2010-04-25 doc/thesis/thesis.tex
more work on the lit review, and some cleanup

@@
 \documentclass{article}
-\usepackage[left=1cm,top=2cm,right=1cm,nohead,nofoot]{geometry}
-
-%\usepackage{times}
+\usepackage[left=3cm,top=2cm,right=3cm,nohead,nofoot]{geometry}
+
+\usepackage{times}
 %\usepackage{tikz}
@@
 
-%\doublespacing
+\doublespacing
 
@@
 typically employed today. Then again, depending on the model there may be between
-seven \cite{Riera2004} and fifteen \cite{Zheng2002} parameters per voxel to
+seven \cite{Riera2004} and fifteen \cite{Zheng2005} parameters per voxel to
 be optimized. Clearly there is a significant risk of error due to variance
@@
 EPI imaging
-\cite{Buxton1999}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
-between DeHb/O2Hb and neural activity is non-trivial. Intuitively, increased 
+\cite{Buxton1998}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
+between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
+Intuitively, increased 
 metabolism will increase Deoxyhemoglobin, however blood vessels are quick
 to compensate by increasing local blood flow. Increased inflow will of course
-preceed increased outflow, and increased inflow is encouraged by loosening 
+preceed increased outflow, and increased inflow is accomplished by loosening 
 capilary beds. Both of these factors drive increased storage capacity.
@@
 This was the impetus
-for the ground breaking balloon model (\cite{Buxton1999}) and windkessel
+for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
 model (\cite{Mandeville1999}). These models derive from first principals
-the increased deoxyhemoglobin ratio and volume of capillaries based on flow.
+the increased deoxyhemoglobin ratio and volume of capillaries based on a given flow.
 These were the first two attempts to quantitatively account for the shape of the 
-BOLD signal as a consequence of the lag between the cerebral blood flow (CBF) 
-and the cerebral blood volume (CBV). In fact \cite{Buxton1999} went to far as
-to show that a simple, well chosen blood flow waveform coupled with a simple
-cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
+BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
+and the cerebral blood flow (CBF). In fact \cite{Buxton1998} went to far as
+to show that a simple, well chosen blood flow waveform coupled with a square 
+wave cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
 model, could fully account for the BOLD signal.
 
+Although \cite{Buxton1998} showed that a well chosen flow waveform could 
+explain much of the BOLD signal, there was still a matter of proposing a
+realistic waveform for the CBF and for the CMRO2. \cite{friston2000} gave
+a reasonable and simple
+expressoin for CBF input,$f$, based on a flow inducing signal, $s$, 
+\begin{eqnarray}
+\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
+\dot{f} &=& s
+\end{eqnarray}
+where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
+are both time constants. In \cite{Buxton2004} the final piece of the simple balloon
+model was put into place, by describing the CMRO2 as a constant multiple of
+the CBF (the inflow of blood). This completed the basic balloon model, and 
+was well summarized in \cite{Riera2004}. 
+\begin{eqnarray}
+\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\end{eqnarray}
+where $v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
+local deoxyhemoglobin/oxygenated hemoglobin ratio, $E_0$ is the resting metabolic
+rate and $\alpha$ is Grubb's parameter controling the balloon model. 
+\cite{Obata2004} refined the readout equation of the BOLD signal based on the
+deoxyhemoglobin content (q) and local blood volume (v), resulting in the
+final BOLD equation:
+\begin{eqnarray}
+y   &=& V_0((k_1 + k_2)(1-q) - (k_2 + k_3)(1-v))\\
+k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8\\
+K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57\\
+k_3 &=& \epsilon_0 - 1 = .43
+\end{eqnarray}
+Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
+deoxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
+change in relaxation rate with change in blood oxygenation, and
+$\epsilon_0 = 1.43$ is the 
+ratio of signal MR from intravascular to extravascular at rest. Although,
+obviously these constants change with experiment ($TE$, $\nu_0$, $r_0$),
+patient, and brain 
+region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are used
+as constant ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$).
+While this model is in a sense complete, it is far from perfect. The major
+problem often brought up with this version of the BOLD model is that it
+does not represent the so called "post-stimulus undershoot" well.
+
 Although complex
-models for metabolism exist (\cite{Zheng2002}), most recent studies have
+models for metabolism exist (\cite{Zheng2005}), most recent studies have
 focused on their ability to explain the prolonged BOLD post stimulus 
@@
 and later \cite{Riera2004} showed that the main portion of the signal may be 
-accurately estimated by a simple blood flow locked expression of the CMRO2. 
+accurately estimated by a simple blood flow locked expression of the CMRO2.. 
 Although \cite{Deneux2006} did not deal extensively with prolonged post
 stimulus undershoot, the comparisons made in that publication showed minimal
-improvement from separate expressions of CMRO2. For some time there was 
+improvement from separate expressions of CMRO2, in comparison to the much 
+increased complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
+terms, first proposed in \cite{Buxton2004}, that a slowed return to baseline for
+the BOLD signal is possible to model. However, viscoelastic effects primarily
+control CBV, which some studies have claimed returns to baseline \emph{before}
+the BOLD signal. If this is the case, then CBV cannot be the cause of the 
+post-stimulus undershoot. Even now there is some debate about the cause of 
+the post-stimulus undershoot. The BOLD post-stimulus undershoot is common,
+and typically lasts longer for longer activations (\cite{Chen2009}), and in
+fact varies across regions of the brain (\cite{Mandeville2001}, \cite{Yacoub2006})
 debate regarding the cause for the post-stimulus undershoot often seen in 
@@
 showing CBV and CBF returning to baseline well before the BOLD signal. In fact
-this result seemed to be confirmed by several studes: \cite{Frahm2008}, \cite{Donahue2008}, 
+this result seemed to be confirmed by several studes: \cite{Frahm2008}, \cite{Donahue2009}, 
 \cite{Lu2004}.
@@
 complex CMRO2 inclusive models, which was one of the major findings of \cite{Deneux2006}. 
-This was reiterated in \cite{Chan2009}, which concluded that many of
+This was reiterated in \cite{Chen2009}, which concluded that many of
 the studies claiming a CBV recovery well before the BOLD signal 
-(\cite{Frahm1996}, \cite{Kruger1996}, \cite{Lu2004}) 
+(\cite{Frahm1996}, \cite{Kruger1998}, \cite{Lu2004}) 
 could actually have been detecting arterial blood flow, which does not significantly effect the
@@
 however, much of the BOLD post stimulus undershoot may be explained by a 
-prolonged CBV undershoot, as shown in \cite{Chan2009}. Because of the greatly
+prolonged CBV undershoot, as shown in \cite{Chen2009}. Because of the greatly
 increased complexity of CMRO2 modeling, compared with the relative simplicity
@@
 past decade and a half, but there is still some question about the origins
-of certain signal features. In the seminal paper \cite{Buxton1999}, the original
+of certain signal features. In the seminal paper \cite{Buxton1998}, the original
 Balloon model was proposed. Buxton et. al showed that all of the features
@@
 the work of recharging is done, there may be a prolonged undershoot lasting
-as much as 90 seconds \cite{spatial_undershoot}, though the reason for this
-is debated \cite{origin_poststim}.
+as much as 90 seconds \cite{Yacoub2006}, though the reason for this
+is debated \cite{Chen2009}.
 
@@
 The model we will use is the model proposed by Buxton Et. Al. and later used
-in Riera et. al. \cite{Riera}. The model has four state variables, $s$, $f$,
+in Riera et. al. \cite{Riera2004}. The model has four state variables, $s$, $f$,
 $v$, $q$, representing flow inducing signal, cerebral blood flow, cerebral
@@
 
-\bibliographystyle{abbrvnat}
-\bibliography{references}
+\bibliographystyle{apalike}
+%\bibliographystyle{abbrvnat}
+%\bibliographystyle{abbrv}
+\bibliography{library}
 

### 2010-04-25 doc/thesis/thesis.tex
sort of finished with the balloon model part

@@
 typically employed today. Then again, depending on the model there may be between
-seven \cite{Riera2004} and fifteen \cite{Zheng2005} parameters per voxel to
+seven \cite{Riera2004} and 50 \cite{Behzadi2005} parameters per voxel to
 be optimized. Clearly there is a significant risk of error due to variance
@@
 region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are used
-as constant ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$).
+as constants ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$) in 1.5 Tesla studies..
 While this model is in a sense complete, it is far from perfect. The major
@@
 does not represent the so called "post-stimulus undershoot" well.
-
-Although complex
-models for metabolism exist (\cite{Zheng2005}), most recent studies have
-focused on their ability to explain the prolonged BOLD post stimulus 
-undershoot \cite{Donahue2009}, \cite{Yacoub2006}. This is because \cite{Buxton2004}
+The post-stimulus undershoot is the name for a prolonged sub-normal
+BOLD response for a period of 10 to 60 seconds after stimulus has
+ceased (\cite{Chen2009}, \cite{Mandeville1999a}).
+
+There are two theories for the cause of the post stimulus undershoot. Recall
+that a lower than base signal means that there is an increased deoxyhemoglobin
+content in the voxel. The first and simplest explanation is that the post-stimulus
+undershoot is caused by a prolonged increase in CMRO2 after CBV and CBF
+have returned to their base levels. This theory is justified by quite a few
+studies that show CBV and CBF returning to the baseline before the BOLD signal
+(\cite{Frahm2008}, \cite{Donahue2009}, \cite{Buxton2004}, \cite{Lu2004},
+\cite{Shen2008}). Unfortunately, because of limitations on FMRI and en vivo
+CBV/CBF measurement techniques it is difficult to isolate whether CBF and
+CBV truly have returned to their baseline. Other research seems to indicate
+that there can be a prolonged residual supernormal CBV (\cite{Mandeville1999a}, 
+\cite{Behzadi2005}, \cite{Chen2009a}), although none of these papers completely
+rule out the possiblity of increased CMRO2. Additionally, in \cite{Yacoub2006}, it 
+was found that the post-stimulus undershoot varried across regions of the brain, 
+which could further explain the contradictions found elsewhere. \cite{Chen2009a}
+makes a compelling case that most of the post stimulus undershoot could be 
+explained be a prolonged CBV increase, and a prolonged CBF undershoot, and that
+many of the previous measurements showing a quick recovery of CBV 
+may have been dominated arterial CBV's return to baseline.
+
+Because of the significant possility of a completely independent CMRO2,
+extremely complex models for metabolism exist (\cite{Zheng2005}), although 
+most recent studies have
+focuse on their ability to explain the prolonged BOLD post stimulus 
+undershoot \cite{Zheng2005}, \cite{Buxton2004}. This is because \cite{Buxton2004}
 and later \cite{Riera2004} showed that the main portion of the signal may be 
@@
 the BOLD signal is possible to model. However, viscoelastic effects primarily
-control CBV, which some studies have claimed returns to baseline \emph{before}
-the BOLD signal. If this is the case, then CBV cannot be the cause of the 
-post-stimulus undershoot. Even now there is some debate about the cause of 
-the post-stimulus undershoot. The BOLD post-stimulus undershoot is common,
-and typically lasts longer for longer activations (\cite{Chen2009}), and in
-fact varies across regions of the brain (\cite{Mandeville2001}, \cite{Yacoub2006})
-debate regarding the cause for the post-stimulus undershoot often seen in 
-FMRI studies. The problem stemmed from the fact that there were many studies
-showing CBV and CBF returning to baseline well before the BOLD signal. In fact
-this result seemed to be confirmed by several studes: \cite{Frahm2008}, \cite{Donahue2009}, 
-\cite{Lu2004}.
-
-%here
-
-In fact, the original Buxton
-paper demonstrated how minimal changes in the flow expression could
-result in vastly different recovery times of the baseline BOLD signal, and
-in \cite{Mandeville1999} a cause for delayed CBF recovery was proposed, based
-on hysterous in the compliance of capillaries. \cite{Behzadi2005} provided
-additional incite into this possibility, going so far as proposing a simple
-yet complete model and explanation for delayed compliance of neural capillaries.
-Those simple changes can result in error levels on par with the more 
-complex CMRO2 inclusive models, which was one of the major findings of \cite{Deneux2006}. 
-This was reiterated in \cite{Chen2009}, which concluded that many of
-the studies claiming a CBV recovery well before the BOLD signal 
-(\cite{Frahm1996}, \cite{Kruger1998}, \cite{Lu2004}) 
-could actually have been detecting arterial blood flow, which does not significantly effect the
-BOLD signal. 
-In the end, it may be impossible to refute the
-possibility of a prolonged period of increased metabolic activity after activation,
-however, much of the BOLD post stimulus undershoot may be explained by a 
-prolonged CBV undershoot, as shown in \cite{Chen2009}. Because of the greatly
-increased complexity of CMRO2 modeling, compared with the relative simplicity
-of the Behzadi model, in most cases it is not worth using the more advanced CMRO2
-model unless the metabolic rate is an end goal.
-
-The physiology that leads to the BOLD signal has been well studied over the 
-past decade and a half, but there is still some question about the origins
-of certain signal features. In the seminal paper \cite{Buxton1998}, the original
-Balloon model was proposed. Buxton et. al showed that all of the features
-of the BOLD signal may be explained by lag between ... . While some uncertainty remains (for instance in
-relation to the BOLD post-stimulus undershoot <papers relating to back and
-forth on post-stimulus undershoot>) generally the model fits reality very 
-well. As with every piece of tissue in the body, the brain requires oxygen
-to extract energy from glucose in the blood. This process of removing oxygen
-from red blood cells sets into motion a chain of events that locally 
-alters the composition of the blood. Inactive neurons can be thought of
-as a slingshot cocked and ready to fire. As soon as a signal moves up
-the axon and causes the neuron to fire (and thus changes the state
-of a membrane at that location), ions quickly move across the altered
-membrane to compensate for a high charge imbalance between the interior
-and exterior of the cell. This process, similar to allowing a stretched
-rubber band to contract leaves the system at a lower energy state. This cannot
-last though, because the neuron needs to be ready to fire again rather quickly.
-To return the cell to a ready position the membrane becomes impermeable
-to the ions again, and then begins pumping ions back into the cell. This
-process takes a large amount of energy, whereas the actual firing takes
-very little energy. Thus, after firing, glucose is burned, removing oxygen
-from the blood and thus causing a dip in the ratio of oxygenated hemoglobin to 
-de-oxygenated hemoglobin. This is the first potentially measurable effect
-that MRI can see, although it usually lasts fewer than 2 seconds making it
-difficult for FMRI to catch. As a result of the decreased amount of local oxygen,
-the capillaries compensate by increasing blood flow to that region. Because,
-of the quickly increased blood flow into local capillaries, the blood
-volume increases in addition to the local oxygen content. This leads to a Windkessel
-effect which further lags the normalization of oxygen content. The effect of
-all this is an overshoot in the oxygen content above the initial level. After
-the work of recharging is done, there may be a prolonged undershoot lasting
-as much as 90 seconds \cite{Yacoub2006}, though the reason for this
-is debated \cite{Chen2009}.
-
-A large number of models have been proposed for the BOLD signal with varying
-amounts of complexity. The simplest model is the so called "Balloon Model"
-which first proposed the windkessel effect as a factor the BOLD response.
-The model we will use is the model proposed by Buxton Et. Al. and later used
-in Riera et. al. \cite{Riera2004}. The model has four state variables, $s$, $f$,
-$v$, $q$, representing flow inducing signal, cerebral blood flow, cerebral
-blood volume and deoxyhemaglobin to hemaglobin ratio, respectively. The 
-state variables change over time, given by the state evolution equations:
-\begin{equation}
-\dot{s} = \epsilon u -  
-\end{equation}
-
-\begin{equation}
-\dot{f} = s
-\end{equation}
-
-\begin{equation}
-\dot{v} = 
-\end{equation}
-
-\begin{equation}
-\dot{q} = 
-\end{equation}
-
-Additional papers have added such things as
-....
-
-All these effects are generally accepted as the cause of the BOLD signal, 
-but FMRI doesn't detect this happening in one neuron, but rather as the 
+control CBV, which, as mentioned already, many studies have claimed cannot be 
+responsible for the BOLD post-stimulus undershoot. Another extensive model
+that attempts to quantify the post-stimulus undershoot is the compliance model 
+proposed by \cite{Behzadi2005}. Alhough through a somewhat different means than
+the \cite{Zheng2005} and \cite{Buxton2004} papers, its possible the increased
+model flexibility ultimately is the key reason for the improvements, as opposed
+to increased plausibility. Because of these controversies, and because this is
+the first time a particle filter has been used for this problem in this way, 
+our aim to keep the model simple was best met by using the original balloon
+model with the addition of the visco-elastic effects from \cite{Buxton2004}.
+
+In the end, its possible thay the balloon model may be as complex as models
+may get before error becomes dominated by model variance, so proving that
+Behzadi's Arterial Compliance model is quantitatively better than Buxton's 
+advanced balloon model \cite{Buxton2004} or that either is better than 
+the extremely detailed model found in \cite{Zheng2005}. The reality is that
+measurement noise alone may dominate any difference between these models.
+
+\subsection{Noise}
+Thus, despite some discrepencies, the cause of the BOLD signal is
+relatively well known.  but FMRI doesn't detect this happening in one neuron,
+but rather as the 
 aggregate over millions of cells. Though local neurons act
@@
 A particularly insidious type of noise present in FMRI is a low frequncey
-drift, characterized by a Weiner process. Though not present in all
-regions, it is prevalent enough to cause problems \cite{detrend}. It is still not
+drift, first characterized by a Weiner process in \cite{Riera2004}. 
+Though not present in all regions, it is prevalent enough to cause significant
+inference problems \cite{Tanabe2002}. It is still not
 clear where exactly this noise comes although it is possible it is 
 the result of magnets heating up, or some distortion in magnetic
-fields. It is clear that this drift signal is not due to a true physiological
-effects however, given its presence in cadavers and phantoms\cite{drift}.
+fields \cite{Smith2007}. It is clear that this drift signal is not soley
+due to a true physiological
+effects however, given its presence in cadavers and phantoms \cite{Smith1999}.
+Additionally, since it is standard operating procedure to perform
+coregistration between volumes, it is unlikely that movement was
+the cause of the drift either.
 

### 2010-04-26 doc/thesis/thesis.tex
talk about priors/parameter values

@@
 our aim to keep the model simple was best met by using the original balloon
-model with the addition of the visco-elastic effects from \cite{Buxton2004}.
-
-In the end, its possible thay the balloon model may be as complex as models
-may get before error becomes dominated by model variance, so proving that
+model with the possible addition of the visco-elastic effects from \cite{Buxton2004}.
+
+Even more advanced versions of the Balloon model exist. In fact \cite{Buxton2004}
+introduced several additional state variables, including the CMRO2, the O2 extraction
+fraction, which is closely related to CMR02, and the neural response, which 
+essentially starts high and decays toward some steady state value. The 
+neural response is intended to emulate neural habituation, wherein neurons
+become less sensitive to a prolonged stimulus. This goes back to the issue
+of drastically increasing the model variance, for a minor decrease in model
+bias, \cite{Deneux2006}. The parameters of the model have decently reliable
+expected values by now. Although \cite{Buxton1998} and \cite{Friston2000}
+both propose reasonable values for the model parameters, \cite{Friston2002} 
+was the first paper to actually use an algorithm and test those parameters:
+\begin{eqnarray}
+epsilon &=& .54 \pm .1    \nonumber \\
+tau_s & =& 1.54 \pm .25   \nonumber \\
+tau_f & =& 2.46 \pm .25   \nonumber \\
+tau_0 & =& .98 \pm .25    \nonumber \\
+alpha & =& .33 \pm .45    \nonumber \\
+E_0   & =& .34 \pm .1     \nonumber \\
+V_0  & = & .03 (not estimated) \nonumber
+\end{eqnarray}
+later \cite{Hu2009} used an Unscented Kalman Filter to confirm these findings.
+On the other hand \cite{Johnston2008} used a hybrid particle filter Expectation
+Maximization algorithm to calculate parameters simultaneously with state.
+His findings differed greatly from Friston et. al.:
+\begin{eqnarray}
+epsilon &=& .069 \pm .014    \nonumber \\
+tau_s & =& 4.98 \pm 1.07  \nonumber \\
+tau_f & =& 8.31 \pm 1.51   \nonumber \\
+tau_0 & =& 8.38 \pm 1.5    \nonumber \\
+alpha & =& .189 \pm .004   \nonumber \\
+E_0   & =& .635 \pm .072     \nonumber \\
+V_0   & =& .0149 \pm .006     \nonumber 
+\end{eqnarray}
+On the one hand, Johnston et. al. is a pseudo-maximum likelihood estimator, and 
+it depends very little on the priors. On the other hand, the results are very 
+different from \cite{Buxton1998}, \cite{Hu2009}, \cite{Friston2000}, and 
+\cite{Friston2002}. Then again, the papers mentioned all have a common root,
+and all used essentially the same priors, and all use Bayesian estimation
+that depends greatly on the priors. It would be interesting to see what
+effect different prior's would have on the results.
+
+In the end, its possible that the balloon model may be as complex as the
+measurement rate and error size will allow before the error becomes dominated 
+by model variance, so proving that
 Behzadi's Arterial Compliance model is quantitatively better than Buxton's 
 advanced balloon model \cite{Buxton2004} or that either is better than 
-the extremely detailed model found in \cite{Zheng2005}. The reality is that
+the extremely detailed model found in \cite{Zheng2005} may be impossible. 
+The reality is that
 measurement noise alone may dominate any difference between these models.

### 2010-04-28 doc/thesis/thesis.tex
small amount of work on previous studies

@@
 \usepackage{cancel}
+\usepackage{hyperref}
 
@@
 most recent studies have
-focuse on their ability to explain the prolonged BOLD post stimulus 
+focused on their ability to explain the prolonged BOLD post stimulus 
 undershoot \cite{Zheng2005}, \cite{Buxton2004}. This is because \cite{Buxton2004}
@@
 that attempts to quantify the post-stimulus undershoot is the compliance model 
-proposed by \cite{Behzadi2005}. Alhough through a somewhat different means than
+proposed by \cite{Behzadi2005}. Although through a somewhat different means than
 the \cite{Zheng2005} and \cite{Buxton2004} papers, its possible the increased
@@
 fraction, which is closely related to CMR02, and the neural response, which 
-essentially starts high and decays toward some steady state value. The 
+causes the stimulus to decay toward some steady state value. The 
 neural response is intended to emulate neural habituation, wherein neurons
-become less sensitive to a prolonged stimulus. This goes back to the issue
-of drastically increasing the model variance, for a minor decrease in model
-bias, \cite{Deneux2006}. The parameters of the model have decently reliable
-expected values by now. Although \cite{Buxton1998} and \cite{Friston2000}
-both propose reasonable values for the model parameters, \cite{Friston2002} 
-was the first paper to actually use an algorithm and test those parameters:
+become less sensitive to a prolonged stimulus. While these advanced may be more
+capable of capturing a more exact version of the BOLD signal, the difference
+between the model will often times be below the noise floor. In essence this 
+is a classic bias-variance
+dilemma: at some point increased model flexibility, and thus variance, is not
+worth the decrease in model bias. For now it remains to be seen where this
+line may be drawn in the BOLD signal, although \cite{deneux2006} does not
+show a significant improvement from the additional parameters added in 
+\cite{Buxton2004}.
+
+\subsection{Previous Studies of Parameters}
+There have been quite a few efforts to quantify the parameters of the
+various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
+both proposed physiologically reasonable values for the model parameters, 
+\cite{Friston2002} was the first paper to calculate the parameters based 
+on actual FMRI data. In that paper, Friston et. al. used a variation of
+Expectation Maximization to find a normal distribution for the parameters:
 \begin{eqnarray}
-epsilon &=& .54 \pm .1    \nonumber \\
-tau_s & =& 1.54 \pm .25   \nonumber \\
-tau_f & =& 2.46 \pm .25   \nonumber \\
-tau_0 & =& .98 \pm .25    \nonumber \\
-alpha & =& .33 \pm .45    \nonumber \\
-E_0   & =& .34 \pm .1     \nonumber \\
-V_0  & = & .03 (not estimated) \nonumber
+\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
+\tau_s & =&  N(1.54, .25^2)   \nonumber \\
+\tau_f & =&  N(2.46, .25^2)   \nonumber \\
+\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
+\alpha & =&  N(.33 , .45^2 )   \nonumber \\
+E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
+V_0  & = &  .03 (not\ estimated) \nonumber
 \end{eqnarray}
+Since then, several other methods of have been used to calculate
+BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
+likelihood method for innovation processes was used as described by
+\cite{Ozaki1994}
 later \cite{Hu2009} used an Unscented Kalman Filter to confirm these findings.
-On the other hand \cite{Johnston2008} used a hybrid particle filter Expectation
-Maximization algorithm to calculate parameters simultaneously with state.
-His findings differed greatly from Friston et. al.:
+Both \cite{Hu2009} and \cite{Friston2002} came to results very similar to the
+expected values.
+On the other hand \cite{Johnston2008} used a hybrid particle filter Expectation-
+Maximization algorithm to calculate parameters and state simultaneously.
+His derived parameter distributions differed greatly from Friston et. al.:
 \begin{eqnarray}
@@
 \end{eqnarray}
+%here
 On the one hand, Johnston et. al. is a pseudo-maximum likelihood estimator, and 
@@
 different from \cite{Buxton1998}, \cite{Hu2009}, \cite{Friston2000}, and 
-\cite{Friston2002}. Then again, the papers mentioned all have a common root,
+\cite{Friston2002}. All four of the papers just mentioned all have a common root,
 and all used essentially the same priors, and all use Bayesian estimation
-that depends greatly on the priors. It would be interesting to see what
-effect different prior's would have on the results.
+that depends greatly on the priors. In this case, it is possible that priors
+biased the results, preventing accurate measurements of the parameters. Its also
+possible that the parameters don't lend themselves to learning very well. 
+It would be interesting to see what
+effect different priors would have on the results in \cite{Friston2002}.
 

### 2010-04-28 doc/thesis/thesis.tex
emph{sort} of done with the lit review, but i need to review it...

@@
 BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
-likelihood method for innovation processes was used as described by
-\cite{Ozaki1994}
-later \cite{Hu2009} used an Unscented Kalman Filter to confirm these findings.
-Both \cite{Hu2009} and \cite{Friston2002} came to results very similar to the
-expected values.
-On the other hand \cite{Johnston2008} used a hybrid particle filter Expectation-
-Maximization algorithm to calculate parameters and state simultaneously.
-His derived parameter distributions differed greatly from Friston et. al.:
+likelihood method for innovation processes was used, as described by
+\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
+Kalman filter, to break the time series into a series of innovations,
+for which Maximum Likelihood was performed. While this is in some since the
+"right" way to find the solution, it comes with several caveats. First, every
+step in parameter space requires a recalculation of all the state variables. With
+two or three parameters this is fine, more than that, and calculations could go on
+indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
+will only be optimal in that case. Third, depending on the nonlinearities present
+in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
+Unscented Kalman Filter over all the parameters and state variables to find the 
+parameter set/variable time series. While this method has the drawback of not necessarily
+being optimal, unfortunately there is no general optimal solution to non-linear non
+Gaussian models. Hu et. al.'s technique also will run significantly faster than
+ML based techniques, since it does not require recalculating the entire timeseries
+for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
+very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
+One potential problem with all these techniques are that the depend heavily on the
+priors. The starting point of the parameters could have a huge impact on the
+results, and while one can be hopeful that this isn't the reason for the agreement
+between \cite{Friston2000} and later results, there is no way to know.
+
+In \cite{Jonston2008}, a hybrid particle filter/gradient
+descent algorithm was used to simultaneously derive the static (classicly called
+parameters) and dynamic parameters (classically known as state variables).
+Essentially a particle filter is used to calculate the state variables at some
+time, and then the estimated distribution of the particles was used to find
+the most likely set of parameters that would give that distribution of state variables.
+\cite{Johnston2008} comes to a very different set of parameter estimates as compared
+to the original \cite{Friston2000} guesses:
 \begin{eqnarray}
@@
 \end{eqnarray}
-%here
-On the one hand, Johnston et. al. is a pseudo-maximum likelihood estimator, and 
-it depends very little on the priors. On the other hand, the results are very 
-different from \cite{Buxton1998}, \cite{Hu2009}, \cite{Friston2000}, and 
-\cite{Friston2002}. All four of the papers just mentioned all have a common root,
-and all used essentially the same priors, and all use Bayesian estimation
-that depends greatly on the priors. In this case, it is possible that priors
-biased the results, preventing accurate measurements of the parameters. Its also
-possible that the parameters don't lend themselves to learning very well. 
-It would be interesting to see what
-effect different priors would have on the results in \cite{Friston2002}.
-
-In the end, its possible that the balloon model may be as complex as the
-measurement rate and error size will allow before the error becomes dominated 
-by model variance, so proving that
-Behzadi's Arterial Compliance model is quantitatively better than Buxton's 
-advanced balloon model \cite{Buxton2004} or that either is better than 
-the extremely detailed model found in \cite{Zheng2005} may be impossible. 
-The reality is that
-measurement noise alone may dominate any difference between these models.
+Notably, the time constants are significantly longer. This could be a result of
+some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
+later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
+depended less on the priors.
+
+Its possible, although unlikely that the balloon model is not possible to learn
+without detrimental cost in variance error. I do not think this is the case, however,
+and I will work to dispell this possibility in the results with simulated time
+series.
 
@@
 inference problems \cite{Tanabe2002}. It is still not
-clear where exactly this noise comes although it is possible it is 
+clear where exactly this noise comes from, although it is possible it is 
 the result of magnets heating up, or some distortion in magnetic
 fields \cite{Smith2007}. It is clear that this drift signal is not soley
-due to a true physiological
-effects however, given its presence in cadavers and phantoms \cite{Smith1999}.
+due to a physiological
+effects, given its presence in cadavers and phantoms \cite{Smith1999}.
 Additionally, since it is standard operating procedure to perform
 coregistration between volumes, it is unlikely that movement was
-the cause of the drift either.
+the cause of the drift either (otherwise the effect would not be 
+widely reported).
 
@@
 \subsection{Basic Statistical Parametric Mapping}
-The most basic method of analyzing FMRI data is through a basic T-test
+Although not strictly the same thing as parameter calculation from 
+FMRI, activation detection is very similar. In fact, estimation of 
+parameters is somewhat a generalization of the idea of activation detection.
+Thus, it is important to draw a comparison between the methods proposed
+in this thesis with existing methods of activation detection.
+
+The most basic method of analyzing FMRI data is through a standard T-test
 between "resting state" and "active state" samples. This is done by 
 taking the average and variance of the inactive period, and the 
-active period separately then treating them both as gaussian distributions.
+period during which the stimulus was activate separately then treating 
+them both as gaussian distributions.
 If they are in fact Gaussian distributions, then a basic t-test will
-give the likelihood that the samples came from the same distribution
-(the p-value). Of course, this test is fraught with problems; even if
+give the probability that the samples came from the same distribution
+(the null hypothesis). Of course, this test is fraught with problems; even if
 the drift mentioned earlier has been removed, there is little reason
@@
 even if the noise were Gaussian, a t-test with a p-value of .05 over
-5000 or more samples is on average going to generate $.05*5000$ false
+50000 or more samples is on average going to generate $.05*50000$ false
 positives. To compensate for this, bonferoni correction, also known as
@@
 \subsection{General Linear Model}
-The most used form of FMRI analysis is still based on Statistical Parametric
+The most used form of FMRI analysis is Statistical Parametric
 Mapping, but is able to account for several different levels or types
-of stimulus. By adding a General Linear Model to the analysis, the
-output signal timeseries (what FMRI detects) is regressed over the weighted
-sum of the various confound's timeseries. The equation for a general linear
-model is then
+of stimulus (see \cite{Hofmann1997}). By using hierarchical models
+the output signal timeseries is considered the weighted
+sum of the various input timeseries. Essentially every experimental factor
+is considered as another level inputs (ex. which patient is an input). 
+The equation for a general linear model is then
 \begin{equation}
@@
 where $Y(t)$ is the smoothed or detrended timeseries of measurements,
-$X(t)$ is a row vector of stimuli, $\beta$ is a column vector of weights,
+$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
 and $\epsilon$ is the error. Thus for every time, the measurement is
 assumed to be a weighted sum of the inputs plus some error. The calculation
-of $\beta$ then is merely a gradient descent search to minimize the
-mean squared error.
+of $\beta$ then performed using a maximum likelihood or gradient descent search 
+to minimize the error.
 
@@
 
-Of course, a square wave input is not going to result in a square wave
+It is well known of course that a square wave stimulus does not result in a square wave
 in the activation of brain regions. Thus, various methods are used to 
-smooth $X(t)$ through time, and bandlimit the input. The best technique
-is convolving the stimulus input with a hemodynamic response function,
-which mimicks the basic shape of BOLD activation, including a delay
-due to rise time and fall time. This hemodynamic signal is static however,
-so every region of the brain gets the same design matrices (X(t)), 
-although the weights of various stimulus or confounds are allowed to vary.
+smooth $X(t)$'s related to actual input (as opposed to experimental confounds
+such as patient number, age etc) through time, and bandlimit the input. 
+The best technique is convolving the stimulus input with a hemodynamic 
+response function, which mimicks the basic shape of BOLD activation, including a delay
+due to rise time and fall time. The downside of this method, however is that 
+the hemodynamic signal is static, meaning the same hemodynamic function is
+used for every region of the brain, and $Y(t)$ must be a constant
+scalar multiple of $X(t)$ to be appropiately identified.
 
@@
 
-Ultimately, activation due to a particular stimuli is decided by the 
-$\beta$ value corresponding to that stimuli's column of $X(t)$.
-<Need to check this>
-The null hypothesis as to whether the outcome was random is then
-based on a t-test of the $\epsilon(t)$ timeseries.
-
-\subsection{Whats wrong with these techniques}
-There are a few problems with the techniques mentioned in the 
-previous sections. First, they essentially ignore prior knowledge about
-the system. Although the most advanced form of the general linear model includes
-a "Hemodynamic Response Function," that hemodynamic response function is
-static across every region of the brain. It is well known that capillary beds
-are not uniform and so blood perfusion cannot possible be static across the
-brain. Thus, if extra information were available a-priori, that information
-could not be incorporated without modifications to the General Linear Model.
-Similarly heart rate could not be added either. It would obviously be advantageous
-to have true physiological parameters as entry points for these various other
-model parameters. The physiological models for the BOLD signal are quite good
-and based on realistic physics. While the exact connection between a stimulus
-and the flow inducing signal is not precisely known, model fits are actually
-quite good \cite{nonlinearmodels}. Regardless, being based on some real 
-physiological 
-parameter would allow for the establishment of reasonable priors and decrease
-model variance without breaking a sweat. Of course, using real parameters 
-has the additional
-bonus of potentially providing information about physical pathologies. It
+The GLM is extremely powerful at determining the linear dependence of
+a set of regressors on the output. Unfortunately, there is significant evidence
+that many of these dependencies are nonlinear, which means they may be
+difficult or impossible to detect using linear techniques. Additionally, 
+the lack of flexibility in delay variables certainly hinders SPM's
+ability to locate voxels that do not conform to the canonical hemodynamic 
+response function. The very fact that different hemodynamic response functions
+work better for different regions/experiments is a testament to the need
+for increased flexibility.
+
+Linear models also ignore prior knowledge about the system. Combined FMRI
+CBF or CBV imaging methods are available and could shed much light on
+neural activation; however there is no way of combining that data into a
+linear model. Although the most advanced form of the general linear model includes
+a Hemodynamic Response Function, it is static across every region of the brain;
+yet it is well known that capillary beds are not uniform and so blood perfusion 
+cannot possibly be static across the brain. Its possible that activation is
+every bit as strong in regions that aren't found by SPM, but that there are
+nonlinearities that do not match the canonical HRF; resulting in signficant bias error. 
+It is not uncommon
+for data to be thrown out in FMRI studies because no significant activation
+has been seen. This practice highlights the fact that strictly linear
+approaches are insufficiently flexible to account for relatively common
+variations of neural activation.
+
+<Image with two different $\alpha$s>
+
+<image comparing the results of 10\% changes in various signals>
+
+In general activation detection type methods also don't have the ability 
+to find pathologies based on state variables or parameters.  It
 is quite possible that physical properties such as decreased compliance of
@@
 seen in a T1 or T2 map. In essence, this could make FMRI a much more 
-useful clinical tool than it is now. The other problem with linear models
-is that they are a linear fit to a nonlinear signal. It is not uncommon
-for data to be thrown out in FMRI studies because no significant activation
-has been seen. However, if, for whatever reason, the BOLD response
-was acting more nonlinear than in other patients it would be completely
-possible for SPM to miss that activation.
-
-<Image with two different $\alpha$s>
-
-<image comparing the results of 10\% changes in various signals>
-
-Secondly, these methods are still based on t-tests, which notoriously
-lack robustness to non-Gaussian noise. While different techniques 
-exist for imposing Gaussianity, those techniques are incapable of
-discriminating noise from signal. There is no way to know how much 
-signal is removed by various smoothing techniques, or even if entire 
-regions have been smoothed into oblivion. Instead of extensively filtering data to 
-remove noise, the analysis method itself must be robust a wide range of noise,
-which is why we propose here the use of particle filters.
+useful clinical tool than it is now.
+
+Finally, all these techniques require the noise to be Gaussian to reach
+an optimal solution. In fact there is no known optimal solution to
+nonlinear models with non-Guassian noise, so obviously some assumptions
+are going to have to be made to reach a solution. However, it would be
+nice to have an algorithm that is robust to these effects, and could
+still give a good solution when Gaussianity is violated.
 

### 2010-06-23 doc/thesis/thesis.tex
added some bullets
more work on the thesis.tex, kalman filter stuff

@@
 \usepackage{times}
-%\usepackage{tikz}
+\usepackage[siunitx]{circuitikz}
 %\usetikzlibrary{trees}
@@
 
+\section{Unscented Kalman Filter}
+The unscented kalman filter is a powerful bayes-gaussian filter that attempts
+to model the posterior distribution of dynamical systems as a multivariate
+Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
+Filter by allowing the state update to be a function, $g$,
+
+\begin{eqnarray}
+X(t) = g(u(t), X(t-1))\\
+Y(t) = h(X(t)
+\end{eqnarray}
+
+In order to estimate the posterior at $t$, a deterministic set of points (often 2 
+per dimension, plus 1 at the mean)
+weighted according to a Gaussian estimate of $X(t-1)$ are passed through
+the update equation. This set of points are then used to estimate the 
+mean and covariance of $X(t)$. The benefit of this, is that it requires
+no Jacobian and only a few extra calculations to get a decent esimate of
+a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
+closed form solution, and finding the Jacobian is impossible without approximations. Although 
+\cite{Riera2004} claims to use a Jacobian for their solution, they are
+in fact using the Jacobian of $\dot{s}$, $\dot{q}$, $\dot{f}$, $\dot{v}$,
+rather than $s$, $q$, $f$, and $v$. This is important because the Extended
+Kalman filter depends on the jacobian to map a gaussian through the 
+advancement of time. Thus the popular Extended Kalman Filter won't work
+in this case, whereas the Unscented Kalman Filter still does.
+
+The difficulty with the Kalman filter is that it is not well suited
+to parameter estimation. If we wanted to find the mean of a static
+parameters, such as $V_0$, it would be necessary to guess a mean 
+and variance for all time. This is because the Kalman filter is primarily
+concenrned with estimating the output, and is not really concerned
+with estimating the underlying state, $X(t)$. As such estimated 
+output, $\hat{Y}(t)$, is simply a weighted sum of the measurement,
+$Y(t)$, and the expected value of the output based on the propagated
+state variables, $\hat{X}(t)$. As such, when the  the filter will attenuate measurements
+the vastly disagree with $E[h(\hat{X}(t))]$, but eventually
+
+The difficulty
+of using a Kalman Filter, however, is that it assumes a multivariate 
+Gaussian for the state variables, $X(t-1)$. The problem with this is that 
+when a system is significantly nonlinear, the state at 
+$X(t)$ will almost certainly be non-Gaussian, and thus estimating
+$X(t+1)$ with $X(t)$ as a multivariate Gaussian will perpetually introduce
+error in the distribution. Furthermore, it is not really known what 
+sort of underlying distributions may exist in such a mixed biological,
+mechanical, chemical system such as the brain. Assuming the parameters
+all to be Gaussian may in fact be a gross error. On the other hand, if
+the our tests indicate a Gaussian distribution is a good fit, it would
+be a wise plan to go back and consider the use of a modified Unscented 
+Kalman Filter rather
+than the full Particle Filter we discuss in the next section.
+
 \section{Proposed Approach}
@@
 that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
-Considering that $y$ contains measurement noise and noise in $x$ can drive
-changes in $y$, it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
+Considering that $y$ contains measurement noise as well as noise from $x$,
+ it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
 but a true posterior. 
-To begin with, the particle filter starts with a proposal distribution, and $N_p$
+To begin with, the particle filter starts with a prior distribution, and $N_p$
 particles need to be drawn from that distribution, $\alpha(x)$:
@@
 
+\section{Simple Example}
+So a simple example may be used to demonstrate the proposed use of a particle filter.
+Thus, let us consider a simple half wave rectifier circuit, shown in figure 
+\ref{fig:HalfWaveRectifier}
+
+\begin{figure}
+\begin{center}
+\begin{circuitikz}[scale=2, american]
+\draw
+ (0,0)  node[transformer core] (T) {}
+ (T.A1) -- (-1,0)
+ (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
+ (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
+ (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
+ (T.base) node {G};
+\end{circuitikz}
+\label{fig:HalfWaveRectifier}
+\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
+gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
+$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
+\end{center}
+\end{figure}
+
+Thus when $u(t)G$ is less than $v_t$, the circuit will discharge the capacitor,
+but when $u(t)G$ is greater than $v_t$, the capacitor will charge up. Unlike
+the normal rectifier circuit, we'll deal with a generic input
+Let $V_C$ be the
+
 \section{Methods}
@@
 
+For simulated data, single timeseries:
+
 For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
@@
 \begin{enumerate}
-\item Comparison with a linear system with similar number of degrees of freedom
 \item Ground truth vs. Estimated signal during particle filter run
@@
 \item True Parameters vs. Final Parameter Sets
-\item Variance of final parameters when faced with same ground truth, different noise
 \item Variance of final parameters when faced with same ground truth, different noise
@@
 \end{enumerate}
+
+For Simulated Data, Full Volume:
 

### 2010-06-23 doc/thesis/thesis.tex
sort of done with example circuit particle filter.

@@
 \section{Unscented Kalman Filter}
-The unscented kalman filter is a powerful bayes-gaussian filter that attempts
+The unscented kalman filter is a powerful gaussian/bayes filter that attempts
 to model the posterior distribution of dynamical systems as a multivariate
@@
 
-In order to estimate the posterior at $t$, a deterministic set of points (often 2 
-per dimension, plus 1 at the mean)
+In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
+(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
 weighted according to a Gaussian estimate of $X(t-1)$ are passed through
@@
 
-The difficulty with the Kalman filter is that it is not well suited
-to parameter estimation. If we wanted to find the mean of a static
-parameters, such as $V_0$, it would be necessary to guess a mean 
-and variance for all time. This is because the Kalman filter is primarily
-concenrned with estimating the output, and is not really concerned
-with estimating the underlying state, $X(t)$. As such estimated 
-output, $\hat{Y}(t)$, is simply a weighted sum of the measurement,
-$Y(t)$, and the expected value of the output based on the propagated
-state variables, $\hat{X}(t)$. As such, when the  the filter will attenuate measurements
-the vastly disagree with $E[h(\hat{X}(t))]$, but eventually
+\begin{figure}
+\label{fig:EKFWorking}
+\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun}}
+\end{figure}
 
@@
 Kalman Filter rather
-than the full Particle Filter we discuss in the next section.
+than the full Particle Filter we discuss in the next section. These are 
+non-trivial issues given that the assumption of Gaussianity is what allows
+the UKF to estimate the posterior using only the first and second moments;
+two parameters that don't uniquely describe most distributions.
+
+Another potential problem with the UKF is that typically the sigma points,
+used to estimate the posterior probability, $P(X(t) | X(t-1), u(t))$, are
+located on the main axes. As a result, the covariances are not allowed to 
+be effected by the state transition the same way the variances are. While
+this may be reasonable in low-dimensional systems, high demsional systems
+have a much greater potential for interplay between the variables. While
+this problem is relatively easy to fix (by using more sigma points off
+the main axes), there is a definite cost in complexity.
+
+Ultimately, the nonlinearities of the BOLD model make Gaussian estimates
+too unrealistic for practical use. As \ref{fig:Transformation} shows, even 
+holding all other parameters steady, the nonlinearities due to $q$ and $v$
+alone are enough to give a non-gaussian result from a two variables 
+gaussian input.
+
+\begin{figure}
+
+\label{fig:Transformation}
+\caption{a) A Two-Variables gaussian input of $q$ and $v$, b) The 
+resulting distribution after being evolved for two seconds. c) The
+estimated Gaussian based on the mean and covariance of b). d) The
+residual between the two distributions.}
+\end{figure}
 
@@
 a particle filter along with the BOLD model to derive physical 
-parameters. In so doing, we hope to be able to show that neuronal
-efficacy, $\epsilon$ is a suitable variable for estimating voxel 
+parameters. In so doing, we hope to be able to show that one or more
+parameters are a suitable replacement for estimating voxel 
 activation from a standard FMRI image. We also hope to show that 
@@
 
-Although not generally necessary for particle filters, we will make a few
+Although not strictly necessary for particle filters, we will make a few
 assumptions based on the particular type of systems faced in biological 
@@
 seconds, when unfrozen the system would continue as if nothing happend. 
-Few biological systems are predictible enough for them to be summarized
-by a time varying function. Although the heart may seem like an obvious
-exception, period between heartbeats vary often enough that prediction
-would necessate another state-space model. In short, we
+Few biological systems are predictable enough for them to be summarized
+by a time varying function. Even systems that one might assume work in that
+way are actually much more complicated. The heart may seem like an obvious
+exception, however, the period between heartbeats vary often enough that prediction
+would necessitate another state-space model. In short, we
 assume no parameters are time varying, because not enough information exists to
@@
 Third, we will assume noise is additive, and that $\nu_x$ may be projected into
-a weiner, or other summing process that is additive with $g$ and $\nu_y$, which
+a Weiner, or other summing process that is additive with $g$ and $\nu_y$, which
 will be named $\nu_d$.
@@
 
-\section{Simple Example}
-So a simple example may be used to demonstrate the proposed use of a particle filter.
+\section{Simple, Nonlinear Example}
+A simple example may be used to demonstrate the proposed use of a particle filter.
 Thus, let us consider a simple half wave rectifier circuit, shown in figure 
@@
  (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
- (T.base) node {G};
+ (T.base) node {G}
+ (T.B1) to[open, *-*, v=$V_1$] (T.B2); 
 \end{circuitikz}
@@
 
-Thus when $u(t)G$ is less than $v_t$, the circuit will discharge the capacitor,
-but when $u(t)G$ is greater than $v_t$, the capacitor will charge up. Unlike
-the normal rectifier circuit, we'll deal with a generic input
-Let $V_C$ be the
+The half wave rectifier circuit smoothes the gaps between high voltage
+with a capacitor. Thu, when $u(t)G$ is less than $v_t$, the circuit will 
+discharge the capacitor and maintain a non-zero voltage,
+but when $u(t)G$ is greater than $v_t$, the output voltage will be set
+by $u(t)G$ and the capacitor will charge up. We will assume a very simple
+model for all the components, ignoring the complex nonlinear behavior
+that can occur in a diode. Although rectifiers are typically thought
+of as receiving an AC circuit 60 Hz, we will ignore such specifics and 
+assume the voltage across the output of the transformer is simply a scalar
+multiple of the input voltage.  As discussed in the \ref{Sec:ParticleFilter} any
+variable with uncertainty must be part of the state variable. Therefore
+the state variable will be: $X(t) = \{G, v_t, C, R_m, v_y\}$. Of course,
+$u(t)$ cannot be allowed to be a square wave in such a system, since that
+such a signal would never get across the transformer and regardless it
+would necessitate an unrealistic infinite current across the capacitor.
+The state equations would then be
+
+\begin{equation}
+v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
+        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
+        v_y(t-1)\left(1 - \frac{1}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
+    \end{cases} 
+\end{equation}
+
+To run the particle filter is relatively easy then, since there exists 
+a recursive definition of the dynamic state variable, $v_y$. To start
+with, an initial distribution must be assumed and while at first a 
+Gaussian seems like a good idea, all the static state variables are strictly
+positive and thus not well suited to the Gaussian. In this case then,
+it would be wise to start with a Gamma distribution, and just be wary of
+any standard deviation that gets larger than the prior mean. We will define
+the gamma distribution as follows:
+
+\begin{equation}
+X \sim Gamma(k, \theta) \rightarrow f(x) = x^{k-1}\frac{e^{-x/\theta}}{\theta^k\Gamma(k)}
+\end{equation}
+
+where $\Gamma$ is the gamma function.
+The in some ways the margin for error is decided by the weighting function, which
+here will define as $W(V_y, v_y)$, where $V_y$ is the actual measurement, and 
+$v_y$ is the estimate by a particular particle. The choice of this function is difficult,
+and although the Gaussian is typically used, we found the exponential helpful
+in dealing with particle deprivation.  The algorithm will then look like the following,
+
+\begin{algorithmic}
+\STATE Initialize $N_p$ Particles
+\FOR{$i$ in $N_p$}
+    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
+    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
+    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
+    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
+    \STATE $v_y = 0$, (Assume the system has been off for a long time)
+    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
+    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
+\ENDFOR
+\FOR{$t$ in Set of Measurement Times}
+    \FOR{$i$ in $N_p$}
+        \STATE $v_y(t) = f(v_y(t-1, u(t))$
+        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
+    \ENDFOR
+\ENDFOR
+
+\end{algorithmic}
 

### 2010-06-24 doc/thesis/thesis.tex
more work on the exmaples, background knowledge, particle filter.

@@
 
+In order to characterize the noise in the FMRI we took a few 
+sample time-series during resting state, which should theoretically be all noise and 
+used Q-Q plots to analyze the type of noise present. \ref{fig:QQDC} shows the 
+results of a Q-Q plots between the samples and a Gaussian of mean and variance based
+on the samples. Because it the noise is often considered to have Weiner noise with
+Gaussian steps, \ref{fig:QQDelta} shows a Q-Q plot using the steps, rather than 
+absolute values.
+\begin{figure}
+\label{fig:QQDC}
+\caption{Q-Q Plots of resting state data}
+\end{figure}
+
+\begin{figure}
+\label{fig:QQDelta}
+\caption{Q-Q Plots of resting state data, using the steps as the samples}
+\end{figure}
+
+Finally, to determine whether much of the autocorrelation was removed by taking
+steps we plotted the autocorrelation of each sample location, before for the
+actual value (\ref{fig:CorrDC}), and the steps between values (\ref{fig:CorrDelta}).
+Ideally, if the steps are Gaussian, they should be independent and have a
+flat autocorrelation; although in real life this is far too much to expect.
+
+\begin{figure}
+\label{fig:CorrDC}
+\caption{Autocorrelation of the absolute value of various time-series}
+\end{figure}
+
+\begin{figure}
+\label{fig:CorrDelta}
+\caption{Autocorrelation of the steps of various time-series}
+\end{figure}
+
+<conclusion about the noise>
+
+\subsection{Analysis of the BOLD Signal}
+Several studies have endeavored to analyze the BOLD equations, which
+we will review here. The most obvious analysis is to calculate the
+steady state signal, when $u$ is held constant long enough for
+all the transients to settle out.
+
+\begin{eqnarray}
+s_{ss} &=& 0 \nonumber \\
+f_{ss} &=& \tau_f\epsilon u + 1\nonumber \\
+v_{ss} &=& (\tau_f\epsilon u + 1)^\alpha\nonumber \\
+q_{ss} &=& \frac{(\tau_f\epsilon u + 1)^\alpha}{E_0}(1-(1-E_0)^{1/(\tau_f\epsilon u + 1)})\nonumber \\
+y_{ss} &=& V_0((k_1+k_2)(1-q_{ss}) - (k_2+k_3)(1-v_{ss}))
+\label{steadystate}
+\end{eqnarray}
+
+[Image with two different $\alpha$'s]
+
+[image comparing the results of 10\% changes in various signals]
+
 \section{Current Techniques}
@@
 sum of the various input timeseries. Essentially every experimental factor
-is considered as another level inputs (ex. which patient is an input). 
+is considered as another level inputs (ex. different type of stimulus,
+different run with the same patient, different patient). 
 The equation for a general linear model is then
@@
 and $\epsilon$ is the error. Thus for every time, the measurement is
-assumed to be a weighted sum of the inputs plus some error. The calculation
-of $\beta$ then performed using a maximum likelihood or gradient descent search 
+assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
+of $\beta$ is then performed using a maximum likelihood or gradient descent search 
 to minimize the error.
@@
 in the activation of brain regions. Thus, various methods are used to 
-smooth $X(t)$'s related to actual input (as opposed to experimental confounds
-such as patient number, age etc) through time, and bandlimit the input. 
-The best technique is convolving the stimulus input with a hemodynamic 
-response function, which mimicks the basic shape of BOLD activation, including a delay
+smooth the columns of $X$, and thus bandlimit the input. 
+The best technique is convolving the stimulus input with a Hemodynamic 
+Response Function (HRF), which mimics the basic shape of BOLD activation, including a delay
 due to rise time and fall time. The downside of this method, however is that 
-the hemodynamic signal is static, meaning the same hemodynamic function is
-used for every region of the brain, and $Y(t)$ must be a constant
-scalar multiple of $X(t)$ to be appropiately identified.
+the shape of the Hemodynamic Signal is static, meaning the same Hemodynamic Function is
+used for every region of the brain, and $Y(t)$ must linear combination 
+of scalar the columns of $X$ to be appropriately identified. Additionally, 
+it is well known that different Hemodynamic Response Functions are necessary for different 
+regions of the brain. The "Canonical" HRF that is most used, has been fitted
+for the visual cortex. Thus the shape of the HRF may differ significantly in
+terms of onset and fall time in other areas of the brain. The inability to
+fit parameters other than scale certainly hinders SPM's
+ability to locate voxels that do not conform to the "canonical" hemodynamic 
+response function. If there are different HRF's for different regions 
+of the brain, might there also be more subtle differences even within
+a particular region? It seems reasonable to think so. As a result, the 
+most common use of SPM will be heavily biased toward the visual cortex.
 
@@
 that many of these dependencies are nonlinear, which means they may be
-difficult or impossible to detect using linear techniques. Additionally, 
-the lack of flexibility in delay variables certainly hinders SPM's
-ability to locate voxels that do not conform to the canonical hemodynamic 
-response function. The very fact that different hemodynamic response functions
-work better for different regions/experiments is a testament to the need
-for increased flexibility.
-
-Linear models also ignore prior knowledge about the system. Combined FMRI
-CBF or CBV imaging methods are available and could shed much light on
-neural activation; however there is no way of combining that data into a
-linear model. Although the most advanced form of the general linear model includes
-a Hemodynamic Response Function, it is static across every region of the brain;
-yet it is well known that capillary beds are not uniform and so blood perfusion 
-cannot possibly be static across the brain. Its possible that activation is
-every bit as strong in regions that aren't found by SPM, but that there are
-nonlinearities that do not match the canonical HRF; resulting in signficant bias error. 
-It is not uncommon
-for data to be thrown out in FMRI studies because no significant activation
-has been seen. This practice highlights the fact that strictly linear
-approaches are insufficiently flexible to account for relatively common
-variations of neural activation.
-
-<Image with two different $\alpha$s>
-
-<image comparing the results of 10\% changes in various signals>
+difficult or impossible to detect using linear techniques. This presents
+a significant and unknown problem that is often left un-addressed in 
+neural studies.
+
+A static linear model is also unable to incorporate other forms of physiological
+data that may be available. Combined FMRI
+CBF or CBV imaging methods are not uncommon and they could shed much light on
+neural activation. However, there is no way of actively combining that data 
+into a static HRF. The fact of the matter is that the relation between the 
+BOLD signal and CBV/CBF simply cannot be described by any linear relationship.
+While it may be possible to determine some regionally varying set of parameters
+for the HRF based on CBV/CBF, the lack of physiologically inspired parameters
+impairs this. The importance of regional differences in the HRF cannot be
+over-stated; it is well known that capillary beds are far from uniformly distributed
+and thus blood perfusion and regional oxygenation cannot possibly be uniform
+either.
+
+Finally, all these techniques require the noise to be Gaussian to reach
+an optimal solution. In fact there is no known optimal solution to
+nonlinear models with non-Guassian noise, so obviously some assumptions
+are going to have to be made to reach a solution. However, it would be
+nice to have an algorithm that is robust to these effects, and could
+still give a good solution when Gaussianity is violated.
+
+Because of these limitations, its entirely possible that activation exists
+in regions that SPM doesn't detect, but because the activation does not conform
+to the set HRF, it is impossible for these signals to be detected. Perhaps
+because of this, it is not uncommon for data to be thrown out or averaged
+together in FMRI studies because 
+no significant activation could be seen in a single run \cite{Riera2004}
+\cite{Johnston2008}. This practice highlights 
+limitations in strictly linear approaches; and suggests that a single HRF is 
+insufficiently flexible to account for relatively common variations of neural activation.
 
@@
 
-Finally, all these techniques require the noise to be Gaussian to reach
-an optimal solution. In fact there is no known optimal solution to
-nonlinear models with non-Guassian noise, so obviously some assumptions
-are going to have to be made to reach a solution. However, it would be
-nice to have an algorithm that is robust to these effects, and could
-still give a good solution when Gaussianity is violated.
-
 \section{Unscented Kalman Filter}
@@
 \begin{eqnarray}
-X(t) = g(u(t), X(t-1))\\
-Y(t) = h(X(t)
+X(t) &=& g(u(t), X(t-1))\\
+Y(t) &=& h(X(t)
 \end{eqnarray}
@@
 closed form solution, and finding the Jacobian is impossible without approximations. Although 
-\cite{Riera2004} claims to use a Jacobian for their solution, they are
-in fact using the Jacobian of $\dot{s}$, $\dot{q}$, $\dot{f}$, $\dot{v}$,
-rather than $s$, $q$, $f$, and $v$. This is important because the Extended
-Kalman filter depends on the jacobian to map a gaussian through the 
+\cite{Riera2004}, \cite{Hu2009} mention a Jacobian of the BOLD response, this is
+not strictly the case and is rather $\frac{\partial J}{\partial t}$
+rather than a true Jacobian.  This is important because the Extended
+Kalman filter depends on the Jacobian to map a Gaussian through the 
 advancement of time. Thus the popular Extended Kalman Filter won't work
-in this case, whereas the Unscented Kalman Filter still does.
+in this case, whereas the Unscented Kalman Filter still does. In fact
+\cite{Hu2009} uses the UKF to perform a similar type of analysis to
+the one performed in this work.
 
@@
 \end{figure}
+
+\subsection{Conclusion}
 
@@
 \subsubsection{Resampling}
+\label{Sub:Resampling}
 As a consequence 
@@
 with a maximum variance of twice the original variance. All the assumptions made originally
-for the particle filter hold, and all of the parameters may be distringuished based on
+for the particle filter hold, and all of the parameters may be distinguished based on
 the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
@@
 
-<frequncy response graphs, highlighting noise frequency range and signal frequency range>
+<frequency response graphs, highlighting noise frequency range and signal frequency range>
 
 \subsubsection{Weighting Function}
-Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting funciton,
+Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting function,
 is based on an unknown distribution, it is necessary to decide on a function
 that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
-needs to centered at zero and have a scale comparable to the signal levels.
-While a Gaussian function is the natural choice, we also wanted to try a distribution
-with wider tails, so that outliers don't completely destroy particle's weights.
-Therefore, we tried three weighting functions; Gaussian, exponential and 
-the Cauchy distribution. The standard deviations (or scale) was set to 
-$\sigma_y/5$, where $\sigma_y$ is the variance of all $y[0:\infty]$. Of course 
-since the particle filter requires a weighting function to run, this means that
-before the particle filter starts all the measurements have to be in. In cases where
-this is impossible, a heuristic based on a small sample may work just as well.
-
-The shape of the weighting function is extremely important, because it essentially decides
-the rejection rate of particles. A very thin gaussian probability 
-distribution function has nice properties, but thin tails. As a result, large
-outliers in the measurement vector could easily force all the particles to
-have near 0 weights, thus forcing the particle filter to converge improperly. 
-On the other end of the spectrum, a cauchy PDF, has relatively fat tails, and
-may not weight central particles high enough, preventing the particles from 
-converging at a reasonable rate. Exponetials have the benefit of having an 
-extremely smooth drop to zero, and a slope of 1 at the origin. Having a non-zero
-slope at the origin is beneficial because it discriminates all the way up until
-the measured and predicted $y$ are the same. The importance of the weighting function
-cannot be overstated, as this is the primary factor in deciding the rate at
-which the particle filter converges.
-
-Obviously the optimal $\hat{Pr}(y(T) | x(T))$ is the true $Pr(y(T) | x(T))$; however
-since that is unknown, I tested multiple different distributions. As stated previous,
-because the exact scale of $Pr(y(T) | x(T))$ it not even known, the exponential distribution
-has the added benefit of having a negative slope all the way form zeo to infinity. Thus
-even if the scale is completely wrong, particles will still be well differentiated.
-
-<Q-Q plot of real fmri data with Gaussian, DC>
-<Q-Q plot of real fmri data with Gaussian, deltas>
+needs to be centered at zero and have a scale comparable to the signal levels.
+Obviously if the actual noise present in $y(t)$ were to be known, then that would
+be the best distribution for $\Omega$. In that case, particles that fell
+far out on that distribution would be statistically impossible representations
+of the system, and it would be completely reasonable to throw such particles away.
+While a Gaussian function is the natural choice, because this distribution
+and weight are unknown we wanted to try distributions
+with wider tails, so that outliers don't completely destroy particle's weights
+(and thus convergence proceeds more slowly).
+
+Another natural choice might be one of the robust estimator weight functions, for
+example the Huber or bi-square. For the purpose of this work we will stick
+with long tailed distributions, however it is worth noting that long tails
+may not be the optimal choice in all, or even this situation. The justification
+for long tailed distributions is that we believe the noise to be long tailed,
+and the variance of the noise is not well known.
+
+Therefore, we tried three weighting functions based on three distributions: Gaussian, 
+Laplace and the Cauchy. The standard deviation of the distribution is extremely
+important to the convergence of particle filter. A standard deviation that is 
+too large will not allow the distribution to converge in any reasonable number of 
+measurements. A standard deviation below the standard deviation of the noise 
+will cause the algorithm to throw out perfectly acceptable particles. 
+The weighting function ultimately will shape the output distribution, $P[y]$, into that
+distribution, however the distribution of $x$ will still approach a reasonable
+estimate of its true distribution. Even an overly wide weighting function, will
+allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
+parameters of the "real" posterior distribution, though the scale parameters may 
+be overly large.
+
+A reasonable method of setting the standard deviation of $\Omega$ may be by 
+taking a small sample from "resting" data and using the sample standard deviation.
+Since this is the first attempt at using particle filters for modeling the 
+BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
+because it gives a more consistency and control. Of course this could be taken
+further, by testing the sample data against a set of stock distributions
+and choosing the best fit. This of course depends on having enough samples
+to make a reasonable inference, which may not always exist.
 
@@
         u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
-        v_y(t-1)\left(1 - \frac{1}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
+        v_y(t-1)\left(1 - \frac{\delta t}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
     \end{cases} 
@@
 The in some ways the margin for error is decided by the weighting function, which
-here will define as $W(V_y, v_y)$, where $V_y$ is the actual measurement, and 
-$v_y$ is the estimate by a particular particle. The choice of this function is difficult,
+here will define as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
+estimate based on all the particles, and 
+$v_{yi}$ is the estimate by a particular (i$^{\text{th}}$) particle. The choice of this function is difficult,
 and although the Gaussian is typically used, we found the exponential helpful
@@
 \begin{algorithmic}
-\STATE Initialize $N_p$ Particles
+\STATE Initialize $N_p$ Particles:
 \FOR{$i$ in $N_p$}
@@
 \ENDFOR
+\STATE Run the Filter:
 \FOR{$t$ in Set of Measurement Times}
     \FOR{$i$ in $N_p$}
-        \STATE $v_y(t) = f(v_y(t-1, u(t))$
+        \STATE $v_{yi}(t) = f(v_{yi}(t-1), u(t))$
+        \STATE (All other members of $X_i(t)$ remain the same)
         \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
@@
 \ENDFOR
-
 \end{algorithmic}
+
+Initially the particles will have the same output, $0$, however, as $u(t)$
+changes, the response of each particle to that input will result in different
+outputs. Particles that have a $v_{yi}$ near $V_y$ will be weighted higher,
+and others farther away will be weighted lower. As the particle filter
+runs, weights will compound resulting in a distribution that asymptotically
+approaches the true joint distribution of the $X(t)$.  Of course, as we
+mentioned in \ref{Sub:Resampling}, particles weighted zero do not significantly
+contribute to the empirical distribution, so re-sampling may be necessary.
 

### 2010-06-25 doc/thesis/thesis.tex
"finished" kalman filter section

@@
-\documentclass{article}
-\usepackage[left=3cm,top=2cm,right=3cm,nohead,nofoot]{geometry}
+\documentclass{report}
+\usepackage[left=3cm,top=2cm,right=3cm]{geometry}
 
@@
 \usepackage{cancel}
-\usepackage{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 
@@
 \begin{document}
-\section{Introduction}
+\chapter{Introduction}
 Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
@@
 
-\subsection{FMRI}
+\section{FMRI}
 Magnetic Resonance Imaging, MRI is a method of building 3D images
@@
 
-\subsection{BOLD Physiology}
+\section{BOLD Physiology}
+\label{sec:BOLD Physiology}
 It is well known that the two types of hemoglobin act as a contrast agents in 
@@
 explain much of the BOLD signal, there was still a matter of proposing a
-realistic waveform for the CBF and for the CMRO2. \cite{friston2000} gave
+realistic waveform for the CBF and for the CMRO2. \cite{Friston2000} gave
 a reasonable and simple
@@
 worth the decrease in model bias. For now it remains to be seen where this
-line may be drawn in the BOLD signal, although \cite{deneux2006} does not
+line may be drawn in the BOLD signal, although \cite{Deneux2006} does not
 show a significant improvement from the additional parameters added in 
@@
 
-\subsection{Previous Studies of Parameters}
+\section{Previous Studies of Parameters}
 There have been quite a few efforts to quantify the parameters of the
@@
 
-In \cite{Jonston2008}, a hybrid particle filter/gradient
+In \cite{Johnston2008}, a hybrid particle filter/gradient
 descent algorithm was used to simultaneously derive the static (classicly called
@@
 
-\subsection{Noise}
+\section{Noise}
 Thus, despite some discrepencies, the cause of the BOLD signal is
@@
 Additionally, since it is standard operating procedure to perform
-coregistration between volumes, it is unlikely that movement was
+co-registration between volumes, it is unlikely that movement was
 the cause of the drift either (otherwise the effect would not be 
@@
 sample time-series during resting state, which should theoretically be all noise and 
-used Q-Q plots to analyze the type of noise present. \ref{fig:QQDC} shows the 
-results of a Q-Q plots between the samples and a Gaussian of mean and variance based
-on the samples. Because it the noise is often considered to have Weiner noise with
-Gaussian steps, \ref{fig:QQDelta} shows a Q-Q plot using the steps, rather than 
-absolute values.
+used Q-Q plots to analyze the type of noise present. Additionally, the autocorrelation
+was also calculated since most methods (including the one used in this paper)
+assume the noise is independent from time and
+identically distributed.  If these assumptions fail, then error rate will
+be higher than expected. \autoref{fig:QQDC} shows the 
+results with a regression line fit to the points. A set of points that are exactly 
+Gaussian should all fall on the same line.  Because it the noise is often considered 
+to have Weiner noise with Gaussian steps, as first described in \cite{Riera2003}, we
+also performed the same tests again, but on the steps, rather than the 
+direct measurements. \autoref{fig:QQDelta} shows the analysis of the same
+set of time-series as \autoref{fig:QQDC}, after being converted to steps.
+Finally, as we will discuss in \autoref{sec:Methods Preprocessing}, some
+amount of drift can be removed by subtracting off a spline from the timeseries.
+
 \begin{figure}
+\caption{Q-Q Plots of resting state data}
 \label{fig:QQDC}
-\caption{Q-Q Plots of resting state data}
 \end{figure}
@@
 \begin{figure}
+\caption{Q-Q Plots of resting state data, using the steps as the samples}
 \label{fig:QQDelta}
-\caption{Q-Q Plots of resting state data, using the steps as the samples}
 \end{figure}
 
-Finally, to determine whether much of the autocorrelation was removed by taking
-steps we plotted the autocorrelation of each sample location, before for the
-actual value (\ref{fig:CorrDC}), and the steps between values (\ref{fig:CorrDelta}).
-Ideally, if the steps are Gaussian, they should be independent and have a
-flat autocorrelation; although in real life this is far too much to expect.
-
-\begin{figure}
-\label{fig:CorrDC}
-\caption{Autocorrelation of the absolute value of various time-series}
-\end{figure}
-
-\begin{figure}
-\label{fig:CorrDelta}
-\caption{Autocorrelation of the steps of various time-series}
-\end{figure}
-
 <conclusion about the noise>
 
-\subsection{Analysis of the BOLD Signal}
+\section{Analysis of the BOLD Signal}
 Several studies have endeavored to analyze the BOLD equations, which
@@
 y_{ss} &=& V_0((k_1+k_2)(1-q_{ss}) - (k_2+k_3)(1-v_{ss}))
-\label{steadystate}
+\label{eq:steadystate}
 \end{eqnarray}
@@
 
-\section{Current Techniques}
-\subsection{Basic Statistical Parametric Mapping}
+\chapter{Current Techniques}
+\section{Basic Statistical Parametric Mapping}
 Although not strictly the same thing as parameter calculation from 
@@
 <Do I need to mathematically define all this?> The derivation and application
-of random field theory, and its use can be found in various papers \cite{univ_mult_rft}.
-
-\subsection{General Linear Model}
+of random field theory, and its use can be found in various papers \cite{Worsley2004}.
+
+\section{General Linear Model}
 The most used form of FMRI analysis is Statistical Parametric
@@
 
-<Image of GLM>
+[Image of GLM]
 
@@
 
-<Image of Hemodynamic Response Function>
+[Image of Hemodynamic Response Function]
 
@@
 \begin{figure}
+\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun2005}}
 \label{fig:EKFWorking}
-\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun}}
 \end{figure}
@@
 mechanical, chemical system such as the brain. Assuming the parameters
-all to be Gaussian may in fact be a gross error. On the other hand, if
-the our tests indicate a Gaussian distribution is a good fit, it would
-be a wise plan to go back and consider the use of a modified Unscented 
-Kalman Filter rather
-than the full Particle Filter we discuss in the next section. These are 
+all to be Gaussian may in fact be a gross error. On the other hand, for
+small variances and short time steps the gaussiann distribution is a good 
+fit, and so in some limited cases the Unscented Kalman Filter could work
+well. These are 
 non-trivial issues given that the assumption of Gaussianity is what allows
@@
 two parameters that don't uniquely describe most distributions.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c || c |}
+\hline 
+Parameter & Run 1 \\
+\hline
+$\tau_0$ & .98  \\
+$\alpha$ & .33 \\
+$E_0$ & .34  \\
+$V_0$ & .03  \\
+$\tau_s$ & 1.54  \\
+$\tau_f$ & 2.46  \\
+$\epsilon$ & .54  \\
+$V_t$ & N(1, .3)  \\
+$Q_t$ & N(1, .3)  \\
+$S_t$ & N(1, .3) \\
+$F_t$ & N(1, .3) \\
+\hline
+\end{tabular}
+\caption{Parameters used to test Gaussianity of variables after being transitioned through
+the BOLD model}
+\label{tab:steptable} 
+\end{table}
+
+To determine the amount of error incurred in a Gaussian estimate during
+a typical sample period, the states of BOLD equations were assigned according
+to four dimensional Gaussian. The states were then propagated through two
+seconds of simulation (a typical TR in FMRI) and then the resulting
+marginal distributions were compared with a Gaussian distribution. The 
+purpose is to determine the degree to which the results of simulation will
+result in non-Gaussian output, given a Gaussian input. This also demonstrates
+the degree of nonlinearity present in the system. The parameters used are 
+shown in \autoref{tab:steptable}
+
+Notably $s_t$ has intentionally been set to a non-equilibrium, but physiologically
+plausible value. The value of $u$ is left at zero the entire time, so the 
+system will decay naturally (see \autoref{sec:BOLD Physiology}), though initializing 
+$s$ at a non-zero level will drive the system for several seconds. 
+\autoref{fig:transp1s} shows the results when the system is essentially
+left on for 100 milliseconds after setting the variables according to \autoref{tab:steptable}.
+The Q-Q plots fit very will with a Gaussian, demonstrating that at this short
+time interval nonlinearities have not yet begun to effect the distribution.
+However, \autoref{fig:trans1s} us the result after 1 second, which is faster
+than most FMRI scanners are capable of sampling at. At that range the tails 
+of the distributions for $v$ and $q$ are clearly starting to deviate from the
+Gaussian distribution. As a result the uncertainty in $y$ is deviating from
+the Gaussian distribution as well. This is important, because although 
+approximating the distribution with a Gaussian based on the first two
+moments will work in the short run, there will be residual error in the
+distribution.
+
+On the other hand, this effect is more limited if the initial variance is somewhat
+smaller. In tests with those cases, it took much longer for the nonlinearities to skew
+the distribution. That result could be encouraging to those looking to use the
+UKF, if the distributions are kept relatively thin.
 
@@
 be effected by the state transition the same way the variances are. While
-this may be reasonable in low-dimensional systems, high demsional systems
+this may be reasonable in low-dimensional systems, high dimensional systems
 have a much greater potential for interplay between the variables. While
@@
 
-Ultimately, the nonlinearities of the BOLD model make Gaussian estimates
-too unrealistic for practical use. As \ref{fig:Transformation} shows, even 
-holding all other parameters steady, the nonlinearities due to $q$ and $v$
-alone are enough to give a non-gaussian result from a two variables 
-gaussian input.
+Ultimately, there is a distinct possibility that the nonlinearities of 
+the BOLD model make Gaussian estimates unrealistic and thus less effective.
+More advanced tests where static variables such as $\alpha$ are varied
+as well could shed even more light on the issue. The trouble with using
+the UKF then to estimate parameters is that all eleven members of $X$ would
+be treated like a single joint Gaussian distribution which certainly 
+compound the issues of nonlinearity seen in \autoref{fig:transp1s} and 
+\autoref{fig:trans1s}
 
 \begin{figure}
-
-\label{fig:Transformation}
-\caption{a) A Two-Variables gaussian input of $q$ and $v$, b) The 
-resulting distribution after being evolved for two seconds. c) The
-estimated Gaussian based on the mean and covariance of b). d) The
-residual between the two distributions.}
+\centering
+\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
+\caption{Distributions of state variables after simulating for .1s}
+\label{fig:transp1s}
 \end{figure}
 
-\subsection{Conclusion}
-
-\section{Proposed Approach}
-\subsection{Goal}
-The ultimate goal of this project is to provide a new set of tools
-for analyzing FMRI data. Whereas SPM techniques have been highly 
-successful at finding macroscopic regions of activation, linear 
-modeling can carry significant bias error due to lack of model
-flexibility. While adding parameters can significantly increase
-error due to model variance, this effect is mitigated by the fact
-that we plan to use a model that is based on first principals. The
-purpose of this paper is thus to evaluate the potential of using
-a particle filter along with the BOLD model to derive physical 
-parameters. In so doing, we hope to be able to show that one or more
-parameters are a suitable replacement for estimating voxel 
-activation from a standard FMRI image. We also hope to show that 
-estimated posterior distribution of the parameters, derived from
-the particle filter, is able to provide an accurate measure of the
-confidence interval.
-
-\subsection{Introduction to Particle Filters}
+\begin{figure}
+\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
+\caption{Distributions of state variables after simulating for 1s}
+\label{fig:trans1s}
+\end{figure}
+
+\section{Conclusion}
+These other methods ....
+
+\chapter{Particle Filters}
+\section{Introduction}
 Particle filters, a type of Sequential Monte Carlo (SMC) methods
@@
 tracking for non-linear, non-gaussian systems.
-%todo: might need to reference attempts to model a system of our type with
-%unscented kalman filter
-\subsubsection{Model}
+
+\subsection{Model}
+\label{sec:Particle Filter Model}
 The idea of the particle
@@
 \begin{equation}
-\label{stateass}
 \dot{x}(t) = f(x(t), u(t))
-\end{equation}
-
-\begin{equation}
-\label{measass}
+\label{eq:stateass}
+\end{equation}
+
+\begin{equation}
 y(t) = g(x(t)) + \nu_y + \nu_d
+\label{eq:measass}
 \end{equation}
@@
 have biological and non-biological sources. MR can lead to both types of noise, 
-as demonstrated in \cite{drift}. Meanwhile changes in metabolism, heart rate, or
+as demonstrated in \cite{Smith1999}. Meanwhile changes in metabolism, heart rate, or
 other biochemical intervention could all lead to either $\nu_d$ or $\nu_y$.
 
-\subsubsection{Prior}
+\subsection{Prior}
 
@@
 
-\subsubsection{Weighting}
+\subsection{Weighting}
 When a measurement becomes available it is incorporated into the probability.
@@
 \begin{equation}
-\label{weightfunc}
 w_i(T) \propto \frac{\hat{Pr}(x_i[0:T] | y[0:T])}{q(x_i[0:T] | y[0:T])}
+\label{eq:weightfunc}
 \end{equation}
@@
 
-Using the assumption that $y(t)$ is fully constrained by $x(t)$ \eqref{measass},
-and that $x(t)$ is fully constrained by $x(t-1)$ \eqref{stateass}, we are able to
+Using the assumption that $y(t)$ is fully constrained by $x(t)$ \autoref{eq:measass},
+and that $x(t)$ is fully constrained by $x(t-1)$ \autoref{eq:stateass}, we are able to
 make the reasonably good assumptions that:
@@
 
-Plugging these simplifications into \eqref{weightfunc} leads to:
+Plugging these simplifications into \autoref{eq:weightfunc} leads to:
 \begin{eqnarray}
-\label{weightevolve}
 w_i(T) & \propto & \frac{\hat{Pr}(y(T)| x(T)) \cancel{\hat{Pr}(x(T) | x(T-1))} \hat{Pr}(x[0:T-1] | y[0:T-1])}
@@
 & \propto & w_i(T-1)\hat{Pr}(y(T)| x(T)) 
+\label{eq:weightevolve}
 \end{eqnarray}
@@
 
-\subsubsection{Basic Particle Filter Algorithm}
+\subsection{Basic Particle Filter Algorithm}
 From the definition of $w_i$, the algorithm sequential importance sampling (SIS) 
@@
 
-\subsubsection{Resampling}
-\label{Sub:Resampling}
+\subsection{Resampling}
+\label{sec:Particle Filter Resampling}
 As a consequence 
@@
 \begin{equation}
-\label{neff}
 \hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
+\label{eq:neff}
 \end{equation}
@@
 
-    \STATE Calculate $N_{eff}$ with \eqref{neff}
+    \STATE Calculate $N_{eff}$ with \autoref{eq:neff}
     \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
@@
 
-\subsection{Choosing $\hat{Pr}(y(T) | x(T))$}
+\chapter{Proposed Approach}
+\section{Goal}
+The ultimate goal of this project is to provide a new set of tools
+for analyzing FMRI data. Whereas SPM techniques have been highly 
+successful at finding macroscopic regions of activation, linear 
+modeling can carry significant bias error due to lack of model
+flexibility. While adding parameters can significantly increase
+error due to model variance, this effect is mitigated by the fact
+that we plan to use a model that is based on first principals. The
+purpose of this paper is thus to evaluate the potential of using
+a particle filter along with the BOLD model to derive physical 
+parameters. In so doing, we hope to be able to show that one or more
+parameters are a suitable replacement for estimating voxel 
+activation from a standard FMRI image. We also hope to show that 
+estimated posterior distribution of the parameters, derived from
+the particle filter, is able to provide an accurate measure of the
+confidence interval.
+
+\section{Choosing $\hat{Pr}(y(T) | x(T))$}
 Choosing a representation of an unknown distribution is certainly tricky,
@@
 
-\subsubsection{Classical De-trending}
+\subsection{Classical De-trending}
 The non-stationary
 aspect of a Weiner process as with $\nu_d$ is difficult to compensate for, and so various methods
-have been developed to compensate for it. \cite{detrend} and \cite{drift} have
+have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
 demonstrated that this component is prevalent, and may in fact be a characteristic
@@
 signal. There is no shortage of theories on the "best" method of detrending, however
-a head to head comparison, \cite{detrend}, showed that in most cases subtracting off
+a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
 a spline works the best. The benefit of the spline versus wavelets, high pass 
@@
 
-\subsubsection{Delta Based Inference}
+\subsection{Delta Based Inference}
 I also propose and test a different method of dealing with the so called "drift". 
@@
 \begin{equation}
-\label{measass}
 \Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
+\label{eq:measass_delta}
 \end{equation}
@@
 
-\subsubsection{Weighting Function}
+\subsection{Weighting Function}
 Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting function,
@@
 \section{Simple, Nonlinear Example}
-A simple example may be used to demonstrate the proposed use of a particle filter.
-Thus, let us consider a simple half wave rectifier circuit, shown in figure 
-\ref{fig:HalfWaveRectifier}
+A typical half wave rectifier takes a AC voltage circuit and removes
+one half (say the negative half) of the signal. The resulting waveform
+is still not DC, however it is then possible to use a capacitor to 
+smooth the signal into something similar to DC, as shown in \autoref{fig:HalfWaveIO}.
+There are other, more
+complex circuits that convert the negative portion into positive and
+waste less energy but but here we will keep the system simple.
+Thus, let us consider a simple half wave rectifier circuit, shown in 
+\autoref{fig:HalfWaveRectifier}.
+
+The half wave rectifier circuit smoothes the gaps between high voltage
+with a capacitor. Thus, when $u(t)G$ is less than $v_t$, the circuit will 
+discharge the capacitor and maintain a non-zero voltage,
+but when $u(t)G$ is greater than $v_t$, the output voltage will be set
+by $u(t)G$ and the capacitor will charge up. We will assume a very simple
+model for all the components, ignoring the complex nonlinear behavior
+that can occur in a diode.
 
 \begin{figure}
-\begin{center}
+\centering
 \begin{circuitikz}[scale=2, american]
@@
 \end{circuitikz}
-\label{fig:HalfWaveRectifier}
 \caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
@@
 $C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
-\end{center}
+\label{fig:HalfWaveRectifier}
 \end{figure}
 
-The half wave rectifier circuit smoothes the gaps between high voltage
-with a capacitor. Thu, when $u(t)G$ is less than $v_t$, the circuit will 
-discharge the capacitor and maintain a non-zero voltage,
-but when $u(t)G$ is greater than $v_t$, the output voltage will be set
-by $u(t)G$ and the capacitor will charge up. We will assume a very simple
-model for all the components, ignoring the complex nonlinear behavior
-that can occur in a diode. Although rectifiers are typically thought
+\begin{figure}
+\centering
+\caption{Example Input/Output of the Half Wave Rectifier}
+\label{fig:HalfWaveIO}
+\end{figure}
+
+Although rectifiers are typically thought
 of as receiving an AC circuit 60 Hz, we will ignore such specifics and 
 assume the voltage across the output of the transformer is simply a scalar
-multiple of the input voltage.  As discussed in the \ref{Sec:ParticleFilter} any
-variable with uncertainty must be part of the state variable. Therefore
+multiple of the input voltage.  As discussed in the \autoref{sec:Particle Filter Model}
+any variable with uncertainty must be part of the state variable. Therefore
 the state variable will be: $X(t) = \{G, v_t, C, R_m, v_y\}$. Of course,
@@
 approaches the true joint distribution of the $X(t)$.  Of course, as we
-mentioned in \ref{Sub:Resampling}, particles weighted zero do not significantly
+mentioned in \autoref{sec:Particle Filter Resampling}, particles weighted zero do not significantly
 contribute to the empirical distribution, so re-sampling may be necessary.
 
-\section{Methods}
+\chapter{Methods}
 This paper describes two types of experiments; first we will cover
@@
 
-\subsection{Preprocessing}
+\section{Preprocessing}
+\label{sec:Methods Preprocessing}
 As discussed in the section on de-trending, the normal pipeline for analyzing
@@
 
-\subsection{Simulation}
+\section{Simulation}
 We performed two types of simulations. First, we simulated a single BOLD time-series based
@@
 
-\subsection{Real Data}
+\section{Real Data}
 Finally, we also performed inference based on real FMRI data. The scanner we used...
@@
 
-\section{Results}
-\subsection{Single Time-Series Simulation}
+\chapter{Results}
+\section{Single Time-Series Simulation}
 
@@
 %note to self, epsilon should probably be uniform between 0 and something
-\subsection{Simulated Volume}
+\section{Simulated Volume}
 \begin{enumerate}
@@
 
-\subsection{FMRI Data}
+\section{FMRI Data}
 ....
@@
 
-\section{Conclusion}
+\chapter{Conclusion}
 

### 2010-06-26 doc/thesis/thesis.tex
got q-q plots in

--- 2010-06-23 doc/thesis/thesis.tex
@@
 \usepackage{graphicx}
+\usepackage{subfigure}
 \usepackage{cancel}
-\usepackage{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 
@@
 \begin{document}
-\section{Introduction}
-Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
-of neural activity. Despite its rather limited temporal resolution, FMRI is still
-the best way of measuring neural activity for the majority of the brain.
-Whereas other methods
-of analyzing neural signals can be invasive or difficult to acquire, 
-FMRI is relatively quick and cheap, and its analysis straight forward.
-Because of these benefits, FMRI continues to be crucial to the study of human 
-cognition. Despite its prevalence, there have been relatively few developments
-in the actual analysis of FMRI images. A steady stream of studies have built
-on the original BOLD signal derivation first described in \cite{Ogawa}, 
-from the Baloon model first proposed by \cite{Buxton1998}
-all the way to full fully autonomous system of equations \cite{Riera2004}. And while
-there have been numerous forks in the model, enough in fact to make an entire paper
-studying the differences,\cite{Deneux2006}, it is widely known that all these
-models have quantitatively less bias error than General Linear Model which is
-typically employed today. Then again, depending on the model there may be between
-seven \cite{Riera2004} and 50 \cite{Behzadi2005} parameters per voxel to
-be optimized. Clearly there is a significant risk of error due to variance
-with so many degrees of freedom, not to mention a significantly increased
-computation cost. In this thesis I demonstrate the use of a 
-particle filter as a means of addressing these problems.
+\title{BOLD Parameter Estimation Using Particle Filters}
+\author{Micah Chambers\\Virginia Polytechnic Institute and State University}
+\date{}
+\maketitle
 
-FMRI images as a method of detecting neural activation is based on 
-temporal changes of the Blood Oxygen Level Dependent (BOLD) signal.
-The BOLD signal is caused by minute changes in the ratio of Deoxygenated
-Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain.
-Because Deoxygenated hemoglobin is paramagnetic, higher concentrations
-attenuate the signal when using T2 weighted imaging techniques, such 
-as Echo Planar Imaging (EPI) which is used in FMRI. When axons becomes active,
-a large amount of ions quickly flow out of the cell. In order for the action
-potential to be used again, an active pumping process moves ions back into the
-axon. This process of recharging the axon takes a large amout of enengy, which 
-naturally uses oxygen. On a massive scale (cubic millimiter) this activation/recharge
-process is happening all
-the time; however, it happens at a much higher rate when a portion of
-the brain is very active. Thus, blood vessels in a very active area will 
-tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin,
-resulting in lower FMRI signal. However, to compensate for activation, muscles that
-control blood vessels relax in that region to allow more blow flow, which 
-in fact results in a higher concentration of oxygenated hemoglobin. Thus,
-increased activation actually tends to \emph{increase} the MR signal in
-comparison with the base level. It is this overcompensation that is the 
-primary signal detected with FMRI imaging. This cascade of events
-can, as a consequence of increased activity, increase the local metabolism, 
-blood flow, blood volume, and oxygenated hemoglobin; though not necessarily
-in sync. The lag between these
-various factors is what causes many of the complexities of the BOLD signal.
+\tableofcontents
 
-\subsection{FMRI}
-Magnetic Resonance Imaging, MRI is a method of building 3D images
-non-invasively, based on the difference between nuclear spin
-relaxation times in various molecules. Initially the entity being
-imaged is brought into a large magnetic field which aligns the spins
-of molecules in the same direction; radio frequency (RF) signals may
-then be used to excite nuclear spin away from the steady alignment. 
-As the nuclei precess back to their original orientation, they resonate
-at the same RF frequency of their original excitation. Conveniently, the
-excitation of nuclear spins return their original state at different
-rates, called the T1 relaxation time, depending on the properties of 
-the material excited. Additionally, the
-coherence of the spins also decay differently (and quite a bit faster
-than T1) based on the properties of the region that has been excited.
-This gives two primary methods of contrasting substances,
-which is the basis of T1 and T2 weighted images. Additionally, there
-dephasing occurs at two different rates, the T2 relaxation time,
-which is impossible to recover from, and T2$^*$ relaxation, which is
-much faster, but possible to recover from with an inversion pulse.
-Oftentimes T1 relaxation times can be on the order of seconds if 
-a significant excitation pulse is applied. 
-In order to rapidly acquire entire brain images, as is done in Functional 
-MRI, a single large excitation pulse is applied to the entire brain,
-and the entire volume is acquired in a single T1 relaxation period. 
-Because the entire k-space (spatial-frequency) volume is acquired 
-from a single excitation, the signal to noise ration is very low
-in this type  of imaging (Echo Planar Imaging).
-
-Increasing the spatial resolution of EPI imaging necessarily 
-requires more time or faster magnetic field switching. Increasing
-magnet switching rates though is difficult, because it can result in
-more artifacts, or even lower signal to noise ratios. The result is
-that at \emph{best} FMRI is capable of 1 second temporal resolution. 
-Additionally, the means that each voxel of the image will contain 
-the sum of a large amount neurons, capillaries and veins. Thus, the
-FMRI signal, which is sensitive to the chemical composition of 
-materials, is summing up the composition of various types of tissue
-in addition to the blood, whose composition is what we actually care about.
-In particular, the presence of Deoxyhemoglobin, Hemoglobin whose
-oxygen has been used by a metabolic process, has a decreased magnetic
-response compared to Oxygenated Hemoglobin. Thus, capillaries near
-very active cells will typically have a higher Deoxyhemoglobin content and
-lower signal, and regions with lower activity will have a lower 
-Deoxyhamoglobin content and thus higher signal.
-Unfortunately, as mentioned previously, blood is only a small part of
-each voxel, which means that a single EPI image doesn't tell much about
-Deoxyhemoglobin content. However assuming blood is the only thing changing
-in the short term, percent difference from a baseline signal \emph{will}
-tell us something. FMRI analysis is thus necessarily performed on the percent change
-from the baseline. Luckily the assumption that tissue content does not change in
-the short run is actually pretty good, although other factors can pollute
-the baseline signal, as we will discuss later.
-
-\subsection{BOLD Physiology}
-It is well known that the two types of hemoglobin act as a contrast agents in 
-EPI imaging
-\cite{Buxton1998}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
-between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
-Intuitively, increased 
-metabolism will increase Deoxyhemoglobin, however blood vessels are quick
-to compensate by increasing local blood flow. Increased inflow will of course
-preceed increased outflow, and increased inflow is accomplished by loosening 
-capilary beds. Both of these factors drive increased storage capacity.
-Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
-Hemoglobin, increased volume of blood can certainly effect this ratio if 
-metabolism doesn't exactly match the increased inflow of oxygenated blood.
-This was the impetus
-for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
-model (\cite{Mandeville1999}). These models derive from first principals
-the increased deoxyhemoglobin ratio and volume of capillaries based on a given flow.
-These were the first two attempts to quantitatively account for the shape of the 
-BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
-and the cerebral blood flow (CBF). In fact \cite{Buxton1998} went to far as
-to show that a simple, well chosen blood flow waveform coupled with a square 
-wave cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
-model, could fully account for the BOLD signal.
-
-Although \cite{Buxton1998} showed that a well chosen flow waveform could 
-explain much of the BOLD signal, there was still a matter of proposing a
-realistic waveform for the CBF and for the CMRO2. \cite{friston2000} gave
-a reasonable and simple
-expressoin for CBF input,$f$, based on a flow inducing signal, $s$, 
-\begin{eqnarray}
-\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
-\dot{f} &=& s
-\end{eqnarray}
-where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
-are both time constants. In \cite{Buxton2004} the final piece of the simple balloon
-model was put into place, by describing the CMRO2 as a constant multiple of
-the CBF (the inflow of blood). This completed the basic balloon model, and 
-was well summarized in \cite{Riera2004}. 
-\begin{eqnarray}
-\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
-\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
-\end{eqnarray}
-where $v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
-local deoxyhemoglobin/oxygenated hemoglobin ratio, $E_0$ is the resting metabolic
-rate and $\alpha$ is Grubb's parameter controling the balloon model. 
-\cite{Obata2004} refined the readout equation of the BOLD signal based on the
-deoxyhemoglobin content (q) and local blood volume (v), resulting in the
-final BOLD equation:
-\begin{eqnarray}
-y   &=& V_0((k_1 + k_2)(1-q) - (k_2 + k_3)(1-v))\\
-k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8\\
-K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57\\
-k_3 &=& \epsilon_0 - 1 = .43
-\end{eqnarray}
-Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
-deoxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
-change in relaxation rate with change in blood oxygenation, and
-$\epsilon_0 = 1.43$ is the 
-ratio of signal MR from intravascular to extravascular at rest. Although,
-obviously these constants change with experiment ($TE$, $\nu_0$, $r_0$),
-patient, and brain 
-region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are used
-as constants ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$) in 1.5 Tesla studies..
-While this model is in a sense complete, it is far from perfect. The major
-problem often brought up with this version of the BOLD model is that it
-does not represent the so called "post-stimulus undershoot" well.
-The post-stimulus undershoot is the name for a prolonged sub-normal
-BOLD response for a period of 10 to 60 seconds after stimulus has
-ceased (\cite{Chen2009}, \cite{Mandeville1999a}).
-
-There are two theories for the cause of the post stimulus undershoot. Recall
-that a lower than base signal means that there is an increased deoxyhemoglobin
-content in the voxel. The first and simplest explanation is that the post-stimulus
-undershoot is caused by a prolonged increase in CMRO2 after CBV and CBF
-have returned to their base levels. This theory is justified by quite a few
-studies that show CBV and CBF returning to the baseline before the BOLD signal
-(\cite{Frahm2008}, \cite{Donahue2009}, \cite{Buxton2004}, \cite{Lu2004},
-\cite{Shen2008}). Unfortunately, because of limitations on FMRI and en vivo
-CBV/CBF measurement techniques it is difficult to isolate whether CBF and
-CBV truly have returned to their baseline. Other research seems to indicate
-that there can be a prolonged residual supernormal CBV (\cite{Mandeville1999a}, 
-\cite{Behzadi2005}, \cite{Chen2009a}), although none of these papers completely
-rule out the possiblity of increased CMRO2. Additionally, in \cite{Yacoub2006}, it 
-was found that the post-stimulus undershoot varried across regions of the brain, 
-which could further explain the contradictions found elsewhere. \cite{Chen2009a}
-makes a compelling case that most of the post stimulus undershoot could be 
-explained be a prolonged CBV increase, and a prolonged CBF undershoot, and that
-many of the previous measurements showing a quick recovery of CBV 
-may have been dominated arterial CBV's return to baseline.
-
-Because of the significant possility of a completely independent CMRO2,
-extremely complex models for metabolism exist (\cite{Zheng2005}), although 
-most recent studies have
-focused on their ability to explain the prolonged BOLD post stimulus 
-undershoot \cite{Zheng2005}, \cite{Buxton2004}. This is because \cite{Buxton2004}
-and later \cite{Riera2004} showed that the main portion of the signal may be 
-accurately estimated by a simple blood flow locked expression of the CMRO2.. 
-Although \cite{Deneux2006} did not deal extensively with prolonged post
-stimulus undershoot, the comparisons made in that publication showed minimal
-improvement from separate expressions of CMRO2, in comparison to the much 
-increased complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
-terms, first proposed in \cite{Buxton2004}, that a slowed return to baseline for
-the BOLD signal is possible to model. However, viscoelastic effects primarily
-control CBV, which, as mentioned already, many studies have claimed cannot be 
-responsible for the BOLD post-stimulus undershoot. Another extensive model
-that attempts to quantify the post-stimulus undershoot is the compliance model 
-proposed by \cite{Behzadi2005}. Although through a somewhat different means than
-the \cite{Zheng2005} and \cite{Buxton2004} papers, its possible the increased
-model flexibility ultimately is the key reason for the improvements, as opposed
-to increased plausibility. Because of these controversies, and because this is
-the first time a particle filter has been used for this problem in this way, 
-our aim to keep the model simple was best met by using the original balloon
-model with the possible addition of the visco-elastic effects from \cite{Buxton2004}.
-
-Even more advanced versions of the Balloon model exist. In fact \cite{Buxton2004}
-introduced several additional state variables, including the CMRO2, the O2 extraction
-fraction, which is closely related to CMR02, and the neural response, which 
-causes the stimulus to decay toward some steady state value. The 
-neural response is intended to emulate neural habituation, wherein neurons
-become less sensitive to a prolonged stimulus. While these advanced may be more
-capable of capturing a more exact version of the BOLD signal, the difference
-between the model will often times be below the noise floor. In essence this 
-is a classic bias-variance
-dilemma: at some point increased model flexibility, and thus variance, is not
-worth the decrease in model bias. For now it remains to be seen where this
-line may be drawn in the BOLD signal, although \cite{deneux2006} does not
-show a significant improvement from the additional parameters added in 
-\cite{Buxton2004}.
-
-\subsection{Previous Studies of Parameters}
-There have been quite a few efforts to quantify the parameters of the
-various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
-both proposed physiologically reasonable values for the model parameters, 
-\cite{Friston2002} was the first paper to calculate the parameters based 
-on actual FMRI data. In that paper, Friston et. al. used a variation of
-Expectation Maximization to find a normal distribution for the parameters:
-\begin{eqnarray}
-\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
-\tau_s & =&  N(1.54, .25^2)   \nonumber \\
-\tau_f & =&  N(2.46, .25^2)   \nonumber \\
-\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
-\alpha & =&  N(.33 , .45^2 )   \nonumber \\
-E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
-V_0  & = &  .03 (not\ estimated) \nonumber
-\end{eqnarray}
-Since then, several other methods of have been used to calculate
-BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
-likelihood method for innovation processes was used, as described by
-\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
-Kalman filter, to break the time series into a series of innovations,
-for which Maximum Likelihood was performed. While this is in some since the
-"right" way to find the solution, it comes with several caveats. First, every
-step in parameter space requires a recalculation of all the state variables. With
-two or three parameters this is fine, more than that, and calculations could go on
-indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
-will only be optimal in that case. Third, depending on the nonlinearities present
-in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
-Unscented Kalman Filter over all the parameters and state variables to find the 
-parameter set/variable time series. While this method has the drawback of not necessarily
-being optimal, unfortunately there is no general optimal solution to non-linear non
-Gaussian models. Hu et. al.'s technique also will run significantly faster than
-ML based techniques, since it does not require recalculating the entire timeseries
-for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
-very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
-One potential problem with all these techniques are that the depend heavily on the
-priors. The starting point of the parameters could have a huge impact on the
-results, and while one can be hopeful that this isn't the reason for the agreement
-between \cite{Friston2000} and later results, there is no way to know.
-
-In \cite{Jonston2008}, a hybrid particle filter/gradient
-descent algorithm was used to simultaneously derive the static (classicly called
-parameters) and dynamic parameters (classically known as state variables).
-Essentially a particle filter is used to calculate the state variables at some
-time, and then the estimated distribution of the particles was used to find
-the most likely set of parameters that would give that distribution of state variables.
-\cite{Johnston2008} comes to a very different set of parameter estimates as compared
-to the original \cite{Friston2000} guesses:
-\begin{eqnarray}
-epsilon &=& .069 \pm .014    \nonumber \\
-tau_s & =& 4.98 \pm 1.07  \nonumber \\
-tau_f & =& 8.31 \pm 1.51   \nonumber \\
-tau_0 & =& 8.38 \pm 1.5    \nonumber \\
-alpha & =& .189 \pm .004   \nonumber \\
-E_0   & =& .635 \pm .072     \nonumber \\
-V_0   & =& .0149 \pm .006     \nonumber 
-\end{eqnarray}
-Notably, the time constants are significantly longer. This could be a result of
-some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
-later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
-depended less on the priors.
-
-Its possible, although unlikely that the balloon model is not possible to learn
-without detrimental cost in variance error. I do not think this is the case, however,
-and I will work to dispell this possibility in the results with simulated time
-series.
-
-\subsection{Noise}
-Thus, despite some discrepencies, the cause of the BOLD signal is
-relatively well known.  but FMRI doesn't detect this happening in one neuron,
-but rather as the 
-aggregate over millions of cells. Though local neurons act
-"together" (i.e. around the same time), the density of neurons, the
-density of capillaries, and slight differences in activation across 
-a particular voxel can all lead to signal attenuation or noise. 
-A particularly insidious type of noise present in FMRI is a low frequncey
-drift, first characterized by a Weiner process in \cite{Riera2004}. 
-Though not present in all regions, it is prevalent enough to cause significant
-inference problems \cite{Tanabe2002}. It is still not
-clear where exactly this noise comes from, although it is possible it is 
-the result of magnets heating up, or some distortion in magnetic
-fields \cite{Smith2007}. It is clear that this drift signal is not soley
-due to a physiological
-effects, given its presence in cadavers and phantoms \cite{Smith1999}.
-Additionally, since it is standard operating procedure to perform
-coregistration between volumes, it is unlikely that movement was
-the cause of the drift either (otherwise the effect would not be 
-widely reported).
-
-\section{Current Techniques}
-\subsection{Basic Statistical Parametric Mapping}
-Although not strictly the same thing as parameter calculation from 
-FMRI, activation detection is very similar. In fact, estimation of 
-parameters is somewhat a generalization of the idea of activation detection.
-Thus, it is important to draw a comparison between the methods proposed
-in this thesis with existing methods of activation detection.
-
-The most basic method of analyzing FMRI data is through a standard T-test
-between "resting state" and "active state" samples. This is done by 
-taking the average and variance of the inactive period, and the 
-period during which the stimulus was activate separately then treating 
-them both as gaussian distributions.
-If they are in fact Gaussian distributions, then a basic t-test will
-give the probability that the samples came from the same distribution
-(the null hypothesis). Of course, this test is fraught with problems; even if
-the drift mentioned earlier has been removed, there is little reason
-to believe that the noise is Gaussian, or even stable. Additionally, 
-even if the noise were Gaussian, a t-test with a p-value of .05 over
-50000 or more samples is on average going to generate $.05*50000$ false
-positives. To compensate for this, bonferoni correction, also known as
-multiple comparison tests are performed; essentially p-values are 
-divided by the number of independent
-tests being run. This, however, leads to extremely low p-values, so
-low that it would be impossible for any biological system to satisfy. To
-compensate, a Gaussian kernel is applied to the image, thus reducing
-variance (and thus separating the active and inactive distributions)
-as well as decreasing the effective number of voxels. Since t-tests are
-now no longer being applied to n <I need to define n> independent voxels,
-the factor by which the p-value must be divided by can be decreased.
-<Do I need to mathematically define all this?> The derivation and application
-of random field theory, and its use can be found in various papers \cite{univ_mult_rft}.
-
-\subsection{General Linear Model}
-The most used form of FMRI analysis is Statistical Parametric
-Mapping, but is able to account for several different levels or types
-of stimulus (see \cite{Hofmann1997}). By using hierarchical models
-the output signal timeseries is considered the weighted
-sum of the various input timeseries. Essentially every experimental factor
-is considered as another level inputs (ex. which patient is an input). 
-The equation for a general linear model is then
-\begin{equation}
-Y(t) = X(t)\beta + \epsilon(t)
-\end{equation}
-where $Y(t)$ is the smoothed or detrended timeseries of measurements,
-$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
-and $\epsilon$ is the error. Thus for every time, the measurement is
-assumed to be a weighted sum of the inputs plus some error. The calculation
-of $\beta$ then performed using a maximum likelihood or gradient descent search 
-to minimize the error.
-
-<Image of GLM>
-
-It is well known of course that a square wave stimulus does not result in a square wave
-in the activation of brain regions. Thus, various methods are used to 
-smooth $X(t)$'s related to actual input (as opposed to experimental confounds
-such as patient number, age etc) through time, and bandlimit the input. 
-The best technique is convolving the stimulus input with a hemodynamic 
-response function, which mimicks the basic shape of BOLD activation, including a delay
-due to rise time and fall time. The downside of this method, however is that 
-the hemodynamic signal is static, meaning the same hemodynamic function is
-used for every region of the brain, and $Y(t)$ must be a constant
-scalar multiple of $X(t)$ to be appropiately identified.
-
-<Image of Hemodynamic Response Function>
-
-The GLM is extremely powerful at determining the linear dependence of
-a set of regressors on the output. Unfortunately, there is significant evidence
-that many of these dependencies are nonlinear, which means they may be
-difficult or impossible to detect using linear techniques. Additionally, 
-the lack of flexibility in delay variables certainly hinders SPM's
-ability to locate voxels that do not conform to the canonical hemodynamic 
-response function. The very fact that different hemodynamic response functions
-work better for different regions/experiments is a testament to the need
-for increased flexibility.
-
-Linear models also ignore prior knowledge about the system. Combined FMRI
-CBF or CBV imaging methods are available and could shed much light on
-neural activation; however there is no way of combining that data into a
-linear model. Although the most advanced form of the general linear model includes
-a Hemodynamic Response Function, it is static across every region of the brain;
-yet it is well known that capillary beds are not uniform and so blood perfusion 
-cannot possibly be static across the brain. Its possible that activation is
-every bit as strong in regions that aren't found by SPM, but that there are
-nonlinearities that do not match the canonical HRF; resulting in signficant bias error. 
-It is not uncommon
-for data to be thrown out in FMRI studies because no significant activation
-has been seen. This practice highlights the fact that strictly linear
-approaches are insufficiently flexible to account for relatively common
-variations of neural activation.
-
-<Image with two different $\alpha$s>
-
-<image comparing the results of 10\% changes in various signals>
-
-In general activation detection type methods also don't have the ability 
-to find pathologies based on state variables or parameters.  It
-is quite possible that physical properties such as decreased compliance of
-blood vessels could indicate a neurological condition that is not easily
-seen in a T1 or T2 map. In essence, this could make FMRI a much more 
-useful clinical tool than it is now.
-
-Finally, all these techniques require the noise to be Gaussian to reach
-an optimal solution. In fact there is no known optimal solution to
-nonlinear models with non-Guassian noise, so obviously some assumptions
-are going to have to be made to reach a solution. However, it would be
-nice to have an algorithm that is robust to these effects, and could
-still give a good solution when Gaussianity is violated.
-
-\section{Unscented Kalman Filter}
-The unscented kalman filter is a powerful gaussian/bayes filter that attempts
-to model the posterior distribution of dynamical systems as a multivariate
-Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
-Filter by allowing the state update to be a function, $g$,
-
-\begin{eqnarray}
-X(t) = g(u(t), X(t-1))\\
-Y(t) = h(X(t)
-\end{eqnarray}
-
-In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
-(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
-weighted according to a Gaussian estimate of $X(t-1)$ are passed through
-the update equation. This set of points are then used to estimate the 
-mean and covariance of $X(t)$. The benefit of this, is that it requires
-no Jacobian and only a few extra calculations to get a decent esimate of
-a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
-closed form solution, and finding the Jacobian is impossible without approximations. Although 
-\cite{Riera2004} claims to use a Jacobian for their solution, they are
-in fact using the Jacobian of $\dot{s}$, $\dot{q}$, $\dot{f}$, $\dot{v}$,
-rather than $s$, $q$, $f$, and $v$. This is important because the Extended
-Kalman filter depends on the jacobian to map a gaussian through the 
-advancement of time. Thus the popular Extended Kalman Filter won't work
-in this case, whereas the Unscented Kalman Filter still does.
-
-\begin{figure}
-\label{fig:EKFWorking}
-\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun}}
-\end{figure}
-
-The difficulty
-of using a Kalman Filter, however, is that it assumes a multivariate 
-Gaussian for the state variables, $X(t-1)$. The problem with this is that 
-when a system is significantly nonlinear, the state at 
-$X(t)$ will almost certainly be non-Gaussian, and thus estimating
-$X(t+1)$ with $X(t)$ as a multivariate Gaussian will perpetually introduce
-error in the distribution. Furthermore, it is not really known what 
-sort of underlying distributions may exist in such a mixed biological,
-mechanical, chemical system such as the brain. Assuming the parameters
-all to be Gaussian may in fact be a gross error. On the other hand, if
-the our tests indicate a Gaussian distribution is a good fit, it would
-be a wise plan to go back and consider the use of a modified Unscented 
-Kalman Filter rather
-than the full Particle Filter we discuss in the next section. These are 
-non-trivial issues given that the assumption of Gaussianity is what allows
-the UKF to estimate the posterior using only the first and second moments;
-two parameters that don't uniquely describe most distributions.
-
-Another potential problem with the UKF is that typically the sigma points,
-used to estimate the posterior probability, $P(X(t) | X(t-1), u(t))$, are
-located on the main axes. As a result, the covariances are not allowed to 
-be effected by the state transition the same way the variances are. While
-this may be reasonable in low-dimensional systems, high demsional systems
-have a much greater potential for interplay between the variables. While
-this problem is relatively easy to fix (by using more sigma points off
-the main axes), there is a definite cost in complexity.
-
-Ultimately, the nonlinearities of the BOLD model make Gaussian estimates
-too unrealistic for practical use. As \ref{fig:Transformation} shows, even 
-holding all other parameters steady, the nonlinearities due to $q$ and $v$
-alone are enough to give a non-gaussian result from a two variables 
-gaussian input.
-
-\begin{figure}
-
-\label{fig:Transformation}
-\caption{a) A Two-Variables gaussian input of $q$ and $v$, b) The 
-resulting distribution after being evolved for two seconds. c) The
-estimated Gaussian based on the mean and covariance of b). d) The
-residual between the two distributions.}
-\end{figure}
-
-\section{Proposed Approach}
-\subsection{Goal}
-The ultimate goal of this project is to provide a new set of tools
-for analyzing FMRI data. Whereas SPM techniques have been highly 
-successful at finding macroscopic regions of activation, linear 
-modeling can carry significant bias error due to lack of model
-flexibility. While adding parameters can significantly increase
-error due to model variance, this effect is mitigated by the fact
-that we plan to use a model that is based on first principals. The
-purpose of this paper is thus to evaluate the potential of using
-a particle filter along with the BOLD model to derive physical 
-parameters. In so doing, we hope to be able to show that one or more
-parameters are a suitable replacement for estimating voxel 
-activation from a standard FMRI image. We also hope to show that 
-estimated posterior distribution of the parameters, derived from
-the particle filter, is able to provide an accurate measure of the
-confidence interval.
-
-\subsection{Introduction to Particle Filters}
-Particle filters, a type of Sequential Monte Carlo (SMC) methods
-are a powerful way of estimating the posterior probability distribution
-of a set of parameters give a timeseries of measurements. Unlike Markov 
-Chain Monte Carlo estimation, Sequential Monte-Carlo methods are designed
-to be used with parameters that vary with time. Unlike variations of the
-Kalman filter, particle filters do not make the assumption that noise
-is Gaussian. Thus particle filters are often the best solution to bayesian 
-tracking for non-linear, non-gaussian systems. 
-%todo: might need to reference attempts to model a system of our type with
-%unscented kalman filter
-\subsubsection{Model}
-The idea of the particle
-filter is to start with a wide mixture PDF of possible parameter sets, 
-and then, as measurements come in, to weight more heavily parameter sets 
-that tend to give good estimations of the measurements. The reliance on
-an initial mixture PDF can introduce bias; however, this effect can be
-minimized by alterring the initial weights in the mixture pdf. Of course
-every gradient descent must choose starting points and it is often quite
-easy to establish a reasonable range of parameters, especially when the
-model being used has a physical meaning. Suppose a set or stream of measurements
-are given, $\{y(t), t = 1, 2, 3, ... T\}$, where $T$ is permitted to go
- to infinity. Then the goal is to find the 
-parameters, $\hat{\theta}$, and underlying state time series, $\hat{x}[0:T]$
-that minimize the difference between $\hat{y}[0:T]$
-and $y[0:T]$. In our case, we will assume that we know the form
-of the model, which is based on first principals, and that
-there is some true $\theta$ and a true time-series of underlying
-state variable, $x[0:T]$ that drives $y[0:T]$. Assuming a model form 
-such as we do here reduces model variance, potentially at the cost of increased
-bias (or systematic) error. We will assume a basic state space model:
-
-\begin{equation}
-\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x)
-\end{equation}
-
-\begin{equation}
-y(t) = g(t, x(t), u(t), \theta, \nu_y)
-\end{equation}
-
-Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
-constants, $u(t)$ is a stimulus, $y(t)$ an observation, and
-$\nu_x$ and $\nu_y$ are random variates. Obviously any one of these could
-be a vector, so for instance $u(t)$ could encode multiple types of stimuli.
-
-Although not strictly necessary for particle filters, we will make a few
-assumptions based on the particular type of systems faced in biological 
-processes. First, the systems are assumed to be time invariant. This 
-assumption is based on the idea that if you froze the system for $\Delta t$
-seconds, when unfrozen the system would continue as if nothing happend. 
-Few biological systems are predictable enough for them to be summarized
-by a time varying function. Even systems that one might assume work in that
-way are actually much more complicated. The heart may seem like an obvious
-exception, however, the period between heartbeats vary often enough that prediction
-would necessitate another state-space model. In short, we
-assume no parameters are time varying, because not enough information exists to
-describe any of theme in that way. Luckily particle filters are capable 
-of dealing with non-white, non-Gaussian noise, so unanticipated influence
-may be re-factored as noise. Secondly we assume that input cannot directly
-influence the output, which in the case of the BOLD signal is a good assumption.
-%Third, through some sort of preprocessing, we will assume that $\nu_d$ can be
-%decreased or completely removed, since it 
-Third, we will assume noise is additive, and that $\nu_x$ may be projected into
-a Weiner, or other summing process that is additive with $g$ and $\nu_y$, which
-will be named $\nu_d$.
-Finally, $x(t)$ will encapsulate $\theta$, the unknown model constants, which
-means that the vector $\dot{x}$ will always have members
-that are 0. The results of these assumptions are a simplified version of the
-state space equations:
-
-\begin{equation}
-\label{stateass}
-\dot{x}(t) = f(x(t), u(t))
-\end{equation}
-
-\begin{equation}
-\label{measass}
-y(t) = g(x(t)) + \nu_y + \nu_d
-\end{equation}
-
-Because $\nu_d$ is something akin to and additive Weiner process $y[0:T]$, it 
-will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
-in $y[0:T]$. For some of the tests, I will use de-trending methods to reduce the effects of 
-$\nu_d$, the remainder of which will be re-factored into $\nu_y$. Both $\nu_d$ and $\nu_y$
-have biological and non-biological sources. MR can lead to both types of noise, 
-as demonstrated in \cite{drift}. Meanwhile changes in metabolism, heart rate, or
-other biochemical intervention could all lead to either $\nu_d$ or $\nu_y$.
-
-\subsubsection{Prior}
-
-The goal of the particle filter is to evolve a probably distribution 
-$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
-that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
-Considering that $y$ contains measurement noise as well as noise from $x$,
- it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
-but a true posterior. 
-To begin with, the particle filter starts with a prior distribution, and $N_p$
-particles need to be drawn from that distribution, $\alpha(x)$:
-\begin{equation}
-\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
-\end{equation}
-
-Where $N_p$ is the number of particles or points used to describe the prior 
-using a Mixture PDF. 
-\begin{equation}
-\hat{Pr}(x(0) = \hat{x}) = \sum_{i=1}^{N_p} w_i\delta(\hat{x} - x_i(0) ) dx
-\end{equation}
-Where $\delta(x-x_0)$ is 1 if and only if $x = x_0$ (the Kronecker delta function).
-
-If a true prior is preferred, then the weights
-should all be $1/N_p$, and since $x_i$ was drawn from the prior, this will
-be an approximation of the prior distribution. If a relatively flat prior is 
-preferred, then each particle's weight could be divided by the density, $\alpha(x_i)$,
-which creates a flat prior with support points in the region of $\alpha(x)$. Either
-way, $\alpha(x)$ should be much broader than the true posterior, $Pr(x(0))$, since the
-choice of support points is crucial to the convergence of any sampling importance
-algorithm. 
-For the BOLD signal all the parameters have been studied and have relatively well
-known mean and variance, so a prior could be very helpful. We ran simulations for
-both normalized and un-normalized priors, although we believe in cases such as this,
-where a good prior exists, it should be used. 
-For strictly positive parameters (members of $x$) we used a gamma distribution,
-whereas for parameters that could be negative, we used a Gaussian distribution. In
-both cases standard deviations twice that found in previous studies were used.
-
-Note that all the probabilities implicitly depend on $u[0:T]$, so those terms 
-will be left off for simplicity.
-Once the probability, $\hat{Pr}(x(T) | x[0:T-1], y[0:T-1])$ has been found
-(initially this is just Mixture approximating the prior since no measurements are 
-available and no previous probabilities are available), its possible to approximate
-the probability for short times between times when measurement is available, by shifting
-the probability according the progression of the state equations. This is only 
-an approximate, since integrating $\nu_d$ should increase uncertainty as
-time without a measurement passes.
-
-\begin{equation}
-\hat{Pr}(x(T+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
-\end{equation}
-
-\subsubsection{Weighting}
-When a measurement becomes available it is incorporated into the probability.
-This process of incorporating new data is called sequential importance sampling,
-and eventually causes the probability to converge. The weight is defined
-as
-
-\begin{equation}
-\label{weightfunc}
-w_i(T) \propto \frac{\hat{Pr}(x_i[0:T] | y[0:T])}{q(x_i[0:T] | y[0:T])}
-\end{equation}
-
-where $q$ is called an \emph{importance density}, meaning it decides where
-the support points for $x(T)$ are located. To remove the bias due to the
-location of the support points, we divide by $q(x_i[0:T] | y[0:T])$. By dividing by 
-the posterior density of the support points (particles), the effect of the particle 
-distribution may be removed from the posterior density. As a result the weight
-is dependent solely based
-on $\hat{Pr}(x_i[0:T] | y[0:T])$, the probability of the $i^{th}$ particle's measurements
-being different from $y[0:T]$ due to noise alone.
-An example of an importance density would be drawing a large
-number of points from the standard normal, $N(0,1)$ and then weighting each point, $l$ by 
-$1/\beta(l), \beta \sim N(0,1)$. Of course if there is a far off peak in
-the posterior that $q$ does not allocate support points in, there will 
-be a quantization error, and that part of the density can't be modeled. This is why
-it is absolutely necessary that $q$ covers $\hat{Pr}(x_i[0:T] | y[0:T])$.
-
-$q(x_i[0:T] | y[0:T])$ may be simplified by assuming that $y(T)$ doesn't contain 
-any information about $x(T-1)$, which is more practical since knowledge of future
-measurements is impractical.
-
-\begin{eqnarray}
-q(x[0:T] | y[0:T]) & = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T]) \nonumber \\
-& = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T-1]) \nonumber \\
-& = & q(x(T) | x(T-1), y[0:T])q(x[0:T-1] | y[0:T-1])
-\end{eqnarray}
-
-In this paper we will use 
-$q(x_i(T) | x_i(T-1), y[0:T]) =  \hat{Pr}(x_i(T) | x_i(T-1))$,
-based on the Markov assumption, and the belief that the state space model is 
-able to approximate the true state. This means that prior to re-weighting 
-particles, the particles will be distributed the same as the previous time but
-moved forward according to the integration of $f(x(t), u(t))$.
-
-In addition to $q(x_i(T) | x_i[0:T-1], y[0:T])$, the weight is also based on $Pr(x_i[0:K] | y[0:K])$,
-which may be broken up as follows.
-\begin{eqnarray}
-\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y[0:T], x[0:T])}{\hat{Pr}(y[0:T])} \nonumber \\
- & = & \frac{\hat{Pr}(y(T), x[0:T] | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}}{\hat{Pr}(y(T) | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}} \nonumber \\
- & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x[0:T] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1]) } \nonumber \\
- & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-\end{eqnarray}
-
-Using the assumption that $y(t)$ is fully constrained by $x(t)$ \eqref{measass},
-and that $x(t)$ is fully constrained by $x(t-1)$ \eqref{stateass}, we are able to
-make the reasonably good assumptions that:
-\begin{equation}
-\hat{Pr}(y(T) | x[0:T], y[0:T-1]) = \hat{Pr}(y(T) | x(T))
-\end{equation}
-
-\begin{equation}
-\hat{Pr}(x(T) | x[0:T], y[0:T-1]) = \hat{Pr}(x(T) | x(T-1))
-\end{equation}
-
-Additionally, for the particle filter $y(T)$ and $y[0:T-1]$ are 
-given, and therefore constant across all particles. Thus $\hat{Pr}(x[0:T] | y[0:T])$
-may be simplified to:
-\begin{eqnarray}
-\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) 
-            \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-& = & \frac{\hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-& \propto & \hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])
-\end{eqnarray}
-
-Plugging these simplifications into \eqref{weightfunc} leads to:
-\begin{eqnarray}
-\label{weightevolve}
-w_i(T) & \propto & \frac{\hat{Pr}(y(T)| x(T)) \cancel{\hat{Pr}(x(T) | x(T-1))} \hat{Pr}(x[0:T-1] | y[0:T-1])}
-                         {\cancel{\hat{Pr}(x_i(T) | x_i(T-1))}q(x[0:T-1] | y[0:T-1])} \nonumber \\
-& \propto & w_i(T-1)\hat{Pr}(y(T)| x(T)) 
-\end{eqnarray}
-
-Thus, by making the following relatively weak assumptions, evolving a posterior
-density  is easy and requires almost no knowledge of noise distribution.
-\begin{enumerate}
-\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ provide 
-a sufficiently flexible model to encapsulate the true time series.
-\item $E[\nu_d] = 0$ and $E[\nu_y] = 0$, and $\nu_x = d\nu_d$, $\nu_y$ are stationary
-\item The PDF $q(x_i(0))$ (the prior) fully covers $Pr(x_i(0))$
-\item Markov Assumption: $Pr(x(T) | x[0:T]) = Pr(x(T) | x(T-1))$
-\item $q(x[0:T-1] | y[0:T]) = q(x[0:T-1] | y[0:T-1])$
-\end{enumerate}
-
-\subsubsection{Basic Particle Filter Algorithm}
-From the definition of $w_i$, the algorithm sequential importance sampling (SIS) 
-is relatively simple.
-
-\begin{algorithmic}
-\STATE Initialize $N_p$ Particles: 
-        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
-\STATE $T$ = \{Set of Measurement Times\}
-\FOR{$t$ in $T$}
-    \FOR{$i$ in $N_p$}
-        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
-    \ENDFOR
-\ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
-\end{algorithmic}
-The result is then a discrete approximation of the posterior distribution.
-
-\subsubsection{Resampling}
-As a consequence 
-of the wide prior distribution (required for a proper discretization of a continuous
-distribution), there will be many particles with insignificant weights. While this does help
-describe the tails of the distribution very well, it means that only a small portion of the
-computation will be spent describing the most probable region. Ideally every particle would 
-equally decrease the entropy of the distribution, thus the lower the variance of the weights,
-the more efficiently the discrete distribution is in describing the continuous distribution. 
-A common measure of "Particle Degeneracy" is the effective number of particles, described
-in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
-Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
-of each particle. Of course the true weight is unknown, so a heuristic approximating 
-$N_{eff}$ is used:
-\begin{equation}
-\label{neff}
-\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
-\end{equation}
-Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
-$N_{eff}$ drops precipitously.  To alleviate this problem
-a common technique known as resampling must be applied. The idea of re-sampling is to 
-draw from the approximate posterior, thus generating a replica of the posterior with 
-a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
-$N_p$ points are drawn from the posterior,
-\begin{equation}
-\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
-\end{equation}
-then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
-still limited to the original particles, the number of unique particles can only go down.
-This effect, often dubbed "particle impoverishment" can result in excessive quantization
-errors in the final distribution. However, there is a solution. Instead of sampling from the
-discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
-that distribution. Because the distribution is continuous, there is no way for a collapse
-of the particles to occur. The question then, is how to decide on the smoothing kernel. 
-Often times the easiest way to sample from the continous distribution is to break the 
-re-sampling down into two steps. First a member of the discrete distribution is randomly
-selected based on the weights, and then based on the smoothing a nearby state variable 
-is selected. The process of the selection will be defined as:
-\begin{equation}
-\chi_i = x_i + h\sigma \epsilon
-\end{equation}
-Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
-and $\epsilon$ is drawn from the chosen kernel.
-It has been proven that when all the elements of the mixture
-have the same weight, as is the case after basic resampling, the kernel that minimizes the 
-MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
-Particle Filters, C Musso, N Oudjane and F LeGrand). 
-\begin{equation}
-K = \left\{
-\begin{array}{lr}
-\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
-0 & otherwise
-\end{array}\right.
-\end{equation}
-%<more here>
-
-If the noise is assumed to be Gaussian then it is possible to further optimize. 
-Thus we let $h$ be defines as:
-\begin{eqnarray}
-h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}
-\end{eqnarray}
-and although it is very possible the underlying noise is non-gaussian, the Gaussian
-may work, but sub-optimally. It has been proposed that (Monte Carlo Approximations for
-General State-Space Models, markus Hurzeler and Hans R. Kunsch) if the underlying 
-distribution is non-Gaussian, then using this bandwidth will oversmooth. 
-In reality over smoothing
-should not be too great an issue because the smoothing is only being applied to find new
-particles. If the distribution is over smoothed then the algorithm may not converge as rapidly;
-however, because the bandwidth is still based on particle variance, which will decay as 
-particles are ruled out, it is still able to converge. In fact over smoothing is preferrable
-to under smoothing, since the latter would result in false negatives, but the previous only
-results in a slower decay of the variance. 
-At the same time, as $n_x$, the number of dimensions in
-$x$, goes to infinity, the standard deviation based approximation becomes less effective
-(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
-tracking, sanjeev arulampalam, simon maskell, neil gordon...).  Because of the high dimensionality of our system,
-and limited measurements, it is helpful to have a broader bandwidth to explore the distribution. 
-Nevertheless, because 
-of the potentially wide smoothing factor applied by regularized resampling, performing this
-step at every measurement would allow particles a great deal of mobility. This mobility is
-the enemy of convergence, which is why regularized resampling should only be done when
-$\hat{N}_{eff}$ drops very low (say less than 50). Other than the periodic regularized
-resampling then, the regularized particle filter is nearly identical to the basic sampling
-importance sampling filter (SIS).
-
- \begin{algorithmic}
-\STATE Initialize $N_p$ Particles: 
-        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
-\STATE $T$ = \{Set of Measurement Times\}
-\FOR{$t$ in $T$}
-    \FOR{$i$ in $N_p$}
-        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
-    \ENDFOR
-
-    \STATE Calculate $N_{eff}$ with \eqref{neff}
-    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
-        \STATE Calculate empirical $\sigma$ 
-        \STATE $h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}$
-        \STATE Redraw particles using (stratified) basic resampling
-        \FOR{$i$ in $N_p$}
-            \STATE Draw $\epsilon \sim K$
-            \STATE $x_i = x_i + h \sigma \epsilon$
-        \ENDFOR
-    \ENDIF
-\ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
-
- \end{algorithmic}
-
-The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
-nearby which allow for more accurate estimation near points of high likelihood. 
-As the variance of the estimated
-$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
-will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.
-
-\subsection{Choosing $\hat{Pr}(y(T) | x(T))$}
-Choosing a representation of an unknown distribution is certainly tricky,
-and so the fact that $\hat{Pr}(y(T) | x(T)) = \nu_d + \nu_y$ means that
-there is a significant piece of the algorithm that is based primarily conjecture.
-Studies of the noise in FMRI typically attribute noise to a Gaussian random
-variable or an additive noise process with Gaussian steps.
-
-\subsubsection{Classical De-trending}
-The non-stationary
-aspect of a Weiner process as with $\nu_d$ is difficult to compensate for, and so various methods
-have been developed to compensate for it. \cite{detrend} and \cite{drift} have
-demonstrated that this component is prevalent, and may in fact be a characteristic
-of FMRI. In some studies, as many as half the voxels benefit from detrending, meaning
-that this is certainly a serious barrier to inference. All the existing methods are performed
-during the preprocessing stage, rather than as an integral part of analyzing the BOLD
-signal. There is no shortage of theories on the "best" method of detrending, however
-a head to head comparison, \cite{detrend}, showed that in most cases subtracting off
-a spline works the best. The benefit of the spline versus wavelets, high pass 
-filtering or other DC removal techniques is that the frequency response is not set.
-A spline is able to move quickly when the signal is moving quickly, and move more
-slowly when the signal moves more slowly. That said, the spline will still remove some
-amount of signal, just like all of these methods.
-
-<image of de-spline'd lines with "true" lines>
-
-\subsubsection{Delta Based Inference}
-I also propose and test a different method of dealing with the so called "drift". 
-Instead of comparing the direct output of the particle filter with the direct
-measurement, the algorith compares the change in signal over a single TR. 
-In most signal processing cases this would foolish, but that is because the 
-general assumption that all noise is high frequency is not the case here.
-In fact, every pipeline for the analysis of BOLD signal uses a high pass filter,
-but low poss filters are rarely applied, because it is a well known fact that 
-most of the signal is in the high frequency range and most of the noise is actually 
-in the low frequency range. The particle filter is an 
-extremely robust method of inference, and so I would assert that the particle
-filter ought to be given as \emph{raw} data as possible. While taking direct measurements
-without de-trending would give awful results, using the difference removes the 
-DC component and turns a Weiner process into a Gaussian random variable.
-
-\begin{equation}
-\label{measass}
-\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
-\end{equation}
-
-Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
-$\nu_d$ is some other additive process, the difference will still be one of a few stable
-distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
-with a maximum variance of twice the original variance. All the assumptions made originally
-for the particle filter hold, and all of the parameters may be distringuished based on
-the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
-rather than string of direct readings.
-
-<frequncy response graphs, highlighting noise frequency range and signal frequency range>
-
-\subsubsection{Weighting Function}
-Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting funciton,
-is based on an unknown distribution, it is necessary to decide on a function
-that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
-needs to centered at zero and have a scale comparable to the signal levels.
-While a Gaussian function is the natural choice, we also wanted to try a distribution
-with wider tails, so that outliers don't completely destroy particle's weights.
-Therefore, we tried three weighting functions; Gaussian, exponential and 
-the Cauchy distribution. The standard deviations (or scale) was set to 
-$\sigma_y/5$, where $\sigma_y$ is the variance of all $y[0:\infty]$. Of course 
-since the particle filter requires a weighting function to run, this means that
-before the particle filter starts all the measurements have to be in. In cases where
-this is impossible, a heuristic based on a small sample may work just as well.
-
-The shape of the weighting function is extremely important, because it essentially decides
-the rejection rate of particles. A very thin gaussian probability 
-distribution function has nice properties, but thin tails. As a result, large
-outliers in the measurement vector could easily force all the particles to
-have near 0 weights, thus forcing the particle filter to converge improperly. 
-On the other end of the spectrum, a cauchy PDF, has relatively fat tails, and
-may not weight central particles high enough, preventing the particles from 
-converging at a reasonable rate. Exponetials have the benefit of having an 
-extremely smooth drop to zero, and a slope of 1 at the origin. Having a non-zero
-slope at the origin is beneficial because it discriminates all the way up until
-the measured and predicted $y$ are the same. The importance of the weighting function
-cannot be overstated, as this is the primary factor in deciding the rate at
-which the particle filter converges.
-
-Obviously the optimal $\hat{Pr}(y(T) | x(T))$ is the true $Pr(y(T) | x(T))$; however
-since that is unknown, I tested multiple different distributions. As stated previous,
-because the exact scale of $Pr(y(T) | x(T))$ it not even known, the exponential distribution
-has the added benefit of having a negative slope all the way form zeo to infinity. Thus
-even if the scale is completely wrong, particles will still be well differentiated.
-
-<Q-Q plot of real fmri data with Gaussian, DC>
-<Q-Q plot of real fmri data with Gaussian, deltas>
-
-\section{Simple, Nonlinear Example}
-A simple example may be used to demonstrate the proposed use of a particle filter.
-Thus, let us consider a simple half wave rectifier circuit, shown in figure 
-\ref{fig:HalfWaveRectifier}
-
-\begin{figure}
-\begin{center}
-\begin{circuitikz}[scale=2, american]
-\draw
- (0,0)  node[transformer core] (T) {}
- (T.A1) -- (-1,0)
- (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
- (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
- (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
- (T.base) node {G}
- (T.B1) to[open, *-*, v=$V_1$] (T.B2); 
-\end{circuitikz}
-\label{fig:HalfWaveRectifier}
-\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
-gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
-$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
-\end{center}
-\end{figure}
-
-The half wave rectifier circuit smoothes the gaps between high voltage
-with a capacitor. Thu, when $u(t)G$ is less than $v_t$, the circuit will 
-discharge the capacitor and maintain a non-zero voltage,
-but when $u(t)G$ is greater than $v_t$, the output voltage will be set
-by $u(t)G$ and the capacitor will charge up. We will assume a very simple
-model for all the components, ignoring the complex nonlinear behavior
-that can occur in a diode. Although rectifiers are typically thought
-of as receiving an AC circuit 60 Hz, we will ignore such specifics and 
-assume the voltage across the output of the transformer is simply a scalar
-multiple of the input voltage.  As discussed in the \ref{Sec:ParticleFilter} any
-variable with uncertainty must be part of the state variable. Therefore
-the state variable will be: $X(t) = \{G, v_t, C, R_m, v_y\}$. Of course,
-$u(t)$ cannot be allowed to be a square wave in such a system, since that
-such a signal would never get across the transformer and regardless it
-would necessitate an unrealistic infinite current across the capacitor.
-The state equations would then be
-
-\begin{equation}
-v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
-        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
-        v_y(t-1)\left(1 - \frac{1}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
-    \end{cases} 
-\end{equation}
-
-To run the particle filter is relatively easy then, since there exists 
-a recursive definition of the dynamic state variable, $v_y$. To start
-with, an initial distribution must be assumed and while at first a 
-Gaussian seems like a good idea, all the static state variables are strictly
-positive and thus not well suited to the Gaussian. In this case then,
-it would be wise to start with a Gamma distribution, and just be wary of
-any standard deviation that gets larger than the prior mean. We will define
-the gamma distribution as follows:
-
-\begin{equation}
-X \sim Gamma(k, \theta) \rightarrow f(x) = x^{k-1}\frac{e^{-x/\theta}}{\theta^k\Gamma(k)}
-\end{equation}
-
-where $\Gamma$ is the gamma function.
-The in some ways the margin for error is decided by the weighting function, which
-here will define as $W(V_y, v_y)$, where $V_y$ is the actual measurement, and 
-$v_y$ is the estimate by a particular particle. The choice of this function is difficult,
-and although the Gaussian is typically used, we found the exponential helpful
-in dealing with particle deprivation.  The algorithm will then look like the following,
-
-\begin{algorithmic}
-\STATE Initialize $N_p$ Particles
-\FOR{$i$ in $N_p$}
-    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
-    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
-    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
-    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
-    \STATE $v_y = 0$, (Assume the system has been off for a long time)
-    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
-    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
-\ENDFOR
-\FOR{$t$ in Set of Measurement Times}
-    \FOR{$i$ in $N_p$}
-        \STATE $v_y(t) = f(v_y(t-1, u(t))$
-        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
-    \ENDFOR
-\ENDFOR
-
-\end{algorithmic}
-
-\section{Methods}
-This paper describes two types of experiments; first we will cover
-simulations which have the benefit of a ground truth, then we will
-cover the methods used in the use of the particle filter on real data.
-
-\subsection{Preprocessing}
-As discussed in the section on de-trending, the normal pipeline for analyzing
-FMRI involves a great deal of preprocessing. In this paper we make an effort to
-minimize any type of preprocessing that will degrade the signal. 
-After FMRI data has been acquired it is always necessary to modify the
-data in some way to make different runs comparable. Because FMRI signal
-levels are unit-less, at the very least it is necessary to convert
-the data into \% difference from the baseline. This process removes no data
-from signal since it merely subtracting then dividing by a constant. This
-is the signal that was input into the delta based particle filter.
-Of course there are much more advanced ways of performing this task.
-The generally accepted standard is actually to use a high pass filter, although the
-cutoff frequency is application dependent and often applied haphazardly.
-The high pass filter thus removes the DC component of the signal, and 
-some amount of the so called "drift". The problem with this method is that it is
-not adaptive to the input. Huge variations in drift frequencies can exist 
-in a single time-series. Thus, a single cutoff frequency could miss a significant
-drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
-set too high. This is why, as I mentioned in the De-trending section, a spline
-based detrending method will generally give better results.
-
-For simulated and real images (tests with multiple time-series), tests were 
-also run with and without Gaussian filtering with sigma of %not sure todo
-were run, since it is standard
-practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
-a spatial filter such as Gaussian filtering increased SNR but can also lead to less
-precision in the output maps.
-
-\subsection{Simulation}
-We performed two types of simulations. First, we simulated a single BOLD time-series based
-on a random chosen set of model parameters. This process was relatively straight forward
-given the state-space equations for the BOLD signal. After a "true" signal was generated,
-we then added a carrier level, since BOLD is typically measured as a \% difference from the
-base level. Finally, we added Gaussian noise, and a Weiner Process to the clean signal. The
-variance of the Gaussian noise may be expressed in terms of the desired noise SNR, $R$ as:
-\begin{equation}
-var(y_{noisy}) = var(y) / R
-\end{equation}
-Since SNR doesn't have quite the same meaning for a Wiener process based noise, the variance 
-of the Gaussian steps was set to be:
-\begin{equation}
-var(y_{noisy}) = var(y) / (4R)
-\end{equation}
-Once this noisy simulated time series was generated, the exact same particle filter algorithm
-that would later be run on full sized images, was run on this single voxel image. We ran
-a series of tests to determine the convergence rate of the particle filter, the number
-of particles that were required, how weighting functions compared, how different de-trending
-methods compared with each other and, finally the variance of the result. By running the exact
-time-series with different noise realizations, it was possible to determine the model variance.
-As the reader may know, the error of an estimator may be calculated as:
-\begin{equation}
-MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
-\end{equation}
-The variance is an expression of how much the result would change for different noise realizations,
-whereas the bias is an expression of how well the model matches the true underlying model. In
-this case, because the same model is being used in the particle filter and underlying simulation,
-the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
-underlying state space equation, there will be some amount of bias error, but assuming that the
-noise is similar to the noise used in these tests, the model variance will actually be about the
-same. Thus calculating the model variance is extremely helpful in calculating how well determined
-our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
-thousands present in a real image, makes it easier to 
-compare the output with the ground truth, with various parameters.
-
-Second we used a modified version of the FSL tool 
-POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
-by creating a random image, smoothing it with a large Gaussian kernel, then thresholding
-the results. Finally connected regions were each given a set of parameters from a finite
-list of randomly chosen parameter sets. The result was a four dimensional (length x width
-x height x parameter) image with spatially varying parameters. Time-series of activation level
-was generated for each set of parameters, then activation levels were fed into POSSUM's 
-function for generating frequency domain data. The patche for POSSUM will be made available.
-For each time-series in the simulate FMRI image, the final \emph{static} parameters are saved
-into a parameter map. This parameter map may then be compared to the map used to generate the 
-simulated data; additionally a new simulation using the calculated parameters may also be 
-generated to test the functional difference between the two maps. This would give an absolute 
-quantitative difference between the two parameter sets irrespective to parameter slopiness.
-So for instance, if $V_0$ is halved, $\epsilon$ doubling may very well give a similar result.
-In this case the \% difference between the parameters will be large in each case, but the functional
-difference between the parameters will not be great. This is obviously a bad situation, which
-is why we wanted to test for it.
-
-\subsection{Real Data}
-Finally, we also performed inference based on real FMRI data. The scanner we used...
-%... more specifics...
-
-The final result from calculating parameters with the real data was similar to that
-from the results from the POSSUM simulated data. The difference being that there was
-no ground truth the check it with.
-
-\section{Results}
-\subsection{Single Time-Series Simulation}
-
-Graphs:
-
-For simulated data, single timeseries:
-
-For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
-\{100, 500, 1000\} particles
-\begin{enumerate}
-\item Ground truth vs. Estimated signal during particle filter run
-\item Ground truth vs. Estimated signal with final parameter set
-\item True Parameters vs. Final Parameter Sets
-\item Variance of final parameters when faced with same ground truth, different noise
-\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
-\item Estimator Variance based on different noise runs
-\item Final Particle Distribution
-\end{enumerate}
-
-For Simulated Data, Full Volume:
-
-%note to self, epsilon should probably be uniform between 0 and something
-\subsection{Simulated Volume}
-\begin{enumerate}
-\item Parameter Map 
-\item Error map of parameters
-\item Histogram of \%errors between parameters
-\item Activation Map based on a single region with high $\epsilon$, compared with linear
-\end{enumerate}
-
-\subsection{FMRI Data}
-....
-
-image comparing epsilon-map with GLM activation map
-
-\section{Conclusion}
+\include{intro}
+\include{background}
+\include{particlefilters}
+\include{methods}
+\include{results}
+\include{conclusion}
 

### 2010-06-28 doc/thesis/background.tex
updated images, more work on background methods

\chapter{Current Techniques}
The ultimate goal of this project is to provide a new set of tools
for analyzing FMRI data. Whereas SPM techniques have been highly 
successful at finding macroscopic regions of activation, linear 
modeling can carry significant bias error due to lack of model
flexibility. While adding parameters can significantly increase
error due to model variance, this effect should be mitigated first
by the use of a highly constrained model based on first principals
and also by the
calculation of a full posterior distribution, rather than a single
estimate. The
purpose of this paper is thus to evaluate the potential of using
a particle filter along with the BOLD model to derive physical 
parameters. In so doing, we hope to be able to show that one or more
parameters are a suitable replacement for estimating voxel 
activation from a standard FMRI image. We also hope to show that 
estimated posterior distribution of the parameters, derived from
the particle filter, is able to provide an accurate measure of the
confidence interval.

\section{Nonlinear Least Squares}
\label{sec:Nonlinear Least Squares}
Although there are certainly benefits to using a derived model, rather
than a purely empirical model, there are serious implications. The
first problem is that all the powerful classical techniques of gradient
descent are off limits; since the model is a true nonlinear dynamical
system with no closed form solution. The implication of this is that
the calculation of a Jacobian for residuals won't work; and thus powerful
techniques such as the Gauss-Newton method, which are helpful in many nonlinear
problems, are off limits. Additionally, a gradient descent is difficult to
perform without the ability to form partials of the output with respect
to all the parameters.

Although anything requiring a Jacobian is out, there are other heuristic
techniques that could potentially illuminate the BOLD response. 
Simulated Annealing (SA) is a common method of optimizing high dimensional
regression problems. The idea is to pick a random start, and then
at each iteration pick a random nearby point, and if that point is
below some energy constraint (energy is a function of the residual), 
called the temperature, the algorithm moves
to that point and continues with the next iteration. The temperature
is slowly lowered until no nearby points below the temperature can
be found (or the temperature drops below the current point). There are
variations of this, for instance it is common to require every movement
to be in the downward direction (in terms of energy). Like most nonlinear
optimization problems, there is no guarantee of an optimal solution,
although the longer the algorithm is allowed to run, the better the solution
will get. Since every step requires an entirely new run of the 
BOLD model, it can be extremely time consuming, which is why we are
not using it here.

\begin{algorithm}
\caption{Simulated Annealing Algorithm}
\label{alg:Simulated Annealing}
\begin{algorithmic}
\STATE Initialize $\Theta$, or if there exists a decent estimate start there
\STATE Initialize temperature, T to value above initial energy
\WHILE{$E(\Theta) < T$}
    \REPEAT
        \STATE Pick $\theta$ near $\Theta$
        \STATE Calculate energy, $E$, of $\theta$
    \UNTIL{$E > T$}
    \STATE Move to new estimate: set $\Theta = \theta$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

[simulated annealing image?]

Another potential method of interest is the use of Genetic Algorithms
(GA). Genetic algorithms are similar to Simulated Annealing, in
that they randomly move to better solutions based on a cost function.
However; in genetic algorithms a single point estimate isn't used. Instead
a population of estimates is generated, each with distinct parameters,
and then each set of parameters is rated with a fitness function. Parameter
sets that are good get a higher "fitness"; then new parameter sets are generated by 
randomly combining pieces of the old parameter sets. The pieces are typically
chosen at a rate proportional to the fitness of the donor; thus "fit"
parameter sets tend to pass on their properties. In addition to this,
random mutations may be introduced that come from no existing parent. 
The new generation is then rated with the fitness function again, and
the entire process starts over. The stop condition for a genetic algorithm
is typically based on some threshold for fitness or a maximum number 
of generations. This is problematic, since it doesn't really provide 
any guarantee of even reaching a local minima, although in practice
it can be quite effective.

[genetic algorithm picture]

\begin{algorithm}
\caption{Genetic Algorithm}
\label{alg:Genetic Algorithm}
\begin{algorithmic}
\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
\FOR{G generations}
    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
    \FOR{$i$ in $N$}
        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
        \STATE At some low probability change one or two parameters in $\Theta_i$
    \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

Although both these methods can be highly effective, they have the downside of
requiring large amounts of computation time. In this case of the BOLD model,
each time the energy or fitness needs to be calculated, a large number of cycles
must be spent re-simulating the BOLD model for the set of parameters. As we'll
discuss in \autoref{sec:Particle Filters}, the Particle Filter method is able
to circumvent this re-calculation to some degree.

It would not be unreasonable at this point to back off and work with a linearized
or more static version of the model. This is the approach taken by the
standard General Linear Model discussed in \autoref{sec:Current Techniques General Linear Model}.

\section{Basic Statistical Parametric Mapping}
Although not strictly the same thing as parameter calculation from 
FMRI, activation detection is very similar. In fact, estimation of 
parameters is somewhat a generalization of the idea of activation detection.
Therefore, it is important to draw a comparison between the methods proposed
in this thesis with existing methods of activation detection.

The most basic method of analyzing FMRI data is through a standard T-test
between "resting state" and "active state" samples. This is done by 
taking the average and variance of the inactive period, and the 
period during which the stimulus was activate separately then treating 
them both as gaussian distributions.
If they are in fact Gaussian distributions, then a basic t-test will
give the probability that the samples came from the same distribution
(the null hypothesis). Of course, this test is fraught with problems; even if
the drift mentioned earlier has been removed, there is little reason
to believe that the noise is Gaussian, or even stable. Additionally, 
even if the noise were Gaussian, a t-test with a p-value of .05 over
50000 or more samples is on average going to generate $.05*50000$ false
positives. To compensate for this, bonferoni correction, also known as
multiple comparison tests are performed; essentially p-values are 
divided by the number of independent
tests being run. This, however, leads to extremely low p-values, so
low that it would be impossible for any biological system to satisfy. To
compensate, a Gaussian kernel is applied to the image, thus reducing
variance (and thus separating the active and inactive distributions)
as well as decreasing the effective number of voxels. Since t-tests are
now no longer being applied to n <I need to define n> independent voxels,
the factor by which the p-value must be divided by can be decreased.
<Do I need to mathematically define all this?> The derivation and application
of random field theory, and its use can be found in various papers \cite{Worsley2004}.

\subsection{General Linear Model}
\label{sec:Current Techniques General Linear Model}
The most used form of FMRI analysis is Statistical Parametric
Mapping, but is able to account for several different levels or types
of stimulus (see \cite{Hofmann1997}). By using hierarchical models
the output signal timeseries is considered the weighted
sum of the various input timeseries. Essentially every experimental factor
is considered as another level inputs (ex. different type of stimulus,
different run with the same patient, different patient). 
The equation for a general linear model is then
\begin{equation}
Y(t) = X(t)\beta + \epsilon(t)
\end{equation}
where $Y(t)$ is the smoothed or detrended timeseries of measurements,
$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
and $\epsilon$ is the error. Thus for every time, the measurement is
assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
of $\beta$ is then performed using a maximum likelihood or gradient descent search 
to minimize the error.

\begin{figure}
\caption{GLM todo}
\label{fig:GLM}
\end{figure}

It is well known of course that a square wave stimulus does not result in a square wave
in the activation of brain regions. Thus, various methods are used to 
smooth the columns of $X$, and thus bandlimit the input. 
The best technique is convolving the stimulus input with a Hemodynamic 
Response Function (HRF), which mimics the basic shape of BOLD activation, including a delay
due to rise time and fall time. The downside of this method, however is that 
the shape of the Hemodynamic Signal is static, meaning the same Hemodynamic Function is
used for every region of the brain, and $Y(t)$ must linear combination 
of scalar the columns of $X$ to be appropriately identified. Additionally, 
it is well known that different Hemodynamic Response Functions are necessary for different 
regions of the brain. The "Canonical" HRF that is most used, has been fitted
for the visual cortex. Thus the shape of the HRF may differ significantly in
terms of onset and fall time in other areas of the brain \cite{Handwerker2004}. 
The inability to
fit parameters other than scale certainly hinders SPM's
ability to locate voxels that do not conform to the "canonical" hemodynamic 
response function. If there are different HRF's for different regions 
of the brain, might there also be more subtle differences even within
a particular region? It seems reasonable to think so. As a result, the 
most common use of SPM will be heavily biased toward the visual cortex.

\begin{figure}
\caption{Hemodynamic Response Function todo}
\label{fig:HRF}
\end{figure}

The GLM is extremely powerful at determining the linear dependence of
a set of regressors on the output. Unfortunately, there is significant evidence
that many of these dependencies are nonlinear, which means they may be
difficult or impossible to detect using linear techniques. This presents
a significant and unknown problem that is often left un-addressed in 
neural studies.

A static linear model is also unable to incorporate other forms of physiological
data that may be available. Combined FMRI
CBF or CBV imaging methods are not uncommon and they could shed much light on
neural activation. However, there is no way of actively combining that data 
into a static HRF. The fact of the matter is that the relation between the 
BOLD signal and CBV/CBF simply cannot be described by any linear relationship.
While it may be possible to determine some regionally varying set of parameters
for the HRF based on CBV/CBF, the lack of physiologically inspired parameters
impairs this. The importance of regional differences in the HRF cannot be
over-stated; it is well known that capillary beds are far from uniformly distributed
and thus blood perfusion and regional oxygenation cannot possibly be uniform
either.

Finally, all these techniques require the noise to be Gaussian to reach
an optimal solution. In fact there is no known optimal solution to
nonlinear models with non-Guassian noise, so obviously some assumptions
are going to have to be made to reach a solution. However, it would be
nice to have an algorithm that is robust to these effects, and could
still give a good solution when Gaussianity is violated.

Because of these limitations, its entirely possible that activation exists
in regions that SPM doesn't detect, but because the activation does not conform
to the set HRF, it is impossible for these signals to be detected. Perhaps
because of this, it is not uncommon for data to be thrown out or averaged
together in FMRI studies because 
no significant activation could be seen in a single run \cite{Riera2004}
\cite{Johnston2008}. This practice highlights 
limitations in strictly linear approaches; and suggests that a single HRF is 
insufficiently flexible to account for relatively common variations of neural activation
\cite{Handwerker2004}.

In general activation detection type methods also don't have the ability 
to find pathologies based on state variables or parameters.  It
is quite possible that physical properties such as decreased compliance of
blood vessels could indicate a neurological condition that is not easily
seen in a T1 or T2 map. In essence, this could make FMRI a much more 
useful clinical tool than it is now.

\section{Unscented Kalman Filter}
Although the classical methods mentioned in \autoref{sec:Nonlinear Least Squares}
won't work for the nonlinear model presented in \autoref{sec:BOLD Physiology},
there is another Bayesian technique that is worth considering.
The Unscented Kalman Filter (UKF) is a powerful Gaussian/Bayes filter that attempts
to model the posterior distribution of dynamical systems as a multivariate
Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
Filter by allowing the state update to be a function, $g$,

\begin{eqnarray}
X(t) &=& g(u(t), X(t-1))\\
Y(t) &=& h(X(t)
\end{eqnarray}

In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
(often 2 per dimension, plus 1 at the mode of the multivariate distribution)
weighted according to a Gaussian estimate of $X(t-1)$ are passed through
the update equation. This set of points are then used to estimate the 
mean and covariance of $X(t)$. The benefit of this, is that it requires
no Jacobian and only a few extra calculations to get a decent esimate of
a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
closed form solution, and finding the Jacobian is impossible without approximations. Although 
\cite{Riera2004}, \cite{Hu2009} mention a Jacobian of the BOLD response, this is
not strictly the case and is rather $\frac{\partial J}{\partial t}$
rather than a true Jacobian.  This is important because the Extended
Kalman filter depends on the Jacobian to map a Gaussian through the 
advancement of time. Thus the popular Extended Kalman Filter won't work
in this case, whereas the Unscented Kalman Filter still does. In fact
\cite{Hu2009} uses the UKF to perform a similar type of analysis to
the one performed in this work.

\begin{figure}
\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun2005}}
\label{fig:EKFWorking}
\end{figure}

The difficulty
of using a Kalman Filter, however, is that it assumes a multivariate 
Gaussian for the state variables, $X(t-1)$. The problem with this is that 
when a system is significantly nonlinear, the state at 
$X(t)$ will almost certainly be non-Gaussian, and thus estimating
$X(t+1)$ with $X(t)$ as a multivariate Gaussian will perpetually introduce
error in the distribution. Furthermore, it is not really known what 
sort of underlying distributions may exist in such a mixed biological,
mechanical, chemical system such as the brain. Assuming the parameters
all to be Gaussian may in fact be a gross error. On the other hand, for
small variances and short time steps the gaussiann distribution is a good 
fit, and so in some limited cases the Unscented Kalman Filter could work
well. These are 
non-trivial issues given that the assumption of Gaussianity is what allows
the UKF to estimate the posterior using only the first and second moments;
two parameters that don't uniquely describe most distributions.

\begin{table}[t]
\centering
\begin{tabular}{|c || c |}
\hline 
Parameter & Run 1 \\
\hline
$\tau_0$ & .98  \\
$\alpha$ & .33 \\
$E_0$ & .34  \\
$V_0$ & .03  \\
$\tau_s$ & 1.54  \\
$\tau_f$ & 2.46  \\
$\epsilon$ & .54  \\
$V_t$ & N(1, .3)  \\
$Q_t$ & N(1, .3)  \\
$S_t$ & N(1, .3) \\
$F_t$ & N(1, .3) \\
\hline
\end{tabular}
\caption{Parameters used to test Gaussianity of variables after being transitioned through
the BOLD model}
\label{tab:steptable} 
\end{table}

To determine the amount of error incurred in a Gaussian estimate during
a typical sample period, the states of BOLD equations were assigned according
to four dimensional Gaussian. The states were then propagated through two
seconds of simulation (a typical TR in FMRI) and then the resulting
marginal distributions were compared with a Gaussian distribution. The 
purpose is to determine the degree to which the results of simulation will
result in non-Gaussian output, given a Gaussian input. This also demonstrates
the degree of nonlinearity present in the system. The parameters used are 
shown in \autoref{tab:steptable}

Notably $s_t$ has intentionally been set to a non-equilibrium, but physiologically
plausible value. The value of $u$ is left at zero the entire time, so the 
system will decay naturally (see \autoref{sec:BOLD Physiology}), though initializing 
$s$ at a non-zero level will drive the system for several seconds. 
\autoref{fig:transp1s} shows the results when the system is essentially
left on for 100 milliseconds after setting the variables according to \autoref{tab:steptable}.
The Q-Q plots fit very will with a Gaussian, demonstrating that at this short
time interval nonlinearities have not yet begun to effect the distribution.
However, \autoref{fig:trans1s} us the result after 1 second, which is faster
than most FMRI scanners are capable of sampling at. At that range the tails 
of the distributions for $v$ and $q$ are clearly starting to deviate from the
Gaussian distribution. As a result the uncertainty in $y$ is deviating from
the Gaussian distribution as well. This is important, because although 
approximating the distribution with a Gaussian based on the first two
moments will work in the short run, there will be residual error in the
distribution.

On the other hand, this effect is more limited if the initial variance is somewhat
smaller. In tests with those cases, it took much longer for the nonlinearities to skew
the distribution. That result could be encouraging to those looking to use the
UKF, if the distributions are kept relatively thin.

Another potential problem with the UKF is that typically the sigma points,
used to estimate the posterior probability, $P(X(t) | X(t-1), u(t))$, are
located on the main axes. As a result, the covariances are not allowed to 
be effected by the state transition the same way the variances are. While
this may be reasonable in low-dimensional systems, high dimensional systems
have a much greater potential for interplay between the variables. While
this problem is relatively easy to fix (by using more sigma points off
the main axes), there is a definite cost in complexity.

Ultimately, there is a distinct possibility that the nonlinearities of 
the BOLD model make Gaussian estimates unrealistic and thus less effective.
More advanced tests where static variables such as $\alpha$ are varied
as well could shed even more light on the issue. The trouble with using
the UKF then to estimate parameters is that all eleven members of $X$ would
be treated like a single joint Gaussian distribution which certainly 
compound the issues of nonlinearity seen in \autoref{fig:transp1s} and 
\autoref{fig:trans1s}

\begin{figure}
\centering
\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
\caption{Distributions of state variables after simulating for .1s}
\label{fig:transp1s}
\end{figure}

\begin{figure}
\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
\caption{Distributions of state variables after simulating for 1s}
\label{fig:trans1s}
\end{figure}

\section{Variational Filtering}
\cite{Friston2008}, \cite{Friston2008b}

\section{Hybrid Methods}
A large number of hybrid methods have been tried, in order to elicit 
parameter estimates from the BOLD model. \cite{Vakorin2007},
\cite{Johnston2007}, \cite{Hu2009}

\section{Conclusion}
In summary, several other approaches could be taken

### 2010-06-28 doc/thesis/intro.tex
updated images, more work on background methods

\chapter{Introduction}
Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
of neural activity. Despite its rather limited temporal resolution, FMRI is still
the best way of measuring neural activity for the majority of the brain.
Whereas other methods
of analyzing neural signals can be invasive or difficult to acquire, 
FMRI is relatively quick and cheap, and its analysis straight forward.
Because of these benefits, FMRI continues to be crucial to the study of human 
cognition. Despite its prevalence, there have been relatively few developments
in the actual analysis of FMRI images. A steady stream of studies have built
on the original BOLD signal derivation first described in \cite{Ogawa}, 
from the Baloon model first proposed by \cite{Buxton1998}
all the way to full fully autonomous system of equations \cite{Riera2004}. And while
there have been numerous forks in the model, enough in fact to make an entire paper
studying the differences,\cite{Deneux2006}, it is widely known that all these
models have quantitatively less bias error than General Linear Model which is
typically employed today. Then again, depending on the model there may be between
seven \cite{Riera2004} and 50 \cite{Behzadi2005} parameters per voxel to
be optimized. Clearly there is a significant risk of error due to variance
with so many degrees of freedom, not to mention a significantly increased
computation cost. In this thesis I demonstrate the use of a 
particle filter as a means of addressing these problems.

FMRI images as a method of detecting neural activation is based on 
temporal changes of the Blood Oxygen Level Dependent (BOLD) signal.
The BOLD signal is caused by minute changes in the ratio of Deoxygenated
Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain.
Because Deoxygenated hemoglobin is paramagnetic, higher concentrations
attenuate the signal when using T2 weighted imaging techniques, such 
as Echo Planar Imaging (EPI) which is used in FMRI. When axons becomes active,
a large amount of ions quickly flow out of the cell. In order for the action
potential to be used again, an active pumping process moves ions back into the
axon. This process of recharging the axon takes a large amout of enengy, which 
naturally uses oxygen. On a massive scale (cubic millimiter) this activation/recharge
process is happening all
the time; however, it happens at a much higher rate when a portion of
the brain is very active. Thus, blood vessels in a very active area will 
tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin,
resulting in lower FMRI signal. However, to compensate for activation, muscles that
control blood vessels relax in that region to allow more blow flow, which 
in fact results in a higher concentration of oxygenated hemoglobin. Thus,
increased activation actually tends to \emph{increase} the MR signal in
comparison with the base level. It is this overcompensation that is the 
primary signal detected with FMRI imaging. This cascade of events
can, as a consequence of increased activity, increase the local metabolism, 
blood flow, blood volume, and oxygenated hemoglobin; though not necessarily
in sync. The lag between these
various factors is what causes many of the complexities of the BOLD signal.

\section{FMRI}
Magnetic Resonance Imaging, MRI is a method of building 3D images
non-invasively, based on the difference between nuclear spin
relaxation times in various molecules. Initially the entity being
imaged is brought into a large magnetic field which aligns the spins
of molecules in the same direction; radio frequency (RF) signals may
then be used to excite nuclear spin away from the steady alignment. 
As the nuclei precess back to their original orientation, they resonate
at the same RF frequency of their original excitation. Conveniently, the
excitation of nuclear spins return their original state at different
rates, called the T1 relaxation time, depending on the properties of 
the material excited. Additionally, the
coherence of the spins also decay differently (and quite a bit faster
than T1) based on the properties of the region that has been excited.
This gives two primary methods of contrasting substances,
which is the basis of T1 and T2 weighted images. Additionally, there
dephasing occurs at two different rates, the T2 relaxation time,
which is impossible to recover from, and T2$^*$ relaxation, which is
much faster, but possible to recover from with an inversion pulse.
Oftentimes T1 relaxation times can be on the order of seconds if 
a significant excitation pulse is applied. 
In order to rapidly acquire entire brain images, as is done in Functional 
MRI, a single large excitation pulse is applied to the entire brain,
and the entire volume is acquired in a single T1 relaxation period. 
Because the entire k-space (spatial-frequency) volume is acquired 
from a single excitation, the signal to noise ration is very low
in this type  of imaging (Echo Planar Imaging).

Increasing the spatial resolution of EPI imaging necessarily 
requires more time or faster magnetic field switching. Increasing
magnet switching rates though is difficult, because it can result in
more artifacts, or even lower signal to noise ratios. The result is
that at \emph{best} FMRI is capable of 1 second temporal resolution. 
Additionally, the means that each voxel of the image will contain 
the sum of a large amount neurons, capillaries and veins. Thus, the
FMRI signal, which is sensitive to the chemical composition of 
materials, is summing up the composition of various types of tissue
in addition to the blood, whose composition is what we actually care about.
In particular, the presence of Deoxyhemoglobin, Hemoglobin whose
oxygen has been used by a metabolic process, has a decreased magnetic
response compared to Oxygenated Hemoglobin. Thus, capillaries near
very active cells will typically have a higher Deoxyhemoglobin content and
lower signal, and regions with lower activity will have a lower 
Deoxyhamoglobin content and thus higher signal.
Unfortunately, as mentioned previously, blood is only a small part of
each voxel, which means that a single EPI image doesn't tell much about
Deoxyhemoglobin content. However assuming blood is the only thing changing
in the short term, percent difference from a baseline signal \emph{will}
tell us something. FMRI analysis is thus necessarily performed on the percent change
from the baseline. Luckily the assumption that tissue content does not change in
the short run is actually pretty good, although other factors can pollute
the baseline signal, as we will discuss later.

\section{BOLD Physiology}
\label{sec:BOLD Physiology}
It is well known that the two types of hemoglobin act as a contrast agents in 
EPI imaging
\cite{Buxton1998}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
Intuitively, increased 
metabolism will increase Deoxyhemoglobin, however blood vessels are quick
to compensate by increasing local blood flow. Increased inflow will of course
preceed increased outflow, and increased inflow is accomplished by loosening 
capilary beds. Both of these factors drive increased storage capacity.
Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
Hemoglobin, increased volume of blood can certainly effect this ratio if 
metabolism doesn't exactly match the increased inflow of oxygenated blood.
This was the impetus
for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
model (\cite{Mandeville1999}). These models derive from first principals
the increased deoxyhemoglobin ratio and volume of capillaries based on a given flow.
These were the first two attempts to quantitatively account for the shape of the 
BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
and the cerebral blood flow (CBF). In fact \cite{Buxton1998} went to far as
to show that a simple, well chosen blood flow waveform coupled with a square 
wave cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
model, could fully account for the BOLD signal.

Although \cite{Buxton1998} showed that a well chosen flow waveform could 
explain much of the BOLD signal, there was still a matter of proposing a
realistic waveform for the CBF and for the CMRO2. \cite{Friston2000} gave
a reasonable and simple
expressoin for CBF input,$f$, based on a flow inducing signal, $s$, 
\begin{eqnarray}
\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
\dot{f} &=& s
\end{eqnarray}
where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
are both time constants. In \cite{Buxton2004} the final piece of the simple balloon
model was put into place, by describing the CMRO2 as a constant multiple of
the CBF (the inflow of blood). This completed the basic balloon model, and 
was well summarized in \cite{Riera2004}. 
\begin{eqnarray}
\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
\end{eqnarray}
where $v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
local deoxyhemoglobin/oxygenated hemoglobin ratio, $E_0$ is the resting metabolic
rate and $\alpha$ is Grubb's parameter controling the balloon model. 
\cite{Obata2004} refined the readout equation of the BOLD signal based on the
deoxyhemoglobin content (q) and local blood volume (v), resulting in the
final BOLD equation:
\begin{eqnarray}
y   &=& V_0((k_1 + k_2)(1-q) - (k_2 + k_3)(1-v))\\
k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8\\
K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57\\
k_3 &=& \epsilon_0 - 1 = .43
\end{eqnarray}
Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
deoxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
change in relaxation rate with change in blood oxygenation, and
$\epsilon_0 = 1.43$ is the 
ratio of signal MR from intravascular to extravascular at rest. Although,
obviously these constants change with experiment ($TE$, $\nu_0$, $r_0$),
patient, and brain 
region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are used
as constants ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$) in 1.5 Tesla studies..
While this model is in a sense complete, it is far from perfect. The major
problem often brought up with this version of the BOLD model is that it
does not represent the so called "post-stimulus undershoot" well.
The post-stimulus undershoot is the name for a prolonged sub-normal
BOLD response for a period of 10 to 60 seconds after stimulus has
ceased (\cite{Chen2009}, \cite{Mandeville1999a}).

There are two theories for the cause of the post stimulus undershoot. Recall
that a lower than base signal means that there is an increased deoxyhemoglobin
content in the voxel. The first and simplest explanation is that the post-stimulus
undershoot is caused by a prolonged increase in CMRO2 after CBV and CBF
have returned to their base levels. This theory is justified by quite a few
studies that show CBV and CBF returning to the baseline before the BOLD signal
(\cite{Frahm2008}, \cite{Donahue2009}, \cite{Buxton2004}, \cite{Lu2004},
\cite{Shen2008}). Unfortunately, because of limitations on FMRI and en vivo
CBV/CBF measurement techniques it is difficult to isolate whether CBF and
CBV truly have returned to their baseline. Other research seems to indicate
that there can be a prolonged residual supernormal CBV (\cite{Mandeville1999a}, 
\cite{Behzadi2005}, \cite{Chen2009a}), although none of these papers completely
rule out the possiblity of increased CMRO2. Additionally, in \cite{Yacoub2006}, it 
was found that the post-stimulus undershoot varried across regions of the brain, 
which could further explain the contradictions found elsewhere. \cite{Chen2009a}
makes a compelling case that most of the post stimulus undershoot could be 
explained be a prolonged CBV increase, and a prolonged CBF undershoot, and that
many of the previous measurements showing a quick recovery of CBV 
may have been dominated arterial CBV's return to baseline.

Because of the significant possility of a completely independent CMRO2,
extremely complex models for metabolism exist (\cite{Zheng2005}), although 
most recent studies have
focused on their ability to explain the prolonged BOLD post stimulus 
undershoot \cite{Zheng2005}, \cite{Buxton2004}. This is because \cite{Buxton2004}
and later \cite{Riera2004} showed that the main portion of the signal may be 
accurately estimated by a simple blood flow locked expression of the CMRO2.. 
Although \cite{Deneux2006} did not deal extensively with prolonged post
stimulus undershoot, the comparisons made in that publication showed minimal
improvement from separate expressions of CMRO2, in comparison to the much 
increased complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
terms, first proposed in \cite{Buxton2004}, that a slowed return to baseline for
the BOLD signal is possible to model. However, viscoelastic effects primarily
control CBV, which, as mentioned already, many studies have claimed cannot be 
responsible for the BOLD post-stimulus undershoot. Another extensive model
that attempts to quantify the post-stimulus undershoot is the compliance model 
proposed by \cite{Behzadi2005}. Although through a somewhat different means than
the \cite{Zheng2005} and \cite{Buxton2004} papers, its possible the increased
model flexibility ultimately is the key reason for the improvements, as opposed
to increased plausibility. Because of these controversies, and because this is
the first time a particle filter has been used for this problem in this way, 
our aim to keep the model simple was best met by using the original balloon
model with the possible addition of the visco-elastic effects from \cite{Buxton2004}.

Even more advanced versions of the Balloon model exist. In fact \cite{Buxton2004}
introduced several additional state variables, including the CMRO2, the O2 extraction
fraction, which is closely related to CMR02, and the neural response, which 
causes the stimulus to decay toward some steady state value. The 
neural response is intended to emulate neural habituation, wherein neurons
become less sensitive to a prolonged stimulus. While these advanced may be more
capable of capturing a more exact version of the BOLD signal, the difference
between the model will often times be below the noise floor. In essence this 
is a classic bias-variance
dilemma: at some point increased model flexibility, and thus variance, is not
worth the decrease in model bias. For now it remains to be seen where this
line may be drawn in the BOLD signal, although \cite{Deneux2006} does not
show a significant improvement from the additional parameters added in 
\cite{Buxton2004}.

\section{Previous Studies of Parameters}
There have been quite a few efforts to quantify the parameters of the
various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
both proposed physiologically reasonable values for the model parameters, 
\cite{Friston2002} was the first paper to calculate the parameters based 
on actual FMRI data. In that paper, Friston et. al. used a variation of
Expectation Maximization to find a normal distribution for the parameters:
\begin{eqnarray}
\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
\tau_s & =&  N(1.54, .25^2)   \nonumber \\
\tau_f & =&  N(2.46, .25^2)   \nonumber \\
\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
\alpha & =&  N(.33 , .45^2 )   \nonumber \\
E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
V_0  & = &  .03 (not\ estimated) \nonumber
\end{eqnarray}
Since then, several other methods of have been used to calculate
BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
likelihood method for innovation processes was used, as described by
\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
Kalman filter, to break the time series into a series of innovations,
for which Maximum Likelihood was performed. While this is in some since the
"right" way to find the solution, it comes with several caveats. First, every
step in parameter space requires a recalculation of all the state variables. With
two or three parameters this is fine, more than that, and calculations could go on
indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
will only be optimal in that case. Third, depending on the nonlinearities present
in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
Unscented Kalman Filter over all the parameters and state variables to find the 
parameter set/variable time series. While this method has the drawback of not necessarily
being optimal, unfortunately there is no general optimal solution to non-linear non
Gaussian models. Hu et. al.'s technique also will run significantly faster than
ML based techniques, since it does not require recalculating the entire timeseries
for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
One potential problem with all these techniques are that the depend heavily on the
priors. The starting point of the parameters could have a huge impact on the
results, and while one can be hopeful that this isn't the reason for the agreement
between \cite{Friston2000} and later results, there is no way to know.

In \cite{Johnston2008}, a hybrid particle filter/gradient
descent algorithm was used to simultaneously derive the static (classicly called
parameters) and dynamic parameters (classically known as state variables).
Essentially a particle filter is used to calculate the state variables at some
time, and then the estimated distribution of the particles was used to find
the most likely set of parameters that would give that distribution of state variables.
\cite{Johnston2008} comes to a very different set of parameter estimates as compared
to the original \cite{Friston2000} guesses:
\begin{eqnarray}
epsilon &=& .069 \pm .014    \nonumber \\
tau_s & =& 4.98 \pm 1.07  \nonumber \\
tau_f & =& 8.31 \pm 1.51   \nonumber \\
tau_0 & =& 8.38 \pm 1.5    \nonumber \\
alpha & =& .189 \pm .004   \nonumber \\
E_0   & =& .635 \pm .072     \nonumber \\
V_0   & =& .0149 \pm .006     \nonumber 
\end{eqnarray}
Notably, the time constants are significantly longer. This could be a result of
some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
depended less on the priors.

Its possible, although unlikely that the balloon model is not possible to learn
without detrimental cost in variance error. I do not think this is the case, however,
and I will work to dispell this possibility in the results with simulated time
series.

\section{Noise}
Thus, despite some small discrepancies, the cause of the BOLD signal is
relatively well known. However, FMRI doesn't detect this happening in one neuron 
or one capillary bed, but rather as the 
aggregate over a space of several millimeters. Though local neurons act
"together" (i.e. around the same time), the density of neurons, the
density of capillaries, and slight differences in activation across 
a particular voxel can all lead to signal attenuation or noise.

A particularly insidious type of noise present in FMRI is a low frequency
drift, often characterized as a Wiener process (\cite{Riera2004}). 
Though not present in all regions, as much as ten to fifteen percent
of voxels can be affected, thus it is prevalent enough to cause significant
inference problems \cite{Tanabe2002}. It is still not
clear what exactly causes this noise noise comes from, although it is possible it is 
the result of magnets heating up, or some distortion in magnetic
fields \cite{Smith2007}. It is clear that this drift signal is not solely
due to a physiological effects, given its presence in cadavers and phantoms 
\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
more prevalent at interfaces between regions,  although
by no means limited to such areas. Often times the noise is written off
as movement, however given that co-registration of volumes to a single time
point is standard, this seems unlikely. Regardless, the problem mandates
the use of some sort of high pass filter to make models useful at all
\cite{Smith2007}.

Given the complexity of the noise, it seemed prudent to characterize
it. In order to do so we took 
sample time-series during resting state, which should theoretically be all noise, and 
analyzed the type of noise present. Both Q-Q plots, and the autocorrelation
are useful in determining the noise distribution observed.
Because most methods (including the one used in this paper)
assume the noise realization are independent of each other, the auto-
correlation is of particular interest. Gaussianity is also a common
assumption made in studies of FMRI data, though that assumption is not
made here. Anytime the assumptions are violated, the error rate will be 
significantly harder to predict and may in fact skyrocket.
Because the noise is often considered 
to have Wiener noise with Gaussian steps, as first described in \cite{Riera2003}, we
also performed the same tests again, but on the steps, rather than the 
direct measurements.

Finally, removal of the so called "drift" is often
performed with some variation of a high pass filter, 
so we checked the noise distribution after applying such a filter (in
this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
Here we wanted to know whether the subtraction of a spline from the
data will result in complete removal of Wiener noise, and whether the resulting
timeseries better conforms to the common assumption of Gaussianity.

\autoref{fig:QQDC} shows the 
results with a regression line fit to the points on the Q-Q plot.
Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
x-axis and the points
on the y-axis come from the same \emph{type} of distribution then all the points will
fall on a line. Differences in the variance will cause the line to have a slope
other than 1, while differences in the expected value will cause the fitted line
to be shifted. In these Q-Q plots, the points are being compared to the standard
Gaussian distribution, so the quality of the line fit determines how closely the points
fit the Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
normalized (changed to percent difference).
\begin{figure}
\centering
\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_29_49_9}}
\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_34_43_24}}
\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_22_38_23}}
\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_37_29_24}}

%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}

\caption{Q-Q Plots of normalized resting state data}
\label{fig:QQDC}
\end{figure}

\begin{figure}
\centering
\subfigure[]{\label{fig:QQDDlta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
\subfigure[]{\label{fig:QQDDlta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
\subfigure[]{\label{fig:QQDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
\subfigure[]{\label{fig:QQDDlta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
\label{fig:QQDelta}
\end{figure}

A Wiener process should still conform to the normal distribution, albeit with a 
variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
are relatively well described by a Gaussian process with a small autocorrelation, 
\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} aren't. In particular the tails don't
seem to fit, which could be well explained by the much more significant drift in the
signal. Thus, its possible a Wiener process does not well describe the error, since overall
the distribution does not tend toward a Gaussian. Still, the steps still seem to
be relatively normal, meaning at the very least that the process is close to Wiener.
As one would expect most of the autocorrelation disappears for the step data, indicating
that the steps are relatively independent. Therefore, the assumption of I.I.D. Normal 
steps is relatively good, even though the accumulated result seems to be less varied than
Wiener.

\begin{figure}
\centering
\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
\caption{Q-Q Plots of resting state data, after the de-trending}
\label{fig:QQSpline}
\end{figure}

[todo analysis after detrending]

\section{Analysis of the BOLD Signal}
Several studies have endeavored to analyze the BOLD equations, which
we will review here. The most obvious analysis is to calculate the
steady state signal, when $u$ is held constant long enough for
all the transients to settle out.

\begin{eqnarray}
s_{ss} &=& 0 \nonumber \\
f_{ss} &=& \tau_f\epsilon u + 1\nonumber \\
v_{ss} &=& (\tau_f\epsilon u + 1)^\alpha\nonumber \\
q_{ss} &=& \frac{(\tau_f\epsilon u + 1)^\alpha}{E_0}(1-(1-E_0)^{1/(\tau_f\epsilon u + 1)})\nonumber \\
y_{ss} &=& V_0((k_1+k_2)(1-q_{ss}) - (k_2+k_3)(1-v_{ss}))
\label{eq:steadystate}
\end{eqnarray}

todo
[Image with two different $\alpha$'s]

[image comparing the results of 10\% changes in various signals]

### 2010-06-28 doc/thesis/methods.tex
updated images, more work on background methods

\chapter{Methods}
This paper describes two types of experiments; first we will cover
simulations, which of course have the benefit of a ground truth, second
we will cover the methods for real data.

\section{Choosing $\hat{Pr}(y(T) | x(T))$}
Choosing a representation of an unknown distribution is certainly tricky,
and so the fact that $\hat{Pr}(y(T) | x(T)) = \nu_d + \nu_y$ means that
there is a significant piece of the algorithm that is based primarily conjecture.
Studies of the noise in FMRI typically attribute noise to a Gaussian random
variable or an additive noise process with Gaussian steps.

\subsection{Classical De-trending}
The non-stationary
aspect of a Weiner process as with $\nu_d$ is difficult to compensate for, and so various methods
have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
demonstrated that this component is prevalent, and may in fact be a characteristic
of FMRI. In some studies, as many as half the voxels benefit from detrending, meaning
that this is certainly a serious barrier to inference. All the existing methods are performed
during the preprocessing stage, rather than as an integral part of analyzing the BOLD
signal. There is no shortage of theories on the "best" method of detrending; however
a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
a spline works the best. The benefit of the spline versus wavelets, high pass 
filtering or other DC removal techniques is that the frequency response is not set.
Rather, the spline is adaptive to the input, having a low cut off if the signal's 
median stays constant but a high cut off frequency if the signal's median tends to
shift heavily over the course of time. In spite of this, the spline will still remove some
amount of signal, just like all of these methods, which is why the the method proposed
in \autoref{sec:Methods Delta Based Inference} was considered.

When using the median-based spline techniques, the resulting signal is will almost
certainly end with a median of zero; in other words half the signal will be above zero,
half below. The problem with this is this is not the "natural" BOLD signal. More specifically,
when the signal is inactive, the BOLD response \emph{should} be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
After removing the spline, the BOLD resting state will be below 0\%.
This is troublesome because it reduces the power the BOLD model. One obvious solution
is to add an arbitrary constant to each BOLD response. Of course this won't scale well
to whole brain analysis. A more effective technique is adding a DC gain to the BOLD
model. Essentially every measurement would then have a constant level added to it. 
Like all the other model parameters, enough measurements should allow the correct
value to be found. On the downside adding another dimension increases the
complexity of the model, for a variable that is relatively obvious to a human.

Thus, a more conventional approach was used. A robust estimator of scale, was used
to determine roughly where the base level was. The Median Absolute Deviation (MAD)
proved to be extremely accurate in determining how much to shift the signal up
by. We tested both methods during the course of analysis, and found that the increase 
in model complexity far outweighted the slight increase in flexibility. Its 
possible that a more accurate method may exist; however, for this case the 
MAD seemed to work perfectly, as \autoref{fig:DesplineQuality} shows.

\begin{figure}
\caption{image of de-spline'd lines with "true" lines}
\label{fig:DesplineQuality}
\end{figure}

\subsection{Delta Based Inference}
\label{sec:Methods Delta Based Inference}
I also propose and test a different method of dealing with the so called "drift". 
Instead of comparing the direct output of the particle filter with the direct
measurement, the algorith compares the change in signal over a single TR. 
In most signal processing cases this would foolish, but that is because the 
general assumption that all noise is high frequency is not the case here.
In fact, every pipeline for the analysis of BOLD signal uses a high pass filter,
but low poss filters are rarely applied, because it is a well known fact that 
most of the signal is in the high frequency range and most of the noise is actually 
in the low frequency range. The particle filter is an 
extremely robust method of inference, and so I would assert that the particle
filter ought to be given as \emph{raw} data as possible. While taking direct measurements
without de-trending would give awful results, using the difference removes the 
DC component and turns a Weiner process into a Gaussian random variable.

\begin{equation}
\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
\label{eq:measass_delta}
\end{equation}

Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
$\nu_d$ is some other additive process, the difference will still be one of a few stable
distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
with a maximum variance of twice the original variance. All the assumptions made originally
for the particle filter hold, and all of the parameters may be distinguished based on
the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
rather than string of direct readings.

\begin{figure}
\label{fig:FrequencyGraphs}
\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
\end{figure}

\section{Preprocessing}
\label{sec:Methods Preprocessing}
As discussed in the section on de-trending, the normal pipeline for analyzing
FMRI involves a great deal of preprocessing. In this paper we make an effort to
minimize any type of preprocessing that will degrade the signal. 
After FMRI data has been acquired it is always necessary to modify the
data in some way to make different runs comparable. Because FMRI signal
levels are unit-less, at the very least it is necessary to convert
the data into \% difference from the baseline. This process removes no data
from signal since it merely subtracting then dividing by a constant. This
is the signal that was input into the delta based particle filter.
Of course there are much more advanced ways of performing this task.
The generally accepted standard is actually to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.
The high pass filter thus removes the DC component of the signal, and 
some amount of the so called "drift". The problem with this method is that it is
not adaptive to the input. Huge variations in drift frequencies can exist 
in a single time-series. Thus, a single cutoff frequency could miss a significant
drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
set too high. This is why, as I mentioned in the De-trending section, a spline
based detrending method will generally give better results.

For simulated and real images (tests with multiple time-series), tests were 
also run with and without Gaussian filtering with sigma of %not sure todo
were run, since it is standard
practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
a spatial filter such as Gaussian filtering increased SNR but can also lead to less
precision in the output maps.

\section{Simulation}
We performed two types of simulations. First, we simulated a single BOLD time-series based
on a random chosen set of model parameters. This process was relatively straight forward
given the state-space equations for the BOLD signal. After a "true" signal was generated,
we then added a carrier level, since BOLD is typically measured as a \% difference from the
base level. Finally, we added Gaussian noise, and a Weiner Process to the clean signal. The
variance of the Gaussian noise may be expressed in terms of the desired noise SNR, $R$ as:
\begin{equation}
var(y_{noisy}) = var(y) / R
\end{equation}
Since SNR doesn't have quite the same meaning for a Wiener process based noise, the variance 
of the Gaussian steps was set to be:
\begin{equation}
var(y_{noisy}) = var(y) / (4R)
\end{equation}
Once this noisy simulated time series was generated, the exact same particle filter algorithm
that would later be run on full sized images, was run on this single voxel image. We ran
a series of tests to determine the convergence rate of the particle filter, the number
of particles that were required, how weighting functions compared, how different de-trending
methods compared with each other and, finally the variance of the result. By running the exact
time-series with different noise realizations, it was possible to determine the model variance.
As the reader may know, the error of an estimator may be calculated as:
\begin{equation}
MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
\end{equation}
The variance is an expression of how much the result would change for different noise realizations,
whereas the bias is an expression of how well the model matches the true underlying model. In
this case, because the same model is being used in the particle filter and underlying simulation,
the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
underlying state space equation, there will be some amount of bias error, but assuming that the
noise is similar to the noise used in these tests, the model variance will actually be about the
same. Thus calculating the model variance is extremely helpful in calculating how well determined
our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
thousands present in a real image, makes it easier to 
compare the output with the ground truth, with various parameters.

Second we used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by creating a random image, smoothing it with a large Gaussian kernel, then thresholding
the results. Finally connected regions were each given a set of parameters from a finite
list of randomly chosen parameter sets. The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Time-series of activation level
was generated for each set of parameters, then activation levels were fed into POSSUM's 
function for generating frequency domain data. The patche for POSSUM will be made available.
For each time-series in the simulate FMRI image, the final \emph{static} parameters are saved
into a parameter map. This parameter map may then be compared to the map used to generate the 
simulated data; additionally a new simulation using the calculated parameters may also be 
generated to test the functional difference between the two maps. This would give an absolute 
quantitative difference between the two parameter sets irrespective to parameter slopiness.
So for instance, if $V_0$ is halved, $\epsilon$ doubling may very well give a similar result.
In this case the \% difference between the parameters will be large in each case, but the functional
difference between the parameters will not be great. This is obviously a bad situation, which
is why we wanted to test for it.

\section{Real Data}
Finally, we also performed inference based on real FMRI data. The scanner we used...
%... more specifics...

The final result from calculating parameters with the real data was similar to that
from the results from the POSSUM simulated data. The difference being that there was
no ground truth the check it with.

### 2010-06-28 doc/thesis/particlefilters.tex
updated images, more work on background methods

\chapter{Particle Filters}
\label{sec:Particle Filter}
\section{Introduction}
Particle filters, a type of Sequential Monte Carlo (SMC) methods
are a powerful way of estimating the posterior probability distribution
of a set of parameters give a timeseries of measurements. Unlike Markov 
Chain Monte Carlo estimation, Sequential Monte-Carlo methods are designed
to be used with parameters that vary with time. Unlike variations of the
Kalman filter, particle filters do not make the assumption that noise
is Gaussian. Thus particle filters are often the best solution to bayesian 
tracking for non-linear, non-gaussian systems.

\subsection{Model}
\label{sec:Particle Filter Model}
The idea of the particle
filter is to start with a wide mixture PDF of possible parameter sets, 
and then, as measurements come in, to weight more heavily parameter sets 
that tend to give good estimations of the measurements. The reliance on
an initial mixture PDF can introduce bias; however, this effect can be
minimized by alterring the initial weights in the mixture pdf. Of course
every gradient descent must choose starting points and it is often quite
easy to establish a reasonable range of parameters, especially when the
model being used has a physical meaning. Suppose a set or stream of measurements
are given, $\{y(t), t = 1, 2, 3, ... T\}$, where $T$ is permitted to go
 to infinity. Then the goal is to find the 
parameters, $\hat{\theta}$, and underlying state time series, $\hat{x}[0:T]$
that minimize the difference between $\hat{y}[0:T]$
and $y[0:T]$. In our case, we will assume that we know the form
of the model, which is based on first principals, and that
there is some true $\theta$ and a true time-series of underlying
state variable, $x[0:T]$ that drives $y[0:T]$. Assuming a model form 
such as we do here reduces model variance, potentially at the cost of increased
bias (or systematic) error. We will assume a basic state space model:

\begin{equation}
\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x)
\end{equation}

\begin{equation}
y(t) = g(t, x(t), u(t), \theta, \nu_y)
\end{equation}

Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
constants, $u(t)$ is a stimulus, $y(t)$ an observation, and
$\nu_x$ and $\nu_y$ are random variates. Obviously any one of these could
be a vector, so for instance $u(t)$ could encode multiple types of stimuli.

Although not strictly necessary for particle filters, we will make a few
assumptions based on the particular type of systems faced in biological 
processes. First, the systems are assumed to be time invariant. This 
assumption is based on the idea that if you froze the system for $\Delta t$
seconds, when unfrozen the system would continue as if nothing happend. 
Few biological systems are predictable enough for them to be summarized
by a time varying function. Even systems that one might assume work in that
way are actually much more complicated. The heart may seem like an obvious
exception, however, the period between heartbeats vary often enough that prediction
would necessitate another state-space model. In short, we
assume no parameters are time varying, because not enough information exists to
describe any of theme in that way. Luckily particle filters are capable 
of dealing with non-white, non-Gaussian noise, so unanticipated influence
may be re-factored as noise. Secondly we assume that input cannot directly
influence the output, which in the case of the BOLD signal is a good assumption.
%Third, through some sort of preprocessing, we will assume that $\nu_d$ can be
%decreased or completely removed, since it 
Third, we will assume noise is additive, and that $\nu_x$ may be projected into
a Weiner, or other summing process that is additive with $g$ and $\nu_y$, which
will be named $\nu_d$.
Finally, $x(t)$ will encapsulate $\theta$, the unknown model constants, which
means that the vector $\dot{x}$ will always have members
that are 0. The results of these assumptions are a simplified version of the
state space equations:

\begin{equation}
\dot{x}(t) = f(x(t), u(t))
\label{eq:stateass}
\end{equation}

\begin{equation}
y(t) = g(x(t)) + \nu_y + \nu_d
\label{eq:measass}
\end{equation}

Because $\nu_d$ is something akin to and additive Weiner process $y[0:T]$, it 
will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
in $y[0:T]$. For some of the tests, I will use de-trending methods to reduce the effects of 
$\nu_d$, the remainder of which will be re-factored into $\nu_y$. Both $\nu_d$ and $\nu_y$
have biological and non-biological sources. MR can lead to both types of noise, 
as demonstrated in \cite{Smith1999}. Meanwhile changes in metabolism, heart rate, or
other biochemical intervention could all lead to either $\nu_d$ or $\nu_y$.

\subsection{Prior}

The goal of the particle filter is to evolve a probably distribution 
$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
Considering that $y$ contains measurement noise as well as noise from $x$,
 it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
but a true posterior. 
To begin with, the particle filter starts with a prior distribution, and $N_p$
particles need to be drawn from that distribution, $\alpha(x)$:
\begin{equation}
\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
\end{equation}

Where $N_p$ is the number of particles or points used to describe the prior 
using a Mixture PDF. 
\begin{equation}
\hat{Pr}(x(0) = \hat{x}) = \sum_{i=1}^{N_p} w_i\delta(\hat{x} - x_i(0) ) dx
\end{equation}
Where $\delta(x-x_0)$ is 1 if and only if $x = x_0$ (the Kronecker delta function).

If a true prior is preferred, then the weights
should all be $1/N_p$, and since $x_i$ was drawn from the prior, this will
be an approximation of the prior distribution. If a relatively flat prior is 
preferred, then each particle's weight could be divided by the density, $\alpha(x_i)$,
which creates a flat prior with support points in the region of $\alpha(x)$. Either
way, $\alpha(x)$ should be much broader than the true posterior, $Pr(x(0))$, since the
choice of support points is crucial to the convergence of any sampling importance
algorithm. 
For the BOLD signal all the parameters have been studied and have relatively well
known mean and variance, so a prior could be very helpful. We ran simulations for
both normalized and un-normalized priors, although we believe in cases such as this,
where a good prior exists, it should be used. 
For strictly positive parameters (members of $x$) we used a gamma distribution,
whereas for parameters that could be negative, we used a Gaussian distribution. In
both cases standard deviations twice that found in previous studies were used.

Note that all the probabilities implicitly depend on $u[0:T]$, so those terms 
will be left off for simplicity.
Once the probability, $\hat{Pr}(x(T) | x[0:T-1], y[0:T-1])$ has been found
(initially this is just Mixture approximating the prior since no measurements are 
available and no previous probabilities are available), its possible to approximate
the probability for short times between times when measurement is available, by shifting
the probability according the progression of the state equations. This is only 
an approximate, since integrating $\nu_d$ should increase uncertainty as
time without a measurement passes.

\begin{equation}
\hat{Pr}(x(T+\Delta t)) \approx 
\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
\end{equation}

\subsection{Weighting}
When a measurement becomes available it is incorporated into the probability.
This process of incorporating new data is called sequential importance sampling,
and eventually causes the probability to converge. The weight is defined
as

\begin{equation}
w_i(T) \propto \frac{\hat{Pr}(x_i[0:T] | y[0:T])}{q(x_i[0:T] | y[0:T])}
\label{eq:weightfunc}
\end{equation}

where $q$ is called an \emph{importance density}, meaning it decides where
the support points for $x(T)$ are located. To remove the bias due to the
location of the support points, we divide by $q(x_i[0:T] | y[0:T])$. By dividing by 
the posterior density of the support points (particles), the effect of the particle 
distribution may be removed from the posterior density. As a result the weight
is dependent solely based
on $\hat{Pr}(x_i[0:T] | y[0:T])$, the probability of the $i^{th}$ particle's measurements
being different from $y[0:T]$ due to noise alone.
An example of an importance density would be drawing a large
number of points from the standard normal, $N(0,1)$ and then weighting each point, $l$ by 
$1/\beta(l), \beta \sim N(0,1)$. Of course if there is a far off peak in
the posterior that $q$ does not allocate support points in, there will 
be a quantization error, and that part of the density can't be modeled. This is why
it is absolutely necessary that $q$ covers $\hat{Pr}(x_i[0:T] | y[0:T])$.

$q(x_i[0:T] | y[0:T])$ may be simplified by assuming that $y(T)$ doesn't contain 
any information about $x(T-1)$, which is more practical since knowledge of future
measurements is impractical.

\begin{eqnarray}
q(x[0:T] | y[0:T]) & = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T]) \nonumber \\
& = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T-1]) \nonumber \\
& = & q(x(T) | x(T-1), y[0:T])q(x[0:T-1] | y[0:T-1])
\end{eqnarray}

In this paper we will use 
$q(x_i(T) | x_i(T-1), y[0:T]) =  \hat{Pr}(x_i(T) | x_i(T-1))$,
based on the Markov assumption, and the belief that the state space model is 
able to approximate the true state. This means that prior to re-weighting 
particles, the particles will be distributed the same as the previous time but
moved forward according to the integration of $f(x(t), u(t))$.

In addition to $q(x_i(T) | x_i[0:T-1], y[0:T])$, the weight is also based on $Pr(x_i[0:K] | y[0:K])$,
which may be broken up as follows.
\begin{eqnarray}
\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y[0:T], x[0:T])}{\hat{Pr}(y[0:T])} \nonumber \\
 & = & \frac{\hat{Pr}(y(T), x[0:T] | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}}{\hat{Pr}(y(T) | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}} \nonumber \\
 & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x[0:T] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1]) } \nonumber \\
 & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
\end{eqnarray}

Using the assumption that $y(t)$ is fully constrained by $x(t)$ \autoref{eq:measass},
and that $x(t)$ is fully constrained by $x(t-1)$ \autoref{eq:stateass}, we are able to
make the reasonably good assumptions that:
\begin{equation}
\hat{Pr}(y(T) | x[0:T], y[0:T-1]) = \hat{Pr}(y(T) | x(T))
\end{equation}

\begin{equation}
\hat{Pr}(x(T) | x[0:T], y[0:T-1]) = \hat{Pr}(x(T) | x(T-1))
\end{equation}

Additionally, for the particle filter $y(T)$ and $y[0:T-1]$ are 
given, and therefore constant across all particles. Thus $\hat{Pr}(x[0:T] | y[0:T])$
may be simplified to:
\begin{eqnarray}
\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) 
            \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
& = & \frac{\hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
& \propto & \hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])
\end{eqnarray}

Plugging these simplifications into \autoref{eq:weightfunc} leads to:
\begin{eqnarray}
w_i(T) & \propto & \frac{\hat{Pr}(y(T)| x(T)) \cancel{\hat{Pr}(x(T) | x(T-1))} \hat{Pr}(x[0:T-1] | y[0:T-1])}
                         {\cancel{\hat{Pr}(x_i(T) | x_i(T-1))}q(x[0:T-1] | y[0:T-1])} \nonumber \\
& \propto & w_i(T-1)\hat{Pr}(y(T)| x(T)) 
\label{eq:weightevolve}
\end{eqnarray}

Thus, by making the following relatively weak assumptions, evolving a posterior
density  is easy and requires almost no knowledge of noise distribution.
\begin{enumerate}
\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ provide 
a sufficiently flexible model to encapsulate the true time series.
\item $E[\nu_d] = 0$ and $E[\nu_y] = 0$, and $\nu_x = d\nu_d$, $\nu_y$ are stationary
\item The PDF $q(x_i(0))$ (the prior) fully covers $Pr(x_i(0))$
\item Markov Assumption: $Pr(x(T) | x[0:T]) = Pr(x(T) | x(T-1))$
\item $q(x[0:T-1] | y[0:T]) = q(x[0:T-1] | y[0:T-1])$
\end{enumerate}

\subsection{Basic Particle Filter Algorithm}
From the definition of $w_i$, the algorithm sequential importance sampling (SIS) 
is relatively simple.

\begin{algorithmic}
\STATE Initialize $N_p$ Particles: 
        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
\STATE $T$ = \{Set of Measurement Times\}
\FOR{$t$ in $T$}
    \FOR{$i$ in $N_p$}
        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
    \ENDFOR
\ENDFOR

\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
\end{algorithmic}
The result is then a discrete approximation of the posterior distribution.

\subsection{Resampling}
\label{sec:Particle Filter Resampling}
As a consequence 
of the wide prior distribution (required for a proper discretization of a continuous
distribution), there will be many particles with insignificant weights. While this does help
describe the tails of the distribution very well, it means that only a small portion of the
computation will be spent describing the most probable region. Ideally every particle would 
equally decrease the entropy of the distribution, thus the lower the variance of the weights,
the more efficiently the discrete distribution is in describing the continuous distribution. 
A common measure of "Particle Degeneracy" is the effective number of particles, described
in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
of each particle. Of course the true weight is unknown, so a heuristic approximating 
$N_{eff}$ is used:
\begin{equation}
\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
\label{eq:neff}
\end{equation}
Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
$N_{eff}$ drops precipitously.  To alleviate this problem
a common technique known as resampling must be applied. The idea of re-sampling is to 
draw from the approximate posterior, thus generating a replica of the posterior with 
a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
$N_p$ points are drawn from the posterior,
\begin{equation}
\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
\end{equation}
then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
still limited to the original particles, the number of unique particles can only go down.
This effect, often dubbed "particle impoverishment" can result in excessive quantization
errors in the final distribution. However, there is a solution. Instead of sampling from the
discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
that distribution. Because the distribution is continuous, there is no way for a collapse
of the particles to occur. The question then, is how to decide on the smoothing kernel. 
Often times the easiest way to sample from the continous distribution is to break the 
re-sampling down into two steps. First a member of the discrete distribution is randomly
selected based on the weights, and then based on the smoothing a nearby state variable 
is selected. The process of the selection will be defined as:
\begin{equation}
\chi_i = x_i + h\sigma \epsilon
\end{equation}
Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
and $\epsilon$ is drawn from the chosen kernel.
It has been proven that when all the elements of the mixture
have the same weight, as is the case after basic resampling, the kernel that minimizes the 
MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
Particle Filters, C Musso, N Oudjane and F LeGrand). 
\begin{equation}
K = \left\{
\begin{array}{lr}
\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
0 & otherwise
\end{array}\right.
\end{equation}
%<more here>

\subsection{Weighting Function}
Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting function,
is based on an unknown distribution, it is necessary to decide on a function
that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
needs to be centered at zero and have a scale comparable to the signal levels.
Obviously if the actual noise present in $y(t)$ were to be known, then that would
be the best distribution for $\Omega$. In that case, particles that fell
far out on that distribution would be statistically impossible representations
of the system, and it would be completely reasonable to throw such particles away.
While a Gaussian function is the natural choice, because this distribution
and weight are unknown we wanted to try distributions
with wider tails, so that outliers don't completely destroy particle's weights
(and thus convergence proceeds more slowly).

Another natural choice might be one of the robust estimator weight functions, for
example the Huber or bi-square. For the purpose of this work we will stick
with long tailed distributions, however it is worth noting that long tails
may not be the optimal choice in all, or even this situation. The justification
for long tailed distributions is that we believe the noise to be long tailed,
and the variance of the noise is not well known.

Therefore, we tried three weighting functions based on three distributions: Gaussian, 
Laplace and the Cauchy. The standard deviation of the distribution is extremely
important to the convergence of particle filter. A standard deviation that is 
too large will not allow the distribution to converge in any reasonable number of 
measurements. A standard deviation below the standard deviation of the noise 
will cause the algorithm to throw out perfectly acceptable particles. 
The weighting function ultimately will shape the output distribution, $P[y]$, into that
distribution, however the distribution of $x$ will still approach a reasonable
estimate of its true distribution. Even an overly wide weighting function, will
allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
parameters of the "real" posterior distribution, though the scale parameters may 
be overly large.

A reasonable method of setting the standard deviation of $\Omega$ may be by 
taking a small sample from "resting" data and using the sample standard deviation.
Since this is the first attempt at using particle filters for modeling the 
BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
because it gives a more consistency and control. Of course this could be taken
further, by testing the sample data against a set of stock distributions
and choosing the best fit. This of course depends on having enough samples
to make a reasonable inference, which may not always exist.

\section{Simple, Nonlinear Example}
A typical half wave rectifier takes a AC voltage circuit and removes
one half (say the negative half) of the signal. The resulting waveform
is still not DC, however it is then possible to use a capacitor to 
smooth the signal into something similar to DC, as shown in \autoref{fig:HalfWaveIO}.
There are other, more
complex circuits that convert the negative portion into positive and
waste less energy but but here we will keep the system simple.
Thus, let us consider a simple half wave rectifier circuit, shown in 
\autoref{fig:HalfWaveRectifier}.

The half wave rectifier circuit smoothes the gaps between high voltage
with a capacitor. Thus, when $u(t)G$ is less than $v_t$, the circuit will 
discharge the capacitor and maintain a non-zero voltage,
but when $u(t)G$ is greater than $v_t$, the output voltage will be set
by $u(t)G$ and the capacitor will charge up. We will assume a very simple
model for all the components, ignoring the complex nonlinear behavior
that can occur in a diode.

\begin{figure}
\centering
\begin{circuitikz}[scale=2, american]
\draw
 (0,0)  node[transformer core] (T) {}
 (T.A1) -- (-1,0)
 (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
 (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
 (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
 (T.base) node {G}
 (T.B1) to[open, *-*, v=$V_1$] (T.B2); 
\end{circuitikz}
\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
\label{fig:HalfWaveRectifier}
\end{figure}

\begin{figure}
\centering
\caption{Example Input/Output of the Half Wave Rectifier}
\label{fig:HalfWaveIO}
\end{figure}

Although rectifiers are typically thought
of as receiving an AC circuit 60 Hz, we will ignore such specifics and 
assume the voltage across the output of the transformer is simply a scalar
multiple of the input voltage.  As discussed in the \autoref{sec:Particle Filter Model}
any variable with uncertainty must be part of the state variable. Therefore
the state variable will be: $X(t) = \{G, v_t, C, R_m, v_y\}$. Of course,
$u(t)$ cannot be allowed to be a square wave in such a system, since that
such a signal would never get across the transformer and regardless it
would necessitate an unrealistic infinite current across the capacitor.
The state equations would then be

\begin{equation}
v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
        v_y(t-1)\left(1 - \frac{\delta t}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
    \end{cases} 
\end{equation}

To run the particle filter is relatively easy then, since there exists 
a recursive definition of the dynamic state variable, $v_y$. To start
with, an initial distribution must be assumed and while at first a 
Gaussian seems like a good idea, all the static state variables are strictly
positive and thus not well suited to the Gaussian. In this case then,
it would be wise to start with a Gamma distribution, and just be wary of
any standard deviation that gets larger than the prior mean. We will define
the gamma distribution as follows:

\begin{equation}
X \sim Gamma(k, \theta) \rightarrow f(x) = x^{k-1}\frac{e^{-x/\theta}}{\theta^k\Gamma(k)}
\end{equation}

where $\Gamma$ is the gamma function.
The in some ways the margin for error is decided by the weighting function, which
here will define as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
estimate based on all the particles, and 
$v_{yi}$ is the estimate by a particular (i$^{\text{th}}$) particle. The choice of this function is difficult,
and although the Gaussian is typically used, we found the exponential helpful
in dealing with particle deprivation.  The algorithm will then look like the following,

\begin{algorithmic}
\STATE Initialize $N_p$ Particles:
\FOR{$i$ in $N_p$}
    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
    \STATE $v_y = 0$, (Assume the system has been off for a long time)
    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
\ENDFOR
\STATE Run the Filter:
\FOR{$t$ in Set of Measurement Times}
    \FOR{$i$ in $N_p$}
        \STATE $v_{yi}(t) = f(v_{yi}(t-1), u(t))$
        \STATE (All other members of $X_i(t)$ remain the same)
        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
    \ENDFOR
\ENDFOR
\end{algorithmic}

Initially the particles will have the same output, $0$, however, as $u(t)$
changes, the response of each particle to that input will result in different
outputs. Particles that have a $v_{yi}$ near $V_y$ will be weighted higher,
and others farther away will be weighted lower. As the particle filter
runs, weights will compound resulting in a distribution that asymptotically
approaches the true joint distribution of the $X(t)$.  Of course, as we
mentioned in \autoref{sec:Particle Filter Resampling}, particles weighted zero do not significantly
contribute to the empirical distribution, so re-sampling may be necessary.
If the noise is assumed to be Gaussian then it is possible to further optimize. 
Thus we let $h$ be defines as:
\begin{eqnarray}
h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}
\end{eqnarray}
and although it is very possible the underlying noise is non-gaussian, the Gaussian
may work, but sub-optimally. It has been proposed that (Monte Carlo Approximations for
General State-Space Models, markus Hurzeler and Hans R. Kunsch) if the underlying 
distribution is non-Gaussian, then using this bandwidth will oversmooth. 
In reality over smoothing
should not be too great an issue because the smoothing is only being applied to find new
particles. If the distribution is over smoothed then the algorithm may not converge as rapidly;
however, because the bandwidth is still based on particle variance, which will decay as 
particles are ruled out, it is still able to converge. In fact over smoothing is preferrable
to under smoothing, since the latter would result in false negatives, but the previous only
results in a slower decay of the variance. 
At the same time, as $n_x$, the number of dimensions in
$x$, goes to infinity, the standard deviation based approximation becomes less effective
(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
tracking, sanjeev arulampalam, simon maskell, neil gordon...).  Because of the high dimensionality of our system,
and limited measurements, it is helpful to have a broader bandwidth to explore the distribution. 
Nevertheless, because 
of the potentially wide smoothing factor applied by regularized resampling, performing this
step at every measurement would allow particles a great deal of mobility. This mobility is
the enemy of convergence, which is why regularized resampling should only be done when
$\hat{N}_{eff}$ drops very low (say less than 50). Other than the periodic regularized
resampling then, the regularized particle filter is nearly identical to the basic sampling
importance sampling filter (SIS).

 \begin{algorithmic}
\STATE Initialize $N_p$ Particles: 
        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
\STATE $T$ = \{Set of Measurement Times\}
\FOR{$t$ in $T$}
    \FOR{$i$ in $N_p$}
        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
    \ENDFOR

    \STATE Calculate $N_{eff}$ with \autoref{eq:neff}
    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
        \STATE Calculate empirical $\sigma$ 
        \STATE $h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}$
        \STATE Redraw particles using (stratified) basic resampling
        \FOR{$i$ in $N_p$}
            \STATE Draw $\epsilon \sim K$
            \STATE $x_i = x_i + h \sigma \epsilon$
        \ENDFOR
    \ENDIF
\ENDFOR

\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$

 \end{algorithmic}

The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
nearby which allow for more accurate estimation near points of high likelihood. 
As the variance of the estimated
$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.

### 2010-06-28 doc/thesis/results.tex
updated images, more work on background methods

\chapter{Results}
\section{Single Time-Series Simulation}

Graphs:

For simulated data, single timeseries:

For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
\{100, 500, 1000\} particles
\begin{enumerate}
\item Ground truth vs. Estimated signal during particle filter run
\item Ground truth vs. Estimated signal with final parameter set
\item True Parameters vs. Final Parameter Sets
\item Variance of final parameters when faced with same ground truth, different noise
\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
\item Estimator Variance based on different noise runs
\item Final Particle Distribution
\end{enumerate}

For Simulated Data, Full Volume:

%note to self, epsilon should probably be uniform between 0 and something
\section{Simulated Volume}
\begin{enumerate}
\item Parameter Map 
\item Error map of parameters
\item Histogram of \%errors between parameters
\item Activation Map based on a single region with high $\epsilon$, compared with linear
\end{enumerate}

\section{FMRI Data}
....

image comparing epsilon-map with GLM activation map

### 2010-06-28 doc/thesis/background.tex
revising the intro

@@
 \chapter{Current Techniques}
-The ultimate goal of this project is to provide a new set of tools
-for analyzing FMRI data. Whereas SPM techniques have been highly 
-successful at finding macroscopic regions of activation, linear 
-modeling can carry significant bias error due to lack of model
+The ultimate purpose of this work is to provide a new set of tools
+for analyzing FMRI data. Whereas esisting techniques have been 
+successful at finding macroscopic regions of activation, :hese
+techniques are not robust to significant noise and thus linear 
+modeling carries a significant bias error due to lack of model
 flexibility. While adding parameters can significantly increase
@@
 confidence interval.
-
-\section{Nonlinear Least Squares}
-\label{sec:Nonlinear Least Squares}
-Although there are certainly benefits to using a derived model, rather
-than a purely empirical model, there are serious implications. The
-first problem is that all the powerful classical techniques of gradient
-descent are off limits; since the model is a true nonlinear dynamical
-system with no closed form solution. The implication of this is that
-the calculation of a Jacobian for residuals won't work; and thus powerful
-techniques such as the Gauss-Newton method, which are helpful in many nonlinear
-problems, are off limits. Additionally, a gradient descent is difficult to
-perform without the ability to form partials of the output with respect
-to all the parameters.
-
-Although anything requiring a Jacobian is out, there are other heuristic
-techniques that could potentially illuminate the BOLD response. 
-Simulated Annealing (SA) is a common method of optimizing high dimensional
-regression problems. The idea is to pick a random start, and then
-at each iteration pick a random nearby point, and if that point is
-below some energy constraint (energy is a function of the residual), 
-called the temperature, the algorithm moves
-to that point and continues with the next iteration. The temperature
-is slowly lowered until no nearby points below the temperature can
-be found (or the temperature drops below the current point). There are
-variations of this, for instance it is common to require every movement
-to be in the downward direction (in terms of energy). Like most nonlinear
-optimization problems, there is no guarantee of an optimal solution,
-although the longer the algorithm is allowed to run, the better the solution
-will get. Since every step requires an entirely new run of the 
-BOLD model, it can be extremely time consuming, which is why we are
-not using it here.
-
-\begin{algorithm}
-\caption{Simulated Annealing Algorithm}
-\label{alg:Simulated Annealing}
-\begin{algorithmic}
-\STATE Initialize $\Theta$, or if there exists a decent estimate start there
-\STATE Initialize temperature, T to value above initial energy
-\WHILE{$E(\Theta) < T$}
-    \REPEAT
-        \STATE Pick $\theta$ near $\Theta$
-        \STATE Calculate energy, $E$, of $\theta$
-    \UNTIL{$E > T$}
-    \STATE Move to new estimate: set $\Theta = \theta$
-\ENDWHILE
-\end{algorithmic}
-\end{algorithm}
-
-[simulated annealing image?]
-
-Another potential method of interest is the use of Genetic Algorithms
-(GA). Genetic algorithms are similar to Simulated Annealing, in
-that they randomly move to better solutions based on a cost function.
-However; in genetic algorithms a single point estimate isn't used. Instead
-a population of estimates is generated, each with distinct parameters,
-and then each set of parameters is rated with a fitness function. Parameter
-sets that are good get a higher "fitness"; then new parameter sets are generated by 
-randomly combining pieces of the old parameter sets. The pieces are typically
-chosen at a rate proportional to the fitness of the donor; thus "fit"
-parameter sets tend to pass on their properties. In addition to this,
-random mutations may be introduced that come from no existing parent. 
-The new generation is then rated with the fitness function again, and
-the entire process starts over. The stop condition for a genetic algorithm
-is typically based on some threshold for fitness or a maximum number 
-of generations. This is problematic, since it doesn't really provide 
-any guarantee of even reaching a local minima, although in practice
-it can be quite effective.
-
-[genetic algorithm picture]
-
-\begin{algorithm}
-\caption{Genetic Algorithm}
-\label{alg:Genetic Algorithm}
-\begin{algorithmic}
-\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
-\FOR{G generations}
-    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
-    \FOR{$i$ in $N$}
-        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
-        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
-        \STATE At some low probability change one or two parameters in $\Theta_i$
-    \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-Although both these methods can be highly effective, they have the downside of
-requiring large amounts of computation time. In this case of the BOLD model,
-each time the energy or fitness needs to be calculated, a large number of cycles
-must be spent re-simulating the BOLD model for the set of parameters. As we'll
-discuss in \autoref{sec:Particle Filters}, the Particle Filter method is able
-to circumvent this re-calculation to some degree.
-
-It would not be unreasonable at this point to back off and work with a linearized
-or more static version of the model. This is the approach taken by the
-standard General Linear Model discussed in \autoref{sec:Current Techniques General Linear Model}.
 
@@
 useful clinical tool than it is now.
+
+\section{Nonlinear Least Squares}
+\label{sec:Nonlinear Least Squares}
+Rather than localizing activation, by using the physiologically
+plausible BOLD model, it is possible to determine the values of
+governing parameters.
+
+Although there are certainly benefits to using a derived model, rather
+than a purely empirical model, there are serious implications. The
+first problem is that all the powerful classical techniques of gradient
+descent are off limits; since the model is a true nonlinear dynamical
+system with no closed form solution. The implication of this is that
+the calculation of a Jacobian for residuals won't work; and thus powerful
+techniques such as the Gauss-Newton method, which are helpful in many nonlinear
+problems, are off limits. Additionally, a gradient descent is difficult to
+perform without the ability to form partials of the output with respect
+to all the parameters.
+
+Although anything requiring a Jacobian is out, there are other heuristic
+techniques that could potentially illuminate the BOLD response. 
+Simulated Annealing (SA) is a common method of optimizing high dimensional
+regression problems. The idea is to pick a random start, and then
+at each iteration pick a random nearby point, and if that point is
+below some energy constraint (energy is a function of the residual), 
+called the temperature, the algorithm moves
+to that point and continues with the next iteration. The temperature
+is slowly lowered until no nearby points below the temperature can
+be found (or the temperature drops below the current point). There are
+variations of this, for instance it is common to require every movement
+to be in the downward direction (in terms of energy). Like most nonlinear
+optimization problems, there is no guarantee of an optimal solution,
+although the longer the algorithm is allowed to run, the better the solution
+will get. Since every step requires an entirely new run of the 
+BOLD model, it can be extremely time consuming, which is why we are
+not using it here.
+
+\begin{algorithm}
+\caption{Simulated Annealing Algorithm}
+\label{alg:Simulated Annealing}
+\begin{algorithmic}
+\STATE Initialize $\Theta$, or if there exists a decent estimate start there
+\STATE Initialize temperature, T to value above initial energy
+\WHILE{$E(\Theta) < T$}
+    \REPEAT
+        \STATE Pick $\theta$ near $\Theta$
+        \STATE Calculate energy, $E$, of $\theta$
+    \UNTIL{$E > T$}
+    \STATE Move to new estimate: set $\Theta = \theta$
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
+
+[simulated annealing image?]
+
+Another potential method of interest is the use of Genetic Algorithms
+(GA). Genetic algorithms are similar to Simulated Annealing, in
+that they randomly move to better solutions based on a cost function.
+However; in genetic algorithms a single point estimate isn't used. Instead
+a population of estimates is generated, each with distinct parameters,
+and then each set of parameters is rated with a fitness function. Parameter
+sets that are good get a higher "fitness"; then new parameter sets are generated by 
+randomly combining pieces of the old parameter sets. The pieces are typically
+chosen at a rate proportional to the fitness of the donor; thus "fit"
+parameter sets tend to pass on their properties. In addition to this,
+random mutations may be introduced that come from no existing parent. 
+The new generation is then rated with the fitness function again, and
+the entire process starts over. The stop condition for a genetic algorithm
+is typically based on some threshold for fitness or a maximum number 
+of generations. This is problematic, since it doesn't really provide 
+any guarantee of even reaching a local minima, although in practice
+it can be quite effective.
+
+[genetic algorithm picture]
+
+\begin{algorithm}
+\caption{Genetic Algorithm}
+\label{alg:Genetic Algorithm}
+\begin{algorithmic}
+\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
+\FOR{G generations}
+    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
+    \FOR{$i$ in $N$}
+        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
+        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
+        \STATE At some low probability change one or two parameters in $\Theta_i$
+    \ENDFOR
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+Although both these methods can be highly effective, they have the downside of
+requiring very high computation time. In this case of the BOLD model,
+each time the energy or fitness needs to be calculated, a large number of cycles
+must be spent re-simulating the BOLD model for the set of parameters. As we'll
+discuss in \autoref{sec:Particle Filters}, the Particle Filter method is able
+to circumvent this re-calculation to some degree.
+
+It would not be unreasonable at this point to back off and work with a linearized
+or more static version of the model. This is the approach taken by the
+standard General Linear Model discussed in \autoref{sec:Current Techniques General Linear Model}.
 

### 2010-06-28 doc/thesis/intro.tex
revising the intro

@@
 \chapter{Introduction}
-Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
-of neural activity. Despite its rather limited temporal resolution, FMRI is still
-the best way of measuring neural activity for the majority of the brain.
-Whereas other methods
+For the past twenty years Functional Magnetic Resonance Imaging (FMRI) 
+has been at the forefront of cognitive research. Despite it's
+limited temporal resolution, FMRI is the standard tool for localizing 
+neural activation.  Whereas other methods
 of analyzing neural signals can be invasive or difficult to acquire, 
-FMRI is relatively quick and cheap, and its analysis straight forward.
-Because of these benefits, FMRI continues to be crucial to the study of human 
-cognition. Despite its prevalence, there have been relatively few developments
-in the actual analysis of FMRI images. A steady stream of studies have built
-on the original BOLD signal derivation first described in \cite{Ogawa}, 
-from the Baloon model first proposed by \cite{Buxton1998}
-all the way to full fully autonomous system of equations \cite{Riera2004}. And while
-there have been numerous forks in the model, enough in fact to make an entire paper
-studying the differences,\cite{Deneux2006}, it is widely known that all these
-models have quantitatively less bias error than General Linear Model which is
-typically employed today. Then again, depending on the model there may be between
-seven \cite{Riera2004} and 50 \cite{Behzadi2005} parameters per voxel to
-be optimized. Clearly there is a significant risk of error due to variance
-with so many degrees of freedom, not to mention a significantly increased
-computation cost. In this thesis I demonstrate the use of a 
-particle filter as a means of addressing these problems.
+FMRI is relatively quick and cheap, and its analysis relatively straight forward.
+In the past fifteen years, a steady stream of studies have built
+on the original Blood Oxygen Level Dependent (BOLD) signal 
+derivation first described by \cite{Ogawa}.
+The seminal work by \cite{Buxton1998} attempted to explain the
+time evolution of the BOLD signal using a windkessel model to
+describe the time local changes in Deoxygenated Hemoglobin content.
+Various papers made substantial improvements to this model until
+\cite{Buxton2004} brought all the changes together into a single complete 
+set of equations. And while there have been numerous adaptations in the model, 
+many of them summarized in \cite{Deneux2006}, even the most basic version
+has been shown to have less bias error than the convolution based 
+"Canonical Hemodynamic Model" \cite{Deneux2006}, \cite{Handwerker2004}.
+On the other hand many of the BOLD signal models have far more parameters
+than a simple scaling parameter. In fact, the number of parameters
+range from seven \cite{Riera2004} to 50 \cite{Behzadi2005} per
+signal time course; a signal which may be as short as 100 samples long.
+Thus, even in a small FMRI image of the brain (20x20x20), the number of parameters 
+could easily exceed 10,000. Clearly this number of parameters presents
+a significant risk of being under-determined and could suffer catastrophic
+variance error; to make no mention of the computation cost. 
+In this work a method of countering these problem is presented
+with the use of a particle filter.
 
-FMRI images as a method of detecting neural activation is based on 
-temporal changes of the Blood Oxygen Level Dependent (BOLD) signal.
+\section{Overview}
+\label{sec:Introduction Overview}
+Detecting neural activity using the changes in FMRI images is based on 
+the so called Blood Oxygen Level Dependent (BOLD) signal.
 The BOLD signal is caused by minute changes in the ratio of Deoxygenated
 Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain.
-Because Deoxygenated hemoglobin is paramagnetic, higher concentrations
-attenuate the signal when using T2 weighted imaging techniques, such 
-as Echo Planar Imaging (EPI) which is used in FMRI. When axons becomes active,
-a large amount of ions quickly flow out of the cell. In order for the action
-potential to be used again, an active pumping process moves ions back into the
-axon. This process of recharging the axon takes a large amout of enengy, which 
-naturally uses oxygen. On a massive scale (cubic millimiter) this activation/recharge
-process is happening all
-the time; however, it happens at a much higher rate when a portion of
-the brain is very active. Thus, blood vessels in a very active area will 
-tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin,
-resulting in lower FMRI signal. However, to compensate for activation, muscles that
-control blood vessels relax in that region to allow more blow flow, which 
-in fact results in a higher concentration of oxygenated hemoglobin. Thus,
-increased activation actually tends to \emph{increase} the MR signal in
-comparison with the base level. It is this overcompensation that is the 
+Because Deoxygenated Hemoglobin (DHb) is paramagnetic, higher concentrations
+attenuate the signal detected during T2-weighted Magnetic Resonance Imaging (MRI)
+techniques. The most common FMRI imaging technique, due to its rapid repetition 
+time (TR), is Echo Planar Imaging (EPI). When axons becomes active,
+a large amount of ions quickly flow out of the cell. In order for this
+action potential to occur again (and thus for the neuron to fire again),
+an active pumping process must move ions back into the
+axon. This process of recharging the axon requires extra energy, which temporarily
+increases the metabolic rate of oxygen. On a massive scale (cubic millimeter) 
+this activation/recharge process happens continuously. However, when a 
+particular region of the brain is very active, the action potentials
+occur significantly more often, resulting in significant local increase
+of the 
+Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, blood vessels in a very 
+active area will 
+tend to have less oxygenated hemoglobin (due to the increased rate at which
+oxygen is being consumed), and more deoxygenated hemoglobin,
+resulting in an attenuated FMRI signal. In compensation for 
+activation, muscles that
+control blood vessels relax in that region to allow more blood flow,
+which actually overcompensates.
+This ultimately results in lower than average concentration of 
+deoxyhemoglobin. Thus, the BOLD signal consists of a short initial
+dip in the MR signal, followed by a prolonged increase in signal
+that slowly settles out. It is this overcompensation that is the 
 primary signal detected with FMRI imaging. This cascade of events
-can, as a consequence of increased activity, increase the local metabolism, 
-blood flow, blood volume, and oxygenated hemoglobin; though not necessarily
-in sync. The lag between these
-various factors is what causes many of the complexities of the BOLD signal.
+is believed to consist of increased the local metabolism, 
+blood flow, blood volume, and oxygenated hemoglobin. The differences
+in onsets of these various effects is what causes the overcompensation
+that is observable in FMRI. Unfortunately, FMRI has no inherent unit
+of measurement, and thus signal levels are all relative: within a particular
+person, scanner and run.
 
 \section{FMRI}
-Magnetic Resonance Imaging, MRI is a method of building 3D images
+Magnetic Resonance Imaging, MRI, is a method of building 3D images
 non-invasively, based on the difference between nuclear spin
-relaxation times in various molecules. Initially the entity being
-imaged is brought into a large magnetic field which aligns the spins
-of molecules in the same direction; radio frequency (RF) signals may
-then be used to excite nuclear spin away from the steady alignment. 
-As the nuclei precess back to their original orientation, they resonate
-at the same RF frequency of their original excitation. Conveniently, the
+relaxation times in various molecules. First, the subject 
+is brought into a large magnetic field which causes nuclear spins
+to align. Radio Frequency (RF) signals may
+then be used to excite nuclear spin away from the base alignment. 
+As the nuclei precess back to the alignment of the magnetic
+field, they emit detectable RF signals. Conveniently, the
 excitation of nuclear spins return their original state at different
-rates, called the T1 relaxation time, depending on the properties of 
-the material excited. Additionally, the
+rates, called the T1 relaxation time, depending on the atoms excited.
+Additionally, the
 coherence of the spins also decay differently (and quite a bit faster
-than T1) based on the properties of the region that has been excited.
+than T1 relaxation) based on the properties of the region.
 This gives two primary methods of contrasting substances,
-which is the basis of T1 and T2 weighted images. Additionally, there
+which form the basis of T1 and T2 weighted images. Additionally, 
 dephasing occurs at two different rates, the T2 relaxation time,
-which is impossible to recover from, and T2$^*$ relaxation, which is
-much faster, but possible to recover from with an inversion pulse.
-Oftentimes T1 relaxation times can be on the order of seconds if 
-a significant excitation pulse is applied. 
-In order to rapidly acquire entire brain images, as is done in Functional 
+which is unrecoverable, and T2$^*$ relaxation, which is
+much faster, but possible to recover from via special RF signals.
+T1 relaxation times are typically on the order of seconds if 
+a sufficiently strong excitation was applied. 
+In order to rapidly acquire entire brain images, as done in Functional 
 MRI, a single large excitation pulse is applied to the entire brain,
@@
 Because the entire k-space (spatial-frequency) volume is acquired 
-from a single excitation, the signal to noise ration is very low
+from a single excitation, the signal-to-noise-ratio is very low
 in this type of imaging (Echo Planar Imaging).
@@
 that at \emph{best} FMRI is capable of 1 second temporal resolution. 
-Additionally, the means that each voxel of the image will contain 
-the sum of a large amount neurons, capillaries and veins. Thus, the
-FMRI signal, which is sensitive to the chemical composition of 
-materials, is summing up the composition of various types of tissue
-in addition to the blood, whose composition is what we actually care about.
-In particular, the presence of Deoxyhemoglobin, Hemoglobin whose
-oxygen has been used by a metabolic process, has a decreased magnetic
-response compared to Oxygenated Hemoglobin. Thus, capillaries near
-very active cells will typically have a higher Deoxyhemoglobin content and
-lower signal, and regions with lower activity will have a lower 
-Deoxyhamoglobin content and thus higher signal.
-Unfortunately, as mentioned previously, blood is only a small part of
-each voxel, which means that a single EPI image doesn't tell much about
-Deoxyhemoglobin content. However assuming blood is the only thing changing
-in the short term, percent difference from a baseline signal \emph{will}
-tell us something. FMRI analysis is thus necessarily performed on the percent change
-from the baseline. Luckily the assumption that tissue content does not change in
-the short run is actually pretty good, although other factors can pollute
-the baseline signal, as we will discuss later.
+The signal is further diluted because each voxel contains
+the signal from a large number of neurons, capillaries and veins. 
+Thus, the FMRI signal, which is sensitive to the chemical composition of 
+materials, is the average signal from various types of tissue
+in addition to the blood. As mentioned in \autoref{sec:Introduction Overview},
+and explored in depth in \autoref{sec:BOLD Physiology},
+the usefulness of FMRI comes from the discerning of changes in 
+Deoxyhemoglobin/Oxyhemoglobin. Therefore, it is necessary to assume
+that in the short term the only chemical changes will be in
+capillary beds feeding neurons. In practice this may not be the case, for
+instance near significant veins, and it may explain some of the
+noise seen in FMRI imaging (see \autoref{sec:Introduction Noise}. 
+Because MRI lacks units and certain
+areas will have a higher base MR signal, all FMRI studies deal with
+percent change from the base signal; rather than raw values. This
+also removes most of the structural data which is not helpful 
+in determining neural activity.
 
@@
 metabolism will increase Deoxyhemoglobin, however blood vessels are quick
-to compensate by increasing local blood flow. Increased inflow will of course
-preceed increased outflow, and increased inflow is accomplished by loosening 
-capilary beds. Both of these factors drive increased storage capacity.
+to compensate by increasing local blood flow. Increased inflow, accomplished by loosening 
+capillary beds, precedes increased outflow, driving increased 
+blood storage capacity.
 Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
-Hemoglobin, increased volume of blood can certainly effect this ratio if 
+Hemoglobin, increased volume of blood can effect this ratio if 
 metabolism doesn't exactly match the increased inflow of oxygenated blood.
@@
 for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
-model (\cite{Mandeville1999}). These models derive from first principals
-the increased deoxyhemoglobin ratio and volume of capillaries based on a given flow.
+model (\cite{Mandeville1999}). These works derive from first principals
+the changes in deoxyhemoglobin ratio and volume of capillaries based on a given flow.
 These were the first two attempts to quantitatively account for the shape of the 
 BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
-and the cerebral blood flow (CBF). In fact \cite{Buxton1998} went to far as
+and the inward cerebral blood flow (CBF). In fact \cite{Buxton1998} went so far as
 to show that a simple, well chosen blood flow waveform coupled with a square 
@@
 
-Although \cite{Buxton1998} showed that a well chosen flow waveform could 
-explain much of the BOLD signal, there was still a matter of proposing a
+Although \cite{Buxton1998} demonstrated that a well chosen flow waveform could 
+explain most features of the BOLD signal, there was still a matter of proposing a
 realistic waveform for the CBF and for the CMRO2. \cite{Friston2000} gave
 a reasonable and simple
-expressoin for CBF input,$f$, based on a flow inducing signal, $s$, 
+expression for CBF input,$f$, based on a flow inducing signal, $s$, 
 \begin{eqnarray}
@@
 \dot{f} &=& s
+\label{eq:bold1}
 \end{eqnarray}
 where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
-are both time constants. In \cite{Buxton2004} the final piece of the simple balloon
+are both time constants. In \cite{Buxton2004} the final piece of the balloon
 model was put into place, by describing the CMRO2 as a constant multiple of
 the CBF (the inflow of blood). This completed the basic balloon model, and 
-was well summarized in \cite{Riera2004}. 
+was well summarized in \cite{Riera2003}. 
 \begin{eqnarray}
@@
 \dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\label{eq:bold2}
 \end{eqnarray}
-where $v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
+where $v$ is normalized cerebral blood volume (CBV), $q$ is the normalized
 local deoxyhemoglobin/oxygenated hemoglobin ratio, $E_0$ is the resting metabolic
-rate and $\alpha$ is Grubb's parameter controling the balloon model. 
+rate and $\alpha$ is Grubb's parameter controlling the balloon model. 
 \cite{Obata2004} refined the readout equation of the BOLD signal based on the
@@
 Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
-deoxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
+de-oxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
 change in relaxation rate with change in blood oxygenation, and
@@
 ratio of signal MR from intravascular to extravascular at rest. Although,
-obviously these constants change with experiment ($TE$, $\nu_0$, $r_0$),
+these constants change with experiment ($TE$, $\nu_0$, $r_0$),
 patient, and brain 
-region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are used
-as constants ($k_1 + k_2 = 3.4$, and $k_2+k_3 = 1$) in 1.5 Tesla studies..
-While this model is in a sense complete, it is far from perfect. The major
-problem often brought up with this version of the BOLD model is that it
-does not represent the so called "post-stimulus undershoot" well.
-The post-stimulus undershoot is the name for a prolonged sub-normal
+region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are 
+taken as the constants $a_1 = k_1 + k_2 = 3.4$, and $a_2 = k_2+k_3 = 1$ in 
+studies using 1.5 Tesla scanners.
+While this model is very accurate, it is not perfect.
+
+\section{Post Stimulus Undershoot}
+The primary problem with the version of the BOLD model described by 
+\autoref{eq:bold1}
+and \autoref{eq:bold2} is that it does not represent the so called 
+"post-stimulus undershoot" well.
+The "post-stimulus" undershoot is the name for a prolonged subnormal
 BOLD response for a period of 10 to 60 seconds after stimulus has
@@
 CBV/CBF measurement techniques it is difficult to isolate whether CBF and
-CBV truly have returned to their baseline. Other research seems to indicate
-that there can be a prolonged residual supernormal CBV (\cite{Mandeville1999a}, 
+CBV truly have returned to their baseline. Other studies seems to indicate
+that there can be a prolonged supernormal CBV (\cite{Mandeville1999a}, 
 \cite{Behzadi2005}, \cite{Chen2009a}), although none of these papers completely
-rule out the possiblity of increased CMRO2. Additionally, in \cite{Yacoub2006}, it 
-was found that the post-stimulus undershoot varried across regions of the brain, 
-which could further explain the contradictions found elsewhere. \cite{Chen2009a}
-makes a compelling case that most of the post stimulus undershoot could be 
-explained be a prolonged CBV increase, and a prolonged CBF undershoot, and that
+rule out the possibility of increased CMRO2. The discrepancies may in part
+be explained by a spatial dependence in the post-stimulus undershoot; described
+by \cite{Yacoub2006}. \cite{Chen2009}
+makes a compelling case that most of the post stimulus undershoot can be 
+explained by combination of a prolonged CBV increase, and a prolonged CBF 
+undershoot, and that
 many of the previous measurements showing a quick recovery of CBV 
-may have been dominated arterial CBV's return to baseline.
+were in fact showing a return to baseline by arterial CBV.
 
-Because of the significant possility of a completely independent CMRO2,
-extremely complex models for metabolism exist (\cite{Zheng2005}), although 
-most recent studies have
-focused on their ability to explain the prolonged BOLD post stimulus 
-undershoot \cite{Zheng2005}, \cite{Buxton2004}. This is because \cite{Buxton2004}
-and later \cite{Riera2004} showed that the main portion of the signal may be 
-accurately estimated by a simple blood flow locked expression of the CMRO2.. 
+Regardless of the probability that CMRO2 and CBF are detached,
+research into the post-stimulus undershoot has driven the creation
+of much more in depth models. In particular \cite{Zheng2005} accounted
+for much more of the physiology present by adding a compartmental model;
+the resulting fit the BOLD undershoot better than the standard Balloon
+model. 
+for all of which aim to both increase overall
+plausibility while simultaneously enhancing 
+estimation of the post-stimulus undershoot (\cite{Zheng2005}, \cite{Buxton2004}).
+As mentioned in \autoref{sec:BOLD Physiology}, the primary BOLD 
+signal may be accurately estimated by a simple blood flow locked 
+expression of the CMRO2 (\cite{Buxton2004}, \cite{Riera2004}).  
 Although \cite{Deneux2006} did not deal extensively with prolonged post
-stimulus undershoot, the comparisons made in that publication showed minimal
-improvement from separate expressions of CMRO2, in comparison to the much 
-increased complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
+stimulus undershoot, the showed incremental improvements
+in quality from separate expressions of CMRO2, at the cost of extensive
+increases complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
 terms, first proposed in \cite{Buxton2004}, that a slowed return to baseline for
@@
 \cite{Buxton2004}.
+
+These 
+controversies are both important and contentious because they determine 
+the validity of the models being used. Additionally as the next
 

### 2010-06-28 doc/thesis/background.tex
more rework of intro

@@
 confidence interval.
+
+\section{Previous Studies of Parameters}
+There have been quite a few efforts to quantify the parameters of the
+various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
+both proposed physiologically reasonable values for the model parameters, 
+\cite{Friston2002} was the first paper to calculate the parameters based 
+on actual FMRI data. In that paper, Friston et. al. used a variation of
+Expectation Maximization to find a normal distribution for the parameters:
+\begin{eqnarray}
+\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
+\tau_s & =&  N(1.54, .25^2)   \nonumber \\
+\tau_f & =&  N(2.46, .25^2)   \nonumber \\
+\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
+\alpha & =&  N(.33 , .45^2 )   \nonumber \\
+E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
+V_0  & = &  .03 (not\ estimated) \nonumber
+\end{eqnarray}
+Since then, several other methods of have been used to calculate
+BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
+likelihood method for innovation processes was used, as described by
+\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
+Kalman filter, to break the time series into a series of innovations,
+for which Maximum Likelihood was performed. While this is in some since the
+"right" way to find the solution, it comes with several caveats. First, every
+step in parameter space requires a recalculation of all the state variables. With
+two or three parameters this is fine, more than that, and calculations could go on
+indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
+will only be optimal in that case. Third, depending on the nonlinearities present
+in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
+Unscented Kalman Filter over all the parameters and state variables to find the 
+parameter set/variable time series. While this method has the drawback of not necessarily
+being optimal, unfortunately there is no general optimal solution to non-linear non
+Gaussian models. Hu et. al.'s technique also will run significantly faster than
+ML based techniques, since it does not require recalculating the entire timeseries
+for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
+very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
+One potential problem with all these techniques are that the depend heavily on the
+priors. The starting point of the parameters could have a huge impact on the
+results, and while one can be hopeful that this isn't the reason for the agreement
+between \cite{Friston2000} and later results, there is no way to know.
+
+In \cite{Johnston2008}, a hybrid particle filter/gradient
+descent algorithm was used to simultaneously derive the static (classicly called
+parameters) and dynamic parameters (classically known as state variables).
+Essentially a particle filter is used to calculate the state variables at some
+time, and then the estimated distribution of the particles was used to find
+the most likely set of parameters that would give that distribution of state variables.
+\cite{Johnston2008} comes to a very different set of parameter estimates as compared
+to the original \cite{Friston2000} guesses:
+\begin{eqnarray}
+epsilon &=& .069 \pm .014    \nonumber \\
+tau_s & =& 4.98 \pm 1.07  \nonumber \\
+tau_f & =& 8.31 \pm 1.51   \nonumber \\
+tau_0 & =& 8.38 \pm 1.5    \nonumber \\
+alpha & =& .189 \pm .004   \nonumber \\
+E_0   & =& .635 \pm .072     \nonumber \\
+V_0   & =& .0149 \pm .006     \nonumber 
+\end{eqnarray}
+Notably, the time constants are significantly longer. This could be a result of
+some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
+later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
+depended less on the priors.
+
+Its possible, although unlikely that the balloon model is not possible to learn
+without detrimental cost in variance error. I do not think this is the case, however,
+and I will work to dispell this possibility in the results with simulated time
+series.
 

### 2010-06-28 doc/thesis/intro.tex
more rework of intro

@@
 expression for CBF input,$f$, based on a flow inducing signal, $s$, 
+in combination with the original balloon model
+where $v$ is normalized cerebral blood volume (CBV), $q$ is the normalized
+local deoxyhemoglobin/oxygenated hemoglobin ratio.
 \begin{eqnarray}
@@
 \dot{f} &=& s
-\label{eq:bold1}
+\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\label{eq:bold}
 \end{eqnarray}
 where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
-are both time constants. In \cite{Buxton2004} the final piece of the balloon
-model was put into place, by describing the CMRO2 as a constant multiple of
-the CBF (the inflow of blood). This completed the basic balloon model, and 
-was well summarized in \cite{Riera2003}. 
-\begin{eqnarray}
-\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
-\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
-\label{eq:bold2}
-\end{eqnarray}
-where $v$ is normalized cerebral blood volume (CBV), $q$ is the normalized
-local deoxyhemoglobin/oxygenated hemoglobin ratio, $E_0$ is the resting metabolic
+are both time constants, $E_0$ is the resting metabolic
 rate and $\alpha$ is Grubb's parameter controlling the balloon model.
-\cite{Obata2004} refined the readout equation of the BOLD signal based on the
+
+This completed the basic balloon model, and was well summarized again
+in \cite{Riera2003}.  \cite{Obata2004} refined the readout equation 
+of the BOLD signal based on the
 deoxyhemoglobin content (q) and local blood volume (v), resulting in the
@@
 k_3 &=& \epsilon_0 - 1 = .43
+\label{eq:boldout}
 \end{eqnarray}
@@
 \section{Post Stimulus Undershoot}
-The primary problem with the version of the BOLD model described by 
-\autoref{eq:bold1}
-and \autoref{eq:bold2} is that it does not represent the so called 
-"post-stimulus undershoot" well.
+Although the most widely used, the BOLD model described in \autoref{eq:bold}
+and \autoref{eq:boldout} have been extensively added on to. The most
+significant feature missing from the original model is the 
+"post-stimulus undershoot".
 The "post-stimulus" undershoot is the name for a prolonged subnormal
@@
 
-There are two theories for the cause of the post stimulus undershoot. Recall
+Because \autoref{eq:bold} is not capable of producing such a prolonged undershoot,
+additional factors must exist.  Two theories exist for the post stimulus undershoot.
+Recall
 that a lower than base signal means that there is an increased deoxyhemoglobin
@@
 (\cite{Frahm2008}, \cite{Donahue2009}, \cite{Buxton2004}, \cite{Lu2004},
-\cite{Shen2008}). Unfortunately, because of limitations on FMRI and en vivo
+\cite{Shen2008}). Unfortunately, because of limitations on FMRI and in vivo
 CBV/CBF measurement techniques it is difficult to isolate whether CBF and
@@
 Regardless of the probability that CMRO2 and CBF are detached,
-research into the post-stimulus undershoot has driven the creation
-of much more in depth models. In particular \cite{Zheng2005} accounted
-for much more of the physiology present by adding a compartmental model;
-the resulting fit the BOLD undershoot better than the standard Balloon
-model. 
-for all of which aim to both increase overall
-plausibility while simultaneously enhancing 
-estimation of the post-stimulus undershoot (\cite{Zheng2005}, \cite{Buxton2004}).
-As mentioned in \autoref{sec:BOLD Physiology}, the primary BOLD 
-signal may be accurately estimated by a simple blood flow locked 
-expression of the CMRO2 (\cite{Buxton2004}, \cite{Riera2004}).  
-Although \cite{Deneux2006} did not deal extensively with prolonged post
-stimulus undershoot, the showed incremental improvements
-in quality from separate expressions of CMRO2, at the cost of extensive
-increases complexity. \cite{Deneux2006} did show that by simply adding viscoelastic
-terms, first proposed in \cite{Buxton2004}, that a slowed return to baseline for
-the BOLD signal is possible to model. However, viscoelastic effects primarily
-control CBV, which, as mentioned already, many studies have claimed cannot be 
-responsible for the BOLD post-stimulus undershoot. Another extensive model
-that attempts to quantify the post-stimulus undershoot is the compliance model 
-proposed by \cite{Behzadi2005}. Although through a somewhat different means than
-the \cite{Zheng2005} and \cite{Buxton2004} papers, its possible the increased
-model flexibility ultimately is the key reason for the improvements, as opposed
-to increased plausibility. Because of these controversies, and because this is
-the first time a particle filter has been used for this problem in this way, 
-our aim to keep the model simple was best met by using the original balloon
-model with the possible addition of the visco-elastic effects from \cite{Buxton2004}.
+research into the post-stimulus undershoot has led to the creation
+of much more in depth models. In \cite{Zhen2002} additional state
+variables model oxygen transport, whereas \cite{Buxton2004} models
+CMR02 from a higher level, and somewhat more simply; though it 
+still adds several new parameters. \cite{Behzadi2005}
+introduces nonlinearities into the CBF equations as a method to
+explain the post-stimulus undershoot, which falls in line with a 
+prolonged increase in CBF observed in \cite{Chen2009}. Similarly
+\cite{Zheng2005} adds additional compartments to model the BOLD signal
+that result from venous and arterial blood. 
+\cite{Deneux2006} compared these various models and though it did 
+not deal extensively with the 
+post-stimulus undershoot, it did show incremental improvements
+in quality from additional parameters (over the basic Balloon model);
+though at the cost of greatly increased complexity.
+Importantly,\cite{Deneux2006} did show that by 
+simply adding viscoelastic terms from \cite{Buxton2004}, a slowed return 
+to baseline is possible to model, without greatly increasing
+complexity. Regardless, because of many of these models are extremely
+complex, and the benefits have yet to be proved, in this work the simple
+Balloon model will be used. Simplicity is even more important because
+it is the first time a particle filter has been used to calculate the
+BOLD parameters. Of course, future works could certainly benefit from
+the more advanced models, especially with the addition of viscoelastic
+effects.
 
-Even more advanced versions of the Balloon model exist. In fact \cite{Buxton2004}
-introduced several additional state variables, including the CMRO2, the O2 extraction
-fraction, which is closely related to CMR02, and the neural response, which 
-causes the stimulus to decay toward some steady state value. The 
-neural response is intended to emulate neural habituation, wherein neurons
-become less sensitive to a prolonged stimulus. While these advanced may be more
-capable of capturing a more exact version of the BOLD signal, the difference
-between the model will often times be below the noise floor. In essence this 
-is a classic bias-variance
-dilemma: at some point increased model flexibility, and thus variance, is not
-worth the decrease in model bias. For now it remains to be seen where this
-line may be drawn in the BOLD signal, although \cite{Deneux2006} does not
-show a significant improvement from the additional parameters added in 
-\cite{Buxton2004}.
-
-These 
-controversies are both important and contentious because they determine 
-the validity of the models being used. Additionally as the next
-
-\section{Previous Studies of Parameters}
-There have been quite a few efforts to quantify the parameters of the
-various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
-both proposed physiologically reasonable values for the model parameters, 
-\cite{Friston2002} was the first paper to calculate the parameters based 
-on actual FMRI data. In that paper, Friston et. al. used a variation of
-Expectation Maximization to find a normal distribution for the parameters:
-\begin{eqnarray}
-\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
-\tau_s & =&  N(1.54, .25^2)   \nonumber \\
-\tau_f & =&  N(2.46, .25^2)   \nonumber \\
-\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
-\alpha & =&  N(.33 , .45^2 )   \nonumber \\
-E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
-V_0  & = &  .03 (not\ estimated) \nonumber
-\end{eqnarray}
-Since then, several other methods of have been used to calculate
-BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
-likelihood method for innovation processes was used, as described by
-\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
-Kalman filter, to break the time series into a series of innovations,
-for which Maximum Likelihood was performed. While this is in some since the
-"right" way to find the solution, it comes with several caveats. First, every
-step in parameter space requires a recalculation of all the state variables. With
-two or three parameters this is fine, more than that, and calculations could go on
-indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
-will only be optimal in that case. Third, depending on the nonlinearities present
-in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
-Unscented Kalman Filter over all the parameters and state variables to find the 
-parameter set/variable time series. While this method has the drawback of not necessarily
-being optimal, unfortunately there is no general optimal solution to non-linear non
-Gaussian models. Hu et. al.'s technique also will run significantly faster than
-ML based techniques, since it does not require recalculating the entire timeseries
-for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
-very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
-One potential problem with all these techniques are that the depend heavily on the
-priors. The starting point of the parameters could have a huge impact on the
-results, and while one can be hopeful that this isn't the reason for the agreement
-between \cite{Friston2000} and later results, there is no way to know.
-
-In \cite{Johnston2008}, a hybrid particle filter/gradient
-descent algorithm was used to simultaneously derive the static (classicly called
-parameters) and dynamic parameters (classically known as state variables).
-Essentially a particle filter is used to calculate the state variables at some
-time, and then the estimated distribution of the particles was used to find
-the most likely set of parameters that would give that distribution of state variables.
-\cite{Johnston2008} comes to a very different set of parameter estimates as compared
-to the original \cite{Friston2000} guesses:
-\begin{eqnarray}
-epsilon &=& .069 \pm .014    \nonumber \\
-tau_s & =& 4.98 \pm 1.07  \nonumber \\
-tau_f & =& 8.31 \pm 1.51   \nonumber \\
-tau_0 & =& 8.38 \pm 1.5    \nonumber \\
-alpha & =& .189 \pm .004   \nonumber \\
-E_0   & =& .635 \pm .072     \nonumber \\
-V_0   & =& .0149 \pm .006     \nonumber 
-\end{eqnarray}
-Notably, the time constants are significantly longer. This could be a result of
-some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
-later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
-depended less on the priors.
-
-Its possible, although unlikely that the balloon model is not possible to learn
-without detrimental cost in variance error. I do not think this is the case, however,
-and I will work to dispell this possibility in the results with simulated time
-series.
+In summary, there have been extensive attempts to refine the Balloon,
+many of them extremely precise. However, the trade off 
+between increased complexity and increased flexibility is extremely 
+important. Given that the basic BOLD model has 7 parameters, for this
+work, where computation time is very important, that model is
+the best fit.
 
 \section{Noise}
-Thus, despite some small discrepancies, the cause of the BOLD signal is
-relatively well known. However, FMRI doesn't detect this happening in one neuron 
-or one capillary bed, but rather as the 
-aggregate over a space of several millimeters. Though local neurons act
+As demonstrated in \autoref{sec:BOLD Phsyiology} the BOLD response has been
+extensively studied and despite minor discrepancies, the cause of the BOLD 
+signal is well known. However, as FMRI detects an  
+aggregate signal over a space of several millimeters, there are
+plenty of sources of noise. Though local neurons act
 "together" (i.e. around the same time), the density of neurons, the
 density of capillaries, and slight differences in activation across 
-a particular voxel can all lead to signal attenuation or noise.
+a particular voxel can all lead to signal attenuation and noise.
 
-A particularly insidious type of noise present in FMRI is a low frequency
+A particularly difficult form of noise present in FMRI is a low frequency
 drift, often characterized as a Wiener process (\cite{Riera2004}). 
 Though not present in all regions, as much as ten to fifteen percent
-of voxels can be affected, thus it is prevalent enough to cause significant
-inference problems \cite{Tanabe2002}. It is still not
+of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
+inference problems \cite{Smith2007}. It is still not
 clear what exactly causes this noise noise comes from, although it is possible it is 
@@
 \cite{Smith1999}. Interestingly, it is usually spatially correlated, and
-more prevalent at interfaces between regions,  although
-by no means limited to such areas. Often times the noise is written off
-as movement, however given that co-registration of volumes to a single time
+more prevalent at interfaces between regions, though 
+by no means limited to such areas. Though one potential source
+could be slight movement, given that co-registration of volumes to a single time
 point is standard, this seems unlikely. Regardless, the problem mandates
-the use of some sort of high pass filter to make models useful at all
+the use of a high pass filter to make inference reasonably powered
 \cite{Smith2007}.
 
-Given the complexity of the noise, it seemed prudent to characterize
-it. In order to do so we took 
-sample time-series during resting state, which should theoretically be all noise, and 
-analyzed the type of noise present. Both Q-Q plots, and the autocorrelation
+In order to characterize the noise, resting state data was analyzed.
+During resting state, the patient is shown no images, and is asked
+to avoid movement or extensive thought; theoretically rending an 
+image composed entirely of baseline signals and noise.
+%here todo
+ Both Q-Q plots, and the autocorrelation
 are useful in determining the noise distribution observed.

### 2010-06-29 doc/thesis/background.tex
done with properties I think

@@
-\chapter{Current Techniques}
-The ultimate purpose of this work is to provide a new set of tools
-for analyzing FMRI data. Whereas esisting techniques have been 
-successful at finding macroscopic regions of activation, :hese
-techniques are not robust to significant noise and thus linear 
-modeling carries a significant bias error due to lack of model
-flexibility. While adding parameters can significantly increase
-error due to model variance, this effect should be mitigated first
-by the use of a highly constrained model based on first principals
-and also by the
-calculation of a full posterior distribution, rather than a single
-estimate. The
-purpose of this paper is thus to evaluate the potential of using
-a particle filter along with the BOLD model to derive physical 
-parameters. In so doing, we hope to be able to show that one or more
-parameters are a suitable replacement for estimating voxel 
-activation from a standard FMRI image. We also hope to show that 
-estimated posterior distribution of the parameters, derived from
-the particle filter, is able to provide an accurate measure of the
-confidence interval.
+\chapter{Prior Works}
+The ultimate purpose of this work is to provide a new tool
+for analyzing FMRI data.
 
-\section{Previous Studies of Parameters}
+Currently, FMRI is used to determine the location of responses
+to stimuli. The method of finding activation is discussed in 
+\autoref{sec:Statistical Parameteric Mapping}. The goal of this 
+work is to move away from the question "Is this region active"
+and instead ask "What does the activation look like". Answering
+this question necessitates more complex models and certainly
+will result in longer run times. Already there have been several other
+attempts to model the BOLD response; these works will be discussed
+in [all other sections basically].
+
+\section{Statistical Parametric Mapping}
+Although not strictly the same as parameter calculation from 
+FMRI, activation detection is similar and worth discussing. Estimation of 
+parameters is a generalization of the idea of activation detection.
+Given the popularity of Statistical Parametric Mapping (SPM) 
+it is important to draw a distinction between the methods proposed
+in this work.
+
+\subsection{Square Wave Model}
+\label{sec:Square Wave}
+The most basic method of analyzing FMRI data is through a standard T-test
+between "resting state" and "active state" samples. Simply put, the
+mean is calculated separately for non-stimulus and stimulus time intervals.
+A classic t-test may then be applied, giving the probability that the
+distributions actually have the same mean. Because of the correlated
+noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
+to apply some sort of high-pass filter to the data. Without applying
+such a filter, P values must be set extraordinarily high to prevent
+false positives \cite{Smith2007}. If there truly is signal
+due to stimuli, the distributions will not actually be independent
+Gaussians because activation does not fit a square wave 
+(\autoref{sec:BOLD Physiology}). For this reason other methods
+are often more used, as discussed in 
+\autoref{sec:Current Techniques General Linear Model}.
+
+\subsection{Random Field Theory}
+SPM methods make significant use of T-Tests across large regions;
+however, such T-tests work slightly differently than a single
+individual test. A t-test with a p-value of $.05$ over a modestly sized FMRI image,
+say 10,000 voxels, will on average generate 500 false
+positives. This is called the multiple comparison problem. Traditional
+Bonferroni Correction
+deals with this by requiring each test to pass with P
+value of $\frac{.05}{10000}$. The probability of 
+a single false positive would then $.05$. Unfortunately this leads to unrealistically
+low p-values; so low that it would be impossible for any biological system to satisfy. To
+compensate, a Gaussian kernel is applied to smooth the image. This has
+the benefit of reducing the noise variance and 
+decreasing the effective number of independent measurements. Because
+the number of independent measurements is smaller, Bonferroni correction
+can theoretically be applied with a lower scaling factor than
+the original voxel count \cite{Worsley2004}. A side effect of this,
+a single voxel activation is virtually impossible to detect.
+
+\subsection{General Linear Model}
+\label{sec:Current Techniques General Linear Model}
+The most common FMRI analysis technique is SPM, though there
+are more advanced versions that the simple square wave
+method discussed in \autoref{sec:Square Wave}. Hierarchal
+Models are one important improvement that allows researchers
+to combine data across multiple runs, patients and stimuli
+(see \cite{Hofmann1997} for more on Hierarchical Modeling). 
+Hierarchical Models
+concatenate all the data into a single dataset, then perform
+a linear fit between a design matrix and the data. The design matrix
+encapsulates all known experimental factors such as stimuli,
+young/old, etc.  The equation for a general linear model is:
+
+\begin{equation}
+Y(t) = X(t)\beta + \epsilon(t)
+\end{equation}
+
+where $Y(t)$ is the smoothed or de-trended time course of measurements,
+$X(t)$ is the design matrix, $\beta$ is a column vector of weights,
+and $\epsilon$ is the error. Thus for every time, the measurement is
+assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
+of $\beta$ is then performed using a maximum likelihood or gradient descent search 
+to minimize the error.
+
+As mentioned previously, a square wave stimulus 
+does not result in a square wave in the activation of brain regions. 
+The BOLD signal is in fact a significantly smoothed version of the 
+stimuli. As such, when fitting an FMRI time course to the input,
+the input ($X$'s columns) is usually smoothed to reduce bias error. 
+The best method, that maintains a linear fit, is convolving the input 
+with a Hemodynamic 
+Response Function (HRF). The so called "Hemodynamic Response Function"
+mimics the basic shape of BOLD activation, including a delay
+due to rise time and fall time. The fitting 
+process is then a least squares fit over the space of the vector $\beta$. 
+Therefore $Y(t)$ is estimated as a linear combination 
+of the columns of $X$.
+
+Smoothing the input with a single HRF poses certain problems.
+It is well known that different Hemodynamic Response Functions are necessary 
+for different regions of the brain. The "Canonical" HRF that is most used, 
+has been optimized for the visual cortex. As \autoref{sec:Post Stimulus Undershoot}
+discussed, there are certainly variations in the shape of the
+BOLD response, both between brain regions and patients.  \cite{Handwerker2004}
+discusses the implications of choosing an incorrect HRF, the most important
+of which is a definite increase in false negatives. While an atlas
+of Hemodynamic Response Functions for each region could definitely 
+mitigate this risk, it does not deal with variation between patients.
+Thus, the inability to fit parameters other than scale definitely hinders 
+analysis of activation. As a result, there is significant bias error
+when using a linear fit to localize activation. Notably, studies
+will also have a bias toward the visual cortex, precisely because
+it is so well studied.
+
+\begin{figure}
+\caption{Hemodynamic Response Function todo}
+\label{fig:HRF}
+\end{figure}
+
+\subsection{Hierarchical Linear Models}
+As mentioned previously and discussed extensively
+in \cite{Friston2002} and \cite{Hoffman1997}, hierarchical models may be 
+applied to account for systematic differences between subjects. For instance, if the
+study happens to be a mix between young and old, incorporating
+that into the model is wise, regardless if that is the purpose of the test. 
+The reason to do this is to account for additional variance that may not
+be present in similar studies. The Hierarchical form used by
+\cite{Friston2002} is shown in \autoref{eq:Hiearchical}.
+
+\begin{eqnarray}
+\label{eq:Hierarchical}
+Y(t) = X_1(t)\theta_1 + \epsilon_1(t)            \nonumber \\
+\theta_1(t) = X_2(t)\theta_2 + \epsilon_2(t)     \nonumber \\
+...                                              \nonumber \\
+\theta_{n-1}(t) = X_n(t)\theta_n + \epsilon_n(t) 
+\end{eqnarray}
+
+The Empirical Bayes algorithm is used in both in both \cite{Friston2002} 
+and \cite{Hoffman1997}. As a consequence, point estimators 
+are used for each $\theta$, rather than the full distributions.
+
+\subsection{Conclusion}
+In all, the GLM is extremely useful for determining linear 
+dependence of a set of regressors on the output. Unfortunately, as discussed in
+\autoref{sec:BOLD Analysis} there are significant nonlinearities 
+that almost certainly cause false negatives in the Statistical Parametric
+Maps. Unfortunately nonlinear analyses have only recently become feasible,
+so the scale of the problem is still unknown. The problem is 
+highlighted by the relatively common scenario where no significant
+activation can be found in a single run \cite{Riera2004}
+\cite{Johnston2008}.
+
+The static nature of the  linear model also limits its inference power. 
+Besides not permitting HRF differences between patients, there is no
+reasonable way to incorporate other forms of physiological
+data. Combined FMRI CBF or CBV imaging methods are rapidly getting better,
+as seen in \cite{Chan2009}. These techniques could shed light on
+neural activation by providing extra measurements, yet a 
+physiologically reasonable model is necessary for this.
+In reverse, activation detection methods also don't have the ability 
+to identify pathologies based on state variables or parameters. For
+example decreased compliance of
+blood vessels could indicate, or even cause, a neurological condition that 
+is not easily seen in other imaging modalities. Thus, the benefits
+of physiologically meaningful models are manifold.
+
+\section{Previous Studies of the BOLD Model}
+Unlike Statistical Parametric Mapping, the techniques described in this
+section are all attempts to learn some version of the BOLD model. 
+Of particular interest are the calculated distributions for the 
+BOLD parameters, since they are used as prior distributions for
+the particles filter, described in \autoref{sec:Methods}.
+
+%todo
 There have been quite a few efforts to quantify the parameters of the
@@
 Expectation Maximization to find a normal distribution for the parameters:
-\begin{eqnarray}
-\epsilon &=& N(.54 , .1 ^2 )  \nonumber \\
-\tau_s & =&  N(1.54, .25^2)   \nonumber \\
-\tau_f & =&  N(2.46, .25^2)   \nonumber \\
-\tau_0 & =&  N(.98 , .25^2 )   \nonumber \\
-\alpha & =&  N(.33 , .45^2 )   \nonumber \\
-E_0   & =&  N(.34 ,  .1 ^2 )   \nonumber \\
-V_0  & = &  .03 (not\ estimated) \nonumber
-\end{eqnarray}
 Since then, several other methods of have been used to calculate
@@
 to the original \cite{Friston2000} guesses:
-\begin{eqnarray}
-epsilon &=& .069 \pm .014    \nonumber \\
-tau_s & =& 4.98 \pm 1.07  \nonumber \\
-tau_f & =& 8.31 \pm 1.51   \nonumber \\
-tau_0 & =& 8.38 \pm 1.5    \nonumber \\
-alpha & =& .189 \pm .004   \nonumber \\
-E_0   & =& .635 \pm .072     \nonumber \\
-V_0   & =& .0149 \pm .006     \nonumber 
-\end{eqnarray}
 Notably, the time constants are significantly longer. This could be a result of
@@
 series.
-
-\section{Basic Statistical Parametric Mapping}
-Although not strictly the same thing as parameter calculation from 
-FMRI, activation detection is very similar. In fact, estimation of 
-parameters is somewhat a generalization of the idea of activation detection.
-Therefore, it is important to draw a comparison between the methods proposed
-in this thesis with existing methods of activation detection.
-
-The most basic method of analyzing FMRI data is through a standard T-test
-between "resting state" and "active state" samples. This is done by 
-taking the average and variance of the inactive period, and the 
-period during which the stimulus was activate separately then treating 
-them both as gaussian distributions.
-If they are in fact Gaussian distributions, then a basic t-test will
-give the probability that the samples came from the same distribution
-(the null hypothesis). Of course, this test is fraught with problems; even if
-the drift mentioned earlier has been removed, there is little reason
-to believe that the noise is Gaussian, or even stable. Additionally, 
-even if the noise were Gaussian, a t-test with a p-value of .05 over
-50000 or more samples is on average going to generate $.05*50000$ false
-positives. To compensate for this, bonferoni correction, also known as
-multiple comparison tests are performed; essentially p-values are 
-divided by the number of independent
-tests being run. This, however, leads to extremely low p-values, so
-low that it would be impossible for any biological system to satisfy. To
-compensate, a Gaussian kernel is applied to the image, thus reducing
-variance (and thus separating the active and inactive distributions)
-as well as decreasing the effective number of voxels. Since t-tests are
-now no longer being applied to n <I need to define n> independent voxels,
-the factor by which the p-value must be divided by can be decreased.
-<Do I need to mathematically define all this?> The derivation and application
-of random field theory, and its use can be found in various papers \cite{Worsley2004}.
-
-\subsection{General Linear Model}
-\label{sec:Current Techniques General Linear Model}
-The most used form of FMRI analysis is Statistical Parametric
-Mapping, but is able to account for several different levels or types
-of stimulus (see \cite{Hofmann1997}). By using hierarchical models
-the output signal timeseries is considered the weighted
-sum of the various input timeseries. Essentially every experimental factor
-is considered as another level inputs (ex. different type of stimulus,
-different run with the same patient, different patient). 
-The equation for a general linear model is then
-\begin{equation}
-Y(t) = X(t)\beta + \epsilon(t)
-\end{equation}
-where $Y(t)$ is the smoothed or detrended timeseries of measurements,
-$X(t)$ is a row vector of input, $\beta$ is a column vector of weights,
-and $\epsilon$ is the error. Thus for every time, the measurement is
-assumed to be a weighted sum of the columns of $X$ plus some error. The calculation
-of $\beta$ is then performed using a maximum likelihood or gradient descent search 
-to minimize the error.
-
-\begin{figure}
-\caption{GLM todo}
-\label{fig:GLM}
-\end{figure}
-
-It is well known of course that a square wave stimulus does not result in a square wave
-in the activation of brain regions. Thus, various methods are used to 
-smooth the columns of $X$, and thus bandlimit the input. 
-The best technique is convolving the stimulus input with a Hemodynamic 
-Response Function (HRF), which mimics the basic shape of BOLD activation, including a delay
-due to rise time and fall time. The downside of this method, however is that 
-the shape of the Hemodynamic Signal is static, meaning the same Hemodynamic Function is
-used for every region of the brain, and $Y(t)$ must linear combination 
-of scalar the columns of $X$ to be appropriately identified. Additionally, 
-it is well known that different Hemodynamic Response Functions are necessary for different 
-regions of the brain. The "Canonical" HRF that is most used, has been fitted
-for the visual cortex. Thus the shape of the HRF may differ significantly in
-terms of onset and fall time in other areas of the brain \cite{Handwerker2004}. 
-The inability to
-fit parameters other than scale certainly hinders SPM's
-ability to locate voxels that do not conform to the "canonical" hemodynamic 
-response function. If there are different HRF's for different regions 
-of the brain, might there also be more subtle differences even within
-a particular region? It seems reasonable to think so. As a result, the 
-most common use of SPM will be heavily biased toward the visual cortex.
-
-\begin{figure}
-\caption{Hemodynamic Response Function todo}
-\label{fig:HRF}
-\end{figure}
-
-The GLM is extremely powerful at determining the linear dependence of
-a set of regressors on the output. Unfortunately, there is significant evidence
-that many of these dependencies are nonlinear, which means they may be
-difficult or impossible to detect using linear techniques. This presents
-a significant and unknown problem that is often left un-addressed in 
-neural studies.
-
-A static linear model is also unable to incorporate other forms of physiological
-data that may be available. Combined FMRI
-CBF or CBV imaging methods are not uncommon and they could shed much light on
-neural activation. However, there is no way of actively combining that data 
-into a static HRF. The fact of the matter is that the relation between the 
-BOLD signal and CBV/CBF simply cannot be described by any linear relationship.
-While it may be possible to determine some regionally varying set of parameters
-for the HRF based on CBV/CBF, the lack of physiologically inspired parameters
-impairs this. The importance of regional differences in the HRF cannot be
-over-stated; it is well known that capillary beds are far from uniformly distributed
-and thus blood perfusion and regional oxygenation cannot possibly be uniform
-either.
-
-Finally, all these techniques require the noise to be Gaussian to reach
-an optimal solution. In fact there is no known optimal solution to
-nonlinear models with non-Guassian noise, so obviously some assumptions
-are going to have to be made to reach a solution. However, it would be
-nice to have an algorithm that is robust to these effects, and could
-still give a good solution when Gaussianity is violated.
-
-Because of these limitations, its entirely possible that activation exists
-in regions that SPM doesn't detect, but because the activation does not conform
-to the set HRF, it is impossible for these signals to be detected. Perhaps
-because of this, it is not uncommon for data to be thrown out or averaged
-together in FMRI studies because 
-no significant activation could be seen in a single run \cite{Riera2004}
-\cite{Johnston2008}. This practice highlights 
-limitations in strictly linear approaches; and suggests that a single HRF is 
-insufficiently flexible to account for relatively common variations of neural activation
-\cite{Handwerker2004}.
-
-In general activation detection type methods also don't have the ability 
-to find pathologies based on state variables or parameters.  It
-is quite possible that physical properties such as decreased compliance of
-blood vessels could indicate a neurological condition that is not easily
-seen in a T1 or T2 map. In essence, this could make FMRI a much more 
-useful clinical tool than it is now.
 

### 2010-06-29 doc/thesis/intro.tex
done with properties I think

@@
 FMRI is relatively quick and cheap, and its analysis relatively straight forward.
+
+By modeling the governing equations behind the neural response that
+drives FMRI, it is possible to make the analyses of FMRI even more
+powerful. The underlying state equations hold important information
+about how individual brain regions react to stimuli. The model parameters
+on the other hand, hold important information about the patients individual
+physiology including existing and future pathologies. In short,
+the long chain of events driving observable FMRI signal contains more
+information than which regions respond to which stimuli.
+
 In the past fifteen years, a steady stream of studies have built
@@
 \dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
-\dot{f} &=& s
+\dot{f} &=& s\\
 \dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
@@
 \section{Post Stimulus Undershoot}
+\label{sec:Post Stimulus Undershoot}
 Although the most widely used, the BOLD model described in \autoref{eq:bold}
@@
 signal is well known. However, as FMRI detects an  
-aggregate signal over a space of several millimeters, there are
+aggregate signal over the space of cubic centimeters, there are
 plenty of sources of noise. Though local neurons act
@@
 
-In order to characterize the noise, resting state data was analyzed.
-During resting state, the patient is shown no images, and is asked
-to avoid movement or extensive thought; theoretically rending an 
-image composed entirely of baseline signals and noise.
-%here todo
- Both Q-Q plots, and the autocorrelation
-are useful in determining the noise distribution observed.
+In order to characterize the noise, I analyzed resting state data.
+During resting state, the patient is shown no images, and he is asked
+to avoid movement and complex thought, to the best of his abilities.
+Some limitations exist, for instance EPI sequences are extremely loud
+which could cause some stimulus in the auditory cortex, and of course
+very long sequences could give the patient time for his mind to wander.
+Overall though there should be no activation, and thus the signal will
+consist entirely of noise. Therefore resting state data is perfect
+for analyzing the noise distribution. The locations were chosen from
+various points around the brain, all in grey matter voxels (see \autoref{sec:methods}
+for a discussion on how grey matter voxels were found). The time
+series were also chosen because they were representative of different types
+of noise I found in the resting state data.
+
 Because most methods (including the one used in this paper)
-assume the noise realization are independent of each other, the auto-
-correlation is of particular interest. Gaussianity is also a common
+assume the noise realizations are independent of each other, the auto-
+correlation is of particular interest (which is a necessary but not
+sufficient condition for independence). Gaussianity is also a common
 assumption made in studies of FMRI data, though that assumption is not
-made here. Anytime the assumptions are violated, the error rate will be 
-significantly harder to predict and may in fact skyrocket.
-Because the noise is often considered 
-to have Wiener noise with Gaussian steps, as first described in \cite{Riera2003}, we
-also performed the same tests again, but on the steps, rather than the 
-direct measurements.
+made in this work. Regardless, comparing the distribution to a Gaussian
+is informative, so I used Q-Q plots to compare example data with the
+Normal distribution. Additionally, in FMRI data the noise is often considered 
+to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
+characterized by steps that are Gaussian; in other words the difference between
+any two samples is Normally distributed. The simulations discussed in 
+\autoref{sec:Methods Single Voxel Simulation} make use of this, 
+by adding a Wiener random process to the overall signal. To determine
+whether the noise is in fact Wiener, the distribution of 
+the difference between adjacent measurements was plotted against
+a Gaussian. Wiener processes further assume that the steps are independent;
+so the steps should have near flat autocorrelation functions.
 
@@
 performed with some variation of a high pass filter, 
-so we checked the noise distribution after applying such a filter (in
+so I checked the noise distribution after applying such a filter (in
 this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
-Here we wanted to know whether the subtraction of a spline from the
-data will result in complete removal of Wiener noise, and whether the resulting
-timeseries better conforms to the common assumption of Gaussianity.
+Here I wanted to know how effective my high pass filter at the removal of
+Wiener noise.
 
@@
 
-A Wiener process should still conform to the normal distribution, albeit with a 
+A Wiener process should still conform to the normal distribution, with a 
 variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
 are relatively well described by a Gaussian process with a small autocorrelation, 
-\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} aren't. In particular the tails don't
-seem to fit, which could be well explained by the much more significant drift in the
-signal. Thus, its possible a Wiener process does not well describe the error, since overall
-the distribution does not tend toward a Gaussian. Still, the steps still seem to
-be relatively normal, meaning at the very least that the process is close to Wiener.
-As one would expect most of the autocorrelation disappears for the step data, indicating
-that the steps are relatively independent. Therefore, the assumption of I.I.D. Normal 
-steps is relatively good, even though the accumulated result seems to be less varied than
-Wiener.
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
+do not seem to fit the Gaussian well. Also note the significant autocorrelation in
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
+Gaussian white noise.  On the other hand, the steps do conform rather
+closely to the normal distribution.
+As expected, most of the autocorrelation disappears for the step data. Given
+that the steps seem to fit the Normal distribution, the low autocorrelation
+indicates that the steps could be Independently Distributed. 
+Therefore, the assumption that the noise follows a Wiener process seems to
+be correct.
 
@@
 \centering
-\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
-\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
-\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
-\subfigure[]{\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
+\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
+\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
+\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
+\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
 \caption{Q-Q Plots of resting state data, after the de-trending}
@@
 
-[todo analysis after detrending]
+De-trending the time-series by subtracting a spline fit to the distribution
+removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
+though clearly not perfectly. Though the distributions still do not exactly fit
+the Normal, \autoref{fig:QQs:D} clearly is much improved compared to \autoref{fig:QQDC:D}.
+In all, the de-trending is effectively removing Wiener noise.
 
-\section{Analysis of the BOLD Signal}
-Several studies have endeavored to analyze the BOLD equations, which
-we will review here. The most obvious analysis is to calculate the
-steady state signal, when $u$ is held constant long enough for
-all the transients to settle out.
+\section{Properties of the BOLD Model}
+\label{sec:BOLD Analysis}
+Since the first complete BOLD model was proposed by \cite{Friston2002}, 
+several studies have endeavored to analyze properties of that model. 
+The most important property is that the system is dissipative, and given
+enough time will converge to a constant value. This is found simply by
+analyzing the eigenvalues of the Jacobian of the state equations, 
+(\cite{Deneux2006}, \cite{Hu2009}). The steady state of the Balloon
+model equations gives:
 
@@
 
-todo
+In real FMRI data, there is a significant nonlinearity in response; with short sequences
+responding disproportionately strong (\cite{Birn2001}, \cite{Wager2005}, \cite{Deneux2006}).
+This nonlinearity is accounted for in the Balloon model, although \cite{Deneux2006}
+shows that if there will be large variance in the length of signals, 
+modeling Neural Habituation may be necessary to fully capture the range of responses. 
+Stimuli that last longer than 4 seconds 
+tend to be more linear, which is why block designs are so well accounted for
+by the General Linear Model (\cite{Birn2001}, \cite{Deneux2006}). For this 
+paper we will use only a simple version of the Balloon model, described by
+\autoref{eq:bold}, to keep the solution tractable.
+
+Another interesting result of \cite{Deneux2006} was the sensitivity analysis.
+There it was found that many of the parameters compensate for each other,
+implying that exact inference of parameters may not be possible without
+constraint.
+
+Several studies have been able to calculate distributions for the 
+Balloon parameters, those results are summarized in \autoref{tab:Params}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c || c | c | c |}
+\hline 
+Parameter  & \cite{Friston2002} & \cite{Johnston2008} & \cite{Vakorin2007} \\
+\hline
+$\epsilon$ &  $N(.54 , .1 ^2)$  & $.069 \pm .014 $ & (NC)\\
+$\tau_s  $ &  $N(1.54, .25^2)$  & $4.98 \pm 1.07 $ & $2.2$\\
+$\tau_f  $ &  $N(2.46, .25^2)$  & $8.31 \pm 1.51 $ & $.45$\\
+$\tau_0  $ &  $N(.98 , .25^2)$  & $8.38 \pm 1.5  $ & $.94$\\
+$\alpha  $ &  $N(.33 , .45^2)$  & $.189 \pm .004 $ & $.4$ (NC) \\
+$E_0     $ &  $N(.34 , .1 ^2)$  & $.635 \pm .072 $ & $.6$ (NC) \\
+$V_0     $ &  $.03$ (NC)        & $.0149 \pm .006$ & (NC)\\
+\hline
+\end{tabular}
+\caption{Parameters found by various studies. (NC) indicates that the value
+wasn't calculated. \cite{Vakorin2007} made use of the values from \cite{Friston2002}
+where not explicitly stated}
+\label{tab:Params} 
+\end{table}
+
+Necessary?
 [Image with two different $\alpha$'s]
-
 [image comparing the results of 10\% changes in various signals]

### 2010-06-30 doc/thesis/background.tex
variation filtering dropped (DCM), background down.

@@
 
-\subsection{Square Wave Model}
+\subsection{Classical Activation Detection}
 \label{sec:Square Wave}
@@
 
-\section{Previous Studies of the BOLD Model}
+\section{Solving the BOLD Model}
 Unlike Statistical Parametric Mapping, the techniques described in this
 section are all attempts to learn some version of the BOLD model. 
-Of particular interest are the calculated distributions for the 
-BOLD parameters, since they are used as prior distributions for
-the particles filter, described in \autoref{sec:Methods}.
-
-%todo
 There have been quite a few efforts to quantify the parameters of the
@@
 \cite{Friston2002} was the first paper to calculate the parameters based 
-on actual FMRI data. In that paper, Friston et. al. used a variation of
-Expectation Maximization to find a normal distribution for the parameters:
-Since then, several other methods of have been used to calculate
-BOLD parameters from FMRI timecourses. In \cite{Riera2004}, a maximum
-likelihood method for innovation processes was used, as described by
-\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
-Kalman filter, to break the time series into a series of innovations,
-for which Maximum Likelihood was performed. While this is in some since the
-"right" way to find the solution, it comes with several caveats. First, every
-step in parameter space requires a recalculation of all the state variables. With
-two or three parameters this is fine, more than that, and calculations could go on
-indefinitely. Second, it still assumes the parameters and noise are Gaussian, and
-will only be optimal in that case. Third, depending on the nonlinearities present
-in the system, local minima may be extremely common. Later \cite{Hu2009} used an 
-Unscented Kalman Filter over all the parameters and state variables to find the 
-parameter set/variable time series. While this method has the drawback of not necessarily
-being optimal, unfortunately there is no general optimal solution to non-linear non
-Gaussian models. Hu et. al.'s technique also will run significantly faster than
-ML based techniques, since it does not require recalculating the entire timeseries
-for every step in parameters. Both \cite{Hu2009} and \cite{Friston2002} came to results 
-very similar to the expected values stated in \cite{Buxton1998} and \cite{Friston2000}.
-One potential problem with all these techniques are that the depend heavily on the
-priors. The starting point of the parameters could have a huge impact on the
-results, and while one can be hopeful that this isn't the reason for the agreement
-between \cite{Friston2000} and later results, there is no way to know.
+on actual FMRI data. However, in that case, the voxels were chosen from
+regions that were detected as active by the GLM. It is therefore possible
+that the parameters are biased toward parameters that fit the linear model.
 
-In \cite{Johnston2008}, a hybrid particle filter/gradient
-descent algorithm was used to simultaneously derive the static (classicly called
-parameters) and dynamic parameters (classically known as state variables).
-Essentially a particle filter is used to calculate the state variables at some
-time, and then the estimated distribution of the particles was used to find
-the most likely set of parameters that would give that distribution of state variables.
-\cite{Johnston2008} comes to a very different set of parameter estimates as compared
-to the original \cite{Friston2000} guesses:
-Notably, the time constants are significantly longer. This could be a result of
-some preprocessing to the stimulus timeseries performed in \cite{Friston2002} and
-later works but not in \cite{Johnston2008}, or it could be that \cite{Johnston2008}
-depended less on the priors.
+\subsection{Linear Approximation}
+In \cite{Friston2002b}, a novel combination of linear and nonlinear modeling
+was used to generate parameter estimates. Because it is impossible to 
+calculate the partial derivative of
+the output with respect to parameters, $\frac{\partial \theta}{\theta}$, 
+\cite{Friston2002b} approximates the partial. At each step of the Expectation-
+Maximization algorithm, the differential equation is integrated, calculating
+the residuals. Then, for each $\theta_i$ surrounding the current estimate of 
+$\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
+of the Volterra Kernel is relatively quick, and, since it is linear there
+is an analytical solution. Thus, once the Volterra Kernel exists, the value
+of $y$ at that measurement point, for that $\theta_i$ can easily be found. 
+Thus, one part of the partial is calculated at one point using
+$\frac{y(t) - y_i(t)}{\theta - \theta_i}$. This is repeated for
+every dimension of $\theta$ to give the full $\frac{\partial y}{\theta}(t)$.
+This is only at time $t$ though, and it must be repeated for every
+measurement. Finally the full derivate matrix is filled out, and
+the next step in the E-M algorithm can proceed. The full E-M algorithm
+for estimating the states is notation-heavy and can be found in \cite{Friston2002b}.
 
-Its possible, although unlikely that the balloon model is not possible to learn
-without detrimental cost in variance error. I do not think this is the case, however,
-and I will work to dispell this possibility in the results with simulated time
-series.
+Although this is certainly an interesting method of performing non-linear
+regression, there are a few caveats. First, the partials are numerical
+approximations, based on an approximate values (using Volterra-Kernels) of $y$.
+Importantly, the Volterra-Expansion of $y$ is not able to model interactions
+between state variables; interactions that were found to cause interesting
+behavior in this work. In \autoref{sec:Unscented Kalman Filter}, for
+the purpose of demonstrating nonlinearities in the signal, it was 
+necessary to propagate state variables through 1 second of simulation.
+For the distribution of state variables mentioned in \autoref{tab:steptable}
+it was necessary to use step sizes smaller than $.001$ to prevent 
+unpredictable behavior. This is using the average parameters, thus the
+only complication came from the interplay between states, which can 
+certainly lead to interesting behavior. Additionally, all the tests performed
+in \cite{Friston2002b} were on regions found to be active by the General
+Linear Model. For this reason, the reliability of the approximation is unknown
+for regions that are active but
+sufficiently nonlinear to avoid detection by conventional tests.
 
-\section{Nonlinear Least Squares}
+\subsection{Nonlinear Least Squares}
 \label{sec:Nonlinear Least Squares}
@@
 
-It would not be unreasonable at this point to back off and work with a linearized
-or more static version of the model. This is the approach taken by the
-standard General Linear Model discussed in \autoref{sec:Current Techniques General Linear Model}.
-
-\section{Unscented Kalman Filter}
-Although the classical methods mentioned in \autoref{sec:Nonlinear Least Squares}
-won't work for the nonlinear model presented in \autoref{sec:BOLD Physiology},
-there is another Bayesian technique that is worth considering.
+\subsection{Unscented Kalman Filter}
+\label{sec:Unscented Kalman Filter}
 The Unscented Kalman Filter (UKF) is a powerful Gaussian/Bayes filter that attempts
@@
 $\epsilon$ & .54  \\
-$V_t$ & N(1, .3)  \\
-$Q_t$ & N(1, .3)  \\
-$S_t$ & N(1, .3) \\
-$F_t$ & N(1, .3) \\
+$V_t$ & N(1, .09)  \\
+$Q_t$ & N(1, .09)  \\
+$S_t$ & N(1, .09) \\
+$F_t$ & N(1, .09) \\
 \hline
@@
 
-\section{Variational Filtering}
-\cite{Friston2008}, \cite{Friston2008b}
+\section{Hybrid Methods}
+In \cite{Riera2004}, a maximum
+likelihood method for innovation processes was used, as described by
+\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
+Kalman filter, to break the time series into a series of innovations,
+for which Maximum Likelihood was performed. While this is probably the most
+likely solution to give the correct output, there are a few problems. First, every
+step in parameter space requires a recalculation of all the state variables. With
+two or three parameters this is fine, more than that, and calculations can become
+intractable. Additionally, there is no way to ensure a global minimum is reached.
+Although this is also true of the methods in \autoref{sec:Nonlinear Least Squares},
+those methods contained random elements designed to overcome this issue.
+[results?]
 
-\section{Hybrid Methods}
-A large number of hybrid methods have been tried, in order to elicit 
-parameter estimates from the BOLD model. \cite{Vakorin2007},
-\cite{Johnston2007}, \cite{Hu2009}
+In \cite{Johnston2008}, a hybrid particle filter/gradient
+descent algorithm was used to simultaneously derive the static and dynamic 
+parameters, (classically known as parameters and state variables, respectively).
+Essentially a particle filter is used to calculate the state variables at each
+time; then the estimated distribution of the particles is used to find
+the most likely set of parameters that would give that distribution of state variables.
+This process is repeated until the parameters converge. Interestingly
+\cite{Johnston2008} comes to a very different set of parameter estimates as compared
+to the original \cite{Friston2000} estimates (\autoref{tab:Params}). In fact the results 
+are significantly different from virtually every other study. The most obvious discrepancy
+is the significantly longer time constants, $\tau_f$, $\tau_s$ and $\tau_0$. 
+While of course this could be poor convergence of the algorithm, there is another other possibility.
+Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Least Squares},
+the algorithm described in \cite{Johnston2008} does not depend on prior distributions.
+It is possible then that the bias toward the prior in other methods screwed the results. 
+While \cite{Johnston2008} is certainly in the minority; further exhaustive studies 
+of the parameters, using unbiased techniques may be called for.
+
+[graph of friston parameters vs. johnston]
 

### 2010-06-30 doc/thesis/conclusion.tex
variation filtering dropped (DCM), background down.

\chapter{Conclusion}

\section{Future Work}
*A more extensive investigation of the parameters distribution
*Using a compliance model
*restrict parameters more (may not be possible)

### 2010-06-30 doc/thesis/background.tex
just a note added
conclusion in background

@@
 \chapter{Prior Works}
+\label{sec:Prior Works}
 The ultimate purpose of this work is to provide a new tool
@@
 
+[might want to mention dynamic causal modeling]
+
 \section{Conclusion}
-In summary, several other approaches could be taken
+Clearly there is no ideal solution to solving this system of nonlinear
+equations. Exhaustive search type methods such as those employed by
+\cite{Johnston2008} and \cite{Vakorin2007} can have extremely long
+run times even for a single voxel. While Volterra models are an interesting
+solution,there have not yet been exhaustive tests to determine whether
+such approximations work well throughout state space. The most promising
+method of those reviewed here is the Kalman filter based method. It is
+able to maintain a very fast runtime while still approaching the solution.
+While the reliance on a Gaussian estimate to the true posterior distribution could
+cause problems, some modifications could make it extremely powerful. 
+The particle filter method proposed in the next section bears a strong 
+resemblance to the unscented Kalman filter; albeit with more point estimating
+the posterior.

### 2010-06-30 doc/thesis/intro.tex
conclusion in background

@@
 \section{Noise}
+\label{sec:Introduction Noise}
 As demonstrated in \autoref{sec:BOLD Phsyiology} the BOLD response has been

### 2010-06-30 doc/thesis/particlefilters.tex
conclusion in background

@@
 \section{Introduction}
-Particle filters, a type of Sequential Monte Carlo (SMC) methods
+Particle filters, a type of Sequential Monte Carlo (SMC) methods,
 are a powerful way of estimating the posterior probability distribution

### 2010-06-30 doc/thesis/particlefilters.tex
started cleaning up notation in particle filter

--- 2010-06-28 doc/thesis/particlefilters.tex
@@
 \section{Introduction}
-Particle filters, a type of Sequential Monte Carlo (SMC) methods
+Particle filters, a type of Sequential Monte Carlo (SMC) methods,
 are a powerful way of estimating the posterior probability distribution
-of a set of parameters give a timeseries of measurements. Unlike Markov 
-Chain Monte Carlo estimation, Sequential Monte-Carlo methods are designed
-to be used with parameters that vary with time. Unlike variations of the
-Kalman filter, particle filters do not make the assumption that noise
-is Gaussian. Thus particle filters are often the best solution to bayesian 
-tracking for non-linear, non-gaussian systems.
+a set of parameters given a timeseries of measurements. Unlike Markov 
+Chain Monte Carlo (MCMC) estimation, particle filters are designed for 
+time-varying random variables. The idea behind particle filters is
+extremely similar to Kalman Filters; however, unlike Kalman Filters,
+distributions are stored as an empircal distribution rather than the
+as the first two moments of a Gaussian. Thus particle filters are 
+preferred when the model is nonlinear, and thus non-gaussian.
 
@@
 \label{sec:Particle Filter Model}
-The idea of the particle
-filter is to start with a wide mixture PDF of possible parameter sets, 
-and then, as measurements come in, to weight more heavily parameter sets 
-that tend to give good estimations of the measurements. The reliance on
-an initial mixture PDF can introduce bias; however, this effect can be
-minimized by alterring the initial weights in the mixture pdf. Of course
-every gradient descent must choose starting points and it is often quite
-easy to establish a reasonable range of parameters, especially when the
-model being used has a physical meaning. Suppose a set or stream of measurements
-are given, $\{y(t), t = 1, 2, 3, ... T\}$, where $T$ is permitted to go
- to infinity. Then the goal is to find the 
-parameters, $\hat{\theta}$, and underlying state time series, $\hat{x}[0:T]$
-that minimize the difference between $\hat{y}[0:T]$
-and $y[0:T]$. In our case, we will assume that we know the form
-of the model, which is based on first principals, and that
-there is some true $\theta$ and a true time-series of underlying
-state variable, $x[0:T]$ that drives $y[0:T]$. Assuming a model form 
-such as we do here reduces model variance, potentially at the cost of increased
-bias (or systematic) error. We will assume a basic state space model:
+The idea of the particle filter is to build an empirical distribution
+out of a large number of parameter \emph{sets}, called particles. Each
+particle contains all the parameters and states needed to propagate
+the model forward.  The particle filter begins with a wide distribution 
+(called the Prior Distribution)
+of possible particles and then, as measurements come in, weights 
+particles based on the quality of their output estimates. Thus parameter sets 
+that tend to give good estimations of the measurements get weighted higher
+than parameter sets that give poor estimates. Although the reliance on
+a prior distribution can be troublesome, when the system being modeled
+has physical meaning, establishing reasonable ranges for parameters can 
+actually be quite easy. Optimizing the prior can be more difficult though,
+unless the system has been extensively studied.
 
-\begin{equation}
-\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x)
-\end{equation}
+Suppose a discrete set or stream of measurements are given, 
+$\{y_k, k = 1, 2, 3, ... K\}$, where $K$ is infinite for a stream. 
+Suppose also that there is a hidden set of state variables,
+denoted $\{x_k, k = 1, 2, 3, ... K\}$ the dictates the movement
+of $y_k$. The goal of the particle filter is to estimate the 
+\emph{distribution} of the
+true parameters $\Theta$ that dictates the movement of $x_k$. Additionally
+the particle filter also estimates the distribution of $x_k$.
+The only difference between the members of parameter vector
+$\Theta$ and the members of $x_k$ is that the memebers of
+$\Theta$ have no known update equation. Members of both vectors
+are permitted to have some time varying noise, although this
+is often not explicitly modeled. The generic, continuous, nonlinear
+system definition is shown in \autoref{eq:GenericNonlinear}.
 
-\begin{equation}
+\begin{eqnarray}
+\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x) \nonumber \\
 y(t) = g(t, x(t), u(t), \theta, \nu_y)
-\end{equation}
+\label{eq:GenericNonlinear}
+\end{eqnarray}
 
-Where $x(t)$ is a vector of state variables, $\theta$ is a vector of system
-constants, $u(t)$ is a stimulus, $y(t)$ an observation, and
-$\nu_x$ and $\nu_y$ are random variates. Obviously any one of these could
-be a vector, so for instance $u(t)$ could encode multiple types of stimuli.
+$x(t)$ is vector of state variables, $\theta$ is a vector of system
+constants, $u(t)$ is an input, $y(t)$ the observation, and
+$\nu_x$ and $\nu_y$ are random variates. Although any of these
+variables could be a vector, for the sake of simplicity only
+$\theta$ and $x(t)$ will be considered as such.
 
-Although not strictly necessary for particle filters, we will make a few
-assumptions based on the particular type of systems faced in biological 
-processes. First, the systems are assumed to be time invariant. This 
-assumption is based on the idea that if you froze the system for $\Delta t$
+Although not strictly for particle filters in general, a few 
+assumptions are made for this work. First, the systems are assumed to be 
+time invariant. This 
+assumption is based on the idea that if you paused the system for $\Delta t$
 seconds, when unfrozen the system would continue as if nothing happend. 
 Few biological systems are predictable enough for them to be summarized
-by a time varying function. Even systems that one might assume work in that
-way are actually much more complicated. The heart may seem like an obvious
-exception, however, the period between heartbeats vary often enough that prediction
-would necessitate another state-space model. In short, we
-assume no parameters are time varying, because not enough information exists to
-describe any of theme in that way. Luckily particle filters are capable 
-of dealing with non-white, non-Gaussian noise, so unanticipated influence
-may be re-factored as noise. Secondly we assume that input cannot directly
+by a time varying function, least of all the brain. While heart beats are certainly
+periodic and have an effect on the BOLD signal, the period varies too much
+for the system to be considered varying with time. 
+Next, its assumes that input cannot directly
 influence the output, which in the case of the BOLD signal is a good assumption.
-%Third, through some sort of preprocessing, we will assume that $\nu_d$ can be
-%decreased or completely removed, since it 
-Third, we will assume noise is additive, and that $\nu_x$ may be projected into
-a Weiner, or other summing process that is additive with $g$ and $\nu_y$, which
-will be named $\nu_d$.
-Finally, $x(t)$ will encapsulate $\theta$, the unknown model constants, which
-means that the vector $\dot{x}$ will always have members
-that are 0. The results of these assumptions are a simplified version of the
+Also, time is treated as discrete and noise is considered to be additive.
+Finally, because the only difference between the members of $x(t)$ and 
+$\theta$ is an update function, from now on $x$ will contain 
+$\theta$. The assumptions now allow for a simplified version of the
 state space equations:
 
-\begin{equation}
-\dot{x}(t) = f(x(t), u(t))
-\label{eq:stateass}
-\end{equation}
-
-\begin{equation}
-y(t) = g(x(t)) + \nu_y + \nu_d
+\begin{eqnarray}
+\dot{x}_k = f(x_{k-1}, u_k) + \nu_x 
+\label{eq:stateass}\\
+y_k = g(x_k) + \nu_y 
 \label{eq:measass}
-\end{equation}
-
-Because $\nu_d$ is something akin to and additive Weiner process $y[0:T]$, it 
-will include low frequency noise. $\nu_y$ on the other hand will cause i.i.d. noise
-in $y[0:T]$. For some of the tests, I will use de-trending methods to reduce the effects of 
-$\nu_d$, the remainder of which will be re-factored into $\nu_y$. Both $\nu_d$ and $\nu_y$
-have biological and non-biological sources. MR can lead to both types of noise, 
-as demonstrated in \cite{Smith1999}. Meanwhile changes in metabolism, heart rate, or
-other biochemical intervention could all lead to either $\nu_d$ or $\nu_y$.
+\end{eqnarray}
 
 \subsection{Prior}
-
 The goal of the particle filter is to evolve a probably distribution 

### 2010-06-30 doc/thesis/methods.tex
conclusion in background
work on particle filter.

@@
 \chapter{Methods}
-This paper describes two types of experiments; first we will cover
-simulations, which of course have the benefit of a ground truth, second
-we will cover the methods for real data.
+\section{Prior Distribution}
+For the BOLD signal all the parameters have been studied and have relatively well
+known mean and variance, so a prior is helpful. 
+For strictly positive parameters (members of $x$) we used a gamma distribution,
+whereas for parameters that could be negative, we used a Gaussian distribution. In
+both cases standard deviations twice that found in previous studies were used.
 

### 2010-06-30 doc/thesis/particlefilters.tex
work on particle filter.

@@
 
-\subsection{Model}
+\section{Model}
 \label{sec:Particle Filter Model}
@@
 
-Suppose a discrete set or stream of measurements are given, 
-$\{y_k, k = 1, 2, 3, ... K\}$, where $K$ is infinite for a stream. 
+Suppose set or stream of measurements at discrete times are given, 
+$\{Y_k, k = 1, 2, 3, ... K\}$, where $K$ is infinite for a stream. 
+Because $k$ is a discrete time, let $t_k$ define the continuous
+time of $k$
 Suppose also that there is a hidden set of state variables,
-denoted $\{x_k, k = 1, 2, 3, ... K\}$ the dictates the movement
-of $y_k$. The goal of the particle filter is to estimate the 
+$X(t)$ that dictates the movement
+of $Y(t)$, although for most of the time dealings will be
+with $X_k = X(t_k)$. The goal of the particle filter is to estimate the 
 \emph{distribution} of the
-true parameters $\Theta$ that dictates the movement of $x_k$. Additionally
-the particle filter also estimates the distribution of $x_k$.
+true parameters $\Theta$ that dictates the movement of $X(t)$.
+The model also permits random motion in $X(t)$, so the 
+particle filter also estimates the distribution of $X(t)$.
 The only difference between the members of parameter vector
-$\Theta$ and the members of $x_k$ is that the memebers of
+$\Theta$ and those of $X(t)$ is that the memebers of
 $\Theta$ have no known update equation. Members of both vectors
-are permitted to have some time varying noise, although this
-is often not explicitly modeled. The generic, continuous, nonlinear
+are permitted to have some noise, although this
+may not be explicitly stated in the model. The generic, continuous, nonlinear
 system definition is shown in \autoref{eq:GenericNonlinear}.
@@
 \begin{eqnarray}
-\dot{x}(t) = f(t, x(t), u(t), \theta, \nu_x) \nonumber \\
-y(t) = g(t, x(t), u(t), \theta, \nu_y)
+\dot{X}(t) = f(t, X(t), u(t), \theta, \nu_x) \nonumber \\
+Y(t) = g(t, X(t), u(t), \theta, \nu_y)
 \label{eq:GenericNonlinear}
@@
 
-$x(t)$ is vector of state variables, $\theta$ is a vector of system
-constants, $u(t)$ is an input, $y(t)$ the observation, and
+$X(t)$ is vector of state variables, $\Theta$ is a vector of system
+constants, $u(t)$ is an input, $Y(t)$ the observation, and
 $\nu_x$ and $\nu_y$ are random variates. Although any of these
 variables could be a vector, for the sake of simplicity only
-$\theta$ and $x(t)$ will be considered as such.
+$\Theta$ and $X(t)$ will be considered as such.
 
-Although not strictly for particle filters in general, a few 
+Although not necessary for particle filters in general, a few  simplifying
 assumptions are made for this work. First, the systems are assumed to be 
@@
 influence the output, which in the case of the BOLD signal is a good assumption.
-Also, time is treated as discrete and noise is considered to be additive.
-Finally, because the only difference between the members of $x(t)$ and 
-$\theta$ is an update function, from now on $x$ will contain 
-$\theta$. The assumptions now allow for a simplified version of the
+Also, noise is considered to be additive.
+Finally, because the only difference between the members of $X(t)$ and 
+$\Theta$ is an update function, from now on $x$ will contain 
+$\Theta$. The assumptions now allow for a simplified version of the
 state space equations:
@@
 \begin{eqnarray}
-\dot{x}_k = f(x_{k-1}, u_k) + \nu_x 
+\dot{X}_k = f(X_{k-1}, u_k) + \nu_x 
 \label{eq:stateass}\\
-y_k = g(x_k) + \nu_y 
+Y_k = g(X_k) + \nu_y 
 \label{eq:measass}
@@
 
-\subsection{Prior}
-The goal of the particle filter is to evolve a probably distribution 
-$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
-that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
-Considering that $y$ contains measurement noise as well as noise from $x$,
- it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
-but a true posterior. 
-To begin with, the particle filter starts with a prior distribution, and $N_p$
-particles need to be drawn from that distribution, $\alpha(x)$:
+\section{Derivation}
+The goal of the particle filter is to evolve an empirical distribution 
+$P(x_k | u_{0:k}, Y_{0:k})$,
+that asymptotically approaches the true probability distribution $P(X_k | u_{0:k})$.
+Note that capital $X$ will be used as the actual realizations of 
+the state variable, whereas $x$ will denote estimates of $X$.
+Additionally, the notation $a:b$ indicates the set $[a,b]$,
+as in $u_{a:b}$, which would indicate all the inputs from time $a$ to time $b$.
+Considering the noise present in $X$,
+ $P(X_k | u_{0:k})$ is not a single true value but probability distribution.
+
+To begin with, the particle filter must be given a prior distribution, from
+which the initial $N_p$ particles are drawn. A particle contains a weight
+as well as an estimate of $X_k$, which as already stated, contains every
+variable needed to run the model. Then the prior is generated from a 
+given distribution, $\alpha(X)$, by:
+
 \begin{equation}
-\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
+\{[x^i_0,w^i] : x^i_0 \sim \alpha(X), w^i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
 \end{equation}
@@
 using a Mixture PDF. 
+Note that any exponents will be explicitly labeled as such, to avoid confusion with
+the particle numbering scheme.
+
+Therefore, after the particle have been generated they should approximate $\alpha(X)$:
+
 \begin{equation}
-\hat{Pr}(x(0) = \hat{x}) = \sum_{i=1}^{N_p} w_i\delta(\hat{x} - x_i(0) ) dx
+\alpha(X) \approx P(x_0) = \sum_{i=0}^{N_p} w^i\delta(X - x^i_0 ) dx
 \end{equation}
@@
 
-If a true prior is preferred, then the weights
-should all be $1/N_p$, and since $x_i$ was drawn from the prior, this will
-be an approximation of the prior distribution. If a relatively flat prior is 
-preferred, then each particle's weight could be divided by the density, $\alpha(x_i)$,
-which creates a flat prior with support points in the region of $\alpha(x)$. Either
-way, $\alpha(x)$ should be much broader than the true posterior, $Pr(x(0))$, since the
-choice of support points is crucial to the convergence of any sampling importance
-algorithm. 
-For the BOLD signal all the parameters have been studied and have relatively well
-known mean and variance, so a prior could be very helpful. We ran simulations for
-both normalized and un-normalized priors, although we believe in cases such as this,
-where a good prior exists, it should be used. 
-For strictly positive parameters (members of $x$) we used a gamma distribution,
-whereas for parameters that could be negative, we used a Gaussian distribution. In
-both cases standard deviations twice that found in previous studies were used.
+If a flat prior is 
+preferred, then each particle's weight could be scaled to the reciprocal of the
+density at the particle: 
+\begin{equation}
+w^i = \frac{1}{\alpha(x^i_0)}
+\end{equation}
+Whether or not to flatten the prior is a design decision. The reason this might 
+be preferred over a direct
+uniform distribution is that the distribution width will inherently scale 
+for increased particle counts although some distributions
+flatten out better than others. Either way, $\alpha(X)$ \emph{must} be
+wide enough to incorporate any posterior that arises. If the prior is
+not sufficiently dense, the particle filter can compensate, if it is
+not sufficiently wide the particle filter won't converge.
 
-Note that all the probabilities implicitly depend on $u[0:T]$, so those terms 
-will be left off for simplicity.
-Once the probability, $\hat{Pr}(x(T) | x[0:T-1], y[0:T-1])$ has been found
-(initially this is just Mixture approximating the prior since no measurements are 
-available and no previous probabilities are available), its possible to approximate
-the probability for short times between times when measurement is available, by shifting
-the probability according the progression of the state equations. This is only 
-an approximate, since integrating $\nu_d$ should increase uncertainty as
-time without a measurement passes.
+%todo, put back?
+%Once the probability, $P(x_k | x_{0:k-1}, y_{0:k-1})$ has been found
+%(initially this is just Mixture approximating the prior since no measurements are 
+%available and no previous probabilities are available), its possible to approximate
+%the probability between times when measurement is available, by shifting
+%the probability according the progression of the state equations. This is only 
+%an approximate, since integrating $\nu_d$ should increase uncertainty as
+%time without a measurement passes. 
+%
+%\begin{equation}
+%P(x(T+\Delta t)) \approx 
+%\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
+%\end{equation}
 
+\subsection{Weighting}
+For all the following areas, the probabilities implicitly depend on $u_{0:k}$, 
+so those terms are left off for simplicity.
+
+Whenever a measurement becomes available it permits refinement of the
+posterior density.
+This process of incorporating new data is called sequential importance sampling,
+and eventually allows convergence. The weight is defined as
 \begin{equation}
-\hat{Pr}(x(T+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
+w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}
+\label{eq:weightfunc}
+\end{equation}
+where $q$ is called an \emph{importance density}. The importance density
+is the density of the points, thus by dividing by this value, the weight
+should not depend on the location of the estimation points, but rather
+only on $P(x^i_{0:k} | y_{0:k})$, the probability of that particle
+being correct given all the measurements up to time $k$. 
+Of course if there is a far off peak in
+the posterior that $q$ does not have support points in, there will 
+be quantization errors, and that part of the density can't be modeled. This is why
+it is absolutely necessary that $q$ fully covers $P(x^i_{0:k} | y_{0:k})$.
+
+It is helpful
+to consider how the importance density affects the initial distribution. 
+In the initial distribution, the weights are all the same; and for
+the sake of argument, let them all be scaled up to 1. Then
+\begin{equation}
+w^i_k q(x^i_{0:k} | y_{0:k}) = q(x^i_{0:k} | y_{0:k}) = P(x^i_{0:k} | y_{0:k})
+\end{equation}
+the estimated probability, $P(x^i_{0:k} | y_{0:k})$ depends only on the 
+way the particles are distributed. As new measurements are incorporated,
+the weight will accumulate probabilities through time, which will be discussed
+next.
+
+\section{Calculating Weights}
+To calculate the weight of a particular particle, it is necessary to 
+calculate both $q(x^i_{0:k} | y_{0:k})$ and $P(x^i_{0:k} | y_{0:k})$.
+Note that $q(x^i_{0:k} | y_{0:k})$ may be simplified by assuming that 
+$y_k$ doesn't contain any information about $x_{k-1}$. Technically this 
+could be false; since later measurements may shed light on currently hidden
+changes in $x$. For practical applications though it is helpful assumption.
+\begin{equation}
+q(x^i_{0:k} | y_{0:k}) = q(x^i_{0:k} | y_{0:k-1})
+\label{eq:QAssump}
+\end{equation}
+The choice of the importance density is another design decision; however
+it is common to use the integrated state equations. 
+Although other importance density functions exist; for the particle filter
+used here, the standard importance density will be used: the modeled
+prior.
+\begin{equation}
+q(x_k | x_{k-1}, y_{0:k}) =  P(x_k | x_{k-1})
+\label{eq:ImportanceDensity}
+\end{equation}
+The benefit of this choice for importance density is that an approximation for
+$P(x_k | x_{k-1})$ is freely available: its simply the set of particles propagated
+forward in time using the state equations. Additionally it makes
+updating weights extremely simple, as seen in \autoref{eq:weightevolve}.
+
+The $q(x^i_{0:k} | y_{0:k})$ may then be simplified:
+\begin{equation}
+\begin{array}{cclr}
+q(x_{0:k} | y_{0:k}) & = & q(x_k | x_{0:k-1}, y_{0:k})q(x_{0:k-1} | y_{0:k}) &  \\
+& = & q(x_k | x_{0:k-1}, y_{0:k})q(x_{0:k-1} | y_{0:k-1})  & \text{[\autoref{eq:QAssump}]} \\
+& = & q(x_k | x_{k-1}, y_{0:k})q(x_{0:k-1} | y_{0:k-1})  & \text{[Markov Property]}\\
+& = & P(x_k | x_{k-1})q(x_{0:k-1} | y_{0:k-1})  & \text{[\autoref{eq:ImportanceDensity}]}
+\end{array}
 \end{equation}
 
-\subsection{Weighting}
-When a measurement becomes available it is incorporated into the probability.
-This process of incorporating new data is called sequential importance sampling,
-and eventually causes the probability to converge. The weight is defined
-as
+Calculating $P(x_{0:k} | y_{0:k})$ is a bit more involved. 
+First, using the assumption that the distribution of $y_k$ is 
+fully constrained by $x_k$, and that $x_k$ is similarly fully 
+constrained by $x_{k-1}$, we are able to make the very good assumptions that:
+\begin{eqnarray}
+P(y_k | x_{0:k}, y_{0:k-1}) &=& P(y_k | x_k) \nonumber \\
+P(x_k | x_{0:k}, y_{0:k-1}) &=& P(x_k | x_{k-1})
+\label{eq:MarkovProperty}
+\end{eqnarray}
+These are of course just re-statements of the state equations assumed by \autoref{eq:stateass}
+and \autoref{eq:measass}.
 
+Additionally, for the particle filter $y_k$ and $y_{0:k-1}$ are 
+constant across all particles, thus $P(y_k| y_{0:k-1})$ can
+be dropped when the equality is changed to a proportion. 
+Using these properties, $P(x^i_{0:k} | y_{0:k})$ may be broken up as follows 
+(mostly using Bayes' Theorem):
 \begin{equation}
-w_i(T) \propto \frac{\hat{Pr}(x_i[0:T] | y[0:T])}{q(x_i[0:T] | y[0:T])}
-\label{eq:weightfunc}
+\begin{array}{lclr}
+P(x_{0:k} | y_{0:k}) & = & \frac{P(y_{0:k}, x_{0:k})}{P(y_{0:k})} & \\
+ & = & \frac{P(y_k, x_{0:k} | y_{0:k-1}) \cancel{P(y_{0:k-1})}}{P(y_k | y_{0:k-1}) \cancel{P(y_{0:k-1})}} & \\
+ & = & \frac{P(y_k| x_{0:k}, y_{0:k-1}) P(x_{0:k} | y_{0:k-1})}{P(y_k | y_{0:k-1}) } & \\
+ & = & \frac{P(y_k| x_{0:k}, y_{0:k-1}) P(x_k | x_{0:k-1}, y_{0:k-1}) P(x_{0:k-1} | y_{0:k-1})}{P(y_k | y_{0:k-1})} &  \\
+& = & \frac{P(y_k| x_k) P(x_k | x_{k-1}) P(x_{0:k-1} | y_{0:k-1})}{P(y_k | y_{0:k-1})}  & [\text{\autoref{eq:MarkovProperty}}]\\
+& \propto & P(y_k| x_k) P(x_k | x_{k-1}) P(x_{0:k-1} | y_{0:k-1}) & [P(y_k|y_{0:k-1}) \text{ is constant}]
+ \end{array}
+ \label{eq:UpdateBayes}
 \end{equation}
-
-where $q$ is called an \emph{importance density}, meaning it decides where
-the support points for $x(T)$ are located. To remove the bias due to the
-location of the support points, we divide by $q(x_i[0:T] | y[0:T])$. By dividing by 
-the posterior density of the support points (particles), the effect of the particle 
-distribution may be removed from the posterior density. As a result the weight
-is dependent solely based
-on $\hat{Pr}(x_i[0:T] | y[0:T])$, the probability of the $i^{th}$ particle's measurements
-being different from $y[0:T]$ due to noise alone.
-An example of an importance density would be drawing a large
-number of points from the standard normal, $N(0,1)$ and then weighting each point, $l$ by 
-$1/\beta(l), \beta \sim N(0,1)$. Of course if there is a far off peak in
-the posterior that $q$ does not allocate support points in, there will 
-be a quantization error, and that part of the density can't be modeled. This is why
-it is absolutely necessary that $q$ covers $\hat{Pr}(x_i[0:T] | y[0:T])$.
-
-$q(x_i[0:T] | y[0:T])$ may be simplified by assuming that $y(T)$ doesn't contain 
-any information about $x(T-1)$, which is more practical since knowledge of future
-measurements is impractical.
-
+Plugging \autoref{eq:ImportanceDensity} and the result of \autoref{eq:UpdateBayes}
+into \autoref{eq:weightfunc} leads to:
 \begin{eqnarray}
-q(x[0:T] | y[0:T]) & = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T]) \nonumber \\
-& = & q(x(T) | x[0:T-1], y[0:T])q(x[0:T-1] | y[0:T-1]) \nonumber \\
-& = & q(x(T) | x(T-1), y[0:T])q(x[0:T-1] | y[0:T-1])
-\end{eqnarray}
-
-In this paper we will use 
-$q(x_i(T) | x_i(T-1), y[0:T]) =  \hat{Pr}(x_i(T) | x_i(T-1))$,
-based on the Markov assumption, and the belief that the state space model is 
-able to approximate the true state. This means that prior to re-weighting 
-particles, the particles will be distributed the same as the previous time but
-moved forward according to the integration of $f(x(t), u(t))$.
-
-In addition to $q(x_i(T) | x_i[0:T-1], y[0:T])$, the weight is also based on $Pr(x_i[0:K] | y[0:K])$,
-which may be broken up as follows.
-\begin{eqnarray}
-\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y[0:T], x[0:T])}{\hat{Pr}(y[0:T])} \nonumber \\
- & = & \frac{\hat{Pr}(y(T), x[0:T] | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}}{\hat{Pr}(y(T) | y[0:T-1]) \cancel{\hat{Pr}(y[0:T-1])}} \nonumber \\
- & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x[0:T] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1]) } \nonumber \\
- & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-\end{eqnarray}
-
-Using the assumption that $y(t)$ is fully constrained by $x(t)$ \autoref{eq:measass},
-and that $x(t)$ is fully constrained by $x(t-1)$ \autoref{eq:stateass}, we are able to
-make the reasonably good assumptions that:
-\begin{equation}
-\hat{Pr}(y(T) | x[0:T], y[0:T-1]) = \hat{Pr}(y(T) | x(T))
-\end{equation}
-
-\begin{equation}
-\hat{Pr}(x(T) | x[0:T], y[0:T-1]) = \hat{Pr}(x(T) | x(T-1))
-\end{equation}
-
-Additionally, for the particle filter $y(T)$ and $y[0:T-1]$ are 
-given, and therefore constant across all particles. Thus $\hat{Pr}(x[0:T] | y[0:T])$
-may be simplified to:
-\begin{eqnarray}
-\hat{Pr}(x[0:T] | y[0:T]) & = & \frac{\hat{Pr}(y(T)| x[0:T], y[0:T-1]) \hat{Pr}(x(T) | x[0:T-1], y[0:T-1]) 
-            \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-& = & \frac{\hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])}{\hat{Pr}(y(T) | y[0:T-1])} \nonumber \\
-& \propto & \hat{Pr}(y(T)| x(T)) \hat{Pr}(x(T) | x(T-1)) \hat{Pr}(x[0:T-1] | y[0:T-1])
-\end{eqnarray}
-
-Plugging these simplifications into \autoref{eq:weightfunc} leads to:
-\begin{eqnarray}
-w_i(T) & \propto & \frac{\hat{Pr}(y(T)| x(T)) \cancel{\hat{Pr}(x(T) | x(T-1))} \hat{Pr}(x[0:T-1] | y[0:T-1])}
-                         {\cancel{\hat{Pr}(x_i(T) | x_i(T-1))}q(x[0:T-1] | y[0:T-1])} \nonumber \\
-& \propto & w_i(T-1)\hat{Pr}(y(T)| x(T)) 
+w^i_k & \propto & \frac{P(y_k| x^i_k) \cancel{P(x^i_k | x^i_{k-1})} P(x^i_{0:k-1} | y_{0:k-1})}
+                         {\cancel{P(x^i_k | x^i_{k-1})}q(x^i_{0:k-1} | y_{0:k-1})} \nonumber \\
+& \propto & w^i_{k-1}P(y_k| x_k) 
 \label{eq:weightevolve}
@@
 
-Thus, by making the following relatively weak assumptions, evolving a posterior
-density  is easy and requires almost no knowledge of noise distribution.
+Thus, by making the following relatively easy assumptions, evolving a posterior
+density  requires no knowledge of noise distribution.
 \begin{enumerate}
-\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ provide 
-a sufficiently flexible model to encapsulate the true time series.
-\item $E[\nu_d] = 0$ and $E[\nu_y] = 0$, and $\nu_x = d\nu_d$, $\nu_y$ are stationary
-\item The PDF $q(x_i(0))$ (the prior) fully covers $Pr(x_i(0))$
-\item Markov Assumption: $Pr(x(T) | x[0:T]) = Pr(x(T) | x(T-1))$
-\item $q(x[0:T-1] | y[0:T]) = q(x[0:T-1] | y[0:T-1])$
+\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ 
+\item The PDF $q(x_i(0))$ (the prior) fully covers $P(x_i(0))$
+\item Markov Property: $P(x_k | x_{0:k-1}) = Pr(x_k | x_{k-1})$
+\item $q(x_{0:k-1} | y_{0:k}) = q(x_{0:k-1} | y_{0:k-1})$
 \end{enumerate}
@@
 \subsection{Basic Particle Filter Algorithm}
-From the definition of $w_i$, the algorithm sequential importance sampling (SIS) 
+From the definition of $w_i$, the algorithm to calculate
+an approximation of $P(X(t_k) | Y_{0:k})$ or $P(X(t_k + \delta t) | Y_{0:k})$
 is relatively simple.
 
+\begin{algorithm}
+\caption{Sequential Importance Sampling}
 \begin{algorithmic}
-\STATE Initialize $N_p$ Particles: 
-        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
-\STATE $T$ = \{Set of Measurement Times\}
-\FOR{$t$ in $T$}
-    \FOR{$i$ in $N_p$}
-        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
+\STATE Initialize Particles:
+\FOR{$i$ : each of $N_p$ particles }
+    \STATE $x^i_0  \sim \alpha(X)$
+    \STATE $w^i_0 = \frac{1}{N_p}$
+\ENDFOR
+\FOR{$k$ : each measurement}
+    \FOR{$i$ : each particle }
+        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
     \ENDFOR
 \ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+\STATE $P(x(t_k+\Delta t)) \approx 
+\sum_{i=0}^{N_p} w^i_k \delta\left(x - (x^i_k + \int_{t_k}^{t_k+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
 \end{algorithmic}
-The result is then a discrete approximation of the posterior distribution.
+\end{algorithm}
 
@@
 distribution), there will be many particles with insignificant weights. While this does help
-describe the tails of the distribution very well, it means that only a small portion of the
-computation will be spent describing the most probable region. Ideally every particle would 
-equally decrease the entropy of the distribution, thus the lower the variance of the weights,
-the more efficiently the discrete distribution is in describing the continuous distribution. 
-A common measure of "Particle Degeneracy" is the effective number of particles, described
-in (Bergman "Navigation and Tracking Applications", 1999, J S Liu and R Chen "Sequential 
-Monte Carlo Methods for Dynamical Systems", 1998), which is based on the "true weight"
-of each particle. Of course the true weight is unknown, so a heuristic approximating 
-$N_{eff}$ is used:
+describe the tails of the distribution, it means a lot of computation will be based.
+Instead, it would be preferable if most of the computation is spent on the most probable regions.
+Ideally the computation time spent on tails would be proportional to the actual size of the
+tails. In this case particle locations would match the true posterior and all weights would
+be equal.  The case where a large number of the weights have become irrelevantly small
+is called "particle degeneracy". In  \cite{Liu1998b}
+an ideal calculation of the "effective" number of particles is found based on the 
+particles' "true weight". However, given that only an approximation for the true weight 
+exists, they also provide a simple heuristic calculation of $N_{eff}$.
 \begin{equation}
-\hat{N}_{eff} \approx \frac{N_p}{\sum_{i=1}^{N_p} w_i^2}
+N_{eff} \approx \frac{\sum_{i=0}^{N_p} w_i}{\sum_{i=0}^{N_p} w_i^2}
 \label{eq:neff}
@@
 $N_{eff}$ drops precipitously.  To alleviate this problem
-a common technique known as resampling must be applied. The idea of re-sampling is to 
+a common technique known as resampling may be applied. The idea of resampling is to 
 draw from the approximate posterior, thus generating a replica of the posterior with 
-a support more suited to the distribution. Thus, if weights are all set to $1/N_p$, and 
-$N_p$ points are drawn from the posterior,
+a better support. Therefore, a new set of particles may be drawn from the empirical
+distribution as follows:
 \begin{equation}
-\hat{\chi}_j \sim \left(\sum_{i=1}^{N_p} w_i(t)\delta(x - x_i(t))\right), j \in \{1, ..., N_p\}
+\hat{x}_j \sim \left(\sum_{i=0}^{N_p} w^i_k\delta(x - x^i_k)\right)
 \end{equation}
-then $\hat{\chi} \sim \hat{x}$ should hold. Unfortunately, this isn't necessarily the truth: since the support is
+
+\begin{algorithm}
+\caption{Stratified Resampling Algorithm}
+\begin{algorithmic}
+
+\end{algorithmic}
+\end{algorithm}
+
+Unfortunately, this isn't necessarily the truth: since the support is
 still limited to the original particles, the number of unique particles can only go down.
@@
 \subsection{Weighting Function}
-Because $\hat{Pr}(y(T) | x(T))$, what I will call the weighting function,
+Because $P(y_k | x(T))$, what I will call the weighting function,
 is based on an unknown distribution, it is necessary to decide on a function
-that will approximate $\hat{Pr}(y(t) | x(T))$. Obviously the function, $\omega(y(t), f(x(t))$
+that will approximate $P(y_k | x(T))$. Obviously the function, $\omega(y_k, f(x(t))$
 needs to be centered at zero and have a scale comparable to the signal levels.
-Obviously if the actual noise present in $y(t)$ were to be known, then that would
+Obviously if the actual noise present in $y_k$ were to be known, then that would
 be the best distribution for $\Omega$. In that case, particles that fell
@@
         \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w_i(t) = w_i(t-1)\hat{Pr}(y(t) | x(t))$
+        \STATE $w_i(t) = w_i(t-1)P(y(t) | x(t))$
     \ENDFOR
@@
 
-\STATE At $t + \Delta t$, $t \in T$, $\hat{Pr}(x(t+\Delta t)) \approx 
+\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
 \sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$

### 2010-06-30 doc/thesis/particlefilters.tex
close to finished with the particlefilters

@@
 \begin{algorithm}
-\caption{Stratified Resampling Algorithm}
+\caption{Resampling Algorithm}
 \begin{algorithmic}
+\STATE Calculate total weight, $W = \sum_{i=0}^{N_p} w^i$
+\FORALL{$0 < i < N_p$}
+    \STATE Draw $V$ from uniform range $[0, W]$
+    \STATE $C = W_t$
+    \FORALL{$0 < j < N_p$ and $C < V$}
+        \STATE $C = C - w^j$
+    \ENDFOR
+    \STATE Add $[x^j, \frac{1}{N_p}]$ to the new distribution
+\ENDFOR
+\STATE 
+\end{algorithmic}
+\label{alg:Resampling}
+\end{algorithm}
+For infinite particles this new distribution will match the old.
+Unfortunately, this isn't the truth in practice: since the support is
+still limited to the original particles, the number of \emph{unique} particles can only go down.
+This effect, dubbed "particle impoverishment" can result in excessive quantization
+errors in the final distribution. However, there is a solution. Instead of sampling from the
+discrete distribution, a smoothing kernel is applied, and particles are drawn from
+that distribution. Because the distribution is continuous, particle impoverishment
+cannot occur. The easiest way to sample from the continous distribution is to break the 
+re-sampling down into two steps. After calculating an estimate of the scale of the original
+distribution, \autoref{alg:Resampling} is performed. Next, a distribution is generated
+based on the original distributions, although it should have a somewhat smaller variance.
+Finally, for each particle in the resampled distribution, a sample is drawn from the smoothing 
+distribution and added to the particle.  The process of the selection will be defined as:
 
+\begin{equation}
+x_i = x_i + h\sigma \epsilon
+\end{equation}
+
+Where $h$ is an optional bandwidth, $\sigma$ is the standard deviation such that 
+$\sigma \sigma^T = cov(x)$
+and $\epsilon$ is drawn from the chosen kernel. \cite{Musso2001a} goes into
+significant 
+depth and proves the optimality of the Epanechnikov Kernel for reducing 
+MSE between the original and resampled distributions. However, \cite{Musso2001a}
+also espouses the usefullness of the Gaussian Kernel, due to the ease
+drawing samples from it, which for this work was more important.
+
+\begin{algorithm}
+\caption{Regularized Resampling Algorithm}
+\begin{algorithmic}
+\STATE Calculate Covariance, $C$, of empirical distribution, $\hat{x}$
+\STATE Find $D$ such that $DD^T = C$
+\STATE Resample $\hat{x}$ using \autoref{alg:Resampling}
+\FOR{$0 < i < N_p$}
+    \STATE Draw $\epsilon$ from the standard normal, same dimensionality as $X$
+    \STATE $x^i = x^i + hD\epsilon$
+\ENDFOR
 \end{algorithmic}
+\label{alg:Resampling}
 \end{algorithm}
 
-Unfortunately, this isn't necessarily the truth: since the support is
-still limited to the original particles, the number of unique particles can only go down.
-This effect, often dubbed "particle impoverishment" can result in excessive quantization
-errors in the final distribution. However, there is a solution. Instead of sampling from the
-discrete distribution, a smoothing kernel is applied, and $\hat{\chi}_j$ are drawn from
-that distribution. Because the distribution is continuous, there is no way for a collapse
-of the particles to occur. The question then, is how to decide on the smoothing kernel. 
-Often times the easiest way to sample from the continous distribution is to break the 
-re-sampling down into two steps. First a member of the discrete distribution is randomly
-selected based on the weights, and then based on the smoothing a nearby state variable 
-is selected. The process of the selection will be defined as:
-\begin{equation}
-\chi_i = x_i + h\sigma \epsilon
-\end{equation}
-Where $h$ is the bandwidth, $\sigma$ is the standard deviation such that $\sigma \sigma^T = cov(x)$
-and $\epsilon$ is drawn from the chosen kernel.
-It has been proven that when all the elements of the mixture
-have the same weight, as is the case after basic resampling, the kernel that minimizes the 
-MSE between the estimated and true posterior is the Epanechnikov Kernel (cite Improving Regularised
-Particle Filters, C Musso, N Oudjane and F LeGrand). 
-\begin{equation}
-K = \left\{
-\begin{array}{lr}
-\frac{n_x+2}{2c_{n_x}}(1-\|x\|^2) & if\ \|x\| < 1\\
-0 & otherwise
-\end{array}\right.
-\end{equation}
-%<more here>
+Thus, with regularized resampling, it is possible to prevent both
+particle degeneracy as well as particle impoverishment. An additional
+bonus is that as the particle filter converges, the density of particles
+in the area of the solution goes up. This has a similar effect to 
+simulated annealing where, as the algorithm approaches the end, the
+random steps get smaller and smaller. On the down side,
+if for some reason the real solution was not covered by the new support 
+generated by the resampling algorithm the algorithm could
 

### 2010-07-02 doc/thesis/methods.tex
more work on the thesis though methods is messy

@@
 
-\section{Choosing $\hat{Pr}(y(T) | x(T))$}
+\section{Model}
+As originally written in \autoref{sec:BOLD Physiology} the state variables
+for the BOLD model are as follows:
+\begin{eqnarray}
+\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
+\dot{f} &=& s\\
+\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\end{eqnarray}
+Because there are potentially
+
+\section{Choosing ${Pr}(y(T) | x(T))$}
 Choosing a representation of an unknown distribution is certainly tricky,
@@
 variable or an additive noise process with Gaussian steps.
+
+Another natural choice might be one of the robust estimator weight functions, for
+example the Huber or bi-square. For the purpose of this work we will stick
+with long tailed distributions, however it is worth noting that long tails
+may not be the optimal choice in all, or even this situation. The justification
+for long tailed distributions is that we believe the noise to be long tailed,
+and the variance of the noise is not well known.
+
+Therefore, we tried three weighting functions based on three distributions: Gaussian, 
+Laplace and the Cauchy. The standard deviation of the distribution is extremely
+important to the convergence of particle filter. A standard deviation that is 
+too large will not allow the distribution to converge in any reasonable number of 
+measurements. A standard deviation below the standard deviation of the noise 
+will cause the algorithm to throw out perfectly acceptable particles. 
+The weighting function ultimately will shape the output distribution, $P[y]$, into that
+distribution, however the distribution of $x$ will still approach a reasonable
+estimate of its true distribution. Even an overly wide weighting function, will
+allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
+parameters of the "real" posterior distribution, though the scale parameters may 
+be overly large.
+
+A reasonable method of setting the standard deviation of $\Omega$ may be by 
+taking a small sample from "resting" data and using the sample standard deviation.
+Since this is the first attempt at using particle filters for modeling the 
+BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
+because it gives a more consistency and control. Of course this could be taken
+further, by testing the sample data against a set of stock distributions
+and choosing the best fit. This of course depends on having enough samples
+to make a reasonable inference, which may not always exist.
 

### 2010-07-02 doc/thesis/particlefilters.tex
more work on the thesis though methods is messy

@@
 
-\subsection{Resampling}
+\section{Resampling}
 \label{sec:Particle Filter Resampling}
@@
 discrete distribution, a smoothing kernel is applied, and particles are drawn from
-that distribution. Because the distribution is continuous, particle impoverishment
-cannot occur. The easiest way to sample from the continous distribution is to break the 
+that distribution. Because it is continuous, particle impoverishment
+cannot occur. The easiest way to sample from the continuous distribution is to break the 
 re-sampling down into two steps. After calculating an estimate of the scale of the original
-distribution, \autoref{alg:Resampling} is performed. Next, a distribution is generated
-based on the original distributions, although it should have a somewhat smaller variance.
-Finally, for each particle in the resampled distribution, a sample is drawn from the smoothing 
-distribution and added to the particle.  The process of the selection will be defined as:
+distribution, algorithm \autoref{alg:Resampling} is performed. Next, a distribution is generated
+based on the variance of the original distributions.
+Finally, for each particle in the discretely re-sampled distribution, a sample is drawn from 
+the smoothing 
+distribution and added to the particle.  The process regularization is defined as:
 
@@
 MSE between the original and resampled distributions. However, \cite{Musso2001a}
-also espouses the usefullness of the Gaussian Kernel, due to the ease
+also espouses the usefulness of the Gaussian Kernel, due to the ease
 drawing samples from it, which for this work was more important.
@@
 \STATE Find $D$ such that $DD^T = C$
-\STATE Resample $\hat{x}$ using \autoref{alg:Resampling}
+\STATE Resample $\hat{x}$ using algorithm \autoref{alg:Resampling}
 \FOR{$0 < i < N_p$}
@@
 \end{algorithmic}
-\label{alg:Resampling}
+\label{alg:RegResampling}
 \end{algorithm}
+
+It has been proposed by \cite{Hurzeler1998} that if the underlying 
+distribution is non-Gaussian, then using the original bandwidth will over-smooth. 
+In reality, over smoothing
+will only become an issue if re-sampling is performed over and over. Thus
+if
+resampling is performed at every step then this could certainly caused problems.
+If the distribution is over-smoothed then the algorithm may not converge as rapidly;
+however, because the bandwidth is still based on particle variance, which decays as 
+particles are ruled out, it is still able to converge. In fact, over-smoothing is preferable
+to under smoothing, since over-smoothing simply slows convergence while 
+under-smoothing could result in false negatives.
+At the same time, as the number of dimensions in the state vector 
+increases, the standard deviation based approximation becomes less effective
+ (\cite{Arulampalam2002a}). Because of the high dimensionality of the BOLD model,
+and limited measurements, it is helpful to have a broader bandwidth to explore the distribution.
+
+\begin{algorithm}
+\caption{Regularized Particle Filter}
+\begin{algorithmic}
+\STATE Initialize Particles:
+\FOR{$i$ : each of $N_p$ particles }
+    \STATE $x^i_0  \sim \alpha(X)$
+    \STATE $w^i_0 = \frac{1}{N_p}$
+\ENDFOR
+\FOR{$k$ : each measurement}
+    \FOR{$i$ : each particle }
+        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
+    \ENDFOR
+
+    \STATE Calculate $N_{eff}$ with \autoref{eq:neff}
+    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
+        \STATE Resample using algorithm \autoref{alg:RegResampling}
+    \ENDIF
+\ENDFOR
+
+\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+ \end{algorithmic}
+ \end{algorithm}
+
+Nevertheless, because 
+of the potentially wide smoothing factor applied by regularized resampling, performing this
+step at every measurement would allow particles a great deal of mobility. This mobility is
+the opposite of convergence, which is why regularized resampling should only be done when
+$N_{eff}$ drops (say less than 50). Other than the periodic regularized
+resampling then, the regularized particle filter is nearly identical to the basic sampling
+importance sampling filter (SIS).
 
@@
 simulated annealing where, as the algorithm approaches the end, the
-random steps get smaller and smaller. On the down side,
-if for some reason the real solution was not covered by the new support 
-generated by the resampling algorithm the algorithm could
+random steps get smaller and smaller. There is a potential risk when
+resampling. If for some reason the solution is not covered by the 
+new support the algorithm will fail to converge properly.
 
-\subsection{Weighting Function}
-Because $P(y_k | x(T))$, what I will call the weighting function,
-is based on an unknown distribution, it is necessary to decide on a function
-that will approximate $P(y_k | x(T))$. Obviously the function, $\omega(y_k, f(x(t))$
-needs to be centered at zero and have a scale comparable to the signal levels.
-Obviously if the actual noise present in $y_k$ were to be known, then that would
-be the best distribution for $\Omega$. In that case, particles that fell
-far out on that distribution would be statistically impossible representations
-of the system, and it would be completely reasonable to throw such particles away.
-While a Gaussian function is the natural choice, because this distribution
-and weight are unknown we wanted to try distributions
-with wider tails, so that outliers don't completely destroy particle's weights
-(and thus convergence proceeds more slowly).
+The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
+or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
+nearby which allow for more accurate estimation near points of high likelihood. 
+As the variance of the estimated
+$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
+will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.
 
-Another natural choice might be one of the robust estimator weight functions, for
-example the Huber or bi-square. For the purpose of this work we will stick
-with long tailed distributions, however it is worth noting that long tails
-may not be the optimal choice in all, or even this situation. The justification
-for long tailed distributions is that we believe the noise to be long tailed,
-and the variance of the noise is not well known.
-
-Therefore, we tried three weighting functions based on three distributions: Gaussian, 
-Laplace and the Cauchy. The standard deviation of the distribution is extremely
-important to the convergence of particle filter. A standard deviation that is 
-too large will not allow the distribution to converge in any reasonable number of 
-measurements. A standard deviation below the standard deviation of the noise 
-will cause the algorithm to throw out perfectly acceptable particles. 
-The weighting function ultimately will shape the output distribution, $P[y]$, into that
-distribution, however the distribution of $x$ will still approach a reasonable
-estimate of its true distribution. Even an overly wide weighting function, will
-allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
-parameters of the "real" posterior distribution, though the scale parameters may 
-be overly large.
-
-A reasonable method of setting the standard deviation of $\Omega$ may be by 
-taking a small sample from "resting" data and using the sample standard deviation.
-Since this is the first attempt at using particle filters for modeling the 
-BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
-because it gives a more consistency and control. Of course this could be taken
-further, by testing the sample data against a set of stock distributions
-and choosing the best fit. This of course depends on having enough samples
-to make a reasonable inference, which may not always exist.
+\section{Weighting Function}
+Because the distribution of $\nu_y$ in \autoref{eq:measass} is unknown,
+it is necessary to choose a distribution for this. This distribution,
+is extremely important because $\nu_y \sim P(y_k | x(T))$, which is used
+for updating weights. Ideally this weighting function would exactly 
+match the measurement error in
+the output. It is logical to assume this error is centered at zero and 
+has a scale comparable to the signal levels.
+While a Gaussian function is the traditional choice, I wanted to try other
+distributions as well, given the unpredictable nature of the noise present in FMRI.
+The choice of this function will be discussed further in \autoref{sec:Methods
+Weighting Function}.
 
@@
 
-Although rectifiers are typically thought
-of as receiving an AC circuit 60 Hz, we will ignore such specifics and 
-assume the voltage across the output of the transformer is simply a scalar
-multiple of the input voltage.  As discussed in the \autoref{sec:Particle Filter Model}
+For this example it will be assumed that the transformer simply scales
+the input by a constant factor, $G$. As discussed in the \autoref{sec:Particle Filter Model}
 any variable with uncertainty must be part of the state variable. Therefore
-the state variable will be: $X(t) = \{G, v_t, C, R_m, v_y\}$. Of course,
-$u(t)$ cannot be allowed to be a square wave in such a system, since that
-such a signal would never get across the transformer and regardless it
-would necessitate an unrealistic infinite current across the capacitor.
+the state variable is defined as: $X(t) = \{G, v_t, C, R_m, v_y\}$. 
 The state equations would then be
@@
 Gaussian seems like a good idea, all the static state variables are strictly
-positive and thus not well suited to the Gaussian. In this case then,
-it would be wise to start with a Gamma distribution, and just be wary of
-any standard deviation that gets larger than the prior mean. We will define
-the gamma distribution as follows:
+positive and thus not well suited to the Gaussian. Thus, instead the prior 
+will start with a Gamma distribution. The gamma is defined as follows:
 
@@
 where $\Gamma$ is the gamma function.
-The in some ways the margin for error is decided by the weighting function, which
-here will define as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
+The margin for error is decided by the weighting function, which
+will be defined as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
 estimate based on all the particles, and 
-$v_{yi}$ is the estimate by a particular (i$^{\text{th}}$) particle. The choice of this function is difficult,
-and although the Gaussian is typically used, we found the exponential helpful
-in dealing with particle deprivation.  The algorithm will then look like the following,
+$v_{yi}$ is the estimate by a particular ($i^{th}$) particle. The choice of this function is difficult,
+and although the Gaussian is typically used, in practice I found the exponential helpful
+in preventing particle deprivation. The algorithm is then,
 
@@
 runs, weights will compound resulting in a distribution that asymptotically
-approaches the true joint distribution of the $X(t)$.  Of course, as we
-mentioned in \autoref{sec:Particle Filter Resampling}, particles weighted zero do not significantly
+approaches the true joint distribution of the $X(t)$.  Of course, as I
+mentioned in \autoref{sec:Particle Filter Resampling}, particles weighted 
+zero do not significantly
 contribute to the empirical distribution, so re-sampling may be necessary.
-If the noise is assumed to be Gaussian then it is possible to further optimize. 
-Thus we let $h$ be defines as:
-\begin{eqnarray}
-h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}
-\end{eqnarray}
-and although it is very possible the underlying noise is non-gaussian, the Gaussian
-may work, but sub-optimally. It has been proposed that (Monte Carlo Approximations for
-General State-Space Models, markus Hurzeler and Hans R. Kunsch) if the underlying 
-distribution is non-Gaussian, then using this bandwidth will oversmooth. 
-In reality over smoothing
-should not be too great an issue because the smoothing is only being applied to find new
-particles. If the distribution is over smoothed then the algorithm may not converge as rapidly;
-however, because the bandwidth is still based on particle variance, which will decay as 
-particles are ruled out, it is still able to converge. In fact over smoothing is preferrable
-to under smoothing, since the latter would result in false negatives, but the previous only
-results in a slower decay of the variance. 
-At the same time, as $n_x$, the number of dimensions in
-$x$, goes to infinity, the standard deviation based approximation becomes less effective
-(cite a Tutorial on Particle Filters for on-line non-linear non-gaussian bayesian
-tracking, sanjeev arulampalam, simon maskell, neil gordon...).  Because of the high dimensionality of our system,
-and limited measurements, it is helpful to have a broader bandwidth to explore the distribution. 
-Nevertheless, because 
-of the potentially wide smoothing factor applied by regularized resampling, performing this
-step at every measurement would allow particles a great deal of mobility. This mobility is
-the enemy of convergence, which is why regularized resampling should only be done when
-$\hat{N}_{eff}$ drops very low (say less than 50). Other than the periodic regularized
-resampling then, the regularized particle filter is nearly identical to the basic sampling
-importance sampling filter (SIS).
-
- \begin{algorithmic}
-\STATE Initialize $N_p$ Particles: 
-        $\{x_i(0),w_i(0) : x_i(0) \sim \alpha(x), w_i(0) = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$
-\STATE $T$ = \{Set of Measurement Times\}
-\FOR{$t$ in $T$}
-    \FOR{$i$ in $N_p$}
-        \STATE $x_i(t) = x_i(t-1) + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w_i(t) = w_i(t-1)P(y(t) | x(t))$
-    \ENDFOR
-
-    \STATE Calculate $N_{eff}$ with \autoref{eq:neff}
-    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
-        \STATE Calculate empirical $\sigma$ 
-        \STATE $h = [N_s8c^{-1}_{n_x}(n_x + 4)(2\sqrt{\pi})^{n_x}]^{\frac{1}{n_x +4}}$
-        \STATE Redraw particles using (stratified) basic resampling
-        \FOR{$i$ in $N_p$}
-            \STATE Draw $\epsilon \sim K$
-            \STATE $x_i = x_i + h \sigma \epsilon$
-        \ENDFOR
-    \ENDIF
-\ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
-
- \end{algorithmic}
-
-The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
-nearby which allow for more accurate estimation near points of high likelihood. 
-As the variance of the estimated
-$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
-will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.

### 2010-07-02 doc/thesis/background.tex
started cleaning up notation in particle filter
more work....

@@
 both proposed physiologically reasonable values for the model parameters, 
-\cite{Friston2002} was the first paper to calculate the parameters based 
+\cite{Friston2002b} was the first paper to calculate the parameters based 
 on actual FMRI data. However, in that case, the voxels were chosen from
@@
 
-\section{Hybrid Methods}
+\subsection{Hybrid Methods}
 In \cite{Riera2004}, a maximum

### 2010-07-02 doc/thesis/intro.tex
more work....

@@
 paper we will use only a simple version of the Balloon model, described by
-\autoref{eq:bold}, to keep the solution tractable.
+\autoref{eq:bold}, to keep the solution tractable, and because it is the most
+well studied.
 
 Another interesting result of \cite{Deneux2006} was the sensitivity analysis.
-There it was found that many of the parameters compensate for each other,
+There it was found that the parameters are far from perpendicular,
 implying that exact inference of parameters may not be possible without
-constraint.
+constraint. This could explain the extreme discrepencies in \autoref{tab:Params}.
 
@@
 Balloon parameters, those results are summarized in \autoref{tab:Params}
+
+%note to self, friston2002b's parameters are from a picture
 
@@
 \hline 
-Parameter  & \cite{Friston2002} & \cite{Johnston2008} & \cite{Vakorin2007} \\
+Parameter  & \cite{Friston2002} & \cite{Johnston2008} & \cite{Vakorin2007} & \cite{Deneux2006}\\
 \hline
-$\epsilon$ &  $N(.54 , .1 ^2)$  & $.069 \pm .014 $ & (NC)\\
-$\tau_s  $ &  $N(1.54, .25^2)$  & $4.98 \pm 1.07 $ & $2.2$\\
-$\tau_f  $ &  $N(2.46, .25^2)$  & $8.31 \pm 1.51 $ & $.45$\\
-$\tau_0  $ &  $N(.98 , .25^2)$  & $8.38 \pm 1.5  $ & $.94$\\
-$\alpha  $ &  $N(.33 , .45^2)$  & $.189 \pm .004 $ & $.4$ (NC) \\
-$E_0     $ &  $N(.34 , .1 ^2)$  & $.635 \pm .072 $ & $.6$ (NC) \\
-$V_0     $ &  $.03$ (NC)        & $.0149 \pm .006$ & (NC)\\
+$\epsilon$ &  $N(.54 , .1 ^2)$  & $.069 \pm .014 $ & (NC) & .89\\
+$\tau_s  $ &  $N(1.54, .25^2)$  & $4.98 \pm 1.07 $ & $2.2$ & 2.04 \\
+$\tau_f  $ &  $N(2.46, .25^2)$  & $8.31 \pm 1.51 $ & $.45$ & 5.26\\
+$\tau_0  $ &  $N(.98 , .25^2)$  & $8.38 \pm 1.5  $ & $.94$ & .27\\
+$\alpha  $ &  $N(.33 , .45^2)$  & $.189 \pm .004 $ & $.4$ (NC) & .63 \\
+$E_0     $ &  $N(.34 , .1 ^2)$  & $.635 \pm .072 $ & $.6$ (NC) & .33\\
+$V_0     $ &  $.03$ (NC)        & $.0149 \pm .006$ & (NC) & .16\\
 \hline

### 2010-07-02 doc/thesis/methods.tex
more work....

@@
 \chapter{Methods}
+Although the particle filter I used for this work is a standard Regularized
+Particle filter, as described in \cite{Arulampalam2002a}, optimizing the 
+particle filter for use with FMRI data is non-trivial.
+
 \section{Prior Distribution}
-For the BOLD signal all the parameters have been studied and have relatively well
-known mean and variance, so a prior is helpful. 
-For strictly positive parameters (members of $x$) we used a gamma distribution,
-whereas for parameters that could be negative, we used a Gaussian distribution. In
-both cases standard deviations twice that found in previous studies were used.
+For the BOLD model described in \autoref{sec:BOLD Physiology}, several
+different studies have endeavored to calculate parameters. The results
+of these studies may be found in \autoref{tab:Params}, and the methods 
+used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
+\cite{Friston2000} only studied small regions; and most research with
+reasonable speed (including \cite{Friston2002b}) used these results as 
+the source for their priors. 
+The one exception is \cite{Johnston2008}, which came to an extremely different
+set of distributions. For a particle filter, the choice of a prior is
+the single most important design choice. Consequently for this work,
+I wanted to be conservative; which meant going with the accepted result,
+\cite{Friston2000}. This constrains the usefulness of the model to
+areas that fall within the prior distribution, yet will allow results
+to be comparable to other works. There is a significant need for better
+estimates of the physiological parameters; and, while physical experiments
+may not be possible, it would not be unreasonable to do a study with
+exhaustive simulated annealing or hill climbing tests for multiple
+regions and multiple patients. The purpose for this work is to determine
+the fitness of particle filters for this task; because with good 
+priors it can be extremely fast. Ultimately this algorithm may become
+more useful as more studies are done on BOLD model that don't require
+priors.
+
+Therefore, because \cite{Friston2000} represents the current best
+knowledge about the BOLD parameters, it forms the basis
+for the prior distributions in this paper.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c || c | c |}
+\hline 
+Parameter & Distribution & $\mu$ & $\sigma$ \\
+\hline
+$\tau_0$ & Gamma & .98 & .5  \\
+$\alpha$ & Gamma & .33 & .045\\
+$E_0$    & Gamma & .34 & .03  \\
+$V_0$    & Gamma & .03 & .006 \\
+$\tau_s$ & Gamma & 1.54  & .5 \\
+$\tau_f$ & Gamma & 2.46  & .5 \\
+$\epsilon$ & Gamma & .64 & .4 \\
+\hline
+\end{tabular}
+\caption{}
+\label{tab:Prior} 
+\end{table}
+
+Note that although the mean remains the same for all the 
+parameters other than $\epsilon$, the standard deviation is set
+much higher to account for the disagreement between studies
+(\autoref{tab:Params}). 
+Because all the parameters are taken to be strictly positive, and the
+standard deviations are approaching the mean, I used a gamma distribution.
+This prevents the Gaussian from placing parameters in the nonsensical 
+territory of negative activation, or negative time constants.
 
@@
 \end{eqnarray}
-Because there are potentially
+The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
+included noise in the update of $x$, however that is not included here.
+The reason for the difference is that cloud of particles is, to some extent,
+able to model that noise. It is common, however, to actually model that noise
+in particle filters by adding a random value to each updated state variable. 
+Because the purpose of this particle filter is to learn the underlying distribution
+of the static parameters, rather than precisely model the time course of the 
+in the dynamic parameters ($\{s,f,v,q\}$) this noise is left out. Because the
+BOLD model is dissipative, when no stimuli are applied, all the particles 
+will decay to ($\{0,1,1,1\}$) which they should. If evolution noise were added, the 
+particles would then weight particles based on the results of that noise, rather
+than on the quality of the static parameters. Typical particle filter uses
+also use this state noise as an exploratory measure; however given the 
+high dimensionality of the system, a vast number of particles would be 
+necessary for this to come to fruition. The noise in the BOLD signal is such
+that several particles all tend to be equally correct; the particle filter will
+still leave variance in the solution; which turned out to be the case.
+
+Thus, at each time step, the states were linearized, and each state variable,
+$s$, $f$, $v$, $q$, were changed according their previous rate of change.
+Because of the difficulty of solving a system of nonlinear equations, I 
+did not use the typical Runge-Kutta 4/5 technique to optimize step sizes. 
+The cost of missing a feature in these differential equations typically leads
+to non-real numbers shortly down the road. Because the non-real numbers do 
+not come until the solution has had the chance to update two or three more
+times, taking long steps can result in catastrophic and un-recoverable errors. 
+Typically a step size of .001 was used, after finding that even .01 can
+at times lead to the state equations careening out of control.
+
+\section{Resampling}
+The algorithm for resampling is described in \autoref{sec:Particle Filter Resampling}.
+The primary design decision for resampling is the regularizing
+kernel. As mentioned previously, the Gaussian kernel is extremely convenient,
+because it is so easy to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
+as long as resampling is kept as a last resort, some amount of over-smoothing
+won't impair convergence. Therefore, for this work I chose a Gaussian kernel of
+bandwidth equal to the original distribution's covariance. Obviously this will
+apply a rather large amount of smoothing to the distribution; however, on average
+resampling is only applied every 10 to 20 measurements, and because randomization
+is being applied to model updates this gives the filter some mobility.
+
+The question of how often to resample is complex. The benefit of resampling
+is that particles that were minimally important can give much more 
+granularity in the middle of the distribution. At the same time, resampling
+right after a batch of outliers can result in an incorrect distribution.
+Because the memory of particle weights is not particularly long, "good" particles
+could be removed by resampling very often. Additionally, because regularized
+resampling is being with a large bandwidth, I did not want the particle filter
+to be prevented from converging by resampling too often. Therefore, only
+when the effective sample size dropped below 25 for two measurement points
+in a row did I resample. This seemed to give good results, and avoided 
+temporary drops in the $N_{eff}$ caused by outliers.
 

### 2010-07-02 doc/thesis/intro.tex
almost done with methods

@@
 \centering
-\subfigure[]{\label{fig:QQDDlta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
-\subfigure[]{\label{fig:QQDDlta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
-\subfigure[]{\label{fig:QQDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
-\subfigure[]{\label{fig:QQDDlta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
+\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
+\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
+\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
+\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
 \caption{Q-Q Plots of resting state data, using the BOLD signal changes}
@@
 \centering
-\begin{tabular}{|c || c | c | c |}
+\begin{tabular}{|c || c | c | c | c|}
 \hline 

### 2010-07-02 doc/thesis/methods.tex
almost done with methods

@@
 \centering
-\begin{tabular}{|c || c | c |}
+\begin{tabular}{|c || c | c | c |}
 \hline 
@@
 
-\section{Choosing ${Pr}(y(T) | x(T))$}
-Choosing a representation of an unknown distribution is certainly tricky,
-and so the fact that $\hat{Pr}(y(T) | x(T)) = \nu_d + \nu_y$ means that
-there is a significant piece of the algorithm that is based primarily conjecture.
-Studies of the noise in FMRI typically attribute noise to a Gaussian random
-variable or an additive noise process with Gaussian steps.
+\section{Choosing $P(y_k | x_k)$}
+Choosing a representation of an unknown distribution tricky, and while
+the usual choice is a Gaussian distribution, there are reasons why it
+may not be the best.  Studies of the noise in FMRI typically characterise
+noise as an additive noise process with Gaussian steps; a Wiener process.
 
-Another natural choice might be one of the robust estimator weight functions, for
-example the Huber or bi-square. For the purpose of this work we will stick
-with long tailed distributions, however it is worth noting that long tails
-may not be the optimal choice in all, or even this situation. The justification
-for long tailed distributions is that we believe the noise to be long tailed,
-and the variance of the noise is not well known.
+As noted in \autoref{sec:Introduction Noise}, the noise is not strictly Gaussian,
+nor is it strictly Wiener. As with any unknown noise however, it is necessary 
+to make some assumption. If the weighting function ($P(y_k | x_k)$) exactly
+matches the measurement error, then the ideal particle filter will result.
+Particles with $x_k$'s that estimate $y_k$ far out on the weighting
+function will quickly have weights near 0. Thus, an weighting function that
+exactly matches $P(Y(t) | X(t))$ will easily, and correctly throw out incorrect
+particles. However assuming that this function is not going to be found 
+for arbitrary voxels, it is necessary to choose a function to approximate this 
+probability. The cost of choosing an overly broad distribution for this
+function is an overly broad $P(x_k | y_k)$; and ultimately an overly
+broad $P(x_k | y_{0:k})$. On the other hand, an overly thin distribution
+will lead to, at best, an overly thin representation of these same 
+distributions; but at wost will lead to particle deprivation (all particles
+being zero-weighted). If the \emph{true} distribution were found, particle
+deprivation would indicate that that time series is not being affected
+in any way describable by the prior distribution/model. Particle deprivation
+due to an under-variant weighting function gives no information. Thus,
+the cost of under variance is significantly higher; a fact that definitely needs
+to be taken into account.
 
-Therefore, we tried three weighting functions based on three distributions: Gaussian, 
-Laplace and the Cauchy. The standard deviation of the distribution is extremely
-important to the convergence of particle filter. A standard deviation that is 
-too large will not allow the distribution to converge in any reasonable number of 
-measurements. A standard deviation below the standard deviation of the noise 
-will cause the algorithm to throw out perfectly acceptable particles. 
-The weighting function ultimately will shape the output distribution, $P[y]$, into that
-distribution, however the distribution of $x$ will still approach a reasonable
-estimate of its true distribution. Even an overly wide weighting function, will
-allow the Gaussian Mixture estimate of $X$ to converge to the correct location 
-parameters of the "real" posterior distribution, though the scale parameters may 
-be overly large.
+Because of the need to err on the side of caution I tested several weighting
+functions. In addition to the Gaussian I also tested the Laplace and Cauchy
+distributions, both of which have much wider tales than the Gaussian. The
+benefit of the wider tailed distributions is that they don't down-weight
+particles quite as fast. Additionally, the Laplace distribution has the
+benefit of having a non-zero slope at the origin. This means that even
+if the distribution is made overly broad, it will still distinguish between
+particles that are near the origin. Results with real data compare 
+the three distribution in \autoref{sec:Results Weights}.
 
-A reasonable method of setting the standard deviation of $\Omega$ may be by 
+After trial and error, I chose the weight standard deviation to be $.005$. Obviously
+the case of the Cauchy, this is simply the scale parameter. While I 
+did attempt to automatically set the standard deviation, I found that the algorithm
+became too unpredictable. Essentially, if the weight function isn't
+fixed across voxels, very noisy time series with no actual signal will 
+converge to nonsensical results. 
+In the future, it may be possible to set the standard deviation by
 taking a small sample from "resting" data and using the sample standard deviation.
 Since this is the first attempt at using particle filters for modeling the 
-BOLD model, in this work we set the standard deviation manually at <weight standard dev>,
-because it gives a more consistency and control. Of course this could be taken
-further, by testing the sample data against a set of stock distributions
-and choosing the best fit. This of course depends on having enough samples
-to make a reasonable inference, which may not always exist.
+BOLD model, in this work I set the standard deviation manually at $.005$,
+because it gives a more consistency and control.
 
@@
 The non-stationary
-aspect of a Weiner process as with $\nu_d$ is difficult to compensate for, and so various methods
+aspect of a Weiner process, presumably the result of integrating some
+$\nu_x$ is difficult to compensate for, and so various methods
 have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
-demonstrated that this component is prevalent, and may in fact be a characteristic
-of FMRI. In some studies, as many as half the voxels benefit from detrending, meaning
-that this is certainly a serious barrier to inference. All the existing methods are performed
+demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
+of FMRI. In some studies, as many as half the voxels benefit from detrending, presenting
+a serious barrier to inference (\cite{Smith2007}). All the existing methods are performed
 during the preprocessing stage, rather than as an integral part of analyzing the BOLD
 signal. There is no shortage of theories on the "best" method of detrending; however
-a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
+in a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
 a spline works the best. The benefit of the spline versus wavelets, high pass 
@@
 shift heavily over the course of time. In spite of this, the spline will still remove some
-amount of signal, just like all of these methods, which is why the the method proposed
-in \autoref{sec:Methods Delta Based Inference} was considered.
+amount of signal, just like all of these methods, which is why I considered the method proposed
+in \autoref{sec:Methods Delta Based Inference}.
 
-When using the median-based spline techniques, the resulting signal is will almost
-certainly end with a median of zero; in other words half the signal will be above zero,
-half below. The problem with this is this is not the "natural" BOLD signal. More specifically,
+The method I used to calculate the spline was picking one knot for every 15
+measurements in an image. Thus a 10 minute session at a repetition time of 
+2.1 seconds would have 19 knots. The knot first and last knots were each 
+given half the number of samples as the rest of the knots; which were all 
+located at the center of their sample group. The median of each sample group
+was then taken and used as the magnitude for the group. Taking the median 
+versus the mean seemed to work better, given the presence of outliers. It seems
+that there is potential to optimize the spline further using a canonical 
+HRF to find resting points; however, for this to work the experiment would have
+to be designed with this in mind.
+
+When using the median-based spline techniques, the normalized signal will,
+by definition have a median near zero. 
+The problem with this is this is not the "natural" BOLD signal. More specifically,
 when the signal is inactive, the BOLD response \emph{should} be at 0\% change from
@@
 After removing the spline, the BOLD resting state will be below 0\%.
-This is troublesome because it reduces the power the BOLD model. One obvious solution
-is to add an arbitrary constant to each BOLD response. Of course this won't scale well
-to whole brain analysis. A more effective technique is adding a DC gain to the BOLD
-model. Essentially every measurement would then have a constant level added to it. 
-Like all the other model parameters, enough measurements should allow the correct
-value to be found. On the downside adding another dimension increases the
-complexity of the model, for a variable that is relatively obvious to a human.
+This is problematic because it reduces the ability of an algorithm to learn.
+One quick and dirty solution is to add an arbitrary constant to each BOLD response. 
+Of course this won't scale to whole-brain analysis, so a more effective technique 
+is adding a DC gain parameter to the BOLD model. 
+Like all the other model parameters, given enough measurements, correct
+value may be found. On the downside, adding another dimension increases the
+complexity of the model, for a variable that is relatively obvious by direct
+observation.
 
-Thus, a more conventional approach was used. A robust estimator of scale, was used
-to determine roughly where the base level was. The Median Absolute Deviation (MAD)
+Thus, I used a more conventional approach to deal with this. To determine
+the DC gain to be added to the signal I used a robust estimator 
+of scale. The Median Absolute Deviation (MAD)
 proved to be extremely accurate in determining how much to shift the signal up
-by. We tested both methods during the course of analysis, and found that the increase 
-in model complexity far outweighted the slight increase in flexibility. Its 
+by. I tested both methods during the course of analysis, and found that the increase 
+in model complexity far outweighed the slight increase in flexibility. Its 
 possible that a more accurate method may exist; however, for this case the 
-MAD seemed to work perfectly, as \autoref{fig:DesplineQuality} shows.
+MAD seems to work as it should, as \autoref{fig:DesplineQuality} shows.
+
+\begin{equation}
+y_{\text{gain}, 0:K} = 2\text{median}_{i=0:K}(y_i - \text{median}(y_{0:K}))
+\end{equation}
 
@@
 
+A serious concern with adding and subtracting arbitrary values to 
+real data is whether this will create false positives. This is a legitimate
+concern; however, all the additions and subtractions done in this section
+have been very low frequency changes and should not substantially change
+the BOLD response, being a relatively high frequency signal. The typical
+method of preprocessing is a high-pass filter with a cutoff of around 30
+seconds; which should about match the spline with knots every fifteen 
+measurements. \cite{Tanabe2002} found that splines tended to far outperform the
+high-pass filter method.
+
 \subsection{Delta Based Inference}
 \label{sec:Methods Delta Based Inference}
-I also propose and test a different method of dealing with the so called "drift". 
+The alternative to these sorts of low frequency manipulation is to
+go around the problem in another way. Here, I propose and test a 
+different method of dealing with the so called "drift". 
 Instead of comparing the direct output of the particle filter with the direct
-measurement, the algorith compares the change in signal over a single TR. 
-In most signal processing cases this would foolish, but that is because the 
-general assumption that all noise is high frequency is not the case here.
-In fact, every pipeline for the analysis of BOLD signal uses a high pass filter,
-but low poss filters are rarely applied, because it is a well known fact that 
-most of the signal is in the high frequency range and most of the noise is actually 
-in the low frequency range. The particle filter is an 
-extremely robust method of inference, and so I would assert that the particle
-filter ought to be given as \emph{raw} data as possible. While taking direct measurements
+measurement, the algorithm would compare the change in signal over a single TR,
+with the result of integrating the model for the same period. 
+In most signal processing cases this would be foolish, but that is because the 
+general assumption is that all noise is high frequency. Considering 
+the fact that every BOLD analysis pipeline uses a high pass filter,
+whereas low poss temporal filter are rarely applied, it makes sense
+that a derivative type method could work. The benefit of particle filters
+is that they are extremely robust method of inference, and I would assert 
+that the particle filter ought to be given as \emph{raw} data as possible. 
+While taking direct measurements
 without de-trending would give awful results, using the difference removes the 
-DC component and turns a Weiner process into a Gaussian random variable.
+DC component and turns what is usually assumed to be a Weiner process into 
+a simple Gaussian random variable.
 
@@
 
-Because $\nu_d$ is a Weiner process, then $\nu_d(t) - \nu_d(t-1)$ is simply a Gaussian step. If
-$\nu_d$ is some other additive process, the difference will still be one of a few stable
-distributions. If $\nu_y$ is i.i.d. then the resulting distribution will still be zero mean
-with a maximum variance of twice the original variance. All the assumptions made originally
-for the particle filter hold, and all of the parameters may be distinguished based on
-the step sizes, thus it is not unreasonable to attempt to match the string of step sizes
+Even if $\nu_d$ is some other additive process, the difference will still be closer
+to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
+in \autoref{fig:QQDDelta} in \autoref{sec:Introduction Noise}. 
+ All the assumptions made originally
+for the particle filter still hold, and all of the parameters may be distinguished based on
+the step sizes, thus it is not unreasonable to consider matching the string of step sizes
 rather than string of direct readings.
@@
 FMRI involves a great deal of preprocessing. In this paper we make an effort to
-minimize any type of preprocessing that will degrade the signal. 
+minimize any type of preprocessing that will modify the signal. 
 After FMRI data has been acquired it is always necessary to modify the

### 2010-07-03 doc/thesis/background.tex
little more on methods

@@
 \subsection{Random Field Theory}
+\label{sec:RFT}
 SPM methods make significant use of T-Tests across large regions;

### 2010-07-03 doc/thesis/methods.tex
little more on methods

@@
 territory of negative activation, or negative time constants.
+
+Another aspect of the prior is using enough particles to get a 
+sufficiently dense approximation of the prior. For 7 dimensions,
+getting a dense prior is extremely difficult. Insufficiently
+dense particles will result in inconsistent results, of course the
+processing time will scale up directly with the number of particles.
+A dense initial estimate is important so that some particles land
+near the solution; but as the variance decreases the number of 
+particles needed decreases as well. Thus, as a heuristic, initially
+the number of particles is set to 28,000, but after resampling,
+the number of particles is dropped to 1,000. Typically during the 
+first few measurements the variance drops precipitously since most particles
+don't describe the system well. The particles that are left are in a
+much more compact location, allowing them to be estimated with 
+significantly fewer particles. These numbers aren't set in stone,
+and depending on the complexity of the system or desired accuracy
+they could be changed; however, they seem to be the minimum that
+will give consistent results.
 
@@
 \subsection{Classical De-trending}
+\label{sec:Detrend}
 The non-stationary
@@
 As discussed in the section on de-trending, the normal pipeline for analyzing
-FMRI involves a great deal of preprocessing. In this paper we make an effort to
-minimize any type of preprocessing that will modify the signal. 
-After FMRI data has been acquired it is always necessary to modify the
-data in some way to make different runs comparable. Because FMRI signal
-levels are unit-less, at the very least it is necessary to convert
-the data into \% difference from the baseline. This process removes no data
+FMRI involves a great deal of preprocessing. The first and most important
+task is motion correction. To do this, a single volume in time is chosen, and
+volumes at every other time are registered to this one volume. This takes care
+of motion by the patient as well as some small changes in the magnetic
+fields that cause the image to shift. After this, a structural image of the
+patient is often registered to an atlas so that a grey-matter mask may
+be created. Since all the "activation" should occur in grey matter voxels,
+it made sense to restrict analysis to only these sections; although
+there may be partially grey-matter areas that are missed. This significantly
+eases computation time though.
+
+In normal statistical parametric mapping, a gaussian smoothing
+filter is applied across the image as discussed in \autoref{sec:RFT}.
+After this, detrending is performed discussed by \autoref{sec:Detrend}.
+Recall that FMRI signal levels are unit-less and even though detrending isn't
+always necessary, at the very least it is necessary to convert
+the data into \% difference from the baseline. Changing to 
+\% difference removes no real data
 from signal since it merely subtracting then dividing by a constant. This

### 2010-07-03 doc/thesis/methods.tex
mostly done with methods for real this time

@@
 Recall that FMRI signal levels are unit-less and even though detrending isn't
-always necessary, at the very least it is necessary to convert
-the data into \% difference from the baseline. Changing to 
-\% difference removes no real data
-from signal since it merely subtracting then dividing by a constant. This
-is the signal that was input into the delta based particle filter.
-Of course there are much more advanced ways of performing this task.
-The generally accepted standard is actually to use a high pass filter, although the
+always necessary, at the very least the data must be converted 
+into \% difference from the baseline. Changing to 
+\% difference removes no real information
+from signal. This is the signal that was input into the delta based 
+particle filter. Of course, most of the time analysis is performed on the
+direct signal; which mandates the removal of low frequency drift.
+The generally accepted method is to use a high pass filter, although the
 cutoff frequency is application dependent and often applied haphazardly.
-The high pass filter thus removes the DC component of the signal, and 
-some amount of the so called "drift". The problem with this method is that it is
-not adaptive to the input. Huge variations in drift frequencies can exist 
-in a single time-series. Thus, a single cutoff frequency could miss a significant
-drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
-set too high. This is why, as I mentioned in the De-trending section, a spline
-based detrending method will generally give better results.
 
-For simulated and real images (tests with multiple time-series), tests were 
-also run with and without Gaussian filtering with sigma of %not sure todo
-were run, since it is standard
-practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
-a spatial filter such as Gaussian filtering increased SNR but can also lead to less
-precision in the output maps.
+%For simulated and real images (tests with multiple time-series), tests were 
+%also run with and without Gaussian filtering with sigma of %not sure todo
+%were run, since it is standard
+%practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
+%a spatial filter such as Gaussian filtering increased SNR but can also lead to less
+%precision in the output maps.
 
 \section{Simulation}
-We performed two types of simulations. First, we simulated a single BOLD time-series based
-on a random chosen set of model parameters. This process was relatively straight forward
+I performed two types of simulations. First, I simulated a single BOLD time-series based
+on a randomly selected set of model parameters. This process was relatively straight forward,
 given the state-space equations for the BOLD signal. After a "true" signal was generated,
-we then added a carrier level, since BOLD is typically measured as a \% difference from the
-base level. Finally, we added Gaussian noise, and a Weiner Process to the clean signal. The
-variance of the Gaussian noise may be expressed in terms of the desired noise SNR, $R$ as:
-\begin{equation}
-var(y_{noisy}) = var(y) / R
-\end{equation}
-Since SNR doesn't have quite the same meaning for a Wiener process based noise, the variance 
-of the Gaussian steps was set to be:
-\begin{equation}
-var(y_{noisy}) = var(y) / (4R)
-\end{equation}
+I then added a carrier level, since BOLD is typically measured as a \% difference from the
+base level. I tested the particle filter with two types of noise; independent and identically
+distributed gaussian noise and weiner drift. For both I tested with various standard deviations.
+
 Once this noisy simulated time series was generated, the exact same particle filter algorithm
-that would later be run on full sized images, was run on this single voxel image. We ran
+that would later be run on full sized images, was run on this single voxel image. I ran
 a series of tests to determine the convergence rate of the particle filter, the number
@@
 
-Second we used a modified version of the FSL tool 
+Second I used a modified version of the FSL tool 
 POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
-by creating a random image, smoothing it with a large Gaussian kernel, then thresholding
-the results. Finally connected regions were each given a set of parameters from a finite
-list of randomly chosen parameter sets. The result was a four dimensional (length x width
-x height x parameter) image with spatially varying parameters. Time-series of activation level
-was generated for each set of parameters, then activation levels were fed into POSSUM's 
-function for generating frequency domain data. The patche for POSSUM will be made available.
-For each time-series in the simulate FMRI image, the final \emph{static} parameters are saved
+by taking an existing activation map and assigning discrete parameter sets to each region.
+The result was a four dimensional (length x width
+x height x parameter) image with spatially varying parameters. Possum was then modified
+to take a parameter map and generate activation levels depending on the parameters at that
+point. The patch for POSSUM will be made available. As an unfortunate side effect of 
+not using Possums' original activation scale, I manually added 750 to the total level of
+simulated Possum images. This is because the BOLD \% difference levels were in the range
+of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
+this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
+For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
 into a parameter map. This parameter map may then be compared to the map used to generate the 
 simulated data; additionally a new simulation using the calculated parameters may also be 
-generated to test the functional difference between the two maps. This would give an absolute 
-quantitative difference between the two parameter sets irrespective to parameter slopiness.
-So for instance, if $V_0$ is halved, $\epsilon$ doubling may very well give a similar result.
-In this case the \% difference between the parameters will be large in each case, but the functional
-difference between the parameters will not be great. This is obviously a bad situation, which
-is why we wanted to test for it.
+generated to test the difference in activation levels between the real parameters and the
+estimated ones. Since it is clear that the parameters are not fully orthogonal 
+(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
+but have different parameters. This way, an absolute 
+quantitative difference between the two parameter sets may be found.
 
@@
 
-The final result from calculating parameters with the real data was similar to that
-from the results from the POSSUM simulated data. The difference being that there was
-no ground truth the check it with.
+Before performing tests on a full image, I tested the results of the particle filter
+on regions deemed active and non-active by statistical parametric mapping.

### 2010-07-03 doc/thesis/methods.tex
some changes to methods

@@
 kernel. As mentioned previously, the Gaussian kernel is extremely convenient,
-because it is so easy to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
+because it is very simple to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
 as long as resampling is kept as a last resort, some amount of over-smoothing
@@
 apply a rather large amount of smoothing to the distribution; however, on average
-resampling is only applied every 10 to 20 measurements, and because randomization
+resampling is only applied every 20 to 30 measurements, and because randomization
 is being applied to model updates this gives the filter some mobility.
 
-The question of how often to resample is complex. The benefit of resampling
-is that particles that were minimally important can give much more 
-granularity in the middle of the distribution. At the same time, resampling
-right after a batch of outliers can result in an incorrect distribution.
-Because the memory of particle weights is not particularly long, "good" particles
-could be removed by resampling very often. Additionally, because regularized
-resampling is being with a large bandwidth, I did not want the particle filter
-to be prevented from converging by resampling too often. Therefore, only
-when the effective sample size dropped below 25 for two measurement points
-in a row did I resample. This seemed to give good results, and avoided 
+Because the regularized resample will almost certainly over-smooth,
+it is necessary to wait until the $N_{eff}$ drops relatively low
+(say below 25) before resampling is performed. As an additional 
+measure against sharp drops in the $N_{eff}$ that may quickly raise
+back up, resampling is only performed when two consecutive low
+$N_{eff}$'s are found. 
+The danger in waiting longer to resample is that if there were no
+particles in the vicinity of the solution, then particle deprivation
+can occur.  When this happens, the resampling
+stage will have an inappropriately small variance which will lead to an 
+inappropriately small support for the distribution. After this, the 
+distribution will be unrecoverable. Thus, when the $N_{eff}$ gets down
+to extremely low values, usually this is an indication of particle
+deprivation; something that is also usually accompanied by extremely
+fast drops in variance. When this happens, it is preferable 
+to return to a previous, wider estimate of scale for the regularization
+and continue from there. Thus to keep particle deprivation from 
+affecting the results; a previous covariance matrix is 
+used as the regularization bandwidth. In my algorithm I use
+the last covariance matrix for which the $E_{eff}$ was greater than
+the threshold. This guarantees that there was at least some amount of 
+particle diversity, and also helps prevent converging prematurely.
+
+When some sort of particle deprivation has not occurred, regularized
+resampling with a large bandwidth will slow down convergence. Thus
+as mentioned previously, to prevent over smoothing, the particle
+filter algorithm used here waits until 
+the effective sample size drops below 25 for two measurement points
+in a row. This seemed to give good results, and avoids
 temporary drops in the $N_{eff}$ caused by outliers.

### 2010-07-05 doc/thesis/results.tex
work on particle filter.
almost done with methods
some changes to methods
some mild additions to methods and results

@@
 \chapter{Results}
+\section{Single-Voxel Simulation}
+The results
+
+\section{Single-Voxel Analysis}
+This section discusses the results when the particle filter was
+applied on a single voxel. The parameters are the same as
+those used later for entire image analysis; however, the results
+are more in-depth.
+
+The run-time for a single voxel depends on the several factors. First, the
+overall length of the signal being analyzed. For 1000 measurements it takes
+about 10 to 15 minutes. On the other hand, in real circumstances the
+length is only around 150 measurements. The number of  integration
+can certainly make a large difference, however dropping below 1000 (.001 seconds)
+is definitely not recommended.
+
+For a period I considered 1000 to be a
+fine number; however when generating simulated data I found that every once
+in a while 1000 was not enough. This is problematic in the actual particle
+filter since, given the large number of simultaneous integrations taking 
+place, its likely that a few particles will fail and be weighted at 0 because
+of this problem. Additionally, because the typical case where a failure would
+occur is at fast moving times/parameters the particles all tend to fail together.
+The result is particle deprivation - no particles with non-zero weights remain.
+The other possible outcome is that low time constant particles get pruned resulting
+in excessively smoothed estimates for the time series'. Its possible that a
+kind of stop-gap measure could be put into place; wherein particles that are
+about to be set to NaN are integrated again with finer grained steps. However
+many times the non-real results don't occur until several time steps after the 
+numbers get strange. So for instance, the timestep was too long, allowing 
+$f$ to go negative, resulting in extremely large values of $q$. There are many
+different ways where this sort of event can occur, and unfortunately sometimes
+there is no way to get back to before the state starting going out of control.
+
+Another crucial factor to run time is how long before the first re-sampling 
+occurs. Because the prior is represented initially with significantly more
+particles, if the model fits very well, or for some reason the effective
+number of particles stays high, resampling could take a long time to occur.
+When this happens the particles filter can take a factor of 10 longer to run.
+However, if the particle count isn't initially set high, there is a much larger
+chance of particle deprivation occurring. Since there is no real way to know
+how long it will take to resmaple the first time, there is little the 
+algorithm can do to fix this. On the other end of the specture, if the time
+series don't match the model at all, particle deprivation will occur extremely
+quickly, and even the recovery technique discussed in \autoref{sec:Resampling}
+won't help. The upshot of this is that the particle filter is able to 
+identify these sections very quickly, and thus not waste much time there.
+Of course, if a fat-tailed distribution is used for the weighting function,
+or the standard deviation of a Gaussian weighting function is very large,
+the particle filter will simply converge to meaningless values.
+
+\begin{figure}
+\caption{Particle Filter converging to values that make little sense,
+because the voxel did not correlate with the input in any known way}
+\end{figure}
+
+\section{Weighting Function Comparison}
+\label{sec:Results Weights}
+
 \section{Single Time-Series Simulation}
@@
 
+Final parameter distribution among active regions.
+Q-Q plots?
+
 \section{FMRI Data}

### 2010-07-05 doc/thesis/methods.tex
some mild additions to methods and results
more work....

@@
 \section{Resampling}
+\label{sec:Resampling}
 The algorithm for resampling is described in \autoref{sec:Particle Filter Resampling}.
@@
 %precision in the output maps.
-
-\section{Simulation}
-I performed two types of simulations. First, I simulated a single BOLD time-series based
-on a randomly selected set of model parameters. This process was relatively straight forward,
-given the state-space equations for the BOLD signal. After a "true" signal was generated,
-I then added a carrier level, since BOLD is typically measured as a \% difference from the
-base level. I tested the particle filter with two types of noise; independent and identically
-distributed gaussian noise and weiner drift. For both I tested with various standard deviations.
-
-Once this noisy simulated time series was generated, the exact same particle filter algorithm
-that would later be run on full sized images, was run on this single voxel image. I ran
-a series of tests to determine the convergence rate of the particle filter, the number
-of particles that were required, how weighting functions compared, how different de-trending
-methods compared with each other and, finally the variance of the result. By running the exact
-time-series with different noise realizations, it was possible to determine the model variance.
-As the reader may know, the error of an estimator may be calculated as:
-\begin{equation}
-MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
-\end{equation}
-The variance is an expression of how much the result would change for different noise realizations,
-whereas the bias is an expression of how well the model matches the true underlying model. In
-this case, because the same model is being used in the particle filter and underlying simulation,
-the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
-underlying state space equation, there will be some amount of bias error, but assuming that the
-noise is similar to the noise used in these tests, the model variance will actually be about the
-same. Thus calculating the model variance is extremely helpful in calculating how well determined
-our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
-thousands present in a real image, makes it easier to 
-compare the output with the ground truth, with various parameters.
-
-Second I used a modified version of the FSL tool 
-POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
-by taking an existing activation map and assigning discrete parameter sets to each region.
-The result was a four dimensional (length x width
-x height x parameter) image with spatially varying parameters. Possum was then modified
-to take a parameter map and generate activation levels depending on the parameters at that
-point. The patch for POSSUM will be made available. As an unfortunate side effect of 
-not using Possums' original activation scale, I manually added 750 to the total level of
-simulated Possum images. This is because the BOLD \% difference levels were in the range
-of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
-this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
-For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
-into a parameter map. This parameter map may then be compared to the map used to generate the 
-simulated data; additionally a new simulation using the calculated parameters may also be 
-generated to test the difference in activation levels between the real parameters and the
-estimated ones. Since it is clear that the parameters are not fully orthogonal 
-(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
-but have different parameters. This way, an absolute 
-quantitative difference between the two parameter sets may be found.
-
-\section{Real Data}
-Finally, we also performed inference based on real FMRI data. The scanner we used...
-%... more specifics...
-
-Before performing tests on a full image, I tested the results of the particle filter
-on regions deemed active and non-active by statistical parametric mapping.

### 2010-07-05 doc/thesis/particlefilters.tex
almost done with methods
more work....

@@
 Because the distribution of $\nu_y$ in \autoref{eq:measass} is unknown,
-it is necessary to choose a distribution for this. This distribution,
+it is necessary to choose a distribution for this. This distribution
 is extremely important because $\nu_y \sim P(y_k | x(T))$, which is used
@@
 Weighting Function}.
+
+\begin{figure}
+\includegraphics[width=16cm]{images/particle_filter}
+\caption{Particle Filter progression, note that the initial support is flat; the particles
+are equally spaced between -10 and 10}
+\end{figure}
 

### 2010-07-05 doc/thesis/results.tex
more work....

@@
 \chapter{Results}
+
+\section{Simulation}
+I performed two types of simulations. First, I simulated a single BOLD time-series based
+on a randomly selected set of model parameters. This process was relatively straight forward,
+given the state-space equations for the BOLD signal. After a "true" signal was generated,
+I then added a carrier level, since BOLD is typically measured as a \% difference from the
+base level. I tested the particle filter with two types of noise; independent and identically
+distributed gaussian noise and weiner drift. For both I tested with various standard deviations.
+
+Once this noisy simulated time series was generated, the exact same particle filter algorithm
+that would later be run on full sized images, was run on this single voxel image. I ran
+a series of tests to determine the convergence rate of the particle filter, the number
+of particles that were required, how weighting functions compared, how different de-trending
+methods compared with each other and, finally the variance of the result. By running the exact
+time-series with different noise realizations, it was possible to determine the model variance.
+As the reader may know, the error of an estimator may be calculated as:
+\begin{equation}
+MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
+\end{equation}
+The variance is an expression of how much the result would change for different noise realizations,
+whereas the bias is an expression of how well the model matches the true underlying model. In
+this case, because the same model is being used in the particle filter and underlying simulation,
+the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
+underlying state space equation, there will be some amount of bias error, but assuming that the
+noise is similar to the noise used in these tests, the model variance will actually be about the
+same. Thus calculating the model variance is extremely helpful in calculating how well determined
+our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
+thousands present in a real image, makes it easier to 
+compare the output with the ground truth, with various parameters.
+
+Second I used a modified version of the FSL tool 
+POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
+by taking an existing activation map and assigning discrete parameter sets to each region.
+The result was a four dimensional (length x width
+x height x parameter) image with spatially varying parameters. Possum was then modified
+to take a parameter map and generate activation levels depending on the parameters at that
+point. The patch for POSSUM will be made available. As an unfortunate side effect of 
+not using Possums' original activation scale, I manually added 750 to the total level of
+simulated Possum images. This is because the BOLD \% difference levels were in the range
+of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
+this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
+For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
+into a parameter map. This parameter map may then be compared to the map used to generate the 
+simulated data; additionally a new simulation using the calculated parameters may also be 
+generated to test the difference in activation levels between the real parameters and the
+estimated ones. Since it is clear that the parameters are not fully orthogonal 
+(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
+but have different parameters. This way, an absolute 
+quantitative difference between the two parameter sets may be found.
+
+\section{Real Data}
+Finally, we also performed inference based on real FMRI data. The scanner we used...
+%... more specifics...
+
+Before performing tests on a full image, I tested the results of the particle filter
+on regions deemed active and non-active by statistical parametric mapping. 
 \section{Single-Voxel Simulation}

### 2010-07-05 doc/thesis/examples/particle.py

@@ -43,6 +43,7 @@ def smooth(pos, weights):
 #time 1
 P.figure(1)
 P.subplot(411)
+P.title("Update 1")
 xvals = scipy.linspace(-10,10,num=1500)
 measmean = 1
 meas = normalize([norm.pdf(i, measmean, 2) for i in xvals])
@@ -58,6 +59,7 @@ P.legend(["Measurement", "State Predict", "Updated State", "Support"], loc=2)
 
 #time 2
 P.subplot(412)
+P.title("Update 2")
 #meas = normalize([norm.pdf(i, measmean, 2) for i in xvals])
 P.plot([val for val in xvals], meas)
 
@@ -71,7 +73,7 @@ P.legend(["Measurement", "State Predict", "Updated State", "Support"], loc=2)
 
 #time 3
 P.subplot(413)
-P.title("Resampled Support")
+P.title("Resample Support")
 P.plot([val for val in xvals], state)
 part_points = [sample(xvals, state) for val in xvals]
 state = [1./len(xvals)  for val in xvals]
@@ -81,6 +83,7 @@ P.legend(["Posterior Distribution", "New Support"], loc=2)
 
 #time 4
 P.subplot(414)
+P.title("Update 3")
 tmpx, tmpy = smooth(part_points, state)
 tmpy = normalize(tmpy)
 meas = normalize([norm.pdf(i, measmean-2, 2) for i in xvals])

### 2010-07-10 doc/thesis/methods.tex
added quite a few plots. Should be able to write results for single voxel simulations.

@@
 \section{Prior Distribution}
+\label{sec:PriorDist}
 For the BOLD model described in \autoref{sec:BOLD Physiology}, several
@@
 The one exception is \cite{Johnston2008}, which came to an extremely different
-set of distributions. For a particle filter, the choice of a prior is
+distributions. For a particle filter, the choice of a prior is
 the single most important design choice. Consequently for this work,
@@
 
-Therefore, because \cite{Friston2000} represents the current best
-knowledge about the BOLD parameters, it forms the basis
-for the prior distributions in this paper.
+There is an interesting anomaly with the priors found in virtually all
+the works that characterized the parameters, except \cite{Johnston2008}.
+The BOLD signal is universally recognized to be around $2-3\%$, maybe
+reaching $5\%$ in extreme activation. Yet using the mean priors
+from \cite{Friston2000}, the signal response for a $.1$ second
+impulse only reaches maybe half a percent, as \autoref{fig:MeanResponse}
+shows.
+
+\begin{figure}
+\centering
+\label{fig:MeanResponse}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response}
+\caption{Response to $.1s$ impulses with the mean parameters from \cite{Friston2000}}
+\end{figure}
+
+While this could be the result of a stimulus
+being too short to lead to strong activation, a similar stimulus
+scheme in real data showed a significantly larger response than 
+half a percent as well. In fact, after applying de-trending,
+converting the image to percent-difference, and removing 
+outliers ($ BOLD > 10\% \text{ or } BOLD < -10\%$) the total variance
+across all \emph{active} voxels was still around .02, indicating
+that in active voxels a signal peaking below .005 seems unlikely. 
+Of course, if more restriction were placed on the outliers, its possible
+this standard deviations could be brought down. 
+The parameter estimates by \cite{Johnston2008} are even more 
+confusing, with peaks of well below $.1\%$ (\autoref{fig:MeanResponseJ}).
+
+\begin{figure}
+\centering
+\label{fig:MeanResponse}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response_johnston}
+\caption{Response to $.1s$ impulses with the mean parameters from \cite{Johnston2008}}
+\end{figure}
+
+Its likely that these differences are due to some difference in preprocessing,
+although in \cite{Deneux2006} the signals were found to be peaking around
+$1\%$, unlike \cite{Friston2000} which shows signals peaking at up to
+$3\%$ or $4\%$. In my own tests, it seemed necessary for $\epsilon$ to
+reach well over $1.5$ and $V_0$ to reach more than $.4$ to reach these
+peaks; of course other methods may be equally able. 
+Therefore, to account for these discrepancies, somewhat broader
+distributions are used than the numbers used in \cite{Friston2000}
+(which are widely used, \cite{Hu2009}). The 
+priors used in the particle filter may be found in \autoref{tab:Prior}.
 
@@
 \hline
-$\tau_0$ & Gamma & .98 & .5  \\
+$\tau_0$ & Gamma & .98 & .25 \\
 $\alpha$ & Gamma & .33 & .045\\
 $E_0$    & Gamma & .34 & .03  \\
-$V_0$    & Gamma & .03 & .006 \\
-$\tau_s$ & Gamma & 1.54  & .5 \\
-$\tau_f$ & Gamma & 2.46  & .5 \\
-$\epsilon$ & Gamma & .64 & .4 \\
+$V_0$    & Gamma & .04 & .03 \\
+$\tau_s$ & Gamma & 1.54  & .25\\
+$\tau_f$ & Gamma & 2.46  & .25\\
+$\epsilon$ & Gamma & .7  & .6 \\
 \hline

### 2010-07-10 doc/thesis/results.tex
added quite a few plots. Should be able to write results for single voxel simulations.

@@
 I performed two types of simulations. First, I simulated a single BOLD time-series based
-on a randomly selected set of model parameters. This process was relatively straight forward,
+on a randomly selected set of model parameters. Second I used a modified version of the
+FSL tool possum to generate a noisy FMRI image.
+
+\subsection{Single Timeseries}
+This process was relatively straight forward,
 given the state-space equations for the BOLD signal. After a "true" signal was generated,
-I then added a carrier level, since BOLD is typically measured as a \% difference from the
-base level. I tested the particle filter with two types of noise; independent and identically
-distributed gaussian noise and weiner drift. For both I tested with various standard deviations.
+I then added a identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
+process with Gaussian I.I.D. steps. Finally a carrier level was added, since BOLD is typically
+measured as a \% difference from the
+base level. Adding a carrier level meant that the exact same algorithm which was used for 
+real data could be used for the simulated data.
 
-Once this noisy simulated time series was generated, the exact same particle filter algorithm
-that would later be run on full sized images, was run on this single voxel image. I ran
+Once this noisy simulated time series was generated, the particle filter algorithm
+ was run on this single timeseries image. In order to find the best setting for the
+particle filter I ran quite a few tests. Here I will include two sets of tests 
+to determine the power of the particle filter in modeling. The first test is
+low noise $\{\sigma_y = .001, \sigma_x = .0005\}$, while the second set of
+tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
+measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
+parameters ($\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8$).
+The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
+thus the particle filter was not centered over the correct values.
+
+\subsubsection{Low Noise Simulation}
+For the low noise case, the ten realization are shown in \autoref{fig:LowNoiseRealization}.
+\begin{figure}
+\label{fig:LowNoiseRealization}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_lownoise}
+\caption{Test Signals with low noise compared to the clean signal}
+\end{figure}
+
+The resulting fits compared to the \emph{true} BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
+The fits are good, but the noise is relatively low. The preprocessed signal compared
+to the true BOLD signal is shown in \autoref{fig:PreprocessedLowNoise}.
+
+\begin{figure}
+\label{fig:PreprocessedLowNoise}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_lownoise}
+\caption{A comparison of the preprocessed signals for the low noise case.}
+\end{figure}
+
+The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Preprocessing}.
+Obviously the spline struggles a little bit at the end, as the graph shows, although 
+overall the fit is pretty well. Finally, the set of fits to the preprocessed data, and in 
+essence to the "true" BOLD signal is shown in \autoref{fig:FitComparisonLowNoise}.
+
+\begin{figure}
+\label{fig:FitComparisonLowNoise}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_lownoise}
+\caption{A comparison of the fitted signals for the low noise case.}
+\end{figure}
+
+It becomes clear now the reason for the term particle filter. In several locations the output
+of the particle filter looks like a filtered version of the input. For instance toward the
+end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
+this point, the algorithm had converged sufficiently to know that that wasn't possible.
+A similar circumstance occurs at 100 seconds in. In general the fits are good, although because
+of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
+the peak a bit. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
+\hline 
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) &$\sqrt{MSE}$\\
+\hline 
+\rowcolor[gray]{.8}
+1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+\hline 
+\hline 
+1.22214 & 0.3449 & 0.33462 & 0.07138 & 1.60446 & 2.27529 & 1.59454 & 5.10189  &  0.00321067  & 0.00987647  \\
+1.37493 & 0.33183 & 0.36296 & 0.07327 & 1.64076 & 2.10299 & 1.57627 & 5.11867 &  0.00305475  & 0.00993227  \\
+1.16604 & 0.32205 & 0.34057 & 0.08217 & 1.64768 & 2.35351 & 1.24515 & 5.16722 &  0.00328932  & 0.00967961  \\
+1.23176 & 0.32707 & 0.34028 & 0.07961 & 1.62698 & 2.18519 & 1.30326 & 5.04394 &  0.00284719  & 0.00912005  \\
+1.1832 & 0.31787 & 0.34718 & 0.0821 & 1.54961 & 2.29115 & 1.27817 & 5.02396   &  0.00300634  & 0.00971258  \\
+1.1424 & 0.33395 & 0.34725 & 0.07366 & 1.62208 & 2.29084 & 1.4025 & 5.05531   &  0.00283287  & 0.00948483  \\
+1.30041 & 0.3596 & 0.35643 & 0.07679 & 1.56406 & 2.13234 & 1.60338 & 4.99681  &  0.00302802  & 0.01021885  \\
+1.24008 & 0.34601 & 0.33978 & 0.08903 & 1.64989 & 2.23655 & 1.29004 & 5.12651 &  0.00304378  & 0.01007964  \\
+1.1709 & 0.32739 & 0.34644 & 0.08255 & 1.53734 & 2.28262 & 1.37828 & 4.99087  &  0.00334488  & 0.01032886  \\
+1.18967 & 0.34344 & 0.33554 & 0.07976 & 1.53582 & 2.30746 & 1.42774 & 5.03295 &  0.00317542  & 0.01001503  \\
+1.184 & 0.34053 & 0.35017 & 0.08917 & 1.61025 & 2.27926 & 1.16448 & 5.07352   &  0.00288855  & 0.00950536  \\
+\hline                                                                           
+1.21868 & 0.33588 & 0.34557 & 0.07995 & 1.59899 & 2.24884 & 1.38762 & 5.06651 & 0.00306562     & 0.00981396 \\
+\hline 
+\end{tabular}
+\caption{Estimated Parameters on 10 different runs with low noise. First row is the true parameters,
+last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
+residuals. Essentially this is the is the MSE between the estimated signal and the noisy signal which 
+was available to the particle filter. Square root of MSE is the actual \emph{error}, with the true signal,
+which of course was not available to the particle filter.}
+\label{tab:LowNoiseResults} 
+\end{table}
+
+There are a few things worth noting here. First the time constants vary greatly across
+runs with different noise realizations, yet the sum of the individual time constants
+, $\tau_f$, $\tau_s$ and $\tau_0$ seem to be more consistent. In general the 
+time constants are falling short of the true time constant. This could be a limitation
+based on the prior distribution (which notably has mean values below the values used
+in the simulation) or it could be caused by the delayed benefit of having a 
+correct time constant. Even if one particle has a better time constant than another, if
+the difference isn't severe, by the time this makes a difference in the weight, the 
+particle with the better $\tau$ will have been weighted based on other parameters several
+times. Another interesting result in the huge variation in the levels of $V_0$. In general,
+with this admittedly small amount of noise, it would appear that the relation between a set
+of parameters/stimuli with the output is not injective. In other words, it would appear
+that a timeseries is not unique to a set of parameters. This is good justification that 
+simultaneous blood volume or tagged flow calculations with the conventional FMRI 
+could definitely benefit the model.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline 
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline 
+\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
+$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
+\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
+$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
+\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
+$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
+\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+\hline 
+\end{tabular}
+\caption{Typical Covariance matrix of the parameters at the end of a run.}
+\label{tab:CovSim} 
+\end{table}
+
+\begin{figure}[H]
+\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
+{\includegraphics[trim=7cm 4cm 7cm 4cm, width=15cm]{images/converge_lownoise1}}\\
+
+\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
+{\includegraphics[trim=7cm 4cm 7cm 1cm, width=15cm]{images/converge_lownoise2}}\\
+\end{figure}
+
+\begin{figure}
+\subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
+{\includegraphics[trim=7cm 2cm 7cm 2cm, width=15cm]{images/converge_lownoise3}}
+\label{fig:LowNoiseHist}
+\end{figure}
+The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the $\tau$ parameters are interchangeable
+as far as the BOLD signal goes. Notice the covariance of $\tau_f$ and $\tau_0$ is $-0.019$ whereas
+the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly they are
+related. The convergence properties of the first run in \autoref{tab:LowNoiseResults} may be
+viewed in \autoref{fig:LowNoiseHist}.
+
+\subsubsection{High Noise Simulation}
+For the high noise simulation, the exact same procedure was performed again, but with 
+measurement error and drift standard
+deviations an order of magnitude higher. The noisy signals versus the base signal are shown
+in \autoref{fig:HighNoiseRealization} and the preprocessed signals which the particle
+filter attempt to fit are in \autoref{fig:PreprocessedHighNoise}.
+\begin{figure}[H]
+\label{fig:LowNoiseRealization}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_highnoise}
+\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
+\end{figure}
+\begin{figure}[H]
+\label{fig:PreprocessedHighNoise}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_highnoise}
+\caption{A comparison of the preprocessed signals for the high noise case.}
+\end{figure}
+
+The results of the particle filter, for each of the ten runs may be seen in 
+\autoref{fig:FitComparisonHighNoise}. Clearly the preprocessing led the algorithm
+to somewhat higher activation levels, and it would appear that the subtleties of
+different time constants are also lost for most of the runs, although a few seem
+to manage quite accurate results.
+
+\begin{figure}[H]
+\label{fig:FitComparisonHighNoise}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise}
+\caption{A comparison of the fitted signals for the high noise case.}
+\end{figure}
+\begin{figure}[H]
+\label{fig:NoiseComparisonJustTwo}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/highnoise_56_noise}
+\caption{Two particular preprocessed noise realizations for the high noise case.}
+\end{figure}
+\begin{figure}[H]
+\label{fig:FitComparisonHighNoiseJust2}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise_just2}
+\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
+\end{figure}
+
+It is interesting to consider how the preprocessing and noise may effect the
+parameters of the fitting model. For instance run 1 in \autoref{fig:NoiseComparisonJustTwo}
+certainly seems to be more biased toward higher peaks than run 2. There also appears
+to be more drift than the 20 points per knot could fit, which explains the 
+prolonged increase at 170 seconds. Despite run 1's bias toward higher peaks, for
+some reason the particle filter was able to get a much better match for the
+modeled post-stimulus undershoot (which is likely significantly shorter than it
+should be). Ultimately the results are pretty good. \autoref{tab:SimMSE} shows
+the mean squared error for all two runs, and highlights the two runs analyzed here and 
+in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
+
+\begin{figure}[H]
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
+{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run5_1}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run5_2}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run5_3}}
+\label{fig:ConvergenceRuns1}
+\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\end{figure}
+
+There are a number of interesting convergence properties of the
+particle filter when more noise is present, as both \autoref{fig:ConvergenceRuns1} and
+\autoref{fig:ConvergenceRuns2} show. Obviously the particle filter seems to converge
+significantly faster, as points tend to be further out on the weighting function. This
+also causes significantly more resampling which is the explanation for the seeming
+jumps in resolution that occur from time to time.
+
+\begin{figure}[H]
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
+{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run6_1}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run6_2}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run6_3}}
+\label{fig:ConvergenceRuns2}
+\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\end{figure}
+
+The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
+\hline 
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) & $\sqrt{MSE}$\\
+\hline 
+\rowcolor[gray]{.8}
+1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+\hline 
+\hline 
+1.19002 & 0.23495 & 0.42228 & 0.128 & 1.01468 & 2.47795 & 1.11677 & 4.68265   &0.0140648  &0.01572919   \\
+0.9721 & 0.21902 & 0.30505 & 0.061 & 0.57801 & 1.99596 & 3.46135 & 3.54608    &0.0137314  &0.01377685   \\
+1.57947 & 0.14153 & 0.33798 & 0.1079 & 0.5843 & 2.12475 & 1.78343 & 4.28852   &0.01275372 &0.01577449   \\
+1.10937 & 0.2374 & 0.53491 & 0.0351 & 1.21862 & 3.07365 & 2.35039 & 5.40164   &0.0167258  &0.0115433    \\
+1.10712 & 0.27535 & 0.33651 & 0.03161 & 1.50567 & 2.65181 & 4.19099 & 5.2646  &0.01369793 &0.01221607   \\
+0.58026 & 0.47931 & 0.41354 & 0.11894 & 0.97563 & 3.69018 & 1.00084 & 5.24607 &0.01149511 &0.01315637   \\
+\rowcolor[rgb]{.9,.5,.5}
+1.29515 & 0.25957 & 0.27559 & 0.25952 & 1.70265 & 2.8458 & 0.66172 & 5.84361  &0.01555035 &0.01789781   \\
+\rowcolor[rgb]{.5,.5,.9}
+1.5185 & 0.21987 & 0.28348 & 0.07417 & 0.88822 & 3.07706 & 1.73929 & 5.48378  &0.01205351 &0.01246202   \\
+0.68736 & 0.3283 & 0.39786 & 0.15614 & 1.07781 & 3.1158 & 0.66432 & 4.88097   &0.01510364 &0.01257713   \\
+1.01703 & 0.28497 & 0.34741 & 0.05672 & 1.58774 & 2.65157 & 2.28519 & 5.25634 &0.012493   &0.01343459   \\
+0.99247 & 0.29795 & 0.32207 & 0.20943 & 0.42757 & 2.21081 & 1.01674 & 3.63085 &0.01216522 &0.01505545   \\
+\hline                                                                          
+1.09535 & 0.27075 & 0.36152 & 0.11259 & 1.05099 & 2.71958 & 1.84282 & 4.86592 &0.01362132  &0.01396575\\
+\hline 
+\end{tabular}
+\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
+Note also that the blue row is Run 1 and the red row is Run 2, as used named this section}
+\label{tab:HighNoiseResults} 
+\end{table}
+
 a series of tests to determine the convergence rate of the particle filter, the number
@@
 
-Another crucial factor to run time is how long before the first re-sampling 
+Another crucial factor for run time is how long before the first re-sampling 
 occurs. Because the prior is represented initially with significantly more
@@
 how long it will take to resmaple the first time, there is little the 
-algorithm can do to fix this. On the other end of the specture, if the time
-series don't match the model at all, particle deprivation will occur extremely
-quickly, and even the recovery technique discussed in \autoref{sec:Resampling}
-won't help. The upshot of this is that the particle filter is able to 
+algorithm can do to fix this (except perhaps forcing resampling after
+some period of time). On the other end of the spectrum, if the time
+series doesn't match the model at all, particle deprivation will occur extremely
+quickly.  The upshot of this is that the particle filter is able to 
 identify these sections very quickly, and thus not waste much time there.
+The difficulty though, is the regions in between the perfect fit and the
+awful fit. If particle deprivation does occur, did it occur randomly to an
+activated region or did it occur inevitably because the region doesn't fit.
+The reason for having so many initial particles is to give density to the
+distribution to make false negatives more rare. Perhaps the correct method
+is to still set the initial particles very high, but if for a long time no
+resampling occurs, halve the standard deviation of the measurement error.
+Thus, if the weighting function is not discriminating enough, force it 
+to be more picky about the results. Of course this could lead to particle
+deprivation as well if the standard deviation is brought down too quickly.
 Of course, if a fat-tailed distribution is used for the weighting function,
-or the standard deviation of a Gaussian weighting function is very large,
-the particle filter will simply converge to meaningless values.
+or the standard deviation of a Gaussian weighting function is sufficiently large,
+the particle filter will simply converge to meaningless values. The question
+of whether
 
 \begin{figure}
+\label{fig:badfit}
+\includegraphics[trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
 \caption{Particle Filter converging to values that make little sense,
@@
 \end{figure}
+
+\begin{figure}
+\label{fig:highhard}
+\includegraphics[trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
+\caption{Example of a timeseries that laplace performs better on than
+gaussian. The large spike at the beginning can quickly eliminate all particles}
+
+\end{figure}
+
+For this reason, it is important to threshold based on the MSE, and the
+impulse response. Naturally the MSE is intended to catch very voxels with a
+very poor fit,
+but like \autoref{fig:badfit}, the algorithm may have found a line
+down the middle that gives a decent mean squared error. In this case
+the excessive smoothing and low actual activation level will easily
+be distinguished by finding the peak response to single short pulse.
+
+The choice of a prior, as discussed previously, is extremely important. While a
+prior may have the potential to give good results, being a monte-carlo algorithm
+there is the possibility for inconsistencies. Thus, increasing the variance
+of the time-constants may allow additional flexibility, it will also cause
+additional model variance. Case in point, the exact same algorithm run twice
+with standard deviations of $.35, .35, .35$ for the time constants resulted in two
+very different fits, see \autoref{fig:param1_var}.
+
+\begin{figure}
+\label{fig:badfit_param1}
+\subfigure[]{\includegraphics{images/badfit_param1}}
+\subfigure[]{\includegraphics{images/goodfit_param1}}
+\caption{The same priors gave rise to both fits. Params: ($\tau_0 \alpha E_0 V_0 \tau_f \tau_s \epsilon$)
+for the good fit: $(1.40 .32 0.34 0.04  0.96 2.76)$, and for the poor fit:
+$(2.28 0.32 0.33 0.017 0.67 2.88 1.36)$.}
+\end{figure}
+
+For this reason, I actually lowered the standard deviattions of the time
+constants to prevent over-smoothing. This resulted in the more consistent, but
+slightly worse results seen in \autoref{fig:param2_var}
+
+\begin{figure}
+\label{fig:param2_var}
+\subfigure[]{\includegraphics{images/param2a}}
+\subfigure[]{\includegraphics{images/param2b}}
+\caption{A poor fit, using the same parameters as }
+\end{figure}
+
+\section{Particle Collapse Recovery}
+This heuristic was initially intended to deal with cases where the particle
+filter prematurely converged; and to reverse this effect. In reality, however
+it did not perform well, and so it was not used in full-brain analysis. Ultimately
+it did not prevent collapse of the distribution, and it often caused dramatic
+increased in the variance of parameter estimates. It also pronged the analysis
+time of inactive time series. Essentially the effect was exactly opposite of what
+one would hope.
 

### 2010-07-11 doc/thesis/results.tex
simulated single voxel analysis done except for a little more discussion

@@
-\chapter{Results}
+\chapter{Simulation}
 
-\section{Simulation}
+\section{Single Timeseries}
 I performed two types of simulations. First, I simulated a single BOLD time-series based
@@
 
-\subsection{Single Timeseries}
 This process was relatively straight forward,
@@
 
-\subsubsection{Low Noise Simulation}
+\subsection{Low Noise}
 For the low noise case, the ten realization are shown in \autoref{fig:LowNoiseRealization}.
@@
 \end{tabular}
-\caption{Estimated Parameters on 10 different runs with low noise. First row is the true parameters,
+\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
 last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
@@
 
-\subsubsection{High Noise Simulation}
+\subsection{High Noise}
+\label{sec:HighNoise}
 For the high noise simulation, the exact same procedure was performed again, but with 
@@
 \label{fig:ConvergenceRuns1}
-\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
 \end{figure}
@@
 \caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
-Note also that the blue row is Run 1 and the red row is Run 2, as used named this section}
+Note also that the red row is Run 1 and the blue row is Run 2, as used named this section}
 \label{tab:HighNoiseResults} 
@@
 
-a series of tests to determine the convergence rate of the particle filter, the number
-of particles that were required, how weighting functions compared, how different de-trending
-methods compared with each other and, finally the variance of the result. By running the exact
-time-series with different noise realizations, it was possible to determine the model variance.
-As the reader may know, the error of an estimator may be calculated as:
-\begin{equation}
-MSE(\Theta) = Var(\Theta) + Bias(\Theta)^2
-\end{equation}
-The variance is an expression of how much the result would change for different noise realizations,
-whereas the bias is an expression of how well the model matches the true underlying model. In
-this case, because the same model is being used in the particle filter and underlying simulation,
-the bias is actually zero. Obviously when this is calculated using \emph{real} data with an unknown
-underlying state space equation, there will be some amount of bias error, but assuming that the
-noise is similar to the noise used in these tests, the model variance will actually be about the
-same. Thus calculating the model variance is extremely helpful in calculating how well determined
-our model is, and how consistent it will be for real data. A single timeseries, as opposed to the
-thousands present in a real image, makes it easier to 
-compare the output with the ground truth, with various parameters.
+\subsection{Pure-Noise, low magnitude}
+The two final single-voxel test, forces the particle filter to attempt to "learn" a noise-only
+timeseries. The noise used will be the same as that from the \autoref{sec:HighNoise},
+$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
+stimulus/input will be set to 0 the entire time. For the particle filter, on the other hand,
+the stimulus will be left as in the previous two sections. The questions is if the particle
+filter is able to appropriately differentiate from a signal with noise. The noise timeseries
+are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:NoiseOnlyPrep}.
+\begin{figure}[H]
+\label{fig:NoiseOnly}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
+\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
+\end{figure}
+\begin{figure}[H]
+\label{fig:PreprocessedNoiseOnly}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
+\caption{A comparison of the preprocessed signals for the signal-free case.}
+\end{figure}
 
-Second I used a modified version of the FSL tool 
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
+\hline 
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSE}$   \\
+\hline 
+1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
+ 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
+ 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
+ 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
+ 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
+ 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
+ 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
+ 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
+ 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
+ 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
+ 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
+\hline                                                                
+1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
+\hline 
+\end{tabular}
+\caption{Estimated Parameters on 10 different runs with low noise. }
+\label{tab:NoiseOnlyResults} 
+\end{table}
+
+The line fits are shown in \autoref{fig:fits_noiseonly}. Now the
+low MSE and reasonableness of the parameters is cause for concern. Additionally, the precision
+with which the convergence occurred is enviable, compared to the actual signals.
+However, the impulse response, which has less importance in nonlinear systems, could
+be usefull for separating out good voxels from bad. Note that in \autoref{fig:PreprocessedNoiseOnly}
+peaks at well below $.005$, the standard deviation of the weighting function. Thus, the reason
+the fits are in such agreement is that no convergence occurred at all, as 
+further shown by \autoref{JustNoiseConvergence}. Thus one more test is necessary on
+a signal with enough noise to actually be considered a signal. But first,
+there is still the matter of identifying this noise-only voxel. Clearly MSE and
+the parameters themselves don't have the information to differentiate in this case.
+Thus the answer falls on either the final estimated variance (which will be high
+because no particles strayed far from the center) or the impulse response.
+I prefer the impulse response because it doesn't depend on something that 
+may be less than dependable, namelay the rate of convergence.
+
+\begin{figure}[H]
+\label{fig:fits_noiseonly}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
+\caption{Fits to the non-active, low noise signal. Note that the line is thick because all
+the fits overlap. This is all 1 fitted lines.}
+\end{figure}
+
+\begin{figure}[H]
+\label{fig:justnoise_fit_0}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
+\caption{Fit from a single particle filter run, with the noise input. }
+\end{figure} %uses allnoise/ALLNOISE-0-w0
+
+\begin{figure}[H]
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
+{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
+\label{fig:JustNoiseConvergence}
+\caption{Converging histogram for parameters when the signal consists purely of low level noise. 
+Same run as \autoref{fig:justnoise_fit_0}}
+\end{figure}
+
+To give a frame of reference to the impulse response level, the impulse response of every
+parameter estimated, includeing ones from \autoref{sec:BigNoise}, are shown in \autoref{tab:ImpulseResponses}.
+
+\subsection{Pure Noise, High Magnitude}
+\label{sec:PureNoiseHighMag}
+
+\begin{figure}[H]
+\label{fig:fits_noiseonly_high}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
+\caption{Fits to the non-active, high noise signal. Note that the line is thick because all
+the fits overlap. This is all 11 fitted lines.}
+\end{figure}
+
+\begin{figure}[H]
+\label{fig:justbignoise_fit_0}
+\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
+\caption{Fit from a single particle filter run, with the noise input. }
+\end{figure} %uses allnoise/ALLNOISE-0-w0
+
+\begin{figure}[H]
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
+{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
+\end{figure}
+\begin{figure}[H]
+\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
+\label{fig:JustNoiseConvergenceLarge}
+\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
+the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
+\end{figure}
+
+\subsection{Single Voxel Review}
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c | c | c | c |}
+\hline 
+& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
+\hline
+Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
+\hline 
+\hline       
+1 &    0.0115482 &       0.015527 &        0.00138579 &      0.00843435   \\
+2 &    0.0116813 &       0.0126847 &       0.00152434 &      0.00495585   \\
+3 &    0.011314 &        0.0123151 &       0.0014101 &       0.00667397  \\
+4 &    0.0107892 &       0.0114702 &       0.00119941 &      0.0038459    \\
+5 &    0.0112809 &       0.0143503 &       0.00145239 &      0.00811513   \\
+6 &    0.0112021 &       0.0113265 &       0.00137792 &      0.00662327   \\
+7 &    0.011943 &        0.0192599 &       0.00141883 &      0.00734546   \\
+8 &    0.0118222 &       0.010144 &        0.00162352 &      0.00817847   \\
+9 &    0.0120212 &       0.0122466 &       0.00146133 &      0.00843033   \\
+10 &   0.0115679 &       0.0156513 &       0.00165983 &      0.00819587   \\
+11 &   0.0110983 &       0.0103732 &       0.001415 &        0.00886845  \\
+\hline
+Mean& 0.0114789417183 &  0.0132135264575 & 0.00144804221451 &0.00724245910533 \\
+\hline
+\end{tabular}
+\caption{Estimated Impulse response on 11 different for each of the cases covered in this chapter.}
+\label{tab:SignleVoxelActivationComparison} 
+\end{table}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c | c | c | c |}
+\hline 
+& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
+\hline
+Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
+\hline 
+\hline       
+1 &   0.00321067 & 0.0140648  &   0.00166666 &  0.01576746 \\
+2 &   0.00305475 & 0.0137314  &   0.00158551 &  0.0155104 \\
+3 &   0.00328932 & 0.01275372  &  0.00165055 &  0.01395712 \\
+4 &   0.00284719 & 0.0167258  &   0.00150604 &  0.01381387 \\
+5 &   0.00300634 & 0.01369793  &  0.00151513 &  0.01435028 \\
+6 &   0.00283287 & 0.01149511  &  0.00155645 &  0.01350408 \\
+7 &   0.00302802 & 0.01555035  &  0.00158526 &  0.01300628 \\
+8 &   0.00304378 & 0.01205351  &  0.00164299 &  0.01544561 \\
+9 &   0.00334488 & 0.01510364  &  0.0016793 &   0.01484681 \\
+10 &  0.00317542 & 0.012493  &    0.00182622 &  0.01312527 \\
+11 &  0.00288855 & 0.01216522  &  0.00195009 &  0.01445691 \\
+\hline
+Mean& 0.00306562 & 0.01362132  &  0.00165129 &  0.01434401 \\
+\hline
+\end{tabular}
+\caption{MSE with response to the preprocessed, noisy signal for 11 runs for each of the cases covered in this chapter.}
+\label{tab:SingleVoxelMSEComparison} 
+\end{table}
+
+Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, it is
+clear that distinguishing between these cases is a difficult prospect. Although
+the impulse response, which is determined by the peak response to a very short input,
+is is slightly different, there is no reason to believe that a noise distribution
+could not get a higher impulse response. The $\sqrt{MSE}$ is also not going to be
+able to distinguish between the high-noise case and the no signal case. Therefore, there
+is a noise ceiling above which it will be impossible to determine model validity. The fact
+that the particle filter is able to get close to the "true" signal is sadly irrelevant.
+Ultimately, a threshold well below those shown in the high noise case 
+will have to be placed on the error, if any confidence is going to be placed on the results.
+
+Each test took approximately 60 seconds on an 8 core Xeon (specifics?).
+
+\section{Multi-voxel Simulation}
+To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
 POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated

### 2010-07-15 doc/thesis/results.tex
added the heatmaps, and updated the charts

@@
 \centering
-\begin{tabular}{|c | c | c | c | c |}
+\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
 \hline 
-& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
+& \multicolumn{4}{|c|}{Signal} & \multicolumn{4}{|c|}{No Signal}\\
 \hline
-Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
+& \multicolumn{2}{|c|}{Low Noise} & \multicolumn{2}{|c|}{High Noise} 
+& \multicolumn{2}{|c|}{$\sigma_y = .001, \sigma_x = .0005$} 
+& \multicolumn{2}{|c|}{$\sigma_y = .01, \sigma_x = .005$}\\
+\hline 
+&$\sqrt{MSE}$ & $N\sqrt{MSE}$ &
+ $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
+ $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
+ $\sqrt{MSE}$ & $N\sqrt{MSE}$ \\
 \hline       
 \hline       
-1 &    0.0115482 &       0.015527 &        0.00138579 &      0.00843435   \\
-2 &    0.0116813 &       0.0126847 &       0.00152434 &      0.00495585   \\
-3 &    0.011314 &        0.0123151 &       0.0014101 &       0.00667397  \\
-4 &    0.0107892 &       0.0114702 &       0.00119941 &      0.0038459    \\
-5 &    0.0112809 &       0.0143503 &       0.00145239 &      0.00811513   \\
-6 &    0.0112021 &       0.0113265 &       0.00137792 &      0.00662327   \\
-7 &    0.011943 &        0.0192599 &       0.00141883 &      0.00734546   \\
-8 &    0.0118222 &       0.010144 &        0.00162352 &      0.00817847   \\
-9 &    0.0120212 &       0.0122466 &       0.00146133 &      0.00843033   \\
-10 &   0.0115679 &       0.0156513 &       0.00165983 &      0.00819587   \\
-11 &   0.0110983 &       0.0103732 &       0.001415 &        0.00886845  \\
+1 & 0.003211 & 0.478015  & 0.014065  & 1.038941    &0.001667 & 1.295007 & 0.015114 & 1.33641 \\
+2 & 0.003055 & 0.531771  & 0.013731  & 0.951648  &  0.001586 & 1.301750 & 0.015510 & 1.336667 \\
+3 & 0.003289 & 0.545798  & 0.012754 &  0.995392  &  0.001651 & 1.26287 &  0.013957 & 1.159567 \\
+4 & 0.002847 & 0.498237  & 0.016726  & 1.161286  &  0.001506 & 1.431959 & 0.013814 & 1.099876 \\
+5 & 0.003006 & 0.468048  & 0.013698 &  1.039724   & 0.001515 & 1.256645 & 0.014350 & 1.201072 \\
+6 & 0.002833 & 0.459001  & 0.011495 &  1.002143  &  0.001557 & 1.270797 & 0.013504 & 1.045886 \\
+7 & 0.003028 & 0.460962  & 0.015550 &  1.088472   & 0.001585 & 1.154406 & 0.013006 & 1.205429 \\
+8 & 0.003044 & 0.518376  & 0.012054 &  1.059620  &  0.001643 & 1.274564 & 0.015446 & 1.122502 \\
+9 & 0.003345 & 0.525303  & 0.015104 &  1.015703  &  0.001679 & 1.320244 & 0.014847 & 1.086366 \\
+10 & 0.003175 &0.492111  & 0.012493  & 1.189964  &  0.001826 & 1.344562 & 0.013125 & 1.221350 \\
+11 & 0.002889 &0.490919  & 0.012165 &  0.953676  &  0.001950 & 1.325223 & 0.014457 & 1.117367 \\
 \hline
-Mean& 0.0114789417183 &  0.0132135264575 & 0.00144804221451 &0.00724245910533 \\
+mean & 0.003066 & 0.497140  & 0.013621 &  1.04514   &  0.001651 & 1.294366 & 0.014285 & 1.175681 \\
 \hline
 \end{tabular}
-\caption{Estimated Impulse response on 11 different for each of the cases covered in this chapter.}
-\label{tab:SignleVoxelActivationComparison} 
-\end{table}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|c | c | c | c | c |}
-\hline 
-& \multicolumn{2}{|c|}{Signal} & \multicolumn{2}{|c|}{No Signal}\\
-\hline
-Run  & Low Noise & High Noise & $\sigma_y=.001, \sigma_x=.0005$ & $\sigma_y=.01, \sigma_x=.005$\\
-\hline 
-\hline       
-1 &   0.00321067 & 0.0140648  &   0.00166666 &  0.01576746 \\
-2 &   0.00305475 & 0.0137314  &   0.00158551 &  0.0155104 \\
-3 &   0.00328932 & 0.01275372  &  0.00165055 &  0.01395712 \\
-4 &   0.00284719 & 0.0167258  &   0.00150604 &  0.01381387 \\
-5 &   0.00300634 & 0.01369793  &  0.00151513 &  0.01435028 \\
-6 &   0.00283287 & 0.01149511  &  0.00155645 &  0.01350408 \\
-7 &   0.00302802 & 0.01555035  &  0.00158526 &  0.01300628 \\
-8 &   0.00304378 & 0.01205351  &  0.00164299 &  0.01544561 \\
-9 &   0.00334488 & 0.01510364  &  0.0016793 &   0.01484681 \\
-10 &  0.00317542 & 0.012493  &    0.00182622 &  0.01312527 \\
-11 &  0.00288855 & 0.01216522  &  0.00195009 &  0.01445691 \\
-\hline
-Mean& 0.00306562 & 0.01362132  &  0.00165129 &  0.01434401 \\
-\hline
-\end{tabular}
-\caption{MSE with response to the preprocessed, noisy signal for 11 runs for each of the cases covered in this chapter.}
-\label{tab:SingleVoxelMSEComparison} 
+\caption{$\sqrt{MSE}$ and the normalized $\sqrt{MSE}$, $\left(\sqrt{MSE}/(\text{Median Absolute Deviation})\right)$  
+for different configurations.} \label{tab:SignleVoxelActivationComparison} 
 \end{table}

### 2010-07-15 doc/thesis/results.tex
updated some discussion about the first low noise results

@@
 \chapter{Simulation}
+Two levels are discussed in this section. First the simulation of single time series.
+This is necessary to investigate the power of particle filters in 
+identifying the BOLD model in a noisy environment. If the particle filter is unable
+to get back to the "true" time-series in spite of noise then the particle filter
+will not be useful. Additionally, determination of reasonable parameters from the
+time-series is also desirable. The second simulation technique will be on a 
+large group of voxels. Although the current technique treats each voxel separately,
+the "Possum" FMRI simulation tool models noise more realistically, and provides additional
+incite into the applicability of particle filters in modeling the BOLD signal on a large
+scale.
 
 \section{Single Timeseries}
-I performed two types of simulations. First, I simulated a single BOLD time-series based
-on a randomly selected set of model parameters. Second I used a modified version of the
-FSL tool possum to generate a noisy FMRI image.
+Given the state-space equations for the BOLD signal, simulating a single time 
+series is relatively straight forward. After a "true" signal is generated,
+I then add identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
+process with Gaussian, I.I.D. steps. Finally a "carrier level" is added, since BOLD is typically
+measured as a \% difference from the baseline. Although technically the particle filter
+algorithm will immediately remove this, and calculate the \% difference, adding a carrier 
+level means that the exact same algorithm which will is used for 
+simulated data may be used for the real data.
 
-This process was relatively straight forward,
-given the state-space equations for the BOLD signal. After a "true" signal was generated,
-I then added a identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian I.I.D. steps. Finally a carrier level was added, since BOLD is typically
-measured as a \% difference from the
-base level. Adding a carrier level meant that the exact same algorithm which was used for 
-real data could be used for the simulated data.
+Once this noisy simulated time series has been generated, the particle filter algorithm
+ is then run on this single time series "image". Here I will include two sets of tests 
+to determine the power of the particle filter in modeling. The first set are 
+used to demonstrate the ability of the particle filter to find the most likely
+set of parameters/state variables over the course of the run. The second type
+of test will investigate the problem of false-positives. As the first
+round of tests will show, given that there is an underlying BOLD signal, 
+the particle filter is excellent at finding the most probable cause of 
+the observed signal; however, given enough noise it will also determine
+the most likely underlying signal; which may not be flat. Thus, it is 
+important to investigate methods of determining false positives using 
+average residuals.
 
-Once this noisy simulated time series was generated, the particle filter algorithm
- was run on this single timeseries image. In order to find the best setting for the
-particle filter I ran quite a few tests. Here I will include two sets of tests 
-to determine the power of the particle filter in modeling. The first test is
-low noise $\{\sigma_y = .001, \sigma_x = .0005\}$, while the second set of
-tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
-measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
-parameters ($\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8$).
-The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
-thus the particle filter was not centered over the correct values.
+\subsection{Signal with Low Noise}
+%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
+%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
+%parameters .
+%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
+%thus the particle filter was not centered over the correct values.
 
-\subsection{Low Noise}
-For the low noise case, the ten realization are shown in \autoref{fig:LowNoiseRealization}.
+To begin the single voxel simulation; I generated a signal using the following parameters:
+$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
+These same parameters will used again in the high noise case. The noise realizations were
+then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
+$\sigma_x$ of $.0005$. The measurement noise as well as the steps of the drift
+were taken to be Gaussian
+For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
 \begin{figure}
@@
 \end{figure}
-
 The resulting fits compared to the \emph{true} BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
-The fits are good, but the noise is relatively low. The preprocessed signal compared
-to the true BOLD signal is shown in \autoref{fig:PreprocessedLowNoise}.
-
+While the error between the true signal and the estimated signal are low, the noise
+is also relatively low. The actual signal that was delivered into the particle filter
+is the result of some preprocessing to remove drift, as described in 
+\autoref{sec:Preprocessing}. The preprocessed signal is compared
+to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
 \begin{figure}
@@
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_lownoise}
-\caption{A comparison of the preprocessed signals for the low noise case.}
+\caption{A comparison of the preprocessed signals for the low noise case. This is the
+noisy input to the actual particle filter algorithm.}
 \end{figure}
-
 The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Preprocessing}.
-Obviously the spline struggles a little bit at the end, as the graph shows, although 
-overall the fit is pretty well. Finally, the set of fits to the preprocessed data, and in 
-essence to the "true" BOLD signal is shown in \autoref{fig:FitComparisonLowNoise}.
-
+Obviously the spline struggles to match the signal near the end, as the graph shows, although 
+overall the preprocessing seems to have been successful at removing trends. Finally, the 
+set of fits to the preprocessed data, and in 
+the true BOLD signal is shown in \autoref{fig:FitComparisonLowNoise}.
 \begin{figure}
@@
 \end{figure}
-
-It becomes clear now the reason for the term particle filter. In several locations the output
+The filtering property of the particle filter is clear from the results here.  In several locations the output
 of the particle filter looks like a filtered version of the input. For instance toward the
 end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
-this point, the algorithm had converged sufficiently to know that that wasn't possible.
-A similar circumstance occurs at 100 seconds in. In general the fits are good, although because
+this point, the algorithm had converged sufficiently to know that that such a movement in
+signal isn't possible. A similar circumstance occurs at 100 seconds in. 
+In general the fits are good, although because
 of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
-the peak a bit. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
+the peaks. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
 \begin{table}[t]
@@
 \end{table}
-
-There are a few things worth noting here. First the time constants vary greatly across
+There are a few results worth noting here. First the time constants vary greatly across
 runs with different noise realizations, yet the sum of the individual time constants
-, $\tau_f$, $\tau_s$ and $\tau_0$ seem to be more consistent. In general the 
+($\tau_f$, $\tau_s$ and $\tau_0$) seem to be more consistent. In general the 
 time constants are falling short of the true time constant. This could be a limitation
-based on the prior distribution (which notably has mean values below the values used
-in the simulation) or it could be caused by the delayed benefit of having a 
-correct time constant. Even if one particle has a better time constant than another, if
-the difference isn't severe, by the time this makes a difference in the weight, the 
+based on the prior distribution (which notably has an initial mean below the values used
+in the simulation) or it could be caused by the dissassociated benefits of a correct time
+constant with correct output. It often takes several measurement periods before a difference
+in time constants becomes relevent. It is possible that in that time other parameters 
+may be more relevent. As a result, by the time the time constant has made a substantial
+difference in the weight, the 
 particle with the better $\tau$ will have been weighted based on other parameters several
@@
 with this admittedly small amount of noise, it would appear that the relation between a set
-of parameters/stimuli with the output is not injective. In other words, it would appear
-that a timeseries is not unique to a set of parameters. This is good justification that 
+of parameters/stimuli and the output is not injective. In other words, it would appear
+that a timeseries is not unique to a single set of parameters. This is good justification that 
 simultaneous blood volume or tagged flow calculations with the conventional FMRI 
 could definitely benefit the model. 
-
 \begin{table}[t]
@@
 as far as the BOLD signal goes. Notice the covariance of $\tau_f$ and $\tau_0$ is $-0.019$ whereas
-the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly they are
-related. The convergence properties of the first run in \autoref{tab:LowNoiseResults} may be
+the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
+some measure of ill-defined behavior between the differen time-constants, as well as several
+other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} may be
 viewed in \autoref{fig:LowNoiseHist}.
+
+TODO more about convergence
 

### 2010-07-16 doc/thesis/results.tex
adjusted the results

@@
 \section{Simulated Volume}
-\begin{enumerate}
-\item Parameter Map 
-\item Error map of parameters
-\item Histogram of \%errors between parameters
-\item Activation Map based on a single region with high $\epsilon$, compared with linear
-\end{enumerate}
 
-Final parameter distribution among active regions.
-Q-Q plots?
+\chapter{Real Data}
+\begin{figure}
+\label{fig:hm_canon_spm}
+\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
+\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=.85]{images/spm_hm_x}}
+\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=.85]{images/spm_hm_y}}
+\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=.85]{images/spm_hm_z}}
+\subfigure{\label{fig:scale_spm} \includegraphics[scale=.3]{images/scale1}}
+\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_spm_x} \autoref{fig:hm_canon_spm_y} 
+         \autoref{fig:hm_canon_spm_x}), as well as a series of axial slices, \autoref{fig:hm_spm}. Units
+         of activation are in Student's T-scores. Higher indicates higher assurance that the signal cannot
+         have occurred through noise alone.}
+\end{figure}
 
-\section{FMRI Data}
-....
+\begin{figure}
+\label{fig:hm_canon_pfilter}
+\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
+\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=.85]{images/pfilter_hm_x}}
+\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=.85]{images/pfilter_hm_y}}
+\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=.85]{images/pfilter_hm_z}}
+\subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.3]{images/scale2}}
+\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
+         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
+         whereas the worst levels could go higher than 100 (not shown).}
+\end{figure}
 

### 2010-07-16 doc/thesis/thesis.tex
updated images, more work on background methods
done with properties I think
added quite a few plots. Should be able to write results for single voxel simulations.
adjusted the results

@@
-\documentclass{report}
-\usepackage[left=3cm,top=2cm,right=3cm]{geometry}
+% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
+%   TITLE: LaTeX report template for ETDs in LaTeX
+%  AUTHOR: Neill Kipp, nkipp@vt.edu
+%     URL: http://etd.vt.edu/latex/
+% SAVE AS: etd.tex
+% REVISED: September 6, 1997
+% 
+%\documentclass[12pt,dvips]{report}
+\documentclass[12pt]{report}
+\setlength{\textwidth}{6.5in}
+\setlength{\textheight}{8.5in}
+\setlength{\evensidemargin}{0in}
+\setlength{\oddsidemargin}{0in}
+\setlength{\topmargin}{0in}
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{0.1in}
+
+%\usepackage[left=3cm,top=2cm,right=3cm]{geometry}
+
+\usepackage{colortbl}
 \usepackage{times}
@@
 \usepackage{amssymb}
-\usepackage{algorithmic}
 \usepackage{setspace}
@@
 \usepackage[colorlinks=true]{hyperref}
+
+\usepackage{algorithm}
+\numberwithin{algorithm}{chapter}
+\usepackage{algorithmic}
 
@@
 \begin{document}
-\title{BOLD Parameter Estimation Using Particle Filters}
-\author{Micah Chambers\\Virginia Polytechnic Institute and State University}
-\date{}
-\maketitle
+\thispagestyle{empty}
+\pagenumbering{roman}
+\begin{center}
+
+% TITLE
+{\Large 
+BOLD Parameter Estimation Using Particle Filters
+}
+
+\vfill
+
+Micah C. Chambers
+
+\vfill
+
+Thesis submitted to the Faculty of the \\
+Virginia Polytechnic Institute and State University \\
+in partial fulfillment of the requirements for the degree of
+
+\vfill
+
+Master of Science \\
+in \\
+Electical Engineering
+
+\vfill
+
+Chris L. Wyatt, Chair \\
+William T. Baumann\\
+Aloysius. A. Beex \\
+Daniel J. Stilwell
+
+\vfill
+
+August 10, 2010\\
+Blacksburg, Virginia
+
+\vfill
+
+Keywords: BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics
+\\
+Copyright 2010, Micah C. Chambers
+
+\end{center}
+
+\pagebreak
+
+\thispagestyle{empty}
+\begin{center}
+
+{\large BOLD Parameter Estimation Using Particle 
+Filters }
+
+\vfill
+
+Micah C. Chambers
+
+\vfill
+
+(ABSTRACT)
+
+\vfill
+
+\end{center}
+
+Abstract Abstract Abstract ...
+
+\vfill
+
+% GRANT INFORMATION
+
+%That this work received support from the Southeastern Universities
+%Research Association (SURA) ``Monticello Library Project'' is purely
+%coincidental.
+
+\pagebreak
+
+% Dedication and Acknowledgments are both optional
+% \chapter*{Dedication}
+% \chapter*{Acknowledgments}
 
 \tableofcontents
+\pagebreak
+
+\listoffigures
+\pagebreak
+
+\listoftables
+\pagebreak
+
+\pagenumbering{arabic}
+\pagestyle{myheadings}
 
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{ieeetr}
+%\bibliographystyle{apalike}
 %\bibliographystyle{abbrvnat}
@@
 %model.
+%
+
+%\chapter{Introduction}
+%\markright{Albert J. Kippleby \hfill Chapter 1. Introduction \hfill}
+%
+%William Shakespeare has profoundly affected the field of literature
+%worldwide.  In the United States there was a surge of Shakespearean
+%literature starting in the 1960s, with the opening of the Montgomery
+%Shakespearean festival and continuing into the present ...
+
+%%%%%%%%%%%%%%%%%
+%
+% Include an EPS figure with this command:
+%   \epsffile{filename.eps}
+%
+
+%%%%%%%%%%%%%%%%
+%
+% Do tables like this:
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
+%%   TITLE: LaTeX report template for ETDs in LaTeX
+%%  AUTHOR: Neill Kipp, nkipp@vt.edu
+%%     URL: http://etd.vt.edu/latex/
+%% SAVE AS: etd.tex
+%% REVISED: September 6, 1997
+%% 
+%
+%% Instructions: Remove the data from this document and replace it with your own,
+%% keeping the style and formatting information intact.  More instructions
+%% appear on the Web site listed above.
+%
+%\documentclass[12pt,dvips]{report}
+%
+%\setlength{\textwidth}{6.5in}
+%\setlength{\textheight}{8.5in}
+%\setlength{\evensidemargin}{0in}
+%\setlength{\oddsidemargin}{0in}
+%\setlength{\topmargin}{0in}
+%
+%\setlength{\parindent}{0pt}
+%\setlength{\parskip}{0.1in}
+%
+%% Uncomment for double-spaced document.
+%% \renewcommand{\baselinestretch}{2}
+%
+%% \usepackage{epsf}
+%
+%\begin{document}
+%
+%\thispagestyle{empty}
+%\pagenumbering{roman}
+%\begin{center}
+%
+%% TITLE
+%{\Large 
+%Use of Metaphor in Shakespeare's Plays and its Potential
+%Application in Twenty-first Century Literature
+%}
+%
+%\vfill
+%
+%Albert J. Kippleby
+%
+%\vfill
+%
+%Dissertation submitted to the Faculty of the \\
+%Virginia Polytechnic Institute and State University \\
+%in partial fulfillment of the requirements for the degree of
+%
+%\vfill
+%
+%Doctor of Philosophy \\
+%in \\
+%Literature and Technology
+%
+%\vfill
+%
+%Neill A. Kipp, Chair \\
+%Emilio J. Arce \\
+%Scott A. Guyer \\
+%Laura Weiss
+%
+%\vfill
+%
+%July 16, 1997 \\
+%Blacksburg, Virginia
+%
+%\vfill
+%
+%Keywords: Metaphysics, Information Retrieval, Spacecraft
+%\\
+%Copyright 1997, Albert J. Kippleby
+%
+%\end{center}
+%
+%\pagebreak
+%
+%\thispagestyle{empty}
+%\begin{center}
+%
+%{\large Use of Metaphor in Shekespeare's Plays and its Potential
+%Application in Twenty-first Century Literature}
+%
+%\vfill
+%
+%Albert J. Kippleby
+%
+%\vfill
+%
+%(ABSTRACT)
+%
+%\vfill
+%
+%\end{center}
+%
+%The need for concrete examples increases when technology becomes
+%difficult to explain.  In documentation for computer systems
+%especially, we see a wide audience of field experts attempting to
+%comprehend documentation for computer software and hardware of which
+%they should only require a cursory understanding.  Additionally, as
+%the pace of the information age quickens we see document authors
+%struggle for \textit{examplia-concretes} with wide applicability, and
+%consistently rely on excerpts from Shakespearean literature as a
+%public-domain source for their various explications.
+%
+%We predict the twenty-first century will be no different.  Actuarial
+%studies show explosion in the information industry such that four out
+%of five persons will be \textit{bona fide} electronic document
+%authors; many of those will have one or more college degrees.  We
+%prove through computer simulation \textsc{Machinum Simitatores} that
+%authors of twenty-first century literature will be affected by these
+%examples and will include metaphor with Shakespearean source into
+%their writing with increasing frequency.
+%
+%\vfill
+%
+%% GRANT INFORMATION
+%
+%That this work received support from the Southeastern Universities
+%Research Association (SURA) ``Monticello Library Project'' is purely
+%coincidental.
+%
+%\pagebreak
+%
+%% Dedication and Acknowledgments are both optional
+%% \chapter*{Dedication}
+%% \chapter*{Acknowledgments}
+%
+%\tableofcontents
+%\pagebreak
+%
+%\listoffigures
+%\pagebreak
+%
+%\listoftables
+%\pagebreak
+%
+%\pagenumbering{arabic}
+%\pagestyle{myheadings}
+%
+%\chapter{Introduction}
+%\markright{Albert J. Kippleby \hfill Chapter 1. Introduction \hfill}
+%
+%William Shakespeare has profoundly affected the field of literature
+%worldwide.  In the United States there was a surge of Shakespearean
+%literature starting in the 1960s, with the opening of the Montgomery
+%Shakespearean festival and continuing into the present ...
+%
+%%%%%%%%%%%%%%%%%%
+%%
+%% Include an EPS figure with this command:
+%%   \epsffile{filename.eps}
+%%
+%
+%%%%%%%%%%%%%%%%%
+%%
+%% Do tables like this:
+%
+% \begin{table}
+% \caption{The Graduate School wants captions above the tables.}
+%\begin{center}
+% \begin{tabular}{ccc}
+% x & 1 & 2 \\ \hline
+% 1 & 1 & 2 \\
+% 2 & 2 & 4 \\ \hline
+% \end{tabular}
+%\end{center}
+% \end{table}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%% If you are using BibTeX, uncomment the following:
+%% \thebibliography
+%%
+%% Otherwise, uncomment the following:
+%% \chapter*{Bibliography}
+%
+%% \appendix
+%
+%% In LaTeX, each appendix is a "chapter"
+%% \chapter{Program Source}
+%
+%
+%% Finally, the VITA
+%\chapter*{Vita}
+%
+%Albert was born on a sunny day...
+%
+%\end{document}

### 2010-07-18 doc/thesis/results.tex
additions to results

@@
 
+\begin{figure}
+\label{fig:hm_canon_pfilter85}
+\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=.85]{images/pfilter_hm85_x}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=.85]{images/pfilter_hm85_y}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=.85]{images/pfilter_hm85_z}}
+\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.3]{images/scale3}}
+\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
+         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
+         whereas the worst levels could go higher than 100 (not shown).}
+\end{figure}
+
+\begin{figure}
+\label{fig:comp1}
+\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_pfilter_37_14_7}}\\
+\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
+\caption{}
+\end{figure}
+
+\begin{figure}
+\label{fig:comp2}
+\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_pfilter_34_12_7}}\\
+\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
+\caption{}
+\end{figure}
+
+\begin{figure}
+\label{fig:comp3}
+\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_pfilter_23_21_7}}\\
+\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
+\caption{}
+\end{figure}
+
+\begin{figure}
+\label{fig:comp4}
+\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
+\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
+\caption{}
+\end{figure}
+
+\begin{figure}
+\label{fig:comp5}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_25_34_25}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
+\caption{}
+\end{figure}
+
 image comparing epsilon-map with GLM activation map

### 2010-07-18 doc/thesis/results.tex
more in the results

--- 2010-07-16 doc/thesis/results.tex
@@
 \chapter{Simulation}
-Two levels are discussed in this section. First the simulation of single time series.
-This is necessary to investigate the power of particle filters in 
+Two levels of simulation are discussed in this section. First the simulation of single time series
+and second, a slice of simulated FMRI time series. The single time series tests
+are necessary to investigate the power of particle filters in 
 identifying the BOLD model in a noisy environment. If the particle filter is unable
@@
 will not be useful. Additionally, determination of reasonable parameters from the
-time-series is also desirable. The second simulation technique will be on a 
-large group of voxels. Although the current technique treats each voxel separately,
-the "Possum" FMRI simulation tool models noise more realistically, and provides additional
-incite into the applicability of particle filters in modeling the BOLD signal on a large
-scale.
+time-series is also desirable. The reason to then simulate an entire slice is that
+the "Possum" FMRI simulation tool models noise more realistically, for instance by
+adding spatial correlation. Additionally 
+it provides insight into the applicability of particle filters in modeling the BOLD 
+signal on a large scale.
 
-\section{Single Timeseries}
+\section{Single Time Series}
 Given the state-space equations for the BOLD signal, simulating a single time 
 series is relatively straight forward. After a "true" signal is generated,
-I then add identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian, I.I.D. steps. Finally a "carrier level" is added, since BOLD is typically
-measured as a \% difference from the baseline. Although technically the particle filter
-algorithm will immediately remove this, and calculate the \% difference, adding a carrier 
-level means that the exact same algorithm which will is used for 
-simulated data may be used for the real data.
+identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
+process with Gaussian I.I.D. steps are added to the true signal. Finally a 
+"carrier level" is added, since BOLD is typically
+measured as a \% difference from the baseline. The particle filter
+algorithm will immediately remove this by calculating the \% difference, 
+but adding a carrier level means that the exact same algorithm which 
+will is used for simulated data may be used for the real data.
 
@@
 the particle filter is excellent at finding the most probable cause of 
-the observed signal; however, given enough noise it will also determine
-the most likely underlying signal; which may not be flat. Thus, it is 
+the observed signal. However, as discussed in the false positive tests,
+even if there is no true underlying signal, the particle filter will still determine
+the most likely BOLD signal; which may not be flat. Thus, it is 
 important to investigate methods of determining false positives using 
-average residuals.
+average residuals. This problem could in actuality never come up, for instance
+if the number of measurements is very large or if the noise peaks are not
+particularly large.
 
@@
 %thus the particle filter was not centered over the correct values. 
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
+\caption{Test Signals with low noise compared to the clean signal.}
+\label{fig:LowNoiseRealization}
+\end{figure}
+
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
+\caption{A comparison of the preprocessed signals for the low noise case. This is the
+noisy input to the actual particle filter algorithm.}
+\label{fig:PreprocessedLowNoise}
+\end{figure}
+
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
+\caption{A comparison of the fitted signals for the low noise case.}
+\label{fig:FitComparisonLowNoise}
+\end{figure}
 
@@
 $\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
-These same parameters will used again in the high noise case. The noise realizations were
+These same parameters will be used again in the high noise case. The noise realizations were
 then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
@@
 For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
-\begin{figure}
-\label{fig:LowNoiseRealization}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_lownoise}
-\caption{Test Signals with low noise compared to the clean signal}
-\end{figure}
-The resulting fits compared to the \emph{true} BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
-While the error between the true signal and the estimated signal are low, the noise
-is also relatively low. The actual signal that was delivered into the particle filter
+The actual signal that was delivered into the particle filter
 is the result of some preprocessing to remove drift, as described in 
-\autoref{sec:Preprocessing}. The preprocessed signal is compared
+\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
+may have some effect on the resulting fit; thus the preprocessed signal is compared
 to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
-\begin{figure}
-\label{fig:PreprocessedLowNoise}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_lownoise}
-\caption{A comparison of the preprocessed signals for the low noise case. This is the
-noisy input to the actual particle filter algorithm.}
-\end{figure}
-The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Preprocessing}.
+The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
 Obviously the spline struggles to match the signal near the end, as the graph shows, although 
-overall the preprocessing seems to have been successful at removing trends. Finally, the 
-set of fits to the preprocessed data, and in 
-the true BOLD signal is shown in \autoref{fig:FitComparisonLowNoise}.
-\begin{figure}
-\label{fig:FitComparisonLowNoise}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_lownoise}
-\caption{A comparison of the fitted signals for the low noise case.}
-\end{figure}
+overall, the preprocessing seems to have been successful at removing trends. Finally, the 
+set of fits to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
 The filtering property of the particle filter is clear from the results here.  In several locations the output
 of the particle filter looks like a filtered version of the input. For instance toward the
-end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
-this point, the algorithm had converged sufficiently to know that that such a movement in
+end, the estimates stay flat in spite of the preprocessed data drifting off. By
+this point, the algorithm has converged sufficiently to know that that such a movement in
 signal isn't possible. A similar circumstance occurs at 100 seconds in. 
-In general the fits are good, although because
-of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
-the peaks. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
+In general the fits are good, although the bias introduced by pre-processing clearly
+results in a slightly larger peak signal. The final parameter sets are shown in 
+\autoref{tab:LowNoiseResults}.
+
 \begin{table}[t]
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) &$\sqrt{MSE}$\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
 \hline 
@@
 \hline 
-1.22214 & 0.3449 & 0.33462 & 0.07138 & 1.60446 & 2.27529 & 1.59454 & 5.10189  &  0.00321067  & 0.00987647  \\
-1.37493 & 0.33183 & 0.36296 & 0.07327 & 1.64076 & 2.10299 & 1.57627 & 5.11867 &  0.00305475  & 0.00993227  \\
-1.16604 & 0.32205 & 0.34057 & 0.08217 & 1.64768 & 2.35351 & 1.24515 & 5.16722 &  0.00328932  & 0.00967961  \\
-1.23176 & 0.32707 & 0.34028 & 0.07961 & 1.62698 & 2.18519 & 1.30326 & 5.04394 &  0.00284719  & 0.00912005  \\
-1.1832 & 0.31787 & 0.34718 & 0.0821 & 1.54961 & 2.29115 & 1.27817 & 5.02396   &  0.00300634  & 0.00971258  \\
-1.1424 & 0.33395 & 0.34725 & 0.07366 & 1.62208 & 2.29084 & 1.4025 & 5.05531   &  0.00283287  & 0.00948483  \\
-1.30041 & 0.3596 & 0.35643 & 0.07679 & 1.56406 & 2.13234 & 1.60338 & 4.99681  &  0.00302802  & 0.01021885  \\
-1.24008 & 0.34601 & 0.33978 & 0.08903 & 1.64989 & 2.23655 & 1.29004 & 5.12651 &  0.00304378  & 0.01007964  \\
-1.1709 & 0.32739 & 0.34644 & 0.08255 & 1.53734 & 2.28262 & 1.37828 & 4.99087  &  0.00334488  & 0.01032886  \\
-1.18967 & 0.34344 & 0.33554 & 0.07976 & 1.53582 & 2.30746 & 1.42774 & 5.03295 &  0.00317542  & 0.01001503  \\
-1.184 & 0.34053 & 0.35017 & 0.08917 & 1.61025 & 2.27926 & 1.16448 & 5.07352   &  0.00288855  & 0.00950536  \\
+1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
+1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
+1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
+1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
+1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
+1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
+1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
+1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
+1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
+1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
+1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
 \hline                                                                           
-1.21868 & 0.33588 & 0.34557 & 0.07995 & 1.59899 & 2.24884 & 1.38762 & 5.06651 & 0.00306562     & 0.00981396 \\
+1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
 \hline 
@@
 \caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
-last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
-residuals. Essentially this is the is the MSE between the estimated signal and the noisy signal which 
-was available to the particle filter. Square root of MSE is the actual \emph{error}, with the true signal,
-which of course was not available to the particle filter.}
+last is mean over the 10 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
+residuals. Although in real data I won't distinguish between residual and error, here I will.
+$\sqrt{MSE}$ is the mean squared difference between the true signal and the estimated signal.
+This is the actual error, due to a combination of noise and preprocessing. }
 \label{tab:LowNoiseResults} 
 \end{table}
-There are a few results worth noting here. First the time constants vary greatly across
-runs with different noise realizations, yet the sum of the individual time constants
-($\tau_f$, $\tau_s$ and $\tau_0$) seem to be more consistent. In general the 
-time constants are falling short of the true time constant. This could be a limitation
-based on the prior distribution (which notably has an initial mean below the values used
-in the simulation) or it could be caused by the dissassociated benefits of a correct time
-constant with correct output. It often takes several measurement periods before a difference
-in time constants becomes relevent. It is possible that in that time other parameters 
-may be more relevent. As a result, by the time the time constant has made a substantial
-difference in the weight, the 
-particle with the better $\tau$ will have been weighted based on other parameters several
-times. Another interesting result in the huge variation in the levels of $V_0$. In general,
-with this admittedly small amount of noise, it would appear that the relation between a set
-of parameters/stimuli and the output is not injective. In other words, it would appear
-that a timeseries is not unique to a single set of parameters. This is good justification that 
-simultaneous blood volume or tagged flow calculations with the conventional FMRI 
-could definitely benefit the model. 
+
 \begin{table}[t]
@@
 
-\begin{figure}[H]
+\begin{figure}
 \subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
-{\includegraphics[trim=7cm 4cm 7cm 4cm, width=15cm]{images/converge_lownoise1}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise1}}\\
 
 \subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
-{\includegraphics[trim=7cm 4cm 7cm 1cm, width=15cm]{images/converge_lownoise2}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
 \end{figure}
@@
 \subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
-{\includegraphics[trim=7cm 2cm 7cm 2cm, width=15cm]{images/converge_lownoise3}}
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
 \label{fig:LowNoiseHist}
 \end{figure}
-The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the $\tau$ parameters are interchangeable
-as far as the BOLD signal goes. Notice the covariance of $\tau_f$ and $\tau_0$ is $-0.019$ whereas
+
+There are a few results worth noting here. First the time constants vary greatly across
+runs with different noise realizations, yet the sum of the individual time constants
+($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
+time constants are falling short of the true time constant. This could be a limitation
+based on the prior distribution (which notably has an initial mean below the true values) 
+or it could be caused by the disassociated benefits of a correct time
+constant with correct output. It often takes several measurement periods before a difference
+in time constants becomes relevant to the weight. Its also possible that output is insensitive
+to changes in the time constants and that varying the activation durations would 
+cause the variations to be more visible. Another interesting result in the huge variation 
+in the levels of $V_0$. In general,
+with this admittedly small amount of noise, it would appear that the relation between a set
+of parameters/stimuli and the output is not injective. In other words,  a time
+series is not unique to a single set of parameters. This is good justification that 
+simultaneous blood volume or tagged flow calculations with the conventional FMRI 
+could benefit the model.
+
+The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the 
+$\tau$ parameters are interchangeable
+as far as the BOLD signal is concerned. Notice the covariance of $\tau_f$ and $\tau_0$
+is $-0.019$ whereas
 the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
-some measure of ill-defined behavior between the differen time-constants, as well as several
-other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} may be
-viewed in \autoref{fig:LowNoiseHist}.
-
-TODO more about convergence
+some measure of ill-defined behavior between the different time-constants, as well as several
+other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
+demonstrates the migration of parameters to match the estimates. Given the relatively high
+variance of parameters, its likely that more measurements or perhaps more variation
+in the stimuli could further differentiate the parameters. Its also notable that
+the final mean of the parameters may not be the best point estimator, and that perhaps
+using the mode would be a better way to estimate the output. Regardless, the time series'
+generated from the estimated mean seems to match the correct signal very well 
+(\autoref{fig:FitComparisonLowNoise}).
 
@@
 \label{sec:HighNoise}
-For the high noise simulation, the exact same procedure was performed again, but with 
-measurement error and drift standard
-deviations an order of magnitude higher. The noisy signals versus the base signal are shown
-in \autoref{fig:HighNoiseRealization} and the preprocessed signals which the particle
-filter attempt to fit are in \autoref{fig:PreprocessedHighNoise}.
-\begin{figure}[H]
-\label{fig:LowNoiseRealization}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_highnoise}
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
 \caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
+\label{fig:HighNoiseRealization}
 \end{figure}
-\begin{figure}[H]
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
+\caption{A comparison of the preprocessed signals for the high noise case.}
 \label{fig:PreprocessedHighNoise}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_highnoise}
-\caption{A comparison of the preprocessed signals for the high noise case.}
+\end{figure}
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
+\caption{A comparison of the fitted signals for the high noise case.}
+\label{fig:FitComparisonHighNoise}
+\end{figure}
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
+\caption{Two particular preprocessed noise realizations for the high noise case.}
+\label{fig:NoiseComparisonJustTwo}
 \end{figure}
 
-The results of the particle filter, for each of the ten runs may be seen in 
-\autoref{fig:FitComparisonHighNoise}. Clearly the preprocessing led the algorithm
-to somewhat higher activation levels, and it would appear that the subtleties of
-different time constants are also lost for most of the runs, although a few seem
-to manage quite accurate results.
-
-\begin{figure}[H]
-\label{fig:FitComparisonHighNoise}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise}
-\caption{A comparison of the fitted signals for the high noise case.}
-\end{figure}
-\begin{figure}[H]
-\label{fig:NoiseComparisonJustTwo}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/highnoise_56_noise}
-\caption{Two particular preprocessed noise realizations for the high noise case.}
-\end{figure}
-\begin{figure}[H]
+\begin{figure}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise_just2}
+\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
 \label{fig:FitComparisonHighNoiseJust2}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise_just2}
-\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
 \end{figure}
 
-It is interesting to consider how the preprocessing and noise may effect the
-parameters of the fitting model. For instance run 1 in \autoref{fig:NoiseComparisonJustTwo}
-certainly seems to be more biased toward higher peaks than run 2. There also appears
-to be more drift than the 20 points per knot could fit, which explains the 
-prolonged increase at 170 seconds. Despite run 1's bias toward higher peaks, for
-some reason the particle filter was able to get a much better match for the
-modeled post-stimulus undershoot (which is likely significantly shorter than it
-should be). Ultimately the results are pretty good. \autoref{tab:SimMSE} shows
-the mean squared error for all two runs, and highlights the two runs analyzed here and 
-in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
-
-\begin{figure}[H]
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run5_1}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run5_2}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run5_3}}
-\label{fig:ConvergenceRuns1}
-\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
-\end{figure}
-
-There are a number of interesting convergence properties of the
-particle filter when more noise is present, as both \autoref{fig:ConvergenceRuns1} and
-\autoref{fig:ConvergenceRuns2} show. Obviously the particle filter seems to converge
-significantly faster, as points tend to be further out on the weighting function. This
-also causes significantly more resampling which is the explanation for the seeming
-jumps in resolution that occur from time to time.
-
-\begin{figure}[H]
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run6_1}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run6_2}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run6_3}}
-\label{fig:ConvergenceRuns2}
-\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
-\end{figure}
-
-The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
+For the high noise simulation, the exact same procedure was performed as with the lower
+noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
+respectively. This an order of magnitude higher than the previous tests, and indeed the
+noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
+As before, the results of preprocessing are helpful in identifying the reason for
+bias in the estimated output signal; thus the preprocessed signals are shown in 
+\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
+for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
+As before, the preprocessing led the algorithm
+to somewhat higher activation levels, and it would appear that the subtleties of
+different time constants are being lost due to the noise. This is easily seen
+by observing the post stimulus undershoots; many of the estimated time series
+have no post-stimulus undershoot but rather decay directly to the base level.
 
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) & $\sqrt{MSE}$\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$  & $\sqrt{MSE}$\\
 \hline 
@@
 \hline 
-1.19002 & 0.23495 & 0.42228 & 0.128 & 1.01468 & 2.47795 & 1.11677 & 4.68265   &0.0140648  &0.01572919   \\
-0.9721 & 0.21902 & 0.30505 & 0.061 & 0.57801 & 1.99596 & 3.46135 & 3.54608    &0.0137314  &0.01377685   \\
-1.57947 & 0.14153 & 0.33798 & 0.1079 & 0.5843 & 2.12475 & 1.78343 & 4.28852   &0.01275372 &0.01577449   \\
-1.10937 & 0.2374 & 0.53491 & 0.0351 & 1.21862 & 3.07365 & 2.35039 & 5.40164   &0.0167258  &0.0115433    \\
-1.10712 & 0.27535 & 0.33651 & 0.03161 & 1.50567 & 2.65181 & 4.19099 & 5.2646  &0.01369793 &0.01221607   \\
-0.58026 & 0.47931 & 0.41354 & 0.11894 & 0.97563 & 3.69018 & 1.00084 & 5.24607 &0.01149511 &0.01315637   \\
+1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 &0.01573   \\
+0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 &0.01378   \\
+1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 &0.01577   \\
+1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 &0.01154    \\
+1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 &0.01222   \\
+0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 &0.01316   \\
 \rowcolor[rgb]{.9,.5,.5}
-1.29515 & 0.25957 & 0.27559 & 0.25952 & 1.70265 & 2.8458 & 0.66172 & 5.84361  &0.01555035 &0.01789781   \\
+1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 &0.01790   \\
 \rowcolor[rgb]{.5,.5,.9}
-1.5185 & 0.21987 & 0.28348 & 0.07417 & 0.88822 & 3.07706 & 1.73929 & 5.48378  &0.01205351 &0.01246202   \\
-0.68736 & 0.3283 & 0.39786 & 0.15614 & 1.07781 & 3.1158 & 0.66432 & 4.88097   &0.01510364 &0.01257713   \\
-1.01703 & 0.28497 & 0.34741 & 0.05672 & 1.58774 & 2.65157 & 2.28519 & 5.25634 &0.012493   &0.01343459   \\
-0.99247 & 0.29795 & 0.32207 & 0.20943 & 0.42757 & 2.21081 & 1.01674 & 3.63085 &0.01216522 &0.01505545   \\
+1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 &0.01246   \\
+0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 &0.01258   \\
+1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 &0.01343   \\
+0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 &0.01506   \\
 \hline                                                                          
-1.09535 & 0.27075 & 0.36152 & 0.11259 & 1.05099 & 2.71958 & 1.84282 & 4.86592 &0.01362132  &0.01396575\\
+1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 &0.01397\\
 \hline 
@@
 \caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
-Note also that the red row is Run 1 and the blue row is Run 2, as used named this section}
+Note also that the red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
+and \autoref{fig:FitComparisonHighNoiseJust2}}
 \label{tab:HighNoiseResults} 
@@
 
+It is interesting to consider how the preprocessing and noise may effect the
+parameters of the fitting model. Consider two runs, shown in 
+\autoref{fig:NoiseComparisonJustTwo}, out of the previous eleven. 
+Run 1 clearly has higher peaks, in spite of the fact that both signals are based
+on the same underlying signal. There also appears
+to be more drift than the 20 measurements per knot could fit, which explains run 1's
+prolonged increase at 170 seconds in. Interestingly, run 1 and run 2 have different
+aspects of the signal that match better. Run 2 had a much better match to the 
+peaks, when compared to the true signal, yet run 1 matched the post-stimulus
+undershoot better. Its worth noting however, that the post-stimulus undershoots
+here are much shorted than the observed prolonged post-stimulus undershoot
+discussed in \autoref{sec:Post Stimulus Undershoot}. 
+Despite the rather significant amount of noise, the ultimate results are 
+actually rather good. \autoref{tab:HighNoiseResults} shows
+the square-root mean-squared-error for all eleven runs, and highlights the two runs analyzed 
+in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
+
+\begin{figure}
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_1}}\\
+\end{figure}
+\begin{figure}
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_2}}\\
+\end{figure}
+\begin{figure}
+\subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_3}}
+\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\label{fig:ConvergenceRuns1}
+\end{figure}
+
+\begin{figure}
+\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_1}}\\
+\end{figure}
+\begin{figure}
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 2] 
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_2}}\\
+\end{figure}
+\begin{figure}
+\subfigure[$Q$, $S$, $F$, $BOLD$, Run 2 ]
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_3}}
+\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\label{fig:ConvergenceRuns2}
+\end{figure}
+
+There are a number of interesting convergence properties of the
+particle filter when more noise is present, as both \autoref{fig:ConvergenceRuns1} and
+\autoref{fig:ConvergenceRuns2} show. The particle filter seems to converge
+significantly faster; as points tend to be further out on the weighting function. This
+also causes significantly more resampling which is the explanation for the percieved
+jumps in resolution that occur from time to time. Here it is clear that the mode and
+the mean will not be substantially different. Because of the increased noise, its likely
+that the weighting function is not sufficiently wide to account for the measurement noise.
+
+The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
+Clearly the additional noise have resulted in much more sporadic results. This is often the 
+result when the particle filter converges too fast, a result of the weighting functions' variance
+being significantly smaller than the measurement noise ($.005$ vs. $.01$). The square-root
+MSE certainly suffers due to this effect.
+
 \subsection{Pure-Noise, low magnitude}
-The two final single-voxel test, forces the particle filter to attempt to "learn" a noise-only
-timeseries. The noise used will be the same as that from the \autoref{sec:HighNoise},
+The two final single-voxel tests force the particle filter to attempt to "learn" a noise-only
+time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
 $\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
-stimulus/input will be set to 0 the entire time. For the particle filter, on the other hand,
-the stimulus will be left as in the previous two sections. The questions is if the particle
-filter is able to appropriately differentiate from a signal with noise. The noise timeseries
+stimulus/input will be set to 0 the entire time. In effect it is a region of the brain for
+which no visual stimulus induces activity. For the particle filter, on the other hand,
+the stimulus will be left as in the previous two sections. The question is if the particle
+filter is able to appropriately differentiate from a signal with noise. The noise time series'
 are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:NoiseOnlyPrep}.
 \begin{figure}[H]
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
+\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
 \label{fig:NoiseOnly}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
-\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
 \end{figure}
 \begin{figure}[H]
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
+\caption{A comparison of the preprocessed signals for the signal-free case.}
 \label{fig:PreprocessedNoiseOnly}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
-\caption{A comparison of the preprocessed signals for the signal-free case.}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:fits_noiseonly}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
 \caption{Fits to the non-active, low noise signal. Note that the line is thick because all
 the fits overlap. This is all 1 fitted lines.}
+\label{fig:fits_noiseonly}
 \end{figure}
@@
 \begin{figure}[H]
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
+\caption{Fit from a single particle filter run, with the noise input. }
 \label{fig:justnoise_fit_0}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
-\caption{Fit from a single particle filter run, with the noise input. }
 \end{figure} %uses allnoise/ALLNOISE-0-w0
@@
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
 \end{figure}
@@
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
 \end{figure}
@@
 \subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
-\label{fig:JustNoiseConvergence}
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
 \caption{Converging histogram for parameters when the signal consists purely of low level noise. 
 Same run as \autoref{fig:justnoise_fit_0}}
+\label{fig:JustNoiseConvergence}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:fits_noiseonly_high}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
 \caption{Fits to the non-active, high noise signal. Note that the line is thick because all
 the fits overlap. This is all 11 fitted lines.}
+\label{fig:fits_noiseonly_high}
 \end{figure}
@@
 \begin{figure}[H]
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
+\caption{Fit from a single particle filter run, with the noise input. }
 \label{fig:justbignoise_fit_0}
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
-\caption{Fit from a single particle filter run, with the noise input. }
 \end{figure} %uses allnoise/ALLNOISE-0-w0
@@
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
 \end{figure}
@@
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
 \end{figure}
@@
 \subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
-\label{fig:JustNoiseConvergenceLarge}
+{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
 \caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
 the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
+\label{fig:JustNoiseConvergenceLarge}
 \end{figure}
@@
 \begin{figure}
-\label{fig:badfit}
-\includegraphics[trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
+\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
 \caption{Particle Filter converging to values that make little sense,
 because the voxel did not correlate with the input in any known way}
+\label{fig:badfit}
 \end{figure}
@@
 \begin{figure}
-\label{fig:highhard}
-\includegraphics[trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
+\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
 \caption{Example of a timeseries that laplace performs better on than
 gaussian. The large spike at the beginning can quickly eliminate all particles} 
+\label{fig:highhard}
 
@@
 \begin{figure}
-\label{fig:badfit_param1}
 \subfigure[]{\includegraphics{images/badfit_param1}}
@@
 $(2.28 0.32 0.33 0.017 0.67 2.88 1.36)$.}
+\label{fig:badfit_param1}
 \end{figure}
@@
 \begin{figure}
-\label{fig:param2_var}
 \subfigure[]{\includegraphics{images/param2a}}
@@
 \caption{A poor fit, using the same parameters as }
+\label{fig:param2_var}
 \end{figure}
@@
 one would hope.
-
-\section{Weighting Function Comparison}
-\label{sec:Results Weights}
 
@@
 \begin{figure}
-\label{fig:hm_canon_spm}
 \subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
@@
          have occurred through noise alone.}
+\label{fig:hm_canon_spm}
 \end{figure}
@@
 \begin{figure}
-\label{fig:hm_canon_pfilter}
 \subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
@@
          whereas the worst levels could go higher than 100 (not shown).}
+\label{fig:hm_canon_pfilter}
+\end{figure}
+
+\begin{figure}
+\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=.85]{images/pfilter_hm85_x}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=.85]{images/pfilter_hm85_y}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=.85]{images/pfilter_hm85_z}}
+\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.3]{images/scale3}}
+\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
+         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
+         whereas the worst levels could go higher than 100 (not shown).}
+\label{fig:hm_canon_pfilter85}
+\end{figure}
+
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_pfilter_37_14_7}}\\
+\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
+\caption{}
+\label{fig:comp1}
+\end{figure}
+
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_pfilter_34_12_7}}\\
+\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
+\caption{}
+\label{fig:comp2}
+\end{figure}
+
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_pfilter_23_21_7}}\\
+\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
+\caption{}
+\label{fig:comp3}
+\end{figure}
+
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
+\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
+\caption{}
+\label{fig:comp4}
+\end{figure}
+
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_25_34_25}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
+\caption{}
+\label{fig:comp5}
 \end{figure}

### 2010-07-18 doc/thesis/results.tex
"done" with single timeseries simulations, on the the real results next then back to the slice sim

@@
 
+%LOW NOISE SECTION, with a signal
 \subsection{Signal with Low Noise}
@@
 \caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
-last is mean over the 10 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
+last is mean over the 11 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
 residuals. Although in real data I won't distinguish between residual and error, here I will.
@@
 
+%HIGH NOISE SECTION, with a signal
 \subsection{High Noise}
@@
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_1}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_1}}\\
 \end{figure}
@@
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_2}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
 \end{figure}
@@
 \subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_3}}
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
 \caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
@@
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_1}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_1}}\\
 \end{figure}
@@
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 2] 
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_2}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_2}}\\
 \end{figure}
@@
 \subfigure[$Q$, $S$, $F$, $BOLD$, Run 2 ]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_3}}
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
 \caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
@@
 
+%NO SIGNAL, LOW NOISE
 \subsection{Pure-Noise, low magnitude}
+\label{sec:PureNoiseLowMag}
 The two final single-voxel tests force the particle filter to attempt to "learn" a noise-only
@@
 which no visual stimulus induces activity. For the particle filter, on the other hand,
-the stimulus will be left as in the previous two sections. The question is if the particle
-filter is able to appropriately differentiate from a signal with noise. The noise time series'
-are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:NoiseOnlyPrep}.
+the stimulus will be left as in the previous two sections. The question then is how the
+particle filter will respond to signal that does not correlate with the input, and how
+the output may be differentiated from a time series that does. The time series'
+are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:PreprocessedNoiseOnly}.
+The line fits are shown in \autoref{fig:fits_noiseonly}.
+
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/realization_noiseonly}
 \caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
@@
 \end{figure}
+
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
-\caption{A comparison of the preprocessed signals for the signal-free case.}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/preprocessed_noiseonly}
+\caption{A comparison of the preprocessed signals for the signal-free case. (
+$\sigma_y = .01, \sigma_x = .005$)}
 \label{fig:PreprocessedNoiseOnly}
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSE}$   \\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSR}$   \\
 \hline 
@@
 \end{tabular}
-\caption{Estimated Parameters on 10 different runs with low noise. }
+\caption{Estimated Parameters on 11 different runs with low noise and no signal present.}
 \label{tab:NoiseOnlyResults} 
@@
 
-The line fits are shown in \autoref{fig:fits_noiseonly}. Now the
-low MSE and reasonableness of the parameters is cause for concern. Additionally, the precision
-with which the convergence occurred is enviable, compared to the actual signals.
-However, the impulse response, which has less importance in nonlinear systems, could
-be usefull for separating out good voxels from bad. Note that in \autoref{fig:PreprocessedNoiseOnly}
-peaks at well below $.005$, the standard deviation of the weighting function. Thus, the reason
-the fits are in such agreement is that no convergence occurred at all, as 
-further shown by \autoref{JustNoiseConvergence}. Thus one more test is necessary on
-a signal with enough noise to actually be considered a signal. But first,
-there is still the matter of identifying this noise-only voxel. Clearly MSE and
-the parameters themselves don't have the information to differentiate in this case.
-Thus the answer falls on either the final estimated variance (which will be high
-because no particles strayed far from the center) or the impulse response.
-I prefer the impulse response because it doesn't depend on something that 
-may be less than dependable, namelay the rate of convergence.
-
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
 \caption{Fits to the non-active, low noise signal. Note that the line is thick because all
-the fits overlap. This is all 1 fitted lines.}
+the fits overlap. This is all 11 fitted lines.}
 \label{fig:fits_noiseonly}
@@
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justnoise_fit_0}
 \caption{Fit from a single particle filter run, with the noise input. }
@@
 \begin{figure}[H]
+\centering
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_1}}\\
 \end{figure}
+
 \begin{figure}[H]
+\centering
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_2}}\\
 \end{figure}
+
 \begin{figure}[H]
+\centering
 \subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_3}}
 \caption{Converging histogram for parameters when the signal consists purely of low level noise. 
@@
 
-To give a frame of reference to the impulse response level, the impulse response of every
-parameter estimated, includeing ones from \autoref{sec:BigNoise}, are shown in \autoref{tab:ImpulseResponses}.
+From the results it is clear that the parameters have not taken any visible turns into
+unrealistic territory. It appears individual parameters will not be a distinguishing factor
+in determining the validity of the model. Additionally the raw square-root mean-squared-residual
+also appears to be lower than the SMSR of the low noise simulation from earlier. 
+The precision with which the convergence occurred is actually the best seen in any
+simulation.
 
+However, the reason for low error may in fact be simply the fact that the over all
+signal is significantly lower than any other simulation seen in other simulations.
+The peaks never even cross 1\% difference, with most staying even below .5\% 
+(\autoref{fig:PreprocessedNoiseOnly}). Because the signal is generally staying
+well within the range of $.005$, which is the standard deviation of the weighting function
+its unlikely that any particles will have a substantially different wieght. Thus, the reason
+the fits are in such agreement is that no particles gained increased weight, and so
+no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
+Thus we have now seen how the particle filter responds when the signal is completely
+random and low level, however there is also some possibility that the signal will
+be random, yet active. Thus one more test is necessary on
+a signal with enough noise to actually be considered a signal.
+
+%NO SIGNAL, VERY HIGH NOISE
 \subsection{Pure Noise, High Magnitude}
 \label{sec:PureNoiseHighMag}
+To determine how the particle filter will respond to active, yet unrelated
+portions of the brain, in this section random data will again be used, but
+with even higher standard deviations. To simulate this case another
+pure noise signal was generated using a $\sigma_x$ of $.1$ and a $\sigma_y$ of $.05$.
+
+As before, the convergence all seems to follow a similar path, leaving almost no
+variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
+the peaks are comparable to peaks in a "real" timeseries. Interestingly the algorithm
+suffered from almost constant particle deprevation, meaning that the heuristic
+for rescuing the particle filter from particle deprivation, discuessed in
+\autoref{sec:Resampling} in fact was working against the proper course of action. Of course
+ the proper course of action in this case would be for the particle filter to fail,
+ since no particle can really properly estimate a random sequence. When this mechanism
+ was removed, all 11 runs stopped due to particle deprevation (all weights hit zero). 
+ The problem with allowing particle deprivation to occur is that it can \emph{rarely}
+ occur if the prior happens to not be dense enough in the area of the solution. Thus,
+ there is an issue of false positives vs. false negatives. A middle ground, such as 
+ allowing deprivation to occur if the recovery mechanism has already triggered once
+ may be an acceptable compromise, however it would be difficult to strike the 
+ right balance. Instead for the purposes here we depend on the square root MSR to 
+ evaluate if the results were reasonable or not.
 
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly_high}
 \caption{Fits to the non-active, high noise signal. Note that the line is thick because all
@@
 \begin{figure}[H]
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justbignoise_fit_0}
 \caption{Fit from a single particle filter run, with the noise input. }
@@
 \begin{figure}[H]
+\centering
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
+{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
 \end{figure}
@@
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
+{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
 \end{figure}
@@
 \subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
+{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
 \caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
@@
 \end{figure}
+
+From \autoref{fig:justbignoise_fit_0} it is clear that, despite the fact that the input 
+is competely detached from the signal, a non-zero response by the particle filter does
+lead to decreased MSR. This is primarily due to the preprocessing which forces the signal to
+have a non-zero mean. While this scheme does create a type of artificial activation for
+the particle filter, in reality the MSR will still be less than if the signal was left
+with zero mean, and the particle filter stayed flat. At best the particle filter will 
+over-smooth creating a line that does down the center of the noisy signal. In such a case
+the MSR will be the same as the alternative.
 
@@
 \hline 
-&$\sqrt{MSE}$ & $N\sqrt{MSE}$ &
- $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
- $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
- $\sqrt{MSE}$ & $N\sqrt{MSE}$ \\
+&$\sqrt{MSR}$ & $N\sqrt{MSR}$ &
+ $\sqrt{MSR}$ & $N\sqrt{MSR}$ &
+ $\sqrt{MSR}$ & $N\sqrt{MSR}$ &
+ $\sqrt{MSR}$ & $N\sqrt{MSR}$ \\
 \hline       
@@
 \end{tabular}
-\caption{$\sqrt{MSE}$ and the normalized $\sqrt{MSE}$, $\left(\sqrt{MSE}/(\text{Median Absolute Deviation})\right)$  
-for different configurations.} \label{tab:SignleVoxelActivationComparison} 
+\caption{$\sqrt{MSE}$ and the normalized $\sqrt{MSE}$, 
+    $\left(\sqrt{MSE}/(\text{Median Absolute Deviation})\right)$ for different configurations.} 
+\label{tab:SignleVoxelActivationComparison} 
 \end{table}
 
+The first step in determining the validity of a model to provide and order of quality
+to rate the results by. Unfortunately, because of the variablity in the signal levels the
+raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
+just because the square-root mean-squared-residual is low doesn't mean the fit is
+"good". Therefore, instead we will normalize the signal based on a value proportional to
+the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
+peaks and troughs, rather than using MAX and MIN values, I used a robust estimator of 
+scale, the Median-Absolute-Deviation (MAD). Technically this is an estimator of the standard
+deviation, however in this case it is also a good estimator of the scale of the input
+signal. Thus, each point in the estimate as well as the original signal may be divided by
+this, leading to a normalized signal. Instead, however, I simply divided the 
+$\sqrt{MSR}$ by the MAD, which is functionally equivalent but leaves the  BOLD
+levels intact for further analysis. The normalized square root MSR values are 
+shown in \autoref{tab:SignleVoxelActivationComparison}.
+
 Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, it is
-clear that distinguishing between these cases is a difficult prospect. Although
-the impulse response, which is determined by the peak response to a very short input,
-is is slightly different, there is no reason to believe that a noise distribution
-could not get a higher impulse response. The $\sqrt{MSE}$ is also not going to be
-able to distinguish between the high-noise case and the no signal case. Therefore, there
-is a noise ceiling above which it will be impossible to determine model validity. The fact
-that the particle filter is able to get close to the "true" signal is sadly irrelevant.
+clear that distinguishing between these cases is a difficult prospect. One possibility
+is to just allow the particle filter to fail when particle deprivation occurs; however,
+as discussed in the previous section, that could open the door to more false negatives.
+Naturally there is a noise ceiling above which it will be impossible to determine model 
+validity. The fact
+that the particle filter is able to get close to the "true" signal is sadly irrelevant
+when the noise is this high.
 Ultimately, a threshold well below those shown in the high noise case 
-will have to be placed on the error, if any confidence is going to be placed on the results.
-
-Each test took approximately 60 seconds on an 8 core Xeon (specifics?).
+will have to be placed on the error, if any confidence is going to be placed in
+the results. Luckily the "high noise" case featured in \autoref{sec:HighNoise}
+is actually noisier than typical FMRI.
 

### 2010-07-18 doc/thesis/results.tex
split up results

@@
 quantitative difference between the two parameter sets may be found.
-
-\section{Real Data}
-Finally, we also performed inference based on real FMRI data. The scanner we used...
-%... more specifics...
-
-Before performing tests on a full image, I tested the results of the particle filter
-on regions deemed active and non-active by statistical parametric mapping. 
-\section{Single-Voxel Simulation}
-The results
-
-\section{Single-Voxel Analysis}
-This section discusses the results when the particle filter was
-applied on a single voxel. The parameters are the same as
-those used later for entire image analysis; however, the results
-are more in-depth.
-
-The run-time for a single voxel depends on the several factors. First, the
-overall length of the signal being analyzed. For 1000 measurements it takes
-about 10 to 15 minutes. On the other hand, in real circumstances the
-length is only around 150 measurements. The number of  integration
-can certainly make a large difference, however dropping below 1000 (.001 seconds)
-is definitely not recommended.
-
-For a period I considered 1000 to be a
-fine number; however when generating simulated data I found that every once
-in a while 1000 was not enough. This is problematic in the actual particle
-filter since, given the large number of simultaneous integrations taking 
-place, its likely that a few particles will fail and be weighted at 0 because
-of this problem. Additionally, because the typical case where a failure would
-occur is at fast moving times/parameters the particles all tend to fail together.
-The result is particle deprivation - no particles with non-zero weights remain.
-The other possible outcome is that low time constant particles get pruned resulting
-in excessively smoothed estimates for the time series'. Its possible that a
-kind of stop-gap measure could be put into place; wherein particles that are
-about to be set to NaN are integrated again with finer grained steps. However
-many times the non-real results don't occur until several time steps after the 
-numbers get strange. So for instance, the timestep was too long, allowing 
-$f$ to go negative, resulting in extremely large values of $q$. There are many
-different ways where this sort of event can occur, and unfortunately sometimes
-there is no way to get back to before the state starting going out of control.
-
-Another crucial factor for run time is how long before the first re-sampling 
-occurs. Because the prior is represented initially with significantly more
-particles, if the model fits very well, or for some reason the effective
-number of particles stays high, resampling could take a long time to occur.
-When this happens the particles filter can take a factor of 10 longer to run.
-However, if the particle count isn't initially set high, there is a much larger
-chance of particle deprivation occurring. Since there is no real way to know
-how long it will take to resmaple the first time, there is little the 
-algorithm can do to fix this (except perhaps forcing resampling after
-some period of time). On the other end of the spectrum, if the time
-series doesn't match the model at all, particle deprivation will occur extremely
-quickly.  The upshot of this is that the particle filter is able to 
-identify these sections very quickly, and thus not waste much time there.
-The difficulty though, is the regions in between the perfect fit and the
-awful fit. If particle deprivation does occur, did it occur randomly to an
-activated region or did it occur inevitably because the region doesn't fit.
-The reason for having so many initial particles is to give density to the
-distribution to make false negatives more rare. Perhaps the correct method
-is to still set the initial particles very high, but if for a long time no
-resampling occurs, halve the standard deviation of the measurement error.
-Thus, if the weighting function is not discriminating enough, force it 
-to be more picky about the results. Of course this could lead to particle
-deprivation as well if the standard deviation is brought down too quickly.
-Of course, if a fat-tailed distribution is used for the weighting function,
-or the standard deviation of a Gaussian weighting function is sufficiently large,
-the particle filter will simply converge to meaningless values. The question
-of whether
-
-\begin{figure}
-\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
-\caption{Particle Filter converging to values that make little sense,
-because the voxel did not correlate with the input in any known way}
-\label{fig:badfit}
-\end{figure}
-
-\begin{figure}
-\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
-\caption{Example of a timeseries that laplace performs better on than
-gaussian. The large spike at the beginning can quickly eliminate all particles} 
-\label{fig:highhard}
-
-\end{figure}
-
-For this reason, it is important to threshold based on the MSE, and the
-impulse response. Naturally the MSE is intended to catch very voxels with a
-very poor fit,
-but like \autoref{fig:badfit}, the algorithm may have found a line
-down the middle that gives a decent mean squared error. In this case
-the excessive smoothing and low actual activation level will easily
-be distinguished by finding the peak response to single short pulse.
-
-The choice of a prior, as discussed previously, is extremely important. While a
-prior may have the potential to give good results, being a monte-carlo algorithm
-there is the possibility for inconsistencies. Thus, increasing the variance
-of the time-constants may allow additional flexibility, it will also cause
-additional model variance. Case in point, the exact same algorithm run twice
-with standard deviations of $.35, .35, .35$ for the time constants resulted in two
-very different fits, see \autoref{fig:param1_var}.
-
-\begin{figure}
-\subfigure[]{\includegraphics{images/badfit_param1}}
-\subfigure[]{\includegraphics{images/goodfit_param1}}
-\caption{The same priors gave rise to both fits. Params: ($\tau_0 \alpha E_0 V_0 \tau_f \tau_s \epsilon$)
-for the good fit: $(1.40 .32 0.34 0.04  0.96 2.76)$, and for the poor fit:
-$(2.28 0.32 0.33 0.017 0.67 2.88 1.36)$.}
-\label{fig:badfit_param1}
-\end{figure}
-
-For this reason, I actually lowered the standard deviattions of the time
-constants to prevent over-smoothing. This resulted in the more consistent, but
-slightly worse results seen in \autoref{fig:param2_var}
-
-\begin{figure}
-\subfigure[]{\includegraphics{images/param2a}}
-\subfigure[]{\includegraphics{images/param2b}}
-\caption{A poor fit, using the same parameters as }
-\label{fig:param2_var}
-\end{figure}
-
-\section{Particle Collapse Recovery}
-This heuristic was initially intended to deal with cases where the particle
-filter prematurely converged; and to reverse this effect. In reality, however
-it did not perform well, and so it was not used in full-brain analysis. Ultimately
-it did not prevent collapse of the distribution, and it often caused dramatic
-increased in the variance of parameter estimates. It also pronged the analysis
-time of inactive time series. Essentially the effect was exactly opposite of what
-one would hope.
-
-\section{Single Time-Series Simulation}
-
-Graphs:
-
-For simulated data, single timeseries:
-
-For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
-\{100, 500, 1000\} particles
-\begin{enumerate}
-\item Ground truth vs. Estimated signal during particle filter run
-\item Ground truth vs. Estimated signal with final parameter set
-\item True Parameters vs. Final Parameter Sets
-\item Variance of final parameters when faced with same ground truth, different noise
-\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
-\item Estimator Variance based on different noise runs
-\item Final Particle Distribution
-\end{enumerate}
-
-For Simulated Data, Full Volume:
-
-%note to self, epsilon should probably be uniform between 0 and something
-\section{Simulated Volume}
-
-\chapter{Real Data}
-\begin{figure}
-\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
-\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=.85]{images/spm_hm_x}}
-\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=.85]{images/spm_hm_y}}
-\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=.85]{images/spm_hm_z}}
-\subfigure{\label{fig:scale_spm} \includegraphics[scale=.3]{images/scale1}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_spm_x} \autoref{fig:hm_canon_spm_y} 
-         \autoref{fig:hm_canon_spm_x}), as well as a series of axial slices, \autoref{fig:hm_spm}. Units
-         of activation are in Student's T-scores. Higher indicates higher assurance that the signal cannot
-         have occurred through noise alone.}
-\label{fig:hm_canon_spm}
-\end{figure}
-
-\begin{figure}
-\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=.85]{images/pfilter_hm_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=.85]{images/pfilter_hm_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=.85]{images/pfilter_hm_z}}
-\subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.3]{images/scale2}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
-         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
-         whereas the worst levels could go higher than 100 (not shown).}
-\label{fig:hm_canon_pfilter}
-\end{figure}
-
-\begin{figure}
-\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=.85]{images/pfilter_hm85_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=.85]{images/pfilter_hm85_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=.85]{images/pfilter_hm85_z}}
-\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.3]{images/scale3}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
-         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
-         whereas the worst levels could go higher than 100 (not shown).}
-\label{fig:hm_canon_pfilter85}
-\end{figure}
-
-\begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_pfilter_37_14_7}}\\
-\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
-\caption{}
-\label{fig:comp1}
-\end{figure}
-
-\begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_pfilter_34_12_7}}\\
-\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
-\caption{}
-\label{fig:comp2}
-\end{figure}
-
-\begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_pfilter_23_21_7}}\\
-\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
-\caption{}
-\label{fig:comp3}
-\end{figure}
-
-\begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
-\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
-\caption{}
-\label{fig:comp4}
-\end{figure}
-
-\begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_25_34_25}}\\
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{}
-\label{fig:comp5}
-\end{figure}
-
-image comparing epsilon-map with GLM activation map

### 2010-07-18 doc/thesis/resultsB.tex
split up results

\chapter{Real Data}
Finally, we also performed inference on a real FMRI scan. The scanner we used...
%... more specifics...

Before performing tests on a full image, I tested the results of the particle filter
on regions deemed active and non-active by statistical parametric mapping. 
This served the purpose of tuning the priors to reasonable levels etc.
After that, the particle filter was applied to every voxel in an FMRI 
image individualy, and the results were put into a parameter map 
image. Additionally, the resulting MSE and square-root MSE were placed
into images for further analysis.

\section{Single-Voxel Analysis}
This section discusses the results when the particle filter was
applied on a single voxel. The parameters are the same as
those used later for entire image analysis; however, the results
are more in-depth.

The run-time for a single voxel depends on the several factors. First, the
overall length of the signal being analyzed. For 1000 measurements it takes
about 10 to 15 minutes. On the other hand, in real circumstances the
length is only around 150 measurements. The number of  integration
can certainly make a large difference, however dropping below 1000 (.001 seconds)
is definitely not recommended.

For a period I considered 1000 to be a
fine number; however when generating simulated data I found that every once
in a while 1000 was not enough. This is problematic in the actual particle
filter since, given the large number of simultaneous integrations taking 
place, its likely that a few particles will fail and be weighted at 0 because
of this problem. Additionally, because the typical case where a failure would
occur is at fast moving times/parameters the particles all tend to fail together.
The result is particle deprivation - no particles with non-zero weights remain.
The other possible outcome is that low time constant particles get pruned resulting
in excessively smoothed estimates for the time series'. Its possible that a
kind of stop-gap measure could be put into place; wherein particles that are
about to be set to NaN are integrated again with finer grained steps. However
many times the non-real results don't occur until several time steps after the 
numbers get strange. So for instance, the timestep was too long, allowing 
$f$ to go negative, resulting in extremely large values of $q$. There are many
different ways where this sort of event can occur, and unfortunately sometimes
there is no way to get back to before the state starting going out of control.

Another crucial factor for run time is how long before the first re-sampling 
occurs. Because the prior is represented initially with significantly more
particles, if the model fits very well, or for some reason the effective
number of particles stays high, resampling could take a long time to occur.
When this happens the particles filter can take a factor of 10 longer to run.
However, if the particle count isn't initially set high, there is a much larger
chance of particle deprivation occurring. Since there is no real way to know
how long it will take to resmaple the first time, there is little the 
algorithm can do to fix this (except perhaps forcing resampling after
some period of time). On the other end of the spectrum, if the time
series doesn't match the model at all, particle deprivation will occur extremely
quickly.  The upshot of this is that the particle filter is able to 
identify these sections very quickly, and thus not waste much time there.
The difficulty though, is the regions in between the perfect fit and the
awful fit. If particle deprivation does occur, did it occur randomly to an
activated region or did it occur inevitably because the region doesn't fit.
The reason for having so many initial particles is to give density to the
distribution to make false negatives more rare. Perhaps the correct method
is to still set the initial particles very high, but if for a long time no
resampling occurs, halve the standard deviation of the measurement error.
Thus, if the weighting function is not discriminating enough, force it 
to be more picky about the results. Of course this could lead to particle
deprivation as well if the standard deviation is brought down too quickly.
Of course, if a fat-tailed distribution is used for the weighting function,
or the standard deviation of a Gaussian weighting function is sufficiently large,
the particle filter will simply converge to meaningless values. The question
of whether

\begin{figure}
\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
\caption{Particle Filter converging to values that make little sense,
because the voxel did not correlate with the input in any known way}
\label{fig:badfit}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
\caption{Example of a timeseries that laplace performs better on than
gaussian. The large spike at the beginning can quickly eliminate all particles} 
\label{fig:highhard}

\end{figure}

For this reason, it is important to threshold based on the MSE, and the
impulse response. Naturally the MSE is intended to catch very voxels with a
very poor fit,
but like \autoref{fig:badfit}, the algorithm may have found a line
down the middle that gives a decent mean squared error. In this case
the excessive smoothing and low actual activation level will easily
be distinguished by finding the peak response to single short pulse.

The choice of a prior, as discussed previously, is extremely important. While a
prior may have the potential to give good results, being a monte-carlo algorithm
there is the possibility for inconsistencies. Thus, increasing the variance
of the time-constants may allow additional flexibility, it will also cause
additional model variance. Case in point, the exact same algorithm run twice
with standard deviations of $.35, .35, .35$ for the time constants resulted in two
very different fits, see \autoref{fig:param1_var}.

\begin{figure}
\subfigure[]{\includegraphics{images/badfit_param1}}
\subfigure[]{\includegraphics{images/goodfit_param1}}
\caption{The same priors gave rise to both fits. Params: ($\tau_0 \alpha E_0 V_0 \tau_f \tau_s \epsilon$)
for the good fit: $(1.40 .32 0.34 0.04  0.96 2.76)$, and for the poor fit:
$(2.28 0.32 0.33 0.017 0.67 2.88 1.36)$.}
\label{fig:badfit_param1}
\end{figure}

For this reason, I actually lowered the standard deviattions of the time
constants to prevent over-smoothing. This resulted in the more consistent, but
slightly worse results seen in \autoref{fig:param2_var}

\begin{figure}
\subfigure[]{\includegraphics{images/param2a}}
\subfigure[]{\includegraphics{images/param2b}}
\caption{A poor fit, using the same parameters as }
\label{fig:param2_var}
\end{figure}

\section{Particle Collapse Recovery}
This heuristic was initially intended to deal with cases where the particle
filter prematurely converged; and to reverse this effect. In reality, however
it did not perform well, and so it was not used in full-brain analysis. Ultimately
it did not prevent collapse of the distribution, and it often caused dramatic
increased in the variance of parameter estimates. It also pronged the analysis
time of inactive time series. Essentially the effect was exactly opposite of what
one would hope.

\section{Single Time-Series Simulation}

Graphs:

For simulated data, single timeseries:

For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
\{100, 500, 1000\} particles
\begin{enumerate}
\item Ground truth vs. Estimated signal during particle filter run
\item Ground truth vs. Estimated signal with final parameter set
\item True Parameters vs. Final Parameter Sets
\item Variance of final parameters when faced with same ground truth, different noise
\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
\item Estimator Variance based on different noise runs
\item Final Particle Distribution
\end{enumerate}

For Simulated Data, Full Volume:

%note to self, epsilon should probably be uniform between 0 and something
\section{Simulated Volume}

\chapter{Real Data}
\begin{figure}
\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=.85]{images/spm_hm_x}}
\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=.85]{images/spm_hm_y}}
\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=.85]{images/spm_hm_z}}
\subfigure{\label{fig:scale_spm} \includegraphics[scale=.3]{images/scale1}}
\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_spm_x} \autoref{fig:hm_canon_spm_y} 
         \autoref{fig:hm_canon_spm_x}), as well as a series of axial slices, \autoref{fig:hm_spm}. Units
         of activation are in Student's T-scores. Higher indicates higher assurance that the signal cannot
         have occurred through noise alone.}
\label{fig:hm_canon_spm}
\end{figure}

\begin{figure}
\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=.85]{images/pfilter_hm_x}}
\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=.85]{images/pfilter_hm_y}}
\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=.85]{images/pfilter_hm_z}}
\subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.3]{images/scale2}}
\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
         whereas the worst levels could go higher than 100 (not shown).}
\label{fig:hm_canon_pfilter}
\end{figure}

\begin{figure}
\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=.85]{images/pfilter_hm85_x}}
\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=.85]{images/pfilter_hm85_y}}
\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=.85]{images/pfilter_hm85_z}}
\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.3]{images/scale3}}
\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
         whereas the worst levels could go higher than 100 (not shown).}
\label{fig:hm_canon_pfilter85}
\end{figure}

\begin{figure}
\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_pfilter_37_14_7}}\\
\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
\caption{}
\label{fig:comp1}
\end{figure}

\begin{figure}
\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_pfilter_34_12_7}}\\
\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
\caption{}
\label{fig:comp2}
\end{figure}

\begin{figure}
\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_pfilter_23_21_7}}\\
\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
\caption{}
\label{fig:comp3}
\end{figure}

\begin{figure}
\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
\caption{}
\label{fig:comp4}
\end{figure}

\begin{figure}
\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_25_34_25}}\\
\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
\caption{}
\label{fig:comp5}
\end{figure}

image comparing epsilon-map with GLM activation map

### 2010-07-19 doc/thesis/methods.tex
more to the results

@@
 
+\section{Particle Filter Settings}
+The run-time for a single voxel depends on the several factors. First, the
+overall length of the signal being analyzed. For 1000 measurements it takes
+about 6 minutes. On the other hand, in real circumstances the
+length is only around 150 measurements and takes around 1 minute. The size of 
+local linearization steps can certainly also make a large difference; and 
+although it can be tempting to decrease the resolution, going above .001 seconds
+per step is not recommended. In most cases millisecond resolution
+is fine; however, when generating simulated data I found that every once
+in a while this was not enough. This is problematic in the actual particle
+filter since, given the large number of simultaneous integrations taking 
+place, its probable that a few particles will fail and be unfairly thrown away.
+There are two possible outcomes of this. The typical case 
+is at some point where are the particles are in a similar location and 
+so many or all particles fail together.
+The result is then particle deprivation - no particles with non-zero weights remain.
+At best this will cause a convergence to a less than optimal posterior 
+probability, at worse at will result in a false negative. 
+The other possible outcome is that low time constant (or fast moving)
+particles get pruned, leaving an excessively smoothed estimate of the output.
+
+Although runge-kutta solvers could be applied, this will still not
+guarantee that features aren't missed. Its possible that a
+kind of stop-gap measure could be put into place; wherein particles that are
+about to be set to NaN are integrated again with finer grained steps. However
+many times the non-real results don't occur until several time steps after the 
+numbers become unrealistic. So for instance, the time step was too long, allowing 
+$f$ to go negative, resulting in extremely large values of $q$. There are many
+different ways where this sort of event can occur, and unfortunately sometimes
+there is no way to get back to before the state starting going out of control.
+
+Another crucial factor for run time is how long before the first re-sampling 
+occurs. Because the prior is represented initially with significantly more
+particles, if the model fits very well, or for some reason the effective
+number of particles stays high, resampling could take a long time to occur.
+For this reason, rather than allowing the particle filter to continue on 
+with this large number of particles, after 20 seconds have passed the
+algorithm forces resampling. Of course the choice of 20 seconds is somewhat
+arbitrary, but at the very least this gives some time for the particle to
+be re-weighted before the new distribution is drawn. At the very least the
+distribution will be a thinned out version of the prior.
+
 %For simulated and real images (tests with multiple time-series), tests were 
@@
 %precision in the output maps.
+
+\section{FMRI Configuration}
+For the FMRI data discussed in \autoref{sec:RealData}, tests were 
+performed on a right handed volunteer using a Gradient EPI sequence
+on a 1.5T GE scanner with a 2.1s TR. Slices were 5mm thick and echo
+time was (todo).

### 2010-07-19 doc/thesis/results.tex
more to the results

@@
 it provides insight into the applicability of particle filters in modeling the BOLD 
-signal on a large scale.
+signal on a large scale. Both sorts of tests are usefull in determining the
+consistency of the particle filter. For the single voxel tests, each test will
+actually be performed eleven different times, each with a new noise realization.
+If runs fail to have consistent estimates of the underlying BOLD signal, then
+this would indicate a large variance error in the model. Similarly, in the
+Possum simulation there are actually only four discrete parameter sets driving the time-series.
+Nevertheless each voxel will have a different tissue composition and a different
+noise realization. Therefore the Possum simulation should also provide a good
+test of the consistency of the BOLD model.
 
@@
 
+There were two primary purposes of these tests. First given the nature of monte-carlo
+techniques it is important to ensure consistency of results. Although the parameter
+sets were somewhat inconsistent, the quality of the output was actually quite consistent
+across eleven runs. The second purpose was of course to determine how the particle 
+filter responded to different signal to noise ratios. Although the direct $\sqrt{MSR}$ is not
+a good measure of quality, the normalized $\sqrt{MSR}$ is. Drawing a line at a particular
+threshold may be difficult, however from the tests it seems that normalized $\sqrt{MSR}$
+below $1$ may be a good start.
+
 \section{Multi-voxel Simulation}

### 2010-07-19 doc/thesis/resultsB.tex
more to the results

@@
-\chapter{Real Data}
-Finally, we also performed inference on a real FMRI scan. The scanner we used...
+%\chapter{Real Data}
+%Finally, we also performed inference on a real FMRI scan. The scanner we used...
 %... more specifics...
 
-Before performing tests on a full image, I tested the results of the particle filter
-on regions deemed active and non-active by statistical parametric mapping. 
-This served the purpose of tuning the priors to reasonable levels etc.
-After that, the particle filter was applied to every voxel in an FMRI 
-image individualy, and the results were put into a parameter map 
-image. Additionally, the resulting MSE and square-root MSE were placed
-into images for further analysis.
+% TODO include single?
+%Before performing tests on a full image, I the particle filter
+%on regions deemed active and inactive by statistical parametric mapping
+%(SPM). This served the purpose of adjusting the priors as well as the 
+%preprocessing based on real world signals. This was actually done before
+%the simulations, and then results were carried back the simulations 
+%to check consistency.
+%After work adjusting parameters, most importantly the weighting function and the 
+%priors, particle filter was applied to every voxel in an FMRI image.
+%The results of this large scale analysis was a parameter map which was
+%then used to calculate normalized square-root MSE image. 
+%
+%\section{Single-Voxel Analysis}
+%The choice of a prior, as discussed previously, is extremely important. While a
+%prior may have the potential to give good results, being a monte-carlo algorithm
+%there is the possibility for inconsistencies. Thus, increasing the variance
+%of the time-constants may allow additional flexibility, it will also cause
+%additional model variance. Before running on a full volume I adjusted the 
+%prior to ensure that the same input would give the same output 100 times in a 
+%row. While this may seem like a given, with a random drawing of the prior,
+%this can be difficult. Case in point, the exact same algorithm run twice
+%with the time constants all having standard deviations of $.35$ resulted in two
+%very different fits, shown in \autoref{fig:badfit_param1}.
+%
+%\begin{figure}
+%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/badfit_param1}}
+%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/goodfit_param1}}
+%\caption{The same priors gave rise to both fits.}
+%\label{fig:badfit_param1}
+%\end{figure}
+%
+%For this reason, I actually lowered the standard deviations of the time
+%constants to prevent over-smoothing. This resulted in more consistent,
+%though potentially slightly worse fits, two examples of which are 
+%shown in \autoref{fig:param2_var}. 
+%
+%\begin{figure}
+%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/param2a}}
+%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/param2b}}
+%\caption{A poor fit, using the same parameters as }
+%\label{fig:param2_var}
+%\end{figure}
+%
+%todo: stats of the 100 fits?
+%
+%\section{Single Time-Series Simulation}
+%
+\chapter{Real Data}
+\label{sec:RealData}
+Modeling the BOLD response is of course not of much use if it is only done in a 
+single voxel. Although this algorithm will hopefully lead to more novel methods 
+of analysis, the standard use for modeling the BOLD signal is to locate "activation".
+By activation I really mean areas where the input correlates with the BOLD response, 
+thus indicating that the stimuli is driving some sort of neural activity. Of course 
+much more is going on behind the scenes, brain regions are stimulating other brain
+regions, but before that can be analyzed we must validate the ability of the particle
+filter to be able to accurately find and model activation. Once areas where the BOLD 
+model may be accurately estimated are found, integrating the model will allow for accurate
+simulation between measurements, which then may be used for more advanced analysis.
+Again though, it all begins with localizing the first activation regions in the
+chain. Therefore here I compare the output of the particle filter based modeling
+solution to the conventional SPM method.
 
-\section{Single-Voxel Analysis}
-This section discusses the results when the particle filter was
-applied on a single voxel. The parameters are the same as
-those used later for entire image analysis; however, the results
-are more in-depth.
+In reality, the SPM8 method is a different animal from the algorithm described here.
+First SPM preprocesses the image by spatially smoothing the FMRI image (here with 
+an $8mm x 8mm x 8mm$ Gaussian kernel), whereas
+this is not done in the particle filter algorithm. Additionally, a spline
+is used to de-trend, rather than SPM8's high pass filter with a cut
+off based on a globally estimated autocorrelation. Thus the preprocessing pipelines 
+are different, and the output of SPM8 is single t-statistic, whereas the output 
+of the particle filter is a posterior probability distribution of the parameters
+at every single voxel. To validate the quality of the particle filter results though
+it is necessary to compare the location of
+"activated" voxels, and the goodness of fit provided by each method.
 
-The run-time for a single voxel depends on the several factors. First, the
-overall length of the signal being analyzed. For 1000 measurements it takes
-about 10 to 15 minutes. On the other hand, in real circumstances the
-length is only around 150 measurements. The number of  integration
-can certainly make a large difference, however dropping below 1000 (.001 seconds)
-is definitely not recommended.
+The results from SPM8 are shown in \autoref{fig:hm_canon_spm}, and the results from 
+the particle filter are shown in \autoref{hm_canon_pfilter}.
 
-For a period I considered 1000 to be a
-fine number; however when generating simulated data I found that every once
-in a while 1000 was not enough. This is problematic in the actual particle
-filter since, given the large number of simultaneous integrations taking 
-place, its likely that a few particles will fail and be weighted at 0 because
-of this problem. Additionally, because the typical case where a failure would
-occur is at fast moving times/parameters the particles all tend to fail together.
-The result is particle deprivation - no particles with non-zero weights remain.
-The other possible outcome is that low time constant particles get pruned resulting
-in excessively smoothed estimates for the time series'. Its possible that a
-kind of stop-gap measure could be put into place; wherein particles that are
-about to be set to NaN are integrated again with finer grained steps. However
-many times the non-real results don't occur until several time steps after the 
-numbers get strange. So for instance, the timestep was too long, allowing 
-$f$ to go negative, resulting in extremely large values of $q$. There are many
-different ways where this sort of event can occur, and unfortunately sometimes
-there is no way to get back to before the state starting going out of control.
-
-Another crucial factor for run time is how long before the first re-sampling 
-occurs. Because the prior is represented initially with significantly more
-particles, if the model fits very well, or for some reason the effective
-number of particles stays high, resampling could take a long time to occur.
-When this happens the particles filter can take a factor of 10 longer to run.
-However, if the particle count isn't initially set high, there is a much larger
-chance of particle deprivation occurring. Since there is no real way to know
-how long it will take to resmaple the first time, there is little the 
-algorithm can do to fix this (except perhaps forcing resampling after
-some period of time). On the other end of the spectrum, if the time
-series doesn't match the model at all, particle deprivation will occur extremely
-quickly.  The upshot of this is that the particle filter is able to 
-identify these sections very quickly, and thus not waste much time there.
-The difficulty though, is the regions in between the perfect fit and the
-awful fit. If particle deprivation does occur, did it occur randomly to an
-activated region or did it occur inevitably because the region doesn't fit.
-The reason for having so many initial particles is to give density to the
-distribution to make false negatives more rare. Perhaps the correct method
-is to still set the initial particles very high, but if for a long time no
-resampling occurs, halve the standard deviation of the measurement error.
-Thus, if the weighting function is not discriminating enough, force it 
-to be more picky about the results. Of course this could lead to particle
-deprivation as well if the standard deviation is brought down too quickly.
-Of course, if a fat-tailed distribution is used for the weighting function,
-or the standard deviation of a Gaussian weighting function is sufficiently large,
-the particle filter will simply converge to meaningless values. The question
-of whether
-
-\begin{figure}
-\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/inactive_illogical}
-\caption{Particle Filter converging to values that make little sense,
-because the voxel did not correlate with the input in any known way}
-\label{fig:badfit}
-\end{figure}
-
-\begin{figure}
-\includegraphics[clip=true,trim=10cm 4cm 10cm 4cm, width=16cm]{images/active_difficult}
-\caption{Example of a timeseries that laplace performs better on than
-gaussian. The large spike at the beginning can quickly eliminate all particles} 
-\label{fig:highhard}
-
-\end{figure}
-
-For this reason, it is important to threshold based on the MSE, and the
-impulse response. Naturally the MSE is intended to catch very voxels with a
-very poor fit,
-but like \autoref{fig:badfit}, the algorithm may have found a line
-down the middle that gives a decent mean squared error. In this case
-the excessive smoothing and low actual activation level will easily
-be distinguished by finding the peak response to single short pulse.
-
-The choice of a prior, as discussed previously, is extremely important. While a
-prior may have the potential to give good results, being a monte-carlo algorithm
-there is the possibility for inconsistencies. Thus, increasing the variance
-of the time-constants may allow additional flexibility, it will also cause
-additional model variance. Case in point, the exact same algorithm run twice
-with standard deviations of $.35, .35, .35$ for the time constants resulted in two
-very different fits, see \autoref{fig:param1_var}.
-
-\begin{figure}
-\subfigure[]{\includegraphics{images/badfit_param1}}
-\subfigure[]{\includegraphics{images/goodfit_param1}}
-\caption{The same priors gave rise to both fits. Params: ($\tau_0 \alpha E_0 V_0 \tau_f \tau_s \epsilon$)
-for the good fit: $(1.40 .32 0.34 0.04  0.96 2.76)$, and for the poor fit:
-$(2.28 0.32 0.33 0.017 0.67 2.88 1.36)$.}
-\label{fig:badfit_param1}
-\end{figure}
-
-For this reason, I actually lowered the standard deviattions of the time
-constants to prevent over-smoothing. This resulted in the more consistent, but
-slightly worse results seen in \autoref{fig:param2_var}
-
-\begin{figure}
-\subfigure[]{\includegraphics{images/param2a}}
-\subfigure[]{\includegraphics{images/param2b}}
-\caption{A poor fit, using the same parameters as }
-\label{fig:param2_var}
-\end{figure}
-
-\section{Particle Collapse Recovery}
-This heuristic was initially intended to deal with cases where the particle
-filter prematurely converged; and to reverse this effect. In reality, however
-it did not perform well, and so it was not used in full-brain analysis. Ultimately
-it did not prevent collapse of the distribution, and it often caused dramatic
-increased in the variance of parameter estimates. It also pronged the analysis
-time of inactive time series. Essentially the effect was exactly opposite of what
-one would hope.
-
-\section{Single Time-Series Simulation}
-
-Graphs:
-
-For simulated data, single timeseries:
-
-For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
-\{100, 500, 1000\} particles
-\begin{enumerate}
-\item Ground truth vs. Estimated signal during particle filter run
-\item Ground truth vs. Estimated signal with final parameter set
-\item True Parameters vs. Final Parameter Sets
-\item Variance of final parameters when faced with same ground truth, different noise
-\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
-\item Estimator Variance based on different noise runs
-\item Final Particle Distribution
-\end{enumerate}
-
-For Simulated Data, Full Volume:
-
-%note to self, epsilon should probably be uniform between 0 and something
-\section{Simulated Volume}
-
-\chapter{Real Data}
 \begin{figure}
@@
 \end{figure}
-
-image comparing epsilon-map with GLM activation map

### 2010-07-19 doc/thesis/background.tex
more work....
mostly done, just a bit of cleanup

@@
 to stimuli. The method of finding activation is discussed in 
-\autoref{sec:Statistical Parameteric Mapping}. The goal of this 
+\autoref{sec:Statistical Parametric Mapping}. The goal of this 
 work is to move away from the question "Is this region active"
@@
 attempts to model the BOLD response; these works will be discussed
-in [all other sections basically].
+in in the next few sections.
 
 \section{Statistical Parametric Mapping}
+\label{sec:Statistical Parametric Mapping}
 Although not strictly the same as parameter calculation from 
@@
 \subsection{Conclusion}
+\label{sec:BackgroundConclusion}
 In all, the GLM is extremely useful for determining linear 
@@
 
-[simulated annealing image?]
+%[simulated annealing image?] todo
 
@@
 
-[genetic algorithm picture]
+%[genetic algorithm picture] todo
 
@@
 \begin{figure}
-\caption{Examples of update distributions, using an Kalman Filter, \cite{Thrun2005}}
+\includegraphics[width=16cm]{images/kalman}
+%\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
+\caption{Example updates of a distribution using Kalman Filter, \cite{Thrun2005}}
 \label{fig:EKFWorking}

### 2010-07-19 doc/thesis/conclusion.tex
mostly done, just a bit of cleanup

@@
 \chapter{Conclusion}
+In this paper I have proposed and demonstrated the use of a particle filter
+for the estimation of BOLD model parameters. In doing so, it has been possible
+to estimate the BOLD time-series and judge the quality of the estimated time-
+series with respect to the true BOLD signal. As result, it is possible to use
+the particle filter to localize areas where a known stimuli most directly drives
+neural activation. In the future it will be possible to use the estimated states
+($v,q,s,f$) to drive other models and learn more about how regions of the brain
+interact.
 
-\section{Future Work}
-*A more extensive investigation of the parameters distribution
-*Using a compliance model
-*restrict parameters more (may not be possible)
+A limitation often reached in previous works was an inconsistency of
+parameter estimates, most likely because of covariance between the model parameters.
+Although individual studies got consistent results, those results often differed
+widely from other similar studies. The reason for this is rather clear from the
+simulation results in \autoref{sec:SimLowNoise}. There is a significant amount of
+trade off between parameters to the point that a signal set of parameters
+is most likely not possible to derive from the BOLD response alone. It
+will therefore be beneficial to combine BOLD studies with cerebral blood
+flow or cerebral blood volume studies to gain multiple more measurements and
+further constrain the model. That said, the benefit of the particle filter
+is that it provides a full posterior distribution at the final time step.
+As such the true solution should be encoded in the particle filter's final 
+distribution given priors that encompass the true parameters. This is beneficial
+in two ways; first, if, after the fact, some parameter becomes known 
+from outside observation, it is then possible to construct a new probability
+conditional on the new observation. Secondly the results from multiple runs
+may be reasonably compiled together, using the final distribution from the previous
+run as the prior distribution of the next run. Rather than simply providing a 
+staring point for parameters to converge from, it in fact continues convergence
+from the previous stopping point.
+
+Although many versions of the BOLD model exist, and it is tempting to use 
+more detailed models; from the results found here the issue of bias error
+from the BOLD model is not the biggest concern. Clearly reducing variance
+in parameters should be the first concern. 
+One future study that would be extremely beneficial in this way  would
+be an extensive study of what the priors should truly be. Although \cite{Friston2000}
+gives an estimate of what is thought to be "reasonable" values, and later studies
+published their estimate of the distributions, given the interplay between
+parameters it is very unlikely that these priors are true to actual distribution
+that occurs in vivo. As I mentioned previously, the addition of simultaneous
+flow or volume measurements are another potentially powerful way to further confine the model,
+and thus deal with the elasticity of parameters. As I mentioned in 
+\autoref{sec:BackgroundConclusion}, a chief advantage of using physiologically
+plausible models is that such data may in fact be added with relative ease.
+
+In conclusion, using particle filters to estimate the BOLD response are 
+an extremely powerful method of fitting to noisy data. The technique also
+holds great promise as extensible platform to build more advanced models
+and techniques on top of. Integrating further information is necessary to
+move beyond the traditional Statistical Parametric Mapping and moving
+toward biologically and medically relevant FMRI scanning techniques.

### 2010-07-19 doc/thesis/intro.tex
added quite a few plots. Should be able to write results for single voxel simulations.
mostly done, just a bit of cleanup

@@
 \label{sec:Introduction Noise}
-As demonstrated in \autoref{sec:BOLD Phsyiology} the BOLD response has been
+As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
 extensively studied and despite minor discrepancies, the cause of the BOLD 
@@
 for analyzing the noise distribution. The locations were chosen from
-various points around the brain, all in grey matter voxels (see \autoref{sec:methods}
-for a discussion on how grey matter voxels were found). The time
+various points around the brain, all in grey matter voxels. The time
 series were also chosen because they were representative of different types
@@
 any two samples is Normally distributed. The simulations discussed in 
-\autoref{sec:Methods Single Voxel Simulation} make use of this, 
+\autoref{sec:Single Voxel Simulation} make use of this, 
 by adding a Wiener random process to the overall signal. To determine
@@
 \hline 
-Parameter  & \cite{Friston2002} & \cite{Johnston2008} & \cite{Vakorin2007} & \cite{Deneux2006}\\
+Parameter  & \cite{Friston2000} & \cite{Johnston2008} & \cite{Vakorin2007} & \cite{Deneux2006}\\
 \hline
-$\epsilon$ &  $N(.54 , .1 ^2)$  & $.069 \pm .014 $ & (NC) & .89\\
-$\tau_s  $ &  $N(1.54, .25^2)$  & $4.98 \pm 1.07 $ & $2.2$ & 2.04 \\
-$\tau_f  $ &  $N(2.46, .25^2)$  & $8.31 \pm 1.51 $ & $.45$ & 5.26\\
 $\tau_0  $ &  $N(.98 , .25^2)$  & $8.38 \pm 1.5  $ & $.94$ & .27\\
@@
 $V_0     $ &  $.03$ (NC)        & $.0149 \pm .006$ & (NC) & .16\\
+$\tau_s  $ &  $N(1.54, .25^2)$  & $4.98 \pm 1.07 $ & $2.2$ & 2.04 \\
+$\tau_f  $ &  $N(2.46, .25^2)$  & $8.31 \pm 1.51 $ & $.45$ & 5.26\\
+$\epsilon$ &  $N(.54 , .1 ^2)$  & $.069 \pm .014 $ & (NC) & .89\\
 \hline

### 2010-07-19 doc/thesis/resultsB.tex
mostly done, just a bit of cleanup

@@
 of analysis, the standard use for modeling the BOLD signal is to locate "activation".
-By activation I really mean areas where the input correlates with the BOLD response, 
-thus indicating that the stimuli is driving some sort of neural activity. Of course 
-much more is going on behind the scenes, brain regions are stimulating other brain
-regions, but before that can be analyzed we must validate the ability of the particle
-filter to be able to accurately find and model activation. Once areas where the BOLD 
+Activation is essentially defined as areas where the input seems to directly drive
+the BOLD response, as opposed to there being intermediate factors controlling it. 
+Once areas where the BOLD 
 model may be accurately estimated are found, integrating the model will allow for accurate
-simulation between measurements, which then may be used for more advanced analysis.
-Again though, it all begins with localizing the first activation regions in the
-chain. Therefore here I compare the output of the particle filter based modeling
-solution to the conventional SPM method.
+estimation of the state between measurements, which could then be used for more 
+advanced analysis, for instance of areas that are being driven by other brain regions.
+It all begins with localizing the first activation regions in the
+chain. Therefore this section compares the output of the particle filter 
+with conventional SPM.
 
-In reality, the SPM8 method is a different animal from the algorithm described here.
-First SPM preprocesses the image by spatially smoothing the FMRI image (here with 
-an $8mm x 8mm x 8mm$ Gaussian kernel), whereas
+In reality, the statistical parametric mapping method is very different from the 
+particle filter algorithm described in this work.
+SPM preprocesses the image by spatially smoothing the FMRI image (in this section 
+SPM8 was used with an $8mm x 8mm x 8mm$ Gaussian kernel), whereas
 this is not done in the particle filter algorithm. Additionally, a spline
-is used to de-trend, rather than SPM8's high pass filter with a cut
-off based on a globally estimated autocorrelation. Thus the preprocessing pipelines 
-are different, and the output of SPM8 is single t-statistic, whereas the output 
+was used to de-trend, rather than SPM8's high pass filter (with a cut
+off based on a globally estimated autocorrelation). Thus the preprocessing pipelines 
+are different; but the output of SPM8 is also different. Whereas SPM outputs
+a t-statistic for each voxel, the output 
 of the particle filter is a posterior probability distribution of the parameters
-at every single voxel. To validate the quality of the particle filter results though
-it is necessary to compare the location of
-"activated" voxels, and the goodness of fit provided by each method.
+at every voxel. To validate the quality of the particle filter results though
+it is necessary to compare the location and fit of the particle filter parameter
+mapping with SPM's T-score map.
+
+\section{Results}
 
 The results from SPM8 are shown in \autoref{fig:hm_canon_spm}, and the results from 
-the particle filter are shown in \autoref{hm_canon_pfilter}.
+the particle filter are shown in \autoref{hm_canon_pfilter}. Note again that the
+output is actually a different scale. Whereas the levels in the SPM map are T-Scores
+indicating the probability that the result cannot be explained from a random signal,
+the result of the particle filter is simply a normalized square-root of the mean-squared-
+residual. Thus, lower is better, and zero is a perfect fit in the particle filter whereas
+infinity is a perfect fit in \autoref{fig:hm_canon_spm}. Note that it would seem that there
+are quite a few dispersed areas with decent fits in \autoref{fig:hm_canon_pfilter}. Also,
+interestingly the lowest square-root MSE (SMSE) was just above $.7$. 
+For the sake of completeness \autoref{fig:hm_canon_pfilter85} shows the same image but
+thresheld at $.85$ rather than $1$. Obviously this image seems like a closer match to
+the results of SPM, however, the results of \autoref{fig:hm_canon_pfilter} are interesting.
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
@@
 \subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
-\caption{}
+\caption{Section 1, Estimated vs. Actual BOLD response}
 \label{fig:comp1}
@@
 \subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
-\caption{}
+\caption{Section 2, Estimated vs. Actual BOLD response}
 \label{fig:comp2}
 \end{figure}
+
+I chose several voxels to discuss further from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
+The first voxel, labeled 1, had a very high T-score as well as a very low SMSE (around $.7$). Thus, the
+fit should be very good in both the SPM and particle filter output; this comparison is shown in 
+\autoref{fig:comp1}. Recall that SPM worked on a slightly less noisy time series because of the
+spatial smoothing; this explains the lack of the sharp peak in the particle filter's preprocessed
+data. Regardless, as expected, both fits are very good.
+
+I chose the second voxel (\autoref{fig:comp2}) because it was active in SPM and it would appear to be in a prime
+location to be active in the SPM image (given the results in the surrounding voxels). The fit,
+however shows just why the voxel did not have a particularly good SMSE. Although certain peaks 
+seem to fit rather well the input doesn't seem to correlate that well with the BOLD response. For
+instance 75 seconds in, the stimulus is clearly not present, and so the signal should drop off; yet
+it doesn't. This is a good example of a voxel that seems to be reacting to something in addition to
+the direct stimulus. Even in what should be a rather smooth time-series in SPM, the BOLD signal
+looks very noisy.
+
+The third voxel, compared in \autoref{fig:comp3}, was far away from any other active voxels 
+and yet had a very low (around $.7$) SMSE.
+In both preprocessed time-series the input is extremely noisy, yet by the normalized SMSE
+the response is good. This could be an example of a false positive, although the general trend
+of the response seems to follow the estimated response.
+
+Whereas the third voxel was a region that was considered "active" by the particle filter
+algorithm but not SPM; the fourth voxel is the reverse. The particle filter did not
+find a good fit, but in SPM8 the voxel had a decently high T-score. \autoref{fig:comp4} shows
+the line fit for each algorithm, and neither are particularly good; though in a few areas
+the upward or downward movements seem to be correlated. Regardless, this would seem to be
+a false positive from SPM8.
+
+The fifth, and last single voxel to be analyzed is one more area that had a low SMSE according
+to the particle filter, and yet did not show up in SPM. The line fit is shown in \autoref{fig:comp5},
+and unlike voxel 3 is certainly \emph{not} a false positive. Regions such as this are a very 
+good example of why spatial smoothing can be problematic. By having a more adaptive model, and
+thus reducing bias error, it is possible to get a very good fit to single voxels such as this,
+and thus the need for spatial smoothing is somewhat reduced.
 
@@
 \subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
-\caption{}
+\caption{Section 3, Estimated vs. Actual BOLD response}
 \label{fig:comp3}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
 \subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
-\caption{}
+\caption{Section 4, Estimated vs. Actual BOLD response}
 \label{fig:comp4}
@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{}
+\caption{Section 5, Estimated vs. Actual BOLD response}
 \label{fig:comp5}
 \end{figure}
+
+\section{Discussion}
+From the maps generated, the similarities to SPM8's results are encouraging. At the 
+very least the output of the particle filter seems to meet the quality of SPM. The normalization
+of the MSE is certainly necessary. Although there is no hard threshold on the 
+normalized SMSE, it would appear that the normalization is providing a reasonable
+ordering of regression quality. The question of what threshold to use is a difficult one.
+The actual division of the Square Root of the MSE by an estimate for the standard deviation
+of the signal would seem to mandate a threshold less than one. However, recall that the 
+same media-absolute-deviation was added to the de-trended signal as a DC-gain. This means
+that a flat, no-response bold signal will actually have a SMSE greater than the standard 
+deviation of the signal. Then again, the SMSE may be reduced for any point where the 
+model spikes up due to stimuli. Ultimately if the model were to run right down the 
+center of the distribution the SMSE would be equal to the MAD, and would have a 
+normalized SMSE of 1. Thus, again, a normalized SMSE (NSMSE) of 1 is a good start for the 
+threshold. The question then, is what the probability of getting NSMSE below some 
+threshold randomly; which would depend on factors such as the amount of stimuli
+present in the signal. For now I would say that .85 seems to be a decent threshold,
+although it may be a bit conservative.
+
+Other measures of fitness could also be used such as mean of the absolute error or 
+mutual information. Although least squares is the best linear unbiased estimator,
+this signal is heavily nonlinear, so other estimators may exceed the OLS in terms
+of likelihood. Regardless, the NSMSE seems to provide a very good solution at this
+point, giving comparable results to SPM.

### 2010-07-19 doc/thesis/background.tex
More clean up.

@@
 \begin{figure}
-\caption{Hemodynamic Response Function todo}
+\centering
+\includegraphics[scale=.7]{images/HRF}
+\caption{Canonical Hemodynamic Response Function}
 \label{fig:HRF}
@@
 As mentioned previously and discussed extensively
-in \cite{Friston2002} and \cite{Hoffman1997}, hierarchical models may be 
+in \cite{Friston2002} and \cite{Hofmann1997}, hierarchical models may be 
 applied to account for systematic differences between subjects. For instance, if the
@@
 be present in similar studies. The Hierarchical form used by
-\cite{Friston2002} is shown in \autoref{eq:Hiearchical}.
+\cite{Friston2002} is shown in \autoref{eq:Hierarchical}.
 
@@
 The Empirical Bayes algorithm is used in both in both \cite{Friston2002} 
-and \cite{Hoffman1997}. As a consequence, point estimators 
+and \cite{Hofmann1997}. As a consequence, point estimators 
 are used for each $\theta$, rather than the full distributions.
 
-\subsection{Conclusion}
+\subsection{Discussion}
 \label{sec:BackgroundConclusion}
@@
 data. Combined FMRI CBF or CBV imaging methods are rapidly getting better,
-as seen in \cite{Chan2009}. These techniques could shed light on
+as seen in \cite{Chen2009}. These techniques could shed light on
 neural activation by providing extra measurements, yet a 
@@
 each time the energy or fitness needs to be calculated, a large number of cycles
-must be spent re-simulating the BOLD model for the set of parameters. As we'll
-discuss in \autoref{sec:Particle Filters}, the Particle Filter method is able
+must be spent re-simulating the BOLD model for the set of parameters. As I'll
+discuss in \autoref{sec:Particle Filter}, the Particle Filter method is able
 to circumvent this re-calculation to some degree.
@@
 those methods contained random elements designed to overcome this issue.
-[results?]
 
@@
 While of course this could be poor convergence of the algorithm, there is another other possibility.
-Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Least Squares},
+Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Nonlinear Least Squares},
 the algorithm described in \cite{Johnston2008} does not depend on prior distributions.
@@
 While \cite{Johnston2008} is certainly in the minority; further exhaustive studies 
-of the parameters, using unbiased techniques may be called for.
-
-[graph of friston parameters vs. johnston]
-
-[might want to mention dynamic causal modeling]
+of the parameters, using unbiased techniques may be called for. A further 
+comparison between the distributions found in \cite{Johnston2008} and 
+\cite{Friston2000} will be discussed in \autoref{sec:PriorDist}, including
+graphs: \autoref{fig:MeanResponseF}, \autoref{fig:MeanResponseJ}.
 

### 2010-07-19 doc/thesis/intro.tex
More clean up.

@@
 research into the post-stimulus undershoot has led to the creation
-of much more in depth models. In \cite{Zhen2002} additional state
+of much more in depth models. In \cite{Zheng2002} additional state
 variables model oxygen transport, whereas \cite{Buxton2004} models
@@
 \end{table}
-
-Necessary?
-[Image with two different $\alpha$'s]
-[image comparing the results of 10\% changes in various signals]

### 2010-07-19 doc/thesis/methods.tex
More clean up.

@@
 from \cite{Friston2000}, the signal response for a $.1$ second
-impulse only reaches maybe half a percent, as \autoref{fig:MeanResponse}
+impulse only reaches maybe half a percent, as \autoref{fig:MeanResponseF}
 shows.
@@
 \centering
-\label{fig:MeanResponse}
+\label{fig:MeanResponseF}
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response}
@@
 \centering
-\label{fig:MeanResponse}
+\label{fig:MeanResponseJ}
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response_johnston}
@@
 \section{Choosing $P(y_k | x_k)$}
+\label{sec:Methods Weighting Function}
 Choosing a representation of an unknown distribution tricky, and while
@@
 if the distribution is made overly broad, it will still distinguish between
-particles that are near the origin. Results with real data compare 
-the three distribution in \autoref{sec:Results Weights}.
+particles that are near the origin.
 
@@
 to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
-in \autoref{fig:QQDDelta} in \autoref{sec:Introduction Noise}. 
+in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
  All the assumptions made originally

### 2010-07-19 doc/thesis/results.tex
mostly done, just a bit of cleanup
need to go back and add some stuff to the slice sim

@@
 \section{Single Time Series}
+\label{sec:Single Voxel Simulation}
 Given the state-space equations for the BOLD signal, simulating a single time 
@@
 \subsection{Signal with Low Noise}
+\label{sec:SimLowNoise}
 %tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
@@
 this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
+Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
+program the true signal-to-noise ratio was much lower, as soon in the fit in \autoref{fig:simslicefit1}.
+
+\begin{figure}
+%includegraphics[]{images/simslicefit1}
+\caption{A single time-series fit in the simulated slice.}
+\labe{fig:simslicefit1}
+\end{figure}
+
 For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
@@
 simulated data; additionally a new simulation using the calculated parameters may also be 
-generated to test the difference in activation levels between the real parameters and the
+generated to test the difference in BOLD levels between the real parameters and the
 estimated ones. Since it is clear that the parameters are not fully orthogonal 
 (\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
-but have different parameters. This way, an absolute 
-quantitative difference between the two parameter sets may be found.
+yet different. This way, a quantitative difference between the two parameter sets 
+may be found.
+
+\begin{figure}
+\includegraphics[]{images/simslice_hm}
+\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
+\label{fig:simslice_hm}
+\end{figure}
+
+\begin{figure}
+\includegraphics[]{images/simslice_hm85}
+\caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
+Darker indicates lower error.}
+\label{fig:simslice_hm85}
+\end{figure}
+
+\begin{figure}
+\includegraphics[]{images/parameters_sec1}
+\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
+Limited to areas with residual less than $.85$.}
+\label{fig:parameters_sec1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[]{images/parameters_sec2}
+\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
+Limited to areas with residual less than $.85$.}
+\label{fig:parameters_sec2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[]{images/parameters_sec3}
+\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$. 
+Limited to areas with residual less than $.85$.}
+\label{fig:parameters_sec3}
+\end{figure}

### 2010-07-19 doc/thesis/results.tex
first draft done

@@
 Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
-program the true signal-to-noise ratio was much lower, as soon in the fit in \autoref{fig:simslicefit1}.
+program the true signal-to-noise ratio was much lower, as seen in the fit in \autoref{fig:simslicefit1}.
 
-\begin{figure}
-%includegraphics[]{images/simslicefit1}
+\begin{figure}[H]
+\centering
+\includegraphics[width=15cm]{images/simslice_noise}
 \caption{A single time-series fit in the simulated slice.}
-\labe{fig:simslicefit1}
+\label{fig:simslicefit1}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[]{images/simslice_hm}
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{images/simslice_mask}
+\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
+\label{fig:simslice_mask}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{images/simslice_hm}
 \caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
@@
 
-\begin{figure}
-\includegraphics[]{images/simslice_hm85}
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{images/simslice_hm85}
 \caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
@@
 
-\begin{figure}
-\includegraphics[]{images/parameters_sec1}
-\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
-Limited to areas with residual less than $.85$.}
-\label{fig:parameters_sec1}
+The regions are numbered according to \autoref{fig:simslice_mask}; the parameters
+for each region may be found in \autoref{tab:simsliceparams}.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c |c | c | c | c | c | c | c |}
+\hline 
+Region & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  \\
+\hline 
+1 & 1.454& .321& .369& .036& .994& 2.774& 1.348\\
+2 &1.151& .353& .380& .026& 1.98& 2.333& 1.645 \\
+3 &1.951& .317& .348& .027& 1.657& 3.719& .757 \\
+4 &1.203 & .310& .326& .036& 2.168& 2.272& .086\\
+\hline
+\end{tabular}
+\caption{Actual parameters for each regions in the simulated slice.}
+\label{tab:simsliceparams} 
+\end{table}
+%last (bottom right), 1.203, .310, .326, .036, 2.168, 2.272, .086, 3.4, 1
+
+Note that region 4 has a very low $\epsilon$. For this reason, the only areas
+with significantly decent estimates of the BOLD time series are 1,2 and 3. With
+such a low $\epsilon$ region 4 was below the noise threshold. Notice that the
+regions 1 2 an 3 clearly stick out in the $\sqrt{MSR}$ map, indicating that the
+particle filter was succesfull in matching those regions.
+
+\begin{figure} %bottom left
+\centering
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist1}
+\caption{Histogram of estimated parameters in section 1 with $\sqrt{MSR}$ less than $.85$.}
+\label{fig:slicesim_hist1}
 \end{figure}
 
-\begin{figure}
-\includegraphics[]{images/parameters_sec2}
-\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
-Limited to areas with residual less than $.85$.}
-\label{fig:parameters_sec2}
+\begin{figure} %top left
+\centering
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
+\caption{Histogram of estimated parameters in section 2 with $\sqrt{MSR}$ less than $.85$.}
+\label{fig:slicesim_hist2}
 \end{figure}
 
-\begin{figure}
-\includegraphics[]{images/parameters_sec3}
-\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$. 
-Limited to areas with residual less than $.85$.}
-\label{fig:parameters_sec3}
+\begin{figure} %top right
+\centering
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
+\caption{Histogram of estimated parameters in section 3 with  $\sqrt{MSR}$ less than $.85$.}
+\label{fig:slicesim_hist3}
 \end{figure}
+
+It is again clear that a single point estimate of the parameters is not feasible,
+but rather a distribution of parameters is the best any modeling algorithm
+is going to be capable of for this version of the BOLD signal. In spite of
+this, and relatively heavy noise, the fact that is that the fits were
+good. Clearly, from the normalized $\sqrt{MSR}$ it is possible to identify
+ \"active" regions; or regions that respond the way the BOLD model expects.
+Although some of the fits in the  \"active" regions did not have good regression,
+this is not necessarily the work of false negatives. Instead, because possum
+simulates different tissues, it is more likeley that those regions non
+grey-matter areas. Essentially these areas should not have reacted very strongly
+because in the real world such regions wouldn't. By the fact that it 
+was relatively easy to identify the regions of activation from the simulated 
+results, something similar should happen with real FMRI data, which is discussed
+in the next session. which

### 2010-07-20 doc/thesis/intro.tex
additions to intro

@@
 \chapter{Introduction}
+Traditional methods of analyzing images produced by Function Magnetic Resonance 
+Imaging (FMRI) use regression techniques with very few degrees of freedom. 
+Though adding more degrees of freedom naturally mandates more computation,
+in this thesis I will discuss a Sequential Monte Carlo method of fitting a nonlinear
+model with seven degrees of freedom at 
+a computation cost that would allow real time calculations for a multiple voxels
+(4 dimensional pixels). More practically, the method described here forms a completely
+separate method of detecting neural activity in comparison to the traditionally
+used Statistical Parametric Mapping (SPM). Though more computationally intense,
+this method is capable of modeling nonlinear effects, is conceptually simpler
+and provides more detailed output. Additionally,
+by using a separate particle filter for each voxel's time
+series it should be feasible to calculate parameters and make real-time predictions
+for small neural regions, a feature which could be useful towards real time FMRI 
+(\cite{DeCharms2005}).
+
 For the past twenty years Functional Magnetic Resonance Imaging (FMRI) 
@@
 of analyzing neural signals can be invasive or difficult to acquire, 
-FMRI is relatively quick and cheap, and its analysis relatively straight forward.
-
+FMRI is quick and cheap, and its analysis straight forward.
 By modeling the governing equations behind the neural response that
-drives FMRI, it is possible to make the analyses of FMRI even more
-powerful. The underlying state equations hold important information
+drives FMRI, it is possible to increase the power of FMRI.
+The underlying state equations hold important information
 about how individual brain regions react to stimuli. The model parameters
@@
 physiology including existing and future pathologies. In short,
-the long chain of events driving observable FMRI signal contains more
-information than which regions respond to which stimuli.
+the long chain of events driving FMRI signals contain information 
+beyond correlation with stimuli.
 
@@
 time evolution of the BOLD signal using a windkessel model to
-describe the time local changes in Deoxygenated Hemoglobin content.
+describe the local changes in Deoxygenated Hemoglobin content.
 Various papers made substantial improvements to this model until
@@
 set of equations. And while there have been numerous adaptations in the model, 
-many of them summarized in \cite{Deneux2006}, even the most basic version
-has been shown to have less bias error than the convolution based 
-"Canonical Hemodynamic Model" \cite{Deneux2006}, \cite{Handwerker2004}.
-On the other hand many of the BOLD signal models have far more parameters
-than a simple scaling parameter. In fact, the number of parameters
-range from seven \cite{Riera2004} to 50 \cite{Behzadi2005} per
-signal time course; a signal which may be as short as 100 samples long.
-Thus, even in a small FMRI image of the brain (20x20x20), the number of parameters 
-could easily exceed 10,000. Clearly this number of parameters presents
-a significant risk of being under-determined and could suffer catastrophic
-variance error; to make no mention of the computation cost. 
-In this work a method of countering these problem is presented
-with the use of a particle filter.
+many of them summarized in \cite{Deneux2006}, even the basic versions
+have less bias error than the empirically driven \emph{Canonical Hemodynamic Model}
+\cite{Deneux2006}, \cite{Handwerker2004}.
+On the other hand many of the BOLD signal models have numbers
+of parameters ranging from seven \cite{Riera2004} to 50 \cite{Behzadi2005} 
+for a signal as short as 100 samples long. Clearly this number of parameters presents
+a significant risk of being under-determined and having high computation cost. 
+In this work, only the simplest physiologically inspired model will be
+used (with 7 parameters), and steps will be taken to make the most of computation
+time.
 

### 2010-07-21 doc/thesis/background.tex
corrected the intro, and moved the noise description out to methods

@@
-\chapter{Prior Works}
+\chapter{Methods of Analyzing FMRI}
 \label{sec:Prior Works}
@@
 The most basic method of analyzing FMRI data is through a standard T-test
-between "resting state" and "active state" samples. Simply put, the
+between resting state and active state samples. Simply put, the
 mean is calculated separately for non-stimulus and stimulus time intervals.
@@
 with a Hemodynamic 
-Response Function (HRF). The so called "Hemodynamic Response Function"
+Response Function (HRF). The \emph{Hemodynamic Response Function}
 mimics the basic shape of BOLD activation, including a delay
@@
 It is well known that different Hemodynamic Response Functions are necessary 
-for different regions of the brain. The "Canonical" HRF that is most used, 
+for different regions of the brain. The Canonical HRF that is most used, 
 has been optimized for the visual cortex. As \autoref{sec:Post Stimulus Undershoot}
@@
 and then each set of parameters is rated with a fitness function. Parameter
-sets that are good get a higher "fitness"; then new parameter sets are generated by 
+sets that are good get a higher weight; then new parameter sets are generated by 
 randomly combining pieces of the old parameter sets. The pieces are typically
-chosen at a rate proportional to the fitness of the donor; thus "fit"
+chosen at a rate proportional to the fitness of the donor; thus fit
 parameter sets tend to pass on their properties. In addition to this,
@@
 
-In order to estimate the posterior at $t$, a deterministic set of "sigma" points 
+In order to estimate the posterior at $t$, a deterministic set of sigma points 
 (often 2 per dimension, plus 1 at the mode of the multivariate distribution)

### 2010-07-21 doc/thesis/conclusion.tex
corrected the intro, and moved the noise description out to methods

@@
 \chapter{Conclusion}
+\label{sec:Conclusion}
 In this paper I have proposed and demonstrated the use of a particle filter
@@
 conditional on the new observation. Secondly the results from multiple runs
-may be reasonably compiled together, using the final distribution from the previous
+may be reasonably concatenated, using the final distribution from the previous
 run as the prior distribution of the next run. Rather than simply providing a 
@@
 more detailed models; from the results found here the issue of bias error
-from the BOLD model is not the biggest concern. Clearly reducing variance
-in parameters should be the first concern. 
+from the BOLD model is not the biggest concern. Clearly clarifying the distributions
+of parameters for the prior should be the first concern. 
 One future study that would be extremely beneficial in this way  would
 be an extensive study of what the priors should truly be. Although \cite{Friston2000}
-gives an estimate of what is thought to be "reasonable" values, and later studies
+gives an estimate of what is thought to be reasonable values, and later studies
 published their estimate of the distributions, given the interplay between

### 2010-07-21 doc/thesis/intro.tex
corrected the intro, and moved the noise description out to methods

@@
 \chapter{Introduction}
-Traditional methods of analyzing images produced by Function Magnetic Resonance 
+Traditional methods of analyzing timeseries images produced by 
+Function Magnetic Resonance 
 Imaging (FMRI) use regression techniques with very few degrees of freedom. 
@@
 model with seven degrees of freedom at 
-a computation cost that would allow real time calculations for a multiple voxels
-(4 dimensional pixels). More practically, the method described here forms a completely
+a computation cost that would still allow real time calculations for multiple voxels.
+More practically, the method described here forms a completely
 separate method of detecting neural activity in comparison to the traditionally
@@
 by using a separate particle filter for each voxel's time
-series it should be feasible to calculate parameters and make real-time predictions
+series it is possible to estimate parameters and make real-time predictions
 for small neural regions, a feature which could be useful towards real time FMRI 
-(\cite{DeCharms2005}).
+(\cite{DeCharms2005}). Future works will also benefit from the ability to 
+apply conditions to the posterior distribution in post-processing without
+having to re-run the algorithm; for instance in light of additional physiological
+constraints. Modeling the BOLD response as a nonlinear system is the
+best way to determine the correlation of stimulus sequence with the BOLD
+response; yet in the past doing this on a large scale has been far too
+computationally taxing. The solution used here takes approximately 40 seconds
+for a single voxel's time series (5 minutes in length, with a quad core machine).
 
-For the past twenty years Functional Magnetic Resonance Imaging (FMRI) 
+This thesis is organized as follows. In the introduction I will introduce
+FMRI, the method by which neural time changing data is detected. This section
+will also describe the basic form of the BOLD model - which drives the 
+detectable changes in MR signal. \autoref{sec:Prior Works} will discuss other
+methods of analyzing FMRI images as well as other techniques that have
+been, or could be applied to the nonlinear regression model described here. 
+\autoref{sec:Particle Filter} derives the particle filter using Bayesian 
+statistics and discusses some practical elements of implementing the 
+particle filter algorithm. \autoref{sec:Methods} then goes into further
+detail about the specific particle filter configuration used in this work.
+This section also describes the pre-processing used before the particle
+filter is applied. The results are described separately for simulated data
+and real FMRI data in \autoref{sec:SimulationResults} and \autoref{sec:RealData},
+respectively. Finally in \autoref{sec:Conclusion} there is a discussion of
+the usefulness and implications of this technique as well as recommendations
+for direction of future works.
+
+\section{Historic Context}
+For the past twenty years, Functional Magnetic Resonance Imaging (FMRI) 
 has been at the forefront of cognitive research. Despite it's
@@
 significant feature missing from the original model is the 
-"post-stimulus undershoot".
-The "post-stimulus" undershoot is the name for a prolonged subnormal
+post-stimulus undershoot.
+The post-stimulus undershoot is the term used for a prolonged subnormal
 BOLD response for a period of 10 to 60 seconds after stimulus has
@@
 the best fit.
-
-\section{Noise}
-\label{sec:Introduction Noise}
-As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
-extensively studied and despite minor discrepancies, the cause of the BOLD 
-signal is well known. However, as FMRI detects an  
-aggregate signal over the space of cubic centimeters, there are
-plenty of sources of noise. Though local neurons act
-"together" (i.e. around the same time), the density of neurons, the
-density of capillaries, and slight differences in activation across 
-a particular voxel can all lead to signal attenuation and noise.
-
-A particularly difficult form of noise present in FMRI is a low frequency
-drift, often characterized as a Wiener process (\cite{Riera2004}). 
-Though not present in all regions, as much as ten to fifteen percent
-of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
-inference problems \cite{Smith2007}. It is still not
-clear what exactly causes this noise noise comes from, although it is possible it is 
-the result of magnets heating up, or some distortion in magnetic
-fields \cite{Smith2007}. It is clear that this drift signal is not solely
-due to a physiological effects, given its presence in cadavers and phantoms 
-\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
-more prevalent at interfaces between regions, though 
-by no means limited to such areas. Though one potential source
-could be slight movement, given that co-registration of volumes to a single time
-point is standard, this seems unlikely. Regardless, the problem mandates
-the use of a high pass filter to make inference reasonably powered
-\cite{Smith2007}.
-
-In order to characterize the noise, I analyzed resting state data.
-During resting state, the patient is shown no images, and he is asked
-to avoid movement and complex thought, to the best of his abilities.
-Some limitations exist, for instance EPI sequences are extremely loud
-which could cause some stimulus in the auditory cortex, and of course
-very long sequences could give the patient time for his mind to wander.
-Overall though there should be no activation, and thus the signal will
-consist entirely of noise. Therefore resting state data is perfect
-for analyzing the noise distribution. The locations were chosen from
-various points around the brain, all in grey matter voxels. The time
-series were also chosen because they were representative of different types
-of noise I found in the resting state data.
-
-Because most methods (including the one used in this paper)
-assume the noise realizations are independent of each other, the auto-
-correlation is of particular interest (which is a necessary but not
-sufficient condition for independence). Gaussianity is also a common
-assumption made in studies of FMRI data, though that assumption is not
-made in this work. Regardless, comparing the distribution to a Gaussian
-is informative, so I used Q-Q plots to compare example data with the
-Normal distribution. Additionally, in FMRI data the noise is often considered 
-to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
-characterized by steps that are Gaussian; in other words the difference between
-any two samples is Normally distributed. The simulations discussed in 
-\autoref{sec:Single Voxel Simulation} make use of this, 
-by adding a Wiener random process to the overall signal. To determine
-whether the noise is in fact Wiener, the distribution of 
-the difference between adjacent measurements was plotted against
-a Gaussian. Wiener processes further assume that the steps are independent;
-so the steps should have near flat autocorrelation functions.
-
-Finally, removal of the so called "drift" is often
-performed with some variation of a high pass filter, 
-so I checked the noise distribution after applying such a filter (in
-this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
-Here I wanted to know how effective my high pass filter at the removal of
-Wiener noise.
-
-\autoref{fig:QQDC} shows the 
-results with a regression line fit to the points on the Q-Q plot.
-Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
-x-axis and the points
-on the y-axis come from the same \emph{type} of distribution then all the points will
-fall on a line. Differences in the variance will cause the line to have a slope
-other than 1, while differences in the expected value will cause the fitted line
-to be shifted. In these Q-Q plots, the points are being compared to the standard
-Gaussian distribution, so the quality of the line fit determines how closely the points
-fit the Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
-normalized (changed to percent difference).
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_29_49_9}}
-\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_34_43_24}}
-\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_22_38_23}}
-\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_37_29_24}}
-
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}
-
-\caption{Q-Q Plots of normalized resting state data}
-\label{fig:QQDC}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
-\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
-\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
-\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
-\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
-\label{fig:QQDelta}
-\end{figure}
-
-A Wiener process should still conform to the normal distribution, with a 
-variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
-are relatively well described by a Gaussian process with a small autocorrelation, 
-\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
-do not seem to fit the Gaussian well. Also note the significant autocorrelation in
-\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
-Gaussian white noise.  On the other hand, the steps do conform rather
-closely to the normal distribution.
-As expected, most of the autocorrelation disappears for the step data. Given
-that the steps seem to fit the Normal distribution, the low autocorrelation
-indicates that the steps could be Independently Distributed. 
-Therefore, the assumption that the noise follows a Wiener process seems to
-be correct.
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
-\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
-\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
-\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
-\caption{Q-Q Plots of resting state data, after the de-trending}
-\label{fig:QQSpline}
-\end{figure}
-
-De-trending the time-series by subtracting a spline fit to the distribution
-removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
-though clearly not perfectly. Though the distributions still do not exactly fit
-the Normal, \autoref{fig:QQs:D} clearly is much improved compared to \autoref{fig:QQDC:D}.
-In all, the de-trending is effectively removing Wiener noise.
 

### 2010-07-21 doc/thesis/methods.tex
corrected the intro, and moved the noise description out to methods

@@
 \chapter{Methods}
+\label{sec:Methods}
 Although the particle filter I used for this work is a standard Regularized
@@
 \end{tabular}
-\caption{}
+\caption{Prior distributions used in the particle filter.}
 \label{tab:Prior} 
@@
 In the future, it may be possible to set the standard deviation by
-taking a small sample from "resting" data and using the sample standard deviation.
+taking a small sample from resting data and using the sample standard deviation.
 Since this is the first attempt at using particle filters for modeling the 
@@
 
-\subsection{Classical De-trending}
+\section{BOLD Noise}
+\label{sec:Introduction Noise}
+As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
+extensively studied and despite minor discrepancies, the cause of the BOLD 
+signal is well known. However, as FMRI detects an  
+aggregate signal over the space of cubic centimeters, there are
+plenty of sources of noise. Though local neurons act
+together (i.e. around the same time), the density of neurons, the
+density of capillaries, and slight differences in activation across 
+a particular voxel can all lead to signal attenuation and noise.
+
+A particularly difficult form of noise present in FMRI is a low frequency
+drift, often characterized as a Wiener process (\cite{Riera2004}). 
+Though not present in all regions, as much as ten to fifteen percent
+of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
+inference problems \cite{Smith2007}. It is still not
+clear what exactly causes this noise noise comes from, although it is possible it is 
+the result of magnets heating up, or some distortion in magnetic
+fields \cite{Smith2007}. It is clear that this drift signal is not solely
+due to a physiological effects, given its presence in cadavers and phantoms 
+\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
+more prevalent at interfaces between regions, though 
+by no means limited to such areas. Though one potential source
+could be slight movement, given that co-registration of volumes to a single time
+point is standard, this seems unlikely. Regardless, the problem mandates
+the use of a high pass filter to make inference reasonably powered
+\cite{Smith2007}.
+
+In order to characterize the noise, I analyzed resting state data.
+During resting state, the patient is shown no images, and he is asked
+to avoid movement and complex thought, to the best of his abilities.
+Some limitations exist, for instance EPI sequences are extremely loud
+which could cause some stimulus in the auditory cortex, and of course
+very long sequences could give the patient time for his mind to wander.
+Overall though there should be no activation, and thus the signal will
+consist entirely of noise. Therefore resting state data is perfect
+for analyzing the noise distribution. The locations were chosen from
+various points around the brain, all in grey matter voxels. The time
+series were also chosen because they were representative of different types
+of noise I found in the resting state data.
+
+Because most methods (including the one used in this paper)
+assume the noise realizations are independent of each other, the auto-
+correlation is of particular interest (which is a necessary but not
+sufficient condition for independence). Gaussianity is also a common
+assumption made in studies of FMRI data, though that assumption is not
+made in this work. Regardless, comparing the distribution to a Gaussian
+is informative, so I used Q-Q plots to compare example data with the
+Normal distribution. Additionally, in FMRI data the noise is often considered 
+to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
+characterized by steps that are Gaussian; in other words the difference between
+any two samples is Normally distributed. The simulations discussed in 
+\autoref{sec:Single Voxel Simulation} make use of this, 
+by adding a Wiener random process to the overall signal. To determine
+whether the noise is in fact Wiener, the distribution of 
+the difference between adjacent measurements was plotted against
+a Gaussian. Wiener processes further assume that the steps are independent;
+so the steps should have near flat autocorrelation functions.
+
+Finally, removal of the drift is often
+performed with some variation of a high pass filter, 
+so I checked the noise distribution after applying such a filter (in
+this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
+Here I wanted to know how effective my high pass filter at the removal of
+Wiener noise.
+
+\autoref{fig:QQDC} shows the 
+results with a regression line fit to the points on the Q-Q plot.
+Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
+x-axis and the points
+on the y-axis come from the same \emph{type} of distribution then all the points will
+fall on a line. Differences in the variance will cause the line to have a slope
+other than 1, while differences in the expected value will cause the fitted line
+to be shifted. In these Q-Q plots, the points are being compared to the standard
+Gaussian distribution, so the quality of the line fit determines how closely the points
+fit the Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
+normalized (changed to percent difference).
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_29_49_9}}
+\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_34_43_24}}
+\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_22_38_23}}
+\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_37_29_24}}
+
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}
+
+\caption{Q-Q Plots of normalized resting state data}
+\label{fig:QQDC}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
+\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
+\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
+\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
+\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
+\label{fig:QQDelta}
+\end{figure}
+
+A Wiener process should still conform to the normal distribution, with a 
+variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
+are relatively well described by a Gaussian process with a small autocorrelation, 
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
+do not seem to fit the Gaussian well. Also note the significant autocorrelation in
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
+Gaussian white noise.  On the other hand, the steps do conform rather
+closely to the normal distribution.
+As expected, most of the autocorrelation disappears for the step data. Given
+that the steps seem to fit the Normal distribution, the low autocorrelation
+indicates that the steps could be Independently Distributed. 
+Therefore, the assumption that the noise follows a Wiener process seems to
+be correct.
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
+\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
+\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
+\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
+\caption{Q-Q Plots of resting state data, after the de-trending}
+\label{fig:QQSpline}
+\end{figure}
+
+De-trending the time-series by subtracting a spline fit to the distribution
+removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
+though clearly not perfectly. Though the distributions still do not exactly fit
+the Normal, \autoref{fig:QQs:D} clearly is much improved compared to \autoref{fig:QQDC:D}.
+In all, the de-trending is effectively removing Wiener noise.
+
+\subsection{Detrending}
 \label{sec:Detrend}
@@
 during the preprocessing stage, rather than as an integral part of analyzing the BOLD
-signal. There is no shortage of theories on the "best" method of detrending; however
+signal. There is no shortage of detrending methods; however
 in a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
@@
 
-The method I used to calculate the spline was picking one knot for every 15
+The method I used to calculate the spline was picking one knot for every 20
 measurements in an image. Thus a 10 minute session at a repetition time of 
@@
 by definition have a median near zero. 
-The problem with this is this is not the "natural" BOLD signal. More specifically,
-when the signal is inactive, the BOLD response \emph{should} be at 0\% change from
+The problem with this is this is not the natural state of the BOLD signal. More specifically,
+when the signal is inactive, the BOLD response should be at 0\% change from
 the base level; activation may then increase, or for short periods decrease from this base.
@@
 \begin{figure}
-\caption{image of de-spline'd lines with "true" lines}
+\caption{Image of de-spline'd lines with true lines}
 \label{fig:DesplineQuality}
@@
 go around the problem in another way. Here, I propose and test a 
-different method of dealing with the so called "drift". 
+different method of dealing with the drift. 
 Instead of comparing the direct output of the particle filter with the direct
@@
 As discussed in the section on de-trending, the normal pipeline for analyzing
-FMRI involves a great deal of preprocessing. The first and most important
+FMRI involves a several preprocessing steps. The first and most important
 task is motion correction. To do this, a single volume in time is chosen, and
-volumes at every other time are registered to this one volume. This takes care
-of motion by the patient as well as some small changes in the magnetic
-fields that cause the image to shift. After this, a structural image of the
-patient is often registered to an atlas so that a grey-matter mask may
-be created. Since all the "activation" should occur in grey matter voxels,
-it made sense to restrict analysis to only these sections; although
-there may be partially grey-matter areas that are missed. This significantly
-eases computation time though.
-
-In normal statistical parametric mapping, a gaussian smoothing
+volumes at every other time are registered to this one volume. This corrects
+for motion by the patient as well as small changes in the magnetic
+fields that cause the image to shift. 
+In conventional statistical parametric mapping, a gaussian smoothing
 filter is applied across the image as discussed in \autoref{sec:RFT}.
-After this, detrending is performed discussed by \autoref{sec:Detrend}.
-Recall that FMRI signal levels are unit-less and even though detrending isn't
+After this, detrending is performed as discussed by \autoref{sec:Detrend}.
+Recall that FMRI signal levels are unit-less and though detrending is not
 always necessary, at the very least the data must be converted 

### 2010-07-21 doc/thesis/particlefilters.tex
More clean up.
corrected the intro, and moved the noise description out to methods

@@
 be equal.  The case where a large number of the weights have become irrelevantly small
-is called "particle degeneracy". In  \cite{Liu1998b}
-an ideal calculation of the "effective" number of particles is found based on the 
-particles' "true weight". However, given that only an approximation for the true weight 
+is called \emph{particle degeneracy}. In  \cite{Liu1998b}
+an ideal calculation of the effective number of particles is found based on the 
+particles' true weight. However, given that only an approximation for the true weight 
 exists, they also provide a simple heuristic calculation of $N_{eff}$.
@@
 still limited to the original particles, the number of \emph{unique} particles can only go down.
-This effect, dubbed "particle impoverishment" can result in excessive quantization
+This effect, dubbed \emph{particle impoverishment} can result in excessive quantization
 errors in the final distribution. However, there is a solution. Instead of sampling from the
@@
 The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
+or a genetic algorithm. Versions of $x$ that are fit (give good measurements) spawn more children 
 nearby which allow for more accurate estimation near points of high likelihood. 
@@
 distributions as well, given the unpredictable nature of the noise present in FMRI.
-The choice of this function will be discussed further in \autoref{sec:Methods
-Weighting Function}.
+The choice of this function will be discussed further in \autoref{sec:Methods Weighting Function}.
 

### 2010-07-21 doc/thesis/results.tex
corrected the intro, and moved the noise description out to methods

@@
 \chapter{Simulation}
+\label{sec:SimulationResults}
 Two levels of simulation are discussed in this section. First the simulation of single time series
@@
 identifying the BOLD model in a noisy environment. If the particle filter is unable
-to get back to the "true" time-series in spite of noise then the particle filter
+to get back to the \emph{true} time-series in spite of noise then the particle filter
 will not be useful. Additionally, determination of reasonable parameters from the
 time-series is also desirable. The reason to then simulate an entire slice is that
-the "Possum" FMRI simulation tool models noise more realistically, for instance by
-adding spatial correlation. Additionally 
+Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM)
+models noise more realistically, for instance by noise in the frequency domain. Additionally 
 it provides insight into the applicability of particle filters in modeling the BOLD 
@@
 Given the state-space equations for the BOLD signal, simulating a single time 
-series is relatively straight forward. After a "true" signal is generated,
+series is relatively straight forward. After a true signal is generated,
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
 process with Gaussian I.I.D. steps are added to the true signal. Finally a 
-"carrier level" is added, since BOLD is typically
+carrier level is added, since BOLD is typically
 measured as a \% difference from the baseline. The particle filter
@@
 Once this noisy simulated time series has been generated, the particle filter algorithm
- is then run on this single time series "image". Here I will include two sets of tests 
+ is then run on this single time series image. Here I will include two sets of tests 
 to determine the power of the particle filter in modeling. The first set are 
@@
 results in a slightly larger peak signal. The final parameter sets are shown in 
-\autoref{tab:LowNoiseResults}.
+\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
+I will define the residual as the square root of the mean squared residual, 
+\begin{equation}
+\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
+\end{equation}
+where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
+sample at time $k$. Note the subtle distinction between this and the error
+which I will define as 
+\begin{equation}
+\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
+\end{equation}
+where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
+output. Throughout this section and the next, error and residual of a time series 
+will be referencing these values.
 
@@
 \caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
-last is mean over the 11 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
-residuals. Although in real data I won't distinguish between residual and error, here I will.
-$\sqrt{MSE}$ is the mean squared difference between the true signal and the estimated signal.
-This is the actual error, due to a combination of noise and preprocessing. }
+last is mean over the 11 runs. The $\sqrt{MSR}$ is residual and $\sqrt{MSE}$ is the error.}
 \label{tab:LowNoiseResults} 
@@
 result when the particle filter converges too fast, a result of the weighting functions' variance
-being significantly smaller than the measurement noise ($.005$ vs. $.01$). The square-root
-MSE certainly suffers due to this effect.
+being significantly smaller than the measurement noise ($.005$ vs. $.01$). The error
+certainly suffers due to this effect.
 
@@
 \label{sec:PureNoiseLowMag}
-The two final single-voxel tests force the particle filter to attempt to "learn" a noise-only
+The two final single-voxel tests force the particle filter to attempt to learn a noise-only
 time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
@@
 variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
-the peaks are comparable to peaks in a "real" timeseries. Interestingly the algorithm
+the peaks are comparable to peaks in a real timeseries. Interestingly the algorithm
 suffered from almost constant particle deprevation, meaning that the heuristic
@@
 just because the square-root mean-squared-residual is low doesn't mean the fit is
-"good". Therefore, instead we will normalize the signal based on a value proportional to
+good. Therefore, instead we will normalize the signal based on a value proportional to
 the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
@@
 validity. The fact
-that the particle filter is able to get close to the "true" signal is sadly irrelevant
+that the particle filter is able to get close to the true signal is sadly irrelevant
 when the noise is this high.
@@
 will have to be placed on the error, if any confidence is going to be placed in
-the results. Luckily the "high noise" case featured in \autoref{sec:HighNoise}
+the results. Luckily the high noise case featured in \autoref{sec:HighNoise}
 is actually noisier than typical FMRI.
@@
 good. Clearly, from the normalized $\sqrt{MSR}$ it is possible to identify
- \"active" regions; or regions that respond the way the BOLD model expects.
-Although some of the fits in the  \"active" regions did not have good regression,
+ active regions; or regions that respond the way the BOLD model expects.
+Although some of the fits in the  active regions did not have good regression,
 this is not necessarily the work of false negatives. Instead, because possum

### 2010-07-21 doc/thesis/resultsB.tex
More clean up.
corrected the intro, and moved the noise description out to methods

@@
 single voxel. Although this algorithm will hopefully lead to more novel methods 
-of analysis, the standard use for modeling the BOLD signal is to locate "activation".
+of analysis, the standard use for modeling the BOLD signal is to locate activation.
 Activation is essentially defined as areas where the input seems to directly drive
@@
 The results from SPM8 are shown in \autoref{fig:hm_canon_spm}, and the results from 
-the particle filter are shown in \autoref{hm_canon_pfilter}. Note again that the
+the particle filter are shown in \autoref{fig:hm_canon_pfilter}. Note again that the
 output is actually a different scale. Whereas the levels in the SPM map are T-Scores
@@
 are quite a few dispersed areas with decent fits in \autoref{fig:hm_canon_pfilter}. Also,
-interestingly the lowest square-root MSE (SMSE) was just above $.7$. 
+interestingly the lowest error was just above $.7$. 
 For the sake of completeness \autoref{fig:hm_canon_pfilter85} shows the same image but
@@
          \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
+         Units of match is normalized error. The lowest (best) levels were $.7$,
          whereas the worst levels could go higher than 100 (not shown).}
@@
 I chose several voxels to discuss further from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
-The first voxel, labeled 1, had a very high T-score as well as a very low SMSE (around $.7$). Thus, the
+The first voxel, labeled 1, had a very high T-score as well as a low error (around $.7$). Thus, the
 fit should be very good in both the SPM and particle filter output; this comparison is shown in 
@@
 location to be active in the SPM image (given the results in the surrounding voxels). The fit,
-however shows just why the voxel did not have a particularly good SMSE. Although certain peaks 
+however shows just why the voxel did not have a particularly good error. Although certain peaks 
 seem to fit rather well the input doesn't seem to correlate that well with the BOLD response. For
@@
 The third voxel, compared in \autoref{fig:comp3}, was far away from any other active voxels 
-and yet had a very low (around $.7$) SMSE.
-In both preprocessed time-series the input is extremely noisy, yet by the normalized SMSE
+and yet had a very low (around $.7$) residual.
+In both preprocessed time-series the input is extremely noisy, yet by the normalized residual 
 the response is good. This could be an example of a false positive, although the general trend
@@
 
-Whereas the third voxel was a region that was considered "active" by the particle filter
+Whereas the third voxel was a region that was considered active by the particle filter
 algorithm but not SPM; the fourth voxel is the reverse. The particle filter did not
@@
 
-The fifth, and last single voxel to be analyzed is one more area that had a low SMSE according
+The fifth, and last single voxel to be analyzed is one more area that had a low residual according
 to the particle filter, and yet did not show up in SPM. The line fit is shown in \autoref{fig:comp5},
@@
 very least the output of the particle filter seems to meet the quality of SPM. The normalization
-of the MSE is certainly necessary. Although there is no hard threshold on the 
-normalized SMSE, it would appear that the normalization is providing a reasonable
+of the residual is certainly necessary. Although there is no hard threshold on the 
+normalized residual, it would appear that the normalization is providing a reasonable
 ordering of regression quality. The question of what threshold to use is a difficult one.
-The actual division of the Square Root of the MSE by an estimate for the standard deviation
+The actual division of the residual by an estimate for the standard deviation
 of the signal would seem to mandate a threshold less than one. However, recall that the 
 same media-absolute-deviation was added to the de-trended signal as a DC-gain. This means
-that a flat, no-response bold signal will actually have a SMSE greater than the standard 
-deviation of the signal. Then again, the SMSE may be reduced for any point where the 
+that a flat, no-response bold signal will actually have a residual greater than the standard 
+deviation of the signal. Then again, the residual may be reduced for any point where the 
 model spikes up due to stimuli. Ultimately if the model were to run right down the 
-center of the distribution the SMSE would be equal to the MAD, and would have a 
-normalized SMSE of 1. Thus, again, a normalized SMSE (NSMSE) of 1 is a good start for the 
-threshold. The question then, is what the probability of getting NSMSE below some 
-threshold randomly; which would depend on factors such as the amount of stimuli
-present in the signal. For now I would say that .85 seems to be a decent threshold,
-although it may be a bit conservative.
-
+center of the distribution the residual would be equal to the MAD, and would have a 
+normalized residual of 1. Thus, again, a normalized residual of 1 is at least as 
+good as a flat line at the median. 
 Other measures of fitness could also be used such as mean of the absolute error or 
@@
 this signal is heavily nonlinear, so other estimators may exceed the OLS in terms
-of likelihood. Regardless, the NSMSE seems to provide a very good solution at this
-point, giving comparable results to SPM.
+of likelihood. Regardless, the normalized residual provides a good solution at this
+point, comparable to SPM.

### 2010-07-21 doc/thesis/thesis.tex
more in the results
split up results
corrected the intro, and moved the noise description out to methods

@@
 
-\doublespacing
+%\doublespacing
 
@@
 {\Large 
-BOLD Parameter Estimation Using Particle Filters
+Full Volume Blood-Oxygen-Level-Dependent-Signal Parameter Estimation Using Particle Filters
 }
@@
 
-Abstract Abstract Abstract ...
+Traditional methods of analyzing FMRI images use a linear combination of
+just a few static regressors. This works demonstrates an alternative
+approach using a physiologically inspired nonlinear model. By using a 
+particle filter to optimize the model parameters, the computation time
+is kept low without requiring a linearization of the noise in the state
+variables. The activation results show regions similar those found in 
+SPM; however, there are some notable regions not detected by 
+SPM. Though the parameters selected by the particle filter based approach
+presented here are more than sufficient to predict the BOLD response,
+more model constraints are needed to uniquely identify a single set
+of parameters. This ill-posed nature explains the large discrepancies
+found in other research that attempt to characterize the model parameters.
+Despite this, the reliance on a single point estimator of parameters is
+unnecessary, given the output of the particle filter is a full posterior
+probability at the final time step. For this reason the final distribution 
+of parameters may still be medically relevant. For this same reason,
+additional model constrains may be performed in post-processing by a simple 
+application of conditional probability. Concluding, this work presents
+not just a viable alternative to the traditional method of detecting
+activation, but an extensible technique of estimating the joint BOLD
+parameter distribution.
 
@@
 \include{results}
+\include{resultsB}
 \include{conclusion}

### 2010-07-21 doc/thesis/background.tex
removed lots of weasel words...

@@
 attempts to model the BOLD response; these works will be discussed
-in in the next few sections.
+in this chapter.
 
@@
 does not result in a square wave in the activation of brain regions. 
-The BOLD signal is in fact a significantly smoothed version of the 
+The BOLD signal is in fact a smoothed version of the 
 stimuli. As such, when fitting an FMRI time course to the input,
@@
 \label{sec:BackgroundConclusion}
-In all, the GLM is extremely useful for determining linear 
+In all, the GLM is useful for determining linear 
 dependence of a set of regressors on the output. Unfortunately, as discussed in
@@
 so the scale of the problem is still unknown. The problem is 
-highlighted by the relatively common scenario where no significant
+highlighted by the common scenario where no significant
 activation can be found in a single run \cite{Riera2004}
@@
 section are all attempts to learn some version of the BOLD model. 
-There have been quite a few efforts to quantify the parameters of the
-various BOLD models.  Although \cite{Buxton1998} and \cite{Friston2000}
+Although \cite{Buxton1998} and \cite{Friston2000}
 both proposed physiologically reasonable values for the model parameters, 
@@
 $\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
-of the Volterra Kernel is relatively quick, and, since it is linear there
+of the Volterra Kernel is quick, and, since it is linear there
 is an analytical solution. Thus, once the Volterra Kernel exists, the value
@@
 the calculation of a Jacobian for residuals won't work; and thus powerful
-techniques such as the Gauss-Newton method, which are helpful in many nonlinear
+techniques such as the Gauss-Newton method, which are helpful in most nonlinear
 problems, are off limits. Additionally, a gradient descent is difficult to
@@
 is typically based on some threshold for fitness or a maximum number 
-of generations. This is problematic, since it doesn't really provide 
-any guarantee of even reaching a local minima, although in practice
-it can be quite effective.
-
+of generations. As with all nonlinear searches, unless the function is
+known to be convex, there is no guarantee that a global minimum has been
+reached.
 %[genetic algorithm picture] todo
@@
 mean and covariance of $X(t)$. The benefit of this, is that it requires
-no Jacobian and only a few extra calculations to get a decent esimate of
+no Jacobian and only a few extra calculations to get a decent estimate of
 a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
@@
 of using a Kalman Filter, however, is that it assumes a multivariate 
-Gaussian for the state variables, $X(t-1)$. The problem with this is that 
-when a system is significantly nonlinear, the state at 
-$X(t)$ will almost certainly be non-Gaussian, and thus estimating
-$X(t+1)$ with $X(t)$ as a multivariate Gaussian will perpetually introduce
-error in the distribution. Furthermore, it is not really known what 
+Gaussian for the state variables, $X(t-1)$. The more nonlinear the system
+gets the more likely that the Gaussian will be insufficient to describe
+the distribution, $X(t)$. When this occurs, every step from  $X(t+1)$ 
+to $X(t)$ will introduce additional
+error in the posterior distribution. Furthermore, it is not really known what 
 sort of underlying distributions may exist in such a mixed biological,
@@
 all to be Gaussian may in fact be a gross error. On the other hand, for
-small variances and short time steps the gaussiann distribution is a good 
+small variances and short time steps the Gaussian distribution is a good 
 fit, and so in some limited cases the Unscented Kalman Filter could work
@@
 $s$ at a non-zero level will drive the system for several seconds. 
-\autoref{fig:transp1s} shows the results when the system is essentially
+\autoref{fig:transp1s} shows the results when the system is 
 left on for 100 milliseconds after setting the variables according to \autoref{tab:steptable}.
-The Q-Q plots fit very will with a Gaussian, demonstrating that at this short
+The Q-Q plots fit will with a Gaussian, demonstrating that at this short
 time interval nonlinearities have not yet begun to effect the distribution.
@@
 than most FMRI scanners are capable of sampling at. At that range the tails 
-of the distributions for $v$ and $q$ are clearly starting to deviate from the
+of the distributions for $v$ and $q$ are starting to deviate from the
 Gaussian distribution. As a result the uncertainty in $y$ is deviating from
@@
 have a much greater potential for interplay between the variables. While
-this problem is relatively easy to fix (by using more sigma points off
+this problem is simple to fix (by using more sigma points off
 the main axes), there is a definite cost in complexity.
@@
 parameters, (classically known as parameters and state variables, respectively).
-Essentially a particle filter is used to calculate the state variables at each
+A particle filter is used to calculate the state variables at each
 time; then the estimated distribution of the particles is used to find
@@
 are significantly different from virtually every other study. The most obvious discrepancy
-is the significantly longer time constants, $\tau_f$, $\tau_s$ and $\tau_0$. 
+is the larger time constants, $\tau_f$, $\tau_s$ and $\tau_0$. 
 While of course this could be poor convergence of the algorithm, there is another other possibility.
@@
 \section{Conclusion}
-Clearly there is no ideal solution to solving this system of nonlinear
+There is no ideal solution to solving this system of nonlinear
 equations. Exhaustive search type methods such as those employed by
-\cite{Johnston2008} and \cite{Vakorin2007} can have extremely long
+\cite{Johnston2008} and \cite{Vakorin2007} have long
 run times even for a single voxel. While Volterra models are an interesting
@@
 method of those reviewed here is the Kalman filter based method. It is
-able to maintain a very fast runtime while still approaching the solution.
+able to maintain a fast runtime while still approaching the solution.
 While the reliance on a Gaussian estimate to the true posterior distribution could
-cause problems, some modifications could make it extremely powerful. 
+cause problems, some modifications could make it powerful. 
 The particle filter method proposed in the next section bears a strong 

### 2010-07-21 doc/thesis/conclusion.tex
removed lots of weasel words...

@@
 more detailed models; from the results found here the issue of bias error
-from the BOLD model is not the biggest concern. Clearly clarifying the distributions
-of parameters for the prior should be the first concern. 
-One future study that would be extremely beneficial in this way  would
+from the BOLD model is not the biggest concern. Clarifying the distributions
+of parameters for the prior should be the first concern; currently no multi-patient
+full-volume studies have been done to estimate parameters.
+One future study that would be beneficial in this way  would
 be an extensive study of what the priors should truly be. Although \cite{Friston2000}
@@
 published their estimate of the distributions, given the interplay between
-parameters it is very unlikely that these priors are true to actual distribution
+parameters it is unlikely that these priors are true to actual distribution
 that occurs in vivo. As I mentioned previously, the addition of simultaneous
@@
 In conclusion, using particle filters to estimate the BOLD response are 
-an extremely powerful method of fitting to noisy data. The technique also
+a powerful method of fitting to noisy data. The technique also
 holds great promise as extensible platform to build more advanced models

### 2010-07-21 doc/thesis/intro.tex
removed lots of weasel words...

@@
 Function Magnetic Resonance 
-Imaging (FMRI) use regression techniques with very few degrees of freedom. 
+Imaging (FMRI) use regression techniques with only linear scaling parameters. 
 Though adding more degrees of freedom naturally mandates more computation,
@@
 describe the local changes in Deoxygenated Hemoglobin content.
-Various papers made substantial improvements to this model until
-\cite{Buxton2004} brought all the changes together into a single complete 
+Incremental improvements were made to this model until
+\cite{Friston2000} brought all the changes together into a single complete 
 set of equations. And while there have been numerous adaptations in the model, 
@@
 \cite{Deneux2006}, \cite{Handwerker2004}.
-On the other hand many of the BOLD signal models have numbers
+On the other hand BOLD signal models have numbers
 of parameters ranging from seven \cite{Riera2004} to 50 \cite{Behzadi2005} 
-for a signal as short as 100 samples long. Clearly this number of parameters presents
+for a signal as short as 100 samples long. This number of parameters presents
 a significant risk of being under-determined and having high computation cost. 
@@
 this activation/recharge process happens continuously. However, when a 
-particular region of the brain is very active, the action potentials
-occur significantly more often, resulting in significant local increase
-of the 
-Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, blood vessels in a very 
+particular region of the brain is significantly active, the action potentials
+occur more often, resulting in a local increase of the 
+Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, blood vessels in an
 active area will 
@@
 blood flow, blood volume, and oxygenated hemoglobin. The differences
-in onsets of these various effects is what causes the overcompensation
+in onsets of these effects is what causes the overcompensation
 that is observable in FMRI. Unfortunately, FMRI has no inherent unit
@@
 non-invasively, based on the difference between nuclear spin
-relaxation times in various molecules. First, the subject 
+relaxation times in different molecules. First, the subject 
 is brought into a large magnetic field which causes nuclear spins
@@
 Additionally, the
-coherence of the spins also decay differently (and quite a bit faster
+coherence of the spins also decay differently (and roughly an order of 
+magnitude faster
 than T1 relaxation) based on the properties of the region.
@@
 Because the entire k-space (spatial-frequency) volume is acquired 
-from a single excitation, the signal-to-noise-ratio is very low
+from a single excitation, the signal-to-noise-ratio is low
 in this type of imaging (Echo Planar Imaging).
@@
 studies using 1.5 Tesla scanners.
-While this model is very accurate, it is not perfect.
+While this model is more accurate than the static Hemodynamic Model used in SPM,
+it is not perfect.
 
@@
 undershoot is caused by a prolonged increase in CMRO2 after CBV and CBF
-have returned to their base levels. This theory is justified by quite a few
+have returned to their base levels. This theory is justified by 
 studies that show CBV and CBF returning to the baseline before the BOLD signal
@@
 undershoot, and that
-many of the previous measurements showing a quick recovery of CBV 
+the previous measurements showing a quick recovery of CBV 
 were in fact showing a return to baseline by arterial CBV.
@@
 CMR02 from a higher level, and somewhat more simply; though it 
-still adds several new parameters. \cite{Behzadi2005}
+still adds 9 new parameters. \cite{Behzadi2005}
 introduces nonlinearities into the CBF equations as a method to
@@
 that result from venous and arterial blood. 
-\cite{Deneux2006} compared these various models and though it did 
+\cite{Deneux2006} compared these models and though it did 
 not deal extensively with the 
@@
 to baseline is possible to model, without greatly increasing
-complexity. Regardless, because of many of these models are extremely
-complex, and the benefits have yet to be proved, in this work the simple
-Balloon model will be used. Simplicity is even more important because
+complexity. Regardless, because these models are more 
+complex, and the parameters are not well characterized, in this work the simple
+Balloon model is used. Simplicity is even more important because
 it is the first time a particle filter has been used to calculate the
@@
 
-In summary, there have been extensive attempts to refine the Balloon,
-many of them extremely precise. However, the trade off 
-between increased complexity and increased flexibility is extremely 
-important. Given that the basic BOLD model has 7 parameters, for this
-work, where computation time is very important, that model is
-the best fit.
+In summary, there have been extensive refinements to the Balloon
+model,however, the increased complexity and lack of known priors 
+make these models less desirable. Therefor, in this work
+work, where computation time is especially important, only the 
+most basic model is used.
 
@@
 Since the first complete BOLD model was proposed by \cite{Friston2002}, 
-several studies have endeavored to analyze properties of that model. 
+several studies have analyzed its properties. 
 The most important property is that the system is dissipative, and given
@@
 
-Several studies have been able to calculate distributions for the 
-Balloon parameters, those results are summarized in \autoref{tab:Params}
+A summary of results from previous studies of the Balloon Parameters
+are shown in \autoref{tab:Params}.
 

### 2010-07-21 doc/thesis/methods.tex
removed lots of weasel words...

@@
 regions and multiple patients. The purpose for this work is to determine
-the fitness of particle filters for this task; because with good 
-priors it can be extremely fast. Ultimately this algorithm may become
+the fitness of particle filters for this task. Ultimately this algorithm may become
 more useful as more studies are done on BOLD model that don't require
@@
 being too short to lead to strong activation, a similar stimulus
-scheme in real data showed a significantly larger response than 
+scheme in real data showed a much larger response than 
 half a percent as well. In fact, after applying de-trending,
@@
 sufficiently dense approximation of the prior. For 7 dimensions,
-getting a dense prior is extremely difficult. Insufficiently
+getting a dense prior is difficult. Insufficiently
 dense particles will result in inconsistent results, of course the
@@
 particles would then weight particles based on the results of that noise, rather
-than on the quality of the static parameters. Typical particle filter uses
-also use this state noise as an exploratory measure; however given the 
-high dimensionality of the system, a vast number of particles would be 
-necessary for this to come to fruition. The noise in the BOLD signal is such
-that several particles all tend to be equally correct; the particle filter will
-still leave variance in the solution; which turned out to be the case.
+than on the quality of the static parameters. Typical particle filters 
+also use this state noise as an exploratory measure; however this method is
+less well suited when the system is dissipative.
 
@@
 The primary design decision for resampling is the regularizing
-kernel. As mentioned previously, the Gaussian kernel is extremely convenient,
-because it is very simple to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
+kernel. As mentioned previously, the Gaussian kernel is convenient,
+because it is simple to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
 as long as resampling is kept as a last resort, some amount of over-smoothing
@@
 
-Because the regularized resample will almost certainly over-smooth,
+Because the regularized resampler will almost certainly over-smooth,
 it is necessary to wait until the $N_{eff}$ drops relatively low
@@
 distribution will be unrecoverable. Thus, when the $N_{eff}$ gets down
-to extremely low values, usually this is an indication of particle
-deprivation; something that is also usually accompanied by extremely
-fast drops in variance. When this happens, it is preferable 
+to extremely low values (below 5), usually this is an indication of particle
+deprivation; something that is also usually accompanied by 
+fast drops in variance. When this happens, it is often preferable 
 to return to a previous, wider estimate of scale for the regularization
@@
 due to an under-variant weighting function gives no information. Thus,
-the cost of under variance is significantly higher; a fact that definitely needs
+the cost of a under estimating variance is higher; a fact that definitely needs
 to be taken into account.
@@
 to avoid movement and complex thought, to the best of his abilities.
-Some limitations exist, for instance EPI sequences are extremely loud
+Some limitations exist, for instance EPI sequences are noisy
 which could cause some stimulus in the auditory cortex, and of course
-very long sequences could give the patient time for his mind to wander.
+long sequences could give the patient time for his mind to wander.
 Overall though there should be no activation, and thus the signal will
@@
 for analyzing the noise distribution. The locations were chosen from
-various points around the brain, all in grey matter voxels. The time
+points all around the brain, all in grey matter voxels. The time
 series were also chosen because they were representative of different types
@@
 variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
-are relatively well described by a Gaussian process with a small autocorrelation, 
+are well described by a Gaussian process with a small autocorrelation, 
 \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
@@
 removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
-though clearly not perfectly. Though the distributions still do not exactly fit
-the Normal, \autoref{fig:QQs:D} clearly is much improved compared to \autoref{fig:QQDC:D}.
+though not perfectly. Though the distributions still do not exactly fit
+the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
 In all, the de-trending is effectively removing Wiener noise.
@@
 aspect of a Weiner process, presumably the result of integrating some
-$\nu_x$ is difficult to compensate for, and so various methods
+$\nu_x$ is difficult to compensate for, and so methods
 have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
@@
 value may be found. On the downside, adding another dimension increases the
-complexity of the model, for a variable that is relatively obvious by direct
+complexity of the model, for a variable that is obvious by direct
 observation.
@@
 of scale. The Median Absolute Deviation (MAD)
-proved to be extremely accurate in determining how much to shift the signal up
+proved to be accurate in determining how much to shift the signal up
 by. I tested both methods during the course of analysis, and found that the increase 
@@
 concern; however, all the additions and subtractions done in this section
-have been very low frequency changes and should not substantially change
-the BOLD response, being a relatively high frequency signal. The typical
+have been low frequency changes and should not substantially change
+the BOLD response, being a higher frequency signal. The typical
 method of preprocessing is a high-pass filter with a cutoff of around 30
@@
 
-\subsection{Delta Based Inference}
-\label{sec:Methods Delta Based Inference}
-The alternative to these sorts of low frequency manipulation is to
-go around the problem in another way. Here, I propose and test a 
-different method of dealing with the drift. 
-Instead of comparing the direct output of the particle filter with the direct
-measurement, the algorithm would compare the change in signal over a single TR,
-with the result of integrating the model for the same period. 
-In most signal processing cases this would be foolish, but that is because the 
-general assumption is that all noise is high frequency. Considering 
-the fact that every BOLD analysis pipeline uses a high pass filter,
-whereas low poss temporal filter are rarely applied, it makes sense
-that a derivative type method could work. The benefit of particle filters
-is that they are extremely robust method of inference, and I would assert 
-that the particle filter ought to be given as \emph{raw} data as possible. 
-While taking direct measurements
-without de-trending would give awful results, using the difference removes the 
-DC component and turns what is usually assumed to be a Weiner process into 
-a simple Gaussian random variable.
-
-\begin{equation}
-\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
-\label{eq:measass_delta}
-\end{equation}
-
-Even if $\nu_d$ is some other additive process, the difference will still be closer
-to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
-in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
- All the assumptions made originally
-for the particle filter still hold, and all of the parameters may be distinguished based on
-the step sizes, thus it is not unreasonable to consider matching the string of step sizes
-rather than string of direct readings.
-
-\begin{figure}
-\label{fig:FrequencyGraphs}
-\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
-\end{figure}
+%\subsection{Linearizing Noise}
+%\label{sec:Methods Delta Based Inference}
+%The alternative to these sorts of low frequency manipulation is to
+%go around the problem in another way. Here, I propose a 
+%different method of dealing with the drift. 
+%Instead of comparing the direct output of the particle filter with the direct
+%measurement, the algorithm would compare the change in signal over a single TR,
+%with the result of integrating the model for the same period. 
+%In most signal processing cases this would be foolish, but that is because the 
+%general assumption is that all noise is high frequency. Considering 
+%the fact that every BOLD analysis pipeline uses a high pass filter,
+%whereas low poss temporal filter are rarely applied, it makes sense
+%that a derivative type method could work. The benefit of particle filters
+%is that they are a robust method of inference, and I would assert 
+%that the particle filter ought to be given as \emph{raw} data as possible. 
+%While taking direct measurements
+%without de-trending would give awful results, using the difference removes the 
+%DC component and turns what is usually assumed to be a Weiner process into 
+%a simple Gaussian random variable. 
+%
+%\begin{equation}
+%\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
+%\label{eq:measass_delta}
+%\end{equation}
+%
+%Even if $\nu_d$ is some other additive process, the difference will still be closer
+%to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
+%in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
+% All the assumptions made originally
+%for the particle filter still hold, and all of the parameters may be distinguished based on
+%the step sizes, thus it is not unreasonable to consider matching the string of step sizes
+%rather than string of direct readings. 
+%
+%\begin{figure}
+%\label{fig:FrequencyGraphs}
+%\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
+%\end{figure}
 
@@
 There are two possible outcomes of this. The typical case 
-is at some point where are the particles are in a similar location and 
-so many or all particles fail together.
+is at some point where all the particles are in a similar location and 
+so all the particles fail together.
 The result is then particle deprivation - no particles with non-zero weights remain.
@@
 about to be set to NaN are integrated again with finer grained steps. However
-many times the non-real results don't occur until several time steps after the 
+oftentimes the non-real results don't occur until several time steps after the 
 numbers become unrealistic. So for instance, the time step was too long, allowing 
@@
 occurs. Because the prior is represented initially with significantly more
-particles, if the model fits very well, or for some reason the effective
+particles, if for some reason the effective
 number of particles stays high, resampling could take a long time to occur.
@@
 arbitrary, but at the very least this gives some time for the particle to
-be re-weighted before the new distribution is drawn. At the very least the
-distribution will be a thinned out version of the prior.
+be re-weighted before the new distribution is drawn. Still the
+distribution will then be a thinned out version of the prior.
 

### 2010-07-21 doc/thesis/particlefilters.tex
removed lots of weasel words...

@@
 time-varying random variables. The idea behind particle filters is
-extremely similar to Kalman Filters; however, unlike Kalman Filters,
+similar to Kalman Filters; however, unlike Kalman Filters,
 distributions are stored as an empircal distribution rather than the
@@
 a prior distribution can be troublesome, when the system being modeled
-has physical meaning, establishing reasonable ranges for parameters can 
-actually be quite easy. Optimizing the prior can be more difficult though,
+has physical meaning, establishing reasonable ranges for parameters may be 
+quite easy. Optimizing the prior can be more difficult though,
 unless the system has been extensively studied.
@@
 assumption is based on the idea that if you paused the system for $\Delta t$
-seconds, when unfrozen the system would continue as if nothing happend. 
+seconds, when unfrozen the system would continue as if nothing happened. 
 Few biological systems are predictable enough for them to be summarized
@@
 forward in time using the state equations. Additionally it makes
-updating weights extremely simple, as seen in \autoref{eq:weightevolve}.
+updating weights simple, as seen in \autoref{eq:weightevolve}.
 
@@
 fully constrained by $x_k$, and that $x_k$ is similarly fully 
-constrained by $x_{k-1}$, we are able to make the very good assumptions that:
+constrained by $x_{k-1}$, we are able to make the good assumptions that:
 \begin{eqnarray}
@@
 Using these properties, $P(x^i_{0:k} | y_{0:k})$ may be broken up as follows 
-(mostly using Bayes' Theorem):
+(primarily using Bayes' Theorem):
 \begin{equation}
@@
 
-Thus, by making the following relatively easy assumptions, evolving a posterior
+Thus, by making the following simple assumptions, evolving a posterior
 density  requires no knowledge of noise distribution.
@@
 an approximation of $P(X(t_k) | Y_{0:k})$ or $P(X(t_k + \delta t) | Y_{0:k})$
-is relatively simple.
+is simple.
 
@@
 of the wide prior distribution (required for a proper discretization of a continuous
-distribution), there will be many particles with insignificant weights. While this does help
+distribution), there will be a significant proportion of particles with insignificant weights. 
+While this does help
 describe the tails of the distribution, it means a lot of computation will be based.
@@
 it is necessary to choose a distribution for this. This distribution
-is extremely important because $\nu_y \sim P(y_k | x(T))$, which is used
+is important because $\nu_y \sim P(y_k | x(T))$, which is used
 for updating weights. Ideally this weighting function would exactly 
@@
 has a scale comparable to the signal levels.
-While a Gaussian function is the traditional choice, I wanted to try other
-distributions as well, given the unpredictable nature of the noise present in FMRI.
+While a Gaussian function is the traditional choice, there are other reasonable
+distributions, given the unpredictable nature of the noise present in FMRI.
 The choice of this function will be discussed further in \autoref{sec:Methods Weighting Function}.
@@
 but when $u(t)G$ is greater than $v_t$, the output voltage will be set
-by $u(t)G$ and the capacitor will charge up. We will assume a very simple
+by $u(t)G$ and the capacitor will charge up. We will assume a simple
 model for all the components, ignoring the complex nonlinear behavior
@@
 
-To run the particle filter is relatively easy then, since there exists 
+To run the particle filter is easy, since there exists 
 a recursive definition of the dynamic state variable, $v_y$. To start

### 2010-07-21 doc/thesis/results.tex
removed lots of weasel words...

@@
 signal isn't possible. A similar circumstance occurs at 100 seconds in. 
-In general the fits are good, although the bias introduced by pre-processing clearly
+In general the fits are good, although the bias introduced by pre-processing 
 results in a slightly larger peak signal. The final parameter sets are shown in 
@@
 \autoref{fig:NoiseComparisonJustTwo}, out of the previous eleven. 
-Run 1 clearly has higher peaks, in spite of the fact that both signals are based
+Run 1 has higher peaks, in spite of the fact that both signals are based
 on the same underlying signal. There also appears
@@
 such a low $\epsilon$ region 4 was below the noise threshold. Notice that the
-regions 1 2 an 3 clearly stick out in the $\sqrt{MSR}$ map, indicating that the
+regions 1 2 an 3 stick out in the $\sqrt{MSR}$ map, indicating that the
 particle filter was succesfull in matching those regions.

### 2010-07-21 doc/thesis/resultsB.tex
removed lots of weasel words...

@@
 of analysis, the standard use for modeling the BOLD signal is to locate activation.
-Activation is essentially defined as areas where the input seems to directly drive
+Activation is defined as areas where the input seems to directly drive
 the BOLD response, as opposed to there being intermediate factors controlling it. 
@@
 are quite a few dispersed areas with decent fits in \autoref{fig:hm_canon_pfilter}. Also,
-interestingly the lowest error was just above $.7$. 
+the lowest error was just above $.7$. 
 For the sake of completeness \autoref{fig:hm_canon_pfilter85} shows the same image but
@@
 seem to fit rather well the input doesn't seem to correlate that well with the BOLD response. For
-instance 75 seconds in, the stimulus is clearly not present, and so the signal should drop off; yet
+instance 75 seconds in, the stimulus is not present, and so the signal should drop off; yet
 it doesn't. This is a good example of a voxel that seems to be reacting to something in addition to

### 2010-07-22 doc/thesis/methods.tex
fixed a lot of images etc.

@@
 could be slight movement, given that co-registration of volumes to a single time
-point is standard, this seems unlikely. Regardless, the problem mandates
+point is standard, this is unlikely. Regardless, the problem mandates
 the use of a high pass filter to make inference reasonably powered
@@
 \centering
-\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_29_49_9}}
-\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_34_43_24}}
-\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_22_38_23}}
-\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_37_29_24}}
+\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_29_49_9}}
+\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_34_43_24}}
+\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_22_38_23}}
+\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_37_29_24}}
 
@@
 \centering
-\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
-\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
-\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
-\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
+\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_29_49_9}}
+\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_34_43_24}}
+\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_22_38_23}}
+\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
 \caption{Q-Q Plots of resting state data, using the BOLD signal changes}
@@
 indicates that the steps could be Independently Distributed. 
-Therefore, the assumption that the noise follows a Wiener process seems to
-be correct.
+Therefore, the noise does come close to a Wiener process.
 
@@
 \centering
-\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
-\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
-\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
-\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
+\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_29_49_9}}
+\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_34_43_24}}
+\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_22_38_23}}
+\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
 \caption{Q-Q Plots of resting state data, after the de-trending}
@@
 shift heavily over the course of time. In spite of this, the spline will still remove some
-amount of signal, just like all of these methods, which is why I considered the method proposed
-in \autoref{sec:Methods Delta Based Inference}.
+amount of signal.
 
@@
 was then taken and used as the magnitude for the group. Taking the median 
-versus the mean seemed to work better, given the presence of outliers. It seems
-that there is potential to optimize the spline further using a canonical 
+versus the mean seemed to work better, given the presence of outliers. 
+There is potential to optimize the spline further using a canonical 
 HRF to find resting points; however, for this to work the experiment would have
@@
 possible that a more accurate method may exist; however, for this case the 
-MAD seems to work as it should, as \autoref{fig:DesplineQuality} shows.
+MAD works well, as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.
 
@@
 \end{equation}
-
-\begin{figure}
-\caption{Image of de-spline'd lines with true lines}
-\label{fig:DesplineQuality}
-\end{figure}
 

### 2010-07-22 doc/thesis/results.tex
fixed a lot of images etc.

@@
  may be an acceptable compromise, however it would be difficult to strike the 
- right balance. Instead for the purposes here we depend on the square root MSR to 
+ right balance. Instead for the purposes here we depend on the residual to 
  evaluate if the results were reasonable or not.
@@
 raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
-just because the square-root mean-squared-residual is low doesn't mean the fit is
-good. Therefore, instead we will normalize the signal based on a value proportional to
+a low residual does not necessarily indicate a good fit. 
+Therefore, instead we will normalize the signal based on a value proportional to
 the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
 peaks and troughs, rather than using MAX and MIN values, I used a robust estimator of 
-scale, the Median-Absolute-Deviation (MAD). Technically this is an estimator of the standard
-deviation, however in this case it is also a good estimator of the scale of the input
-signal. Thus, each point in the estimate as well as the original signal may be divided by
-this, leading to a normalized signal. Instead, however, I simply divided the 
-$\sqrt{MSR}$ by the MAD, which is functionally equivalent but leaves the  BOLD
-levels intact for further analysis. The normalized square root MSR values are 
+scale, the Median-Absolute-Deviation (MAD). This is an estimator of the standard
+deviation, and good estimator of the scale of the input signal. 
+Thus, each point in the estimate as well as the original signal may be divided by
+this, leading to a normalized signal. The normalized residual values are 
 shown in \autoref{tab:SignleVoxelActivationComparison}.
@@
 \centering
-\includegraphics[width=10cm]{images/simslice_mask}
-\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
-\label{fig:simslice_mask}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=10cm]{images/simslice_hm}
-\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
+\subfigure[Region labels for simulated slice.]{
+\includegraphics[height=9cm]{images/simregions.png} \label{fig:simslice_mask} }
+\subfigure[Residual Heat Map.]{\includegraphics[height=9cm]{images/sim_hm}}
+\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9cm]{images/sim_gm}}
+\caption{Comparison of activation with greymatter, parameter regions.}
 \label{fig:simslice_hm}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=10cm]{images/simslice_hm85}
-\caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
-Darker indicates lower error.}
-\label{fig:simslice_hm85}
 \end{figure}

### 2010-07-22 doc/thesis/resultsB.tex
fixed a lot of images etc.

@@
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.66]{images/spm_hm}}
-\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=.85]{images/spm_hm_x}}
-\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=.85]{images/spm_hm_y}}
-\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=.85]{images/spm_hm_z}}
-\subfigure{\label{fig:scale_spm} \includegraphics[scale=.3]{images/scale1}}
+\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=3]{images/spm_hm}}
+\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
+\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
+\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
+\subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
 \caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_spm_x} \autoref{fig:hm_canon_spm_y} 
@@
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.66]{images/pfilter_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=.85]{images/pfilter_hm_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=.85]{images/pfilter_hm_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=.85]{images/pfilter_hm_z}}
-\subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.3]{images/scale2}}
+\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=3]{images/pfilter_hm}}
+\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=3]{images/pfilter_hm_x}}
+\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=3]{images/pfilter_hm_y}}
+\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=3]{images/pfilter_hm_z}}
+\subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
 \caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
@@
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.66]{images/pfilter85_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=.85]{images/pfilter_hm85_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=.85]{images/pfilter_hm85_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=.85]{images/pfilter_hm85_z}}
-\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.3]{images/scale3}}
+\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=3]{images/pfilter85_hm}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=3]{images/pfilter_hm85_x}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=3]{images/pfilter_hm85_y}}
+\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=3]{images/pfilter_hm85_z}}
+\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.5]{images/scale3}}
 \caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 

### 2010-07-22 doc/thesis/results.tex
got all the new graphics in

@@
 \hline 
-&$\sqrt{MSR}$ & $N\sqrt{MSR}$ &
- $\sqrt{MSR}$ & $N\sqrt{MSR}$ &
- $\sqrt{MSR}$ & $N\sqrt{MSR}$ &
- $\sqrt{MSR}$ & $N\sqrt{MSR}$ \\
+& $M.I.$ & N. Res. &
+  $M.I.$ & N. Res. &
+  $M.I.$ & N. Res. &
+  $M.I.$ & N. Res. \\
 \hline       
 \hline       
-1 & 0.003211 & 0.478015  & 0.014065  & 1.038941    &0.001667 & 1.295007 & 0.015114 & 1.33641 \\
-2 & 0.003055 & 0.531771  & 0.013731  & 0.951648  &  0.001586 & 1.301750 & 0.015510 & 1.336667 \\
-3 & 0.003289 & 0.545798  & 0.012754 &  0.995392  &  0.001651 & 1.26287 &  0.013957 & 1.159567 \\
-4 & 0.002847 & 0.498237  & 0.016726  & 1.161286  &  0.001506 & 1.431959 & 0.013814 & 1.099876 \\
-5 & 0.003006 & 0.468048  & 0.013698 &  1.039724   & 0.001515 & 1.256645 & 0.014350 & 1.201072 \\
-6 & 0.002833 & 0.459001  & 0.011495 &  1.002143  &  0.001557 & 1.270797 & 0.013504 & 1.045886 \\
-7 & 0.003028 & 0.460962  & 0.015550 &  1.088472   & 0.001585 & 1.154406 & 0.013006 & 1.205429 \\
-8 & 0.003044 & 0.518376  & 0.012054 &  1.059620  &  0.001643 & 1.274564 & 0.015446 & 1.122502 \\
-9 & 0.003345 & 0.525303  & 0.015104 &  1.015703  &  0.001679 & 1.320244 & 0.014847 & 1.086366 \\
-10 & 0.003175 &0.492111  & 0.012493  & 1.189964  &  0.001826 & 1.344562 & 0.013125 & 1.221350 \\
-11 & 0.002889 &0.490919  & 0.012165 &  0.953676  &  0.001950 & 1.325223 & 0.014457 & 1.117367 \\
+1 & 0.86687 &0.47801 &0.09077 &1.03894 &0.06326  &  1.29501 &0.03024   &1.33641 \\
+2 & 0.93975 &0.53177 &0.13767 &0.95165 &-0.01075  & 1.30175 &-0.02677  &1.33667 \\
+3 & 0.82382 &0.5458  &0.13505 &0.99539 &0.02345  &  1.26287 &-0.0111   &1.15957 \\
+4 & 0.94661 &0.49824 &0.04341 &1.16129 &-0.00906  & 1.43196 &0.00147   &1.09988 \\
+5 & 0.94281 &0.46805 &0.13718 &1.03972 &0.00663  &  1.25664 &-0.00204  &1.20107 \\
+6 & 0.92539 &0.459   &0.12337 &1.00214 &-0.00816  & 1.2708 & 0.01775   &1.04589 \\
+7 & 0.98892 &0.46096 &0.15381 &1.08847 &0.02664  &  1.15441 &0.03163   &1.20543 \\
+8 & 0.98796 &0.51838 &0.11325 &1.05962 &0.03285  &  1.27456 &0.01951   &1.1225 \\
+9 & 0.8804  &0.5253  &0.09669 &1.0157  &0.01628  &  1.32024 &0.01039   &1.08637 \\
+10 & 0.88721 &0.49211 &0.18339 &1.18996 &0.00407  &  1.34456 &0.00508   &1.22135 \\
+11 & 0.96644 &0.49092 &0.10949 &0.95368 &0.03323  &  1.32522 &-0.01284  &1.11737 \\
 \hline                                                                           
-mean & 0.003066 & 0.497140  & 0.013621 &  1.04514   &  0.001651 & 1.294366 & 0.014285 & 1.175681 \\
+min  & 0.92329 &0.49714 &0.12037 &1.04514 &0.01622  &  1.29437 &0.00576   &1.17568 \\
+\hline                                                                           
+max & 0.96644 &0.459   &0.10949 &0.95165 &-0.01075  & 1.15441 &-0.01284  &1.04589 \\
+\hline                                                                           
+mean  & 0.86687 &0.5458  &0.09077 &1.18996 &0.06326  &  1.43196 &0.03024   &1.33667 \\
 \hline
 \end{tabular}
-\caption{$\sqrt{MSE}$ and the normalized $\sqrt{MSE}$, 
-    $\left(\sqrt{MSE}/(\text{Median Absolute Deviation})\right)$ for different configurations.} 
+\caption{Mutual Information and the normalized $\sqrt{MSE}$, 
+    for signal/noise configurations.}
 \label{tab:SignleVoxelActivationComparison} 
@@
 \subfigure[Region labels for simulated slice.]{
-\includegraphics[height=9cm]{images/simregions.png} \label{fig:simslice_mask} }
-\subfigure[Residual Heat Map.]{\includegraphics[height=9cm]{images/sim_hm}}
-\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9cm]{images/sim_gm}}
+\includegraphics[height=9.5cm]{images/simregions.png} \label{fig:simslice_mask} }
+\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9.5cm]{images/sim_gm}}
+\subfigure[Residual Map. Scale is Normalized $\sqrt{MSE}$.]
+{\includegraphics[height=9cm]{images/sim_hm}}
+\subfigure[Mutual Information Map. Scale is bits.]
+{\includegraphics[height=9cm]{images/sim_hm_mi}}
 \caption{Comparison of activation with greymatter, parameter regions.}

### 2010-07-22 doc/thesis/resultsB.tex
got all the new graphics in

@@
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=3]{images/spm_hm}}
+\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
 \subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
@@
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=3]{images/pfilter_hm}}
+\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.85]{images/pfilter_hm}}
 \subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=3]{images/pfilter_hm_x}}
@@
 \subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+\caption{Sagittal, coronal and axial (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
          \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized error. The lowest (best) levels were $.7$,
-         whereas the worst levels could go higher than 100 (not shown).}
+         Units of match is normalized residual. The lowest (best) levels were $.7$.
+         The highest error shown is $1$.}
 \label{fig:hm_canon_pfilter}
@@
 
+%\begin{figure}[H]
+%\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.75]{images/pfilter85_hm}}
+%\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=3]{images/pfilter_hm85_x}}
+%\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=3]{images/pfilter_hm85_y}}
+%\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=3]{images/pfilter_hm85_z}}
+%\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.5]{images/scale3}}
+%\caption{Sagittal, coronal and axial (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+%         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
+%         Units of match is normalized residual. The lowest (best) levels were $.7$.
+%         The highest error shown is $.85$.}
+%\label{fig:hm_canon_pfilter85}
+%\end{figure}
+
 \begin{figure}[H]
-\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=3]{images/pfilter85_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_x} \includegraphics[scale=3]{images/pfilter_hm85_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_y} \includegraphics[scale=3]{images/pfilter_hm85_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter85_z} \includegraphics[scale=3]{images/pfilter_hm85_z}}
-\subfigure{\label{fig:scale_pfilter85} \includegraphics[scale=.5]{images/scale3}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
+\subfigure[]{\label{fig:hm_mi} \includegraphics[scale=.85]{images/mi_hm}}
+\subfigure[]{\label{fig:hm_canon_mi_x} \includegraphics[scale=3]{images/mi_hm_x}}
+\subfigure[]{\label{fig:hm_canon_mi_y} \includegraphics[scale=3]{images/mi_hm_y}}
+\subfigure[]{\label{fig:hm_canon_mi_z} \includegraphics[scale=3]{images/mi_hm_z}}
+\subfigure{\label{fig:scale_mi} \includegraphics[scale=.5]{images/scale4}}
+\caption{Sagittal, coronal and axial (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
          \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized $\sqrt{MSE}$. The lowest (best) levels were $.7$,
-         whereas the worst levels could go higher than 100 (not shown).}
+         Units of match is bits (standard for base-2 Mutual Information). The highest (best) levels are 
+         $.42$. The worst shown is $.1$.}
 \label{fig:hm_canon_pfilter85}
@@
 
+% or in original coordinates 29-9-13
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_25_34_25}}\\
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{Section 5, Estimated vs. Actual BOLD response}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_29_9_13}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_29_9_13}}
+\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, T-Value
+of $4.17399$ and $MSE$ of $1.14171$.}
+\label{fig:comp5}
+\end{figure}
+
+%23-10-18 or in original coordinates 36-17-19
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_pfilter_36_17_19}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
+\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
+\label{fig:comp5}
+\end{figure}
+
+%16-19-0 or in original coordinates 29-26-1
+\begin{figure}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/7_pfilter_29_26_1}}\\
+%\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
+\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
 \label{fig:comp5}

### 2010-07-22 doc/thesis/results.tex
fixed up results.

@@
 Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM)
-models noise more realistically, for instance by noise in the frequency domain. Additionally 
-it provides insight into the applicability of particle filters in modeling the BOLD 
+models noise more realistically, for instance by adding noise in the frequency domain. 
+It also provides insight into the applicability of particle filters in modeling the BOLD 
 signal on a large scale. Both sorts of tests are usefull in determining the
@@
 Given the state-space equations for the BOLD signal, simulating a single time 
-series is relatively straight forward. After a true signal is generated,
+series is relatively straight forward. After generating a true signal,
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian I.I.D. steps are added to the true signal. Finally a 
+process with Gaussian I.I.D. steps wered added to this true signal. Finally a 
 carrier level is added, since BOLD is typically
@@
 algorithm will immediately remove this by calculating the \% difference, 
-but adding a carrier level means that the exact same algorithm which 
-will is used for simulated data may be used for the real data.
+but adding a carrier level meant that the exact same algorithm used 
+for simulated data could be used for the real data.
 
-Once this noisy simulated time series has been generated, the particle filter algorithm
- is then run on this single time series image. Here I will include two sets of tests 
+Once this noisy simulated time series was generated, the particle filter algorithm
+ was then run on single time series image. Here I will include two sets of tests 
 to determine the power of the particle filter in modeling. The first set are 
@@
 the observed signal. However, as discussed in the false positive tests,
-even if there is no true underlying signal, the particle filter will still determine
+even if there is no underlying signal, the particle filter will still determine
 the most likely BOLD signal; which may not be flat. Thus, it is 
 important to investigate methods of determining false positives using 
-average residuals. This problem could in actuality never come up, for instance
-if the number of measurements is very large or if the noise peaks are not
-particularly large.
+other methods.
 
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_lownoise}
 \caption{Test Signals with low noise compared to the clean signal.}
@@
 \end{figure}
-
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_lownoise}
 \caption{A comparison of the preprocessed signals for the low noise case. This is the
@@
 \end{figure}
-
 \begin{figure}
@@
 $\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
-These same parameters will be used again in the high noise case. The noise realizations were
-then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
-$\sigma_x$ of $.0005$. The measurement noise as well as the steps of the drift
+These same parameters were also used in the high noise case. The noise realizations were
+then generated based on a measurement noise ($\sigma_y$) of $.001$ and a drift standard deviation
+($\sigma_x$) of $.0005$. The measurement noise as well as the steps of the drift
 were taken to be Gaussian
 For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
-The actual signal that was delivered into the particle filter
-is the result of some preprocessing to remove drift, as described in 
+The actual signal delivered into the particle filter
+was the result of some preprocessing to remove drift, as described in 
 \autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
@@
 The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
-Obviously the spline struggles to match the signal near the end, as the graph shows, although 
+The data shows that the spline struggles to match the signal near the end, although 
 overall, the preprocessing seems to have been successful at removing trends. Finally, the 
@@
 signal isn't possible. A similar circumstance occurs at 100 seconds in. 
-In general the fits are good, although the bias introduced by pre-processing 
+It is also worth noting the bias introduced by pre-processing 
 results in a slightly larger peak signal. The final parameter sets are shown in 
 \autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
-I will define the residual as the square root of the mean squared residual, 
+the residual will be defined as the square root of the mean squared residual, 
 \begin{equation}
@@
 sample at time $k$. Note the subtle distinction between this and the error
-which I will define as 
+which is defines as 
 \begin{equation}
@@
 where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
-output. Throughout this section and the next, error and residual of a time series 
-will be referencing these values.
+output. Throughout this section and the next, the terms \emph{error} and \emph{residual}
+will be referencing these values, for an entire timeseries.
 
@@
 
-There are a few results worth noting here. First the time constants vary greatly across
+There are a few results worth noting in the final parameter estimates (which 
+are the mean of the particle filter's posterior distribution). First the time constants vary 
+greatly across
 runs with different noise realizations, yet the sum of the individual time constants
 ($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
-time constants are falling short of the true time constant. This could be a limitation
+time constants fall short of the true time constant. This could be a limitation
 based on the prior distribution (which notably has an initial mean below the true values) 
 or it could be caused by the disassociated benefits of a correct time
-constant with correct output. It often takes several measurement periods before a difference
+constant with regards to a correct output. It often takes several measurement periods before 
+a difference
 in time constants becomes relevant to the weight. Its also possible that output is insensitive
-to changes in the time constants and that varying the activation durations would 
-cause the variations to be more visible. Another interesting result in the huge variation 
-in the levels of $V_0$. In general,
+to differences in the time constants; althouhg varying the activation durations could induce
+better estimates.  The large variation of $V_0$ are also interesting, although as one final
+covariance matrix shows, there is a definite negative correlation between $\epsilon$ and $V_0$.
+In general,
 with this admittedly small amount of noise, it would appear that the relation between a set
-of parameters/stimuli and the output is not injective. In other words,  a time
+of parameters/stimuli and the output is not injective; a time
 series is not unique to a single set of parameters. This is good justification that 
@@
 
-The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the 
-$\tau$ parameters are interchangeable
-as far as the BOLD signal is concerned. Notice the covariance of $\tau_f$ and $\tau_0$
+The covariance matrix (\autoref{tab:CovSim}) provides a great deal of insight into 
+the way parameters evolve in the particle filter. The fact that there are so many 
+covariance elements that are as big or bigger than variance elements indicates just
+loose the parameters set is. Every single parameters, including $\alpha$ seems to
+have a non-zero covariance. Why $\alpha$ and $\epsilon$ would be positively related
+is one of the most interesting element. Given the difficulty in mapping out a 
+seven dimensional space, it is difficult to know why such correlations exist. 
+ Notice the covariance of $\tau_f$ and $\tau_0$
 is $-0.019$ whereas
 the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
-some measure of ill-defined behavior between the different time-constants, as well as several
-other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
+some measure of ill-defined behavior between the different time-constants, which is to
+be expected given the number steps before a stimuli affects the output. 
+The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
 demonstrates the migration of parameters to match the estimates. Given the relatively high
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_highnoise}
 \caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_highnoise}
 \caption{A comparison of the preprocessed signals for the high noise case.}
@@
 noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
-respectively. This an order of magnitude higher than the previous tests, and indeed the
+respectively. This is an order of magnitude higher than the previous tests, and indeed the
 noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
@@
 result when the particle filter converges too fast, a result of the weighting functions' variance
-being significantly smaller than the measurement noise ($.005$ vs. $.01$). The error
+being smaller than the measurement noise ($.005$ vs. $.01$). The error
 certainly suffers due to this effect.
@@
 $\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
-stimulus/input will be set to 0 the entire time. In effect it is a region of the brain for
-which no visual stimulus induces activity. For the particle filter, on the other hand,
-the stimulus will be left as in the previous two sections. The question then is how the
+stimulus/input neuronal efficiency ($\epsilon$) will be set to 0. In effect it is a region of the brain for
+which no visual stimulus induces activity. The question then is how the
 particle filter will respond to signal that does not correlate with the input, and how
@@
 \centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/realization_noiseonly}
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/realization_noiseonly}
 \caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
@@
 \centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/preprocessed_noiseonly}
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/preprocessed_noiseonly}
 \caption{A comparison of the preprocessed signals for the signal-free case. (
@@
 \hline 
-1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
+1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & todo \\
  0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
@@
 
-From the results it is clear that the parameters have not taken any visible turns into
-unrealistic territory. It appears individual parameters will not be a distinguishing factor
-in determining the validity of the model. Additionally the raw square-root mean-squared-residual
-also appears to be lower than the SMSR of the low noise simulation from earlier. 
+The data shows that the parameters have not taken any visible turns into
+unrealistic territory; individual parameters will not be a distinguishing factor
+in determining the validity of the model. Additionally the residual 
+also appears to be lower than the residual of the low noise simulation from earlier. 
 The precision with which the convergence occurred is actually the best seen in any
-simulation.
-
-However, the reason for low error may in fact be simply the fact that the over all
-signal is significantly lower than any other simulation seen in other simulations.
-The peaks never even cross 1\% difference, with most staying even below .5\% 
-(\autoref{fig:PreprocessedNoiseOnly}). Because the signal is generally staying
+simulation.  The reason for the low error is that the overall
+signal is significantly smaller than any other simulation in this thesis.
+The peaks never even 1\% difference, with most staying even below .5\% 
+(\autoref{fig:PreprocessedNoiseOnly}). Because the stayed 
 well within the range of $.005$, which is the standard deviation of the weighting function
-its unlikely that any particles will have a substantially different wieght. Thus, the reason
+the particles did not ever get substantially different weights. Thus, the reason
 the fits are in such agreement is that no particles gained increased weight, and so
 no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
+This is good evidence that a per time-series weighting function could benefit
+even out the convergence rate; and thus allow better determination between signals
+like this and low noise case from earlier.
 Thus we have now seen how the particle filter responds when the signal is completely
 random and low level, however there is also some possibility that the signal will
-be random, yet active. Thus one more test is necessary on
-a signal with enough noise to actually be considered a signal.
+be random, yet active. Thus one more test is necessary; time series with no 
+underlying signal (as far as responding stimuli) but with enough noise to 
+actually be considered a signal.
 
@@
 the peaks are comparable to peaks in a real timeseries. Interestingly the algorithm
-suffered from almost constant particle deprevation, meaning that the heuristic
-for rescuing the particle filter from particle deprivation, discuessed in
-\autoref{sec:Resampling} in fact was working against the proper course of action. Of course
- the proper course of action in this case would be for the particle filter to fail,
+suffered from almost constant particle deprivation, meaning that the heuristic
+for rescuing the particle filter from particle deprivation, discussed in
+\autoref{sec:Resampling} in fact was working against the proper course of action. 
+ The proper course of action in this case would be for the particle filter to fail,
  since no particle can really properly estimate a random sequence. When this mechanism
- was removed, all 11 runs stopped due to particle deprevation (all weights hit zero). 
+ was removed, all 11 runs stopped due to particle deprivation (all weights hit zero). 
  The problem with allowing particle deprivation to occur is that it can \emph{rarely}
@@
  may be an acceptable compromise, however it would be difficult to strike the 
- right balance. Instead for the purposes here we depend on the residual to 
- evaluate if the results were reasonable or not.
+ right balance. Instead, for the purposes here, I depend on other methods, described
+ later, evaluate if the results were reasonable or not.
 
@@
 
-\begin{figure}[H]
-\centering
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
-{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
-\end{figure}
-\begin{figure}[H]
-\subfigure[$Q$, $S$, $F$, $BOLD$ ]
-{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
-\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
-the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
-\label{fig:JustNoiseConvergenceLarge}
-\end{figure}
+%\begin{figure}[H]
+%\centering
+%\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
+%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
+%\end{figure}
+%\begin{figure}[H]
+%\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
+%\end{figure}
+%\begin{figure}[H]
+%\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+%{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
+%\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
+%the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
+%\label{fig:JustNoiseConvergenceLarge}
+%\end{figure}
 
@@
 is competely detached from the signal, a non-zero response by the particle filter does
-lead to decreased MSR. This is primarily due to the preprocessing which forces the signal to
-have a non-zero mean. While this scheme does create a type of artificial activation for
-the particle filter, in reality the MSR will still be less than if the signal was left
+lead to decreased residual. This is primarily due to the preprocessing which forces the signal to
+have a non-zero mean. While this scheme could create a type of artificial activation for
+the particle filter, in reality the residual will still be less than if the signal were left
 with zero mean, and the particle filter stayed flat. At best the particle filter will 
-over-smooth creating a line that does down the center of the noisy signal. In such a case
-the MSR will be the same as the alternative.
+over-smooth, estimating a centerline to noisy signal. In such a case
+the residual will be the same as the alternative, although the mutual information will 
+be different.
 
@@
 \caption{Mutual Information and the normalized $\sqrt{MSE}$, 
-    for signal/noise configurations.}
+    for signal/noise configurations. TODO FIX MIN/MAX/MEAN}
 \label{tab:SignleVoxelActivationComparison} 
@@
 
-The first step in determining the validity of a model to provide and order of quality
-to rate the results by. Unfortunately, because of the variablity in the signal levels the
+The first step in determining the validity of a model is to provide some order of quality
+to rate the results by. Unfortunately, because of the variability in the signal levels the
 raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
 a low residual does not necessarily indicate a good fit. 
-Therefore, instead we will normalize the signal based on a value proportional to
+Therefore, instead I  normalized the signal based on a value proportional to
 the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
-peaks and troughs, rather than using MAX and MIN values, I used a robust estimator of 
-scale, the Median-Absolute-Deviation (MAD). This is an estimator of the standard
-deviation, and good estimator of the scale of the input signal. 
-Thus, each point in the estimate as well as the original signal may be divided by
-this, leading to a normalized signal. The normalized residual values are 
-shown in \autoref{tab:SignleVoxelActivationComparison}.
+peaks and troughs, rather than using MAX and MIN values to estimate the scale of the 
+signal, I used a robust estimator of 
+scale; the Median-Absolute-Deviation (MAD). This is an estimator of the standard
+deviation, and thus a good estimator of the scale of the input signal. 
+The normalized residual values are 
+shown in \autoref{tab:SignleVoxelActivationComparison}. A second potential method of 
+gauging performance is mutual information. Mutual information is a method of measuring
+the interdependence of two random variables. If two signals are truly independent,
+then the mutual information will be zero. Although ideally suited to Bernoulli 
+random variables where there are a finite number of options, by using histograms
+it is possible to derive a joint distribution of two signals. The algorithm for 
+mutual information is based on that joint distribution:
+\begin{equation}
+\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\left)
+\end{equation}
+Unfortunately the number of bins causing bias in the output, thus to correct for this,
+I subtracted the estimated bias:
+\begin{equation}
+\text{bias} = \frac{N_{bins}}{2Nlog(2)}
+\end{equation}
+where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
+the mutual information estimates in this work 6 bins are used for the marginal
+distribution of each signal. Additionally, throughout log base 2 will be used.
+This leads to 36 total bins in the joint, so the bias is: 
+\begin{equation}
+\text{bias} = \frac{18}{N}
+\end{equation}
+After calculating the empirical mutual information, I subtracted the bias from
+the estimated mutual information. Note that this can result in negative mutual
+information, which should not technically be possible; so any negative mutual information
+was taken as 0.
 
-Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, it is
-clear that distinguishing between these cases is a difficult prospect. One possibility
-is to just allow the particle filter to fail when particle deprivation occurs; however,
-as discussed in the previous section, that could open the door to more false negatives.
-Naturally there is a noise ceiling above which it will be impossible to determine model 
-validity. The fact
-that the particle filter is able to get close to the true signal is sadly irrelevant
-when the noise is this high.
-Ultimately, a threshold well below those shown in the high noise case 
-will have to be placed on the error, if any confidence is going to be placed in
-the results. Luckily the high noise case featured in \autoref{sec:HighNoise}
-is actually noisier than typical FMRI.
+Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, 
+distinguishing between these cases with normalized residual alone is difficult.
+On the other hand the minimum mutual information in the high noise case was $.09077$,
+and the maximum mutual information in the no signal, high noise case is $.03163$. This
+is a decent break, and should provide a means to  detect what would otherwise be a 
+false positive by the normalized residual. That said, mutual information is in some 
+ways more permissive than residual based metrics. For instance, the addition of an arbitrary
+constant to a random variable does not effect mutual information. For this reason, 
+throughout the rest of the results, both values are reports.
 
@@
 techniques it is important to ensure consistency of results. Although the parameter
-sets were somewhat inconsistent, the quality of the output was actually quite consistent
+sets were somewhat inconsistent, the quality of the fit was actually quite consistent
 across eleven runs. The second purpose was of course to determine how the particle 
-filter responded to different signal to noise ratios. Although the direct $\sqrt{MSR}$ is not
-a good measure of quality, the normalized $\sqrt{MSR}$ is. Drawing a line at a particular
-threshold may be difficult, however from the tests it seems that normalized $\sqrt{MSR}$
-below $1$ may be a good start.
+filter responded to different signal to noise ratios.
 
 \section{Multi-voxel Simulation}
-To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
+To test the usefulness of the particle filter for I used a modified version of the FSL tool 
 POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
@@
 of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
-this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
+this has no effect on the quality of the simulation and should not effect the 
+underlying parameters (other than perhaps $\epsilon$ and $V_0$). 
 Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
@@
 generated to test the difference in BOLD levels between the real parameters and the
-estimated ones. Since it is clear that the parameters are not fully orthogonal 
+estimated ones. Since it is the parameters are far from orthogonal 
 (\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
-yet different. This way, a quantitative difference between the two parameter sets 
-may be found.
+despite large difference. This way, a quantitative difference between the two parameter sets 
+was found.
 
@@
 Note that region 4 has a very low $\epsilon$. For this reason, the only areas
-with significantly decent estimates of the BOLD time series are 1,2 and 3. With
-such a low $\epsilon$ region 4 was below the noise threshold. Notice that the
-regions 1 2 an 3 stick out in the $\sqrt{MSR}$ map, indicating that the
-particle filter was succesfull in matching those regions.
+with significant estimates of the BOLD time series were 1,2 and 3. With
+such a low $\epsilon$, region 4 was below the noise threshold. Notice that the
+regions 1 2 an 3 stick out in both the $\sqrt{MSR}$ and the mutual information
+map, indicating that the
+particle filter was successful in matching those regions.
 
@@
 \end{figure}
+todo regenerate histograms
 
-It is again clear that a single point estimate of the parameters is not feasible,
-but rather a distribution of parameters is the best any modeling algorithm
-is going to be capable of for this version of the BOLD signal. In spite of
-this, and relatively heavy noise, the fact that is that the fits were
-good. Clearly, from the normalized $\sqrt{MSR}$ it is possible to identify
- active regions; or regions that respond the way the BOLD model expects.
-Although some of the fits in the  active regions did not have good regression,
-this is not necessarily the work of false negatives. Instead, because possum
-simulates different tissues, it is more likeley that those regions non
-grey-matter areas. Essentially these areas should not have reacted very strongly
-because in the real world such regions wouldn't. By the fact that it 
-was relatively easy to identify the regions of activation from the simulated 
-results, something similar should happen with real FMRI data, which is discussed
-in the next session. which
+The histograms again demonstrate that a single point estimate of the parameters 
+is elusive for the current model. However the data clearly show the power
+of the particle filter at identifying regions of activation; which usually
+performed with statistical parametric maps. The thresholds applied to this
+slice, both for mutual information and residuals is arbitrary, and needs further
+research. Tighter thresholds would certainly remove the false positives present
+in the images. Interestingly the false positives present in the Mutual Information
+map are different from those in the residual map. For this reason, it is 
+possible that the best solution would make use of both metrics. Although
+at first glance it would appear that there are false negatives; this is
+not actually the case. POSSUM simulates different tissues, and white matter
+does not typically  have a BOLD response. This is why there are holes in regions
+2 and 3. These results certainly indicate that the particle filter is effective
+at regressing against a noisy signal. Therefore the next section moves forward
+with real FMRI data.

### 2010-07-22 doc/thesis/results.tex
fixed the histograms for slicesim

@@
 \hline 
-1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & todo \\
- 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
- 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
- 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
- 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
- 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
- 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
- 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
- 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
- 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
- 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
+1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 &   0.00167  \\
+ 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00159   \\
+ 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165   \\
+ 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00151   \\
+ 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00152   \\
+ 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00156   \\
+ 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00159   \\
+ 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164   \\
+ 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.00168   \\
+ 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00183   \\
+ 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195   \\
 \hline                                                                  
-1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
+1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165 \\
 \hline
@@
 \hline       
-1 & 0.86687 &0.47801 &0.09077 &1.03894 &0.06326  &  1.29501 &0.03024   &1.33641 \\
-2 & 0.93975 &0.53177 &0.13767 &0.95165 &-0.01075  & 1.30175 &-0.02677  &1.33667 \\
-3 & 0.82382 &0.5458  &0.13505 &0.99539 &0.02345  &  1.26287 &-0.0111   &1.15957 \\
-4 & 0.94661 &0.49824 &0.04341 &1.16129 &-0.00906  & 1.43196 &0.00147   &1.09988 \\
-5 & 0.94281 &0.46805 &0.13718 &1.03972 &0.00663  &  1.25664 &-0.00204  &1.20107 \\
-6 & 0.92539 &0.459   &0.12337 &1.00214 &-0.00816  & 1.2708 & 0.01775   &1.04589 \\
-7 & 0.98892 &0.46096 &0.15381 &1.08847 &0.02664  &  1.15441 &0.03163   &1.20543 \\
-8 & 0.98796 &0.51838 &0.11325 &1.05962 &0.03285  &  1.27456 &0.01951   &1.1225 \\
-9 & 0.8804  &0.5253  &0.09669 &1.0157  &0.01628  &  1.32024 &0.01039   &1.08637 \\
-10 & 0.88721 &0.49211 &0.18339 &1.18996 &0.00407  &  1.34456 &0.00508   &1.22135 \\
-11 & 0.96644 &0.49092 &0.10949 &0.95368 &0.03323  &  1.32522 &-0.01284  &1.11737 \\
+1 &   0.86687  &0.47801 &  0.09077  &1.03894 &  0.06326  &  1.29501 & 0.03024  &1.33641 \\
+2 &   0.93975  &0.53177 &  0.13767  &0.95165 &  -0.01075  & 1.30175 & -0.02677 &1.33667 \\
+3 &   0.82382  &0.5458  &  0.13505  &0.99539 &  0.02345  &  1.26287 & -0.0111  &1.15957 \\
+4 &   0.94661  &0.49824 &  0.04341  &1.16129 &  -0.00906  & 1.43196 & 0.00147  &1.09988 \\
+5 &   0.94281  &0.46805 &  0.13718  &1.03972 &  0.00663  &  1.25664 & -0.00204 &1.20107 \\
+6 &   0.92539  &0.459   &  0.12337  &1.00214 &  -0.00816  & 1.2708 &  0.01775  &1.04589 \\
+7 &   0.98892  &0.46096 &  0.15381  &1.08847 &  0.02664  &  1.15441 & 0.03163  &1.20543 \\
+8 &   0.98796  &0.51838 &  0.11325  &1.05962 &  0.03285  &  1.27456 & 0.01951  &1.1225 \\
+9 &   0.8804   &0.5253  &  0.09669  &1.0157  &  0.01628  &  1.32024 & 0.01039  &1.08637 \\
+10 &  0.88721   &0.49211 & 0.18339  &1.18996 &  0.00407  &  1.34456 & 0.00508  &1.22135 \\
+11 &  0.96644   &0.49092 & 0.10949  &0.95368 &  0.03323  &  1.32522 & -0.01284 &1.11737 \\
 \hline                                                                         
-min  & 0.92329 &0.49714 &0.12037 &1.04514 &0.01622  &  1.29437 &0.00576   &1.17568 \\
+mean &0.92329  &0.49714 & 0.12037   &1.04514 &  0.01622  &  1.29437 & 0.00576  &1.17568 \\
 \hline                                                                           
-max & 0.96644 &0.459   &0.10949 &0.95165 &-0.01075  & 1.15441 &-0.01284  &1.04589 \\
+min &  0.82382  &0.459   &  0.04341  &0.95165 & -0.01075  & 1.15441 & -0.02677 &1.04589 \\
 \hline                                                                         
-mean  & 0.86687 &0.5458  &0.09077 &1.18996 &0.06326  &  1.43196 &0.03024   &1.33667 \\
+max &  0.98892  &0.5458  & 0.18339   &1.18996 & 0.06326  &  1.43196 & 0.03163  &1.33667 \\
 \hline
@@
 \caption{Mutual Information and the normalized $\sqrt{MSE}$, 
-    for signal/noise configurations. TODO FIX MIN/MAX/MEAN}
+    for signal/noise configurations.}
 \label{tab:SignleVoxelActivationComparison} 
@@
 \begin{equation}
-\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\left)
+\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\right)
 \end{equation}
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist1}
-\caption{Histogram of estimated parameters in section 1 with $\sqrt{MSR}$ less than $.85$.}
+\caption{Histogram of estimated parameters in section 1 in voxels with mutual information greater
+than $.14$}
 \label{fig:slicesim_hist1}
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
-\caption{Histogram of estimated parameters in section 2 with $\sqrt{MSR}$ less than $.85$.}
+\caption{Histogram of estimated parameters in section 2 in voxels with mutual information greater
+than $.14$}
 \label{fig:slicesim_hist2}
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
-\caption{Histogram of estimated parameters in section 3 with  $\sqrt{MSR}$ less than $.85$.}
+\caption{Histogram of estimated parameters in section 3 in voxels with mutual information greater
+than $.14$}
 \label{fig:slicesim_hist3}

### 2010-07-22 doc/thesis/resultsB.tex
fixed up results.
fixed the histograms for slicesim

@@
 
-In reality, the statistical parametric mapping method is very different from the 
+In reality, the purpose of statistical parametric mapping method is 
+very different from the 
 particle filter algorithm described in this work.
@@
 
-The results from SPM8 are shown in \autoref{fig:hm_canon_spm}, and the results from 
-the particle filter are shown in \autoref{fig:hm_canon_pfilter}. Note again that the
-output is actually a different scale. Whereas the levels in the SPM map are T-Scores
-indicating the probability that the result cannot be explained from a random signal,
-the result of the particle filter is simply a normalized square-root of the mean-squared-
-residual. Thus, lower is better, and zero is a perfect fit in the particle filter whereas
-infinity is a perfect fit in \autoref{fig:hm_canon_spm}. Note that it would seem that there
-are quite a few dispersed areas with decent fits in \autoref{fig:hm_canon_pfilter}. Also,
-the lowest error was just above $.7$. 
-For the sake of completeness \autoref{fig:hm_canon_pfilter85} shows the same image but
-thresheld at $.85$ rather than $1$. Obviously this image seems like a closer match to
-the results of SPM, however, the results of \autoref{fig:hm_canon_pfilter} are interesting.
+The results from T-values from SPM8 are shown in \autoref{fig:hm_canon_spm} (threshold of
+4), and the results from 
+the particle filter are shown in \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_pfilter_mi}.
+Note that the scales for all three images are different, because the metrics are different.
+SPM measures using T-Tests to determine the likelihood of a false positive. \autoref{fig:hm_canon_pfilter}
+uses simple normalized residuals, meaning that lower indicates less error.  
+\autoref{fig:hm_canon_pfilter_mi} measures in terms of the dependence between the 
+measured signal and the estimated signal; thus higher indicates a better fit. The particle
+filter data shows a large number of false positives, however application of a threshold
+of $.85$ on the residual map removes these false positives. Similarly, in the mutual information
+map, the false positives may be elminated by upping the threshold to $.15$. However, just
+because the results disagree with SPM does not necessarily mean they are false positives.
+SPM operates on smoothed data (8mm x 8mm x 8mm), so there are certainly active 
+areas that have been missed because of the smoothing.
 
@@
          $.42$. The worst shown is $.1$.}
-\label{fig:hm_canon_pfilter85}
+\label{fig:hm_canon_pfilter_mi}
 \end{figure}

### 2010-07-22 doc/thesis/results.tex
re-write real data section

@@
 \subsection{Single Voxel Review}
+\label{sec:SingleVoxelReview}
 \begin{table}[t]
@@
 Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, 
-distinguishing between these cases with normalized residual alone is difficult.
-On the other hand the minimum mutual information in the high noise case was $.09077$,
-and the maximum mutual information in the no signal, high noise case is $.03163$. This
-is a decent break, and should provide a means to  detect what would otherwise be a 
-false positive by the normalized residual. That said, mutual information is in some 
-ways more permissive than residual based metrics. For instance, the addition of an arbitrary
-constant to a random variable does not effect mutual information. For this reason, 
-throughout the rest of the results, both values are reports.
+distinguishing between these cases with ether normalized residual or mutual information
+is not clear cut. While the average mutual information is in fact more than 10 times
+greater than the average mutual information in the two non-signal cases, the maximum
+mutual information of the low noise/no signal case exceeds the minimum M.I. of the high noise/signal
+case. The Low Noise/ signal determination is easier to make; given
+the minimum mutual information is above $.8$ and the maximum normalized residual is below $.6$.
+However, it is worth noting that the worst case scenario  for mutual information (maximum) in the
+low noise/no signal case does not concur with the worse case (minimum) normalized residual. 
+There is no reason why this has to be the case, it is beneficial. In
+other words, if it were necessary to make a statement that
+a particular voxel active or inactive; the accuracy would improve if both techniques
+were used with loose restrictions, but a voxel would only be considered active if the
+voxel passed both tests.
 
-There were two primary purposes of these tests. First given the nature of monte-carlo
+There were two primary purposes of these tests. First, given the nature of monte-carlo
 techniques it is important to ensure consistency of results. Although the parameter
@@
 \end{figure}
-todo regenerate histograms
 
@@
 is elusive for the current model. However the data clearly show the power
-of the particle filter at identifying regions of activation; which usually
+of the particle filter at identifying regions of activation; which is usually
 performed with statistical parametric maps. The thresholds applied to this
 slice, both for mutual information and residuals is arbitrary, and needs further
-research. Tighter thresholds would certainly remove the false positives present
-in the images. Interestingly the false positives present in the Mutual Information
-map are different from those in the residual map. For this reason, it is 
-possible that the best solution would make use of both metrics. Although
-at first glance it would appear that there are false negatives; this is
+research. Tighter thresholds removes the false positives present
+in the images at the cost of false negatives. Interestingly the false 
+positives present in the Mutual Information
+map are different from those in the residual map. This furthers the argument from
+\autoref{sec:SingleVoxelReview} for combining the two metrics to 
+increase power. Although
+at first glance it would appear that there are false negatives in the 
+\autoref{fig:simslice_hm}; this is
 not actually the case. POSSUM simulates different tissues, and white matter

### 2010-07-22 doc/thesis/resultsB.tex
re-write real data section

@@
 \label{sec:RealData}
-Modeling the BOLD response is of course not of much use if it is only done in a 
-single voxel. Although this algorithm will hopefully lead to more novel methods 
+Modeling the BOLD response is of course not of much use if does not work for
+real FMRI.  Although this algorithm will hopefully lead to more novel methods 
 of analysis, the standard use for modeling the BOLD signal is to locate activation.
-Activation is defined as areas where the input seems to directly drive
-the BOLD response, as opposed to there being intermediate factors controlling it. 
+Activation is defined as areas where the input is the primary drive
+for the BOLD response, as opposed to intermediate factors controlling it. 
 Once areas where the BOLD 
@@
 
-In reality, the purpose of statistical parametric mapping method is 
-very different from the 
-particle filter algorithm described in this work.
-SPM preprocesses the image by spatially smoothing the FMRI image (in this section 
+The data described in this section are fundamentally different in several 
+ways. SPM preprocesses the image by spatially smoothing the FMRI image (in this section 
 SPM8 was used with an $8mm x 8mm x 8mm$ Gaussian kernel), whereas
@@
 of the particle filter is a posterior probability distribution of the parameters
-at every voxel. To validate the quality of the particle filter results though
-it is necessary to compare the location and fit of the particle filter parameter
-mapping with SPM's T-score map.
+at every voxel. To validate the quality of the particle filter results 
+it is necessary to compare the both the location and the fit calculated by the particle filter
+with SPM's location and fit.
 
 \section{Results}
-
 The results from T-values from SPM8 are shown in \autoref{fig:hm_canon_spm} (threshold of
@@
 I chose several voxels to discuss further from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
-The first voxel, labeled 1, had a very high T-score as well as a low error (around $.7$). Thus, the
+The first voxel, labeled 1, had a very high T-score, high mutual information,  as well as 
+a low error (around $.7$). Thus, the
 fit should be very good in both the SPM and particle filter output; this comparison is shown in 
@@
 spatial smoothing; this explains the lack of the sharp peak in the particle filter's preprocessed
-data. Regardless, as expected, both fits are very good.
+data. Regardless, as expected, both work.
 
@@
 location to be active in the SPM image (given the results in the surrounding voxels). The fit,
-however shows just why the voxel did not have a particularly good error. Although certain peaks 
-seem to fit rather well the input doesn't seem to correlate that well with the BOLD response. For
-instance 75 seconds in, the stimulus is not present, and so the signal should drop off; yet
-it doesn't. This is a good example of a voxel that seems to be reacting to something in addition to
-the direct stimulus. Even in what should be a rather smooth time-series in SPM, the BOLD signal
-looks very noisy.
+however shows just why the residual was high and the mutual information was low. This is a
+prime example of a false positive due to the large smoothing kernel applied by SPM. 
+For instance at 75 seconds, the stimulus is not present, and so the signal should drop off; yet
+it doesn't. While a few peaks seem to match, most of the signal does not correlate with the
+expected state progression. This voxel is not being driven direcly by the input, although it
+may be gated or driven through intermediate region.
 
 The third voxel, compared in \autoref{fig:comp3}, was far away from any other active voxels 
-and yet had a very low (around $.7$) residual.
+and yet had a very low (around $.7$) residual. At the same time, the mutual information was
+high enough to make the point suspect. Although the estimated signal seems to run down the
+middle of the measurement signal; it is clear that there is a significant amount of noise
+present in the signal that is not being explained by the particle filter, 
 In both preprocessed time-series the input is extremely noisy, yet by the normalized residual 
-the response is good. This could be an example of a false positive, although the general trend
-of the response seems to follow the estimated response.
+the response is good. This is an example of a false positive from the residual metric.
 
-Whereas the third voxel was a region that was considered active by the particle filter
-algorithm but not SPM; the fourth voxel is the reverse. The particle filter did not
-find a good fit, but in SPM8 the voxel had a decently high T-score. \autoref{fig:comp4} shows
-the line fit for each algorithm, and neither are particularly good; though in a few areas
-the upward or downward movements seem to be correlated. Regardless, this would seem to be
-a false positive from SPM8.
+The fourth voxel (\autoref{fig:comp4}selected for anyalysis had a relatively high residual, and was not picked
+for activation by SPM. On the other hand, the mutual information was above the threshold
+for the image (todo). It is simple to see why the residual was not acceptable in this case.
+At the same time, the peaks do seem to correlate with the measurement peaks. Regardless,
+this is an example of a  false positive from the mutual information metric.
 
-The fifth, and last single voxel to be analyzed is one more area that had a low residual according
-to the particle filter, and yet did not show up in SPM. The line fit is shown in \autoref{fig:comp5},
-and unlike voxel 3 is certainly \emph{not} a false positive. Regions such as this are a very 
-good example of why spatial smoothing can be problematic. By having a more adaptive model, and
-thus reducing bias error, it is possible to get a very good fit to single voxels such as this,
-and thus the need for spatial smoothing is somewhat reduced.
+The fifth voxel (\autoref{fig:comp5}) is an example of a region with increased activation
+due to smoothing. The fit provided by the BOLD particle filter is not very good, and the time
+series input to SPM is far less noisy. Regardless this is another case where the peaks
+seem to match, yet nothing else does. Take for instance the measurements from $30$ seconds
+to $40$ seconds. At that point there is a significant spike in the BOLD signal, yet the
+actual measured signal declines. This is an indication that the BOLD signal is not directly
+being driven by the input. It is likeley that the cause of the activation in this reason is
+not true activation, but recieved from surrounding active regions through spatial smoothing.
 
@@
 \caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
-\label{fig:comp5}
+\label{fig:comp6}
 \end{figure}
@@
 \caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
-\label{fig:comp5}
+\label{fig:comp7}
 \end{figure}
+
+Voxel 6 (\autoref{comp6}) is another region that is active according to both metrics used
+in this work, yet was missed by SPM. The time series in \autoref{comp6} shows an extremely
+good fit, perhaps the best in this batch, for the particle filter. In contrast, the
+fit for SPM is abysmal. In spite of the fact that several other areas around are active
+as well, the smoothing seems to have completely wiped out activation in this voxel.
+
+Finally voxel 7 is a completely ambiguous time series. While the thresholds applied here
+are empirical, this voxel is on the borderline of the thresholds for both mutual information
+and the residual. The signal itself is extremely noisy, it should the voxel should be rejected.
+However this case is reminiscent of the pure-noise tests performed in the previous chapter. 
+This time series an extremely good example of the danger of false positives. In spite of the
+fact that the signal oscillates far faster than the BOLD estimate, because of that, it is somehow
+able to maintaint a mutual information value of $.1052$ and a residual of just $1.09992$.
+As such, a clear method of detecting these sorts of regions is necessary if the results
+of the particle filter algorithm are toe be trusted.
 
@@
 normalized residual, it would appear that the normalization is providing a reasonable
-ordering of regression quality. The question of what threshold to use is a difficult one.
-The actual division of the residual by an estimate for the standard deviation
-of the signal would seem to mandate a threshold less than one. However, recall that the 
-same media-absolute-deviation was added to the de-trended signal as a DC-gain. This means
-that a flat, no-response bold signal will actually have a residual greater than the standard 
-deviation of the signal. Then again, the residual may be reduced for any point where the 
-model spikes up due to stimuli. Ultimately if the model were to run right down the 
-center of the distribution the residual would be equal to the MAD, and would have a 
-normalized residual of 1. Thus, again, a normalized residual of 1 is at least as 
-good as a flat line at the median. 
-Other measures of fitness could also be used such as mean of the absolute error or 
-mutual information. Although least squares is the best linear unbiased estimator,
-this signal is heavily nonlinear, so other estimators may exceed the OLS in terms
-of likelihood. Regardless, the normalized residual provides a good solution at this
-point, comparable to SPM.
+ordering of regression quality. Mutual information also shows promising results.
+Many of the differences between SPM and the particle filter seem to be driven
+more by the pre-processing methods than the regression methd. The reason for 
+SPM applying such extrensive smoothing is to combat false positives. While this is
+extremely important, regions such as \autoref{comp6} show the problem with doing 
+so. In all, the particle filter method does very good job of estimating parameters that
+fit the measurements; and at the very least this method is a workable, albiet
+computationally intensive, alternative to SPM.

### 2010-07-22 doc/thesis/resultsB.tex
added parameter map and parameter hist

@@
 fact that the signal oscillates far faster than the BOLD estimate, because of that, it is somehow
-able to maintaint a mutual information value of $.1052$ and a residual of just $1.09992$.
+able to maintain a mutual information value of $.1052$ and a residual of just $1.09992$.
 As such, a clear method of detecting these sorts of regions is necessary if the results
 of the particle filter algorithm are toe be trusted.
+
+\section{Parameter Estimates}
+Although the parameters are not uniquely identifiable by a single time-series, that
+does not mean estimating them is not without benefit. The parameters still contain useful
+information about the system. Additionally, as an aggregate they form a distribution of
+the feasible parameters parameters for a particular patient. Therefore \autoref{fig:parammaps}
+contains the parameter maps for the system and \autoref{} is a histogram across all voxels
+for which the mutual information was greater than $.15$. As before, the threshold is not
+scientifically derived, yet in tests this threshold provided a decent balance to remove most
+of the questionably active voxels in the system.
+
+\begin{figure}
+\centering
+\subfigure[$\tau_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0000_6}}
+\subfigure[$\tau_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0000_7}}
+\subfigure[$\alpha$, Slice 6]{\includegraphics[width=7cm]{images/param_0001_6}}
+\subfigure[$\alpha$, Slice 7]{\includegraphics[width=7cm]{images/param_0001_7}}
+\subfigure[$E_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0002_6}}
+\subfigure[$E_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0002_7}}
+\subfigure[$V_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0003_6}}
+\subfigure[$V_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0003_7}}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[$\tau_f$, Slice 6]{\includegraphics[width=7cm]{images/param_0004_6}}
+\subfigure[$\tau_f$, Slice 7]{\includegraphics[width=7cm]{images/param_0004_7}}
+\subfigure[$\tau_s$, Slice 6]{\includegraphics[width=7cm]{images/param_0005_6}}
+\subfigure[$\tau_s$, Slice 7]{\includegraphics[width=7cm]{images/param_0005_7}}
+\subfigure[$\epsilon$, Slice 6]{\includegraphics[width=7cm]{images/param_0006_6}}
+\subfigure[$\epsilon$, Slice 7]{\includegraphics[width=7cm]{images/param_0006_7}}
+\caption{Parameter Map of two regions of interest ($M.I. > .15$).}
+\label{fig:ParameterMap}
+\end{figure}
+
+Regions with poor fit cannot have reliable parameter estimates because the input did not meaningfully
+correlate input with output. Therefore before creating the maps,
+each parameters map was masked to regions with mutual information greater than $.15$. I chose mutual
+information over the residual because in the maps comparing regression fitness, the mutual information
+maps had more coherency and less randomness. Additionally, in the single voxel tests (\autoref{sec:SingleVoxelReview}) 
+there was more separation between the no signal case and the signal case than there was in the 
+residual metric.
+
+\begin{figure}
+\centering
+\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=16cm]{images/realhist}
+\caption{Histogram of parameters in active regions ($M.I. > .15$).}
+\label{fig:FinalHist}
+\end{figure}
+
+The parameter maps show consistency parameter across active regions, however, 
+the histogram of parameter estimates across all regions with $M.I. > .15$ is the
+more interesting result (\autoref{fig:FinalHist}).
 
@@
 SPM applying such extrensive smoothing is to combat false positives. While this is
-extremely important, regions such as \autoref{comp6} show the problem with doing 
+extremely important, regions such as \autoref{fig:comp6} show the problem with doing 
 so. In all, the particle filter method does very good job of estimating parameters that

### 2010-07-23 doc/thesis/resultsB.tex
more images, finalized parameter map output.
removes some duplicate labels

@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_pfilter_36_17_19}}\\
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
+\subfigure[Particle Filter]{\label{fig:comp6pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_pfilter_36_17_19}}\\
+\subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
 \caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/7_pfilter_29_26_1}}\\
-%\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
+\subfigure[Particle Filter]{\label{fig:comp7pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/7_pfilter_29_26_1}}\\
+%\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
 \caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
@@
 
-Voxel 6 (\autoref{comp6}) is another region that is active according to both metrics used
-in this work, yet was missed by SPM. The time series in \autoref{comp6} shows an extremely
+Voxel 6 (\autoref{fig:comp6}) is another region that is active according to both metrics used
+in this work, yet was missed by SPM. The time series in \autoref{fig:comp6} shows an extremely
 good fit, perhaps the best in this batch, for the particle filter. In contrast, the
@@
 information about the system. Additionally, as an aggregate they form a distribution of
-the feasible parameters parameters for a particular patient. Therefore \autoref{fig:parammaps}
-contains the parameter maps for the system and \autoref{} is a histogram across all voxels
+the feasible parameters parameters for a particular patient. Therefore 
+\autoref{fig:pmap0} through \autoref{fig:pmap6}
+contain the parameter maps for the system and \autoref{fig:realhist} is a histogram across all voxels
 for which the mutual information was greater than $.15$. As before, the threshold is not
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
-\subfigure[$\tau_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0000_6}}
-\subfigure[$\tau_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0000_7}}
-\subfigure[$\alpha$, Slice 6]{\includegraphics[width=7cm]{images/param_0001_6}}
-\subfigure[$\alpha$, Slice 7]{\includegraphics[width=7cm]{images/param_0001_7}}
-\subfigure[$E_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0002_6}}
-\subfigure[$E_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0002_7}}
-\subfigure[$V_0$, Slice 6]{\includegraphics[width=7cm]{images/param_0003_6}}
-\subfigure[$V_0$, Slice 7]{\includegraphics[width=7cm]{images/param_0003_7}}
+\subfigure{\includegraphics[width=13cm]{images/pmap0}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_00}}
+\caption{$\tau_0$ Estimates}
+\label{fig:pmap0}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[H]
 \centering
-\subfigure[$\tau_f$, Slice 6]{\includegraphics[width=7cm]{images/param_0004_6}}
-\subfigure[$\tau_f$, Slice 7]{\includegraphics[width=7cm]{images/param_0004_7}}
-\subfigure[$\tau_s$, Slice 6]{\includegraphics[width=7cm]{images/param_0005_6}}
-\subfigure[$\tau_s$, Slice 7]{\includegraphics[width=7cm]{images/param_0005_7}}
-\subfigure[$\epsilon$, Slice 6]{\includegraphics[width=7cm]{images/param_0006_6}}
-\subfigure[$\epsilon$, Slice 7]{\includegraphics[width=7cm]{images/param_0006_7}}
-\caption{Parameter Map of two regions of interest ($M.I. > .15$).}
-\label{fig:ParameterMap}
+\subfigure{\includegraphics[width=13cm]{images/pmap1}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_01}}
+\caption{$\alpha$ Estimates}
+\label{fig:pmap1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[width=13cm]{images/pmap2}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_02}}
+\caption{$E_0$ Estimates}
+\label{fig:pmap2}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[width=13cm]{images/pmap3}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_03}}
+\caption{$V_0$ Estimates}
+\label{fig:pmap3}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[width=13cm]{images/pmap4}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_04}}
+\caption{$\tau_f$ Estimates}
+\label{fig:pmap4}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[width=13cm]{images/pmap5}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_05}}
+\caption{$\tau_s$ Estimates}
+\label{fig:pmap5}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[width=13cm]{images/pmap6}}
+\subfigure{\includegraphics[scale=.5]{images/pscale_06}}
+\caption{$\epsilon$ Estimates}
+\label{fig:pmap6}
 \end{figure}
@@
 \caption{Histogram of parameters in active regions ($M.I. > .15$).}
-\label{fig:FinalHist}
+\label{fig:realhist}
 \end{figure}
@@
 the histogram of parameter estimates across all regions with $M.I. > .15$ is the
-more interesting result (\autoref{fig:FinalHist}).
+more interesting result (\autoref{fig:realhist}).
 

### 2010-07-23 doc/thesis/results.tex
added parameter map and parameter hist
Couple Corrections to figure captions

@@
 \caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
-Note also that the red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
+The red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
 and \autoref{fig:FitComparisonHighNoiseJust2}}

### 2010-07-23 doc/thesis/resultsB.tex
Couple Corrections to figure captions

@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_29_9_13}}
-\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, T-Value
-of $4.17399$ and $MSE$ of $1.14171$.}
+\caption{Section 5, Estimated vs. Actual BOLD Response}
+%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, T-Value
+%of $4.17399$ and $MSE$ of $1.14171$.}
 \label{fig:comp5}
@@
 \subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
-\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
+\caption{Section 6, Estimated vs. Actual BOLD Response}
+%\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
 \label{fig:comp6}
@@
 %\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
+\caption{Section 6, Estimated vs. Actual BOLD Response}
+%\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
 \label{fig:comp7}

### 2010-07-23 doc/thesis/intro.tex
fixed a lot of images etc.
changes in introduction primarily

@@
 Function Magnetic Resonance 
-Imaging (FMRI) use regression techniques with only linear scaling parameters. 
-Though adding more degrees of freedom naturally mandates more computation,
+Imaging (FMRI) perform regression using the linear combination of explanatory variables. 
+Though adding nonlinear degrees of freedom naturally mandates more computation,
 in this thesis I will discuss a Sequential Monte Carlo method of fitting a nonlinear
-model with seven degrees of freedom at 
-a computation cost that would still allow real time calculations for multiple voxels.
-More practically, the method described here forms a completely
-separate method of detecting neural activity in comparison to the traditionally
+model a computation cost that would still allow real time calculations for multiple voxels.
+More practically, this method is an alternative
+method of detecting neural activity from the traditionally
 used Statistical Parametric Mapping (SPM). Though more computationally intense,
@@
 and provides more detailed output. Additionally,
-by using a separate particle filter for each voxel's time
-series it is possible to estimate parameters and make real-time predictions
+by using a separate particle filter for each single time series it 
+is possible to estimate parameters and make real-time predictions
 for small neural regions, a feature which could be useful towards real time FMRI 
@@
 apply conditions to the posterior distribution in post-processing without
-having to re-run the algorithm; for instance in light of additional physiological
+having to recalculating the parameters; for instance to impose of additional 
+physiological
 constraints. Modeling the BOLD response as a nonlinear system is the
@@
 detail about the specific particle filter configuration used in this work.
-This section also describes the pre-processing used before the particle
-filter is applied. The results are described separately for simulated data
+This section also describes the pre-processing pipeline. 
+The results are described separately for simulated data
 and real FMRI data in \autoref{sec:SimulationResults} and \autoref{sec:RealData},
@@
 the usefulness and implications of this technique as well as recommendations
-for direction of future works.
+for the direction of future works.
 
@@
 Detecting neural activity using the changes in FMRI images is based on 
-the so called Blood Oxygen Level Dependent (BOLD) signal.
+the Blood Oxygen Level Dependent (BOLD) signal.
 The BOLD signal is caused by minute changes in the ratio of Deoxygenated
@@
 occur more often, resulting in a local increase of the 
-Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, blood vessels in an
-active area will 
-tend to have less oxygenated hemoglobin (due to the increased rate at which
-oxygen is being consumed), and more deoxygenated hemoglobin,
-resulting in an attenuated FMRI signal. In compensation for 
-activation, muscles that
-control blood vessels relax in that region to allow more blood flow,
+Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, if everything else 
+stay the same, blood vessels in an active area will 
+tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin
+(due to the increased rate at which oxygen is being consumed).
+This would then result in an attenuated FMRI signal. However, to
+compensate for activation, muscles that
+control blood vessels relax allowing increased blood flow,
 which actually overcompensates.
@@
 dip in the MR signal, followed by a prolonged increase in signal
-that slowly settles out. It is this overcompensation that is the 
-primary signal detected with FMRI imaging. This cascade of events
-is believed to consist of increased the local metabolism, 
+that slowly settles out. It is the overcompensation that provides
+the  primary signal detected with FMRI. This cascade of events
+is believed to drive a prolonged increased in local metabolism, 
 blood flow, blood volume, and oxygenated hemoglobin. The differences
-in onsets of these effects is what causes the overcompensation
-that is observable in FMRI. Unfortunately, FMRI has no inherent unit
-of measurement, and thus signal levels are all relative: within a particular
+in onsets and recovery times of these variables are what causes the 
+distinguishing characteristics of the BOLD signal. Unfortunately, 
+FMRI signal levels are all relative: within a particular
 person, scanner and run.
@@
 field, they emit detectable RF signals. Conveniently, the
-excitation of nuclear spins return their original state at different
+nuclear spins return their original state at different
 rates, called the T1 relaxation time, depending on the atoms excited.
@@
 T1 relaxation times are typically on the order of seconds if 
-a sufficiently strong excitation was applied. 
+a sufficiently strong excitation was applied, whereas T2 relaxation
+times are usually less than 100ms. 
 In order to rapidly acquire entire brain images, as done in Functional 
@@
 from a single excitation, the signal-to-noise-ratio is low
-in this type of imaging (Echo Planar Imaging).
+in EPI.
 
-Increasing the spatial resolution of EPI imaging necessarily 
+Increasing the spatial resolution of EPI necessarily 
 requires more time or faster magnetic field switching. Increasing
-magnet switching rates though is difficult, because it can result in
-more artifacts, or even lower signal to noise ratios. The result is
-that at \emph{best} FMRI is capable of 1 second temporal resolution. 
-The signal is further diluted because each voxel contains
-the signal from a large number of neurons, capillaries and veins. 
+magnet switching rates can result in
+more artifacts and lower signal to noise ratios. The result is
+that at best FMRI is capable of 1 second temporal resolution. 
+The signal is also diluted because each voxel contains
+a variety of neurons, capillaries and veins. 
 Thus, the FMRI signal, which is sensitive to the chemical composition of 
@@
 and explored in depth in \autoref{sec:BOLD Physiology},
-the usefulness of FMRI comes from the discerning of changes in 
+the usefulness of FMRI comes from discerning of changes in 
 Deoxyhemoglobin/Oxyhemoglobin. Therefore, it is necessary to assume
-that in the short term the only chemical changes will be in
+that the only chemical changes will be in
 capillary beds feeding neurons. In practice this may not be the case, for
-instance near significant veins, and it may explain some of the
+instance near large veins, and it may explain some of the
 noise seen in FMRI imaging (see \autoref{sec:Introduction Noise}. 
-Because MRI lacks units and certain
+Because MRI lacks is unitless and certain
 areas will have a higher base MR signal, all FMRI studies deal with
-percent change from the base signal; rather than raw values. This
-also removes most of the structural data which is not helpful 
-in determining neural activity.
+percent change from the base signal; rather than raw values.
 
@@
 capillary beds, precedes increased outflow, driving increased 
-blood storage capacity.
+blood storage.
 Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
-Hemoglobin, increased volume of blood can effect this ratio if 
+Hemoglobin, increased blood volume effects this ratio if 
 metabolism doesn't exactly match the increased inflow of oxygenated blood.
@@
 for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
-model (\cite{Mandeville1999}). These works derive from first principals
-the changes in deoxyhemoglobin ratio and volume of capillaries based on a given flow.
+model (\cite{Mandeville1999}). These works derive, from first principals,
+the changes in deoxyhemoglobin ratio and volume of capillaries given flow waveform.
 These were the first two attempts to quantitatively account for the shape of the 
 BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
-and the inward cerebral blood flow (CBF). In fact \cite{Buxton1998} went so far as
-to show that a simple, well chosen blood flow waveform coupled with a square 
-wave cerebral metabolic rate of oxygen (CMRO2) curve, in the context of a balloon 
-model, could fully account for the BOLD signal.
+and the inward cerebral blood flow (CBF).
 
 Although \cite{Buxton1998} demonstrated that a well chosen flow waveform could 
-explain most features of the BOLD signal, there was still a matter of proposing a
-realistic waveform for the CBF and for the CMRO2. \cite{Friston2000} gave
-a reasonable and simple
-expression for CBF input,$f$, based on a flow inducing signal, $s$, 
-in combination with the original balloon model
-where $v$ is normalized cerebral blood volume (CBV), $q$ is the normalized
-local deoxyhemoglobin/oxygenated hemoglobin ratio.
+explain most features of the BOLD signal, it stopped short of proposing a
+realistic waveform for the CBF and CMRO2. In \cite{Friston2000} Friston et. al.
+gave a reasonable and simple
+expression for CBF input based on a flow inducing signal.
+Additionally, in the same work Friston et. al. proposed a simple method
+of estimating metabolic rate: as a direct function of the inward blood flow.
+By combining these equations with the balloon model from \cite{Buxton1998},
+it is possible to predict the BOLD signal directly from a stimulus time course.
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\epsilon$ is a neuronal efficiency term, $u(t)$ is a stimulus, and $\tau_f$, $\tau_s$
-are both time constants, $E_0$ is the resting metabolic
-rate and $\alpha$ is Grubb's parameter controlling the balloon model.
+where $s$ is a flow inducing signal, $f$ is the input blood flow (CBF),
+$v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
+local deoxyhemoglobin. The 
+parameters controlling blood flow are $\epsilon$, which is a neuronal 
+efficiency term, $u(t)$, which is the stimulus, and $\tau_f$, $\tau_s$ 
+which are time constants. The parameters for the evolution of blood 
+volume are $E_0$ which the resting metabolic
+rate and $\alpha$ which is Grubb's parameter controlling the balloon model. 
+$\tau_0$ is a single time constant controlling the speed of $v$ and $q$.
 
-This completed the basic balloon model, and was well summarized again
+This completed balloon model was well summarized again
 in \cite{Riera2003}.  \cite{Obata2004} refined the readout equation 
@@
 $\epsilon_0 = 1.43$ is the 
-ratio of signal MR from intravascular to extravascular at rest. Although,
+ratio of signal MR from intravascular to extravascular regions at rest. Although,
 these constants change with experiment ($TE$, $\nu_0$, $r_0$),
@@
 While this model is more accurate than the static Hemodynamic Model used in SPM,
-it is not perfect.
+there are other additions which give it more degrees of freedom.
 
@@
 Although the most widely used, the BOLD model described in \autoref{eq:bold}
-and \autoref{eq:boldout} have been extensively added on to. The most
+and \autoref{eq:boldout} have been extended in various fashions. The most
 significant feature missing from the original model is the 
@@
 Because \autoref{eq:bold} is not capable of producing such a prolonged undershoot,
-additional factors must exist.  Two theories exist for the post stimulus undershoot.
-Recall
+additional factors must be at play. Two prominent theories exist to explain the post 
+stimulus undershoot.  Recall
 that a lower than base signal means that there is an increased deoxyhemoglobin
@@
 CBV/CBF measurement techniques it is difficult to isolate whether CBF and
-CBV truly have returned to their baseline. Other studies seems to indicate
+CBV truly have returned to their baseline. Other studies indicate
 that there can be a prolonged supernormal CBV (\cite{Mandeville1999a}, 
@@
 the previous measurements showing a quick recovery of CBV 
-were in fact showing a return to baseline by arterial CBV.
+were in fact showing a return to baseline by arterial CBV (which
+has little effect on the BOLD signal).
 
@@
 prolonged increase in CBF observed in \cite{Chen2009}. Similarly
-\cite{Zheng2005} adds additional compartments to model the BOLD signal
-that result from venous and arterial blood. 
-\cite{Deneux2006} compared these models and though it did 
-not deal extensively with the 
+\cite{Zheng2005} adds additional compartments to model 
+venous and arterial blood. 
+In \cite{Deneux2006} Deneux et. al. compared these models and though 
+that work did not deal extensively with the 
 post-stimulus undershoot, it did show incremental improvements
-in quality from additional parameters (over the basic Balloon model);
-though at the cost of greatly increased complexity.
+in quality from the additional parameters. 
 Importantly,\cite{Deneux2006} did show that by 
@@
 complex, and the parameters are not well characterized, in this work the simple
-Balloon model is used. Simplicity is even more important because
-it is the first time a particle filter has been used to calculate the
-BOLD parameters. Of course, future works could certainly benefit from
-the more advanced models, especially with the addition of viscoelastic
-effects.
+Balloon model is used.
 
@@
 model,however, the increased complexity and lack of known priors 
-make these models less desirable. Therefor, in this work
-work, where computation time is especially important, only the 
-most basic model is used.
+make these models less desirable. Additional degrees of freedom 
+could also make parameter estimation intractable.
 
@@
 enough time will converge to a constant value. This is found simply by
-analyzing the eigenvalues of the Jacobian of the state equations, 
+analyzing the eigenvalues of the state equation Jacobian, 
 (\cite{Deneux2006}, \cite{Hu2009}). The steady state of the Balloon
@@
 This nonlinearity is accounted for in the Balloon model, although \cite{Deneux2006}
-shows that if there will be large variance in the length of signals, 
-modeling Neural Habituation may be necessary to fully capture the range of responses. 
-Stimuli that last longer than 4 seconds 
+shows that when duration of stimuli varies greatly,
+modeling Neural Habituation is necessary to fully capture the range of responses. 
+In both \cite{Birn2001} and \cite{Deneux2006} it was found that 
+Stimuli lasting longer than 4 seconds 
 tend to be more linear, which is why block designs are so well accounted for
-by the General Linear Model (\cite{Birn2001}, \cite{Deneux2006}). For this 
-paper we will use only a simple version of the Balloon model, described by
-\autoref{eq:bold}, to keep the solution tractable, and because it is the most
-well studied.
+by the General Linear Model (see \autoref{sec:Current Techniques General Linear Model}).
 
@@
 There it was found that the parameters are far from perpendicular,
-implying that exact inference of parameters may not be possible without
-constraint. This could explain the extreme discrepencies in \autoref{tab:Params}.
-
-A summary of results from previous studies of the Balloon Parameters
-are shown in \autoref{tab:Params}.
+and that very different parameters could give nearly identical BOLD output.
+The means that that without constraining parameter values, they may not be 
+precisely ascertainable. This could explain discrepancies in previous studies
+(\autoref{tab:Params}).
 
 %note to self, friston2002b's parameters are from a picture
-
 \begin{table}[t]

### 2010-07-24 doc/thesis/background.tex
changes in introduction primarily
Cleaned up particle filter and background.

@@
 \label{sec:Prior Works}
-The ultimate purpose of this work is to provide a new tool
-for analyzing FMRI data.
-
 Currently, FMRI is used to determine the location of responses
@@
 and instead ask "What does the activation look like". Answering
-this question necessitates more complex models and certainly
-will result in longer run times. Already there have been several other
+this question necessitates more complex models and require
+more computation. Already there have been several other
 attempts to model the BOLD response; these works will be discussed
@@
 FMRI, activation detection is similar and worth discussing. Estimation of 
-parameters is a generalization of the idea of activation detection.
+parameters may be considered generalization of the idea 
+of activation detection.
 Given the popularity of Statistical Parametric Mapping (SPM) 
-it is important to draw a distinction between the methods proposed
+it is important to draw a distinction between it and the methods proposed
 in this work.
@@
 \label{sec:Square Wave}
-The most basic method of analyzing FMRI data is through a standard T-test
+The most basic method of analyzing FMRI data is by standard T-test
 between resting state and active state samples. Simply put, the
@@
 however, such T-tests work slightly differently than a single
-individual test. A t-test with a p-value of $.05$ over a modestly sized FMRI image,
-say 10,000 voxels, will on average generate 500 false
+individual test. A t-test with a p-value of $.05$ over 
+10,000 voxels, will on average generate 500 false
 positives. This is called the multiple comparison problem. Traditional
@@
 can theoretically be applied with a lower scaling factor than
-the original voxel count \cite{Worsley2004}. A side effect of this,
+the original voxel count \cite{Worsley2004}. As side effect of this,
 a single voxel activation is virtually impossible to detect.
@@
 The most common FMRI analysis technique is SPM, though there
-are more advanced versions that the simple square wave
+are more advanced versions than the simple square wave
 method discussed in \autoref{sec:Square Wave}. Hierarchal
@@
 encapsulates all known experimental factors such as stimuli,
-young/old, etc.  The equation for a general linear model is:
+young/old, etc.  The general linear model is defined as:
 
@@
 to minimize the error.
+
+\begin{figure}
+\centering
+\includegraphics[scale=.7]{images/HRF}
+\caption{Canonical Hemodynamic Response Function}
+\label{fig:HRF}
+\end{figure}
 
@@
 discussed, there are certainly variations in the shape of the
-BOLD response, both between brain regions and patients.  \cite{Handwerker2004}
-discusses the implications of choosing an incorrect HRF, the most important
+BOLD response, over both brain regions and patients. Handwerker et. al. (\cite{Handwerker2004})
+discuessed the implications of choosing an incorrect HRF, the most important
 of which is a definite increase in false negatives. While an atlas
 of Hemodynamic Response Functions for each region could definitely 
-mitigate this risk, it does not deal with variation between patients.
-Thus, the inability to fit parameters other than scale definitely hinders 
-analysis of activation. As a result, there is significant bias error
-when using a linear fit to localize activation. Notably, studies
-will also have a bias toward the visual cortex, precisely because
-it is so well studied.
-
-\begin{figure}
-\centering
-\includegraphics[scale=.7]{images/HRF}
-\caption{Canonical Hemodynamic Response Function}
-\label{fig:HRF}
-\end{figure}
+mitigate this risk, it does not account for variation between patients.
+Thus, the inability to fit parameters other than scale hinders 
+analysis of activation.
 
@@
 applied to account for systematic differences between subjects. For instance, if the
-study happens to be a mix between young and old, incorporating
-that into the model is wise, regardless if that is the purpose of the test. 
+study mixes young and old, incorporating
+that into the model is wise, regardless of whether it is the goal of the test. 
 The reason to do this is to account for additional variance that may not
-be present in similar studies. The Hierarchical form used by
+be otherwise explainable.  The Hierarchical form used by
 \cite{Friston2002} is shown in \autoref{eq:Hierarchical}.
@@
 The Empirical Bayes algorithm is used in both in both \cite{Friston2002} 
-and \cite{Hofmann1997}. As a consequence, point estimators 
-are used for each $\theta$, rather than the full distributions.
+and \cite{Hofmann1997}. Note that in Empirical Bayes point estimators 
+are used for each $\theta$, rather than full distributions.
 
@@
 
-The static nature of the  linear model also limits its inference power. 
-Besides not permitting HRF differences between patients, there is no
+The static nature of the  linear model also limits its potential use. 
+Besides not allowing HRF differences between patients, there is no
 reasonable way to incorporate other forms of physiological
-data. Combined FMRI CBF or CBV imaging methods are rapidly getting better,
+data. Combined FMRI CBF or CBV imaging methods are improving,
 as seen in \cite{Chen2009}. These techniques could shed light on
 neural activation by providing extra measurements, yet a 
-physiologically reasonable model is necessary for this.
-In reverse, activation detection methods also don't have the ability 
+physiologically reasonable model is necessary to incorporate this extra data.
+Activation detection methods also don't have the ability 
 to identify pathologies based on state variables or parameters. For
-example decreased compliance of
+example, decreased compliance of
 blood vessels could indicate, or even cause, a neurological condition that 
-is not easily seen in other imaging modalities. Thus, the benefits
-of physiologically meaningful models are manifold.
+is not easily seen in other imaging modalities.
 
-\section{Solving the BOLD Model}
+\section{Approaches to the Balloon Model}
 Unlike Statistical Parametric Mapping, the techniques described in this
@@
 In \cite{Friston2002b}, a novel combination of linear and nonlinear modeling
-was used to generate parameter estimates. Because it is impossible to 
-calculate the partial derivative of
-the output with respect to parameters, $\frac{\partial \theta}{\theta}$, 
-\cite{Friston2002b} approximates the partial. At each step of the Expectation-
-Maximization algorithm, the differential equation is integrated, calculating
+was used to generate parameter estimates. Because there is no closed 
+form solution to the ballon model, it impossible to 
+calculate the Jacobian matrix $\frac{\partial Y}{\theta}$. Thus 
+\cite{Friston2002b} approximates the partial using a Volterra Kernel. 
+At each step of the Expectation-
+Maximization (EM) algorithm, the differential equation is integrated, calculating
 the residuals. Then, for each $\theta_i$ surrounding the current estimate of 
 $\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
-of the Volterra Kernel is quick, and, since it is linear there
-is an analytical solution. Thus, once the Volterra Kernel exists, the value
-of $y$ at that measurement point, for that $\theta_i$ can easily be found. 
-Thus, one part of the partial is calculated at one point using
-$\frac{y(t) - y_i(t)}{\theta - \theta_i}$. This is repeated for
-every dimension of $\theta$ to give the full $\frac{\partial y}{\theta}(t)$.
-This is only at time $t$ though, and it must be repeated for every
-measurement. Finally the full derivate matrix is filled out, and
-the next step in the E-M algorithm can proceed. The full E-M algorithm
+of the Volterra Kernel is quick, and, since there
+is an analytical solution calculating partial derivatives is easy. 
+ The full E-M algorithm
 for estimating the states is notation-heavy and can be found in \cite{Friston2002b}.
 
-Although this is certainly an interesting method of performing non-linear
-regression, there are a few caveats. First, the partials are numerical
-approximations, based on an approximate values (using Volterra-Kernels) of $y$.
+There are a few caveats with this method of optimization. First, the 
+partials are based on an approximate values (using Volterra-Kernels) of $y$.
 Importantly, the Volterra-Expansion of $y$ is not able to model interactions
-between state variables; interactions that were found to cause interesting
-behavior in this work. In \autoref{sec:Unscented Kalman Filter}, for
-the purpose of demonstrating nonlinearities in the signal, it was 
-necessary to propagate state variables through 1 second of simulation.
-For the distribution of state variables mentioned in \autoref{tab:steptable}
-it was necessary to use step sizes smaller than $.001$ to prevent 
-unpredictable behavior. This is using the average parameters, thus the
-only complication came from the interplay between states, which can 
-certainly lead to interesting behavior. Additionally, all the tests performed
+between state variables (\cite{Friston2002b}); which certainly increases error.
+Additionally, all the tests performed
 in \cite{Friston2002b} were on regions found to be active by the General
@@
 \label{sec:Nonlinear Least Squares}
-Rather than localizing activation, by using the physiologically
-plausible BOLD model, it is possible to determine the values of
-governing parameters.
-
 Although there are certainly benefits to using a derived model, rather
@@
 descent are off limits; since the model is a true nonlinear dynamical
-system with no closed form solution. The implication of this is that
-the calculation of a Jacobian for residuals won't work; and thus powerful
-techniques such as the Gauss-Newton method, which are helpful in most nonlinear
-problems, are off limits. Additionally, a gradient descent is difficult to
-perform without the ability to form partials of the output with respect
+system with no closed form solution (although \autoref{sec:Linear Approximation}
+circumvented this by calculating a Volterra Series approximation). Without
+a Jacobian for residuals, the Gauss-Newton method
+is impossible. Additionally, a gradient descent is slow
+without the ability to form partials of the output with respect
 to all the parameters.
 
-Although anything requiring a Jacobian is out, there are other heuristic
-techniques that could potentially illuminate the BOLD response. 
+Still, there are other heuristic algorithms
+that could illuminate the BOLD response. 
 Simulated Annealing (SA) is a common method of optimizing high dimensional
-regression problems. The idea is to pick a random start, and then
-at each iteration pick a random nearby point, and if that point is
+regression problems. First the program selects a random starting point, then
+at each iteration it selects a random point in the neighborhood. If the 
+that new point is 
 below some energy constraint (energy is a function of the residual), 
@@
 is slowly lowered until no nearby points below the temperature can
-be found (or the temperature drops below the current point). There are
+be found. There are
 variations of this, for instance it is common to require every movement
@@
 optimization problems, there is no guarantee of an optimal solution,
-although the longer the algorithm is allowed to run, the better the solution
-will get. Since every step requires an entirely new run of the 
-BOLD model, it can be extremely time consuming, which is why we are
-not using it here.
+although the longer the algorithm is allowed to run, the better the solution.
+Since every step requires a re-integration of the  
+balloon model, it can be extremely time consuming, which is why I did 
+not use it here.
 
@@
 
-Another potential method of interest is the use of Genetic Algorithms
-(GA). Genetic algorithms are similar to Simulated Annealing, in
+Genetic algorithms (GA) are similar to Simulated Annealing, in
 that they randomly move to better solutions based on a cost function.
@@
 and then each set of parameters is rated with a fitness function. Parameter
-sets that are good get a higher weight; then new parameter sets are generated by 
-randomly combining pieces of the old parameter sets. The pieces are typically
-chosen at a rate proportional to the fitness of the donor; thus fit
+sets that are good get a higher weight. New parameter sets are generated by 
+randomly combining pieces of the old parameter sets. The pieces are 
+chosen at a rate proportional to the fitness of the donor; thus good
 parameter sets tend to pass on their properties. In addition to this,
-random mutations may be introduced that come from no existing parent. 
-The new generation is then rated with the fitness function again, and
+random mutations are introduced that come from no existing parent. 
+The fitness function is then used to weight the new generation, and
 the entire process starts over. The stop condition for a genetic algorithm
 is typically based on some threshold for fitness or a maximum number 
-of generations. As with all nonlinear searches, unless the function is
-known to be convex, there is no guarantee that a global minimum has been
-reached.
+of generations. As with simulated annealing (or any generic non-linear optimization),
+there is no guarantee that a global minimum has been reached.
 %[genetic algorithm picture] todo
@@
         \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
-        \STATE At some low probability change one or two parameters in $\Theta_i$
+        \STATE With low probability introduce random mutations to parameters in $\Theta_i$
     \ENDFOR
@@
 Although both these methods can be highly effective, they have the downside of
-requiring very high computation time. In this case of the BOLD model,
+requiring high computation time. In the case of the BOLD model,
 each time the energy or fitness needs to be calculated, a large number of cycles
@@
 no Jacobian and only a few extra calculations to get a decent estimate of
-a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
-closed form solution, and finding the Jacobian is impossible without approximations. Although 
-\cite{Riera2004}, \cite{Hu2009} mention a Jacobian of the BOLD response, this is
-not strictly the case and is rather $\frac{\partial J}{\partial t}$
-rather than a true Jacobian.  This is important because the Extended
-Kalman filter depends on the Jacobian to map a Gaussian through the 
-advancement of time. Thus the popular Extended Kalman Filter won't work
-in this case, whereas the Unscented Kalman Filter still does. In fact
-\cite{Hu2009} uses the UKF to perform a similar type of analysis to
-the one performed in this work.
+a posterior Gaussian. Hu et. al. used the UKF to  perform a similar type of analysis to
+the one performed in this work (\cite{Hu2009}).
 
@@
 mechanical, chemical system such as the brain. Assuming the parameters
-all to be Gaussian may in fact be a gross error. On the other hand, for
+all to be Gaussian could be a gross error. On the other hand, for
 small variances and short time steps the Gaussian distribution is a good 
-fit, and so in some limited cases the Unscented Kalman Filter could work
-well. These are 
-non-trivial issues given that the assumption of Gaussianity is what allows
+fit, and so in some cases the Unscented Kalman Filter could work quite
+well. The assumption of Gaussianity is what allows
 the UKF to estimate the posterior using only the first and second moments;
-two parameters that don't uniquely describe most distributions.
+however, if this assumption is violated significant error will result.
 
@@
 To determine the amount of error incurred in a Gaussian estimate during
-a typical sample period, the states of BOLD equations were assigned according
-to four dimensional Gaussian. The states were then propagated through two
-seconds of simulation (a typical TR in FMRI) and then the resulting
-marginal distributions were compared with a Gaussian distribution. The 
-purpose is to determine the degree to which the results of simulation will
-result in non-Gaussian output, given a Gaussian input. This also demonstrates
+a typical sample period, I assigned the states of BOLD equations according
+to a four dimensional Gaussian. I then propagated the states through 
+two seconds of simulation (a typical TR in FMRI) and plotted the resulting
+marginal distributions against a Gaussian distribution. This also demonstrates
 the degree of nonlinearity present in the system. The parameters used are 
@@
 
-Notably $s_t$ has intentionally been set to a non-equilibrium, but physiologically
-plausible value. The value of $u$ is left at zero the entire time, so the 
-system will decay naturally (see \autoref{sec:BOLD Physiology}), though initializing 
-$s$ at a non-zero level will drive the system for several seconds. 
-\autoref{fig:transp1s} shows the results when the system is 
-left on for 100 milliseconds after setting the variables according to \autoref{tab:steptable}.
+In order to drive the system without input I intentionally set $s_t$ a non-steady
+state, but physiologically plausible, value. The value of $u$ is left at zero 
+the entire time, so the 
+system would decay naturally (see \autoref{sec:BOLD Physiology})
+\autoref{fig:transp1s} shows the results when the system after the
+system ran for 100 milliseconds. 
 The Q-Q plots fit will with a Gaussian, demonstrating that at this short
 time interval nonlinearities have not yet begun to effect the distribution.
-However, \autoref{fig:trans1s} us the result after 1 second, which is faster
+However, \autoref{fig:trans1s} shows the result after 1 second, which is faster
 than most FMRI scanners are capable of sampling at. At that range the tails 
-of the distributions for $v$ and $q$ are starting to deviate from the
-Gaussian distribution. As a result the uncertainty in $y$ is deviating from
+of the distributions for $v$ and $q$ started to deviate from the
+Gaussian distribution. As a result the uncertainty in $y$ deviated from
 the Gaussian distribution as well. This is important, because although 
 approximating the distribution with a Gaussian based on the first two
-moments will work in the short run, there will be residual error in the
-distribution.
+moments will worked in the short run, within a period less than typical
+measurement rates the distribution deviated from a Gaussian substantially.
 
-On the other hand, this effect is more limited if the initial variance is somewhat
-smaller. In tests with those cases, it took much longer for the nonlinearities to skew
-the distribution. That result could be encouraging to those looking to use the
-UKF, if the distributions are kept relatively thin.
-
-Another potential problem with the UKF is that typically the sigma points,
-used to estimate the posterior probability, $P(X(t) | X(t-1), u(t))$, are
-located on the main axes. As a result, the covariances are not allowed to 
-be effected by the state transition the same way the variances are. While
-this may be reasonable in low-dimensional systems, high dimensional systems
-have a much greater potential for interplay between the variables. While
-this problem is simple to fix (by using more sigma points off
-the main axes), there is a definite cost in complexity.
-
-Ultimately, there is a distinct possibility that the nonlinearities of 
-the BOLD model make Gaussian estimates unrealistic and thus less effective.
-More advanced tests where static variables such as $\alpha$ are varied
-as well could shed even more light on the issue. The trouble with using
-the UKF then to estimate parameters is that all eleven members of $X$ would
-be treated like a single joint Gaussian distribution which certainly 
-compound the issues of nonlinearity seen in \autoref{fig:transp1s} and 
-\autoref{fig:trans1s}
+Thus, even without introducing variation in the model parameters,
+a distinct nonlinearity and non-Gaussianity was present. 
+More advanced tests where parameters (especially $\alpha$) are varied
+would certainly introduce more error into the Gaussian estimate. 
+For this reason, estimating the posterior distribution using only 
+the first two moments, which is the basis for the UKF, is not wise.
 
@@
 \subsection{Hybrid Methods}
-In \cite{Riera2004}, a maximum
+In \cite{Riera2003}, a maximum
 likelihood method for innovation processes was used, as described by
-\cite{Ozaki1994}. \cite{Ozaki1994} uses a similar construction to a 
-Kalman filter, to break the time series into a series of innovations,
-for which Maximum Likelihood was performed. While this is probably the most
-likely solution to give the correct output, there are a few problems. First, every
-step in parameter space requires a recalculation of all the state variables. With
-two or three parameters this is fine, more than that, and calculations can become
-intractable. Additionally, there is no way to ensure a global minimum is reached.
-Although this is also true of the methods in \autoref{sec:Nonlinear Least Squares},
-those methods contained random elements designed to overcome this issue.
+\cite{Ozaki1994}. Ozaki et. al. used a similar construction to a 
+Kalman filter to filter the BOLD signal. 
+The method used in \cite{Riera2003} performs maximum likelihood on
+the innovations rather than the absolute signal levels. 
+By using a Local Linearization Filter,
+innovation noise such as noise in $\dot{f}$ is turned into simple 
+Gaussian noise. This approach is useful when the DC portion of the signal is
+of no use; however, it depends greatly on the type of signal. I found that
+this method was not very effective for the BOLD signal, when I used it 
+with the particle filter approach discussed in the next chapter.
 
@@
 the algorithm described in \cite{Johnston2008} does not depend on prior distributions.
-It is possible then that the bias toward the prior in other methods screwed the results. 
+It is possible then that the bias toward the prior in other methods skewed their results. 
 While \cite{Johnston2008} is certainly in the minority; further exhaustive studies 
@@
 \section{Conclusion}
-There is no ideal solution to solving this system of nonlinear
+Currently there is no ideal solution to solving this system of nonlinear
 equations. Exhaustive search type methods such as those employed by
@@
 The particle filter method proposed in the next section bears a strong 
-resemblance to the unscented Kalman filter; albeit with more point estimating
+resemblance to the unscented Kalman filter; albeit with more points estimating
 the posterior.

### 2010-07-24 doc/thesis/particlefilters.tex
fixed a lot of images etc.
Cleaned up particle filter and background.

@@
 Particle filters, a type of Sequential Monte Carlo (SMC) methods,
-are a powerful way of estimating the posterior probability distribution
-a set of parameters given a timeseries of measurements. Unlike Markov 
+are a powerful method for estimating the posterior probability distribution
+ of parameters given a timeseries of measurements. Unlike Markov 
 Chain Monte Carlo (MCMC) estimation, particle filters are designed for 
@@
 distributions are stored as an empircal distribution rather than the
-as the first two moments of a Gaussian. Thus particle filters are 
-preferred when the model is nonlinear, and thus non-gaussian.
+as the first two moments of a Gaussian. Thus, particle filters are 
+preferred when the model is nonlinear, and non-gaussian. This section
+is largely based on \cite{Arulampalam2002a} and \cite{Thrun2005}.
 
@@
 The idea of the particle filter is to build an empirical distribution
-out of a large number of parameter \emph{sets}, called particles. Each
+out of a large number of parameter sets, called particles. Each
 particle contains all the parameters and states needed to propagate
@@
 than parameter sets that give poor estimates. Although the reliance on
-a prior distribution can be troublesome, when the system being modeled
+a prior distribution are often troublesome, when the system being modeled
 has physical meaning, establishing reasonable ranges for parameters may be 
@@
 
-Suppose set or stream of measurements at discrete times are given, 
-$\{Y_k, k = 1, 2, 3, ... K\}$, where $K$ is infinite for a stream. 
+Suppose a set or stream of measurements at discrete times are given, 
+$\{Y_k, k = 1, 2, 3, ... K\}$, where $K$ may be infinite. 
 Because $k$ is a discrete time, let $t_k$ define the continuous
-time of $k$
+time of $k$.
 Suppose also that there is a hidden set of state variables,
-$X(t)$ that dictates the movement
-of $Y(t)$, although for most of the time dealings will be
+$X(t)$ that drives the value of $Y(t)$. Throughout this section
 with $X_k = X(t_k)$. The goal of the particle filter is to estimate the 
-\emph{distribution} of the
+distribution of the
 true parameters $\Theta$ that dictates the movement of $X(t)$.
@@
 
-Although not necessary for particle filters in general, a few  simplifying
-assumptions are made for this work. First, the systems are assumed to be 
+I will also make a few  simplifying assumptions for this work. 
+First, the systems are assumed to be 
 time invariant. This 
@@
 influence the output, which in the case of the BOLD signal is a good assumption.
-Also, noise is considered to be additive.
 Finally, because the only difference between the members of $X(t)$ and 
-$\Theta$ is an update function, from now on $x$ will contain 
+$\Theta$ is an update function, from now on $X$ will contain 
 $\Theta$. The assumptions now allow for a simplified version of the
@@
 \begin{eqnarray}
-\dot{X}_k = f(X_{k-1}, u_k) + \nu_x 
+\dot{X}_k = f(X_{k-1}, u_k, \nu_x)
 \label{eq:stateass}\\
-Y_k = g(X_k) + \nu_y 
+Y_k = g(X_k, \nu_y)
 \label{eq:measass}
@@
 which the initial $N_p$ particles are drawn. A particle contains a weight
-as well as an estimate of $X_k$, which as already stated, contains every
+as well as an estimate of $X_k$, which as previously mentioned, contains every
 variable needed to run the model. Then the prior is generated from a 
@@
 
-%todo, put back?
-%Once the probability, $P(x_k | x_{0:k-1}, y_{0:k-1})$ has been found
-%(initially this is just Mixture approximating the prior since no measurements are 
-%available and no previous probabilities are available), its possible to approximate
-%the probability between times when measurement is available, by shifting
-%the probability according the progression of the state equations. This is only 
-%an approximate, since integrating $\nu_d$ should increase uncertainty as
-%time without a measurement passes. 
-%
-%\begin{equation}
-%P(x(T+\Delta t)) \approx 
-%\sum_{i=1}^{N_p} w_i\delta\left(x - (x_i(T) + \int_T^{T+\Delta} \dot{x}_i(t) dt) \right)
-%\end{equation}
-
 \subsection{Weighting}
@@
 being correct given all the measurements up to time $k$. 
-Of course if there is a far off peak in
+Note that if there is a far off peak in
 the posterior that $q$ does not have support points in, there will 
-be quantization errors, and that part of the density can't be modeled. This is why
+be quantization errors, and that part of the density cannot be modeled. This is why
 it is absolutely necessary that $q$ fully covers $P(x^i_{0:k} | y_{0:k})$.
@@
 could be false; since later measurements may shed light on currently hidden
-changes in $x$. For practical applications though it is helpful assumption.
+changes in $x$. For practical applications though it is a helpful assumption.
 \begin{equation}
@@
 Although other importance density functions exist; for the particle filter
-used here, the standard importance density will be used: the modeled
+used here, the standard importance density will be used: the model
 prior.
@@
 fully constrained by $x_k$, and that $x_k$ is similarly fully 
-constrained by $x_{k-1}$, we are able to make the good assumptions that:
+constrained by $x_{k-1}$, I make the good assumptions that:
 \begin{eqnarray}
@@
 of the wide prior distribution (required for a proper discretization of a continuous
-distribution), there will be a significant proportion of particles with insignificant weights. 
+distribution), in short order there will be a significant proportion of particles 
+with insignificant weights. 
 While this does help
-describe the tails of the distribution, it means a lot of computation will be based.
+describe the tails of the distribution, it means a great deal of computation will be wasted.
 Instead, it would be preferable if most of the computation is spent on the most probable regions.
@@
 be equal.  The case where a large number of the weights have become irrelevantly small
-is called \emph{particle degeneracy}. In  \cite{Liu1998b}
+is called particle degeneracy. In  \cite{Liu1998b}
 an ideal calculation of the effective number of particles is found based on the 
@@
 still limited to the original particles, the number of \emph{unique} particles can only go down.
-This effect, dubbed \emph{particle impoverishment} can result in excessive quantization
+This effect, dubbed particle impoverishment can result in excessive quantization
 errors in the final distribution. However, there is a solution. Instead of sampling from the
@@
 the smoothing 
-distribution and added to the particle.  The process regularization is defined as:
+distribution and added to the particle.  The regularization process is defined as:
 
@@
 In reality, over smoothing
-will only become an issue if re-sampling is performed over and over. Thus
-if
-resampling is performed at every step then this could certainly caused problems.
+will only become an issue if re-sampling is performed very often. Thus
+if resampling is performed at every step then this could certainly cause problems.
 If the distribution is over-smoothed then the algorithm may not converge as rapidly;
@@
 to under smoothing, since over-smoothing simply slows convergence while 
-under-smoothing could result in false negatives.
-At the same time, as the number of dimensions in the state vector 
-increases, the standard deviation based approximation becomes less effective
- (\cite{Arulampalam2002a}). Because of the high dimensionality of the BOLD model,
+under-smoothing could leave gaps in the distribution.
+Moreover, because of the high dimensionality of the BOLD model,
 and limited measurements, it is helpful to have a broader bandwidth to explore the distribution.
@@
 of the potentially wide smoothing factor applied by regularized resampling, performing this
-step at every measurement would allow particles a great deal of mobility. This mobility is
-the opposite of convergence, which is why regularized resampling should only be done when
-$N_{eff}$ drops (say less than 50). Other than the periodic regularized
-resampling then, the regularized particle filter is nearly identical to the basic sampling
+step at every measurement would allow particles a great deal of mobility. This mobility could
+hinder convergence, which is why regularized resampling should only be done when
+$N_{eff}$ drops off (less than 50). Other than the periodic regularized
+resampling, the regularized particle filter is identical to the basic sampling
 importance sampling filter (SIS).
 
-Thus, with regularized resampling, it is possible to prevent both
+With regularized resampling, it is possible to prevent both
 particle degeneracy as well as particle impoverishment. An additional
@@
 simulated annealing where, as the algorithm approaches the end, the
-random steps get smaller and smaller. There is a potential risk when
+random steps get smaller and smaller. However, there is risk in 
 resampling. If for some reason the solution is not covered by the 
-new support the algorithm will fail to converge properly.
-
+new support the algorithm may not be able to reach the true value. 
 The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
@@
 $x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
-will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.
+will approach the width of the underlying uncertainty.
 
@@
 for updating weights. Ideally this weighting function would exactly 
-match the measurement error in
-the output. It is logical to assume this error is centered at zero and 
+match the measurement error in the output. 
+Typically it is assumed that this error is additive, centered at zero and 
 has a scale comparable to the signal levels.
@@
 complex circuits that convert the negative portion into positive and
-waste less energy but but here we will keep the system simple.
+waste less energy but here I will keep the system simple.
 Thus, let us consider a simple half wave rectifier circuit, shown in 
@@
 but when $u(t)G$ is greater than $v_t$, the output voltage will be set
-by $u(t)G$ and the capacitor will charge up. We will assume a simple
+by $u(t)G$ and the capacitor will charge up. I will assume a simple
 model for all the components, ignoring the complex nonlinear behavior
@@
 
-For this example it will be assumed that the transformer simply scales
+For this example I will assume that the transformer simply scales
 the input by a constant factor, $G$. As discussed in the \autoref{sec:Particle Filter Model}

### 2010-07-24 doc/thesis/particlefilters.tex
some changes to the abstract

@@
 and others farther away will be weighted lower. As the particle filter
-runs, weights will compound resulting in a distribution that asymptotically
-approaches the true joint distribution of the $X(t)$.  Of course, as I
-mentioned in \autoref{sec:Particle Filter Resampling}, particles weighted 
-zero do not significantly
-contribute to the empirical distribution, so re-sampling may be necessary.
+runs, weights will compound converging to a distribution that asymptotically
+approaches the true joint distribution of the $X(t)$.  As I
+mentioned in \autoref{sec:Particle Filter Resampling}, particles with 
+weights approaching zero do not significantly
+contribute to the empirical distribution, so re-sampling will be necessary.

### 2010-07-24 doc/thesis/thesis.tex
some changes to the abstract

--- 2010-07-16 doc/thesis/thesis.tex
@@
 
-Abstract Abstract Abstract ...
+Traditional methods of analyzing FMRI images use a linear combination of
+just a few static regressors. This works demonstrates an alternative
+approach using a physiologically inspired nonlinear model. By using a 
+particle filter to optimize the model parameters, the computation time
+is kept reasonable without requiring a linearization of the noise in the state
+variables. The activation results show regions similar those found in 
+SPM; however, there are some notable regions not detected by 
+SPM. Though the parameters selected by the particle filter based approach
+are more than sufficient to predict the BOLD response,
+more model constraints are needed to uniquely identify a single set
+of parameters. This ill-posed nature explains the large discrepancies
+found in other research that attempted to characterize the model parameters.
+For this reason the final distribution of parameters is more medically relevant
+than a single estimate. Because the output of the particle filter is 
+a full posterior probability, the reliance on the mean to estimate 
+parameters is not strictly necessary. This work presents
+not just a viable alternative to the traditional method of detecting
+activation, but an extensible technique of estimating the joint BOLD
+parameter distribution.
 

### 2010-08-06 doc/thesis/conclusion.tex
fixed up results.
added parameter map and parameter hist
changes in introduction primarily
some changes to the abstract
fixed some captions, and updated conclusion. Various cleanup in methods.

@@
 \label{sec:Conclusion}
-In this paper I have proposed and demonstrated the use of a particle filter
-for the estimation of BOLD model parameters. In doing so, it has been possible
-to estimate the BOLD time-series and judge the quality of the estimated time-
-series with respect to the true BOLD signal. As result, it is possible to use
+This work has proposed and demonstrated the use of a particle filter
+for the estimation of BOLD model parameters. Because many of the parameters
+were under constrained from the point of view of the BOLD response, a
+full posterior estimate was a more logical approach to parameter estimation
+than typical point estimates of parameters. The particle filter also converged
+quickly to correct output estimates, and thus provides a framework
+for future real-time FMRI experiments. Final BOLD estimates were also comparable
+to the results of SPM.
+
+* Overview of results
+** Parameters under-constrained
+** Output estimates good
+
+* Pros/Cons
+** Pro
+*** Non-parametric
+*** Real time
+*** Full Posterior
+*** More intuitive
+** Cons
+*** Computation longer than SPM
+*** Harder to interpret
+
+* Enhancements/Dehancements 
+** flatten prior
+** \% difference using spline?
+** "smart knots"
+** DC gain as a parameter
+** linearizing
+
+* Future works
+** Extensive test of quality of volterra kernel estimate from friston
+** Better de-trending
+** Automatic estimation of measurement error.
+
+As result, it is possible to use
 the particle filter to localize areas where a known stimuli most directly drives
@@
 
+Automatic detection of the noise level in the signal, to get a decent wieghting
+function. Large scale activation to get more parameter estimates. 
+Viscoelastic effects from \cite{Buston2004}, discussed in \cite{Deneux2006}.
+
 In conclusion, using particle filters to estimate the BOLD response are 

### 2010-08-06 doc/thesis/methods.tex
fixed some captions, and updated conclusion. Various cleanup in methods.

@@
 \label{sec:Methods}
-Although the particle filter I used for this work is a standard Regularized
+Although the particle filter  is a standard Regularized
 Particle filter, as described in \cite{Arulampalam2002a}, optimizing the 
@@
 used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
-\cite{Friston2000} only studied small regions; and most research with
-reasonable speed (including \cite{Friston2002b}) used these results as 
+\cite{Friston2000} only studied regions deemed active by the General 
+Linear Model; and most other research (including \cite{Friston2002b}) used these results as 
 the source for their priors. 
-The one exception is \cite{Johnston2008}, which came to an extremely different
+The one exception is \cite{Johnston2008}, which came to a extremely different
 distributions. For a particle filter, the choice of a prior is
-the single most important design choice. Consequently for this work,
-I wanted to be conservative; which meant going with the accepted result,
+the single most important design choice. A very wide prior will require
+more particles to be sufficiently dense, a very thin prior may miss
+the true parameters. Consequently, for this work it was natural
+to use priors that will give results consistent with similar tests. accepted result,
 \cite{Friston2000}. This constrains the usefulness of the model to

### 2010-08-06 doc/thesis/resultsB.tex
fixed some captions, and updated conclusion. Various cleanup in methods.

@@
 \subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
-\caption{Sagittal, coronal and axial slices of SPM results (\autoref{fig:hm_canon_spm_x} \autoref{fig:hm_canon_spm_y} 
-         \autoref{fig:hm_canon_spm_x}), as well as a series of axial slices, \autoref{fig:hm_spm}. Units
-         of activation are in Student's T-scores. Higher indicates higher assurance that the signal cannot
-         have occurred through noise alone.}
+\caption{SPM results. Units of activation are in Student's T-scores; higher indicates higher 
+        assurance that the signal cannot have occurred through noise alone. 
+        Sagittal, coronal and axial slices  are 
+        \autoref{fig:hm_canon_spm_x}, \autoref{fig:hm_canon_spm_y}, and
+         \autoref{fig:hm_canon_spm_z}, respectively. A series of axial slices are
+         shown in \autoref{fig:hm_spm}. }
 \label{fig:hm_canon_spm}
@@
 \subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
-\caption{Sagittal, coronal and axial (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
-         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
-         Units of match is normalized residual. The lowest (best) levels were $.7$.
-         The highest error shown is $1$.}
+\caption{Particle Filter results measured in normalized $\sqrt{MSE}$. 
+        Units of match is normalized residual where the  the lowest (best) levels shown are
+        $.7$ and the highest error shown (threshold) is $1$.
+        Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_pfilter_x},
+        \autoref{fig:hm_canon_pfilter_y}, and 
+         \autoref{fig:hm_canon_pfilter_z}, respectively. A series of axial slices are shown in
+         \autoref{fig:hm_pfilter}.  }
 \label{fig:hm_canon_pfilter}
@@
 \subfigure{\label{fig:scale_mi} \includegraphics[scale=.5]{images/scale4}}
-\caption{Sagittal, coronal and axial (\autoref{fig:hm_canon_pfilter_x} \autoref{fig:hm_canon_pfilter_y} 
-         \autoref{fig:hm_canon_pfilter_x}), as well as a series of axial slices, \autoref{fig:hm_pfilter}. 
+\caption{Particle Filter results measured in mutual information. 
          Units of match is bits (standard for base-2 Mutual Information). The highest (best) levels are 
-         $.42$. The worst shown is $.1$.}
+         $.42$ and the worst shown (threshold) is $.1$.
+        Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_mi_x},
+        \autoref{fig:hm_canon_mi_y}, and \autoref{fig:hm_canon_mi_z}, respectively. 
+        A series of axial slices are shown in \autoref{fig:hm_mi}.  }
 \label{fig:hm_canon_pfilter_mi}

### 2010-08-07 doc/thesis/background.tex
fixes in methods and results. Also big additions to conclusion

@@
 
-\subsection{Linear Approximation}
+\subsection{Polynomial Approximation}
+\label{sec:Background Linear Approximation}
 In \cite{Friston2002b}, a novel combination of linear and nonlinear modeling
@@
 descent are off limits; since the model is a true nonlinear dynamical
-system with no closed form solution (although \autoref{sec:Linear Approximation}
+system with no closed form solution (although \autoref{sec:Background Linear Approximation}
 circumvented this by calculating a Volterra Series approximation). Without
@@
 
+\subsection{Previous Particle Filter Approaches}
+In \cite{Murray2008} a particle filter based approach was used to integrate
+the BOLD equations. The method used in that work focused primarily on estimating
+the BOLD output and state equations as a nonlinear stochastic differential 
+equation. The primary difference between \cite{Murray2008} and this work is that
+\cite{Murray2008} took the parameters as a given. Thus, differences in the BOLD output
+were taken to be primarily driven by stochastic changes in the underlying state
+equations. Because the parameters were not allowed to change, the estimate of 
+the BOLD signal was not very good. The fact that the
+differences in BOLD response cannot be explained solely by stochastic elements 
+is important, however. The particle filter library created
+in that work, $dysii$, forms the basis for the particle filter used in this work, 
+and was extremely well designed. The work also clearly presents the particle filter;
+both its derivation and use.
+
 \section{Conclusion}

### 2010-08-07 doc/thesis/conclusion.tex
fixes in methods and results. Also big additions to conclusion

@@
 
-* Overview of results
-** Parameters under-constrained
-** Output estimates good
+% Overview of results
+%% Parameters under-constrained
+%% Output estimates good
+\section{Review of Results}
+The results unequivocally show that the parameters of the BOLD model are under
+constrained. While unsurprising given the sensitivity analysis in \cite{Deneux2006},
+it is still an important limitation when calculating parameters. Thus, other methods
+that depend on a single point estimate of the parameters, such as kalman filters
+or even least squares are limited to estimating the BOLD signal, but the estimate
+of the underlying parameters and state variables are suspect at best. Similarly, as the
+histograms in \autoref{sec:Multi-voxel Simulation} and \autoref{sec:Real Data Parameter Estimates}
+show, using the mean to estimate parameters also makes little sense in the particle filter
+case. Attributing the variance of estimate to the underlying parameter distribution would
+also be a mistake. While it is definitely true that parameters vary from person to person
+and region to region, the estimated distributions in \autoref{sec:Real Data Parameter Estimates} 
+are smoothed versions of those distributions. For this reason, any analysis of parameters should
+take the entire estimated posterior distribution into account, and perform F-tests on that
+distribution, rather than an idealized Gaussian. In this sense, particle filters represent
+an important step forward in BOLD parameter estimation. It is now clear that representing
+the uncertainty in parameters as a Gaussian is insufficient; so using a particle filter
+or Bayesian estimate of the posterior is not simply an enhancement, but a necessary precaution.
 
-* Pros/Cons
-** Pro
-*** Non-parametric
-*** Real time
-*** Full Posterior
-*** More intuitive
-** Cons
-*** Computation longer than SPM
-*** Harder to interpret
+Although point estimates of parameters are not dependable estimates of the 
+true parameters, they are by no means usefulness. Analysis of differences in parameters
+could be medically relevant. Additionally, a single estimate for parameters is still able
+to form a good estimate of the BOLD signal. This factor adds bridge to 
+earlier types of activation tests, such as the statistical
+parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
+those of mutual information closely resembled the results of SPM. While the activation
+tests were more sensitive, they were also more prone to false positive. However, it is
+worth noting that certain enhancements could be useful in reducing false positives; for instance
+by using the maximum likelihood of the final distribution, or even the median. The 
+usefulness of these techniques depend on the degree to which the algorithm converged, 
+and thus on the experimental design. Future work may shed further light on these
+techniques.
 
-* Enhancements/Dehancements 
-** flatten prior
-** \% difference using spline?
-** "smart knots"
-** DC gain as a parameter
-** linearizing
+% Pros/Cons
+%% Cons
+%%% Computation longer than SPM
+%%% Harder to interpret
+%% Pro
+%%% Non-parametric
+%%% Real time
+%%% Full Posterior
+%%% More intuitive
+\section{Particle Filter Approach}
+The Particle Filter algorithm was originally designed for on-line parameter 
+estimation. For this reason, there is no guarantee of optimality or even 
+convergence for finite measurements. However, for the BOLD nonlinear ODE there
+can be no guarantee of an optimal solution. A guaranteed local minimum,
+which nonlinear least squares provides, however would be helpful. One difficulty
+with the use of a particle filter with finite data is the calculation of a good
+weight function $P(y_k | x_k)$  that will converge at a decent rate. If the weight
+function does not sufficiently differentiate particles, the final distribution will 
+no be significantly different from the prior distribution. On the other hand, if the 
+weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
+measurements it is better to let the algorithm take longer to converge, because the
+convergence will be better (more accurate). The particle filter takes longer
+to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
+however, it is free from the uncertainty of whether a quadratic approximation is 
+sufficient for the BOLD model.
 
-* Future works
-** Extensive test of quality of volterra kernel estimate from friston
-** Better de-trending
-** Automatic estimation of measurement error.
+The particle filter certainly has significant advantages over other estimation procedures
+discussed in \autoref{sec:Prior Works}. The most important advantage is that it provides
+an estimate of the posterior probability, rather than a single estimate. While it is natural
+to want a simple estimate of parameters, such an estimate is impossible with this particular
+model. The results are more difficult to interpret, but this is a necessity. The fact that 
+the final distribution is not dependent on any particular distribution is also advantageous.
+Of particular note; the final distribution does not need to conform to any parametric
+distribution. While the particle filter was not fast for full brain calculations, its speed
+was sufficient on a quad core machine to perform real time calculations of small regions
+(approximate run time .27 seconds per voxel-measurement). Today it would be possible
+ to perform real time analysis of 10 voxels on an average quad core. The algorithm also scales
+well and does not require burdensome amounts of memory (approximately 11 megabytes). 
+For this reason this algorithm is perfect for extension to vector or video card
+based processors.
+
+A more practical concern with the particle filter is that it is mathematically
+simple. Only a basic understanding of Bayesian statistics is needed to understand
+how the particle filter works. This is in contrast to the Volterra based approach
+of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
+are needed for the particle filter. It is a common problem in traditional statistics 
+for assumptions to be unrealistic which means the results may be invalid or at 
+least in question. 
+Fewer and more realistic assumptions mean that the particle filter is more robust
+to unforeseen difficulties in FMRI data.
+
+% Enhancements/Dehancements 
+%% flatten prior
+%% \% difference using spline?
+%% "smart knots"
+%% DC gain as a parameter
+%% linearizing
+
+% Future works
+%% Extensive test of quality of volterra kernel estimate from friston
+%% Better de-trending
+%% Automatic estimation of measurement error.
 
@@
 function. Large scale activation to get more parameter estimates. 
-Viscoelastic effects from \cite{Buston2004}, discussed in \cite{Deneux2006}.
+Viscoelastic effects from \cite{Buxton2004}, discussed in \cite{Deneux2006}.
 

### 2010-08-07 doc/thesis/methods.tex
fixes in methods and results. Also big additions to conclusion

@@
 \cite{Friston2000} only studied regions deemed active by the General 
-Linear Model; and most other research (including \cite{Friston2002b}) used these results as 
+Linear Model; and most other research (including \cite{Friston2001}) used these results as 
 the source for their priors. 

### 2010-08-07 doc/thesis/results.tex
fixes in methods and results. Also big additions to conclusion

@@
 \section{Multi-voxel Simulation}
+\label{sec:Multi-voxel Simulation}
 To test the usefulness of the particle filter for I used a modified version of the FSL tool 

### 2010-08-07 doc/thesis/resultsB.tex
fixes in methods and results. Also big additions to conclusion

@@
 \section{Parameter Estimates}
+\label{sec:Real Data Parameter Estimates}
 Although the parameters are not uniquely identifiable by a single time-series, that

### 2010-08-07 doc/thesis/conclusion.tex
more work with the conclusion, just need to do future works

@@
 %% linearizing
+\section{Particle Filter Variations}
+There are quite a few areas where the particle filter could be tuned. While
+some of these areas were touched on \autoref{sec:Methods}, certain parts were
+not used in the analysis and so weren't mentioned.
+
+\subsection{Prior Distribution}
+Starting with a flattened
+prior (by setting weights to be non-uniform after the initial particles have
+been drawn) was not used in the final analysis. The reason for not flattening
+the prior was a practical issue with weighting points in a 
+7 dimensional joint distribution. Often differences in particle weights approached machine 
+precision, meaning that the flattened prior was actually far from flat. Instead
+the distribution became unpredictable, which made the results unpredictable. This
+is quantization issue that will exist unless the prior were made to be deterministic
+or the initial number particles was raised to computationally prohibitive levels.
+
+Wider prior
+
+\subsection{Detrending}
+A large number of de-trending methods were tested, but all had some major problem. 
+One viable method that could be utilized given the right experimental design is detrending
+based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
+constant to be added to the pre-processed signal for the BOLD model to fit properly.
+The disadvantage of this approach is that it could hide long fall times by normalizing them
+out. It also requires periodic breaks in the stimulus, which could reduce value of those
+samples for fitting purposes.
+
+Another potential way to deal with detrending is to 
+linearize. This would use the delta between measurements for fitting rather than the 
+direct value. This has the advantage of not requiring detrending and thus 
+gives nearly raw data to the particle filter
+for processing. The effectiveness of this method depends directly on the type of stimulus.
+In an event driven stimulus with sparse stimuli this could be extremely effective because
+the DC level has minimal data; however the longer high levels are held the less effective
+this will be. On the other hand, splines will also reduce the plateaus, so its possible
+linearizing may be just as effective. I found in tests that large drift-low white noise
+type signals performed much better with a linearization approach, as one might expect. 
+The results were equal to or worse in real data; but that was only for one particular
+stimulus sequence.
+
+A subtle difference that could be made in detrending is how the percent difference
+is calculated. In this work the spline was subtracted, and then the result
+was divided by the average of the initial signal. A more correct method may be dividing
+by the spline value at that particular point. The reason for not dividing by the spline
+at that point is that it could be less stable, and in a sense the trend has already
+been removed. The difference is not particularly large though 
+(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.
+
+Finally, rather than adding a constant to the detrended data before
+applying the particle filter, a DC gain parameter could be added to the model. In
+tests, this could work well, however, it often did not. The problem with this
+could simply be the addition of another degree of freedom without any increase in
+the number of particles. Increasing the number of particles further though
+can be computationally intractable. Thus, while this is in a sense the correct
+solution, it is an impractical one.
 
@@
 %% Automatic estimation of measurement error.
+%todo
 

### 2010-08-09 doc/thesis/conclusion.tex
added a signle section title and recompiled
updated conclusion

@@
 Although point estimates of parameters are not dependable estimates of the 
-true parameters, they are by no means usefulness. Analysis of differences in parameters
+true parameters, they are by no means useless. Analysis of differences in parameters
 could be medically relevant. Additionally, a single estimate for parameters is still able
-to form a good estimate of the BOLD signal. This factor adds bridge to 
+to form a good estimate of the BOLD signal. This factor forms a bridge to 
 earlier types of activation tests, such as the statistical
@@
 %%% More intuitive
-\section{Particle Filter Approach}
+\section{Review of Particle Filter Approach}
 The Particle Filter algorithm was originally designed for on-line parameter 
 estimation. For this reason, there is no guarantee of optimality or even 
-convergence for finite measurements. However, for the BOLD nonlinear ODE there
-can be no guarantee of an optimal solution. A guaranteed local minimum,
-which nonlinear least squares provides, however would be helpful. One difficulty
+convergence for finite measurements. However, for the BOLD nonlinear ODE 
+this is less of a concern than it might first appear to be. For this
+particular problem there can be no guarantee of a global minimum, and although
+other techniques guarantee a local minimum, the particle filter doesn't settle 
+to one of these local minima because it can settle into multiple ones. This
+difference typifies the major difference between the particle filter and 
+competing approaches.
+
+One difficulty
 with the use of a particle filter with finite data is the calculation of a good
@@
 
-A more practical concern with the particle filter is that it is mathematically
-simple. Only a basic understanding of Bayesian statistics is needed to understand
+A more practical benefit with the particle filter is that it is mathematically
+simple. An understanding of Bayesian statistics is all that is necessary to understand
 how the particle filter works. This is in contrast to the Volterra based approach
@@
 %% Automatic estimation of measurement error.
-%todo
-
-As result, it is possible to use
-the particle filter to localize areas where a known stimuli most directly drives
-neural activation. In the future it will be possible to use the estimated states
-($v,q,s,f$) to drive other models and learn more about how regions of the brain
-interact.
+%% Using s as the input to other regions. 
+%% Extensive parameter estimation to find a correct prior
+%% Increased constraints on parameters using physiological data
+%% Incorporating Blood Flow/Volume data as separate measurement
+\section{Future Works}
 

### 2010-08-11 doc/thesis/conclusion.tex
added a discussion and future work section. discussion and conclusion need more work.

--- 2010-08-07 doc/thesis/conclusion.tex
@@
 full posterior estimate was a more logical approach to parameter estimation
-than typical point estimates of parameters. The particle filter also converged
+than a single point estimate of the parameters. The particle filter also converged
 quickly to correct output estimates, and thus provides a framework
@@
 to the results of SPM.
-
-% Overview of results
-%% Parameters under-constrained
-%% Output estimates good
-\section{Review of Results}
-The results unequivocally show that the parameters of the BOLD model are under
-constrained. While unsurprising given the sensitivity analysis in \cite{Deneux2006},
-it is still an important limitation when calculating parameters. Thus, other methods
-that depend on a single point estimate of the parameters, such as kalman filters
-or even least squares are limited to estimating the BOLD signal, but the estimate
-of the underlying parameters and state variables are suspect at best. Similarly, as the
-histograms in \autoref{sec:Multi-voxel Simulation} and \autoref{sec:Real Data Parameter Estimates}
-show, using the mean to estimate parameters also makes little sense in the particle filter
-case. Attributing the variance of estimate to the underlying parameter distribution would
-also be a mistake. While it is definitely true that parameters vary from person to person
-and region to region, the estimated distributions in \autoref{sec:Real Data Parameter Estimates} 
-are smoothed versions of those distributions. For this reason, any analysis of parameters should
-take the entire estimated posterior distribution into account, and perform F-tests on that
-distribution, rather than an idealized Gaussian. In this sense, particle filters represent
-an important step forward in BOLD parameter estimation. It is now clear that representing
-the uncertainty in parameters as a Gaussian is insufficient; so using a particle filter
-or Bayesian estimate of the posterior is not simply an enhancement, but a necessary precaution.
-
-Although point estimates of parameters are not dependable estimates of the 
-true parameters, they are by no means usefulness. Analysis of differences in parameters
-could be medically relevant. Additionally, a single estimate for parameters is still able
-to form a good estimate of the BOLD signal. This factor adds bridge to 
-earlier types of activation tests, such as the statistical
-parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
-those of mutual information closely resembled the results of SPM. While the activation
-tests were more sensitive, they were also more prone to false positive. However, it is
-worth noting that certain enhancements could be useful in reducing false positives; for instance
-by using the maximum likelihood of the final distribution, or even the median. The 
-usefulness of these techniques depend on the degree to which the algorithm converged, 
-and thus on the experimental design. Future work may shed further light on these
-techniques.
-
-% Pros/Cons
-%% Cons
-%%% Computation longer than SPM
-%%% Harder to interpret
-%% Pro
-%%% Non-parametric
-%%% Real time
-%%% Full Posterior
-%%% More intuitive
-\section{Particle Filter Approach}
-The Particle Filter algorithm was originally designed for on-line parameter 
-estimation. For this reason, there is no guarantee of optimality or even 
-convergence for finite measurements. However, for the BOLD nonlinear ODE there
-can be no guarantee of an optimal solution. A guaranteed local minimum,
-which nonlinear least squares provides, however would be helpful. One difficulty
-with the use of a particle filter with finite data is the calculation of a good
-weight function $P(y_k | x_k)$  that will converge at a decent rate. If the weight
-function does not sufficiently differentiate particles, the final distribution will 
-no be significantly different from the prior distribution. On the other hand, if the 
-weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
-measurements it is better to let the algorithm take longer to converge, because the
-convergence will be better (more accurate). The particle filter takes longer
-to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
-however, it is free from the uncertainty of whether a quadratic approximation is 
-sufficient for the BOLD model.
-
-The particle filter certainly has significant advantages over other estimation procedures
-discussed in \autoref{sec:Prior Works}. The most important advantage is that it provides
-an estimate of the posterior probability, rather than a single estimate. While it is natural
-to want a simple estimate of parameters, such an estimate is impossible with this particular
-model. The results are more difficult to interpret, but this is a necessity. The fact that 
-the final distribution is not dependent on any particular distribution is also advantageous.
-Of particular note; the final distribution does not need to conform to any parametric
-distribution. While the particle filter was not fast for full brain calculations, its speed
-was sufficient on a quad core machine to perform real time calculations of small regions
-(approximate run time .27 seconds per voxel-measurement). Today it would be possible
- to perform real time analysis of 10 voxels on an average quad core. The algorithm also scales
-well and does not require burdensome amounts of memory (approximately 11 megabytes). 
-For this reason this algorithm is perfect for extension to vector or video card
-based processors.
-
-A more practical concern with the particle filter is that it is mathematically
-simple. Only a basic understanding of Bayesian statistics is needed to understand
-how the particle filter works. This is in contrast to the Volterra based approach
-of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
-are needed for the particle filter. It is a common problem in traditional statistics 
-for assumptions to be unrealistic which means the results may be invalid or at 
-least in question. 
-Fewer and more realistic assumptions mean that the particle filter is more robust
-to unforeseen difficulties in FMRI data.
-
-% Enhancements/Dehancements 
-%% flatten prior
-%% \% difference using spline?
-%% "smart knots"
-%% DC gain as a parameter
-%% linearizing
-
-% Future works
-%% Extensive test of quality of volterra kernel estimate from friston
-%% Better de-trending
-%% Automatic estimation of measurement error.
-
-As result, it is possible to use
-the particle filter to localize areas where a known stimuli most directly drives
-neural activation. In the future it will be possible to use the estimated states
-($v,q,s,f$) to drive other models and learn more about how regions of the brain
-interact.
-
-A limitation often reached in previous works was an inconsistency of
-parameter estimates, most likely because of covariance between the model parameters.
-Although individual studies got consistent results, those results often differed
-widely from other similar studies. The reason for this is rather clear from the
-simulation results in \autoref{sec:SimLowNoise}. There is a significant amount of
-trade off between parameters to the point that a signal set of parameters
-is most likely not possible to derive from the BOLD response alone. It
-will therefore be beneficial to combine BOLD studies with cerebral blood
-flow or cerebral blood volume studies to gain multiple more measurements and
-further constrain the model. That said, the benefit of the particle filter
-is that it provides a full posterior distribution at the final time step.
-As such the true solution should be encoded in the particle filter's final 
-distribution given priors that encompass the true parameters. This is beneficial
-in two ways; first, if, after the fact, some parameter becomes known 
-from outside observation, it is then possible to construct a new probability
-conditional on the new observation. Secondly the results from multiple runs
-may be reasonably concatenated, using the final distribution from the previous
-run as the prior distribution of the next run. Rather than simply providing a 
-staring point for parameters to converge from, it in fact continues convergence
-from the previous stopping point.
-
-Although many versions of the BOLD model exist, and it is tempting to use 
-more detailed models; from the results found here the issue of bias error
-from the BOLD model is not the biggest concern. Clarifying the distributions
-of parameters for the prior should be the first concern; currently no multi-patient
-full-volume studies have been done to estimate parameters.
-One future study that would be beneficial in this way  would
-be an extensive study of what the priors should truly be. Although \cite{Friston2000}
-gives an estimate of what is thought to be reasonable values, and later studies
-published their estimate of the distributions, given the interplay between
-parameters it is unlikely that these priors are true to actual distribution
-that occurs in vivo. As I mentioned previously, the addition of simultaneous
-flow or volume measurements are another potentially powerful way to further confine the model,
-and thus deal with the elasticity of parameters. As I mentioned in 
-\autoref{sec:BackgroundConclusion}, a chief advantage of using physiologically
-plausible models is that such data may in fact be added with relative ease.
-
-Automatic detection of the noise level in the signal, to get a decent wieghting
-function. Large scale activation to get more parameter estimates. 
-Viscoelastic effects from \cite{Buxton2004}, discussed in \cite{Deneux2006}.
-
-In conclusion, using particle filters to estimate the BOLD response are 
-a powerful method of fitting to noisy data. The technique also
-holds great promise as extensible platform to build more advanced models
-and techniques on top of. Integrating further information is necessary to
-move beyond the traditional Statistical Parametric Mapping and moving
-toward biologically and medically relevant FMRI scanning techniques.

### 2010-08-11 doc/thesis/discussion.tex
added a discussion and future work section. discussion and conclusion need more work.

\chapter{Discussion}
% Overview of results
%% Parameters under-constrained
%% Output estimates good
The results unequivocally show that the parameters of the BOLD model are under
constrained. While unsurprising given the sensitivity analysis in \cite{Deneux2006},
it is still an important limitation when calculating parameters. Thus, other methods
that depend on a point estimate of the parameters, such as least squares 
or kalman filters (which uses the first two moments) are limited. While estimates of
the  BOLD signal may still be correct, the estimate
of the underlying parameters and state variables are suspect. Similarly, as the
histograms in \autoref{sec:Multi-voxel Simulation} and \autoref{sec:Real Data Parameter Estimates}
show, using the mean to estimate parameters also makes little sense in the particle filter
case. Attributing the variance of estimate to the underlying parameter distribution would
also be a mistake; given that variance persisted even in simulated data. While parameters 
definitely vary from person to person and cortex to cortex, the estimated distributions 
in \autoref{sec:Real Data Parameter Estimates} are wider than the true underlying 
distribution. For this reason and because of the nonlinear relation between
parameters, any analysis should take the entire posterior estimate into account, 
rather than relying on an idealized Gaussian. In this sense, particle filters represent
an important step forward in BOLD parameter estimation. Representing
the uncertainty in parameters as a Gaussian is insufficient; so using a particle filter
or Bayesian estimate of the posterior is not simply an enhancement, but a necessary precaution.

Although point estimates of parameters are not dependable estimates of the 
true parameters, they still useful. Analysis of differences in parameters
could be medically relevant for instance. Additionally, a single estimate for parameters is still able
to form an accurate estimate of the BOLD signal. This trait bridges the gap with earlier types 
of activation tests, such as the statistical
parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information, closely resembled the results of SPM. While the activation
tests were more sensitive, they were also more prone to false positive. However, it is
worth noting that certain enhancements could be useful in reducing false positives; for instance
by using the maximum likelihood, or median of the final distribution.  Future work may 
shed further light on these techniques.

% Pros/Cons
%% Cons
%%% Computation longer than SPM
%%% Harder to interpret
%% Pro
%%% Non-parametric
%%% Real time
%%% Full Posterior
%%% More intuitive
\section{Particle Filter Review}
The Particle Filter algorithm was originally designed for on-line parameter 
estimation. For this reason, there is no guarantee of optimality or even 
convergence for finite measurements. However, for the BOLD nonlinear ODE 
this is less of a concern than it might first appear to be. For this
particular problem there can be no guarantee of a global minimum, and although
other techniques guarantee a local minimum, the particle filter doesn't settle 
to one of these local minima because it can settle into multiple ones. This
difference typifies the major difference between the particle filter and 
competing approaches.

One difficulty
with the use of a particle filter with a finite number of datapoint is the calculation of a good
weight function $P(y_k | x_k)$  that will converge reasonably quickly. If the weight
function does not sufficiently differentiate particles, the final distribution will 
no be significantly different from the prior distribution. On the other hand, if the 
weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
measurements it is better to let the algorithm take longer to converge, because the
convergence will be better (more accurate). The particle filter takes longer
to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
however, it is free from the uncertainty of whether a quadratic approximation is 
sufficient for the BOLD model.

The particle filter certainly has significant advantages over other estimation procedures
discussed in \autoref{sec:Prior Works}. The most important advantage is that it provides
an estimate of the posterior probability, rather than a single estimate. While it is natural
to want a simple estimate of parameters, such an estimate is impossible with this particular
model. The results are more difficult to interpret, but this is a necessity. The fact that 
the final distribution is not dependent on any particular distribution is also advantageous.
Of particular note; the final distribution does not need to conform to any parametric
distribution. While the particle filter was not fast for full brain calculations, its speed
was sufficient on a quad core machine to perform real time calculations of small regions
(approximate run time .27 seconds per voxel-measurement). Today it would be possible
 to perform real time analysis of 10 voxels on an average quad core. The algorithm also scales
well and does not require burdensome amounts of memory (approximately 11 megabytes). 
For this reason this algorithm is perfect for extension to vector or video card
based processors.

A more practical benefit with the particle filter is that it is mathematically
simple. An understanding of Bayesian statistics is all that is necessary to understand
how the particle filter works. This is in contrast to the Volterra based approach
of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
are needed for the particle filter. It is a common problem in traditional statistics 
for assumptions to be unrealistic which means the results may be invalid or at 
least in question. 
Fewer and more realistic assumptions mean that the particle filter is more robust
to unforeseen difficulties in FMRI data.

\chapter{Future Work}
% Enhancements/Dehancements 
%% prior
%%% flatten prior
%%% Larger variance
%% Detrending
%%% \% difference using spline?
%%% "smart knots"
%%% DC gain as a parameter
%%% linearizing
%% Experimental Changes
%%% Variations in stimuli
%%% longer timeseries or artificial
%%% Automatic estimation of measurement error.
\section{Improving the Particle Filter}
\label{sec:Particle Filter Variations}

There are a few areas where future research may improve upon the current
work. The first is modifying the Prior 
Distribution. The choice in this paper to use the distribution calculated in
\cite{Friston2000}, may not be optimal and so is open to modification. The
second major area is in preprocessing, in particular the method used to remove
drift in the FMRI signal. This is an area of ongoing research and so future
algorithms could be strengthened by applying the latest technique rather than a 
spline. Third, the experiment itself may be modified to improve results.
And the last method is improving the , by using a more
advanced version.

\subsection{Prior Distribution}
The prior distribution used in this work is listed in \autoref{tab:Prior}, 
and is based on the findings of \cite{Friston2000}. Unfortunately, that
result depended on a quadratic approximation (Volterra Series) which has
not been extensively tested (at least not in a published work). As such,
the prior is probably not complete. Additionally, the current prior is
based only around the BOLD output, which is why it is capable of generating
good estimates of the BOLD signal. However, it would be advantagous to have
in-vivo estimates of the actual parameters. Better estimates could be found
using population studies with additional measurements as discussed in 
\autoref{sec:Sideways Measurements}. Changes to this prior could
boost the power of the BOLD particle filter algorithm described here.

Starting with a flattened
prior (by setting weights to be non-uniform after the initial particles have
been drawn) was not used in the final analysis. The reason for not flattening
the prior was a practical issue with weighting points in a 
7 dimensional joint distribution. Often differences in particle weights approached machine 
precision, meaning that the flattened prior was actually far from flat. Instead
the distribution became unpredictable, which made the results unpredictable. This
is quantization issue that will exist unless the prior were made to be deterministic
or the initial number particles was raised to computationally prohibitive levels.

Another improvement that could be made to the prior is increasing the initial variance.
This would allow the particle filter to learn a wider range of parameters; at the cost
of decreased convergence rate and support density. In tests, both real and simulated,
increasing the variance tended to allow the particle filter to over-smooth 
(ex \autoref{fig:badfit_param1}). In 
essence the particle filter tended to converge to a mean where the time constants
dominated the results by smoothing out all the peaks and troughs. The result is that
no other parameters alter the estimated BOLD signal. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/badfit_param1}
\caption{Large variance in time constants over-smoothing.}
\label{fig:badfit_param1}
\end{figure}
In \autoref{fig:badfit_param1}, the particle filter still had not fully converged, and
given more measurements it could still converge to a better estimate. Thus, increases
in the prior variance necessarily must be accompanied by more measurements,
although re-presenting data could be used to artificially extend the measurements.

\subsection{Detrending}
A large number of de-trending methods were tested, but all had some major problem. 
One viable method that could be utilized given the right experimental design is detrending
based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
constant to be added to the pre-processed signal for the BOLD model to fit properly.
The disadvantage of this approach is that it could hide long fall times by normalizing them
out. It also requires periodic breaks in the stimulus, which could reduce value of those
samples for fitting purposes.

Another potential way to deal with detrending is to 
linearize. This would use the delta between measurements for fitting rather than the 
direct value. This has the advantage of not requiring detrending and thus 
gives nearly raw data to the particle filter
for processing. The effectiveness of this method depends directly on the type of stimulus.
In an event driven stimulus with sparse stimuli this could be extremely effective because
the DC level has minimal data; however the longer high levels are held the less effective
this will be. On the other hand, splines will also reduce the plateaus, so its possible
linearizing may be just as effective. I found in tests that large drift-low white noise
type signals performed much better with a linearization approach, as one might expect. 
The results were equal to or worse in real data; but that was only for one particular
stimulus sequence.

A subtle difference that could be made in detrending is how the percent difference
is calculated. In this work the spline was subtracted, and then the result
was divided by the average of the initial signal. A more correct method may be dividing
by the spline value at that particular point. The reason for not dividing by the spline
at that point is that it could be less stable, and in a sense the trend has already
been removed. The difference is not particularly large though 
(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.

Finally, rather than adding a constant to the detrended data before
applying the particle filter, a DC gain parameter could be added to the model. In
tests, this could work well, however, it often did not. The problem with this
could simply be the addition of another degree of freedom without any increase in
the number of particles. Increasing the number of particles further though
can be computationally intractable. Thus, while this is in a sense the correct
solution, it is an impractical one.

\subsection{Experimental Changes}
\label{sec:Sideways Measurements}
One definite way of improving the results of the particle filter is more 
measurements. While increasing the sample rate of FMRI scanners may
not be possible, simultaneous measurements of volume and flow is
possible \cite{Hu2009}, albeit at 9T in a cat. Just one of those measurements
though would be extremely powerful when incorporated into the BOLD
model. By adding another measurement, especially for one of the other state
variables (like blood volume and blood flow), the
variance in the parameter estimates would certainly drop. Of course, more measurements
may be gained by simply performing longer FMRI tests, or concatenating several 
together. Even more basic, its possible to artificially increase the number of 
measurements by presenting them the time-series several times. 
This activity is similar to the process used in neural networks,
and gives the particle filter longer to converge to the optimal estimate. On the 
downside, this increases runtime and artificially reduces variance. 
Thus it does not necessarily improve the results. In tests, it did reduce the
final covariance but did not lead to better estimates.

Additionally, given the non-linearities in the system, differences in stimuli
can make a huge difference in the observability of parameters. The mentality 
for using a physiologically based nonlinear model for BOLD signal is to model 
those nonlinearities. Its logical then that nonlinear parameters may not
be identifiable when the input is primarily an impulse response. Therefore,
a wide range of inputs may shed further light on the parameters than found 
in this work.

As discussed in \autoref{sec:Methods Weighting Function}, choosing a weighting
function is difficult. Automatically estimating measurement error could improve
the quality of the particle filter results. Although I made some attempts to do
this, finding a generic, consistent solution is complex, and often depends on the
experimental design. This problem strikes at the very of heart of the difficulty
in analyzing FMRI, and so it is not likely to be solved soon.

% Future works
%% Using s as the input to other regions. 
%% Comparison of posterior of parameters
\section{Applications}
There are a number of advantages to the particle filter approach presented
here. In the past FMRI data has been analyzed strictly for determining
correlation between a stimulus and response. With this new method
the correlation is simply a means to determining the joint posterior distribution
of the parameters. While only regions that correlate with the input will 
be calculable, this method exploits that correlation to constrain the 
prior distribution of the parameters. The resulting distribution, while difficult 
to visualize because of the high-dimensionality, could nevertheless be 
correlated with neural pathologies. Despite the fact that the parameters
are under-determined, the final distribution is still a reduction in the
uncertainty of the parameters. This reduction in uncertainty naturally
contains information which may be exploited based on non-parametric 
statistical analyses.

Because of the limitations present in every image modality in neuroimaging,
its becoming increasingly clear that combining data from multiple sources
will be necessary to push neurology forward. In order to do so however, the
output of each source needs to fully represent what information that source
can provide. Combining the sources using Bayesian statistics is promising
yet often difficult because full probability distributions are hard to come by.
However, in this case, the particle filter provides a full posterior which
is extremely versatile. Therefore, future works will easily be able to
plug in data from multiple sources if they all output Bayesian posteriors.
For instance, if two different modalities have calculated the probability
distribution of neural efficiency, those two beliefs may be combined into
one conditional belief for the probability of neural efficiency.

An advantageous aspect of using a physiological model such as this, is that
it permits estimates of otherwise hidden parameters. In particular the BOLD
model gives an estimate of the value of the flow inducing signal, $s$. Having
this value available opens up new avenues for determining inter-regional
dependencies. All the regions for which the value of $s$ are trustworthy
correlate closely with the original stimuli, so determining the connection
between their values of $s$ would be less than edifying. Instead, the values
of $s$, which in some sense is the activation in a particular region, could
be used to drive other inputs. Thus, the particle filter could be re-run
with the time-series of a particular voxel's $s$ value as an additional stimulus.
In this way, it could be possible to determine chains of events. This is just
one possible benefit being able to determine the time course of the hidden
state variables present in the BOLD equations; the potential benefits of being
able to determine this information is limitless. Of course, improving the quality
of the parameter estimates is necessary if such information is to be fully
trusted.

### 2010-08-11 doc/thesis/conclusion.tex
more thesis gah

@@
 \label{sec:Conclusion}
-This work has proposed and demonstrated the use of a particle filter
-for the estimation of BOLD model parameters. Because many of the parameters
-were under constrained from the point of view of the BOLD response, a
-full posterior estimate was a more logical approach to parameter estimation
-than a single point estimate of the parameters. The particle filter also converged
-quickly to correct output estimates, and thus provides a framework
-for future real-time FMRI experiments. Final BOLD estimates were also comparable
-to the results of SPM.
+This work has demonstrated the use of the particle filter to
+learn parameters of the BOLD model. Since the inception of the
+BOLD equations, many attempts have been made to learn the parameters,
+given FMRI data. These attempts have typically either been extremely 
+slow or made extensive approximations. While the particle filter method
+is not quick, 60 seconds to analyze each voxel is well within the capabilities
+of the typical research lab.
+
+One significant finding of this work, that would not be clear without 
+calculating a full posterior distribution, is the interplay 
+between the parameters. The results of simulations clearly demonstrate
+that identifying a single set of parameters is not possible with this 
+model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
+at this, the current work clearly demonstrates this fact. Therefore,
+any single estimate of parameters is insufficient for analysis. As such,
+approaches to the BOLD model that do not treat the parameters as distributions
+will not be able to overcome the inherent indeterminability of the
+parameters. Besides the Unscented Kalman Filter, this is the only approach
+that accomplishes this task without recalculating the parameters over and 
+over to build a distribution. The Unscented Kalman Filter, though, is limited
+to Gaussian estimation.

### 2010-08-11 doc/thesis/discussion.tex
more thesis gah

@@
 those of mutual information, closely resembled the results of SPM. While the activation
-tests were more sensitive, they were also more prone to false positive. However, it is
+tests were more sensitive, there were some false positive, though at this point it is
+difficult to quantify. It is
 worth noting that certain enhancements could be useful in reducing false positives; for instance
@@
 for assumptions to be unrealistic which means the results may be invalid or at 
-least in question. 
-Fewer and more realistic assumptions mean that the particle filter is more robust
-to unforeseen difficulties in FMRI data.
+least in question.  Fewer and more realistic assumptions mean that the particle 
+filter is more robust to unforeseen difficulties in FMRI data.
 

### 2010-08-11 doc/thesis/resultsB.tex
more thesis gah

@@
 with SPM's location and fit.
+
+\section{FMRI Configuration}
+For the FMRI data discussed in \autoref{sec:RealData}, tests were 
+performed on a right handed volunteer using a Gradient EPI sequence
+on a 1.5T GE scanner with a 2.1s TR. Slices were 5mm thick and echo
+time was (TODO). The subject was then presented with either a single or
+double flash, at randomly dispersed time points over the course of
+5 minutes, 10 seconds. The FMRI began 18.9 seconds
+before the beginning of the experiment, to allow for transients 
+in to image to settle out. The timing of the flashes are shown in 
+\autoref{tab:flashes}, with time 0 corresponding to the beginning of
+the 10th TR (so 9 images were dropped).
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c || c |}
+\hline 
+Single Flash Times & Double Flash Times \\
+\hline 
+.7 & 2\\
+2.7 & 21\\
+5.7 & 28\\
+12.7 & 52\\
+\hline 
+\end{tabular}
+\caption{Timing of single and double flashes.}
+\label{tab:flashes} 
+\end{table}
+
+After dropping the first 9 volume, each of the remaining volumes was
+co-registered with the first volume in time. At this point the SPM method
+diverges from the experimental method. For the particle filter, detrending 
+was then applied as discussed in \autoref{sec:Detrend} and the resulting
+data was processed with the particle filter. To generate the SPM output
+in this section, the co-registered data was passed through a high pass
+filter with a cut off of (TODO). The data was then smoothed spatially 
+with a Gaussian kernel (FWHM 8x8x8 voxels). A Statistical Parametric Map
+was then generated using SPM8. All settings used in SPM8 were the defaults
+for analyzing FMRI data.
 

### 2010-08-11 doc/thesis/conclusion.tex
more in the conclusion

@@
 learn parameters of the BOLD model. Since the inception of the
-BOLD equations, many attempts have been made to learn the parameters,
-given FMRI data. These attempts have typically either been extremely 
-slow or made extensive approximations. While the particle filter method
+BOLD equations, many attempts have been made use FMRI data to 
+to learn these parameters. These attempts have typically either been extremely 
+slow or relied on extensive assumptions. While the particle filter method
 is not quick, 60 seconds to analyze each voxel is well within the capabilities
-of the typical research lab.
+of the typical research lab. Previous attempts have also treated
+the problem as if there were a single solution.
 
@@
 parameters. Besides the Unscented Kalman Filter, this is the only approach
-that accomplishes this task without recalculating the parameters over and 
-over to build a distribution. The Unscented Kalman Filter, though, is limited
-to Gaussian estimation.
+that accomplishes this task without repeatedly calculating the parameters 
+to build a distribution. The Unscented Kalman Filter, of course, is limited
+to Gaussian estimation which limits its ability to estimate the posterior.
+
+The primary reason this method has not been used before is the high 
+dimensionality of the system. In \cite{Murray2008} the idea of learning
+the parameters is floated as the correct method, yet 
+learning 7 parameters with a Monte Carlo method is often intractable. The 
+concern was that so many parameters creates a prohibitively large 
+search space, that requires too many particle to learn properly. To overcome
+this difficulty, instead of starting the algorithm with 1000 particles,
+the initial number of particles was set to 16,000. This meant that
+when the search space was the largest, the number of particles was
+sufficiently dense to represent the prior distribution. Then resampling
+was performed for the first time, the number of particles was dropped, 
+since the weight of most the particles had dropped to 0. The result is
+a particle filter algorithm that spends computing resources only where
+they are really needed.
+
+Many of the approaches to determining areas of neural activation aim
+to be Bayesian, yet ultimately must make limiting assumptions about
+the distributions that turn the approach into classical statistics. 
+This approach is a true Bayesian non-parametric approach that given
+enough measurements will approach the true probability distribution
+of the parameters. While the conclusion that the parameters are not
+uniquely identifiable is a disappointing in light years of research 
+attempting to learn these parameters, it in fact further demonstrates
+the need to estimate probability distribution rather than a single 
+parameter estimate.

### 2010-08-11 doc/thesis/conclusion.tex
conclusion-ed

@@
 of the parameters. While the conclusion that the parameters are not
-uniquely identifiable is a disappointing in light years of research 
+uniquely identifiable is a disappointing one in light years of research 
 attempting to learn these parameters, it in fact further demonstrates
 the need to estimate probability distribution rather than a single 
-parameter estimate.
+parameter estimate. At the same time, the output of the particle filter
+is still about to give good estimates of the BOLD output, and not
+dependent on heavy Gaussian smoothing. Therefore, all the current 
+experimental paradigms to determine regional activity are still viable.
+To borrow a computer term, the particle filter algorithm is backward 
+compatible with the current methods.
+
+Concluding, the particle filter provides comparable performance with
+conventional tests, but also provides a platform for a wide range of
+future uses beyond what is possible with the methods applied in the
+past.

### 2010-08-12 doc/paper/paper.tex
added the paper for my research

\documentclass{article}
\usepackage[left=1cm,top=2cm,right=1cm,nohead,nofoot]{geometry}

%\usepackage{times}
%\usepackage{tikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}

%\doublespacing

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\qed}{\\$\Box$}
\newcommand{\qle}{\stackrel{?}{\le}}
\newcommand{\qeq}{\stackrel{?}{=}}
\newcommand{\closure}{\overline}
\newcommand{\intersect}{\cap}
\newcommand{\union}{\cup}
\newcommand{\nullset}{\emptyset}
\newcommand{\minus}{\ \backslash\ }

\begin{comment}

:Author: Micah Chambers
%:Title: Homework 7 for Real Analysis

\end{comment}

%\author{Micah Chambers}
%\title{Homework 8}

\begin{document}

\section*{Introduction to Modeling}
The short term goal of current neurological models is to
learn something about the system being modeled.
While control (for instance in prosthetics), or predication
(such as BlueBrain) may some day be feasible, in the
short run we still have a great deal to learn about how our
brain works. By learning about specific systems (the dynamics
of a single person's brain), we may be able to diagnose diseases,
find their causes and even someday rectify the cause. Of course
by learning about various specific systems we can also learn 
about the general way our brains work. The reason then, for 
studying the BOLD model then is
to be able to derive the various system parameters and understand
why one brain region (say a voxel) reacts in one way, when another
acts completely differently.

While the General Linear Model has been extremely effective at
gleaning information about which voxels are "active" and which
are not, its purpose is only to say which regions are "statistically
significant". However, considering that voxels consist of millions
of neurons, the reality is that a
region could be active, but at some other level or time constant
than expected; and be completely missed by the linear model. 
Studies showing activation
in Cadavers of Phantoms highlight the inherant problem with the 
General Linear Model: a problem that mandates multiple
comparison testing with very high thresholds to ensure any 
sort of face validity \citep{drift}.
By moving to models that actually "learn" the 
parameters of the model, model error will be drastically
reduced allowing for much more reasonable statistical thresholds.

Of course, it is well known that many of the assumptions 
that the GLM makes about BOLD activation do not hold. 
The General Linear Model ignores
the fact that the BOLD response is nonlinear, contains non-gaussian
noise and most importantly that brain networks are 
so called "Small World Networks" \citep{smallworld}, \citep{noise}. 
Merely accounting for
nonlinearities in the time-series of the BOLD response can
result in significantly better estimation as shown in \citep{nonlinearmodels}.
But even more crucial is the recent movements away from mass-
univariate models. Current mass-univariate models 
assume uniform connectivity from input to every voxel; yet
this is obviously not the case. While
violating linearity assumptions or Gaussian assumptions 
requires us to be more conservative, the lack of any method
to switch brain circuits on and off means that any model
we make for brain regions would not be Turing complete;
in effect implying that humans are incapable of 
thought \citep{turing}.

So called Dynamic Causal Modeling is the beginning of 
the next phase of neurology. DCM is the first 
brain study to show any significant connection between 
diffusion tensor imaging and actual function layout in
the human brain \citep{dti_dcm}. By incorporating connections
between regions and a realistic activation model, DCM
corrects two of the largest problems with the GLM. While
there are other techniques that may be useful in the future,
they either lack physiological analogs (Artificial Neural
Networks) or are extremely computationally expensive 
(mutual information).  Given that much of the 
potential of the GLM has been exhausted, and that DCM 
is one of the first well defined methods capable of learning 
complicated brain circuits DCM is crucial to the future
of FMRI and our understanding of the human brain.

\section*{Results}
The particle filter shows great promise in being able 
to learn a variety of different regional activation
parameters. We have performed tests with a random binary
pulse train as input to a simulated region and then tested
the particle filter's ability to estimate the signal
parameters. Additionally random gaussian noise with an
Signal-To-Noise ratio of 2.0 was added. Figure 1 shows
the pulse train, the clean signal and the noisy signal, which
was the input to the particle filter. A random set of 
system parameters was chosen to simulate the region,
which the particle filter had no knowledge of other than
a mean and rough variance.
The tests were performed using 10,000 particles and took
approximately 4 minutes to compute.

%figure 1
\begin{figure}
\includegraphics[width=8in,height=4in]{noise.png}
\caption{Simulated time-series for a single region. The red line
shows the input stimulus, the blue line shows the base BOLD 
response, and the green line is the BOLD response with Gaussian
white noise added at SNR of 2}
\label{fig:noise}
\end{figure}

%figure 2
\begin{figure}
\includegraphics[width=8in,height=4in]{bold.png}
\caption{Particle Filter Results. The blue line is the "true"
(simulated) BOLD response, the red line
is the output of the particle filter with 2 standard deviations
in green. Note that the error bar for time 0 is outside the scope
of the image and is approximately $\pm .003$.}
\label{fig:bold}
\end{figure}

%figure 3
\begin{figure}
\includegraphics[width=8in,height=4in]{state.png}
\caption{Particle Filter Results. The blue line is the "true"
parameter and the green line is the estimated value for the parameter.
Note that the timescale for $S_t$ is different to highlight an typical
stimulus response. Again 2 standard deviations are shown.}
\label{fig:state}
\end{figure}

Figure 2 shows the estimated BOLD response versus the simulated
data (without noise). It is clear that as the estimated BOLD
response converges to the true BOLD as time progresses. This
is because more and more potential states are eliminated by system's
response to new input sequences. Convergence depends highly on 
the input signal and the weighting function thus choice of both
are extremely important. We have found that an impulse train
(emulated with .1 second width square waves) gives a very good
response. It is also important to choose an input with
some longer hold times, at least as long as than the 
expected time constants, which in this case was around 
8 seconds. The weighting function we have chosen is based
on the exponential distribution, with a variance equal to
the RMS of the signal. While it is important to converge
by the end of the time-series, it is often the case that 
converging too fast will harm the state estimation by over-
emphasizing the measurements of single time points. The rate
of convergence is well within control of the user of the 
particle filter based on the weighting function and the frequency
of re-sampling. Note that resampling can cause quantization
errors due to the nature of estimating a long tailed 
distribution with a finite number of samples.

Figure 3 shows the estimated versus simulated values for
several key parameters. Notice how the variance drops as
the particle filter continues. The parameters shown are
$V_0$ which is a scaling parameter, $\tau_s$, and $\tau_f$
which are both time constants and $s_t$ which is a hidden
state variable, estimating the flow inducing signal. Note that
even though some of the parameters don't converge to the exact
value, that the estimated $s_t$ still matches the true
$s_t$ relatively well. It is often the cast that although a
few parameters don't converge to their true value, one 
parameter may pick up the slack for another. It is of
note that $\tau_s$ and $\tau_f$ do converge at least close
to their true values, because they cannot be determined
from steady state.

\bibliographystyle{abbrvnat}
\bibliography{proposal_micah}

\end{document}

### 2010-08-12 code/BoldModel.cpp
fixed all the isnan isinf stuff with the boost library version.

@@ -5,6 +5,8 @@
 #include <indii/ml/aux/GaussianPdf.hpp>
 #include <indii/ml/aux/DiracPdf.hpp>
 
+#include <boost/math/special_functions/fpclassify.hpp>
+
 #include <vector>
 #include <cmath>
 #include <ctime>
@@ -147,7 +149,7 @@ int BoldModel::transition(aux::vector& dustin, const double time,
      * inf - inf = nan, so nan or inf in any member 
      * will cause this to fail
     */
-    if(isnan(tmp) || isinf(tmp)) {
+    if(boost::math::isnan<double>(tmp) || boost::math::isinf<double>(tmp)) {
         dustin = defaultstate;
         return -1;
     }

### 2010-08-12 code/brainid.cpp

@@ -9,6 +9,8 @@
 #include <itkImageSliceIteratorWithIndex.h>
 #include <itkMetaDataObject.h>
 
+
+
 #include <indii/ml/filter/ParticleFilter.hpp>
 #include <indii/ml/filter/StratifiedParticleResampler.hpp>
 #include <indii/ml/aux/GaussianPdf.hpp>
@@ -17,6 +19,8 @@
 
 #include <indii/ml/aux/Almost2Norm.hpp>
 #include <indii/ml/aux/AlmostGaussianKernel.hpp>
+
+#include <boost/math/special_functions/fpclassify.hpp>
 
 #include <boost/archive/binary_iarchive.hpp>
 #include <boost/archive/binary_oarchive.hpp>
@@ -531,7 +535,7 @@ int main(int argc, char* argv[])
             //check for errors, could be caused by total collapse of particles,
             //for instance if all the particles go to an unreasonable value like 
             //inf/nan/neg
-            if(isnan(ess) || isinf(ess)) {
+            if(boost::math::isnan<double>(ess) || boost::math::isinf<double>(ess)) {
                 *out << std::endl << "Total Weight: "
                             << filter.getFilteredState().getDistributedTotalWeight()
                             << endl;

### 2010-08-12 code/include/BoldPF.h

@@ -28,6 +28,8 @@
 #include "boost/numeric/bindings/traits/ublas_vector.hpp"
 #include "boost/numeric/bindings/traits/ublas_symmetric.hpp"
 #include "boost/numeric/bindings/lapack/lapack.hpp"
+#include "boost/numeric/bindings/lapack/lapack.hpp"
+#include <boost/math/special_functions/fpclassify.hpp>
 
 namespace aux = indii::ml::aux;
 
@@ -184,7 +186,7 @@ aux::matrix calcCov(indii::ml::aux::DiracMixturePdf& p)
     for(unsigned int i = 0 ; i < p.getSize() ;i++) {
         if(p.getWeight(i) > 0)
             sum += p.getWeight(i)*outer_prod(p.get(i)-mu, p.get(i)-mu);
-        if(isnan(sum(8,8))) {
+        if(boost::math::isnan<double>(sum(8,8))) {
             std::cerr << "Sum" << std::endl;
             outputMatrix(std::cerr, sum);
             std::cerr << std::endl << "old" << std::endl;
@@ -344,7 +346,7 @@ int BoldPF::run(void* pass = NULL)
             //check for errors, could be caused by total collapse of particles,
             //for instance if all the particles go to an unreasonable value like 
             //inf/nan/neg
-            if(isnan(ess)) {
+            if(boost::math::isnan<double>(ess)) {
                 *debug << "\n" << "Error! ESS was " << ess << "\n";
                 status = ERROR;
                 break;

### 2010-08-12 code/makeest.cpp

@@ -9,6 +9,9 @@
 #include <string>
 #include <sstream>
 
+#include <boost/math/special_functions/fpclassify.hpp>
+
+
 template <typename T>
 void writeImage(std::string base, std::string name, typename T::Pointer in)
 {
@@ -142,7 +145,7 @@ int main(int argc, char* argv[])
         it.GoToBegin();
         while(!it.IsAtEnd()) {
             while(!it.IsAtEndOfLine()) {
-                if(!isnan(it.Get())) {
+                if(!boost::math::isnan<float>(it.Get())) {
                     index3 = it.GetIndex();
                     for(uint32_t l = 0 ; l < 3 ; l++)
                         index4[l] = index3[l];

### 2010-08-12 code/parammap_eval.cpp

@@ -1,3 +1,5 @@
+#include <boost/math/special_functions/fpclassify.hpp>
+
 #include <itkOrientedImage.h>
 #include <itkImageFileWriter.h>
 #include <itkImageFileReader.h>
@@ -224,8 +226,10 @@ Image4DType::Pointer simulate(Image4DType::Pointer paramImg, Label3DType::Pointe
 
                 /*If it is time to sample, do so*/
                 while(dt_l*index4[3] <= rt) {
-                    if(isnan(readout(state, params)) || isinf(readout(state,params))) 
+                    if(boost::math::isnan<double>(readout(state, params)) || 
+                                boost::math::isinf<double>(readout(state,params))) {
                         throw(-2);
+                    }
                     out->SetPixel(index4, readout(state, params));
                     index4[3]++;
                 }

### 2010-08-12 external/dysii/src/indii/ml/aux/DistributedPartitioner.cpp

@@ -6,6 +6,9 @@
 #include <vector>
 #include <algorithm>
 
+//for isnan
+#include "boost/math/special_functions/fpclassify.hpp"
+
 using namespace indii::ml::aux;
 
 DistributedPartitioner::DistributedPartitioner(const unsigned int nth) :
@@ -85,10 +88,10 @@ bool DistributedPartitioner::init(DiracMixturePdf* p,
       }
       boost::mpi::broadcast(world, value, guesser);
 
-      if (value != value) {
+      if (boost::math::isnan<double>(value)) {
         guesser++;
       }
-    } while ( (value != value) && guesser < size);
+    } while (boost::math::isnan<double>(value) && guesser < size);
     
     /* if we've managed to find a value... */
     if (guesser < size) {

### 2010-08-12 external/dysii/src/indii/ml/aux/GaussianPdf.cpp

@@ -9,6 +9,7 @@
 #include "boost/numeric/bindings/traits/ublas_vector.hpp"
 #include "boost/numeric/bindings/traits/ublas_symmetric.hpp"
 #include "boost/numeric/bindings/lapack/lapack.hpp"
+#include "boost/math/special_functions/fpclassify.hpp"
 
 #include <assert.h>
 
@@ -167,7 +168,7 @@ double GaussianPdf::densityAt(const vector& x) {
   }
   
   double p = ZI * exp(-0.5 * exponent);
-  if (p != p) {
+  if (boost::math::isnan<double>(p)) {
     p = 0.0;
   }
   

### 2010-08-13 doc/thesis/results.tex
changes to thesis

@@
 will not be useful. Additionally, determination of reasonable parameters from the
-time-series is also desirable. The reason to then simulate an entire slice is that
-Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM)
-models noise more realistically, for instance by adding noise in the frequency domain. 
-It also provides insight into the applicability of particle filters in modeling the BOLD 
+time-series is also desirable. An entire slice is then simulated using 
+Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM) which
+models noise more realistically. POSSUM also provides insight into the 
+applicability of particle filters in modeling the BOLD 
 signal on a large scale. Both sorts of tests are usefull in determining the

### 2010-08-14 doc/thesis/results.tex
fixes to results

@@
 Given the state-space equations for the BOLD signal, simulating a single time 
-series is relatively straight forward. After generating a true signal,
+series was relatively straight forward. After generating a true signal,
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian I.I.D. steps wered added to this true signal. Finally a 
-carrier level is added, since BOLD is typically
+process with Gaussian I.I.D. steps were added to this true signal. Finally a 
+carrier level was added, since BOLD is typically
 measured as a \% difference from the baseline. The particle filter
@@
 Once this noisy simulated time series was generated, the particle filter algorithm
- was then run on single time series image. Here I will include two sets of tests 
-to determine the power of the particle filter in modeling. The first set are 
+ was run on the single time series image. Here I include four sets of tests 
+to determine the power of the particle filter in modeling. The first two were
 used to demonstrate the ability of the particle filter to find the most likely
-set of parameters/state variables over the course of the run. The second type
-of test will investigate the problem of false-positives. As the first
-round of tests will show, given that there is an underlying BOLD signal, 
+set of parameters/state variables over the course of the run. The second two 
+tests investigated the problem of false-positives. As the first
+round of tests show, given that there is an underlying BOLD signal, 
 the particle filter is excellent at finding the most probable cause of 
 the observed signal. However, as discussed in the false positive tests,
-even if there is no underlying signal, the particle filter will still determine
-the most likely BOLD signal; which may not be flat. Thus, it is 
+even if when there was no underlying signal, the particle filter still determined
+the most likely BOLD signal; which was not 0. Thus, it was
 important to investigate methods of determining false positives using 
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_lownoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
 \caption{Test Signals with low noise compared to the clean signal.}
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_lownoise}
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
 \caption{A comparison of the preprocessed signals for the low noise case. This is the
@@
 ($\sigma_x$) of $.0005$. The measurement noise as well as the steps of the drift
-were taken to be Gaussian
+were taken to be Gaussian.
 For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
 The actual signal delivered into the particle filter
-was the result of some preprocessing to remove drift, as described in 
+was the result of preprocessing to remove drift, as described in 
 \autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
@@
 to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
-The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
 The data shows that the spline struggles to match the signal near the end, although 
-overall, the preprocessing seems to have been successful at removing trends. Finally, the 
-set of fits to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
-The filtering property of the particle filter is clear from the results here.  In several locations the output
+overall, the preprocessing seems to have been successful at removing trends. While finding
+a better preprocessing algorithm could be useful, non will ever be perfect. In this case
+the cause of the problem was the final stimuli that pushed the final median well above the base
+level. The fitting lines to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
+ In several locations the output
 of the particle filter looks like a filtered version of the input. For instance toward the
 end, the estimates stay flat in spite of the preprocessed data drifting off. By
-this point, the algorithm has converged sufficiently to know that that such a movement in
-signal isn't possible. A similar circumstance occurs at 100 seconds in. 
-It is also worth noting the bias introduced by pre-processing 
+this point, the algorithm has converged sufficiently to prevent inexplicable movement.
+A similar circumstance occurs at 100 seconds. Note that bias introduced by pre-processing 
 results in a slightly larger peak signal. The final parameter sets are shown in 
 \autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
-the residual will be defined as the square root of the mean squared residual, 
+the term \emph{residual} will be defined as the square root of the mean squared residual, 
 \begin{equation}
@@
 where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
-sample at time $k$. Note the subtle distinction between this and the error
-which is defines as 
+sample at time $k$. Note the subtle distinction between this and what will be 
+termed the \emph{error} which is defines as 
 \begin{equation}
@@
 where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
-output. Throughout this section and the next, the terms \emph{error} and \emph{residual}
-will be referencing these values, for an entire timeseries.
+signal. Throughout this section and the next, the terms error and residual
+will be referencing these values (for a particular timeseries).
 
@@
 
-\begin{table}[t]
-\centering
-\begin{tabular}{|c | c  c  c  c  c  c  c |}
-\hline 
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-\hline 
-\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
-$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
-\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
-$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
-\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
-$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
-\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
-\hline 
-\end{tabular}
-\caption{Typical Covariance matrix of the parameters at the end of a run.}
-\label{tab:CovSim} 
-\end{table}
+%\begin{table}[t]
+%\centering
+%\begin{tabular}{|c | c  c  c  c  c  c  c |}
+%\hline 
+%  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+%\hline 
+%\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
+%$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
+%\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
+%$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
+%\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
+%$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
+%\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+%\hline 
+%\end{tabular}
+%\caption{Typical Covariance matrix of the parameters at the end of a run.}
+%\label{tab:CovSim} 
+%\end{table}
 
@@
 
-There are a few results worth noting in the final parameter estimates (which 
-are the mean of the particle filter's posterior distribution). First the time constants vary 
+There are a few important results in the final parameter estimates (which 
+are the mean of the particle filter's posterior distribution). First, the time constants vary 
 greatly across
@@
 based on the prior distribution (which notably has an initial mean below the true values) 
-or it could be caused by the disassociated benefits of a correct time
-constant with regards to a correct output. It often takes several measurement periods before 
-a difference
-in time constants becomes relevant to the weight. Its also possible that output is insensitive
-to differences in the time constants; althouhg varying the activation durations could induce
-better estimates.  The large variation of $V_0$ are also interesting, although as one final
-covariance matrix shows, there is a definite negative correlation between $\epsilon$ and $V_0$.
-In general,
-with this admittedly small amount of noise, it would appear that the relation between a set
-of parameters/stimuli and the output is not injective; a time
-series is not unique to a single set of parameters. This is good justification that 
-simultaneous blood volume or tagged flow calculations with the conventional FMRI 
-could benefit the model.
-
-The covariance matrix (\autoref{tab:CovSim}) provides a great deal of insight into 
-the way parameters evolve in the particle filter. The fact that there are so many 
-covariance elements that are as big or bigger than variance elements indicates just
-loose the parameters set is. Every single parameters, including $\alpha$ seems to
-have a non-zero covariance. Why $\alpha$ and $\epsilon$ would be positively related
-is one of the most interesting element. Given the difficulty in mapping out a 
-seven dimensional space, it is difficult to know why such correlations exist. 
- Notice the covariance of $\tau_f$ and $\tau_0$
-is $-0.019$ whereas
-the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
-some measure of ill-defined behavior between the different time-constants, which is to
-be expected given the number steps before a stimuli affects the output. 
+or it could be other parameters are compensating. Its also possible that output is insensitive
+to differences in the time constants; although varying the activation durations could induce
+better estimates.  
 The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
-demonstrates the migration of parameters to match the estimates. Given the relatively high
-variance of parameters, its likely that more measurements or perhaps more variation
-in the stimuli could further differentiate the parameters. Its also notable that
-the final mean of the parameters may not be the best point estimator, and that perhaps
-using the mode would be a better way to estimate the output. Regardless, the time series'
+demonstrates the migration of parameters to match the estimates. In spite of the
+rather wide final parameter variances, the time series'
 generated from the estimated mean seems to match the correct signal very well 
 (\autoref{fig:FitComparisonLowNoise}).
+
+%The large variation of $V_0$ are also interesting, although as the final
+%covariance matrix of the first run shows, there is a significant correlation between $\epsilon$ and $V_0$.
+%In general, with even a small amount of noise, parameters are ill defined. Thus at
+%least for this set of stimuli the parameters are not uniquely identifiable. 
+%
+%The covariance matrix (\autoref{tab:CovSim}) provides a great deal of insight into 
+%the way parameters evolve in the particle filter. The number of 
+%covariance elements that exceed the variance elements indicates the
+%significant interplay between parameters. Every single parameters, including $\alpha$, 
+%has a non-zero covariance. Why $\alpha$ and $\epsilon$ would be positively related
+%is the most interesting element. The answer of how exactly each parameter
+%compensates for other parameters is complex.  Notice the covariance of $\tau_f$ and $\tau_0$
+%is $-0.019$ whereas the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. 
+%%stopped here
+%As Clearly there is some measure of ill-defined behavior between the different time-constants, which is to
+%be expected given the number steps before a stimuli affects the output. 
+%The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
+%demonstrates the migration of parameters to match the estimates. Given the relatively high
+%variance of parameters, its likely that more measurements or perhaps more variation
+%in the stimuli could further differentiate the parameters. Its also notable that
+%the final mean of the parameters may not be the best point estimator, and that perhaps
+%using the mode would be a better way to estimate the output. Regardless, the time series'
+%generated from the estimated mean seems to match the correct signal very well 
+%(\autoref{fig:FitComparisonLowNoise}).
 
@@
 
+\subsection{Very Long FMRI Run}
+\begin{figure}[H]
+\centering
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/longfit}}
+\caption{BOLD estimate converging for a very long FMRI run.}
+\label{fig:long_preprocessed}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/longfit_500}}
+\caption{BOLD estimate converging for a very long FMRI run, first 500 seconds.}
+\label{fig:long_preprocessed}
+\end{figure}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline 
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline 
+\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
+$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
+\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
+$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
+\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
+$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
+\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+\hline 
+\end{tabular}
+\caption{Typical Covariance matrix of the parameters at the end of a run.}
+\label{tab:long_cov} 
+\end{table}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline 
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline 
+\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
+$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
+\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
+$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
+\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
+$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
+\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+\hline 
+\end{tabular}
+\caption{Typical Covariance matrix of the parameters at the end of a run.}
+\label{tab:long_corr} 
+\end{table}
+
+To test how well the particle filter would do given much more data, and 
+determine the inherent variance in the model parameters, a very long simulation
+with randomly generated impulse input was created. The preprocessed timeseries
+and the final estimate are shown in \autoref{fig:long_preprocessed}, 
+ and the final covariance matrix of the parameters is shown in \autoref{tab:long_cov}.
+Note that even though the
+BOLD response converged well (\autoref{fig:long_converge}),
+the parameters still have significant correlation (\autoref{tab:long_corr}). This points
+to a certain amount of indeterminability of the parameters.
+
 \subsection{Single Voxel Review}

### 2010-08-14 doc/thesis/results.tex
Results sort of done

@@
 \centering
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/longfit}}
+\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
 \caption{BOLD estimate converging for a very long FMRI run.}
-\label{fig:long_preprocessed}
+\label{fig:long_converge}
 \end{figure}
@@
 \centering
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/longfit_500}}
+\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
 \caption{BOLD estimate converging for a very long FMRI run, first 500 seconds.}
-\label{fig:long_preprocessed}
+\label{fig:long_converge_500}
 \end{figure}
@@
 \hline 
-\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
-$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
-\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
-$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
-\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
-$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
-\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
+$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
+\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
+$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
+$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
+\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
 \hline 
 \end{tabular}
-\caption{Typical Covariance matrix of the parameters at the end of a run.}
+\caption{Covariance matrix of the parameters at the end of a run.}
 \label{tab:long_cov} 
@@
 
-\begin{table}[t]
-\centering
 \begin{tabular}{|c | c  c  c  c  c  c  c |}
@@
 \hline 
-\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
-$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
-\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
-$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
-\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
-$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
-\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
+\rowcolor[gray]{.8} $\tau_0$  & & & & & & & \\
+$\alpha$                      & 0.889884 & & & & & & \\
+\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & & \\
+$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & & \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & & \\
+$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & & \\
+\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 & \\
 \hline 
 \end{tabular}
-\caption{Typical Covariance matrix of the parameters at the end of a run.}
+\begin{table}[t]
+\centering
+
+\caption{Correlation of parameter estimates at the end of a run.}
 \label{tab:long_corr} 
@@
 \label{sec:SingleVoxelReview}
+
+The first step in determining the validity of a model is to provide some order of quality
+to rate the results by. Unfortunately, because of the variability in the signal levels the
+raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
+a low residual does not necessarily indicate a good fit. 
+Therefore, instead I  normalized the signal based on a value proportional to
+the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
+peaks and troughs, rather than using MAX and MIN values to estimate the scale of the 
+signal, I used a robust estimator of 
+scale; the Median-Absolute-Deviation (MAD). This is an estimator of the standard
+deviation, and thus a good estimator of the scale of the input signal. 
+The normalized residual values are 
+shown in \autoref{tab:SignleVoxelActivationComparison}. A second potential method of 
+gauging performance is mutual information. Mutual information is a method of measuring
+the interdependence of two random variables. If two signals are truly independent,
+then the mutual information will be zero. Although ideally suited to Bernoulli 
+random variables where there are a finite number of options, by using histograms
+it is possible to derive a joint distribution of two signals. The algorithm for 
+mutual information is based on that joint distribution:
+\begin{equation}
+\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\right)
+\end{equation}
+Unfortunately the number of bins causing bias in the output, thus to correct for this,
+I subtracted the estimated bias:
+\begin{equation}
+\text{bias} = \frac{N_{bins}}{2Nlog(2)}
+\end{equation}
+where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
+the mutual information estimates in this work 6 bins are used for the marginal
+distribution of each signal. Additionally, throughout log base 2 will be used.
+This leads to 36 total bins in the joint, so the bias is: 
+\begin{equation}
+\text{bias} = \frac{18}{N}
+\end{equation}
+After calculating the empirical mutual information, I subtracted the bias from
+the estimated mutual information. Note that this can result in negative mutual
+information, which should not technically be possible; so any negative mutual information
+was taken as 0.
+
 \begin{table}[t]
@@
 
-The first step in determining the validity of a model is to provide some order of quality
-to rate the results by. Unfortunately, because of the variability in the signal levels the
-raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
-a low residual does not necessarily indicate a good fit. 
-Therefore, instead I  normalized the signal based on a value proportional to
-the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
-peaks and troughs, rather than using MAX and MIN values to estimate the scale of the 
-signal, I used a robust estimator of 
-scale; the Median-Absolute-Deviation (MAD). This is an estimator of the standard
-deviation, and thus a good estimator of the scale of the input signal. 
-The normalized residual values are 
-shown in \autoref{tab:SignleVoxelActivationComparison}. A second potential method of 
-gauging performance is mutual information. Mutual information is a method of measuring
-the interdependence of two random variables. If two signals are truly independent,
-then the mutual information will be zero. Although ideally suited to Bernoulli 
-random variables where there are a finite number of options, by using histograms
-it is possible to derive a joint distribution of two signals. The algorithm for 
-mutual information is based on that joint distribution:
-\begin{equation}
-\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\right)
-\end{equation}
-Unfortunately the number of bins causing bias in the output, thus to correct for this,
-I subtracted the estimated bias:
-\begin{equation}
-\text{bias} = \frac{N_{bins}}{2Nlog(2)}
-\end{equation}
-where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
-the mutual information estimates in this work 6 bins are used for the marginal
-distribution of each signal. Additionally, throughout log base 2 will be used.
-This leads to 36 total bins in the joint, so the bias is: 
-\begin{equation}
-\text{bias} = \frac{18}{N}
-\end{equation}
-After calculating the empirical mutual information, I subtracted the bias from
-the estimated mutual information. Note that this can result in negative mutual
-information, which should not technically be possible; so any negative mutual information
-was taken as 0.
-
 Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, 
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering

### 2010-08-14 doc/thesis/methods.tex
more thesis gah
reviewed methods

@@
 particle filter for use with FMRI data is non-trivial.
+
+\section{Model}
+As originally written in \autoref{sec:BOLD Physiology} the state variables
+for the BOLD model are as follows:
+\begin{eqnarray}
+\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
+\dot{f} &=& s\\
+\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\end{eqnarray}
+The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
+included noise in the update of $x$, however that is not included here.
+The reason for the difference is that cloud of particles is, to some extent,
+able to account for that noise. It is common, however, to model that noise
+in particle filters by adding a random value to each updated state variable. 
+Because the purpose of this particle filter is to learn the underlying distribution
+of the static parameters, rather than precisely model the time course of the 
+in the dynamic parameters ($\{s,f,v,q\}$) this noise is left out. It also helps
+that detrending is applied before the particle filter and that the
+BOLD model is dissipative. When no stimuli are applied, all the particles 
+decay to ($\{0,1,1,1\}$). Typical particle filters 
+also use this state noise as an exploratory measure; however this method is
+less necessary when good priors are available.
+
+Typically a step size of .001 was used, after finding that even .01 can
+at times lead to the state equations careening out of control.
+
+\section{Preprocessing}
+\label{sec:Methods Preprocessing}
+The normal pipeline for analyzing
+FMRI involves a several preprocessing steps. The first and most important
+task is motion correction. To do this, a single volume in time is chosen, and
+volumes at every other time are registered to this one volume. This corrects
+for motion by the patient as well as small changes in the magnetic
+fields that cause the image to shift. 
+In conventional statistical parametric mapping, a Gaussian smoothing
+filter is applied across the image as discussed in \autoref{sec:RFT}.
+After this, detrending is performed which is discussed in \autoref{sec:Detrend}.
+Recall that FMRI signal levels are unit-less and though detrending is not
+always necessary, the data must always be converted 
+into \% difference from baseline. 
+The generally accepted method is to use a high pass filter, although the
+cutoff frequency is application dependent and often applied haphazardly.
+Before going into the detrending used in this work, it is necessary to 
+discuss the type of noise present in FMRI.
+
+\subsection{BOLD Noise}
+\label{sec:Introduction Noise}
+As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
+extensively studied and despite minor discrepancies, the cause of the BOLD 
+signal is well known. However, as FMRI detects an  
+aggregate signal over the space of cubic centimeters, there are
+plenty of noise sources . Though local neurons act
+together (i.e. around the same time), the density of neurons, the
+density of capillaries, and slight differences in activation across 
+a particular voxel can all lead to signal attenuation and noise.
+
+A particularly difficult form of noise present in FMRI is a low frequency
+drift, often characterized as a Wiener process (\cite{Riera2004}). 
+Though not present in all regions, as many as ten to fifteen percent
+of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
+inference problems \cite{Smith2007}. It is still not
+clear what exactly causes this noise, although one possibility is 
+the temperature difference in scanner magnetic coils\cite{Smith2007}. 
+It is clear that this drift signal is not solely
+due to a physiological effects, given its presence in cadavers and phantoms 
+\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
+more prevalent at interfaces between regions. Though one potential source
+could be slight movement, co-registration is standard, making this unlikely. 
+Regardless, the problem mandates the use of a high pass filter \cite{Smith2007}.
+
+In order to characterize the noise, I analyzed resting state data.
+During resting state, the patient is shown no images, and he is asked
+to avoid movement and complex thought.  Overall though there should be 
+very little activation, and thus the signal consists entirely of noise. 
+Therefore resting state data is perfect for analyzing noise. 
+The locations were chosen from points all around the brain, 
+all in grey matter voxels. These time
+series were chosen because they were representative of different types
+of noise found in the resting state data.
+
+TODO give scanner details
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_29_49_9}}
+\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_34_43_24}}
+\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_22_38_23}}
+\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_37_29_24}}
+
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
+%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}
+
+\caption{Q-Q Plots of normalized resting state data}
+\label{fig:QQDC}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_29_49_9}}
+\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_34_43_24}}
+\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_22_38_23}}
+\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
+\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
+\label{fig:QQDelta}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_29_49_9}}
+\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_34_43_24}}
+\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_22_38_23}}
+\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
+\caption{Q-Q Plots of resting state data, after the de-trending}
+\label{fig:QQSpline}
+\end{figure}
+
+Because most methods (including the one used in this paper)
+assume the noise realizations are independent of each other, the auto-
+correlation is of particular interest (which is a necessary but not
+sufficient condition for independence). Gaussianity is also a common
+assumption made in studies of FMRI data, though that assumption is not
+needed in this work. Regardless, comparing the distribution to a Gaussian
+is informative, so Q-Q plots are used to compare example data with the
+Normal distribution. Additionally, in FMRI data the noise is often considered 
+to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
+characterized by steps that are Gaussian and independent. The simulations discussed in 
+\autoref{sec:Single Voxel Simulation} make use of this, 
+by adding a Wiener random process to the overall signal. To determine
+whether the noise is in fact Wiener, the distribution of 
+the steps were plotted against a Gaussian.
+
+Finally, removal of the drift is often performed with a high pass filter,
+so analyzed the distribution after subtracting of a spline, (see \autoref{sec:Methods Preprocessing}).
+
+\autoref{fig:QQDC} shows the 
+results with a regression line fit to the points.
+Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
+x-axis and the points
+on the y-axis come from the same type of distribution, then all the points will
+be collinear. Differences in the variance will cause the line to have a slope
+other than 1, while differences in the expected value will cause the fitted line
+to be shifted. In these Q-Q plots, the points are being compared to the standard
+Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
+normalized (changed to percent difference).
+
+Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
+are well described by a Gaussian process with a small autocorrelation, 
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
+do not seem to fit the Gaussian well. Also note the significant autocorrelation in
+\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
+Gaussian white noise.  On the other hand, the steps do conform rather
+closely to the normal distribution.
+As expected, most of the autocorrelation disappears for the step data. Given
+that the steps seem to fit the Normal distribution, the low autocorrelation
+indicates that the steps could be Independently Distributed. 
+Therefore, the noise does seem to come close to a Wiener process.
+
+De-trending the time-series by subtracting a spline fit to the distribution
+removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
+though not perfectly. Though the distributions still do not exactly fit
+the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
+In all, the de-trending is effectively removing Wiener noise.
+
+\subsection{Detrending}
+\label{sec:Detrend}
+The non-stationary
+aspect of a Weiner process, presumably the result of integrating some
+$\nu_x$ is difficult to compensate for, and so many methods
+have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
+demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
+of FMRI. It has been reported that in some studies as many as half the voxels 
+benefited from detrending (\cite{Smith2007}). In a head to head comparison, 
+\cite{Tanabe2002}, showed that in most cases subtracting off
+a spline worked the best. The benefit of the spline versus wavelets, high pass 
+filtering or other DC removal techniques is that the frequency response is not set.
+Rather, the spline is adaptive to the input. Unfortunately no method will 
+perfectly remove noise, and no method will leave the signal untouched.
+
+The method I used to calculate the spline was picking one knot for every 20
+measurements in an image. Thus a 10 minute session at a repetition time of 
+2.1 seconds would have 19 knots. The knot first and last knots were each 
+given half the number of samples as the rest of the knots; which were all 
+located at the center of their sample group. The median of each sample group
+was then taken and used as the magnitude for the group. Taking the median 
+versus the mean seemed to work better, given the presence of outliers. 
+There is potential to optimize the spline further using a canonical 
+HRF to find resting points; however, for this to work the experiment would have
+to be designed with this in mind.
+
+Problematically, after removing the DC component of the signal,
+by definition the signal will have a median near zero. 
+Unfortunately this is not the natural state of the BOLD signal. More specifically,
+when the signal is inactive, the BOLD response should be at 0\% change from
+the base level; activation may then increase, or for short periods decrease from this base.
+Because most of the BOLD signal is above baseline, after removing the spline
+the BOLD resting state will be below 0\%.  This reduces the ability of an algorithm to learn.
+One method of accounting for this is to simply add a DC gain model parameter.
+Like all the other model parameters, with enough measurements, a viable
+parameter would fall. Yet adding another dimension increases the
+complexity of the model, when the parameter is relatively easy to estimate
+by visual inspection.  In this work a simpler approach was used. To determine
+the DC gain I used a robust estimator of scale. The Median Absolute Deviation (MAD)
+proved to be accurate in determining how much to shift the signal up
+by. I tested both methods during the course of analysis, and found that the increase 
+in model complexity far outweighed the slight increase in flexibility. Other
+methods may work better, however the MAD worked well, 
+as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.
+
+\begin{equation}
+y_{\text{gain}, 0:K} = 1.4826\text{median}_{i=0:K}(y_i - \text{median}(y_{0:K}))
+\end{equation}
+
+A serious concern when adding and subtracting arbitrary values to 
+real data is whether this will create false positives. This is a legitimate
+concern; however, a boosted response does not effect how well the BOLD model 
+predicts the actual measurements.
+
+%\subsection{Linearizing Noise}
+%\label{sec:Methods Delta Based Inference}
+%The alternative to these sorts of low frequency manipulation is to
+%go around the problem in another way. Here, I propose a 
+%different method of dealing with the drift. 
+%Instead of comparing the direct output of the particle filter with the direct
+%measurement, the algorithm would compare the change in signal over a single TR,
+%with the result of integrating the model for the same period. 
+%In most signal processing cases this would be foolish, but that is because the 
+%general assumption is that all noise is high frequency. Considering 
+%the fact that every BOLD analysis pipeline uses a high pass filter,
+%whereas low poss temporal filter are rarely applied, it makes sense
+%that a derivative type method could work. The benefit of particle filters
+%is that they are a robust method of inference, and I would assert 
+%that the particle filter ought to be given as \emph{raw} data as possible. 
+%While taking direct measurements
+%without de-trending would give awful results, using the difference removes the 
+%DC component and turns what is usually assumed to be a Weiner process into 
+%a simple Gaussian random variable. 
+%
+%\begin{equation}
+%\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
+%\label{eq:measass_delta}
+%\end{equation}
+%
+%Even if $\nu_d$ is some other additive process, the difference will still be closer
+%to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
+%in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
+% All the assumptions made originally
+%for the particle filter still hold, and all of the parameters may be distinguished based on
+%the step sizes, thus it is not unreasonable to consider matching the string of step sizes
+%rather than string of direct readings. 
+%
+%\begin{figure}
+%\label{fig:FrequencyGraphs}
+%\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
+%\end{figure}
+
+\section{Particle Filter Settings}
+There are quite a few options when using a particle filter; those
+options will be discussed in this section.
 
@@
 the single most important design choice. A very wide prior will require
-more particles to be sufficiently dense, a very thin prior may miss
+more particles to be sufficiently dense, a very thin (low variance) prior may miss
 the true parameters. Consequently, for this work it was natural
-to use priors that will give results consistent with similar tests. accepted result,
+to use priors that will give results consistent with previous works, 
 \cite{Friston2000}. This constrains the usefulness of the model to
@@
 exhaustive simulated annealing or hill climbing tests for multiple
-regions and multiple patients. The purpose for this work is to determine
-the fitness of particle filters for this task. Ultimately this algorithm may become
-more useful as more studies are done on BOLD model that don't require
-priors.
+regions and multiple patients.
 
@@
 getting a dense prior is difficult. Insufficiently
-dense particles will result in inconsistent results, of course the
+dense particles will result in inconsistent results. Of course the
 processing time will scale up directly with the number of particles.
@@
 particles needed decreases as well. Thus, as a heuristic, initially
-the number of particles is set to 28,000, but after resampling,
-the number of particles is dropped to 1,000. Typically during the 
-first few measurements the variance drops precipitously since most particles
-don't describe the system well. The particles that are left are in a
+the number of particles was set to 16,000, but after resampling,
+the number of particles was dropped to 1,000. Typically during the 
+first few measurements the variance dropped precipitously because
+most particles were far from a solution.  The particles that are left are in a
 much more compact location, allowing them to be estimated with 
@@
 
-\section{Model}
-As originally written in \autoref{sec:BOLD Physiology} the state variables
-for the BOLD model are as follows:
-\begin{eqnarray}
-\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
-\dot{f} &=& s\\
-\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
-\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
-\end{eqnarray}
-The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
-included noise in the update of $x$, however that is not included here.
-The reason for the difference is that cloud of particles is, to some extent,
-able to model that noise. It is common, however, to actually model that noise
-in particle filters by adding a random value to each updated state variable. 
-Because the purpose of this particle filter is to learn the underlying distribution
-of the static parameters, rather than precisely model the time course of the 
-in the dynamic parameters ($\{s,f,v,q\}$) this noise is left out. Because the
-BOLD model is dissipative, when no stimuli are applied, all the particles 
-will decay to ($\{0,1,1,1\}$) which they should. If evolution noise were added, the 
-particles would then weight particles based on the results of that noise, rather
-than on the quality of the static parameters. Typical particle filters 
-also use this state noise as an exploratory measure; however this method is
-less well suited when the system is dissipative.
-
-Thus, at each time step, the states were linearized, and each state variable,
-$s$, $f$, $v$, $q$, were changed according their previous rate of change.
-Because of the difficulty of solving a system of nonlinear equations, I 
-did not use the typical Runge-Kutta 4/5 technique to optimize step sizes. 
-The cost of missing a feature in these differential equations typically leads
-to non-real numbers shortly down the road. Because the non-real numbers do 
-not come until the solution has had the chance to update two or three more
-times, taking long steps can result in catastrophic and un-recoverable errors. 
-Typically a step size of .001 was used, after finding that even .01 can
-at times lead to the state equations careening out of control.
-
 \section{Resampling}
@@
 The algorithm for resampling is described in \autoref{sec:Particle Filter Resampling}.
-The primary design decision for resampling is the regularizing
-kernel. As mentioned previously, the Gaussian kernel is convenient,
-because it is simple to sample from. As discussed in \autoref{sec:Particle Filter Resampling},
-as long as resampling is kept as a last resort, some amount of over-smoothing
-won't impair convergence. Therefore, for this work I chose a Gaussian kernel of
+When regularizing, the Gaussian kernel is convenient,
+because it is simple to sample from and long tailed.
+As discussed in \autoref{sec:Particle Filter Resampling},
+as long as resampling is kept as a last resort, some over-smoothing
+doesn't impair convergence. Therefore, for this work I chose a Gaussian kernel of
 bandwidth equal to the original distribution's covariance. Obviously this will
@@
 
-Because the regularized resampler will almost certainly over-smooth,
-it is necessary to wait until the $N_{eff}$ drops relatively low
-(say below 25) before resampling is performed. As an additional 
-measure against sharp drops in the $N_{eff}$ that may quickly raise
-back up, resampling is only performed when two consecutive low
-$N_{eff}$'s are found. 
-The danger in waiting longer to resample is that if there were no
-particles in the vicinity of the solution, then particle deprivation
-can occur.  When this happens, the resampling
-stage will have an inappropriately small variance which will lead to an 
-inappropriately small support for the distribution. After this, the 
-distribution will be unrecoverable. Thus, when the $N_{eff}$ gets down
-to extremely low values (below 5), usually this is an indication of particle
-deprivation; something that is also usually accompanied by 
-fast drops in variance. When this happens, it is often preferable 
-to return to a previous, wider estimate of scale for the regularization
-and continue from there. Thus to keep particle deprivation from 
-affecting the results; a previous covariance matrix is 
-used as the regularization bandwidth. In my algorithm I use
-the last covariance matrix for which the $E_{eff}$ was greater than
-the threshold. This guarantees that there was at least some amount of 
-particle diversity, and also helps prevent converging prematurely.
-
-When some sort of particle deprivation has not occurred, regularized
-resampling with a large bandwidth will slow down convergence. Thus
-as mentioned previously, to prevent over smoothing, the particle
-filter algorithm used here waits until 
-the effective sample size drops below 25 for two measurement points
-in a row. This seemed to give good results, and avoids
-temporary drops in the $N_{eff}$ caused by outliers.
+Re-sampling is a not strictly necessary, but increases the effectiveness
+of the particle filter by adjusting the support to emphasize areas
+of higher probability. Re-sampling is slow because it requires re-drawing
+all the particles. It also closes off avenues of investigation, and is
+designed to over-smooth to prevent overly thinning the support. For all these
+reasons, resampling was only performed when the $N_{eff}$ dropped below
+50 (for 1000 particles).  As a measure against sharp drops in the $N_{eff}$ 
+caused by a large spike in error, resampling was only performed when 
+two consecutive low ($<50$) $N_{eff}$'s were found.
 
@@
 \label{sec:Methods Weighting Function}
-Choosing a representation of an unknown distribution tricky, and while
-the usual choice is a Gaussian distribution, there are reasons why it
-may not be the best.  Studies of the noise in FMRI typically characterise
-noise as an additive noise process with Gaussian steps; a Wiener process.
-
+The choice of $P(y_k | x_k)$ is the second most important design decision, behind 
+the prior. While the conventional choice for an unknown distribution is the 
+Gaussian, there are reasons why it may not be the best in this case.  
 As noted in \autoref{sec:Introduction Noise}, the noise is not strictly Gaussian,
@@
 matches the measurement error, then the ideal particle filter will result.
-Particles with $x_k$'s that estimate $y_k$ far out on the weighting
-function will quickly have weights near 0. Thus, an weighting function that
+Particles with $x_k$'s that repeatedly estimate $y_k$ with large residual 
+will quickly have weights near 0. Thus, a weighting function that
 exactly matches $P(Y(t) | X(t))$ will easily, and correctly throw out incorrect
-particles. However assuming that this function is not going to be found 
-for arbitrary voxels, it is necessary to choose a function to approximate this 
-probability. The cost of choosing an overly broad distribution for this
-function is an overly broad $P(x_k | y_k)$; and ultimately an overly
-broad $P(x_k | y_{0:k})$. On the other hand, an overly thin distribution
-will lead to, at best, an overly thin representation of these same 
-distributions; but at wost will lead to particle deprivation (all particles
-being zero-weighted). If the \emph{true} distribution were found, particle
-deprivation would indicate that that time series is not being affected
-in any way describable by the prior distribution/model. Particle deprivation
-due to an under-variant weighting function gives no information. Thus,
-the cost of a under estimating variance is higher; a fact that definitely needs
-to be taken into account.
+particles.  The cost of choosing an overly broad distribution for this
+function is slow convergence.  On the other hand, an overly thin 
+distribution will lead to particle deprivation (all particles
+being zero-weighted).  I tested several weighting
+functions: in addition to the Gaussian I also tested the Laplace and Cauchy
+distributions, both of which have much wider tales than the Gaussian. 
+Wider tailed distributions don't down-weight
+particles as fast; and converge more slowly (and perhaps more accurately). 
+The Laplace distribution also has the
+benefit of a non-zero slope at the origin; which means that even
+it will distinguish between particles even near the origin.
 
-Because of the need to err on the side of caution I tested several weighting
-functions. In addition to the Gaussian I also tested the Laplace and Cauchy
-distributions, both of which have much wider tales than the Gaussian. The
-benefit of the wider tailed distributions is that they don't down-weight
-particles quite as fast. Additionally, the Laplace distribution has the
-benefit of having a non-zero slope at the origin. This means that even
-if the distribution is made overly broad, it will still distinguish between
-particles that are near the origin.
-
-After trial and error, I chose the weight standard deviation to be $.005$. Obviously
-the case of the Cauchy, this is simply the scale parameter. While I 
-did attempt to automatically set the standard deviation, I found that the algorithm
-became too unpredictable. Essentially, if the weight function isn't
-fixed across voxels, very noisy time series with no actual signal will 
-converge to nonsensical results. 
+After trial and error, for this work I chose a zero-mean Gaussian with standard deviation 
+of $.005$. While I made some attempts to automatically set the standard 
+deviation, results were often unpredictable. If the weight function and scale aren't
+fixed across voxels, very noisy time series with no actual signal 
+converged to nonsensical results. 
 In the future, it may be possible to set the standard deviation by
@@
 BOLD model, in this work I set the standard deviation manually at $.005$,
-because it gives a more consistency and control.
+because it gave the best consistency.
 
-\section{BOLD Noise}
-\label{sec:Introduction Noise}
-As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
-extensively studied and despite minor discrepancies, the cause of the BOLD 
-signal is well known. However, as FMRI detects an  
-aggregate signal over the space of cubic centimeters, there are
-plenty of sources of noise. Though local neurons act
-together (i.e. around the same time), the density of neurons, the
-density of capillaries, and slight differences in activation across 
-a particular voxel can all lead to signal attenuation and noise.
-
-A particularly difficult form of noise present in FMRI is a low frequency
-drift, often characterized as a Wiener process (\cite{Riera2004}). 
-Though not present in all regions, as much as ten to fifteen percent
-of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
-inference problems \cite{Smith2007}. It is still not
-clear what exactly causes this noise noise comes from, although it is possible it is 
-the result of magnets heating up, or some distortion in magnetic
-fields \cite{Smith2007}. It is clear that this drift signal is not solely
-due to a physiological effects, given its presence in cadavers and phantoms 
-\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
-more prevalent at interfaces between regions, though 
-by no means limited to such areas. Though one potential source
-could be slight movement, given that co-registration of volumes to a single time
-point is standard, this is unlikely. Regardless, the problem mandates
-the use of a high pass filter to make inference reasonably powered
-\cite{Smith2007}.
-
-In order to characterize the noise, I analyzed resting state data.
-During resting state, the patient is shown no images, and he is asked
-to avoid movement and complex thought, to the best of his abilities.
-Some limitations exist, for instance EPI sequences are noisy
-which could cause some stimulus in the auditory cortex, and of course
-long sequences could give the patient time for his mind to wander.
-Overall though there should be no activation, and thus the signal will
-consist entirely of noise. Therefore resting state data is perfect
-for analyzing the noise distribution. The locations were chosen from
-points all around the brain, all in grey matter voxels. The time
-series were also chosen because they were representative of different types
-of noise I found in the resting state data.
-
-Because most methods (including the one used in this paper)
-assume the noise realizations are independent of each other, the auto-
-correlation is of particular interest (which is a necessary but not
-sufficient condition for independence). Gaussianity is also a common
-assumption made in studies of FMRI data, though that assumption is not
-made in this work. Regardless, comparing the distribution to a Gaussian
-is informative, so I used Q-Q plots to compare example data with the
-Normal distribution. Additionally, in FMRI data the noise is often considered 
-to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
-characterized by steps that are Gaussian; in other words the difference between
-any two samples is Normally distributed. The simulations discussed in 
-\autoref{sec:Single Voxel Simulation} make use of this, 
-by adding a Wiener random process to the overall signal. To determine
-whether the noise is in fact Wiener, the distribution of 
-the difference between adjacent measurements was plotted against
-a Gaussian. Wiener processes further assume that the steps are independent;
-so the steps should have near flat autocorrelation functions.
-
-Finally, removal of the drift is often
-performed with some variation of a high pass filter, 
-so I checked the noise distribution after applying such a filter (in
-this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
-Here I wanted to know how effective my high pass filter at the removal of
-Wiener noise.
-
-\autoref{fig:QQDC} shows the 
-results with a regression line fit to the points on the Q-Q plot.
-Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
-x-axis and the points
-on the y-axis come from the same \emph{type} of distribution then all the points will
-fall on a line. Differences in the variance will cause the line to have a slope
-other than 1, while differences in the expected value will cause the fitted line
-to be shifted. In these Q-Q plots, the points are being compared to the standard
-Gaussian distribution, so the quality of the line fit determines how closely the points
-fit the Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
-normalized (changed to percent difference).
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_29_49_9}}
-\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_34_43_24}}
-\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_22_38_23}}
-\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_37_29_24}}
-
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}
-
-\caption{Q-Q Plots of normalized resting state data}
-\label{fig:QQDC}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_29_49_9}}
-\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_34_43_24}}
-\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_22_38_23}}
-\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
-\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
-\label{fig:QQDelta}
-\end{figure}
-
-A Wiener process should still conform to the normal distribution, with a 
-variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
-are well described by a Gaussian process with a small autocorrelation, 
-\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
-do not seem to fit the Gaussian well. Also note the significant autocorrelation in
-\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
-Gaussian white noise.  On the other hand, the steps do conform rather
-closely to the normal distribution.
-As expected, most of the autocorrelation disappears for the step data. Given
-that the steps seem to fit the Normal distribution, the low autocorrelation
-indicates that the steps could be Independently Distributed. 
-Therefore, the noise does come close to a Wiener process.
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_29_49_9}}
-\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_34_43_24}}
-\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_22_38_23}}
-\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
-\caption{Q-Q Plots of resting state data, after the de-trending}
-\label{fig:QQSpline}
-\end{figure}
-
-De-trending the time-series by subtracting a spline fit to the distribution
-removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
-though not perfectly. Though the distributions still do not exactly fit
-the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
-In all, the de-trending is effectively removing Wiener noise.
-
-\subsection{Detrending}
-\label{sec:Detrend}
-The non-stationary
-aspect of a Weiner process, presumably the result of integrating some
-$\nu_x$ is difficult to compensate for, and so methods
-have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
-demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
-of FMRI. In some studies, as many as half the voxels benefit from detrending, presenting
-a serious barrier to inference (\cite{Smith2007}). All the existing methods are performed
-during the preprocessing stage, rather than as an integral part of analyzing the BOLD
-signal. There is no shortage of detrending methods; however
-in a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
-a spline works the best. The benefit of the spline versus wavelets, high pass 
-filtering or other DC removal techniques is that the frequency response is not set.
-Rather, the spline is adaptive to the input, having a low cut off if the signal's 
-median stays constant but a high cut off frequency if the signal's median tends to
-shift heavily over the course of time. In spite of this, the spline will still remove some
-amount of signal.
-
-The method I used to calculate the spline was picking one knot for every 20
-measurements in an image. Thus a 10 minute session at a repetition time of 
-2.1 seconds would have 19 knots. The knot first and last knots were each 
-given half the number of samples as the rest of the knots; which were all 
-located at the center of their sample group. The median of each sample group
-was then taken and used as the magnitude for the group. Taking the median 
-versus the mean seemed to work better, given the presence of outliers. 
-There is potential to optimize the spline further using a canonical 
-HRF to find resting points; however, for this to work the experiment would have
-to be designed with this in mind.
-
-When using the median-based spline techniques, the normalized signal will,
-by definition have a median near zero. 
-The problem with this is this is not the natural state of the BOLD signal. More specifically,
-when the signal is inactive, the BOLD response should be at 0\% change from
-the base level; activation may then increase, or for short periods decrease from this base.
-After removing the spline, the BOLD resting state will be below 0\%.
-This is problematic because it reduces the ability of an algorithm to learn.
-One quick and dirty solution is to add an arbitrary constant to each BOLD response. 
-Of course this won't scale to whole-brain analysis, so a more effective technique 
-is adding a DC gain parameter to the BOLD model. 
-Like all the other model parameters, given enough measurements, correct
-value may be found. On the downside, adding another dimension increases the
-complexity of the model, for a variable that is obvious by direct
-observation.
-
-Thus, I used a more conventional approach to deal with this. To determine
-the DC gain to be added to the signal I used a robust estimator 
-of scale. The Median Absolute Deviation (MAD)
-proved to be accurate in determining how much to shift the signal up
-by. I tested both methods during the course of analysis, and found that the increase 
-in model complexity far outweighed the slight increase in flexibility. Its 
-possible that a more accurate method may exist; however, for this case the 
-MAD works well, as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.
-
-\begin{equation}
-y_{\text{gain}, 0:K} = 2\text{median}_{i=0:K}(y_i - \text{median}(y_{0:K}))
-\end{equation}
-
-A serious concern with adding and subtracting arbitrary values to 
-real data is whether this will create false positives. This is a legitimate
-concern; however, all the additions and subtractions done in this section
-have been low frequency changes and should not substantially change
-the BOLD response, being a higher frequency signal. The typical
-method of preprocessing is a high-pass filter with a cutoff of around 30
-seconds; which should about match the spline with knots every fifteen 
-measurements. \cite{Tanabe2002} found that splines tended to far outperform the
-high-pass filter method.
-
-%\subsection{Linearizing Noise}
-%\label{sec:Methods Delta Based Inference}
-%The alternative to these sorts of low frequency manipulation is to
-%go around the problem in another way. Here, I propose a 
-%different method of dealing with the drift. 
-%Instead of comparing the direct output of the particle filter with the direct
-%measurement, the algorithm would compare the change in signal over a single TR,
-%with the result of integrating the model for the same period. 
-%In most signal processing cases this would be foolish, but that is because the 
-%general assumption is that all noise is high frequency. Considering 
-%the fact that every BOLD analysis pipeline uses a high pass filter,
-%whereas low poss temporal filter are rarely applied, it makes sense
-%that a derivative type method could work. The benefit of particle filters
-%is that they are a robust method of inference, and I would assert 
-%that the particle filter ought to be given as \emph{raw} data as possible. 
-%While taking direct measurements
-%without de-trending would give awful results, using the difference removes the 
-%DC component and turns what is usually assumed to be a Weiner process into 
-%a simple Gaussian random variable. 
-%
-%\begin{equation}
-%\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
-%\label{eq:measass_delta}
-%\end{equation}
-%
-%Even if $\nu_d$ is some other additive process, the difference will still be closer
-%to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
-%in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
-% All the assumptions made originally
-%for the particle filter still hold, and all of the parameters may be distinguished based on
-%the step sizes, thus it is not unreasonable to consider matching the string of step sizes
-%rather than string of direct readings. 
-%
-%\begin{figure}
-%\label{fig:FrequencyGraphs}
-%\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
-%\end{figure}
-
-\section{Preprocessing}
-\label{sec:Methods Preprocessing}
-As discussed in the section on de-trending, the normal pipeline for analyzing
-FMRI involves a several preprocessing steps. The first and most important
-task is motion correction. To do this, a single volume in time is chosen, and
-volumes at every other time are registered to this one volume. This corrects
-for motion by the patient as well as small changes in the magnetic
-fields that cause the image to shift. 
-In conventional statistical parametric mapping, a gaussian smoothing
-filter is applied across the image as discussed in \autoref{sec:RFT}.
-After this, detrending is performed as discussed by \autoref{sec:Detrend}.
-Recall that FMRI signal levels are unit-less and though detrending is not
-always necessary, at the very least the data must be converted 
-into \% difference from the baseline. Changing to 
-\% difference removes no real information
-from signal. This is the signal that was input into the delta based 
-particle filter. Of course, most of the time analysis is performed on the
-direct signal; which mandates the removal of low frequency drift.
-The generally accepted method is to use a high pass filter, although the
-cutoff frequency is application dependent and often applied haphazardly.
-
-\section{Particle Filter Settings}
+\subsection{Runtime}
 The run-time for a single voxel depends on the several factors. First, the
@@
 about 6 minutes. On the other hand, in real circumstances the
-length is only around 150 measurements and takes around 1 minute. The size of 
-local linearization steps can certainly also make a large difference; and 
-although it can be tempting to decrease the resolution, going above .001 seconds
+length is only around 150 measurements and takes around 40 seconds (for 1000 
+particles, 1500 integration points and a Quad Core CPU). The size of 
+local linearization steps are also crucial although going above .001 seconds
 per step is not recommended. In most cases millisecond resolution
-is fine; however, when generating simulated data I found that every once
-in a while this was not enough. This is problematic in the actual particle
+is fine; however, when generating simulated data I found that at times it was
+still not enough every once. This is problematic in the actual particle
 filter since, given the large number of simultaneous integrations taking 
 place, its probable that a few particles will fail and be unfairly thrown away.
-There are two possible outcomes of this. The typical case 
-is at some point where all the particles are in a similar location and 
-so all the particles fail together.
-The result is then particle deprivation - no particles with non-zero weights remain.
-At best this will cause a convergence to a less than optimal posterior 
-probability, at worse at will result in a false negative. 
-The other possible outcome is that low time constant (or fast moving)
-particles get pruned, leaving an excessively smoothed estimate of the output.
-
-Although runge-kutta solvers could be applied, this will still not
-guarantee that features aren't missed. Its possible that a
-kind of stop-gap measure could be put into place; wherein particles that are
-about to be set to NaN are integrated again with finer grained steps. However
-oftentimes the non-real results don't occur until several time steps after the 
-numbers become unrealistic. So for instance, the time step was too long, allowing 
-$f$ to go negative, resulting in extremely large values of $q$. There are many
-different ways where this sort of event can occur, and unfortunately sometimes
-there is no way to get back to before the state starting going out of control.
+To prevent such events, 1500 integration points per second were used throughout
+the tests.
 
@@
 with this large number of particles, after 20 seconds have passed the
-algorithm forces resampling. Of course the choice of 20 seconds is somewhat
-arbitrary, but at the very least this gives some time for the particle to
-be re-weighted before the new distribution is drawn. Still the
-distribution will then be a thinned out version of the prior.
-
-%For simulated and real images (tests with multiple time-series), tests were 
-%also run with and without Gaussian filtering with sigma of %not sure todo
-%were run, since it is standard
-%practice to apply a Gaussian spatial filter to the images at each timestep. Obviously
-%a spatial filter such as Gaussian filtering increased SNR but can also lead to less
-%precision in the output maps.
-
-\section{FMRI Configuration}
-For the FMRI data discussed in \autoref{sec:RealData}, tests were 
-performed on a right handed volunteer using a Gradient EPI sequence
-on a 1.5T GE scanner with a 2.1s TR. Slices were 5mm thick and echo
-time was (todo).
+algorithm forces resampling. The choice of 20 seconds is 
+arbitrary, but at the very least it gives a more optimized version of the
+prior.

### 2010-08-14 doc/thesis/background.tex
more thesis gah
more on the results. Updated the HRF and applied pierces recommendations

@@
-\chapter{Methods of Analyzing FMRI}
+\chapter{Alternative Techniques}
 \label{sec:Prior Works}
@@
 \autoref{sec:Statistical Parametric Mapping}. The goal of this 
-work is to move away from the question "Is this region active"
-and instead ask "What does the activation look like". Answering
-this question necessitates more complex models and require
+work is to move away from simple localization, and move toward
+characterizing the response curve. Doing so 
+necessitates more complex models and requires
 more computation. Already there have been several other
@@
 \includegraphics[scale=.7]{images/HRF}
-\caption{Canonical Hemodynamic Response Function}
+\caption{Canonical Hemodynamic Response Function (y-axis unts are arbitrary because 
+normalization is performed).}
 \label{fig:HRF}
@@
 It is well known that different Hemodynamic Response Functions are necessary 
-for different regions of the brain. The Canonical HRF that is most used, 
-has been optimized for the visual cortex. As \autoref{sec:Post Stimulus Undershoot}
-discussed, there are certainly variations in the shape of the
-BOLD response, over both brain regions and patients. Handwerker et. al. (\cite{Handwerker2004})
+for different regions of the brain \cite{Handwerker2004}. The Canonical HRF that is most often used, 
+has been optimized for the visual cortex. As discussed in \autoref{sec:Post Stimulus Undershoot},
+there are certainly variations in the shape of the
+BOLD response, over both brain regions and patients. Handwerker et. al. \cite{Handwerker2004}
 discuessed the implications of choosing an incorrect HRF, the most important
@@
 Unlike Statistical Parametric Mapping, the techniques described in this
-section are all attempts to learn some version of the BOLD model. 
+section are all attempts to regress some version of the BOLD model. 
 Although \cite{Buxton1998} and \cite{Friston2000}
@@
 form solution to the ballon model, it impossible to 
-calculate the Jacobian matrix $\frac{\partial Y}{\theta}$. Thus 
+calculate the Jacobian matrix $\frac{\partial Y}{\partial \theta}$. Thus 
 \cite{Friston2002b} approximates the partial using a Volterra Kernel. 
-At each step of the Expectation-
-Maximization (EM) algorithm, the differential equation is integrated, calculating
-the residuals. Then, for each $\theta_i$ surrounding the current estimate of 
+At each step of the Expectation-Maximization (EM) algorithm, 
+the differential equation is integrated and the residuals
+calculated. Then, for each $\theta_i$ surrounding the current estimate of 
 $\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
@@
 Importantly, the Volterra-Expansion of $y$ is not able to model interactions
-between state variables (\cite{Friston2002b}); which certainly increases error.
+between state variables \cite{Friston2002b}; which certainly increases error.
 Additionally, all the tests performed
@@
 the update equation. This set of points are then used to estimate the 
-mean and covariance of $X(t)$. The benefit of this, is that it requires
+mean and covariance of $X(t)$. The benefit of this is that it requires
 no Jacobian and only a few extra calculations to get a decent estimate of
 a posterior Gaussian. Hu et. al. used the UKF to  perform a similar type of analysis to
-the one performed in this work (\cite{Hu2009}).
+the one performed in this work \cite{Hu2009}.
 

### 2010-08-14 doc/thesis/intro.tex
more on the results. Updated the HRF and applied pierces recommendations

@@
 Traditional methods of analyzing timeseries images produced by 
-Function Magnetic Resonance 
+Functional Magnetic Resonance 
 Imaging (FMRI) perform regression using the linear combination of explanatory variables. 
@@
 for small neural regions, a feature which could be useful towards real time FMRI 
-(\cite{DeCharms2005}). Future works will also benefit from the ability to 
+\cite{DeCharms2005}. Future works will also benefit from the ability to 
 apply conditions to the posterior distribution in post-processing without
-having to recalculating the parameters; for instance to impose of additional 
+recalculating parameters; for instance to impose additional 
 physiological
@@
 computationally taxing. The solution used here takes approximately 40 seconds
-for a single voxel's time series (5 minutes in length, with a quad core machine).
+for a single voxel's time series (5 minutes in length, with Core 2 Duo Q6600).
 
@@
 For the past twenty years, Functional Magnetic Resonance Imaging (FMRI) 
-has been at the forefront of cognitive research. Despite it's
+has been at the forefront of cognitive research. Despite its
 limited temporal resolution, FMRI is the standard tool for localizing 
@@
 have less bias error than the empirically driven \emph{Canonical Hemodynamic Model}
-\cite{Deneux2006}, \cite{Handwerker2004}.
+\cite{Deneux2006,Handwerker2004}.
 On the other hand BOLD signal models have numbers
@@
 The BOLD signal is caused by minute changes in the ratio of Deoxygenated
-Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain.
+Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain \cite{Ogawa}.
 Because Deoxygenated Hemoglobin (DHb) is paramagnetic, higher concentrations
@@
 noise seen in FMRI imaging (see \autoref{sec:Introduction Noise}. 
-Because MRI lacks is unitless and certain
+Because MRI is unitless and certain
 areas will have a higher base MR signal, all FMRI studies deal with
@@
 EPI imaging
-\cite{Buxton1998}, \cite{WEISSKOFF1994}, \cite{Ogawa}, however the connection
+\cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
 between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
@@
 Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
-Hemoglobin, increased blood volume effects this ratio if 
+Hemoglobin, increased blood volume affects this ratio if 
 metabolism doesn't exactly match the increased inflow of oxygenated blood.
 This was the impetus
-for the ground breaking balloon model (\cite{Buxton1998}) and windkessel
-model (\cite{Mandeville1999}). These works derive, from first principals,
+for the ground breaking balloon model \cite{Buxton1998} and windkessel
+model \cite{Mandeville1999}. These works derive, from first principals,
 the changes in deoxyhemoglobin ratio and volume of capillaries given flow waveform.
@@
 This completed balloon model was well summarized again
-in \cite{Riera2003}.  \cite{Obata2004} refined the readout equation 
+in \cite{Riera2003}. Ogawa et. al. \cite{Obata2004} refined the readout equation 
 of the BOLD signal based on the
@@
 Although the most widely used, the BOLD model described in \autoref{eq:bold}
-and \autoref{eq:boldout} have been extended in various fashions. The most
+and \autoref{eq:boldout} has been extended in various fashions. The most
 significant feature missing from the original model is the 
@@
 BOLD response for a period of 10 to 60 seconds after stimulus has
-ceased (\cite{Chen2009}, \cite{Mandeville1999a}).
+ceased \cite{Chen2009,Mandeville1999a}.
 
@@
 studies that show CBV and CBF returning to the baseline before the BOLD signal
-(\cite{Frahm2008}, \cite{Donahue2009}, \cite{Buxton2004}, \cite{Lu2004},
-\cite{Shen2008}). Unfortunately, because of limitations on FMRI and in vivo
+\cite{Frahm2008, Donahue2009, Buxton2004, Lu2004, Shen2008}. 
+Unfortunately, because of limitations on FMRI and in vivo
 CBV/CBF measurement techniques it is difficult to isolate whether CBF and
 CBV truly have returned to their baseline. Other studies indicate
-that there can be a prolonged supernormal CBV (\cite{Mandeville1999a}, 
-\cite{Behzadi2005}, \cite{Chen2009a}), although none of these papers completely
+that there can be a prolonged supernormal CBV \cite{Mandeville1999a,
+Behzadi2005, Chen2009a}, although none of these papers completely
 rule out the possibility of increased CMRO2. The discrepancies may in part
 be explained by a spatial dependence in the post-stimulus undershoot; described
-by \cite{Yacoub2006}. \cite{Chen2009}
-makes a compelling case that most of the post stimulus undershoot can be 
+by \cite{Yacoub2006}. In \cite{Chen2009}
+a compelling case is made that most of the post stimulus undershoot can be 
 explained by combination of a prolonged CBV increase, and a prolonged CBF 
@@
 in quality from the additional parameters. 
-Importantly,\cite{Deneux2006} did show that by 
+Importantly, \cite{Deneux2006} did show that by 
 simply adding viscoelastic terms from \cite{Buxton2004}, a slowed return 
@@
 In summary, there have been extensive refinements to the Balloon
-model,however, the increased complexity and lack of known priors 
-make these models less desirable. Additional degrees of freedom 
+model; however, the increased complexity and lack of known priors 
+make these models difficult to work with. Additional degrees of freedom 
 could also make parameter estimation intractable.
@@
 analyzing the eigenvalues of the state equation Jacobian, 
-(\cite{Deneux2006}, \cite{Hu2009}). The steady state of the Balloon
+\cite{Deneux2006, Hu2009}. The steady state of the Balloon
 model equations gives:
@@
 
+where the parameters are all the same as in \autoref{eq:bold}
+
 In real FMRI data, there is a significant nonlinearity in response; with short sequences
-responding disproportionately strong (\cite{Birn2001}, \cite{Wager2005}, \cite{Deneux2006}).
+responding disproportionately strongly \cite{Birn2001, Wager2005, Deneux2006}.
 This nonlinearity is accounted for in the Balloon model, although \cite{Deneux2006}
@@
 In both \cite{Birn2001} and \cite{Deneux2006} it was found that 
-Stimuli lasting longer than 4 seconds 
+stimuli lasting longer than 4 seconds 
 tend to be more linear, which is why block designs are so well accounted for

### 2010-08-14 doc/thesis/methods.tex
more on the results. Updated the HRF and applied pierces recommendations

@@
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response}
-\caption{Response to $.1s$ impulses with the mean parameters from \cite{Friston2000}}
+\caption{Response to $0.1s$ impulses with the mean parameters from \cite{Friston2000}}
 \end{figure}
@@
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response_johnston}
-\caption{Response to $.1s$ impulses with the mean parameters from \cite{Johnston2008}}
+\caption{Response to $0.1s$ impulses with the mean parameters from \cite{Johnston2008}}
 \end{figure}

### 2010-08-14 doc/thesis/particlefilters.tex
more on the results. Updated the HRF and applied pierces recommendations

@@
 \section{Introduction}
-Particle filters, a type of Sequential Monte Carlo (SMC) methods,
+Particle filters, a type of Sequential Monte Carlo (SMC) method,
 are a powerful method for estimating the posterior probability distribution
@@
 than parameter sets that give poor estimates. Although the reliance on
-a prior distribution are often troublesome, when the system being modeled
+a prior distribution is often troublesome, when the system being modeled
 has physical meaning, establishing reasonable ranges for parameters may be 
-quite easy. Optimizing the prior can be more difficult though,
+quite easy. Optimizing the prior distribution can be more difficult,
 unless the system has been extensively studied.

### 2010-08-14 doc/thesis/results.tex
more on the results. Updated the HRF and applied pierces recommendations

@@
 \label{sec:SimulationResults}
-Two levels of simulation are discussed in this section. First the simulation of single time series
-and second, a slice of simulated FMRI time series. The single time series tests
-are necessary to investigate the power of particle filters in 
-identifying the BOLD model in a noisy environment. If the particle filter is unable
-to get back to the \emph{true} time-series in spite of noise then the particle filter
-will not be useful. Additionally, determination of reasonable parameters from the
-time-series is also desirable. An entire slice is then simulated using 
+Two levels of simulation are discussed in this section; first of single voxels,
+and second of a single slice (64x64 voxels).  The single time series tests
+investigate the ability of particle filters to estimate the BOLD model 
+in a noisy environment. If the particle filter were unable
+to identify the \emph{true} time-series in spite of noise, then it 
+would not be a good choice for identifying parameters. Single voxel tests were
+performed eleven different times, each with a new noise realization.
+If runs failed to have consistent estimates of the underlying BOLD signal, then
+that would indicate a significant variance error in the model.
+
+The slice simulation was performed using 
 Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM) which
-models noise more realistically. POSSUM also provides insight into the 
-applicability of particle filters in modeling the BOLD 
-signal on a large scale. Both sorts of tests are usefull in determining the
-consistency of the particle filter. For the single voxel tests, each test will
-actually be performed eleven different times, each with a new noise realization.
-If runs fail to have consistent estimates of the underlying BOLD signal, then
-this would indicate a large variance error in the model. Similarly, in the
-Possum simulation there are actually only four discrete parameter sets driving the time-series.
-Nevertheless each voxel will have a different tissue composition and a different
-noise realization. Therefore the Possum simulation should also provide a good
-test of the consistency of the BOLD model.
+models noise realistically. POSSUM demonstrates the 
+particle filter modeling the BOLD signal on a large scale. 
+In the POSSUM simulation there were four
+discrete parameter sets driving the timeseries', however each voxel had a 
+different tissue composition and a different
+noise realization. Therefore the POSSUM simulation was 
+a good test of the applicability of particle filters for performing
+whole brain analysis.
 
@@
 Given the state-space equations for the BOLD signal, simulating a single time 
-series was relatively straight forward. After generating a true signal,
+series was straight forward. After generating a true signal,
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
@@
 measured as a \% difference from the baseline. The particle filter
-algorithm will immediately remove this by calculating the \% difference, 
+algorithm immediately removes this by calculating the \% difference, 
 but adding a carrier level meant that the exact same algorithm used 
@@
 Once this noisy simulated time series was generated, the particle filter algorithm
- was run on the single time series image. Here I include four sets of tests 
-to determine the power of the particle filter in modeling. The first two were
+ was run on the single time series image. Five sets of tests were used 
+to determine the power of the particle filter in modeling. The first two (\autoref{sec:SimLowNoise}
+and \autoref{sec:SimHighNoise}) were
 used to demonstrate the ability of the particle filter to find the most likely
 set of parameters/state variables over the course of the run. The second two 
-tests investigated the problem of false-positives. As the first
-round of tests show, given that there is an underlying BOLD signal, 
+tests (\autoref{sec:PureNoiseLowMag} and \autoref{sec:PureNoiseHighMag}) 
+investigated the problem of false-positives. As the first
+round of tests will show, given that there is an underlying BOLD signal, 
 the particle filter is excellent at finding the most probable cause of 
-the observed signal. However, as discussed in the false positive tests,
-even if when there was no underlying signal, the particle filter still determined
-the most likely BOLD signal; which was not 0. Thus, it was
-important to investigate methods of determining false positives using 
-other methods.
+the observed signal. As discussed in \autoref{sec:PureNoiseLowMag} and 
+\autoref{sec:PureNoiseHighMag}, even when there was no underlying signal, 
+the particle filter still converged. Thus, it was 
+important to investigate methods of identifying false positives.
+
+The final single-voxel test was used to determine the effectiveness of the
+particle filter for a longer FMRI session. The final covariance and correlation
+matrices are also discussed.
 
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
+\centering
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_lownoise}
 \caption{Test Signals with low noise compared to the clean signal.}
@@
 \end{figure}
+
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
-\caption{A comparison of the preprocessed signals for the low noise case. This is the
+\centering
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_lownoise}
+\caption{A comparison of the preprocessed signals for the low noise case. The
 noisy input to the actual particle filter algorithm.}
@@
 \end{figure}
+
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
+\centering
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_lownoise}
 \caption{A comparison of the fitted signals for the low noise case.}
@@
 To begin the single voxel simulation; I generated a signal using the following parameters:
-$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
-These same parameters were also used in the high noise case. The noise realizations were
-then generated based on a measurement noise ($\sigma_y$) of $.001$ and a drift standard deviation
-($\sigma_x$) of $.0005$. The measurement noise as well as the steps of the drift
+$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
+These same parameters were used for all the tests in this chapter. The noise realizations were
+then generated based on a measurement noise ($\sigma_y$) of $0.001$ and a drift standard deviation
+($\sigma_x$) of $0.0005$. The measurement noise as well as the steps of the drift
 were taken to be Gaussian.
@@
 \autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
-may have some effect on the resulting fit; thus the preprocessed signal is compared
+could have some effect on the resulting fit; thus the preprocessed signal is compared
 to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
 The data shows that the spline struggles to match the signal near the end, although 
-overall, the preprocessing seems to have been successful at removing trends. While finding
-a better preprocessing algorithm could be useful, non will ever be perfect. In this case
+overall, the preprocessing seems to have been successful at removing trends.  In this case
 the cause of the problem was the final stimuli that pushed the final median well above the base
-level. The fitting lines to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
- In several locations the output
-of the particle filter looks like a filtered version of the input. For instance toward the
-end, the estimates stay flat in spite of the preprocessed data drifting off. By
+level. The estimated BOLD signal is compared to the true BOLD signal in \autoref{fig:FitComparisonLowNoise}.
+At several time points the particle filter successfully filtered the input. 
+For instance, in the last 30 seconds
+the estimates stay flat in spite of the preprocessed data drifting off. By
 this point, the algorithm has converged sufficiently to prevent inexplicable movement.
-A similar circumstance occurs at 100 seconds. Note that bias introduced by pre-processing 
+A similar circumstance occurs at 100 seconds in. Note that bias introduced by pre-processing 
 results in a slightly larger peak signal. The final parameter sets are shown in 
@@
 \end{equation}
-where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the output
-sample at time $k$. Note the subtle distinction between this and what will be 
+where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
+sampled at time $k$. Note the subtle distinction between this and what will be 
 termed the \emph{error} which is defines as 
@@
 signal. Throughout this section and the next, the terms error and residual
-will be referencing these values (for a particular timeseries).
+will be referencing these values.
 
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
 \hline 
 \rowcolor[gray]{.8}
-1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
 \hline 
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
-last is mean over the 11 runs. The $\sqrt{MSR}$ is residual and $\sqrt{MSE}$ is the error.}
+\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, and last is the average.}
 \label{tab:LowNoiseResults} 
@@
 
-\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
+\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $v$ of the first run, low noise simulation.]
 {\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
@@
 \begin{figure}
-\subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
+\subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
 {\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
@@
 based on the prior distribution (which notably has an initial mean below the true values) 
-or it could be other parameters are compensating. Its also possible that output is insensitive
-to differences in the time constants; although varying the activation durations could induce
-better estimates.  
+or it could be other parameters are compensating. It is also possible that output is insensitive
+to differences in the time constants.  
 The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
 demonstrates the migration of parameters to match the estimates. In spite of the
-rather wide final parameter variances, the time series'
-generated from the estimated mean seems to match the correct signal very well 
+wide final parameter variances, the time series'
+generated from the estimated mean match the correct signal very well 
 (\autoref{fig:FitComparisonLowNoise}).
@@
 \subsection{High Noise}
-\label{sec:HighNoise}
+\label{sec:SimHighNoise}
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/realization_highnoise}
-\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_highnoise}
+\caption{Test Signals with high noise ($\sigma_x = 0.01, \sigma_y=0.005$) compared to the clean signal.}
 \label{fig:HighNoiseRealization}
 \end{figure}
+
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/preprocessed_highnoise}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_highnoise}
 \caption{A comparison of the preprocessed signals for the high noise case.}
@@
 \end{figure}
+
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise}
 \caption{A comparison of the fitted signals for the high noise case.}
@@
 \end{figure}
+
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/highnoise_56_noise}
 \caption{Two particular preprocessed noise realizations for the high noise case.}
@@
 \begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise_just2}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise_just2}
 \caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
@@
 
-For the high noise simulation, the exact same procedure was performed as with the lower
-noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
+For the high noise simulation, the exact same procedure was followed as in \autoref{sec:SimLowNoise}
+except that $\sigma_y$ and $\sigma_x$ were set to $0.01$ and $0.005$,
 respectively. This is an order of magnitude higher than the previous tests, and indeed the
@@
 for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
-As before, the preprocessing led the algorithm
+The preprocessing again led the estimate
 to somewhat higher activation levels, and it would appear that the subtleties of
-different time constants are being lost due to the noise. This is easily seen
-by observing the post stimulus undershoots; many of the estimated time series
-have no post-stimulus undershoot but rather decay directly to the base level.
+different time constants are being lost due to the noise.
 
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$  & $\sqrt{MSE}$\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual  & Error\\
 \hline 
 \rowcolor[gray]{.8}
-1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
 \hline 
@@
 \end{tabular}
-\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
-The red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
-and \autoref{fig:FitComparisonHighNoiseJust2}}
+\caption{Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
+last row contains the mean.
+The red row is Run 1 and the blue row is Run 2 from  \autoref{fig:NoiseComparisonJustTwo}
+and \autoref{fig:FitComparisonHighNoiseJust2}, respectively.}
 \label{tab:HighNoiseResults} 
@@
 
-It is interesting to consider how the preprocessing and noise may effect the
-parameters of the fitting model. Consider two runs, shown in 
-\autoref{fig:NoiseComparisonJustTwo}, out of the previous eleven. 
+Two runs are shown in more detail in \autoref{fig:NoiseComparisonJustTwo}.
 Run 1 has higher peaks, in spite of the fact that both signals are based
@@
 to be more drift than the 20 measurements per knot could fit, which explains run 1's
-prolonged increase at 170 seconds in. Interestingly, run 1 and run 2 have different
-aspects of the signal that match better. Run 2 had a much better match to the 
+prolonged increase at 170 seconds. Interestingly, run 1 and run 2 emphasize
+different aspects of the signal. Run 2 had a much better match to the 
 peaks, when compared to the true signal, yet run 1 matched the post-stimulus
-undershoot better. Its worth noting however, that the post-stimulus undershoots
-here are much shorted than the observed prolonged post-stimulus undershoot
+undershoot better. Its worth noting that the post-stimulus undershoots
+here are much shorter than the observed prolonged post-stimulus undershoot
 discussed in \autoref{sec:Post Stimulus Undershoot}. 
 Despite the rather significant amount of noise, the ultimate results are 
-actually rather good. \autoref{tab:HighNoiseResults} shows
-the square-root mean-squared-error for all eleven runs, and highlights the two runs analyzed 
+actually quite good. \autoref{tab:HighNoiseResults} shows
+the error  for all eleven runs, and highlights the two runs analyzed 
 in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
@@
 \end{figure}
+
 \begin{figure}
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 1] 
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
 \end{figure}
+
 \begin{figure}
-\subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
+\subfigure[$q$, $s$, $f$, $BOLD$, Run 1 ]
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
@@
 \begin{figure}
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 2] 
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 2] 
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_2}}\\
@@
 \begin{figure}
-\subfigure[$Q$, $S$, $F$, $BOLD$, Run 2 ]
+\subfigure[$q$, $s$, $f$, $BOLD$, Run 2 ]
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
@@
 result when the particle filter converges too fast, a result of the weighting functions' variance
-being smaller than the measurement noise ($.005$ vs. $.01$). The error
+being smaller than the measurement noise ($0.005$ vs. $0.01$). The error
 certainly suffers due to this effect.
@@
 \label{sec:PureNoiseLowMag}
-The two final single-voxel tests force the particle filter to attempt to learn a noise-only
-time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
-$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
+The next two single-voxel tests force the particle filter to attempt to learn a noise-only
+time series. In this first test the noise used is the same as that from the \autoref{sec:SimHighNoise},
+$\sigma_x = 0.01, \sigma_y = 0.005$. The parameters will be set the same as well, but the
 stimulus/input neuronal efficiency ($\epsilon$) will be set to 0. In effect it is a region of the brain for
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/realization_noiseonly}
-\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
+\caption{Time-series lacking any real signal. With, ($\sigma_x = 0.01, \sigma_y=0.005$).}
 \label{fig:NoiseOnly}
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/preprocessed_noiseonly}
-\caption{A comparison of the preprocessed signals for the signal-free case. (
-$\sigma_y = .01, \sigma_x = .005$)}
+\caption{A comparison of the preprocessed signals for the signal-free case (
+$\sigma_y = 0.01, \sigma_x = 0.005$).}
 \label{fig:PreprocessedNoiseOnly}
@@
 \hline 
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSR}$   \\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  Residual   \\
 \hline 
@@
 \caption{Fits to the non-active, low noise signal. Note that the line is thick because all
-the fits overlap. This is all 11 fitted lines.}
+the fits overlap. All 11 fitted lines.}
 \label{fig:fits_noiseonly}
@@
 \centering
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$] 
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_2}}\\
@@
 \centering
-\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+\subfigure[$q$, $s$, $f$, $BOLD$ ]
 {\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_3}}
@@
 (\autoref{fig:PreprocessedNoiseOnly}). Because the stayed 
-well within the range of $.005$, which is the standard deviation of the weighting function
+well within the range of $0.005$, which is the standard deviation of the weighting function
 the particles did not ever get substantially different weights. Thus, the reason
@@
 with even higher standard deviations. To simulate this case another
-pure noise signal was generated using a $\sigma_x$ of $.1$ and a $\sigma_y$ of $.05$.
+pure noise signal was generated using a $\sigma_x$ of $.1$ and a $\sigma_y$ of $0.05$.
 
@@
 \caption{Fits to the non-active, high noise signal. Note that the line is thick because all
-the fits overlap. This is all 11 fitted lines.}
+the fits overlap. All 11 fitted lines.}
 \label{fig:fits_noiseonly_high}
@@
 %\begin{figure}[H]
-%\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
+%\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$] 
 %{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
@@
 %\begin{figure}[H]
-%\subfigure[$Q$, $S$, $F$, $BOLD$ ]
+%\subfigure[$q$, $s$, $f$, $BOLD$ ]
 %{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
@@
 to rate the results by. Unfortunately, because of the variability in the signal levels the
-raw $\sqrt{MSR}$ cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
+raw residual cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
 a low residual does not necessarily indicate a good fit. 
@@
 & \multicolumn{2}{|c|}{Low Noise} & \multicolumn{2}{|c|}{High Noise} 
-& \multicolumn{2}{|c|}{$\sigma_y = .001, \sigma_x = .0005$} 
-& \multicolumn{2}{|c|}{$\sigma_y = .01, \sigma_x = .005$}\\
+& \multicolumn{2}{|c|}{$\sigma_y = 0.001, \sigma_x = 0.0005$} 
+& \multicolumn{2}{|c|}{$\sigma_y = 0.01, \sigma_x = 0.005$}\\
 \hline 
@@
 
-Comparing the results of \autoref{sec:HighNoise} and \autoref{sec:PureNoiseHighMag}, 
+Comparing the results of \autoref{sec:SimHighNoise} and \autoref{sec:PureNoiseHighMag}, 
 distinguishing between these cases with ether normalized residual or mutual information
@@
 \hline 
-1 & 1.454& .321& .369& .036& .994& 2.774& 1.348\\
-2 &1.151& .353& .380& .026& 1.98& 2.333& 1.645 \\
-3 &1.951& .317& .348& .027& 1.657& 3.719& .757 \\
-4 &1.203 & .310& .326& .036& 2.168& 2.272& .086\\
+1 & 1.454& 0.321& 0.369& 0.036& 0.994& 2.774& 1.348\\
+2 &1.151&  0.353& 0.380& 0.026& 1.98&  2.333& 1.645 \\
+3 &1.951&  0.317& 0.348& 0.027& 1.657& 3.719& 0.757 \\
+4 &1.203 & 0.310& 0.326& 0.036& 2.168& 2.272& 0.086\\
 \hline
@@
 such a low $\epsilon$, region 4 was below the noise threshold. Notice that the
-regions 1 2 an 3 stick out in both the $\sqrt{MSR}$ and the mutual information
+regions 1 2 an 3 stick out in both the residual and the mutual information
 map, indicating that the
@@
 \caption{Histogram of estimated parameters in section 1 in voxels with mutual information greater
-than $.14$}
+than $0.14$}
 \label{fig:slicesim_hist1}
@@
 \caption{Histogram of estimated parameters in section 2 in voxels with mutual information greater
-than $.14$}
+than $0.14$}
 \label{fig:slicesim_hist2}
@@
 \caption{Histogram of estimated parameters in section 3 in voxels with mutual information greater
-than $.14$}
+than $0.14$}
 \label{fig:slicesim_hist3}

### 2010-08-14 doc/thesis/thesis.tex
added a discussion and future work section. discussion and conclusion need more work.
updated title
fixes to results
more on the results. Updated the HRF and applied pierces recommendations

@@
 
+ \usepackage[hang,small,bf]{caption}
+
 \usepackage{colortbl}
@@
 \usepackage{cancel}
-\usepackage[colorlinks=true]{hyperref}
+\usepackage[colorlinks=true,linkcolor=black]{hyperref}
+\def\chapterautorefname{Chapter}
+\def\sectionautorefname{Section}
+\def\subsectionautorefname{Section}
 
@@
 {\Large 
-Full Volume Blood-Oxygen-Level-Dependent-Signal Parameter Estimation Using Particle Filters
+Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters
 }
@@
 in \\
-Electical Engineering
+Electrical Engineering
 
@@
 particle filter to optimize the model parameters, the computation time
-is kept reasonable without requiring a linearization of the noise in the state
+is kept below a minute per voxel without requiring a linearization 
+of the noise in the state
 variables. The activation results show regions similar those found in 
@@
 \include{resultsB}
+\include{discussion}
 \include{conclusion}
 
-\bibliographystyle{ieeetr}
+\bibliographystyle{plain}
 %\bibliographystyle{apalike}

### 2010-08-15 doc/thesis/results.tex
more work on the results

@@
 $\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
-These same parameters were used for all the tests in this chapter. The noise realizations were
+These same parameters were used for all the tests in this chapter (except when the signal
+consists entirely of noise). The noise realizations were
 then generated based on a measurement noise ($\sigma_y$) of $0.001$ and a drift standard deviation
@@
 respectively. This is an order of magnitude higher than the previous tests, and indeed the
-noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
-As before, the results of preprocessing are helpful in identifying the reason for
-bias in the estimated output signal; thus the preprocessed signals are shown in 
-\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
+noise appears to dominate the output, as \autoref{fig:HighNoiseRealization} shows. 
+The preprocessed signals are shown in 
+\autoref{fig:PreprocessedHighNoise} and the results of the particle filter
 for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
-The preprocessing again led the estimate
+The noise and preprocessing again led the estimate
 to somewhat higher activation levels, and it would appear that the subtleties of
-different time constants are being lost due to the noise.
+different time constants are being lost in the noise.
 
@@
 peaks, when compared to the true signal, yet run 1 matched the post-stimulus
-undershoot better. Its worth noting that the post-stimulus undershoots
-here are much shorter than the observed prolonged post-stimulus undershoot
-discussed in \autoref{sec:Post Stimulus Undershoot}. 
-Despite the rather significant amount of noise, the ultimate results are 
+undershoot better. Despite the rather significant amount of noise, the ultimate results are 
 actually quite good. \autoref{tab:HighNoiseResults} shows
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 1] 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[$q$, $s$, $f$, $BOLD$, Run 1 ]
@@
 
-\begin{figure}
+\begin{figure}[H]
 \subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
@@
 \end{figure}
-\begin{figure}
+
+\begin{figure}[H]
 \subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 2] 
@@
 \end{figure}
-\begin{figure}
+
+\begin{figure}[H]
 \subfigure[$q$, $s$, $f$, $BOLD$, Run 2 ]
@@
 
-There are a number of interesting convergence properties of the
-particle filter when more noise is present, as both \autoref{fig:ConvergenceRuns1} and
-\autoref{fig:ConvergenceRuns2} show. The particle filter seems to converge
-significantly faster; as points tend to be further out on the weighting function. This
-also causes significantly more resampling which is the explanation for the percieved
-jumps in resolution that occur from time to time. Here it is clear that the mode and
-the mean will not be substantially different. Because of the increased noise, its likely
-that the weighting function is not sufficiently wide to account for the measurement noise.
-
-The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.
-Clearly the additional noise have resulted in much more sporadic results. This is often the 
-result when the particle filter converges too fast, a result of the weighting functions' variance
-being smaller than the measurement noise ($0.005$ vs. $0.01$). The error
-certainly suffers due to this effect.
+The particles converged much faster when more noise was present (\autoref{fig:ConvergenceRuns1}, 
+\autoref{fig:ConvergenceRuns2}). This also caused significantly more resampling which 
+is the explanation for the percieved jumps in the histograms. 
+The parameters arrived at for all 11 runs are shown in \autoref{tab:HighNoiseResults}.
+Clearly the additional noise has resulted in much more sporadic results. This is often the 
+result when the particle filter converges too fast, in this case the result of the 
+weighting functions' variance being smaller than the measurement noise ($0.005$ vs. $0.01$). 
+The error suffers due to this effect.
 
 %NO SIGNAL, LOW NOISE
-\subsection{Pure-Noise, low magnitude}
+\subsection{Pure Noise, Low magnitude}
 \label{sec:PureNoiseLowMag}
-The next two single-voxel tests force the particle filter to attempt to learn a noise-only
+The next two single-voxel tests forced the particle filter to attempt to learn a noise-only
 time series. In this first test the noise used is the same as that from the \autoref{sec:SimHighNoise},
-$\sigma_x = 0.01, \sigma_y = 0.005$. The parameters will be set the same as well, but the
-stimulus/input neuronal efficiency ($\epsilon$) will be set to 0. In effect it is a region of the brain for
-which no visual stimulus induces activity. The question then is how the
-particle filter will respond to signal that does not correlate with the input, and how
-the output may be differentiated from a time series that does. The time series'
-are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:PreprocessedNoiseOnly}.
+$\sigma_x = 0.01, \sigma_y = 0.005$. The parameters were set the same as well, but the
+stimulus neuronal efficiency ($\epsilon$) was set to 0. In effect it simulates a region 
+of the brain that does respond to the stimulus. The importance of this test is
+to determine how the
+particle filter responds to a signal that is not correlated with the input. Additionally,
+this test determines how the output may be differentiated from the output of the previous
+two sections. The time series'
+are shown in \autoref{fig:NoiseOnly} and the preprocessed versions are shown 
+in \autoref{fig:PreprocessedNoiseOnly}.
 The line fits are shown in \autoref{fig:fits_noiseonly}.
@@
 
-The data shows that the parameters have not taken any visible turns into
-unrealistic territory; individual parameters will not be a distinguishing factor
-in determining the validity of the model. Additionally the residual 
-also appears to be lower than the residual of the low noise simulation from earlier. 
+The data shows that the parameters did not converge to any unreasonable
+values; individual parameters could not be used as a distinguishing factor
+to determine the validity of the model. Additionally the residual 
+also appears to be lower than the residual of the low noise simulation from \autoref{sec:SimLowNoise}. 
 The precision with which the convergence occurred is actually the best seen in any

### 2010-08-15 doc/thesis/particlefilters.tex
another pass through single voxel sim

@@
 
-\section{Derivation}
+\section{Sequential Importance Sampling}
 The goal of the particle filter is to evolve an empirical distribution 
@@
 
-\section{Calculating Weights}
+\subsection{Calculating Weights}
 To calculate the weight of a particular particle, it is necessary to 
@@
 \item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ 
-\item The PDF $q(x_i(0))$ (the prior) fully covers $P(x_i(0))$
+\item The prior distribution PDF, $q(x_i(0))$, covers $P(x_i(0))$
 \item Markov Property: $P(x_k | x_{0:k-1}) = Pr(x_k | x_{k-1})$
@@
 
-\subsection{Basic Particle Filter Algorithm}
+\subsection{Basic Algorithm}
 From the definition of $w_i$, the algorithm to calculate
 an approximation of $P(X(t_k) | Y_{0:k})$ or $P(X(t_k + \delta t) | Y_{0:k})$
-is simple.
+is simple. This basic form of the particle filter is given in 
+algorithm \autoref{alg:BasicParticleFilter}.
 
@@
 \end{algorithmic}
+\label{alg:BasicParticleFilter}
 \end{algorithm}
 
-\section{Resampling}
+\section{Sequential Importance Resampling}
 \label{sec:Particle Filter Resampling}

### 2010-08-15 doc/thesis/results.tex
another pass through single voxel sim

@@
 and second of a single slice (64x64 voxels).  The single time series tests
-investigate the ability of particle filters to estimate the BOLD model 
-in a noisy environment. If the particle filter were unable
-to identify the \emph{true} time-series in spite of noise, then it 
-would not be a good choice for identifying parameters. Single voxel tests were
+investigate the ability of particle filters to estimate parameters of the BOLD model 
+in a noisy environment. Single voxel tests were
 performed eleven different times, each with a new noise realization.
-If runs failed to have consistent estimates of the underlying BOLD signal, then
-that would indicate a significant variance error in the model.
+Repeating tests with different noise realizations demonstrates the 
+particle filter's resiliance to noise and explores the variance of
+the model.
 
@@
 In the POSSUM simulation there were four
-discrete parameter sets driving the timeseries', however each voxel had a 
+discrete parameter sets driving the timeseries, although each voxel had a 
 different tissue composition and a different
@@
 whole brain analysis.
+
+Note that except for \autoref{sec:VeryLongSim} the same stimulus sequence
+that was used in real data (\autoref{sec:RealData}) was used in the simulations. This sequence may
+be found in \autoref{tab:flashes}, but because of the availability of ground
+truth in simulations is not displayed in this section.
 
@@
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian I.I.D. steps were added to this true signal. Finally a 
+process with Gaussian I.I.D. steps were added to the true signal. Finally a 
 carrier level was added, since BOLD is typically
@@
 Once this noisy simulated time series was generated, the particle filter algorithm
- was run on the single time series image. Five sets of tests were used 
-to determine the power of the particle filter in modeling. The first two (\autoref{sec:SimLowNoise}
-and \autoref{sec:SimHighNoise}) were
-used to demonstrate the ability of the particle filter to find the most likely
-set of parameters/state variables over the course of the run. The second two 
+was run on the single time series image. Five sets of tests 
+determine the power of the particle filter in modeling. 
+The first test demonstrates the particle filter on a very long FMRI
+run, and discusses the inherent learn-ability of the BOLD model 
+(\autoref{sec:VeryLongSim}). 
+The next two (\autoref{sec:SimLowNoise} and \autoref{sec:SimHighNoise}) 
+demonstrate the ability of the particle filter to find the most likely
+set of parameters/state variables over the course of a run identical to the
+real run in \autoref{sec:RealData}. The last two 
 tests (\autoref{sec:PureNoiseLowMag} and \autoref{sec:PureNoiseHighMag}) 
-investigated the problem of false-positives. As the first
-round of tests will show, given that there is an underlying BOLD signal, 
+investigate the problem of false-positives. As the first
+round of tests show, given the presence of an underlying BOLD signal, 
 the particle filter is excellent at finding the most probable cause of 
 the observed signal. As discussed in \autoref{sec:PureNoiseLowMag} and 
-\autoref{sec:PureNoiseHighMag}, even when there was no underlying signal, 
-the particle filter still converged. Thus, it was 
-important to investigate methods of identifying false positives.
+\autoref{sec:PureNoiseHighMag}, even when there is no underlying signal, 
+the particle filter may still converge. Because of this problem, it was 
+necessary to investigate methods of identifying false positives.
 
-The final single-voxel test was used to determine the effectiveness of the
-particle filter for a longer FMRI session. The final covariance and correlation
-matrices are also discussed.
+\subsection{Ideal FMRI Run}
+\label{sec:VeryLongSim}
+
+\begin{figure}
+\centering
+\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
+\caption{BOLD estimate converging for a very long FMRI run. Darker bars indicate
+bins with more regions. }
+\label{fig:long_converge}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
+\caption{BOLD estimate converging for a very long FMRI run, first 500 seconds. Darker bars
+indicate bins with more particles.}
+\label{fig:long_converge_500}
+\end{figure}
+
+\begin{table}[t]
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline 
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline 
+\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
+$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
+\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
+$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
+$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
+\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
+\hline 
+\end{tabular}
+\caption{Covariance matrix of the parameters at the end of \autoref{fig:long_converge}.}
+\label{tab:long_cov} 
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline 
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline 
+\rowcolor[gray]{.8} $\tau_0$  & & & & & & & \\
+$\alpha$                      & 0.889884 & & & & & & \\
+\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & & \\
+$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & & \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & & \\
+$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & & \\
+\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 & \\
+\hline 
+\end{tabular}
+\caption{Correlation of parameter estimates at the end of \autoref{fig:long_converge}.}
+\label{tab:long_corr} 
+\end{table}
+
+To begin the single voxel simulation; I generated a signal using the following parameters:
+$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
+These same parameters were used throughout this chapter. Noise
+was generated based on measurement noise ($\sigma_y$) of $0.001$ and drift standard deviation
+($\sigma_x$) of $0.0005$. The measurement noise as well as the steps of the drift
+were taken to be Gaussian. The actual signal delivered into the particle filter
+was the result of preprocessing to remove drift, as described in 
+\autoref{sec:Methods Preprocessing}.
+
+To test how well the particle filter would do plenty of data, and 
+determine the inherent variance in the model parameters, a very long simulation
+with randomly generated impulse stimulus was created. The preprocessed timeseries
+and the final estimate are shown in \autoref{fig:long_converge};
+the final covariance matrix of the parameters is in \autoref{tab:long_cov}.
+Note that even though the BOLD response converged well (\autoref{fig:long_converge}),
+the parameters still have significant correlation (\autoref{tab:long_corr}). 
+This is in spite of the noisy input (green line in \autoref{fig:long_converge}).
+Based on the histograms in the first 500 seconds, the parameters converged to
+their final values well before the end.
+Although this is only a single test, the correlation (\autoref{tab:long_corr}) 
+of the parameters 
+indicates significant parameter indeterminability. When the input
+consists entirely of impulses, the best parameters are not one particular
+set, but a spectrum. Note that the correlation is in parameters whose priors
+are completely independent. It is possible that varying the type of input could
+give improved results, although informal tests did not show significant difference.
 
 %LOW NOISE SECTION, with a signal
-\subsection{Signal with Low Noise}
+\subsection{Simulation with Low Noise}
 \label{sec:SimLowNoise}
@@
 \end{figure}
-
-To begin the single voxel simulation; I generated a signal using the following parameters:
-$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
-These same parameters were used for all the tests in this chapter (except when the signal
-consists entirely of noise). The noise realizations were
-then generated based on a measurement noise ($\sigma_y$) of $0.001$ and a drift standard deviation
-($\sigma_x$) of $0.0005$. The measurement noise as well as the steps of the drift
-were taken to be Gaussian.
-For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
-The actual signal delivered into the particle filter
-was the result of preprocessing to remove drift, as described in 
-\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
-could have some effect on the resulting fit; thus the preprocessed signal is compared
-to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
-The data shows that the spline struggles to match the signal near the end, although 
-overall, the preprocessing seems to have been successful at removing trends.  In this case
-the cause of the problem was the final stimuli that pushed the final median well above the base
-level. The estimated BOLD signal is compared to the true BOLD signal in \autoref{fig:FitComparisonLowNoise}.
-At several time points the particle filter successfully filtered the input. 
-For instance, in the last 30 seconds
-the estimates stay flat in spite of the preprocessed data drifting off. By
-this point, the algorithm has converged sufficiently to prevent inexplicable movement.
-A similar circumstance occurs at 100 seconds in. Note that bias introduced by pre-processing 
-results in a slightly larger peak signal. The final parameter sets are shown in 
-\autoref{tab:LowNoiseResults}. Note that throughout the results, unless otherwise specified,
-the term \emph{residual} will be defined as the square root of the mean squared residual, 
-\begin{equation}
-\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
-\end{equation}
-where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
-sampled at time $k$. Note the subtle distinction between this and what will be 
-termed the \emph{error} which is defines as 
-\begin{equation}
-\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
-\end{equation}
-where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
-signal. Throughout this section and the next, the terms error and residual
-will be referencing these values.
 
@@
 \end{table}
+
+The following tests all use less data, on par with the average length of an FMRI
+session.  For this low noise case, the eleven realizations are shown in 
+\autoref{fig:LowNoiseRealization}.
+The bias introduced into the signal by preprocessing 
+could have some effect on the resulting fit; thus the preprocessed signal is compared
+to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
+Overall, \autoref{fig:PreprocessedLowNoise} shows that the preprocessing was 
+effective at removing trends. The spline did struggle to match the signal 
+torward the end. This slight drift effect was caused by the final median being 
+well above the base
+level. The estimated BOLD signal is compared to the true BOLD signal in \autoref{fig:FitComparisonLowNoise}.
+At several time points the particle filter successfully filtered the input. 
+For instance, in the last 30 seconds
+the estimates stay flat in spite of the preprocessed data drifting off. By
+this point, the algorithm had converged sufficiently to prevent such inexplicable movement.
+A similar circumstance occurs at 100 seconds in. A combination of 
+noise and preprocessing biased the results toward a peak signal above the true peaks. 
+The final parameter sets are shown in 
+\autoref{tab:LowNoiseResults}.
+
+Note that throughout the results, unless otherwise specified,
+the term \emph{residual} will be defined as the square root of the mean squared residual, 
+\begin{equation}
+\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
+\end{equation}
+where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
+sampled at time $k$. Note the subtle distinction between this and what will be 
+termed the \emph{error} which is defines as 
+\begin{equation}
+\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
+\end{equation}
+where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
+signal. Throughout this section and the next, the terms error and residual
+will be referencing these values.
 
@@
 \label{fig:LowNoiseHist}
+\caption{Convergence of the first run from \autoref{tab:LowNoiseResults}. The bars represent
+a histogram, where darker bars indicate more particles in that bin.}
 \end{figure}
@@
 greatly across
-runs with different noise realizations, yet the sum of the individual time constants
-($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
-time constants fall short of the true time constant. This could be a limitation
+runs, yet the sum of the individual time constants
+($\tau_f$, $\tau_s$ and $\tau_0$) were more consistent. In general the 
+time constants fell short of the true time constant. This could be a limitation
 based on the prior distribution (which notably has an initial mean below the true values) 
-or it could be other parameters are compensating. It is also possible that output is insensitive
-to differences in the time constants.  
+or it could be that other parameters compensated. It is also possible that the output is insensitive
+to small differences in the time constants.  
 The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
-demonstrates the migration of parameters to match the estimates. In spite of the
-wide final parameter variances, the time series'
-generated from the estimated mean match the correct signal very well 
+demonstrates the migration of parameters through the run. In spite of the
+significant differences in parameter estimates, the estimated BOLD
+consistently performed well 
 (\autoref{fig:FitComparisonLowNoise}).
@@
 \label{sec:SimHighNoise}
-\begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_highnoise}
-\caption{Test Signals with high noise ($\sigma_x = 0.01, \sigma_y=0.005$) compared to the clean signal.}
-\label{fig:HighNoiseRealization}
-\end{figure}
+%\begin{figure}
+%\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_highnoise}
+%\caption{Test Signals with high noise ($\sigma_x = 0.01, \sigma_y=0.005$) compared to the clean signal.}
+%\label{fig:HighNoiseRealization}
+%\end{figure}
 
@@
 respectively. This is an order of magnitude higher than the previous tests, and indeed the
-noise appears to dominate the output, as \autoref{fig:HighNoiseRealization} shows. 
-The preprocessed signals are shown in 
-\autoref{fig:PreprocessedHighNoise} and the results of the particle filter
+noise appears to dominate the output, as \autoref{fig:PreprocessedHighNoise} shows. 
+The results of the particle filter
 for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
-The noise and preprocessing again led the estimate
-to somewhat higher activation levels, and it would appear that the subtleties of
-different time constants are being lost in the noise.
+The noise and preprocessing again led the estimates
+to higher peak activation levels, and the subtleties of different time constants 
+are being lost in the noise.
 
@@
 
-Two runs are shown in more detail in \autoref{fig:NoiseComparisonJustTwo}.
-Run 1 has higher peaks, in spite of the fact that both signals are based
-on the same underlying signal. There also appears
-to be more drift than the 20 measurements per knot could fit, which explains run 1's
-prolonged increase at 170 seconds. Interestingly, run 1 and run 2 emphasize
+\autoref{fig:NoiseComparisonJustTwo} shows two of the runs in more detail.
+There appears
+to be more drift than the 20 measurements per knot could fit, which explains 
+the prolonged increase at 170 seconds in Run 1; although such areas
+permeate the preprocessed signals.
+Interestingly, Run 1 and Run 2 emphasize
 different aspects of the signal. Run 2 had a much better match to the 
-peaks, when compared to the true signal, yet run 1 matched the post-stimulus
-undershoot better. Despite the rather significant amount of noise, the ultimate results are 
-actually quite good. \autoref{tab:HighNoiseResults} shows
+peaks, when compared to the true signal, yet Run 1 matched the post-stimulus
+undershoot better.  \autoref{tab:HighNoiseResults} shows
 the error  for all eleven runs, and highlights the two runs analyzed 
@@
 The parameters arrived at for all 11 runs are shown in \autoref{tab:HighNoiseResults}.
-Clearly the additional noise has resulted in much more sporadic results. This is often the 
+Clearly the additional noise resulted in much more sporadic results. This is often the 
 result when the particle filter converges too fast, in this case the result of the 
-weighting functions' variance being smaller than the measurement noise ($0.005$ vs. $0.01$). 
-The error suffers due to this effect.
+weighting function's variance being smaller than the measurement noise ($0.005$ vs. $0.01$). 
+The error clearly suffers due to this effect (\autoref{tab:HighNoiseResults}).
 
@@
 The next two single-voxel tests forced the particle filter to attempt to learn a noise-only
-time series. In this first test the noise used is the same as that from the \autoref{sec:SimHighNoise},
-$\sigma_x = 0.01, \sigma_y = 0.005$. The parameters were set the same as well, but the
-stimulus neuronal efficiency ($\epsilon$) was set to 0. In effect it simulates a region 
-of the brain that does respond to the stimulus. The importance of this test is
-to determine how the
-particle filter responds to a signal that is not correlated with the input. Additionally,
-this test determines how the output may be differentiated from the output of the previous
-two sections. The time series'
-are shown in \autoref{fig:NoiseOnly} and the preprocessed versions are shown 
-in \autoref{fig:PreprocessedNoiseOnly}.
-The line fits are shown in \autoref{fig:fits_noiseonly}.
+time series. In this test the noise was the same as that from the \autoref{sec:SimHighNoise},
+$\sigma_x = 0.01, \sigma_y = 0.005$. The stimulus neuronal efficiency ($\epsilon$) was set 
+to 0, in effect simulating a brain region with no response to the stimuli.
+This test was used to determine how the output of a pure noise timeseries
+is different from that of a simple noisy signal (as in the previous two sections). 
+The preprocessed signals are shown in \autoref{fig:PreprocessedNoiseOnly}
+and line fit for each run is shown in \autoref{fig:fits_noiseonly}.
+
+%\begin{figure}[H]
+%\centering
+%\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/realization_noiseonly}
+%\caption{Time-series lacking any real signal. With, ($\sigma_x = 0.01, \sigma_y=0.005$).}
+%\label{fig:NoiseOnly}
+%\end{figure}
 
@@
 \centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/realization_noiseonly}
-\caption{Time-series lacking any real signal. With, ($\sigma_x = 0.01, \sigma_y=0.005$).}
-\label{fig:NoiseOnly}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=7cm]{images/preprocessed_noiseonly}
-\caption{A comparison of the preprocessed signals for the signal-free case (
+\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=15cm]{images/preprocessed_noiseonly}
+\caption{Comparison of the preprocessed signals for the low noise signal-free case (
 $\sigma_y = 0.01, \sigma_x = 0.005$).}
@@
 \caption{Fits to the non-active, low noise signal. Note that the line is thick because all
-the fits overlap. All 11 fitted lines.}
+the estimates overlap. All 11 fitted lines.}
 \label{fig:fits_noiseonly}
@@
 
-The data shows that the parameters did not converge to any unreasonable
-values; individual parameters could not be used as a distinguishing factor
-to determine the validity of the model. Additionally the residual 
-also appears to be lower than the residual of the low noise simulation from \autoref{sec:SimLowNoise}. 
-The precision with which the convergence occurred is actually the best seen in any
-simulation.  The reason for the low error is that the overall
-signal is significantly smaller than any other simulation in this thesis.
-The peaks never even 1\% difference, with most staying even below .5\% 
-(\autoref{fig:PreprocessedNoiseOnly}). Because the stayed 
-well within the range of $0.005$, which is the standard deviation of the weighting function
-the particles did not ever get substantially different weights. Thus, the reason
-the fits are in such agreement is that no particles gained increased weight, and so
-no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
-This is good evidence that a per time-series weighting function could benefit
-even out the convergence rate; and thus allow better determination between signals
-like this and low noise case from earlier.
-Thus we have now seen how the particle filter responds when the signal is completely
-random and low level, however there is also some possibility that the signal will
-be random, yet active. Thus one more test is necessary; time series with no 
-underlying signal (as far as responding stimuli) but with enough noise to 
-actually be considered a signal.
+The data shows that the parameters did really not converge
+\autoref{fig:JustNoiseConvergence}.
+The peaks never even reached 1\% difference
+(\autoref{fig:PreprocessedNoiseOnly}) so  the signal  stayed 
+well within the range of $0.005$, the standard deviation of the weighting function.
+Note that the residuals
+were actually lower than the residuals in the low noise simulation from \autoref{sec:SimLowNoise}
+and the parameter estimates were extremely consistent across 11 runs. 
+The low residual was caused by the overall
+signal being significantly smaller than any previous simulation.
+From these results it would seem that lack of convergence could be
+a differentiating factor from a signal with an actual signal; however,
+the next section casts doubt on this possibility.
 
@@
 \label{sec:PureNoiseHighMag}
-To determine how the particle filter will respond to active, yet unrelated
-portions of the brain, in this section random data will again be used, but
-with even higher standard deviations. To simulate this case another
-pure noise signal was generated using a $\sigma_x$ of $.1$ and a $\sigma_y$ of $0.05$.
+To determine how the particle filter responds to active, yet unrelated
+portions of the brain, this section repeats the previous with much higher
+noise peaks. To simulate this case another
+pure noise signal was generated using $\sigma_x$ of $.1$ and $\sigma_y$ of $0.05$.
 
-As before, the convergence all seems to follow a similar path, leaving almost no
-variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
-the peaks are comparable to peaks in a real timeseries. Interestingly the algorithm
-suffered from almost constant particle deprivation, meaning that the heuristic
+As before, the convergence all follows a similar path, leaving almost no
+variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). 
+Interestingly the algorithm suffered from almost constant particle deprivation, 
+meaning that the heuristic
 for rescuing the particle filter from particle deprivation, discussed in
-\autoref{sec:Resampling} in fact was working against the proper course of action. 
+\autoref{sec:Resampling}, in fact was working against the proper course of action. 
 The proper course of action in this case would be for the particle filter to fail,
@@
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
- occur if the prior happens to not be dense enough in the area of the solution. Thus,
- there is an issue of false positives vs. false negatives. A middle ground, such as 
- allowing deprivation to occur if the recovery mechanism has already triggered once
- may be an acceptable compromise, however it would be difficult to strike the 
- right balance. Instead, for the purposes here, I depend on other methods, described
- later, evaluate if the results were reasonable or not.
+occur if the prior happens to miss the solution.
 
@@
 \centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly_high}
-\caption{Fits to the non-active, high noise signal. Note that the line is thick because all
-the fits overlap. All 11 fitted lines.}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3cm,width=15cm]{images/fits_noiseonly_high}
+\caption{BOLD estimates for the non-active, high noise signal. Note the line thickness is caused
+by all the estimates overlapping.}
 \label{fig:fits_noiseonly_high}
@@
 
-From \autoref{fig:justbignoise_fit_0} it is clear that, despite the fact that the input 
-is competely detached from the signal, a non-zero response by the particle filter does
-lead to decreased residual. This is primarily due to the preprocessing which forces the signal to
-have a non-zero mean. While this scheme could create a type of artificial activation for
-the particle filter, in reality the residual will still be less than if the signal were left
-with zero mean, and the particle filter stayed flat. At best the particle filter will 
-over-smooth, estimating a centerline to noisy signal. In such a case
-the residual will be the same as the alternative, although the mutual information will 
-be different.
-
-\subsection{Very Long FMRI Run}
-\begin{figure}[H]
-\centering
-\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
-\caption{BOLD estimate converging for a very long FMRI run.}
-\label{fig:long_converge}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
-\caption{BOLD estimate converging for a very long FMRI run, first 500 seconds.}
-\label{fig:long_converge_500}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|c | c  c  c  c  c  c  c |}
-\hline 
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-\hline 
-\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
-$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
-\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
-$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
-$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
-\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
-\hline 
-\end{tabular}
-\caption{Covariance matrix of the parameters at the end of a run.}
-\label{tab:long_cov} 
-\end{table}
-
-\begin{tabular}{|c | c  c  c  c  c  c  c |}
-\hline 
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-\hline 
-\rowcolor[gray]{.8} $\tau_0$  & & & & & & & \\
-$\alpha$                      & 0.889884 & & & & & & \\
-\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & & \\
-$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & & \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & & \\
-$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & & \\
-\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 & \\
-\hline 
-\end{tabular}
-\begin{table}[t]
-\centering
-
-\caption{Correlation of parameter estimates at the end of a run.}
-\label{tab:long_corr} 
-\end{table}
-
-To test how well the particle filter would do given much more data, and 
-determine the inherent variance in the model parameters, a very long simulation
-with randomly generated impulse input was created. The preprocessed timeseries
-and the final estimate are shown in \autoref{fig:long_preprocessed}, 
- and the final covariance matrix of the parameters is shown in \autoref{tab:long_cov}.
-Note that even though the
-BOLD response converged well (\autoref{fig:long_converge}),
-the parameters still have significant correlation (\autoref{tab:long_corr}). This points
-to a certain amount of indeterminability of the parameters.
+Because of the preprocessing steps, the pure noise signal can look markedly like a real
+signal (\autoref{fig:justbignoise_fit_0}). The preprocessing also causes the particle filter
+to converge to a non-zero response in spite of the fact that the input does not correlate
+with the stimuli in any way (\autoref{fits_noiseonly_high}).
 

### 2010-08-15 doc/thesis/results.tex
low/high noise signal case reviewed

@@
 
-To test how well the particle filter would do plenty of data, and 
+To test how well the particle filter would do with plenty of data, and 
 determine the inherent variance in the model parameters, a very long simulation
-with randomly generated impulse stimulus was created. The preprocessed timeseries
+with randomly generated impulse stimuli was created. The preprocessed timeseries
 and the final estimate are shown in \autoref{fig:long_converge};
@@
 the parameters still have significant correlation (\autoref{tab:long_corr}). 
-This is in spite of the noisy input (green line in \autoref{fig:long_converge}).
 Based on the histograms in the first 500 seconds, the parameters converged to
@@
 give improved results, although informal tests did not show significant difference.
+In spite of the noisy input (green line in \autoref{fig:long_converge}), the estimates
+of the BOLD were actually very close to the true (noise-free) BOLD signal.
 
@@
 
-The following tests all use less data, on par with the average length of an FMRI
-session.  For this low noise case, the eleven realizations are shown in 
+\begin{figure}
+\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise1}}\\
+
+\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $v$ of the first run, low noise simulation.]
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
+\end{figure}
+
+\begin{figure}
+\subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
+\caption{Convergence of the first run from \autoref{tab:LowNoiseResults}. The bars represent
+a histogram, where darker bars indicate more particles in that bin.}
+\label{fig:LowNoiseHist}
+\end{figure}
+
+The following tests all use less data (shorter sequences), on par with 
+the average length of an FMRI session. For this low noise case 
+$(\sigma_y = 0.001,\ \sigma_x = .0005)$, the eleven realizations are shown in 
 \autoref{fig:LowNoiseRealization}.
@@
 Overall, \autoref{fig:PreprocessedLowNoise} shows that the preprocessing was 
-effective at removing trends. The spline did struggle to match the signal 
+effective at removing trends, although the spline did struggle to match the signal 
 torward the end. This slight drift effect was caused by the final median being 
 well above the base
-level. The estimated BOLD signal is compared to the true BOLD signal in \autoref{fig:FitComparisonLowNoise}.
-At several time points the particle filter successfully filtered the input. 
+level. 
+At several time points the particle filter successfully filtered the input
+(\autoref{fig:FitComparisonLowNoise}).
 For instance, in the last 30 seconds
@@
 this point, the algorithm had converged sufficiently to prevent such inexplicable movement.
-A similar circumstance occurs at 100 seconds in. A combination of 
+A similar circumstance occurs at around 100 seconds in. A combination of 
 noise and preprocessing biased the results toward a peak signal above the true peaks. 
@@
 %\end{table}
-
-\begin{figure}
-\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise1}}\\
-
-\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $v$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
-\end{figure}
-
-\begin{figure}
-\subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
-\label{fig:LowNoiseHist}
-\caption{Convergence of the first run from \autoref{tab:LowNoiseResults}. The bars represent
-a histogram, where darker bars indicate more particles in that bin.}
-\end{figure}
 
@@
 
-The particles converged much faster when more noise was present (\autoref{fig:ConvergenceRuns1}, 
-\autoref{fig:ConvergenceRuns2}). This also caused significantly more resampling which 
+The particles converged much faster when more noise was present (\autoref{fig:LowNoiseHist} vs. 
+\autoref{fig:ConvergenceRuns1}, \autoref{fig:ConvergenceRuns2}). 
+This also caused significantly more resampling which 
 is the explanation for the percieved jumps in the histograms. 

### 2010-08-15 doc/thesis/results.tex
mostly finished with simulation results

@@
 
-\subsection{Ideal FMRI Run}
+\subsection{Ideal Analysis}
 \label{sec:VeryLongSim}
@@
 \subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
+{\label{fig:LowNoiseHistc} \includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
 \caption{Convergence of the first run from \autoref{tab:LowNoiseResults}. The bars represent
@@
 %HIGH NOISE SECTION, with a signal
-\subsection{High Noise}
+\subsection{Simulation with High Noise}
 \label{sec:SimHighNoise}
@@
 \subfigure[$q$, $s$, $f$, $BOLD$, Run 1 ]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
+{\label{fig:ConvergenceRuns1c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
 \caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
@@
 \subfigure[$q$, $s$, $f$, $BOLD$, Run 2 ]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
+{\label{fig:ConvergenceRuns2c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
 \caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
@@
 This also caused significantly more resampling which 
-is the explanation for the percieved jumps in the histograms. 
-The parameters arrived at for all 11 runs are shown in \autoref{tab:HighNoiseResults}.
-Clearly the additional noise resulted in much more sporadic results. This is often the 
+is the explanation for the perceived jumps in the histograms. 
+Clearly the additional noise resulted in much more sporadic results (\autoref{tab:HighNoiseResults}). 
+This is often the 
 result when the particle filter converges too fast, in this case the result of the 
@@
 The error clearly suffers due to this effect (\autoref{tab:HighNoiseResults}).
+Note that neither Run 1 or Run 2 estimated the underlying state well (\autoref{fig:ConvergenceRuns1c}
+and \autoref{fig:ConvergenceRuns2c}), whereas
+in the previous test the particle filter was extremely successful in this area (\autoref{fig:LowNoiseHistc}).
 
@@
 To determine how the particle filter responds to active, yet unrelated
-portions of the brain, this section repeats the previous with much higher
-noise peaks. To simulate this case another
-pure noise signal was generated using $\sigma_x$ of $.1$ and $\sigma_y$ of $0.05$.
+portions of the brain, this section repeats the test of \autoref{sec:PureNoiseLowMag} 
+with much higher noise peaks. To simulate this case another
+pure noise signal was generated using $\sigma_x$ of $0.1$ and $\sigma_y$ of $0.05$.
 
@@
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
-occur if the prior happens to miss the solution.
+occur in good data if the prior distribution does not cover a solution.
+
+Because of the preprocessing steps, the pure noise signal can look markedly like a real
+signal (\autoref{fig:justbignoise_fit_0}). The preprocessing causes the particle filter
+to converge to a non-zero response in spite of the fact that the input does not correlate
+with the stimuli in any way (\autoref{fig:fits_noiseonly_high}).
 
@@
 
-Because of the preprocessing steps, the pure noise signal can look markedly like a real
-signal (\autoref{fig:justbignoise_fit_0}). The preprocessing also causes the particle filter
-to converge to a non-zero response in spite of the fact that the input does not correlate
-with the stimuli in any way (\autoref{fits_noiseonly_high}).
-
 \subsection{Single Voxel Review}
 \label{sec:SingleVoxelReview}
-
 The first step in determining the validity of a model is to provide some order of quality
-to rate the results by. Unfortunately, because of the variability in the signal levels the
-raw residual cannot perform this task. As demonstrated by \autoref{sec:PureNoiseLowMag},
+to rate the results by. Unfortunately, because of the variability in the signal levels, the
+raw residual cannot perform this task. As demonstrated by \autoref{tab:NoiseOnlyResults},
 a low residual does not necessarily indicate a good fit. 
@@
 signal, I used a robust estimator of 
-scale; the Median-Absolute-Deviation (MAD). This is an estimator of the standard
+scale; the Median-Absolute-Deviation (MAD) which is defined in equation \autoref{eq:mad}.
+This is an estimator of the standard
 deviation, and thus a good estimator of the scale of the input signal. 
 The normalized residual values are 
-shown in \autoref{tab:SignleVoxelActivationComparison}. A second potential method of 
+shown in \autoref{tab:SingleVoxelActivationComparison}. A second potential method of 
 gauging performance is mutual information. Mutual information is a method of measuring
 the interdependence of two random variables. If two signals are truly independent,
-then the mutual information will be zero. Although ideally suited to Bernoulli 
-random variables where there are a finite number of options, by using histograms
+then the mutual information will be zero. Although ideally suited to discrete
+distributions, by using histograms
 it is possible to derive a joint distribution of two signals. The algorithm for 
@@
 \end{equation}
-Unfortunately the number of bins causing bias in the output, thus to correct for this,
+Unfortunately the number of bins causes bias in the output, thus to correct for this,
 I subtracted the estimated bias:
@@
 \end{equation}
-After calculating the empirical mutual information, I subtracted the bias from
-the estimated mutual information. Note that this can result in negative mutual
+ Note that subtracting the bias can result in negative mutual
 information, which should not technically be possible; so any negative mutual information
@@
     for signal/noise configurations.}
-\label{tab:SignleVoxelActivationComparison} 
+\label{tab:SingleVoxelActivationComparison} 
 \end{table}
 
-Comparing the results of \autoref{sec:SimHighNoise} and \autoref{sec:PureNoiseHighMag}, 
+Comparing the results of \autoref{sec:SimHighNoise} and \autoref{sec:PureNoiseHighMag} 
+in \autoref{tab:SingleVoxelActivationComparison}, 
 distinguishing between these cases with ether normalized residual or mutual information
 is not clear cut. While the average mutual information is in fact more than 10 times
-greater than the average mutual information in the two non-signal cases, the maximum
+the average mutual information in the two non-signal cases, the maximum
 mutual information of the low noise/no signal case exceeds the minimum M.I. of the high noise/signal
@@
 low noise/no signal case does not concur with the worse case (minimum) normalized residual. 
-There is no reason why this has to be the case, it is beneficial. In
+There is no reason why this has to be the case, but it is beneficial. In
 other words, if it were necessary to make a statement that
-a particular voxel active or inactive; the accuracy would improve if both techniques
-were used with loose restrictions, but a voxel would only be considered active if the
-voxel passed both tests.
+a particular voxel were active or inactive; the accuracy would improve if both techniques
+were used with loose restrictions.
 
@@
 techniques it is important to ensure consistency of results. Although the parameter
-sets were somewhat inconsistent, the quality of the fit was actually quite consistent
-across eleven runs. The second purpose was of course to determine how the particle 
-filter responded to different signal to noise ratios.
+sets were inconsistent, the quality of the fit was actually quite consistent
+over eleven runs. The second purpose was of course to determine how the particle 
+filter responded to different signal to noise ratios. From \autoref{sec:SimLowNoise},
+the particle filter performed extremely well in filtering out nonsensical measurements.
 
@@
 \label{sec:Multi-voxel Simulation}
-To test the usefulness of the particle filter for I used a modified version of the FSL tool 
+To test the usefulness of the particle filter on a larger scale, I used a modified version 
+of the FSL tool 
 POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
@@
 to take a parameter map and generate activation levels depending on the parameters at that
-point. The patch for POSSUM will be made available. As an unfortunate side effect of 
-not using Possums' original activation scale, I manually added 750 to the total level of
-simulated Possum images. This is because the BOLD \% difference levels were in the range
-of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
-this has no effect on the quality of the simulation and should not effect the 
-underlying parameters (other than perhaps $\epsilon$ and $V_0$). 
-Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
-program the true signal-to-noise ratio was much lower, as seen in the fit in \autoref{fig:simslicefit1}.
+point. The patch for POSSUM will be made available. As side effect of 
+not using POSSUM's original activation scale, the BOLD \% difference levels 
+were in the range of 50 - 100\% from the base, about 5 times as large as they 
+should have been. I therefore manually added 750 to the total level of
+simulated POSSUM images, putting results into the 1 - 5\% range. This 
+had no effect on the quality of the simulation and should not effect the 
+underlying parameters (other than $V_0$). 
+Also worth noting is that the noise level was set to an SNR of 20, due to changes 
+in the program the true signal-to-noise ratio was much lower, as seen in the 
+fit in \autoref{fig:simslicefit1}.
 
+%todo where is this from?
 \begin{figure}
 \centering
-\includegraphics[width=15cm]{images/simslice_noise}
-\caption{A single time-series fit in the simulated slice.}
+\includegraphics[clip=true,trim=5cm 2cm 4.5cm 2cm,width=15cm]{images/simslice_noise}
+\caption{A single time-series fit in the simulated slice. Preprocessed BOLD is
+the noisy response generated by POSSUM.}
 \label{fig:simslicefit1}
@@
 
-For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
-into a parameter map. This parameter map may then be compared to the map used to generate the 
-simulated data; additionally a new simulation using the calculated parameters may also be 
+For each time-series in the simulated FMRI image, the final parameters were saved
+into a parameter map. This parameter map could then be compared to the map used to generate the 
+simulated data; additionally a new simulation using the calculated parameters could also be 
 generated to test the difference in BOLD levels between the real parameters and the
-estimated ones. Since it is the parameters are far from orthogonal 
-(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
-despite large difference. This way, a quantitative difference between the two parameter sets 
-was found.
+estimated ones. Since the parameters were far from orthogonal 
+\cite{Deneux2006}, this provided a quantitative difference between the two parameter sets.
 
@@
 
+%todo quantify the number of false positives, add extra Heatmap removing false positives
 The histograms again demonstrate that a single point estimate of the parameters 
-is elusive for the current model. However the data clearly show the power
+is elusive for this set of parameters. However the data clearly show the power
 of the particle filter at identifying regions of activation; which is usually
@@
 research. Tighter thresholds removes the false positives present
-in the images at the cost of false negatives. Interestingly the false 
-positives present in the Mutual Information
+in the images at the cost of false negatives. As noted in \autoref{sec:SingleVoxelReview}
+the false positives present in the Mutual Information
 map are different from those in the residual map. This furthers the argument from
-\autoref{sec:SingleVoxelReview} for combining the two metrics to 
-increase power. Although
+for combining the two metrics to increase power. Although
 at first glance it would appear that there are false negatives in the 
@@
 does not typically  have a BOLD response. This is why there are holes in regions
-2 and 3. These results certainly indicate that the particle filter is effective
-at regressing against a noisy signal. Therefore the next section moves forward
-with real FMRI data.
+2 and 3. These results indicate that the particle filter is effective
+at regressing against a noisy signal.

### 2010-08-15 doc/thesis/resultsB.tex
mostly finished with simulation results

@@
-%\chapter{Real Data}
-%Finally, we also performed inference on a real FMRI scan. The scanner we used...
-%... more specifics...
-
-% TODO include single?
-%Before performing tests on a full image, I the particle filter
-%on regions deemed active and inactive by statistical parametric mapping
-%(SPM). This served the purpose of adjusting the priors as well as the 
-%preprocessing based on real world signals. This was actually done before
-%the simulations, and then results were carried back the simulations 
-%to check consistency.
-%After work adjusting parameters, most importantly the weighting function and the 
-%priors, particle filter was applied to every voxel in an FMRI image.
-%The results of this large scale analysis was a parameter map which was
-%then used to calculate normalized square-root MSE image. 
-%
-%\section{Single-Voxel Analysis}
-%The choice of a prior, as discussed previously, is extremely important. While a
-%prior may have the potential to give good results, being a monte-carlo algorithm
-%there is the possibility for inconsistencies. Thus, increasing the variance
-%of the time-constants may allow additional flexibility, it will also cause
-%additional model variance. Before running on a full volume I adjusted the 
-%prior to ensure that the same input would give the same output 100 times in a 
-%row. While this may seem like a given, with a random drawing of the prior,
-%this can be difficult. Case in point, the exact same algorithm run twice
-%with the time constants all having standard deviations of $.35$ resulted in two
-%very different fits, shown in \autoref{fig:badfit_param1}.
-%
-%\begin{figure}
-%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/badfit_param1}}
-%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/goodfit_param1}}
-%\caption{The same priors gave rise to both fits.}
-%\label{fig:badfit_param1}
-%\end{figure}
-%
-%For this reason, I actually lowered the standard deviations of the time
-%constants to prevent over-smoothing. This resulted in more consistent,
-%though potentially slightly worse fits, two examples of which are 
-%shown in \autoref{fig:param2_var}. 
-%
-%\begin{figure}
-%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/param2a}}
-%\subfigure{\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/param2b}}
-%\caption{A poor fit, using the same parameters as }
-%\label{fig:param2_var}
-%\end{figure}
-%
-%todo: stats of the 100 fits?
-%
-%\section{Single Time-Series Simulation}
-%
 \chapter{Real Data}
 \label{sec:RealData}
-Modeling the BOLD response is of course not of much use if does not work for
-real FMRI.  Although this algorithm will hopefully lead to more novel methods 
-of analysis, the standard use for modeling the BOLD signal is to locate activation.
-Activation is defined as areas where the input is the primary drive
-for the BOLD response, as opposed to intermediate factors controlling it. 
-Once areas where the BOLD 
-model may be accurately estimated are found, integrating the model will allow for accurate
-estimation of the state between measurements, which could then be used for more 
-advanced analysis, for instance of areas that are being driven by other brain regions.
-It all begins with localizing the first activation regions in the
-chain. Therefore this section compares the output of the particle filter 
-with conventional SPM.
+The ultimate goal of modeling the BOLD response is to estimate parameters
+in real FMRI data. The most basic use of modeling the BOLD signal is to
+locate activation. A voxel is considered active when the stimulus 
+is the primary drive for the BOLD response. This is in contrast to 
+most voxels which have are controlled by intermediate or complete unrelated factors. 
+Inactive regions cannot be modeled because their input is unknown, therefore
+parameter estimates in such regions are impossible. 
+Because SPM is the de facto standard for localizing activation,
+this section compares its output with that of the particle filter.
 
-The data described in this section are fundamentally different in several 
-ways. SPM preprocesses the image by spatially smoothing the FMRI image (in this section 
-SPM8 was used with an $8mm x 8mm x 8mm$ Gaussian kernel), whereas
-this is not done in the particle filter algorithm. Additionally, a spline
-was used to de-trend, rather than SPM8's high pass filter (with a cut
-off based on a globally estimated autocorrelation). Thus the preprocessing pipelines 
-are different; but the output of SPM8 is also different. Whereas SPM outputs
-a t-statistic for each voxel, the output 
-of the particle filter is a posterior probability distribution of the parameters
-at every voxel. To validate the quality of the particle filter results 
-it is necessary to compare the both the location and the fit calculated by the particle filter
-with SPM's location and fit.
+%todo check SPM smoothing parameter
+Note that SPM must pre-process the image with a spatial smoothing 
+filter. For this work SPM8 was smoothed with a
+$8\text{mm} \times  8\text{mm} \times 8\text{mm}$ Full-Width Half Maximum
+(FWHM) Gaussian kernel.  
+Additionally, SPM8 applied a high pass filter (with a cut
+off based on a globally estimated autocorrelation). Thus the preprocessing pipeline
+of SPM is very different from that of the particle filter. SPM8 also outputs
+a T-statistic for each voxel, whereas the 
+particle filter's primary output is a posterior probability distribution of the parameters
+at every voxel. To validate the quality of the particle filter, the results 
+were compared with SPM, both in terms of the location and the fit.
 
+%todo check scanner information/stimulus
 \section{FMRI Configuration}
@@
 
+%todo the stimulus
 \begin{table}[t]
@@
 
-After dropping the first 9 volume, each of the remaining volumes was
-co-registered with the first volume in time. At this point the SPM method
+%spm options
+After dropping the first 9 volumes, each of the remaining volumes was
+co-registered with the first. At this point the SPM method
 diverges from the experimental method. For the particle filter, detrending 
@@
 data was processed with the particle filter. To generate the SPM output
-in this section, the co-registered data was passed through a high pass
-filter with a cut off of (TODO). The data was then smoothed spatially 
-with a Gaussian kernel (FWHM 8x8x8 voxels). A Statistical Parametric Map
-was then generated using SPM8. All settings used in SPM8 were the defaults
+in this section, the co-registered data was spatially smoothed,
+and then filtered with an adaptive cut off (which is built into the
+SPM analysis). For the SPM analysis, the Canonical HRF was used
+and model derivatives were included. All settings used in SPM8 were the defaults
 for analyzing FMRI data.

### 2010-08-15 doc/thesis/results.tex
added some more images and fixes to results

@@
 to take a parameter map and generate activation levels depending on the parameters at that
-point. The patch for POSSUM will be made available. As side effect of 
-not using POSSUM's original activation scale, the BOLD \% difference levels 
-were in the range of 50 - 100\% from the base, about 5 times as large as they 
-should have been. I therefore manually added 750 to the total level of
-simulated POSSUM images, putting results into the 1 - 5\% range. This 
-had no effect on the quality of the simulation and should not effect the 
-underlying parameters (other than $V_0$). 
-Also worth noting is that the noise level was set to an SNR of 20, due to changes 
+point. The patch for POSSUM will be made available. It is worth noting is that the noise 
+level was set to an SNR of 20, but due to changes 
 in the program the true signal-to-noise ratio was much lower, as seen in the 
-fit in \autoref{fig:simslicefit1}.
+in \autoref{fig:SimSNRhm}. The mean SNR for each region was calculated only for the
+voxels with a Signal-To-Noise ratio above $0.1$.
 
-%todo where is this from?
 \begin{figure}
 \centering
-\includegraphics[clip=true,trim=5cm 2cm 4.5cm 2cm,width=15cm]{images/simslice_noise}
-\caption{A single time-series fit in the simulated slice. Preprocessed BOLD is
-the noisy response generated by POSSUM.}
-\label{fig:simslicefit1}
+\subfigure{
+\includegraphics[width=10cm]{images/snr_hm.png}}
+\subfigure{\includegraphics[width=1cm]{images/scale5.png}}
+\caption{SNR Map of POSSUM simulated data. Region 1 mean SNR was $0.8$, Region 2 mean SNR was $0.97$, and 
+Region 3 mean SNR was $0.39$.}
+\label{fig:SimSNRhm}
 \end{figure}
@@
 \subfigure[Grey matter map for simulate region.]{\includegraphics[height=9.5cm]{images/sim_gm}}
-\subfigure[Residual Map. Scale is Normalized $\sqrt{MSE}$.]
+\subfigure[Residual Map. Scale is Normalized $\sqrt{MSE}$. Lower (yellow) is better.]
 {\includegraphics[height=9cm]{images/sim_hm}}
-\subfigure[Mutual Information Map. Scale is bits.]
+\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
 {\includegraphics[height=9cm]{images/sim_hm_mi}}
@@
 \end{table}
-%last (bottom right), 1.203, .310, .326, .036, 2.168, 2.272, .086, 3.4, 1
 
@@
 such a low $\epsilon$, region 4 was below the noise threshold. Notice that the
-regions 1 2 an 3 stick out in both the residual and the mutual information
-map, indicating that the
-particle filter was successful in matching those regions.
+regions 1, 2 an 3 stick out in both the residual and the mutual information
+map, indicating that the particle filter was successful in matching those regions. 
+Mutual information was an extremely succesfull metric, with the exception of 
+a few false positives. Thus, another heatmap (\autoref{fig:sim_hm2}) is 
+helpful to show what threshold would remove those false positives.
+
+\begin{figure}
+\centering
+\subfigure{
+\includegraphics[width=10cm]{images/sim_hm_15_6.png}}
+\subfigure{\includegraphics[width=1cm]{images/scale6.png}}
+\caption{More stringent mutual information heatmap. Higher (yellow) is better.}
+\label{fig:SimSNRhm}
+\end{figure}
 

### 2010-08-15 doc/thesis/resultsB.tex
added some more images and fixes to results

@@
 
-%todo check scanner information/stimulus
+%scanner info?
 \section{FMRI Configuration}
 For the FMRI data discussed in \autoref{sec:RealData}, tests were 
-performed on a right handed volunteer using a Gradient EPI sequence
-on a 1.5T GE scanner with a 2.1s TR. Slices were 5mm thick and echo
-time was (TODO). The subject was then presented with either a single or
-double flash, at randomly dispersed time points over the course of
-5 minutes, 10 seconds. The FMRI began 18.9 seconds
+performed on a right handed volunteer using a GE SIGNA HDx 1.5 Tesla
+scanner with a single echo EPI sequence. Slice spacing was
+5mm, and pixel sizes were 3.75mm.  Repitition Time was 2.1s, 
+Echo Time was 40 ms and the imaging frequency was set to $63.854$MHz.
+The image resolution was $64 \times 64 \times 28$. 
+The subject was presented with either a single or
+double flash and was asked to respond with a right handed
+or lefted handed finger tap, respectively. The FMRI began 18.9 seconds
 before the beginning of the experiment, to allow for transients 
-in to image to settle out. The timing of the flashes are shown in 
-\autoref{tab:flashes}, with time 0 corresponding to the beginning of
+in to image to settle out. The timing of the flashes are shown 
+below with time 0 corresponding to the beginning of
 the 10th TR (so 9 images were dropped).
 
-%todo the stimulus
-\begin{table}[t]
-\centering
-\begin{tabular}{|c || c |}
-\hline 
-Single Flash Times & Double Flash Times \\
-\hline 
-.7 & 2\\
-2.7 & 21\\
-5.7 & 28\\
-12.7 & 52\\
-\hline 
-\end{tabular}
-\caption{Timing of single and double flashes.}
-\label{tab:flashes} 
-\end{table}
+The timing of the single flashes were:
+
+1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350, 
+61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063, 
+90.169, 96.994, 110.644, 117.469, 120.881, 130.988, 132.694, 154.875, 
+158.288, 161.700, 165.113, 168.525, 176.925, 178.631, 183.750, 190.575, 
+204.225, 205.931, 211.050, 216.038, 222.863, 226.275, 236.513, 248.456, 
+255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394, 
+299.381
+
+and the timing of the double flashes were:
+
+0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756, 
+91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000, 
+134.400, 136.106, 141.225, 144.638, 156.581, 159.994, 166.819, 171.806, 
+175.219, 185.456, 188.869, 202.519, 212.756, 217.744, 221.156, 227.981, 
+229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506, 
+272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500
 
@@
 SPM analysis). For the SPM analysis, the Canonical HRF was used
-and model derivatives were included. All settings used in SPM8 were the defaults
+and model time derivatives were included although not in the contrast
+vector. All other settings used in SPM8 were left at default
 for analyzing FMRI data.
@@
 
+%37-14-7
 \begin{figure}
@@
 \subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
-\caption{Section 1, Estimated vs. Actual BOLD response}
+\caption{Section 1, Estimated vs. Actual BOLD response. T-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 
+%34-12-7 
 \begin{figure}
@@
 \subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
-\caption{Section 2, Estimated vs. Actual BOLD response}
+\caption{Section 2, Estimated vs. Actual BOLD response. T-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 
-The fourth voxel (\autoref{fig:comp4}selected for anyalysis had a relatively high residual, and was not picked
-for activation by SPM. On the other hand, the mutual information was above the threshold
-for the image (todo). It is simple to see why the residual was not acceptable in this case.
+The fourth voxel (\autoref{fig:comp4}) selected for anyalysis had a relatively high residual, and was not picked
+for activation by SPM. The mutual information was only $0.06$
+for the image. It is simple to see why the residual was not acceptable in this case.
 At the same time, the peaks do seem to correlate with the measurement peaks. Regardless,
@@
 
+%23-21-7
 \begin{figure}
@@
 \subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
-\caption{Section 3, Estimated vs. Actual BOLD response}
+\caption{Section 3, Estimated vs. Actual BOLD response. T-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 
+%33-30-4
 \begin{figure}[H]
@@
 \subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
-\caption{Section 4, Estimated vs. Actual BOLD response}
+\caption{Section 4, Estimated vs. Actual BOLD response. T-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_29_9_13}}
-\caption{Section 5, Estimated vs. Actual BOLD Response}
+\caption{Section 5, Estimated vs. Actual BOLD Response. T-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
 %\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, T-Value
@@
 \subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
-\caption{Section 6, Estimated vs. Actual BOLD Response}
+\caption{Section 6, Estimated vs. Actual BOLD Response. T-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 %\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{Section 6, Estimated vs. Actual BOLD Response}
+\caption{Section 6, Estimated vs. Actual BOLD Response. T-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }

### 2010-08-15 doc/thesis/methods.tex
mostly finished with simulation results
more work

@@
 y_{\text{gain}, 0:K} = 1.4826\text{median}_{i=0:K}(y_i - \text{median}(y_{0:K}))
+\label{eq:mad}
 \end{equation}
@@
 
-\section{Prior Distribution}
+\subsection{Prior Distribution}
 \label{sec:PriorDist}
@@
 reaching $5\%$ in extreme activation. Yet using the mean priors
-from \cite{Friston2000}, the signal response for a $.1$ second
-impulse only reaches maybe half a percent, as \autoref{fig:MeanResponseF}
+from \cite{Friston2000}, the signal response for a $0.1$ second
+impulse only reaches half a percent, as \autoref{fig:MeanResponseF}
 shows.
@@
 \centering
-\label{fig:MeanResponseF}
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response}
 \caption{Response to $0.1s$ impulses with the mean parameters from \cite{Friston2000}}
+\label{fig:MeanResponseF}
 \end{figure}
@@
 \centering
-\label{fig:MeanResponseJ}
 \includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response_johnston}
 \caption{Response to $0.1s$ impulses with the mean parameters from \cite{Johnston2008}}
+\label{fig:MeanResponseJ}
 \end{figure}
@@
 
-\section{Resampling}
+\subsection{Resampling}
 \label{sec:Resampling}
@@
 
-\section{Choosing $P(y_k | x_k)$}
+\subsection{Choosing $P(y_k | x_k)$}
 \label{sec:Methods Weighting Function}

### 2010-08-15 doc/thesis/results.tex
more work

@@
 that was used in real data (\autoref{sec:RealData}) was used in the simulations. This sequence may
-be found in \autoref{tab:flashes}, but because of the availability of ground
+be found in \autoref{sec:ExperimentConfig}, but because of the availability of ground
 truth in simulations is not displayed in this section.

### 2010-08-15 doc/thesis/resultsB.tex
more work

@@
 %scanner info?
-\section{FMRI Configuration}
+\section{Experiment Configuration}
+\label{sec:ExperimentConfig}
 For the FMRI data discussed in \autoref{sec:RealData}, tests were 

### 2010-08-15 doc/thesis/methods.tex
removed the todo items

@@
 
-Typically a step size of .001 was used, after finding that even .01 can
-at times lead to the state equations careening out of control.
+For all the analyses  in this work, $1400$ integration points
+per second were used.  Typically a step size of $0.001$ was sufficient,
+however, from time to time $0.001$ can still be too high for the
+BOLD model.
 
@@
 
-TODO give scanner details
+The resting state was gathered in the exact same way as the data in 
+\autoref{sec:ExperimentConfig}, except without the stimuli.
 
@@
 particles, 1500 integration points and a Quad Core CPU). The size of 
-local linearization steps are also crucial although going above .001 seconds
+local linearization steps are also crucial although going above $0.001$ seconds
 per step is not recommended. In most cases millisecond resolution

### 2010-08-15 doc/thesis/results.tex
removed the todo items
removed a question mark
spell check and added some more stats

@@
 Repeating tests with different noise realizations demonstrates the 
-particle filter's resiliance to noise and explores the variance of
+particle filter's resilience to noise and explores the variance of
 the model.
@@
 effective at removing trends, although the spline did struggle to match the signal 
-torward the end. This slight drift effect was caused by the final median being 
+toward the end. This slight drift effect was caused by the final median being 
 well above the base
@@
 map, indicating that the particle filter was successful in matching those regions. 
-Mutual information was an extremely succesfull metric, with the exception of 
+Mutual information was an extremely successful metric, with the exception of 
 a few false positives. Thus, another heatmap (\autoref{fig:sim_hm2}) is 
@@
 \caption{More stringent mutual information heatmap. Higher (yellow) is better.}
-\label{fig:SimSNRhm}
+\label{fig:sim_hm2}
 \end{figure}
@@
 
-%todo quantify the number of false positives, add extra Heatmap removing false positives
+There were 1479 voxels analyzed, and there 11 were false positives, according
+the mutual information threshold of $0.1$. Raising the threshold to $0.15$ reduced
+the number of false positives to $0$. Quantifying false negatives is harder, because
+of the presence of white matter.
+
 The histograms again demonstrate that a single point estimate of the parameters 

### 2010-08-15 doc/thesis/background.tex
more work
removed the todo items
lots of cleanup

@@
 
-\subsection{Classical Activation Detection}
-\label{sec:Square Wave}
-The most basic method of analyzing FMRI data is by standard T-test
-between resting state and active state samples. Simply put, the
-mean is calculated separately for non-stimulus and stimulus time intervals.
-A classic t-test may then be applied, giving the probability that the
-distributions actually have the same mean. Because of the correlated
-noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
-to apply some sort of high-pass filter to the data. Without applying
-such a filter, P values must be set extraordinarily high to prevent
-false positives \cite{Smith2007}. If there truly is signal
-due to stimuli, the distributions will not actually be independent
-Gaussians because activation does not fit a square wave 
-(\autoref{sec:BOLD Physiology}). For this reason other methods
-are often more used, as discussed in 
-\autoref{sec:Current Techniques General Linear Model}.
-
 \subsection{Random Field Theory}
 \label{sec:RFT}
-SPM methods make significant use of T-Tests across large regions;
-however, such T-tests work slightly differently than a single
-individual test. A t-test with a p-value of $.05$ over 
+SPM methods make significant use of $t$-Tests across large regions;
+however, such $t$-tests work slightly differently than a single
+individual test. A $t$-test with a p-value of $0.05$ over 
 10,000 voxels, will on average generate 500 false
@@
 deals with this by requiring each test to pass with P
-value of $\frac{.05}{10000}$. The probability of 
-a single false positive would then $.05$. Unfortunately this leads to unrealistically
+value of $\frac{0.05}{10000}$. The probability of 
+a single false positive would then $0.05$. Unfortunately this leads to unrealistically
 low p-values; so low that it would be impossible for any biological system to satisfy. To
@@
 can theoretically be applied with a lower scaling factor than
-the original voxel count \cite{Worsley2004}. As side effect of this,
-a single voxel activation is virtually impossible to detect.
+the original voxel count \cite{Worsley2004}. As side effect of this 
+extreme smoothing, a single voxel activation is very hard to detect.
+
+\subsection{Classical Activation Detection}
+\label{sec:Square Wave}
+The most basic method of analyzing FMRI data is by standard $t$-test
+between resting state and active state samples. Simply put, the
+mean is calculated separately for non-stimulus and stimulus time intervals.
+A classic $t$-test may then be applied, giving the probability that the
+distributions actually have the same mean. Because of the correlated
+noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
+to apply some sort of high-pass filter to the data. Without applying
+such a filter, P values must be set extraordinarily high to prevent
+false positives \cite{Smith2007}. Because the BOLD response is known
+not to fit a square wave this method is rarely used
+(\autoref{sec:BOLD Physiology}). For this reason other methods
+are often more used, as discussed in 
+\autoref{sec:Current Techniques General Linear Model}.
 
@@
 \label{sec:Current Techniques General Linear Model}
-The most common FMRI analysis technique is SPM, though there
+The most common FMRI analysis technique is called SPM, though there
 are more advanced versions than the simple square wave
@@
 \includegraphics[scale=.7]{images/HRF}
-\caption{Canonical Hemodynamic Response Function (y-axis unts are arbitrary because 
+\caption{Canonical Hemodynamic Response Function (y-axis units are arbitrary because 
 normalization is performed).}
@@
 has been optimized for the visual cortex. As discussed in \autoref{sec:Post Stimulus Undershoot},
-there are certainly variations in the shape of the
-BOLD response, over both brain regions and patients. Handwerker et. al. \cite{Handwerker2004}
-discuessed the implications of choosing an incorrect HRF, the most important
-of which is a definite increase in false negatives. While an atlas
+there are definitely variations in the shape of the
+BOLD response, over both brain regions and patients. Handwerker et. al. 
+discussed the implications of choosing an incorrect HRF, the most important
+of which is a definite increase in false negatives \cite{Handwerker2004}. While an atlas
 of Hemodynamic Response Functions for each region could definitely 
 mitigate this risk, it does not account for variation between patients.
-Thus, the inability to fit parameters other than scale hinders 
-analysis of activation.
 
@@
 As mentioned previously and discussed extensively
-in \cite{Friston2002} and \cite{Hofmann1997}, hierarchical models may be 
-applied to account for systematic differences between subjects. For instance, if the
+in Friston et al. and Hoffman et al., hierarchical models may be 
+applied to account for systematic differences between subjects
+\cite{Friston2002, Hofmann1997}. For instance, if the
 study mixes young and old, incorporating
@@
 be otherwise explainable.  The Hierarchical form used by
-\cite{Friston2002} is shown in \autoref{eq:Hierarchical}.
+Friston et al. is shown in \autoref{eq:Hierarchical} \cite{Friston2002}.
 
@@
 \label{eq:Hierarchical}
-Y(t) = X_1(t)\theta_1 + \epsilon_1(t)            \nonumber \\
-\theta_1(t) = X_2(t)\theta_2 + \epsilon_2(t)     \nonumber \\
-...                                              \nonumber \\
-\theta_{n-1}(t) = X_n(t)\theta_n + \epsilon_n(t) 
+Y(t) & = & X_1(t)\theta_1 + \epsilon_1(t)            \nonumber \\
+\theta_1(t) &=& X_2(t)\theta_2 + \epsilon_2(t)     \nonumber \\
+ & ... &                                             \nonumber \\
+\theta_{n-1}(t)& =& X_n(t)\theta_n + \epsilon_n(t) 
 \end{eqnarray}
 
-The Empirical Bayes algorithm is used in both in both \cite{Friston2002} 
-and \cite{Hofmann1997}. Note that in Empirical Bayes point estimators 
-are used for each $\theta$, rather than full distributions.
+The Empirical Bayes algorithm was used by both Friston et al. and Hofmann et al. 
+\cite{Friston2002, Hofmann1997}. Note that in Empirical Bayes point estimators 
+are used for each $\theta$, rather than full distributions or the first two moments.
 
@@
 highlighted by the common scenario where no significant
-activation can be found in a single run \cite{Riera2004}
-\cite{Johnston2008}.
+activation can be found in a single run \cite{Riera2004, Johnston2008}.
 
@@
 data. Combined FMRI CBF or CBV imaging methods are improving,
-as seen in \cite{Chen2009}. These techniques could shed light on
+as seen in Chen et al. \cite{Chen2009}. These techniques could shed light on
 neural activation by providing extra measurements, yet a 
@@
 section are all attempts to regress some version of the BOLD model. 
-Although \cite{Buxton1998} and \cite{Friston2000}
+Although Buxton et al. and Friston et al. 
 both proposed physiologically reasonable values for the model parameters, 
-\cite{Friston2002b} was the first paper to calculate the parameters based 
-on actual FMRI data. However, in that case, the voxels were chosen from
+Friston et al. in 2002 was the first paper to calculate the parameters based 
+on actual FMRI data \cite{Buxton1998, Friston2000, Friston2002b}. 
+However, in that case, the voxels were chosen from
 regions that were detected as active by the GLM. It is therefore possible
@@
 \label{sec:Background Linear Approximation}
-In \cite{Friston2002b}, a novel combination of linear and nonlinear modeling
-was used to generate parameter estimates. Because there is no closed 
+In Friston et al., a novel combination of linear and nonlinear modeling
+was used to generate parameter estimates \cite{Friston2002b}. 
+Although this method was restricted to regions determined to be 
+active by the GLM, there is no reason it could not be applied more broadly.
+Because there is no closed 
 form solution to the ballon model, it impossible to 
 calculate the Jacobian matrix $\frac{\partial Y}{\partial \theta}$. Thus 
-\cite{Friston2002b} approximates the partial using a Volterra Kernel. 
+Friston et al. approximated the partial using a Volterra Kernel \cite{Friston2002b}. 
 At each step of the Expectation-Maximization (EM) algorithm, 
-the differential equation is integrated and the residuals
-calculated. Then, for each $\theta_i$ surrounding the current estimate of 
-$\theta$, a Volterra-Kernel expansion of the output $y$ is generated. Generation
+the differential equation was integrated and the residuals
+calculated. Then, for each $\theta_i$ surrounding the estimate of 
+$\theta$, a Volterra-Kernel expansion of the output $y$ was generated. Generation
 of the Volterra Kernel is quick, and, since there
-is an analytical solution calculating partial derivatives is easy. 
- The full E-M algorithm
-for estimating the states is notation-heavy and can be found in \cite{Friston2002b}.
+is an analytical solution, calculating partial derivatives is easy. 
+The full E-M algorithm for estimating the states is notation-heavy 
+ and can be found in Friston et al. \cite{Friston2002b}.
 
@@
 Importantly, the Volterra-Expansion of $y$ is not able to model interactions
-between state variables \cite{Friston2002b}; which certainly increases error.
-Additionally, all the tests performed
-in \cite{Friston2002b} were on regions found to be active by the General
-Linear Model. For this reason, the reliability of the approximation is unknown
-for regions that are active but
-sufficiently nonlinear to avoid detection by conventional tests.
+between state variables; which certainly increases error \cite{Friston2002b}.
+Additionally, to date no extensive review of the accuracy of the 
+Volterra expansion has been performed. Finally all the tests performed
+by Friston et al. were on regions found to be active by the General
+Linear Model \cite{Friston2002b}. For this reason, the reliability of 
+the approximation is unknown for regions that are active but sufficiently 
+nonlinear to avoid detection by conventional tests.
 
@@
 than a purely empirical model, there are serious implications. The
-first problem is that all the powerful classical techniques of gradient
+first problem is that all the powerful linear techniques of gradient
 descent are off limits; since the model is a true nonlinear dynamical
@@
 at each iteration it selects a random point in the neighborhood. If the 
-that new point is 
+new point is 
 below some energy constraint (energy is a function of the residual), 
@@
 
-%[simulated annealing image?] todo
-
 Genetic algorithms (GA) are similar to Simulated Annealing, in
 that they randomly move to better solutions based on a cost function.
-However; in genetic algorithms a single point estimate isn't used. Instead
+However, in genetic algorithms, a single point estimate isn't used. Instead
 a population of estimates is generated, each with distinct parameters,
@@
 there is no guarantee that a global minimum has been reached.
-%[genetic algorithm picture] todo
 
@@
 discuss in \autoref{sec:Particle Filter}, the Particle Filter method is able
-to circumvent this re-calculation to some degree.
+to circumvent this re-calculation to some degree. Its worth noting though, that 
+to beat the particle filter algorithm discussed in \autoref{sec:Particle Filter} 
+these would have to converge in less than 1000 simulations (which is far less than
+1000 generations in GA).
 
@@
 %\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption{Example updates of a distribution using Kalman Filter, \cite{Thrun2005}}
+\caption{Example updates of a distribution using the Kalman Filter, \cite{Thrun2005}}
 \label{fig:EKFWorking}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption{Distributions of state variables after simulating for .1s}
+\caption{Distributions of state variables after simulating for $0.1$s}
 \label{fig:transp1s}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
-\caption{Distributions of state variables after simulating for 1s}
+\caption{Distributions of state variables after simulating for $1$s}
 \label{fig:trans1s}
@@
 \subsection{Hybrid Methods}
-In \cite{Riera2003}, a maximum
+In Riera et al. , a maximum
 likelihood method for innovation processes was used, as described by
-\cite{Ozaki1994}. Ozaki et. al. used a similar construction to a 
-Kalman filter to filter the BOLD signal. 
+Ozakai \cite{Riera2003,Ozaki1994}. Ozaki used a similar construction to a 
+Kalman filter to filter nonlinear signals \cite{Ozakai1994}. 
 The method used in \cite{Riera2003} performs maximum likelihood on
@@
 comparison between the distributions found in \cite{Johnston2008} and 
-\cite{Friston2000} will be discussed in \autoref{sec:PriorDist}, including
-graphs: \autoref{fig:MeanResponseF}, \autoref{fig:MeanResponseJ}.
+\cite{Friston2000} will be discussed in \autoref{sec:PriorDist}.
 

### 2010-08-15 doc/thesis/intro.tex
lots of cleanup

@@
 Imaging (FMRI) perform regression using the linear combination of explanatory variables. 
-Though adding nonlinear degrees of freedom naturally mandates more computation,
-in this thesis I will discuss a Sequential Monte Carlo method of fitting a nonlinear
-model a computation cost that would still allow real time calculations for multiple voxels.
+Though adding degrees of freedom naturally mandates more computation,
+in this thesis I will demonstrate the use of the Particle Filters to 
+estimate the governing parameters of the BOLD model at a computation cost 
+that would still allow real time calculations for multiple voxels.
 More practically, this method is an alternative
@@
 physiological
-constraints. Modeling the BOLD response as a nonlinear system is the
+constraints. Modeling the BOLD (Blood-Oxygen-Level-Dependent) response 
+as a nonlinear system is the
 best way to determine the correlation of stimulus sequence with the BOLD
@@
 computationally taxing. The solution used here takes approximately 40 seconds
-for a single voxel's time series (5 minutes in length, with Core 2 Duo Q6600).
+for a single 5 minute time series (with Core 2 Duo Q6600).
 
@@
 and real FMRI data in \autoref{sec:SimulationResults} and \autoref{sec:RealData},
-respectively. Finally in \autoref{sec:Conclusion} there is a discussion of
-the usefulness and implications of this technique as well as recommendations
-for the direction of future works.
+respectively. In \autoref{sec:Discussion} the results and their implications
+are further discussed. Future improvements to this technique, as well
+as applications are then explored in \autoref{sec:FutureWork}.
+Finally in \autoref{sec:Conclusion} the significant findings and 
+impact of this work are reviewed.
 
@@
 on the original Blood Oxygen Level Dependent (BOLD) signal 
-derivation first described by \cite{Ogawa}.
-The seminal work by \cite{Buxton1998} attempted to explain the
+derivation first described by Ogawa et al. \cite{Ogawa}.
+The seminal work by Buxton et al. attempted to explain the
 time evolution of the BOLD signal using a windkessel model to
-describe the local changes in Deoxygenated Hemoglobin content.
-Incremental improvements were made to this model until
-\cite{Friston2000} brought all the changes together into a single complete 
-set of equations. And while there have been numerous adaptations in the model, 
-many of them summarized in \cite{Deneux2006}, even the basic versions
+describe the local changes in Deoxygenated Hemoglobin content \cite{Buxton1998}.
+Incremental improvements were made to this model until Friston et al.
+brought all the changes together into a single complete 
+set of equations \cite{Friston2000}. And while there have been numerous adaptations in the model, 
+many of them summarized by Deneux et al., even the basic versions
 have less bias error than the empirically driven \emph{Canonical Hemodynamic Model}
@@
 nuclear spins return their original state at different
-rates, called the T1 relaxation time, depending on the atoms excited.
+rates, called the T1 relaxation time, depending on the material excited.
 Additionally, the
@@
 T1 relaxation times are typically on the order of seconds if 
-a sufficiently strong excitation was applied, whereas T2 relaxation
+a sufficiently strong excitation is applied, whereas T2$^*$ relaxation
 times are usually less than 100ms. 
@@
 Because the entire k-space (spatial-frequency) volume is acquired 
-from a single excitation, the signal-to-noise-ratio is low
+from a single excitation, the signal-to-noise-ratio (SNR) is very low
 in EPI.
@@
 It is well known that the two types of hemoglobin act as a contrast agents in 
-EPI imaging
-\cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
+EPI imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
 between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
@@
 model \cite{Mandeville1999}. These works derive, from first principals,
-the changes in deoxyhemoglobin ratio and volume of capillaries given flow waveform.
+the changes in deoxyhemoglobin ratio and volume of capillaries given an inflow waveform.
 These were the first two attempts to quantitatively account for the shape of the 
@@
 
-Although \cite{Buxton1998} demonstrated that a well chosen flow waveform could 
+Although Buxton et al. demonstrated that a well chosen flow waveform could 
 explain most features of the BOLD signal, it stopped short of proposing a
-realistic waveform for the CBF and CMRO2. In \cite{Friston2000} Friston et. al.
+realistic waveform for the CBF and CMRO2 \cite{Buxton1998}. Friston et al. 
 gave a reasonable and simple
-expression for CBF input based on a flow inducing signal.
-Additionally, in the same work Friston et. al. proposed a simple method
-of estimating metabolic rate: as a direct function of the inward blood flow.
-By combining these equations with the balloon model from \cite{Buxton1998},
-it is possible to predict the BOLD signal directly from a stimulus time course.
+expression for CBF input based on a flow inducing signal
+and in the same work proposed a simple method
+of estimating metabolic rate: as a direct function of the inward blood flow \cite{Friston2000}.
+By combining these equations with the balloon model from Buxton et al.,
+it is possible to predict the BOLD signal directly from a stimulus time course:
 \begin{eqnarray}
@@
 parameters controlling blood flow are $\epsilon$, which is a neuronal 
-efficiency term, $u(t)$, which is the stimulus, and $\tau_f$, $\tau_s$ 
+efficiency term, $u(t)$ which is the stimulus, and $\tau_f$, $\tau_s$ 
 which are time constants. The parameters for the evolution of blood 
@@
 
-This completed balloon model was well summarized again
-in \cite{Riera2003}. Ogawa et. al. \cite{Obata2004} refined the readout equation 
+This completed balloon model was assembled and analyzed
+by Riera et al. \cite{Riera2003}. Obata refined the readout equation 
 of the BOLD signal based on the
 deoxyhemoglobin content (q) and local blood volume (v), resulting in the
-final BOLD equation:
+final BOLD equation \cite{Obata2004}.
 \begin{eqnarray}
@@
 patient, and brain 
-region ($E_0$, $r_0$), often the estimated values taken from \cite{Obata2004} are 
+region ($E_0$, $r_0$), often the estimated values by Obata are 
 taken as the constants $a_1 = k_1 + k_2 = 3.4$, and $a_2 = k_2+k_3 = 1$ in 
-studies using 1.5 Tesla scanners.
+studies using 1.5 Tesla scanners \cite{Obata2004}.
 While this model is more accurate than the static Hemodynamic Model used in SPM,
@@
 Because \autoref{eq:bold} is not capable of producing such a prolonged undershoot,
-additional factors must be at play. Two prominent theories exist to explain the post 
+additional factors must be present. Two prominent theories exist to explain the post 
 stimulus undershoot.  Recall
@@
 CBV truly have returned to their baseline. Other studies indicate
-that there can be a prolonged supernormal CBV \cite{Mandeville1999a,
-Behzadi2005, Chen2009a}, although none of these papers completely
-rule out the possibility of increased CMRO2. The discrepancies may in part
+that there can be a prolonged supernormal CBV, although none of these papers completely
+rule out the possibility of increased CMRO2  \cite{Mandeville1999a,
+Behzadi2005, Chen2009a}. The discrepancies may in part
 be explained by a spatial dependence in the post-stimulus undershoot; described
-by \cite{Yacoub2006}. In \cite{Chen2009}
+by Yacoub et al. \cite{Yacoub2006}. In Chen et al. 
 a compelling case is made that most of the post stimulus undershoot can be 
 explained by combination of a prolonged CBV increase, and a prolonged CBF 
-undershoot, and that
+undershoot \cite{Chen2009}. Additionally it was proposed that
 the previous measurements showing a quick recovery of CBV 
@@
 research into the post-stimulus undershoot has led to the creation
-of much more in depth models. In \cite{Zheng2002} additional state
-variables model oxygen transport, whereas \cite{Buxton2004} models
+of much more in depth models. In Zhen et al. additional state
+variables model oxygen transport, whereas Buxton et al. models
 CMR02 from a higher level, and somewhat more simply; though it 
-still adds 9 new parameters. \cite{Behzadi2005}
-introduces nonlinearities into the CBF equations as a method to
+still adds 9 new parameters \cite{Zheng2002, Buxton2004}. Behzadi et al. 
+introduced nonlinearities into the CBF equations as a method to
 explain the post-stimulus undershoot, which falls in line with a 
-prolonged increase in CBF observed in \cite{Chen2009}. Similarly
-\cite{Zheng2005} adds additional compartments to model 
-venous and arterial blood. 
-In \cite{Deneux2006} Deneux et. al. compared these models and though 
+prolonged increase in CBF observed by Chen et al. \cite{Behzadi2005, Chen2009}. 
+Similarly Zheng et al. added additional compartments for 
+venous and arterial blood \cite{Zheng2005}. 
+Deneux et. al. compared these models and though 
 that work did not deal extensively with the 
 post-stimulus undershoot, it did show incremental improvements
-in quality from the additional parameters. 
-Importantly, \cite{Deneux2006} did show that by 
-simply adding viscoelastic terms from \cite{Buxton2004}, a slowed return 
+in quality from the additional parameters \cite{Deneux2006}.  
+Importantly, Deneux et al. did show that by 
+simply adding viscoelastic terms from Buxton et al. a slowed return 
 to baseline is possible to model, without greatly increasing
-complexity. Regardless, because these models are more 
+complexity \cite{Deneux2006, Buxton2004}. Regardless, because these models are more 
 complex, and the parameters are not well characterized, in this work the simple
@@
 \label{sec:BOLD Analysis}
-Since the first complete BOLD model was proposed by \cite{Friston2002}, 
-several studies have analyzed its properties. 
+Since the first complete BOLD model was proposed by Friston et al.,  
+several studies have analyzed its properties \cite{Friston2002}.  
 The most important property is that the system is dissipative, and given
@@
 responding disproportionately strongly \cite{Birn2001, Wager2005, Deneux2006}.
-This nonlinearity is accounted for in the Balloon model, although \cite{Deneux2006}
-shows that when duration of stimuli varies greatly,
-modeling Neural Habituation is necessary to fully capture the range of responses. 
-In both \cite{Birn2001} and \cite{Deneux2006} it was found that 
+This nonlinearity is accounted for in the Balloon model, although Deneux et al.
+showed that when duration of stimuli varies greatly,
+modeling Neural Habituation is necessary to fully capture the range of responses \cite{Deneux2006}. 
+Both Birn et al. and Deneux et al. found that 
 stimuli lasting longer than 4 seconds 
 tend to be more linear, which is why block designs are so well accounted for
-by the General Linear Model (see \autoref{sec:Current Techniques General Linear Model}).
+by the General Linear Model (see \autoref{sec:Current Techniques General Linear Model})
+\cite{Birn2001, Deneux2006}.
 
-Another interesting result of \cite{Deneux2006} was the sensitivity analysis.
+Another interesting result from Deneux et al. was the sensitivity analysis \cite{Deneux2006}.
 There it was found that the parameters are far from perpendicular,
@@
 \hline 
-Parameter  & \cite{Friston2000} & \cite{Johnston2008} & \cite{Vakorin2007} & \cite{Deneux2006}\\
+Parameter  & Friston et al. \cite{Friston2000} & Jonston et al. \cite{Johnston2008} 
+    & Vakorin et al. \cite{Vakorin2007} & Deneux et al. \cite{Deneux2006}\\
 \hline
@@
 \end{tabular}
-\caption{Parameters found by various studies. (NC) indicates that the value
+\caption{Parameters found in various studies. (NC) indicates that the value
 wasn't calculated. \cite{Vakorin2007} made use of the values from \cite{Friston2002}

### 2010-08-15 doc/thesis/resultsB.tex
removed the todo items
lots of cleanup

@@
 
-%todo check SPM smoothing parameter
 Note that SPM must pre-process the image with a spatial smoothing 
@@
 scanner with a single echo EPI sequence. Slice spacing was
-5mm, and pixel sizes were 3.75mm.  Repitition Time was 2.1s, 
+5mm, and pixel sizes were 3.75mm.  Repetition Time was 2.1s, 
 Echo Time was 40 ms and the imaging frequency was set to $63.854$MHz.
@@
 double flash and was asked to respond with a right handed
-or lefted handed finger tap, respectively. The FMRI began 18.9 seconds
+or left handed finger tap, respectively. The FMRI began 18.9 seconds
 before the beginning of the experiment, to allow for transients 
@@
 \subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
-\caption{Particle Filter results measured in normalized $\sqrt{MSE}$. 
-        Units of match is normalized residual where the  the lowest (best) levels shown are
-        $.7$ and the highest error shown (threshold) is $1$.
+\caption{Particle Filter results measured in normalized error. 
+        Units of match is normalized residual where the lowest (best) levels shown are
+        $0.7$ and the highest error shown (threshold) is $1$.
         Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_pfilter_x},
@@
          Units of match is bits (standard for base-2 Mutual Information). The highest (best) levels are 
-         $.42$ and the worst shown (threshold) is $.1$.
-        Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_mi_x},
+         $0.42$ and the worst shown (threshold) is $0.1$.
+        Sagittal, coronal and axial slices are shown in \autoref{fig:hm_canon_mi_x},
         \autoref{fig:hm_canon_mi_y}, and \autoref{fig:hm_canon_mi_z}, respectively. 
@@
 it doesn't. While a few peaks seem to match, most of the signal does not correlate with the
-expected state progression. This voxel is not being driven direcly by the input, although it
+expected state progression. This voxel is not being driven directly by the input, although it
 may be gated or driven through intermediate region.
@@
 
-The fourth voxel (\autoref{fig:comp4}) selected for anyalysis had a relatively high residual, and was not picked
+The fourth voxel (\autoref{fig:comp4}) selected for analysis had a relatively high residual, and was not picked
 for activation by SPM. The mutual information was only $0.06$
@@
 actual measured signal declines. This is an indication that the BOLD signal is not directly
-being driven by the input. It is likeley that the cause of the activation in this reason is
-not true activation, but recieved from surrounding active regions through spatial smoothing.
+being driven by the input. It is likely that the cause of the activation in this reason is
+not true activation, but received from surrounding active regions through spatial smoothing.
 
@@
 Many of the differences between SPM and the particle filter seem to be driven
-more by the pre-processing methods than the regression methd. The reason for 
-SPM applying such extrensive smoothing is to combat false positives. While this is
+more by the pre-processing methods than the regression method. The reason for 
+SPM applying such extensive smoothing is to combat false positives. While this is
 extremely important, regions such as \autoref{fig:comp6} show the problem with doing 
 so. In all, the particle filter method does very good job of estimating parameters that
-fit the measurements; and at the very least this method is a workable, albiet
+fit the measurements; and at the very least this method is a workable, albeit
 computationally intensive, alternative to SPM.

### 2010-08-15 doc/thesis/thesis.tex
more work on the results
lots of cleanup

@@
 \usepackage{cancel}
-\usepackage[colorlinks=true,linkcolor=black]{hyperref}
+\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
 \def\chapterautorefname{Chapter}
@@
 
-August 10, 2010\\
+August 30, 2010\\
 Blacksburg, Virginia
@@
 
-Keywords: BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics
+Keywords: BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
 \\
@@
 
-{\large BOLD Parameter Estimation Using Particle 
-Filters }
+{\large Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
 
@@
 a full posterior probability, the reliance on the mean to estimate 
-parameters is not strictly necessary. This work presents
+parameters is unnecessary. This work presents
 not just a viable alternative to the traditional method of detecting
-activation, but an extensible technique of estimating the joint BOLD
-parameter distribution.
+activation, but an extensible technique of estimating the joint Probability
+Distribution Function of the BOLD parameters.
 
@@
 % \chapter*{Dedication}
-% \chapter*{Acknowledgments}
+\chapter*{Acknowledgments}
+
+This thesis would not have been possible without the guidance, advice and
+patience of my advisor, Chris Wyatt. I knew nothing about being a researcher
+or about medical imaging when I came to him, and he taught me to read broadly
+and to always dig deeper. Thank you Dr. Wyatt.
+
+I would also like to thank William Baumann for his help and inspiration. I
+have learned an incredible amount reading papers with you and Dr. Wyatt.
+
+To my parents, for all the psychological (and financial) help along the way,
+thank you so much. When things were the darkest you were there for me, I couldn't
+ask for better parents. Love you guys.
+
+Finally I would like to thank my friends, for the support, for the prayer,
+and, yes, for the occasional stress-relieving game of Halo, SOASE, and every
+type of SC.  You guys are the best.
 
@@
 \bibliographystyle{plain}
+%\bibliographystyle{abbrvnat}
 %\bibliographystyle{apalike}
@@
 \end{document}
-
-%trash
-%stream of new developments in modeling the BOLD signal, nothing has have been
-%a plethora of new studies pushing advancements 
-%the choice of analysis tools is still relatively limited. Every widely
-%available analysis tool is based on parametric modeling, which requires
-%prior knowledge of the noise distribution. Indeed, the linear methods
-%usually applied are known not to be robust to non-gaussian, non-white
-%noise, which is why so many pre-processing steps are necessary.
-%These limitations cast a long shadow over any experiment that claims to be
-%able to reject the null hypothesis, no matter the P-value. While obviously
-%there is a correlation between the BOLD signal and actual activity, its
-%possible the current parametric tests are over-estimating the actual amount
-%of activity. Its also possible that the current methods underestimate
-%actual activity. In this paper we propose a robust model-based approach to 
-%activation detection that makes no assumptions about the underlying noise
-%model.
-%
 

### 2010-08-16 doc/thesis/background.tex
SAVE

@@
 Because there is no closed 
-form solution to the ballon model, it impossible to 
+form solution to the balloon model, it impossible to 
 calculate the Jacobian matrix $\frac{\partial Y}{\partial \theta}$. Thus 
@@
 Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
-Filter by allowing the state update to be a function, $g$,
+Filter by allowing the state update and output functions to be arbitrary functions.
 
@@
 X(t) &=& g(u(t), X(t-1))\\
-Y(t) &=& h(X(t)
+Y(t) &=& h(X(t))
 \end{eqnarray}
@@
 sort of underlying distributions may exist in such a mixed biological,
-mechanical, chemical system such as the brain. Assuming the parameters
-all to be Gaussian could be a gross error. On the other hand, for
+mechanical, chemical system such as the brain.
+ The assumption of Gaussianity is what allows
+the UKF to estimate the posterior using only the first and second moments;
+however, if this assumption is violated significant error will result.
+ On the other hand, for
 small variances and short time steps the Gaussian distribution is a good 
 fit, and so in some cases the Unscented Kalman Filter could work quite
-well. The assumption of Gaussianity is what allows
-the UKF to estimate the posterior using only the first and second moments;
-however, if this assumption is violated significant error will result.
-
+well. 
 \begin{table}[t]
@@
 In order to drive the system without input I intentionally set $s_t$ a non-steady
-state, but physiologically plausible, value. The value of $u$ is left at zero 
+state, but physiologically plausible, value. The value of $u$ was left at zero 
 the entire time, so the 
 system would decay naturally (see \autoref{sec:BOLD Physiology})
-\autoref{fig:transp1s} shows the results when the system after the
+\autoref{fig:transp1s} shows the results after the
 system ran for 100 milliseconds. 
@@
 Ozakai \cite{Riera2003,Ozaki1994}. Ozaki used a similar construction to a 
-Kalman filter to filter nonlinear signals \cite{Ozakai1994}. 
-The method used in \cite{Riera2003} performs maximum likelihood on
-the innovations rather than the absolute signal levels. 
+Kalman filter on nonlinear signals \cite{Ozaki1994}. 
+The method used by Riera et al. used this method to perform maximum likelihood on
+the innovations rather than the absolute signal levels \cite{Riera2003}. 
 By using a Local Linearization Filter,
@@
 this method was not very effective for the BOLD signal, when I used it 
-with the particle filter approach discussed in the next chapter.
+with the particle filter approach discussed in the next chapter; although its
+effectiveness depended greatly on the stimulus.
 
-In \cite{Johnston2008}, a hybrid particle filter/gradient
+In Johnston et al., a hybrid particle filter/gradient
 descent algorithm was used to simultaneously derive the static and dynamic 
-parameters, (classically known as parameters and state variables, respectively).
-A particle filter is used to calculate the state variables at each
-time; then the estimated distribution of the particles is used to find
+parameters, (classically known as parameters and state variables, respectively)
+ \cite{Johnston2008}.
+A particle filter was used to calculate the state variables at each
+time; then the estimated distribution of the particles was used to find
 the most likely set of parameters that would give that distribution of state variables.
-This process is repeated until the parameters converge. Interestingly
-\cite{Johnston2008} comes to a very different set of parameter estimates as compared
-to the original \cite{Friston2000} estimates (\autoref{tab:Params}). In fact the results 
+This process was repeated until the parameters converged. Interestingly Johnston et al.
+came to a very different set of parameter estimates as compared
+to the original Friston et al. estimates (\autoref{tab:Params})
+\cite{Johnston2008, Friston2000}. In fact the results 
 are significantly different from virtually every other study. The most obvious discrepancy
@@
 Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Nonlinear Least Squares},
-the algorithm described in \cite{Johnston2008} does not depend on prior distributions.
+the algorithm described in Johnston et al. does not depend on prior distributions \cite{Johnston2008}.
 It is possible then that the bias toward the prior in other methods skewed their results. 
-While \cite{Johnston2008} is certainly in the minority; further exhaustive studies 
-of the parameters, using unbiased techniques may be called for. A further 
-comparison between the distributions found in \cite{Johnston2008} and 
-\cite{Friston2000} will be discussed in \autoref{sec:PriorDist}.
+While Johnston et al. is certainly in the minority; further exhaustive studies 
+of the parameters, using unbiased techniques may be called for \cite{Johnston2008}. A further 
+comparison between the distributions found in Johnston et al. and Friston et al. 
+will be discussed in \autoref{sec:PriorDist} \cite{Johnston2008, Friston2000}.
+
+In Vakorin et al., a combination of Genetic Algorithms and Simulated Annealing were
+used to estimate not only the parameters, but the true stimuli driving the BOLD
+signal. This addresses the inherent uncertainty of exactly where and when 
+stimuli actually get applied. Unfortunately this algorithm was extremely slow.
 
 \subsection{Previous Particle Filter Approaches}
-In \cite{Murray2008} a particle filter based approach was used to integrate
-the BOLD equations. The method used in that work focused primarily on estimating
+In a PhD thesis, Murray used a particle filter based approach was used to integrate
+the BOLD equations \cite{Murray2008}. The method used in that work focused primarily on estimating
 the BOLD output and state equations as a nonlinear stochastic differential 
-equation. The primary difference between \cite{Murray2008} and this work is that
-\cite{Murray2008} took the parameters as a given. Thus, differences in the BOLD output
+equation. The primary difference between that work and this is that
+Murray took the parameters as a given \cite{Murray2008}. Thus, differences in the BOLD output
 were taken to be primarily driven by stochastic changes in the underlying state
@@
 differences in BOLD response cannot be explained solely by stochastic elements 
-is important, however. The particle filter library created
-in that work, $dysii$, forms the basis for the particle filter used in this work, 
-and was extremely well designed. The work also clearly presents the particle filter;
+is important, however. The filtering framework created
+in that work, dysii, forms the basis for the particle filter used in this work, 
+and was very designed. The work also clearly presents the particle filter;
 both its derivation and use.
@@
 equations. Exhaustive search type methods such as those employed by
-\cite{Johnston2008} and \cite{Vakorin2007} have long
-run times even for a single voxel. While Volterra models are an interesting
+Johnston et al. and Vakorin et al. have long
+run times even for a single voxel \cite{Johnston2008, Vakorin2007}. 
+While Volterra models are an interesting
 solution,there have not yet been exhaustive tests to determine whether
@@
 able to maintain a fast runtime while still approaching the solution.
-While the reliance on a Gaussian estimate to the true posterior distribution could
-cause problems, some modifications could make it powerful. 
+The reliance on a Gaussian estimate to the true posterior distribution could
+cause problems however. 
 The particle filter method proposed in the next section bears a strong 
-resemblance to the unscented Kalman filter; albeit with more points estimating
-the posterior.
+resemblance to the unscented Kalman filter; albeit without the Gaussian 
+assumptions.

### 2010-08-16 doc/thesis/discussion.tex
lots of cleanup
SAVE

--- 2010-08-11 doc/thesis/discussion.tex
@@
 \chapter{Discussion}
+\label{sec:Discussion}
+\section{Review of Results}
 % Overview of results
@@
 %% Output estimates good
-The results unequivocally show that the parameters of the BOLD model are under
-constrained. While unsurprising given the sensitivity analysis in \cite{Deneux2006},
-it is still an important limitation when calculating parameters. Thus, other methods
+There are two possible, but unlikely causes for the large variance in
+parameter estimates. First, it could be that the particle filter is
+incapable of learning the model. However, given the high quality
+of BOLD estimates, this is almost certainly not the case. The other 
+possible explanation is that the given stimulus is not sufficiently
+varied to bring out all the properties of the system. This also is unlikely
+because of results in \autoref{sec:VeryLongSim} which used a random impulse
+stimulus sequence. The high correlation of the parameters in \autoref{sec:VeryLongSim},
+point to a significant trade off in parameters.
+
+Thus the results show that the parameters of the BOLD model are under
+constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
+it is a notable conclusion \cite{Deneux2006}. Thus, other methods
 that depend on a point estimate of the parameters, such as least squares 
-or kalman filters (which uses the first two moments) are limited. While estimates of
+or Kalman filters (which uses the first two moments) are limited. While estimates of
 the  BOLD signal may still be correct, the estimate
-of the underlying parameters and state variables are suspect. Similarly, as the
-histograms in \autoref{sec:Multi-voxel Simulation} and \autoref{sec:Real Data Parameter Estimates}
-show, using the mean to estimate parameters also makes little sense in the particle filter
-case. Attributing the variance of estimate to the underlying parameter distribution would
-also be a mistake; given that variance persisted even in simulated data. While parameters 
-definitely vary from person to person and cortex to cortex, the estimated distributions 
-in \autoref{sec:Real Data Parameter Estimates} are wider than the true underlying 
-distribution. For this reason and because of the nonlinear relation between
-parameters, any analysis should take the entire posterior estimate into account, 
-rather than relying on an idealized Gaussian. In this sense, particle filters represent
-an important step forward in BOLD parameter estimation. Representing
-the uncertainty in parameters as a Gaussian is insufficient; so using a particle filter
-or Bayesian estimate of the posterior is not simply an enhancement, but a necessary precaution.
+of the underlying parameters and state variables cannot be described without using
+a joint probability distribution function. In this sense, particle 
+filters represent an important step forward in BOLD parameter 
+estimation. Representing uncertainty with a mean
+and variance is insufficient; so using a particle filter
+or Bayesian estimate of the posterior is not simply an enhancement, 
+but a necessary precaution.
 
-Although point estimates of parameters are not dependable estimates of the 
-true parameters, they still useful. Analysis of differences in parameters
-could be medically relevant for instance. Additionally, a single estimate for parameters is still able
-to form an accurate estimate of the BOLD signal. This trait bridges the gap with earlier types 
-of activation tests, such as the statistical
-parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
+Because of the ill-defined nature of the parameters, the results 
+of Friston et al. may have been wider than true parameter distribution
+\cite{Friston2002}. Any algorithm that attempts to calculate parameters
+of the BOLD model should be run multiple times, to check for consistency.
+
+As the results in \autoref{sec:RealData} show, the heatmaps, especially
 those of mutual information, closely resembled the results of SPM. While the activation
-tests were more sensitive, they were also more prone to false positive. However, it is
-worth noting that certain enhancements could be useful in reducing false positives; for instance
-by using the maximum likelihood, or median of the final distribution.  Future work may 
-shed further light on these techniques.
+tests were more sensitive, there were some additional false positives, 
+though at this problem is difficult to quantify. Regions with M.I. above 
+$0.15$ consistently fit the FMRI data well, and simulations showed that
+the particle filter performed extremely well in the face of significant
+noise. In sum, the estimated BOLD output remained consistent despite
+large swings in parameters.
 
@@
 particular problem there can be no guarantee of a global minimum, and although
-other techniques guarantee a local minimum, the particle filter doesn't settle 
-to one of these local minima because it can settle into multiple ones. This
-difference typifies the major difference between the particle filter and 
-competing approaches.
+other techniques guarantee a local minimum, the particle filter in some
+sense can settle into multiple local minima.
 
 One difficulty
-with the use of a particle filter with a finite number of datapoint is the calculation of a good
-weight function $P(y_k | x_k)$  that will converge reasonably quickly. If the weight
-function does not sufficiently differentiate particles, the final distribution will 
-no be significantly different from the prior distribution. On the other hand, if the 
-weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
+with the use of a particle filter when given a finite number of measurements is finding
+a good weight function $P(y_k | x_k)$. This is more important for a finite
+number of measurements because $P(y_k | x_k)$ needs to converge in finite time.
+In spite of this potential problem, \autoref{sec:VeryLongSim} managed
+to converge in less than 500 seconds.  Given sufficient
 measurements it is better to let the algorithm take longer to converge, because the
-convergence will be better (more accurate). The particle filter takes longer
+convergence will be less prone to particle deprivation. The particle filter takes longer
 to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
@@
 
-The particle filter certainly has significant advantages over other estimation procedures
+The particle filter also has advantages over other estimation procedures
 discussed in \autoref{sec:Prior Works}. The most important advantage is that it provides
-an estimate of the posterior probability, rather than a single estimate. While it is natural
-to want a simple estimate of parameters, such an estimate is impossible with this particular
-model. The results are more difficult to interpret, but this is a necessity. The fact that 
-the final distribution is not dependent on any particular distribution is also advantageous.
-Of particular note; the final distribution does not need to conform to any parametric
-distribution. While the particle filter was not fast for full brain calculations, its speed
+an estimate of the posterior probability, rather than a single estimate. While researchers
+often want a simple estimate of parameters, such an estimate is impossible with this particular
+model. The fact that the final distribution does not need to conform to a
+parametric distribution is also advantageous, given the nonlinearities in the system.
+While the particle filter took a day to calculate for full brain calculations, its speed
 was sufficient on a quad core machine to perform real time calculations of small regions
@@
 well and does not require burdensome amounts of memory (approximately 11 megabytes).
-For this reason this algorithm is perfect for extension to vector or video card
-based processors.
 
@@
 how the particle filter works. This is in contrast to the Volterra based approach
-of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
-are needed for the particle filter. It is a common problem in traditional statistics 
-for assumptions to be unrealistic which means the results may be invalid or at 
-least in question. 
-Fewer and more realistic assumptions mean that the particle filter is more robust
-to unforeseen difficulties in FMRI data.
+of, which can be quite complex. Additionally very few assumptions
+are needed for the particle filter; which reduces the risk of those assumptions
+being violated. Fewer and more realistic assumptions make the particle 
+filter more robust to unforeseen difficulties in FMRI data.
 
 \chapter{Future Work}
+\label{sec:FutureWork}
 % Enhancements/Dehancements 
@@
 advanced version.
+
+Collapse parameters
 

### 2010-08-16 doc/thesis/methods.tex
SAVE

@@
 Although the particle filter  is a standard Regularized
-Particle filter, as described in \cite{Arulampalam2002a}, optimizing the 
-particle filter for use with FMRI data is non-trivial.
+Particle filter, as described by Arulampalam et al., optimizing the 
+particle filter for use with FMRI data is non-trivial \cite{Arulampalam2002a}.
 
@@
 Because the purpose of this particle filter is to learn the underlying distribution
-of the static parameters, rather than precisely model the time course of the 
-in the dynamic parameters ($\{s,f,v,q\}$) this noise is left out. It also helps
+of the static parameters, rather than precisely modeling the time course of the 
+in the dynamic state variables $\{s,f,v,q\}$ this noise is left out. It also helps
 that detrending is applied before the particle filter and that the
@@
 per second were used.  Typically a step size of $0.001$ was sufficient,
-however, from time to time $0.001$ can still be too high for the
+however, from time to time $0.001$ can still result in problems for the
 BOLD model.
@@
 task is motion correction. To do this, a single volume in time is chosen, and
-volumes at every other time are registered to this one volume. This corrects
+volumes at every other time are realigned to best match the target volume. This corrects
 for motion by the patient as well as small changes in the magnetic
@@
 A particularly difficult form of noise present in FMRI is a low frequency
-drift, often characterized as a Wiener process (\cite{Riera2004}). 
+drift, often characterized as a Wiener process \cite{Riera2004}. 
 Though not present in all regions, as many as ten to fifteen percent
-of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
+of voxels can be affected \cite{Tanabe2002}, thus it is prevalent enough to cause significant
 inference problems \cite{Smith2007}. It is still not
 clear what exactly causes this noise, although one possibility is 
-the temperature difference in scanner magnetic coils\cite{Smith2007}. 
+the temperature difference in scanner magnetic coils \cite{Smith2007}. 
 It is clear that this drift signal is not solely
@@
 \subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
-\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
+\caption{Q-Q Plots of resting state data, using the BOLD signal steps.}
 \label{fig:QQDelta}
@@
 Because most methods (including the one used in this paper)
-assume the noise realizations are independent of each other, the auto-
-correlation is of particular interest (which is a necessary but not
+assume the noise realizations are independent of each other, the autocorrelation 
+is of particular interest (which is a necessary but not
 sufficient condition for independence). Gaussianity is also a common
@@
 whether the noise is in fact Wiener, the distribution of 
-the steps were plotted against a Gaussian.
+the steps were also plotted against a Gaussian.
 
 Finally, removal of the drift is often performed with a high pass filter,
-so analyzed the distribution after subtracting of a spline, (see \autoref{sec:Methods Preprocessing}).
+so I also analyzed the distribution after subtracting a spline, (see \autoref{sec:Methods Preprocessing}).
 
@@
 Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
-x-axis and the points
-on the y-axis come from the same type of distribution, then all the points will
+x and y axes come from the same type of distribution, then all the points will
 be collinear. Differences in the variance will cause the line to have a slope
@@
 to be shifted. In these Q-Q plots, the points are being compared to the standard
-Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
+Gaussian distribution. Note that the points have all been 
 normalized (changed to percent difference).
 
-Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
-are well described by a Gaussian process with a small autocorrelation, 
+\autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
+are well described by a Gaussian process with a small autocorrelation, but 
 \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
@@
 indicates that the steps could be Independently Distributed. 
-Therefore, the noise does seem to come close to a Wiener process.
+Therefore, the noise is consistent with a Wiener process.
 
@@
 $\nu_x$ is difficult to compensate for, and so many methods
-have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
+have been developed to compensate for it. Tanabe et al. and Smith et al. 
 demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
-of FMRI. It has been reported that in some studies as many as half the voxels 
-benefited from detrending (\cite{Smith2007}). In a head to head comparison, 
-\cite{Tanabe2002}, showed that in most cases subtracting off
-a spline worked the best. The benefit of the spline versus wavelets, high pass 
+of FMRI \cite{Tanabe2002, Smith1999}. It has been reported that in some studies as 
+many as half the voxels 
+benefited from detrending \cite{Smith2007}. In a head to head comparison, 
+Tanabe et al.  showed that in most cases subtracting off
+a spline worked the best \cite{Tanabe2002}. 
+The benefit of the spline versus wavelets, high pass 
 filtering or other DC removal techniques is that the frequency response is not set.
@@
 measurements in an image. Thus a 10 minute session at a repetition time of 
-2.1 seconds would have 19 knots. The knot first and last knots were each 
+2.1 seconds would have 19 knots. The first and last knots were each 
 given half the number of samples as the rest of the knots; which were all 
@@
 There is potential to optimize the spline further using a canonical 
-HRF to find resting points; however, for this to work the experiment would have
+HRF to find resting points; however, the experiment would have
 to be designed with this in mind.
@@
 Because most of the BOLD signal is above baseline, after removing the spline
-the BOLD resting state will be below 0\%.  This reduces the ability of an algorithm to learn.
+the resting state will be below 0\%.  
 One method of accounting for this is to simply add a DC gain model parameter.
-Like all the other model parameters, with enough measurements, a viable
-parameter would fall. Yet adding another dimension increases the
+Like all the other model parameters, with enough measurements, the particle filter
+would be able to settle on a good estimate. Yet adding another dimension increases the
 complexity of the model, when the parameter is relatively easy to estimate
@@
 \begin{equation}
-y_{\text{gain}, 0:K} = 1.4826\text{median}_{i=0:K}(y_i - \text{median}(y_{0:K}))
+y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
 \label{eq:mad}
@@
 
-A serious concern when adding and subtracting arbitrary values to 
+A serious concern when adding constant values to 
 real data is whether this will create false positives. This is a legitimate
 concern; however, a boosted response does not effect how well the BOLD model 
-predicts the actual measurements.
+predicts the actual measurements; and as mentioned before, the DC signal of FMRI
+is never used.
 
@@
 \section{Particle Filter Settings}
-There are quite a few options when using a particle filter; those
-options will be discussed in this section.
+There are quite a few options when using a particle filter; therefore this
+section will discuss the design choices made in this work.
 
@@
 used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
-\cite{Friston2000} only studied regions deemed active by the General 
+Friston et al. only studied regions deemed active by the General 
 Linear Model; and most other research (including \cite{Friston2001}) used these results as 
-the source for their priors. 
-The one exception is \cite{Johnston2008}, which came to a extremely different
-distributions. For a particle filter, the choice of a prior is
-the single most important design choice. A very wide prior will require
+the source for their priors \cite{Friston2000}. 
+The one exception is to this is Johnston et al. which came to a extremely different
+distributions \cite{Johnston2008}. For a particle filter, the choice of a prior is
+the single most important design decision. A very wide prior will require
 more particles to be sufficiently dense, a very thin (low variance) prior may miss
 the true parameters. Consequently, for this work it was natural
-to use priors that will give results consistent with previous works, 
+to use priors that would give results consistent with previous works
 \cite{Friston2000}. This constrains the usefulness of the model to
-areas that fall within the prior distribution, yet will allow results
+areas that fall within the prior distribution, yet allowed results
 to be comparable to other works. There is a significant need for better
@@
 exhaustive simulated annealing or hill climbing tests for multiple
-regions and multiple patients.
-
-There is an interesting anomaly with the priors found in virtually all
-the works that characterized the parameters, except \cite{Johnston2008}.
-The BOLD signal is universally recognized to be around $2-3\%$, maybe
-reaching $5\%$ in extreme activation. Yet using the mean priors
-from \cite{Friston2000}, the signal response for a $0.1$ second
-impulse only reaches half a percent, as \autoref{fig:MeanResponseF}
-shows.
+regions and multiple patients. That said, a significant amount of
+indeterminability exists in the parameters, which is discussed in 
+\autoref{sec:Single Voxel Simulation}
 
@@
 \centering
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response}
+\includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response}
 \caption{Response to $0.1s$ impulses with the mean parameters from \cite{Friston2000}}
@@
 
-While this could be the result of a stimulus
-being too short to lead to strong activation, a similar stimulus
-scheme in real data showed a much larger response than 
-half a percent as well. In fact, after applying de-trending,
-converting the image to percent-difference, and removing 
-outliers ($ BOLD > 10\% \text{ or } BOLD < -10\%$) the total variance
-across all \emph{active} voxels was still around .02, indicating
-that in active voxels a signal peaking below .005 seems unlikely. 
-Of course, if more restriction were placed on the outliers, its possible
-this standard deviations could be brought down. 
-The parameter estimates by \cite{Johnston2008} are even more 
-confusing, with peaks of well below $.1\%$ (\autoref{fig:MeanResponseJ}).
-
 \begin{figure}
 \centering
-\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/mean_response_johnston}
+\includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response_johnston}
 \caption{Response to $0.1s$ impulses with the mean parameters from \cite{Johnston2008}}
@@
 
-Its likely that these differences are due to some difference in preprocessing,
-although in \cite{Deneux2006} the signals were found to be peaking around
-$1\%$, unlike \cite{Friston2000} which shows signals peaking at up to
-$3\%$ or $4\%$. In my own tests, it seemed necessary for $\epsilon$ to
-reach well over $1.5$ and $V_0$ to reach more than $.4$ to reach these
-peaks; of course other methods may be equally able. 
-Therefore, to account for these discrepancies, somewhat broader
-distributions are used than the numbers used in \cite{Friston2000}
-(which are widely used, \cite{Hu2009}). The 
-priors used in the particle filter may be found in \autoref{tab:Prior}.
+The differences between the parameter estimates of Johnston et al. and 
+Friston et al. are clearly visible in \autoref{fig:MeanResponseJ} and 
+\autoref{fig:MeanResponseF} \cite{Johnston2008,Friston2002}. 
+Therefore, to account for these discrepancies, I used tests on
+real data to adjust the distributions from those arrived at in 
+Friston et al. \cite{Friston2002}. In tests on real data, the Johnston et al.
+distributions never converged to meaningful BOLD estimates.
+The priors used in the particle filter may be found in \autoref{tab:Prior}.
 
@@
 Note that although the mean remains the same for all the 
-parameters other than $\epsilon$, the standard deviation is set
-much higher to account for the disagreement between studies
-(\autoref{tab:Params}). 
-Because all the parameters are taken to be strictly positive, and the
-standard deviations are approaching the mean, I used a gamma distribution.
-This prevents the Gaussian from placing parameters in the nonsensical 
-territory of negative activation, or negative time constants.
+parameters other than $\epsilon$, the standard deviation were
+adjusted based on tests with real data. Also, the state variables
+were all assumed to be at rest at the beginning of the simulation.
 
-Another aspect of the prior is using enough particles to get a 
+It is also import to use enough particles to obtain a 
 sufficiently dense approximation of the prior. For 7 dimensions,
@@
 the number of particles was dropped to 1,000. Typically during the 
-first few measurements the variance dropped precipitously because
-most particles were far from a solution.  The particles that are left are in a
+first few measurements the variance drops precipitously because
+most particles are far from a solution.  The particles that are left are in a
 much more compact location, allowing them to be estimated with 
@@
 doesn't impair convergence. Therefore, for this work I chose a Gaussian kernel of
-bandwidth equal to the original distribution's covariance. Obviously this will
-apply a rather large amount of smoothing to the distribution; however, on average
-resampling is only applied every 20 to 30 measurements, and because randomization
+bandwidth equal to the original distribution's covariance. This 
+applies a large amount of smoothing to the distribution; however, on average
+resampling was only applied every 20 to 30 measurements, and because randomization
 is being applied to model updates this gives the filter some mobility.
 
-Re-sampling is a not strictly necessary, but increases the effectiveness
+Resampling is a not strictly necessary, but it increases the effectiveness
 of the particle filter by adjusting the support to emphasize areas
-of higher probability. Re-sampling is slow because it requires re-drawing
+of higher probability. Resampling is slow because it requires redrawing
 all the particles. It also closes off avenues of investigation, and is
@@
 two consecutive low ($<50$) $N_{eff}$'s were found.
+
+\subsection{Particle Deprivation}
+\label{sec:Particle Deprivation}
+To prevent particle deprivation, which may happen even with a good
+estimate of the prior distribution, a method of rescuing the particle
+filter from this state was used. When particle deprivation was detected,
+resampling was performed with a saved version of the covariance matrix (from
+the previous time step). This allowed for the particles to be re-scattered
+without having to go all the way back to the prior distribution. Particle 
+deprivation was defined by $N_{eff}$ dropping to $1$.
 
@@
 distribution will lead to particle deprivation (all particles
-being zero-weighted).  I tested several weighting
-functions: in addition to the Gaussian I also tested the Laplace and Cauchy
-distributions, both of which have much wider tales than the Gaussian. 
+being zero-weighted). Three weighting functions were 
+tested. In addition to the Gaussian I also tested the Laplace and Cauchy
+distributions, both of which have wide tales. 
 Wider tailed distributions don't down-weight
-particles as fast; and converge more slowly (and perhaps more accurately). 
+particles as fast; and thus converge more slowly. 
 The Laplace distribution also has the
-benefit of a non-zero slope at the origin; which means that even
-it will distinguish between particles even near the origin.
+benefit of a non-zero slope at the origin; meaning it 
+differentiates between particles even near the origin.
 
-After trial and error, for this work I chose a zero-mean Gaussian with standard deviation 
-of $.005$. While I made some attempts to automatically set the standard 
-deviation, results were often unpredictable. If the weight function and scale aren't
-fixed across voxels, very noisy time series with no actual signal 
-converged to nonsensical results. 
-In the future, it may be possible to set the standard deviation by
+After trial and error, for this work the Gaussian with a standard deviation 
+of $0.005$ was chosen. Although using an adaptive weighting function
+might be better, in tests this often led to unpredictable results. 
+With some work it may be possible to set the standard deviation by
 taking a small sample from resting data and using the sample standard deviation.
-Since this is the first attempt at using particle filters for modeling the 
-BOLD model, in this work I set the standard deviation manually at $.005$,
+In this work, however, I set the standard deviation manually at $0.005$,
 because it gave the best consistency.
@@
 overall length of the signal being analyzed. For 1000 measurements it takes
-about 6 minutes. On the other hand, in real circumstances the
-length is only around 150 measurements and takes around 40 seconds (for 1000 
-particles, 1500 integration points and a Quad Core CPU). The size of 
-local linearization steps are also crucial although going above $0.001$ seconds
-per step is not recommended. In most cases millisecond resolution
+approximately 6 minutes. On the other hand, in real circumstances the
+length tends to be around 150 measurements which takes around 40 seconds (for 1000 
+particles, 1400 integration points and a Quad Core CPU). The size of 
+integration steps are also crucial, but using step sizes above 
+$0.001$ seconds is not recommended. In most cases millisecond resolution
 is fine; however, when generating simulated data I found that at times it was
-still not enough every once. This is problematic in the actual particle
+still not enough. This is problematic in the actual particle
 filter since, given the large number of simultaneous integrations taking 
 place, its probable that a few particles will fail and be unfairly thrown away.
-To prevent such events, 1500 integration points per second were used throughout
+To prevent such events, 1400 integration points per second were used throughout
 the tests.
 
-Another crucial factor for run time is how long before the first re-sampling 
-occurs. Because the prior is represented initially with significantly more
+Because the prior is initially represented with significantly more
 particles, if for some reason the effective
@@
 algorithm forces resampling. The choice of 20 seconds is 
-arbitrary, but at the very least it gives a more optimized version of the
-prior.
+arbitrary, but at the very least it the particles some time to be reweighted.

### 2010-08-16 doc/thesis/particlefilters.tex
SAVE

@@
 preferred when the model is nonlinear, and non-gaussian. This section
-is largely based on \cite{Arulampalam2002a} and \cite{Thrun2005}.
+is largely based on Arulampalam et al. and Thrun et al. \cite{Arulampalam2002a, Thrun2005}.
 
@@
 than parameter sets that give poor estimates. Although the reliance on
-a prior distribution is often troublesome, when the system being modeled
+a prior distribution could be problematic, when the system being modeled
 has physical meaning, establishing reasonable ranges for parameters may be 
@@
 Finally, because the only difference between the members of $X(t)$ and 
-$\Theta$ is an update function, from now on $X$ will contain 
+$\Theta$ is an update function, from now on $X$ will encapsulate
 $\Theta$. The assumptions now allow for a simplified version of the
@@
 wide enough to incorporate any posterior that arises. If the prior is
-not sufficiently dense, the particle filter can compensate, if it is
+not sufficiently dense, the particle filter may be able to compensate, if it is
 not sufficiently wide the particle filter won't converge.
@@
 
-\subsection{Basic Algorithm}
 From the definition of $w_i$, the algorithm to calculate
@@
 \label{sec:Particle Filter Resampling}
+
+\begin{figure}
+\includegraphics[width=16cm]{images/particle_filter}
+\caption{Particle Filter progression, note that the initial support is flat; the particles
+are equally spaced between -10 and 10}
+\end{figure}
 As a consequence 
 of the wide prior distribution (required for a proper discretization of a continuous
-distribution), in short order there will be a significant proportion of particles 
-with insignificant weights. 
+distribution), a significant proportion of particles will quickly
+develop insignificant weights. 
 While this does help
@@
 tails. In this case particle locations would match the true posterior and all weights would
-be equal.  The case where a large number of the weights have become irrelevantly small
-is called particle degeneracy. In  \cite{Liu1998b}
+be equal.  The case where a large number of the weights have become extremely small
+is called particle degeneracy. In  Lui et al. 
 an ideal calculation of the effective number of particles is found based on the 
-particles' true weight. However, given that only an approximation for the true weight 
+particles' true weight \cite{Liu1998b}. However, given that only an approximation 
+for the true weight 
 exists, they also provide a simple heuristic calculation of $N_{eff}$.
@@
 \end{equation}
-Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
-$N_{eff}$ drops precipitously.  To alleviate this problem
-a common technique known as resampling may be applied. The idea of resampling is to 
+Unless the prior is particularly accurate,
+$N_{eff}$ drops very quickly.  To alleviate particle degeneracy
+a common technique known as resampling is often applied. The idea of resampling is to 
 draw from the approximate posterior, thus generating a replica of the posterior with 
@@
 $\sigma \sigma^T = cov(x)$
-and $\epsilon$ is drawn from the chosen kernel. \cite{Musso2001a} goes into
-significant 
-depth and proves the optimality of the Epanechnikov Kernel for reducing 
-MSE between the original and resampled distributions. However, \cite{Musso2001a}
-also espouses the usefulness of the Gaussian Kernel, due to the ease
-drawing samples from it, which for this work was more important.
+and $\epsilon$ is drawn from the chosen kernel. The choice of a kernel is
+complex, although it has been proven that the optimal kernel for reducing MSE between
+the original and resampled distributions is 
+is the Epanechnikov Kernel \cite{Musso2001a}. However, Musso et al. 
+also espoused the usefulness of the Gaussian Kernel, due to the ease
+drawing samples from it, which for this work was more important \cite{Musso2001a}.
 
@@
 
-It has been proposed by \cite{Hurzeler1998} that if the underlying 
-distribution is non-Gaussian, then using the original bandwidth will over-smooth. 
-In reality, over smoothing
-will only become an issue if re-sampling is performed very often. Thus
-if resampling is performed at every step then this could certainly cause problems.
-If the distribution is over-smoothed then the algorithm may not converge as rapidly;
-however, because the bandwidth is still based on particle variance, which decays as 
-particles are ruled out, it is still able to converge. In fact, over-smoothing is preferable
+Hurzeler et al. demonstrated that if the underlying 
+distribution is non-Gaussian, then using the original bandwidth may over-smooth
+\cite{Hurzeler1998}.  In reality, over smoothing
+will only become an issue if resampling is performed often. Thus
+if resampling is performed at every step then this would certainly cause problems.
+If the distribution is over-smoothed then the algorithm may not converge as rapidly,
+or at all. However, because the bandwidth is still based on particle variance, 
+which should decay as particles are ruled out, the particle filter is still able to converge. 
+In fact, over-smoothing is preferable
 to under smoothing, since over-smoothing simply slows convergence while 
@@
 
-Nevertheless, because 
+Because 
 of the potentially wide smoothing factor applied by regularized resampling, performing this
-step at every measurement would allow particles a great deal of mobility. This mobility could
-hinder convergence, which is why regularized resampling should only be done when
-$N_{eff}$ drops off (less than 50). Other than the periodic regularized
+step at every measurement would allow particles a great deal of mobility. Therefore
+in this work regularized resampling was only performed when $N_{eff}$ dropped
+below 50. Other than the periodic regularized
 resampling, the regularized particle filter is identical to the basic sampling
@@
 With regularized resampling, it is possible to prevent both
-particle degeneracy as well as particle impoverishment. An additional
-bonus is that as the particle filter converges, the density of particles
+particle degeneracy as well as particle impoverishment. Additionally,
+as the particle filter converges, the density of particles
 in the area of the solution goes up. This has a similar effect to 
 simulated annealing where, as the algorithm approaches the end, the
-random steps get smaller and smaller. However, there is risk in 
-resampling. If for some reason the solution is not covered by the 
-new support the algorithm may not be able to reach the true value. 
+random steps get smaller and smaller. Note however, that resampling
+carries certain risks.  If for some reason the solution is not covered by the 
+new support, the algorithm may not be able to reach the true value. 
 The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
@@
 match the measurement error in the output. 
-Typically it is assumed that this error is additive, centered at zero and 
-has a scale comparable to the signal levels.
 While a Gaussian function is the traditional choice, there are other reasonable
-distributions, given the unpredictable nature of the noise present in FMRI.
+distributions, given the nature of the noise present in FMRI.
 The choice of this function will be discussed further in \autoref{sec:Methods Weighting Function}.
 
-\begin{figure}
-\includegraphics[width=16cm]{images/particle_filter}
-\caption{Particle Filter progression, note that the initial support is flat; the particles
-are equally spaced between -10 and 10}
-\end{figure}
-
-\section{Simple, Nonlinear Example}
-A typical half wave rectifier takes a AC voltage circuit and removes
-one half (say the negative half) of the signal. The resulting waveform
-is still not DC, however it is then possible to use a capacitor to 
-smooth the signal into something similar to DC, as shown in \autoref{fig:HalfWaveIO}.
-There are other, more
-complex circuits that convert the negative portion into positive and
-waste less energy but here I will keep the system simple.
-Thus, let us consider a simple half wave rectifier circuit, shown in 
-\autoref{fig:HalfWaveRectifier}.
-
-The half wave rectifier circuit smoothes the gaps between high voltage
-with a capacitor. Thus, when $u(t)G$ is less than $v_t$, the circuit will 
-discharge the capacitor and maintain a non-zero voltage,
-but when $u(t)G$ is greater than $v_t$, the output voltage will be set
-by $u(t)G$ and the capacitor will charge up. I will assume a simple
-model for all the components, ignoring the complex nonlinear behavior
-that can occur in a diode.
-
-\begin{figure}
-\centering
-\begin{circuitikz}[scale=2, american]
-\draw
- (0,0)  node[transformer core] (T) {}
- (T.A1) -- (-1,0)
- (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
- (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
- (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
- (T.base) node {G}
- (T.B1) to[open, *-*, v=$V_1$] (T.B2); 
-\end{circuitikz}
-\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
-gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
-$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
-\label{fig:HalfWaveRectifier}
-\end{figure}
-
-\begin{figure}
-\centering
-\caption{Example Input/Output of the Half Wave Rectifier}
-\label{fig:HalfWaveIO}
-\end{figure}
-
-For this example I will assume that the transformer simply scales
-the input by a constant factor, $G$. As discussed in the \autoref{sec:Particle Filter Model}
-any variable with uncertainty must be part of the state variable. Therefore
-the state variable is defined as: $X(t) = \{G, v_t, C, R_m, v_y\}$. 
-The state equations would then be
-
-\begin{equation}
-v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
-        u(t)G & \text{ if }  u(t)G-v_y \ge v_t\\
-        v_y(t-1)\left(1 - \frac{\delta t}{R_mC}\right) & \text{ if }  u(t)G-v_y < v_t
-    \end{cases} 
-\end{equation}
-
-To run the particle filter is easy, since there exists 
-a recursive definition of the dynamic state variable, $v_y$. To start
-with, an initial distribution must be assumed and while at first a 
-Gaussian seems like a good idea, all the static state variables are strictly
-positive and thus not well suited to the Gaussian. Thus, instead the prior 
-will start with a Gamma distribution. The gamma is defined as follows:
-
-\begin{equation}
-X \sim Gamma(k, \theta) \rightarrow f(x) = x^{k-1}\frac{e^{-x/\theta}}{\theta^k\Gamma(k)}
-\end{equation}
-
-where $\Gamma$ is the gamma function.
-The margin for error is decided by the weighting function, which
-will be defined as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
-estimate based on all the particles, and 
-$v_{yi}$ is the estimate by a particular ($i^{th}$) particle. The choice of this function is difficult,
-and although the Gaussian is typically used, in practice I found the exponential helpful
-in preventing particle deprivation. The algorithm is then,
-
-\begin{algorithmic}
-\STATE Initialize $N_p$ Particles:
-\FOR{$i$ in $N_p$}
-    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
-    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
-    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
-    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
-    \STATE $v_y = 0$, (Assume the system has been off for a long time)
-    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
-    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
-\ENDFOR
-\STATE Run the Filter:
-\FOR{$t$ in Set of Measurement Times}
-    \FOR{$i$ in $N_p$}
-        \STATE $v_{yi}(t) = f(v_{yi}(t-1), u(t))$
-        \STATE (All other members of $X_i(t)$ remain the same)
-        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
-    \ENDFOR
-\ENDFOR
-\end{algorithmic}
-
-Initially the particles will have the same output, $0$, however, as $u(t)$
-changes, the response of each particle to that input will result in different
-outputs. Particles that have a $v_{yi}$ near $V_y$ will be weighted higher,
-and others farther away will be weighted lower. As the particle filter
-runs, weights will compound converging to a distribution that asymptotically
-approaches the true joint distribution of the $X(t)$.  As I
-mentioned in \autoref{sec:Particle Filter Resampling}, particles with 
-weights approaching zero do not significantly
-contribute to the empirical distribution, so re-sampling will be necessary.
+%\section{Piecewise Example}
+%A half wave rectifier takes a AC voltage circuit and removes
+%the negative half  of the signal. The resulting waveform
+%is still not DC, however it is then possible to use a capacitor to 
+%smooth the signal into something similar to DC, as shown in \autoref{fig:HalfWaveIO}.
+%There are other, more
+%complex circuits that convert the negative portion into positive and
+%waste less energy but here I will keep the system simple.
+%Thus, consider a simple half wave rectifier circuit, shown in 
+%\autoref{fig:HalfWaveRectifier}.
+%
+%The half wave rectifier circuit smoothes the gaps between high voltage
+%with a capacitor. Thus, when $u(t)G$ is less than $v_t$, the circuit will 
+%discharge the capacitor and maintain a non-zero voltage,
+%but when $u(t)G$ is greater than $v_t$, the output voltage will be set
+%by $u(t)G$ and the capacitor will charge up. I will assume a simple
+%model for all the components, ignoring the complex nonlinear behavior
+%that can occur in the components. Also note that the 
+%
+%\begin{figure}
+%\centering
+%\begin{circuitikz}[scale=2, american]
+%\draw
+% (0,0)  node[transformer core] (T) {}
+% (T.A1) -- (-1,0)
+% (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
+% (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
+% (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
+% (T.base) node {G}
+% (T.B1) to[open, *-*, v=$V_1$] (T.B2); 
+%\end{circuitikz}
+%\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
+%gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
+%$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
+%\label{fig:HalfWaveRectifier}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\caption{Example Input/Output of the Half Wave Rectifier}
+%\label{fig:HalfWaveIO}
+%\end{figure}
+%
+%Consider the piecewise function in \autoref{eq:example}. 
+%As discussed in the \autoref{sec:Particle Filter Model}
+%any variable with uncertainty must be part of the state variable. Therefore,
+%even though $G$, $C$, $R_m$ and $V_t$ are constants, they are still
+%included in the state variable: $X(t) = \{G, V_t, C, R_m, v_y\}$. 
+%
+%\begin{equation}
+%v_y(t)  = f(v_y(t-1, u(t)) =  \begin{cases} 
+%        u(t)G & \text{ if }  u(t)G-v_y \ge V_t\\
+%        v_y(t-1)\left(1 - \frac{\delta t}{R_mC}\right) & \text{ if }  u(t)G-v_y < V_t
+%    \end{cases} 
+%\end{equation}
+%
+%To run the particle filter is easy, since there exists 
+%a recursive definition of the dynamic state variable, $v_y$.
+%To start with, an initial distribution must be used; which for
+%this case will be:
+%Gaussian seems like a good idea, all the static state variables are strictly
+%positive and thus not well suited to the Gaussian. Thus, instead the prior 
+%will start with a Gamma distribution. The gamma is defined as follows:
+%
+%\begin{equation}
+%X \sim Gamma(k, \theta) \rightarrow f(x) = x^{k-1}\frac{e^{-x/\theta}}{\theta^k\Gamma(k)}
+%\end{equation}
+%
+%where $\Gamma$ is the gamma function.
+%The margin for error is decided by the weighting function, which
+%will be defined as $W(V_y, v_{yi})$, where $V_y$ is the actual measurement, $v_y$ is the 
+%estimate based on all the particles, and 
+%$v_{yi}$ is the estimate by a particular ($i^{th}$) particle. The choice of this function is difficult,
+%and although the Gaussian is typically used, in practice I found the exponential helpful
+%in preventing particle deprivation. The algorithm is then,
+%
+%\begin{algorithmic}
+%\STATE Initialize $N_p$ Particles:
+%\FOR{$i$ in $N_p$}
+%    \STATE $G \sim Gamma(\frac{\mu^2_G}{\sigma^2_G}, \frac{\sigma^2_G}{\mu_G})$
+%    \STATE $v_t \sim Gamma(\frac{\mu^2_{v_t}}{\sigma^2_{v_t}}, \frac{\sigma^2_{v_t}}{\mu_{v_t}})$
+%    \STATE $C \sim Gamma(\frac{\mu^2_C}{\sigma^2_C}, \frac{\sigma^2_C}{\mu_C})$
+%    \STATE $R_m \sim Gamma(\frac{\mu^2_R}{\sigma^2_R}, \frac{\sigma^2_R}{\mu_R})$
+%    \STATE $v_y = 0$, (Assume the system has been off for a long time)
+%    \STATE let $X_i(0) = \{G, v_t, C, R_m, v_y\}$
+%    \STATE let $w_i(0) = 1$ or to make a flat prior, $w_i(0) = \frac{1}{Pr(X_i(0))}$ 
+%\ENDFOR
+%\STATE Run the Filter:
+%\FOR{$t$ in Set of Measurement Times}
+%    \FOR{$i$ in $N_p$}
+%        \STATE $v_{yi}(t) = f(v_{yi}(t-1), u(t))$
+%        \STATE (All other members of $X_i(t)$ remain the same)
+%        \STATE $w_i(t) = w_i(t-1)W(V_y(t), v_y(t))$
+%    \ENDFOR
+%\ENDFOR
+%\end{algorithmic}
+%
+%Initially the particles will have the same output, $0$, however, as $u(t)$
+%changes, the response of each particle to that input will result in different
+%outputs. Particles that have a $v_{yi}$ near $V_y$ will be weighted higher,
+%and others farther away will be weighted lower. As the particle filter
+%runs, weights will compound converging to a distribution that asymptotically
+%approaches the true joint distribution of the $X(t)$.  As I
+%mentioned in \autoref{sec:Particle Filter Resampling}, particles with 
+%weights approaching zero do not significantly
+%contribute to the empirical distribution, so re-sampling will be necessary.

### 2010-08-16 doc/thesis/results.tex
lots of cleanup
SAVE

@@
 Once this noisy simulated time series was generated, the particle filter algorithm
-was run on the single time series image. Five sets of tests 
-determine the power of the particle filter in modeling. 
+was run on the single time series image. Five sets of simulation tests 
+were performed on the particle filter. 
 The first test demonstrates the particle filter on a very long FMRI
@@
 the final covariance matrix of the parameters is in \autoref{tab:long_cov}.
-Note that even though the BOLD response converged well (\autoref{fig:long_converge}),
+Note that even though the BOLD response converged (\autoref{fig:long_converge},
+\autoref{fig:long_converge_500}),
 the parameters still have significant correlation (\autoref{tab:long_corr}). 
@@
 consists entirely of impulses, the best parameters are not one particular
-set, but a spectrum. Note that the correlation is in parameters whose priors
+set, but a joint distribution. Note that the correlation is in parameters whose priors
 are completely independent. It is possible that varying the type of input could
@@
 runs, yet the sum of the individual time constants
-($\tau_f$, $\tau_s$ and $\tau_0$) were more consistent. In general the 
+($\tau_f$, $\tau_s$ and $\tau_0$) were more consistent. On average the 
 time constants fell short of the true time constant. This could be a limitation
-based on the prior distribution (which notably has an initial mean below the true values) 
+based on the prior distribution (which notably had an initial mean below the true values) 
 or it could be that other parameters compensated. It is also possible that the output is insensitive
@@
 to higher peak activation levels, and the subtleties of different time constants 
-are being lost in the noise.
+were lost in the noise.
 
@@
 The error clearly suffers due to this effect (\autoref{tab:HighNoiseResults}).
-Note that neither Run 1 or Run 2 estimated the underlying state well (\autoref{fig:ConvergenceRuns1c}
+Note that neither Run 1 nor Run 2 estimated the underlying state well (\autoref{fig:ConvergenceRuns1c}
 and \autoref{fig:ConvergenceRuns2c}), whereas
@@
 
-The data shows that the parameters did really not converge
-\autoref{fig:JustNoiseConvergence}.
+The data shows that the parameters did not converge
+(\autoref{fig:JustNoiseConvergence}).
 The peaks never even reached 1\% difference
@@
 signal being significantly smaller than any previous simulation.
-From these results it would seem that lack of convergence could be
-a differentiating factor from a signal with an actual signal; however,
-the next section casts doubt on this possibility.
+This run would benefit greatly from an adaptive weight function, since
+such a much tighter weighting function would have forced the particle
+filter to fail.
 
@@
 for rescuing the particle filter from particle deprivation, discussed in
-\autoref{sec:Resampling}, in fact was working against the proper course of action. 
-The proper course of action in this case would be for the particle filter to fail,
-since no particle can really properly estimate a random sequence. When this mechanism
+\autoref{sec:Particle Deprivation} was used when it shouldn't have been.
+ When this mechanism
 was removed, all 11 runs stopped due to particle deprivation (all weights hit zero). 
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
-occur in good data if the prior distribution does not cover a solution.
+occur in otherwise good data if resampling is performed at the wrong moment.
 
@@
 \label{sec:SingleVoxelReview}
-The first step in determining the validity of a model is to provide some order of quality
-to rate the results by. Unfortunately, because of the variability in the signal levels, the
-raw residual cannot perform this task. As demonstrated by \autoref{tab:NoiseOnlyResults},
+Because of the variability in the signal levels, the
+raw residual cannot be used to rate the fit. As demonstrated by \autoref{tab:NoiseOnlyResults},
 a low residual does not necessarily indicate a good fit. 
-Therefore, instead I  normalized the signal based on a value proportional to
-the magnitude of the signal. Considering the tendency of FMRI noise to have large unexplainable
-peaks and troughs, rather than using MAX and MIN values to estimate the scale of the 
-signal, I used a robust estimator of 
-scale; the Median-Absolute-Deviation (MAD) which is defined in equation \autoref{eq:mad}.
+Therefore, a normalized version of the residual was used. To normalize the
+residual, both the estimated and preprocessed BOLD signals were divided by an estimator
+of scale and then the new residual was calculated. 
+Considering the tendency of FMRI to have large unexplainable
+peaks and troughs, the Median-Absolute-Deviation (MAD) was used (\autoref{eq:mad}).
 This is an estimator of the standard
@@
 where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
-the mutual information estimates in this work 6 bins are used for the marginal
+the mutual information estimates in this work 6 bins were used for the marginal
 distribution of each signal. Additionally, throughout log base 2 will be used.
@@
 \end{tabular}
-\caption{Mutual Information and the normalized $\sqrt{MSE}$, 
+\caption{Mutual Information and the normalized error, 
     for signal/noise configurations.}
@@
 distinguishing between these cases with ether normalized residual or mutual information
-is not clear cut. While the average mutual information is in fact more than 10 times
+is not clear cut. While the average mutual information is more than 10 times
 the average mutual information in the two non-signal cases, the maximum
 mutual information of the low noise/no signal case exceeds the minimum M.I. of the high noise/signal
-case. The Low Noise/ signal determination is easier to make; given
+case. The Low Noise/signal determination is easier to make; given
 the minimum mutual information is above $.8$ and the maximum normalized residual is below $.6$.
 However, it is worth noting that the worst case scenario  for mutual information (maximum) in the
-low noise/no signal case does not concur with the worse case (minimum) normalized residual. 
-There is no reason why this has to be the case, but it is beneficial. In
+low noise/no signal case does not coincide with the worse case (minimum) normalized residual. 
+There is no reason why this has to be the case, but it could be beneficial. In
 other words, if it were necessary to make a statement that
 a particular voxel were active or inactive; the accuracy would improve if both techniques
-were used with loose restrictions.
+were used with loose restrictions. In all, both M.I. and the normalized residual 
+provide a good measure of performance.
 
@@
 techniques it is important to ensure consistency of results. Although the parameter
-sets were inconsistent, the quality of the fit was actually quite consistent
-over eleven runs. The second purpose was of course to determine how the particle 
-filter responded to different signal to noise ratios. From \autoref{sec:SimLowNoise},
-the particle filter performed extremely well in filtering out nonsensical measurements.
+sets were inconsistent, the quality of the fit for the BOLD output was often 
+accurate even when noise was drastically increased. The second purpose was to validate
+Mutual Information and the normalized residual as measurements of output quality. 
+Although not perfect, both methods do provide a decent metric.
 
@@
 \label{sec:Multi-voxel Simulation}
+
+\begin{figure}
+\centering
+\subfigure[Region labels for simulated slice.]{
+\includegraphics[height=9.5cm]{images/simregions.png} \label{fig:simslice_mask} }
+\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9.5cm]{images/sim_gm}}
+\subfigure[Residual Map. Scale is Normalized error. Lower (yellow) is better.]
+{\label{fig:simslice_hm_res} \includegraphics[height=9cm]{images/sim_hm}}
+\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
+{\label{fig:sim_hm_mi} \includegraphics[height=9cm]{images/sim_hm_mi}}
+\caption{Comparison of activation with greymatter, parameter regions.}
+\label{fig:simslice_hm}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{
+\includegraphics[width=10cm]{images/snr_hm.png}}
+\subfigure{\includegraphics[width=1cm]{images/scale5.png}}
+\caption{SNR Map of POSSUM simulated data. Region 1 mean SNR was $0.8$, Region 2 mean SNR was $0.97$, and 
+Region 3 mean SNR was $0.39$.}
+\label{fig:SimSNRhm}
+\end{figure}
+
 To test the usefulness of the particle filter on a larger scale, I used a modified version 
@@
 
-\begin{figure}
-\centering
-\subfigure{
-\includegraphics[width=10cm]{images/snr_hm.png}}
-\subfigure{\includegraphics[width=1cm]{images/scale5.png}}
-\caption{SNR Map of POSSUM simulated data. Region 1 mean SNR was $0.8$, Region 2 mean SNR was $0.97$, and 
-Region 3 mean SNR was $0.39$.}
-\label{fig:SimSNRhm}
-\end{figure}
-
 For each time-series in the simulated FMRI image, the final parameters were saved
 into a parameter map. This parameter map could then be compared to the map used to generate the 
-simulated data; additionally a new simulation using the calculated parameters could also be 
+simulated data. Additionally a new simulation using the calculated parameters could also be 
 generated to test the difference in BOLD levels between the real parameters and the
-estimated ones. Since the parameters were far from orthogonal 
-\cite{Deneux2006}, this provided a quantitative difference between the two parameter sets.
-
-\begin{figure}[H]
-\centering
-\subfigure[Region labels for simulated slice.]{
-\includegraphics[height=9.5cm]{images/simregions.png} \label{fig:simslice_mask} }
-\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9.5cm]{images/sim_gm}}
-\subfigure[Residual Map. Scale is Normalized $\sqrt{MSE}$. Lower (yellow) is better.]
-{\includegraphics[height=9cm]{images/sim_hm}}
-\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
-{\includegraphics[height=9cm]{images/sim_hm_mi}}
-\caption{Comparison of activation with greymatter, parameter regions.}
-\label{fig:simslice_hm}
-\end{figure}
+estimated ones. Since the parameters were far from orthogonal, 
+ this provided a quantitative difference between the two parameter sets \cite{Deneux2006}.
 
@@
 
-Note that region 4 has a very low $\epsilon$. For this reason, the only areas
-with significant estimates of the BOLD time series were 1,2 and 3. With
-such a low $\epsilon$, region 4 was below the noise threshold. Notice that the
+Note that region 4 had a very low $\epsilon$, putting it below the noise
+threshold. For this reason, the only areas
+with significant estimates of the BOLD time series were 1,2 and 3. 
+Notice that the
 regions 1, 2 an 3 stick out in both the residual and the mutual information
@@
 Mutual information was an extremely successful metric, with the exception of 
-a few false positives. Thus, another heatmap (\autoref{fig:sim_hm2}) is 
-helpful to show what threshold would remove those false positives.
-
+a few false positives. \autoref{fig:sim_hm2} shows what the heat map looks
+like when the threshold is raised from $0.1$ to $0.15$.
 \begin{figure}
@@
 \caption{Histogram of estimated parameters in section 1 in voxels with mutual information greater
-than $0.14$}
+than $0.15$}
 \label{fig:slicesim_hist1}
@@
 \caption{Histogram of estimated parameters in section 2 in voxels with mutual information greater
-than $0.14$}
+than $0.15$}
 \label{fig:slicesim_hist2}
@@
 \caption{Histogram of estimated parameters in section 3 in voxels with mutual information greater
-than $0.14$}
+than $0.15$}
 \label{fig:slicesim_hist3}
@@
 
-There were 1479 voxels analyzed, and there 11 were false positives, according
-the mutual information threshold of $0.1$. Raising the threshold to $0.15$ reduced
-the number of false positives to $0$. Quantifying false negatives is harder, because
-of the presence of white matter.
+If the thresholds are left at the values of \autoref{fig:simslice_hm}, then
+the normalized residual gave $2/1479$  false positives and mutual information gave
+$11/1479$. However, normalized residual also clearly missed some active voxels, although
+false negatives are harder to quantify because of the presence of white matter. 
+Raising the threshold to $0.15$ eliminated the false positives in the 
+Mutual Information map, although some active voxels were also removed.
 
@@
 is elusive for this set of parameters. However the data clearly show the power
-of the particle filter at identifying regions of activation; which is usually
-performed with statistical parametric maps. The thresholds applied to this
-slice, both for mutual information and residuals is arbitrary, and needs further
-research. Tighter thresholds removes the false positives present
-in the images at the cost of false negatives. As noted in \autoref{sec:SingleVoxelReview}
+of the particle filter at identifying regions of activation.
+ The thresholds applied to this
+slice, both for mutual information and residuals are arbitrary. Applying a
+threshold is helpful for visualization, but is rarely useful for further analysis.
+As noted in \autoref{sec:SingleVoxelReview},
 the false positives present in the Mutual Information
-map are different from those in the residual map. This furthers the argument from
+map are different from those in the residual map. This furthers the argument 
 for combining the two metrics to increase power. Although
-at first glance it would appear that there are false negatives in the 
-\autoref{fig:simslice_hm}; this is
+at first glance it would appear that there are false negatives in 
+\autoref{fig:sim_hm_mi}; this is
 not actually the case. POSSUM simulates different tissues, and white matter
-does not typically  have a BOLD response. This is why there are holes in regions
-2 and 3. These results indicate that the particle filter is effective
+does not typically  have a BOLD response. This explains some of the holes in regions
+2 and 3. These results certainly indicate that the particle filter is effective
 at regressing against a noisy signal.

### 2010-08-16 doc/thesis/resultsB.tex
SAVE

@@
 is the primary drive for the BOLD response. This is in contrast to 
-most voxels which have are controlled by intermediate or complete unrelated factors. 
+most voxels which are controlled by intermediate or completely unrelated factors. 
 Inactive regions cannot be modeled because their input is unknown, therefore
@@
 of SPM is very different from that of the particle filter. SPM8 also outputs
-a T-statistic for each voxel, whereas the 
+a $t$-statistic for each voxel, whereas the 
 particle filter's primary output is a posterior probability distribution of the parameters
@@
 After dropping the first 9 volumes, each of the remaining volumes was
-co-registered with the first. At this point the SPM method
+co-registered with the new first volume. At this point the SPM method
 diverges from the experimental method. For the particle filter, detrending 
@@
 and model time derivatives were included although not in the contrast
-vector. All other settings used in SPM8 were left at default
+vector. All other settings used in SPM8 were left at the default
 for analyzing FMRI data.
@@
 \section{Results}
-The results from T-values from SPM8 are shown in \autoref{fig:hm_canon_spm} (threshold of
+The $t$-values from SPM8 are shown in \autoref{fig:hm_canon_spm} (threshold of
 4), and the results from 
@@
 Note that the scales for all three images are different, because the metrics are different.
-SPM measures using T-Tests to determine the likelihood of a false positive. \autoref{fig:hm_canon_pfilter}
+SPM measures using $t$-tests to determine the likelihood of a false positive. \autoref{fig:hm_canon_pfilter}
 uses simple normalized residuals, meaning that lower indicates less error.  
-\autoref{fig:hm_canon_pfilter_mi} measures in terms of the dependence between the 
+\autoref{fig:hm_canon_pfilter_mi} measures in terms of mutual dependence between the 
 measured signal and the estimated signal; thus higher indicates a better fit. The particle
@@
 SPM operates on smoothed data (8mm x 8mm x 8mm), so there are certainly active 
-areas that have been missed because of the smoothing.
+areas that have been missed because of the smoothing. Note that throughout this 
+section yellow indicates a better fit.
 
 \begin{figure}[H]
+\centering
 \subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
@@
 \subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
-\caption{SPM results. Units of activation are in Student's T-scores; higher indicates higher 
+\caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher 
         assurance that the signal cannot have occurred through noise alone. 
@@
 \begin{figure}[H]
+\centering
 \subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.85]{images/pfilter_hm}}
@@
 %\begin{figure}[H]
+%\centering
 %\subfigure[]{\label{fig:hm_pfilter85} \includegraphics[scale=.75]{images/pfilter85_hm}}
@@
 \begin{figure}[H]
+\centering
 \subfigure[]{\label{fig:hm_mi} \includegraphics[scale=.85]{images/mi_hm}}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_pfilter_37_14_7}}\\
-\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/1_spm_37_14_7}}
-\caption{Section 1, Estimated vs. Actual BOLD response. T-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/1_pfilter_37_14_7}}\\
+\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/1_spm_37_14_7}}
+\caption{Section 1, Estimated vs. Actual BOLD response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_pfilter_34_12_7}}\\
-\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/2_spm_34_12_7}}
-\caption{Section 2, Estimated vs. Actual BOLD response. T-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/2_pfilter_34_12_7}}\\
+\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14.5cm]{images/2_spm_34_12_7}}
+\caption{Section 2, Estimated vs. Actual BOLD response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 
-I chose several voxels to discuss further from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
-The first voxel, labeled 1, had a very high T-score, high mutual information,  as well as 
-a low error (around $.7$). Thus, the
-fit should be very good in both the SPM and particle filter output; this comparison is shown in 
-\autoref{fig:comp1}. Recall that SPM worked on a slightly less noisy time series because of the
-spatial smoothing; this explains the lack of the sharp peak in the particle filter's preprocessed
-data. Regardless, as expected, both work.
-
-I chose the second voxel (\autoref{fig:comp2}) because it was active in SPM and it would appear to be in a prime
-location to be active in the SPM image (given the results in the surrounding voxels). The fit,
-however shows just why the residual was high and the mutual information was low. This is a
-prime example of a false positive due to the large smoothing kernel applied by SPM. 
-For instance at 75 seconds, the stimulus is not present, and so the signal should drop off; yet
-it doesn't. While a few peaks seem to match, most of the signal does not correlate with the
-expected state progression. This voxel is not being driven directly by the input, although it
-may be gated or driven through intermediate region.
-
-The third voxel, compared in \autoref{fig:comp3}, was far away from any other active voxels 
-and yet had a very low (around $.7$) residual. At the same time, the mutual information was
-high enough to make the point suspect. Although the estimated signal seems to run down the
-middle of the measurement signal; it is clear that there is a significant amount of noise
-present in the signal that is not being explained by the particle filter, 
-In both preprocessed time-series the input is extremely noisy, yet by the normalized residual 
-the response is good. This is an example of a false positive from the residual metric.
-
-The fourth voxel (\autoref{fig:comp4}) selected for analysis had a relatively high residual, and was not picked
-for activation by SPM. The mutual information was only $0.06$
-for the image. It is simple to see why the residual was not acceptable in this case.
-At the same time, the peaks do seem to correlate with the measurement peaks. Regardless,
-this is an example of a  false positive from the mutual information metric.
-
-The fifth voxel (\autoref{fig:comp5}) is an example of a region with increased activation
-due to smoothing. The fit provided by the BOLD particle filter is not very good, and the time
-series input to SPM is far less noisy. Regardless this is another case where the peaks
-seem to match, yet nothing else does. Take for instance the measurements from $30$ seconds
-to $40$ seconds. At that point there is a significant spike in the BOLD signal, yet the
-actual measured signal declines. This is an indication that the BOLD signal is not directly
-being driven by the input. It is likely that the cause of the activation in this reason is
-not true activation, but received from surrounding active regions through spatial smoothing.
-
 %23-21-7
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_pfilter_23_21_7}}\\
-\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/3_spm_23_21_7}}
-\caption{Section 3, Estimated vs. Actual BOLD response. T-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/3_pfilter_23_21_7}}\\
+\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/3_spm_23_21_7}}
+\caption{Section 3, Estimated vs. Actual BOLD response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 \begin{figure}[H]
-\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_pfilter_26_15_7}}\\
-\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/4_spm_26_15_7}}
-\caption{Section 4, Estimated vs. Actual BOLD response. T-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
+\centering
+\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/4_pfilter_26_15_7}}\\
+\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/4_spm_26_15_7}}
+\caption{Section 4, Estimated vs. Actual BOLD response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_pfilter_29_9_13}}\\
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_29_9_13}}
-\caption{Section 5, Estimated vs. Actual BOLD Response. T-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
-%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, T-Value
+\centering
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_pfilter_29_9_13}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_29_9_13}}
+\caption{Section 5, Estimated vs. Actual BOLD Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
+%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
 %of $4.17399$ and $MSE$ of $1.14171$.}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp6pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_pfilter_36_17_19}}\\
-\subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/6_spm_36_17_19}}
-\caption{Section 6, Estimated vs. Actual BOLD Response. T-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp6pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/6_pfilter_36_17_19}}\\
+\subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/6_spm_36_17_19}}
+\caption{Section 6, Estimated vs. Actual BOLD Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 \begin{figure}
-\subfigure[Particle Filter]{\label{fig:comp7pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/7_pfilter_29_26_1}}\\
-%\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=15cm]{images/5_spm_25_34_25}}
-\caption{Section 6, Estimated vs. Actual BOLD Response. T-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp7pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/7_pfilter_29_26_1}}\\
+%\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_25_34_25}}
+\caption{Section 7, Estimated vs. Actual BOLD Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
@@
 
-Voxel 6 (\autoref{fig:comp6}) is another region that is active according to both metrics used
-in this work, yet was missed by SPM. The time series in \autoref{fig:comp6} shows an extremely
-good fit, perhaps the best in this batch, for the particle filter. In contrast, the
-fit for SPM is abysmal. In spite of the fact that several other areas around are active
-as well, the smoothing seems to have completely wiped out activation in this voxel.
+I chose several voxels to further investigate from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
 
-Finally voxel 7 is a completely ambiguous time series. While the thresholds applied here
-are empirical, this voxel is on the borderline of the thresholds for both mutual information
-and the residual. The signal itself is extremely noisy, it should the voxel should be rejected.
-However this case is reminiscent of the pure-noise tests performed in the previous chapter. 
-This time series an extremely good example of the danger of false positives. In spite of the
-fact that the signal oscillates far faster than the BOLD estimate, because of that, it is somehow
-able to maintain a mutual information value of $.1052$ and a residual of just $1.09992$.
-As such, a clear method of detecting these sorts of regions is necessary if the results
-of the particle filter algorithm are toe be trusted.
+Section 1 (\autoref{fig:comp1}) had a very high $t$-score, high mutual information and
+a low error. Visual inspection makes it obvious that this voxel fits the BOLD model. A region
+such as this would be called strong activation by Neurologists. The
+reason for the cleaner looking signal in SPM is the additional Gaussian spatial smoothing
+applied in preprocessing. This
+has allowed for an extremely good fit in SPM, although the particle filter handles the noise
+well.
+
+Section 2 is a difficult voxel to assess (\autoref{fig:comp2}). 
+While certain peaks seem to correlate, over all 
+the particle filter input signal is extremely noisy. 
+This would appear to be a false positive on SPM's part because of the Gaussian smoothing.
+This voxel is not being driven directly by the input, although it
+could be gated or driven through intermediate region.
+
+Section 3 and 4 are both examples of nearly pure-noise signals, at least from the standpoint
+of the given stimulus (\autoref{fig:comp3}, \autoref{fig:comp4}). 
+These two voxels are false positives by the normalized residual.
+Note that in both cases the Mutual Information was extremely low, meaning that
+according the M.I. these were not good fits. Both section 3 and 4 have 
+significant peaks, begging the question of whether they may be active,
+but not correlated directly with the input.
+
+It is questionable whether the fifth voxel should be considered inside the
+brain; however, like section 2, SPM deemed it active (\autoref{fig:comp5}). 
+Also like section 2, the most likely cause of the problem is the smoothing
+applied for SPM. There are certain peaks that certainly match the
+BOLD signal, yet that is only a small part of the whole. 
+Take for instance the measurements between $30$ and $40$ seconds.
+In that interval there is a significant spike in the BOLD signal, yet the
+actual measured signal declines. This is an indication that the BOLD signal is not directly
+being driven by the input. The SPM fit is not particularly good either,
+with the $t$-score barely above the threshold of $4$.
+
+Section 6 (\autoref{fig:comp6}) is a clear example of a false negative 
+from SPM. The BOLD model fit is extremely good, yet the SPM signal looks completely
+like noise. One possible reason for this is that the low peak signal ($<0.03$),
+made it more susceptible to being smoothed out.
+
+Section 7 is an ambiguous time series. While the thresholds applied here
+are empirical, this voxel is on the borderline for both mutual information
+and the normalized residual. 
+This case is reminiscent of the pure-noise tests performed in the previous chapter,
+and is a perfect example of the false positive problem. 
+The signal oscillates far faster than the BOLD estimate, yet 
+somehow the mutual information reaches $.1052$ and the normalized residual is just below $1.1$.
 
@@
 contain the parameter maps for the system and \autoref{fig:realhist} is a histogram across all voxels
-for which the mutual information was greater than $.15$. As before, the threshold is not
+for which the mutual information was greater than $0.15$. As before, the threshold is not
 scientifically derived, yet in tests this threshold provided a decent balance to remove most
@@
 Regions with poor fit cannot have reliable parameter estimates because the input did not meaningfully
-correlate input with output. Therefore before creating the maps,
-each parameters map was masked to regions with mutual information greater than $.15$. I chose mutual
-information over the residual because in the maps comparing regression fitness, the mutual information
-maps had more coherency and less randomness. Additionally, in the single voxel tests (\autoref{sec:SingleVoxelReview}) 
-there was more separation between the no signal case and the signal case than there was in the 
-residual metric.
+correlate input with output, which is why a threshold was applied. 
+Mutual information was chosen over the residual because the
+maps had more coherency, and had less error in slice simulations (\autoref{fig:sim_hm2} vs. 
+\autoref{fig:simslice_hm_res}).
 
@@
 
-The parameter maps show consistency parameter across active regions, however, 
-the histogram of parameter estimates across all regions with $M.I. > .15$ is the
-more interesting result (\autoref{fig:realhist}).
+The histograms show variance higher than the prior, indicating that
+parameters are migrating quite a bit during particle filter convergence. 
+It can be said that all of these regions had relatively accurate BOLD 
+measurements, and yet the final parameter estimates vary widely.
 
 \section{Discussion}
-From the maps generated, the similarities to SPM8's results are encouraging. At the 
-very least the output of the particle filter seems to meet the quality of SPM. The normalization
-of the residual is certainly necessary. Although there is no hard threshold on the 
-normalized residual, it would appear that the normalization is providing a reasonable
-ordering of regression quality. Mutual information also shows promising results.
-Many of the differences between SPM and the particle filter seem to be driven
-more by the pre-processing methods than the regression method. The reason for 
-SPM applying such extensive smoothing is to combat false positives. While this is
-extremely important, regions such as \autoref{fig:comp6} show the problem with doing 
-so. In all, the particle filter method does very good job of estimating parameters that
-fit the measurements; and at the very least this method is a workable, albeit
-computationally intensive, alternative to SPM.
+The activation maps generated by the particle filter are very similar
+to those produced by SPM. This indicates that the particle filter is
+successfully estimating the BOLD output, in spite of the lack of spatial
+smoothing. Comparing the mutual information metric with the normalized residual,
+M.I. gives much better clusters, although that doesn't necessarily mean it
+is better. Although many of the differences between SPM and the particle filter 
+were driven by differences in preprocessing, it is worth noting that 
+the preprocessing applied by SPM are necessary for SPM to give decent results.
+Without spatial smoothing SPM is unable to cope with additional noise, and
+gives significant numbers of false positives.
+
+Although the parameter maps have some spatial coherence, the histograms
+of the parameters still indicate that the parameters are under-determined.

### 2010-08-16 doc/thesis/conclusion.tex
more on the results. Updated the HRF and applied pierces recommendations
"done"

@@
-\chapter{Conclusion}
+\chapter{Conclusions}
 \label{sec:Conclusion}
@@
 slow or relied on extensive assumptions. While the particle filter method
-is not quick, 60 seconds to analyze each voxel is well within the capabilities
+is not quick, 40 seconds to analyze each voxel is well within the capabilities
 of the typical research lab. Previous attempts have also treated
@@
 that identifying a single set of parameters is not possible with this 
-model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
-at this, the current work clearly demonstrates this fact. Therefore,
+model. Although sensitivity tests in Deneux et al. certainly hinted
+at this, the current work clearly demonstrates this fact \cite{Deneux2006}. Therefore,
 any single estimate of parameters is insufficient for analysis. As such,
-approaches to the BOLD model that do not treat the parameters as distributions
+identification of the BOLD model that do not treat the parameters as distributions
 will not be able to overcome the inherent indeterminability of the
@@
 The primary reason this method has not been used before is the high 
-dimensionality of the system. In \cite{Murray2008} the idea of learning
+dimensionality of the system. In Murray, 2008 the idea of learning
 the parameters is floated as the correct method, yet 
-learning 7 parameters with a Monte Carlo method is often intractable. The 
+learning 7 parameters with a Monte Carlo method was deemed intractable
+\cite{Murray2008}. The 
 concern was that so many parameters creates a prohibitively large 
@@
 
-Many of the approaches to determining areas of neural activation aim
+Many of the methods to determining areas of neural activation aim
 to be Bayesian, yet ultimately must make limiting assumptions about
-the distributions that turn the approach into classical statistics. 
-This approach is a true Bayesian non-parametric approach that given
+the distributions that remove many of the benefits. 
+The particle filter is a true Bayesian non-parametric algorithm that, given
 enough measurements will approach the true probability distribution
@@
 uniquely identifiable is a disappointing one in light years of research 
-attempting to learn these parameters, it in fact further demonstrates
+attempting to learn these them, it in fact further demonstrates
 the need to estimate probability distribution rather than a single 
 parameter estimate. At the same time, the output of the particle filter
-is still about to give good estimates of the BOLD output, and not
+is still able to give good estimates of the BOLD output, and is not
 dependent on heavy Gaussian smoothing. Therefore, all the current 
@@
 To borrow a computer term, the particle filter algorithm is backward 
-compatible with the current methods.
+compatible with the current methods. Indeed, the particle filter excelled
+at determining activation in very low SNR simulations 
+(\autoref{sec:Multi-voxel Simulation}). Not only that, but the particle
+filter algorithm identified areas of activation that were completely
+missed by SPM (\autoref{fig:comp6}).
 

### 2010-08-16 doc/thesis/discussion.tex
"done"

@@
 %% Output estimates good
-There are two possible, but unlikely causes for the large variance in
+There are two possible causes for the large variance in
 parameter estimates. First, it could be that the particle filter is
@@
 possible explanation is that the given stimulus is not sufficiently
-varied to bring out all the properties of the system. This also is unlikely
-because of results in \autoref{sec:VeryLongSim} which used a random impulse
-stimulus sequence. The high correlation of the parameters in \autoref{sec:VeryLongSim},
-point to a significant trade off in parameters.
+varied to bring out all the properties of the system. While it is 
+possible that this is the case, it is unlikely that even the best 
+stimulus sequence could differentiate all the parameters. The fact
+that \autoref{sec:VeryLongSim}, which used randomly spaced impulses,
+still had a high correlation definitely indicates that the parameters
+are ill-defined.
 
-Thus the results show that the parameters of the BOLD model are under
+Thus there is little doubt that the parameters of the BOLD model are under
 constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
-it is a notable conclusion \cite{Deneux2006}. Thus, other methods
+it is a notable conclusion \cite{Deneux2006}. Other methods
 that depend on a point estimate of the parameters, such as least squares 
-or Kalman filters (which uses the first two moments) are limited. While estimates of
-the  BOLD signal may still be correct, the estimate
-of the underlying parameters and state variables cannot be described without using
+or Kalman filters (which uses the first two moments) are limited in their
+capacity to estimate such parameters. While estimates of
+the  BOLD signal may still be correct, the 
+underlying parameters and state variables cannot be described without using
 a joint probability distribution function. In this sense, particle 
@@
 of Friston et al. may have been wider than true parameter distribution
-\cite{Friston2002}. Any algorithm that attempts to calculate parameters
-of the BOLD model should be run multiple times, to check for consistency.
+\cite{Friston2002}. Future works that attempt to calculate parameters
+of the BOLD model should be run multiple times to ensure consistency,
+which most likely cannot be attained. This result is particularly
+troublesome given the number of papers that use those parameters for
+a prior distribution.
 
@@
 tests were more sensitive, there were some additional false positives, 
-though at this problem is difficult to quantify. Regions with M.I. above 
+though the problem is difficult to quantify. Regions with M.I. above 
 $0.15$ consistently fit the FMRI data well, and simulations showed that
@@
 particular problem there can be no guarantee of a global minimum, and although
-other techniques guarantee a local minimum, the particle filter in some
-sense can settle into multiple local minima.
+other techniques guarantee a local minimum, tests show that the particle
+filter did converge relatively quickly \autoref{sec:VeryLongSim}.
 
@@
 simple. An understanding of Bayesian statistics is all that is necessary to understand
-how the particle filter works. This is in contrast to the Volterra based approach
-of, which can be quite complex. Additionally very few assumptions
-are needed for the particle filter; which reduces the risk of those assumptions
-being violated. Fewer and more realistic assumptions make the particle 
+how the particle filter works.  Additionally very few assumptions
+are needed for the particle filter. Few assumptions  reduces the risk of those assumptions
+being violated. Fewer and more realistic assumptions also make the particle 
 filter more robust to unforeseen difficulties in FMRI data.
+
+%\section{Particle Filter Options}
+%There are a large number of options available in the particle filter, that 
+%were of little use. 
+%\section{Didn't Work}
+%A more subtle difference that could be made in detrending is how the percent difference
+%is calculated. In this work the spline was subtracted, and then the result
+%was divided by the average of the initial signal. A more correct method may be dividing
+%by the spline value at that particular point. The reason for not dividing by the spline
+%at that point is that it could be less stable, and in a sense the trend has already
+%been removed. The difference is not particularly large though 
+%(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.
+
+%Finally, rather than adding a constant to the detrended data before
+%applying the particle filter, a DC gain parameter could be added to the model. In
+%tests, this could work well, however, it often did not. The problem with this
+%could simply be the addition of another degree of freedom without any increase in
+%the number of particles. Increasing the number of particles further though
+%can be computationally intractable. Thus, while this is in a sense the correct
+%solution, it is an impractical one.
+
+%Starting with a flattened
+%prior (by setting weights to be non-uniform after the initial particles have
+%been drawn) was not used in the final analysis. The reason for not flattening
+%the prior was a practical issue with weighting points in a 
+%7 dimensional joint distribution. Often differences in particle weights approached machine 
+%precision, meaning that the flattened prior was actually far from flat. Instead
+%the distribution and results became unpredictable. This
+%is quantization issue that will exist unless the prior were made to be deterministic
+%or the initial number particles was raised to computationally prohibitive levels.
+
+%Another improvement that could be made to the prior is increasing the initial variance.
+%This would allow the particle filter to learn a wider range of parameters; at the cost
+%of decreased convergence rate and support density. In tests, both real and simulated,
+%increasing the variance tended to allow the particle filter to over-smooth 
+%(ex \autoref{fig:badfit_param1}). In 
+%essence the particle filter tended to converge to a mean where the time constants
+%dominated the results by smoothing out all the peaks and troughs. The result is that
+%no other parameters alter the estimated BOLD signal. 
+%\begin{figure}
+%\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/badfit_param1}
+%\caption{Large variance in time constants over-smoothing.}
+%\label{fig:badfit_param1}
+%\end{figure}
+%In \autoref{fig:badfit_param1}, the particle filter still had not fully converged, and
+%given more measurements it could still converge to a better estimate. Thus, increases
+%in the prior variance necessarily must be accompanied by more measurements,
+%although re-presenting data could be used to artificially extend the measurements.
+
+%Even more basic, its possible to artificially increase the number of 
+%measurements by presenting them the time-series several times. 
+%This activity is similar to the process used in neural networks,
+%and gives the particle filter longer to converge to the optimal estimate. On the 
+%downside, this increases runtime and artificially reduces variance. 
+%Thus it does not necessarily improve the results. In tests, it did reduce the
+%final covariance but did not lead to better estimates.
 
@@
 %%% Automatic estimation of measurement error.
-\section{Improving the Particle Filter}
+\section{Algorithm Improvements}
 \label{sec:Particle Filter Variations}
+There are a few areas where future research may improve upon the current
+work. The first is modifying the prior 
+distribution. The prior distribution used in this work is listed in \autoref{tab:Prior}, 
+and is based on the findings of Friston et al. \cite{Friston2000}. Unfortunately, that
+result depended on a quadratic approximation (Volterra Series) which has
+not been extensively tested (at least not in a published work). 
+Additionally, the current prior is
+based only around the BOLD output, which is why it is good at generating
+estimates of the BOLD signal, but weak in estimating parameters. 
+The best solution would be to have
+in-vivo estimates of the actual parameters, although this
+is unlikely to happen. Instead, better estimates could be found
+using population studies with additional measurements as discussed in 
+\autoref{sec:Sideways Measurements}. Better knowledge of the prior distribution
+would make it possible to decrease variance of certain 
+parameters, reducing the mobility of the parameters.
 
-There are a few areas where future research may improve upon the current
-work. The first is modifying the Prior 
-Distribution. The choice in this paper to use the distribution calculated in
-\cite{Friston2000}, may not be optimal and so is open to modification. The
-second major area is in preprocessing, in particular the method used to remove
-drift in the FMRI signal. This is an area of ongoing research and so future
-algorithms could be strengthened by applying the latest technique rather than a 
-spline. Third, the experiment itself may be modified to improve results.
-And the last method is improving the , by using a more
-advanced version.
-
-Collapse parameters
-
-\subsection{Prior Distribution}
-The prior distribution used in this work is listed in \autoref{tab:Prior}, 
-and is based on the findings of \cite{Friston2000}. Unfortunately, that
-result depended on a quadratic approximation (Volterra Series) which has
-not been extensively tested (at least not in a published work). As such,
-the prior is probably not complete. Additionally, the current prior is
-based only around the BOLD output, which is why it is capable of generating
-good estimates of the BOLD signal. However, it would be advantagous to have
-in-vivo estimates of the actual parameters. Better estimates could be found
-using population studies with additional measurements as discussed in 
-\autoref{sec:Sideways Measurements}. Changes to this prior could
-boost the power of the BOLD particle filter algorithm described here.
-
-Starting with a flattened
-prior (by setting weights to be non-uniform after the initial particles have
-been drawn) was not used in the final analysis. The reason for not flattening
-the prior was a practical issue with weighting points in a 
-7 dimensional joint distribution. Often differences in particle weights approached machine 
-precision, meaning that the flattened prior was actually far from flat. Instead
-the distribution became unpredictable, which made the results unpredictable. This
-is quantization issue that will exist unless the prior were made to be deterministic
-or the initial number particles was raised to computationally prohibitive levels.
-
-Another improvement that could be made to the prior is increasing the initial variance.
-This would allow the particle filter to learn a wider range of parameters; at the cost
-of decreased convergence rate and support density. In tests, both real and simulated,
-increasing the variance tended to allow the particle filter to over-smooth 
-(ex \autoref{fig:badfit_param1}). In 
-essence the particle filter tended to converge to a mean where the time constants
-dominated the results by smoothing out all the peaks and troughs. The result is that
-no other parameters alter the estimated BOLD signal. 
-\begin{figure}
-\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/badfit_param1}
-\caption{Large variance in time constants over-smoothing.}
-\label{fig:badfit_param1}
-\end{figure}
-In \autoref{fig:badfit_param1}, the particle filter still had not fully converged, and
-given more measurements it could still converge to a better estimate. Thus, increases
-in the prior variance necessarily must be accompanied by more measurements,
-although re-presenting data could be used to artificially extend the measurements.
-
-\subsection{Detrending}
-A large number of de-trending methods were tested, but all had some major problem. 
-One viable method that could be utilized given the right experimental design is detrending
+Another major area for improvement is removing drift. 
+A large number of detrending methods were tested, but every method was in some way limited. 
+One viable method that could be utilized, given the right experimental design, is detrending
 based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
@@
 
-Another potential way to deal with detrending is to 
-linearize. This would use the delta between measurements for fitting rather than the 
+In terms of dealing with drift, linearizing is another possibility.
+This would use the delta between measurements for fitting rather than the 
 direct value. This has the advantage of not requiring detrending and thus 
@@
 for processing. The effectiveness of this method depends directly on the type of stimulus.
-In an event driven stimulus with sparse stimuli this could be extremely effective because
-the DC level has minimal data; however the longer high levels are held the less effective
-this will be. On the other hand, splines will also reduce the plateaus, so its possible
-linearizing may be just as effective. I found in tests that large drift-low white noise
+In a test with rapid impulse stimuli, this could be extremely effective because
+the DC level has minimal data; conversely prolonged flat level levels make this less effective.
+In tests I found that large drift-low white noise
 type signals performed much better with a linearization approach, as one might expect. 
-The results were equal to or worse in real data; but that was only for one particular
-stimulus sequence.
+For the stimulus sequence used in \autoref{sec:RealData}, the results tended to be worse
+than using spline detrending.
 
-A subtle difference that could be made in detrending is how the percent difference
-is calculated. In this work the spline was subtracted, and then the result
-was divided by the average of the initial signal. A more correct method may be dividing
-by the spline value at that particular point. The reason for not dividing by the spline
-at that point is that it could be less stable, and in a sense the trend has already
-been removed. The difference is not particularly large though 
-(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.
-
-Finally, rather than adding a constant to the detrended data before
-applying the particle filter, a DC gain parameter could be added to the model. In
-tests, this could work well, however, it often did not. The problem with this
-could simply be the addition of another degree of freedom without any increase in
-the number of particles. Increasing the number of particles further though
-can be computationally intractable. Thus, while this is in a sense the correct
-solution, it is an impractical one.
+As discussed in \autoref{sec:Methods Weighting Function}, choosing a weighting
+function is difficult, and the optimal solution varies based on the input.
+Automatically estimating measurement error could improve
+the quality of the particle filter results. Although I made some attempts to do
+this, finding a generic, consistent solution is complex, and often depends on the
+experimental design. Part of the problem is that SNR varies greatly across
+the brain, and there is no way to actively measure noise without also knowing
+the underlying signal. One possible solution is having the particle filter automatically
+set the weight based on the total particle weight, or on the prevalence of resampling.
 
@@
 \label{sec:Sideways Measurements}
-One definite way of improving the results of the particle filter is more 
+One definite way of improving the results of the particle filter is additional
 measurements. While increasing the sample rate of FMRI scanners may
 not be possible, simultaneous measurements of volume and flow is
-possible \cite{Hu2009}, albeit at 9T in a cat. Just one of those measurements
+possible, albeit at 9T in a cat \cite{Hu2009}. Just one of those measurements
 though would be extremely powerful when incorporated into the BOLD
-model. By adding another measurement, especially for one of the other state
-variables (like blood volume and blood flow), the
-variance in the parameter estimates would certainly drop. Of course, more measurements
-may be gained by simply performing longer FMRI tests, or concatenating several 
-together. Even more basic, its possible to artificially increase the number of 
-measurements by presenting them the time-series several times. 
-This activity is similar to the process used in neural networks,
-and gives the particle filter longer to converge to the optimal estimate. On the 
-downside, this increases runtime and artificially reduces variance. 
-Thus it does not necessarily improve the results. In tests, it did reduce the
-final covariance but did not lead to better estimates.
+model. By adding another measurement, the
+variance in the parameter estimates would significantly drop. The fact
+that such a measurement would be closer to the true stimulus would make it
+all the more powerful. 
+A more conventional method of adding measurements is to 
+simply perform longer FMRI tests. Although this wouldn't be 
+groundbreaking, it would certainly increase the ability of the 
+particle filter to develop the posterior distribution. 
+It may also be possible to concatenate multiple FMRI tests together.
+It should be possible to link the multiple runs simply through
+the final parameter distribution from the previous run.
 
-Additionally, given the non-linearities in the system, differences in stimuli
-can make a huge difference in the observability of parameters. The mentality 
+Given the non-linearities in the system, differences in stimuli
+could make a large difference in the observability of parameters. The mentality 
 for using a physiologically based nonlinear model for BOLD signal is to model 
-those nonlinearities. Its logical then that nonlinear parameters may not
-be identifiable when the input is primarily an impulse response. Therefore,
-a wide range of inputs may shed further light on the parameters than found 
-in this work.
-
-As discussed in \autoref{sec:Methods Weighting Function}, choosing a weighting
-function is difficult. Automatically estimating measurement error could improve
-the quality of the particle filter results. Although I made some attempts to do
-this, finding a generic, consistent solution is complex, and often depends on the
-experimental design. This problem strikes at the very of heart of the difficulty
-in analyzing FMRI, and so it is not likely to be solved soon.
+those nonlinearities. Its logical then that certain nonlinear parameters may not
+be identifiable when the input is primarily an impulse response. It 
+has been reported that short responses are disproportionately large 
+in FMRI data \cite{Miller2001, Deneux2006}. Therefore,
+a wide range of activation hold time may shed further light on 
+the parameters' distribution as well as the validity of the BOLD model.
 
@@
 %% Comparison of posterior of parameters
-\section{Applications}
+\section{Future Applications}
 There are a number of advantages to the particle filter approach presented
-here. In the past FMRI data has been analyzed strictly for determining
+here. In the past, FMRI data has been analyzed strictly for determining
 correlation between a stimulus and response. With this new method
@@
 to visualize because of the high-dimensionality, could nevertheless be 
-correlated with neural pathologies. Despite the fact that the parameters
+correlated with neural pathologies. In spite of the fact that the parameters
 are under-determined, the final distribution is still a reduction in the
-uncertainty of the parameters. This reduction in uncertainty naturally
-contains information which may be exploited based on non-parametric 
-statistical analyses.
+uncertainty of the parameters. The availability of a full posterior distribution
+opens up many avenues for further inquiry, for instance
+to compare normal vs. symptomatic
+populations. This would be especially useful in patients whose symptoms
+do not include structural changes in brain. This would effectively be
+a way to differentiate the way the brain operates.
 
-Because of the limitations present in every image modality in neuroimaging,
+Because of limitations present in every imaging modality,
 its becoming increasingly clear that combining data from multiple sources
-will be necessary to push neurology forward. In order to do so however, the
+will be necessary to push Neurology forward. In order to do so however, the
 output of each source needs to fully represent what information that source
@@
 this value available opens up new avenues for determining inter-regional
-dependencies. All the regions for which the value of $s$ are trustworthy
-correlate closely with the original stimuli, so determining the connection
-between their values of $s$ would be less than edifying. Instead, the values
-of $s$, which in some sense is the activation in a particular region, could
-be used to drive other inputs. Thus, the particle filter could be re-run
+dependencies. The values of $s$ could serve as a proxy for activation in 
+that region and thus could be used to drive other inputs regions' $u(t)$. 
+Thus, the particle filter could be re-run
 with the time-series of a particular voxel's $s$ value as an additional stimulus.
@@
 state variables present in the BOLD equations; the potential benefits of being
-able to determine this information is limitless. Of course, improving the quality
-of the parameter estimates is necessary if such information is to be fully
-trusted.
+able to determine this information is limitless. Of course the quality of these
+estimates will continue to improve with the model priors.

### 2010-08-16 doc/thesis/conclusion.tex
small edits

@@
 identification of the BOLD model that do not treat the parameters as distributions
-will not be able to overcome the inherent indeterminability of the
+will not be able to overcome the inherent ill-posed nature of the
 parameters. Besides the Unscented Kalman Filter, this is the only approach
 that accomplishes this task without repeatedly calculating the parameters 
-to build a distribution. The Unscented Kalman Filter, of course, is limited
+to build a distribution. The Unscented Kalman Filter though, is restricted
 to Gaussian estimation which limits its ability to estimate the posterior.
@@
 concern was that so many parameters creates a prohibitively large 
-search space, that requires too many particle to learn properly. To overcome
+search space, which requires too many particle to learn properly. To overcome
 this difficulty, instead of starting the algorithm with 1000 particles,
@@
 when the search space was the largest, the number of particles was
-sufficiently dense to represent the prior distribution. Then resampling
-was performed for the first time, the number of particles was dropped, 
-since the weight of most the particles had dropped to 0. The result is
+sufficiently dense to represent the prior distribution. The result is
 a particle filter algorithm that spends computing resources only where
-they are really needed.
+it is really needed.
 
-Many of the methods to determining areas of neural activation aim
-to be Bayesian, yet ultimately must make limiting assumptions about
-the distributions that remove many of the benefits. 
 The particle filter is a true Bayesian non-parametric algorithm that, given
 enough measurements will approach the true probability distribution
-of the parameters. While the conclusion that the parameters are not
-uniquely identifiable is a disappointing one in light years of research 
-attempting to learn these them, it in fact further demonstrates
+of the parameters. The conclusion that the parameters are not
+uniquely identifiable is a disappointing one in light of the plethora
+of research papers that have attempted to learn them. However, the fact
+that the parameters are ill-posed further demonstrates
 the need to estimate probability distribution rather than a single 
@@
 
+The physiological plausibility is also of great value to future
+research. Unlike traditional methods, use of the BOLD model allows 
+the incorporation of outside, physiological knowledge in regression.
+If, for instance, it becomes possible to measure $v$ in vivo that data
+could instantly be added to the model to improve regression. Data can
+also be retrieved from the model. BOLD parameter distributions could 
+for the first time be used as diagnosis tools. In the past the fact 
+that the parameters are
+ill-posed would have prevented such research. However, now that a full
+distribution is available, comparing parameter estimates is actually
+feasible.
+
 Concluding, the particle filter provides comparable performance with

### 2010-08-16 doc/thesis/discussion.tex
small edits

@@
 varied to bring out all the properties of the system. While it is 
-possible that this is the case, it is unlikely that even the best 
-stimulus sequence could differentiate all the parameters. The fact
-that \autoref{sec:VeryLongSim}, which used randomly spaced impulses,
-still had a high correlation definitely indicates that the parameters
+possible that this is the case, it is unlikely that any
+stimulus sequence could differentiate all the parameters. 
+Its also extremely intesting that the histograms of parameters' mean
+approach the width of the prior, even when the particle filter converges
+proprerly to far thinner distributions. Its possible the prior 
+distribution is far wider than 
+it needs to be. Finally, the fact that \autoref{sec:VeryLongSim}, which 
+used randomly spaced impulses,
+still had a high correlation further indicates that the parameters
 are ill-defined.
@@
 not been extensively tested (at least not in a published work). 
-Additionally, the current prior is
-based only around the BOLD output, which is why it is good at generating
-estimates of the BOLD signal, but weak in estimating parameters. 
+This is in additoin to the fact that the parameters are ill-defined, and
+thus the parameters found in Friston et al. are most likely far too wide.
+The prior is also
+based completely on the BOLD output, which is why it is good at generating
+estimates of the BOLD signal, but weak at estimating parameters. 
 The best solution would be to have
@@
 A large number of detrending methods were tested, but every method was in some way limited. 
-One viable method that could be utilized, given the right experimental design, is detrending
-based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
-constant to be added to the pre-processed signal for the BOLD model to fit properly.
-The disadvantage of this approach is that it could hide long fall times by normalizing them
-out. It also requires periodic breaks in the stimulus, which could reduce value of those
-samples for fitting purposes.
+One potentially viable method that could be utilized, given the right 
+experimental design, is detrending based on intervals of low activity. This 
+has the advantage that it wouldn't require an arbitrary
+constant to be added to the pre-processed signal for the
+BOLD model to fit properly.  The disadvantage of this approach is 
+that it could hide long fall times by normalizing them
+out. It also requires periodic breaks in the stimulus, 
+which could reduce value of those samples for fitting purposes.
 
 In terms of dealing with drift, linearizing is another possibility.
-This would use the delta between measurements for fitting rather than the 
+This would use the delta between measurements for fitting, rather than the 
 direct value. This has the advantage of not requiring detrending and thus 
-gives nearly raw data to the particle filter
-for processing. The effectiveness of this method depends directly on the type of stimulus.
+gives nearly raw data to the particle filter for processing. 
+The effectiveness of this method depends directly on the type of stimulus.
 In a test with rapid impulse stimuli, this could be extremely effective because
-the DC level has minimal data; conversely prolonged flat level levels make this less effective.
-In tests I found that large drift-low white noise
-type signals performed much better with a linearization approach, as one might expect. 
+the DC level has minimal data; conversely prolonged flat levels make this less effective.
+In tests I found that large drift/low white noise
+type signals performed much better with a linearization approach.
 For the stimulus sequence used in \autoref{sec:RealData}, the results tended to be worse
@@
 
-\subsection{Experimental Changes}
+\section{Experimental Design Improvements}
 \label{sec:Sideways Measurements}
-One definite way of improving the results of the particle filter is additional
-measurements. While increasing the sample rate of FMRI scanners may
+One definite way of improving the results of the particle filter is through
+additional measurements. While increasing the sample rate of FMRI scanners may
 not be possible, simultaneous measurements of volume and flow is

### 2010-08-16 doc/thesis/results.tex
small edits

@@
 \subfigure[Region labels for simulated slice.]{
-\includegraphics[height=9.5cm]{images/simregions.png} \label{fig:simslice_mask} }
-\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9.5cm]{images/sim_gm}}
+\includegraphics[height=9cm]{images/simregions.png} \label{fig:simslice_mask} }
+\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9cm]{images/sim_gm}}
 \subfigure[Residual Map. Scale is Normalized error. Lower (yellow) is better.]
-{\label{fig:simslice_hm_res} \includegraphics[height=9cm]{images/sim_hm}}
+{\label{fig:simslice_hm_res} \includegraphics[height=8cm]{images/sim_hm}}
 \subfigure[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
-{\label{fig:sim_hm_mi} \includegraphics[height=9cm]{images/sim_hm_mi}}
+{\label{fig:sim_hm_mi} \includegraphics[height=8cm]{images/sim_hm_mi}}
 \caption{Comparison of activation with greymatter, parameter regions.}
@@
 of the particle filter at identifying regions of activation.
- The thresholds applied to this
-slice, both for mutual information and residuals are arbitrary. Applying a
-threshold is helpful for visualization, but is rarely useful for further analysis.
+The thresholds applied to this slice, both for mutual information and 
+residuals somewhat arbitrary. Applying a
+threshold is helpful for visualization but is rarely useful for further analysis.
 As noted in \autoref{sec:SingleVoxelReview},

### 2010-08-16 doc/thesis/background.tex
used diction a little then gave up

@@
-\chapter{Alternative Techniques}
+\chapter{Alternate Methods}
 \label{sec:Prior Works}
@@
 FMRI, activation detection is similar and worth discussing. Estimation of 
-parameters may be considered generalization of the idea 
+parameters is a generalization of the idea 
 of activation detection.
@@
 value of $\frac{0.05}{10000}$. The probability of 
-a single false positive would then $0.05$. Unfortunately this leads to unrealistically
+a single false positive would then be $0.05$. Unfortunately this leads to unrealistically
 low p-values; so low that it would be impossible for any biological system to satisfy. To
@@
 decreasing the effective number of independent measurements. Because
-the number of independent measurements is smaller, Bonferroni correction
+the number of \emph{independent} measurements is smaller, Bonferroni correction
 can theoretically be applied with a lower scaling factor than
@@
 noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
-to apply some sort of high-pass filter to the data. Without applying
+to apply a high-pass filter to the data. Without applying
 such a filter, P values must be set extraordinarily high to prevent
@@
 \label{sec:Current Techniques General Linear Model}
-The most common FMRI analysis technique is called SPM, though there
-are more advanced versions than the simple square wave
-method discussed in \autoref{sec:Square Wave}. Hierarchal
+The most common FMRI analysis tool is SPM, which generates
+statistical parametric map using more advanced methods than a simple
+square wave. Hierarchal
 Models are one important improvement that allows researchers

### 2010-08-16 doc/thesis/methods.tex
saw a typo and had to fix

@@
 will quickly have weights near 0. Thus, a weighting function that
-exactly matches $P(Y(t) | X(t))$ will easily, and correctly throw out incorrect
+exactly matches $P(Y(t) | X(t))$ will easily and correctly throw out invalid
 particles.  The cost of choosing an overly broad distribution for this
@@
 
-After trial and error, for this work the Gaussian with a standard deviation 
-of $0.005$ was chosen. Although using an adaptive weighting function
+After trial and error, for this work I chose the Gaussian with a standard deviation 
+of $0.005$. Although using an adaptive weighting function
 might be better, in tests this often led to unpredictable results. 
@@
 algorithm forces resampling. The choice of 20 seconds is 
-arbitrary, but at the very least it the particles some time to be reweighted.
+arbitrary, but at the very least it gives the particles time to be weighted.

### 2010-08-16 doc/thesis/conclusion.tex
spelling fixes etc.

@@
 dimensionality of the system. In Murray, 2008 the idea of learning
-the parameters is floated as the correct method, yet 
+the parameters is floated as the best method, yet
 learning 7 parameters with a Monte Carlo method was deemed intractable
@@
 
-The particle filter is a true Bayesian non-parametric algorithm that, given
+The particle filter is a Bayesian non-parametric algorithm that, given
 enough measurements will approach the true probability distribution
 of the parameters. The conclusion that the parameters are not
-uniquely identifiable is a disappointing one in light of the plethora
+uniquely identifiable is disappointing in light of the plethora
 of research papers that have attempted to learn them. However, the fact
 that the parameters are ill-posed further demonstrates
-the need to estimate probability distribution rather than a single 
+the need to estimate a probability distribution rather than a single
 parameter estimate. At the same time, the output of the particle filter
@@
 conventional tests, but also provides a platform for a wide range of
-future uses beyond what is possible with the methods applied in the
-past.
+future uses beyond what was possible with the methods that have been
+used before.

### 2010-08-16 doc/thesis/discussion.tex
spelling fixes etc.

@@
 stimulus sequence could differentiate all the parameters.
-Its also extremely intesting that the histograms of parameters' mean
+Its also extremely interesting that the histograms of parameters' mean
 approach the width of the prior, even when the particle filter converges
-proprerly to far thinner distributions. Its possible the prior 
+properly to far thinner distributions. Its possible the prior
 distribution is far wider than
@@
 it is a notable conclusion \cite{Deneux2006}. Other methods
-that depend on a point estimate of the parameters, such as least squares 
+that calculate a point estimate of the parameters, such as least squares
 or Kalman filters (which uses the first two moments) are limited in their
@@
 
-As the results in \autoref{sec:RealData} show, the heatmaps, especially
+As the results in \autoref{sec:RealData} show, the heat maps, especially
 those of mutual information, closely resembled the results of SPM. While the activation
@@
 
-A more practical benefit with the particle filter is that it is mathematically
-simple. An understanding of Bayesian statistics is all that is necessary to understand
-how the particle filter works.  Additionally very few assumptions
-are needed for the particle filter. Few assumptions  reduces the risk of those assumptions
-being violated. Fewer and more realistic assumptions also make the particle 
-filter more robust to unforeseen difficulties in FMRI data.
+%A practical benefit with the particle filter is that it is mathematically
+%simple. An understanding of Bayesian statistics is all that is necessary to understand
+%how the particle filter works.  Additionally very few assumptions
+%are needed for the particle filter. Few assumptions  reduces the risk of those assumptions
+%being violated. Fewer and more realistic assumptions also make the particle
+%filter more robust to unforeseen difficulties in FMRI data.
 
@@
 not been extensively tested (at least not in a published work).
-This is in additoin to the fact that the parameters are ill-defined, and
+This is in addition to the fact that the parameters are ill-defined, and
 thus the parameters found in Friston et al. are most likely far too wide.
@@
 constant to be added to the pre-processed signal for the
-BOLD model to fit properly.  The disadvantage of this approach is 
+BOLD model to fit properly. It would also minimize the actual
+removed. The disadvantage of this approach is
 that it could hide long fall times by normalizing them
@@
 
-In terms of dealing with drift, linearizing is another possibility.
+Linearizing is another possible method of dealing with drift.
 This would use the delta between measurements for fitting, rather than the
@@
 In a test with rapid impulse stimuli, this could be extremely effective because
-the DC level has minimal data; conversely prolonged flat levels make this less effective.
+the DC signal has minimal data; conversely prolonged flat levels make this less effective.
 In tests I found that large drift/low white noise
@@
 possible, albeit at 9T in a cat \cite{Hu2009}. Just one of those measurements
-though would be extremely powerful when incorporated into the BOLD
+would be extremely powerful when incorporated into the BOLD
 model. By adding another measurement, the
@@
 that such a measurement would be closer to the true stimulus would make it
-all the more powerful. 
+all the more useful.
 A more conventional method of adding measurements is to
 simply perform longer FMRI tests. Although this wouldn't be
-groundbreaking, it would certainly increase the ability of the 
+ground breaking, it would certainly increase the ability of the
 particle filter to develop the posterior distribution.
-It may also be possible to concatenate multiple FMRI tests together.
-It should be possible to link the multiple runs simply through
-the final parameter distribution from the previous run.
+It may also be possible to concatenate multiple FMRI tests together
+by using the parameter estimates from previous tests as the
+prior for later tests.
 
@@
 populations. This would be especially useful in patients whose symptoms
-do not include structural changes in brain. This would effectively be
-a way to differentiate the way the brain operates.
+are caused by functional difference in the brain, rather than structural differences.
+This would effectively be
+a way to differentiate the way the brains operate.
 
@@
 distribution of neural efficiency, those two beliefs may be combined into
-one conditional belief for the probability of neural efficiency.
+one conditional belief for a new conditional probability of neural efficiency.
 

### 2010-08-16 doc/thesis/methods.tex
spelling fixes etc.

@@
 included noise in the update of $x$, however that is not included here.
-The reason for the difference is that cloud of particles is, to some extent,
+The reason for the difference is that the cloud of particles is, to some extent,
 able to account for that noise. It is common, however, to model that noise
@@
 BOLD model is dissipative. When no stimuli are applied, all the particles
-decay to ($\{0,1,1,1\}$). Typical particle filters 
-also use this state noise as an exploratory measure; however this method is
-less necessary when good priors are available.
+decay to $\{0,1,1,1\}$. Typical particle filters
+also use this state noise as an exploratory measure; however, this method is
+less necessary when good priors are available, and when regularized particle filtering
+is being used.
 
-For all the analyses  in this work, $1400$ integration points
+For all the analysis  in this work, $1400$ integration points
 per second were used.  Typically a step size of $0.001$ was sufficient,
@@
 task is motion correction. To do this, a single volume in time is chosen, and
-volumes at every other time are realigned to best match the target volume. This corrects
+volumes at every other time point are realigned to best match the target volume. This corrects
 for motion by the patient as well as small changes in the magnetic
@@
 filter is applied across the image as discussed in \autoref{sec:RFT}.
-After this, detrending is performed which is discussed in \autoref{sec:Detrend}.
+After this detrending is performed, which is discussed in \autoref{sec:Detrend}.
 Recall that FMRI signal levels are unit-less and though detrending is not
@@
 aggregate signal over the space of cubic centimeters, there are
-plenty of noise sources . Though local neurons act
+plenty of noise sources. Though local neurons act
 together (i.e. around the same time), the density of neurons, the
@@
 During resting state, the patient is shown no images, and he is asked
-to avoid movement and complex thought.  Overall though there should be 
+to avoid movement and complex thought.  Overall there should be
 very little activation, and thus the signal consists entirely of noise.
@@
 is informative, so Q-Q plots are used to compare example data with the
-Normal distribution. Additionally, in FMRI data the noise is often considered 
+Normal distribution. Additionally, in FMRI data, the noise is often considered
 to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
-characterized by steps that are Gaussian and independent. The simulations discussed in 
+characterized by steps that are independent and Gaussian. The simulations discussed in
 \autoref{sec:Single Voxel Simulation} make use of this,
@@
 
-De-trending the time-series by subtracting a spline fit to the distribution
+Detrending the time-series by subtracting a spline fit to the distribution
 removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
@@
 the Normal, \autoref{fig:QQs:D} is much improved compared to \autoref{fig:QQDC:D}.
-In all, the de-trending is effectively removing Wiener noise.
+In all, the detrending is effectively removing Wiener noise.
 
@@
 aspect of a Weiner process, presumably the result of integrating some
-$\nu_x$ is difficult to compensate for, and so many methods
+$\nu_x$, is difficult to compensate for, and so many methods
 have been developed to compensate for it. Tanabe et al. and Smith et al.
@@
 a spline worked the best \cite{Tanabe2002}.
-The benefit of the spline versus wavelets, high pass 
-filtering or other DC removal techniques is that the frequency response is not set.
-Rather, the spline is adaptive to the input. Unfortunately no method will 
+Unfortunately no method will
 perfectly remove noise, and no method will leave the signal untouched.
 
-The method I used to calculate the spline was picking one knot for every 20
+The method I used to calculate the spline was one knot for every 20
 measurements in an image. Thus a 10 minute session at a repetition time of
@@
 would be able to settle on a good estimate. Yet adding another dimension increases the
-complexity of the model, when the parameter is relatively easy to estimate
+complexity of the model, for a parameter that is relatively easy to estimate
 by visual inspection.  In this work a simpler approach was used. To determine
@@
 
-%\subsection{Linearizing Noise}
-%\label{sec:Methods Delta Based Inference}
-%The alternative to these sorts of low frequency manipulation is to
-%go around the problem in another way. Here, I propose a 
-%different method of dealing with the drift. 
-%Instead of comparing the direct output of the particle filter with the direct
-%measurement, the algorithm would compare the change in signal over a single TR,
-%with the result of integrating the model for the same period. 
-%In most signal processing cases this would be foolish, but that is because the 
-%general assumption is that all noise is high frequency. Considering 
-%the fact that every BOLD analysis pipeline uses a high pass filter,
-%whereas low poss temporal filter are rarely applied, it makes sense
-%that a derivative type method could work. The benefit of particle filters
-%is that they are a robust method of inference, and I would assert 
-%that the particle filter ought to be given as \emph{raw} data as possible. 
-%While taking direct measurements
-%without de-trending would give awful results, using the difference removes the 
-%DC component and turns what is usually assumed to be a Weiner process into 
-%a simple Gaussian random variable. 
-%
-%\begin{equation}
-%\Delta y = y(t) - y(t-1) = g(x(t)) - g(x(t-1)) + \nu_y(t) - \nu_y(t-1) + \nu_d(t) - \nu(t-1)
-%\label{eq:measass_delta}
-%\end{equation}
-%
-%Even if $\nu_d$ is some other additive process, the difference will still be closer
-%to I.I.D. than a Wiener process, as the autocorrelation of the $\delta y$ shows
-%in \autoref{fig:QQDelta} in \autoref{sec:Introduction Noise}. 
-% All the assumptions made originally
-%for the particle filter still hold, and all of the parameters may be distinguished based on
-%the step sizes, thus it is not unreasonable to consider matching the string of step sizes
-%rather than string of direct readings. 
-%
-%\begin{figure}
-%\label{fig:FrequencyGraphs}
-%\caption{frequency response graphs, highlighting noise frequency range and signal frequency range}
-%\end{figure}
-
 \section{Particle Filter Settings}
@@
 Friston et al. only studied regions deemed active by the General
-Linear Model; and most other research (including \cite{Friston2001}) used these results as 
-the source for their priors \cite{Friston2000}. 
+Linear Model; and most other research used these results as
+the source for their priors \cite{Friston2002}.
 The one exception is to this is Johnston et al. which came to a extremely different
@@
 to use priors that would give results consistent with previous works
-\cite{Friston2000}. This constrains the usefulness of the model to
+\cite{Friston2002}. This constrains the usefulness of the model to
 areas that fall within the prior distribution, yet allowed results
@@
 regions and multiple patients. That said, a significant amount of
-indeterminability exists in the parameters, which is discussed in 
+indeterminateness exists in the parameters, which is discussed in
 \autoref{sec:Single Voxel Simulation}
@@
 processing time will scale up directly with the number of particles.
-A dense initial estimate is important so that some particles land
+A dense initial estimate is important to ensure that some particles land
 near the solution; but as the variance decreases the number of
@@
 of higher probability. Resampling is slow because it requires redrawing
-all the particles. It also closes off avenues of investigation, and is
+all the particles. It also closes off avenues of investigation, but is
 designed to over-smooth to prevent overly thinning the support. For all these
@@
 \label{sec:Particle Deprivation}
-To prevent particle deprivation, which may happen even with a good
+To prevent particle deprivation (all weights dropping to 0), which may happen even with a good
 estimate of the prior distribution, a method of rescuing the particle
-filter from this state was used. When particle deprivation was detected,
+filter from this state was implemented. When particle deprivation was detected,
 resampling was performed with a saved version of the covariance matrix (from
@@
 
-After trial and error, for this work I chose the Gaussian with a standard deviation 
+After trial and error, for this thesis I chose the Gaussian with a standard deviation
 of $0.005$. Although using an adaptive weighting function
@@
 particles, 1400 integration points and a Quad Core CPU). The size of
-integration steps are also crucial, but using step sizes above 
+integration steps are also crucial. Using step sizes above
 $0.001$ seconds is not recommended. In most cases millisecond resolution

### 2010-08-16 doc/thesis/results.tex
FINAL!
spelling fixes etc.

@@
 Although this is only a single test, the correlation (\autoref{tab:long_corr})
-of the parameters 
-indicates significant parameter indeterminability. When the input
+of the parameters increases the probability that the parameters are
+ill-defined. When the input
 consists entirely of impulses, the best parameters are not one particular
@@
 \begin{equation}
-\text{Residual} = \sqrt{\sum (\hat{y}_k - y_k)^2}
+\text{Residual} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - y_k)^2}
 \end{equation}
@@
 \begin{equation}
-\text{Error} = \sqrt{\sum (\hat{y}_k - Y_k)^2}
+\text{Error} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - Y_k)^2}
 \end{equation}

### 2010-08-17 doc/thesis/thesis.tex
fixed typos

@@
 Traditional methods of analyzing FMRI images use a linear combination of
-just a few static regressors. This works demonstrates an alternative
+just a few static regressors. This work demonstrates an alternative
 approach using a physiologically inspired nonlinear model. By using a 
@@
 of the noise in the state
-variables. The activation results show regions similar those found in 
+variables. The activation results show regions similar to those found in 
 SPM; however, there are some notable regions not detected by 
@@
 not just a viable alternative to the traditional method of detecting
-activation, but an extensible technique of estimating the joint Probability
-Distribution Function of the BOLD parameters.
+activation, but an extensible technique of estimating the joint probability
+distribution function of the BOLD parameters.
 

### 2010-08-18 doc/paper/examples/kalman.py
Initialized paper with this document. Will remove stuff from here.

@@ -1,0 +1,86 @@
+#!/usr/bin/python
+
+from scipy.stats.distributions import norm, cauchy
+import scipy
+import numpy 
+import pylab as P
+from math import sqrt
+
+def plus(a,b):
+    return a+b
+
+#time 1
+P.subplot(311)
+xvals = scipy.linspace(-10,10,num=1500)
+measmean = 2
+meas = [norm.pdf(i, measmean, .5) for i in xvals]
+P.plot([-val for val in xvals], meas)
+
+state = [cauchy.pdf(i, -2, 1)/2. +cauchy.pdf(i, 2, 1)/2. for i in xvals]
+P.plot([-val for val in xvals], state)
+
+statew = [var/sum(state) for var in state]
+statemu = numpy.average(xvals, weights=statew)
+statestd = sqrt(sum([pair1[1]*(pair1[0] - statemu)**2 for pair1 in zip(xvals, statew)]))
+statest = [norm.pdf(i, statemu, statestd) for i in xvals]
+P.plot([-val for val in xvals], statest)
+
+combined_mean = (statemu + measmean)/2.
+combined_var = sum([var[0]*(var[1] - combined_mean)**2 for var in zip(statest, xvals)])
+combined_var = combined_var/sum(statest) + sum([var[0]*(var[1] - combined_mean)**2 for var in zip(meas, xvals)])/sum(meas)
+print combined_var/2, sqrt(combined_var/2)
+state = [norm.pdf(i, combined_mean, sqrt(combined_var/2)) for i in xvals]
+P.plot([-val for val in xvals], state)
+P.legend(["Measurement", "State Predict", "State Precit (Norm)", "Updated State"])
+
+#time 2
+P.subplot(312)
+xvals = scipy.linspace(-10,10,num=1500)
+measmean = 2
+meas = [norm.pdf(i, measmean, .5) for i in xvals]
+P.plot([10 - val for val in xvals], meas)
+
+state2 = map(plus, state, [norm.pdf(i, 0, 20) for i in xvals])
+state = state2
+P.plot([10-val for val in xvals], state)
+
+statew = [var/sum(state) for var in state]
+statemu = numpy.average(xvals, weights=statew)
+statestd = sqrt(sum([pair1[1]*(pair1[0] - statemu)**2 for pair1 in zip(xvals, statew)]))
+statest = [norm.pdf(i, statemu, statestd) for i in xvals]
+P.plot([10-val for val in xvals], statest)
+
[…]

### 2010-08-18 doc/paper/paper.tex
got formatting right for paper

@@
-\documentclass{article}
-\usepackage[left=1cm,top=2cm,right=1cm,nohead,nofoot]{geometry}
+%Journal Alticle Layout by Micah Chambers
+%using the IEE bare_jrnl.tex file at 
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
+%% and
+%% http://www.ieee.org/
 
-%\usepackage{times}
-%\usepackage{tikz}
-%\usetikzlibrary{trees}
-%\usetikzlibrary{shapes,snakes}
-\usepackage{verbatim}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{setspace}
-\usepackage{graphicx}
+\documentclass[journal]{./IEEEtran}
 
-%\doublespacing
+\usepackage{cite}
 
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\Q}{\mathbb{Q}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\qed}{\\$\Box$}
-\newcommand{\qle}{\stackrel{?}{\le}}
-\newcommand{\qeq}{\stackrel{?}{=}}
-\newcommand{\closure}{\overline}
-\newcommand{\intersect}{\cap}
-\newcommand{\union}{\cup}
-\newcommand{\nullset}{\emptyset}
-\newcommand{\minus}{\ \backslash\ }
+\usepackage[pdftex]{graphicx}
+\graphicspath{{pdf/}{png/}}
 
-\begin{comment}
+\usepackage[cmex10]{amsmath}
+\interdisplaylinepenalty=2500
 
-:Author: Micah Chambers
-%:Title: Homework 7 for Real Analysis
+\usepackage{algorithmic}
+\usepackage{array}
 
-\end{comment}
+\usepackage[caption=false,font=footnotesize]{subfig}
+\usepackage{fixltx2e}
 
-%\author{Micah Chambers}
-%\title{Homework 8}
+\usepackage[colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
+\usepackage{url}
+\hyphenation{op-tical net-works semi-conduc-tor}
+
+%\begin{comment}
+%:Author: Micah Chambers
+%:Institution: Virginia Tech University
+%\end{comment}
+
+%\newcommand{\gettitle}{\text{Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}}
 
@@
 
-\section*{Introduction to Modeling}
-The short term goal of current neurological models is to
-learn something about the system being modeled.
-While control (for instance in prosthetics), or predication
-(such as BlueBrain) may some day be feasible, in the
-short run we still have a great deal to learn about how our
-brain works. By learning about specific systems (the dynamics
-of a single person's brain), we may be able to diagnose diseases,
-find their causes and even someday rectify the cause. Of course
-by learning about various specific systems we can also learn 
-about the general way our brains work. The reason then, for 
-studying the BOLD model then is
-to be able to derive the various system parameters and understand
-why one brain region (say a voxel) reacts in one way, when another
-acts completely differently.
+\title{Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
 
-While the General Linear Model has been extremely effective at
-gleaning information about which voxels are "active" and which
-are not, its purpose is only to say which regions are "statistically
-significant". However, considering that voxels consist of millions
-of neurons, the reality is that a
-region could be active, but at some other level or time constant
-than expected; and be completely missed by the linear model. 
-Studies showing activation
-in Cadavers of Phantoms highlight the inherant problem with the 
-General Linear Model: a problem that mandates multiple
-comparison testing with very high thresholds to ensure any 
-sort of face validity \citep{drift}.
-By moving to models that actually "learn" the 
-parameters of the model, model error will be drastically
-reduced allowing for much more reasonable statistical thresholds.
+\author{{Micah~Chambers}% <-this % stops a space
+\thanks{M. Chambers is with the Department
+of Electrical and Computer Engineering, Virginia Tech University, Blacksburg,
+VA, 24060 USA e-mail: (micahc@vt.edu}% <-this % stops a space
+\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}
 
-Of course, it is well known that many of the assumptions 
-that the GLM makes about BOLD activation do not hold. 
-The General Linear Model ignores
-the fact that the BOLD response is nonlinear, contains non-gaussian
-noise and most importantly that brain networks are 
-so called "Small World Networks" \citep{smallworld}, \citep{noise}. 
-Merely accounting for
-nonlinearities in the time-series of the BOLD response can
-result in significantly better estimation as shown in \citep{nonlinearmodels}.
-But even more crucial is the recent movements away from mass-
-univariate models. Current mass-univariate models 
-assume uniform connectivity from input to every voxel; yet
-this is obviously not the case. While
-violating linearity assumptions or Gaussian assumptions 
-requires us to be more conservative, the lack of any method
-to switch brain circuits on and off means that any model
-we make for brain regions would not be Turing complete;
-in effect implying that humans are incapable of 
-thought \citep{turing}.
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
+{Chambers: Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
+% The only time the second header will appear is for the odd numbered pages
+% after the title page when using the twoside option.
+% 
+% *** Note that you probably will NOT want to include the author's ***
+% *** name in the headers of peer review papers.                   ***
+% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
+% you desire.
 
-So called Dynamic Causal Modeling is the beginning of 
-the next phase of neurology. DCM is the first 
-brain study to show any significant connection between 
-diffusion tensor imaging and actual function layout in
-the human brain \citep{dti_dcm}. By incorporating connections
-between regions and a realistic activation model, DCM
-corrects two of the largest problems with the GLM. While
-there are other techniques that may be useful in the future,
-they either lack physiological analogs (Artificial Neural
-Networks) or are extremely computationally expensive 
-(mutual information).  Given that much of the 
-potential of the GLM has been exhausted, and that DCM 
-is one of the first well defined methods capable of learning 
-complicated brain circuits DCM is crucial to the future
-of FMRI and our understanding of the human brain.
+\maketitle
 
-\section*{Results}
-The particle filter shows great promise in being able 
-to learn a variety of different regional activation
-parameters. We have performed tests with a random binary
-pulse train as input to a simulated region and then tested
-the particle filter's ability to estimate the signal
-parameters. Additionally random gaussian noise with an
-Signal-To-Noise ratio of 2.0 was added. Figure 1 shows
-the pulse train, the clean signal and the noisy signal, which
-was the input to the particle filter. A random set of 
-system parameters was chosen to simulate the region,
-which the particle filter had no knowledge of other than
-a mean and rough variance.
-The tests were performed using 10,000 particles and took
-approximately 4 minutes to compute.
+\begin{abstract}
+Traditional methods of analyzing FMRI images use a linear combination of
+just a few static regressors. This works demonstrates an alternative
+approach using a physiologically inspired nonlinear model. By using a 
+particle filter to optimize the model parameters, the computation time
+is kept below a minute per voxel without requiring a linearization 
+of the noise in the state
+variables. The activation results show regions similar those found in 
+SPM; however, there are some notable regions not detected by 
+SPM. Though the parameters selected by the particle filter based approach
+are more than sufficient to predict the BOLD response,
+more model constraints are needed to uniquely identify a single set
+of parameters. This ill-posed nature explains the large discrepancies
+found in other research that attempted to characterize the model parameters.
+For this reason the final distribution of parameters is more medically relevant
+than a single estimate. Because the output of the particle filter is 
+a full posterior probability, the reliance on the mean to estimate 
+parameters is unnecessary. This work presents
+not just a viable alternative to the traditional method of detecting
+activation, but an extensible technique of estimating the joint Probability
+Distribution Function of the BOLD parameters.
+\end{abstract}
 
-%figure 1
-\begin{figure}
-\includegraphics[width=8in,height=4in]{noise.png}
-\caption{Simulated time-series for a single region. The red line
-shows the input stimulus, the blue line shows the base BOLD 
-response, and the green line is the BOLD response with Gaussian
-white noise added at SNR of 2}
-\label{fig:noise}
-\end{figure}
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
+\end{IEEEkeywords}
 
-%figure 2
-\begin{figure}
-\includegraphics[width=8in,height=4in]{bold.png}
-\caption{Particle Filter Results. The blue line is the "true"
-(simulated) BOLD response, the red line
-is the output of the particle filter with 2 standard deviations
-in green. Note that the error bar for time 0 is outside the scope
-of the image and is approximately $\pm .003$.}
-\label{fig:bold}
-\end{figure}
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
+%
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
 
-%figure 3
-\begin{figure}
-\includegraphics[width=8in,height=4in]{state.png}
-\caption{Particle Filter Results. The blue line is the "true"
-parameter and the green line is the estimated value for the parameter.
-Note that the timescale for $S_t$ is different to highlight an typical
-stimulus response. Again 2 standard deviations are shown.}
-\label{fig:state}
-\end{figure}
+\section{Introduction}
+% The very first letter is a 2 line initial drop letter followed
+% by the rest of the first word in caps.
+% 
+% form to use if the first word consists of a single letter:
+% \IEEEPARstart{A}{demo} file is ....
+% 
+% form to use if you need the single drop letter followed by
+% normal text (unknown if ever used by IEEE):
+% \IEEEPARstart{A}{}demo file is ....
+% 
+% Some journals put the first two words in caps:
+% \IEEEPARstart{T}{his demo} file is ....
+% 
+% Here we have the typical use of a "T" for an initial drop letter
+% and "HIS" in caps to complete the first word.
+\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
+for IEEE journal papers produced under \LaTeX\ using
+IEEEtran.cls version 1.7 and later.
+% You must have at least 2 lines in the paragraph with the drop letter
+% (should never be an issue)
+I wish you the best of success.
 
-Figure 2 shows the estimated BOLD response versus the simulated
-data (without noise). It is clear that as the estimated BOLD
-response converges to the true BOLD as time progresses. This
-is because more and more potential states are eliminated by system's
-response to new input sequences. Convergence depends highly on 
-the input signal and the weighting function thus choice of both
-are extremely important. We have found that an impulse train
-(emulated with .1 second width square waves) gives a very good
-response. It is also important to choose an input with
-some longer hold times, at least as long as than the 
-expected time constants, which in this case was around 
-8 seconds. The weighting function we have chosen is based
-on the exponential distribution, with a variance equal to
-the RMS of the signal. While it is important to converge
-by the end of the time-series, it is often the case that 
-converging too fast will harm the state estimation by over-
-emphasizing the measurements of single time points. The rate
-of convergence is well within control of the user of the 
-particle filter based on the weighting function and the frequency
-of re-sampling. Note that resampling can cause quantization
-errors due to the nature of estimating a long tailed 
-distribution with a finite number of samples.
+\appendices
+\section{Particle Filter Algorithm}
+Appendix one text goes here.
 
-Figure 3 shows the estimated versus simulated values for
-several key parameters. Notice how the variance drops as
-the particle filter continues. The parameters shown are
-$V_0$ which is a scaling parameter, $\tau_s$, and $\tau_f$
-which are both time constants and $s_t$ which is a hidden
-state variable, estimating the flow inducing signal. Note that
-even though some of the parameters don't converge to the exact
-value, that the estimated $s_t$ still matches the true
-$s_t$ relatively well. It is often the cast that although a
-few parameters don't converge to their true value, one 
-parameter may pick up the slack for another. It is of
-note that $\tau_s$ and $\tau_f$ do converge at least close
-to their true values, because they cannot be determined
-from steady state.
+% you can choose not to have a title for an appendix
+% if you want by leaving the argument blank
+\section{}
+Appendix two text goes here.
 
-\bibliographystyle{abbrvnat}
-\bibliography{proposal_micah}
+% use section* for acknowledgement
+%\section*{Acknowledgment}
+
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,./library}
 

### 2010-08-19 doc/paper/main.tex
added a makefile and edited the paper

--- 2010-08-18 doc/paper/paper.tex
@@
 not just a viable alternative to the traditional method of detecting
-activation, but an extensible technique of estimating the joint Probability
-Distribution Function of the BOLD parameters.
+activation, but an extensible technique of estimating the joint probability
+distribution of the BOLD parameters.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:Introduction}
 % The very first letter is a 2 line initial drop letter followed
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-for IEEE journal papers produced under \LaTeX\ using
-IEEEtran.cls version 1.7 and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
+\IEEEPARstart{T}{raditional} methods of analyzing 
+Functional Magnetic Resonance Imaging (FMRI)
+time series perform regression using a linear 
+combination of static explanatory variables. 
+In this paper I will demonstrate the use of the Particle 
+Filters as a means of estimating the governing parameters of the 
+nonlinear BOLD model at a computation cost 
+that would still allow real time calculations for multiple voxels.
+Practically, this paper presents an alternative
+to detecting neural activity from the traditionally
+used Statistical Parametric Mapping (SPM). Though more computationally intense,
+this method is capable of modeling nonlinear effects, 
+and provides significantly more detailed output. 
+By using a separate particle filter for each single time series it 
+is possible to estimate parameters and make real-time predictions
+for small neural regions, a feature which could be useful towards real time FMRI 
+\cite{DeCharms2005}. Future works will also benefit from the ability to 
+apply conditions to the posterior distribution in post-processing without
+recalculating parameters; for instance to impose additional 
+physiological
+constraints. Modeling the BOLD (Blood-Oxygen-Level-Dependent) response 
+as a nonlinear system is the
+best way to determine the correlation of stimulus sequence with the BOLD
+response; yet in the past doing this on a large scale has been far too
+computationally taxing. The solution used here takes approximately 40 seconds
+for a single 5 minute time series (with Core 2 Duo Q6600).
+
+%This article is organized as follows. The rest of the introduction will
+%overview similar efforts and discuss their strengths ans weaknesses.
+%\autoref{sec:Bold Model} will describe the genesis of the BOLD model,
+%and its current state. \autoref{sec:Particle Filter} will explain the 
+%particle filter and how it is being used in this case. 
+%\auroref{sec:Methods} will describe in detail the experimental
+%design, algorithm set up, and the preprocessing that was applied
+%for the particle filter algorithm.
+%The results are explored separately for simulated data
+%and real FMRI data in \autoref{sec:SimulationResults} and 
+%\autoref{sec:RealData}, respectively. 
+%In \autoref{sec:Discussion} the results and their implications 
+%for future works are interpreted.
 

### 2010-08-20 doc/paper/main.tex
More in the introduction

@@
 combination of static explanatory variables. 
-In this paper I will demonstrate the use of the Particle 
-Filters as a means of estimating the governing parameters of the 
-nonlinear BOLD model at a computation cost 
-that would still allow real time calculations for multiple voxels.
-Practically, this paper presents an alternative
-to detecting neural activity from the traditionally
-used Statistical Parametric Mapping (SPM). Though more computationally intense,
-this method is capable of modeling nonlinear effects, 
-and provides significantly more detailed output. 
-By using a separate particle filter for each single time series it 
-is possible to estimate parameters and make real-time predictions
-for small neural regions, a feature which could be useful towards real time FMRI 
-\cite{DeCharms2005}. Future works will also benefit from the ability to 
-apply conditions to the posterior distribution in post-processing without
-recalculating parameters; for instance to impose additional 
-physiological
-constraints. Modeling the BOLD (Blood-Oxygen-Level-Dependent) response 
-as a nonlinear system is the
-best way to determine the correlation of stimulus sequence with the BOLD
-response; yet in the past doing this on a large scale has been far too
-computationally taxing. The solution used here takes approximately 40 seconds
-for a single 5 minute time series (with Core 2 Duo Q6600).
+However, the static nature of the General Linear Model (GLM)
+limits its potential use. It is well known that different 
+Hemodynamic Response Functions are necessary 
+for different regions of the brain to prevent excessive false
+negatives \cite{Handwerker2004}. Additionally quite a few studies
+have reported spatially varying nonlinearities \cite{Wager2005,Birn2001}.
+Besides not allowing HRF differences between patients, there is no
+reasonable way to incorporate other forms of physiological
+data. Combined FMRI CBF or CBV imaging methods are improving,
+as seen in Chen et al. \cite{Chen2009}. These techniques could shed light on
+neural activation by providing extra measurements, yet a 
+physiologically reasonable model is necessary to incorporate this extra data.
+Activation detection methods also don't have the ability 
+to identify pathologies based on state variables or parameters. For
+example, decreased compliance of blood vessels could indicate, or 
+even cause, a neurological condition that 
+is not easily seen in other imaging modalities.
+
+It is well known that the changes in deoxy-hemoglobin content is the
+primary driver of short term changes in MR signal for FMRI imaging
+techniques \cite{Buxton1998, WEISSKOFF1994, Ogawa, Obata2004}. For
+this reason, the nonlinear state equations that change Deoxy-Hb (DHb) 
+content have been heavily studied, and are well-characterized. Attempts
+at learning the parameters of the BOLD model have also been actively
+studied but significantly less successful. The most wide-spread method
+of calculating the parameters, from two papers by Friston et al., 
+are part of the SPM toolkit; though this algorithm is rarely used 
+as well \cite{Friston2000, Friston2002b}. 
+This method depends on a quadratic convolution (two term Volterra Kernel)
+based estimation of the BOLD output to obtain partials with respect 
+to parameters. Problematically, quality of the Volterra estimate
+is not well known, so it is difficult to quantify error. This method 
+also doesn't account for noise in the parameters or the state variables.
+These shortcomings have led to several other separate attempts. In particular 
+Riera et al. linearized the noise and thus performed regression between
+the model and steps of the BOLD output \cite{Riera2003}. 
+In Vakorin et al., a combination of Genetic Algorithms and Simulated 
+Annealing were used to estimate not only the parameters, but the 
+true stimuli driving the BOLD signal \cite{Vakorin2007}. 
+This addresses the inherent uncertainty of exactly where and when 
+stimuli actually get applied. Unfortunately this algorithm was extremely slow.
+In Johnston et al. a particle filter was applied with the purpose of
+estimating the output distribution, which was then used to optimize
+parameters by means of least squares based techniques \cite{Johnston2008}. 
+Murray used a particle filter, similar to \cite{Johnston2008}, but
+held the model parameters constant; thus all error in the output
+was treated as the result of noise in the underlying parameters 
+\cite{Murray2008}.  Hu et al. used an unscented kalman filter (UKF) 
+was used to optimize parameters \cite{Hu2009}.
+
+\section{Methods}
+\label{sec:Methods}
+\begin{itemize}
+\item BOLD Model
+\item Particle Filter Description 
+\item Description of Simulation
+\item Description of Real FMRI data
+\end{itemize}
+
+\section{Results}
+\label{sec:Results}
+\begin{itemize}
+\item Simulation Results - Possum Only
+\begin{itemize}
+    \item Histogram of parameters for each region
+    \item Final histogram of one run?
+    \item Comparison of MI with SNR
+\end{itemize}
+\item FMRI Results 
+\begin{itemize}
+    \item Activation maps comparing M.I. with SPM
+\end{itemize}
+\end{itemize}
+
+\section{Discussion}
+\label{sec:Discussion}
+
+\begin{itemize}
+\item Parameters ill-defined
+\end{itemize}
+
+\section{Conclusion}
+\label{sec:Conclusion}
 

### 2010-08-20 doc/presentation/main.tex
some changes to presentation

--- 2010-03-01 doc/presentation/thesis_pres.tex
@@
 \documentclass{beamer}
-\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5,debug]{beamerposter}
+\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5]{beamerposter}
 
@@
 {
-  \usetheme{VT}
+  \usetheme{Antibes}
   % or ...
 
-  \setbeamercovered{transparent}
+  %\setbeamercovered{transparent}
   % or whatever (possibly just delete it)
@@
 
-\usefonttheme[onlylarge]{structurebold}
-\usecolortheme{crane}
-\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
-\setbeamertemplate{navigation symbols}{}
-\setbeamercovered{transparent}
-
-\usepackage[english]{babel}
-% or whatever
-
-\usepackage[latin1]{inputenc}
-% or whatever
+%\usefonttheme[onlylarge]{structurebold}
+%\usecolortheme{crane}
+%\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
+%\setbeamertemplate{navigation symbols}{}
+%\setbeamercovered{transparent}
 
@@
 %\usepackage{times}
-\usepackage[T1]{fontenc}
-\usepackage[small]{caption}
+%\usepackage[T1]{fontenc}
+%\usepackage[small]{caption}
 \usepackage{algorithmic}
 
-\title{BOLD Parameter Estimation using Sequential Monte Carlo Methods}
+\title{Full Brain Blood-Oxygen-Level-Dependent Signal \\
+Parameter Estimation using Sequential Monte Carlo Methods}
 
@@
 \author{Micah Chambers}
-
-\institute {Virginia Tech Bioimaging Systems Lab}
+\institute{Virginia Tech Bioimaging Systems Lab}
 
@@
 
-\section{FMRI Review}
+\section{Background}
 \begin{frame}{The BOLD Response}

### 2010-08-21 doc/presentation/main.tex
Work on the first few slides

@@
 \documentclass{beamer}
-\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5]{beamerposter}
+%\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5]{beamerposter} 
+%\setbeameroption{show notes on second screen}
 
@@
 
-\section{Background}
-\begin{frame}{The BOLD Response}
+\section{Introduction}
+\begin{frame}{Balloon Model}
 \begin{figure}
@@
 \end{figure}
+\note{Rates range from below 1 Hz to 5 to 6 Hundred Hz}
+\note{Change in average firing rates increase metabolism and sympathetic response causes
+        increased inflow of oxygenated blood.}
+\note{In the version of the Balloon model I used, the flow is locked to metabolism.}
+\note{According the balloon model this increases the local blood volume, and the outflow pressure.}
+\note{So the signal is altered by changes in the ratio of Deoxyhemoglobin and Oxygenated Hemoglobin,
+        which are a function of Volume and Oxygenation.}
+\end{frame}
+
+\begin{frame}{Equations}
+  \begin{itemize}
+    \item Flow Inducing Signal
+    $$\dot{s}(t) = \epsilon u(t) - s(t)/\tau_s - (f(t)/\tau_f - 1)$$
+    \item Normalized Cerebral Blood Inflow:
+    $$\dot{f}(t) = s $$
+    \item Normalized Cerebral Blood Volume:
+    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
+    \item Normalized Deoxyhaemoglobin Content:
+    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
+            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
+    \item Hemodynamic Response - BOLD Signal
+    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
+  \end{itemize}
+  \note{System is dissipative:}
+  \note{If no input is provided, it will rather quickly decay.}
+  \note{Constant input leads to steady state response.}
+  \note{First term in $\dot{q}$ is the metabolism term}
+  \note{Outflow is the $v^{1/\alpha}$ term
+  \note{Significant nonlinearity.}
+\end{frame}
+
+\begin{frame}{Solutions: Approximate Method}
+  \begin{itemize}
+    \item Canonical Hemodynamic Response Function
+    \note{Linear Approximation}
+    \begin{itemize}
+        \item Maximum Likelihood Possible
+        \item Inflexible - even to changes in onset time
+    \end{itemize}
+
+    \item Volterra Kernel 
+    \note{Quadratic Approximation}
+    \begin{itemize}
+        \item Quadratic Convolution used to approximate Jacobian Matrix.
+        \note{Because of the nonlinearities in the BOLD model, step size for
+                integrating would otherwise be extremely small}
+        \item Volterra approximation quality is not known.
+        \note{Its possible this isn't even know-able}
+    \end{itemize}
+    \note{Volterra Approximation depends on sparsity etc}
+  \end{itemize}
+    \begin{center}
+    \includegraphics[width=10cm,height=4cm]{HRF}
+    \end{center}
 \end{frame}
@@
 \section{Nonlinear Regression}
-\begin{frame}{Equations}
-  \begin{itemize}
-    \item Normalized Cerebral Blood Flow:
-    $$\ddot{f}(t) = \epsilon u(t) - \dot{f}(t)/\tau_s - (f(t)/\tau_f - 1)$$
-    \item Normalized Cerebral Blood Volume:
-    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
-    \item Normalized Deoxyhaemoglobin Content:
-    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
-            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
-    \item Hemodynamic Response - BOLD Signal
-    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
-  \end{itemize}
-\end{frame}
 

### 2010-08-21 doc/presentation/main.tex
finished the friston methods

@@
 
-\usepackage{graphics}
+\usepackage[pdftex]{graphicx}
+\graphicspath{{pdf/}{png/}}
 %\usepackage{times}
@@
 
-\pgfdeclareimage[width=1.5cm]{university-logo}{logo}
-\logo{\pgfuseimage{university-logo}}
+%\pgfdeclareimage[width=1.5cm]{university-logo}{logo}
+\logo{\includegraphics[width=1.5cm]{logo}}
 
@@
 \begin{figure}
-\includegraphics[scale=.23]{model}
+\includegraphics[width=10cm]{model}
 \caption{
@@
 
-\begin{frame}{Equations}
+\begin{frame}{BOLD State Equations}
   \begin{itemize}
@@
     $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
-    \item Normalized Deoxyhaemoglobin Content:
-    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
+    \item Normalized Deoxyhaemoglobin Content:\\
+    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1-(1-E_0)^{1/f(t)})}{E_0} -
             \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
@@
   \note{Constant input leads to steady state response.}
-  \note{First term in $\dot{q}$ is the metabolism term}
-  \note{Outflow is the $v^{1/\alpha}$ term
+  \note{First term in $\dot{q}$ is the metabolism term.}
+  \note{Outflow is the $v^{1/\alpha}$ term}
   \note{Significant nonlinearity.}
@@
   \begin{itemize}
-    \item Canonical Hemodynamic Response Function
-    \note{Linear Approximation}
-    \begin{itemize}
-        \item Maximum Likelihood Possible
-        \item Inflexible - even to changes in onset time
-    \end{itemize}
-
-    \item Volterra Kernel 
+    \item 2nd Order Volterra Kernel 
     \note{Quadratic Approximation}
@@
         \item Volterra approximation quality is not known.
-        \note{Its possible this isn't even know-able}
     \end{itemize}
     \note{Volterra Approximation depends on sparsity etc}
+    {\footnotesize
+    $$y(t) = k_0 + \int_{-\infty}^{\infty} k_1(s_1) x(t-s_1) ds_1
+        + \int_{-\infty}^{\infty} k_2(s_1,s_2) x(t-s_1)x(t-s_2) ds_1 ds_2$$
+        \note{Its possible this isn't even knowable}
+    }
+    \item Canonical Hemodynamic Response Function
+    \note{Linear Approximation}
+    \begin{itemize}
+        \item No Parameters Estimated
+        \item Maximum Likelihood Possible
+        \item Inflexible - even to changes in onset time
     \end{itemize}
     \begin{center}
-    \includegraphics[width=10cm,height=4cm]{HRF}
+    \includegraphics[clip=true,trim=4.88cm 0cm 5cm 1.5cm,height=2.7cm,width=8cm]{HRF}
     \end{center}
-\end{frame}
 
-\begin{frame}{BOLD Signal Properties}
-  \begin{itemize}
-    \item Exact variables and parameters are unknown and are
-        difficult to calculate.
-    \item Significant Amount of Lag between activation
-        and a measurable output - can be as much as 8 seconds.
-    \item Slow Temporal Resolution
-    \item Noise characterized by brownian motion, which clashes with low 
-        frequency elements.
-    \begin{center}
-    \includegraphics[width=10cm,height=4cm]{signal_fft.pdf}
-    \end{center}
   \end{itemize}
 \end{frame}
-
-\begin{frame}{Preprocessing}
-  \begin{itemize}
-    \item Low Pass Filter (Gaussian Filter, not recommended)
-    \item Drift Removal (not always performed)
-    \begin{itemize}
-        \item High Pass Filter
-        \item Linear 
-        \item Quadratic
-        \item Wavelet 
-        \item Spline (Which I am using)
-    \end{itemize}
-  \end{itemize}
-  \begin{figure}
-    \includegraphics[width=10cm]{detrending_effect.png}
-    \caption{
-        \tiny
-        \cite{detrending}
-    }
-  \end{figure}
-\end{frame}
-
-\section{Statistical Parametric Mapping}
-\begin{frame}{Method}
-  \begin{figure}
-    \includegraphics[width=11cm]{glm_pipeline.png}
-    \caption{
-        \tiny
-        \cite{spm_pipeline}
-    }
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Limitations}
-\begin{columns}
-  \column{2.5in}
-  \begin{itemize}
-    \item Linear, for a signal which is known to be nonlinear
-    \item Essentially the weighted sum of a set of "expected" responses.
-    \item Parametric
-    \begin{itemize}
-        \item Forced to make assumptions about underlying distributions
-        \item No time-scaling.
-    \end{itemize}
-  \end{itemize}
-  \column{3in}
-  \begin{figure}
-    \includegraphics[width=2.5in]{glm.png}
-    \caption{
-        \tiny
-        \cite{spm_pipeline}
-    }
-  \end{figure}
-\end{columns}
-\end{frame}
-
-\section{Nonlinear Regression}
-
-\begin{frame}{Model Comparison}
-  \begin{figure}
-    \includegraphics[width=5in,height=2.4in]{comparison.png}
-    \caption{
-        \tiny
-        \cite{comparison}
-    }
-  \end{figure}
-\end{frame}
-
-\section{Parameter Identification}
-\begin{frame}{Particle Filters}
-\begin{itemize}
-    \item Non-parametric, no assumptions are violated
-    \item Model based, fit parameters to input, constrained by physical variables
-    \item Fits a mixture PDF to the posterior of all parameters
-    \item Non-trivial computation cost
-    \item I use a Regularized Particle Filter
-    \begin{enumerate}
-        \item Regularized Re-sampling prevents particles from de-generating into a 
-            small number of unique particles
-        \item Allows distributions to move more freely
-    \end{enumerate}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Particle Filter}
-    \begin{columns}
-    \column{1.8in}
-    \begin{itemize}
-        \item $S_t = \{p_{0,t}, ... , p_{N,t}\}$, the set of particles
-        \item $w_{i,t}$, weight of particle $p_{i,t}$
-        \item $y_t$, measurement at time $t$, there is not a $y_t$ for every $t$.
-        \item $f(p_{i,t},y_t)$, weighting function 
-        \item $s(p_{i,t})$, step function 
-    \end{itemize}
-    \column{3.2in}
-    \begin{algorithmic}
-     \STATE Draw $S_0$ from prior distribution 
-     \FOR{$t = 0:t_{step}:t_{end}$}
-        \FOR{each $p_{i,t-1} \in S_{t-1}$}
-           \STATE $p_{i,t} = s(p_{i,t-1})$ 
-           \IF{There is a measurement at time $t$}
-               \FOR{every $p_{i,t}$}
-                    \STATE $w_{i,t} = w_{i,t-1}f(p_{i,t},y_t)$
-               \ENDFOR 
-               \STATE Resample if weights are unevenly distributed\\
-           \ENDIF
-        \ENDFOR
-     \ENDFOR
-%                    \item If a weights are not evenly distributed enough, then resample
-     \end{algorithmic}
-     \end{columns}
-\end{frame}
-
-\subsection{Single-Region}
-\begin{frame}{Single Timeseries Results}
-  \begin{figure}
-    \includegraphics[height=3in]{noise.png}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Single Timeseries Results, Measurement Convergence}
-  \begin{figure}
-    \includegraphics[height=3in]{bold.png}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Single Timeseries Results, State Convergence}
-  \begin{figure}
-    \includegraphics[height=3in]{state.png}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Factors Affecting Convergence}
-    \begin{enumerate}
-        \item Weighting function
-            \begin{itemize}
-                \item Needs to be continuous and defined for any input, should go to 0
-                \item Too wide a weighting function results in under-sensitivity, 
-                            slow or no convergence
-                \item Too thin a weighting function reduces robustness to noise
-            \end{itemize}
-        \item How often re-sampling is done, re-sampling should be minimized
-            \begin{itemize}
-                \item Stratified Resampling can result in truncated tails on posterior
-                \item Regularized Resampling can result in reduced robustness to noise 
-            \end{itemize}
-        \item Number of particles
-            \begin{itemize}
-                \item More particles give higher fidelity of posterior
-            \end{itemize}
-    \end{enumerate}
-\end{frame}
-
-\subsection{Multi-Region}
-\begin{frame}{Parameter Map Generation/Simulation}
-\begin{columns}
-  \column{2.5in}
-  \begin{enumerate}
-    \item Generate a parameter map, with a set of parameters for each voxel
-    \item Simulate every set of parameters, and use as input to possum
-    \item Perform preprocessing (de-trend and normalize)
-    \item Run particle filter on every grey matter voxel in image, generating a new
-            parameter map
-  \end{enumerate}
-  \column{3in}
-  \begin{figure}
-    \includegraphics[width=2.8in]{equation.png}
-    \caption{
-    }
-  \end{figure}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Simulation Results, $\tau_0$}
-  \begin{figure}
-    \includegraphics[height=3in]{results.png}
-    \caption{Percent Difference from "Actual"}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Simulation Results, multiple}
-  \begin{figure}
-    \includegraphics[height=3in]{results2.png}
-    \caption{Percent Difference from "Actual"}
-  \end{figure}
-\end{frame}
-
-\section{Conclusion}
-
-% All of the following is optional and typically not needed. 
-\appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{For Further Reading}
+%
+%\begin{frame}{BOLD Signal Properties}
+%  \begin{itemize}
+%    \item Exact variables and parameters are unknown and are
+%        difficult to calculate.
+%    \item Significant Amount of Lag between activation
+%        and a measurable output - can be as much as 8 seconds.
+%    \item Slow Temporal Resolution
+%    \item Noise characterized by brownian motion, which clashes with low 
+%        frequency elements.
+%    \begin{center}
+%    \includegraphics[width=10cm,height=4cm]{signal_fft.pdf}
+%    \end{center}
+%  \end{itemize}
+%\end{frame}
+%
+%\begin{frame}{Preprocessing}
+%  \begin{itemize}
+%    \item Low Pass Filter (Gaussian Filter, not recommended)
+%    \item Drift Removal (not always performed)
+%    \begin{itemize}
+%        \item High Pass Filter
+%        \item Linear 
+%        \item Quadratic
+%        \item Wavelet 
+%        \item Spline (Which I am using)
+%    \end{itemize}
+%  \end{itemize}
+%  \begin{figure}
+%    \includegraphics[width=10cm]{detrending_effect.png}
+%    \caption{
+%        \tiny
+%        \cite{detrending}
+%    }
+%  \end{figure}
+%\end{frame}
+%
+%\section{Statistical Parametric Mapping}
+%\begin{frame}{Method}
+%  \begin{figure}
+%    \includegraphics[width=11cm]{glm_pipeline.png}
+%    \caption{
+%        \tiny
+%        \cite{spm_pipeline}
+%    }
+%  \end{figure}
+%\end{frame}
+%
+%\begin{frame}{Limitations}
+%\begin{columns}
+%  \column{2.5in}
+%  \begin{itemize}
+%    \item Linear, for a signal which is known to be nonlinear
+%    \item Essentially the weighted sum of a set of "expected" responses.
+%    \item Parametric
+%    \begin{itemize}
+%        \item Forced to make assumptions about underlying distributions
+%        \item No time-scaling.
+%    \end{itemize}
+%  \end{itemize}
+%  \column{3in}
+%  \begin{figure}
+%    \includegraphics[width=2.5in]{glm.png}
+%    \caption{
+%        \tiny
+%        \cite{spm_pipeline}
+%    }
+%  \end{figure}
+%\end{columns}
+%\end{frame}
+%
+%\section{Nonlinear Regression}
+%
+%\begin{frame}{Model Comparison}
+%  \begin{figure}
+%    \includegraphics[width=5in,height=2.4in]{comparison.png}
+%    \caption{
+%        \tiny
+%        \cite{comparison}
+%    }
+%  \end{figure}
+%\end{frame}
+%
+%\section{Parameter Identification}
+%\begin{frame}{Particle Filters}
+%\begin{itemize}
+%    \item Non-parametric, no assumptions are violated
+%    \item Model based, fit parameters to input, constrained by physical variables
+%    \item Fits a mixture PDF to the posterior of all parameters
+%    \item Non-trivial computation cost
+%    \item I use a Regularized Particle Filter
+%    \begin{enumerate}
+%        \item Regularized Re-sampling prevents particles from de-generating into a 
+%            small number of unique particles
+%        \item Allows distributions to move more freely
+%    \end{enumerate}
+%\end{itemize}
+%\end{frame}
+%
+%\begin{frame}{Particle Filter}
+%    \begin{columns}
+%    \column{1.8in}
+%    \begin{itemize}
+%        \item $S_t = \{p_{0,t}, ... , p_{N,t}\}$, the set of particles
+%        \item $w_{i,t}$, weight of particle $p_{i,t}$
+%        \item $y_t$, measurement at time $t$, there is not a $y_t$ for every $t$.
+%        \item $f(p_{i,t},y_t)$, weighting function 
+%        \item $s(p_{i,t})$, step function 
+%    \end{itemize}
+%    \column{3.2in}
+%    \begin{algorithmic}
+%     \STATE Draw $S_0$ from prior distribution 
+%     \FOR{$t = 0:t_{step}:t_{end}$}
+%        \FOR{each $p_{i,t-1} \in S_{t-1}$}
+%           \STATE $p_{i,t} = s(p_{i,t-1})$ 
+%           \IF{There is a measurement at time $t$}
+%               \FOR{every $p_{i,t}$}
+%                    \STATE $w_{i,t} = w_{i,t-1}f(p_{i,t},y_t)$
+%               \ENDFOR 
+%               \STATE Resample if weights are unevenly distributed\\
+%           \ENDIF
+%        \ENDFOR
+%     \ENDFOR
+%%                    \item If a weights are not evenly distributed enough, then resample
+%     \end{algorithmic}
+%     \end{columns}
+%\end{frame}
+%
+%\subsection{Single-Region}
+%\begin{frame}{Single Timeseries Results}
+%  \begin{figure}
+%    \includegraphics[height=3in]{noise.png}
+%  \end{figure}
+%\end{frame}
+%
+%\begin{frame}{Single Timeseries Results, Measurement Convergence}
+%  \begin{figure}
+%    \includegraphics[height=3in]{bold.png}
+%  \end{figure}
+%\end{frame}
+%
+%\begin{frame}{Single Timeseries Results, State Convergence}
+%  \begin{figure}
+%    \includegraphics[height=3in]{state.png}
+%  \end{figure}
+%\end{frame}
+%
+%\begin{frame}{Factors Affecting Convergence}
+%    \begin{enumerate}
+%        \item Weighting function
+%            \begin{itemize}
+%                \item Needs to be continuous and defined for any input, should go to 0
+%                \item Too wide a weighting function results in under-sensitivity, 
+%                            slow or no convergence
+%                \item Too thin a weighting function reduces robustness to noise
+%            \end{itemize}
+%        \item How often re-sampling is done, re-sampling should be minimized
+%            \begin{itemize}
+%                \item Stratified Resampling can result in truncated tails on posterior
+%                \item Regularized Resampling can result in reduced robustness to noise 
+%            \end{itemize}
+%        \item Number of particles
+%            \begin{itemize}
+%                \item More particles give higher fidelity of posterior
+%            \end{itemize}
+%    \end{enumerate}
+%\end{frame}
+%
+%\subsection{Multi-Region}
+%\begin{frame}{Parameter Map Generation/Simulation}
+%\begin{columns}
+%  \column{2.5in}
+%  \begin{enumerate}
+%    \item Generate a parameter map, with a set of parameters for each voxel
+%    \item Simulate every set of parameters, and use as input to possum
+%    \item Perform preprocessing (de-trend and normalize)
+%    \item Run particle filter on every grey matter voxel in image, generating a new
+%            parameter map
+%  \end{enumerate}
+%  \column{3in}
+%  \begin{figure}
+%    \includegraphics[width=2.8in]{equation.png}
+%    \caption{
+%    }
+%  \end{figure}
+%\end{columns}
+%\end{frame}
+%
+%\begin{frame}{Simulation Results, $\tau_0$}
+%  \begin{figure}
+%    \includegraphics[height=3in]{results.png}
+%    \caption{Percent Difference from "Actual"}
+%  \end{figure}
+%\end{frame}
+%
+%\begin{frame}{Simulation Results, multiple}
+%  \begin{figure}
+%    \includegraphics[height=3in]{results2.png}
+%    \caption{Percent Difference from "Actual"}
+%  \end{figure}
+%\end{frame}
+%
+%\section{Conclusion}
+%
+%% All of the following is optional and typically not needed. 
+%\appendix
+%\section<presentation>*{\appendixname}
+%\subsection<presentation>*{For Further Reading}
 

### 2010-08-21 doc/presentation/run.sh
got multihead working and added a script to start the show

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+impressive main.pdf

### 2010-08-22 doc/paper/main.tex
added bold section

@@
 also doesn't account for noise in the parameters or the state variables.
+
 These shortcomings have led to several other separate attempts. In particular 
 Riera et al. linearized the noise and thus performed regression between
-the model and steps of the BOLD output \cite{Riera2003}. 
+the model and steps of the BOLD output \cite{Riera2003}. This aproach
+had the benefit of a Jacobian,  but at the same time removes all DC 
+signal. For this reason, the type of activation limits the
+algorithms' ability to learn the model.
 In Vakorin et al., a combination of Genetic Algorithms and Simulated 
@@
 This addresses the inherent uncertainty of exactly where and when 
-stimuli actually get applied. Unfortunately this algorithm was extremely slow.
-In Johnston et al. a particle filter was applied with the purpose of
-estimating the output distribution, which was then used to optimize
-parameters by means of least squares based techniques \cite{Johnston2008}. 
+stimuli actually get applied. Unfortunately this algorithm was extremely
+slow; taking hours or days per single time course.
+In Johnston et al. alternating estimates of $P(X_t | \Theta, Y_t)$ and 
+$P(\Theta | Y_t)$ were calculated, to maximize $P(X_t, \Theta | Y_t)$. 
+$P(X_t | \Theta, Y_t)$ was calculated using a particle filter 
+\cite{Johnston2008}. This however was quite slow, and the results
+were inconsistent with other similar works. 
 Murray used a particle filter, similar to \cite{Johnston2008}, but
@@
 was treated as the result of noise in the underlying parameters 
-\cite{Murray2008}.  Hu et al. used an unscented kalman filter (UKF) 
-was used to optimize parameters \cite{Hu2009}.
+\cite{Murray2008}. Murray's results showed that differences in
+BOLD output cannot be well explained by error in the state variables.
+Hu et al. used an unscented kalman filter (UKF) to estimate parameters,
+and filtered parameter sets inconsistent with the output  \cite{Hu2009}.
+ Furthermore, Birn et al. and Yacoub et al. showed that many
+of the characteristics of the BOLD output depend on the region 
+of interest \cite{Birn2001, Yacoub2006}. For this reason, a single
+Hemodynamic Response Function is not likely to accurately estimate
+activation across the full brain.
+
+With the exception of Hu et al., the works described thus far all
+worked to estimate a single set of parameters that could
+explain the output \cite{Hu2009}. Yet Deneux et al showed that the very similar
+outputs can be achieved with very different parameters \cite{Deneux2006}.
+This explains the inconsistencies in parameters across the different
+studies. Thus, estimating parameters using a least squares
+framework is not likely to be fruitful; and though the posterior
+distributions will vary spatially, that difference will not be
+discernible from a single parameter estimate. The problem with 
+using an Unscented Kalman Filter to approximate the distributions
+is that with a nonlinear, dissipative system, significant non-Gaussian
+effects will be present. In tests, significant non-Gaussian effects 
+began to appear within a second of state integration. With FMRI
+repetition times being well above 1 second, a Gaussian approximation
+will result in significant error. The particle filter, which can be
+thought of an extension of UKF, uses non-parametric distributions
+rather than a Gaussian. For this reason the particle filter, described
+in \autoref{sec:ParticleFilters}, is used in this work.
 
@@
 \label{sec:Methods}
-\begin{itemize}
-\item BOLD Model
-\item Particle Filter Description 
-\item Description of Simulation
-\item Description of Real FMRI data
-\end{itemize}
+\subsection{Bold Model}
+\label{sec:BoldModel}
+In the past fifteen years, a steady stream of studies have built
+on the original Blood Oxygen Level Dependent (BOLD) signal 
+derivation first described by Ogawa et al. \cite{Ogawa}.
+The seminal work by Buxton et al. attempted to explain the
+time evolution of the BOLD signal using a windkessel model to
+describe the local changes in Deoxygenated Hemoglobin content \cite{Buxton1998}.
+Incremental improvements were made to this model until Friston et al.
+brought all the changes together into a single complete 
+set of equations \cite{Friston2000}. And while there have been numerous adaptations in the model, 
+many of them summarized by Deneux et al., even the basic versions
+have less bias error than the empirically driven 
+\emph{Canonical Hemodynamic Model} \cite{Deneux2006,Handwerker2004}.
+On the other hand BOLD signal models have numbers
+of parameters ranging from seven \cite{Riera2003} to 50 \cite{Behzadi2005} 
+for a signal as short as 100 samples long. This number of parameters presents
+a significant risk of being under-determined and having high computation cost. 
+In this work, only the simplest physiologically inspired model will be
+used (with 7 parameters), and steps will be taken to make the most of computation
+time.
+
+It is well known that the two types of hemoglobin act as a contrast agents in 
+EPI imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
+between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
+Intuitively, increased metabolism will increase Deoxyhemoglobin, 
+however blood vessels are quick to compensate by increasing local 
+blood flow. Increased inflow, accomplished by loosening 
+capillary beds, precedes increased outflow, driving increased 
+blood storage.  Since the local MR signal depends on the ratio of 
+Deoxyhemoglobin to Oxygenated Hemoglobin, increased blood volume 
+affects this ratio if metabolism doesn't exactly match the increased 
+inflow of oxygenated blood.  This was the impetus
+for the ground breaking balloon model \cite{Buxton1998} and windkessel
+model \cite{Mandeville1999}. These works derive, from first principals,
+the changes in deoxyhemoglobin ratio and volume of capillaries given an 
+inflow waveform.  These were the first two attempts to quantitatively 
+account for the shape of the BOLD signal as a consequence of the lag 
+between the cerebral blood volume (CBV) and the inward cerebral blood 
+flow (CBF).
+
+Although Buxton et al. demonstrated that a well chosen flow waveform could 
+explain most features of the BOLD signal, it stopped short of proposing a
+realistic waveform for the CBF and CMRO2 \cite{Buxton1998}. Friston et al. 
+gave a reasonable and simple expression for CBF input based on a flow 
+inducing signal and in the same work proposed a simple method
+of estimating metabolic rate: as a linear function of the inward 
+blood flow \cite{Friston2000}. By combining these equations with 
+the balloon model from Buxton et al., it is possible to predict 
+the BOLD signal directly from a stimulus time course:
+\begin{eqnarray}
+\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
+\dot{f} &=& s\\
+\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
+\label{eq:bold}
+\end{eqnarray}
+where $s$ is a flow inducing signal, $f$ is the input blood flow (CBF),
+$v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
+local deoxyhemoglobin. The parameters controlling blood flow are 
+$\epsilon$, which is a neuronal efficiency term, $u(t)$ which is 
+the stimulus, and $\tau_f$, $\tau_s$ which are time constants. 
+The parameters for the evolution of blood volume are $E_0$ which 
+the resting metabolic rate and $\alpha$ which is Grubb's parameter 
+controlling the balloon model.  $\tau_0$ is a single time constant 
+controlling the speed of $v$ and $q$.
+
+This completed balloon model was assembled and analyzed
+by Riera et al. \cite{Riera2003}. Obata refined the readout equation 
+of the BOLD signal based on the deoxyhemoglobin content ($q$) and local 
+blood volume ($v$), resulting in the final BOLD equation \cite{Obata2004}.
+\begin{eqnarray}
+y   &=& V_0((k_1 + k_2)(1-q) - (k_2 + k_3)(1-v))\\
+k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8 \nonumber \\
+K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57 \nonumber \\
+k_3 &=& \epsilon_0 - 1 = .43 \nonumber
+\label{eq:boldout}
+\end{eqnarray}
+Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
+de-oxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
+change in relaxation rate with change in blood oxygenation, and
+$\epsilon_0 = 1.43$ is the ratio of signal MR from intravascular to 
+extravascular regions at rest. Although, these constants change with 
+experiment ($TE$, $\nu_0$, $r_0$), patient, and brain 
+region ($E_0$, $r_0$), often the estimated values by Obata are 
+taken as the constants $a_1 = k_1 + k_2 = 3.4$, and $a_2 = k_2+k_3 = 1$ in 
+studies using 1.5 Tesla scanners \cite{Obata2004}.
+Additional compartments and parameters have been added since 
+Riera et al., however Deneux et al. showed that only the viscoelastic
+effects of Buxton et al. were necessary to model the primary 
+elements of the BOLD response \cite{Riera2003, Deneux2006, Buxton2004}.
+Despite this, only the basic balloon model of Buxton et al. is
+used in this work, although future works may benefit from constraining
+the parameters as described by Deneux et al. \cite{Buxton2004, Deneux2006}.
+
+\subsection{Particle Filter}
+\label{sec:ParticleFilter}
+\subsection{Simulation}
+\label{sec:MethodsSim}
+\subsection{Data Acquisition}
+\label{sec:MethodsData}
 

### 2010-08-22 doc/thesis/background.tex
small fixes, typos, and a strangely copied line

@@
 %\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption{Example updates of a distribution using the Kalman Filter, \cite{Thrun2005}}
+\caption{Example updates of a distribution using the Kalman Filter.}
 \label{fig:EKFWorking}
@@
 used to estimate not only the parameters, but the true stimuli driving the BOLD
-signal. This addresses the inherent uncertainty of exactly where and when 
+signal \cite{Vakorin2007}. 
+This addresses the inherent uncertainty of exactly where and when 
 stimuli actually get applied. Unfortunately this algorithm was extremely slow.

### 2010-08-22 doc/thesis/particlefilters.tex
FINAL!
small fixes, typos, and a strangely copied line

@@
 $\Theta$ is an update function, from now on $X$ will encapsulate
-$\Theta$. The assumptions now allow for a simplified version of the
+$\Theta$. The assumptions allow for a simplified version of the
 state space equations:
@@
 \begin{eqnarray}
-\dot{X}_k = f(X_{k-1}, u_k, \nu_x)
+\dot{X}_k &=& f(X_{k-1}, u_k, \nu_x)
 \label{eq:stateass}\\
-Y_k = g(X_k, \nu_y)
+Y_k &= &g(X_k, \nu_y)
 \label{eq:measass}
@@
 \begin{algorithmic}
-\STATE Calculate total weight, $W = \sum_{i=0}^{N_p} w^i$
+\STATE Calculate total weight, $W_t = \sum_{i=0}^{N_p} w^i$
 \FORALL{$0 < i < N_p$}
@@
  \end{algorithmic}
+ \label{alg:RegularizedParticleFilter}
  \end{algorithm}
@@
 With regularized resampling, it is possible to prevent both
-particle degeneracy as well as particle impoverishment. Additionally,
-as the particle filter converges, the density of particles
-in the area of the solution goes up. This has a similar effect to 
-simulated annealing where, as the algorithm approaches the end, the
-random steps get smaller and smaller. Note however, that resampling
-carries certain risks.  If for some reason the solution is not covered by the 
+particle degeneracy as well as particle impoverishment. 
+Note that resampling carries certain risks.  
+If for some reason the solution is not covered by the 
 new support, the algorithm may not be able to reach the true value. 
-The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
-or a genetic algorithm. Versions of $x$ that are fit (give good measurements) spawn more children 
+The ultimate effect of this regularized resampling is a convergence 
+similar to simulated annealing or a genetic algorithm. 
+Versions of $x$ that are fit (give good measurements) spawn more children 
 nearby which allow for more accurate estimation near points of high likelihood. 
-As the variance of the estimated
-$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
+As the variance of the estimated $x$'s decrease, the radius in 
+which children are spawned also decreases. Eventually the radius
 will approach the width of the underlying uncertainty.

### 2010-08-22 doc/paper/main.tex
removed stuff from thesis that doesn't really belong and moved methods through the particle filter stuff.

@@
 
+\usepackage{algorithm}
 \usepackage{algorithmic}
@@
 experiment ($TE$, $\nu_0$, $r_0$), patient, and brain 
-region ($E_0$, $r_0$), often the estimated values by Obata are 
+region ($E_0$, $r_0$), often the estimated values by Obata et al. are 
 taken as the constants $a_1 = k_1 + k_2 = 3.4$, and $a_2 = k_2+k_3 = 1$ in 
@@
 \subsection{Particle Filter}
-\label{sec:ParticleFilter}
+\label{sec:ParticleFilters}
+Particle filters, a type of Sequential Monte Carlo (SMC) method,
+are a powerful method for estimating the posterior probability distribution
+ of parameters given a time series of measurements. Unlike Markov 
+Chain Monte Carlo (MCMC) estimation, particle filters are designed for 
+time-varying random variables. Unlike the UFK, sample points are kept
+and continuously propagated forward, rather than being used to calculate
+the first two moments. Thus, the posterior distribution is
+stored as an empircal distribution made up of sample points.
+Particle filters are preferred when the model is nonlinear, 
+and non-Gaussian. This section based primarily on 
+Arulampalam et al. and Thrun et al. \cite{Arulampalam2002a, Thrun2005}.
+
+The idea of the particle filter is to build an empirical distribution
+out of a large number of parameter sets, called particles. Each
+particle contains all the parameters and states needed to propagate
+the model forward.  The particle filter begins with a wide distribution 
+(called the Prior Distribution)
+of possible particles and then, as measurements come in, weights 
+particles based on the quality of their output estimates. Thus parameter sets 
+that tend to give good estimations of the measurements get weighted higher
+than parameter sets that give poor estimates. Although the reliance on
+a prior distribution could be problematic, when the system being modeled
+has physical meaning, establishing reasonable ranges for parameters may be 
+quite easy. Optimizing the prior distribution can be more difficult,
+unless the system has been extensively studied.
+
+\subsubsection{Assumptions}
+Suppose a set or stream of measurements at discrete times are given, 
+$\{Y_k, k = 1, 2, 3, ... K\}$, where $K$ may be infinite. 
+Because $k$ is a discrete time, let $t_k$ define the continuous
+time of $k$.
+Suppose also that there is a hidden set of state variables,
+$X(t)$ that drives the value of $Y(t)$. Throughout this section
+with $X_k = X(t_k)$. The goal of the particle filter is to estimate the 
+distribution of the
+true parameters $\Theta$ that dictates the movement of $X(t)$.
+The model also permits random motion in $X(t)$, so the 
+particle filter also estimates the distribution of $X(t)$.
+The only difference between the members of parameter vector
+$\Theta$ and those of $X(t)$ is that the memebers of
+$\Theta$ have no known update equation. Members of both vectors
+are permitted to have some noise, although this
+may not be explicitly stated in the model. The generic, continuous, nonlinear
+system definition is shown in \autoref{eq:GenericNonlinear}.
+
+\begin{eqnarray}
+\dot{X}(t) = f(t, X(t), u(t), \theta, \nu_x) \nonumber \\
+Y(t) = g(t, X(t), u(t), \theta, \nu_y)
+\label{eq:GenericNonlinear}
+\end{eqnarray}
+
+$X(t)$ is vector of state variables, $\Theta$ is a vector of system
+constants, $u(t)$ is an input, $Y(t)$ the observation, and
+$\nu_x$ and $\nu_y$ are random variates. Although any of these
+variables could be a vector, for the sake of simplicity only
+$\Theta$ and $X(t)$ will be considered as such.
+
+I will also make a few  simplifying assumptions for this work. 
+First, the systems are assumed to be 
+time invariant. This 
+assumption is based on the idea that if you paused the system for $\Delta t$
+seconds, when unfrozen the system would continue as if nothing happened. 
+Few biological systems are predictable enough for them to be summarized
+by a time varying function, least of all the brain. While heart beats are certainly
+periodic and have an effect on the BOLD signal, the period varies too much
+for the system to be considered varying with time. 
+Next, its assumes that input cannot directly
+influence the output, which in the case of the BOLD signal is a good assumption.
+Finally, because the only difference between the members of $X(t)$ and 
+$\Theta$ is an update function, from now on $X$ will encapsulate
+$\Theta$. The assumptions allow for a simplified version of the
+state space equations:
+
+\begin{eqnarray}
+\dot{X}_k &=& f(X_{k-1}, u_k, \nu_x)
+\label{eq:stateass}\\
+Y_k &= &g(X_k, \nu_y)
+\label{eq:measass}
+\end{eqnarray}
+
+\subsubsection{Prior Distribution}
+The goal of the particle filter is to evolve an empirical distribution 
+$P(x_k | u_{0:k}, Y_{0:k})$,
+that asymptotically approaches the true probability distribution $P(X_k | u_{0:k})$.
+Note that capital $X$ will be used as the actual realizations of 
+the state variable, whereas $x$ will denote estimates of $X$.
+Additionally, the notation $a:b$ indicates the set $[a,b]$,
+as in $u_{a:b}$, which would indicate all the inputs from time $a$ to time $b$.
+Considering the noise present in $X$,
+ $P(X_k | u_{0:k})$ is not a single true value but probability distribution.
+
+To begin with, the particle filter must be given a prior distribution, from
+which the initial $N_p$ particles are drawn. A particle contains a weight
+as well as an estimate of $X_k$, which as previously mentioned, contains every
+variable needed to run the model. Then the prior is generated from a 
+given distribution, $\alpha(X)$, by:
+
+\begin{equation}
+\{(x^i_0,w^i) : x^i_0 \sim \alpha(X), w^i = 
+            \frac{1}{N_p} : i \in \{1, 2, ... , N_p\} \}
+\end{equation}
+
+Where $N_p$ is the number of particles or points used to 
+describe the prior using a Mixture PDF.  Note that any 
+exponents will be explicitly labeled as such, to avoid 
+confusion with the particle numbering scheme.
+
+Therefore, as $N_p \rightarrow \infty$:
+\begin{equation}
+\alpha(X) \approx P(x_0) = \sum_{i=0}^{N_p} w^i\delta(X - x^i_0 ) dx
+\end{equation}
+Where $\delta(x-x_0)$ is 1 if and only if $x = x_0$
+(the Kronecker delta function).
+$\alpha(X)$, the prior distribution, \emph{must} cover the 
+true posterior distribution in order for the particle filter
+to converge to the true posterior distribution.
+
+\subsubsection{Basic Particle Filter}
+For all the following areas, the probabilities implicitly depend on $u_{0:k}$, 
+so they are left off for simplicity.
+Whenever a measurement becomes available it permits refinement of the
+posterior density.  Incorporating new data is called sequential 
+importance sampling. The weight of a particle in an empirical distribution
+is defined as
+\begin{equation}
+w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}
+\label{eq:weightfunc}
+\end{equation}
+where $q$ is called an \emph{importance density}. The importance density
+is the density of the particles, thus by dividing by this value, the weight
+should not depend on the location of the estimation points, but rather
+only on $P(x^i_{0:k} | y_{0:k})$, the probability of that particle
+being correct given all the measurements up to time $k$. 
+Note that if there is a far off peak in
+the posterior that $q$ does not have support points in, there will 
+be quantization errors, and that part of the density cannot be modeled. 
+This is why $q$ must fully cover $P(x^i_{0:k} | y_{0:k})$.
+
+To update the weight of a particular particle at time $k$, it is necessary to 
+calculate both $q(x^i_{0:k} | y_{0:k})$ and $P(x^i_{0:k} | y_{0:k})$, 
+in \autoref{eq:Support} and \autoref{eq:UpdateBayes} respectively.
+
+Note that $q(x^i_{0:k} | y_{0:k})$ may be simplified by assuming that 
+$y_k$ doesn't contain any additional information about $x_{k-1}$. 
+\begin{equation}
+q(x^i_{0:k} | y_{0:k}) = q(x^i_{0:k} | y_{0:k-1})
+\label{eq:QAssump}
+\end{equation}
+The choice of the importance density, $q(x_k | x_{k-1}, y_{0:k})$ is 
+a design decision; however it is common to use the state model: 
+\begin{equation}
+q(x_k | x_{k-1}, y_{0:k}) =  P(x_k | x_{k-1})
+\label{eq:ImportanceDensity}
+\end{equation}
+This choice for importance density is helpful because 
+$P(x_k | x_{k-1})$ is freely available: its simply the set 
+of particles propagated forward in time using the state 
+equations. Additionally it makes
+updating weights simple, as seen later (\autoref{eq:weightevolve}).
+$q(x^i_{0:k} | y_{0:k})$ may then be simplified:
+\begin{equation}
+q(x_{0:k} | y_{0:k}) = P(x_k | x_{k-1})q(x_{0:k-1} | y_{0:k-1})  
+\label{eq:Support}
+\end{equation}
+
+Calculating $P(x_{0:k} | y_{0:k})$ 
+uses the assumption that the distribution of $y_k$ is 
+fully constrained by $x_k$, and that $x_k$ is similarly fully 
+constrained by $x_{k-1}$, I make the good assumptions that:
+\begin{eqnarray}
+P(y_k | x_{0:k}, y_{0:k-1}) &=& P(y_k | x_k) \nonumber \\
+P(x_k | x_{0:k}, y_{0:k-1}) &=& P(x_k | x_{k-1})
+\label{eq:MarkovProperty}
+\end{eqnarray}
+These are restatements of the state equations assumed by \autoref{eq:stateass}
+and \autoref{eq:measass}.
+Additionally, for the particle filter $y_k$ and $y_{0:k-1}$ are 
+constant across all particles, thus $P(y_k| y_{0:k-1})$ can
+be dropped when the equality is changed to a proportion. 
+Using these properties and Bayes' Theorem, $P(x^i_{0:k} | y_{0:k})$ 
+may be broken up as follows:
+\begin{equation}
+P(x_{0:k} | y_{0:k}) \propto P(y_k| x_k) P(x_k | x_{k-1}) P(x_{0:k-1} | y_{0:k-1}) 
+ \label{eq:UpdateBayes}
+\end{equation}
+
+Plugging \autoref{eq:ImportanceDensity} and the result of \autoref{eq:UpdateBayes}
+into \autoref{eq:weightfunc} leads to:
+\begin{eqnarray}
+w^i_k & \propto & \frac{P(y_k| x^i_k)  P(x^i_{0:k-1} | y_{0:k-1})}
+                         {q(x^i_{0:k-1} | y_{0:k-1})} \nonumber \\
+& \propto & w^i_{k-1}P(y_k| x_k) 
+\label{eq:weightevolve}
+\end{eqnarray}
+
+Thus, by making the following simple assumptions, evolving a posterior
+density  requires no knowledge of noise distribution.
+\begin{enumerate}
+\item $f(t, x(t), u(t)) = f(x(t), u(t))$ and $g(t, x(t), u(t)) = g(x(t))$ 
+\item The prior distribution PDF, $q(x_i(0))$, covers $P(x_i(0))$
+\item Markov Property: $P(x_k | x_{0:k-1}) = Pr(x_k | x_{k-1})$
+\item $q(x_{0:k-1} | y_{0:k}) = q(x_{0:k-1} | y_{0:k-1})$
+\end{enumerate}
+
+From the definition of $w_i$, the algorithm to calculate
+an approximation of $P(X(t_k) | Y_{0:k})$ or $P(X(t_k + \delta t) | Y_{0:k})$
+is simple (see Algorithm \autoref{alg:BasicParticleFilter}).
+
+\subsubsection{Sequential Importance Resampling}
+\label{sec:ParticleFilterResampling}
+As a consequence of the wide prior distribution 
+(required for a proper discretization of a continuous
+distribution), a significant proportion of particles quickly
+develop insignificant weights. 
+Ideally the computation time spent on tails would be proportional to 
+the actual size of the tails.  The case where a large number of the 
+weights have become extremely small
+is called particle degeneracy. In  Lui et al. 
+an ideal measurement of degeneracy based on the particles' true 
+weight is given \cite{Liu1998b}. An approximation of 
+the effective number of particles is also given: 
+\begin{equation}
+N_{eff} \approx \frac{\sum_{i=0}^{N_p} w_i}{\sum_{i=0}^{N_p} w_i^2}
+\label{eq:neff}
+\end{equation}
+To alleviate particle degeneracy, a technique known as resampling 
+may be applied. The idea of resampling is to 
+draw from the approximate posterior, thus generating 
+a replica of the posterior probability with 
+a matching support. Each particle in the new distribution,
+$\hat{x}_j$, is drawn from the old distribution 
+(see Algorithm \autoref{alg:Resampling}):
+\begin{equation}
+\hat{x}_j \sim \left(\sum_{i=0}^{N_p} w^i_k\delta(x - x^i_k)\right)
+\end{equation}
+
+For infinite particles this new distribution will match the old.
+Unfortunately, in practice since the support is
+limited to the original particles, the number of \emph{unique} 
+particles can only decrease.
+This effect, called particle impoverishment, can result in 
+excessive quantization errors in the final distribution. 
+To prevent particle impoverishment, instead of sampling from the
+discrete distribution, a smoothing kernel is applied.
+Because the distribution is continuous, particle impoverishment
+is prevented. This regularization process is defined as:
+\begin{equation}
+x_i = x_i + h\sigma \epsilon
+\end{equation}
+Where $h$ is an optional bandwidth, $\sigma$ is the standard 
+deviation such that $\sigma \sigma^T = cov(x)$
+and $\epsilon$ is drawn from the chosen kernel. In this case
+a Gaussian kernel was used, and the bandwidth was set to 1
+(see \autoref{alg:RegResampling}). All the 
+pseudocode may be found in \autoref{sec:algorithms},
+including the complete regularized particle filter used
+(Algorithm \autoref{alg:RegularizedParticleFilter}).
+
 \subsection{Simulation}
@@
 \appendices
+
 \section{Particle Filter Algorithm}
-Appendix one text goes here.
+\label{sec:algorithms}
+
+\subsection{Sequential Importance Sampling}
+\label{alg:BasicParticleFilter}
+\begin{algorithm}
+\begin{algorithmic}
+\STATE Initialize Particles:
+\FOR{$i$ : each of $N_p$ particles }
+    \STATE $x^i_0  \sim \alpha(X)$
+    \STATE $w^i_0 = \frac{1}{N_p}$
+\ENDFOR
+\FOR{$k$ : each measurement}
+    \FOR{$i$ : each particle }
+        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
+    \ENDFOR
+\ENDFOR
+\begin{equation*}
+\begin{array}{l}
+P(x(t_k+\Delta t)) \approx \\
+\sum_{i=0}^{N_p} w^i_k \delta\left(x - x^i_k - \int_{t_k}^{t_k+\Delta t} f(x(\tau), u(\tau)) d\tau \right) 
+\end{array}
+\end{equation*}
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Resampling Algorithm}
+\label{alg:RegResampling}
+\begin{algorithm}
+\begin{algorithmic}
+\STATE Calculate total weight, $W_t = \sum_{i=0}^{N_p} w^i$
+\FORALL{$0 < i < N_p$}
+    \STATE Draw $V$ from uniform range $[0, W]$
+    \STATE $C = W_t$
+    \FORALL{$0 < j < N_p$ and $C < V$}
+        \STATE $C = C - w^j$
+    \ENDFOR
+    \STATE Add $[x^j, \frac{1}{N_p}]$ to the new distribution
+\ENDFOR
+\STATE 
+\end{algorithmic}
+\label{alg:Resampling}
+\end{algorithm}
+
+\subsection{Regularized Resampling Algorithm}
+\label{alg:RegResampler}
+\begin{algorithm}
+\begin{algorithmic}
+\STATE Calculate Covariance, $C$, of empirical distribution, $\hat{x}$
+\STATE Find $D$ such that $DD^T = C$
+\STATE Resample $\hat{x}$ using algorithm from \autoref{alg:Resampling}
+\FOR{$0 < i < N_p$}
+    \STATE Draw $\epsilon$ from the standard normal, same dimensionality as $X$
+    \STATE $x^i = x^i + hD\epsilon$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Regularized Particle Filter}
+\label{alg:RegularizedParticleFilter}
+\begin{algorithm}
+\begin{algorithmic}
+\STATE Initialize Particles:
+\FOR{$i$ : each of $N_p$ particles }
+    \STATE $x^i_0  \sim \alpha(X)$
+    \STATE $w^i_0 = \frac{1}{N_p}$
+\ENDFOR
+\FOR{$k$ : each measurement}
+    \FOR{$i$ : each particle }
+        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
+    \ENDFOR
+
+    \STATE Calculate $N_{eff}$ with \autoref{eq:neff}
+    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
+        \STATE Resample using algorithm from \autoref{alg:RegResampling}
+    \ENDIF
+\ENDFOR
+
+\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+ \end{algorithmic}
+ \end{algorithm}
 

### 2010-08-23 doc/paper/main.tex
added simulation stuff

@@
 activation, but an extensible technique of estimating the joint probability
-distribution of the BOLD parameters.
+distribution function of the BOLD parameters.
 \end{abstract}
@@
                          {q(x^i_{0:k-1} | y_{0:k-1})} \nonumber \\
-& \propto & w^i_{k-1}P(y_k| x_k) 
+& \propto & w^i_{k-1}P(y_k| x^i_k) 
 \label{eq:weightevolve}
@@
 (Algorithm \autoref{alg:RegularizedParticleFilter}).
+\begin{table}[t]
+\centering
+\begin{tabular}{|c || c | c | c |}
+\hline
+Parameter & Distribution & $\mu$ & $\sigma$ \\
+\hline
+$\tau_0$ & Gamma & .98 & .25 \\
+$\alpha$ & Gamma & .33 & .045\\
+$E_0$    & Gamma & .34 & .03  \\
+$V_0$    & Gamma & .04 & .03 \\
+$\tau_s$ & Gamma & 1.54  & .25\\
+$\tau_f$ & Gamma & 2.46  & .25\\
+$\epsilon$ & Gamma & .7  & .6 \\
+\hline
+\end{tabular}
+\caption{Prior distributions used in the particle filter.}
+\label{tab:Prior}
+\end{table}
+\autoref{eq:weightevolve} still depends on knowing
+$P(y_k| x^i_k)$. Ideally this would exactly match the
+measurement error distribution of the system; however,
+even if this distribution is incorrect, the solution will
+still converge. For this work it was assumed that
+\begin{equation}
+P(y_k | x^i_k) = g(X_k, \nu_y) = g(X_k) + \nu_y\text{, }\nu_y \sim N(0, 0.005^2)
+\end{equation}
+A standard deviation of $0.005$ performed well in the general case, primarily
+because the BOLD signal tends to be in the range of $0.01-0.05$, (1\%-5\%).
+Additionally, 1400 Euler integration points were used per second, 
+the system was assumed to start at rest. The system was initialized
+with 16,000 particles, but during the first resampling, the number
+of particles was dropped to 1,000. To prevent over-smoothing, resampling
+was only performed when $\N_{eff}$ (\autoref{eq:neff}) dropped below
+$50$ for two consecutive measurements.
+
+\subsection{Preprocessing}
+The BOLD response has been
+extensively studied and despite minor discrepancies, the cause of the BOLD
+signal is well known. However, FMRI has an extremely low SNR. 
+A particularly difficult form of noise present 
+in FMRI is a low frequency
+drift, often characterized as a Wiener process \cite{Riera2004}.
+Though not present in all regions, as many as ten to fifteen percent
+of voxels can be affected, thus it is prevalent 
+enough to cause significant inference problems \cite{Tanabe2002, Smith2007}. 
+Though the cause is still not
+clear it is clear that this drift signal is not solely
+due to a physiological effects, given its presence in cadavers and phantoms
+\cite{Smith1999}. The problem mandates the use of a high pass 
+filter \cite{Smith2007}.  In a comparison of high pass filters,
+Tanabe et al.  showed that generally subtracting off
+a spline improved the SNR the most \cite{Tanabe2002}.
+
+To prevent transients at the beginning of the FMRI run, the first 9
+measurements (18.9 seconds) were dropped.
+One knot was used for every 20 time points (at a TR of 2.1), with the first
+knot and last knot taking half as many points as the rest of the knots.
+The median level of each group of measurements was calculated, and the
+knot placed at the average time of the group. After subtracting this spline
+from the timeseries, the resulting timeseries was divided by the original
+sample mean. To compensate for the BOLD signal not generally being zero-mean,
+the Median-Absolute-Deviation was added to the final signal (\autoref{eq:mad}). 
+\begin{equation}
+y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
+\label{eq:mad}
+\end{equation}
 
@@
 \label{sec:MethodsSim}
+Single voxel simulations were performed by propagating the BOLD
+model forward through time with 2000 integration points per
+second, then adding I.I.D. Gaussian noise and a Wiener process
+with I.I.D. steps. 
+Full volume simulation were performed using 
+a modified version of the FSL tool POSSUM. The modifications allowed
+activation levels to be driven by spatially dependent BOLD parameter
+sets. Due to the modifications, the SNR configuration was set to 20,
+although actual SNR's ranged from .5 to 1.5 in active areas.
+
 \subsection{Data Acquisition}
@@
 \label{sec:Results}
+\subsection{Simulation}
+The final
+histogram for a single active region, with a single set of true 
+parameters is shown in \autoref{fig:slicesim_hist2}.
+
+%parameters:
+%&1.151&  0.353& 0.380& 0.026& 1.98&  2.333& 1.645 \\
+\begin{figure*}[!t] %top left
+\centering
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
+\caption{Histogram of estimated parameters. Average SNR in the region was 
+$0.97$ and Mutual Information between simulation and BOLD estimate was greater 
+than $0.15$. TODO: add actual error.}
+\label{fig:slicesim_hist2}
+\end{figure*}
+
 \begin{itemize}
-\item Simulation Results - Possum Only
-\begin{itemize}
-    \item Histogram of parameters for each region
-    \item Final histogram of one run?
-    \item Comparison of MI with SNR
-\end{itemize}
-\item FMRI Results 
-\begin{itemize}
-    \item Activation maps comparing M.I. with SPM
-\end{itemize}
-\end{itemize}
+\item Long Single Result: Correlation Matrix, error in estimate
+\item Possum Results: Histogram of parameters for a region
+\item FMRI Results: Activation maps comparing M.I. with SPM
 

### 2010-08-23 doc/paper/main.tex
updated paper

@@
 
+\usepackage{colortbl}
 \usepackage{algorithm}
@@
 
-%\newcommand{\gettitle}{\text{Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}}
-
 \begin{document}
@@
 
+%TODO change manuscript recieved/revised
 \author{{Micah~Chambers}% <-this % stops a space
@@
 
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
+%TODO change journal name
+\markboth{Journal of ...}%
 {Chambers: Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
@@
 To prevent transients at the beginning of the FMRI run, the first 9
-measurements (18.9 seconds) were dropped.
+measurements (18.9 seconds) were dropped. Each of the remaining 
+images were registered to the new first image, to account for
+movement. 
 One knot was used for every 20 time points (at a TR of 2.1), with the first
@@
 \end{equation}
+
+For SPM, the first 9 measurements were again dropped, and the images
+co-registered across time. The co-registered data was spatially smoothed,
+and then filtered with an adaptive cut off (which is built into the
+SPM analysis). For the SPM analysis, the Canonical HRF was used
+and model time derivatives were included (but not in the contrast
+vector). All other settings used in SPM8 were left at the default
+for analyzing FMRI data.
 
@@
 \label{sec:MethodsData}
+For the FMRI data discussed in \autoref{sec:RealData}, tests were
+performed on a right handed volunteer using a GE SIGNA HDx 1.5 Tesla
+scanner with a single echo EPI sequence. Slice spacing was
+5mm, and pixel sizes were 3.75mm.  Repetition Time was 2.1s,
+Echo Time was 40 ms and the imaging frequency was set to $63.854$MHz.
+The image resolution was $64 \times 64 \times 28$.
+The subject was presented with either a single or
+double flash and was asked to respond with a right handed
+or left handed finger tap, respectively. The FMRI began 18.9 seconds
+before the beginning of the experiment, to allow for transients
+in to image to settle out. The timing of the flashes are shown
+below with time 0 corresponding to the beginning of
+the 10th TR (so 9 images were dropped).
+
+The timing of the single flashes were:
+1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350,
+61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063,
+90.169, 96.994, 110.644, 117.469, 120.881, 130.988, 132.694, 154.875,
+158.288, 161.700, 165.113, 168.525, 176.925, 178.631, 183.750, 190.575,
+204.225, 205.931, 211.050, 216.038, 222.863, 226.275, 236.513, 248.456,
+255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394,
+299.381
+
+Timing of the double flashes were:
+0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756,
+91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000,
+134.400, 136.106, 141.225, 144.638, 156.581, 159.994, 166.819, 171.806,
+175.219, 185.456, 188.869, 202.519, 212.756, 217.744, 221.156, 227.981,
+229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506,
+272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500
 
@@
 \subsection{Simulation}
-The final
-histogram for a single active region, with a single set of true 
-parameters is shown in \autoref{fig:slicesim_hist2}.
+\begin{table*}[t]
+\centering
+\begin{tabular}{|c | c  c  c  c  c  c  |}
+\hline
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ \\
+\hline
+$\alpha$                      & 0.889884 & & & & & \\
+\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & \\
+$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & \\
+$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & \\
+\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 \\
+\hline
+\end{tabular}
+\caption{Correlation of parameter estimates for a 2400 second simulation.}
+\label{tab:long_corr}
+\end{table*}
+
+For a 2400 second simulated FMRI run, a signal was generated using the 
+following parameters:
+$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$,
+with measurement noise ($\sigma_y$) of $0.001$ and drift standard deviation
+($\sigma_x$) of $0.0005$. The minimum and maximum error particles had MSE of $ $ and $ $,
+respectively, while the mean of the particles gave MSE of $ $. %TODO fill in errors
+The correlation matrix for the final parameter estimates is shown in \autoref{tab:long_corr}.
+The significant correlation of parameters indicate that no one estimate of 
+particles dominate the solution. Additionally, given the very low error in the BOLD,
+the energy graph is full of global minima. The final histogram for a single active region, 
+with a single set of true parameters is shown in \autoref{fig:slicesim_hist2}.
+This also indicates that ill-defined nature of the parameters.
 
@@
 
+\begin{figure}[!t]
+\subfloat[Region labels for simulated slice.]
+{\includegraphics[width=2.5in]{images/simregions.png}\label{fig:simslice_mask}}
+\hfil
+\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
+{\label{fig:sim_hm_mi} \includegraphics[width=2.5in]{images/sim_hm_mi}}
+\caption{Comparison of activation with greymatter, parameter regions.}
+\label{fig:simslice_hm}
+\end{figure}
+
 \begin{itemize}
@@
 \item FMRI Results: Activation maps comparing M.I. with SPM
+\end{itemize}
 

### 2010-08-24 doc/paper/main.tex
more changes to methods/results

@@
 \label{sec:MethodsSim}
+\begin{table}[t]
+\centering
+\begin{tabular}{|c |c | c | c | c | c | c | c |}
+\hline
+Region & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  \\
+\hline
+1 & 1.454& 0.321& 0.369& 0.036& 0.994& 2.774& 1.348\\
+2 &1.151&  0.353& 0.380& 0.026& 1.98&  2.333& 1.645 \\
+3 &1.951&  0.317& 0.348& 0.027& 1.657& 3.719& 0.757 \\
+4 &1.203 & 0.310& 0.326& 0.036& 2.168& 2.272& 0.086\\
+\hline
+\end{tabular}
+\caption{Actual parameters for each regions in the POSSUM simulated slice.
+Regions shown in \autoref{fig:simslice_mask}.}
+\label{tab:simsliceparams}
+\end{table}
+
 Single voxel simulations were performed by propagating the BOLD
@@
 second, then adding I.I.D. Gaussian noise and a Wiener process
-with I.I.D. steps. 
-Full volume simulation were performed using 
+with I.I.D. steps. For all single voxel simulations,
+the signal was generated using the following parameters:
+$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, 
+\tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$,
+with measurement noise ($\sigma_y$) of $0.001$ and drift standard deviation
+($\sigma_x$) of $0.0005$. Two separate tests are discussed in the
+results. A single voxel test was performed with for 2400 seconds with random
+stimuli (Poisson distributed, $\lambda = 1$ event per second) lasting 
+$0.1$ seconds. The second single voxel tests discussed were the 310.8
+seconds after preprocessing, and used stimulus matching 
+\autoref{sec:MethodsData}.
+
+Full volume simulations were performed using 
 a modified version of the FSL tool POSSUM. The modifications allowed
@@
 
-For a 2400 second simulated FMRI run, a signal was generated using the 
-following parameters:
-$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$,
-with measurement noise ($\sigma_y$) of $0.001$ and drift standard deviation
-($\sigma_x$) of $0.0005$. The minimum and maximum error particles had MSE of $ $ and $ $,
-respectively, while the mean of the particles gave MSE of $ $. %TODO fill in errors
-The correlation matrix for the final parameter estimates is shown in \autoref{tab:long_corr}.
+%\begin{table}[t]
+%\centering
+%\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
+%\hline
+%$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
+%\hline
+%\rowcolor[gray]{.8}
+%1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+%\hline
+%\hline
+%1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
+%1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
+%1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
+%1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
+%1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
+%1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
+%1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
+%1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
+%1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
+%1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
+%1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
+%\hline
+%1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
+%\hline
+%\end{tabular}
+%\caption{Estimated Parameters on 11 separate 310.8 second runs. 
+%    True value in first row,
+%    and average in last. Residual measured as $\sqrt{MSE}$ with noisy signal,
+%    Error measured as $\sqrt{MSE}$ with true (noiseless) signal.}
+%\label{tab:LowNoiseResults}
+%\end{table}
+
+The minimum and maximum error particles had MSE of $ $ and $ $,
+respectively, while the mean of the particles gave MSE of $ $. 
+%TODO fill in errors
+The correlation matrix for the final parameter estimates is shown in 
+\autoref{tab:long_corr}.
 The significant correlation of parameters indicate that no one estimate of 
-particles dominate the solution. Additionally, given the very low error in the BOLD,
-the energy graph is full of global minima. The final histogram for a single active region, 
+particles dominate the solution. Additionally, given the very low error in 
+the BOLD,
+the energy graph is full of global minima.
+
+The final histogram for a single active region, 
 with a single set of true parameters is shown in \autoref{fig:slicesim_hist2}.

### 2010-08-24 doc/paper/main.tex
simulation results about done, need to change the activation graph.

@@
 \begin{equation}
-y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
+y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - 
+        \text{median}(y_{0:K}))
 \label{eq:mad}
@@
 The timing of the single flashes were:
-1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350,
+\small{1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350, 
 61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063, 
@@
 255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394, 
-299.381
+299.381}
 
 Timing of the double flashes were:
-0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756,
+\small{ 0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756, 
 91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000, 
@@
 229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506, 
-272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500
+272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500 }
 
@@
 
-%\begin{table}[t]
-%\centering
-%\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
-%\hline
-%$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
-%\hline
-%\rowcolor[gray]{.8}
-%1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
-%\hline
-%\hline
-%1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
-%1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
-%1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
-%1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
-%1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
-%1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
-%1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
-%1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
-%1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
-%1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
-%1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
-%\hline
-%1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
-%\hline
-%\end{tabular}
-%\caption{Estimated Parameters on 11 separate 310.8 second runs. 
-%    True value in first row,
-%    and average in last. Residual measured as $\sqrt{MSE}$ with noisy signal,
-%    Error measured as $\sqrt{MSE}$ with true (noiseless) signal.}
-%\label{tab:LowNoiseResults}
-%\end{table}
+%TODO double check the error measurement, add min/max/mean errors for the runs
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\centering
+\begin{tabular}{|c | c | c | c | c | c | c | }
+\hline
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline
+\rowcolor[gray]{.8}
+1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  \\
+\hline
+\hline
+1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 \\
+1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 \\
+1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 \\
+1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 \\
+1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 \\
+1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 \\
+1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 \\
+1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 \\
+1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 \\
+1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 \\
+1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 \\
+\hline
+1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 \\
+\hline
+\end{tabular}
+\caption{Estimated Parameters on 11 separate 310.8 second runs. 
+    True value in first row,
+    and average in last.}
+\label{tab:LowNoiseResults}
+\end{table}
 
-The minimum and maximum error particles had MSE of $ $ and $ $,
-respectively, while the mean of the particles gave MSE of $ $. 
 %TODO fill in errors
-The correlation matrix for the final parameter estimates is shown in 
-\autoref{tab:long_corr}.
-The significant correlation of parameters indicate that no one estimate of 
-particles dominate the solution. Additionally, given the very low error in 
-the BOLD,
-the energy graph is full of global minima.
+The final square root Mean-Squared-Error (between the estimated and 
+noiseless BOLD signal) for the 2400s run was %TODO
+and the final correlation matrix is shown in \autoref{tab:long_corr}.
 
-The final histogram for a single active region, 
-with a single set of true parameters is shown in \autoref{fig:slicesim_hist2}.
-This also indicates that ill-defined nature of the parameters.
+The final parameter estimates of the 310.8s runs are shown in
+\autoref{tab:LowNoiseResults}; the square-root MSE for these
+parameters ranged from $0.002833$ to $0.003345$, with a mean of 
+$0.003066$. Peak signals reached levels of $0.025$.
 
-%parameters:
-%&1.151&  0.353& 0.380& 0.026& 1.98&  2.333& 1.645 \\
+\begin{figure}[!t]
+\subfloat[Region labels for simulated slice.]
+{\includegraphics[width=3cm]{images/simregions.png}\label{fig:simslice_mask}}
+\hfil
+\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
+{\label{fig:sim_hm_mi} \includegraphics[width=4cm]{images/sim_hm_15_6.png}}
+\subfigure{\includegraphics[width=.6cm]{images/scale6.png}}
+\caption{Comparison of activation with greymatter, parameter regions.}
+\label{fig:simslice_hm}
+\end{figure}
+
+%todo fix these histograms with marker for true value and drop A1,A2
 \begin{figure*}[!t] %top left
@@
 $0.97$ and Mutual Information between simulation and BOLD estimate was greater 
-than $0.15$. TODO: add actual error.}
+than $0.15$.}
 \label{fig:slicesim_hist2}
@@
 
-\begin{figure}[!t]
-\subfloat[Region labels for simulated slice.]
-{\includegraphics[width=2.5in]{images/simregions.png}\label{fig:simslice_mask}}
-\hfil
-\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
-{\label{fig:sim_hm_mi} \includegraphics[width=2.5in]{images/sim_hm_mi}}
-\caption{Comparison of activation with greymatter, parameter regions.}
-\label{fig:simslice_hm}
-\end{figure}
+\begin{figure*}[!t] %top left
+\centering
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
+\caption{Histogram of estimated parameters for region 3. Average SNR in the 
+region was 
+$0.39$ and Mutual Information between simulation and BOLD estimate was greater 
+than $0.15$.}
+\label{fig:slicesim_hist3}
+\end{figure*}
 
-\begin{itemize}
-\item Long Single Result: Correlation Matrix, error in estimate
-\item Possum Results: Histogram of parameters for a region
-\item FMRI Results: Activation maps comparing M.I. with SPM
-\end{itemize}
+In the POSSUM simulation, areas passing a threshold of
+mutual information of $0.15$  are shown in 
+\autoref{fig:sim_hm_mi}
+whereas the active region numbers are shown in 
+\autoref{fig:simslice_mask}.
+Note that region 4 had a very small $\epsilon$ and so
+had an extremely low $SNR$ ($<.1$). The histograms of parameter
+estimates for sections
+2 and 3 are shown in \autoref{fig:slicesim_hist2} and 
+\autoref{fig:slicesim_hist3}, respectively. Region 1 is
+left out because of the low number of points and because
+its SNR was between region 2 and region 3 ($0.8$), so it
+was a less interesting result.
+
+\subsection{Real Data}
+\begin{figure*}
+\centering
+\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
+\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
+\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
+\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
+\subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
+\caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher
+        assurance that the signal cannot have occurred through noise alone.
+        Sagittal, coronal and axial slices  are
+        \autoref{fig:hm_canon_spm_x}, \autoref{fig:hm_canon_spm_y}, and
+         \autoref{fig:hm_canon_spm_z}, respectively. A series of axial slices are
+         shown in \autoref{fig:hm_spm}. }
+\label{fig:hm_canon_spm}
+\end{figure*}
 

### 2010-08-24 doc/paper/main.tex
added correct error range for the simulations

@@
 
-%TODO double check the error measurement, add min/max/mean errors for the runs
 \begin{table}[!t]
@@
 
-%TODO fill in errors
 The final square root Mean-Squared-Error (between the estimated and 
-noiseless BOLD signal) for the 2400s run was %TODO
+noiseless BOLD signal) for the 2400s run was $0.00381$, signal
+peaks were $0.035$
 and the final correlation matrix is shown in \autoref{tab:long_corr}.
@@
 \autoref{tab:LowNoiseResults}; the square-root MSE for these
-parameters ranged from $0.002833$ to $0.003345$, with a mean of 
-$0.003066$. Peak signals reached levels of $0.025$.
+parameters ranged from $0.00147$ to $0.0027$, with a mean of 
+$0.00217$. Peak signals reached levels of $0.025$.
 

### 2010-08-24 doc/thesis/results.tex
fixed the error measurements in result
fixed some of the images, and removed redunt ones for POSSUM results.

@@
 \hline
-1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
-1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
-1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
-1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
-1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
-1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
-1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
-1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
-1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
-1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
-1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
+1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
+1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
+1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
+1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
+1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
+1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
+1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
+1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
+1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
+1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
+1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
 \hline                                                                               
-1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
+1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
 \hline
@@
 will be referencing these values.
-
-%\begin{table}[t]
-%\centering
-%\begin{tabular}{|c | c  c  c  c  c  c  c |}
-%\hline
-%  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-%\hline
-%\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
-%$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
-%\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
-%$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
-%\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
-%$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
-%\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
-%\hline
-%\end{tabular}
-%\caption{Typical Covariance matrix of the parameters at the end of a run.}
-%\label{tab:CovSim}
-%\end{table}
 
@@
 \hline
-1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 &0.01573   \\
-0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 &0.01378   \\
-1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 &0.01577   \\
-1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 &0.01154    \\
-1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 &0.01222   \\
-0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 &0.01316   \\
+1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 & 0.00859\\
+0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 & 0.00735\\
+1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 & 0.00951\\
+1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 & 0.00479 \\
+1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 & 0.00475\\
+0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 & 0.00672\\
 \rowcolor[rgb]{.9,.5,.5}
-1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 &0.01790   \\
+1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 & 0.01039\\
 \rowcolor[rgb]{.5,.5,.9}
-1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 &0.01246   \\
-0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 &0.01258   \\
-1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 &0.01343   \\
-0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 &0.01506   \\
+1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 & 0.00655\\
+0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 & 0.0057 \\
+1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 & 0.00582\\
+0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 & 0.00916\\
 \hline
-1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 &0.01397\\
+1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 & 0.00721\\
 \hline
@@
 \subfigure[Region labels for simulated slice.]{
-\includegraphics[height=9cm]{images/simregions.png} \label{fig:simslice_mask} }
-\subfigure[Grey matter map for simulate region.]{\includegraphics[height=9cm]{images/sim_gm}}
+\includegraphics[height=8cm]{images/simregions.png} \label{fig:simslice_mask} }
 \subfigure[Residual Map. Scale is Normalized error. Lower (yellow) is better.]
@@
 {\label{fig:sim_hm_mi} \includegraphics[height=8cm]{images/sim_hm_mi}}
+\subfigure[SNR Map of POSSUM simulated data. Mean SNR per region, Region 1: $0.8$, Region 2: $0.97$, and
+Region 3: $0.39$.]{\includegraphics[height=8cm]{images/snr_hm.png}\label{fig:SimSNRhm}}
+\label{fig:simslice_hm}
 \caption{Comparison of activation with greymatter, parameter regions.}
-\label{fig:simslice_hm}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{
-\includegraphics[width=10cm]{images/snr_hm.png}}
-\subfigure{\includegraphics[width=1cm]{images/scale5.png}}
-\caption{SNR Map of POSSUM simulated data. Region 1 mean SNR was $0.8$, Region 2 mean SNR was $0.97$, and
-Region 3 mean SNR was $0.39$.}
-\label{fig:SimSNRhm}
 \end{figure}

### 2010-08-25 doc/paper/main.tex
results all in

@@
 \centering
-\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{sec2hist}
 \caption{Histogram of estimated parameters. Average SNR in the region was 
@@
 \centering
-\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{sec3hist}
 \caption{Histogram of estimated parameters for region 3. Average SNR in the 
@@
 \centering
-\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
-\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
-\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
-\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
-\subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
+\subfigure{\label{fig:hm_spm} \includegraphics[scale=3]{spm_hm}}
+\subfigure{\label{fig:scale_spm} \includegraphics[scale=.35]{scale1}}
 \caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher
-        assurance that the signal cannot have occurred through noise alone.
-        Sagittal, coronal and axial slices  are
-        \autoref{fig:hm_canon_spm_x}, \autoref{fig:hm_canon_spm_y}, and
-         \autoref{fig:hm_canon_spm_z}, respectively. A series of axial slices are
-         shown in \autoref{fig:hm_spm}. }
+        assurance that the signal cannot have occurred through noise alone.}
 \label{fig:hm_canon_spm}
@@
 
+\begin{figure*}
+\centering
+\subfigure{\label{fig:hm_mi} \includegraphics[scale=3]{hm_mi_strict}}
+\subfigure{\label{fig:scale_mi} \includegraphics[scale=.35]{scale4}}
+\caption{Particle filter results. Units of activation are in mutual information; higher
+    indicates more assured activation.}
+\label{fig:hm_canon_spm}
+\end{figure*}
+
+The result of real experiments concentrates on regions of activation, because
+of the significant possiblity that the parameters are strictly defined. The
+quality of fit used for the particle filter results is mutual information which
+in preliminary testing showed fewer false positives than normalized Mean Squared
+Error.
+
 \section{Discussion}
 \label{sec:Discussion}
-
-\begin{itemize}
-\item Parameters ill-defined
-\end{itemize}
 

### 2010-08-25 doc/presentation/main.tex
added a frame where the graphs of the timeseries will go
got multihead working and added a script to start the show
new additions in nonlinear

@@
 %\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5]{beamerposter}
+
+%\usepackage{pgfpages}
 %\setbeameroption{show notes on second screen}
@@
     \tiny
-    \cite{nonlinearfiltering}
+    \cite{Riera2003}
 }
@@
         which are a function of Volume and Oxygenation.}
+\end{frame}
+
+\begin{frame}{Time Changing Parameters}
+    \note{Should include a picture of the different parts of the BOLD signal}
 \end{frame}
@@
 
-\begin{frame}{Solutions: Approximate Method}
+\section{Regression}
+\begin{frame}{Approximate Method}
   \begin{itemize}
-    \item 2nd Order Volterra Kernel 
+    \item 2nd Order Volterra Kernel \cite{Friston2000}
     \note{Quadratic Approximation}
@@
 \end{frame}
-%
+
+\begin{frame}{Nonlinear Modeling}
+\begin{itemize}
+    \item Local Linearization filter \cite{Riera2003}
+     \note{Jacobian of the $\dot{x}$ is available, making gradient descent possible}
+    \item Genetic Algorithms and Simulated Annealing \cite{Vakorin2007}
+     \note{Genetic Algorithms were used to optimize spike points, simulated annealing for the parameters}
+    \item Particle Filters to estimate $P(X_t | \Theta, Y_T)$ \cite{Murray2009}
+    \begin{itemize}
+        \item ML estimate of $\Theta$ based on $P(X_t | \Theta, Y_t)$ \cite{Johnston2007}
+         \note{Particle Filter used to estimate states, ML of parameters based on the posterior of states and so on}
+    \end{itemize}
+    \item Unscented Kalman Filter to estimate parameters \cite{Hu2009}
+    \item Direct Particle Filter, used in this work
+     \note{Technically Kalman Filters and Particle Filters are Bayesian Filters}
+     \note{Essentially filtering out inconsistent parameters.}
+\end{itemize}
+\end{frame}
+
 %\begin{frame}{BOLD Signal Properties}
@@
 %
-%\begin{frame}{Model Comparison}
-%  \begin{figure}
-%    \includegraphics[width=5in,height=2.4in]{comparison.png}
-%    \caption{
-%        \tiny
-%        \cite{comparison}
-%    }
-%  \end{figure}
-%\end{frame}
 %
@@
 \begin{frame}[allowframebreaks]
-  \bibliographystyle{abbrvnat}
-  \bibliography{references}
+  \bibliographystyle{plain}
+  \bibliography{../library}
 \end{frame}

### 2010-08-25 doc/presentation/main.tex
outline of defense presentation done

@@
 
-\title{Full Brain Blood-Oxygen-Level-Dependent Signal \\
-Parameter Estimation using Sequential Monte Carlo Methods}
+\title{Full Brain BOLD Signal Parameter Estimation \\
+using Particle Filters}
 
@@
 
-%\begin{frame}{BOLD Signal Properties}
-%  \begin{itemize}
-%    \item Exact variables and parameters are unknown and are
-%        difficult to calculate.
-%    \item Significant Amount of Lag between activation
-%        and a measurable output - can be as much as 8 seconds.
-%    \item Slow Temporal Resolution
-%    \item Noise characterized by brownian motion, which clashes with low 
-%        frequency elements.
-%    \begin{center}
-%    \includegraphics[width=10cm,height=4cm]{signal_fft.pdf}
-%    \end{center}
-%  \end{itemize}
-%\end{frame}
-%
-%\begin{frame}{Preprocessing}
-%  \begin{itemize}
-%    \item Low Pass Filter (Gaussian Filter, not recommended)
-%    \item Drift Removal (not always performed)
-%    \begin{itemize}
-%        \item High Pass Filter
-%        \item Linear 
-%        \item Quadratic
-%        \item Wavelet 
-%        \item Spline (Which I am using)
-%    \end{itemize}
-%  \end{itemize}
-%  \begin{figure}
-%    \includegraphics[width=10cm]{detrending_effect.png}
-%    \caption{
-%        \tiny
-%        \cite{detrending}
-%    }
-%  \end{figure}
-%\end{frame}
-%
-%\section{Statistical Parametric Mapping}
-%\begin{frame}{Method}
-%  \begin{figure}
-%    \includegraphics[width=11cm]{glm_pipeline.png}
-%    \caption{
-%        \tiny
-%        \cite{spm_pipeline}
-%    }
-%  \end{figure}
-%\end{frame}
-%
-%\begin{frame}{Limitations}
-%\begin{columns}
-%  \column{2.5in}
-%  \begin{itemize}
-%    \item Linear, for a signal which is known to be nonlinear
-%    \item Essentially the weighted sum of a set of "expected" responses.
-%    \item Parametric
-%    \begin{itemize}
-%        \item Forced to make assumptions about underlying distributions
-%        \item No time-scaling.
-%    \end{itemize}
-%  \end{itemize}
-%  \column{3in}
-%  \begin{figure}
-%    \includegraphics[width=2.5in]{glm.png}
-%    \caption{
-%        \tiny
-%        \cite{spm_pipeline}
-%    }
-%  \end{figure}
-%\end{columns}
-%\end{frame}
-%
-%\section{Nonlinear Regression}
-%
-%
-%\section{Parameter Identification}
-%\begin{frame}{Particle Filters}
-%\begin{itemize}
-%    \item Non-parametric, no assumptions are violated
-%    \item Model based, fit parameters to input, constrained by physical variables
-%    \item Fits a mixture PDF to the posterior of all parameters
-%    \item Non-trivial computation cost
-%    \item I use a Regularized Particle Filter
-%    \begin{enumerate}
-%        \item Regularized Re-sampling prevents particles from de-generating into a 
-%            small number of unique particles
-%        \item Allows distributions to move more freely
-%    \end{enumerate}
-%\end{itemize}
-%\end{frame}
-%
-%\begin{frame}{Particle Filter}
-%    \begin{columns}
-%    \column{1.8in}
-%    \begin{itemize}
-%        \item $S_t = \{p_{0,t}, ... , p_{N,t}\}$, the set of particles
-%        \item $w_{i,t}$, weight of particle $p_{i,t}$
-%        \item $y_t$, measurement at time $t$, there is not a $y_t$ for every $t$.
-%        \item $f(p_{i,t},y_t)$, weighting function 
-%        \item $s(p_{i,t})$, step function 
-%    \end{itemize}
-%    \column{3.2in}
-%    \begin{algorithmic}
-%     \STATE Draw $S_0$ from prior distribution 
-%     \FOR{$t = 0:t_{step}:t_{end}$}
-%        \FOR{each $p_{i,t-1} \in S_{t-1}$}
-%           \STATE $p_{i,t} = s(p_{i,t-1})$ 
-%           \IF{There is a measurement at time $t$}
-%               \FOR{every $p_{i,t}$}
-%                    \STATE $w_{i,t} = w_{i,t-1}f(p_{i,t},y_t)$
-%               \ENDFOR 
-%               \STATE Resample if weights are unevenly distributed\\
-%           \ENDIF
-%        \ENDFOR
-%     \ENDFOR
-%%                    \item If a weights are not evenly distributed enough, then resample
-%     \end{algorithmic}
-%     \end{columns}
-%\end{frame}
-%
-%\subsection{Single-Region}
-%\begin{frame}{Single Timeseries Results}
-%  \begin{figure}
-%    \includegraphics[height=3in]{noise.png}
-%  \end{figure}
-%\end{frame}
-%
-%\begin{frame}{Single Timeseries Results, Measurement Convergence}
-%  \begin{figure}
-%    \includegraphics[height=3in]{bold.png}
-%  \end{figure}
-%\end{frame}
-%
-%\begin{frame}{Single Timeseries Results, State Convergence}
-%  \begin{figure}
-%    \includegraphics[height=3in]{state.png}
-%  \end{figure}
-%\end{frame}
-%
-%\begin{frame}{Factors Affecting Convergence}
-%    \begin{enumerate}
-%        \item Weighting function
-%            \begin{itemize}
-%                \item Needs to be continuous and defined for any input, should go to 0
-%                \item Too wide a weighting function results in under-sensitivity, 
-%                            slow or no convergence
-%                \item Too thin a weighting function reduces robustness to noise
-%            \end{itemize}
-%        \item How often re-sampling is done, re-sampling should be minimized
-%            \begin{itemize}
-%                \item Stratified Resampling can result in truncated tails on posterior
-%                \item Regularized Resampling can result in reduced robustness to noise 
-%            \end{itemize}
-%        \item Number of particles
-%            \begin{itemize}
-%                \item More particles give higher fidelity of posterior
-%            \end{itemize}
-%    \end{enumerate}
-%\end{frame}
-%
-%\subsection{Multi-Region}
-%\begin{frame}{Parameter Map Generation/Simulation}
-%\begin{columns}
-%  \column{2.5in}
-%  \begin{enumerate}
-%    \item Generate a parameter map, with a set of parameters for each voxel
-%    \item Simulate every set of parameters, and use as input to possum
-%    \item Perform preprocessing (de-trend and normalize)
-%    \item Run particle filter on every grey matter voxel in image, generating a new
-%            parameter map
-%  \end{enumerate}
-%  \column{3in}
-%  \begin{figure}
-%    \includegraphics[width=2.8in]{equation.png}
-%    \caption{
-%    }
-%  \end{figure}
-%\end{columns}
-%\end{frame}
-%
-%\begin{frame}{Simulation Results, $\tau_0$}
-%  \begin{figure}
-%    \includegraphics[height=3in]{results.png}
-%    \caption{Percent Difference from "Actual"}
-%  \end{figure}
-%\end{frame}
-%
-%\begin{frame}{Simulation Results, multiple}
-%  \begin{figure}
-%    \includegraphics[height=3in]{results2.png}
-%    \caption{Percent Difference from "Actual"}
-%  \end{figure}
-%\end{frame}
-%
-%\section{Conclusion}
-%
+\section{Particle Filter}
+\begin{frame}{Construction}
+\begin{itemize}
+    \item Initialize Particles with Prior ($\alpha$):\\
+        $$\{(x^i_0,w^i) : x^i_0 \sim \alpha(X), w^i = 
+                \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$$
+        $$\alpha(X) \stackrel{\text{\tiny{k=0}}}{\approx} P(x_k) 
+                = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
+    \item Particle Weights at time $t_k$:\\
+        $$w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}$$
+      \note{$q$ is the density/location of the particle}
+      \note{$p$ is the actual probability }
+    \item Adding Measurements:\\
+        $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
+      \note{Because of the proportion, note that the weights do not
+            have to integrate to 1, and can be extremely large or extremely small}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Details and Example}
+  \begin{columns}
+    \begin{column}{.55\textwidth}
+        \begin{itemize}
+            \item Non-parametric
+            \item Model based
+            \item Online
+            \item Provides Full Posterior Probability
+            \item Non-trivial computation cost
+            \item Regularized Particle Filter Used
+        \end{itemize}
+    \end{column}
+    \begin{column}{.45\textwidth}
+        \begin{figure}
+          \includegraphics[width=\textwidth]{particle_filter}
+        \end{figure}
+    \end{column}
+    \end{columns}
+\end{frame}
+
+%todo I should probably add labels on the axes
+\section{Methods}
+\begin{frame}{FMRI Noise}
+\centering
+
+Resting State Noise
+
+\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009_22_38_23}
+
+Resting State Noise Steps
+
+\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009d_22_38_23}
+
+\note{Tanabe 2002 \autoref{Tanabe2002} 10-15\% of voxels exhibit significant drift}
+\note{Occurs in Cadavers.}
+\end{frame}
+
+\begin{frame}{Preprocessing}
+  \begin{columns}
+    \begin{column}{.7\textwidth}
+        \begin{itemize}
+            \item Drop Inital Volumes (9 Here)
+            \item Realign Over Time
+            \item Detrend
+            \item Gaussian Smoothing (SPM Only)
+            \begin{itemize}
+                \item Imposes Gaussianity
+                \item Increases SNR
+                \item Reduces Bonferroni Correction Requirement
+            \end{itemize}
+        \end{itemize}
+    \end{column}
+
+    \begin{column}{.3\textwidth}
+    %todo graphic of preprocessing
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\section{Results}
+\begin{frame}{Long Run Results}
+%convergence
+%correlation
+\end{frame}
+
+\begin{frame}{Short Run Results}
+%results image
+%table of estimated parameters
+\end{frame}
+
+\begin{frame}{POSSUM Results}
+%mutual information map
+\end{frame}
+
+\begin{frame}{SPM vs. Mutual Information Map}
+
+\end{frame}
+
+\begin{frame}{Factors Affecting Convergence}
+    \begin{itemize}
+        \item Weighting function
+            \begin{itemize}
+                \item Continuous, Long Tailed, Zero-Mean 
+                \begin{itemize}
+                \item Too wide -> under-sensitivity, slow or no convergence
+                \item Too thin -> reduces robustness to noise, particle deprivation
+                \end{itemize}
+            \end{itemize}
+        \item Resampling 
+            \begin{itemize}
+                \item Stratified Resampling can result in truncated tails 
+                        on posterior
+                \item Regularized Resampling can result in over smoothing and 
+                        slow convergence
+            \end{itemize}
+        \item Number of particles
+            \begin{itemize}
+                \item More particles give higher fidelity posterior
+            \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\section{Conclusion}
+\begin{frame}{Key Findings}
+
+\end{frame}
+
+\begin{frame}{Conclusion}
+\end{frame}
+
 %% All of the following is optional and typically not needed. 

### 2010-08-25 doc/presentation/main.tex
basically done except for 2 pictures I need to fix.

@@
 
+\usepackage{colortbl}
 \usepackage[pdftex]{graphicx}
@@
 \usepackage{algorithmic}
+\usepackage{subfigure}
 
@@
 \begin{frame}{Long Run Results}
-%convergence
-%correlation
+\includegraphics[width=\textwidth]{long_converge}
 \end{frame}
 
-\begin{frame}{Short Run Results}
-%results image
-%table of estimated parameters
+\begin{frame}{Long Run Correlation}
+\centering
+Estimated Parameter Covariance
+
+\begin{table}[t]
+\tiny
+\begin{tabular}{|c | c  c  c  c  c  c  c |}
+\hline
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
+\hline
+\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
+$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
+\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
+$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
+$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
+\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
+\hline
+\end{tabular}
+\label{tab:long_cov}
+\end{table}
+
+Estimated Parameter Correlation
+
+\begin{table}[t]
+\tiny
+\begin{tabular}{|c | c  c  c  c  c  c  |}
+\hline
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
+\hline
+\rowcolor[gray]{.8} $\tau_0$  & & & & & & \\
+$\alpha$                      & 0.889884 & & & & & \\
+\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & \\
+$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & \\
+$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & \\
+\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 \\
+\hline
+\end{tabular}
+\note{Correlation of parameter estimates at the end of 
+            \autoref{fig:long_converge}.}
+\label{tab:long_corr}
+\end{table}
 \end{frame}
 
-\begin{frame}{POSSUM Results}
-%mutual information map
+\begin{frame}{Short Run Results A}
+\begin{center}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
+            {preprocessed_lownoise}
+\end{center}
+\end{frame}
+
+\begin{frame}{Short Run Results B}
+\begin{center}
+\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
+            {comparison_lownoise}
+\end{center}
+\end{frame}
+
+\begin{frame}{Short Run Results C}
+\begin{table}[t]
+\tiny
+\centering
+\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
+\hline
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Error & Residual\\
+\hline
+\rowcolor[gray]{.8}
+1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
+\hline
+\hline
+1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
+1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
+1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
+1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
+1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
+1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
+1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
+1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
+1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
+1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
+1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
+\hline                                                                               
+1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
+\hline
+\end{tabular}
+\label{tab:LowNoiseResults}
+\end{table}
+\end{frame}
+
+\begin{frame}{POSSUM Results: Activation}
+\begin{figure}
+\centering
+\subfigure[Regions]{\includegraphics[height=.46\textheight]{simregions}}
+\subfigure[Mutual Information]
+{\includegraphics[height=.46\textheight]{sim_hm_15_6}}
+\subfigure[SNR]{\includegraphics[height=.46\textheight]{snr_hm}}
+\note{Mean SNR per region, Region 1: $0.8$, Region 2: $0.97$, and
+            Region 3: $0.39$.}
+\end{figure}
+\end{frame}
+
+\begin{frame}{POSSUM Results: Histogram}
+\begin{figure}
+%\centering
+\includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=\textwidth]{sec2hist}
+\note{Histogram of estimated parameters. Average SNR in the region was 
+$0.97$ and Mutual Information between simulation and BOLD estimate was greater 
+than $0.15$.}
+\end{figure}
 \end{frame}
@@
 \begin{frame}{SPM vs. Mutual Information Map}
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{figure}
+\setcounter{subfigure}{0}
 
+\subfigure[SPM Results]{\includegraphics[width=.9\textwidth]{spm_hm}}
+\subfigure{\includegraphics[width=.07\textwidth]{scale1}}
+\note{SPM results. Units of activation are in Student's $t$-scores; higher 
+        indicates higher assurance that the signal cannot have occurred 
+        through noise alone.}
+\label{fig:hm_canon_spm}
+\end{figure}
+\end{column}
+
+\begin{column}{.5\textwidth}
+\begin{figure}
+\setcounter{subfigure}{1}
+\subfigure[Particle Filter Results]{\includegraphics[width=.9\textwidth]{hm_mi_strict}}
+\subfigure{\includegraphics[height=1.5cm,width=.07\textwidth]{scale4}}
+\note{Particle filter results. Units of activation are in mutual information;
+    higher indicates more assured activation.}
+\label{blah2}
+\end{figure}
+\end{column}
+\end{columns}
+\end{frame}
+
+%todo fix histogram like I did with the POSSUM results
+\begin{frame}{Particle Filter Results: Histogram}
+\begin{figure}
+\centering
+\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=.8\textwidth]{realhist}
+\note{Histogram of parameters in active regions ($M.I. > .15$).}
+\end{figure}
 \end{frame}
@@
                 \begin{itemize}
-                \item Too wide -> under-sensitivity, slow or no convergence
-                \item Too thin -> reduces robustness to noise, particle deprivation
+                \item Too wide $\rightarrow$ under-sensitivity, 
+                    slow or no convergence
+                \item Too thin $\rightarrow$ reduces robustness 
+                    to noise, particle deprivation
                 \end{itemize}
@@
 \section{Conclusion}
-\begin{frame}{Key Findings}
+\begin{frame}{Conclusion}
+\begin{itemize}
+    \item Summary:
+    \begin{itemize}
+    \item BOLD Parameters Ill-Defined
+    \item Particle Filter Capable of good parameter BOLD estimate with 1000 particles
+    \end{itemize}
+    \item Future Works
+    \begin{itemize}
+        \item Further limitations should be placed on priors, Deneux 
+            et al. \cite{Deneux2006} shows that parameters could imposed.
+        \item Automation of Weighting Function
+        \item Analysis of joint parameter distribution for populations
+        \item Real Time analysis possible for multiple Voxels, similar to
+            De Charms et al. \cite{DeCharms2005}
 
-\end{frame}
-
-\begin{frame}{Conclusion}
+    \end{itemize}
+\end{itemize}
 \end{frame}
@@
 
+\appendix
+\begin{frame}{SPM vs. Mutual Information Map, SPM}
+\begin{figure}
+\setcounter{subfigure}{0}
+\subfigure[SPM Results]{\includegraphics[width=.8\textwidth]{images/spm_hm}}
+\subfigure{\includegraphics[scale=.5]{images/scale1}}
+\note{SPM results. Units of activation are in Student's $t$-scores; higher 
+        indicates higher assurance that the signal cannot have occurred 
+        through noise alone.}
+\end{figure}
+\end{frame}
+
+\begin{frame}{SPM vs. Mutual Information Map, $M.I. > .15$}
+\begin{figure}
+\setcounter{subfigure}{1}
+\subfigure[Particle Filter Results]{\includegraphics[width=.8\textwidth]{hm_mi_strict}}
+\subfigure{\includegraphics[width=.1\textwidth]{scale4}}
+\note{Particle filter results. Units of activation are in mutual information;
+    higher indicates more assured activation.}
+\label{blah2}
+\end{figure}
+\end{frame}
+
+\begin{frame}{SPM vs. Mutual Information Map, $M.I. > .1$}
+\begin{figure}
+\setcounter{subfigure}{2}
+\subfigure[Particle Filter Results]{\includegraphics[width=.8\textwidth]{mi_hm}}
+\subfigure{\includegraphics[width=.1\textwidth]{scale6}}
+\note{Particle filter results. Units of activation are in mutual information;
+    higher indicates more assured activation.}
+\label{blah2}
+\end{figure}
+\end{frame}
+
+\begin{frame}{1: 37-14-7}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/1_pfilter_37_14_7}}
+\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/1_spm_37_14_7}}
+\caption{Section 1, Estimated vs. Actual BOLD response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
+\label{fig:comp1}
+\end{figure}
+\end{frame}
+
+\begin{frame}{2: 34-12-7}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/2_pfilter_34_12_7}}
+\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/2_spm_34_12_7}}
+\caption{Section 2, Estimated vs. Actual BOLD response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
+\label{fig:comp2}
+\end{figure}
+\end{frame}
+
+\begin{frame}{3: 23-21-7}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/3_pfilter_23_21_7}}
+\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/3_spm_23_21_7}}
+\caption{Section 3, Estimated vs. Actual BOLD response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
+\label{fig:comp3}
+\end{figure}
+\end{frame}
+
+\begin{frame}{4: 33-40-4}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/4_pfilter_26_15_7}}
+\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/4_spm_26_15_7}}
+\caption{Section 4, Estimated vs. Actual BOLD response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
+\label{fig:comp4}
+\end{figure}
+\end{frame}
+
+\begin{frame}{5: 29-9-13}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/5_pfilter_29_9_13}}
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/5_spm_29_9_13}}
+\caption{Section 5, Estimated vs. Actual BOLD Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
+%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
+%of $4.17399$ and $MSE$ of $1.14171$.}
+\label{fig:comp5}
+\end{figure}
+\end{frame}
+
+\begin{frame}{6: 36-17-19}
+\setcounter{subfigure}{0}
+\begin{figure}
+\centering
+\subfigure[Particle Filter]{\includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/6_pfilter_36_17_19}}
+\subfigure[SPM]{\includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/6_spm_36_17_19}}
+\caption{Section 6, Estimated vs. Actual BOLD Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
+%\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
+\end{figure}
+\end{frame}
+
 \end{document}

### 2010-08-25 doc/paper/main.tex
mostly done, a few better images from the defense

@@
 \subfloat[Region labels for simulated slice.]
-{\includegraphics[width=3cm]{images/simregions.png}\label{fig:simslice_mask}}
+{\includegraphics[height=4.5cm]{simregions.png}\label{fig:simslice_mask}}
 \hfil
 \subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
-{\label{fig:sim_hm_mi} \includegraphics[width=4cm]{images/sim_hm_15_6.png}}
-\subfigure{\includegraphics[width=.6cm]{images/scale6.png}}
+{\label{fig:sim_hm_mi} \includegraphics[height=4.5cm]{sim_hm_15_6.png}}
 \caption{Comparison of activation with greymatter, parameter regions.}
@@
 \centering
-\subfigure{\label{fig:hm_spm} \includegraphics[scale=3]{spm_hm}}
-\subfigure{\label{fig:scale_spm} \includegraphics[scale=.35]{scale1}}
+\subfloat{\label{fig:hm_spm} \includegraphics[scale=3]{spm_hm}}
+\subfloat{\label{fig:scale_spm} \includegraphics[scale=.35]{scale1}}
 \caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher
@@
 \centering
-\subfigure{\label{fig:hm_mi} \includegraphics[scale=3]{hm_mi_strict}}
-\subfigure{\label{fig:scale_mi} \includegraphics[scale=.35]{scale4}}
+\subfloat{\label{fig:hm_mi} \includegraphics[scale=3]{hm_mi_strict}}
+\subfloat{\label{fig:scale_mi} \includegraphics[scale=.35]{scale4}}
 \caption{Particle filter results. Units of activation are in mutual information; higher
     indicates more assured activation.}
-\label{fig:hm_canon_spm}
+\label{fig:hm_mi_res}
 \end{figure*}

### 2010-08-26 doc/presentation/main.tex
all thats left is adding some images for preprocessing
defense is "done"
couple small changes, question slide
more work on notes

@@
 
-%\usepackage{pgfpages}
+\usepackage{pgfpages}
 %\setbeameroption{show notes on second screen}
@@
 
+%0
 \begin{frame}{Outline}
@@
   % You might wish to add the option [pausesections]
+  \note{Introduction describes FMRI and the Balloon Model}
+  \note{Potential Approaches - describes previous used methods of analysis}
+  \note{Particle Filter - derives and discusses the BOLD model}
+  \note{Methods - The methods used in this work, how to recreate.}
 \end{frame}
@@
 \section{Introduction}
+%2
 \begin{frame}{Balloon Model}
@@
 \end{figure}
-\note{Rates range from below 1 Hz to 5 to 6 Hundred Hz}
+\note{Rates from below 1 Hz to 5 to 6 Hundred Hz}
 \note{Change in average firing rates increase metabolism and sympathetic response causes
@@
 
+%4
 \begin{frame}{Time Changing Parameters}
+    \includegraphics[clip=true,trim=3cm 1cm 3cm 1cm,width=\textwidth]{stateTS}
     \note{Should include a picture of the different parts of the BOLD signal}
@@
 
+%6 
 \begin{frame}{BOLD State Equations}
+  \small
   \begin{itemize}
     \item Flow Inducing Signal
-    $$\dot{s}(t) = \epsilon u(t) - s(t)/\tau_s - (f(t)/\tau_f - 1)$$
+    $$\dot{s}(t) = \epsilon u(t) - s(t)/\tau_s - (f(t)-1)/\tau_f$$
     \item Normalized Cerebral Blood Inflow:
@@
 
-\section{Regression}
+\section{Potential Approaches}
 \begin{frame}{Approximate Method}
@@
 
+\begin{frame}{Particle Filter Algorithm}
+\footnotesize
+\begin{algorithm}
+\begin{algorithmic}
+\STATE Initialize Particles:
+\FOR{$i$ : each of $N_p$ particles }
+    \STATE $x^i_0  \sim \alpha(X)$
+    \STATE $w^i_0 = \frac{1}{N_p}$
+\ENDFOR
+\FOR{$k$ : each measurement}
+    \FOR{$i$ : each particle }
+        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
+        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
+    \ENDFOR
+    \STATE Calculate $N_{eff}$ 
+    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
+        \STATE Regularized Resample 
+    \ENDIF
+\ENDFOR
+
+\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
+\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
+ \end{algorithmic}
+ \end{algorithm}
+\end{frame}
+
 \begin{frame}{Details and Example}
@@
 
-%todo I should probably add labels on the axes
 \section{Methods}
@@
   \begin{columns}
-    \begin{column}{.7\textwidth}
+    \begin{column}{.6\textwidth}
         \begin{itemize}
-            \item Drop Inital Volumes (9 Here)
+            \item Drop Inital Volumes (9, $18.9$s)
             \item Realign Over Time
-            \item Detrend
+            \item Detrend (SPM uses $0.0078125 Hz$ cut off)
             \item Gaussian Smoothing (SPM Only)
@@
 
-    \begin{column}{.3\textwidth}
-    %todo graphic of preprocessing
+    \begin{column}{.4\textwidth}
+        \includegraphics[width=\textwidth]{preprocess}
     \end{column}
@@
 \includegraphics[width=\textwidth]{long_converge}
+\note{Bars drop well before the end}
+\note{Mean is good estimator}
+\note{Mean comes very close to truth}
 \end{frame}
@@
 \begin{table}[t]
-\tiny
 \centering
+\resizebox{\textwidth}{!}{
 \begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
@@
 \end{tabular}
-\label{tab:LowNoiseResults}
+}
 \end{table}
@@
 \subfigure[SNR]{\includegraphics[height=.46\textheight]{snr_hm}}
-\note{Mean SNR per region, Region 1: $0.8$, Region 2: $0.97$, and
-            Region 3: $0.39$.}
+\note{Mean SNR's Region 1: $0.8$, Region 2: $0.97$, and Region 3: $0.39$.}
 \end{figure}
@@
 \includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=\textwidth]{sec2hist}
-\note{Histogram of estimated parameters. Average SNR in the region was 
-$0.97$ and Mutual Information between simulation and BOLD estimate was greater 
-than $0.15$.}
+\note{Histogram of estimated parameters.}
+\note{Average SNR: $0.97$}
+\note{Mutual Information greater than $0.15$.}
 \end{figure}
@@
     higher indicates more assured activation.}
-\label{blah2}
 \end{figure}
@@
 
-%todo fix histogram like I did with the POSSUM results
 \begin{frame}{Particle Filter Results: Histogram}
@@
 
+\section{Conclusion}
 \begin{frame}{Factors Affecting Convergence}
@@
                 \item Continuous, Long Tailed, Zero-Mean 
-                \begin{itemize}
                 \item Too wide $\rightarrow$ under-sensitivity, 
@@
                     to noise, particle deprivation
-                \end{itemize}
             \end{itemize}
@@
 
-\section{Conclusion}
 \begin{frame}{Conclusion}
@@
     \item Particle Filter Capable of good parameter BOLD estimate with 1000 particles
+    \item Mutual Information performs well as estimate of Quality
     \end{itemize}
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}{Questions?}
 \end{frame}

### 2010-08-26 doc/presentation/prep.sh
needs fixes....but whateve

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+xrandr --output LVDS1 --mode 1024x768
+xrandr --output VGA1 --mode 1024x768 --left-of LVDS1

### 2010-08-26 doc/presentation/reverse.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+xrandr --output LVDS1 --mode 1440x900

### 2010-08-26 doc/presentation/run.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-impressive main.pdf
+impressive -c memory --geometry 2048x768 -t None withnotes.pdf

### 2010-08-27 doc/presentation/main.tex
needs fixes....but whateve
some comments
changed the section on bold, need to change the particle filter now

@@
 \section{Introduction}
-%2
 \begin{frame}{Balloon Model}
@@
 
-%4
 \begin{frame}{Time Changing Parameters}
-    \includegraphics[clip=true,trim=3cm 1cm 3cm 1cm,width=\textwidth]{stateTS}
-    \note{Should include a picture of the different parts of the BOLD signal}
+\begin{figure}
+    \includegraphics[width=\textwidth]{graphs}
+\caption{
+    \tiny
+    \cite{Buxton2004}
+}
+\end{figure}
 \end{frame}
 
-%6 
-\begin{frame}{BOLD State Equations}
-  \small
+%more expansion of all the parameters
+\begin{frame}{BOLD State Space Model}
+  \scriptsize
+  \begin{columns}
+  \begin{column}{.7\textwidth}
     \begin{itemize}
-    \item Flow Inducing Signal
-    $$\dot{s}(t) = \epsilon u(t) - s(t)/\tau_s - (f(t)-1)/\tau_f$$
-    \item Normalized Cerebral Blood Inflow:
-    $$\dot{f}(t) = s $$
-    \item Normalized Cerebral Blood Volume:
-    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
-    \item Normalized Deoxyhaemoglobin Content:\\
-    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1-(1-E_0)^{1/f(t)})}{E_0} -
-            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
-    \item Hemodynamic Response - BOLD Signal
+    \item State Vector:
+    \begin{eqnarray}
+    \dot{s}(t) &=& \epsilon u(t) - s(t)/\tau_s - (f(t)-1)/\tau_f \nonumber \\
+    \dot{f}(t) &=& s \nonumber \\
+    \dot{v}(t) &=& (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) \nonumber \\
+    \dot{q}(t) &=& \frac{1}{\tau_0}\left(\frac{f(t)(1-(1-E_0)^{1/f(t)})}{E_0} -
+            \frac{q(t)}{v(t)^{1-1/\alpha}}\right) \nonumber 
+    \end{eqnarray}
+    \item Output:
     $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
     \end{itemize}
-  \note{System is dissipative:}
-  \note{If no input is provided, it will rather quickly decay.}
-  \note{Constant input leads to steady state response.}
-  \note{First term in $\dot{q}$ is the metabolism term.}
-  \note{Outflow is the $v^{1/\alpha}$ term}
-  \note{Significant nonlinearity.}
+  \end{column}
+
+  \begin{column}{.3\textwidth}
+        State Variables:
+        $$s(t), f(t), v(t), q(t)$$
+        Parameters:
+        $$\epsilon, \tau_s, \tau_f, \alpha, \tau_0, E_0, V_0$$
+        Constants:
+        $$a_1, a_2$$
+        Input:
+        $$u(t)$$
+  \end{column}
+  \end{columns}
+  \note{System is dissipative:\\
+  If no input is provided, it will rather quickly decay.\\
+  Constant input leads to steady state response.\\
+  First term in $\dot{q}$ is the metabolism term.\\
+  Outflow is the $v^{1/\alpha}$ term\\
+  Significant nonlinearity.\\
+  $a_1 = (4.3\nu_0TE+\epsilon_025\epsilon_0)E_0$, $a_2 = \epsilon_025\epsilon_0E_0 + .43$}
 \end{frame}
@@
 \section{Particle Filter}
+%frame with example
+
 \begin{frame}{Construction}
 \begin{itemize}
-    \item Initialize Particles with Prior ($\alpha$):\\
-        $$\{(x^i_0,w^i) : x^i_0 \sim \alpha(X), w^i = 
-                \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$$
+    \item Mixture PDF:\\
         $$\alpha(X) \stackrel{\text{\tiny{k=0}}}{\approx} P(x_k) 

### 2010-08-28 doc/presentation/main.tex
more work on the presentation

@@
 
+\usepackage{color}
 \usepackage{colortbl}
@@
 %\usepackage[small]{caption}
+\usepackage{algorithm}
 \usepackage{algorithmic}
 \usepackage{subfigure}
+
+\definecolor{mblue}{RGB}{178,34,34}
+\definecolor{mred}{RGB}{0,0,128}
+\definecolor{mgreen}{RGB}{0,100,0}
 
@@
 %\beamerdefaultoverlayspecification{<+->}
+
+\AtBeginSection{
+\begin{frame}
+\begin{center}
+\structure{\Huge \insertsection}
+\end{center}
+\end{frame}
+}
+
+\AtBeginSubsection{
+\begin{frame}
+\begin{center}
+\structure{\Huge \insertsubsection}
+\end{center}
+\end{frame}
+}
 
@@
 
-\begin{frame}{Time Changing Parameters}
+\begin{frame}{Activation Chain}
 \begin{figure}
@@
     \begin{eqnarray}
-    \dot{s}(t) &=& \epsilon u(t) - s(t)/\tau_s - (f(t)-1)/\tau_f \nonumber \\
-    \dot{f}(t) &=& s \nonumber \\
-    \dot{v}(t) &=& (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) \nonumber \\
-    \dot{q}(t) &=& \frac{1}{\tau_0}\left(\frac{f(t)(1-(1-E_0)^{1/f(t)})}{E_0} -
-            \frac{q(t)}{v(t)^{1-1/\alpha}}\right) \nonumber 
+    \color{mblue}\dot{s}(t) &=& {\color{mred}\epsilon} {\color{mgreen}u(t)} - 
+                \frac{\color{mblue}s(t)}{\color{mred}\tau_s} - \frac{{\color{mblue}f(t)}-1}{\color{mred}\tau_f} \nonumber \\
+    \color{mblue}\dot{f}(t) &=& \color{mblue}s \nonumber \\
+    \color{mblue}\dot{v}(t) &=& \frac{{\color{mblue}f(t)} - 
+                {\color{mblue}v(t)} ^ {1/{\color{mred}\alpha}}}{\color{mred}\tau_0} \nonumber \\
+    \color{mblue}\dot{q}(t) &=& \frac{1}{\color{mred}\tau_0}
+            \left(\frac{{\color{mblue}f(t)}(1-(1-{\color{mred}E_0})^{1/\color{mblue}f(t)})}{\color{mred}E_0} -
+        \frac{\color{mblue}q(t)}{{\color{mblue}v(t)}^{1-1/\color{mred}\alpha}}\right) \nonumber 
     \end{eqnarray}
     \item Output:
-    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
+    $$y(t) = {\color{mred}V_0}(a_1( 1 - {\color{mblue}Q(t)}) - a_2(1 - {\color{mblue}V(t)}))$$
     \end{itemize}
@@
   \begin{column}{.3\textwidth}
-        State Variables:
-        $$s(t), f(t), v(t), q(t)$$
-        Parameters:
-        $$\epsilon, \tau_s, \tau_f, \alpha, \tau_0, E_0, V_0$$
-        Constants:
+    \begin{itemize}
+        \item State Variables:
+        $${\color{mblue}s(t)}, {\color{mblue}f(t)}, {\color{mblue}v(t)}, {\color{mblue}q(t)}$$
+        \item Parameters:
+        $${\color{mred}\epsilon}, {\color{mred}\tau_s}, {\color{mred}\tau_f}, {\color{mred}\alpha}, 
+                    {\color{mred}\tau_0}, {\color{mred}E_0}, {\color{mred}V_0}$$
+        \item Constants:
         $$a_1, a_2$$
-        Input:
-        $$u(t)$$
+        \item Input:
+        $$\color{mgreen}u(t)$$
+    \end{itemize}
   \end{column}
@@
   Outflow is the $v^{1/\alpha}$ term\\
-  Significant nonlinearity.\\
-  $a_1 = (4.3\nu_0TE+\epsilon_025\epsilon_0)E_0$, $a_2 = \epsilon_025\epsilon_0E_0 + .43$}
+  Significant nonlinearity.}
 \end{frame}
 
-\section{Potential Approaches}
-\begin{frame}{Approximate Method}
+\section{Nonlinear Regression}
+\subsection{Prior Works}
+\begin{frame}{Approximation Method}
   \begin{itemize}
@@
 
-\begin{frame}{Nonlinear Modeling}
+\begin{frame}{Nonlinear Methods}
 \begin{itemize}
-    \item Local Linearization filter \cite{Riera2003}
+    \begin{columns}
+    \begin{column}{.45\textwidth}
+    \item Local Linearization filter, \cite{Riera2003}
+    $$f(t)- f(t-1) \sim N(0, \sigma^2)$$
      \note{Jacobian of the $\dot{x}$ is available, making gradient descent possible}
-    \item Genetic Algorithms and Simulated Annealing \cite{Vakorin2007}
+    \item Genetic Algorithms and Simulated Annealing, \cite{Vakorin2007}
      \note{Genetic Algorithms were used to optimize spike points, simulated annealing for the parameters}
@@
     \begin{itemize}
-        \item ML estimate of $\Theta$ based on $P(X_t | \Theta, Y_t)$ \cite{Johnston2007}
+        \item ML estimate of $\Theta$ based on $P(X_t | \Theta, Y_t)$, \cite{Johnston2007}
          \note{Particle Filter used to estimate states, ML of parameters based on the posterior of states and so on}
@@
     \item Unscented Kalman Filter to estimate parameters \cite{Hu2009}
-    \item Direct Particle Filter, used in this work
      \note{Technically Kalman Filters and Particle Filters are Bayesian Filters}
      \note{Essentially filtering out inconsistent parameters.}
+    \end{column}
+
+    \begin{column}{.55\textwidth}
+    \small
+    \begin{figure}
+    \includegraphics[width=.7\textwidth]{simulated_annealing}
+    \caption{Simulated Annealing can escape local minima with chaotic jumps. \cite{Dama}}
+    \end{figure}
+    \begin{figure}
+    \includegraphics[width=.7\textwidth]{kalman}
+    \caption{(a) Initial Belief  (b) Noisy Measurement  (c) incorporated into the belief. \cite{Thrun2005}}
+    \end{figure}
+    \end{column}
+    \end{columns}
 \end{itemize}
@@
 
-\section{Particle Filter}
+\subsection{Particle Filter}
 %frame with example
+\begin{frame}{Example System Identification}
+\begin{itemize}
+\item Given:
+    \begin{itemize}
+    \item Elevation Map (1D)
+    \item Ability To Measure Elevation
+    \end{itemize}
+\item Particle Consists of:
+    \begin{itemize}
+    \item Location (Altered By Speed/Direction) $[0, 20]$
+    \item Direction (Constant, Unknown) $\{Left, Right\}$
+    \end{itemize}
+\end{itemize}
+
+\begin{center}
+\includegraphics[clip=true,trim=2cm 3cm 3cm 3cm,width=.7\textwidth]{setup}
+\end{center}
+\end{frame}
+
+\begin{frame}{Initial Distribution}
+\centering
+\includegraphics[width=\textwidth]{init}\\
+\end{frame}
+
+\begin{frame}{Measurement 1}
+\centering
+\includegraphics[width=\textwidth]{meas1}\\
+\end{frame}
+
+\begin{frame}{Step Forward}
+\centering
+\includegraphics[width=\textwidth]{move1}\\
+\end{frame}
+
+\begin{frame}{Measurement 2}
+\centering
+\includegraphics[width=\textwidth]{meas2}\\
+\end{frame}
+
+\begin{frame}{Step Forward}
+\centering
+\includegraphics[width=\textwidth]{move2}\\
+\end{frame}
 
@@
 \begin{itemize}
-    \item Mixture PDF:\\
-        $$\alpha(X) \stackrel{\text{\tiny{k=0}}}{\approx} P(x_k) 
-                = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
+    \item Mixture PDF at time $k$, with $N_p$ particles:\\
+        $$P(x_k) = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
     \item Particle Weights at time $t_k$:\\
@@
       \note{$p$ is the actual probability }
-    \item Adding Measurements:\\
+    \item Incorporating Measurement:\\
         $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
@@
 
-\begin{frame}{Particle Filter Algorithm}
+\begin{frame}{Particle Filter }
 \footnotesize
-\begin{algorithm}
-\begin{algorithmic}
-\STATE Initialize Particles:
-\FOR{$i$ : each of $N_p$ particles }
-    \STATE $x^i_0  \sim \alpha(X)$
-    \STATE $w^i_0 = \frac{1}{N_p}$
-\ENDFOR
-\FOR{$k$ : each measurement}
-    \FOR{$i$ : each particle }
-        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
-    \ENDFOR
-    \STATE Calculate $N_{eff}$ 
-    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
-        \STATE Regularized Resample 
-    \ENDIF
-\ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
- \end{algorithmic}
- \end{algorithm}
-\end{frame}
-
-\begin{frame}{Details and Example}
-  \begin{columns}
-    \begin{column}{.55\textwidth}
 \begin{itemize}
-            \item Non-parametric
-            \item Model based
-            \item Online
-            \item Provides Full Posterior Probability
-            \item Non-trivial computation cost
-            \item Regularized Particle Filter Used
+    \item 
 \end{itemize}
-    \end{column}
-    \begin{column}{.45\textwidth}
-        \begin{figure}
-          \includegraphics[width=\textwidth]{particle_filter}
-        \end{figure}
-    \end{column}
-    \end{columns}
 \end{frame}
@@
         \begin{itemize}
-            \item Drop Inital Volumes (9, $18.9$s)
+            \item Drop Initial Volumes (9, $18.9$s)
             \item Realign Over Time
@@
 \section{Results}
+\subsection{40 Minute Single Voxel Simulation}
 \begin{frame}{Long Run Results}
@@
 
-\begin{frame}{Long Run Correlation}
+\begin{frame}{Long Run Covariance and Correlation}
 \centering
@@
 
-\begin{frame}{Short Run Results A}
+\subsection{5 Minute, Single Voxel Simulation}
+\begin{frame}{Preprocessed Noisy Simulation vs. Underlying Signal}
 \begin{center}
@@
 
-\begin{frame}{Short Run Results B}
+\begin{frame}{Estimated vs. Underlying Signal}
 \begin{center}
@@
 
-\begin{frame}{Short Run Results C}
+\begin{frame}{Estimated Parameters}
 \begin{table}[t]
@@
 
-\begin{frame}{POSSUM Results: Activation}
+\subsection{POSSUM Simulation}
+\begin{frame}{Mutual Information Compared with SNR}
 \begin{figure}
@@
 %\centering
-\includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=\textwidth]{sec2hist}
+\includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=.9\textwidth]{sec2hist}
 \note{Histogram of estimated parameters.}
@@
 
+\subsection{Real FMRI Data}
 \begin{frame}{SPM vs. Mutual Information Map}

### 2010-08-28 doc/presentation/main.tex
more work getting the presentation togeth

@@
 \definecolor{mgreen}{RGB}{0,100,0}
+\definecolor{myellow}{RGB}{176,169,0}
 
@@
 \begin{figure}
-\includegraphics[width=10cm]{model}
+\includegraphics[width=10cm]{model2}
 \caption{
@@
 \end{figure}
+\note{Its been known for about 20 years that changes in Blood Oxygen Level alter
+        MR signals in certain types of imaging. The basis of contrast in fMRI is 
+        changes in the so-called spin-spin relaxation (or T2$^*$ relaxation time
+        of tissue). The rate of loss in nuclear spin coherence depends on the molecules in the voxel being
+        measured \emph{and the surrounding voxels}. For this reason the signal 
+        detected from BLOOD doesn't just depend on the amount of oxygen, but also the
+        amount of De-oxygenated hemoglobin. The question then, is what drives changes
+        in De-oxygenated Hemoglobin and blood volume?
+
+        Well the brain has no local oxygen storage, so it is completely dependent on 
+        fresh Oxygenated-Hemoglobin. When the brain becomes more active, firing rates
+        for neurons increase by several orders of magnitude, and recharging these neurons
+        requires additional oxygen for matabolism. }
 \note{Rates from below 1 Hz to 5 to 6 Hundred Hz}
@@
     \item Output:
-    $$y(t) = {\color{mred}V_0}(a_1( 1 - {\color{mblue}Q(t)}) - a_2(1 - {\color{mblue}V(t)}))$$
+    $${\color{myellow}y(t)} = {\color{mred}V_0}(a_1( 1 - {\color{mblue}Q(t)}) - a_2(1 - {\color{mblue}V(t)}))$$
     \end{itemize}
@@
     \begin{columns}
-    \begin{column}{.45\textwidth}
+    \begin{column}{.42\textwidth}
     \item Local Linearization filter, \cite{Riera2003}
@@
      \note{Genetic Algorithms were used to optimize spike points, simulated annealing for the parameters}
-    \item Particle Filters to estimate $P(X_t | \Theta, Y_T)$ \cite{Murray2009}
+    \item Particle Filters to estimate States, \cite{Murray2009}
     \begin{itemize}
-        \item ML estimate of $\Theta$ based on $P(X_t | \Theta, Y_t)$, \cite{Johnston2007}
+        \item ML estimate of $\Theta$, \cite{Johnston2007}
          \note{Particle Filter used to estimate states, ML of parameters based on the posterior of states and so on}
     \end{itemize}
-    \item Unscented Kalman Filter to estimate parameters \cite{Hu2009}
+    \item Unscented Kalman Filter \cite{Hu2009}
      \note{Technically Kalman Filters and Particle Filters are Bayesian Filters}
@@
 
-    \begin{column}{.55\textwidth}
-    \small
+    \begin{column}{.58\textwidth}
     \begin{figure}
-    \includegraphics[width=.7\textwidth]{simulated_annealing}
+    \includegraphics[clip=true,trim=0cm 0cm 0cm 1cm,width=.6\textwidth]{simulated_annealing}
     \caption{Simulated Annealing can escape local minima with chaotic jumps. \cite{Dama}}
@@
     \begin{figure}
-    \includegraphics[width=.7\textwidth]{kalman}
-    \caption{(a) Initial Belief  (b) Noisy Measurement  (c) incorporated into the belief. \cite{Thrun2005}}
+    \includegraphics[width=\textwidth]{kalman}
+    \caption{UKF: (a) Initial Belief  (b) Noisy Measurement  (c) incorporated into the belief. \cite{Thrun2005}}
     \end{figure}
@@
     \end{itemize}
-\item Particle Consists of:
+\item Particle Consists of the unknowns:
     \begin{itemize}
-    \item Location (Altered By Speed/Direction) $[0, 20]$
-    \item Direction (Constant, Unknown) $\{Left, Right\}$
+    \item State: Location  $[0, 20]$
+    \item Constant: Direction  $\{Left, Right\}$
     \end{itemize}
@@
 
+\begin{frame}{Particle Filter Concepts}
+\small
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{itemize}
+    \item Weighting Function
+    \item Resampling
+    \begin{itemize}
+        \item Remove Useless Particles
+    \end{itemize}
+    \item Regularized Resampling
+    \begin{itemize}
+        \item Prevent Redundant Particles
+        \item Reduces Quantization Error
+    \end{itemize}
+\end{itemize}
+\end{column}
+
+\begin{column}{.5\textwidth}
+\includegraphics[clip=true,trim=0cm 6cm 0cm 6cm,width=\textwidth]{particle_filter2}
+\end{column}
+\end{columns}
+\end{frame}
+
 \begin{frame}{Construction}
@@
         $$P(x_k) = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
-    \item Particle Weights at time $t_k$:\\
+    \item Weight of Particle $i$, at time $t_k$:\\
         $$w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}$$
@@
       \note{$p$ is the actual probability }
-    \item Incorporating Measurement:\\
+    \item Incorporating a Measurement, $y_k$:\\
         $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
@@
 
-\begin{frame}{Particle Filter }
+\section{Methods}
+\begin{frame}{Particle Filter Setup}
 \footnotesize
 \begin{itemize}
-    \item 
+    \item Weighting function
+        \begin{itemize}
+            \item Continuous, Long Tailed, Zero-Mean 
+            \item Too wide $\rightarrow$ under-sensitivity, 
+                slow or no convergence
+            \item Too thin $\rightarrow$ reduces robustness 
+                to noise, particle deprivation
+            \item For this work, $N(0, 0.005^2)$ used
         \end{itemize}
+    \item Resampling 
+        \begin{itemize}
+            \item Stratified Resampling can result in truncated tails 
+                    on posterior
+            \item Regularized Resampling can result in over smoothing and 
+                    slow convergence
+            \item Resampling Performed Rarely
+        \end{itemize}
+    \item Number of particles
+        \begin{itemize}
+            \item More particles give higher fidelity 
+            \item Large Initial Particle Count (16000)
+        \end{itemize}
+    \item Prior Distribution
+    \begin{itemize}
+            \item Large $\tau$ parameters can oversmooth
+            \item $\epsilon$ needs large variance
+    \end{itemize}
+\end{itemize}
+\note{Because the first 20 measurements drop most weights very quickly this gives a large breadth,
+    but after the initial calculations drops the computation requirements}
+\note{This has not been done before}
 \end{frame}
 
-\section{Methods}
 \begin{frame}{FMRI Noise}
@@
 \section{Conclusion}
-\begin{frame}{Factors Affecting Convergence}
-    \begin{itemize}
-        \item Weighting function
-            \begin{itemize}
-                \item Continuous, Long Tailed, Zero-Mean 
-                \item Too wide $\rightarrow$ under-sensitivity, 
-                    slow or no convergence
-                \item Too thin $\rightarrow$ reduces robustness 
-                    to noise, particle deprivation
-            \end{itemize}
-        \item Resampling 
-            \begin{itemize}
-                \item Stratified Resampling can result in truncated tails 
-                        on posterior
-                \item Regularized Resampling can result in over smoothing and 
-                        slow convergence
-            \end{itemize}
-        \item Number of particles
-            \begin{itemize}
-                \item More particles give higher fidelity posterior
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
 \begin{frame}{Conclusion}
@@
             et al. \cite{Deneux2006} shows that parameters could imposed.
-        \item Automation of Weighting Function
         \item Analysis of joint parameter distribution for populations

### 2010-08-28 doc/presentation/introduction.tex
split some stuff, changed the preprocessing image, and started adding one more picture for the particle filter.

--- 2010-08-28 doc/presentation/main.tex
@@
-%Micah Chambers
-
-\documentclass{beamer}
-%\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5]{beamerposter}
-
-\usepackage{pgfpages}
-%\setbeameroption{show notes on second screen}
-
-\mode<presentation>
-{
-  \usetheme{Antibes}
-  % or ...
-
-  %\setbeamercovered{transparent}
-  % or whatever (possibly just delete it)
-}
-
-%\usefonttheme[onlylarge]{structurebold}
-%\usecolortheme{crane}
-%\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
-%\setbeamertemplate{navigation symbols}{}
-%\setbeamercovered{transparent}
-
-\usepackage{color}
-\usepackage{colortbl}
-\usepackage[pdftex]{graphicx}
-\graphicspath{{pdf/}{png/}}
-%\usepackage{times}
-%\usepackage[T1]{fontenc}
-%\usepackage[small]{caption}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{subfigure}
-
-\definecolor{mblue}{RGB}{178,34,34}
-\definecolor{mred}{RGB}{0,0,128}
-\definecolor{mgreen}{RGB}{0,100,0}
-\definecolor{myellow}{RGB}{176,169,0}
-
-\title{Full Brain BOLD Signal Parameter Estimation \\
-using Particle Filters}
-
-%\subtitle{}
-
-\author{Micah Chambers}
-\institute{Virginia Tech Bioimaging Systems Lab}
-
-\subject{Medical Imaging}
-
-% If you have a file called "university-logo-filename.xxx", where xxx
-% is a graphic format that can be processed by latex or pdflatex,
-% resp., then you can add a logo as follows:
-
-%\pgfdeclareimage[width=1.5cm]{university-logo}{logo}
-\logo{\includegraphics[width=1.5cm]{logo}}
-
-% If you wish to uncover everything in a step-wise fashion, uncomment
-% the following command:
-
-%\beamerdefaultoverlayspecification{<+->}
-
-\AtBeginSection{
-\begin{frame}
-\begin{center}
-\structure{\Huge \insertsection}
-\end{center}
-\end{frame}
-}
-
-\AtBeginSubsection{
-\begin{frame}
-\begin{center}
-\structure{\Huge \insertsubsection}
-\end{center}
-\end{frame}
-}
-
-\begin{document}
-\begin{frame}
-  \titlepage
-\end{frame}
-
-%0
-\begin{frame}{Outline}
-  \tableofcontents
-  % You might wish to add the option [pausesections]
-  \note{Introduction describes FMRI and the Balloon Model}
-  \note{Potential Approaches - describes previous used methods of analysis}
-  \note{Particle Filter - derives and discusses the BOLD model}
-  \note{Methods - The methods used in this work, how to recreate.}
-\end{frame}
-
 \section{Introduction}
@@
 \note{Its been known for about 20 years that changes in Blood Oxygen Level alter
-        MR signals in certain types of imaging. The basis of contrast in fMRI is 
+        T2$^*$ relaxation times. The basis of contrast in fMRI is 
         changes in the so-called spin-spin relaxation (or T2$^*$ relaxation time
@@
 \end{frame}
-
-\section{Nonlinear Regression}
-\subsection{Prior Works}
-\begin{frame}{Approximation Method}
-  \begin{itemize}
-    \item 2nd Order Volterra Kernel \cite{Friston2000}
-    \note{Quadratic Approximation}
-    \begin{itemize}
-        \item Quadratic Convolution used to approximate Jacobian Matrix.
-        \note{Because of the nonlinearities in the BOLD model, step size for
-                integrating would otherwise be extremely small}
-        \item Volterra approximation quality is not known.
-    \end{itemize}
-    \note{Volterra Approximation depends on sparsity etc}
-    {\footnotesize
-    $$y(t) = k_0 + \int_{-\infty}^{\infty} k_1(s_1) x(t-s_1) ds_1
-        + \int_{-\infty}^{\infty} k_2(s_1,s_2) x(t-s_1)x(t-s_2) ds_1 ds_2$$
-        \note{Its possible this isn't even knowable}
-    }
-    \item Canonical Hemodynamic Response Function
-    \note{Linear Approximation}
-    \begin{itemize}
-        \item No Parameters Estimated
-        \item Maximum Likelihood Possible
-        \item Inflexible - even to changes in onset time
-    \end{itemize}
-    \begin{center}
-    \includegraphics[clip=true,trim=4.88cm 0cm 5cm 1.5cm,height=2.7cm,width=8cm]{HRF}
-    \end{center}
-
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Nonlinear Methods}
-\begin{itemize}
-    \begin{columns}
-    \begin{column}{.42\textwidth}
-    \item Local Linearization filter, \cite{Riera2003}
-    $$f(t)- f(t-1) \sim N(0, \sigma^2)$$
-     \note{Jacobian of the $\dot{x}$ is available, making gradient descent possible}
-    \item Genetic Algorithms and Simulated Annealing, \cite{Vakorin2007}
-     \note{Genetic Algorithms were used to optimize spike points, simulated annealing for the parameters}
-    \item Particle Filters to estimate States, \cite{Murray2009}
-    \begin{itemize}
-        \item ML estimate of $\Theta$, \cite{Johnston2007}
-         \note{Particle Filter used to estimate states, ML of parameters based on the posterior of states and so on}
-    \end{itemize}
-    \item Unscented Kalman Filter \cite{Hu2009}
-     \note{Technically Kalman Filters and Particle Filters are Bayesian Filters}
-     \note{Essentially filtering out inconsistent parameters.}
-    \end{column}
-
-    \begin{column}{.58\textwidth}
-    \begin{figure}
-    \includegraphics[clip=true,trim=0cm 0cm 0cm 1cm,width=.6\textwidth]{simulated_annealing}
-    \caption{Simulated Annealing can escape local minima with chaotic jumps. \cite{Dama}}
-    \end{figure}
-    \begin{figure}
-    \includegraphics[width=\textwidth]{kalman}
-    \caption{UKF: (a) Initial Belief  (b) Noisy Measurement  (c) incorporated into the belief. \cite{Thrun2005}}
-    \end{figure}
-    \end{column}
-    \end{columns}
-\end{itemize}
-\end{frame}
-
-\subsection{Particle Filter}
-%frame with example
-\begin{frame}{Example System Identification}
-\begin{itemize}
-\item Given:
-    \begin{itemize}
-    \item Elevation Map (1D)
-    \item Ability To Measure Elevation
-    \end{itemize}
-\item Particle Consists of the unknowns:
-    \begin{itemize}
-    \item State: Location  $[0, 20]$
-    \item Constant: Direction  $\{Left, Right\}$
-    \end{itemize}
-\end{itemize}
-
-\begin{center}
-\includegraphics[clip=true,trim=2cm 3cm 3cm 3cm,width=.7\textwidth]{setup}
-\end{center}
-\end{frame}
-
-\begin{frame}{Initial Distribution}
-\centering
-\includegraphics[width=\textwidth]{init}\\
-\end{frame}
-
-\begin{frame}{Measurement 1}
-\centering
-\includegraphics[width=\textwidth]{meas1}\\
-\end{frame}
-
-\begin{frame}{Step Forward}
-\centering
-\includegraphics[width=\textwidth]{move1}\\
-\end{frame}
-
-\begin{frame}{Measurement 2}
-\centering
-\includegraphics[width=\textwidth]{meas2}\\
-\end{frame}
-
-\begin{frame}{Step Forward}
-\centering
-\includegraphics[width=\textwidth]{move2}\\
-\end{frame}
-
-\begin{frame}{Particle Filter Concepts}
-\small
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{itemize}
-    \item Weighting Function
-    \item Resampling
-    \begin{itemize}
-        \item Remove Useless Particles
-    \end{itemize}
-    \item Regularized Resampling
-    \begin{itemize}
-        \item Prevent Redundant Particles
-        \item Reduces Quantization Error
-    \end{itemize}
-\end{itemize}
-\end{column}
-
-\begin{column}{.5\textwidth}
-\includegraphics[clip=true,trim=0cm 6cm 0cm 6cm,width=\textwidth]{particle_filter2}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Construction}
-\begin{itemize}
-    \item Mixture PDF at time $k$, with $N_p$ particles:\\
-        $$P(x_k) = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
-    \item Weight of Particle $i$, at time $t_k$:\\
-        $$w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}$$
-      \note{$q$ is the density/location of the particle}
-      \note{$p$ is the actual probability }
-    \item Incorporating a Measurement, $y_k$:\\
-        $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
-      \note{Because of the proportion, note that the weights do not
-            have to integrate to 1, and can be extremely large or extremely small}
-\end{itemize}
-\end{frame}
-
-\section{Methods}
-\begin{frame}{Particle Filter Setup}
-\footnotesize
-\begin{itemize}
-    \item Weighting function
-        \begin{itemize}
-            \item Continuous, Long Tailed, Zero-Mean 
-            \item Too wide $\rightarrow$ under-sensitivity, 
-                slow or no convergence
-            \item Too thin $\rightarrow$ reduces robustness 
-                to noise, particle deprivation
-            \item For this work, $N(0, 0.005^2)$ used
-        \end{itemize}
-    \item Resampling 
-        \begin{itemize}
-            \item Stratified Resampling can result in truncated tails 
-                    on posterior
-            \item Regularized Resampling can result in over smoothing and 
-                    slow convergence
-            \item Resampling Performed Rarely
-        \end{itemize}
-    \item Number of particles
-        \begin{itemize}
-            \item More particles give higher fidelity 
-            \item Large Initial Particle Count (16000)
-        \end{itemize}
-    \item Prior Distribution
-    \begin{itemize}
-            \item Large $\tau$ parameters can oversmooth
-            \item $\epsilon$ needs large variance
-    \end{itemize}
-\end{itemize}
-\note{Because the first 20 measurements drop most weights very quickly this gives a large breadth,
-    but after the initial calculations drops the computation requirements}
-\note{This has not been done before}
-\end{frame}
-
-\begin{frame}{FMRI Noise}
-\centering
-
-Resting State Noise
-
-\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009_22_38_23}
-
-Resting State Noise Steps
-
-\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009d_22_38_23}
-
-\note{Tanabe 2002 \autoref{Tanabe2002} 10-15\% of voxels exhibit significant drift}
-\note{Occurs in Cadavers.}
-\end{frame}
-
-\begin{frame}{Preprocessing}
-  \begin{columns}
-    \begin{column}{.6\textwidth}
-        \begin{itemize}
-            \item Drop Initial Volumes (9, $18.9$s)
-            \item Realign Over Time
-            \item Detrend (SPM uses $0.0078125 Hz$ cut off)
-            \item Gaussian Smoothing (SPM Only)
-            \begin{itemize}
-                \item Imposes Gaussianity
-                \item Increases SNR
-                \item Reduces Bonferroni Correction Requirement
-            \end{itemize}
-        \end{itemize}
-    \end{column}
-
-    \begin{column}{.4\textwidth}
-        \includegraphics[width=\textwidth]{preprocess}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\section{Results}
-\subsection{40 Minute Single Voxel Simulation}
-\begin{frame}{Long Run Results}
-\includegraphics[width=\textwidth]{long_converge}
-\note{Bars drop well before the end}
-\note{Mean is good estimator}
-\note{Mean comes very close to truth}
-\end{frame}
-
-\begin{frame}{Long Run Covariance and Correlation}
-\centering
-Estimated Parameter Covariance
-
-\begin{table}[t]
-\tiny
-\begin{tabular}{|c | c  c  c  c  c  c  c |}
-\hline
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-\hline
-\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
-$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
-\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
-$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
-$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
-\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
-\hline
-\end{tabular}
-\label{tab:long_cov}
-\end{table}
-
-Estimated Parameter Correlation
-
-\begin{table}[t]
-\tiny
-\begin{tabular}{|c | c  c  c  c  c  c  |}
-\hline
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
-\hline
-\rowcolor[gray]{.8} $\tau_0$  & & & & & & \\
-$\alpha$                      & 0.889884 & & & & & \\
-\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & \\
-$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & \\
-$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & \\
-\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 \\
-\hline
-\end{tabular}
-\note{Correlation of parameter estimates at the end of 
-            \autoref{fig:long_converge}.}
-\label{tab:long_corr}
-\end{table}
-\end{frame}
-
-\subsection{5 Minute, Single Voxel Simulation}
-\begin{frame}{Preprocessed Noisy Simulation vs. Underlying Signal}
-\begin{center}
-\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
-            {preprocessed_lownoise}
-\end{center}
-\end{frame}
-
-\begin{frame}{Estimated vs. Underlying Signal}
-\begin{center}
-\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
-            {comparison_lownoise}
-\end{center}
-\end{frame}
-
-\begin{frame}{Estimated Parameters}
-\begin{table}[t]
-\centering
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
-\hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Error & Residual\\
-\hline
-\rowcolor[gray]{.8}
-1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
-\hline
-\hline
-1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
-1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
-1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
-1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
-1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
-1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
-1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
-1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
-1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
-1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
-1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
-\hline                                                                               
-1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
-\hline
-\end{tabular}
-}
-\end{table}
-\end{frame}
-
-\subsection{POSSUM Simulation}
-\begin{frame}{Mutual Information Compared with SNR}
-\begin{figure}
-\centering
-\subfigure[Regions]{\includegraphics[height=.46\textheight]{simregions}}
-\subfigure[Mutual Information]
-{\includegraphics[height=.46\textheight]{sim_hm_15_6}}
-\subfigure[SNR]{\includegraphics[height=.46\textheight]{snr_hm}}
-\note{Mean SNR's Region 1: $0.8$, Region 2: $0.97$, and Region 3: $0.39$.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{POSSUM Results: Histogram}
-\begin{figure}
-%\centering
-\includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=.9\textwidth]{sec2hist}
-\note{Histogram of estimated parameters.}
-\note{Average SNR: $0.97$}
-\note{Mutual Information greater than $0.15$.}
-\end{figure}
-\end{frame}
-
-\subsection{Real FMRI Data}
-\begin{frame}{SPM vs. Mutual Information Map}
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{figure}
-\setcounter{subfigure}{0}
-
-\subfigure[SPM Results]{\includegraphics[width=.9\textwidth]{spm_hm}}
-\subfigure{\includegraphics[width=.07\textwidth]{scale1}}
-\note{SPM results. Units of activation are in Student's $t$-scores; higher 
-        indicates higher assurance that the signal cannot have occurred 
-        through noise alone.}
-\label{fig:hm_canon_spm}
-\end{figure}
-\end{column}
-
-\begin{column}{.5\textwidth}
-\begin{figure}
-\setcounter{subfigure}{1}
-\subfigure[Particle Filter Results]{\includegraphics[width=.9\textwidth]{hm_mi_strict}}
-\subfigure{\includegraphics[height=1.5cm,width=.07\textwidth]{scale4}}
-\note{Particle filter results. Units of activation are in mutual information;
-    higher indicates more assured activation.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Particle Filter Results: Histogram}
-\begin{figure}
-\centering
-\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=.8\textwidth]{realhist}
-\note{Histogram of parameters in active regions ($M.I. > .15$).}
-\end{figure}
-\end{frame}
-
-\section{Conclusion}
-\begin{frame}{Conclusion}
-\begin{itemize}
-    \item Summary:
-    \begin{itemize}
-    \item BOLD Parameters Ill-Defined
-    \item Particle Filter Capable of good parameter BOLD estimate with 1000 particles
-    \item Mutual Information performs well as estimate of Quality
-    \end{itemize}
-    \item Future Works
-    \begin{itemize}
-        \item Further limitations should be placed on priors, Deneux 
-            et al. \cite{Deneux2006} shows that parameters could imposed.
-        \item Analysis of joint parameter distribution for populations
-        \item Real Time analysis possible for multiple Voxels, similar to
-            De Charms et al. \cite{DeCharms2005}
-
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Questions?}
-\end{frame}
-
-%% All of the following is optional and typically not needed. 
-%\appendix
-%\section<presentation>*{\appendixname}
-%\subsection<presentation>*{For Further Reading}
-
-\begin{frame}[allowframebreaks]
-  \bibliographystyle{plain}
-  \bibliography{../library}
-\end{frame}
-
-\appendix
-\begin{frame}{SPM vs. Mutual Information Map, SPM}
-\begin{figure}
-\setcounter{subfigure}{0}
-\subfigure[SPM Results]{\includegraphics[width=.8\textwidth]{images/spm_hm}}
-\subfigure{\includegraphics[scale=.5]{images/scale1}}
-\note{SPM results. Units of activation are in Student's $t$-scores; higher 
-        indicates higher assurance that the signal cannot have occurred 
-        through noise alone.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{SPM vs. Mutual Information Map, $M.I. > .15$}
-\begin{figure}
-\setcounter{subfigure}{1}
-\subfigure[Particle Filter Results]{\includegraphics[width=.8\textwidth]{hm_mi_strict}}
-\subfigure{\includegraphics[width=.1\textwidth]{scale4}}
-\note{Particle filter results. Units of activation are in mutual information;
-    higher indicates more assured activation.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{SPM vs. Mutual Information Map, $M.I. > .1$}
-\begin{figure}
-\setcounter{subfigure}{2}
-\subfigure[Particle Filter Results]{\includegraphics[width=.8\textwidth]{mi_hm}}
-\subfigure{\includegraphics[width=.1\textwidth]{scale6}}
-\note{Particle filter results. Units of activation are in mutual information;
-    higher indicates more assured activation.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{1: 37-14-7}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/1_pfilter_37_14_7}}
-\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/1_spm_37_14_7}}
-\caption{Section 1, Estimated vs. Actual BOLD response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
-\label{fig:comp1}
-\end{figure}
-\end{frame}
-
-\begin{frame}{2: 34-12-7}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/2_pfilter_34_12_7}}
-\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/2_spm_34_12_7}}
-\caption{Section 2, Estimated vs. Actual BOLD response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
-\label{fig:comp2}
-\end{figure}
-\end{frame}
-
-\begin{frame}{3: 23-21-7}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/3_pfilter_23_21_7}}
-\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/3_spm_23_21_7}}
-\caption{Section 3, Estimated vs. Actual BOLD response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
-\label{fig:comp3}
-\end{figure}
-\end{frame}
-
-\begin{frame}{4: 33-40-4}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/4_pfilter_26_15_7}}
-\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/4_spm_26_15_7}}
-\caption{Section 4, Estimated vs. Actual BOLD response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
-\label{fig:comp4}
-\end{figure}
-\end{frame}
-
-\begin{frame}{5: 29-9-13}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/5_pfilter_29_9_13}}
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/5_spm_29_9_13}}
-\caption{Section 5, Estimated vs. Actual BOLD Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
-%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
-%of $4.17399$ and $MSE$ of $1.14171$.}
-\label{fig:comp5}
-\end{figure}
-\end{frame}
-
-\begin{frame}{6: 36-17-19}
-\setcounter{subfigure}{0}
-\begin{figure}
-\centering
-\subfigure[Particle Filter]{\includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/6_pfilter_36_17_19}}
-\subfigure[SPM]{\includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=.4\textwidth]{images/6_spm_36_17_19}}
-\caption{Section 6, Estimated vs. Actual BOLD Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
-%\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
-\end{figure}
-\end{frame}
-
-\end{document}

### 2010-08-28 doc/presentation/main.tex
split some stuff, changed the preprocessing image, and started adding one more picture for the particle filter.

--- 2010-08-27 doc/presentation/main.tex
@@
 %\usepackage[small]{caption}
+\usepackage{algorithm}
 \usepackage{algorithmic}
 \usepackage{subfigure}
+
+\usepackage{amsmath}	% math fonts
+\usepackage{amsthm}
+\usepackage{amsfonts}
+
+%plots
+\usepackage{tikz}
+
+\definecolor{mblue}{RGB}{178,34,34}
+\definecolor{mred}{RGB}{0,0,128}
+\definecolor{mgreen}{RGB}{0,100,0}
+\definecolor{myellow}{RGB}{176,169,0}
 
@@
 
+\AtBeginSection{
+\begin{frame}
+\begin{center}
+\structure{\Huge \insertsection}
+\end{center}
+\end{frame}
+}
+
+\AtBeginSubsection{
+\begin{frame}
+\begin{center}
+\structure{\Huge \insertsubsection}
+\end{center}
+\end{frame}
+}
+
 \begin{document}
@@
 
-\section{Introduction}
-\begin{frame}{Balloon Model}
-\begin{figure}
-\includegraphics[width=10cm]{model}
-\caption{
-    \tiny
-    \cite{Riera2003}
-}
-\end{figure}
-\note{Rates from below 1 Hz to 5 to 6 Hundred Hz}
-\note{Change in average firing rates increase metabolism and sympathetic response causes
-        increased inflow of oxygenated blood.}
-\note{In the version of the Balloon model I used, the flow is locked to metabolism.}
-\note{According the balloon model this increases the local blood volume, and the outflow pressure.}
-\note{So the signal is altered by changes in the ratio of Deoxyhemoglobin and Oxygenated Hemoglobin,
-        which are a function of Volume and Oxygenation.}
-\end{frame}
-
-\begin{frame}{Time Changing Parameters}
-\begin{figure}
-    \includegraphics[width=\textwidth]{graphs}
-\caption{
-    \tiny
-    \cite{Buxton2004}
-}
-\end{figure}
-\end{frame}
-
-%more expansion of all the parameters
-\begin{frame}{BOLD State Space Model}
-  \scriptsize
-  \begin{columns}
-  \begin{column}{.7\textwidth}
-    \begin{itemize}
-    \item State Vector:
-    \begin{eqnarray}
-    \dot{s}(t) &=& \epsilon u(t) - s(t)/\tau_s - (f(t)-1)/\tau_f \nonumber \\
-    \dot{f}(t) &=& s \nonumber \\
-    \dot{v}(t) &=& (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) \nonumber \\
-    \dot{q}(t) &=& \frac{1}{\tau_0}\left(\frac{f(t)(1-(1-E_0)^{1/f(t)})}{E_0} -
-            \frac{q(t)}{v(t)^{1-1/\alpha}}\right) \nonumber 
-    \end{eqnarray}
-    \item Output:
-    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
-    \end{itemize}
-  \end{column}
-
-  \begin{column}{.3\textwidth}
-        State Variables:
-        $$s(t), f(t), v(t), q(t)$$
-        Parameters:
-        $$\epsilon, \tau_s, \tau_f, \alpha, \tau_0, E_0, V_0$$
-        Constants:
-        $$a_1, a_2$$
-        Input:
-        $$u(t)$$
-  \end{column}
-  \end{columns}
-  \note{System is dissipative:\\
-  If no input is provided, it will rather quickly decay.\\
-  Constant input leads to steady state response.\\
-  First term in $\dot{q}$ is the metabolism term.\\
-  Outflow is the $v^{1/\alpha}$ term\\
-  Significant nonlinearity.\\
-  $a_1 = (4.3\nu_0TE+\epsilon_025\epsilon_0)E_0$, $a_2 = \epsilon_025\epsilon_0E_0 + .43$}
-\end{frame}
-
-\section{Potential Approaches}
-\begin{frame}{Approximate Method}
-  \begin{itemize}
-    \item 2nd Order Volterra Kernel \cite{Friston2000}
-    \note{Quadratic Approximation}
-    \begin{itemize}
-        \item Quadratic Convolution used to approximate Jacobian Matrix.
-        \note{Because of the nonlinearities in the BOLD model, step size for
-                integrating would otherwise be extremely small}
-        \item Volterra approximation quality is not known.
-    \end{itemize}
-    \note{Volterra Approximation depends on sparsity etc}
-    {\footnotesize
-    $$y(t) = k_0 + \int_{-\infty}^{\infty} k_1(s_1) x(t-s_1) ds_1
-        + \int_{-\infty}^{\infty} k_2(s_1,s_2) x(t-s_1)x(t-s_2) ds_1 ds_2$$
-        \note{Its possible this isn't even knowable}
-    }
-    \item Canonical Hemodynamic Response Function
-    \note{Linear Approximation}
-    \begin{itemize}
-        \item No Parameters Estimated
-        \item Maximum Likelihood Possible
-        \item Inflexible - even to changes in onset time
-    \end{itemize}
-    \begin{center}
-    \includegraphics[clip=true,trim=4.88cm 0cm 5cm 1.5cm,height=2.7cm,width=8cm]{HRF}
-    \end{center}
-
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Nonlinear Modeling}
-\begin{itemize}
-    \item Local Linearization filter \cite{Riera2003}
-     \note{Jacobian of the $\dot{x}$ is available, making gradient descent possible}
-    \item Genetic Algorithms and Simulated Annealing \cite{Vakorin2007}
-     \note{Genetic Algorithms were used to optimize spike points, simulated annealing for the parameters}
-    \item Particle Filters to estimate $P(X_t | \Theta, Y_T)$ \cite{Murray2009}
-    \begin{itemize}
-        \item ML estimate of $\Theta$ based on $P(X_t | \Theta, Y_t)$ \cite{Johnston2007}
-         \note{Particle Filter used to estimate states, ML of parameters based on the posterior of states and so on}
-    \end{itemize}
-    \item Unscented Kalman Filter to estimate parameters \cite{Hu2009}
-    \item Direct Particle Filter, used in this work
-     \note{Technically Kalman Filters and Particle Filters are Bayesian Filters}
-     \note{Essentially filtering out inconsistent parameters.}
-\end{itemize}
-\end{frame}
-
-\section{Particle Filter}
-%frame with example
-
-\begin{frame}{Construction}
-\begin{itemize}
-    \item Mixture PDF:\\
-        $$\alpha(X) \stackrel{\text{\tiny{k=0}}}{\approx} P(x_k) 
-                = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
-    \item Particle Weights at time $t_k$:\\
-        $$w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}$$
-      \note{$q$ is the density/location of the particle}
-      \note{$p$ is the actual probability }
-    \item Adding Measurements:\\
-        $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
-      \note{Because of the proportion, note that the weights do not
-            have to integrate to 1, and can be extremely large or extremely small}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Particle Filter Algorithm}
-\footnotesize
-\begin{algorithm}
-\begin{algorithmic}
-\STATE Initialize Particles:
-\FOR{$i$ : each of $N_p$ particles }
-    \STATE $x^i_0  \sim \alpha(X)$
-    \STATE $w^i_0 = \frac{1}{N_p}$
-\ENDFOR
-\FOR{$k$ : each measurement}
-    \FOR{$i$ : each particle }
-        \STATE $x^i_k = x^i_{k-1} + \int_{t-1}^t f(x(\tau), u(\tau)) d\tau $
-        \STATE $w^i_k = w^i_{k-1}P(y_k | x_k)$
-    \ENDFOR
-    \STATE Calculate $N_{eff}$ 
-    \IF{$N_{eff} < N_R$ (recommend $N_R = min(50, .1N_p)$ )}
-        \STATE Regularized Resample 
-    \ENDIF
-\ENDFOR
-
-\STATE At $t + \Delta t$, $t \in T$, $P(x(t+\Delta t)) \approx 
-\sum_{i=1}^{N_p} w_i(t)\delta\left(x - (x_i(t) + \int_t^{t+\Delta t} f(x(\tau), u(\tau)) d\tau) \right)$
- \end{algorithmic}
- \end{algorithm}
-\end{frame}
-
-\begin{frame}{Details and Example}
-  \begin{columns}
-    \begin{column}{.55\textwidth}
-        \begin{itemize}
-            \item Non-parametric
-            \item Model based
-            \item Online
-            \item Provides Full Posterior Probability
-            \item Non-trivial computation cost
-            \item Regularized Particle Filter Used
-        \end{itemize}
-    \end{column}
-    \begin{column}{.45\textwidth}
-        \begin{figure}
-          \includegraphics[width=\textwidth]{particle_filter}
-        \end{figure}
-    \end{column}
-    \end{columns}
-\end{frame}
-
-\section{Methods}
-\begin{frame}{FMRI Noise}
-\centering
-
-Resting State Noise
-
-\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009_22_38_23}
-
-Resting State Noise Steps
-
-\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009d_22_38_23}
-
-\note{Tanabe 2002 \autoref{Tanabe2002} 10-15\% of voxels exhibit significant drift}
-\note{Occurs in Cadavers.}
-\end{frame}
-
-\begin{frame}{Preprocessing}
-  \begin{columns}
-    \begin{column}{.6\textwidth}
-        \begin{itemize}
-            \item Drop Inital Volumes (9, $18.9$s)
-            \item Realign Over Time
-            \item Detrend (SPM uses $0.0078125 Hz$ cut off)
-            \item Gaussian Smoothing (SPM Only)
-            \begin{itemize}
-                \item Imposes Gaussianity
-                \item Increases SNR
-                \item Reduces Bonferroni Correction Requirement
-            \end{itemize}
-        \end{itemize}
-    \end{column}
-
-    \begin{column}{.4\textwidth}
-        \includegraphics[width=\textwidth]{preprocess}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\section{Results}
-\begin{frame}{Long Run Results}
-\includegraphics[width=\textwidth]{long_converge}
-\note{Bars drop well before the end}
-\note{Mean is good estimator}
-\note{Mean comes very close to truth}
-\end{frame}
-
-\begin{frame}{Long Run Correlation}
-\centering
-Estimated Parameter Covariance
-
-\begin{table}[t]
-\tiny
-\begin{tabular}{|c | c  c  c  c  c  c  c |}
-\hline
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
-\hline
-\rowcolor[gray]{.8} $\tau_0$  & 0.0004334 & 5.2e-05 & -6.95e-05 & 3.3e-06 & 0.0001628 & -2e-07 & 0.0001798 \\
-$\alpha$                      & 5.2e-05 & 7.9e-06 & -6.4e-06 & 3e-07 & 1.04e-05 & -1.92e-05 & 2.58e-05 \\
-\rowcolor[gray]{.8} $E_0$     & -6.95e-05 & -6.4e-06 & 1.9e-05 & -9e-07 & -4.11e-05 & -3.24e-05 & -3.92e-05 \\
-$V_0$                         & 3.3e-06 & 3e-07 & -9e-07 & 1e-07 & 1.1e-06 & 9e-07 & 1e-06 \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.0001628 & 1.04e-05 & -4.11e-05 & 1.1e-06 & 0.0001589 & 0.0001518 & 7.88e-05 \\
-$\tau_f$                      & -2e-07 & -1.92e-05 & -3.24e-05 & 9e-07 & 0.0001518 & 0.0002966 & -2.34e-05 \\
-\rowcolor[gray]{.8} $\epsilon$& 0.0001798 & 2.58e-05 & -3.92e-05 & 1e-06 & 7.88e-05 & -2.34e-05 & 0.0001966 \\
-\hline
-\end{tabular}
-\label{tab:long_cov}
-\end{table}
-
-Estimated Parameter Correlation
-
-\begin{table}[t]
-\tiny
-\begin{tabular}{|c | c  c  c  c  c  c  |}
-\hline
-  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
-\hline
-\rowcolor[gray]{.8} $\tau_0$  & & & & & & \\
-$\alpha$                      & 0.889884 & & & & & \\
-\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & \\
-$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & \\
-\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & \\
-$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & \\
-\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 \\
-\hline
-\end{tabular}
-\note{Correlation of parameter estimates at the end of 
-            \autoref{fig:long_converge}.}
-\label{tab:long_corr}
-\end{table}
-\end{frame}
-
-\begin{frame}{Short Run Results A}
-\begin{center}
-\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
-            {preprocessed_lownoise}
-\end{center}
-\end{frame}
-
-\begin{frame}{Short Run Results B}
-\begin{center}
-\includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=.9\textwidth]
-            {comparison_lownoise}
-\end{center}
-\end{frame}
-
-\begin{frame}{Short Run Results C}
-\begin{table}[t]
-\centering
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
-\hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Error & Residual\\
-\hline
-\rowcolor[gray]{.8}
-1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
-\hline
-\hline
-1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
-1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
-1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
-1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
-1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
-1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
-1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
-1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
-1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
-1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
-1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
-\hline                                                                               
-1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
-\hline
-\end{tabular}
-}
-\end{table}
-\end{frame}
-
-\begin{frame}{POSSUM Results: Activation}
-\begin{figure}
-\centering
-\subfigure[Regions]{\includegraphics[height=.46\textheight]{simregions}}
-\subfigure[Mutual Information]
-{\includegraphics[height=.46\textheight]{sim_hm_15_6}}
-\subfigure[SNR]{\includegraphics[height=.46\textheight]{snr_hm}}
-\note{Mean SNR's Region 1: $0.8$, Region 2: $0.97$, and Region 3: $0.39$.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{POSSUM Results: Histogram}
-\begin{figure}
-%\centering
-\includegraphics[clip=true,trim=5cm 2cm 5cm 2cm, width=\textwidth]{sec2hist}
-\note{Histogram of estimated parameters.}
-\note{Average SNR: $0.97$}
-\note{Mutual Information greater than $0.15$.}
-\end{figure}
-\end{frame}
-
-\begin{frame}{SPM vs. Mutual Information Map}
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{figure}
-\setcounter{subfigure}{0}
-
-\subfigure[SPM Results]{\includegraphics[width=.9\textwidth]{spm_hm}}
-\subfigure{\includegraphics[width=.07\textwidth]{scale1}}
-\note{SPM results. Units of activation are in Student's $t$-scores; higher 
-        indicates higher assurance that the signal cannot have occurred 
-        through noise alone.}
-\label{fig:hm_canon_spm}
-\end{figure}
-\end{column}
-
-\begin{column}{.5\textwidth}
-\begin{figure}
-\setcounter{subfigure}{1}
-\subfigure[Particle Filter Results]{\includegraphics[width=.9\textwidth]{hm_mi_strict}}
-\subfigure{\includegraphics[height=1.5cm,width=.07\textwidth]{scale4}}
-\note{Particle filter results. Units of activation are in mutual information;
-    higher indicates more assured activation.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Particle Filter Results: Histogram}
-\begin{figure}
-\centering
-\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=.8\textwidth]{realhist}
-\note{Histogram of parameters in active regions ($M.I. > .15$).}
-\end{figure}
-\end{frame}
+\include{introduction}
+\include{nonlinear}
+\include{methods}
+\include{results}
 
 \section{Conclusion}
-\begin{frame}{Factors Affecting Convergence}
-    \begin{itemize}
-        \item Weighting function
-            \begin{itemize}
-                \item Continuous, Long Tailed, Zero-Mean 
-                \item Too wide $\rightarrow$ under-sensitivity, 
-                    slow or no convergence
-                \item Too thin $\rightarrow$ reduces robustness 
-                    to noise, particle deprivation
-            \end{itemize}
-        \item Resampling 
-            \begin{itemize}
-                \item Stratified Resampling can result in truncated tails 
-                        on posterior
-                \item Regularized Resampling can result in over smoothing and 
-                        slow convergence
-            \end{itemize}
-        \item Number of particles
-            \begin{itemize}
-                \item More particles give higher fidelity posterior
-            \end{itemize}
-    \end{itemize}
-\end{frame}
-
 \begin{frame}{Conclusion}

### 2010-08-29 doc/paper/main.tex
added discussion

@@
 co-registered across time. The co-registered data was spatially smoothed,
-and then filtered with an adaptive cut off (which is built into the
+and then filtered with a cut off of $0.0078125$ Hz. (which default in
 SPM analysis). For the SPM analysis, the Canonical HRF was used
@@
 
-\section{Conclusion}
-\label{sec:Conclusion}
-
-%This article is organized as follows. The rest of the introduction will
-%overview similar efforts and discuss their strengths ans weaknesses.
-%\autoref{sec:Bold Model} will describe the genesis of the BOLD model,
-%and its current state. \autoref{sec:Particle Filter} will explain the 
-%particle filter and how it is being used in this case. 
-%\auroref{sec:Methods} will describe in detail the experimental
-%design, algorithm set up, and the preprocessing that was applied
-%for the particle filter algorithm.
-%The results are explored separately for simulated data
-%and real FMRI data in \autoref{sec:SimulationResults} and 
-%\autoref{sec:RealData}, respectively. 
-%In \autoref{sec:Discussion} the results and their implications 
-%for future works are interpreted.
+There are two possible causes for the large variance in
+parameter estimates. First, it could be that the particle filter is
+incapable of learning the model. However, given the high quality
+of BOLD estimates, this is almost certainly not the case. The other
+possible explanation is that the given stimulus is not sufficiently
+varied to bring out all the properties of the system. While it is
+possible that this is the case, it is unlikely that any
+stimulus sequence could differentiate all the parameters.
+Its also extremely interesting that the histograms of parameters' mean
+approach the width of the prior, even when the particle filter converges
+properly to far thinner distributions. Its possible the prior
+distribution is far wider than
+it needs to be. Finally, the fact that \autoref{sec:VeryLongSim}, which
+used randomly spaced impulses,
+still had a high correlation further indicates that the parameters
+are ill-defined.
+
+Thus there is little doubt that the parameters of the BOLD model are under
+constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
+it is a notable conclusion \cite{Deneux2006}. Other methods
+that calculate a point estimate of the parameters, such as least squares
+or Kalman filters (which uses the first two moments) are limited in their
+capacity to estimate such parameters. While estimates of
+the  BOLD signal may still be correct, the
+underlying parameters and state variables cannot be described without using
+a joint probability distribution function. In this sense, particle
+filters represent an important step forward in BOLD parameter
+estimation. Representing uncertainty with a mean
+and variance is insufficient; so using a particle filter
+or Bayesian estimate of the posterior is not simply an enhancement,
+but a necessary precaution.
+
+Because of the ill-defined nature of the parameters, the results
+of Friston et al. may have been wider than true parameter distribution
+\cite{Friston2002}. Future works that attempt to calculate parameters
+of the BOLD model should be run multiple times to ensure consistency,
+which most likely cannot be attained. This result is particularly
+troublesome given the number of papers that use those parameters for
+a prior distribution.
+
+As the results in \autoref{sec:RealData} show, the heat maps, especially
+those of mutual information, closely resembled the results of SPM. While the activation
+tests were more sensitive, there were some additional false positives,
+though the problem is difficult to quantify. Regions with M.I. above
+$0.15$ consistently fit the FMRI data well, and simulations showed that
+the particle filter performed extremely well in the face of significant
+noise. In sum, the estimated BOLD output remained consistent despite
+large swings in parameters.
+
+The Particle Filter algorithm was originally designed for on-line parameter
+estimation. For this reason, there is no guarantee of optimality or even
+convergence for finite measurements. However, for the BOLD nonlinear ODE
+this is less of a concern than it might first appear to be. For this
+particular problem there can be no guarantee of a global minimum, and although
+other techniques guarantee a local minimum, tests show that the particle
+filter did converge relatively quickly \autoref{sec:VeryLongSim}.
+
+One difficulty with the use of a particle filter when given a finite number 
+of measurements is finding
+a good weight function $P(y_k | x_k)$. This is more important for a finite
+number of measurements because $P(y_k | x_k)$ needs to converge in finite time.
+In spite of this potential problem, \autoref{sec:VeryLongSim} managed
+to converge in less than 500 seconds.  Given sufficient
+measurements it is better to let the algorithm take longer to converge, because the
+convergence will be less prone to particle deprivation. The particle filter takes longer
+to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
+however, it is free from the uncertainty of whether a quadratic approximation is
+sufficient for the BOLD model.
+
+The particle filter also has advantages over other estimation procedures
+discussed in \autoref{sec:Prior Works}. The most important advantage is that it provides
+an estimate of the posterior probability, rather than a single estimate. While researchers
+often want a simple estimate of parameters, such an estimate is impossible with this particular
+model. The fact that the final distribution does not need to conform to a
+parametric distribution is also advantageous, given the nonlinearities in the system.
+While the particle filter took a day to calculate for full brain calculations, its speed
+was sufficient on a quad core machine to perform real time calculations of small regions
+(approximate run time .27 seconds per voxel-measurement). Today it would be possible
+ to perform real time analysis of 10 voxels on an average quad core. The algorithm also scales
+well and does not require burdensome amounts of memory (approximately 11 megabytes).
+
+There are a number of advantages to the particle filter approach presented
+here. In the past, FMRI data has been analyzed strictly for determining
+correlation between a stimulus and response. With this new method
+the correlation is simply a means to determining the joint posterior distribution
+of the parameters. While only regions that correlate with the input will
+be calculable, this method exploits that correlation to constrain the
+prior distribution of the parameters. The resulting distribution, while difficult
+to visualize because of the high-dimensionality, could nevertheless be
+correlated with neural pathologies. In spite of the fact that the parameters
+are under-determined, the final distribution is still a reduction in the
+uncertainty of the parameters. The availability of a full posterior distribution
+opens up many avenues for further inquiry, for instance
+to compare normal vs. symptomatic
+populations. This would be especially useful in patients whose symptoms
+are caused by functional difference in the brain, rather than structural differences.
+This would effectively be
+a way to differentiate the way the brains operate.
+
+Because of limitations present in every imaging modality,
+its becoming increasingly clear that combining data from multiple sources
+will be necessary to push Neurology forward. In order to do so however, the
+output of each source needs to fully represent what information that source
+can provide. Combining the sources using Bayesian statistics is promising
+yet often difficult because full probability distributions are hard to come by.
+However, in this case, the particle filter provides a full posterior which
+is extremely versatile. Therefore, future works will easily be able to
+plug in data from multiple sources if they all output Bayesian posteriors.
+For instance, if two different modalities have calculated the probability
+distribution of neural efficiency, those two beliefs may be combined into
+one conditional belief for a new conditional probability of neural efficiency.
+
+An advantageous aspect of using a physiological model such as this, is that
+it permits estimates of otherwise hidden parameters. In particular the BOLD
+model gives an estimate of the value of the flow inducing signal, $s$. Having
+this value available opens up new avenues for determining inter-regional
+dependencies. The values of $s$ could serve as a proxy for activation in
+that region and thus could be used to drive other inputs regions' $u(t)$.
+Thus, the particle filter could be re-run
+with the time-series of a particular voxel's $s$ value as an additional stimulus.
+In this way, it could be possible to determine chains of events. This is just
+one possible benefit being able to determine the time course of the hidden
+state variables present in the BOLD equations; the potential benefits of being
+able to determine this information is limitless. Of course the quality of these
+estimates will continue to improve with the model priors.
 

### 2010-08-29 doc/presentation/introduction.tex
notes are in and ready to print

@@
 \end{figure}
-\note{Its been known for about 20 years that changes in Blood Oxygen Level alter
-        T2$^*$ relaxation times. The basis of contrast in fMRI is 
-        changes in the so-called spin-spin relaxation (or T2$^*$ relaxation time
-        of tissue). The rate of loss in nuclear spin coherence depends on the molecules in the voxel being
-        measured \emph{and the surrounding voxels}. For this reason the signal 
-        detected from BLOOD doesn't just depend on the amount of oxygen, but also the
-        amount of De-oxygenated hemoglobin. The question then, is what drives changes
-        in De-oxygenated Hemoglobin and blood volume?
+\note{
+    \begin{itemize}
+        \item FMRI has been around for about 20 years
+        \item 10 Years Since the First BOLD Model
+        \item Basis of FMRI, T2$^*$ relaxation time, blood oxygen levels
+        \item Windkessel/Balloon Model, Volume Changes driving fluctations
+            in blood oxygen 
+        \item Brain has no local oxygen storage, dependent on blood
+        \item brain active, firing rates of neurons increase 10 or 100x
+        \item recharging requires additional oxygen ... metabolism. 
+        \item Increased metabolism drives increased blood flow
+        \item Increased blood flow increases volume
 
-        Well the brain has no local oxygen storage, so it is completely dependent on 
-        fresh Oxygenated-Hemoglobin. When the brain becomes more active, firing rates
-        for neurons increase by several orders of magnitude, and recharging these neurons
-        requires additional oxygen for matabolism. }
-\note{Rates from below 1 Hz to 5 to 6 Hundred Hz}
-\note{Change in average firing rates increase metabolism and sympathetic response causes
-        increased inflow of oxygenated blood.}
-\note{In the version of the Balloon model I used, the flow is locked to metabolism.}
-\note{According the balloon model this increases the local blood volume, and the outflow pressure.}
-\note{So the signal is altered by changes in the ratio of Deoxyhemoglobin and Oxygenated Hemoglobin,
-        which are a function of Volume and Oxygenation.}
+        \item Rates from below 1 Hz to 5 to 6 Hundred Hz
+    \end{itemize}
+}
 \end{frame}
@@
 \end{figure}
+\note{
+    \begin{itemize}
+        \item CMRO2 Increased Metabolism Causes Initial Drop In Oxygen
+        \item CBF Flow Quickly Compensates Increases Oxygen
+        \item CBV Increases, Decides the outflow, local oxygen content
+        \item This delays result in the characteristic BOLD response
+        \item CBF CMRO2 Locked 
+    \end{itemize}
+}
 \end{frame}
@@
   \end{columns}
-  \note{System is dissipative:\\
-  If no input is provided, it will rather quickly decay.\\
-  Constant input leads to steady state response.\\
-  First term in $\dot{q}$ is the metabolism term.\\
-  Outflow is the $v^{1/\alpha}$ term\\
-  Significant nonlinearity.}
+  \note{
+    \tiny
+    \begin{itemize}
+    \item System is dissipative,
+    \item Constant input leads to steady state 
+    \item $u(t)$ is the input
+    \item States
+    \begin{itemize}
+    \item s - Neural Response (Flow Inducing Signal)
+    \item f - Inflow of Oxygenated Blood
+    \item v - Local Blood Volume
+    \item q - Deoxygenated hemoglobin content
+    \end{itemize}
+    \item Parameters
+    \begin{itemize}
+    \item $\epsilon$ - efficiency
+    \item $\tau_s$ - Falloff time for Flow Inducing Signal
+    \item $\tau_f$ - Onset time for Flow Inducing Signal
+    \item $\alpha$ - Grubbs Constant determining outflow volume
+    \item $\tau_0$ - Time Constant for volume, deoxygenated hemoglobin
+    \item $E_0$ - Resting Oxygen Metabolism
+    \item $V_0$ - Resting Metabolism
+    \end{itemize}
+    \item $a_1$,$a_2$ are functions of the imaging modality,
+            and to a vary somewhat with $E_0$. 
+    \item $a_1 = k_1*E_0 + k_2*E_0$
+    \item $a_2 = k_2*E_0 + k_3$
+    \end{itemize}
+  }
 \end{frame}

### 2010-08-29 doc/presentation/main.tex
notes are in and ready to print

@@
 %\setbeameroption{show notes on second screen}
+\setbeameroption{show notes}
 
@@
   % You might wish to add the option [pausesections]
-  \note{Introduction describes FMRI and the Balloon Model}
-  \note{Potential Approaches - describes previous used methods of analysis}
-  \note{Particle Filter - derives and discusses the BOLD model}
-  \note{Methods - The methods used in this work, how to recreate.}
+  \note{
+  \begin{itemize}
+  \item Introduction - FMRI and the Balloon Model
+  \item Nonlinear Regression - overview prior works and the particle filter
+  \item Methods - how the particle was configured, and the preprocessing
+  \item Results - Several different types of simulation and real data 
+  \item Conclusion - Why this is important
+  \end{itemize}
+  }
 \end{frame}
@@
 
-\begin{frame}[allowframebreaks]
-  \bibliographystyle{plain}
-  \bibliography{../library}
+\appendix
+
+\section{Backup}
+\begin{frame}{Balloon Flowchart}
+\begin{figure}
+\includegraphics[width=\textwidth]{backup_balloon}
+\end{figure}
 \end{frame}
 
-\appendix
 \begin{frame}{SPM vs. Mutual Information Map, SPM}
@@
 
+\begin{frame}{Particle Filter Results: Histogram}
+\begin{figure}
+\centering
+\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=.8\textwidth]{realhist}
+\note{Histogram of parameters in active regions ($M.I. > .15$).}
+\end{figure}
+\end{frame}
+
+\section{Bibliography}
+\begin{frame}[allowframebreaks]
+  \bibliographystyle{plain}
+  \bibliography{../library}
+\end{frame}
+
 \end{document}

### 2010-08-30 doc/presentation/run.sh
final

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-impressive -c memory --geometry 2048x768 -t None withnotes.pdf
+impressive -c memory --geometry 1024x768 -t None main.pdf

### 2010-08-31 doc/thesis/intro.tex
added table of acronyms and almost done correcting introduction

@@
 \chapter{Introduction}
+\section{Overview}
 Traditional methods of analyzing timeseries images produced by 
-Functional Magnetic Resonance 
-Imaging (FMRI) perform regression using the linear combination of explanatory variables. 
+\ac{fMRI} perform regression using the linear combination of explanatory variables. 
 Though adding degrees of freedom naturally mandates more computation,
 in this thesis I will demonstrate the use of the Particle Filters to 
-estimate the governing parameters of the BOLD model at a computation cost 
+estimate the governing parameters of the \ac{BOLD} model at a computation cost 
 that would still allow real time calculations for multiple voxels.
@@
 method of detecting neural activity from the traditionally
-used Statistical Parametric Mapping (SPM). Though more computationally intense,
+used \ac{SPM}. Though more computationally intense,
 this method is capable of modeling nonlinear effects, is conceptually simpler
@@
 is possible to estimate parameters and make real-time predictions
-for small neural regions, a feature which could be useful towards real time FMRI 
+for small neural regions, a feature which could be useful towards real time \ac{fMRI}
 \cite{DeCharms2005}. Future works will also benefit from the ability to 
 apply conditions to the posterior distribution in post-processing without
-recalculating parameters; for instance to impose additional 
+recalculating parameters; for instance, to impose additional 
 physiological
-constraints. Modeling the BOLD (Blood-Oxygen-Level-Dependent) response 
+constraints. Modeling the \ac{BOLD} response 
 as a nonlinear system is the
-best way to determine the correlation of stimulus sequence with the BOLD
+best way to determine the correlation of a stimulus sequence with the \ac{BOLD}
 response; yet in the past doing this on a large scale has been far too
@@
 
-This thesis is organized as follows. In the introduction I will introduce
-FMRI, the method by which neural time changing data is detected. This section
-will also describe the basic form of the BOLD model - which drives the 
-detectable changes in MR signal. \autoref{sec:Prior Works} will discuss other
-methods of analyzing FMRI images as well as other techniques that have
+This thesis is organized as follows. The introduction will introduce
+\ac{BOLD}, the method by which neural time changing data is detected. This section
+will also describe the basic form of the \ac{BOLD} model - which drives the 
+detectable changes in \ac{MR} signal. \autoref{sec:Prior Works} will discuss other
+methods of analyzing \ac{fMRI} images as well as other techniques that have
 been, or could be applied to the nonlinear regression model described here. 
@@
 The results are described separately for simulated data
-and real FMRI data in \autoref{sec:SimulationResults} and \autoref{sec:RealData},
+and real \ac{fMRI} data in \autoref{sec:SimulationResults} and \autoref{sec:RealData},
 respectively. In \autoref{sec:Discussion} the results and their implications
@@
 \section{Historic Context}
-For the past twenty years, Functional Magnetic Resonance Imaging (FMRI) 
+\acresetall
+For the past twenty years, \ac{fMRI}
 has been at the forefront of cognitive research. Despite its
-limited temporal resolution, FMRI is the standard tool for localizing 
+limited temporal resolution, \ac{fMRI} is the standard tool for localizing 
 neural activation.  Whereas other methods
 of analyzing neural signals can be invasive or difficult to acquire, 
-FMRI is quick and cheap, and its analysis straight forward.
+\ac{fMRI} is simple to setup quick and inexpensive.
 By modeling the governing equations behind the neural response that
-drives FMRI, it is possible to increase the power of FMRI.
+drives \ac{fMRI}, it is possible to increase the power of \ac{fMRI}.
 The underlying state equations hold important information
@@
 physiology including existing and future pathologies. In short,
-the long chain of events driving FMRI signals contain information 
+the long chain of events driving \ac{fMRI} signals contain information 
 beyond correlation with stimuli.
@@
 In the past fifteen years, a steady stream of studies have built
-on the original Blood Oxygen Level Dependent (BOLD) signal 
+on the original \ac{BOLD} signal 
 derivation first described by Ogawa et al. \cite{Ogawa}.
 The seminal work by Buxton et al. attempted to explain the
-time evolution of the BOLD signal using a windkessel model to
-describe the local changes in Deoxygenated Hemoglobin content \cite{Buxton1998}.
+time evolution of the \ac{BOLD} signal using a windkessel model to
+describe the local changes in \ac{dHb} content \cite{Buxton1998}.
 Incremental improvements were made to this model until Friston et al.
 brought all the changes together into a single complete 
-set of equations \cite{Friston2000}. And while there have been numerous adaptations in the model, 
+set of equations \cite{Friston2000}. While there have been numerous adaptations in the model, 
 many of them summarized by Deneux et al., even the basic versions
-have less bias error than the empirically driven \emph{Canonical Hemodynamic Model}
+have less bias error than the empirically driven Canonical \ac{HRF}
 \cite{Deneux2006,Handwerker2004}.
-On the other hand BOLD signal models have numbers
+On the other hand \ac{BOLD} signal models have numbers
 of parameters ranging from seven \cite{Riera2004} to 50 \cite{Behzadi2005} 
@@
 \label{sec:Introduction Overview}
-Detecting neural activity using the changes in FMRI images is based on 
-the Blood Oxygen Level Dependent (BOLD) signal.
-The BOLD signal is caused by minute changes in the ratio of Deoxygenated
-Hemoglobin to Oxygenated Hemoglobin in blood vessels throughout the brain \cite{Ogawa}.
-Because Deoxygenated Hemoglobin (DHb) is paramagnetic, higher concentrations
-attenuate the signal detected during T2-weighted Magnetic Resonance Imaging (MRI)
-techniques. The most common FMRI imaging technique, due to its rapid repetition 
-time (TR), is Echo Planar Imaging (EPI). When axons becomes active,
-a large amount of ions quickly flow out of the cell. In order for this
+Detecting neural activity using the changes in \ac{fMRI} images is based on 
+the \ac{BOLD} signal.
+The \ac{BOLD} signal is caused by minute changes in the ratio of 
+\ac{dHb} to \ac{O2Hb} in blood vessels throughout the brain \cite{Ogawa}.
+Because \ac{dHb} is paramagnetic, higher concentrations
+attenuate the signal detected during \ac{T2}-weighted \ac{MRI}
+techniques. The most common \ac{fMRI} imaging technique, due to its rapid 
+\ac{TR}, is \ac{EPI}. When axons become active,
+large amounts of ions quickly flow out of the cell. In order for this
 action potential to occur again (and thus for the neuron to fire again),
@@
 occur more often, resulting in a local increase of the 
-Cerebral Metabolic Rate of Oxygen (CMRO2). Thus, if everything else 
-stay the same, blood vessels in an active area will 
-tend to have less oxygenated hemoglobin, and more deoxygenated hemoglobin
+\ac{CMRO2}. Thus, if everything else 
+stays the same, blood vessels in an active area will 
+tend to have less \ac{O2Hb}, and more \ac{dHb}
 (due to the increased rate at which oxygen is being consumed).
-This would then result in an attenuated FMRI signal. However, to
+This would then result in an attenuated \ac{fMRI} signal. However, to
 compensate for activation, muscles that
@@
 This ultimately results in lower than average concentration of 
-deoxyhemoglobin. Thus, the BOLD signal consists of a short initial
-dip in the MR signal, followed by a prolonged increase in signal
+\ac{dHb}. Thus, the \ac{BOLD} signal consists of a short initial
+dip in the \ac{MR} signal, followed by a prolonged increase in signal
 that slowly settles out. It is the overcompensation that provides
-the  primary signal detected with FMRI. This cascade of events
-is believed to drive a prolonged increased in local metabolism, 
-blood flow, blood volume, and oxygenated hemoglobin. The differences
-in onsets and recovery times of these variables are what causes the 
-distinguishing characteristics of the BOLD signal. Unfortunately, 
-FMRI signal levels are all relative: within a particular
+the  primary signal detected with \ac{fMRI}. This cascade of events
+is believed to drive a prolonged increase in local metabolism, 
+blood flow, blood volume, and \ac{O2Hb}. The differences
+in onsets and recovery times of these variables are what cause the 
+distinguishing characteristics of the \ac{BOLD} signal. Unfortunately, 
+\ac{fMRI} signal levels are all relative: within a particular
 person, scanner and run.
 
-\section{FMRI}
-Magnetic Resonance Imaging, MRI, is a method of building 3D images
+\section{\ac{fMRI}}
+\ac{MRI}, is a method of building 3D images
 non-invasively, based on the difference between nuclear spin
@@
 is brought into a large magnetic field which causes nuclear spins
-to align. Radio Frequency (RF) signals may
+to align. \ac{RF} signals may
 then be used to excite nuclear spin away from the base alignment. 
 As the nuclei precess back to the alignment of the magnetic
-field, they emit detectable RF signals. Conveniently, the
+field, they emit detectable \ac{RF} signals. Conveniently, the
 nuclear spins return their original state at different
-rates, called the T1 relaxation time, depending on the material excited.
+rates, called the \ac{T1} relaxation time, depending on the material excited.
 Additionally, the
-coherence of the spins also decay differently (and roughly an order of 
+coherence of the spins also decays differently (and roughly an order of 
 magnitude faster
-than T1 relaxation) based on the properties of the region.
+than \ac{T1} relaxation) based on the properties of the region.
 This gives two primary methods of contrasting substances,
-which form the basis of T1 and T2 weighted images. Additionally, 
-dephasing occurs at two different rates, the T2 relaxation time,
-which is unrecoverable, and T2$^*$ relaxation, which is
-much faster, but possible to recover from via special RF signals.
-T1 relaxation times are typically on the order of seconds if 
-a sufficiently strong excitation is applied, whereas T2$^*$ relaxation
+which form the basis of \ac{T1} and \ac{T2} weighted images. Additionally, 
+dephasing occurs at two different rates, the \ac{T2} relaxation time,
+which is unrecoverable, and \ac{T2}$^*$ relaxation, which is
+much faster, but possible to recover from via special \ac{RF} signals.
+\ac{T1} relaxation times are typically on the order of seconds if 
+a sufficiently strong excitation is applied, whereas \ac{T2}$^*$ relaxation
 times are usually less than 100ms. 
 In order to rapidly acquire entire brain images, as done in Functional 
-MRI, a single large excitation pulse is applied to the entire brain,
-and the entire volume is acquired in a single T1 relaxation period. 
+\ac{MRI}, a single large excitation pulse is applied to the entire brain,
+and the entire volume is acquired in a single \ac{T1} relaxation period. 
 Because the entire k-space (spatial-frequency) volume is acquired 
-from a single excitation, the signal-to-noise-ratio (SNR) is very low
-in EPI.
+from a single excitation, the \ac{SNR} is very low
+in \ac{EPI}.
 
-Increasing the spatial resolution of EPI necessarily 
+Increasing the spatial resolution of \ac{EPI} necessarily 
 requires more time or faster magnetic field switching. Increasing
-magnet switching rates can result in
+switching rates can result in
 more artifacts and lower signal to noise ratios. The result is
-that at best FMRI is capable of 1 second temporal resolution. 
+that at best \ac{fMRI} is capable of 1 second temporal resolution. 
 The signal is also diluted because each voxel contains
 a variety of neurons, capillaries and veins. 
-Thus, the FMRI signal, which is sensitive to the chemical composition of 
+Thus, the \ac{fMRI} signal, which is sensitive to the chemical composition of 
 materials, is the average signal from various types of tissue
@@
 and explored in depth in \autoref{sec:BOLD Physiology},
-the usefulness of FMRI comes from discerning of changes in 
-Deoxyhemoglobin/Oxyhemoglobin. Therefore, it is necessary to assume
+the usefulness of \ac{fMRI} comes from discerning changes in 
+\ac{dHb}/\ac{O2Hb}. Therefore, it is necessary to assume
 that the only chemical changes will be in
@@
 instance near large veins, and it may explain some of the
-noise seen in FMRI imaging (see \autoref{sec:Introduction Noise}. 
-Because MRI is unitless and certain
-areas will have a higher base MR signal, all FMRI studies deal with
+noise seen in \ac{fMRI} imaging (see \autoref{sec:Introduction Noise}). 
+Because \ac{MRI} is unitless and certain
+areas will have a higher base \ac{MR} signal, all \ac{fMRI} studies deal with
 percent change from the base signal; rather than raw values.
 
-\section{BOLD Physiology}
+\section{\ac{BOLD} Physiology}
 \label{sec:BOLD Physiology}
-It is well known that the two types of hemoglobin act as a contrast agents in 
-EPI imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
-between Deoxyhemoglobin/Oxygenated Hemoglobin and neural activity is non-trivial. 
+It is well known that the two types of \ac{Hb} act as a contrast agents in 
+\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
+between \ac{dHb}/\ac{O2Hb} and neural activity is non-trivial. 
 Intuitively, increased 
-metabolism will increase Deoxyhemoglobin, however blood vessels are quick
+metabolism will increase \ac{dHb}, however blood vessels are quick
 to compensate by increasing local blood flow. Increased inflow, accomplished by loosening 
@@
 blood storage.
-Since the local MR signal depends on the ratio of Deoxyhemoglobin to Oxygenated
-Hemoglobin, increased blood volume affects this ratio if 
+Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
+increased blood volume affects this ratio if 
 metabolism doesn't exactly match the increased inflow of oxygenated blood.
@@
 model \cite{Mandeville1999}. These works derive, from first principals,
-the changes in deoxyhemoglobin ratio and volume of capillaries given an inflow waveform.
+the changes in \ac{dHb} ratio and volume of capillaries given an inflow waveform.
 These were the first two attempts to quantitatively account for the shape of the 
-BOLD signal as a consequence of the lag between the cerebral blood volume (CBV) 
-and the inward cerebral blood flow (CBF).
+\ac{BOLD} signal as a consequence of the lag between the \ac{CBV}
+and the inward \ac{CBF}.
 
 Although Buxton et al. demonstrated that a well chosen flow waveform could 
-explain most features of the BOLD signal, it stopped short of proposing a
-realistic waveform for the CBF and CMRO2 \cite{Buxton1998}. Friston et al. 
+explain most features of the \ac{BOLD} signal, it stopped short of proposing a
+realistic waveform for the \ac{CBF} and \ac{CMRO2} \cite{Buxton1998}. Friston et al. 
 gave a reasonable and simple
-expression for CBF input based on a flow inducing signal
+expression for \ac{CBF} input based on a flow inducing signal
 and in the same work proposed a simple method
@@
 By combining these equations with the balloon model from Buxton et al.,
-it is possible to predict the BOLD signal directly from a stimulus time course:
+it is possible to predict the \ac{BOLD} signal directly from a stimulus time course:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $s$ is a flow inducing signal, $f$ is the input blood flow (CBF),
-$v$ is normalized cerebral blood volume (CBV), and $q$ is the normalized
-local deoxyhemoglobin. The 
+where $s$ is a flow inducing signal, $f$ is the input \ac{CBF},
+$v$ is normalized \ac{CBV}, and $q$ is the normalized
+local \ac{dHb} content. The 
 parameters controlling blood flow are $\epsilon$, which is a neuronal 
@@
 by Riera et al. \cite{Riera2003}. Obata refined the readout equation 
-of the BOLD signal based on the
-deoxyhemoglobin content (q) and local blood volume (v), resulting in the
-final BOLD equation \cite{Obata2004}.
+of the \ac{BOLD} signal based on the
+\ac{dHb} content (q) and local blood volume (v), resulting in the
+final \ac{BOLD} equation \cite{Obata2004}.
 \begin{eqnarray}
@@
 de-oxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
-change in relaxation rate with change in blood oxygenation, and
-$\epsilon_0 = 1.43$ is the 
-ratio of signal MR from intravascular to extravascular regions at rest. Although,
+change in relaxation rate with change in blood oxygenation, $TE$ is the
+echo time used by the \ac{EPI} sequence and $\epsilon_0 = 1.43$ is the 
+ratio of signal \ac{MR} from intravascular to extravascular regions at rest. Although,
 these constants change with experiment ($TE$, $\nu_0$, $r_0$),
@@
 studies using 1.5 Tesla scanners \cite{Obata2004}.
-While this model is more accurate than the static Hemodynamic Model used in SPM,
+While this model is more accurate than the static hemodynamic model used in \ac{SPM},
 there are other additions which give it more degrees of freedom.
@@
 \label{sec:Post Stimulus Undershoot}
-Although the most widely used, the BOLD model described in \autoref{eq:bold}
+Although the most widely used, the \ac{BOLD} model described in \autoref{eq:bold}
 and \autoref{eq:boldout} has been extended in various fashions. The most
@@
 The post-stimulus undershoot is the term used for a prolonged subnormal
-BOLD response for a period of 10 to 60 seconds after stimulus has
+\ac{BOLD} response for a period of 10 to 60 seconds after stimulus has
 ceased \cite{Chen2009,Mandeville1999a}.
@@
 stimulus undershoot.  Recall
-that a lower than base signal means that there is an increased deoxyhemoglobin
+that a lower than base signal means that there is an increased \ac{dHb}
 content in the voxel. The first and simplest explanation is that the post-stimulus
-undershoot is caused by a prolonged increase in CMRO2 after CBV and CBF
+undershoot is caused by a prolonged increase in \ac{CMRO2} after \ac{CBV} and \ac{CBF}
 have returned to their base levels. This theory is justified by 
-studies that show CBV and CBF returning to the baseline before the BOLD signal
+studies that show \ac{CBV} and \ac{CBF} returning to the baseline before the \ac{BOLD} signal
 \cite{Frahm2008, Donahue2009, Buxton2004, Lu2004, Shen2008}. 
-Unfortunately, because of limitations on FMRI and in vivo
-CBV/CBF measurement techniques it is difficult to isolate whether CBF and
-CBV truly have returned to their baseline. Other studies indicate
-that there can be a prolonged supernormal CBV, although none of these papers completely
-rule out the possibility of increased CMRO2  \cite{Mandeville1999a,
+Unfortunately, because of limitations on \ac{fMRI} and in vivo
+\ac{CBV}/\ac{CBF} measurement techniques it is difficult to isolate whether \ac{CBF} and
+\ac{CBV} truly have returned to their baseline. Other studies indicate
+that there can be a prolonged supernormal \ac{CBV}, although none of these papers completely
+rule out the possibility of increased \ac{CMRO2}  \cite{Mandeville1999a,
 Behzadi2005, Chen2009a}. The discrepancies may in part
@@
 a compelling case is made that most of the post stimulus undershoot can be 
-explained by combination of a prolonged CBV increase, and a prolonged CBF 
+explained by combination of a prolonged \ac{CBV} increase, and a prolonged \ac{CBF} 
 undershoot \cite{Chen2009}. Additionally it was proposed that
-the previous measurements showing a quick recovery of CBV 
-were in fact showing a return to baseline by arterial CBV (which
-has little effect on the BOLD signal).
+the previous measurements showing a quick recovery of \ac{CBV} 
+were in fact showing a return to baseline by arterial \ac{CBV} (which
+has little effect on the \ac{BOLD} signal).
 
-Regardless of the probability that CMRO2 and CBF are detached,
+Regardless of the probability that \ac{CMRO2} and \ac{CBF} are detached,
 research into the post-stimulus undershoot has led to the creation
@@
 variables model oxygen transport, whereas Buxton et al. models
-CMR02 from a higher level, and somewhat more simply; though it 
+\ac{CMRO2} from a higher level, and somewhat more simply; though it 
 still adds 9 new parameters \cite{Zheng2002, Buxton2004}. Behzadi et al. 
-introduced nonlinearities into the CBF equations as a method to
+introduced nonlinearities into the \ac{CBF} equations as a method to
 explain the post-stimulus undershoot, which falls in line with a 
-prolonged increase in CBF observed by Chen et al. \cite{Behzadi2005, Chen2009}. 
+prolonged increase in \ac{CBF} observed by Chen et al. \cite{Behzadi2005, Chen2009}. 
 Similarly Zheng et al. added additional compartments for 
@@
 
-\section{Properties of the BOLD Model}
+\section{Properties of the \ac{BOLD} Model}
 \label{sec:BOLD Analysis}
-Since the first complete BOLD model was proposed by Friston et al.,  
+Since the first complete \ac{BOLD} model was proposed by Friston et al.,  
 several studies have analyzed its properties \cite{Friston2002}.  
@@
 
-In real FMRI data, there is a significant nonlinearity in response; with short sequences
+In real \ac{fMRI} data, there is a significant nonlinearity in response; with short sequences
 responding disproportionately strongly \cite{Birn2001, Wager2005, Deneux2006}.
@@
 tend to be more linear, which is why block designs are so well accounted for
-by the General Linear Model (see \autoref{sec:Current Techniques General Linear Model})
+by the \ac{GLM} (see \autoref{sec:Current Techniques General Linear Model})
 \cite{Birn2001, Deneux2006}.
@@
 Another interesting result from Deneux et al. was the sensitivity analysis \cite{Deneux2006}.
-There it was found that the parameters are far from perpendicular,
-and that very different parameters could give nearly identical BOLD output.
+That work found that the parameters are far from perpendicular,
+and that very different parameters could give nearly identical \ac{BOLD} output.
 The means that that without constraining parameter values, they may not be 
-precisely ascertainable. This could explain discrepancies in previous studies
-(\autoref{tab:Params}).
+precisely ascertainable. This could explain discrepancies of parameter estimates
+in previous studies (\autoref{tab:Params}).
 
@@
 \hline 
-Parameter  & Friston et al. \cite{Friston2000} & Jonston et al. \cite{Johnston2008} 
+Parameter  & Friston et al. \cite{Friston2002} & Jonston et al. \cite{Johnston2008} 
     & Vakorin et al. \cite{Vakorin2007} & Deneux et al. \cite{Deneux2006}\\
@@
 $\tau_0  $ &  $N(.98 , .25^2)$  & $8.38 \pm 1.5  $ & $.94$ & .27\\
-$\alpha  $ &  $N(.33 , .45^2)$  & $.189 \pm .004 $ & $.4$ (NC) & .63 \\
+$\alpha  $ &  $N(.33 , .045^2)$  & $.189 \pm .004 $ & $.4$ (NC) & .63 \\
 $E_0     $ &  $N(.34 , .1 ^2)$  & $.635 \pm .072 $ & $.6$ (NC) & .33\\

### 2010-08-31 doc/thesis/thesis.tex
added table of acronyms and almost done correcting introduction

@@
 
+\usepackage[printonlyused,withpage]{acronym}
 \usepackage{colortbl}
@@
 
-%\doublespacing
-
 \newcommand{\R}{\mathbb{R}}
@@
 \newcommand{\Z}{\mathbb{Z}}
-\newcommand{\qed}{\\$\Box$}
-\newcommand{\qle}{\stackrel{?}{\le}}
-\newcommand{\qeq}{\stackrel{?}{=}}
-\newcommand{\closure}{\overline}
-\newcommand{\intersect}{\cap}
-\newcommand{\union}{\cup}
-\newcommand{\nullset}{\emptyset}
-\newcommand{\minus}{\ \backslash\ }
 
@@
 
-Traditional methods of analyzing FMRI images use a linear combination of
+Traditional methods of analyzing functional Magnetic Resonance 
+Images use a linear combination of
 just a few static regressors. This work demonstrates an alternative
@@
 variables. The activation results show regions similar to those found in 
-SPM; however, there are some notable regions not detected by 
-SPM. Though the parameters selected by the particle filter based approach
-are more than sufficient to predict the BOLD response,
+Statistical Parametric Mapping; however, there are some notable regions not 
+detected by that technique. Though the parameters selected by the particle filter based approach
+are more than sufficient to predict the Blood-Oxygen-Level-Dependent signal
+response,
 more model constraints are needed to uniquely identify a single set
@@
 activation, but an extensible technique of estimating the joint probability
-distribution function of the BOLD parameters.
+distribution function of the Blood-Oxygen-Level-Dependent Signal parameters.
 
@@
 
+The Neuroscientists at Wake Forest have also been extremely helpful. In 
+particular I would like to thank Paul Laurienti for help gathering fMRI
+data and using SPM.
+
 To my parents, for all the psychological (and financial) help along the way,
-thank you so much. When things were the darkest you were there for me, I couldn't
+thank you so much. You were always there for me, I couldn't
 ask for better parents. Love you guys.
@@
 \listoftables
+\pagebreak
+
+\huge
+\bfseries
+List of Acronyms
+\mdseries
+\normalsize
+\begin{acronym}[CMRO2]
+\acro{BOLD}{Blood-Oxygen-Level-Dependent}
+\acro{CBF}{Cerebral Blood Flow}
+\acro{CBV}{Cerebral Blood Volume}
+\acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
+\acro{dHb}{Deoxygenated Hemoglobin}
+\acro{EPI}{Echo Planar Imaging}
+\acro{fMRI}{Functional Magnetic Resonance Imaging}
+\acro{GLM}{General Linear Model}
+\acro{HRF}{Hemodynamic Response Function}
+\acro{Hb}{Hemoglobin}
+\acro{MR}{Magnetic Resonance}
+\acro{MRI}{Magnetic Resonance Imaging}
+\acro{O2Hb}{Oxygenated Hemoglobin}
+\acro{RF}{Radio Frequency}
+\acro{SNR}{Signal-to-Noise Ratio}
+\acro{SPM}{Statistical Parametric Mapping}
+\acro{T1}{Longitudinal}
+\acro{T2}{Spin-Spin}
+\acro{TR}{Repetition Time}
+\end{acronym}
 \pagebreak

### 2010-09-01 doc/thesis/background.tex
lots of fixes from the corrections

@@
 \label{sec:Prior Works}
-Currently, FMRI is used to determine the location of responses
+Currently, \ac{fMRI} is used to determine the location of responses
 to stimuli. The method of finding activation is discussed in 
@@
 more computation. Already there have been several other
-attempts to model the BOLD response; these works will be discussed
+attempts to model the \ac{BOLD} response; these works will be discussed
 in this chapter.
@@
 Although not strictly the same as parameter calculation from 
-FMRI, activation detection is similar and worth discussing. Estimation of 
+\ac{fMRI}, activation detection is similar and worth discussing. Estimation of 
 parameters is a generalization of the idea 
 of activation detection.
-Given the popularity of Statistical Parametric Mapping (SPM) 
+Given the popularity of \ac{SPM}
 it is important to draw a distinction between it and the methods proposed
@@
 \label{sec:RFT}
-SPM methods make significant use of $t$-Tests across large regions;
+\ac{SPM} methods make significant use of $t$-Tests across large regions;
 however, such $t$-tests work slightly differently than a single
@@
 can theoretically be applied with a lower scaling factor than
-the original voxel count \cite{Worsley2004}. As side effect of this 
-extreme smoothing, a single voxel activation is very hard to detect.
+the original voxel count \cite{Worsley2004}. As a side effect of this 
+extreme smoothing, a single active voxel is very difficult to detect.
 
@@
 \label{sec:Square Wave}
-The most basic method of analyzing FMRI data is by standard $t$-test
+The most basic method of analyzing \ac{fMRI} data is by standard $t$-test
 between resting state and active state samples. Simply put, the
@@
 distributions actually have the same mean. Because of the correlated
-noise present in FMRI (\autoref{sec:Introduction Noise}), it is necessary
+noise present in \ac{fMRI} (\autoref{sec:Introduction Noise}), it is necessary
 to apply a high-pass filter to the data. Without applying
-such a filter, P values must be set extraordinarily high to prevent
-false positives \cite{Smith2007}. Because the BOLD response is known
+such a filter, T-values must be set extraordinarily high to prevent
+false positives \cite{Smith2007}. Because the \ac{BOLD} response is known
 not to fit a square wave this method is rarely used
@@
 \label{sec:Current Techniques General Linear Model}
-The most common FMRI analysis tool is SPM, which generates
+The most common \ac{fMRI} analysis tool is \ac{SPM}, which generates
 statistical parametric map using more advanced methods than a simple
@@
 does not result in a square wave in the activation of brain regions. 
-The BOLD signal is in fact a smoothed version of the 
-stimuli. As such, when fitting an FMRI time course to the input,
+The \ac{BOLD} signal is in fact a smoothed version of the 
+stimuli. As such, when fitting an \ac{fMRI} time course to the input,
 the input ($X$'s columns) is usually smoothed to reduce bias error. 
 The best method, that maintains a linear fit, is convolving the input 
-with a Hemodynamic 
-Response Function (HRF). The \emph{Hemodynamic Response Function}
-mimics the basic shape of BOLD activation, including a delay
+with a \ac{HRF}. The \ac{HRF}
+mimics the basic shape of \ac{BOLD} activation, including a delay
 due to rise time and fall time. The fitting 
@@
 
-Smoothing the input with a single HRF poses certain problems.
+Smoothing the input with a single \ac{HRF} poses certain problems.
 It is well known that different Hemodynamic Response Functions are necessary 
-for different regions of the brain \cite{Handwerker2004}. The Canonical HRF that is most often used, 
+for different regions of the brain \cite{Handwerker2004}. The Canonical 
+\ac{HRF} that is most often used, 
 has been optimized for the visual cortex. As discussed in \autoref{sec:Post Stimulus Undershoot},
 there are definitely variations in the shape of the
-BOLD response, over both brain regions and patients. Handwerker et. al. 
-discussed the implications of choosing an incorrect HRF, the most important
+\ac{BOLD} response, over both brain regions and patients. Handwerker et. al. 
+discussed the implications of choosing an incorrect \ac{HRF}, the most important
 of which is a definite increase in false negatives \cite{Handwerker2004}. While an atlas
@@
 As mentioned previously and discussed extensively
-in Friston et al. and Hoffman et al., hierarchical models may be 
-applied to account for systematic differences between subjects
-\cite{Friston2002, Hofmann1997}. For instance, if the
-study mixes young and old, incorporating
+in Friston et al. \cite{Friston2002a} and Hoffman et al. \cite{Friston2002a},
+hierarchical models may be 
+applied to account for systematic differences between subjects.
+For instance, if the
+study mixes young and old subjects, incorporating
 that into the model is wise, regardless of whether it is the goal of the test. 
@@
 be otherwise explainable.  The Hierarchical form used by
-Friston et al. is shown in \autoref{eq:Hierarchical} \cite{Friston2002}.
+Friston et al. \cite{Friston2002a} the \ac{GLM} which 
+is shown in \autoref{eq:Hierarchical} .
 
@@
 
-The Empirical Bayes algorithm was used by both Friston et al. and Hofmann et al. 
-\cite{Friston2002, Hofmann1997}. Note that in Empirical Bayes point estimators 
-are used for each $\theta$, rather than full distributions or the first two moments.
+The Empirical Bayes algorithm was used by both Friston et al. \cite{Friston2002a}
+and Hofmann et al. \cite{Hofmann1997}. Note that in Empirical Bayes, 
+point estimators are used for each $\theta$, rather than full 
+distributions or the first two moments.
 
@@
 \label{sec:BackgroundConclusion}
-In all, the GLM is useful for determining linear 
+In all, the \ac{GLM} is useful for determining linear 
 dependence of a set of regressors on the output. Unfortunately, as discussed in
@@
 highlighted by the common scenario where no significant
-activation can be found in a single run \cite{Riera2004, Johnston2008}.
+activation can be found in a single run \cite{Riera2003, Johnston2007}.
 
 The static nature of the  linear model also limits its potential use. 
-Besides not allowing HRF differences between patients, there is no
+Besides not allowing \ac{HRF} differences between patients, there is no
 reasonable way to incorporate other forms of physiological
-data. Combined FMRI CBF or CBV imaging methods are improving,
+data. Combined \ac{fMRI} \ac{CBF} or \ac{CBV} imaging methods are improving,
 as seen in Chen et al. \cite{Chen2009}. These techniques could shed light on
@@
 Unlike Statistical Parametric Mapping, the techniques described in this
-section are all attempts to regress some version of the BOLD model. 
-Although Buxton et al. and Friston et al. 
+section are all attempts to regress some version of the \ac{BOLD} model. 
+Although Buxton et al. \cite{Buxton1998} and Friston et al. \cite{Friston2000}
 both proposed physiologically reasonable values for the model parameters, 
-Friston et al. in 2002 was the first paper to calculate the parameters based 
-on actual FMRI data \cite{Buxton1998, Friston2000, Friston2002b}. 
+Friston et al. \cite{Friston2002b} was the first paper to calculate the 
+parameters based on actual \ac{fMRI} data. 
 However, in that case, the voxels were chosen from
-regions that were detected as active by the GLM. It is therefore possible
+regions that were detected as active by the \ac{GLM}. It is therefore possible
 that the parameters are biased toward parameters that fit the linear model.
@@
 \label{sec:Background Linear Approximation}
-In Friston et al., a novel combination of linear and nonlinear modeling
-was used to generate parameter estimates \cite{Friston2002b}. 
-Although this method was restricted to regions determined to be 
-active by the GLM, there is no reason it could not be applied more broadly.
+In Friston et al. \cite{Friston2002b}, a novel combination of 
+linear and nonlinear modeling proposed to generate parameter estimates. 
 Because there is no closed 
-form solution to the balloon model, it impossible to 
+form solution to the balloon model, it is impossible to 
 calculate the Jacobian matrix $\frac{\partial Y}{\partial \theta}$. Thus 
-Friston et al. approximated the partial using a Volterra Kernel \cite{Friston2002b}. 
-At each step of the Expectation-Maximization (EM) algorithm, 
+Friston et al. \cite{Friston2002b} approximated the partial using a
+low-order Volterra Kernel. 
+At each step of the \ac{EM} algorithm, 
 the differential equation was integrated and the residuals
 calculated. Then, for each $\theta_i$ surrounding the estimate of 
-$\theta$, a Volterra-Kernel expansion of the output $y$ was generated. Generation
+$\theta$, a Volterra Kernel expansion of the output $y$ was generated. Generation
 of the Volterra Kernel is quick, and, since there
 is an analytical solution, calculating partial derivatives is easy. 
+Unfortunately, it was found in that work that estimating a Volterra kernel
+took longer than simply integrating the state variables. 
 The full E-M algorithm for estimating the states is notation-heavy 
@@
 There are a few caveats with this method of optimization. First, the 
-partials are based on an approximate values (using Volterra-Kernels) of $y$.
+partials are based on approximate values (using Volterra-Kernels) of $y$.
 Importantly, the Volterra-Expansion of $y$ is not able to model interactions
@@
 Still, there are other heuristic algorithms
-that could illuminate the BOLD response. 
-Simulated Annealing (SA) is a common method of optimizing high dimensional
+that could estimate the \ac{BOLD} response. 
+\ac{SA} is a common method of optimizing high dimensional
 regression problems. First the program selects a random starting point, then
@@
 
-\begin{algorithm}
-\caption{Simulated Annealing Algorithm}
-\label{alg:Simulated Annealing}
-\begin{algorithmic}
-\STATE Initialize $\Theta$, or if there exists a decent estimate start there
-\STATE Initialize temperature, T to value above initial energy
-\WHILE{$E(\Theta) < T$}
-    \REPEAT
-        \STATE Pick $\theta$ near $\Theta$
-        \STATE Calculate energy, $E$, of $\theta$
-    \UNTIL{$E > T$}
-    \STATE Move to new estimate: set $\Theta = \theta$
-\ENDWHILE
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{Simulated Annealing Algorithm}
+%\label{alg:Simulated Annealing}
+%\begin{algorithmic}
+%\STATE Initialize $\Theta$, or if there exists a decent estimate start there
+%\STATE Initialize temperature, T to value above initial energy
+%\WHILE{$E(\Theta) < T$}
+%    \REPEAT
+%        \STATE Pick $\theta$ near $\Theta$
+%        \STATE Calculate energy, $E$, of $\theta$
+%    \UNTIL{$E > T$}
+%    \STATE Move to new estimate: set $\Theta = \theta$
+%\ENDWHILE
+%\end{algorithmic}
+%\end{algorithm}
 
-Genetic algorithms (GA) are similar to Simulated Annealing, in
+\ac{GA} are similar to Simulated Annealing, in
 that they randomly move to better solutions based on a cost function.
@@
 
-\begin{algorithm}
-\caption{Genetic Algorithm}
-\label{alg:Genetic Algorithm}
-\begin{algorithmic}
-\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
-\FOR{G generations}
-    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
-    \FOR{$i$ in $N$}
-        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
-        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
-        \STATE With low probability introduce random mutations to parameters in $\Theta_i$
-    \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{Genetic Algorithm}
+%\label{alg:Genetic Algorithm}
+%\begin{algorithmic}
+%\STATE Initialize $N$ estimates, $E = \{\Theta_0, \Theta_1, ... \Theta_N\}$
+%\FOR{G generations}
+%    \STATE Calculate fitness for each $\Theta$, Ex. for residual $R$, $1/R$ or, $e^{-R}$
+%    \FOR{$i$ in $N$}
+%        \STATE Randomly select two parents (with higher probability for more fit $\Theta$'s)
+%        \STATE Randomly merge parts of the two parents to form a new $\Theta_i$
+%        \STATE With low probability introduce random mutations to parameters in $\Theta_i$
+%    \ENDFOR
+%\ENDFOR
+%\end{algorithmic}
+%\end{algorithm}
 
 Although both these methods can be highly effective, they have the downside of
-requiring high computation time. In the case of the BOLD model,
+requiring high computation time. In the case of the \ac{BOLD} model,
 each time the energy or fitness needs to be calculated, a large number of cycles
-must be spent re-simulating the BOLD model for the set of parameters. As I'll
+must be spent re-simulating the \ac{BOLD} model for the set of parameters. As I'll
 discuss in \autoref{sec:Particle Filter}, the Particle Filter method is able
@@
 these would have to converge in less than 1000 simulations (which is far less than
-1000 generations in GA).
+1000 generations in \ac{GA}).
 
@@
 \label{sec:Unscented Kalman Filter}
-The Unscented Kalman Filter (UKF) is a powerful Gaussian/Bayes filter that attempts
+The \ac{UKF} is a powerful Gaussian/Bayes filter that attempts
 to model the posterior distribution of dynamical systems as a multivariate
-Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
+Gaussian. The \ac{UKF} generalizes the Extended Kalman
 Filter by allowing the state update and output functions to be arbitrary functions.
@@
 no Jacobian and only a few extra calculations to get a decent estimate of
-a posterior Gaussian. Hu et. al. used the UKF to  perform a similar type of analysis to
+a posterior Gaussian. Hu et. al. used the \ac{UKF} to  perform a similar type of analysis to
 the one performed in this work \cite{Hu2009}.
@@
 gets the more likely that the Gaussian will be insufficient to describe
-the distribution, $X(t)$. When this occurs, every step from  $X(t+1)$ 
+the distribution, $X(t)$. When this occurs, every step from  $X(t-1)$ 
 to $X(t)$ will introduce additional
@@
  The assumption of Gaussianity is what allows
-the UKF to estimate the posterior using only the first and second moments;
+the \ac{UKF} to estimate the posterior distribution using only the 
+first and second moments;
 however, if this assumption is violated significant error will result.
@@
 \caption{Parameters used to test Gaussianity of variables after being transitioned through
-the BOLD model}
+the \ac{BOLD} model}
 \label{tab:steptable} 
@@
 To determine the amount of error incurred in a Gaussian estimate during
-a typical sample period, I assigned the states of BOLD equations according
-to a four dimensional Gaussian. I then propagated the states through 
-two seconds of simulation (a typical TR in FMRI) and plotted the resulting
+a typical sample period, the states of \ac{BOLD} equations were assigned according
+to a four dimensional Gaussian. The states were then propagated through 
+two seconds of simulation (a typical \ac{TR} in \ac{fMRI}) and plotted the resulting
 marginal distributions against a Gaussian distribution. This also demonstrates
@@
 
-In order to drive the system without input I intentionally set $s_t$ a non-steady
-state, but physiologically plausible, value. The value of $u$ was left at zero 
+In order to drive the system without input, the initial state was set
+to a non-steady but physically 
+plausable $s_t$ state value. The value of $u$ was left at zero 
 the entire time, so the 
@@
 However, \autoref{fig:trans1s} shows the result after 1 second, which is faster
-than most FMRI scanners are capable of sampling at. At that range the tails 
+than most \ac{fMRI} scanners are capable of sampling at. At that range the tails 
 of the distributions for $v$ and $q$ started to deviate from the
@@
 approximating the distribution with a Gaussian based on the first two
-moments will worked in the short run, within a period less than typical
+moments will work in the short run, within a period less than typical
 measurement rates the distribution deviated from a Gaussian substantially.
@@
 For this reason, estimating the posterior distribution using only 
-the first two moments, which is the basis for the UKF, is not wise.
+the first two moments, which is the basis for the \ac{UKF}, is not wise.
 
@@
 \subsection{Hybrid Methods}
-In Riera et al. , a maximum
+In Riera et al. \cite{Riera2003}, a maximum
 likelihood method for innovation processes was used, as described by
-Ozakai \cite{Riera2003,Ozaki1994}. Ozaki used a similar construction to a 
+Ozakai \cite{Ozaki1994}. Ozaki \cite{Ozaki1994} used a similar construction to a 
 Kalman filter on nonlinear signals \cite{Ozaki1994}. 
-The method used by Riera et al. used this method to perform maximum likelihood on
-the innovations rather than the absolute signal levels \cite{Riera2003}. 
+The method used by Riera et al.\cite{Riera2003} used this method to 
+perform maximum likelihood on
+the innovations rather than the absolute signal levels. 
 By using a Local Linearization Filter,
 innovation noise such as noise in $\dot{f}$ is turned into simple 
-Gaussian noise. This approach is useful when the DC portion of the signal is
+Gaussian noise. This approach is useful when the \ac{DC} portion of the signal is
 of no use; however, it depends greatly on the type of signal. I found that
-this method was not very effective for the BOLD signal, when I used it 
+this method was not very effective for the \ac{BOLD} signal, when I used it 
 with the particle filter approach discussed in the next chapter; although its
@@
 
-In Johnston et al., a hybrid particle filter/gradient
+In Johnston et al. \cite{Johnston2007}, a hybrid particle filter/gradient
 descent algorithm was used to simultaneously derive the static and dynamic 
 parameters, (classically known as parameters and state variables, respectively)
- \cite{Johnston2008}.
+.
 A particle filter was used to calculate the state variables at each
@@
 the most likely set of parameters that would give that distribution of state variables.
-This process was repeated until the parameters converged. Interestingly Johnston et al.
+This process was repeated until the parameters converged. Interestingly Johnston et al. \cite{Johnston2007}
 came to a very different set of parameter estimates as compared
-to the original Friston et al. estimates (\autoref{tab:Params})
-\cite{Johnston2008, Friston2000}. In fact the results 
+to the original Friston et al. \cite{Friston2000} estimates (\autoref{tab:Params}).
+In fact the results 
 are significantly different from virtually every other study. The most obvious discrepancy
@@
 While of course this could be poor convergence of the algorithm, there is another other possibility.
-Unlike all the other methods mentioned, excepting the methods in \autoref{sec:Nonlinear Least Squares},
-the algorithm described in Johnston et al. does not depend on prior distributions \cite{Johnston2008}.
-It is possible then that the bias toward the prior in other methods skewed their results. 
-While Johnston et al. is certainly in the minority; further exhaustive studies 
-of the parameters, using unbiased techniques may be called for \cite{Johnston2008}. A further 
-comparison between the distributions found in Johnston et al. and Friston et al. 
-will be discussed in \autoref{sec:PriorDist} \cite{Johnston2008, Friston2000}.
+Unlike all the other methods mentioned, excepting the methods in 
+\autoref{sec:Nonlinear Least Squares},
+the algorithm described in Johnston et al. \cite{Johnston2007} 
+does not depend on prior distributions.
+It is possible then that the bias toward the prior in other methods 
+skewed their results. 
+While Johnston et al. \cite{Johnston2007} is certainly in the minority;
+further exhaustive studies 
+of the parameters, using unbiased techniques may be called for. A further 
+comparison between the distributions found in Johnston et al.
+\cite{Johnston2007} and Friston et al. \cite{Friston2000} 
+will be discussed in \autoref{sec:PriorDist}. Although Johnston et al.
+\cite{Johnston2007} did not explicity list computation times, from
+this work, a particle filter with 100 particles integrating the 
+BOLD model should take around 5 seconds, repeated 10 times should
+take approximately 50 seconds for the particle filter alone. For
+the least squares searches performed at each iteration, computation
+likely takes around a minute, giving the algorithm a 10 minute 
+computation time per voxel.
 
-In Vakorin et al., a combination of Genetic Algorithms and Simulated Annealing were
-used to estimate not only the parameters, but the true stimuli driving the BOLD
-signal \cite{Vakorin2007}. 
+In Vakorin et al. \cite{Vakorin2007}, a combination of Genetic Algorithms and 
+Simulated Annealing was
+used to estimate not only the parameters, but the true stimuli driving the \ac{BOLD}
+signal. 
 This addresses the inherent uncertainty of exactly where and when 
-stimuli actually get applied. Unfortunately this algorithm was extremely slow.
+stimuli actually get applied. Unfortunately this algorithm took
+in more than 16 hours per voxel.
 
 \subsection{Previous Particle Filter Approaches}
-In a PhD thesis, Murray used a particle filter based approach was used to integrate
-the BOLD equations \cite{Murray2008}. The method used in that work focused primarily on estimating
-the BOLD output and state equations as a nonlinear stochastic differential 
+In his PhD thesis, Murray \cite{Murray2008} used a particle filter based 
+approach to integrate
+the \ac{BOLD} equations. The method used in that work focused primarily on estimating
+the \ac{BOLD} output and state equations as a nonlinear stochastic differential 
 equation. The primary difference between that work and this is that
-Murray took the parameters as a given \cite{Murray2008}. Thus, differences in the BOLD output
+Murray \cite{Murray2008} took the parameters as a given. Thus, differences in the \ac{BOLD} output
 were taken to be primarily driven by stochastic changes in the underlying state
 equations. Because the parameters were not allowed to change, the estimate of 
-the BOLD signal was not very good. The fact that the
-differences in BOLD response cannot be explained solely by stochastic elements 
+the \ac{BOLD} signal was not very good. The fact that the
+differences in \ac{BOLD} response cannot be explained solely by stochastic elements 
 is important, however. The filtering framework created
-in that work, dysii, forms the basis for the particle filter used in this work, 
-and was very designed. The work also clearly presents the particle filter;
+in Murray \cite{Murray2008}, dysii, forms the basis for the particle 
+filter used in this work, 
+and was well designed. The work also clearly presents the particle filter;
 both its derivation and use.
@@
 equations. Exhaustive search type methods such as those employed by
-Johnston et al. and Vakorin et al. have long
-run times even for a single voxel \cite{Johnston2008, Vakorin2007}. 
+Johnston et al. \cite{Johnston2007} and Vakorin et al. \cite{Vakorin2007}
+have greater than 10 minute run times for a single voxel. 
 While Volterra models are an interesting

### 2010-09-01 doc/thesis/intro.tex
lots of fixes from the corrections

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{images/bold}
+\caption[Diagram of BOLD Signal]{A) End of Pre-Stimulus Undershoot, $< 2s$ typical,
+B) End of Stimulus, C) Input Bloodflow Drops from Supernormal to Subnormal,
+$5 s$ after B typical, D) Beginning of Post-stimulus Undershoot, $12 s$ after B
+typical, E) Full return to normal, between $20-30s$ after B. \cite{Behzadi2005}}
+\end{figure}
+
 \section{\ac{fMRI}}
@@
 non-invasively, based on the difference between nuclear spin
-relaxation times in different molecules. First, the subject 
+relaxation times in different molecules. The technique is
+well established and documented in various works \cite{Bushberg2002}. 
+First, the subject 
 is brought into a large magnetic field which causes nuclear spins
@@
 magnitude faster
-than \ac{T1} relaxation) based on the properties of the region.
+than \ac{T1} relaxation) based on the properties of the region 
+\cite{Bushberg2002}.
 This gives two primary methods of contrasting substances,
@@
 \ac{MRI}, a single large excitation pulse is applied to the entire brain,
-and the entire volume is acquired in a single \ac{T1} relaxation period. 
+and the entire volume is acquired in a single \ac{T1} relaxation period
+\cite{Obata2004}. 
 Because the entire k-space (spatial-frequency) volume is acquired 
 from a single excitation, the \ac{SNR} is very low
-in \ac{EPI}.
+in \ac{EPI} \cite{Smith2007}.
 
@@
 more artifacts and lower signal to noise ratios. The result is
-that at best \ac{fMRI} is capable of 1 second temporal resolution. 
+that at best \ac{fMRI} is capable of 1 second temporal resolution
+\cite{Ladd2010, Poser2010}. 
 The signal is also diluted because each voxel contains
-a variety of neurons, capillaries and veins. 
+a variety of neurons, capillaries and veins \cite{Obata2004}. 
 Thus, the \ac{fMRI} signal, which is sensitive to the chemical composition of 
@@
 the usefulness of \ac{fMRI} comes from discerning changes in 
-\ac{dHb}/\ac{O2Hb}. Therefore, it is necessary to assume
+\ac{dHb}/\ac{O2Hb} \cite{Ogawa, Obata2004}. Therefore, it is necessary to assume
 that the only chemical changes will be in
@@
 
-\section{\ac{BOLD} Physiology}
+\section{BOLD Physiology}
 \label{sec:BOLD Physiology}
-It is well known that the two types of \ac{Hb} act as a contrast agents in 
+It is well known that the two types of \ac{Hb} act as contrast agents in 
 \ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
@@
 rule out the possibility of increased \ac{CMRO2}  \cite{Mandeville1999a,
-Behzadi2005, Chen2009a}. The discrepancies may in part
+Behzadi2005, Chen2009}. The discrepancies may in part
 be explained by a spatial dependence in the post-stimulus undershoot; described
@@
 \label{sec:BOLD Analysis}
-Since the first complete \ac{BOLD} model was proposed by Friston et al.,  
-several studies have analyzed its properties \cite{Friston2002}.  
+Since the first complete \ac{BOLD} model was proposed by Friston et al. 
+\cite{Friston2000},  
+several studies have analyzed its properties.  
 The most important property is that the system is dissipative, and given
@@
 \hline 
-Parameter  & Friston et al. \cite{Friston2002} & Jonston et al. \cite{Johnston2008} 
+Parameter  & Friston et al. \cite{Friston2000} & Jonston et al. \cite{Johnston2007} 
     & Vakorin et al. \cite{Vakorin2007} & Deneux et al. \cite{Deneux2006}\\
@@
 \caption{Parameters found in various studies. (NC) indicates that the value
-wasn't calculated. \cite{Vakorin2007} made use of the values from \cite{Friston2002}
+wasn't calculated. \cite{Vakorin2007} made use of the values from \cite{Friston2000}
 where not explicitly stated}

### 2010-09-01 doc/thesis/particlefilters.tex
small typos changed
lots of fixes from the corrections

@@
 similar to Kalman Filters; however, unlike Kalman Filters,
-distributions are stored as an empircal distribution rather than the
+distributions are stored as an empircal distribution rather than 
 as the first two moments of a Gaussian. Thus, particle filters are 
@@
 $\{Y_k, k = 1, 2, 3, ... K\}$, where $K$ may be infinite. 
-Because $k$ is a discrete time, let $t_k$ define the continuous
-time of $k$.
+Let $k$ denote a time index and $t_k$ define the continuous
+time of index $k$.
 Suppose also that there is a hidden set of state variables,
 $X(t)$ that drives the value of $Y(t)$. Throughout this section
-with $X_k = X(t_k)$. The goal of the particle filter is to estimate the 
+$X_k$ will identify $X(t_k)$. The goal of the particle filter is to estimate the 
 distribution of the
@@
 The only difference between the members of parameter vector
-$\Theta$ and those of $X(t)$ is that the memebers of
+$\Theta$ and those of $X(t)$ is that the members of
 $\Theta$ have no known update equation. Members of both vectors
@@
 for the system to be considered varying with time. 
-Next, its assumes that input cannot directly
+Next, it was assumed that input cannot directly
 influence the output, which in the case of the BOLD signal is a good assumption.
@@
 
-To begin with, the particle filter must be given a prior distribution, from
+To begin, the particle filter must be given a prior distribution, from
 which the initial $N_p$ particles are drawn. A particle contains a weight
@@
 
-Where $N_p$ is the number of particles or points used to describe the prior 
+where $N_p$ is the number of particles or points used to describe the prior 
 using a Mixture PDF. 
@@
 
-Therefore, after the particle have been generated they should approximate $\alpha(X)$:
+Therefore, after the particles have been generated they should approximate $\alpha(X)$:
 
@@
 \end{equation}
-Where $\delta(x-x_0)$ is 1 if and only if $x = x_0$ (the Kronecker delta function).
+where $\delta(x-x_0)$ is 1 if and only if $x = x_0$ 
+(the Kronecker delta function).
 
@@
 for increased particle counts although some distributions
-flatten out better than others. Either way, $\alpha(X)$ \emph{must} be
+flatten out better than others. Either way, $\alpha(X)$ must be
 wide enough to incorporate any posterior that arises. If the prior is
@@
 being correct given all the measurements up to time $k$. 
-Note that if there is a far off peak in
+Note that if there is a distant peak in
 the posterior that $q$ does not have support points in, there will 
-be quantization errors, and that part of the density cannot be modeled. This is why
-it is absolutely necessary that $q$ fully covers $P(x^i_{0:k} | y_{0:k})$.
+be quantization errors, and that part of the density cannot be modeled. 
+This is why it is absolutely necessary that $q$ fully covers 
+$P(x^i_{0:k} | y_{0:k})$.
 
@@
 could be false; since later measurements may shed light on currently hidden
-changes in $x$. For practical applications though it is a helpful assumption.
+changes in $x$, however in this work it is a good assumption because the
+input does not immediately effect the BOLD signal. 
 \begin{equation}
@@
 The benefit of this choice for importance density is that an approximation for
-$P(x_k | x_{k-1})$ is freely available: its simply the set of particles propagated
+$P(x_k | x_{k-1})$ is freely available; its simply the set of particles propagated
 forward in time using the state equations. Additionally it makes
@@
 fully constrained by $x_k$, and that $x_k$ is similarly fully 
-constrained by $x_{k-1}$, I make the good assumptions that:
+constrained by $x_{k-1}$, I make the assumptions that:
 \begin{eqnarray}
@@
 \end{eqnarray}
-These are of course just re-statements of the state equations assumed by \autoref{eq:stateass}
-and \autoref{eq:measass}.
+These are of course just re-statements of the state equations assumed 
+by \autoref{eq:stateass} and \autoref{eq:measass}.
 
@@
 Thus, by making the following simple assumptions, evolving a posterior
-density  requires no knowledge of noise distribution.
+density  requires no knowledge of the noise distribution.
 \begin{enumerate}

### 2010-09-02 doc/thesis/background.tex
just need to edit discussion/conclusion, change captions and copy some graphs from defense

@@
 \begin{figure}
-\includegraphics[width=16cm]{images/kalman}
-%\includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption{Example updates of a distribution using the Kalman Filter.}
+\includegraphics[width=16cm]{images/kalman2}
+\caption[Example Kalman Filter Progression]
+{a) Sigma Points are passed through the state/measurement equations
+to get a state prediction (green line). The first two moments of
+the state prediction are then used to estimate the distribution with
+a Gaussian (red line). The distribution centered at the new measurement,
+estimating measurement error (dark blue), is then combined with the existing
+state prediction (red line) to get the new estimated state (light blue).
+The same process is applied in b), and c) with different measurements and
+initial state estimates.}
 \label{fig:EKFWorking}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption{Distributions of state variables after simulating for $0.1$s}
+\caption[Q-Q Plot Of Standard Normal and Result of $0.1$s BOLD Simulation Output]
+{Q-Q Plot Of Standard Normal and Result of $0.1$ BOLD Simulation Output. A
+perfect Gaussian would fall on the green line}
 \label{fig:transp1s}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
-\caption{Distributions of state variables after simulating for $1$s}
+\caption{Q-Q Plot Of Standard Normal and Result of $1$s BOLD Simulation Output}
 \label{fig:trans1s}

### 2010-09-02 doc/thesis/methods.tex
small fix
just need to edit discussion/conclusion, change captions and copy some graphs from defense

@@
 Although the particle filter  is a standard Regularized
-Particle filter, as described by Arulampalam et al., optimizing the
-particle filter for use with FMRI data is non-trivial \cite{Arulampalam2002a}.
+Particle filter, as described by Arulampalam et al.
+ \cite{Arulampalam2002a}, optimizing the
+particle filter for use with \ac{fMRI} data is non-trivial.
 
@@
 As originally written in \autoref{sec:BOLD Physiology} the state variables
-for the BOLD model are as follows:
+for the \ac{BOLD} model are as follows:
 \begin{eqnarray}
@@
 that detrending is applied before the particle filter and that the
-BOLD model is dissipative. When no stimuli are applied, all the particles
+\ac{BOLD} model is dissipative. When no stimuli are applied, all the particles
 decay to $\{0,1,1,1\}$. Typical particle filters
@@
 however, from time to time $0.001$ can still result in problems for the
-BOLD model.
+\ac{BOLD} model.
 
@@
 The normal pipeline for analyzing
-FMRI involves a several preprocessing steps. The first and most important
+\ac{fMRI} involves a several preprocessing steps to condition the signal.
+The first and most important
 task is motion correction. To do this, a single volume in time is chosen, and
@@
 filter is applied across the image as discussed in \autoref{sec:RFT}.
-After this detrending is performed, which is discussed in \autoref{sec:Detrend}.
-Recall that FMRI signal levels are unit-less and though detrending is not
+After this, detrending is performed, which is discussed in \autoref{sec:Detrend}.
+Recall that \ac{fMRI} signal levels are unit-less and though detrending is not
 always necessary, the data must always be converted
@@
 Before going into the detrending used in this work, it is necessary to
-discuss the type of noise present in FMRI.
+discuss the type of noise present in \ac{fMRI}.
 
@@
 \label{sec:Introduction Noise}
-As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
-extensively studied and despite minor discrepancies, the cause of the BOLD
-signal is well known. However, as FMRI detects an
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_29_49_9}}
+\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_34_43_24}}
+\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_22_38_23}}
+\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_37_29_24}}
+\caption[Resting State Noise Analysis.]{Resting state noise analysis.
+From Left to Right:
+(resting state) signal, Q-Q Plot of measurements with the standard Normal,
+Signal auto-regression, Histogram of the Points}
+\label{fig:QQDC}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_29_49_9}}
+\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_34_43_24}}
+\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_22_38_23}}
+\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
+\caption[Resting State Noise Analysis, Steps]{Resting state noise analysis, where
+signal levels are the difference between adjacent signals in \autoref{fig:QQDC}.
+From Left to Right:
+(resting state) signal, Q-Q Plot of measurements with the standard Normal,
+Signal auto-regression, Histogram of the Points}
+\label{fig:QQDelta}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_29_49_9}}
+\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_34_43_24}}
+\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_22_38_23}}
+\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
+\caption[Resting State Noise Analysis, After Subtracting Spline]
+{Resting State Noise Analysis, where
+signal levels are the result of subtracting a spline from the signal in 
+\autoref{fig:QQDC}.
+From Left to Right:
+(resting state) signal, Q-Q Plot of measurements with the standard Normal,
+Signal auto-regression, Histogram of the Points}
+\label{fig:QQSpline}
+\end{figure}
+
+As demonstrated in \autoref{sec:BOLD Physiology} the \ac{BOLD} response has been
+extensively studied and despite minor discrepancies, the cause of the \ac{BOLD}
+signal is well known. However, as \ac{fMRI} detects an
 aggregate signal over the space of cubic centimeters, there are
@@
 density of capillaries, and slight differences in activation across
-a particular voxel can all lead to signal attenuation and noise.
+a particular voxel can all lead to signal attenuation and structured noise.
 
-A particularly difficult form of noise present in FMRI is a low frequency
+A particularly difficult form of noise present in \ac{fMRI} is a low frequency
 drift, often characterized as a Wiener process \cite{Riera2004}.
@@
 It is clear that this drift signal is not solely
-due to a physiological effects, given its presence in cadavers and phantoms
+due to physiological effects, given its presence in cadavers and phantoms
 \cite{Smith1999}. Interestingly, it is usually spatially correlated, and
@@
 
-In order to characterize the noise, I analyzed resting state data.
+In order to characterize the noise, resting state data was analyzed.
 During resting state, the patient is shown no images, and he is asked
@@
 
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_29_49_9}}
-\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_34_43_24}}
-\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_22_38_23}}
-\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009_37_29_24}}
-
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
-%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}
-
-\caption{Q-Q Plots of normalized resting state data}
-\label{fig:QQDC}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_29_49_9}}
-\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_34_43_24}}
-\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_22_38_23}}
-\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
-\caption{Q-Q Plots of resting state data, using the BOLD signal steps.}
-\label{fig:QQDelta}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_29_49_9}}
-\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_34_43_24}}
-\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_22_38_23}}
-\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
-\caption{Q-Q Plots of resting state data, after the de-trending}
-\label{fig:QQSpline}
-\end{figure}
-
 Because most methods (including the one used in this paper)
@@
 sufficient condition for independence). Gaussianity is also a common
-assumption made in studies of FMRI data, though that assumption is not
+assumption made in studies of \ac{fMRI} data, though that assumption is not
 needed in this work. Regardless, comparing the distribution to a Gaussian
 is informative, so Q-Q plots are used to compare example data with the
-Normal distribution. Additionally, in FMRI data, the noise is often considered
+Normal distribution. Additionally, in \ac{fMRI} data, the noise is often considered
 to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
@@
 closely to the normal distribution.
-As expected, most of the autocorrelation disappears for the step data. Given
+As expected, most of the autocorrelation disappears for the step data
+(\autoref{fig:QQDelta}). Given
 that the steps seem to fit the Normal distribution, the low autocorrelation
@@
 Detrending the time-series by subtracting a spline fit to the distribution
-removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
+(\autoref{fig:QQSpline})
+removed much of the autocorrelation present in \autoref{fig:QQDC:C} and 
+\autoref{fig:QQDC:D},
 though not perfectly. Though the distributions still do not exactly fit
@@
 The non-stationary
-aspect of a Weiner process, presumably the result of integrating some
+aspect of a Wiener process, presumably the result of integrating some
 $\nu_x$, is difficult to compensate for, and so many methods
-have been developed to compensate for it. Tanabe et al. and Smith et al.
+have been developed to compensate for it. Tanabe et al. \cite{Tanabe2002}
+and Smith et al. \cite{Smith1999}
 demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
-of FMRI \cite{Tanabe2002, Smith1999}. It has been reported that in some studies as
+of \ac{fMRI}. It has been reported that in some studies as
 many as half the voxels
-benefited from detrending \cite{Smith2007}. In a head to head comparison,
-Tanabe et al.  showed that in most cases subtracting off
-a spline worked the best \cite{Tanabe2002}.
+benefited from detrending \cite{Smith2007}. In comparison of high pass 
+filters, 
+Tanabe et al. \cite{Tanabe2002} showed that in most cases subtracting off
+a spline worked the best.
 Unfortunately no method will
@@
 
-The method I used to calculate the spline was one knot for every 20
+The method used to calculate the spline was one knot for every 20
 measurements in an image. Thus a 10 minute session at a repetition time of
@@
 by definition the signal will have a median near zero.
-Unfortunately this is not the natural state of the BOLD signal. More specifically,
-when the signal is inactive, the BOLD response should be at 0\% change from
+Unfortunately this is not the natural state of the \ac{BOLD} signal. More specifically,
+when the signal is inactive, the \ac{BOLD} response should be at 0\% change from
 the base level; activation may then increase, or for short periods decrease from this base.
-Because most of the BOLD signal is above baseline, after removing the spline
+Because most of the \ac{BOLD} signal is above baseline, after removing the spline
 the resting state will be below 0\%.
@@
 by visual inspection.  In this work a simpler approach was used. To determine
-the DC gain I used a robust estimator of scale. The Median Absolute Deviation (MAD)
-proved to be accurate in determining how much to shift the signal up
-by. I tested both methods during the course of analysis, and found that the increase
+the DC gain a robust estimator of scale was used. The 
+\ac{MAD} proved to be accurate in determining how much to shift the signal up
+by. Both methods analysis were tested, and it was found that the increase
 in model complexity far outweighed the slight increase in flexibility. Other
-methods may work better, however the MAD worked well,
-as \autoref{fig:PreprocessedLowNoise} and \autoref{fig:PreprocessedHighNoise} show.
+methods may work better, however the \ac{MAD} worked well,
+as \autoref{fig:PreprocessedLowNoise} and 
+\autoref{fig:PreprocessedHighNoise} show. The \ac{DC} gain was then set to:
 
@@
 
+For more information on the \ac{MAD}, it is discussed in great detail in
+Iglewicz's text on robust estimation \cite{iglewicz1983robust}.
 A serious concern when adding constant values to
 real data is whether this will create false positives. This is a legitimate
-concern; however, a boosted response does not effect how well the BOLD model
-predicts the actual measurements; and as mentioned before, the DC signal of FMRI
-is never used.
+concern; however, a boosted response does not affect how well the \ac{BOLD} model
+predicts the actual measurements; and as mentioned before, the DC signal 
+of \ac{fMRI} is never used.
 
@@
 \label{sec:PriorDist}
-For the BOLD model described in \autoref{sec:BOLD Physiology}, several
+For the \ac{BOLD} model described in \autoref{sec:BOLD Physiology}, several
 different studies have endeavored to calculate parameters. The results
@@
 used for each may be found in \autoref{sec:Prior Works}. Unfortunately,
-Friston et al. only studied regions deemed active by the General
+Friston et al. \cite{Friston2002b} only studied regions deemed active 
+by the General
 Linear Model; and most other research used these results as
-the source for their priors \cite{Friston2002}.
-The one exception is to this is Johnston et al. which came to a extremely different
-distributions \cite{Johnston2008}. For a particle filter, the choice of a prior is
+the source for their priors.
+The one exception is to this is Johnston et al. \cite{Johnston2007} 
+which came to extremely different distribution. For a particle filter, 
+the choice of a prior is
 the single most important design decision. A very wide prior will require
@@
 to use priors that would give results consistent with previous works
-\cite{Friston2002}. This constrains the usefulness of the model to
+\cite{Friston2002b}. This constrains the usefulness of the model to
 areas that fall within the prior distribution, yet allowed results
@@
 exhaustive simulated annealing or hill climbing tests for multiple
-regions and multiple patients. That said, a significant amount of
-indeterminateness exists in the parameters, which is discussed in
+regions and multiple patients. That said, the parameters may  not be
+strictly identifiable, which is discussed in
 \autoref{sec:Single Voxel Simulation}
@@
 \includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response}
-\caption{Response to $0.1s$ impulses with the mean parameters from \cite{Friston2000}}
+\caption{Response to $0.1s$ impulses with the mean parameters from 
+Friston et al. \cite{Friston2000b}}
 \label{fig:MeanResponseF}
@@
 \includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response_johnston}
-\caption{Response to $0.1s$ impulses with the mean parameters from \cite{Johnston2008}}
+\caption{Response to $0.1s$ impulses with the mean parameters from 
+Johnston et al. \cite{Johnston2007}}
 \label{fig:MeanResponseJ}
@@
 
-The differences between the parameter estimates of Johnston et al. and
-Friston et al. are clearly visible in \autoref{fig:MeanResponseJ} and
-\autoref{fig:MeanResponseF} \cite{Johnston2008,Friston2002}.
-Therefore, to account for these discrepancies, I used tests on
-real data to adjust the distributions from those arrived at in
-Friston et al. \cite{Friston2002}. In tests on real data, the Johnston et al.
-distributions never converged to meaningful BOLD estimates.
+The differences between the parameter estimates of Johnston et al.
+\cite{Johnston2007} and
+Friston et al. \cite{Friston2002b} are clearly visible in 
+\autoref{fig:MeanResponseJ} and \autoref{fig:MeanResponseF}.
+Therefore, to account for these discrepancies, tests on
+real data was used to adjust the distributions from those arrived at in
+Friston et al. \cite{Friston2002b}. In tests on real data, the Johnston et al.
+\cite{Johnston2007}
+distributions never converged to meaningful \ac{BOLD} estimates.
 The priors used in the particle filter may be found in \autoref{tab:Prior}.
@@
 Note that although the mean remains the same for all the
-parameters other than $\epsilon$, the standard deviation were
+parameters other than $\epsilon$, the standard deviations were
 adjusted based on tests with real data. Also, the state variables
@@
 
-It is also import to use enough particles to obtain a
+It is also important to use enough particles to obtain a
 sufficiently dense approximation of the prior. For 7 dimensions,
@@
 as long as resampling is kept as a last resort, some over-smoothing
-doesn't impair convergence. Therefore, for this work I chose a Gaussian kernel of
-bandwidth equal to the original distribution's covariance. This
+doesn't impair convergence. Therefore, for this work a Gaussian kernel of
+bandwidth equal to the original distribution's covariance was used. This
 applies a large amount of smoothing to the distribution; however, on average
-resampling was only applied every 20 to 30 measurements, and because randomization
-is being applied to model updates this gives the filter some mobility.
+resampling was only applied every 20 to 30 measurements. At this rate,
+regularized resampling still gives the filter some ability to explore
+state space, without excessively revering convergence.
 
-Resampling is a not strictly necessary, but it increases the effectiveness
+Resampling is not strictly necessary, but it increases the effectiveness
 of the particle filter by adjusting the support to emphasize areas
@@
 tested. In addition to the Gaussian I also tested the Laplace and Cauchy
-distributions, both of which have wide tales.
+distributions, both of which have heavy tales.
 Wider tailed distributions don't down-weight
@@
 
-After trial and error, for this thesis I chose the Gaussian with a standard deviation
-of $0.005$. Although using an adaptive weighting function
+After trial and error, for this thesis the Gaussian with a standard deviation
+of $0.005$ was used. Although using an adaptive weighting function
 might be better, in tests this often led to unpredictable results.
@@
 taking a small sample from resting data and using the sample standard deviation.
-In this work, however, I set the standard deviation manually at $0.005$,
+In this work, however, the standard deviation was manually set at $0.005$,
 because it gave the best consistency.
@@
 \subsection{Runtime}
-The run-time for a single voxel depends on the several factors. First, the
-overall length of the signal being analyzed. For 1000 measurements it takes
-approximately 6 minutes. On the other hand, in real circumstances the
-length tends to be around 150 measurements which takes around 40 seconds (for 1000
-particles, 1400 integration points and a Quad Core CPU). The size of
-integration steps are also crucial. Using step sizes above
+The run-time for a single voxel depended on several factors. First, the
+overall length of the signal being analyzed directly scaled the run-time. 
+For 1000 measurements it took
+approximately 6 minutes. On the other hand, in real \ac{fMRI} the
+length tends to be around 150 measurements which took around 40 seconds 
+(for 1000 particles, 1400 integration points and a Quad Core CPU). The number of
+integration steps was also crucial. Using step sizes above
 $0.001$ seconds is not recommended. In most cases millisecond resolution
-is fine; however, when generating simulated data I found that at times it was
-still not enough. This is problematic in the actual particle
+was fine; however, when generating simulated data at times this was
+still not enough. This could cause problems in the actual particle
 filter since, given the large number of simultaneous integrations taking
-place, its probable that a few particles will fail and be unfairly thrown away.
+place, its probable that a few particles would fail and be unfairly thrown away.
 To prevent such events, 1400 integration points per second were used throughout
@@
 number of particles stays high, resampling could take a long time to occur.
-For this reason, rather than allowing the particle filter to continue on
+For this reason, rather than allowing the particle filter to continue 
 with this large number of particles, after 20 seconds have passed the

### 2010-09-02 doc/thesis/particlefilters.tex
just need to edit discussion/conclusion, change captions and copy some graphs from defense

@@
 where $N_p$ is the number of particles or points used to describe the prior 
-using a Mixture PDF. 
+using a Mixture \ac{PDF}. 
 Note that any exponents will be explicitly labeled as such, to avoid confusion with
@@
 \begin{equation}
-\alpha(X) \approx P(x_0) = \sum_{i=0}^{N_p} w^i\delta(X - x^i_0 ) dx
+\alpha(X) \approx P(x_0) = \int_{-\infty}^{\infty}
+    \sum_{i=0}^{N_p} w^i\delta(X - x^i_0 ) dx
 \end{equation}
@@
 \end{equation}
-The benefit of this choice for importance density is that an approximation for
-$P(x_k | x_{k-1})$ is freely available; its simply the set of particles propagated
+This choice for importance density is conventient since its simply 
+the set of particles propagated
 forward in time using the state equations. Additionally it makes
-updating weights simple, as seen in \autoref{eq:weightevolve}.
+updating weights simple, as seen later in \autoref{eq:weightevolve}.
 
@@
 \end{equation}
-Plugging \autoref{eq:ImportanceDensity} and the result of \autoref{eq:UpdateBayes}
+Inserting \autoref{eq:ImportanceDensity} and the result of \autoref{eq:UpdateBayes}
 into \autoref{eq:weightfunc} leads to:
@@
 \label{sec:Particle Filter Resampling}
+\begin{figure}
+\includegraphics[width=16cm]{images/particle_filter2}
+\caption[Example Particle Filter Progression]{a) Initial particle locations
+uniformly distributed in [-10, 10], with
+weights set to give the State Prediction (green). When a measurement is
+received it converted to a distribution approximating uncertainty. This
+distribution is then used to alter the weights of the existing particles.
+b) Another measurement is applied similar to a), c) Resampling draws
+new particles in locations according to the Mixture \ac{PDF} approximated 
+by the particle weights. d) Another measurement is applied as before,
+but now with non-uniform particle locations.}
+\end{figure}
 
-\begin{figure}
-\includegraphics[width=16cm]{images/particle_filter}
-\caption{Particle Filter progression, note that the initial support is flat; the particles
-are equally spaced between -10 and 10}
-\end{figure}
 As a consequence 

### 2010-09-02 doc/thesis/results.tex
fixed bad label.
just need to edit discussion/conclusion, change captions and copy some graphs from defense

@@
 and second of a single slice (64x64 voxels).  The single time series tests
-investigate the ability of particle filters to estimate parameters of the BOLD model
+investigate the ability of particle filters to estimate parameters of the \ac{BOLD} model
 in a noisy environment. Single voxel tests were
@@
 models noise realistically. POSSUM demonstrates the
-particle filter modeling the BOLD signal on a large scale.
+particle filter modeling the \ac{BOLD} signal on a large scale.
 In the POSSUM simulation there were four
-discrete parameter sets driving the timeseries, although each voxel had a
+discrete parameter sets driving the time series, although each voxel had a
 different tissue composition and a different
@@
 \label{sec:Single Voxel Simulation}
-Given the state-space equations for the BOLD signal, simulating a single time
-series was straight forward. After generating a true signal,
+Given the state-space equations for the \ac{BOLD} signal, simulating a single time
+series was straightforward. After generating a true signal,
 identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
 process with Gaussian I.I.D. steps were added to the true signal. Finally a
-carrier level was added, since BOLD is typically
+carrier level was added, since \ac{BOLD} is typically
 measured as a \% difference from the baseline. The particle filter
@@
 were performed on the particle filter.
-The first test demonstrates the particle filter on a very long FMRI
-run, and discusses the inherent learn-ability of the BOLD model
+The first test demonstrates the particle filter on a very long \ac{fMRI}
+run, and discusses the inherent learn-ability of the \ac{BOLD} model
 (\autoref{sec:VeryLongSim}).
@@
 investigate the problem of false-positives. As the first
-round of tests show, given the presence of an underlying BOLD signal,
+round of tests show, given the presence of an underlying \ac{BOLD} signal,
 the particle filter is excellent at finding the most probable cause of
@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
-\caption{BOLD estimate converging for a very long FMRI run. Darker bars indicate
+\caption[\ac{BOLD} estimate converging for a very long \ac{fMRI} run.]{\ac{BOLD} estimate 
+converging for a 2400 second \ac{fMRI} run. Darker bars indicate
 bins with more regions. }
@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
-\caption{BOLD estimate converging for a very long FMRI run, first 500 seconds. Darker bars
+\caption{\ac{BOLD} estimate converging for a very long \ac{fMRI} run, first 500 seconds. Darker bars
 indicate bins with more particles.}
@@
 
-To begin the single voxel simulation; I generated a signal using the following parameters:
+To begin the single voxel simulation; a signal using the following parameters
+was generated:
 $\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
@@
 determine the inherent variance in the model parameters, a very long simulation
-with randomly generated impulse stimuli was created. The preprocessed timeseries
+with randomly generated impulse stimuli was created. The preprocessed time series
 and the final estimate are shown in \autoref{fig:long_converge};
 the final covariance matrix of the parameters is in \autoref{tab:long_cov}.
-Note that even though the BOLD response converged (\autoref{fig:long_converge},
+Note that even though the \ac{BOLD} response converged (\autoref{fig:long_converge},
 \autoref{fig:long_converge_500}),
@@
 In spite of the noisy input (green line in \autoref{fig:long_converge}), the estimates
-of the BOLD were actually very close to the true (noise-free) BOLD signal.
+of the \ac{BOLD} were actually very close to the true (noise-free) \ac{BOLD} signal.
 
@@
 \hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \ac{RMSR} & \ac{RMSE} \\
 \hline
@@
 The following tests all use less data (shorter sequences), on par with
-the average length of an FMRI session. For this low noise case
+the average length of an \ac{fMRI} session. For this low noise case
 $(\sigma_y = 0.001,\ \sigma_x = .0005)$, the eleven realizations are shown in
@@
 could have some effect on the resulting fit; thus the preprocessed signal is compared
-to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
+to the true \ac{BOLD} signal in \autoref{fig:PreprocessedLowNoise}.
 Overall, \autoref{fig:PreprocessedLowNoise} shows that the preprocessing was
-effective at removing trends, although the spline did struggle to match the signal
-toward the end. This slight drift effect was caused by the final median being
-well above the base
+effective at removing trends, although the spline did leave some noise 
+structure toward the end (\autore{fig:LowNoiseRealization}). This slight 
+drift effect was caused by the final median being well above the base
 level.
@@
 
-Note that throughout the results, unless otherwise specified,
-the term \emph{residual} will be defined as the square root of the mean squared residual,
+For the purpose of quantifying the quality of fit, the 
+\ac{RMSR} was used:
 \begin{equation}
-\text{Residual} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - y_k)^2}
+\text{RMSR} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - y_k)^2}
 \end{equation}
-where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the preprocessed output
-sampled at time $k$. Note the subtle distinction between this and what will be
-termed the \emph{error} which is defines as
+where $\hat{y}_k$ is the estimated output at time $k$ and $y_k$ is the 
+preprocessed output
+sampled at time $k$. Note the subtle distinction between this and
+the \ac{RMSE}:
 \begin{equation}
-\text{Error} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - Y_k)^2}
+\text{RMSE} = \sqrt{\frac{1}{N}\sum (\hat{y}_k - Y_k)^2}
 \end{equation}
-where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying (free of noise)
-signal. Throughout this section and the next, the terms error and residual
-will be referencing these values.
+where $\hat{y}_k$ is the estimated output, and $Y_k$ is the underlying 
+(free of noise) signal.
 
@@
 runs, yet the sum of the individual time constants
-($\tau_f$, $\tau_s$ and $\tau_0$) were more consistent. On average the
+($\tau_f$, $\tau_s$ and $\tau_0$) was more consistent. On average the
 time constants fell short of the true time constant. This could be a limitation
@@
 to small differences in the time constants.
-The convergence properties of the first run in \autoref{tab:LowNoiseResults}
+The convergence of the first run in \autoref{tab:LowNoiseResults}
 demonstrates the migration of parameters through the run. In spite of the
-significant differences in parameter estimates, the estimated BOLD
+significant differences in parameter estimates, the estimated \ac{BOLD}
 consistently performed well
@@
 \hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual  & Error\\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \ac{RMSR} & \ac{RMSE} \\
 \hline
@@
 
-\autoref{fig:NoiseComparisonJustTwo} shows two of the runs in more detail.
-There appears
-to be more drift than the 20 measurements per knot could fit, which explains
+\autoref{fig:NoiseComparisonJustTwo} shows two runs in more detail; these
+results show that more drift was present that 20 measurements per knot 
+could fit. Insufficient flexibility of the spline explains
 the prolonged increase at 170 seconds in Run 1; although such areas
@@
 undershoot better.  \autoref{tab:HighNoiseResults} shows
-the error  for all eleven runs, and highlights the two runs analyzed
+the \ac{RMSE} for all eleven runs, and highlights the two runs analyzed
 in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
@@
 is the explanation for the perceived jumps in the histograms.
-Clearly the additional noise resulted in much more sporadic results (\autoref{tab:HighNoiseResults}).
+Clearly the additional noise resulted in less consistent results 
+(\autoref{tab:HighNoiseResults}).
 This is often the
@@
 weighting function's variance being smaller than the measurement noise ($0.005$ vs. $0.01$).
-The error clearly suffers due to this effect (\autoref{tab:HighNoiseResults}).
+The \ac{RMSE} clearly suffers due to this effect (\autoref{tab:HighNoiseResults}).
 Note that neither Run 1 nor Run 2 estimated the underlying state well (\autoref{fig:ConvergenceRuns1c}
@@
 to 0, in effect simulating a brain region with no response to the stimuli.
-This test was used to determine how the output of a pure noise timeseries
+This test was used to determine how the output of a pure noise time series
 is different from that of a simple noisy signal (as in the previous two sections).
@@
 \hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  Residual   \\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & \ac{RMSR} \\
 \hline
@@
 well within the range of $0.005$, the standard deviation of the weighting function.
-Note that the residuals
-were actually lower than the residuals in the low noise simulation from \autoref{sec:SimLowNoise}
-and the parameter estimates were extremely consistent across 11 runs.
-The low residual was caused by the overall
+Note that the \acs{RMSR}
+were actually lower than the \acs{RMSR} in the low noise simulation 
+from \autoref{sec:SimLowNoise}
+and the parameter estimates were more consistent across 11 runs.
+The low \ac{RMSR} was caused by the overall
 signal being significantly smaller than any previous simulation.
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3cm,width=15cm]{images/fits_noiseonly_high}
-\caption{BOLD estimates for the non-active, high noise signal. Note the line thickness is caused
+\caption{\ac{BOLD} estimates for the non-active, high noise signal. Note the line thickness is caused
 by all the estimates overlapping.}
@@
 Because of the variability in the signal levels, the
-raw residual cannot be used to rate the fit. As demonstrated by \autoref{tab:NoiseOnlyResults},
-a low residual does not necessarily indicate a good fit.
-Therefore, a normalized version of the residual was used. To normalize the
-residual, both the estimated and preprocessed BOLD signals were divided by an estimator
-of scale and then the new residual was calculated.
-Considering the tendency of FMRI to have large unexplainable
+raw \ac{RMSR} cannot be used to rate the fit. As demonstrated by \autoref{tab:NoiseOnlyResults},
+a low \ac{RMSR} does not necessarily indicate a good fit.
+Therefore, a normalized version of the \ac{RMSR} was used. To normalize the
+residual, both the estimated and preprocessed \ac{BOLD} signals were divided by an estimator
+of scale and then the new \ac{RMSR} was calculated.
+Considering the tendency of \ac{fMRI} to have large unexplainable
 peaks and troughs, the Median-Absolute-Deviation (MAD) was used (\autoref{eq:mad}).
@@
 deviation, and thus a good estimator of the scale of the input signal.
-The normalized residual values are
+The normalized \ac{RMSR} values are
 shown in \autoref{tab:SingleVoxelActivationComparison}. A second potential method of
@@
 \end{tabular}
-\caption{Mutual Information and the normalized error,
+\caption{Mutual Information and the normalized \ac{RMSE},
     for signal/noise configurations.}
@@
 
-Comparing the results of \autoref{sec:SimHighNoise} and \autoref{sec:PureNoiseHighMag}
+Comparing the results of \autoref{sec:SimHighNoise} and 
+\autoref{sec:PureNoiseHighMag}
 in \autoref{tab:SingleVoxelActivationComparison},
-distinguishing between these cases with ether normalized residual or mutual information
-is not clear cut. While the average mutual information is more than 10 times
+distinguishing between these cases with either normalized \ac{RMSR} or 
+mutual information is not clear cut. While the average mutual information 
+is more than 10 times
 the average mutual information in the two non-signal cases, the maximum
-mutual information of the low noise/no signal case exceeds the minimum M.I. of the high noise/signal
+mutual information of the low noise/no signal case exceeds the minimum M.I. 
+of the high noise/signal
 case. The Low Noise/signal determination is easier to make; given
-the minimum mutual information is above $.8$ and the maximum normalized residual is below $.6$.
+the minimum mutual information is above $.8$ and the maximum normalized 
+\ac{RMSR} is below $.6$.
 However, it is worth noting that the worst case scenario  for mutual information (maximum) in the
@@
 
-There were two primary purposes of these tests. First, given the nature of monte-carlo
+There were two primary purposes of these tests. First, given the nature of Monte Carlo
 techniques it is important to ensure consistency of results. Although the parameter
-sets were inconsistent, the quality of the fit for the BOLD output was often
+sets were inconsistent, the quality of the fit for the \ac{BOLD} output was often
 accurate even when noise was drastically increased. The second purpose was to validate
-Mutual Information and the normalized residual as measurements of output quality.
+Mutual Information and the normalized \ac{RMSR} as measurements of output quality.
 Although not perfect, both methods do provide a decent metric.
@@
 \includegraphics[height=8cm]{images/simregions.png} \label{fig:simslice_mask} }
-\subfigure[Residual Map. Scale is Normalized error. Lower (yellow) is better.]
+\subfigure[Normalized \ac{RMSR} Map. Scale is Normalized \ac{RMSE}. Lower (yellow) is better.]
 {\label{fig:simslice_hm_res} \includegraphics[height=8cm]{images/sim_hm}}
@@
 Region 3: $0.39$.]{\includegraphics[height=8cm]{images/snr_hm.png}\label{fig:SimSNRhm}}
+\caption{Comparison of activation with greymatter, parameter regions.}
 \label{fig:simslice_hm}
-\caption{Comparison of activation with greymatter, parameter regions.}
 \end{figure}
 
-To test the usefulness of the particle filter on a larger scale, I used a modified version
+To test the usefulness of the particle filter on a larger scale, a modified version
 of the FSL tool
-POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
+POSSUM was used to generate an entire \ac{fMRI} image from a parameter map. The parameter map was generated
 by taking an existing activation map and assigning discrete parameter sets to each region.
@@
 to take a parameter map and generate activation levels depending on the parameters at that
-point. The patch for POSSUM will be made available. It is worth noting is that the noise
+point. The patch for POSSUM will be made available. Note that the noise
 level was set to an SNR of 20, but due to changes
@@
 
-For each time-series in the simulated FMRI image, the final parameters were saved
+For each time series in the simulated \ac{fMRI} image, the final parameters were saved
 into a parameter map. This parameter map could then be compared to the map used to generate the
 simulated data. Additionally a new simulation using the calculated parameters could also be
-generated to test the difference in BOLD levels between the real parameters and the
+generated to test the difference in \ac{BOLD} levels between the real parameters and the
 estimated ones. Since the parameters were far from orthogonal,
@@
 threshold. For this reason, the only areas
-with significant estimates of the BOLD time series were 1,2 and 3.
+with significant estimates of the \ac{BOLD} time series were 1,2 and 3.
 Notice that the
-regions 1, 2 an 3 stick out in both the residual and the mutual information
+regions 1, 2 an 3 stick out in both the \ac{RMSR} and the mutual information
 map, indicating that the particle filter was successful in matching those regions.
 Mutual information was an extremely successful metric, with the exception of
-a few false positives. \autoref{fig:sim_hm2} shows what the heat map looks
+a few false positives (\autoref{fig:simslice_hm_res}). 
+\autoref{fig:sim_hm2} shows what the heat map looks
 like when the threshold is raised from $0.1$ to $0.15$.
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist1}
-\caption{Histogram of estimated parameters in section 1 in voxels with mutual information greater
-than $0.15$}
-\label{fig:slicesim_hist1}
+\caption{Histogram of estimated parameters in section 1 in voxels 
+with mutual information greater than $0.15$} \label{fig:slicesim_hist1}
 \end{figure}
@@
 If the thresholds are left at the values of \autoref{fig:simslice_hm}, then
-the normalized residual gave $2/1479$  false positives and mutual information gave
-$11/1479$. However, normalized residual also clearly missed some active voxels, although
+the normalized \ac{RMSR} gave $2/1479$  false positives and mutual information gave
+$11/1479$. However, normalized \ac{RMSR} also clearly missed some active voxels, although
 false negatives are harder to quantify because of the presence of white matter.
@@
 The thresholds applied to this slice, both for mutual information and
-residuals somewhat arbitrary. Applying a
+\acs{RMSR} are arbitrary. Applying a
 threshold is helpful for visualization but is rarely useful for further analysis.
@@
 the false positives present in the Mutual Information
-map are different from those in the residual map. This furthers the argument
+map are different from those in the \ac{RMSR} map. This furthers the argument
 for combining the two metrics to increase power. Although
@@
 not actually the case. POSSUM simulates different tissues, and white matter
-does not typically  have a BOLD response. This explains some of the holes in regions
+does not typically  have a \ac{BOLD} response. This explains some of the 
+holes in regions
 2 and 3. These results certainly indicate that the particle filter is effective

### 2010-09-02 doc/thesis/resultsB.tex
just need to edit discussion/conclusion, change captions and copy some graphs from defense

@@
 \label{sec:RealData}
-The ultimate goal of modeling the BOLD response is to estimate parameters
-in real FMRI data. The most basic use of modeling the BOLD signal is to
+The ultimate goal of modeling the \ac{BOLD} response is to estimate parameters
+in real \ac{fMRI} data. The most basic use of modeling the \ac{BOLD} signal is to
 locate activation. A voxel is considered active when the stimulus
-is the primary drive for the BOLD response. This is in contrast to 
+is the primary drive for the \ac{BOLD} response. This is in contrast to
 most voxels which are controlled by intermediate or completely unrelated factors.
@@
 parameter estimates in such regions are impossible.
-Because SPM is the de facto standard for localizing activation,
+Because \ac{SPM} is the de facto standard for localizing activation,
 this section compares its output with that of the particle filter.
 
-Note that SPM must pre-process the image with a spatial smoothing 
+Note that \ac{SPM} must pre-process the image with a spatial smoothing
 filter. For this work SPM8 was smoothed with a
-$8\text{mm} \times  8\text{mm} \times 8\text{mm}$ Full-Width Half Maximum
-(FWHM) Gaussian kernel. 
+$8\text{mm} \times  8 \text{mm} \times 8\text{mm}$ 
+\ac{FWHM} Gaussian kernel.
 Additionally, SPM8 applied a high pass filter (with a cut
 off based on a globally estimated autocorrelation). Thus the preprocessing pipeline
-of SPM is very different from that of the particle filter. SPM8 also outputs
+of \ac{SPM} is very different from that of the particle filter. SPM8 also outputs
 a $t$-statistic for each voxel, whereas the
@@
 at every voxel. To validate the quality of the particle filter, the results
-were compared with SPM, both in terms of the location and the fit.
+were compared with \ac{SPM}, both in terms of the location and the fit.
 
@@
 \label{sec:ExperimentConfig}
-For the FMRI data discussed in \autoref{sec:RealData}, tests were 
+For the \ac{fMRI} data discussed in \autoref{sec:RealData}, tests were
 performed on a right handed volunteer using a GE SIGNA HDx 1.5 Tesla
-scanner with a single echo EPI sequence. Slice spacing was
-5mm, and pixel sizes were 3.75mm.  Repetition Time was 2.1s, 
-Echo Time was 40 ms and the imaging frequency was set to $63.854$MHz.
+scanner with a single echo \ac{EPI} sequence. Slice spacing was
+5 mm, and pixel sizes were 3.75 mm.  Repetition Time was 2.1 s,
+Echo Time was 40 ms and the imaging frequency was set to $63.854$ MHz.
 The image resolution was $64 \times 64 \times 28$.
@@
 double flash and was asked to respond with a right handed
-or left handed finger tap, respectively. The FMRI began 18.9 seconds
-before the beginning of the experiment, to allow for transients 
-in to image to settle out. The timing of the flashes are shown 
+or left handed finger tap, respectively. The \ac{fMRI} began 18.9 seconds
+before the beginning of the experiment, to allow for 
+the Gradient Echo spins to to settle out. 
+The timing of the flashes is shown
 below with time 0 corresponding to the beginning of
-the 10th TR (so 9 images were dropped).
+the 10th \ac{TR} (so 9 images were dropped).
 
@@
 After dropping the first 9 volumes, each of the remaining volumes was
-co-registered with the new first volume. At this point the SPM method
+co-registered with the new first volume. At this point the \ac{SPM} method
 diverges from the experimental method. For the particle filter, detrending
 was then applied as discussed in \autoref{sec:Detrend} and the resulting
-data was processed with the particle filter. To generate the SPM output
+data was processed with the particle filter. To generate the \ac{SPM} output
 in this section, the co-registered data was spatially smoothed,
 and then filtered with an adaptive cut off (which is built into the
-SPM analysis). For the SPM analysis, the Canonical HRF was used
+\ac{SPM} analysis). For the \ac{SPM} analysis, the Canonical \ac{HRF} was used
 and model time derivatives were included although not in the contrast
 vector. All other settings used in SPM8 were left at the default
-for analyzing FMRI data.
+for analyzing \ac{fMRI} data.
 
@@
 Note that the scales for all three images are different, because the metrics are different.
-SPM measures using $t$-tests to determine the likelihood of a false positive. \autoref{fig:hm_canon_pfilter}
+\ac{SPM} measures using $t$-tests to determine the likelihood of a false positive. \autoref{fig:hm_canon_pfilter}
 uses simple normalized residuals, meaning that lower indicates less error.
@@
 filter data shows a large number of false positives, however application of a threshold
-of $.85$ on the residual map removes these false positives. Similarly, in the mutual information
-map, the false positives may be elminated by upping the threshold to $.15$. However, just
-because the results disagree with SPM does not necessarily mean they are false positives.
-SPM operates on smoothed data (8mm x 8mm x 8mm), so there are certainly active 
+of $0.85$ on the residual map removes these false positives. Similarly, in the mutual information
+map, the false positives may be eliminated by upping the threshold to $0.15$. However, just
+because the results disagree with \ac{SPM} does not necessarily mean they are false positives.
+\ac{SPM} operates on smoothed data (8 mm x 8 mm x 8 mm), so there are certainly active
 areas that have been missed because of the smoothing. Note that throughout this
@@
 \subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
-\caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher 
+\caption{\ac{SPM} results. Units of activation are in Student's $t$-scores; higher indicates higher
         assurance that the signal cannot have occurred through noise alone.
@@
 \subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/1_spm_37_14_7}}
-\caption{Section 1, Estimated vs. Actual BOLD response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
+\caption{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 \subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14.5cm]{images/2_spm_34_12_7}}
-\caption{Section 2, Estimated vs. Actual BOLD response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
+\caption{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 \subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/3_spm_23_21_7}}
-\caption{Section 3, Estimated vs. Actual BOLD response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
+\caption{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 \subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/4_spm_26_15_7}}
-\caption{Section 4, Estimated vs. Actual BOLD response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
+\caption{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_29_9_13}}
-\caption{Section 5, Estimated vs. Actual BOLD Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
+\caption{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
 %\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
@@
 \subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/6_spm_36_17_19}}
-\caption{Section 6, Estimated vs. Actual BOLD Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
+\caption{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 %\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_25_34_25}}
-\caption{Section 7, Estimated vs. Actual BOLD Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
+\caption{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
@@
 
-I chose several voxels to further investigate from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
+Several voxels were chosen for further investigation from \autoref{fig:hm_canon_pfilter} and \autoref{fig:hm_canon_spm}.
 
 Section 1 (\autoref{fig:comp1}) had a very high $t$-score, high mutual information and
-a low error. Visual inspection makes it obvious that this voxel fits the BOLD model. A region
-such as this would be called strong activation by Neurologists. The
-reason for the cleaner looking signal in SPM is the additional Gaussian spatial smoothing
+a low error. Visual inspection makes it obvious that this voxel fits the \ac{BOLD} model. A region
+such as this would be called strong activation by neurologists. The
+reason for the cleaner looking signal in \ac{SPM} is the additional Gaussian spatial smoothing
 applied in preprocessing. This
-has allowed for an extremely good fit in SPM, although the particle filter handles the noise
+has allowed for an extremely good fit in \ac{SPM}, although the particle filter handles the noise
 well.
@@
 Section 2 is a difficult voxel to assess (\autoref{fig:comp2}).
-While certain peaks seem to correlate, over all 
+While certain peaks seem to correlate, overall
 the particle filter input signal is extremely noisy.
-This would appear to be a false positive on SPM's part because of the Gaussian smoothing.
+This would appear to be a false positive on \ac{SPM}'s part because of the Gaussian smoothing.
 This voxel is not being driven directly by the input, although it
@@
 Note that in both cases the Mutual Information was extremely low, meaning that
-according the M.I. these were not good fits. Both section 3 and 4 have 
-significant peaks, begging the question of whether they may be active,
+according to the M.I. these were not good fits. Both section 3 and 4 have
+significant peaks, raising the question of whether they may be active,
 but not correlated directly with the input.
@@
 It is questionable whether the fifth voxel should be considered inside the
-brain; however, like section 2, SPM deemed it active (\autoref{fig:comp5}). 
+brain; however, like section 2, \ac{SPM} deemed it active (\autoref{fig:comp5}).
 Also like section 2, the most likely cause of the problem is the smoothing
-applied for SPM. There are certain peaks that certainly match the
-BOLD signal, yet that is only a small part of the whole. 
+applied for \ac{SPM}. There are certain peaks that certainly match the
+\ac{BOLD} signal, yet that is only a small part of the whole.
 Take for instance the measurements between $30$ and $40$ seconds.
-In that interval there is a significant spike in the BOLD signal, yet the
-actual measured signal declines. This is an indication that the BOLD signal is not directly
-being driven by the input. The SPM fit is not particularly good either,
+In that interval there is a significant spike in the \ac{BOLD} signal, yet the
+actual measured signal declines. This is an indication that the \ac{BOLD} signal is not directly
+being driven by the input. The \ac{SPM} fit is not particularly good either,
 with the $t$-score barely above the threshold of $4$.
@@
 Section 6 (\autoref{fig:comp6}) is a clear example of a false negative
-from SPM. The BOLD model fit is extremely good, yet the SPM signal looks completely
+from \ac{SPM}. The \ac{BOLD} model fit is extremely good, yet the \ac{SPM} signal looks completely
 like noise. One possible reason for this is that the low peak signal ($<0.03$),
@@
 and is a perfect example of the false positive problem.
-The signal oscillates far faster than the BOLD estimate, yet 
+The signal oscillates far faster than the \ac{BOLD} estimate, yet
 somehow the mutual information reaches $.1052$ and the normalized residual is just below $1.1$.
@@
 Although the parameters are not uniquely identifiable by a single time-series, that
-does not mean estimating them is not without benefit. The parameters still contain useful
+does not mean the estimates are useless . The parameters still contain useful
 information about the system. Additionally, as an aggregate they form a distribution of
@@
 parameters are migrating quite a bit during particle filter convergence.
-It can be said that all of these regions had relatively accurate BOLD 
+It can be said that all of these regions had relatively accurate \ac{BOLD}
 measurements, and yet the final parameter estimates vary widely.
@@
 The activation maps generated by the particle filter are very similar
-to those produced by SPM. This indicates that the particle filter is
-successfully estimating the BOLD output, in spite of the lack of spatial
+to those produced by \ac{SPM}. This indicates that the particle filter is
+successfully estimating the \ac{BOLD} output, in spite of the lack of spatial
 smoothing. Comparing the mutual information metric with the normalized residual,
 M.I. gives much better clusters, although that doesn't necessarily mean it
-is better. Although many of the differences between SPM and the particle filter 
+is better. Although many of the differences between \ac{SPM} and the particle filter
 were driven by differences in preprocessing, it is worth noting that
-the preprocessing applied by SPM are necessary for SPM to give decent results.
-Without spatial smoothing SPM is unable to cope with additional noise, and
+the preprocessing applied by \ac{SPM} is necessary for \ac{SPM} to give decent results.
+Without spatial smoothing \ac{SPM} is unable to cope with additional noise, and
 gives significant numbers of false positives.
-
 Although the parameter maps have some spatial coherence, the histograms

### 2010-09-02 doc/thesis/background.tex
just need to fix figure captions for multivoxel cases. And fix bib

@@
 \includegraphics[scale=.7]{images/HRF}
-\caption{Canonical Hemodynamic Response Function (y-axis units are arbitrary because 
-normalization is performed).}
+\caption[Canonical Hemodynamic Response Function]
+{Canonical Hemodynamic Response Function, normalization makes $y$ units
+arbitrary}
 \label{fig:HRF}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption[Q-Q Plot Of Standard Normal and Result of $0.1$s BOLD Simulation Output]
-{Q-Q Plot Of Standard Normal and Result of $0.1$ BOLD Simulation Output. A
-perfect Gaussian would fall on the green line}
+\caption[Q-Q Plot Of Standard Normal and Result of $0.1$ s BOLD Simulation Output]
+{Q-Q Plot Of Standard Normal and Result of $0.1$ s BOLD Simulation Output. A
+perfect Gaussian would fall on the green line.}
 \label{fig:transp1s}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
-\caption{Q-Q Plot Of Standard Normal and Result of $1$s BOLD Simulation Output}
+\caption[Q-Q Plot Of Standard Normal and Result of $1$ s BOLD Simulation Output]
+{Q-Q Plot Of Standard Normal and Result of $1$ s BOLD Simulation Output. A
+perfect Gaussian would fall on the green line.}
 \label{fig:trans1s}

### 2010-09-02 doc/thesis/discussion.tex
just need to edit discussion/conclusion, change captions and copy some graphs from defense
just need to fix figure captions for multivoxel cases. And fix bib

@@
-\chapter{Discussion}
+\chapter{Discussion and Conclusions}
 \label{sec:Discussion}
-\section{Review of Results}
+\section{Overview of Results}
+\subsection{Parameter Estimation}
 % Overview of results
@@
 incapable of learning the model. However, given the high quality
-of BOLD estimates, this is almost certainly not the case. The other
+of \ac{BOLD} estimates, this is almost certainly not the case. The other
 possible explanation is that the given stimulus is not sufficiently
@@
 stimulus sequence could differentiate all the parameters.
-Its also extremely interesting that the histograms of parameters' mean
+It is also extremely interesting that the histograms of parameters' mean
 approach the width of the prior, even when the particle filter converges
-properly to far thinner distributions. Its possible the prior
+properly to far thinner distributions. It is possible the prior
 distribution is far wider than
@@
 
-Thus there is little doubt that the parameters of the BOLD model are under
-constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
-it is a notable conclusion \cite{Deneux2006}. Other methods
+Thus there is little doubt that the parameters of the \ac{BOLD} model are under-constrained. 
+While unsurprising given the sensitivity analysis by Deneux et al. \cite{Deneux2006},
+it is a notable conclusion. Other methods
 that calculate a point estimate of the parameters, such as least squares
@@
 capacity to estimate such parameters. While estimates of
-the  BOLD signal may still be correct, the
+the  \ac{BOLD} signal may still be correct, the
 underlying parameters and state variables cannot be described without using
 a joint probability distribution function. In this sense, particle
-filters represent an important step forward in BOLD parameter
+filters represent an important step forward in \ac{BOLD} parameter
 estimation. Representing uncertainty with a mean
@@
 Because of the ill-defined nature of the parameters, the results
-of Friston et al. may have been wider than true parameter distribution
-\cite{Friston2002}. Future works that attempt to calculate parameters
-of the BOLD model should be run multiple times to ensure consistency,
+of Friston et al. \cite{Friston2002b} may have been wider than true parameter distribution
+. Future works that attempt to calculate parameters
+of the \ac{BOLD} model should be run multiple times to ensure consistency,
 which most likely cannot be attained. This result is particularly
@@
 As the results in \autoref{sec:RealData} show, the heat maps, especially
-those of mutual information, closely resembled the results of SPM. While the activation
+those of mutual information, closely resembled the results of \ac{SPM}. While the activation
 tests were more sensitive, there were some additional false positives,
 though the problem is difficult to quantify. Regions with M.I. above
-$0.15$ consistently fit the FMRI data well, and simulations showed that
+$0.15$ consistently fit the \ac{fMRI} data well, and simulations showed that
 the particle filter performed extremely well in the face of significant
-noise. In sum, the estimated BOLD output remained consistent despite
+noise. In sum, the estimated \ac{BOLD} output remained consistent despite
 large swings in parameters.
@@
 %% Cons
-%%% Computation longer than SPM
+%%% Computation longer than \ac{SPM}
 %%% Harder to interpret
@@
 %%% More intuitive
-\section{Particle Filter Review}
+\subsection{Particle Filter Review}
 The Particle Filter algorithm was originally designed for on-line parameter
 estimation. For this reason, there is no guarantee of optimality or even
-convergence for finite measurements. However, for the BOLD nonlinear ODE
+convergence for finite measurements. However, for the \ac{BOLD} nonlinear \ac{ODE}
 this is less of a concern than it might first appear to be. For this
@@
 however, it is free from the uncertainty of whether a quadratic approximation is
-sufficient for the BOLD model.
+sufficient for the \ac{BOLD} model.
 
@@
 %being violated. Fewer and more realistic assumptions also make the particle
-%filter more robust to unforeseen difficulties in FMRI data.
+%filter more robust to unforeseen difficulties in \ac{fMRI} data.
 
@@
 %Finally, rather than adding a constant to the detrended data before
-%applying the particle filter, a DC gain parameter could be added to the model. In
+%applying the particle filter, a \ac{DC} gain parameter could be added to the model. In
 %tests, this could work well, however, it often did not. The problem with this
@@
 %dominated the results by smoothing out all the peaks and troughs. The result is that
-%no other parameters alter the estimated BOLD signal.
+%no other parameters alter the estimated \ac{BOLD} signal.
 %\begin{figure}
@@
 
-\chapter{Future Work}
+\section{Conclusions}
+\label{sec:Conclusion}
+This work has demonstrated the use of the particle filter to
+learn parameters of the \ac{BOLD} model. Since the inception of the
+\ac{BOLD} equations, many attempts have been made use \ac{fMRI} data to
+to learn the model parameters. These attempts have typically either been extremely
+slow or relied on extensive assumptions. While the particle filter method
+is not quick, 40 seconds to analyze each voxel is well within the capabilities
+of the typical research lab. Previous attempts have also treated
+the problem as if there were a single solution.
+
+One significant finding of this work, that would not be clear without
+calculating a full posterior distribution, is the interplay
+between the parameters. The results of simulations clearly demonstrate
+that identifying a single set of parameters is not possible with this
+model. Although sensitivity tests in Deneux et al. \cite{Deneux2006} certainly hinted
+at this, the current work clearly demonstrates this fact. Therefore,
+any single estimate of parameters is insufficient for analysis. As such,
+identification of the \ac{BOLD} model that does not treat the parameters as distributions
+will not be able to overcome the inherent ill-posed nature of the
+parameters. Besides the Unscented Kalman Filter, this is the only approach
+that accomplishes this task without repeatedly calculating the parameters
+to build a distribution. The Unscented Kalman Filter though, is restricted
+to the Gaussian distribution which limits its ability to estimate the posterior.
+
+The primary reason this method has not been used before is the high
+dimensionality of the system. In Murray, 2008 the idea of learning
+the parameters is floated as the best method, yet
+learning 7 parameters with a Monte Carlo method was deemed intractable
+\cite{Murray2008}. The
+concern was that seven parameters creates a prohibitively large
+search space, which would require too many particles to learn properly. To overcome
+this difficulty, instead of starting the algorithm with 1000 particles,
+the initial number of particles was set to 16,000. This meant that
+when the search space was the largest, the number of particles was
+sufficiently dense to represent the prior distribution. The result is
+a particle filter algorithm that spends computing resources only where
+it is really needed.
+
+The particle filter is a Bayesian non-parametric algorithm that, given
+enough measurements will approach the true probability distribution
+of the parameters. The conclusion that the parameters are not
+uniquely identifiable is disappointing in light of the plethora
+of research papers that have attempted to learn them. However, the fact
+that the parameters are not identifiable further demonstrates
+the need to estimate a probability distribution rather than a single
+parameter estimate. At the same time, the output of the particle filter
+is still able to give good estimates of the \ac{BOLD} output, and is not
+dependent on heavy Gaussian smoothing. Therefore, all the current
+experimental paradigms to determine regional activity are still viable.
+To borrow a computer term, the particle filter algorithm is backward
+compatible with the current methods. Indeed, the particle filter excelled
+at determining activation in very low \ac{SNR} simulations
+(\autoref{sec:Multi-voxel Simulation}). Not only that, but the particle
+filter algorithm identified areas of activation that were completely
+missed by \ac{SPM} (\autoref{fig:comp6}).
+
+The physiological plausibility is also of great value to future
+research. Unlike traditional methods, use of the \ac{BOLD} model allows
+the incorporation of outside, physiological knowledge in regression.
+If, for instance, it becomes possible to measure $v$ in vivo that data
+could instantly be added to the model to improve regression. Data can
+also be retrieved from the model. \ac{BOLD} parameter distributions could
+for the first time be used as diagnosis tools. In the past the fact
+that the BOLD model is ill-posed would have prevented such research. 
+However, now that a full
+distribution is available, comparing parameter estimates is actually
+feasible.
+
+Concluding, the particle filter provides comparable performance with
+conventional tests, but also provides a platform for a wide range of
+future uses beyond what was possible with the methods that have been
+used before.
+
+\section{Future Work}
 \label{sec:FutureWork}
@@
 %%% "smart knots"
-%%% DC gain as a parameter
+%%% \ac{DC} gain as a parameter
 %%% linearizing
@@
 %%% Automatic estimation of measurement error.
-\section{Algorithm Improvements}
+\subsection{Algorithm Improvements}
 \label{sec:Particle Filter Variations}
@@
 distribution. The prior distribution used in this work is listed in \autoref{tab:Prior},
-and is based on the findings of Friston et al. \cite{Friston2000}. Unfortunately, that
+and is based on the findings of Friston et al. \cite{Friston2002b}. Unfortunately, that
 result depended on a quadratic approximation (Volterra Series) which has
@@
 This is in addition to the fact that the parameters are ill-defined, and
-thus the parameters found in Friston et al. are most likely far too wide.
+thus the parameters found in Friston et al. \cite{Friston2002b} are most likely far too wide.
 The prior is also
-based completely on the BOLD output, which is why it is good at generating
-estimates of the BOLD signal, but weak at estimating parameters.
+based completely on the \ac{BOLD} output, which is why it is good at generating
+estimates of the \ac{BOLD} signal, but weak at estimating parameters.
 The best solution would be to have
@@
 Another major area for improvement is removing drift.
-A large number of detrending methods were tested, but every method was in some way limited.
+Several variations of spline detrending were tested, but only
+the simple case spline described in \autoref{sec:Detrend}
+gave consistent results.
 One potentially viable method that could be utilized, given the right
@@
 constant to be added to the pre-processed signal for the
-BOLD model to fit properly. It would also minimize the actual
+\ac{BOLD} model to fit properly. It would also minimizes the actual
 removed. The disadvantage of this approach is
@@
 out. It also requires periodic breaks in the stimulus,
-which could reduce value of those samples for fitting purposes.
+which could limit persistent excitation.
 
 Linearizing is another possible method of dealing with drift.
-This would use the delta between measurements for fitting, rather than the
+This would use the slope between measurements for fitting, rather than the
 direct value. This has the advantage of not requiring detrending and thus
@@
 In a test with rapid impulse stimuli, this could be extremely effective because
-the DC signal has minimal data; conversely prolonged flat levels make this less effective.
+the \ac{DC} signal carries less information; conversely prolonged flat levels make this less effective.
 In tests I found that large drift/low white noise
@@
 this, finding a generic, consistent solution is complex, and often depends on the
-experimental design. Part of the problem is that SNR varies greatly across
+experimental design. Part of the problem is that \ac{SNR} varies greatly across
 the brain, and there is no way to actively measure noise without also knowing
@@
 
-\section{Experimental Design Improvements}
+\subsection{Experimental Design Improvements}
 \label{sec:Sideways Measurements}
 One definite way of improving the results of the particle filter is through
-additional measurements. While increasing the sample rate of FMRI scanners may
+additional measurements. While increasing the sample rate of \ac{fMRI} scanners may
 not be possible, simultaneous measurements of volume and flow is
 possible, albeit at 9T in a cat \cite{Hu2009}. Just one of those measurements
-would be extremely powerful when incorporated into the BOLD
+would be extremely powerful when incorporated into the \ac{BOLD}
 model. By adding another measurement, the
@@
 A more conventional method of adding measurements is to
-simply perform longer FMRI tests. Although this wouldn't be
-ground breaking, it would certainly increase the ability of the
+simply perform longer \ac{fMRI} tests. Although this wouldn't be
+groundbreaking, it would certainly increase the ability of the
 particle filter to develop the posterior distribution.
-It may also be possible to concatenate multiple FMRI tests together
-by using the parameter estimates from previous tests as the
-prior for later tests.
+It may also be possible to link \ac{fMRI} tests together
+by using the parameter estimates from previously acquired
+data as the prior for later experiments.
 
@@
 could make a large difference in the observability of parameters. The mentality
-for using a physiologically based nonlinear model for BOLD signal is to model
-those nonlinearities. Its logical then that certain nonlinear parameters may not
+for using a physiologically based nonlinear model for \ac{BOLD} signal is to model
+those nonlinearities. It is logical then that certain nonlinear parameters may not
 be identifiable when the input is primarily an impulse response. It
 has been reported that short responses are disproportionately large
-in FMRI data \cite{Miller2001, Deneux2006}. Therefore,
+in \ac{fMRI} data \cite{Miller2001, Deneux2006}. Therefore,
 a wide range of activation hold time may shed further light on
-the parameters' distribution as well as the validity of the BOLD model.
+the parameters' distribution as well as the validity of the \ac{BOLD} model.
 
@@
 %% Comparison of posterior of parameters
-\section{Future Applications}
+\subsection{Future Applications}
 There are a number of advantages to the particle filter approach presented
-here. In the past, FMRI data has been analyzed strictly for determining
+here. In the past, \ac{fMRI} data has been analyzed strictly for determining
 correlation between a stimulus and response. With this new method
@@
 correlated with neural pathologies. In spite of the fact that the parameters
-are under-determined, the final distribution is still a reduction in the
+are under-determined, the final distribution still reduces the
 uncertainty of the parameters. The availability of a full posterior distribution
@@
 populations. This would be especially useful in patients whose symptoms
-are caused by functional difference in the brain, rather than structural differences.
+are caused by functional differences in the brain, rather than structural differences.
 This would effectively be
@@
 Because of limitations present in every imaging modality,
-its becoming increasingly clear that combining data from multiple sources
-will be necessary to push Neurology forward. In order to do so however, the
+it is becoming increasingly clear that combining data from multiple sources
+will be necessary to push neurology forward. In order to do so however, the
 output of each source needs to fully represent what information that source
@@
 For instance, if two different modalities have calculated the probability
-distribution of neural efficiency, those two beliefs may be combined into
+distribution of neural efficiency, those two beliefs (distributions) may be 
+combined into
 one conditional belief for a new conditional probability of neural efficiency.
@@
 An advantageous aspect of using a physiological model such as this, is that
-it permits estimates of otherwise hidden parameters. In particular the BOLD
+it permits estimates of otherwise hidden parameters. In particular the \ac{BOLD}
 model gives an estimate of the value of the flow inducing signal, $s$. Having
@@
 In this way, it could be possible to determine chains of events. This is just
-one possible benefit being able to determine the time course of the hidden
-state variables present in the BOLD equations; the potential benefits of being
+one possible benefit of being able to determine the time course of the hidden
+state variables present in the \ac{BOLD} equations; the potential benefits of being
 able to determine this information is limitless. Of course the quality of these

### 2010-09-02 doc/thesis/methods.tex
just need to fix figure captions for multivoxel cases. And fix bib

@@
 \includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response}
-\caption{Response to $0.1s$ impulses with the mean parameters from 
-Friston et al. \cite{Friston2000b}}
+\caption{BOLD Response to $0.1$ s impulses with the mean parameters from 
+Friston et al. \cite{Friston2002b}}
 \label{fig:MeanResponseF}
@@
 \includegraphics[trim=6cm 2cm 6cm 2cm,width=15cm]{images/mean_response_johnston}
-\caption{Response to $0.1s$ impulses with the mean parameters from 
+\caption{BOLD Response to $0.1$ s impulses with the mean parameters from 
 Johnston et al. \cite{Johnston2007}}

### 2010-09-02 doc/thesis/results.tex
just need to fix figure captions for multivoxel cases. And fix bib

@@
 converging for a 2400 second \ac{fMRI} run. Darker bars indicate
-bins with more regions. }
+bins with more particles with BOLD estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge}
@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
-\caption{\ac{BOLD} estimate converging for a very long \ac{fMRI} run, first 500 seconds. Darker bars
-indicate bins with more particles.}
+\caption[First 500 seconds of the \ac{BOLD} estimate converging for a very 
+long \ac{fMRI} run.]{First 500 seconds of \ac{BOLD} estimate 
+converging for a 2400 second \ac{fMRI} run. Darker bars indicate
+bins with more particles with BOLD estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge_500}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_lownoise}
-\caption{Test Signals with low noise compared to the clean signal.}
+\caption[Simulated Signal with Low Noise/Drift]
+{Simulated signals with different noise realizations compared to true
+signal. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:LowNoiseRealization}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_lownoise}
-\caption{A comparison of the preprocessed signals for the low noise case. The
-noisy input to the actual particle filter algorithm.}
+\caption[Preprocessed Signal with Low Noise/Drift]
+{Results of preprocessing compared to the true signal. 
+ Knots of spline placed every 20 measurements.
+ ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:PreprocessedLowNoise}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_lownoise}
-\caption{A comparison of the fitted signals for the low noise case.}
+\caption[Results with Low Noise/Drift]
+{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedLowNoise}
+as input.}
 \label{fig:FitComparisonLowNoise}
@@
 \hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \ac{RMSR} & \ac{RMSE} \\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \acs{RMSR} & \acs{RMSE} \\
 \hline
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, and last is the average.}
+\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
+last is the average. $\sum \tau$ is the sum of the estimated time constants.}
 \label{tab:LowNoiseResults}
@@
 \begin{figure}
-\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise1}}\\
-
-\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $v$ of the first run, low noise simulation.]
-{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
+\subfigure
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/converge_lownoise1}}
+\subfigure
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/converge_lownoise2}}
 \end{figure}
@@
 \begin{figure}
-\subfigure[Converging histogram for $q$, $s$, $f$, and $BOLD$ of the first run, low noise simulation.]
-{\label{fig:LowNoiseHistc} \includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
-\caption{Convergence of the first run from \autoref{tab:LowNoiseResults}. The bars represent
-a histogram, where darker bars indicate more particles in that bin.}
+\subfigure
+{\label{fig:LowNoiseHistc} \includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/converge_lownoise3}}
+\caption[Convergence of the parameters from the first run of \autoref{tab:LowNoiseResults}.]
+{Convergence of the parameters from the first run of \autoref{tab:LowNoiseResults}. 
+Order of estimates: $\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f, \epsilon, v,
+q, s, f, BOLD$.  The bars represent
+a histogram, where darker bars indicate more particles with parameters in that bin. The red 
+line is the parameter used to generate the true signal, blue line is the final mean of the
+particles.}
 \label{fig:LowNoiseHist}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_highnoise}
-\caption{A comparison of the preprocessed signals for the high noise case.}
+\caption[Preprocessed Signal with High Noise/Drift]
+{Results of preprocessing compared to the true signal.  
+ Knots of spline placed every 20 measurements. ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
 \label{fig:PreprocessedHighNoise}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise}
-\caption{A comparison of the fitted signals for the high noise case.}
+\caption[Results with High Noise/Drift]
+{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedHighNoise}
+as input.}
 \label{fig:FitComparisonHighNoise}
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
+\caption[Estimated Parameters on 11 different runs with high noise.]
+{Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
 last row contains the mean.
@@
 \begin{figure}[H]
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_1}}\\
+\subfigure
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run5_1}}
+\subfigure
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run5_2}}
 \end{figure}
@@
 \begin{figure}[H]
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 1]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
-\end{figure}
-
-\begin{figure}[H]
-\subfigure[$q$, $s$, $f$, $BOLD$, Run 1 ]
-{\label{fig:ConvergenceRuns1c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
-\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\subfigure
+{\label{fig:ConvergenceRuns1c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run5_3}}
+\caption[Convergence of the parameters during run 1 of \autoref{fig:NoiseComparisonJustTwo}]
+{Convergence of the parameters during run 1 of \autoref{fig:NoiseComparisonJustTwo}.
+Order of estimates: $\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f, \epsilon, v,
+q, s, f, BOLD$.  The bars represent
+a histogram, where darker bars indicate more particles with parameters in that bin. The red 
+line is the parameter used to generate the true signal, blue line is the final mean of the
+particles.}
 \label{fig:ConvergenceRuns1}
@@
 \begin{figure}[H]
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_1}}\\
+\subfigure
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run6_1}}
+\subfigure
+{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run6_2}}
 \end{figure}
@@
 \begin{figure}[H]
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$, Run 2]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_2}}\\
-\end{figure}
-
-\begin{figure}[H]
-\subfigure[$q$, $s$, $f$, $BOLD$, Run 2 ]
-{\label{fig:ConvergenceRuns2c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
-\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
+\subfigure
+{\label{fig:ConvergenceRuns2c} \includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=\textwidth]{images/highnoise_run6_3}}
+\caption[Convergence of the parameters during run 2 of \autoref{fig:NoiseComparisonJustTwo}]
+{Convergence of the parameters during run 2 of \autoref{fig:NoiseComparisonJustTwo}.
+Order of estimates: $\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f, \epsilon, v,
+q, s, f, BOLD$.  The bars represent
+a histogram, where darker bars indicate more particles with parameters in that bin. The red 
+line is the parameter used to generate the true signal, blue line is the final mean of the
+particles.}
 \label{fig:ConvergenceRuns2}
@@
 \includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=15cm]{images/preprocessed_noiseonly}
-\caption{Comparison of the preprocessed signals for the low noise signal-free case (
-$\sigma_y = 0.01, \sigma_x = 0.005$).}
+\caption[Preprocessed Signal for non-active, low noise signal]
+{Comparison of the preprocessed signals for the low noise signal-free case.
+ ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
 \label{fig:PreprocessedNoiseOnly}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
+\caption[Results for non-active, low noise signal]  ($\sigma_x = 0.005$, $\sigma_y = 0.01$)
+{Comparison of the estimated BOLD signal for the low noise signal-free case.
+ ($\sigma_x = 0.005$, $\sigma_y = 0.01$). Note the line thickness is caused
+by all the estimates overlapping.}
+\label{fig:fits_noiseonly}
 \end{figure}
@@
 \hline
-$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & \ac{RMSR} \\
+$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & \acs{RMSR} \\
 \hline
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with low noise and no signal present.}
+\caption{Estimated Parameters on 11 different runs with low noise and no signal present.
+ ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
 \label{tab:NoiseOnlyResults}
@@
 \centering
-\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
-\caption{Fits to the non-active, low noise signal. Note that the line is thick because all
-the estimates overlap. All 11 fitted lines.}
-\label{fig:fits_noiseonly}
-\end{figure}
-
-\begin{figure}[H]
-\centering
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justnoise_fit_0}
-\caption{Fit from a single particle filter run, with the noise input. }
+\caption{Single Fit Results for non-active, low noise signal.}
 \label{fig:justnoise_fit_0}
@@
 \centering
-\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_1}}\\
+\subfigure
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/justnoise_hist_1}}
+\subfigure
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/justnoise_hist_2}}
 \end{figure}
@@
 \centering
-\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $v$]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_2}}\\
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\subfigure[$q$, $s$, $f$, $BOLD$ ]
-{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_3}}
-\caption{Converging histogram for parameters when the signal consists purely of low level noise.
-Same run as \autoref{fig:justnoise_fit_0}}
+\subfigure
+{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/justnoise_hist_3}}
+\caption[Convergence of the parameters during a noise-only run.]
+{Convergence of the parameters when the signal consisted purely of low level noise 
+($\sigma_x = 0.01$, $\sigma_y = 0.005$).
+Order of estimates: $\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f, \epsilon, v,
+q, s, f, BOLD$.  The bars represent
+a histogram, where darker bars indicate more particles with parameters in that bin. The red 
+line is the parameter used to generate the true signal, blue line is the final mean of the
+particles. See \autoref{fig:justnoise_fit_0}}
 \label{fig:JustNoiseConvergence}
@@
 well within the range of $0.005$, the standard deviation of the weighting function.
-Note that the \acs{RMSR}
-were actually lower than the \acs{RMSR} in the low noise simulation 
+Note that the \acp{RMSR}
+were actually lower than the \acp{RMSR} in the low noise simulation 
 from \autoref{sec:SimLowNoise}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3cm,width=15cm]{images/fits_noiseonly_high}
-\caption{\ac{BOLD} estimates for the non-active, high noise signal. Note the line thickness is caused
-by all the estimates overlapping.}
+\caption[Results for non-active, high noise signal.]
+{\ac{BOLD} estimates for the non-active, high noise signal ($\sigma_x = 0.1$, $\sigma_y = 0.05$). 
+Note the line thickness is caused by all the estimates overlapping.}
 \label{fig:fits_noiseonly_high}
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justbignoise_fit_0}
-\caption{Fit from a single particle filter run, with the noise input. }
+\caption{Single Fit Results for non-active, high noise signal.}
 \label{fig:justbignoise_fit_0}
@@
 
-\subsection{Single Voxel Review}
+\subsection{Single Voxel Summary}
 \label{sec:SingleVoxelReview}
@@
 \end{tabular}
-\caption{Mutual Information and the normalized \ac{RMSE},
-    for signal/noise configurations.}
+\caption{Mutual Information and the normalized \ac{RMSE}, for each of the previous
+sections.}
 \label{tab:SingleVoxelActivationComparison}
@@
 The thresholds applied to this slice, both for mutual information and
-\acs{RMSR} are arbitrary. Applying a
+\acp{RMSR} are arbitrary. Applying a
 threshold is helpful for visualization but is rarely useful for further analysis.

### 2010-09-02 doc/thesis/results.tex
finished I think

@@
 The slice simulation was performed using
-Physics-Oriented Simulated Scanner for Understanding MRI (POSSUM) which
-models noise realistically. POSSUM demonstrates the
+\ac{POSSUM} which
+models noise realistically. \ac{POSSUM} demonstrates the
 particle filter modeling the \ac{BOLD} signal on a large scale.
-In the POSSUM simulation there were four
+In the \ac{POSSUM} simulation there were four
 discrete parameter sets driving the time series, although each voxel had a
 different tissue composition and a different
-noise realization. Therefore the POSSUM simulation was
+noise realization. Therefore the \ac{POSSUM} simulation was
 a good test of the applicability of particle filters for performing
@@
 converging for a 2400 second \ac{fMRI} run. Darker bars indicate
-bins with more particles with BOLD estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
+bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge}
@@
 converging for a 2400 second \ac{fMRI} run. Darker bars indicate
-bins with more particles with BOLD estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
+bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge_500}
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
-\caption[Results for non-active, low noise signal]  ($\sigma_x = 0.005$, $\sigma_y = 0.01$)
-{Comparison of the estimated BOLD signal for the low noise signal-free case.
+\caption[Results for non-active, low noise signal]  
+{Comparison of the estimated \ac{BOLD} signal for the low noise signal-free case.
  ($\sigma_x = 0.005$, $\sigma_y = 0.01$). Note the line thickness is caused
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justnoise_fit_0}
-\caption{Single Fit Results for non-active, low noise signal.}
+\caption[Single Fit Results for non-active, low noise signal.]
+{Single Fit Results for non-active, low noise signal
+ ($\sigma_x = 0.005$, $\sigma_y = 0.01$).}
 \label{fig:justnoise_fit_0}
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justbignoise_fit_0}
-\caption{Single Fit Results for non-active, high noise signal.}
+\caption[Single Fit Results for non-active, high noise signal.]
+{Single Fit Results for non-active, high noise signal
+ ($\sigma_x = 0.005$, $\sigma_y = 0.01$).}
 \label{fig:justbignoise_fit_0}
@@
 Considering the tendency of \ac{fMRI} to have large unexplainable
-peaks and troughs, the Median-Absolute-Deviation (MAD) was used (\autoref{eq:mad}).
+peaks and troughs, the \ac{MAD} was used (\autoref{eq:mad}).
 This is an estimator of the standard
@@
 \subfigure[Region labels for simulated slice.]{
-\includegraphics[height=8cm]{images/simregions.png} \label{fig:simslice_mask} }
+\includegraphics[height=9cm]{images/simregions.png} \label{fig:simslice_mask} }
+\subfigure[\ac{SNR} Map of \ac{POSSUM} simulated data. Mean \ac{SNR} per region, 
+Region 1: $0.8$, Region 2: $0.97$, and
+Region 3: $0.39$.]{\includegraphics[height=9cm]{images/snr_hm.png}\label{fig:SimSNRhm}}
+
 \subfigure[Normalized \ac{RMSR} Map. Scale is Normalized \ac{RMSE}. Lower (yellow) is better.]
-{\label{fig:simslice_hm_res} \includegraphics[height=8cm]{images/sim_hm}}
-\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
-{\label{fig:sim_hm_mi} \includegraphics[height=8cm]{images/sim_hm_mi}}
-\subfigure[SNR Map of POSSUM simulated data. Mean SNR per region, Region 1: $0.8$, Region 2: $0.97$, and
-Region 3: $0.39$.]{\includegraphics[height=8cm]{images/snr_hm.png}\label{fig:SimSNRhm}}
-\caption{Comparison of activation with greymatter, parameter regions.}
+{\label{fig:simslice_hm_res} \includegraphics[height=8.5cm]{images/sim_hm}}
+\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) indicates better fit.]
+{\label{fig:sim_hm_mi} \includegraphics[height=8.5cm]{images/sim_hm_mi}}
+
+\caption{Comparison of activation regions, a) grey matter and regions, b) \ac{SNR} of active
+regions, c) \acl{RMSR} Map, d) Mutual Information Map.}
 \label{fig:simslice_hm}
@@
 To test the usefulness of the particle filter on a larger scale, a modified version
-of the FSL tool
-POSSUM was used to generate an entire \ac{fMRI} image from a parameter map. The parameter map was generated
+of the \ac{FSL} tool
+\ac{POSSUM} was used to generate an entire \ac{fMRI} image from a parameter map. The parameter map was generated
 by taking an existing activation map and assigning discrete parameter sets to each region.
@@
 to take a parameter map and generate activation levels depending on the parameters at that
-point. The patch for POSSUM will be made available. Note that the noise
-level was set to an SNR of 20, but due to changes
+point. The patch for \ac{POSSUM} will be made available. Note that the noise
+level was set to an \ac{SNR} of 20, but due to changes
 in the program the true signal-to-noise ratio was much lower, as seen in the
-in \autoref{fig:SimSNRhm}. The mean SNR for each region was calculated only for the
+in \autoref{fig:SimSNRhm}. The mean \ac{SNR} for each region was calculated only for the
 voxels with a Signal-To-Noise ratio above $0.1$.
@@
 \centering
-\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist1}
-\caption{Histogram of estimated parameters in section 1 in voxels 
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist1}
+\caption[Histogram of estimated parameters in section 1]{Histogram of estimated parameters in 
+section 1 in voxels 
 with mutual information greater than $0.15$} \label{fig:slicesim_hist1}
@@
 \centering
-\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
-\caption{Histogram of estimated parameters in section 2 in voxels with mutual information greater
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist2}
+\caption[Histogram of estimated parameters in section 2]
+{Histogram of estimated parameters in section 2 in voxels with mutual information greater
 than $0.15$}
@@
 \centering
-\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
-\caption{Histogram of estimated parameters in section 3 in voxels with mutual information greater
+\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist3}
+\caption[Histogram of estimated parameters in section 3]
+{Histogram of estimated parameters in section 3 in voxels with mutual information greater
 than $0.15$}
@@
 \autoref{fig:sim_hm_mi}; this is
-not actually the case. POSSUM simulates different tissues, and white matter
+not actually the case. \ac{POSSUM} simulates different tissues, and white matter
 does not typically  have a \ac{BOLD} response. This explains some of the 

### 2010-09-02 doc/thesis/resultsB.tex
finished I think

@@
 \centering
-\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
-\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
-\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
-\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
+\subfigure[Axial Slices]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
+\subfigure[Sagittal Slice]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
+\subfigure[Coronal Slice]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
+\subfigure[Axial Slice]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
 \subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
-\caption{\ac{SPM} results. Units of activation are in Student's $t$-scores; higher indicates higher
-        assurance that the signal cannot have occurred through noise alone.
-        Sagittal, coronal and axial slices  are
-        \autoref{fig:hm_canon_spm_x}, \autoref{fig:hm_canon_spm_y}, and
-         \autoref{fig:hm_canon_spm_z}, respectively. A series of axial slices are
-         shown in \autoref{fig:hm_spm}. }
+\caption[\ac{SPM} Results]{\ac{SPM} results. 
+        Units of activation are in Student's $t$-scores; higher indicates higher
+        assurance that the signal cannot have occurred through noise alone.}
 \label{fig:hm_canon_spm}
@@
 \centering
-\subfigure[]{\label{fig:hm_pfilter} \includegraphics[scale=.85]{images/pfilter_hm}}
-\subfigure[]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=3]{images/pfilter_hm_x}}
-\subfigure[]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=3]{images/pfilter_hm_y}}
-\subfigure[]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=3]{images/pfilter_hm_z}}
+\subfigure[Axial Slices]{\label{fig:hm_pfilter} \includegraphics[scale=.85]{images/pfilter_hm}}
+\subfigure[Sagittal Slice]{\label{fig:hm_canon_pfilter_x} \includegraphics[scale=3]{images/pfilter_hm_x}}
+\subfigure[Coronal Slice]{\label{fig:hm_canon_pfilter_y} \includegraphics[scale=3]{images/pfilter_hm_y}}
+\subfigure[Axial Slice]{\label{fig:hm_canon_pfilter_z} \includegraphics[scale=3]{images/pfilter_hm_z}}
 \subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
-\caption{Particle Filter results measured in normalized error.
+\caption[Particle Filter Normalized Residual Results]{Particle Filter results measured in normalized error.
         Units of match is normalized residual where the lowest (best) levels shown are
         $0.7$ and the highest error shown (threshold) is $1$.
-        Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_pfilter_x},
-        \autoref{fig:hm_canon_pfilter_y}, and
-         \autoref{fig:hm_canon_pfilter_z}, respectively. A series of axial slices are shown in
-         \autoref{fig:hm_pfilter}.  }
+        Sagittal, coronal and axial slices are shwon in \autoref{fig:hm_canon_pfilter_x}.}
 \label{fig:hm_canon_pfilter}
@@
 \centering
-\subfigure[]{\label{fig:hm_mi} \includegraphics[scale=.85]{images/mi_hm}}
-\subfigure[]{\label{fig:hm_canon_mi_x} \includegraphics[scale=3]{images/mi_hm_x}}
-\subfigure[]{\label{fig:hm_canon_mi_y} \includegraphics[scale=3]{images/mi_hm_y}}
-\subfigure[]{\label{fig:hm_canon_mi_z} \includegraphics[scale=3]{images/mi_hm_z}}
+\subfigure[Axial Slices]{\label{fig:hm_mi} \includegraphics[scale=.85]{images/mi_hm}}
+\subfigure[Sagittal Slice]{\label{fig:hm_canon_mi_x} \includegraphics[scale=3]{images/mi_hm_x}}
+\subfigure[Coronal Slice]{\label{fig:hm_canon_mi_y} \includegraphics[scale=3]{images/mi_hm_y}}
+\subfigure[Axial Slice]{\label{fig:hm_canon_mi_z} \includegraphics[scale=3]{images/mi_hm_z}}
 \subfigure{\label{fig:scale_mi} \includegraphics[scale=.5]{images/scale4}}
-\caption{Particle Filter results measured in mutual information.
+\caption[Particle Filter Mutual Information Results]
+        {Particle Filter results measured in mutual information.
          Units of match is bits (standard for base-2 Mutual Information). The highest (best) levels are
-         $0.42$ and the worst shown (threshold) is $0.1$.
-        Sagittal, coronal and axial slices are shown in \autoref{fig:hm_canon_mi_x},
-        \autoref{fig:hm_canon_mi_y}, and \autoref{fig:hm_canon_mi_z}, respectively.
-        A series of axial slices are shown in \autoref{fig:hm_mi}.  }
+         $0.42$ and the worst shown (threshold) is $0.1$.}
 \label{fig:hm_canon_pfilter_mi}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/1_pfilter_37_14_7}}\\
-\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/1_spm_37_14_7}}
-\caption{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
+\subfigure[Particle Filter]{\label{fig:comp1pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/1_pfilter_37_14_7}}
+\subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/1_spm_37_14_7}}
+\caption[Time Series Comparison, Section 1]{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/2_pfilter_34_12_7}}\\
-\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14.5cm]{images/2_spm_34_12_7}}
-\caption{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
+\subfigure[Particle Filter]{\label{fig:comp2pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/2_pfilter_34_12_7}}
+\subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/2_spm_34_12_7}}
+\caption[Time Series Comparison, Section 2]{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/3_pfilter_23_21_7}}\\
-\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/3_spm_23_21_7}}
-\caption{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
+\subfigure[Particle Filter]{\label{fig:comp3pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/3_pfilter_23_21_7}}
+\subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/3_spm_23_21_7}}
+\caption[Time Series Comparison, Section 3]{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/4_pfilter_26_15_7}}\\
-\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/4_spm_26_15_7}}
-\caption{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
+\subfigure[Particle Filter]{\label{fig:comp4pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/4_pfilter_26_15_7}}
+\subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/4_spm_26_15_7}}
+\caption[Time Series Comparison, Section 4]{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_pfilter_29_9_13}}\\
-\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_29_9_13}}
-\caption{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
+\subfigure[Particle Filter]{\label{fig:comp5pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_pfilter_29_9_13}}\\
+\subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_29_9_13}}
+\caption[Time Series Comparison, Section 5]{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
 %\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp6pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/6_pfilter_36_17_19}}\\
-\subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/6_spm_36_17_19}}
-\caption{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
+\subfigure[Particle Filter]{\label{fig:comp6pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/6_pfilter_36_17_19}}
+\subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/6_spm_36_17_19}}
+\caption[Time Series Comparison, Section 6]{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 \centering
-\subfigure[Particle Filter]{\label{fig:comp7pfilter} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/7_pfilter_29_26_1}}\\
-%\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 1cm 4cm 1cm,width=14cm]{images/5_spm_25_34_25}}
-\caption{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
+\subfigure[Particle Filter]{\label{fig:comp7pfilter} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/7_pfilter_29_26_1}}
+%\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_25_34_25}}
+\caption[Time Series Comparison, Section 7]{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_00}}
-\caption{$\tau_0$ Estimates}
+\caption{Regional $\tau_0$ Estimates}
 \label{fig:pmap0}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_01}}
-\caption{$\alpha$ Estimates}
+\caption{Regional $\alpha$ Estimates}
 \label{fig:pmap1}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_02}}
-\caption{$E_0$ Estimates}
+\caption{Regional $E_0$ Estimates}
 \label{fig:pmap2}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_03}}
-\caption{$V_0$ Estimates}
+\caption{Regional $V_0$ Estimates}
 \label{fig:pmap3}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_04}}
-\caption{$\tau_f$ Estimates}
+\caption{Regional $\tau_f$ Estimates}
 \label{fig:pmap4}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_05}}
-\caption{$\tau_s$ Estimates}
+\caption{Regional $\tau_s$ Estimates}
 \label{fig:pmap5}
@@
 \subfigure{\includegraphics[scale=.5]{images/pscale_06}}
-\caption{$\epsilon$ Estimates}
+\caption{Regional $\epsilon$ Estimates}
 \label{fig:pmap6}

### 2010-09-02 doc/thesis/background.tex
fixed all the acronyms, think I might be done

@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_point1sec_3sigma.pdf}
-\caption[Q-Q Plot Of Standard Normal and Result of $0.1$ s BOLD Simulation Output]
-{Q-Q Plot Of Standard Normal and Result of $0.1$ s BOLD Simulation Output. A
+\caption[Q-Q Plot Of Standard Normal and Result of $0.1$ s \ac{BOLD} Simulation Output]
+{Q-Q Plot Of Standard Normal and Result of $0.1$ s \ac{BOLD} Simulation Output. A
 perfect Gaussian would fall on the green line.}
@@
 \includegraphics[trim=6cm .75cm 6cm .75cm,width=16cm]{images/gauss_step_1sec_3sigma.pdf}
-\caption[Q-Q Plot Of Standard Normal and Result of $1$ s BOLD Simulation Output]
-{Q-Q Plot Of Standard Normal and Result of $1$ s BOLD Simulation Output. A
+\caption[Q-Q Plot Of Standard Normal and Result of $1$ s \ac{BOLD} Simulation Output]
+{Q-Q Plot Of Standard Normal and Result of $1$ s \ac{BOLD} Simulation Output. A
 perfect Gaussian would fall on the green line.}

### 2010-09-02 doc/thesis/intro.tex
fixed all the acronyms, think I might be done

@@
 \includegraphics[width=.5\textwidth]{images/bold}
-\caption[Diagram of BOLD Signal]{A) End of Pre-Stimulus Undershoot, $< 2s$ typical,
+\caption[Diagram of \ac{BOLD} Signal]{A) End of Pre-Stimulus Undershoot, $< 2s$ typical,
 B) End of Stimulus, C) Input Bloodflow Drops from Supernormal to Subnormal,

### 2010-09-02 doc/thesis/methods.tex
fixed all the acronyms, think I might be done

@@
 
-Problematically, after removing the DC component of the signal,
+Problematically, after removing the \ac{DC} component of the signal,
 by definition the signal will have a median near zero.
@@
 the resting state will be below 0\%.
-One method of accounting for this is to simply add a DC gain model parameter.
+One method of accounting for this is to simply add a \ac{DC} gain model parameter.
 Like all the other model parameters, with enough measurements, the particle filter
@@
 by visual inspection.  In this work a simpler approach was used. To determine
-the DC gain a robust estimator of scale was used. The 
+the \ac{DC} gain a robust estimator of scale was used. The 
 \ac{MAD} proved to be accurate in determining how much to shift the signal up
@@
 concern; however, a boosted response does not affect how well the \ac{BOLD} model
-predicts the actual measurements; and as mentioned before, the DC signal 
+predicts the actual measurements; and as mentioned before, the \ac{DC} signal 
 of \ac{fMRI} is never used.

### 2010-09-02 doc/thesis/particlefilters.tex
just need to fix figure captions for multivoxel cases. And fix bib
fixed all the acronyms, think I might be done

@@
 preferred when the model is nonlinear, and non-gaussian. This section
-is largely based on Arulampalam et al. and Thrun et al. \cite{Arulampalam2002a, Thrun2005}.
+is largely based on Arulampalam et al. and Thrun et al. \cite{Arulampalam2002a, 
+thrun2008probabilistic}.
 
@@
 by a time varying function, least of all the brain. While heart beats are certainly
-periodic and have an effect on the BOLD signal, the period varies too much
+periodic and have an effect on the \ac{BOLD} signal, the period varies too much
 for the system to be considered varying with time. 
 Next, it was assumed that input cannot directly
-influence the output, which in the case of the BOLD signal is a good assumption.
+influence the output, which in the case of the \ac{BOLD} signal is a good assumption.
 Finally, because the only difference between the members of $X(t)$ and 
@@
 changes in $x$, however in this work it is a good assumption because the
-input does not immediately effect the BOLD signal. 
+input does not immediately effect the \ac{BOLD} signal. 
 \begin{equation}
@@
 an ideal calculation of the effective number of particles is found based on the 
-particles' true weight \cite{Liu1998b}. However, given that only an approximation 
+particles' true weight \cite{Liu1998}. However, given that only an approximation 
 for the true weight 
@@
 and $\epsilon$ is drawn from the chosen kernel. The choice of a kernel is
-complex, although it has been proven that the optimal kernel for reducing MSE between
+complex, although it has been proven that the optimal kernel for reducing \ac{MSE} between
 the original and resampled distributions is 
@@
 under-smoothing could leave gaps in the distribution.
-Moreover, because of the high dimensionality of the BOLD model,
+Moreover, because of the high dimensionality of the \ac{BOLD} model,
 and limited measurements, it is helpful to have a broader bandwidth to explore the distribution.
@@
 %\section{Piecewise Example}
-%A half wave rectifier takes a AC voltage circuit and removes
+%A half wave rectifier takes a \ac{AC} voltage circuit and removes
 %the negative half  of the signal. The resulting waveform

### 2010-09-02 doc/thesis/results.tex
fixed all the acronyms, think I might be done

@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
-\caption[\ac{BOLD} estimate converging for a very long \ac{fMRI} run.]{\ac{BOLD} estimate 
-converging for a 2400 second \ac{fMRI} run. Darker bars indicate
-bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
+\caption[\acs{BOLD} estimate converging for a very long \acs{fMRI} run.]{\acs{BOLD} estimate 
+converging for a 2400 second \acs{fMRI} run. Darker bars indicate
+bins with more particles with \acs{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge}
@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
-\caption[First 500 seconds of the \ac{BOLD} estimate converging for a very 
-long \ac{fMRI} run.]{First 500 seconds of \ac{BOLD} estimate 
-converging for a 2400 second \ac{fMRI} run. Darker bars indicate
-bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
+\caption[First 500 seconds of the \acs{BOLD} estimate converging for a very 
+long \acs{fMRI} run.]{First 500 seconds of \acs{BOLD} estimate 
+converging for a 2400 second \acs{fMRI} run. Darker bars indicate
+bins with more particles with \acs{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
 \label{fig:long_converge_500}
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
+\caption[Estimated Parameters with low noise.]
+{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
 last is the average. $\sum \tau$ is the sum of the estimated time constants.}
@@
 \end{tabular}
-\caption{Estimated Parameters on 11 different runs with low noise and no signal present.
+\caption[Estimated Parameters with low noies, and no signal.]
+{Estimated Parameters on 11 different runs with low noise and no signal present.
  ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
@@
 shown in \autoref{tab:SingleVoxelActivationComparison}. A second potential method of
-gauging performance is mutual information. Mutual information is a method of measuring
+gauging performance is \ac{MI}. \acl{MI} is a method of measuring
 the interdependence of two random variables. If two signals are truly independent,
-then the mutual information will be zero. Although ideally suited to discrete
+then the \ac{MI} will be zero. Although ideally suited to discrete
 distributions, by using histograms
 it is possible to derive a joint distribution of two signals. The algorithm for
-mutual information is based on that joint distribution:
+\ac{MI} is based on that joint distribution:
 \begin{equation}
@@
 where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
-the mutual information estimates in this work 6 bins were used for the marginal
+the \ac{MI} estimates in this work 6 bins were used for the marginal
 distribution of each signal. Additionally, throughout log base 2 will be used.
@@
 \end{equation}
- Note that subtracting the bias can result in negative mutual
-information, which should not technically be possible; so any negative mutual information
+ Note that subtracting the bias can result in negative 
+\ac{MI}, which should not technically be possible; so any negative \acl{MI}
 was taken as 0.
@@
 \hline
-& $M.I.$ & N. Res. &
-  $M.I.$ & N. Res. &
-  $M.I.$ & N. Res. &
-  $M.I.$ & N. Res. \\
+& $MI$ & N. Res. &
+  $MI$ & N. Res. &
+  $MI$ & N. Res. &
+  $MI$ & N. Res. \\
 \hline
@@
 \end{tabular}
-\caption{Mutual Information and the normalized \ac{RMSE}, for each of the previous
-sections.}
+\caption{\acs{MI} and the normalized \acs{RMSE}, for each of the previous sections.}
 \label{tab:SingleVoxelActivationComparison}
@@
 distinguishing between these cases with either normalized \ac{RMSR} or 
-mutual information is not clear cut. While the average mutual information 
+\ac{MI} is not clear cut. While the average \ac{MI}
 is more than 10 times
-the average mutual information in the two non-signal cases, the maximum
-mutual information of the low noise/no signal case exceeds the minimum M.I. 
+the average \ac{MI} in the two non-signal cases, the maximum
+\ac{MI} of the low noise/no signal case exceeds the minimum \ac{MI}
 of the high noise/signal
 case. The Low Noise/signal determination is easier to make; given
-the minimum mutual information is above $.8$ and the maximum normalized 
+the minimum \ac{MI} is above $.8$ and the maximum normalized 
 \ac{RMSR} is below $.6$.
-However, it is worth noting that the worst case scenario  for mutual information (maximum) in the
+However, it is worth noting that the worst case scenario  for \ac{MI} (maximum) in the
 low noise/no signal case does not coincide with the worse case (minimum) normalized residual.
@@
 a particular voxel were active or inactive; the accuracy would improve if both techniques
-were used with loose restrictions. In all, both M.I. and the normalized residual
+were used with loose restrictions. In all, both \ac{MI} and the normalized residual
 provide a good measure of performance.
@@
 accurate even when noise was drastically increased. The second purpose was to validate
-Mutual Information and the normalized \ac{RMSR} as measurements of output quality.
+\ac{MI} and the normalized \ac{RMSR} as measurements of output quality.
 Although not perfect, both methods do provide a decent metric.
@@
 {\label{fig:simslice_hm_res} \includegraphics[height=8.5cm]{images/sim_hm}}
-\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) indicates better fit.]
+\subfigure[\ac{MI} Map. Scale is bits. Higher (yellow) indicates better fit.]
 {\label{fig:sim_hm_mi} \includegraphics[height=8.5cm]{images/sim_hm_mi}}
 
-\caption{Comparison of activation regions, a) grey matter and regions, b) \ac{SNR} of active
-regions, c) \acl{RMSR} Map, d) Mutual Information Map.}
+\caption[Comparison of activation regions.]
+{Comparison of activation regions, a) grey matter and regions, b) \acs{SNR} of active
+regions, c) \acs{RMSR} Map, d) \acs{MI} Map.}
 \label{fig:simslice_hm}
@@
 Notice that the
-regions 1, 2 an 3 stick out in both the \ac{RMSR} and the mutual information
+regions 1, 2 an 3 stick out in both the \ac{RMSR} and the \ac{MI}
 map, indicating that the particle filter was successful in matching those regions.
-Mutual information was an extremely successful metric, with the exception of
+\ac{MI} was an extremely successful metric, with the exception of
 a few false positives (\autoref{fig:simslice_hm_res}). 
@@
 \subfigure{\includegraphics[width=1cm]{images/scale6.png}}
-\caption{More stringent mutual information heatmap. Higher (yellow) is better.}
+\caption{More stringent \acs{MI} heatmap. Higher (yellow) is better.}
 \label{fig:sim_hm2}
@@
 section 1 in voxels 
-with mutual information greater than $0.15$} \label{fig:slicesim_hist1}
+with \acs{MI} greater than $0.15$} \label{fig:slicesim_hist1}
 \end{figure}
@@
 \caption[Histogram of estimated parameters in section 2]
-{Histogram of estimated parameters in section 2 in voxels with mutual information greater
+{Histogram of estimated parameters in section 2 in voxels with \acs{MI} greater
 than $0.15$}
@@
 \caption[Histogram of estimated parameters in section 3]
-{Histogram of estimated parameters in section 3 in voxels with mutual information greater
+{Histogram of estimated parameters in section 3 in voxels with \acs{MI} greater
 than $0.15$}
@@
 If the thresholds are left at the values of \autoref{fig:simslice_hm}, then
-the normalized \ac{RMSR} gave $2/1479$  false positives and mutual information gave
+the normalized \ac{RMSR} gave $2/1479$  false positives and \ac{MI} gave
 $11/1479$. However, normalized \ac{RMSR} also clearly missed some active voxels, although
@@
 Raising the threshold to $0.15$ eliminated the false positives in the
-Mutual Information map, although some active voxels were also removed.
+\ac{MI} map, although some active voxels were also removed.
 
@@
 of the particle filter at identifying regions of activation.
-The thresholds applied to this slice, both for mutual information and
+The thresholds applied to this slice, both for \ac{MI} and
 \acp{RMSR} are arbitrary. Applying a
@@
 As noted in \autoref{sec:SingleVoxelReview},
-the false positives present in the Mutual Information
+the false positives present in the \ac{MI}
 map are different from those in the \ac{RMSR} map. This furthers the argument

### 2010-09-02 doc/thesis/resultsB.tex
fixed all the acronyms, think I might be done

@@
 filter data shows a large number of false positives, however application of a threshold
-of $0.85$ on the residual map removes these false positives. Similarly, in the mutual information
+of $0.85$ on the residual map removes these false positives. Similarly, in the \acl{MI}
 map, the false positives may be eliminated by upping the threshold to $0.15$. However, just
@@
 \subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
-\caption[\ac{SPM} Results]{\ac{SPM} results. 
+\caption[\acs{SPM} Results]{\acs{SPM} results. 
         Units of activation are in Student's $t$-scores; higher indicates higher
@@
 \subfigure{\label{fig:scale_mi} \includegraphics[scale=.5]{images/scale4}}
-\caption[Particle Filter Mutual Information Results]
-        {Particle Filter results measured in mutual information.
-         Units of match is bits (standard for base-2 Mutual Information). The highest (best) levels are
+\caption[Particle Filter \acs{MI} Results]
+        {Particle Filter results measured in \ac{MI}.
+         Units of match is bits (standard for base-2 \ac{MI}). The highest (best) levels are
          $0.42$ and the worst shown (threshold) is $0.1$.}
@@
 \subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/1_spm_37_14_7}}
-\caption[Time Series Comparison, Section 1]{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, Mutual Information: $0.33$, Residual: $0.72$.}
+\caption[Time Series Comparison, Section 1]{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, \ac{MI}: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 \subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/2_spm_34_12_7}}
-\caption[Time Series Comparison, Section 2]{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, Mutual Information: $0.04$, Residual: $1.02$.}
+\caption[Time Series Comparison, Section 2]{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, \ac{MI}: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 \subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/3_spm_23_21_7}}
-\caption[Time Series Comparison, Section 3]{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, Mutual Information: $-0.03$, Residual: $0.81$.}
+\caption[Time Series Comparison, Section 3]{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, \ac{MI}: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 \subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/4_spm_26_15_7}}
-\caption[Time Series Comparison, Section 4]{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, Mutual Information: $0.06$, Residual: $0.95$. }
+\caption[Time Series Comparison, Section 4]{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, \ac{MI}: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_29_9_13}}
-\caption[Time Series Comparison, Section 5]{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, Mutual Information: $0.02$, Residual: $1.14$.}
-%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. Mutual Information of $0.0212822$, $t$-Value
+\caption[Time Series Comparison, Section 5]{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, \ac{MI}: $0.02$, Residual: $1.14$.}
+%\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. \ac{MI} of $0.0212822$, $t$-Value
 %of $4.17399$ and $MSE$ of $1.14171$.}
@@
 \subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/6_spm_36_17_19}}
-\caption[Time Series Comparison, Section 6]{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, Mutual Information: $.34$, Residual: $0.78$.}
+\caption[Time Series Comparison, Section 6]{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, \ac{MI}: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 %\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_25_34_25}}
-\caption[Time Series Comparison, Section 7]{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, Mutual Information: $0.11$, Residual: $1.10$.}
+\caption[Time Series Comparison, Section 7]{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, \ac{MI}: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }
@@
 
-Section 1 (\autoref{fig:comp1}) had a very high $t$-score, high mutual information and
+Section 1 (\autoref{fig:comp1}) had a very high $t$-score, high \ac{MI} and
 a low error. Visual inspection makes it obvious that this voxel fits the \ac{BOLD} model. A region
@@
 These two voxels are false positives by the normalized residual.
-Note that in both cases the Mutual Information was extremely low, meaning that
-according to the M.I. these were not good fits. Both section 3 and 4 have
+Note that in both cases the \ac{MI} was extremely low, meaning that
+according to the \ac{MI} these were not good fits. Both section 3 and 4 have
 significant peaks, raising the question of whether they may be active,
@@
 Section 7 is an ambiguous time series. While the thresholds applied here
-are empirical, this voxel is on the borderline for both mutual information
+are empirical, this voxel is on the borderline for both \ac{MI}
 and the normalized residual.
@@
 The signal oscillates far faster than the \ac{BOLD} estimate, yet
-somehow the mutual information reaches $.1052$ and the normalized residual is just below $1.1$.
+somehow the \ac{MI} reaches $.1052$ and the normalized residual is just below $1.1$.
 
@@
 contain the parameter maps for the system and \autoref{fig:realhist} is a histogram across all voxels
-for which the mutual information was greater than $0.15$. As before, the threshold is not
+for which the \ac{MI} was greater than $0.15$. As before, the threshold is not
 scientifically derived, yet in tests this threshold provided a decent balance to remove most
@@
 correlate input with output, which is why a threshold was applied.
-Mutual information was chosen over the residual because the
+\ac{MI} was chosen over the residual because the
 maps had more coherency, and had less error in slice simulations (\autoref{fig:sim_hm2} vs.
@@
 \includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=16cm]{images/realhist}
-\caption{Histogram of parameters in active regions ($M.I. > .15$).}
+\caption{Histogram of parameters in active regions (\acs{MI} $> .15$).}
 \label{fig:realhist}
@@
 successfully estimating the \ac{BOLD} output, in spite of the lack of spatial
-smoothing. Comparing the mutual information metric with the normalized residual,
-M.I. gives much better clusters, although that doesn't necessarily mean it
+smoothing. Comparing the \acl{MI} metric with the normalized residual,
+\ac{MI} gives much better clusters, although that doesn't necessarily mean it
 is better. Although many of the differences between \ac{SPM} and the particle filter

### 2010-09-02 doc/paper/main.tex
small fixes

@@
 $P(X_t | \Theta, Y_t)$ was calculated using a particle filter 
-\cite{Johnston2008}. This however was quite slow, and the results
+\cite{Johnston2007}. This however was quite slow, and the results
 were inconsistent with other similar works. 
-Murray used a particle filter, similar to \cite{Johnston2008}, but
+Murray used a particle filter, similar to \cite{Johnston2007}, but
 held the model parameters constant; thus all error in the output
@@
 explain most features of the BOLD signal, it stopped short of proposing a
-realistic waveform for the CBF and CMRO2 \cite{Buxton1998}. Friston et al. 
+realistic waveform for the CBF and CMRO2 \cite{Buxton1998}. Friston et al. \cite{Friston2000}
 gave a reasonable and simple expression for CBF input based on a flow 
@@
 of estimating metabolic rate: as a linear function of the inward 
-blood flow \cite{Friston2000}. By combining these equations with 
+blood flow. By combining these equations with 
 the balloon model from Buxton et al., it is possible to predict 
@@
 and non-Gaussian. This section based primarily on 
-Arulampalam et al. and Thrun et al. \cite{Arulampalam2002a, Thrun2005}.
+Arulampalam et al. \cite{Arulampalam2002a} and Thrun et al. \cite{thrun2008probabilistic}.
 
@@
 an ideal measurement of degeneracy based on the particles' true 
-weight is given \cite{Liu1998b}. An approximation of 
+weight is given \cite{Liu1998}. An approximation of 
 the effective number of particles is also given: 
@@
 of Friston et al. may have been wider than true parameter distribution
-\cite{Friston2002}. Future works that attempt to calculate parameters
+\cite{Friston2002b}. Future works that attempt to calculate parameters
 of the BOLD model should be run multiple times to ensure consistency,

### 2010-09-03 doc/thesis/resultsB.tex
fixed one graph and added the pdf

@@
 \centering
-\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=16cm]{images/realhist}
-\caption{Histogram of parameters in active regions (\acs{MI} $> .15$).}
+\includegraphics[clip=true,trim=6cm 1cm 5cm 1cm,width=\textwidth]{images/realhist}
+\caption{Histogram of parameters in active regions (\acs{MI} $> .15$) overlayed
+with the parameter estimates from Friston et al. \cite{Friston2002b}}
 \label{fig:realhist}

### 2010-10-21 doc/isbi/main.tex
more on isbi

% Template for ISBI-2011 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage[printonlyused,nolist]{acronym}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black]{hyperref}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{An Analysis of Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
%
% Single address.
% ---------------
\name{Chambers, Micah C}
\address{Virginia Tech\\
    Bradley Department of Electrical and Computer Engineering\\
    302 Whittemore Hall\\
    Blacksburg, VA 24061-0111}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
% More than two addresses
% -----------------------
% \name{Author Name$^{\star \dagger}$ \qquad Author Name$^{\star}$ \qquad Author Name$^{\dagger}$}
%
% \address{$^{\star}$ Affiliation Number One \\
%     $^{\dagger}$}Affiliation Number Two
%

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
The Blood-Oxygen-Level-Dependent (BOLD) Signal that is measured by functional 
magnetic resonance imaging (fMRI)
has been the subject of extensive research since the first development 
of the balloon model. While there would be definite benefits to 
moving from the General Linear Model to a physiologically inspired
BOLD model, significant barriers remain. Optimizing even the simplest
balloon model requires searching in 7 dimensions, and even more flexible
models are available. Whereas traditional methods of analyzing fMRI asks
where activation occurs, BOLD models would ask what the activation 
looks like. Unfortunately, the nonlinear nature of the Balloon Model
makes it difficult to analyze, therefore this work demonstrates the 
use of a particle filter to regresses the simplest form of the BOLD 
model. The results show that the system of equations are not observable,
leading to a large range of valid model parameters. 
\end{abstract}
%
\begin{keywords}
BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
\end{keywords}
%

\begin{acronym}[CMRO2]
\acro{BOLD}{Blood-Oxygen-Level-Dependent}
\acro{CBF}{Cerebral Blood Flow}
\acro{CBV}{Cerebral Blood Volume}
\acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
\acro{DC}{Direct Current}
\acro{dHb}{Deoxygenated Hemoglobin}
\acro{EM}{Expectation-Maximization}
\acro{EPI}{Echo Planar Imaging}
\acro{fMRI}{Functional Magnetic Resonance Imaging}
\acro{FMRIB}{Oxford Centre for Functional MRI of the Brain}
\acro{FSL}{FMRIB Software Library}
\acro{FWHM}{Full-Width Half-Maximum}
\acro{GA}{Genetic Algorithms}
\acro{GLM}{General Linear Model}
\acro{HRF}{Hemodynamic Response Function}
\acro{Hb}{Hemoglobin}
\acro{MAD}{Median Absolute Deviation}
\acro{MI}{Mutual Information}
\acro{MR}{Magnetic Resonance}
\acro{MRI}{Magnetic Resonance Imaging}
\acro{MSE}{Mean Squared Error}
\acro{O2Hb}{Oxygenated Hemoglobin}
\acro{ODE}{Ordinary Differential Equation}
\acro{PDF}{Probability Density Function}
\acro{POSSUM}{Physics-Oriented Simulated Scanner for Understanding MRI}
\acro{RF}{Radio Frequency}
\acro{RMSE}{Root Mean Squared Error}
\acro{RMSR}{Root Mean Squared Residual}
\acro{SA}{Simulated Annealing}
\acro{SNR}{Signal-to-Noise Ratio}
\acro{SPM}{Statistical Parametric Mapping}
\acro{T1}{Longitudinal}
\acro{T2}{Spin-Spin}
\acro{TR}{Repetition Time}
\acro{UKF}{Unscented Kalman Filter}
\end{acronym}
\section{Introduction}
\label{sec:intro}
Traditional methods of analyzing timeseries images produced by 
\ac{fMRI} perform regression using the linear combination of explanatory variables. 
Though moving to a state space model with more more degrees of freedom 
requires additional computation, in this paper Particle Filters will
be used to estimate the governing parameters of the \ac{BOLD} model 
at a computation cost 
that would still allow real time calculations for multiple voxels.
In practical terms, this algorithm could be used as an alternative to \ac{SPM}
for activation detection. Though more computationally intense,
this method is capable of modeling nonlinear effects
and provides more detailed output.

In particular this work focuses
on the observability of the \ac{BOLD} state equations from Friston et al.
\cite{Friston2000}. Because of the nonlinearities in the state equations,
no analytical method exists to determine observability, however little
movement in parameters was necessary to achieve high predictive quality from a
set parameters. High correlation as calculated from the joint \ac{PDF}
also indicates a lack of observability.  Future works will benefit from
restricting parameters and thus reducing the dimensionality of the system.
While this could reduce the plausibility of parameter estimates, it would
allow for better differentiation of parameters for analyzing pathologies
and reduce reduce computation time from the 40 seconds of analysis
per 5 minute timeseries (on a Core 2 Duo Q6600).

\subsection{BOLD Model}
It is well known that the two types of \ac{Hb} act as contrast agents in 
\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
between \ac{dHb}/\ac{O2Hb} and neural activity is non-trivial. 
Intuitively, increased 
metabolism will increase \ac{dHb}, however blood vessels are quick
to compensate by increasing local blood flow. Increased inflow, accomplished by loosening 
capillary beds, precedes increased outflow, driving increased 
blood storage.
Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
increased blood volume affects this ratio if 
metabolism doesn't exactly match the increased inflow of oxygenated blood.
This was the impetus
for the ground breaking balloon model \cite{Buxton1998} and windkessel
model \cite{Mandeville1999}. These works derive, from first principals,
the changes in \ac{dHb} ratio and volume of capillaries given an inflow waveform.
These were the first two attempts to quantitatively account for the shape of the 
\ac{BOLD} signal as a consequence of the lag between the \ac{CBV}
and the inward \ac{CBF}.

Although Buxton et al. demonstrated that a well chosen flow waveform could 
explain most features of the \ac{BOLD} signal, it stopped short of proposing a
realistic waveform for the \ac{CBF} and \ac{CMRO2} \cite{Buxton1998}. Friston et al. 
gave a reasonable and simple
expression for \ac{CBF} input based on a flow inducing signal
and in the same work proposed a simple method
of estimating metabolic rate: as a direct function of the inward blood flow \cite{Friston2000}.
By combining these equations with the balloon model from Buxton et al.,
it is possible to predict the \ac{BOLD} signal directly from a stimulus time course:
\begin{eqnarray}
\dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
\dot{f} &=& s\\
\dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
\dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
\label{eq:bold}
\end{eqnarray}
where $s$ is a flow inducing signal, $f$ is the input \ac{CBF},
$v$ is normalized \ac{CBV}, and $q$ is the normalized
local \ac{dHb} content. The 
parameters controlling blood flow are $\epsilon$, which is a neuronal 
efficiency term, $u(t)$ which is the stimulus, and $\tau_f$, $\tau_s$ 
which are time constants. The parameters for the evolution of blood 
volume are $E_0$ which the resting metabolic
rate and $\alpha$ which is Grubb's parameter controlling the balloon model. 
$\tau_0$ is a single time constant controlling the speed of $v$ and $q$.

Obata refined the readout equation 
of the \ac{BOLD} signal based on the
\ac{dHb} content (q) and local blood volume (v), resulting in the
final \ac{BOLD} measurement \autoref{eq:boldout} \cite{Obata2004}.
\begin{eqnarray}
\label{eq:boldout}
y   &=& V_0((k_1 + k_2)(1-q) - (k_2 + k_3)(1-v))\\
k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8 \nonumber\\
K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57 \nonumber \\
k_3 &=& \epsilon_0 - 1 = .43 \nonumber
\end{eqnarray}
Where $\nu_0 = 40.3 s^{-1}$  is the frequency offset in Hz for fully
de-oxygenated blood (at 1.5T), $r_0 = 25 s^{-1}$  is the slope relating
change in relaxation rate with change in blood oxygenation, $TE$ is the
echo time used by the \ac{EPI} sequence and $\epsilon_0 = 1.43$ is the 
ratio of signal \ac{MR} from intravascular to extravascular regions at rest.

While this model is more accurate than the static hemodynamic model used in \ac{SPM},
there have been significant additions which add more degrees of freedom \cite{Deneux2006}. 
In Deneux et al. it was found that the parameters are far from perpendicular,
and that very different parameters could give nearly identical \ac{BOLD} output;
this work builds on that by calculating a joint distribution for the
model parameters of this basic version of the model\cite{Deneux2006}.

There have been many previous attempts to learn the parameters of the
Balloon model, although none have been extremely successful. 
Friston et al. proposed a novel combination of 
linear and nonlinear modeling to generate parameter estimates
\cite{Friston2002b}.  Thus Friston et al. approximated the Jacobian, 
$\frac{\partial Y}{\partial \theta}$, by approximating the response
using a low-order Volterra Kernel \cite{Friston2002b}. 
Unfortunately, estimating a Volterra kernel
takes longer than simply integrating the state variables. Thus,
this method simplified to a generic nonlinear-least squares algorithm.

In Johnston et al. \cite{Johnston2007}, a hybrid particle filter/gradient
descent algorithm was used to simultaneously derive the static and dynamic 
parameters, (classically known as parameters and state variables, respectively)
A particle filter was used to calculate the state variables at each
time; then the estimated distribution of the particles was used to find
the most likely set of parameters that would give that distribution of state variables.
This process was repeated until the parameters converged. This
method is significantly more complicated than simply using the particle
filter to calculate the parameters; and certainly takes longer to run.

Hu et al. used an Unscented Kalman Filter to simultaneously calculate
the model parameters and state variables \cite{Hu2009}.
However, because the Kalman Filter assumes a multivariate 
Gaussian for the state variables, $X(t-1)$, the posterior 
distribution is limited. Additionally, a nonlinear transformation
of a Gaussian can be significantly non-Gaussian.

In Vakorin et al., a combination of Genetic Algorithms and 
Simulated Annealing was used to estimate not only the parameters, but the true 
stimuli driving the \ac{BOLD} signal  \cite{Vakorin2007}. 
This addresses the inherent uncertainty of exactly where and when 
stimuli actually get applied. Unfortunately this algorithm took
in more than 16 hours per voxel.

In his PhD thesis, Murray \cite{Murray2008} used a particle filter based 
approach to integrate
the \ac{BOLD} equations. The method used in that work focused primarily on estimating
the \ac{BOLD} output and state equations as a nonlinear stochastic differential 
equation. The primary difference between that work and this is that
Murray \cite{Murray2008} took the parameters as a given. Thus, differences in the \ac{BOLD} output
were taken to be primarily driven by stochastic changes in the underlying state
equations. Because the parameters were not allowed to change, the estimate of 
the \ac{BOLD} signal was not very good. The filtering framework created
in Murray \cite{Murray2008}, dysii, forms the basis for the particle 
filter used in this work, 
and was well designed. The work also clearly presents the particle filter;
both its derivation and use.

\section{Methods}
\label{sec:Methods}
\subsection{Particle Filter}
\label{sec:ParticleFilter}
Particle filters, a type of Sequential Monte Carlo (SMC) method,
are a powerful method for on line estimation of the posterior probability 
distribution of parameters given a timeseries and a model. The concept of 
particle filters is similar to Kalman Filters; however,
distributions are stored as an empirical distribution rather than 
as the first two moments of a Gaussian. Thus, particle filters are 
preferred when the model is nonlinear, and by implication non-Gaussian. 
The particle filters are a well established technique and are 
described in great detail in Arulampalam et al. and Thrun et al. 
\cite{Arulampalam2002a, thrun2008probabilistic}.

The particle filter was set with each particle having one possible 
system realization: $\{\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f,
\epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
from the prior, with parameters independently gamma distributed:
$\tau_0 \sim \Gamma(.98, .25)$, 
$\alpha \sim \Gamma(.33, .045)$, $E_0    \sim \Gamma(.34, .03)$,
$V_0    \sim \Gamma(.04, .03)$, $\tau_s \sim \Gamma(1.54, .25)$,
$\tau_f \sim \Gamma(2.46, .25)$, $\epsilon \sim \Gamma(.7, .6)$.
To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
were assumed to be at resting state ($\{0, 1, 1, 1\}$). Although
it is common to add noise to the state during integration (to 
emulate noise in the state updates), this was not performed because 
the cloud of particles is able to account for this to some degree.

\subsection{Configuration}
The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
included noise in the update of $x$, however that is not included here.
The reason for the difference is that the cloud of particles is, to some extent,
able to account for that noise. It is common, however, to model that noise
in particle filters by adding a random value to each updated state variable.
Because the purpose of this particle filter is to learn the underlying distribution
of the static parameters, rather than precisely modeling the time course of the
in the dynamic state variables $\{s,f,v,q\}$ this noise is left out. It also helps
that detrending is applied before the particle filter and that the
\ac{BOLD} model is dissipative. When no stimuli are applied, all the particles
decay to $\{0,1,1,1\}$. Typical particle filters
also use this state noise as an exploratory measure; however, this method is
less necessary when good priors are available, and when regularized particle filtering
is being used.

For all the analysis  in this work, $1400$ integration points
per second were used.  Typically a step size of $0.001$ was sufficient,
however, from time to time $0.001$ can still result in problems for the
\ac{BOLD} model.

\subsection{Preprocessing}
\label{sec:Methods Preprocessing}
The normal pipeline for analyzing
\ac{fMRI} involves a several preprocessing steps to condition the signal.
The first and most important
task is motion correction. To do this, a single volume in time is chosen, and
volumes at every other time point are realigned to best match the target volume. This corrects
for motion by the patient as well as small changes in the magnetic
fields that cause the image to shift.
In conventional statistical parametric mapping, a Gaussian smoothing
filter is applied across the image as discussed in \autoref{sec:RFT}.
After this, detrending is performed, which is discussed in \autoref{sec:Detrend}.
Recall that \ac{fMRI} signal levels are unit-less and though detrending is not
always necessary, the data must always be converted
into \% difference from baseline.
The generally accepted method is to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.
Before going into the detrending used in this work, it is necessary to
discuss the type of noise present in \ac{fMRI}.

The non-stationary
aspect of a Wiener process, presumably the result of integrating some
$\nu_x$, is difficult to compensate for, and so many methods
have been developed to compensate for it. Tanabe et al. \cite{Tanabe2002}
and Smith et al. \cite{Smith1999}
demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
of \ac{fMRI}. It has been reported that in some studies as
many as half the voxels
benefited from detrending \cite{Smith2007}. In comparison of high pass 
filters, 
Tanabe et al. \cite{Tanabe2002} showed that in most cases subtracting off
a spline worked the best.
Unfortunately no method will
perfectly remove noise, and no method will leave the signal untouched.

The method used to calculate the spline was one knot for every 20
measurements in an image. Thus a 10 minute session at a repetition time of
2.1 seconds would have 19 knots. The first and last knots were each
given half the number of samples as the rest of the knots; which were all
located at the center of their sample group. The median of each sample group
was then taken and used as the magnitude for the group. Taking the median
versus the mean seemed to work better, given the presence of outliers.
There is potential to optimize the spline further using a canonical
HRF to find resting points; however, the experiment would have
to be designed with this in mind.

Problematically, after removing the \ac{DC} component of the signal,
by definition the signal will have a median near zero.
Unfortunately this is not the natural state of the \ac{BOLD} signal. More specifically,
when the signal is inactive, the \ac{BOLD} response should be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
Because most of the \ac{BOLD} signal is above baseline, after removing the spline
the resting state will be below 0\%.
One method of accounting for this is to simply add a \ac{DC} gain model parameter.
Like all the other model parameters, with enough measurements, the particle filter
would be able to settle on a good estimate. Yet adding another dimension increases the
complexity of the model, for a parameter that is relatively easy to estimate
by visual inspection.  In this work a simpler approach was used. To determine
the \ac{DC} gain a robust estimator of scale was used. The 
\ac{MAD} proved to be accurate in determining how much to shift the signal up
by. Both methods analysis were tested, and it was found that the increase
in model complexity far outweighed the slight increase in flexibility. Other
methods may work better, however the \ac{MAD} worked well,
as \autoref{fig:PreprocessedLowNoise} and 
\autoref{fig:PreprocessedHighNoise} show. The \ac{DC} gain was then set to:

\begin{equation}
y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
\label{eq:mad}
\end{equation}

For more information on the \ac{MAD}, it is discussed in great detail in
Iglewicz's text on robust estimation \cite{iglewicz1983robust}.
A serious concern when adding constant values to
real data is whether this will create false positives. This is a legitimate
concern; however, a boosted response does not affect how well the \ac{BOLD} model
predicts the actual measurements; and as mentioned before, the \ac{DC} signal 
of \ac{fMRI} is never used.

\section{Results}
\label{sec:Results}

\section{Conclusion}
\label{sec:Conclusion}

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
%  \centerline{\includegraphics[width=8.5cm]{image1}}
%  \vspace{2.0cm}
  \centerline{(a) Result 1}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
%  \centerline{\includegraphics[width=4.0cm]{image3}}
%  \vspace{1.5cm}
  \centerline{(b) Results 3}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
%  \centerline{\includegraphics[width=4.0cm]{image4}}
%  \vspace{1.5cm}
  \centerline{(c) Result 4}\medskip
\end{minipage}
%
\caption{Example of placing a figure with experimental results.}
\label{fig:res}
%
\end{figure}

% To start a new column (but not a new page) and help balance the last-page
% column length use \vfill\pagebreak.
% -------------------------------------------------------------------------
%\vfill
%\pagebreak

\section{References}
\label{sec:ref}

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{/home/micahc/library}

\end{document}

### 2010-10-21 doc/isbi/examples/particle.py

@@ -1,0 +1,105 @@
+#!/usr/bin/python
+
+import sys
+from scipy.stats.distributions import norm, cauchy
+import scipy
+import numpy 
+import pylab as P
+from math import sqrt
+import random
+
+def plus(a,b):
+    return a+b
+def multiply(a,b):
+    return a*b
+
+def normalize(arr):
+    tot = sum(arr)*20/1500
+    return [val/tot for val in arr]
+
+def sample(pos, weights):
+    sm = sum(weights)
+    weights = [w/sm for w in weights]
+    var = random.uniform(0,1)
+    agg = sum(weights)
+    i = 0
+    while agg > var:
+        agg = agg - weights[i]
+        i = i+1
+    return pos[i]
+
+def smooth(pos, weights):
+    tog = sorted(zip(pos, weights), key=lambda val: val[0])
+    xvals = scipy.linspace(-10,10,num=1500)
+    pdf = [0 for val in xvals]
+    j = 0
+    for i in range(0, len(pdf)):
+        while j < len(tog) and xvals[i] > tog[j][0]:
+            pdf[i] = pdf[i] + tog[j][1]
+            j = j + 1
+    gauss = [norm.pdf(i/100., 2, .2) for i in range(0,400)]
+    return (xvals, numpy.convolve(pdf, gauss, mode="same"))
+
+#time 1
+P.figure(1)
+P.subplot(411)
+P.title("Update 1")
+xvals = scipy.linspace(-10,10,num=1500)
+measmean = 1
+meas = normalize([norm.pdf(i, measmean, 2) for i in xvals])
+P.plot([val for val in xvals], meas)
+
+state = normalize([cauchy.pdf(i, -2, 1) +cauchy.pdf(i, 2, 1) for i in xvals])
+P.plot([val for val in xvals], state)
+
+state = normalize(map(multiply, state, meas))
+P.plot([val for val in xvals], state)
+P.hist([val for val in xvals], weights=[-1./3000 for val in xvals], color="b")
+P.legend(["Measurement", "State Predict", "Updated State", "Support"], loc=2)
+
+#time 2
+P.subplot(412)
+P.title("Update 2")
+#meas = normalize([norm.pdf(i, measmean, 2) for i in xvals])
+P.plot([val for val in xvals], meas)
+
+#state = normalize([cauchy.pdf(i, -2, 1) +cauchy.pdf(i, 2, 1) for i in xvals])
+P.plot([val for val in xvals], state)
+
[…]

### 2010-10-22 doc/isbi/main.tex
added some more images. and about half done with ISBI

@@
 \usepackage{spconf,amsmath,graphicx}
+\graphicspath{{pdf/}{png/}}
 \usepackage[printonlyused,nolist]{acronym}
 \usepackage[colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
+\usepackage{colortbl}
 
@@
 \begin{eqnarray}
+\label{eq:bold1}
 \dot{s} &=& \epsilon u(t) - \frac{s}{\tau_s} - \frac{f - 1}{\tau_f} \\
+\label{eq:bold2}
 \dot{f} &=& s\\
+\label{eq:bold3}
 \dot{v} &=& \frac{1}{\tau_0}(f - v^\alpha)\\
+\label{eq:bold4}
 \dot{q} &=& \frac{1}{\tau_0}(\frac{f(1-(1-E_0)^f)}{E_0} - \frac{q}{v^{1-1/\alpha}})
-\label{eq:bold}
 \end{eqnarray}
@@
 To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
-were assumed to be at resting state ($\{0, 1, 1, 1\}$). Although
+were assumed to be at resting state, $\{0, 1, 1, 1\}$. Although
 it is common to add noise to the state during integration (to 
@@
 the cloud of particles is able to account for this to some degree.
+Initially each particle is weighted equally, however
+as new measurements arrive, this effects the probability of particles.
+New measurements
+are incorporated into the weights by \autoref{eq:weightevolve}:
+\begin{equation}
+w^i_k & \propto & w^i_{k-1}P(y_k| x^i_k) 
+\label{eq:weightevolve}
+\end{equation}
+where $i$ is the particle index, and $k$ is the time index. $P(y_k | x^i_k)$
+is the probability of the observed measurement coming from the $i^{th}$ 
+particle and depends on the observation function, \autoref{eq:boldout}.
+$x^i_k$ is calculated from $x^i_{k-1}$ using the balloon model, 
+\autoref{eq:bold1} through \autoref{eq:bold4}.
+For all the analysis  in this work, $1400$ Euler integration points
+per TR ($2.1$ s) were used.
 
-\subsection{Configuration}
-The original assumption regarding particle filter models (\autoref{sec:Particle Filter Model})
-included noise in the update of $x$, however that is not included here.
-The reason for the difference is that the cloud of particles is, to some extent,
-able to account for that noise. It is common, however, to model that noise
-in particle filters by adding a random value to each updated state variable.
-Because the purpose of this particle filter is to learn the underlying distribution
-of the static parameters, rather than precisely modeling the time course of the
-in the dynamic state variables $\{s,f,v,q\}$ this noise is left out. It also helps
-that detrending is applied before the particle filter and that the
-\ac{BOLD} model is dissipative. When no stimuli are applied, all the particles
-decay to $\{0,1,1,1\}$. Typical particle filters
-also use this state noise as an exploratory measure; however, this method is
-less necessary when good priors are available, and when regularized particle filtering
-is being used.
+Because fMRI drift can add structure to the noise, and removing
+drift could adversely affect regression, a spline with 1 knot every
+20 measurements was fitted to and then subtracted from the timeseries
+\autoref{Smith1999, Tanabe2002}. The level was then converted
+to \% difference from the original mean. Since the BOLD signal is 
+predominantly positive, a constant (the Gaussian normed \acl{MAD} 
+of the \% difference signal) was added to each point.
 
-For all the analysis  in this work, $1400$ integration points
-per second were used.  Typically a step size of $0.001$ was sufficient,
-however, from time to time $0.001$ can still result in problems for the
-\ac{BOLD} model.
-
-\subsection{Preprocessing}
-\label{sec:Methods Preprocessing}
-The normal pipeline for analyzing
-\ac{fMRI} involves a several preprocessing steps to condition the signal.
-The first and most important
-task is motion correction. To do this, a single volume in time is chosen, and
-volumes at every other time point are realigned to best match the target volume. This corrects
-for motion by the patient as well as small changes in the magnetic
-fields that cause the image to shift.
-In conventional statistical parametric mapping, a Gaussian smoothing
-filter is applied across the image as discussed in \autoref{sec:RFT}.
-After this, detrending is performed, which is discussed in \autoref{sec:Detrend}.
-Recall that \ac{fMRI} signal levels are unit-less and though detrending is not
-always necessary, the data must always be converted
-into \% difference from baseline.
-The generally accepted method is to use a high pass filter, although the
-cutoff frequency is application dependent and often applied haphazardly.
-Before going into the detrending used in this work, it is necessary to
-discuss the type of noise present in \ac{fMRI}.
-
-The non-stationary
-aspect of a Wiener process, presumably the result of integrating some
-$\nu_x$, is difficult to compensate for, and so many methods
-have been developed to compensate for it. Tanabe et al. \cite{Tanabe2002}
-and Smith et al. \cite{Smith1999}
-demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
-of \ac{fMRI}. It has been reported that in some studies as
-many as half the voxels
-benefited from detrending \cite{Smith2007}. In comparison of high pass 
-filters, 
-Tanabe et al. \cite{Tanabe2002} showed that in most cases subtracting off
-a spline worked the best.
-Unfortunately no method will
-perfectly remove noise, and no method will leave the signal untouched.
-
-The method used to calculate the spline was one knot for every 20
-measurements in an image. Thus a 10 minute session at a repetition time of
-2.1 seconds would have 19 knots. The first and last knots were each
-given half the number of samples as the rest of the knots; which were all
-located at the center of their sample group. The median of each sample group
-was then taken and used as the magnitude for the group. Taking the median
-versus the mean seemed to work better, given the presence of outliers.
-There is potential to optimize the spline further using a canonical
-HRF to find resting points; however, the experiment would have
-to be designed with this in mind.
-
-Problematically, after removing the \ac{DC} component of the signal,
-by definition the signal will have a median near zero.
-Unfortunately this is not the natural state of the \ac{BOLD} signal. More specifically,
-when the signal is inactive, the \ac{BOLD} response should be at 0\% change from
-the base level; activation may then increase, or for short periods decrease from this base.
-Because most of the \ac{BOLD} signal is above baseline, after removing the spline
-the resting state will be below 0\%.
-One method of accounting for this is to simply add a \ac{DC} gain model parameter.
-Like all the other model parameters, with enough measurements, the particle filter
-would be able to settle on a good estimate. Yet adding another dimension increases the
-complexity of the model, for a parameter that is relatively easy to estimate
-by visual inspection.  In this work a simpler approach was used. To determine
-the \ac{DC} gain a robust estimator of scale was used. The 
-\ac{MAD} proved to be accurate in determining how much to shift the signal up
-by. Both methods analysis were tested, and it was found that the increase
-in model complexity far outweighed the slight increase in flexibility. Other
-methods may work better, however the \ac{MAD} worked well,
-as \autoref{fig:PreprocessedLowNoise} and 
-\autoref{fig:PreprocessedHighNoise} show. The \ac{DC} gain was then set to:
-
-\begin{equation}
-y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
-\label{eq:mad}
-\end{equation}
-
-For more information on the \ac{MAD}, it is discussed in great detail in
-Iglewicz's text on robust estimation \cite{iglewicz1983robust}.
-A serious concern when adding constant values to
-real data is whether this will create false positives. This is a legitimate
-concern; however, a boosted response does not affect how well the \ac{BOLD} model
-predicts the actual measurements; and as mentioned before, the \ac{DC} signal 
-of \ac{fMRI} is never used.
+Given the state-space equations for the \ac{BOLD} signal, simulating a single time
+series is straightforward. After generating a true signal,
+identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
+process with Gaussian I.I.D. steps were added to the true signal. Finally a
+carrier level was added, since \ac{BOLD} is typically
+measured as a \% difference from the baseline. The particle filter
+algorithm immediately removes this by calculating the \% difference,
+but adding a carrier level meant that the exact same algorithm used
+for simulated data could be used for the real data. For single voxel
+analysis the noise was kept relatively low ($\sigma = 0.001$  for
+I.I.D. Gaussian Noise, and $\sigma = .0005$ for Wiener Steps) to
+explore the properties of the model. These noise magnitudes were added
+the the BOLD signal which is typically around $0.01-0.03$ (1-3\%). 
+To simulate a slice with POSSUM, the code was modified to
+output activation levels based on sets of parameters rather than
+activation levels. The only noise added was the noise normally
+added by POSSUM which does not include drift. 
+For the POSSUM simulated data, SNR was significantly lower ($<6dB$).
 
@@
 \label{sec:Results}
+A primary advantage of the particle filter is the that calculates
+a complete joint distribution of the parameters and states. This 
+allows for advanced analyses not available to many other methods.
+Thus, the correlation matrix of parameters from a single 40 minute
+simulated fMRI regression are shown in. %\autoref{
+\begin{table}[t]
+\ninept
+  \centering
+\begin{tabular}{|c | c  c  c  c  c  c  |}
+\hline
+  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
+\hline
+$\alpha$                      & 0.880& & & & & \\
+\rowcolor[gray]{.8} $E_0$     & -0.766& -0.523& & & & \\
+$V_0$                         & 0.624& 0.424& -0.796& & & \\
+\rowcolor[gray]{.8} $\tau_s$  & 0.620& 0.295& -0.748& 0.344& & \\
+$\tau_f$                      & 0.000& -0.397& -0.431& 0.196& 0.699& \\
+\rowcolor[gray]{.8} $\epsilon$& 0.616& 0.656& -0.641& 0.285& 0.446& -0.097\\
+\hline
+\end{tabular}
+  \caption{Correlation Matrix of Posterior Distribution for 40 minute simulated BOLD signal.}
+\label{tab:long_corr}
+\end{table}[htb]
+\normalsize
 
-\section{Conclusion}
+For POSSUM simulations, \autoref{fig:PossumResult} shows the
+the mutual information between the estimated timeseries and the 
+preprocessed signal. The estimate was generated from the 
+mean of the particles' parameters at the end of the run.
+
+\begin{figure}[htb]
+\begin{minipage}[b]{.5\linewidth}
+  \centering
+  \centerline{\includegraphics[width=\textwidth]{snr_hm.png}}
+  \centerline{(a) SNR}\medskip
+\end{minipage}
+\hfill
+\begin{minipage}[b]{.49\linewidth}
+  \centering
+  \centerline{\includegraphics[width=\textwidth]{sim_hm_mi.png}}
+  \centerline{(b) Mutual Information}\medskip
+\end{minipage}
+\caption{Results of POSSUM simulated data. $a)$ 
+SNR of POSSUM additive noise. $b)$ Mutual Information between
+the signal predicted by the particle filter and the actual signal.}
+\label{fig:PossumResult}
+\end{figure}
+
+\section{Discussion}
 \label{sec:Conclusion}
 
-% Below is an example of how to insert images. Delete the ``\vspace'' line,
-% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
-% with a suitable PostScript file name.
-% -------------------------------------------------------------------------
-\begin{figure}[htb]
-
-\begin{minipage}[b]{1.0\linewidth}
-  \centering
-%  \centerline{\includegraphics[width=8.5cm]{image1}}
-%  \vspace{2.0cm}
-  \centerline{(a) Result 1}\medskip
-\end{minipage}
-%
-\begin{minipage}[b]{.48\linewidth}
-  \centering
-%  \centerline{\includegraphics[width=4.0cm]{image3}}
-%  \vspace{1.5cm}
-  \centerline{(b) Results 3}\medskip
-\end{minipage}
-\hfill
-\begin{minipage}[b]{0.48\linewidth}
-  \centering
-%  \centerline{\includegraphics[width=4.0cm]{image4}}
-%  \vspace{1.5cm}
-  \centerline{(c) Result 4}\medskip
-\end{minipage}
-%
-\caption{Example of placing a figure with experimental results.}
-\label{fig:res}
-%
-\end{figure}
-
-% To start a new column (but not a new page) and help balance the last-page
-% column length use \vfill\pagebreak.
-% -------------------------------------------------------------------------
-%\vfill
-%\pagebreak
-
-\section{References}
-\label{sec:ref}
-
-% References should be produced using the bibtex program from suitable
-% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
-% style file from IEEE produces unsorted bibliography list.
-% -------------------------------------------------------------------------
 \bibliographystyle{IEEEbib}

### 2010-10-24 doc/isbi/main.tex
added another chart to main.tex
getting close to the end

@@
 \label{sec:Results}
-A primary advantage of the particle filter is the that calculates
+A primary advantage of the particle filter is that it estimates
 a complete joint distribution of the parameters and states. This 
-allows for advanced analyses not available to many other methods.
-Thus, the correlation matrix of parameters from a single 40 minute
-simulated fMRI regression are shown in. %\autoref{
+allows for more detailed analyses not available to many other methods.
+In particular it is possible to calculate the correlation of the
+parameters in the posterior distribution. The correlation matrix
+for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
 \begin{table}[t]
@@
 \end{table}[htb]
-\normalsize
 
@@
 the mutual information between the estimated timeseries and the 
-preprocessed signal. The estimate was generated from the 
+preprocessed signal. Each estimate was generated from the 
 mean of the particles' parameters at the end of the run. 
+Additionally the mean and variance of the parameter estimates
+(mean) across the spatial regions of \autoref{fig:PossumResult} are
+given in \autoref{tab:beforeafter}.
 
@@
 
+\begin{table}[t]
+\ninept
+  \centering
+\begin{tabular}{| c | c | c | c | c | c |}
+\hline
+& Sim Value & \multicolumn{2}{|c|}{Prior} &\multicolumn{2}{|c|}{Posterior} \\
+\hline
+        &       &$\mu$  & $\sigma$& $\mu$& $\sigma$\\
+\hline
+$\tau_0  $& 1.45  & .98   & .25     & 1.07319 & 0.297749  \\
+\rowcolor[gray]{.8} 
+$\alpha  $&0.3    & .33   & .045    & 0.317955& 0.07311  \\
+$E_0     $& 0.47  & .34   & .03     & 0.364195& 0.0457019  \\
+\rowcolor[gray]{.8} 
+$V_0     $& 0.044 & .04   & .03     & 0.110542& 0.0500039  \\
+$\tau_s  $& 1.94  & 1.54  & .25     & 1.86785 & 0.390455  \\
+\rowcolor[gray]{.8} 
+$\tau_f  $& 1.99  & 2.46  & .25     & 2.50166 & 0.348326  \\
+$\epsilon$& 1.8   & .7    & .6      & 1.58716 & 1.14536  \\
+\hline
+\end{tabular}
+\caption{Comparison of the true value (left), the prior distribution
+used to generate the initial particles (center two columns) and the 
+average/variance of the estimates across simulated voxels. Note that
+each voxel used the same underlying parameters (with differences only
+in absolute value and noise realization), and all particle filters
+were initialized with the same prior.}
+\label{tab:beforeafter}
+\end{table}[htb]
+
 \section{Discussion}
 \label{sec:Conclusion}
+The data shows substantial correlation present in the parameter 
+estimates for the 40 minute simulated data. Correlation of 
+constants is a definite sign of the model being under-determined
+and thus not observable. This is in spite of low noise, and the 
+particle filter's solution converging to a Root-Mean-Squared error
+of less than $0.002$.
 
+On a larger scale, the particle filter also performed well, clearly
+locating regions of activation in spite of Signal-to-Noise Ratio 
+well below 6dB. In spite of this, the final parameter estimates 
+varied more than the initial (prior) distribution (\autoref{tab:beforeafter}).
+This effect is slightly deceiving because the distributions
+were far from Gaussian, and many were bi-modal. It is 
+interesting, however, that such a wide range of solutions
+could give good estimates for essentially one BOLD signal.
+
+From the results, it is clear that the BOLD model as presented
+by Friston et al. is not observable, and thus parameter estimates
+suffer from very large variance \cite{Friston2000}. As a result, parameter estimates
+in that work are likely the result in the under-determination of
+the model rather than underlying differences in parameters. 
+There are two possible ways to deal with this problem. First, as
+specified in Deneux et al., certain parameters could be locked into
+place \cite{Deneux2006}. This solution is relatively easy, and would
+also reduce computation time. True in-vivo parameter estimation
+would help to determine proper values for parameters that are to be
+kept constant, although it is a difficult proposition. The other
+possible method is adding another measurement point. Technology exists
+to simultaneously measure fMRI and \ac{CBF}. This additional
+measurement, which is slightly closer to original stimulus, could
+make the Balloon Model observable.
+
+\small
 \bibliographystyle{IEEEbib}

### 2010-10-25 doc/isbi/main.tex
Material in, now time to clean up.

@@
 that would still allow real time calculations for multiple voxels.
-In practical terms, this algorithm could be used as an alternative to \ac{SPM}
-for activation detection. Though more computationally intense,
+The parameters guiding the BOLD model could provide insight into 
+functional differences between patients and brain regions.
+More practically, this algorithm is an alternative to the General
+Linear Model for activation detection. Though more computationally intense,
 this method is capable of modeling nonlinear effects
@@
 
-In particular this work focuses
+This work focuses
 on the observability of the \ac{BOLD} state equations from Friston et al.
@@
 movement in parameters was necessary to achieve high predictive quality from a
-set parameters. High correlation as calculated from the joint \ac{PDF}
+set of parameters. High correlation as calculated from the joint \ac{PDF}
 also indicates a lack of observability.  Future works will benefit from
@@
 \epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
-from the prior, with parameters independently gamma distributed:
+from the prior, however after resampling the number was dropped
+to 1000. The parameters were drawn independently from the gamma distribution:
 $\tau_0 \sim \Gamma(.98, .25)$, 
@@
 the model rather than underlying differences in parameters. 
-There are two possible ways to deal with this problem. First, as
+There are two ways to deal with this problem. First, as
 specified in Deneux et al., certain parameters could be locked into
@@
 would help to determine proper values for parameters that are to be
-kept constant, although it is a difficult proposition. The other
-possible method is adding another measurement point. Technology exists
-to simultaneously measure fMRI and \ac{CBF}. This additional
-measurement, which is slightly closer to original stimulus, could
-make the Balloon Model observable.
+kept constant, although such estimation is not currently possible. 
+A better way of dealing with this problem is adding measurement
+points. Methods of simultaneously measuring fMRI and \ac{CBF}. 
+This additional measurement, which is slightly closer to 
+original stimulus, could make the Balloon Model observable.
+
+Despite the lack of observability, the BOLD estimates generated 
+by the particle filter stayed near the noise level. 
+Thus, for the purpose of estimating BOLD levels between measurements
+the particle filter is sufficient. The on-line nature of the
+particle filter could be utilized, along with BOLD estimates,
+for adaptive control. Thus the stimulus could be
+adapted during the fMRI scan based on BOLD estimates within a small neural
+region. The particle filter also has the distinct benefit of 
+estimating a full posterior distribution. A joint distribution of 
+parameters allows for more advanced analysis; such as calculating
+non-parametric hypothesis testing.  
+While the particle filter took a day to calculate for full brain calculations, its speed
+was sufficient on a quad core machine to perform real time calculations of small regions
+(approximate run time .27 seconds per voxel-measurement).
 

### 2010-10-28 doc/isbi/main.tex
fixes needed draft complete
final ISBI pdf

@@
 % ---------------
-\name{Chambers, Micah C}
+\name{M. Chambers, C. Wyatt}
 \address{Virginia Tech\\
@@
 \begin{abstract}
-The Blood-Oxygen-Level-Dependent (BOLD) Signal that is measured by functional 
+The Blood-Oxygen-Level-Dependent (BOLD) signal that is measured by functional 
 magnetic resonance imaging (fMRI)
 has been the subject of extensive research since the first development 
-of the balloon model. While there would be definite benefits to 
-moving from the General Linear Model to a physiologically inspired
+of the balloon model. While there are definite benefits to 
+moving from the Canonical Hemodynamic Response function to a physiologically inspired
 BOLD model, significant barriers remain. Optimizing even the simplest
-balloon model requires searching in 7 dimensions, and even more flexible
-models are available. Whereas traditional methods of analyzing fMRI asks
-where activation occurs, BOLD models would ask what the activation 
-looks like. Unfortunately, the nonlinear nature of the Balloon Model
+balloon model requires searching within 7 dimensions, and even more complex
+models exist. Whereas traditional methods of analyzing fMRI aims
+to determine where activation occurs, BOLD models seek a parametric representation
+of the signal. Unfortunately, the nonlinear nature of these models
 makes it difficult to analyze, therefore this work demonstrates the 
@@
 model. The results show that the system of equations are not observable,
-leading to a large range of valid model parameters. 
+leading to a large range of parameters that are consistent with the measurements. 
 \end{abstract}
@@
 \begin{keywords}
-BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
+BOLD Response, Functional MRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
 \end{keywords}
@@
 \label{sec:intro}
-Traditional methods of analyzing timeseries images produced by 
-\ac{fMRI} perform regression using the linear combination of explanatory variables. 
+Traditional methods of analyzing 
+\ac{fMRI} images perform regression using the linear combination of explanatory variables. 
 Though moving to a state space model with more more degrees of freedom 
@@
 More practically, this algorithm is an alternative to the General
-Linear Model for activation detection. Though more computationally intense,
+Linear Model for activation detection. Though computation intensive,
 this method is capable of modeling nonlinear effects
@@
 on the observability of the \ac{BOLD} state equations from Friston et al.
-\cite{Friston2000}. Because of the nonlinearities in the state equations,
-no analytical method exists to determine observability, however little
-movement in parameters was necessary to achieve high predictive quality from a
-set of parameters. High correlation as calculated from the joint \ac{PDF}
-also indicates a lack of observability.  Future works will benefit from
-restricting parameters and thus reducing the dimensionality of the system.
-While this could reduce the plausibility of parameter estimates, it would
-allow for better differentiation of parameters for analyzing pathologies
-and reduce reduce computation time from the 40 seconds of analysis
-per 5 minute timeseries (on a Core 2 Duo Q6600).
+\cite{Friston2000}. 
+Because of the nonlinearities in the state equations,
+no analytical method exists to determine observability. Therefore
+Particle Filters are used to build a joint mixture \ac{PDF} of the model parameters. 
+This work will show that this method is sufficient for predicting
+the BOLD signal, without requiring excessive computation. 
+Because particle filters estimate a full posterior distribution, 
+it was possible to calculate correlation of model parameters, an
+indicator of observability.
 
@@
 It is well known that the two types of \ac{Hb} act as contrast agents in 
-\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}, however the connection
+\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}. Despite this, the connection
 between \ac{dHb}/\ac{O2Hb} and neural activity is non-trivial. 
@@
 metabolism will increase \ac{dHb}, however blood vessels are quick
-to compensate by increasing local blood flow. Increased inflow, accomplished by loosening 
-capillary beds, precedes increased outflow, driving increased 
-blood storage.
+to compensate by permitting increased local blood flow. Increased inflow, accomplished by loosening 
+capillary beds, precedes a matching outflow, driving increased 
+blood volume.
 Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
-increased blood volume affects this ratio if 
-metabolism doesn't exactly match the increased inflow of oxygenated blood.
+increased blood volume also affects this ratio. 
 This was the impetus
 for the ground breaking balloon model \cite{Buxton1998} and windkessel
-model \cite{Mandeville1999}. These works derive, from first principals,
-the changes in \ac{dHb} ratio and volume of capillaries given an inflow waveform.
-These were the first two attempts to quantitatively account for the shape of the 
-\ac{BOLD} signal as a consequence of the lag between the \ac{CBV}
-and the inward \ac{CBF}.
+model \cite{Mandeville1999}. These works derive 
+the changes in \ac{dHb} ratio and volume of capillaries given \ac{CBF}
+and \ac{CMRO2} waveforms.
 
@@
 explain most features of the \ac{BOLD} signal, it stopped short of proposing a
-realistic waveform for the \ac{CBF} and \ac{CMRO2} \cite{Buxton1998}. Friston et al. 
-gave a reasonable and simple
-expression for \ac{CBF} input based on a flow inducing signal
-and in the same work proposed a simple method
-of estimating metabolic rate: as a direct function of the inward blood flow \cite{Friston2000}.
+deterministic waveform for the \ac{CBF} and \ac{CMRO2} \cite{Buxton1998}. Friston et al. 
+gave a simple expression for \ac{CBF} input based on a flow inducing signal,
+and proposed that metabolic rate should be a linear function of the \ac{CBF} \cite{Friston2000}.
 By combining these equations with the balloon model from Buxton et al.,
-it is possible to predict the \ac{BOLD} signal directly from a stimulus time course:
+it is possible to predict the \ac{BOLD} signal from a stimulus function:
 \begin{eqnarray}
@@
 which are time constants. The parameters for the evolution of blood 
-volume are $E_0$ which the resting metabolic
+volume are $E_0$ which is the resting metabolic
 rate and $\alpha$ which is Grubb's parameter controlling the balloon model. 
@@
 k_1 &=& 4.3 \times \nu_0 \times E_0 \times TE = 2.8 \nonumber\\
-K_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57 \nonumber \\
+k_2 &=& \epsilon_0 \times r_0 \times E_0 \times TE = .57 \nonumber \\
 k_3 &=& \epsilon_0 - 1 = .43 \nonumber
@@
 ratio of signal \ac{MR} from intravascular to extravascular regions at rest. 
-
-While this model is more accurate than the static hemodynamic model used in \ac{SPM},
-there have been significant additions which add more degrees of freedom \cite{Deneux2006}. 
-In Deneux et al. it was found that the parameters are far from perpendicular,
+In Deneux et al. \cite{Deneux2006} it was found that the parameters are far from perpendicular,
 and that very different parameters could give nearly identical \ac{BOLD} output;
 this work builds on that by calculating a joint distribution for the
-model parameters of this basic version of the model\cite{Deneux2006}.
+model parameters of this version of the model.
 
 There have been many previous attempts to learn the parameters of the
-Balloon model, although none have been extremely successful. 
+balloon model, although none enjoy the ubiquity of the \ac{GLM}. 
 Friston et al. proposed a novel combination of 
@@
 takes longer than simply integrating the state variables. Thus,
-this method simplified to a generic nonlinear-least squares algorithm.
+this method was simplified to a generic nonlinear-least squares algorithm.
 
@@
 descent algorithm was used to simultaneously derive the static and dynamic 
-parameters, (classically known as parameters and state variables, respectively)
-A particle filter was used to calculate the state variables at each
-time; then the estimated distribution of the particles was used to find
-the most likely set of parameters that would give that distribution of state variables.
+parameters, (more generally known as parameters and state variables, respectively).
+A particle filter was used to calculate the time course of the state variables.
+The estimated distribution of the state variables was then used in
+a Maximum Likelihood fashion to find the most likely set of parameters. 
 This process was repeated until the parameters converged. This
-method is significantly more complicated than simply using the particle
-filter to calculate the parameters; and certainly takes longer to run.
+method is  more complex than using the particle
+filter to calculate the parameters, takes longer to run and
+does not account for uncertainty in parameters.
 
@@
 Gaussian for the state variables, $X(t-1)$, the posterior 
-distribution is limited. Additionally, a nonlinear transformation
-of a Gaussian can be significantly non-Gaussian.
+distribution is limited. Additionally, a nonlinear transformation of 
+a Gaussian, 
+such as the state transition function of the Balloon Model,
+can be significantly non-Gaussian.
 
@@
 stimuli driving the \ac{BOLD} signal  \cite{Vakorin2007}. 
-This addresses the inherent uncertainty of exactly where and when 
-stimuli actually get applied. Unfortunately this algorithm took
-in more than 16 hours per voxel.
+This addresses the inherent uncertainty of exactly when 
+stimuli get applied. Unfortunately this algorithm took
+more than 16 hours per voxel to run.
 
@@
 Murray \cite{Murray2008} took the parameters as a given. Thus, differences in the \ac{BOLD} output
-were taken to be primarily driven by stochastic changes in the underlying state
-equations. Because the parameters were not allowed to change, the estimate of 
-the \ac{BOLD} signal was not very good. The filtering framework created
+were assumed to be driven by noise in the underlying state
+equations. Because the parameters were held constant, the 
+\ac{BOLD} estimate was poor. However, the primary purpose of that work
+was to advance understanding of filtering techniques; and the filtering framework created
 in Murray \cite{Murray2008}, dysii, forms the basis for the particle 
-filter used in this work, 
-and was well designed. The work also clearly presents the particle filter;
-both its derivation and use.
+filter used in this work.
 
@@
 distribution of parameters given a timeseries and a model. The concept of 
-particle filters is similar to Kalman Filters; however,
+particle filters is similar to that of the Unscented Kalman Filter; however,
 distributions are stored as an empirical distribution rather than 
 as the first two moments of a Gaussian. Thus, particle filters are 
-preferred when the model is nonlinear, and by implication non-Gaussian. 
+preferred when the model is nonlinear, and non-Gaussian. 
 The particle filters are a well established technique and are 
-described in great detail in Arulampalam et al. and Thrun et al. 
-\cite{Arulampalam2002a, thrun2008probabilistic}.
-
-The particle filter was set with each particle having one possible 
-system realization: $\{\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f,
-\epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
-from the prior, however after resampling the number was dropped
-to 1000. The parameters were drawn independently from the gamma distribution:
-$\tau_0 \sim \Gamma(.98, .25)$, 
-$\alpha \sim \Gamma(.33, .045)$, $E_0    \sim \Gamma(.34, .03)$,
-$V_0    \sim \Gamma(.04, .03)$, $\tau_s \sim \Gamma(1.54, .25)$,
-$\tau_f \sim \Gamma(2.46, .25)$, $\epsilon \sim \Gamma(.7, .6)$.
-To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
-were assumed to be at resting state, $\{0, 1, 1, 1\}$. Although
-it is common to add noise to the state during integration (to 
-emulate noise in the state updates), this was not performed because 
-the cloud of particles is able to account for this to some degree.
-Initially each particle is weighted equally, however
-as new measurements arrive, this effects the probability of particles.
-New measurements
-are incorporated into the weights by \autoref{eq:weightevolve}:
-\begin{equation}
-w^i_k & \propto & w^i_{k-1}P(y_k| x^i_k) 
-\label{eq:weightevolve}
-\end{equation}
-where $i$ is the particle index, and $k$ is the time index. $P(y_k | x^i_k)$
-is the probability of the observed measurement coming from the $i^{th}$ 
-particle and depends on the observation function, \autoref{eq:boldout}.
-$x^i_k$ is calculated from $x^i_{k-1}$ using the balloon model, 
-\autoref{eq:bold1} through \autoref{eq:bold4}.
-For all the analysis  in this work, $1400$ Euler integration points
-per TR ($2.1$ s) were used.
-
-Because fMRI drift can add structure to the noise, and removing
-drift could adversely affect regression, a spline with 1 knot every
-20 measurements was fitted to and then subtracted from the timeseries
-\autoref{Smith1999, Tanabe2002}. The level was then converted
-to \% difference from the original mean. Since the BOLD signal is 
-predominantly positive, a constant (the Gaussian normed \acl{MAD} 
-of the \% difference signal) was added to each point.
-
-Given the state-space equations for the \ac{BOLD} signal, simulating a single time
-series is straightforward. After generating a true signal,
-identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
-process with Gaussian I.I.D. steps were added to the true signal. Finally a
-carrier level was added, since \ac{BOLD} is typically
-measured as a \% difference from the baseline. The particle filter
-algorithm immediately removes this by calculating the \% difference,
-but adding a carrier level meant that the exact same algorithm used
-for simulated data could be used for the real data. For single voxel
-analysis the noise was kept relatively low ($\sigma = 0.001$  for
-I.I.D. Gaussian Noise, and $\sigma = .0005$ for Wiener Steps) to
-explore the properties of the model. These noise magnitudes were added
-the the BOLD signal which is typically around $0.01-0.03$ (1-3\%). 
-To simulate a slice with POSSUM, the code was modified to
-output activation levels based on sets of parameters rather than
-activation levels. The only noise added was the noise normally
-added by POSSUM which does not include drift. 
-For the POSSUM simulated data, SNR was significantly lower ($<6dB$).
-
-\section{Results}
-\label{sec:Results}
-A primary advantage of the particle filter is that it estimates
-a complete joint distribution of the parameters and states. This 
-allows for more detailed analyses not available to many other methods.
-In particular it is possible to calculate the correlation of the
-parameters in the posterior distribution. The correlation matrix
-for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
+described in great detail in Arulampalam et al., Thrun et al. 
+and Murray \cite{Arulampalam2002a, thrun2008probabilistic, Murray2008}.
 \begin{table}[t]
@@
 \label{tab:long_corr}
-\end{table}[htb]
+\end{table}
+The particle filter was set with each particle representing one possible 
+system realization: $\{\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f,
+\epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
+from the prior, however after resampling the number was dropped
+to 1000. This gave the prior sufficient density to
+cover parameter space, but saved computation time once the solution space
+became more compact. 
+The parameters were drawn independently from the gamma distribution:
+$\tau_0 \sim \Gamma(.98, .25)$, 
+$\alpha \sim \Gamma(.33, .045)$, $E_0    \sim \Gamma(.34, .03)$,
+$V_0    \sim \Gamma(.04, .03)$, $\tau_s \sim \Gamma(1.54, .25)$,
+$\tau_f \sim \Gamma(2.46, .25)$, $\epsilon \sim \Gamma(.7, .6)$.
+To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
+were assumed to start at resting state, $\{0, 1, 1, 1\}$. Although
+it is common to add noise to the state during integration (to 
+emulate noise in the state updates), this was not performed because 
+regularization during resampling tended to expand the distribution.
+Additionally the cloud of particles is able to account for some amount
+of stationary randomness.
+Initially each particle is weighted equally, however
+as new measurements arrive, they alter the probability
+of particles being a solution.
+New measurements
+are incorporated into the weights by \autoref{eq:weightevolve}:
+\begin{equation}
+w^i_k & \propto & w^i_{k-1}P(y_k| x^i_k) 
+\label{eq:weightevolve}
+\end{equation}
+where $i$ is the particle index, and $k$ is the time index. $P(y_k | x^i_k)$
+is the probability of the observed measurement coming from the $i^{th}$ 
+particle and depends on the measurement noise and the observation function, 
+\autoref{eq:boldout}.
+$x^i_k$ is calculated from $x^i_{k-1}$ using the balloon model, 
+\autoref{eq:bold1} through \autoref{eq:bold4}.
+For all the analysis  in this work, $1400$ Euler integration points
+per TR ($2.1$ s) were used. When the number of particles with significant
+(non-zero) weights dropped below 50, regularized resampling was 
+applied. Thus, a new set of particles was drawn from the smoothed
+version of the mixture \ac{PDF}.
 
-For POSSUM simulations, \autoref{fig:PossumResult} shows the
-the mutual information between the estimated timeseries and the 
-preprocessed signal. Each estimate was generated from the 
-mean of the particles' parameters at the end of the run. 
-Additionally the mean and variance of the parameter estimates
-(mean) across the spatial regions of \autoref{fig:PossumResult} are
-given in \autoref{tab:beforeafter}.
+Because fMRI drift can add structure to the noise, a spline with 1 knot every
+20 measurements was fitted to and then subtracted from the timeseries
+\cite{Smith1999, Tanabe2002}. The level was then converted
+to \% difference from the original mean. Since the BOLD signal is 
+predominantly positive, a constant (the Gaussian normed \acl{MAD} 
+of the \% difference signal) was added to each point.
 
-\begin{figure}[htb]
+Given the state-space equations for the \ac{BOLD} signal, simulating a single time
+series is straightforward. After generating a true signal,
+identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
+process with Gaussian I.I.D. steps were added to the true signal. Finally a
+carrier level was added, since \ac{BOLD} is 
+measured as a \% difference from the baseline. For single voxel
+analysis, the noise was kept relatively low ($\sigma = 0.001$  for
+I.I.D. Gaussian Noise, and $\sigma = .0005$ for Wiener Steps) to
+explore the properties of the model. This noise was added
+to the the BOLD signal typically peaks at $0.01-0.03$ (1-3\%). 
+To simulate a slice with \ac{POSSUM} \cite{Drobnjak}, it's code was modified to
+output activation levels based on sets of parameters rather than
+activation levels. For the POSSUM simulated data, SNR was significantly 
+lower although the noise did not include drift ($<6dB$).
+
+\section{Results}
+\label{sec:Results}
+A primary advantage of the particle filter is that it estimates
+a complete joint distribution of the parameters and states. This 
+allows for more detailed analyses not available to many other methods.
+In particular it is possible to calculate the correlation of the
+parameters in the posterior distribution. The correlation matrix
+for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
+\begin{figure}
 \begin{minipage}[b]{.5\linewidth}
@@
 \end{figure}
-
 \begin{table}[t]
@@
 \end{tabular}
-\caption{Comparison of the true value (left), the prior distribution
-used to generate the initial particles (center two columns) and the 
-average/variance of the estimates across simulated voxels. Note that
-each voxel used the same underlying parameters (with differences only
-in absolute value and noise realization), and all particle filters
-were initialized with the same prior.}
+\caption{Comparison of the true value (left), the initial (prior) distribution
+(center two columns) and the 
+mean/variance of the estimates across simulated voxels (Right two Columns).} 
 \label{tab:beforeafter}
-\end{table}[htb]
+\end{table}
+For POSSUM simulations, \autoref{fig:PossumResult} shows the
+the mutual information between the estimated timeseries and the 
+preprocessed signal. Note that in order to remove all false
+positives a threshold of $0.15$ bits was applied, 
+the regions of significant SNR are visible despite this.
+Each estimate was generated from the 
+mean of the particles' parameters at the end of the run. 
+Additionally the mean and variance of the parameter estimates
+(mean) across the spatial regions of \autoref{fig:PossumResult} are
+given in \autoref{tab:beforeafter}.
 
@@
 locating regions of activation in spite of Signal-to-Noise Ratio 
-well below 6dB. In spite of this, the final parameter estimates 
+well below 6dB. However, the final parameter estimates 
 varied more than the initial (prior) distribution (\autoref{tab:beforeafter}).
 This effect is slightly deceiving because the distributions
-were far from Gaussian, and many were bi-modal. It is 
-interesting, however, that such a wide range of solutions
-could give good estimates for essentially one BOLD signal.
+were far from Gaussian, and many were bimodal. It is 
+significant, however, that such a wide range of solutions
+could give good estimates for the true BOLD signal.
 
@@
 by Friston et al. is not observable, and thus parameter estimates
-suffer from very large variance \cite{Friston2000}. As a result, parameter estimates
-in that work are likely the result in the under-determination of
-the model rather than underlying differences in parameters. 
+suffer from very large variance \cite{Friston2000}. Therefore
+any work attempting to estimate these parameters from the BOLD
+signal alone is likely measuring the variance due the model
+rather than the variance due to try physiological differences. 
 There are two ways to deal with this problem. First, as
-specified in Deneux et al., certain parameters could be locked into
-place \cite{Deneux2006}. This solution is relatively easy, and would
-also reduce computation time. True in-vivo parameter estimation
-would help to determine proper values for parameters that are to be
-kept constant, although such estimation is not currently possible. 
-A better way of dealing with this problem is adding measurement
-points. Methods of simultaneously measuring fMRI and \ac{CBF}. 
-This additional measurement, which is slightly closer to 
+specified in Deneux et al., certain parameters could be held constant
+\cite{Deneux2006}. This solution is relatively easy, and would
+also reduce computation time. A better way of dealing with this 
+problem is simultaneously measuring fMRI and \ac{CBF}.  
+This additional measurement, which is closer to 
 original stimulus, could make the Balloon Model observable.
@@
 by the particle filter stayed near the noise level. 
-Thus, for the purpose of estimating BOLD levels between measurements
+Thus, for the purpose of estimating BOLD levels between measurements,
 the particle filter is sufficient. The on-line nature of the
-particle filter could be utilized, along with BOLD estimates,
+particle filter could also be utilized, along with BOLD estimates,
 for adaptive control. Thus the stimulus could be
@@
 estimating a full posterior distribution. A joint distribution of 
-parameters allows for more advanced analysis; such as calculating
+parameters allows for more advanced analysis; such as 
 non-parametric hypothesis testing.  
-While the particle filter took a day to calculate for full brain calculations, its speed
+While the particle filter took a day of processing for full brain calculations, its speed
 was sufficient on a quad core machine to perform real time calculations of small regions
@@
 
-\small
 \bibliographystyle{IEEEbib}
-\bibliography{/home/micahc/library}
+\bibliography{library}
 

### 2010-12-13 doc/thesis/background.tex
updated thesis for first ETD

@@
 \caption[Canonical Hemodynamic Response Function]
-{Canonical Hemodynamic Response Function, normalization makes $y$ units
+{Canonical Hemodynamic Response Function. Normalization makes $y$ units
 arbitrary}
@@
 \caption[Example Kalman Filter Progression]
-{a) Sigma Points are passed through the state/measurement equations
+{Example Kalman Filter Progression. a) Sigma Points are passed through the state/measurement equations
 to get a state prediction (green line). The first two moments of
@@
 \caption{Parameters used to test Gaussianity of variables after being transitioned through
-the \ac{BOLD} model}
+the \ac{BOLD} model.}
 \label{tab:steptable} 

### 2010-12-13 doc/thesis/intro.tex
updated thesis for first ETD

--- 2010-09-01 doc/thesis/intro.tex
@@
 \centering
-\includegraphics[width=.5\textwidth]{images/bold}
-\caption[Diagram of BOLD Signal]{A) End of Pre-Stimulus Undershoot, $< 2s$ typical,
+\includegraphics[width=.5\textwidth]{images/bold2}
+\caption[Diagram of \ac{BOLD} Signal]{Diagram of the \ac{BOLD} Sigmal.
+A) End of Pre-Stimulus Undershoot, $< 2s$ typical,
 B) End of Stimulus, C) Input Bloodflow Drops from Supernormal to Subnormal,
 $5 s$ after B typical, D) Beginning of Post-stimulus Undershoot, $12 s$ after B
-typical, E) Full return to normal, between $20-30s$ after B. \cite{Behzadi2005}}
+typical, E) Full return to normal, between $20-30s$ after B.}
 \end{figure}
 
-\section{\ac{fMRI}}
+\section{Functional Magnetic Resonance Imaging}
 \ac{MRI}, is a method of building 3D images
@@
 
-\section{Properties of the \ac{BOLD} Model}
+\section{Properties of the Blood-Oxygen-Level Dependent Model}
 \label{sec:BOLD Analysis}
@@
 \end{tabular}
-\caption{Parameters found in various studies. (NC) indicates that the value
+\caption[Parameters found in various studies.]{Parameters found in various studies. (NC) indicates that the value
 wasn't calculated. \cite{Vakorin2007} made use of the values from \cite{Friston2000}

### 2010-12-13 doc/thesis/methods.tex
more on isbi
updated thesis for first ETD

@@
 \subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009d_37_29_24}}
-\caption[Resting State Noise Analysis, Steps]{Resting state noise analysis, where
+\caption[Resting state noise analysis, where signal levels are the difference between adjacent signals in \autoref{fig:QQDC}]{Resting state noise analysis, where
 signal levels are the difference between adjacent signals in \autoref{fig:QQDC}.
@@
 (resting state) signal, Q-Q Plot of measurements with the standard Normal,
-Signal auto-regression, Histogram of the Points}
+Signal auto-regression, Histogram of the Points.}
 \label{fig:QQDelta}
@@
 \subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 1cm,width=13cm]{images/noise2_0009s_37_29_24}}
-\caption[Resting State Noise Analysis, After Subtracting Spline]
+\caption[Resting State Noise Analysis, where
+signal levels are the result of subtracting a spline from the signal in 
+\autoref{fig:QQDC}]
 {Resting State Noise Analysis, where

### 2010-12-13 doc/thesis/particlefilters.tex
added some more images. and about half done with ISBI
updated thesis for first ETD

@@
                          {\cancel{P(x^i_k | x^i_{k-1})}q(x^i_{0:k-1} | y_{0:k-1})} \nonumber \\
-& \propto & w^i_{k-1}P(y_k| x_k) 
+& \propto & w^i_{k-1}P(y_k| x^i_k) 
 \label{eq:weightevolve}
@@
 \includegraphics[width=16cm]{images/particle_filter2}
-\caption[Example Particle Filter Progression]{a) Initial particle locations
+\caption[Example Particle Filter Progression.]{
+Example Particle Filter Progression. a) Initial particle locations
 uniformly distributed in [-10, 10], with

### 2010-12-13 doc/thesis/results.tex
updated thesis for first ETD

@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
-\caption[\acs{BOLD} estimate converging for a very long \acs{fMRI} run.]{\acs{BOLD} estimate 
+\caption[\acs{BOLD} estimate 
+converging for a 2400 second \acs{fMRI} run.]{\acs{BOLD} estimate 
 converging for a 2400 second \acs{fMRI} run. Darker bars indicate
@@
 \includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
-\caption[First 500 seconds of the \acs{BOLD} estimate converging for a very 
-long \acs{fMRI} run.]{First 500 seconds of \acs{BOLD} estimate 
+\caption[First 500 seconds of \acs{BOLD} estimate 
+converging for a 2400 second \acs{fMRI} run.]{First 500 seconds of \acs{BOLD} estimate 
 converging for a 2400 second \acs{fMRI} run. Darker bars indicate
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/realization_lownoise}
-\caption[Simulated Signal with Low Noise/Drift]
-{Simulated signals with different noise realizations compared to true
+\caption[Simulated signals with different noise realizations compared to true
+signal.] {Simulated signals with different noise realizations compared to true
 signal. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_lownoise}
-\caption[Preprocessed Signal with Low Noise/Drift]
+\caption[Results of preprocessing compared to the true signal.]
 {Results of preprocessing compared to the true signal. 
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_lownoise}
-\caption[Results with Low Noise/Drift]
-{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedLowNoise}
+\caption{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedLowNoise}
 as input.}
@@
 \end{tabular}
-\caption[Estimated Parameters with low noise.]
-{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
+\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
 last is the average. $\sum \tau$ is the sum of the estimated time constants.}
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/preprocessed_highnoise}
-\caption[Preprocessed Signal with High Noise/Drift]
+\caption[Results of preprocessing compared to the true signal.]
 {Results of preprocessing compared to the true signal.  
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3.5cm,width=15cm]{images/comparison_highnoise}
-\caption[Results with High Noise/Drift]
-{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedHighNoise}
+\caption{Results of the particles filter with preprocessed signals from \autoref{fig:PreprocessedHighNoise}
 as input.}
@@
 \end{tabular}
-\caption[Estimated Parameters on 11 different runs with high noise.]
+\caption[Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
+last row contains the mean.]
 {Estimated Parameters on 11 different runs with high noise. First row contains the true parameters,
@@
 \includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=15cm]{images/preprocessed_noiseonly}
-\caption[Preprocessed Signal for non-active, low noise signal]
+\caption[Comparison of the preprocessed signals for the low noise signal-free case.
+($\sigma_x = 0.005$, $\sigma_y = 0.01$)]
 {Comparison of the preprocessed signals for the low noise signal-free case.
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
-\caption[Results for non-active, low noise signal]  
+\caption[Comparison of the estimated \ac{BOLD} signal for the low noise signal-free case.
+($\sigma_x = 0.005$, $\sigma_y = 0.01$)]  
 {Comparison of the estimated \ac{BOLD} signal for the low noise signal-free case.
@@
 \end{tabular}
-\caption[Estimated Parameters with low noies, and no signal.]
-{Estimated Parameters on 11 different runs with low noise and no signal present.
+\caption{Estimated Parameters on 11 different runs with low noise and no signal present.
  ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justnoise_fit_0}
-\caption[Single Fit Results for non-active, low noise signal.]
-{Single Fit Results for non-active, low noise signal
- ($\sigma_x = 0.005$, $\sigma_y = 0.01$).}
+\caption {Single Fit Results for non-active, low noise signal ($\sigma_x = 0.005$, $\sigma_y = 0.01$).}
 \label{fig:justnoise_fit_0}
@@
 {\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=\textwidth]{images/justnoise_hist_3}}
-\caption[Convergence of the parameters during a noise-only run.]
+\caption[Convergence of the parameters when the signal consisted purely of low level noise 
+($\sigma_x = 0.01$, $\sigma_y = 0.005$).]
 {Convergence of the parameters when the signal consisted purely of low level noise 
@@
 \includegraphics[clip=true,trim=6cm 2cm 5cm 3cm,width=15cm]{images/fits_noiseonly_high}
-\caption[Results for non-active, high noise signal.]
+\caption[\ac{BOLD} estimates for the non-active, high noise signal ($\sigma_x = 0.1$, $\sigma_y = 0.05$).]
 {\ac{BOLD} estimates for the non-active, high noise signal ($\sigma_x = 0.1$, $\sigma_y = 0.05$). 
@@
 \includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justbignoise_fit_0}
-\caption[Single Fit Results for non-active, high noise signal.]
-{Single Fit Results for non-active, high noise signal
+\caption{Single Fit Results for non-active, high noise signal
  ($\sigma_x = 0.005$, $\sigma_y = 0.01$).}
@@
 \caption[Comparison of activation regions.]
-{Comparison of activation regions, a) grey matter and regions, b) \acs{SNR} of active
+{Comparison of activation regions. a) grey matter and regions, b) \acs{SNR} of active
 regions, c) \acs{RMSR} Map, d) \acs{MI} Map.}
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist1}
-\caption[Histogram of estimated parameters in section 1]{Histogram of estimated parameters in 
-section 1 in voxels 
+\caption{Histogram of estimated parameters in section 1 in voxels 
 with \acs{MI} greater than $0.15$} \label{fig:slicesim_hist1}
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist2}
-\caption[Histogram of estimated parameters in section 2]
+\caption
 {Histogram of estimated parameters in section 2 in voxels with \acs{MI} greater
@@
 \includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist3}
-\caption[Histogram of estimated parameters in section 3]
+\caption
 {Histogram of estimated parameters in section 3 in voxels with \acs{MI} greater

### 2010-12-13 doc/thesis/resultsB.tex
updated thesis for first ETD

@@
 \subfigure{\label{fig:scale_pfilter} \includegraphics[scale=.5]{images/scale2}}
-\caption[Particle Filter Normalized Residual Results]{Particle Filter results measured in normalized error.
+\caption[Particle Filter results measured in normalized error.]
+    {Particle Filter results measured in normalized error.
         Units of match is normalized residual where the lowest (best) levels shown are
@@
 \subfigure{\label{fig:scale_mi} \includegraphics[scale=.5]{images/scale4}}
-\caption[Particle Filter \acs{MI} Results]
+\caption[Particle Filter results measured in \ac{MI}.]
         {Particle Filter results measured in \ac{MI}.
@@
 \subfigure[SPM]{\label{fig:comp1spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/1_spm_37_14_7}}
-\caption[Time Series Comparison, Section 1]{Section 1, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, \ac{MI}: $0.33$, Residual: $0.72$.}
+\caption[Time Series Comparison, Section 1]{Time Series Comparison, Section 1.
+Estimated vs. Actual \ac{BOLD} response. $t$-Score: $10.71$, \ac{MI}: $0.33$, Residual: $0.72$.}
 \label{fig:comp1}
@@
 \subfigure[SPM]{\label{fig:comp2spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/2_spm_34_12_7}}
-\caption[Time Series Comparison, Section 2]{Section 2, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, \ac{MI}: $0.04$, Residual: $1.02$.}
+\caption[Time Series Comparison, Section 2]{Time Series Comparison, Section 2.
+Estimated vs. Actual \ac{BOLD} response. $t$-Score: $6.97$, \ac{MI}: $0.04$, Residual: $1.02$.}
 \label{fig:comp2}
@@
 \subfigure[SPM]{\label{fig:comp3spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/3_spm_23_21_7}}
-\caption[Time Series Comparison, Section 3]{Section 3, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, \ac{MI}: $-0.03$, Residual: $0.81$.}
+\caption[Time Series Comparison, Section 3]{Time Series Comparison, Section 3.
+Estimated vs. Actual \ac{BOLD} response. $t$-Score: $2.85$, \ac{MI}: $-0.03$, Residual: $0.81$.}
 \label{fig:comp3}
@@
 \subfigure[SPM]{\label{fig:comp4spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/4_spm_26_15_7}}
-\caption[Time Series Comparison, Section 4]{Section 4, Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, \ac{MI}: $0.06$, Residual: $0.95$. }
+\caption[Time Series Comparison, Section 4]{Time Series Comparison, Section 4.
+Estimated vs. Actual \ac{BOLD} response. $t$-Score: $0.50$, \ac{MI}: $0.06$, Residual: $0.95$. }
 \label{fig:comp4}
@@
 \subfigure[SPM]{\label{fig:comp5spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_29_9_13}}
-\caption[Time Series Comparison, Section 5]{Section 5, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, \ac{MI}: $0.02$, Residual: $1.14$.}
+\caption[Time Series Comparison, Section 5]{Time Series Comparison, Section 5.
+Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $4.17$, \ac{MI}: $0.02$, Residual: $1.14$.}
 %\caption{Section 5, Below threshold in both particle filter checks, but above threshold in SPM. \ac{MI} of $0.0212822$, $t$-Value
@@
 \subfigure[SPM]{\label{fig:comp6spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/6_spm_36_17_19}}
-\caption[Time Series Comparison, Section 6]{Section 6, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, \ac{MI}: $.34$, Residual: $0.78$.}
+\caption[Time Series Comparison, Section 6]{Time Series Comparison, Section 6.
+Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $2.49$, \ac{MI}: $.34$, Residual: $0.78$.}
 %\caption{Section 6, MI of $0.335504$, T Value: $2.49154$, normalized error: $0.783348$ Not visible in SPM}
@@
 %\subfigure[SPM]{\label{fig:comp7spm} \includegraphics[clip=true,trim=5cm 2.5cm 4cm 3cm,width=.93\textwidth]{images/5_spm_25_34_25}}
-\caption[Time Series Comparison, Section 7]{Section 7, Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, \ac{MI}: $0.11$, Residual: $1.10$.}
+\caption[Time Series Comparison, Section 7]{Time Series Comparison, Section 7.
+Estimated vs. Actual \ac{BOLD} Response. $t$-Score: $1.32$, \ac{MI}: $0.11$, Residual: $1.10$.}
 %\caption{Section 7, $.1052$ in MI, $1.31534$ in SPM (which left this flat) and $1.09992$ normalized $\sqrt{MSE}$. Only visible in MI map. }

### 2010-12-13 doc/thesis/thesis.tex
lots of fixes from the corrections
just need to edit discussion/conclusion, change captions and copy some graphs from defense
just need to fix figure captions for multivoxel cases. And fix bib
finished I think
fixed all the acronyms, think I might be done
updated thesis for first ETD

@@
 \def\subsectionautorefname{Section}
+\usepackage{chapterbib}
 
@@
 
-August 30, 2010\\
+September 14, 2010\\
 Blacksburg, Virginia
@@
 
-Keywords: BOLD Response, FMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
+Keywords: BOLD Response, fMRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
 \\
@@
 \acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
+\acro{DC}{Direct Current}
 \acro{dHb}{Deoxygenated Hemoglobin}
+\acro{EM}{Expectation-Maximization}
 \acro{EPI}{Echo Planar Imaging}
 \acro{fMRI}{Functional Magnetic Resonance Imaging}
+\acro{FMRIB}{Oxford Centre for Functional MRI of the Brain}
+\acro{FSL}{FMRIB Software Library}
+\acro{FWHM}{Full-Width Half-Maximum}
+\acro{GA}{Genetic Algorithms}
 \acro{GLM}{General Linear Model}
@@
 \acro{Hb}{Hemoglobin}
+\acro{MAD}{Median Absolute Deviation}
+\acro{MI}{Mutual Information}
 \acro{MR}{Magnetic Resonance}
 \acro{MRI}{Magnetic Resonance Imaging}
+\acro{MSE}{Mean Squared Error}
 \acro{O2Hb}{Oxygenated Hemoglobin}
+\acro{ODE}{Ordinary Differential Equation}
+\acro{PDF}{Probability Density Function}
+\acro{POSSUM}{Physics-Oriented Simulated Scanner for Understanding MRI}
 \acro{RF}{Radio Frequency}
+\acro{RMSE}{Root Mean Squared Error}
+\acro{RMSR}{Root Mean Squared Residual}
+\acro{SA}{Simulated Annealing}
 \acro{SNR}{Signal-to-Noise Ratio}
@@
 \acro{TR}{Repetition Time}
+\acro{UKF}{Unscented Kalman Filter}
 \end{acronym}
@@
 \pagestyle{myheadings}
-
+\acresetall
 \include{intro}
@@
 \include{discussion}
-\include{conclusion}
+%\include{conclusion}
 
+\clearpage
+\addcontentsline{toc}{chapter}{Bibliography}
 \bibliographystyle{plain}

### 2010-12-13 doc/thesis/examples/kalman.py

@@ -28,7 +28,7 @@ P.plot([-val for val in xvals], statest)
 combined_mean = (statemu + measmean)/2.
 combined_var = sum([var[0]*(var[1] - combined_mean)**2 for var in zip(statest, xvals)])
 combined_var = combined_var/sum(statest) + sum([var[0]*(var[1] - combined_mean)**2 for var in zip(meas, xvals)])/sum(meas)
-print combined_var/2, sqrt(combined_var/2)
+#print combined_var/2, sqrt(combined_var/2)
 state = [norm.pdf(i, combined_mean, sqrt(combined_var/2)) for i in xvals]
 P.plot([-val for val in xvals], state)
 P.legend(["Measurement", "State Predict", "State Precit (Norm)", "Updated State"])
@@ -53,7 +53,7 @@ P.plot([10-val for val in xvals], statest)
 combined_mean = (statemu + measmean)/2.
 combined_var = sum([var[0]*(var[1] - combined_mean)**2 for var in zip(statest, xvals)])
 combined_var = combined_var/sum(statest) + sum([var[0]*(var[1] - combined_mean)**2 for var in zip(meas, xvals)])/sum(meas)
-print combined_var/2, sqrt(combined_var/2)
+#print combined_var/2, sqrt(combined_var/2)
 state = [norm.pdf(i, combined_mean, sqrt(combined_var/2)) for i in xvals]
 P.plot([10-val for val in xvals], state)
 P.legend(["Measurement", "State Predict", "State Precit (Norm)", "Updated State"])
@@ -78,7 +78,7 @@ P.plot([-3 - val for val in xvals], statest)
 combined_mean = (statemu + measmean)/2.
 combined_var = sum([var[0]*(var[1] - combined_mean)**2 for var in zip(statest, xvals)])
 combined_var = combined_var/sum(statest) + sum([var[0]*(var[1] - combined_mean)**2 for var in zip(meas, xvals)])/sum(meas)
-print combined_var/2, sqrt(combined_var/2)
+#print combined_var/2, sqrt(combined_var/2)
 state = [norm.pdf(i, combined_mean, sqrt(combined_var/2)) for i in xvals]
 P.plot([-3 - val for val in xvals], state)
 P.legend(["Measurement", "State Predict", "State Precit (Norm)", "Updated State"])

### 2011-01-26 doc/isbi/main.tex
updated my ISBI paper

@@
 magnetic resonance imaging (fMRI)
-has been the subject of extensive research since the first development 
-of the balloon model. While there are definite benefits to 
+has been the subject of extensive research since the development 
+of the first balloon model. While there are definite benefits to 
 moving from the Canonical Hemodynamic Response function to a physiologically inspired
-BOLD model, significant barriers remain. Optimizing even the simplest
+BOLD model, significant barriers remain. Optimizing the simplest
 balloon model requires searching within 7 dimensions, and even more complex
-models exist. Whereas traditional methods of analyzing fMRI aims
-to determine where activation occurs, BOLD models seek a parametric representation
-of the signal. Unfortunately, the nonlinear nature of these models
-makes it difficult to analyze, therefore this work demonstrates the 
-use of a particle filter to regresses the simplest form of the BOLD 
-model. The results show that the system of equations are not observable,
+models exist. Additionally, the nonlinear nature of these models
+make them difficult to analyze, therefore this work 
+uses a particle filter to regresses a simple form of the BOLD 
+model. Whereas traditional methods of analyzing fMRI aims
+to determine where activation occurs, BOLD model regression seeks a parametric representation
+of the signal. The results show that the particle filter 
+attains a good fit but that the system of equations are not observable,
 leading to a large range of parameters that are consistent with the measurements. 
@@
 \ac{fMRI} images perform regression using the linear combination of explanatory variables. 
-Though moving to a state space model with more more degrees of freedom 
-requires additional computation, in this paper Particle Filters will
+Moving to a state-space model with more degrees of freedom 
+requires additional computation; however, in this paper Particle Filters will
 be used to estimate the governing parameters of the \ac{BOLD} model 
 at a computation cost 
-that would still allow real time calculations for multiple voxels.
-The parameters guiding the BOLD model could provide insight into 
+that would still permit real time calculations for multiple voxels.
+The parameters guiding the BOLD model may provide insight into 
 functional differences between patients and brain regions.
-More practically, this algorithm is an alternative to the General
-Linear Model for activation detection. Though computation intensive,
+More practically, this algorithm provides a non-parametric
+alternative to the General Linear Model. Though computationally intensive,
 this method is capable of modeling nonlinear effects
@@
 Particle Filters are used to build a joint mixture \ac{PDF} of the model parameters. 
-This work will show that this method is sufficient for predicting
+This work will show that the method is sufficient for predicting
 the BOLD signal, without requiring excessive computation. 
@@
 \ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}. Despite this, the connection
-between \ac{dHb}/\ac{O2Hb} and neural activity is non-trivial. 
+between \ac{dHb}/\ac{O2Hb} and neural activity is complex. 
 Intuitively, increased 
-metabolism will increase \ac{dHb}, however blood vessels are quick
-to compensate by permitting increased local blood flow. Increased inflow, accomplished by loosening 
-capillary beds, precedes a matching outflow, driving increased 
-blood volume.
+metabolism increases \ac{dHb}, however blood vessels quickly
+compensate by permitting increased local blood flow. Increased inflow, 
+precedes a matching outflow, which causes increased blood volume.
 Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
 increased blood volume also affects this ratio. 
-This was the impetus
+This relation of \ac{dHb}, local blood flow and local blood volume,
+was the impetus
 for the ground breaking balloon model \cite{Buxton1998} and windkessel
 model \cite{Mandeville1999}. These works derive 
-the changes in \ac{dHb} ratio and volume of capillaries given \ac{CBF}
-and \ac{CMRO2} waveforms.
+the changes in \ac{dHb}/\ac{O2Hb} and blood volume given waveforms for \ac{CBF}
+and \ac{CMRO2}.
 
@@
 as the first two moments of a Gaussian. Thus, particle filters are 
-preferred when the model is nonlinear, and non-Gaussian. 
+preferred when the model is non-Gaussian and nonlinear. 
 The particle filters are a well established technique and are 
@@
 emulate noise in the state updates), this was not performed because 
-regularization during resampling tended to expand the distribution.
-Additionally the cloud of particles is able to account for some amount
+a) regularization during resampling tended to expand the distribution
+and b) the cloud of particles is able to account for some amount
 of stationary randomness.
@@
 
-Because fMRI drift can add structure to the noise, a spline with 1 knot every
-20 measurements was fitted to and then subtracted from the timeseries
-\cite{Smith1999, Tanabe2002}. The level was then converted
-to \% difference from the original mean. Since the BOLD signal is 
-predominantly positive, a constant (the Gaussian normed \acl{MAD} 
-of the \% difference signal) was added to each point.
+Because fMRI drift can add structure to the noise, a spline fit with 1 knot every
+40 seconds was subtracted from the timeseries
+\cite{Smith1999, Tanabe2002}. Since the BOLD signal is 
+predominantly positive, a constant (the Gaussian-normed \acl{MAD})
+was added to each point.
 
@@
 explore the properties of the model. This noise was added
-to the the BOLD signal typically peaks at $0.01-0.03$ (1-3\%). 
+to the BOLD signal which has a typical RMS of $0.01-0.03$ (1-3\%). 
 To simulate a slice with \ac{POSSUM} \cite{Drobnjak}, it's code was modified to
@@
 preprocessed signal. Note that in order to remove all false
-positives a threshold of $0.15$ bits was applied, 
-the regions of significant SNR are visible despite this.
+positives a threshold of $0.15$ bits was applied.
 Each estimate was generated from the 
@@
 \label{sec:Conclusion}
-The data shows substantial correlation present in the parameter 
+The single voxel simulation data (\autoref{tab:long_corr}) 
+shows substantial correlation present in the parameter 
 estimates for the 40 minute simulated data. Correlation of 
@@
 and thus not observable. This is in spite of low noise, and the 
-particle filter's solution converging to a Root-Mean-Squared error
+particle filter's solution (mean) converging to a Root-Mean-Squared error
 of less than $0.002$.
 
-On a larger scale, the particle filter also performed well, clearly
-locating regions of activation in spite of Signal-to-Noise Ratio 
-well below 6dB. However, the final parameter estimates 
-varied more than the initial (prior) distribution (\autoref{tab:beforeafter}).
+In the \ac{POSSUM} simulation, the particle filter clearly
+located regions of activation in spite of a Signal-to-Noise Ratio 
+well below 6dB. However, the variance of the final parameter estimates 
+was greater than the initial (prior) distribution (\autoref{tab:beforeafter}).
 This effect is slightly deceiving because the distributions
 were far from Gaussian, and many were bimodal. It is 
-significant, however, that such a wide range of solutions
+important, however, that such a wide range of solutions
 could give good estimates for the true BOLD signal.
 
-From the results, it is clear that the BOLD model as presented
+From single voxel and single slice results, 
+there is substantial evidence that the BOLD model as presented
 by Friston et al. is not observable, and thus parameter estimates
-suffer from very large variance \cite{Friston2000}. Therefore
-any work attempting to estimate these parameters from the BOLD
-signal alone is likely measuring the variance due the model
-rather than the variance due to try physiological differences. 
+suffer from very large variance \cite{Friston2000}.  
 There are two ways to deal with this problem. First, as
